### 2020-07-16 Code/Rmd/translationelicitation.Rmd
Second commit

---
title             : "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
shorttitle        : "Translation Elicitation"

author: 
  - name          : "Serene Siow"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Rd, Oxford, Oxfordshire OX2 6GG, UK"
    email         : "serene.siow@psy.ox.ac.edu"
  - name          : "Gonzalo García-Castro"
    affiliation   : "2"
  - name          : "Núria Sebastian-Galles"
    affiliation   : "2"
  - name          : "Kim Plunkett"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Experimental Psychology, Oxford University"
  - id            : "2"
    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"

authornote: |
  SS and GGC share first authorship.

abstract: |
  Many studies in bilingualism have explored how form similarity between translation equivalents (i.e. cognate- ness) impacts language processing. Cognates range from being identical to sharing only a few key features. Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity. In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value. A widely used metric for word similarity is Levenshtein distance [@levenshtein1966binary]. This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other. However, a challenge when defining phonological similarity is that some phonemes (e.g. [k] and [g]) are perceptually more similar than others (e.g. [k] and [o]). The standard Lev- enshtein distance calculation treats each change equally. Thus it often overestimates phonological distance between words. We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g. vowel or consonant insertion) and position in the word. To estimate the weights, we are collecting behavioural data in an auditory translation elicitation task. Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words. The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent. Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation. The higher the probability, the higher the perceptual similarity. Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r prepare}
library(dplyr)
library(tidyr)
library(ggplot2)
library(here)
```

# Introduction

# Methods

## Algorithm

## Behavioural experiment

### Participants

### Stimuli

### Procedure

### Design

### Data analysis

We fit a generalised Bayesian hierarchical model, with accuracy as dependent variable and

## Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2020-07-16 Code/Rmd/translationelicitation.tex
Second commit

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{hyperref}
\hypersetup{
            pdftitle={A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity},
            pdfkeywords={keywords},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Translation Elicitation}
\author{Serene Siow\textsuperscript{1}, Gonzalo García-Castro\textsuperscript{2}, Núria Sebastian-Galles\textsuperscript{2}, \& Kim Plunkett\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Experimental Psychology, Oxford University\\\textsuperscript{2} Center for Brain and Cognition, Universitat Pompeu Fabra}
\authornote{SS and GGC share first authorship.

Correspondence concerning this article should be addressed to Serene Siow, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Rd, Oxford, Oxfordshire OX2 6GG, UK. E-mail: serene.siow@psy.ox.ac.edu}
\keywords{keywords\newline\indent Word count: X}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{lineno}

\linenumbers
\usepackage{csquotes}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi

\title{A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity}

\date{}

\abstract{
Many studies in bilingualism have explored how form similarity between translation equivalents (i.e.~cognate- ness) impacts language processing. Cognates range from being identical to sharing only a few key features. Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity. In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value. A widely used metric for word similarity is Levenshtein distance ({\textbf{???}}). This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other. However, a challenge when defining phonological similarity is that some phonemes (e.g. {[}k{]} and {[}g{]}) are perceptually more similar than others (e.g. {[}k{]} and {[}o{]}). The standard Lev- enshtein distance calculation treats each change equally. Thus it often overestimates phonological distance between words. We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g.~vowel or consonant insertion) and position in the word. To estimate the weights, we are collecting behavioural data in an auditory translation elicitation task. Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words. The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent. Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation. The higher the probability, the higher the perceptual similarity. Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.
}

\begin{document}
\maketitle

\hypertarget{methods}{%
\section{Methods}\label{methods}}

We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

\hypertarget{material}{%
\subsection{Material}\label{material}}

\hypertarget{procedure}{%
\subsection{Procedure}\label{procedure}}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

We used R (Version 3.6.3; R Core Team, 2020) for all our analyses.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-R-base}{}%
R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}

\endgroup

\end{document}

### 2020-07-16 Manuscript/draft.Rmd
Second commit

--- 2020-07-16 Code/Rmd/translationelicitation.Rmd
@@
 ---
-title             : "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
-shorttitle        : "Translation Elicitation"
+title: "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
+output: pdf_document
+---
 
-author: 
-  - name          : "Serene Siow"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Rd, Oxford, Oxfordshire OX2 6GG, UK"
-    email         : "serene.siow@psy.ox.ac.edu"
-  - name          : "Gonzalo García-Castro"
-    affiliation   : "2"
-  - name          : "Núria Sebastian-Galles"
-    affiliation   : "2"
-  - name          : "Kim Plunkett"
-    affiliation   : "1"
-affiliation:
-  - id            : "1"
-    institution   : "Department of Experimental Psychology, Oxford University"
-  - id            : "2"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = TRUE)
+```
 
-authornote: |
-  SS and GGC share first authorship.
+```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-abstract: |
+library(dplyr)
+library(magrittr)
+library(tidyr)
+library(stringr)
+library(citr)
+library(ggplot2)
+library(papaja)
+
+```
+
+# Abstract
+
 Many studies in bilingualism have explored how form similarity between translation equivalents (i.e.
-cognate- ness) impacts language processing.
+cognateness) impacts language processing.
 Cognates range from being identical to sharing only a few key features.
@@
 [k] and [o]).
-The standard Lev- enshtein distance calculation treats each change equally.
+The standard Levenshtein distance calculation treats each change equally.
 Thus it often overestimates phonological distance between words.
@@
 The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent.
-Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation.
+Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding translation.
 The higher the probability, the higher the perceptual similarity.
 Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "keywords"
-wordcount         : "X"
-
-bibliography      : ["r-references.bib"]
-
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            : papaja::apa6_pdf
----
-
-```{r setup, include = FALSE}
-library("papaja")
-```
-
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
-```
-
-```{r prepare}
-library(dplyr)
-library(tidyr)
-library(ggplot2)
-library(here)
-```
 
@@
 
-## Algorithm
+## Computing cognateness: Phonological distance across translation equivalents
 
-## Behavioural experiment
+## Behavioural task: Translation elicitation task
 
-### Participants
+### Sample
+
+We collected data from XXX monolingual participants with no history of hearing, motor, or language impairments.
+XXX were English monolinguals (XXX females, age range = XXX, mean age = XXX), and XXX were Spanish monolinguals (XXX females, age range = XXX, mean age = XXX).
+English-monolingual participants were recruited in UK and Spanish monolinual particpiants were recruited in Spain.
+Participants were contacted via email, provided informed consent prior to any data collection, and were retributed with 7€/6X after completing the experiment.
+This experiment was approved by the Ethical Committee of XXXXXXX, with code XXXXXXXXXXXXXX.
+
+The language profile of participants was assesed using a brief *ad hoc* questionnaire were participants reported previous experience and with any second language, if any.
+Participants were asked to report their proficiency in oral comprehension and orthography in Spanish, Catalan, English, and any other non-native languge, if reported.
+Participants with exposure to a non-native language for longer than XXX or a proficiency in any non-native language higher than 2 in a 5-point lickert scale were discarded.
 
@@
 
+### Design
+
 ### Procedure
 
-### Design
+### Apparatus
 
-### Data analysis
-
-We fit a generalised Bayesian hierarchical model, with accuracy as dependent variable and
-
-## Results
+# Results
 
@@
 
-\newpage
+# Appendix
 
-# References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
+## Appendix 1: Session info
+
+```{r sessionInfo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+sessionInfo()
+
 ```
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup

### 2020-07-16 Manuscript/translationelicitation_manuscript-appendix.Rmd
Second commit

---
title: "Methods"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(here)
library(data.table)
library(lubridate)
library(papaja)

logs <- fread(here("Data", "01_logs.csv"), na.strings = "") %>%
    as_tibble() %>%
    mutate(date = as_date(date))
dat <- fread(here("Data", "01_processed.csv"), na.strings = "") %>% 
    as_tibble()

```

# Participants

Data collection took place from `r format(min(logs$date), "%dth %B, %Y")` to  `r format(max(logs$date), "%dth %B, %Y")`. We collected data from `r nrow(logs)` participants ($M_{age}$ = `r printnum(mean(logs$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(sd(logs$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(min(logs$age, na.rm = TRUE))`-`r printnum(max(logs$age, na.rm = TRUE))`). `r filter(logs, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(logs, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(logs, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(logs, country=="Spain", sex =="Female") %>% nrow()` female). Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits). Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value. Participants were asked to complete the experiment in a quiet place with good internet connection. We exluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(logs, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(logs, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION? (Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan? (Sí/No)". However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(logs, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(logs, country %in% "Spain", !vision))/nrow(filter(logs, country=="Spain")))*100)`%. This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(logs$exclude_trials, na.rm = TRUE)`, `r printnum(100*(sum(logs$exclude_trials, na.rm = TRUE)/nrow(logs)))`%). Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).

### 2020-07-16 Code/00_stimuli.R

@@ -1,277 +1,0 @@
-#### 00_stimuli-amplitude: analyse audio amplitude
-
-#### set up ######################################
-
-# load packages
-library(dplyr)
-library(tidyr)
-library(data.table)
-library(purrr)
-library(stringr)
-library(readxl)
-library(tibble)
-library(ggplot2)
-library(ggridges)
-library(gghighlight)
-library(patchwork)
-library(janitor)
-library(audio)
-library(PraatR)
-library(here)
-
-# set parameters
-time_bin_duration <- 0.001
-# time-step (s), 
-formant_arguments <- list(0.001,  5, 5500, time_bin_duration, 50)
-# include frame number, include time, time decimals, include intensity, intensity decimals, include n formants, freq decimals, include bandwidths
-table_arguments <- list(TRUE, TRUE, 3, TRUE, 5, FALSE, 3, TRUE) 
-# time step (s), pitch floor, pitch ceiling
-pitch_arguments <- list(0.001, 50, 800) 
-
-# set paths
-audio_paths <- list.files(here("Stimuli", "Sounds"), full.names = TRUE)
-formant_paths <- str_replace(audio_paths, "Sounds/", "Formants/") %>% str_replace(., ".wav", ".Formant")
-formant_table_paths <- str_replace(audio_paths, "Sounds/", "/Formants/Tables/") %>% str_replace(., ".wav", "_table.txt")
-pitch_path <- str_replace(audio_paths, "Sounds/", "Pitch/") %>% str_replace(., ".wav", ".Pitch")
-pitch_tier_path <- str_replace(audio_paths, "Sounds/", "Pitch/PitchTiers/") %>% str_replace(., ".wav", ".PitchTier")
-
-#### import data #################################
-trials <- read_xlsx(here("Stimuli", "trials.xlsx"))
-subtlex <- fread(here("Data", "00_subtlex.csv")) %>% as_tibble()
-
-#### analyse frequency (SUBTLEX) #################
-frequency <- left_join(trials, subtlex, by = c("orthography" = "word", "language"))
-
-#### analyse formats #############################
-
-map2(.x = audio_paths, .y = formant_paths,
-    ~praat("To Formant (burg)...",
-           arguments = formant_arguments,
-           input = .x,
-           output = .y,
-           overwrite = TRUE))
-
-map2(.x = formant_paths, .y = formant_table_paths,
-     ~praat("Down to Table...",
-            arguments = table_arguments,
-            input = .x,
[…]

### 2020-07-16 R/01_process.R

@@ -1,0 +1,146 @@
+#### 01_preprocess: Pre-process data ################
+
+#### set up #########################################
+
+# load packages
+library(tidyverse)
+library(lubridate)
+library(purrr)
+library(readxl)
+library(data.table)
+library(stringdist)
+library(mice)
+library(lme4)
+library(car)
+library(janitor)
+library(patchwork)
+library(here)
+
+# create/load functions
+source(here("R", "functions.R"))
+
+# set parameters
+practice_trials <- c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)
+spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
+
+#### import data ####################################
+trials <- read_xlsx(here("Stimuli", "trials.xlsx")) # trial-level data
+
+# participant files
+filenames <- list.files(here("Data", "Raw")) %>% 
+    str_extract(".*?\\_") %>% 
+    str_remove("_")
+
+dat_raw <- list.files(here("Data", "Raw"), full.names = TRUE) %>% 
+    map(fread, na.string = "") %>%  # import participant files
+    map(mutate, participant = as.character(participant)) %>% 
+    set_names(filenames) %>% # label each dataset with the participant's ID
+    bind_rows() %>% # merge datasets
+    as_tibble() %>%
+    clean_names() %>% 
+    # select relevant variables and rename if necessary
+    select(participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
+           matches("language_l|demo_|setup_"), -contains("_rt")) %>% 
+    rename(demo_impairment = demo_language_key_keys) %>% 
+    rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
+    # clean text input by participants (because of typos of need to translate) and redefine location    
+    group_by(participant) %>% 
+    mutate_at(vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), city, setup_location, setup_noise), first_non_na) %>% 
+    mutate(date = max(date, na.rm = TRUE)) %>% 
+    ungroup() %>% 
+    mutate_at(vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), clean_input_text) %>% 
+    mutate(city = str_to_sentence(city),
+           country = ifelse(city %in% spanish_cities, "Spain", "UK"),
+           date = as_datetime(ymd_hms(date)),
[…]

### 2020-07-16 R/03_analyse-rts.R

@@ -1,0 +1,114 @@
+#### 03_analyse-rts: Analyse RTs #########
+
+#### set up ##############################
+
+# load packages
+library(data.table)
+library(dplyr)
+library(purrr)
+library(tibble)
+library(lme4)
+library(broom.mixed)
+library(emmeans)
+library(ggplot2)
+library(wesanderson)
+library(patchwork)
+library(janitor)
+library(here)
+
+# load/create functions
+source(here("R", "functions.R"))
+
+#### load data ###########################
+dat <- fread(here("Data", "04_prepared.csv"), na.strings = "") %>%
+    as_tibble() %>%
+    mutate_at(1:2, factor)
+
+#### LMM: Reaction times #################
+
+fit0 <- lmer(
+    log(typing_onset) ~ frequency_center +
+        (1 | participant) +
+        (1 | trial_id),
+    data = dat
+)
+
+fit1 <- lmer(
+    log(typing_onset) ~ frequency_center*lev_center +
+        (1 + lev_center | participant) +
+        (1 + lev_center | trial_id),
+    data = dat
+)
+
+fit2 <- lmer(
+    log(typing_onset) ~ frequency_center*sim_center +
+        (1 + sim_center | participant) +
+        (1 + sim_center | trial_id),
+    data = dat
+)
+
+#### compare models ########################
+comp <- anova(fit0, fit1, fit2) %>%
+    as_tibble() %>%
+    clean_names() %>%
+    mutate(model = paste0("model", 0:2)) %>%
+    select(model, everything(), p = pr_chisq)
+
+confints <- confint.merMod(fit1, method = "Wald") %>%
+    as_tibble() %>%
+    clean_names()
+
+coefs <- tidy(fit1) %>%
+    arrange(desc(effect)) %>%
+    cbind(confints) %>%
+    filter(effect=="fixed") %>%
+    rename(ci_low = x2_5_percent,
+           ci_hight = x97_5_percent) %>%
+    select(-c(group, effect)) %>%
+    mutate(term = c("Intercept", "Frequency", "Levenshtein", "Frequency x Levenshtein"))
+
+predictions <- ranef(fit1) %>%
+    map(~rename(., intercept = "(Intercept)")) %>%
+    map(~rename(., slope = "lev_center")) %>%
+    
+    map(~rownames_to_column(., "participant")) %>%
+    map(~mutate_at(., 1, as.character))
+
[…]

### 2020-07-16 R/functions.R

@@ -1,0 +1,176 @@
+#### functions.R ############################
+
+# evaluate if x is NOT an element of y
+"%!in%" <- function(x, y) !(x %in% y)
+
+# find first/last non-missing value in vector
+first_non_na <- function(x) ifelse(is.logical(first(x[!is.na(x)])), NA, first(x[!is.na(x)]))
+last_non_na <- function(x) ifelse(is.logical(last(x[!is.na(x)])), NA, last(x[!is.na(x)]))
+add_big_mark <- function(x) format(x, big.mark = ",", scientific = FALSE)
+clean_input_text <- function(x) {
+    {{ x }} %>% 
+        str_to_sentence() %>%
+        str_replace("Lshift|space|minus", "") %>%
+        as.character() %>% 
+        as_tibble(x = .) %>% 
+        mutate(. , value = case_when( 
+            nchar(value) > 12 ~ "Several",
+            value %in% c("Francés", "Frances", "Basic french") ~ "French",
+            value %in% c("Italiano", "Italino") ~ "Italian",
+            value %in% c("Alemán") ~ "German",
+            value %in% c("NrLondon") ~ "London",
+            value %in% c("Rnorthampton") ~ "London",
+            value %in% c("Newcastleminusuponminustyne") ~ "Newcastle upon Tyne",
+            value %in% c("Newcastlespaceuk") ~ "Newcastle",
+            value %in% c("Miltonkeynes") ~ "Milton Keynes",
+            value %in% c("Irthlingboroug") ~ "Irthlingborough",
+            value %in% c("spalding", "berkshire", "bedfordshire") ~ str_to_sentence(value),
+            value %in% c("No", "Cuál", "Space") ~ NA_character_,
+            TRUE ~ value)) %>% 
+        pull(value)
+}
+#### ggplot custom theme ##################################
+theme_custom <- theme(
+    panel.background = element_rect(fill = "transparent"),
+    panel.grid = element_blank(),
+    panel.border = element_rect(fill = "transparent", colour = "black"),
+    text = element_text(colour = "black", size = 12),
+    axis.text = element_text(colour = "black")
+)
+
+#### extract formants #####################################
+extract_formants <- function(
+    file,
+    time_step = 0.001,
+    n = 2,
+    max_freq = 5500,
+    window_length = 0.001,
+    pre = 50,
+    include_frames = TRUE,
+    include_time = TRUE,
[…]

### 2020-07-21 Manuscript/te_manuscript.Rmd
Remove media

---
title             : "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
shorttitle        : "Phoneme-weighted Levenshtein Distance"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Second Author"
    affiliation   : "1,2"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Many studies in bilingualism have explored how form similarity between translation equivalents (i.e. cognate- ness) impacts language processing. Cognates range from being identical to sharing only a few key features. Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity. In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value. A widely used metric for word similarity is Levenshtein distance [@levenshtein1966binary]. This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other. However, a challenge when defining phonological similarity is that some phonemes (e.g. [k] and [g]) are perceptually more similar than others (e.g. [k] and [o]). The standard Lev- enshtein distance calculation treats each change equally. Thus it often overestimates phonological distance between words. We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g. vowel or consonant insertion) and position in the word. To estimate the weights, we are col- lecting behavioural data in an auditory translation elicitation task. Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words. The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent. Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation. The higher the probability, the higher the perceptual similarity. Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

library(tidyverse)
library(here)
library(data.table)
library(lubridate)
library(papaja)

logs <- fread(here("Data", "01_logs.csv"), na.strings = "") %>%
    as_tibble() %>%
    mutate(date = as_date(date))
dat <- fread(here("Data", "01_processed.csv"), na.strings = "") %>% 
    as_tibble()

```

#Methods

## Translation Elicitation task

### Participants

Data collection took place from `r format(min(logs$date), "%dth %B, %Y")` to  `r format(max(logs$date), "%dth %B, %Y")`. We collected data from `r nrow(logs)` participants ($M_{age}$ = `r printnum(mean(logs$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(sd(logs$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(min(logs$age, na.rm = TRUE))`-`r printnum(max(logs$age, na.rm = TRUE))`). `r filter(logs, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(logs, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(logs, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(logs, country=="Spain", sex =="Female") %>% nrow()` female). Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits). Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value. Participants were asked to complete the experiment in a quiet place with good internet connection. We exluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(logs, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(logs, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION? (Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan? (Sí/No)". However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(logs, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(logs, country %in% "Spain", !vision))/nrow(filter(logs, country=="Spain")))*100)`%). This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(logs$exclude_trials, na.rm = TRUE)`).

### Procedure

Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla). First participants were informed about the aims of the study and gave informed consent for participating. Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study. Third, participants completed the experimental task. BEfore the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants). They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible. English participants were randomly assigned to the list of Catalan or Spanish trials. Spanish participants completed the list of Catalan trials.

### Design

Each trial started with a yellow fixation dot presented during one second on the center of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio offset. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Typos were allowed, however. Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish). The task took approximately 15 minutes to be completed.

### Stimuli

## Data analysis

# Results

# Discussion

\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup

### 2020-07-21 Manuscript/te_manuscript.tex
Remove media

--- 2020-07-16 Code/Rmd/translationelicitation.tex
@@
 \makeatother
-\shorttitle{Translation Elicitation}
-\author{Serene Siow\textsuperscript{1}, Gonzalo García-Castro\textsuperscript{2}, Núria Sebastian-Galles\textsuperscript{2}, \& Kim Plunkett\textsuperscript{1}}
-\affiliation{
-\vspace{0.5cm}
-\textsuperscript{1} Department of Experimental Psychology, Oxford University\\\textsuperscript{2} Center for Brain and Cognition, Universitat Pompeu Fabra}
-\authornote{SS and GGC share first authorship.
+\shorttitle{Phoneme-weighted Levenshtein Distance}
+\author{First Author\textsuperscript{1}\ \& Second Author\textsuperscript{1,2}}
+\affiliation{\phantom{a}}
+\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
 
-Correspondence concerning this article should be addressed to Serene Siow, Anna Watts Building, Radcliffe Observatory Quarter, Woodstock Rd, Oxford, Oxfordshire OX2 6GG, UK.
-E-mail: serene.siow@psy.ox.ac.edu}
+Enter author note here.
+
+Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: my@email.com}
 \keywords{keywords\newline\indent Word count: X}
@@
 We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g.~vowel or consonant insertion) and position in the word.
-To estimate the weights, we are collecting behavioural data in an auditory translation elicitation task.
+To estimate the weights, we are col- lecting behavioural data in an auditory translation elicitation task.
 Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words.
@@
 
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
+\#Methods
 
-We report how we determined our sample size, all data exclusions (if any), all manipulations, and all measures in the study.
+\hypertarget{translation-elicitation-task}{%
+\subsection{Translation Elicitation task}\label{translation-elicitation-task}}
 
 \hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
+\subsubsection{Participants}\label{participants}}
 
-\hypertarget{material}{%
-\subsection{Material}\label{material}}
+Data collection took place from 04th June, 2020 to 28th June, 2020.
+We collected data from 105 participants (\(M_{age}\) = 21.58, \(SD_{age}\) = 3.09, \(Range_{age}\) = 2-33).
+72 participants were British English native speakers living in United Kingdom (0 female), and 72 participants were Spanish native speakers living in Spain (0 female).
+Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
+Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
+Participants were asked to complete the experiment in a quiet place with good internet connection.
+We exluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were diagnosed with a language (\emph{n} = 2)\footnote{We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as \enquote{Do you have normal or corrected-to-normal VISION?
+(Yes/No)} in English, and as \enquote{¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
+(Sí/No)}.
+However, the proportion of Spanish participants that reported visual impairment was unplausibly large (\emph{n} = 6, 18.18\%).
+This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.}, or c) did not contribute more than 80\% of valid trials (\emph{n} = 1).
 
 \hypertarget{procedure}{%
-\subsection{Procedure}\label{procedure}}
+\subsubsection{Procedure}\label{procedure}}
+
+Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
+First participants were informed about the aims of the study and gave informed consent for participating.
+Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+Third, participants completed the experimental task.
+BEfore the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
+English participants were randomly assigned to the list of Catalan or Spanish trials.
+Spanish participants completed the list of Catalan trials.
+
+\hypertarget{design}{%
+\subsubsection{Design}\label{design}}
+
+Each trial started with a yellow fixation dot presented during one second on the center of the screen over a black background.
+After one second, the audio started playing while the dot remained being displayed until the audio offset.
+Upon the offset of the fixation point and audio, participants were prompted to write their answer by a \enquote{\textgreater{}} symbol.
+Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
+Participants were allowed to correct their answer.
+Then, participants pressed the RETURN key to start and new trial.
+Participants contributed a total of 9536 trials (5901 in Catalan, 3635 in Spanish).
+The task took approximately 15 minutes to be completed.
+
+\hypertarget{stimuli}{%
+\subsubsection{Stimuli}\label{stimuli}}
 
@@
 \subsection{Data analysis}\label{data-analysis}}
-
-We used R (Version 3.6.3; R Core Team, 2020) for all our analyses.
 
@@
 \hypertarget{refs}{}
-\leavevmode\hypertarget{ref-R-base}{}%
-R Core Team. (2020). \emph{R: A language and environment for statistical computing}. Vienna, Austria: R Foundation for Statistical Computing. Retrieved from \url{https://www.R-project.org/}
 

### 2020-07-21 R/01_process.R

@@ -1,146 +1,0 @@
-#### 01_preprocess: Pre-process data ################
-
-#### set up #########################################
-
-# load packages
-library(tidyverse)
-library(lubridate)
-library(purrr)
-library(readxl)
-library(data.table)
-library(stringdist)
-library(mice)
-library(lme4)
-library(car)
-library(janitor)
-library(patchwork)
-library(here)
-
-# create/load functions
-source(here("R", "functions.R"))
-
-# set parameters
-practice_trials <- c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)
-spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
-
-#### import data ####################################
-trials <- read_xlsx(here("Stimuli", "trials.xlsx")) # trial-level data
-
-# participant files
-filenames <- list.files(here("Data", "Raw")) %>% 
-    str_extract(".*?\\_") %>% 
-    str_remove("_")
-
-dat_raw <- list.files(here("Data", "Raw"), full.names = TRUE) %>% 
-    map(fread, na.string = "") %>%  # import participant files
-    map(mutate, participant = as.character(participant)) %>% 
-    set_names(filenames) %>% # label each dataset with the participant's ID
-    bind_rows() %>% # merge datasets
-    as_tibble() %>%
-    clean_names() %>% 
-    # select relevant variables and rename if necessary
-    select(participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
-           matches("language_l|demo_|setup_"), -contains("_rt")) %>% 
-    rename(demo_impairment = demo_language_key_keys) %>% 
-    rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
-    # clean text input by participants (because of typos of need to translate) and redefine location    
-    group_by(participant) %>% 
-    mutate_at(vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), city, setup_location, setup_noise), first_non_na) %>% 
-    mutate(date = max(date, na.rm = TRUE)) %>% 
-    ungroup() %>% 
-    mutate_at(vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), clean_input_text) %>% 
-    mutate(city = str_to_sentence(city),
-           country = ifelse(city %in% spanish_cities, "Spain", "UK"),
-           date = as_datetime(ymd_hms(date)),
[…]

### 2020-07-21 R/03_analyse-acc.R

@@ -1,113 +1,0 @@
-#### 03_analyse-acc: Analyse accuracy ####
-
-#### set up ##############################
-
-# load packages
-library(tidyverse)
-library(data.table)
-library(lme4)
-library(lmerTest)
-library(broom.mixed)
-library(ggplot2)
-library(wesanderson)
-library(patchwork)
-library(janitor)
-library(here)
-
-# load/create functions
-source(here("R", "functions.R"))
-
-#### load data ###########################
-dat <- fread(here("Data", "04_prepared.csv"), na.strings = "") %>%
-    as_tibble() %>%
-    mutate_at(1:2, factor)
-
-#### LMM: Reaction times #################
-
-fit0 <- lmer(accuracy ~ same_onset + phon_overlap + n_consecutive_phon + phon_neigh + frequency_zipf +
-                 (1 | test_language/participant) +
-                 (1 | trial_id),
-             data = dat)
-
-fit1 <- lmer(accuracy ~ frequency_zipf*sim +
-                 (1 + sim | participant) +
-                 (1 + sim | trial_id), na.action = na.omit,
-             data = dat)
-
-#### compare models ########################
-comp <- anova(fit0, fit1) %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(model = paste0("model", 0:1)) %>%
-    select(model, everything(), p = pr_chisq)
-
-confints <- confint.merMod(fit1, method = "Wald") %>%
-    as_tibble() %>%
-    clean_names()
-
-coefs <- tidy(fit1) %>%
-    arrange(desc(effect)) %>%
-    cbind(confints) %>%
-    filter(effect=="fixed") %>%
-    rename(ci_low = x2_5_percent,
-           ci_hight = x97_5_percent) %>%
-    select(-c(group, effect)) %>%
-    mutate(term = c("Intercept", "Frequency", "Levenshtein", "Frequency \u00D7 Levenshtein"),
-           sig = p.value < 0.05,
-           sig = ifelse(sig, "p < 0.05", "p \u2265 0.05")) %>%
-    clean_names()
-
-predictions <- ranef(fit1) %>%
-    map(~rename(., intercept = "(Intercept)")) %>% 
-    map(~rename(., slope = "sim")) %>%
-    map(~rownames_to_column(., "participant")) %>%
-    map(~mutate_at(., 1, as.character))
-
-dat_predictions <-  left_join(dat, predictions$participant) %>%
[…]

### 2020-07-21 R/03_analyse-rt-bayesian.R

@@ -1,78 +1,0 @@
-#### 03-analyse-rts: Analyse reaction times
-
-#### set up ###############################
-
-# load packages
-library(dplyr)
-library(data.table)
-library(tidyr)
-library(ggplot2)
-library(patchwork)
-library(wesanderson)
-library(brms)
-library(modelr)
-library(tidybayes)
-library(ggridges)
-library(janitor)
-library(here)
-
-# load functions
-source(here("R", "functions.R"))
-
-#### load data #############################
-dat <- fread(here("Data", "04_prepared.csv")) %>%
-    select(participant, trial_id, typing_onset, lev_center, frequency_center) %>%
-    rename(rt = typing_onset,
-           lev = lev_center,
-           frequency = frequency_center)
-
-#### fit model #############################
-fit <- brm(rt ~ lev + frequency, data = dat)
-
-dat_fitted <- tibble(lev = seq(from = min(dat$lev), to = max(dat$lev),  length.out = 101),
-                     frequency = seq(from = min(dat$frequency), to = max(dat$frequency), length.out = 101)) %>%
-    add_fitted_draws(newdata = ., model = fit) %>% 
-    bind_cols()
-dat_predicted <- tibble(lev = seq(from = min(dat$lev), to = max(dat$lev),  length.out = 101),
-                        frequency = seq(from = min(dat$frequency), to = max(dat$frequency), length.out = 101)) %>%
-    add_predicted_draws(newdata = ., model = fit) %>% 
-    bind_cols()
-
-fit_dat <- left_join(dat_fitted, dat_predicted)
-
-#### visualise data ########################
-
-# preditions
-ggplot(fit_dat, aes(lev, .value)) +
-    stat_lineribbon(.width = c(0.67, 0.89, 0.97)) +
-    labs(x = "Levenstein distance", y = "P(RT | Levenshtein)", fill = "Credible interval") +
-    scale_fill_manual(values = wes_palette("Zissou1", 3)) +
-    theme_custom +
-    theme(legend.position = "top")
-
-# coefficients
-fit %>%
-    gather_draws(b_Intercept, b_lev, b_frequency) %>%
-    ggplot(., aes(x = .value, y = .variable, fill = .variable)) +
-    geom_vline(xintercept = 0) +
-    stat_halfeyeh() +
-    geom_vline(xintercept = c(-.8, .8), linetype = "dashed") +
[…]

### 2020-07-21 R/03_analyse-rts.R

@@ -1,114 +1,0 @@
-#### 03_analyse-rts: Analyse RTs #########
-
-#### set up ##############################
-
-# load packages
-library(data.table)
-library(dplyr)
-library(purrr)
-library(tibble)
-library(lme4)
-library(broom.mixed)
-library(emmeans)
-library(ggplot2)
-library(wesanderson)
-library(patchwork)
-library(janitor)
-library(here)
-
-# load/create functions
-source(here("R", "functions.R"))
-
-#### load data ###########################
-dat <- fread(here("Data", "04_prepared.csv"), na.strings = "") %>%
-    as_tibble() %>%
-    mutate_at(1:2, factor)
-
-#### LMM: Reaction times #################
-
-fit0 <- lmer(
-    log(typing_onset) ~ frequency_center +
-        (1 | participant) +
-        (1 | trial_id),
-    data = dat
-)
-
-fit1 <- lmer(
-    log(typing_onset) ~ frequency_center*lev_center +
-        (1 + lev_center | participant) +
-        (1 + lev_center | trial_id),
-    data = dat
-)
-
-fit2 <- lmer(
-    log(typing_onset) ~ frequency_center*sim_center +
-        (1 + sim_center | participant) +
-        (1 + sim_center | trial_id),
-    data = dat
-)
-
-#### compare models ########################
-comp <- anova(fit0, fit1, fit2) %>%
-    as_tibble() %>%
-    clean_names() %>%
-    mutate(model = paste0("model", 0:2)) %>%
-    select(model, everything(), p = pr_chisq)
-
-confints <- confint.merMod(fit1, method = "Wald") %>%
-    as_tibble() %>%
-    clean_names()
-
-coefs <- tidy(fit1) %>%
-    arrange(desc(effect)) %>%
-    cbind(confints) %>%
-    filter(effect=="fixed") %>%
-    rename(ci_low = x2_5_percent,
-           ci_hight = x97_5_percent) %>%
-    select(-c(group, effect)) %>%
-    mutate(term = c("Intercept", "Frequency", "Levenshtein", "Frequency x Levenshtein"))
-
-predictions <- ranef(fit1) %>%
-    map(~rename(., intercept = "(Intercept)")) %>%
-    map(~rename(., slope = "lev_center")) %>%
-    
-    map(~rownames_to_column(., "participant")) %>%
-    map(~mutate_at(., 1, as.character))
-
[…]

### 2020-07-24 Manuscript/te_manuscript.Rmd
Add PTHN as predictor

@@
 
-#Methods
+# Methods
 
@@
 
+The experiment was implemented online using Psychopy/Pavlovia [@peirce2019].
 Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-First participants were informed about the aims of the study and gave informed consent for participating.
+First, participants were informed about the aims of the study and gave informed consent for participating.
 Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
 Third, participants completed the experimental task.
-BEfore the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+Before the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
 They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
@@
 
+We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT).
+We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
+We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance.
+We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct reponses as output.
+Continuous predictors were standadised and categorical predictors were sum-coded [@schad2020].
+Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013].
+To test and account for cross-group differences, we included a random intercept for each group.
+We compared models using Paretho-Smoothed Importance Sampling (PSIS-LOO, @vehtari2016).
+More information about the models and model comparison can be found in Appendix X.
+
+All analyses were performed in R environment [@rcoreteam2019].
+We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] to impute missing data, and the `brms` R package to [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
+
 # Results

### 2020-07-30 Communications/2020-07-30_meeting.Rmd
Build model incrementally

---
title: "Translation Elicitation"
date: "30/07/2020"
output: html_document
---

```{r prepare, include=FALSE}

library(dplyr)
library(papaja)
library(tidyr)
library(tibble)
library(stringr)
library(forcats)
library(ggplot2)
library(here)
library(data.table)
library(tidybayes)
library(gt)
library(lubridate)

participants <- fread(here("Data", "01_participants.csv"), na.strings = "") %>% as_tibble() %>% mutate(date = as_date(date))
dat <- fread(here("Data", "01_processed_coded.csv"), na.strings = "") %>% as_tibble()
posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()
predictions <- fread(here("Results", "posterior_predictions.csv"), na.strings = "") %>% as_tibble()
fit <- readRDS(here("Results", "fit3_phon.rds"))
comparison <- readRDS(here("Results", "model_comparison.rds"))
item_effects <- fread(here("Results", "item_effects.csv"), na.strings = "") %>% as_tibble()
group_effects <- fread(here("Results", "group_effects.csv"), na.strings = "") %>% as_tibble()
corrs <- fread(here("Results", "corrs.csv"), na.strings = "") %>% as_tibble()

valid_participant <- filter(participants, valid_participant) %>% pull(participant)
```

# Behavioural task

## Sample

`r nrow(participants)` participants (mean age = `r printnum(mean(participants$age, na.rm = TRUE))`, *SD* age = `r printnum(sd(participants$age, na.rm = TRUE))`, age range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).

`r filter(participants, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(participants, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(participants, country=="Spain") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(participants, country=="Spain", sex =="Female") %>% nrow()` female).

Exclusion criteria:

a) Oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`)
b) Language impairment (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION? (Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan? (Sí/No)". However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%). This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.]
c) Did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`)

This resulted in `r nrow(participants)-sum(participants$valid_participant)` participants excluded, leaving `r nrow(filter(participants, valid_participant))` valid participants and `r dat %>% filter(response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% nrow() %>% printnum(big.mark = ",")` valid trials.

```{r participants_group, echo=FALSE}

dat %>%
    filter(participant %in% valid_participant, response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% 
    group_by(group, participant) %>% 
    summarise(n_trials = n()) %>% 
    group_by(group)%>%
    summarise(n = n(),
              n_trials = sum(n_trials)) %>% 
    gt() %>% 
    cols_label(group = "Group", n_trials = "Trials")

```

```{r participants_language, echo=FALSE}

participants %>%
    ggplot(aes(l2, fill = l2)) +
    facet_wrap(~group, ncol = 2) +
    geom_bar(stat = "count", show.legend = FALSE) +
    labs(x = "L2", y = "Count", title = "What languages others than you native one do you know?") +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
          axis.text.x = element_text(angle = 330, hjust = 0))

```

```{r participants_prof_l2, echo=FALSE}

participants %>%
    filter(!(l2 %in% "None")) %>% 
    pivot_longer(c(l2oral, l2written), names_to = "modality", values_to = "proficiency") %>%
    mutate(modality = str_to_sentence(str_remove(modality, "l2"))) %>% 
    ggplot(aes(l2, proficiency, fill = l2)) +
    facet_wrap(group~modality) +
    geom_col(show.legend = FALSE) +
    labs(x = "L2", y = "Self rated proficiency (0-5)") +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
          axis.text.x = element_text(angle = 330, hjust = 0))

```

```{r participants_prof_spacat, echo=FALSE}

participants %>%
    pivot_longer(c(spanish_oral, spanish_written, catalan_oral, catalan_written), names_to = "language", values_to = "proficiency") %>%
    separate(language, c("language", "modality"), sep = "_") %>% 
    mutate(modality = str_to_sentence(str_remove(modality, "l2")),
           language = str_to_sentence(language)) %>% 
    ggplot(aes(proficiency)) +
    facet_wrap(group~language~modality, nrow = 3) +
    geom_bar(position = position_dodge(), fill = "black") +
    labs(y = "Count", x = "Self rated proficiency (0-5)") +
    scale_x_continuous(limits = c(0, 5)) +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))

```

## Design

* 0 - 1000 ms: fixation point
* 1000 - 2500 ms: audio
* 2500 ms - Press RETURN key

## Data analysis

We fit Bayesian logistic models to model the probability of participants typing a correct translation in each trial. As predictors, we included the number of more frequent phonological neighbours of the translation (`PTHN`, standardised), whether both translation equivalents started with the same phoneme (`same_onset`, with levels `different`, `same_vowel`, `same_consonant`, helmert contrast comparing different vs. same and same vowel vs. same consonant), the highest number of consecutive phonemes shared (`consecutive_longest`, standardised), and our meausure of phonological similarity (`similarity_phop`, standardised). We first for a null model `fit0_null` that only included PTHN as predictor. We then extended the model incrementally by adding other predictors. The models were finally, we fitted a model that only included `PTHN` and the traditional orthographical Levenshtein similarity (`similarity`, standardised), we test whether our measure predicted participants responses better than just relying on orthographic similarity. The models were implemente in the `brms` R package using the following formulas:

* `fit0_null <- correct | total ~ pthn + (1 + pthn | group/trial_id)`
* `fit1_onset <- correct | total ~ pthn + same_onset + (1 +  same_onset | group/trial_id)`
* `fit2_consec <- correct | total ~ pthn + same_onset + consecutive_longest + (1 + same_onset + consecutive_longest | group/trial_id)`
* `fit3_phon <- correct | total ~ pthn + same_onset + consecutive_longest + similarity_phon + (1 + same_onset + consecutive_longest + similarity_phon | group/trial_id)`
* `fit4_ort <- correct | total ~ pthn + similarity_ort + (1 + similarity_ort | group/trial_id)`

# Results

## Model comparison

```{r waic, echo=FALSE}

waic <- comparison %>%
    unlist() %>%
    as.data.frame() %>% 
    rownames_to_column("model")

colnames(waic) <- c("Model", "WAIC")

waic %>%
    gt()

```

## Posterior distributions

```{r posterior_draws, echo=FALSE}

posterior %>%
    ggplot(aes(.value, .variable)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    stat_slab(fill = "black") + 
    stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.15)) +
    labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
         title = "Posterior distribution of coefficients of fixed effects") +
    scale_colour_brewer(palette = "Blues") +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey"),
          legend.position = "top",
          legend.key = element_rect(fill = "transparent"))

```

## Posterior predictors

```{r posterior_predictions, echo=FALSE}

predictions %>%
    ggplot(aes(similarity_phon, .value, colour = same_onset, fill = same_onset)) +
    facet_grid(consecutive_longest~pthn)+
    stat_summary(fun = mean, geom = "line", size = 1) +
    geom_vline(xintercept = 0, colour= "grey") +
    geom_hline(yintercept = 0.5, colour= "grey") +
   labs(x = "Phonological similarity", y = "Probability of correct translation",
         fill = "Credible interval", colour = "Onset",
         title = "Effect of phonological similarity on probability of correct translation",
         subtitle = "Model predictions are displayed for different quartiles of PTHN") +
    scale_fill_brewer(palette = "Set1") +
    scale_colour_brewer(palette = "Set1") +
    theme_classic() +
    theme(legend.position = "top")

```

## Random effects

### Item-level effects

```{r item_effects, echo=FALSE}

item_effects %>%
  pivot_wider(names_from = estimate, values_from = value) %>% 
  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
  ggplot(aes(estimate, fill = param)) +
  facet_wrap(~param, scales = "free") +
  geom_histogram(bins = 20) +
  geom_rug(alpha = 0.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Estimate", y = "Density", fill = "Parameter") +
  scale_fill_brewer(palette = "Set1") +
  theme_classic() +
  theme(legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))

```

### Group-level effects

```{r goup_effects_table, echo=FALSE}

group_effects %>%
  pivot_wider(names_from = estimate, values_from = value) %>% 
  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  mutate(hdi = paste0("[", lower, ", ", upper, "]")) %>% 
  select(param, group, error, hdi) %>% 
  group_by(param) %>% 
  gt() %>% 
  cols_label(group = "Group", error = "SE", hdi = "95% HDI")

```

```{r group_effects, echo=FALSE}

group_effects %>%
  pivot_wider(names_from = estimate, values_from = value) %>% 
  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
  ggplot(aes(group, estimate)) +
  facet_wrap(~param) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = lower, ymax = upper, colour =  group), width = 0, size = 5) +
  geom_errorbar(aes(ymin = estimate-error, ymax = estimate+error), width = 0, size = 1, alpha = 0.5) +
  geom_point(size = 3) +
  labs(x = "Group", y = "Estimate", fill = "Group") +
  scale_fill_brewer(palette = "Set1") +
  theme_classic() +
  theme(legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))

```

### Correlations

```{r correlations, echo=FALSE}

corrs %>% 
  filter(str_detect(variable, "trial_id")) %>%
  mutate(variable = str_remove(variable, "cor_group:trial_id__"),
         variable = str_replace(variable, "__" , " - ")) %>% 
  ggplot(aes(value, variable, xmin = lower, xmax = upper)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_errorbarh(show.legend = FALSE, height = 0) +
  geom_point(size = 2, show.legend = FALSE) +
  labs(x = "Estimate", y = "Correlation") +
  theme_classic() +
  theme(axis.title.y = element_blank())

```

### 2020-07-30 Manuscript/te_appendix.Rmd
Build model incrementally

---
title: "Appendices"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r prepare}

prior <- fread(here("Results", "prior_predictions.csv"), na.strings = "") %>% as_tibble()
posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()

```

# Appendix 2: Model

Prior predictive checks

Convergence

```{r convergence}

posterior %>%
    filter(.iteration < 2000) %>% 
    ggplot(aes(.iteration, .value, colour = .chain)) +
    facet_wrap(~.variable, scales = "free_y") +
    geom_line() +
    labs(x = "Iteration", y = "Value", colour = "Chain") +
    theme_custom +
    theme(legend.position = "top") +
    ggsave(here("Figures", "convergence.png"))

```

# Appendix 4: Session info

```{r session_info}

sessionInfo()

```

### 2020-07-30 Manuscript/te_manuscript.Rmd
Build model incrementally

@@
 classoption       : "man"
-output            : papaja::apa6_pdf
+output            :
+  papaja::apa6_pdf:
+    latex_engine: "xelatex"
 ---
@@
 ```{r setup, include = FALSE}
-library("papaja")
+library(papaja)
 ```
@@
 
-library(tidyverse)
+library(dplyr)
+library(tidyr)
+library(tibble)
+library(stringr)
+library(ggplot2)
 library(here)
 library(data.table)
+library(tidybayes)
+library(gt)
 library(lubridate)
-library(papaja)
 
-logs <- fread(here("Data", "01_logs.csv"), na.strings = "") %>%
-    as_tibble() %>%
-    mutate(date = as_date(date))
-dat <- fread(here("Data", "01_processed.csv"), na.strings = "") %>% 
-    as_tibble()
+participants <- fread(here("Data", "01_participants.csv"), na.strings = "") %>% as_tibble() %>% mutate(date = as_date(date))
+dat <- fread(here("Data", "01_processed_coded.csv"), na.strings = "") %>% as_tibble()
+posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()
+predictions <- fread(here("Results", "posterior_predictions.csv"), na.strings = "") %>% as_tibble()
+fit <- readRDS(here("Results", "fit3_phon.rds"))
+comparison <- readRDS(here("Results", "model_comparison.rds"))
+item_effects <- fread(here("Results", "item_effects.csv"), na.strings = "") %>% as_tibble()
+group_effects <- fread(here("Results", "group_effects.csv"), na.strings = "") %>% as_tibble()
+corrs <- fread(here("Results", "corrs.csv"), na.strings = "") %>% as_tibble()
 
@@
 
+## Algorithm
+
 ## Translation Elicitation task
@@
 
-Data collection took place from `r format(min(logs$date), "%dth %B, %Y")` to  `r format(max(logs$date), "%dth %B, %Y")`.
-We collected data from `r nrow(logs)` participants ($M_{age}$ = `r printnum(mean(logs$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(sd(logs$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(min(logs$age, na.rm = TRUE))`-`r printnum(max(logs$age, na.rm = TRUE))`).
-`r filter(logs, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(logs, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(logs, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(logs, country=="Spain", sex =="Female") %>% nrow()` female).
+Data collection took place from `r format(min(participants$date), "%dth %B, %Y")` to  `r format(max(participants$date), "%dth %B, %Y")`.
+We collected data from `r nrow(participants)` participants ($M_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(sd(participants$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
+`r filter(participants, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(participants, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(participants, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(participants, country=="Spain", sex =="Female") %>% nrow()` female).
 Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
@@
 Participants were asked to complete the experiment in a quiet place with good internet connection.
-We exluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(logs, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(logs, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION?
+We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION?
 (Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
 (Sí/No)".
-However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(logs, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(logs, country %in% "Spain", !vision))/nrow(filter(logs, country=="Spain")))*100)`%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(logs$exclude_trials, na.rm = TRUE)`).
+However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
+This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`).
 
@@
 
+## Posterior draws
+
+```{r posterior_draws}
+
+posterior %>%
+    ggplot(aes(.value, .variable)) +
+    geom_vline(xintercept = 0, linetype = "dashed") +
+    stat_slab(fill = "black") + 
+    stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.15)) +
+    labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
+         title = "Posterior distribution of coefficients of fixed effects") +
+    scale_colour_brewer(palette = "Blues") +
+    theme_classic() +
+    theme(panel.grid.major.y = element_line(colour = "grey"),
+          legend.position = "top",
+          legend.key = element_rect(fill = "transparent")) +
+    ggsave(here("Figures", "coefs.png"))
+
+```
+
+## Posterior predictions
+
+```{r posterior_predictions}
+
+predictions %>%
+    ggplot(aes(similarity_phon, .value, colour = same_onset, fill = same_onset)) +
+    facet_grid(consecutive_longest~pthn)+
+    stat_summary(fun = mean, geom = "line", size = 1) +
+    geom_vline(xintercept = 0, colour= "grey") +
+    geom_hline(yintercept = 0.5, colour= "grey") +
+   labs(x = "Phonological similarity", y = "Probability of correct translation",
+         fill = "Credible interval", colour = "Onset",
+         title = "Effect of phonological similarity on probability of correct translation",
+         subtitle = "Model predictions are displayed for different quartiles of PTHN") +
+    scale_fill_brewer(palette = "Set1") +
+    scale_colour_brewer(palette = "Set1") +
+    theme_classic() +
+    theme(legend.position = "top") +
+    ggsave(here("Figures", "predictions.png"), height = 4)
+
+```
+
+## Random effects
+
+### Item effects
+
+```{r item_effects}
+
+item_effects %>%
+  pivot_wider(names_from = estimate, values_from = value) %>% 
+  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
+                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
+  ggplot(aes(estimate, fill = param)) +
+  facet_wrap(~param, scales = "free") +
+  geom_histogram(bins = 20) +
+  geom_rug(alpha = 0.1) +
+  geom_vline(xintercept = 0, linetype = "dashed") +
+  labs(x = "Estimate", y = "Density", fill = "Parameter") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_classic() +
+  theme(legend.position = "none",
+        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
+        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
+
+```
+
+### Group effects
+
+```{r goup_effects_table}
+
+group_effects %>%
+  pivot_wider(names_from = estimate, values_from = value) %>% 
+  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
+                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
+  mutate_if(is.numeric, round, 2) %>% 
+  mutate(hdi = paste0("[", lower, ", ", upper, "]")) %>% 
+  select(param, group, error, hdi) %>% 
+  group_by(param) %>% 
+  gt() %>% 
+  cols_label(group = "Group", error = "SE", hdi = "95% HDI")
+
+```
+
+```{r group_effects}
+
+group_effects %>%
+  pivot_wider(names_from = estimate, values_from = value) %>% 
+  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
+                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
+  ggplot(aes(group, estimate)) +
+  facet_wrap(~param) +
+  geom_hline(yintercept = 0) +
+  geom_errorbar(aes(ymin = lower, ymax = upper, colour =  group), width = 0, size = 5) +
+  geom_errorbar(aes(ymin = estimate-error, ymax = estimate+error), width = 0, size = 1, alpha = 0.5) +
+  geom_point(size = 3) +
+  labs(x = "Group", y = "Estimate", fill = "Group") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_classic() +
+  theme(legend.position = "none",
+        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
+        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
+
+```
+### Correlations
+
+```{r correlations}
+
+corrs %>% 
+  filter(str_detect(variable, "trial_id")) %>%
+  mutate(variable = str_remove(variable, "cor_group:trial_id__"),
+         variable = str_replace(variable, "__" , " - ")) %>% 
+  ggplot(aes(value, variable, xmin = lower, xmax = upper)) +
+  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
+  geom_errorbarh(show.legend = FALSE, height = 0) +
+  geom_point(size = 2, show.legend = FALSE) +
+  labs(x = "Estimate", y = "Correlation") +
+  theme_classic() +
+  theme(axis.title.y = element_blank())
+
+```
+
 # Discussion

### 2020-07-30 Manuscript/te_manuscript.tex
Build model incrementally

@@
-\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
 \PassOptionsToPackage{hyphens}{url}
 %
-\documentclass[english,man]{apa6}
+\documentclass[
+  english,
+  man]{apa6}
 \usepackage{lmodern}
@@
 \usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
 \ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
@@
   \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provides euro and other symbols
-\else % if luatex or xelatex
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
   \usepackage{unicode-math}
-  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
 \fi
-% use upquote if available, for straight quotes in verbatim environments
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-% use microtype if available
-\IfFileExists{microtype.sty}{%
+\IfFileExists{microtype.sty}{% use microtype if available
   \usepackage[]{microtype}
@@
 }{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
   \IfFileExists{parskip.sty}{%
@@
     \setlength{\parindent}{0pt}
-\setlength{\parskip}{6pt plus 2pt minus 1pt}
-}
-\usepackage{hyperref}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
 \hypersetup{
   pdftitle={A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity},
+  pdfauthor={First Author1 \& Second Author1,2},
+  pdflang={en-EN},
   pdfkeywords={keywords},
-            pdfborder={0 0 0},
-            breaklinks=true}
-\urlstyle{same}  % don't use monospace font for urls
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
 \usepackage{graphicx,grffile}
@@
 \setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
 \setlength{\emergencystretch}{3em} % prevent overfull lines
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{0}
-
-% set default figure placement to htbp
-\makeatletter
-\def\fps@figure{htbp}
-\makeatother
-
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 % Manuscript styling
@@
   {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
 \makeatother
 \shorttitle{Phoneme-weighted Levenshtein Distance}
-\author{First Author\textsuperscript{1}\ \& Second Author\textsuperscript{1,2}}
-\affiliation{\phantom{a}}
-\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-Enter author note here.
-
-Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: my@email.com}
 \keywords{keywords\newline\indent Word count: X}
@@
 \usepackage{csquotes}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[shorthands=off,main=english]{babel}
-\else
-  % load polyglossia as late as possible as it *could* call bidi if RTL lang (e.g. Hebrew or Arabic)
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
   \usepackage{polyglossia}
   \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
 \fi
@@
 \title{A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity}
+\author{First Author\textsuperscript{1} \& Second Author\textsuperscript{1,2}}
+\date{}
 
-\date{}
+\authornote{
+
+Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+
+Enter author note here.
+
+Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
+
+}
+
+\affiliation{\phantom{0}}
 
@@
 This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other.
-However, a challenge when defining phonological similarity is that some phonemes (e.g.
-{[}k{]} and {[}g{]}) are perceptually more similar than others (e.g.
-{[}k{]} and {[}o{]}).
+However, a challenge when defining phonological similarity is that some phonemes (e.g.~{[}k{]} and {[}g{]}) are perceptually more similar than others (e.g.~{[}k{]} and {[}o{]}).
 The standard Lev- enshtein distance calculation treats each change equally.
@@
 
-\#Methods
+\hypertarget{methods}{%
+\section{Methods}\label{methods}}
+
+\hypertarget{algorithm}{%
+\subsection{Algorithm}\label{algorithm}}
 
@@
 
-Data collection took place from 04th June, 2020 to 28th June, 2020.
-We collected data from 105 participants (\(M_{age}\) = 21.58, \(SD_{age}\) = 3.09, \(Range_{age}\) = 2-33).
-72 participants were British English native speakers living in United Kingdom (0 female), and 72 participants were Spanish native speakers living in Spain (0 female).
+Data collection took place from 04th junio, 2020 to 28th junio, 2020.
+We collected data from 107 participants (\(M_{age}\) = 21.79, \(SD_{age}\) = 2.41, \(Range_{age}\) = 18-33).
+74 participants were British English native speakers living in United Kingdom (49 female), and 74 participants were Spanish native speakers living in Spain (28 female).
 Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
@@
 However, the proportion of Spanish participants that reported visual impairment was unplausibly large (\emph{n} = 6, 18.18\%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.}, or c) did not contribute more than 80\% of valid trials (\emph{n} = 1).
+This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.}, or c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
 
@@
 
+The experiment was implemented online using Psychopy/Pavlovia ({\textbf{???}}).
 Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-First participants were informed about the aims of the study and gave informed consent for participating.
+First, participants were informed about the aims of the study and gave informed consent for participating.
 Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
 Third, participants completed the experimental task.
-BEfore the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+Before the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
 They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
@@
 Then, participants pressed the RETURN key to start and new trial.
-Participants contributed a total of 9536 trials (5901 in Catalan, 3635 in Spanish).
+We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
+Typos were allowed, however.
+Participants contributed a total of 9536 valid trials (5901 in Catalan, 3635 in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 
+We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT).
+We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
+We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance.
+We fit several Bayesian linear models ({\textbf{???}}), each including one of the aforementioned variables as predictor, and the proportion of correct reponses as output.
+Continuous predictors were standadised and categorical predictors were sum-coded ({\textbf{???}}).
+Missing predictor scores were imputed before fitting the models using multiple imputation ({\textbf{???}}).
+To test and account for cross-group differences, we included a random intercept for each group.
+We compared models using Paretho-Smoothed Importance Sampling (PSIS-LOO, ({\textbf{???}})).
+More information about the models and model comparison can be found in Appendix X.
+
+All analyses were performed in R environment ({\textbf{???}}).
+We used the \texttt{tidyverse} family of R packages to process data and to generate figures, the \texttt{mice} R package ({\textbf{???}}) to impute missing data, and the \texttt{brms} R package to ({\textbf{???}}) to fit and compare models, which uses the Stan probabilistic programming language language ({\textbf{???}}).
+
 \hypertarget{results}{%
 \section{Results}\label{results}}
+
+\hypertarget{posterior-draws}{%
+\subsection{Posterior draws}\label{posterior-draws}}
+
+\includegraphics{te_manuscript_files/figure-latex/posterior_draws-1.pdf}
+
+\hypertarget{posterior-predictions}{%
+\subsection{Posterior predictions}\label{posterior-predictions}}
+
+\includegraphics{te_manuscript_files/figure-latex/posterior_predictions-1.pdf}
+
+\hypertarget{random-effects}{%
+\subsection{Random effects}\label{random-effects}}
+
+\hypertarget{item-effects}{%
+\subsubsection{Item effects}\label{item-effects}}
+
+\includegraphics{te_manuscript_files/figure-latex/item_effects-1.pdf}
+
+\hypertarget{group-effects}{%
+\subsubsection{Group effects}\label{group-effects}}
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lrl}
+\toprule
+Group & SE & 95\% HDI \\ 
+\midrule
+\multicolumn{1}{l}{Intercept} \\ 
+\midrule
+ENG-CAT & 1.15 & [-4.27, 0.09] \\ 
+ENG-SPA & 1.08 & [-4.05, 0.15] \\ 
+SPA-CAT & 0.92 & [-2.87, 0.84] \\ 
+\midrule
+\multicolumn{1}{l}{Different onset vs. Same onset} \\ 
+\midrule
+ENG-CAT & 1.73 & [-1.05, 5.49] \\ 
+ENG-SPA & 1.62 & [-1.42, 5.09] \\ 
+SPA-CAT & 1.49 & [-1.3, 4.56] \\ 
+\midrule
+\multicolumn{1}{l}{Same onset: Vowel vs. Consonant} \\ 
+\midrule
+ENG-CAT & 2.58 & [-4.11, 6.94] \\ 
+ENG-SPA & 2.59 & [-3.61, 7.29] \\ 
+SPA-CAT & 1.98 & [-3.59, 4.64] \\ 
+\midrule
+\multicolumn{1}{l}{Consecutive longest} \\ 
+\midrule
+ENG-CAT & 1.09 & [-2.54, 2.18] \\ 
+ENG-SPA & 1.13 & [-3.27, 1.59] \\ 
+SPA-CAT & 1.11 & [-1.73, 2.94] \\ 
+\midrule
+\multicolumn{1}{l}{Phonological similarity} \\ 
+\midrule
+ENG-CAT & 1.05 & [-2.45, 1.96] \\ 
+ENG-SPA & 1.04 & [-2.71, 1.69] \\ 
+SPA-CAT & 1.10 & [-1.62, 2.87] \\ 
+\bottomrule
+\end{longtable}
+
+\includegraphics{te_manuscript_files/figure-latex/group_effects-1.pdf}
+\#\#\# Correlations
+
+\includegraphics{te_manuscript_files/figure-latex/correlations-1.pdf}
 

### 2020-07-30 Code/analysis.R

@@ -3,10 +3,10 @@
 #### set up #########################################
 
 # load packages
-library(dplyr)  # for basically everything
-library(tidyr)
-library(ggplot2)
-library(stringr)
+library(dplyr)      # for manipulating variables
+library(tidyr)      # for reshaping datasets
+library(ggplot2)    # for visualising plots
+library(stringr)    # for manipulating string characters
 library(patchwork)  # for arranging plots
 library(readxl)     # for importing Excel spreadsheets
 library(purrr)      # for working with lists
@@ -28,14 +28,14 @@ set.seed(888) # for reproducibility
 practice_trials <- c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)
 spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
 
-#### import data ####################################
+#### process data ####################################
 
 # participant files
 filenames <- list.files(here("Data", "Raw")) %>% 
     str_extract(".*?\\_") %>% 
     str_remove("_")
 
-dat_raw <- list.files(here("Data", "Raw"), full.names = TRUE) %>% 
+dat_processed <- list.files(here("Data", "Raw"), full.names = TRUE) %>% 
     map(fread, na.string = "") %>%  # import participant files
     map(mutate, participant = as.character(participant)) %>% 
     set_names(filenames) %>% # label each dataset with the participant's ID
@@ -44,12 +44,12 @@ dat_raw <- list.files(here("Data", "Raw"), full.names = TRUE) %>%
     clean_names() %>% 
     # select relevant variables and rename if necessary
     select(participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
-           matches("language_l|demo_|setup_"), -contains("_rt")) %>% 
+           matches("language_l|language_s|language_c|demo_|setup_"), -matches("_rt|_time")) %>% 
     rename(demo_impairment = demo_language_key_keys) %>% 
     rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
     # clean text input by participants (because of typos of need to translate) and redefine location    
     group_by(participant) %>% 
-    mutate_at(vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), city, setup_location, setup_noise), first_non_na) %>% 
[…]

### 2020-08-06 Communications/2020-07-30_meeting.Rmd
Prepare report for meeting

@@
 library(tibble)
+library(janitor)
 library(stringr)
@@
 library(ggplot2)
+library(patchwork)
 library(here)
 library(data.table)
+library(broom)
+library(brms)
+library(effectsize)
+library(pairwiseComparisons)
 library(tidybayes)
@@
 
-participants <- fread(here("Data", "01_participants.csv"), na.strings = "") %>% as_tibble() %>% mutate(date = as_date(date))
-dat <- fread(here("Data", "01_processed_coded.csv"), na.strings = "") %>% as_tibble()
-posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()
-predictions <- fread(here("Results", "posterior_predictions.csv"), na.strings = "") %>% as_tibble()
-fit <- readRDS(here("Results", "fit3_phon.rds"))
-comparison <- readRDS(here("Results", "model_comparison.rds"))
-item_effects <- fread(here("Results", "item_effects.csv"), na.strings = "") %>% as_tibble()
-group_effects <- fread(here("Results", "group_effects.csv"), na.strings = "") %>% as_tibble()
-corrs <- fread(here("Results", "corrs.csv"), na.strings = "") %>% as_tibble()
+participants <- fread(here("Data", "02_participants.csv"), na.strings = c("", "NA")) %>%
+  as_tibble() %>%
+  mutate(date = as_date(date),
+         group = as.factor(group))
+contrasts(participants$group) <-  contr.helmert(3)/2
+
+trials <- fread(here("Data", "trial_stats.csv"),
+                na.strings = c("", "NA")) %>%
+  as_tibble() %>%
+  clean_names()
+
+accuracy <- fread(here("Data", "03_accuracy_coded.csv"),
+                  na.strings = c("", "NA")) %>%
+  as_tibble() %>% 
+  mutate(correct_coded = response_type %in% c("correct", "typo"),
+         correct_coded = as.numeric(correct_coded))
+
+posterior <- fread(here("Results", "posterior_draws.csv"),
+                   na.strings = c("", "NA")) %>%
+  as_tibble()
+
+prior <- fread(here("Results", "prior_draws.csv"),
+               na.strings = c("", "NA")) %>%
+  as_tibble()
+
+predictions <- fread(here("Results", "posterior_predictions.csv"),
+                     na.strings = c("", "NA")) %>%
+  as_tibble()
+
+fit <- readRDS(here("Results", "fit4_phon.rds"))
+loos <- readRDS(here("Results", "loos.rds"))
+r2 <- readRDS(here("Results", "r2.rds"))
+
+item_effects <- fread(here("Results", "item_effects.csv"),
+                      na.strings = c("", "NA")) %>%
+  as_tibble()
+
+group_effects <- fread(here("Results", "group_effects.csv"),
+                       na.strings = c("", "NA")) %>%
+  as_tibble()
+
+corrs <- fread(here("Results", "corrs.csv"),
+               na.strings = c("", "NA")) %>%
+  as_tibble()
 
 valid_participant <- filter(participants, valid_participant) %>% pull(participant)
+
+clean_varnames <- function(x){
+  str_remove({{ x }}, "b_") %>% 
+    str_replace_all(., "pthn", "PTHN") %>% 
+    str_replace_all(., "similarity_phon", "Phon. Similarity") %>% 
+    str_replace_all(., ":", " \U000D7 ")
+}
+
 ```
@@
 
-This resulted in `r nrow(participants)-sum(participants$valid_participant)` participants excluded, leaving `r nrow(filter(participants, valid_participant))` valid participants and `r dat %>% filter(response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% nrow() %>% printnum(big.mark = ",")` valid trials.
+This resulted in `r nrow(participants)-sum(participants$valid_participant)` participants excluded, leaving `r nrow(filter(participants, valid_participant))` valid participants and `r accuracy %>% filter(response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% nrow() %>% printnum(big.mark = ",")` valid trials.
 
-```{r participants_group, echo=FALSE}
-
-dat %>%
-    filter(participant %in% valid_participant, response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% 
-    group_by(group, participant) %>% 
-    summarise(n_trials = n()) %>% 
-    group_by(group)%>%
-    summarise(n = n(),
-              n_trials = sum(n_trials)) %>% 
-    gt() %>% 
-    cols_label(group = "Group", n_trials = "Trials")
-
-```
-
-```{r participants_language, echo=FALSE}
-
-participants %>%
-    ggplot(aes(l2, fill = l2)) +
-    facet_wrap(~group, ncol = 2) +
-    geom_bar(stat = "count", show.legend = FALSE) +
-    labs(x = "L2", y = "Count", title = "What languages others than you native one do you know?") +
-    theme_classic() +
-    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-          axis.text.x = element_text(angle = 330, hjust = 0))
-
-```
-
-```{r participants_prof_l2, echo=FALSE}
+```{r participants_prof_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
   mutate(modality = str_to_sentence(str_remove(modality, "l2"))) %>%
-    ggplot(aes(l2, proficiency, fill = l2)) +
-    facet_wrap(group~modality) +
+  group_by(l2, country, modality) %>% 
+  summarise(mean = mean(proficiency, na.rm = TRUE),
+            n = n(), .groups = "drop") %>%
+  arrange(n) %>% 
+  mutate(l2 = fct_inorder(as.factor(l2))) %>% 
+  ggplot(aes(l2, mean, fill = l2)) +
+  facet_grid(country~modality) +
   geom_col(show.legend = FALSE) +
-    labs(x = "L2", y = "Self rated proficiency (0-5)") +
+  geom_text(aes(y = mean-0.5, label = paste0("n = ", n)), size = 3) +
+  labs(x = "L2", y = "Self-rated proficiency (0-5)") +
+  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
+  scale_y_continuous(limits = c(0, 5)) +
   theme_classic() +
   theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-          axis.text.x = element_text(angle = 330, hjust = 0))
+        axis.title.x = element_blank(),
+        axis.text = element_text(colour = "black"))
 
@@
 
-```{r participants_prof_spacat, echo=FALSE}
+```{r participants_prof_spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
   mutate(modality = str_to_sentence(str_remove(modality, "l2")),
-           language = str_to_sentence(language)) %>% 
-    ggplot(aes(proficiency)) +
-    facet_wrap(group~language~modality, nrow = 3) +
-    geom_bar(position = position_dodge(), fill = "black") +
-    labs(y = "Count", x = "Self rated proficiency (0-5)") +
+         language = as.factor(str_to_sentence(language))) %>% 
+  group_by(language, country, modality, proficiency) %>% 
+  summarise(n = n(), .groups = "drop") %>% 
+  ggplot(aes(x = proficiency, y = n, fill = country)) +
+  facet_grid(language~modality) +
+  geom_col() +
+  labs(y = "Count", x = "Self rated proficiency (0-5)", fill = "Country") +
   scale_x_continuous(limits = c(0, 5)) +
   theme_classic() +
-    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))
+  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
+        axis.text = element_text(colour = "black"))
 
@@
 
+## Stimuli
+
+```{r stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+plot_onset <- trials %>%
+  select(trial_id, word, language, matches("same_onset")) %>% 
+  rename_all(str_remove, "same_") %>% 
+  pivot_longer(matches("onset"), names_to = "language_pair", values_to = "onset") %>%  
+  mutate(language_pair = str_to_upper(str_remove(language_pair, "onset_")),
+         onset = ifelse(str_detect(onset, "same"), "Same", "Different")) %>% 
+  drop_na(onset) %>% 
+  ggplot(aes(onset, colour = language_pair, fill = language_pair)) +
+  facet_wrap(~language_pair) +
+  geom_bar(position = position_dodge(width = 1), na.rm = TRUE) +
+  labs(x = "Onset", y = "Counts",
+       colour = "Language pair", fill = "Language pair") +
+  theme_classic() +
+  theme(legend.position = "none")
+
+plot_subs <- trials %>%
+  select(trial_id, word, language, matches("close_")) %>% 
+  rename_all(str_remove, "close_") %>% 
+  pivot_longer(matches("substitutions"), names_to = "language_pair", values_to = "substitutions") %>% 
+  mutate(language_pair = str_to_upper(str_remove(language_pair, "substitutions_"))) %>% 
+  ggplot(aes(language_pair, substitutions, colour = language_pair, fill = language_pair)) +
+  geom_violin() +
+  stat_summary(fun = "mean", geom = "point", colour = "black", shape = "_", size = 10) +
+  labs(x = "Close substitutions", y = "Phonological similarity",
+       colour = "Language pair", fill = "Language pair") +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_classic() +
+  theme(legend.position = "none",
+        axis.title.x = element_blank())
+
+plot_sim <- trials %>%
+  select(trial_id, word, language, matches("similarity")) %>% 
+  rename_all(str_remove, "similarity_") %>% 
+  rename_at(vars(4:9), ~paste0("phon_", .)) %>% 
+  rename_at(vars(4:9), str_replace, "phon_ort_", "ort_") %>% 
+  pivot_longer(4:9, names_to = c("modality", "language_pair"), values_to = "similarity", names_sep = "_") %>% 
+  pivot_wider(names_from = modality, values_from = similarity) %>% 
+  mutate(language_pair = str_to_upper(language_pair)) %>% 
+  ggplot(aes(ort, phon, colour = language_pair, fill = language_pair)) +
+  geom_point(alpha = 0.5, size = 0.25) +
+  geom_smooth(formula = "y ~ x", method = "lm") +
+  labs(x = "Orthographical similarity", y = "Phonological similarity",
+       colour = "Language pair", fill = "Language pair") +
+  scale_y_continuous(limits = c(0, 1)) +
+  scale_x_continuous(limits = c(0, 1)) +
+  theme_classic()
+
+plot_onset + plot_subs + plot_sim + guide_area() +
+  plot_layout(nrow = 2, guides = "collect") &
+  theme(axis.text = element_text(colour = "black"),
+        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))
+
+```
 ## Data analysis
@@
 We fit Bayesian logistic models to model the probability of participants typing a correct translation in each trial.
-As predictors, we included the number of more frequent phonological neighbours of the translation (`PTHN`, standardised), whether both translation equivalents started with the same phoneme (`same_onset`, with levels `different`, `same_vowel`, `same_consonant`, helmert contrast comparing different vs.
-same and same vowel vs.
-same consonant), the highest number of consecutive phonemes shared (`consecutive_longest`, standardised), and our meausure of phonological similarity (`similarity_phop`, standardised).
+As predictors, we included the number of more frequent phonological neighbours of the translation (`PTHN`, standardised), whether both translation equivalents started with the same phoneme (`same_onset`, with levels `different`, `same_vowel`, `same_consonant`, sum-coded as `different` (-0.5) vs.
+`same` (0.5), the highest number of consecutive phonemes shared (`consecutive_longest`, standardised), and our measure of phonological similarity (`similarity_phop`, standardised).
 We first for a null model `fit0_null` that only included PTHN as predictor.
@@
 The models were finally, we fitted a model that only included `PTHN` and the traditional orthographical Levenshtein similarity (`similarity`, standardised), we test whether our measure predicted participants responses better than just relying on orthographic similarity.
-The models were implemente in the `brms` R package using the following formulas:
+The models were implemented in the `brms` R package using the following formulas:
 
-* `fit0_null <- correct | total ~ pthn + (1 + pthn | group/trial_id)`
-* `fit1_onset <- correct | total ~ pthn + same_onset + (1 +  same_onset | group/trial_id)`
-* `fit2_consec <- correct | total ~ pthn + same_onset + consecutive_longest + (1 + same_onset + consecutive_longest | group/trial_id)`
-* `fit3_phon <- correct | total ~ pthn + same_onset + consecutive_longest + similarity_phon + (1 + same_onset + consecutive_longest + similarity_phon | group/trial_id)`
-* `fit4_ort <- correct | total ~ pthn + similarity_ort + (1 + similarity_ort | group/trial_id)`
+* `fit1_pthn <- correct | trials(total) ~ pthn + (1 + pthn | group)`
+* `fit2_onset <- correct | trials(total) ~ pthn*same_onset + (1 +  pthn*same_onset | group)`
+* `fit3_close <- correct | trials(total) ~ pthn*close_substitutions + (1 + pthn*close_substitutions | group)`
+* `fit4_phon <- correct | trials(total) ~ pthn*similarity_phon + (1 + pthn*similarity_phon | group)`
+* `fit5_ort <- correct | trials(total) ~ pthn*similarity_ort + (1 + pthn*similarity_ort | group)`
 
@@
 
+```{r accuracy_anova, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+
+model <- lm(prop_correct ~ group, data = participants)
+anov <- anova(model) %>% 
+  tidy()
+
+es <- anov$sumsq[1]/(anov$sumsq[1]+anov$sumsq[2])
+pairs <- pairwise_comparisons(data = participants,
+                              x = group,
+                              y = prop_correct, type = "parametric",
+                              p.adjust.method = "bonferroni") %>% 
+  select(-c(label, test.details, p.value.adjustment, significance))
+
+```
+
+We performed a one-way ANOVA to analyse how participants' accuracy (the proportion of correct trials) changed across groups the three groups (ENG-SPA, ENG-CAT, and SPA-CAT).
+The `group` predictor explained a significant proportion of the variance, *F*(`r printnum(anov$df[1])`, `r printnum(anov$df[2])`) = `r printnum(anov$statistic[1])`, `r printp(anov$p.value[1])`, $\hat{\eta}^2$ = `r printnum(es)`).
+Follow-up *t*-tests revealed that Spanish-native speakers translating Catalan words were more accurate than English-native participants translating Spanish words, *t*(`r printnum(pairs$df[3])`) = `r printnum(pairs$t.value[3])`, `r printp(pairs$p.value[3])`, Cohen's *d* = `r printnum(pairs$t.value[3]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`), or Catalan words, *t*(`r printnum(pairs$df[2])`) = `r printnum(pairs$t.value[2])`, `r printp(pairs$p.value[2])`, Cohen's *d* = `r printnum(pairs$t.value[2]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`).
+We did not find evidence against accuracy being similar in both English-native groups, , *t*(`r printnum(pairs$df[1])`) = `r printnum(pairs$t.value[1])`, `r printp(pairs$p.value[1])`, Cohen's *d* = `r printnum(pairs$t.value[1]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`).
+Bonferroni corrections were applied in the aforementioned contrasts.
+
+```{r accuracy, echo=FALSE, fig.cap="Participants' overall accuracy across groups.
+The X-axis represents the native language of participants (English or Spanish) and the language participants listened during the task (Spanish or Catalan).
+The Y-axis represents the proportion of correct.
+Each datapoint corresponds to one participant.
+Horizontal black lines represent group means."}
+
+ggplot(participants, aes(group, prop_correct, fill = group)) +
+  geom_violin(colour = NA) +
+  geom_boxplot(fill = "white", width = 0.1) +
+  labs(x = "Group", y = "Accuracy (% correct trials)", fill = "Group") +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_classic() +
+  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
+        legend.position = "none",
+        axis.title.x = element_blank(),
+        axis.text = element_text(colour = "black"))
+```
+
 ## Model comparison
 
-```{r waic, echo=FALSE}
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
 
-waic <- comparison %>%
-    unlist() %>%
+r2_table <- r2 %>%
+  mutate(r2_ci95 = paste0("[", round(q2_5, 2), ", ", round(q97_5, 2), "]")) %>% 
+  rename(r2 = estimate, r2_se = est_error) %>% 
+  select(model, r2, r2_se, r2_ci95)
+loos$diffs %>% 
   as.data.frame() %>% 
-    rownames_to_column("model")
-
-colnames(waic) <- c("Model", "WAIC")
-
-waic %>%
-    gt()
+  rownames_to_column("model") %>% 
+  select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>% 
+  left_join(r2_table) %>% 
+  relocate(model, r2, r2_se, r2_ci95) %>% 
+  gt() %>% 
+  fmt_number(c(2, 3, 5, 6, 7, 8)) %>% 
+  cols_label(model = "Model", r2 = "R2", r2_se = "SE", r2_ci95 = "95% CrI", elpd_diff = "\U0394 ELPD", se_diff = "\U0394 SE", elpd_loo = "ELPD LOO", se_elpd_loo = "SE ELPD LOO") %>%
+  tab_spanner(label = "R2", columns = matches("r2")) %>% 
+  tab_spanner(label = "PSIS-LOO", columns = matches("loo|diff"))
 
@@
 
-```{r posterior_draws, echo=FALSE}
+```{r posterior_coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+
+summary(fit)$fixed %>%
+  as.data.frame() %>% 
+  rownames_to_column("term") %>% 
+  clean_names() %>% 
+  mutate(ci95 = paste0("[", round(l_95_percent_ci, 2), ", ", round(u_95_percent_ci, 2), "]"),
+         term = clean_varnames(term)) %>% 
+  select(-matches("percent")) %>% 
+  relocate(term, estimate, est_error, ci95) %>% 
+  gt(rowname_col = "term") %>% 
+  fmt_number(columns = vars(estimate, est_error, rhat, bulk_ess, tail_ess)) %>% 
+  cols_label(estimate = "Estimate", est_error = "SE", ci95 = "95% CrI", rhat = "R-hat", bulk_ess = "Bulk ESS", tail_ess = "Tail ESS")
+
+```
+
+```{r posterior_draws, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Posterior distribution of the extended model's (fit3_phon) coefficients.
+The X-axis represents all the values that each coefficient can take.
+The Y-axis represents the probability density at each value, i.e., the probability of the coefficient taking a specific value.
+Grey bars colour-code highest density intervals (HDI) of different width.
+Each interval indicates the narrowest interval that can be define within each distribution capturing a given proportion of the probability mass (e.g., given our data there is a 95% probability that the true value of a coefficient lies within the 95% HDI interval."}
 
 posterior %>% 
+  mutate(.variable = clean_varnames(.variable)) %>% 
   ggplot(aes(.value, .variable)) +
   geom_vline(xintercept = 0, linetype = "dashed") +
-    stat_slab(fill = "black") + 
-    stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.15)) +
+  stat_slab(show.legend = FALSE, fill = "grey30") + 
+  stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.2)) +
   labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
        title = "Posterior distribution of coefficients of fixed effects") +
-    scale_colour_brewer(palette = "Blues") +
+  scale_colour_brewer(palette = "Oranges") +
   theme_classic() +
@@
         legend.position = "top",
-          legend.key = element_rect(fill = "transparent"))
+        legend.key = element_rect(fill = "transparent", colour = NA),
+        axis.text = element_text(colour = "black"))
 
@@
 
-## Posterior predictors
+## Posterior predictions
 
-```{r posterior_predictions, echo=FALSE}
+```{r posterior_predictions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Predicted probability of correct translation across different measures of phonological similarity scores.
+Predictions were generated by sampling 1000 random draws from the posterior distribution of each predictor in the extended model (fit3_phon)."}
 
 predictions %>%
-    ggplot(aes(similarity_phon, .value, colour = same_onset, fill = same_onset)) +
-    facet_grid(consecutive_longest~pthn)+
-    stat_summary(fun = mean, geom = "line", size = 1) +
-    geom_vline(xintercept = 0, colour= "grey") +
-    geom_hline(yintercept = 0.5, colour= "grey") +
+  ggplot(aes(similarity_phon, .value)) +
+  geom_point(data = accuracy, aes(y = correct_coded), shape = "|") +
+  geom_line(aes(group = .draw), alpha = 0.25, colour = "grey") +
+  stat_summary(fun = "mean", geom = "line", colour = "blue") +
   labs(x = "Phonological similarity", y = "Probability of correct translation",
-         fill = "Credible interval", colour = "Onset",
-         title = "Effect of phonological similarity on probability of correct translation",
-         subtitle = "Model predictions are displayed for different quartiles of PTHN") +
-    scale_fill_brewer(palette = "Set1") +
-    scale_colour_brewer(palette = "Set1") +
+       fill = "PTHN", colour = "Group") +
   theme_classic() +
-    theme(legend.position = "top")
-
+  theme(axis.text = element_text(colour = "black"))
 ```
@@
 
-### Item-level effects
+```{r group_effects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Group-level effects.
+We specified a maximal random effects structure.
+As a result, intercepts and slopes were estimated for each group.
+This figure shows how the estimated intercepts and slopes varied across groups.
+There are not any remarkable differences across groups, except for the slope of phonological similarity.
+The accuracy of Spanish-native speakers translating from Catala \\(SPA-CAT\\) was slightly boosted by the phonological similarity between translation equivalents in comparison with the rest of the groups \\(ENG-SPA, ENG-CAT\\)."}
 
-```{r item_effects, echo=FALSE}
-
-item_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  ggplot(aes(estimate, fill = param)) +
+group_effects %>%
+  mutate(param = clean_varnames(param)) %>% 
+  ggplot(aes(.value, group)) +
   facet_wrap(~param, scales = "free") +
-  geom_histogram(bins = 20) +
-  geom_rug(alpha = 0.1) +
+  stat_interval(.width = c(0.95, 0.89, 0.67, 0.50)) +
   geom_vline(xintercept = 0, linetype = "dashed") +
-  labs(x = "Estimate", y = "Density", fill = "Parameter") +
-  scale_fill_brewer(palette = "Set1") +
+  labs(x = "Estimate", y = "Group", colour = "HDI (%)") +
+  scale_colour_brewer() +
   theme_classic() +
-  theme(legend.position = "none",
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
+  theme(legend.position = "top",
+        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+        axis.title.y = element_blank(),
+        axis.text = element_text(colour = "black"))
 
@@
 
-### Group-level effects
+```{r correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Pairwise correlations between group-level effects. This figure displaya"}
 
-```{r goup_effects_table, echo=FALSE}
-
-group_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(hdi = paste0("[", lower, ", ", upper, "]")) %>% 
-  select(param, group, error, hdi) %>% 
-  group_by(param) %>% 
-  gt() %>% 
-  cols_label(group = "Group", error = "SE", hdi = "95% HDI")
+corrs %>% 
+  mutate(.variable = str_remove(.variable, "cor_group__"),
+         .variable = str_replace(.variable, "__" , " - "),
+         .variable = clean_varnames(.variable)) %>% 
+  ggplot(aes(.value, .variable)) +
+  stat_interval(.width = c(0.95, 0.89, 0.67, 0.50)) +
+    geom_vline(xintercept = 0, linetype = "dashed") +
+  labs(x = "Estimate", y = "Group", colour = "HDI (%)") +
+  scale_colour_brewer(palette = "Oranges") +
+  theme_classic() +
+  theme(legend.position = "top",
+        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+        axis.title.y = element_blank(),
+        axis.text = element_text(colour = "black"))
 
 ```
-
-```{r group_effects, echo=FALSE}
-
-group_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  ggplot(aes(group, estimate)) +
-  facet_wrap(~param) +
-  geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = lower, ymax = upper, colour =  group), width = 0, size = 5) +
-  geom_errorbar(aes(ymin = estimate-error, ymax = estimate+error), width = 0, size = 1, alpha = 0.5) +
-  geom_point(size = 3) +
-  labs(x = "Group", y = "Estimate", fill = "Group") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_classic() +
-  theme(legend.position = "none",
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
-
-```
-
-### Correlations
-
-```{r correlations, echo=FALSE}
-
-corrs %>% 
-  filter(str_detect(variable, "trial_id")) %>%
-  mutate(variable = str_remove(variable, "cor_group:trial_id__"),
-         variable = str_replace(variable, "__" , " - ")) %>% 
-  ggplot(aes(value, variable, xmin = lower, xmax = upper)) +
-  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
-  geom_errorbarh(show.legend = FALSE, height = 0) +
-  geom_point(size = 2, show.legend = FALSE) +
-  labs(x = "Estimate", y = "Correlation") +
-  theme_classic() +
-  theme(axis.title.y = element_blank())
-
-```

### 2020-08-06 Manuscript/te_manuscript.Rmd
Prepare report for meeting

@@
 
-Each trial started with a yellow fixation dot presented during one second on the center of the screen over a black background.
+Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
 After one second, the audio started playing while the dot remained being displayed until the audio offset.
@@
 We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance.
-We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct reponses as output.
-Continuous predictors were standadised and categorical predictors were sum-coded [@schad2020].
+We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.
+Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020].
 Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013].

### 2020-08-06 Code/processing.R

@@ -1,0 +1,203 @@
+#### analysius: Process and analyse data #########
+
+#### set up #########################################
+
+# load packages
+library(dplyr)      # for manipulating variables
+library(tidyr)      # for reshaping datasets
+library(ggplot2)    # for visualising plots
+library(stringr)    # for manipulating string characters
+library(readxl)     # for importing Excel spreadsheets
+library(purrr)      # for working with lists
+library(janitor)    # for cleaning variable names
+library(lubridate)  # for working with dates
+library(forcats)    # for working with factors
+library(data.table) # for importing data
+library(stringdist) # for calculating Levenshtein distance
+library(here)       # for locating files in the repository with reproducibility
+
+# create/load functions
+source(here("Code", "functions.R")) # helper functions
+
+# set parameters
+set.seed(888) # for reproducibility
+practice_trials <- c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)
+spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
+
+#### import trial data ###############################
+
+# frequency and phonological neighbourhoods for English (from CLEARPOND)
+clearpond_eng <- fread(here("Data", "clearpond_Eng.csv"), na.strings = "") %>%
+    select(-matches("spa")) %>% 
+    distinct(Word_eng, .keep_all = TRUE)
+clearpond_spa <- fread(here("Data", "clearpond_Spa.csv"), na.strings = "") %>%
+    rename(pthn_spa = PTHN) %>% 
+    distinct(Word, .keep_all = TRUE)
+
+
+# word pair comparison statistics - similarity, same_onset, consecutive_longest, close_substitutions
+# statistics for each language pair are different columns
+# only includes statistics for words that appear in trials for the specific language
+similarity_engspa <- read_excel(here("Data", "Output_similarity.xlsx"), sheet = "English-Spanish") %>%
+    select(-c("Phon_eng", "Phon_spa", "check_diff")) 
+similarity_engcat <- read_excel(here("Data", "Output_similarity.xlsx"), sheet = "English-Catalan") %>%
+    select(-c("Phon_eng", "Phon_cat", "check_diff"))
+similarity_spacat <- read_excel(here("Data", "Output_similarity.xlsx"), sheet = "Spanish-Catalan") %>%
+    select(-c("Phon_cat", "Phon_spa", "check_diff"))
+
[…]

### 2020-08-06 Code/trials_stats.R

@@ -1,0 +1,32 @@
+
+# For compiling the different statistics for Translation Elicitation
+
+library(readxl)
+library(dplyr)
+library(here)
+
+trials <- read_excel(here("Data", "trials.xlsx")) # trial list
+
+# Frequency and phonological neighbourhoods for English (from CLEARPOND)
+clearpond <- read.csv(here("Data", "clearpond_Eng.csv") )
+
+# Word pair comparison statistics - similarity, same_onset, consecutive_longest, close_substitutions
+# Statistics for each language pair are different columns
+# Only includes statistics for words that appear in trials for the specific language
+similarity_engspa <- read_excel(here("Data", "Output_similarity.xlsx"), sheet = "English-Spanish") %>%
+  select(-c("Phon_eng", "Phon_spa", "check_diff")) 
+similarity_engcat <- read_excel(here("Data", "Output_similarity.xlsx"), sheet = "English-Catalan") %>%
+  select(-c("Phon_eng", "Phon_cat", "check_diff"))
+similarity_spacat <- read_excel(here("Data", "Output_similarity.xlsx"), sheet = "Spanish-Catalan") %>%
+  select(-c("Phon_cat", "Phon_spa", "check_diff"))
+
+# replace name to match up to trial list - similarity is calculated for pork
+similarity_engcat$Word_eng[similarity_engcat$Word_eng=="pork"]<-"pig/pork"
+
+
+trials <- left_join(trials, clearpond, by = c("translation_english_orthography" = "Word_eng"))
+trials <- left_join(trials, similarity_engspa, by = c("translation_english_orthography" = "Word_eng", "translation_spanish_orthography" = "Word_spa"))
+trials <- left_join(trials, similarity_engcat, by = c("translation_english_orthography" = "Word_eng", "translation_catalan_orthography" = "Word_cat"))
+trials <- left_join(trials, similarity_spacat, by = c("translation_spanish_orthography" = "Word_spa", "translation_catalan_orthography" = "Word_cat"))
+
+write.table(trials, here("Data", "trial_stats.csv"), sep = ",", dec = ".", row.names = FALSE)

### 2021-02-04 Rmd/te_manuscript.Rmd
Rename Rmd folder

--- 2020-08-06 Manuscript/te_manuscript.Rmd
@@
 
-library(dplyr)
-library(tidyr)
-library(tibble)
-library(stringr)
-library(ggplot2)
-library(here)
+library(tidyverse)
 library(data.table)
@@
 library(lubridate)
+library(here)
 
-participants <- fread(here("Data", "01_participants.csv"), na.strings = "") %>% as_tibble() %>% mutate(date = as_date(date))
-dat <- fread(here("Data", "01_processed_coded.csv"), na.strings = "") %>% as_tibble()
-posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()
-predictions <- fread(here("Results", "posterior_predictions.csv"), na.strings = "") %>% as_tibble()
+participants <- readRDS(here("Results", "participants.rds"))
+stimuli <- readRDS(here("Results", "stimuli.rds"))
+accuracy <- readRDS(here("Results", "accuracy.rds"))
 fit <- readRDS(here("Results", "fit3_phon.rds"))
 comparison <- readRDS(here("Results", "model_comparison.rds"))
-item_effects <- fread(here("Results", "item_effects.csv"), na.strings = "") %>% as_tibble()
-group_effects <- fread(here("Results", "group_effects.csv"), na.strings = "") %>% as_tibble()
-corrs <- fread(here("Results", "corrs.csv"), na.strings = "") %>% as_tibble()
 
@@
 Third, participants completed the experimental task.
-Before the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
 They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
 English participants were randomly assigned to the list of Catalan or Spanish trials.
-Spanish participants completed the list of Catalan trials.
-
-### Design
+Participants in the Catalan list were presented with `r count(stimuli, language)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, language)$n[2]` trials.
 
 Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio offset.
+After one second, the audio started playing while the dot remained being displayed until the audio ended.
 Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
@@
 We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Typos were allowed, however.
+Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
 Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
@@
 ### Stimuli
+
+Participants listened to one audio file in each trial.
+This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infancia of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@boersma2018].
+The average duration of the audios was `r printnum(mean(stimuli$duration), digits = 2)` (*SD* = `r printnum(sd(stimuli$duration), digits = 2)`, *Min* = `r printnum(min(stimuli$duration), digits = 2)`, *Max* = `r printnum(max(stimuli$duration), digits = 2)`).
+The average duration of the Catalan audios was `r printnum(mean(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`), and the average duration of the Spanish audios was `r printnum(mean(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`).
+The average lexical frequency of the words was `r printnum(mean(stimuli$frequency_zipf), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf), digits = 2)`).
+The average lexical frequency of the Catalan audios was `r printnum(mean(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`), and the average lexical frequency of the Spanish audios was `r printnum(mean(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`).
+Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT [@boada2013] and SUBTLEX-ESP [@cuetos2011], respectively, and are expressed as Zipf scores [@van-heuven2012; @zipf1949].
 
@@
 All analyses were performed in R environment [@rcoreteam2019].
-We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] to impute missing data, and the `brms` R package to [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
+We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
+Audio duration, intensity and pitch were extracted using the `PraatR` R package [@albin2014].
 

### 2021-02-04 Communications/2020-07-30_meeting.Rmd
Ingnore files

@@
 a) Oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`)
-b) Language impairment (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION?
-(Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No)".
-However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.]
+b) Language impairment (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct.
+This item was phrased as *Do you have normal or corrected-to-normal VISION?
+(Yes/No)* in English, and as *¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
+(Sí/No)*.
+Surprisingly, the proportion of Spanish participants that reported visual impairment was implausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
+This is possibly due to some participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.]
 c) Did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`)
@@
 
-```{r participants_prof_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants_prof_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Second languages.
+Some participants reported being familiar with one or more langugaes other than English (for UK participants) or Spanish (for participants in Spain).
+We asked them to self-report their oral and written proficiency in those languages in a 0-to-5 scale.
+This graph shows the mean proficiency reported by participants that reported being familiar with each language.
+the number of participants that reported being familiar with each language is reported on top of each bar.
+Data are presented for UK and Spanish participants separately."}
 
@@
   ggplot(aes(l2, mean, fill = l2)) +
-  facet_grid(country~modality) +
+  facet_wrap(country~modality) +
   geom_col(show.legend = FALSE) +
@@
 
-```{r participants_prof_spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants_prof_spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="We also asked participants to rate their oral and written proficiency in the target languages (Spanish and Catalan).
+UK participants were aked about Spanish and Catalan while Spanish participants were aonly asked about Catalan.
+The graph represents how many participants (self-) reported each levels of proficiency (0-5).
+Data are presented separately for UK and Spanish participants."}
 
@@
   ggplot(aes(x = proficiency, y = n, fill = country)) +
-  facet_grid(language~modality) +
+  facet_wrap(language~modality) +
   geom_col() +
@@
 * 1000 - 2500 ms: audio
-* 2500 ms - Press RETURN key
+* 2500 ms - Press RETURN key: type response
 
@@
 
-```{r stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Distribution of trial words from the three groups accross different levels of the predictors."}
 
@@
 plot_subs <- trials %>%
-  select(trial_id, word, language, matches("close_")) %>% 
+  select(trial_id, word, language, matches("close_|similarity"), -matches("ort")) %>% 
   rename_all(str_remove, "close_") %>% 
-  pivot_longer(matches("substitutions"), names_to = "language_pair", values_to = "substitutions") %>% 
+  pivot_longer(matches("substitutions|similarity"), names_to = c("measure", "language_pair"), values_to = "value", names_sep = "_") %>% 
+  pivot_wider(names_from = "measure", values_from = "value") %>% 
   mutate(language_pair = str_to_upper(str_remove(language_pair, "substitutions_"))) %>% 
-  ggplot(aes(language_pair, substitutions, colour = language_pair, fill = language_pair)) +
-  geom_violin() +
-  stat_summary(fun = "mean", geom = "point", colour = "black", shape = "_", size = 10) +
-  labs(x = "Close substitutions", y = "Phonological similarity",
+  ggplot(aes(similarity, substitutions, colour = language_pair, fill = language_pair)) +
+  geom_point(alpha = 0.5, shape = 1) +
+  geom_smooth(formula = "y ~ x", method = "lm") +
+  labs(x = "Close substitutions", y = "Base similarity",
        colour = "Language pair", fill = "Language pair") +
@@
   theme_classic() +
-  theme(legend.position = "none",
-        axis.title.x = element_blank())
+  theme(legend.position = "none")
 
@@
   ggplot(aes(ort, phon, colour = language_pair, fill = language_pair)) +
-  geom_point(alpha = 0.5, size = 0.25) +
+  geom_point(alpha = 0.5, shape = 1) +
   geom_smooth(formula = "y ~ x", method = "lm") +
-  labs(x = "Orthographical similarity", y = "Phonological similarity",
+  labs(x = "Orthographical similarity", y = "Base similarity",
        colour = "Language pair", fill = "Language pair") +
@@
 ```
+
 ## Data analysis
 
-We fit Bayesian logistic models to model the probability of participants typing a correct translation in each trial.
-As predictors, we included the number of more frequent phonological neighbours of the translation (`PTHN`, standardised), whether both translation equivalents started with the same phoneme (`same_onset`, with levels `different`, `same_vowel`, `same_consonant`, sum-coded as `different` (-0.5) vs.
-`same` (0.5), the highest number of consecutive phonemes shared (`consecutive_longest`, standardised), and our measure of phonological similarity (`similarity_phop`, standardised).
-We first for a null model `fit0_null` that only included PTHN as predictor.
-We then extended the model incrementally by adding other predictors.
-The models were finally, we fitted a model that only included `PTHN` and the traditional orthographical Levenshtein similarity (`similarity`, standardised), we test whether our measure predicted participants responses better than just relying on orthographic similarity.
-The models were implemented in the `brms` R package using the following formulas:
+We fit Bayesian logistic models to model the probability of participants typing a correct translation in each trial. As predictors, we included:
+
+* *PTHN*: the number of more frequent phonological neighbours of the translation (`PTHN`, standardised)
+* *Onset*: do both translation equivalents start with the same phoneme? (`same_onset`, with levels `Different`, and `Same`, sum-coded as `different` (-0.5) vs. `same` (0.5)
+* *Close substitutions*: proportion of phoneme matches differing by 1 level on 1 feature (`close_substitutions`, standardised).
+* *Base similarity*: Proportion of phoneme matches with identical codes (`similarity_phon`, standardised).
+
+We first for a null model `fit1_pthn` that only included PTHN as predictor.
+We then extended the model by fitting three other models that included one of the predictors of interest (*Onset*, *Close substitutions*, or *Base similarity*) in interaction with *PTHN*.
+Finally, we fitted a model that included the traditional orthographical Levenshtein similarity (`similarity_ort`, standardised) to test whether our phonology-based measure predicted participants responses better than just relying on orthography.
+The models included a maximal random-effects structure (with random intercepts and slopes for all fixed effects) and were implemented in the `brms` R package using the following formulas:
 
@@
 The `group` predictor explained a significant proportion of the variance, *F*(`r printnum(anov$df[1])`, `r printnum(anov$df[2])`) = `r printnum(anov$statistic[1])`, `r printp(anov$p.value[1])`, $\hat{\eta}^2$ = `r printnum(es)`).
-Follow-up *t*-tests revealed that Spanish-native speakers translating Catalan words were more accurate than English-native participants translating Spanish words, *t*(`r printnum(pairs$df[3])`) = `r printnum(pairs$t.value[3])`, `r printp(pairs$p.value[3])`, Cohen's *d* = `r printnum(pairs$t.value[3]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`), or Catalan words, *t*(`r printnum(pairs$df[2])`) = `r printnum(pairs$t.value[2])`, `r printp(pairs$p.value[2])`, Cohen's *d* = `r printnum(pairs$t.value[2]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`).
+Follow-up *t*-tests revealed that Spanish-native speakers translating Catalan words (ENG-SPA, ENG-CAT), *t*(`r printnum(pairs$df[3])`) = `r printnum(pairs$t.value[3])`, `r printp(pairs$p.value[3])`, Cohen's *d* = `r printnum(pairs$t.value[3]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`), or Catalan words, *t*(`r printnum(pairs$df[2])`) = `r printnum(pairs$t.value[2])`, `r printp(pairs$p.value[2])`, Cohen's *d* = `r printnum(pairs$t.value[2]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`).
 We did not find evidence against accuracy being similar in both English-native groups, , *t*(`r printnum(pairs$df[1])`) = `r printnum(pairs$t.value[1])`, `r printp(pairs$p.value[1])`, Cohen's *d* = `r printnum(pairs$t.value[1]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`).
@@
 
-```{r posterior_coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}
+```{r posterior_coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE, fig.cap="Point estimations (median), standard errors, 95% credible intervals, and diagnostics of the fixed efffects of the model."}
 
@@
         legend.key = element_rect(fill = "transparent", colour = NA),
-        axis.text = element_text(colour = "black"))
+        axis.text = element_text(colour = "black"),
+        axis.title.y = element_blank())
 
@@
 predictions %>%
-  ggplot(aes(similarity_phon, .value)) +
-  geom_point(data = accuracy, aes(y = correct_coded), shape = "|") +
+  ggplot(aes(similarity_phon, .value, colour = group)) +
+  geom_point(data = accuracy, aes(y = correct_coded), shape = 1, alpha = 0.1) +
   geom_line(aes(group = .draw), alpha = 0.25, colour = "grey") +
-  stat_summary(fun = "mean", geom = "line", colour = "blue") +
+  stat_summary(fun = "mean", geom = "line", colour = "black", size = 1) +
   labs(x = "Phonological similarity", y = "Probability of correct translation",
@@
 There are not any remarkable differences across groups, except for the slope of phonological similarity.
-The accuracy of Spanish-native speakers translating from Catala \\(SPA-CAT\\) was slightly boosted by the phonological similarity between translation equivalents in comparison with the rest of the groups \\(ENG-SPA, ENG-CAT\\)."}
+The accuracy of Spanish-native speakers translating from Catala (SPA-CAT) was slightly boosted by the phonological similarity between translation equivalents in comparison with the rest of the groups (ENG-SPA and ENG-CAT)."}
 
@@
 
-```{r correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Pairwise correlations between group-level effects. This figure displaya"}
+```{r correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Pairwise correlations between group-level effects.
+This figure displays the correlation between the posterior draws of all pairs of coefficients in the random effects structure.
+All 95% HDI of the correlation included zero, and no remarkable differences can be found across groups."}
 

### 2021-05-07 jTRACE/compare_jtrace_translation.Rmd
Add files via upload
R script to match jtrace output with translation elicitation task data for analysis

---
title: "Comparing jTRACE Output with Translation Elicitation Task"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggplot2)

```

```{r load jtrace output}
filedir <- "data/" 
# directory with save files (
  ## 1 file per input word, with 1 output word per column
  ## files should be named with input word orthography and language (e.g. "gato_spanish.xlsx")
  ## files should be xlsx (csv doesn't deal well with special characters)
  ## folder should not contain any other files
fnames <- dir(filedir) # extract list of file names in directory

trace_output <- data.frame() # create empty data frame

# for loop to extract highest activations for each input word
for (i in 1:length(fnames)){
  filename <- paste(filedir, fnames[i], sep = "") # get file name
  temp <- read_excel(filename) # read file

  temp <- apply(temp,2,max) # get highest value in each column
  temp <- as.data.frame(temp)
  temp <- cbind(newColName = rownames(temp), temp) # change index column (which has output items) into column 1 for easier organisation
  rownames(temp) <- 1:nrow(temp)
  names(temp)[1] <- "output" # (rename col) output item
  names(temp)[2] <- "activation" # (rename col) max activation value

  temp <- temp %>%
    filter(!output == "cycle") %>% # remove row corresponding to cycle number
    mutate(input = fnames[i]) %>% # specify input word
    arrange(desc(activation))

  trace_output <- bind_rows(trace_output, temp)
}

trace_output <- trace_output %>%
  mutate(input = str_remove(input, ".xlsx")) %>%
  separate(input, c("input", "language"), "_")

```

```{r match word}

# xlsx file with 1 column for TRACE transcriptions and 1 column for orthography (English words) 
df_transcription <- read_excel("trace_transcription.xlsx") %>%
  select(trace_eng, ort_eng) %>%
  unique()

# Match the jTRACE output transcriptions with their orthography
trace_output <- left_join(trace_output, df_transcription, by = c("output" = "trace_eng")) %>%
  rename(output_word = ort_eng)

```

```{r load translation task}

# Cleaned file with participant answers, 1 row per trial per participant
TranslationTask <- read.csv("translation_engdata_cleaned.csv") %>%
  mutate(test_language = tolower(test_language))

# count the number of participants who gave each answer in response to a given trial
df_answer <- TranslationTask %>%
  group_by(test_language, word, answer) %>%
  summarise(n_participant = n()) %>%
  group_by(test_language, word) %>%
  mutate(pct_participant = n_participant/sum(n_participant)*100, .groups = "drop") %>%
  filter(!n_participant == 1) # remove answers only given by 1 participant as idiosyncratic answers?

df_match <- left_join(df_answer, trace_output, by = c("word" = "input", "answer" = "output_word", "test_language" = "language"))

```

```{r plot, warning = FALSE}

ggplot(df_match, aes(x = activation, y = pct_participant)) +
  geom_smooth() +
  geom_point()

```

### 2021-05-20 Rmd/compare_jtrace_translation.Rmd
Integrate jtRACE script by @SiowSerene into the main report, clean repository, update data files.

---
title: "Comparing jTRACE Output with Translation Elicitation Task"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(tidyverse)
library(ggplot2)

```

```{r load jtrace output}
filedir <- "data/" 
# directory with save files (
  ## 1 file per input word, with 1 output word per column
  ## files should be named with input word orthography and language (e.g. "gato_spanish.xlsx")
  ## files should be xlsx (csv doesn't deal well with special characters like ^ and @)
  ## folder should not contain any other files

# Note: it's effortful to have to convert to xlsx (jTRACE doesn't allow direct xlsx export), but it's the most reliable method to preserve special characters I've found so far. Would welcome ideas

# Export jTRACE output with Top 100 items and 100 cycles (arbritary value)
# Max ad-hoc activation

fnames <- dir(filedir) # extract list of file names in directory

trace_output <- data.frame() # create empty data frame

# for loop to extract highest activations for each input word
for (i in 1:length(fnames)){
  filename <- paste(filedir, fnames[i], sep = "") # get file name
  temp <- read_excel(filename) # read file

  temp <- apply(temp,2,max) # get highest value in each column
  temp <- as.data.frame(temp)
  temp <- cbind(newColName = rownames(temp), temp) # change index column (which has output items) into column 1 for easier organisation
  rownames(temp) <- 1:nrow(temp)
  names(temp)[1] <- "output" # (rename col) output item
  names(temp)[2] <- "activation" # (rename col) max activation value

  temp <- temp %>%
    filter(!output == "cycle") %>% # remove row corresponding to cycle number
    mutate(input = fnames[i]) %>% # specify input word
    arrange(desc(activation)) %>%
    filter(!activation <= 0) # remove low activation (maybe even apply a higher threshold? TBD)

  trace_output <- bind_rows(trace_output, temp)
}

trace_output <- trace_output %>%
  mutate(input = str_remove(input, ".xlsx")) %>%
  separate(input, c("input", "language"), "_")

```

```{r match word}

# xlsx file with 1 column for TRACE transcriptions and 1 column for orthography (English words) 
df_transcription <- read_excel("trace_transcription.xlsx") %>%
  select(trace_eng, ort_eng) %>%
  unique()

# Match the jTRACE output transcriptions with their orthography
trace_output <- left_join(trace_output, df_transcription, by = c("output" = "trace_eng")) %>%
  rename(output_word = ort_eng)

```

```{r load translation task}

# Cleaned file with participant answers, 1 row per trial per participant
TranslationTask <- read.csv("translation_engdata_cleaned.csv") %>%
  mutate(test_language = tolower(test_language))

# count the number of participants who gave each answer in response to a given trial
df_answer <- TranslationTask %>%
  group_by(test_language, word, answer) %>%
  summarise(n_participant = n(), .groups = "drop") %>%
  group_by(test_language, word) %>%
  mutate(pct_participant = n_participant/sum(n_participant)*100) #%>%
#  filter(!n_participant == 1) # remove answers only given by 1 participant as idiosyncratic answers?

```

### Predicting participant response by jTRACE

```{r plot, warning = FALSE}

# Predicting participant response by jTRACE

df_match <- full_join(trace_output, df_answer, by = c("input" = "word", "output_word" = "answer", "language" = "test_language")) %>%
  semi_join(., trace_output, by = c("input", "language"))

# answers
print("Presented words and the answer that was given by the most number of participants for that trial")

# If "output" is <NA>, the word is not in the currently-used jTRACE lexicon
df_match %>%
  filter(language == "spanish") %>% 
  arrange(input, desc(pct_participant)) %>%
  group_by(input) %>%
  slice(1) %>%
  select(input, output_word, pct_participant, output, activation) %>%
  arrange(desc(pct_participant))

df_match %>%
  filter(language == "catalan") %>% 
  arrange(input, desc(pct_participant)) %>%
  group_by(input) %>%
  slice(1) %>%
  select(input, output_word, pct_participant, output, activation) %>%
  arrange(desc(pct_participant))

# if no data for a given output word is available in df_answer, it means no participants gave that word as a response - recode as 0
df_match$pct_participant[is.na(df_match$pct_participant)] <- 0

# if no data for a given output word is available in trace output, it means that activation is outside of Top 100 words and/or below 0
df_match$activation[is.na(df_match$activation)] <- 0

# However, this is also constrained by the size of TRACE's lexicon, so subset only the words that do appear in the lexicon
df_match_trace <- semi_join(df_match, df_transcription, by = c("output_word" = "ort_eng"))

# plot
ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
  geom_smooth() +
  geom_point() +
  ggtitle("jTRACE activation against % participants who gave word as response")

# correlation test
cor.test(df_match_trace$activation, df_match_trace$pct_participant)

# plot by word
ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
  geom_smooth() +
  geom_point() +
  xlab("jTRACE activation value for the word") +
  ylab("% participants who responded with the word") +
  ggtitle("jTRACE activation against % participants who gave word as response") +
  facet_wrap(language~input)

```

### 2021-05-20 Rmd/report.Rmd
Integrate jtRACE script by @SiowSerene into the main report, clean repository, update data files.

---
title: "Translation Elicitation"
output:
  html_document:
    df_print: paged
    toc: true
    toc_depth: 2
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = TRUE,
    warning = TRUE, 
    cache.extra = knitr::rand_seed,
    out.width = "100%",
    results = "asis",
    dev.args = list(png = list(type = "cairo"))
)
options(knitr.kable.NA = '-')
```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# load packages
library(tidyverse)
library(gt)
library(ggdist)
library(readxl)
library(tidytext)
library(brms)
library(tidybayes)
library(patchwork)
library(mice)
library(here)

# set params
source(here("R", "utils.R"))

# import data
trials <- readRDS(here("Data", "trials.rds")) %>% 
    mutate(
        test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"),
        zipf = log10(frequency)+3,
        pthn_log = log(pthn),
        overlap_stress = ifelse(overlap_stress==1, "Same", "Different")
    )
accuracy <- readRDS(here("Data", "accuracy.rds"))
d <- readRDS(here("Data", "accuracy.rds")) %>% 
    # typos are considered correct responses
    mutate_at(vars(onset, overlap_stress, group), as.factor) %>% 
    # center predictors
    mutate_at(
        vars(consonant_ratio, vowel_ratio, pthn, frequency),
        function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
    # impute missing data
    mice(m = 5, print = FALSE, method = "pmm") %>% 
    complete() %>% 
    as_tibble() %>% 
    arrange(trial_id, group)

participants <- readRDS(here("Data", "participants.rds")) %>% 
    mutate(valid_participant = ifelse(valid_participant, "Valid", "Not valid"))
fit <- readRDS(here("Results", "fit.rds"))
coefs <- readRDS(here("Results", "coefficients.rds")) %>% 
    select(term, estimate, std.error, conf.low, conf.high)

```

# Experimental task: Translation Elicitation

## Methods

### Participants {.tabset .tab-pills}

#### Summary

```{r participants_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

ggplot(participants, aes(group, fill = valid_participant)) +
    geom_bar() +
    labs(x = "Group", y = "# participants", fill = "Valid?") +
    theme_custom() +
    theme(
        legend.title = element_blank(),
        legend.position = "top",
        axis.title.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )

```

#### Demographic information

```{r participants_demo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants_valid <- filter(participants, valid_participant=="Valid")
p_sex <- ggplot(participants_valid, aes(group, fill = sex)) +
    geom_bar() +
    labs(x = "Group", y = "# participants", fill = "Sex") +
    theme(axis.title.x = element_blank())

p_age <- ggplot(participants_valid, aes(as.factor(age), fill = group, colour = group)) +
    geom_bar() +
    labs(x = "Age (years)", y = "# participants", fill = "Group", colour = "Group")

(p_sex + p_age) &
    plot_layout() &
    plot_annotation() &
    theme_custom() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )

```

#### Second language

```{r participants_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_valid %>% 
    count(group, l2) %>% 
    mutate(l2 = reorder_within(l2, n, within = group)) %>% 
    ggplot(aes(x = reorder(l2, -n), n,  fill = l2)) +
    facet_wrap(~group, scales = "free_x") +
    geom_col() +
    labs(x = "L2", y = "# participants", fill = "L2") +
    scale_x_reordered() +
    theme_custom() +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        axis.text.x = element_text(angle = 90)
    )

```

#### Language profile proficiency

```{r participants_proficiency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_valid %>% 
    select(participant, group, catalan_oral, spanish_oral, catalan_written, spanish_written) %>% 
    mutate_at(vars(catalan_oral, spanish_oral, catalan_written, spanish_written), replace_na, 0) %>% 
    pivot_longer(c(catalan_oral, spanish_oral, catalan_written, spanish_written),
                 names_to = "type", values_to = "score") %>% 
    separate(type, c("language", "modality"), sep = "_") %>% 
    mutate_at(vars(language, modality), str_to_sentence) %>%
    ggplot(aes(as.integer(score), fill = language)) +
    facet_grid(modality~group) +
    geom_bar() +
    labs(y = "# participants", x = "Self-reportd proficiency (0-5)", fill = "Language") +
    scale_x_continuous(limits = c(0, 5)) +
    coord_flip() +
    theme_custom() +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

#### {-}

### Stimuli {.tabset .tab-pill}

#### Lexical frequency

```{r trials_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials %>% 
    ggplot(aes(zipf, fill = test_language)) +
    geom_histogram(bins = 25) +
    labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
    theme_custom() +
    theme(
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

#### PTHN

```{r trials_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

trials %>% 
    count(test_language, pthn) %>% 
    ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
    geom_col() +
    labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
    theme_custom() +
    theme(
        legend.position = "top",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

#### Vowel and consonant similarity

```{r trials_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials %>% 
    select(trial_id, group, ends_with("_ratio")) %>% 
    pivot_longer(ends_with("_ratio"), names_to = "type", values_to = "ratio") %>% 
    mutate(type = str_to_sentence(str_remove(type, "_ratio"))) %>% 
    ggplot(aes(group, ratio, fill = group, colour = group)) +
    facet_wrap(~type) +
    geom_violin() +
    geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
    geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
    labs(x = "Grouop", y = "Similarity", fill = "Test language") +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
    theme_custom() +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

#### Phonological onset

```{r trials_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials %>% 
    select(trial_id, group, onset) %>% 
    ggplot(aes(group, fill = onset, colour = onset)) +
    geom_bar(position = position_dodge(width = 0.9)) +
    labs(x = "Group", y = "# trials", fill = "Phonological onset", colour = "Phonological onset") +
    theme_custom() +
    theme(
        legend.position = "top",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )
```

#### Stress position

```{r trials_stress, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
trials %>% 
    select(trial_id, group, overlap_stress) %>% 
    ggplot(aes(group, fill = overlap_stress, colour = overlap_stress)) +
    geom_bar(position = position_dodge(width = 0.9)) +
    labs(x = "Group", y = "# trials", fill = "Same stress position?", colour = "Same stress position?") +
    theme_custom() +
    theme(
        legend.position = "top",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )
```

#### {-}

### Design

### Data analysis

## Results {.tabset .tab-pill}

### Posterior samples: Fixed coefficients

```{r coefs_fixed_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
summary(fit)$fixed %>% 
    as.data.frame() %>%
    rownames_to_column("term") %>% 
    mutate(
        Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
        `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
        `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
        Est.Error = ifelse(term=="Intercep)", inv_logit_scaled(Est.Error), Est.Error/4)
    ) %>% 
    gt() %>% 
    fmt_percent(2:5) %>% 
    fmt_number(6:6, decimals = 2) %>% 
    fmt_number(7:8, decimals = 0) %>% 
    cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
    cols_label(
        term = md("**Predictor**"),
        Estimate = md("**Mean**"),
        Est.Error = md("**SEM**"),
        "l-95% CI" = md("**95\\% CrI**"),
        Rhat = md("**Rhat**"),
        Bulk_ESS = md("**Bulk ESS**"),
        Tail_ESS = md("**Tail ESS**")
    ) %>% 
    tab_footnote(
        footnote = "Transformed using the inverse logit to get the average probability of correct response",
        locations = cells_body(columns = "term", rows = term=="(Intercept)")
    ) %>% 
    tab_footnote(
        footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
        locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "consonant_ratio", "vowel_ratio", "onset1", "overlap_stress1"))
    ) %>% 
    tab_footnote(
        footnote = "ESS: Effective sample size",
        locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
    ) %>% 
    cols_align(
        align = c("center"),
        columns = 2:4
    )
```

```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
post <- gather_draws(fit, `b_.*`, regex = TRUE)

ggplot(post, aes(.value, fill = .variable)) +
    facet_wrap(~.variable, scales = "free") +
    stat_histinterval() +
    labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
    theme_custom() +
    theme(
        legend.position = "none",
        axis.text.x = element_blank(),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )
```

### Random effects

```{r post_re, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
post_re <- gather_draws(fit, r_participant[participant, .param]) %>% 
    mean_qi() %>% 
    ungroup() %>% 
    left_join(select(participants, participant, group)) %>% 
    arrange(group, participant, .param)

ggplot(post_re, aes(.value, sort(as.factor(participant)), xmin = .lower, xmax = .upper, colour = group)) +
    facet_wrap(~.param) +
    geom_errorbarh(height = 0, alpha = 0.75) +
    geom_point(size = 0.1, shape = 4) +
    labs(x = "Estimate", y = "Participant", colour = "Group") +
    theme_custom() +
    theme(
        legend.position = c(0.65, 0.15),
        legend.direction = "horizontal",
        axis.title.x = element_blank(),
        legend.title = element_blank(),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()
    )
```

### Marginal effects

```{r emmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
nd <- expand_grid(
    pthn = seq(min(d$pthn, na.rm = TRUE), max(d$pthn, na.rm = TRUE), by = 0.1),
    onset = unique(d$onset),
    overlap_stress = unique(d$overlap_stress),
    vowel_ratio = c(-1, 1),
    consonant_ratio = c(-1, 1)
) %>% 
    mutate(frequency = 0)
m <- add_fitted_draws(nd, fit, n = 20, re_formula = NA) %>% 
    mutate_at(vars(ends_with("_ratio")),
              function(x) ifelse(x>0, paste0(x, " SD"), paste0("-", x, " SD")))

ggplot(m, aes(pthn, .value, 
              colour = interaction(consonant_ratio, vowel_ratio, sep = "/"),
              fill = interaction(consonant_ratio, vowel_ratio, sep = "/"))) +
    facet_grid(~overlap_stress~onset) +
    stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
    stat_summary(fun = mean, geom = "line", size = 1) +
    labs(x = "Phonological neighbourhood density\n(higher frequency neighbours)",
         y = "P(Correct)", colour = "Consonant similarity / Vowel similarity",
         fill = "Consonant similarity / Vowel similarity") +
    theme_ggdist() +
    theme(
        legend.position = "top"
    )
```

### {-}

# TRACE simulations  {.tabset .tab-pill}

```{r jtrace_import, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
filedir <- here("Data/jtrace")
# directory with save files (
## 1 file per input word, with 1 output word per column
## files should be named with input word orthography and language (e.g. "gato_spanish.xlsx")
## files should be xlsx (csv doesn't deal well with special characters like ^ and @)
## folder should not contain any other files

# Note: it's effortful to have to convert to xlsx (jTRACE doesn't allow direct xlsx export), but it's the most reliable method to preserve special characters I've found so far. Would welcome ideas

# Export jTRACE output with Top 100 items and 100 cycles (arbritary value)
# Max ad-hoc activation

fnames <- dir(filedir) # extract list of file names in directory

trace_output <- data.frame() # create empty data frame
# for loop to extract highest activations for each input word
for (i in 1:length(fnames)){
    filename <- paste(filedir, fnames[i], sep = "/") # get file name
    temp <- read_excel(filename) # read file

    temp <- apply(temp,2,max) # get highest value in each column
    temp <- as.data.frame(temp)
    temp <- cbind(newColName = rownames(temp), temp) # change index column (which has output items) into column 1 for easier organisation
    rownames(temp) <- 1:nrow(temp)
    names(temp)[1] <- "output" # (rename col) output item
    names(temp)[2] <- "activation" # (rename col) max activation value

    temp <- temp %>%
        filter(!output == "cycle") %>% # remove row corresponding to cycle number
        mutate(input = fnames[i]) %>% # specify input word
        arrange(desc(activation)) %>%
        filter(!activation <= 0) # remove low activation (maybe even apply a higher threshold? TBD)
    trace_output <- bind_rows(trace_output, temp)
}

trace_output <- trace_output %>%
    mutate(input = str_remove(input, ".xlsx")) %>%
    separate(input, c("input", "language"), "_")

# xlsx file with 1 column for TRACE transcriptions and 1 column for orthography (English words) 
df_transcription <- read_excel(here("Stimuli", "trace.xlsx")) %>%
    select(trace_eng, ort_eng) %>%
    unique()

# Match the jTRACE output transcriptions with their orthography
trace_output <- left_join(trace_output, df_transcription, by = c("output" = "trace_eng")) %>%
    rename(output_word = ort_eng) %>% 
    mutate(language = str_to_sentence(language))

# Cleaned file with participant answers, 1 row per trial per participant

# count the number of participants who gave each answer in response to a given trial
df_answer <- accuracy %>%
    rename(answer = input_text) %>% 
    group_by(test_language, word, answer) %>%
    summarise(n_participant = n(), .groups = "drop") %>%
    group_by(test_language, word) %>%
    mutate(pct_participant = n_participant/sum(n_participant)*100) #%>%
#  filter(!n_participant == 1) # remove answers only given by 1 participant as idiosyncratic answers?

```

## Predicting participant response by jTRACE

Presented words and the answer that was given by the most number of participants for that trial.

```{r jtrace_all, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

# Predicting participant response by jTRACE

df_match <- full_join(trace_output, df_answer, by = c("input" = "word", "output_word" = "answer", "language" = "test_language")) %>%
    semi_join(., trace_output, by = c("input", "language"))

# If "output" is <NA>, the word is not in the currently-used jTRACE lexicon
df_match %>%
    filter(language == "Spanish") %>% 
    arrange(input, desc(pct_participant)) %>%
    group_by(input) %>%
    slice(1) %>%
    select(input, output_word, pct_participant, output, activation) %>%
    arrange(desc(pct_participant))

df_match %>%
    filter(language == "Catalan") %>% 
    arrange(input, desc(pct_participant)) %>%
    group_by(input) %>%
    slice(1) %>%
    select(input, output_word, pct_participant, output, activation) %>%
    arrange(desc(pct_participant))

# if no data for a given output word is available in df_answer, it means no participants gave that word as a response - recode as 0
df_match$pct_participant[is.na(df_match$pct_participant)] <- 0

# if no data for a given output word is available in trace output, it means that activation is outside of Top 100 words and/or below 0
df_match$activation[is.na(df_match$activation)] <- 0

# However, this is also constrained by the size of TRACE's lexicon, so subset only the words that do appear in the lexicon
df_match_trace <- semi_join(df_match, df_transcription, by = c("output_word" = "ort_eng"))

```

```{r jtrace_all_plot, echo=FALSE, warning=FALSE}
ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
    geom_smooth() +
    geom_point(shape = 1, stroke = 1, alpha = 0.5) +
    labs(
        x = "jTRACE activation value for the word",
        y = "% participants who responded with the word",
        title = "jTRACE activation against % participants who gave word as response"
    ) +
    theme_custom() +
    theme(
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

## Item-wise analysis

```{r jtrace_word_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# plot by word
df_match_trace %>% 
    mutate(input_language = paste0(input, " (", language, ")")) %>% 
    ggplot(aes(x = activation, y = pct_participant)) +
    facet_wrap(~input_language) +
    geom_smooth() +
    geom_point(shape = 1, stroke = 1, alpha = 0.5) +
    labs(
        x = "jTRACE activation value for the word",
        y = "% participants who responded with the word",
        title = "jTRACE activation against % participants who gave word as response"
    ) +
    theme_custom() +
    theme(
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

## {-}

# References

### 2021-05-21 Rmd/report.Rmd
Add more details to the report files

@@
 library(gt)
+library(lubridate)
 library(ggdist)
@@
 # import data
-trials <- readRDS(here("Data", "trials.rds")) %>% 
-    mutate(
-        test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"),
-        zipf = log10(frequency)+3,
-        pthn_log = log(pthn),
-        overlap_stress = ifelse(overlap_stress==1, "Same", "Different")
-    )
-accuracy <- readRDS(here("Data", "accuracy.rds"))
-d <- readRDS(here("Data", "accuracy.rds")) %>% 
-    # typos are considered correct responses
-    mutate_at(vars(onset, overlap_stress, group), as.factor) %>% 
-    # center predictors
-    mutate_at(
-        vars(consonant_ratio, vowel_ratio, pthn, frequency),
-        function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
-    # impute missing data
-    mice(m = 5, print = FALSE, method = "pmm") %>% 
-    complete() %>% 
-    as_tibble() %>% 
-    arrange(trial_id, group)
+trials <- readRDS(here("Data", "trials.rds"))
+participants <- readRDS(here("Data", "participants.rds"))
+responses <- readRDS(here("Data", "accuracy.rds"))
+jtrace <- readRDS(here("Data", "jtrace.rds"))
 
-participants <- readRDS(here("Data", "participants.rds")) %>% 
-    mutate(valid_participant = ifelse(valid_participant, "Valid", "Not valid"))
-fit <- readRDS(here("Results", "fit.rds"))
-coefs <- readRDS(here("Results", "coefficients.rds")) %>% 
+# import models
+fit_jtrace <- readRDS(here("Results", "fit_jtrace.rds"))
+fit_responses <- readRDS(here("Results", "fit.rds"))
+fit_responses_coefs <- readRDS(here("Results", "coefficients.rds")) %>% 
   select(term, estimate, std.error, conf.low, conf.high)
@@
 
+```{r participants_coded}
+participants_coded <- participants %>% 
+  mutate(valid_participant = ifelse(valid_participant, "Valid", "Not valid"))
+```
+
+We collected data from `r format(as_date(min(participants$date)), "%B %dth %Y")` to `r format(as_date(max(participants$date)), "%B %dth %Y")`.
+We tested `r length(unique(participants$participant))` participants.
+`r length(unique(participants$participant[participants$group=="ENG-SPA"]))` were English natives tested in Spanish, `r length(unique(participants$participant[participants$group=="ENG-CAT"]))` were English natives tested in Catalan, and `r length(unique(participants$participant[participants$group=="SPA-CAT"]))` were Spanish natives tested in Catalan.
+`r sum(!participants$valid_participant)` failed to meet all the inclusion criteria and were excluded from further analyses.
+
+Participants were included if^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct.
+This item was phrased as Do you have normal or corrected-to-normal VISION?
+(Yes/No) in English, and as ¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
+(Sí/No).
+Surprisingly, the proportion of Spanish participants that reported visual impairment was implausibly large (n = 6, 18.18%).
+This is possibly due to some participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision]:
+
+* Aged 18 to 25 years
+* Did not report being fluent in Catalan, Spanish, or Italian
+* Provided at least 80% of the presented trials
+* Did not report motor, auditory or visual (other than wearing glasses) problems
+
+Trials were considered valid if:
+
+* Participant did not leave a blank response
+* Participant took longer than 10 seconds to respond
+
 #### Summary
@@
 
-ggplot(participants, aes(group, fill = valid_participant)) +
+ggplot(participants_coded, aes(group, fill = valid_participant)) +
   geom_bar() +
@@
 
+#### Valid trials
+
+```{r participants_trials, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+participants_coded %>% 
+  filter(valid_participant=="Valid") %>% 
+  select(participant, group, n, n_valid) %>%
+  mutate(valid_prop = n_valid/n) %>% 
+  ggplot(aes(group, valid_prop, fill = group, colour = group)) +
+  geom_violin() +
+  geom_point(shape = 1, stroke = 1, alpha = 0.5, position = position_jitter(width = 0.1),
+             colour = "black") +
+  geom_boxplot(colour = "black", fill = "white", width = 0.05, outlier.color = NA) +
+  labs(x = "Group", y = "Valid trials", fill = "Group", colour = "Group") +
+  scale_y_continuous(limits = c(0.8, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top",
+    axis.title.x = element_blank(),
+    panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
+  ) 
+```
+
 #### Demographic information
@@
 ```{r participants_demo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-participants_valid <- filter(participants, valid_participant=="Valid")
+participants_valid <- filter(participants_coded, valid_participant=="Valid")
 p_sex <- ggplot(participants_valid, aes(group, fill = sex)) +
@@
 
+```{r trials_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
+trials_coded <- trials %>% 
+  mutate(
+    test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"),
+    zipf = log10(frequency)+3,
+    pthn_log = log(pthn),
+    overlap_stress = ifelse(overlap_stress==1, "Same", "Different")
+  )
+```
+
 #### Lexical frequency
@@
 ```{r trials_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials %>% 
+trials_coded %>% 
   ggplot(aes(zipf, fill = test_language)) +
@@
 ```{r trials_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-trials %>% 
+trials_coded %>% 
   count(test_language, pthn) %>% 
   ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
-    geom_col() +
+  geom_col(position = position_dodge()) +
   labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
@@
 ```{r trials_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials %>% 
+trials_coded %>% 
   select(trial_id, group, ends_with("_ratio")) %>% 
@@
 ```{r trials_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials %>% 
+trials_coded %>% 
   select(trial_id, group, onset) %>% 
@@
 ```{r trials_stress, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials %>% 
+trials_coded %>% 
   select(trial_id, group, overlap_stress) %>% 
@@
 
-### Posterior samples: Fixed coefficients
+### Accuracy by item
+
+```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_item <- responses %>%
+  group_by(group, word) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    word_label = reorder_within(word, by = correct_prop, within = group)
+  )
+
+ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5, colour = "red") +
+  geom_errorbar() +
+  geom_point(size = 0.5) +
+  labs(y = "Correct responses", x = "Word", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  )
+
+```
+
+### Accuracy by participant
+
+```{r responses_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_participant <- responses %>%
+  group_by(group, participant) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    participant_label = reorder_within(participant, by = correct_prop, within = group)
+  )
+
+ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5, colour = "red") +
+  geom_errorbar(width = 0.5) +
+  geom_point(size = 1) +
+  labs(y = "Correct responses", x = "Participant", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "top",
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  )
+
+```
+
+### Bayesian regression model {.tabset .tab-pill}
+
+#### Model details
+
+We fitted a Bayesian multilevel regression model using the R package `brms` [@burkner2017brms], with correct responses (`correct`, 0 is incorrect response, 1 if correct response) as the response variable with a Bernoulli distribution and a logit link function.
+We modelled the probability of an average participant providing a correct response, conditional to a set of predictors (aka.
+fixed effects):
+
+* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database.
+* Lexical frequency (`frequency`, 0-Inf): lexical frequency (counts per million) of the target word. Extracted from the SUBTLEX database (via CLEAPOND).
+* Vowel similarity (`vowel_ratio`, 0-1): edit distance between the sequence of vowels of the presented word  and the target word. Computed using custom code.
+* Consonant similarity (`consonant_ratio`, 0-1): edit distance between the sequence of consonants of the presented word and the target word. Computed using custom code.
+* Shared phonological onset (`onset`, 0 or 1): do the presented word and the target word share phonological onset? Manually coded.
+* Share stress position (`overlap_stress`, 0 or 1): are the presented word and the target word stressed at the same position? Manually coded.
+
+We also added participants (`participants`) and words (`word`) as grouping variables (aka., random effects), therefore adjusting the model to each participant and word.
+We specified random intercepts for both grouping variables, and random slopes for the fixed effects within the `participant` grouping variable.
+
+We model was implemented using the following formula:
+
+```
+correct ~ pthn + frequency + consonant_ratio + vowel_ratio + onset + overlap_stress + (1 + pthn + frequency + consonant_ratio + vowel_ratio + onset + overlap_stress | participant) + (1 | trial_id)
+```
+
+The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 intertions each.
+The output of this model is the approximated posterior distribution of all the parameters estimated, therefore indicating the probability of each combination of values of the parameters, conditional to the observed data.
+
+#### Stan code
+
+Stan code generated by `brms`:
+
+```{stan stan_model, echo=TRUE, output.var="model"}
+// generated with brms 2.15.0
+functions {
+ /* compute correlated group-level effects
+  * Args: 
+  *   z: matrix of unscaled group-level effects
+  *   SD: vector of standard deviation parameters
+  *   L: cholesky factor correlation matrix
+  * Returns: 
+  *   matrix of scaled group-level effects
+  */ 
+  matrix scale_r_cor(matrix z, vector SD, matrix L) {
+    // r is stored in another dimension order than z
+    return transpose(diag_pre_multiply(SD, L) * z);
+  }
+}
+data {
+  int<lower=1> N;  // total number of observations
+  int Y[N];  // response variable
+  int<lower=1> K;  // number of population-level effects
+  matrix[N, K] X;  // population-level design matrix
+  // data for group-level effects of ID 1
+  int<lower=1> N_1;  // number of grouping levels
+  int<lower=1> M_1;  // number of coefficients per level
+  int<lower=1> J_1[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_1_1;
+  vector[N] Z_1_2;
+  vector[N] Z_1_3;
+  vector[N] Z_1_4;
+  vector[N] Z_1_5;
+  vector[N] Z_1_6;
+  vector[N] Z_1_7;
+  int<lower=1> NC_1;  // number of group-level correlations
+  // data for group-level effects of ID 2
+  int<lower=1> N_2;  // number of grouping levels
+  int<lower=1> M_2;  // number of coefficients per level
+  int<lower=1> J_2[N];  // grouping indicator per observation
+  // group-level predictor values
+  vector[N] Z_2_1;
+  int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+  int Kc = K - 1;
+  matrix[N, Kc] Xc;  // centered version of X without an intercept
+  vector[Kc] means_X;  // column means of X before centering
+  for (i in 2:K) {
+    means_X[i - 1] = mean(X[, i]);
+    Xc[, i - 1] = X[, i] - means_X[i - 1];
+  }
+}
+parameters {
+  vector[Kc] b;  // population-level effects
+  real Intercept;  // temporary intercept for centered predictors
+  matrix[M_1, N_1] z_1;  // standardized group-level effects
+  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
+  vector[N_2] z_2[M_2];  // standardized group-level effects
+}
+transformed parameters {
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+  matrix[N_1, M_1] r_1;  // actual group-level effects
+  // using vectors speeds up indexing in loops
+  vector[N_1] r_1_1;
+  vector[N_1] r_1_2;
+  vector[N_1] r_1_3;
+  vector[N_1] r_1_4;
+  vector[N_1] r_1_5;
+  vector[N_1] r_1_6;
+  vector[N_1] r_1_7;
+  vector[N_2] r_2_1;  // actual group-level effects
+  sd_1 = rep_vector(1, rows(sd_1));
+  // compute actual group-level effects
+  r_1 = scale_r_cor(z_1, sd_1, L_1);
+  r_1_1 = r_1[, 1];
+  r_1_2 = r_1[, 2];
+  r_1_3 = r_1[, 3];
+  r_1_4 = r_1[, 4];
+  r_1_5 = r_1[, 5];
+  r_1_6 = r_1[, 6];
+  r_1_7 = r_1[, 7];
+  r_2_1 = (sd_2[1] * (z_2[1]));
+}
+model {
+  // likelihood including constants
+  if (!prior_only) {
+    // initialize linear predictor term
+    vector[N] mu = Intercept + rep_vector(0.0, N);
+    for (n in 1:N) {
+      // add more terms to the linear predictor
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n];
+    }
+    target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
+  }
+  // priors including constants
+  target += normal_lpdf(b | 0, 1);
+  target += normal_lpdf(Intercept | 0, 1);
+  target += std_normal_lpdf(to_vector(z_1));
+  target += lkj_corr_cholesky_lpdf(L_1 | 5);
+  target += cauchy_lpdf(sd_2 | 0, 3)
+    - 1 * cauchy_lccdf(0 | 0, 3);
+  target += std_normal_lpdf(z_2[1]);
+}
+generated quantities {
+  // actual population-level intercept
+  real b_Intercept = Intercept - dot_product(means_X, b);
+  // compute group-level correlations
+  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+  vector<lower=-1,upper=1>[NC_1] cor_1;
+  // extract upper diagonal of correlation matrix
+  for (k in 1:M_1) {
+    for (j in 1:(k - 1)) {
+      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+    }
+  }
+}
+```
+
+```{r responses_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+responses_coded <- responses %>% 
+  # typos are considered correct responses
+  mutate_at(vars(onset, overlap_stress, group), as.factor) %>% 
+  # center predictors
+  mutate_at(
+    vars(consonant_ratio, vowel_ratio, pthn, frequency),
+    function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
+  # impute missing data
+  mice(m = 5, print = FALSE, method = "pmm") %>% 
+  complete() %>% 
+  as_tibble() %>% 
+  arrange(trial_id, group)
+```
+
+#### Fixed coefficients
 
 ```{r coefs_fixed_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-summary(fit)$fixed %>% 
+summary(fit_responses)$fixed %>% 
   as.data.frame() %>%
@@
 ```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-post <- gather_draws(fit, `b_.*`, regex = TRUE)
+post <- gather_draws(fit_responses, `b_.*`, regex = TRUE) %>% 
+  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))
 
@@
   facet_wrap(~.variable, scales = "free") +
-    stat_histinterval() +
+  geom_vline(xintercept = 0, colour = "red") +
+  stat_halfeye() +
   labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
@@
     legend.position = "none",
-        axis.text.x = element_blank(),
+    axis.text = element_text(size = 7),
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
@@
 
-### Random effects
+#### Random effects
 
 ```{r post_re, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-post_re <- gather_draws(fit, r_participant[participant, .param]) %>% 
+post_re <- gather_draws(fit_responses, r_participant[participant, .param]) %>% 
   mean_qi() %>% 
@@
 ggplot(post_re, aes(.value, sort(as.factor(participant)), xmin = .lower, xmax = .upper, colour = group)) +
-    facet_wrap(~.param) +
-    geom_errorbarh(height = 0, alpha = 0.75) +
+  facet_grid(group~.param, scales = "free_y") +
+  geom_errorbarh(height = 0.5, alpha = 0.75) +
   geom_point(size = 0.1, shape = 4) +
-    labs(x = "Estimate", y = "Participant", colour = "Group") +
+  geom_vline(xintercept = 0, colour = "red") +
+  labs(x = "Participant-level adjustment", y = "Participant", colour = "Group") +
   theme_custom() +
   theme(
-        legend.position = c(0.65, 0.15),
-        legend.direction = "horizontal",
-        axis.title.x = element_blank(),
+    legend.position = "top",
     legend.title = element_blank(),
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    axis.text.x = element_text(size = 9),
     axis.text.y = element_blank(),
@@
 
-### Marginal effects
+#### Marginal effects
 
@@
 nd <- expand_grid(
-    pthn = seq(min(d$pthn, na.rm = TRUE), max(d$pthn, na.rm = TRUE), by = 0.1),
-    onset = unique(d$onset),
-    overlap_stress = unique(d$overlap_stress),
+  pthn = seq(
+    min(responses_coded$pthn, na.rm = TRUE),
+    max(responses_coded$pthn, na.rm = TRUE),
+    by = 0.1
+  ),
+  onset = unique(responses_coded$onset),
+  overlap_stress = unique(responses_coded$overlap_stress),
   vowel_ratio = c(-1, 1),
@@
   mutate(frequency = 0)
-m <- add_fitted_draws(nd, fit, n = 20, re_formula = NA) %>% 
+
+m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
   mutate_at(vars(ends_with("_ratio")),
@@
   labs(x = "Phonological neighbourhood density\n(higher frequency neighbours)",
-         y = "P(Correct)", colour = "Consonant similarity / Vowel similarity",
+       y = "P(Correct)", colour = "Consonant/Vowel similarity",
        fill = "Consonant similarity / Vowel similarity") +
-    theme_ggdist() +
+  theme_custom() +
   theme(
-        legend.position = "top"
+    legend.position = "top",
+    panel.grid = element_line(colour = "grey", linetype = "dotted")
   )
@@
 
+#### {-}
+
 ### {-}
 
-# TRACE simulations  {.tabset .tab-pill}
+# TRACE simulations
 
-```{r jtrace_import, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-filedir <- here("Data/jtrace")
-# directory with save files (
-## 1 file per input word, with 1 output word per column
-## files should be named with input word orthography and language (e.g. "gato_spanish.xlsx")
-## files should be xlsx (csv doesn't deal well with special characters like ^ and @)
-## folder should not contain any other files
+## TRACE vs. human data {.tabset .tab-pill}
 
-# Note: it's effortful to have to convert to xlsx (jTRACE doesn't allow direct xlsx export), but it's the most reliable method to preserve special characters I've found so far. Would welcome ideas
-
-# Export jTRACE output with Top 100 items and 100 cycles (arbritary value)
-# Max ad-hoc activation
-
-fnames <- dir(filedir) # extract list of file names in directory
-
-trace_output <- data.frame() # create empty data frame
-# for loop to extract highest activations for each input word
-for (i in 1:length(fnames)){
-    filename <- paste(filedir, fnames[i], sep = "/") # get file name
-    temp <- read_excel(filename) # read file
-
-    temp <- apply(temp,2,max) # get highest value in each column
-    temp <- as.data.frame(temp)
-    temp <- cbind(newColName = rownames(temp), temp) # change index column (which has output items) into column 1 for easier organisation
-    rownames(temp) <- 1:nrow(temp)
-    names(temp)[1] <- "output" # (rename col) output item
-    names(temp)[2] <- "activation" # (rename col) max activation value
-
-    temp <- temp %>%
-        filter(!output == "cycle") %>% # remove row corresponding to cycle number
-        mutate(input = fnames[i]) %>% # specify input word
-        arrange(desc(activation)) %>%
-        filter(!activation <= 0) # remove low activation (maybe even apply a higher threshold? TBD)
-    trace_output <- bind_rows(trace_output, temp)
-}
-
-trace_output <- trace_output %>%
-    mutate(input = str_remove(input, ".xlsx")) %>%
-    separate(input, c("input", "language"), "_")
-
-# xlsx file with 1 column for TRACE transcriptions and 1 column for orthography (English words) 
-df_transcription <- read_excel(here("Stimuli", "trace.xlsx")) %>%
-    select(trace_eng, ort_eng) %>%
-    unique()
-
-# Match the jTRACE output transcriptions with their orthography
-trace_output <- left_join(trace_output, df_transcription, by = c("output" = "trace_eng")) %>%
-    rename(output_word = ort_eng) %>% 
-    mutate(language = str_to_sentence(language))
-
-# Cleaned file with participant answers, 1 row per trial per participant
-
-# count the number of participants who gave each answer in response to a given trial
-df_answer <- accuracy %>%
-    rename(answer = input_text) %>% 
-    group_by(test_language, word, answer) %>%
-    summarise(n_participant = n(), .groups = "drop") %>%
-    group_by(test_language, word) %>%
-    mutate(pct_participant = n_participant/sum(n_participant)*100) #%>%
-#  filter(!n_participant == 1) # remove answers only given by 1 participant as idiosyncratic answers?
-
-```
-
-## Predicting participant response by jTRACE
-
-Presented words and the answer that was given by the most number of participants for that trial.
-
-```{r jtrace_all, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
-# Predicting participant response by jTRACE
-
-df_match <- full_join(trace_output, df_answer, by = c("input" = "word", "output_word" = "answer", "language" = "test_language")) %>%
-    semi_join(., trace_output, by = c("input", "language"))
-
-# If "output" is <NA>, the word is not in the currently-used jTRACE lexicon
-df_match %>%
-    filter(language == "Spanish") %>% 
-    arrange(input, desc(pct_participant)) %>%
-    group_by(input) %>%
-    slice(1) %>%
-    select(input, output_word, pct_participant, output, activation) %>%
-    arrange(desc(pct_participant))
-
-df_match %>%
-    filter(language == "Catalan") %>% 
-    arrange(input, desc(pct_participant)) %>%
-    group_by(input) %>%
-    slice(1) %>%
-    select(input, output_word, pct_participant, output, activation) %>%
-    arrange(desc(pct_participant))
-
-# if no data for a given output word is available in df_answer, it means no participants gave that word as a response - recode as 0
-df_match$pct_participant[is.na(df_match$pct_participant)] <- 0
-
-# if no data for a given output word is available in trace output, it means that activation is outside of Top 100 words and/or below 0
-df_match$activation[is.na(df_match$activation)] <- 0
-
-# However, this is also constrained by the size of TRACE's lexicon, so subset only the words that do appear in the lexicon
-df_match_trace <- semi_join(df_match, df_transcription, by = c("output_word" = "ort_eng"))
-
-```
+### TRACE activation vs. responses
 
 ```{r jtrace_all_plot, echo=FALSE, warning=FALSE}
-ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
+ggplot(jtrace, aes(x = activation, y = pct_participant)) +
   geom_smooth() +
@@
 
-## Item-wise analysis
+### By items
 
@@
 # plot by word
-df_match_trace %>% 
+jtrace %>% 
   mutate(input_language = paste0(input, " (", language, ")")) %>% 
@@
 
+### {-}
+
+## Regression model {.tabset .tab-pill}
+
+### Fixed coefficients
+
+```{r jtrace_fixed_table, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+summary(fit_jtrace$fit)$fixed %>% 
+  as.data.frame() %>%
+  rownames_to_column("term") %>% 
+  mutate(
+    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
+    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
+    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
+    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
+  ) %>% 
+  gt() %>% 
+  fmt_number(2:6, decimals = 2) %>% 
+  fmt_number(7:8, decimals = 0) %>% 
+  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
+  cols_label(
+    term = md("**Predictor**"),
+    Estimate = md("**Mean**"),
+    Est.Error = md("**SEM**"),
+    "l-95% CI" = md("**95\\% CrI**"),
+    Rhat = md("**Rhat**"),
+    Bulk_ESS = md("**Bulk ESS**"),
+    Tail_ESS = md("**Tail ESS**")
+  ) %>% 
+  tab_footnote(
+    footnote = "Transformed using the inverse logit to get the average probability of correct response",
+    locations = cells_body(columns = "term", rows = term=="(Intercept)")
+  ) %>% 
+  tab_footnote(
+    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
+    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "consonant_ratio", "vowel_ratio", "onset1", "overlap_stress1") | str_detect(term, "activation"))
+  ) %>% 
+  tab_footnote(
+    footnote = "ESS: Effective sample size",
+    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
+  )
+```
+
+```{r jtrace_fixed_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%"}
+fit_jtrace$fixed %>%
+  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
+  ggplot(aes(.value, fill = .variable)) +
+  facet_wrap(~.variable, scales = "free") +
+  geom_vline(xintercept = 0, colour = "red") +
+  stat_halfeye() +
+  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
+  theme_custom() +
+  theme(
+    legend.position = "none",
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
+    axis.text = element_text(size = 7)
+  )
+```
+
+### Random coefficients
+
+```{r jtrace_post_re, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+re <- ranef(fit_jtrace$fit)$`group:input`
+re <- rbind(Intercept = re[,,1], activation = re[,,2]) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("term") %>% 
+  mutate(
+    term = str_replace(str_remove(term, ".1"), "\\.", "\\-"),
+    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
+    Q2.5 = ifelse(term=="Intercept", inv_logit_scaled(Q2.5), Q2.5/4),
+    Q97.5 = ifelse(term=="Intercept", inv_logit_scaled(Q97.5), Q97.5/4),
+    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
+  ) %>% 
+  separate(term, c("group", "input"), sep = "_")
+re$.param <- rep(c("Intercept", "activation"), each = 9)
+
+ggplot(re, aes(Estimate, input, colour = input)) +
+  facet_wrap(~.param) +
+  geom_vline(xintercept = 0, colour = "red") +
+  geom_errorbarh(aes(xmin = Estimate-Est.Error, xmax = Estimate+Est.Error), height = 0, size = 5, alpha = 0.5) +
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), height = 0) +
+  geom_point() +
+  labs(x = "Estimate", y = "Input", colour = "Input") +
+  theme_custom() +
+  theme(
+    legend.position = "top",
+    legend.direction = "horizontal",
+    axis.title.x = element_blank(),
+    legend.title = element_blank(),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank()
+  )
+```
+
+### {-}
+
+## Candidate pool
+
+```{r jtrace_candidates, echo=FALSE, fig.height=30, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+trials_trace <- trials %>% 
+  as_tibble() %>% 
+  select(group, input = word1, target = word2, phon_trace)
+
+# process TRACE data ----
+trace <- map(
+  list.files(here("Data", "jtrace"), full.names = TRUE),
+  function(x){
+    read_xlsx(x) %>% 
+      pivot_longer(-cycle, names_to = "response", values_to = "activation")
+  }
+) %>%
+  set_names(list.files(here("Data", "jtrace"))) %>% 
+  bind_rows(.id = "target") %>% 
+  separate(target, c("input", "language"), sep = "_") %>% 
+  mutate(language = str_remove(language, ".xlsx"),
+         language = str_to_sentence(language))
+
+# get top candidates
+trace_top <- trace %>% 
+  group_by(input, language, response) %>% 
+  summarise(
+    activation = max(activation, na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  group_by(input, language) %>% 
+  slice_max(activation, n = 10) %>% 
+  ungroup() %>% 
+  mutate(group = ifelse(language=="Catalan", "ENG-CAT", "ENG-SPA")) %>% 
+  select(-language) %>% 
+  # add experimental trials
+  left_join(trials_trace) %>% 
+  mutate(correct = response==phon_trace) %>% 
+  select(group, input, response, activation, correct)
+
+# experimental results ----
+responses_jtrace <- responses %>% 
+  filter(valid_response, group %in% c("ENG-CAT", "ENG-SPA")) %>% 
+  rename(response = input_text, input = word) %>% 
+  count(group, input, response, correct) %>% 
+  group_by(group, input) %>%
+  mutate(total = sum(n), activation = n/total) %>% 
+  ungroup() %>% 
+  group_by(group, input) %>% 
+  slice_max(order_by = activation, n = 10) %>% 
+  ungroup() %>% 
+  select(group, input, response, correct, activation) %>% 
+  filter(input %in% trace_top$input)
+
+# merge everything
+d_jtrace <- list(TRACE = trace_top, Participants = responses_jtrace) %>% 
+  bind_rows(.id = "source") %>% 
+  mutate(
+    # to reorder variable within panels
+    input_label = paste0(input, " (", group, ")"),
+    response_reordered = reorder_within(response, activation, input_label)
+  )
+
+ggplot(d_jtrace, aes(response_reordered, activation, fill = correct)) +
+  facet_wrap(input_label~source, scales = "free_y", ncol = 2) +
+  geom_col(colour = "white") +
+  labs(x = "Top 10 candidates", y = "Max. TRACE activation") +
+  coord_flip() +
+  scale_x_reordered() +
+  # scale_y_continuous(limits = c(0, 1)) +
+  theme_bw() +
+  theme(
+    legend.position = "none"
+  )
+
+```
+
 ## {-}

### 2021-05-21 R/trace.R

@@ -1,0 +1,229 @@
+# jtrace analyses
+
+# set up ----
+library(tidyverse)
+library(brms)
+library(readxl)
+library(tidytext)
+library(janitor)
+library(patchwork)
+library(tidybayes)
+
+# set parameters
+set.seed(888) # for reproducibility
+options(mc.cores = 4, chains = 4, iter = 500, seed = 888)
+
+# import data ----
+trials <- readRDS(here("Data", "trials.rds")) %>% 
+    mutate(
+        test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"),
+        zipf = log10(frequency)+3,
+        pthn_log = log(pthn),
+        overlap_stress = ifelse(overlap_stress==1, "Same", "Different")
+    )
+accuracy <- readRDS(here("Data", "accuracy.rds"))
+
+
+# by Serene (some fixes by Gonzalo) ----
+
+filedir <- here("Data", "jtrace")
+# directory with save files (
+## 1 file per input word, with 1 output word per column
+## files should be named with input word orthography and language (e.g. "gato_spanish.xlsx")
+## files should be xlsx (csv doesn't deal well with special characters like ^ and @)
+## folder should not contain any other files
+
+# Note: it's effortful to have to convert to xlsx (jTRACE doesn't allow direct xlsx export), but it's the most reliable method to preserve special characters I've found so far. Would welcome ideas
+
+# Export jTRACE output with Top 100 items and 100 cycles (arbritary value)
+# Max ad-hoc activation
+
+fnames <- dir(filedir) # extract list of file names in directory
+
+trace_output <- data.frame() # create empty data frame
+# for loop to extract highest activations for each input word
+for (i in 1:length(fnames)){
+    filename <- paste(filedir, fnames[i], sep = "/") # get file name
+    temp <- read_excel(filename) # read file
+    
+    temp <- apply(temp,2,max) # get highest value in each column
+    temp <- as.data.frame(temp)
+    temp <- cbind(newColName = rownames(temp), temp) # change index column (which has output items) into column 1 for easier organisation
+    rownames(temp) <- 1:nrow(temp)
+    names(temp)[1] <- "output" # (rename col) output item
+    names(temp)[2] <- "activation" # (rename col) max activation value
+    
+    temp <- temp %>%
[…]

### 2021-05-21 R/utils.R

@@ -262,4 +262,25 @@ import_trials <- function(path = "Stimuli/trials.xlsx", subtlex = NULL){
     return(trials)
 }
 
+# proportion adjusted from boundary values (Gelman, Hill & Vehtari, 2020)
+prop_adj <- function(x, n){
+    e <- (x+2)/(n+4)
+    return(e)
+}
 
+# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+prop_adj_se <- function(x, n) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    return(se)
+}
+
+# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+prop_adj_ci <- function(x, n, .width = 0.95) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    ci <-  e + qnorm(c((1-.width)/2, (1-(1-.width)/2)))*se
+    ci[1] <- ifelse(ci[1]<0, 0, ci[1]) # truncate at 0
+    ci[2] <- ifelse(ci[2]>1, 1, ci[2]) # truncate at 1
+    return(ci)
+}

### 2021-05-21 Rmd/report.Rmd
Fix model

@@
 )
-options(knitr.kable.NA = '-')
+options(knitr.kable.NA = '-',
+        ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
+        ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
+        ggplot2.continuous.fill = wesanderson::wes_palettes$Darjeeling1,
+        ggplot2.continuous.colour = wesanderson::wes_palettes$Darjeeling1
+)
+
 ```
@@
 library(lubridate)
+library(wesanderson)
 library(ggdist)
@@
 participants <- readRDS(here("Data", "participants.rds"))
-responses <- readRDS(here("Data", "accuracy.rds"))
+responses <- readRDS(here("Data", "responses.rds"))
 jtrace <- readRDS(here("Data", "jtrace.rds"))
@@
 fit_jtrace <- readRDS(here("Results", "fit_jtrace.rds"))
-fit_responses <- readRDS(here("Results", "fit.rds"))
-fit_responses_coefs <- readRDS(here("Results", "coefficients.rds")) %>% 
-  select(term, estimate, std.error, conf.low, conf.high)
+fit_responses <- readRDS(here("Results", "fit_responses.rds"))
 
@@
 
-# Experimental task: Translation Elicitation
+# Experimental task
 
@@
 `r sum(!participants$valid_participant)` failed to meet all the inclusion criteria and were excluded from further analyses.
+The final sample comprised data from `r sum(!participants$valid_participant)` participants, `r sum(participants$valid_participant[participants$group=="ENG-SPA"])` in the ENG-SPA group, `r sum(participants$valid_participant[participants$group=="ENG-CAT"])` in the ENG-CAT group, and `r sum(participants$valid_participant[participants$group=="SPA-CAT"])` in the SPA-CAT group.
 
@@
 
+Participants in the ENG-SPA listened to `r nrow(trials[trials$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `nrow(trials[trials$group=="ENG-CAT",]) trials`.
+Five trials in each condition were practice trials and were excluded from analyses.
+
 ```{r trials_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
@@
 
-#### Stress position
-
-```{r trials_stress, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials_coded %>% 
-  select(trial_id, group, overlap_stress) %>% 
-  ggplot(aes(group, fill = overlap_stress, colour = overlap_stress)) +
-  geom_bar(position = position_dodge(width = 0.9)) +
-  labs(x = "Group", y = "# trials", fill = "Same stress position?", colour = "Same stress position?") +
-  theme_custom() +
-  theme(
-    legend.position = "top",
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  )
-```
-
 #### {-}
@@
 
-### Data analysis
-
-## Results {.tabset .tab-pill}
-
-### Accuracy by item
-
-```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses_item <- responses %>%
-  group_by(group, word) %>%
-  summarise(
-    correct_sum = sum(correct, na.rm = TRUE),
-    correct_n = sum(!is.na(correct), na.rm = TRUE),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    correct_prop = prop_adj(correct_sum, correct_n),
-    correct_se = prop_adj_se(correct_sum, correct_n),
-    word_label = reorder_within(word, by = correct_prop, within = group)
-  )
-
-ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
-  facet_wrap(~group, scales = "free_y") +
-  geom_hline(yintercept = 0.5, colour = "red") +
-  geom_errorbar() +
-  geom_point(size = 0.5) +
-  labs(y = "Correct responses", x = "Word", colour = "Group") +
-  coord_flip() +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme_custom() +
-  theme(
-    axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 9),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-  )
-
-```
-
-### Accuracy by participant
-
-```{r responses_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses_participant <- responses %>%
-  group_by(group, participant) %>%
-  summarise(
-    correct_sum = sum(correct, na.rm = TRUE),
-    correct_n = sum(!is.na(correct), na.rm = TRUE),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    correct_prop = prop_adj(correct_sum, correct_n),
-    correct_se = prop_adj_se(correct_sum, correct_n),
-    participant_label = reorder_within(participant, by = correct_prop, within = group)
-  )
-
-ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
-  facet_wrap(~group, scales = "free_y") +
-  geom_hline(yintercept = 0.5, colour = "red") +
-  geom_errorbar(width = 0.5) +
-  geom_point(size = 1) +
-  labs(y = "Correct responses", x = "Participant", colour = "Group") +
-  coord_flip() +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme_custom() +
-  theme(
-    legend.position = "top",
-    axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 9),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-  )
-
-```
-
-### Bayesian regression model {.tabset .tab-pill}
+### Data analysis {.tabset .tab-pills}
 
@@
 * PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database.
-* Lexical frequency (`frequency`, 0-Inf): lexical frequency (counts per million) of the target word. Extracted from the SUBTLEX database (via CLEAPOND).
 * Vowel similarity (`vowel_ratio`, 0-1): edit distance between the sequence of vowels of the presented word  and the target word. Computed using custom code.
@@
 * Shared phonological onset (`onset`, 0 or 1): do the presented word and the target word share phonological onset? Manually coded.
-* Share stress position (`overlap_stress`, 0 or 1): are the presented word and the target word stressed at the same position? Manually coded.
 
@@
 
+#### {-}
+
+## Results {.tabset .tab-pill}
+
+### Accuracy by item
+
+```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_item <- responses %>%
+  group_by(group, word) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    word_label = reorder_within(word, by = correct_prop, within = group)
+  )
+
+ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5, colour = "black") +
+  geom_errorbar() +
+  geom_point(size = 0.5) +
+  labs(y = "Correct responses", x = "Word", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "accuracy_item.png"))
+
+```
+
+### Accuracy by participant
+
+```{r responses_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_participant <- responses %>%
+  group_by(group, participant) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    participant_label = reorder_within(participant, by = correct_prop, within = group)
+  )
+
+ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5) +
+  geom_errorbar(width = 0.5) +
+  geom_point(size = 1) +
+  labs(y = "Correct responses", x = "Participant", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "accuracy_participant.png"))
+
+```
+
+### Model outputs {.tabset .tab-pill}
+
 ```{r responses_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
@@
     footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "consonant_ratio", "vowel_ratio", "onset1", "overlap_stress1"))
+    locations = cells_body(columns = "term", rows = term %in% c("pthn", "consonant_ratio", "vowel_ratio", "onset1"))
   ) %>% 
@@
   facet_wrap(~.variable, scales = "free") +
-  geom_vline(xintercept = 0, colour = "red") +
+  geom_vline(xintercept = 0) +
   stat_halfeye() +
@@
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  )
+  ) +
+  ggsave(here("Figures", "responses_fixed.png"))
 ```
@@
   geom_point(size = 0.1, shape = 4) +
-  geom_vline(xintercept = 0, colour = "red") +
+  geom_vline(xintercept = 0) +
   labs(x = "Participant-level adjustment", y = "Participant", colour = "Group") +
@@
 
-#### Marginal effects
+#### Marginal effects {.tabset .tab-pills}
 
-```{r emmeans, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+##### By PTHN
+
+```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 nd <- expand_grid(
@@
 m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
-  mutate_at(vars(ends_with("_ratio")),
-            function(x) ifelse(x>0, paste0(x, " SD"), paste0("-", x, " SD")))
+  mutate_at(vars(ends_with("_ratio")), function(x) paste0(x, " SD")) %>% 
+  mutate(
+    overlap_stress = paste0("Stress: ", overlap_stress),
+    onset = paste0("Onset: ", onset)
+  )
 
@@
               fill = interaction(consonant_ratio, vowel_ratio, sep = "/"))) +
-  facet_grid(~overlap_stress~onset) +
-  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
+  facet_wrap(~onset) +
+  # stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
   stat_summary(fun = mean, geom = "line", size = 1) +
@@
        y = "P(Correct)", colour = "Consonant/Vowel similarity",
-       fill = "Consonant similarity / Vowel similarity") +
+       fill = "Consonant similarity/Vowel similarity") +
+  guides(colour = guide_legend(ncol = 2)) +
   theme_custom() +
+  scale_y_continuous(labels = scales::percent) +
   theme(
@@
     panel.grid = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "responses_pthn.png"))
+
+```
+
+##### By consonant similarity
+
+```{r emmeans_consonant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+nd <- expand_grid(
+  pthn = c(-1, 1),
+  onset = unique(responses_coded$onset),
+  overlap_stress = unique(responses_coded$overlap_stress),
+  vowel_ratio = c(-1, 1),
+  consonant_ratio = seq(
+    min(responses_coded$consonant_ratio, na.rm = TRUE),
+    max(responses_coded$consonant_ratio, na.rm = TRUE),
+    by = 0.1
   )
+) %>% 
+  mutate(frequency = 0)
+
+m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
+  mutate_at(vars(vowel_ratio, pthn), function(x) paste0(x, " SD")) %>% 
+  mutate(
+    overlap_stress = paste0("Stress: ", overlap_stress),
+    onset = paste0("Onset: ", onset)
+  )
+
+ggplot(m, aes(consonant_ratio, .value, 
+              colour = interaction(pthn, vowel_ratio, sep = "/"),
+              fill = interaction(pthn, vowel_ratio, sep = "/"))) +
+  facet_wrap(~onset) +
+  # stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
+  stat_summary(fun = mean, geom = "line", size = 1) +
+  labs(x = "Consonant similarity (SD)",
+       y = "P(Correct)", colour = "PTHN/Vowel similarity",
+       fill = "PTHN/Vowel similarity") +
+  guides(colour = guide_legend(ncol = 2)) +
+  theme_custom() +
+  scale_y_continuous(labels = scales::percent) +
+  theme(
+    legend.position = "top",
+    panel.grid = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "responses_consonant.png"))
+
 ```
+
+##### By vowel similarity
+
+```{r emmeans_vowel, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+nd <- expand_grid(
+  pthn = c(-1, 1),
+  onset = unique(responses_coded$onset),
+  overlap_stress = unique(responses_coded$overlap_stress),
+  vowel_ratio = seq(
+    min(responses_coded$vowel_ratio, na.rm = TRUE),
+    max(responses_coded$vowel_ratio, na.rm = TRUE),
+    by = 0.1
+  ),
+  consonant_ratio = c(-1, 1),
+) %>% 
+  mutate(frequency = 0)
+
+m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
+  mutate_at(vars(consonant_ratio, pthn), function(x) paste0(x, " SD")) %>% 
+  mutate(
+    overlap_stress = paste0("Stress: ", overlap_stress),
+    onset = paste0("Onset: ", onset)
+  )
+
+ggplot(m, aes(vowel_ratio, .value, 
+              colour = interaction(pthn, consonant_ratio, sep = "/"),
+              fill = interaction(pthn, consonant_ratio, sep = "/"))) +
+  facet_wrap(~onset) +
+  # stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
+  stat_summary(fun = mean, geom = "line", size = 1) +
+  labs(x = "Vowel similarity (SD)",
+       y = "P(Correct)", colour = "PTHN/Consonant similarity",
+       fill = "PTHN/Consonant similarity") +
+  guides(colour = guide_legend(ncol = 2)) +
+  theme_custom() +
+  scale_y_continuous(labels = scales::percent) +
+  theme(
+    legend.position = "top",
+    panel.grid = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "responses_vowel.png"))
+```
+##### {-}
 
@@
 ggplot(jtrace, aes(x = activation, y = pct_participant)) +
-  geom_smooth() +
+  geom_smooth(colour = wes_palettes$Darjeeling1[1],
+              fill = wes_palettes$Darjeeling1[1],
+              alpha = 0.5) +
   geom_point(shape = 1, stroke = 1, alpha = 0.5) +
@@
     x = "jTRACE activation value for the word",
-    y = "% participants who responded with the word",
-    title = "jTRACE activation against % participants who gave word as response"
+    y = "Participants who responded with the word"
   ) +
+  scale_y_continuous(labels = scales::percent) +
   theme_custom() +
@@
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  )
+  ) +
+  ggsave(here("Figures", "trace_average.png"))
 
@@
   facet_wrap(~input_language) +
-  geom_smooth() +
+  geom_smooth(colour = wes_palettes$Darjeeling1[1],
+              fill = wes_palettes$Darjeeling1[1],
+              alpha = 0.5) +
   geom_point(shape = 1, stroke = 1, alpha = 0.5) +
@@
   geom_errorbarh(aes(xmin = Estimate-Est.Error, xmax = Estimate+Est.Error), height = 0, size = 5, alpha = 0.5) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), height = 0) +
+  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), height = 0, size = 1) +
   geom_point() +

### 2021-05-21 R/analysis.R

@@ -12,11 +12,13 @@ library(mice) # for multiple imputation
 # set parameters
 set.seed(888) # for reproducibility
 options(mc.cores = 4, chains = 4, iter = 500, seed = 888)
+source("R/utils.R")
 
 # import data ----
-d <- readRDS("Data/accuracy.rds") %>% 
+responses <- readRDS("Data/responses.rds") 
     # typos are considered correct responses
-    mutate_at(vars(onset, overlap_stress, group), as.factor) %>% 
+    # transform relative frequency to Zipf score
+    mutate(frequency_zipf = relative_to_zipf(frequency)) %>% 
     # center predictors
     mutate_at(
         vars(consonant_ratio, vowel_ratio, pthn, frequency),
@@ -25,18 +27,17 @@ d <- readRDS("Data/accuracy.rds") %>%
     mice(m = 5, print = FALSE, method = "pmm") %>% 
     complete() %>% 
     as_tibble() %>% 
-    arrange(trial_id, group)
+    arrange(trial_id, group) %>% 
+    mutate_at(vars(onset, overlap_stress, group), as.factor)
 
-contrasts(d$group) <- c(-0.5, -0.5, 1)
-contrasts(d$onset) <- c(-0.5, 0.5)
-contrasts(d$overlap_stress) <- c(-0.5, 0.5)
+contrasts(responses$group) <- c(-0.5, -0.5, 1)
+contrasts(responses$onset) <- c(-0.5, 0.5)
 
 # fit models ----
 # formula
 f <- bf(
-    correct ~ pthn + frequency + consonant_ratio + vowel_ratio + onset + overlap_stress +
-        (1 + pthn + frequency + consonant_ratio + vowel_ratio + onset + overlap_stress | participant) +
-        (1 | trial_id),
+    correct ~ pthn + consonant_ratio + vowel_ratio + onset +
+        (1 + pthn + consonant_ratio + vowel_ratio | participant),
     family = bernoulli(link = "logit")
 )
 # prior
@@ -44,22 +45,14 @@ priors <- c(
     prior(normal(0, 1), class = "Intercept"),
     prior(normal(0, 1), clas = "b"),
     prior(cauchy(0, 3), class= "sd", group = "participant"),
-    prior(constant(1), class = "sd", group = "trial_id"),
     prior(lkj(5), class = "cor")
 )
 # fit model
 fit <- brm(
-    formula = f, data = d, prior = priors,
[…]

### 2021-05-21 R/trace.R

@@ -8,10 +8,18 @@ library(tidytext)
 library(janitor)
 library(patchwork)
 library(tidybayes)
+library(here)
 
 # set parameters
 set.seed(888) # for reproducibility
-options(mc.cores = 4, chains = 4, iter = 500, seed = 888)
+options(
+    mc.cores = 4, chains = 4, iter = 500, seed = 888,
+    ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
+    ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
+    ggplot2.continuous.fill = wesanderson::wes_palettes$Darjeeling1,
+    ggplot2.continuous.colour = wesanderson::wes_palettes$Darjeeling1
+)
+source("R/utils.R")
 
 # import data ----
 trials <- readRDS(here("Data", "trials.rds")) %>% 
@@ -21,7 +29,7 @@ trials <- readRDS(here("Data", "trials.rds")) %>%
         pthn_log = log(pthn),
         overlap_stress = ifelse(overlap_stress==1, "Same", "Different")
     )
-accuracy <- readRDS(here("Data", "accuracy.rds"))
+responses <- readRDS(here("Data", "responses.rds"))
 
 
 # by Serene (some fixes by Gonzalo) ----
@@ -78,7 +86,7 @@ trace_output <- left_join(trace_output, df_transcription, by = c("output" = "tra
 # Cleaned file with participant answers, 1 row per trial per participant
 
 # count the number of participants who gave each answer in response to a given trial
-df_match <- accuracy %>%
+df_match <- responses %>%
     rename(answer = input_text) %>% 
     group_by(test_language, word, answer) %>%
     summarise(n_participant = n(), .groups = "drop") %>%
@@ -112,9 +120,8 @@ d <- df_match %>%
 contrasts(d$onset) <- c(-0.5, 0.5)
 contrasts(d$overlap_stress) <- c(-0.5, 0.5)
 
-f <- bf(pct_participant ~
-            activation*(consonant_ratio + vowel_ratio + onset + overlap_stress + pthn + frequency) +
-            (1 + activation | group/input),
+f <- bf(pct_participant ~ activation*(consonant_ratio + vowel_ratio + onset + pthn) +
+            (1 + activation | input),
         family = zero_one_inflated_beta("logit"))
 
 prior <- c(
@@ -180,7 +187,7 @@ trace_top <- trace %>%
     select(group, input, response, activation, correct)
 
 # experimental results ----
[…]

### 2021-05-25 Rmd/jtrace_report.Rmd
Add model comparisons using LOO-IC

---
title: "jTRACE notbook"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
    echo = FALSE,
    message = TRUE,
    warning = TRUE, 
    cache.extra = knitr::rand_seed,
    out.width = "100%",
    results = "asis",
    dev.args = list(png = list(type = "cairo"))
)
options(knitr.kable.NA = '-')
```

```{r prepare, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

# load packages
library(tidyverse)
library(googlesheets4)
library(gt)

# import data
phonemes <- range_read(
    "1iAK2zF84MqFwdLj-_CpWVC1O2xxmYDztc0Qe_2Sy0ww",
    sheet = "(Gonzalo coding)"
) %>% 
    mutate_at(
        vars(pow, voc, diff, acu, cons, voi, burst),
        as.numeric
    )

lexicon <- sheet_names("1pZE58wHYGFxaIC90hSr-202Qv9go6-VZLixl_lF1QKY")[-1] %>% 
    map(
        function(x){
            range_read(
                ss = "1pZE58wHYGFxaIC90hSr-202Qv9go6-VZLixl_lF1QKY",
                sheet = x
            ) %>% 
                select(trial_id, ipa1, ipa2, word1, word2, jTRACE1, jTRACE2)
        }
    ) %>%
    set_names(sheet_names("1pZE58wHYGFxaIC90hSr-202Qv9go6-VZLixl_lF1QKY")[-1])

```

# Phonemes {.tabset .tab-pills}

## Phoneme inventory {.tabset .tab-pills}

### Consonants

```{r phonemes_consonants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Consonant") %>% 
    select(id, ipa, sampa, trace, english, spanish, catalan, voice, place, manner) %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(everything()) %>% 
    cols_label(
        id = "ID",
        sampa = "SAMPA",
        trace = "jTRACE",
        english = "English",
        spanish = "Spanish",
        catalan = "Catalan",
        voice = "Voice",
        place = "Place",
        manner = "Manner"
    ) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("english"),
			rows = english
		)
	) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("spanish"),
			rows = spanish
		)
	) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("catalan"),
			rows = catalan
		)
	) %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### Vowels

```{r phonemes_vowels, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Vowel") %>% 
    select(id, ipa, sampa, trace, english, spanish, catalan, height, backness, roundness) %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(2:7) %>% 
    cols_label(
        id = "ID",
        sampa = "SAMPA",
        trace = "jTRACE",
        height = "Height",
        backness = "Backness",
        roundness = "Roundness"
    ) %>% 
       tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("spanish"),
			rows = spanish
		)
	) %>% 
    tab_style(
		style = list(
			cell_fill(color = "#00FA9A")
		),
		locations = cells_body(
			columns = c("catalan"),
			rows = catalan
		)
	) %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### {-}

## Features {.tabset .tab-pills}

### Consonants

```{r features_consonants, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Consonant") %>% 
    select(id, ipa, trace, description, pow, cons, voc, diff, acu, voi, burst) %>% 
    gt() %>% 
    cols_label(
        id = "ID",
        ipa = "IPA",
        trace = "jTRACE",
        description = "Description",
        pow = "Power",
        voc = "Vocalic",
        diff = "Diffuseness",
        acu = "Acuteness",
        cons = "Consonantal",
        voi = "Voicing",
        burst = "Burst"
    ) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(2:4) %>%
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### Vowels

```{r features_vowels, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    filter(type=="Vowel") %>% 
    select(id, ipa, trace, description, pow, cons, voc, diff, acu, voi, burst) %>% 
    gt() %>% 
    cols_label(
        id = "ID",
        ipa = "IPA",
        trace = "jTRACE",
        description = "Description",
        pow = "Power",
        voc = "Vocalic",
        diff = "Diffuseness",
        acu = "Acuteness",
        cons = "Consonantal",
        voi = "Voicing",
        burst = "Burst"
    ) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_markdown(2:4) %>%
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

### Feature correlations

```{r features_correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
phonemes %>% 
    select(pow, voc, diff, acu, cons, voi, burst) %>% 
    mutate_all(replace_na, replace = 0) %>%
    as.matrix() %>% 
    heatmap(
        labRow = phonemes$trace,
        labCol = c("Power", "Vocalic", "Difusseness", "Acuteness", "Consonantal", "Voice", "Burst"),
        Colv = NA,
        na.rm = FALSE
    )
```

### {-}

## {-}

# Lexicon {.tabset .tab-pills}

## English-Spanish

```{r lexicon_engspa, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
lexicon$`English-Spanish` %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

## English-Catalan

```{r lexicon_engcat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
lexicon$`English-Catalan` %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

## Spanish-Catalan

```{r lexicon_spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
lexicon$`Spanish-Catalan` %>% 
    gt() %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    tab_options(
        container.overflow.y = TRUE,
        container.height = 500
    )
```

## {-}

# Parameters

# Design

### 2021-05-25 Rmd/report.Rmd
Add model comparisons using LOO-IC

@@
 library(patchwork)
+library(janitor)
 library(mice)
@@
 fit_responses <- readRDS(here("Results", "fit_responses.rds"))
+loo <- readRDS(here("Results", "loo.rds"))
 
@@
   theme(
-    legend.position = "none",
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
@@
 
-Participants in the ENG-SPA listened to `r nrow(trials[trials$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `nrow(trials[trials$group=="ENG-CAT",]) trials`.
+Participants in the ENG-SPA listened to `r nrow(trials[trials$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(trials[trials$group=="ENG-CAT",])` trials.
 Five trials in each condition were practice trials and were excluded from analyses.
@@
 
-We also added participants (`participants`) and words (`word`) as grouping variables (aka., random effects), therefore adjusting the model to each participant and word.
-We specified random intercepts for both grouping variables, and random slopes for the fixed effects within the `participant` grouping variable.
-
-We model was implemented using the following formula:
-
-```
-correct ~ pthn + frequency + consonant_ratio + vowel_ratio + onset + overlap_stress + (1 + pthn + frequency + consonant_ratio + vowel_ratio + onset + overlap_stress | participant) + (1 | trial_id)
-```
-
-The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 intertions each.
+We also added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model.
+We specified random intercepts slopes for all fixed effects within the `participant` grouping variable.
+
+We performed multiple imputation via predictive mean matching using the `mice` function of the `mice` package [@van2011mice] to impute missing values in the response variable or the predictors.
+
+The model was implemented using the following formula:
+
+```
+~ pthn + consonant_ratio + vowel_ratio + onset + (1 + pthn + consonant_ratio + vowel_ratio + onset | participant)
+```
+
+The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 iterations each (1,000 warmups).
 The output of this model is the approximated posterior distribution of all the parameters estimated, therefore indicating the probability of each combination of values of the parameters, conditional to the observed data.
+
+We compared this (extended model) to other models dropping one predictor at a time using Leave-one-out cross-validation (LOO-CV), using the `loo` and `loo_compare` functions of the `brms` package.
+We compared our extended model against the following models:
+
+* fit_2: `~ pthn + consonant_ratio + vowel_ratio + (1 + pthn + consonant_ratio + vowel_ratio | participant)`
+* fit_3: `~ pthn + consonant_ratio + (1 + pthn + consonant_ratio | participant)`
+* fit_4: `~ pthn + (1 + pthn | participant)`
+* fit_5: `~ 1 + (1  | participant)`
 
@@
 
-### Accuracy by item
-
-```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses_item <- responses %>%
-  group_by(group, word) %>%
-  summarise(
-    correct_sum = sum(correct, na.rm = TRUE),
-    correct_n = sum(!is.na(correct), na.rm = TRUE),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    correct_prop = prop_adj(correct_sum, correct_n),
-    correct_se = prop_adj_se(correct_sum, correct_n),
-    word_label = reorder_within(word, by = correct_prop, within = group)
-  )
-
-ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
-  facet_wrap(~group, scales = "free_y") +
-  geom_hline(yintercept = 0.5, colour = "black") +
-  geom_errorbar() +
-  geom_point(size = 0.5) +
-  labs(y = "Correct responses", x = "Word", colour = "Group") +
-  coord_flip() +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme_custom() +
-  theme(
-    legend.position = "none",
-    axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 9),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "accuracy_item.png"))
-
-```
-
-### Accuracy by participant
-
-```{r responses_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses_participant <- responses %>%
-  group_by(group, participant) %>%
-  summarise(
-    correct_sum = sum(correct, na.rm = TRUE),
-    correct_n = sum(!is.na(correct), na.rm = TRUE),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    correct_prop = prop_adj(correct_sum, correct_n),
-    correct_se = prop_adj_se(correct_sum, correct_n),
-    participant_label = reorder_within(participant, by = correct_prop, within = group)
-  )
-
-ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
-  facet_wrap(~group, scales = "free_y") +
-  geom_hline(yintercept = 0.5) +
-  geom_errorbar(width = 0.5) +
-  geom_point(size = 1) +
-  labs(y = "Correct responses", x = "Participant", colour = "Group") +
-  coord_flip() +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme_custom() +
-  theme(
-    legend.position = "none",
-    axis.text.y = element_blank(),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 9),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "accuracy_participant.png"))
-
-```
-
 ### Model outputs {.tabset .tab-pill}
@@
 
+#### Model comparison
+
+Leave-one-out cross-validation.
+The more negative is `elpd_diff` (and the larger its magnitude compared to its corresponding standard error), the better it first the data compared to the null model (`~ 1 + (1 | participant)`).
+The model that fits the data the best is the extended model.
+
+```{r model_comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+map(loo$loos, "estimates") %>% 
+  map(function(x){
+    as.data.frame(x) %>% 
+      rownames_to_column("ic")
+  }
+  ) %>% 
+  bind_rows(.id = "model") %>% 
+  clean_names() %>% 
+  filter(ic %in% c("elpd_loo", "looic")) %>% 
+  pivot_wider(names_from = ic, values_from = c("estimate", "se")) %>% 
+  left_join(rownames_to_column(as.data.frame(loo$comparison, "model"))) %>% 
+  mutate(formula = c(
+    "... + onset + (0 + onset | participant)",
+    "... + vowel_ratio + (0 + vowel_ratio | participant)",
+    "... + consonant_ratio + (0 + consonant_ratio | participant)",
+    "... + pthn + (0 + pthn | participant)",
+    "~ 1 + (1  | participant)"
+  )) %>% 
+  arrange(desc(model)) %>% 
+  select(formula, estimate_elpd_loo, se_elpd_loo, estimate_looic, se_looic, elpd_diff, se_diff) %>%
+  rename(model = formula) %>% 
+  gt() %>% 
+  fmt_number(2:7) %>% 
+  cols_merge(c("estimate_elpd_loo", "se_elpd_loo"), pattern = "{1} ({2})") %>% 
+  cols_merge(c("estimate_looic", "se_looic"), pattern = "{1} ({2})") %>% 
+  cols_merge(c("elpd_diff", "se_diff"), pattern = "{1} ({2})") %>% 
+  cols_label(
+    model = md("**Model**"),
+    estimate_elpd_loo = md("**ELPD<sub>LOO</sub> (SE)**"),
+    estimate_looic = md("**LOO-IC (SE)**"),
+    elpd_diff = md("**ELPD<sub>diff</sub> (SE)**")
+
+  )
+
+```
+
 #### Random effects
@@
 ```
+
 ##### {-}
@@
 
+### Accuracy by item
+
+```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_item <- responses %>%
+  group_by(group, word) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    word_label = reorder_within(word, by = correct_prop, within = group)
+  )
+
+ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5, colour = "black") +
+  geom_errorbar() +
+  geom_point(size = 0.5) +
+  labs(y = "Correct responses", x = "Word", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "accuracy_item.png"))
+
+```
+
+### Accuracy by participant
+
+```{r responses_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_participant <- responses %>%
+  group_by(group, participant) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    participant_label = reorder_within(participant, by = correct_prop, within = group)
+  )
+
+ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5) +
+  geom_errorbar(width = 0.5) +
+  geom_point(size = 1) +
+  labs(y = "Correct responses", x = "Participant", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "accuracy_participant.png"))
+
+```
+
 ### {-}
@@
 
-```{r jtrace_all_plot, echo=FALSE, warning=FALSE}
+```{r jtrace_all_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 ggplot(jtrace, aes(x = activation, y = pct_participant)) +

### 2021-05-25 R/analysis.R

@@ -8,14 +8,15 @@ library(janitor) # for cleaning variable names
 library(brms) # for fitting Bayesian models
 library(tidybayes) # for processing Bayesian models
 library(mice) # for multiple imputation
+library(here) # for locating files
 
 # set parameters
 set.seed(888) # for reproducibility
 options(mc.cores = 4, chains = 4, iter = 500, seed = 888)
-source("R/utils.R")
+source(here("R", "utils.R"))
 
 # import data ----
-responses <- readRDS("Data/responses.rds") 
+responses <- readRDS(here("Data", "responses.rds")) %>% 
     # typos are considered correct responses
     # transform relative frequency to Zipf score
     mutate(frequency_zipf = relative_to_zipf(frequency)) %>% 
@@ -37,22 +38,38 @@ contrasts(responses$onset) <- c(-0.5, 0.5)
 # formula
 f <- bf(
     correct ~ pthn + consonant_ratio + vowel_ratio + onset +
-        (1 + pthn + consonant_ratio + vowel_ratio | participant),
+        (1 + pthn + consonant_ratio + vowel_ratio + onset | participant),
     family = bernoulli(link = "logit")
 )
+f_2 <- bf(correct ~ pthn + consonant_ratio + vowel_ratio + (1 + pthn + consonant_ratio + vowel_ratio | participant), family = bernoulli(link = "logit"))
+f_3 <- bf(correct ~ pthn + consonant_ratio + (1 + pthn + consonant_ratio | participant), family = bernoulli(link = "logit"))
+f_4 <- bf(correct ~ pthn + (1 + pthn | participant), family = bernoulli(link = "logit"))
+f_5 <- bf(correct ~ 1 + (1 | participant), family = bernoulli(link = "logit"))
+
 # prior
 priors <- c(
-    prior(normal(0, 1), class = "Intercept"),
-    prior(normal(0, 1), clas = "b"),
+    prior(normal(0, 3), class = "Intercept"),
+    prior(normal(0, 3), clas = "b"),
     prior(cauchy(0, 3), class= "sd", group = "participant"),
     prior(lkj(5), class = "cor")
 )
-# fit model
-fit <- brm(
-    formula = f, data = responses, prior = priors,
-    backend = "cmdstanr",
-    file = "Results/fit_responses.rds"
-)
+# fit models
[…]

### 2021-05-25 Rmd/report.Rmd
Replace vowel and consonant similarity with Levenshtein distance and add interactions

@@
 fit_jtrace <- readRDS(here("Results", "fit_jtrace.rds"))
-fit_responses <- readRDS(here("Results", "fit_responses.rds"))
+fit_responses <- readRDS(here("Results", "fit_responses_3.rds"))
 loo <- readRDS(here("Results", "loo.rds"))
@@
     pthn_log = log(pthn),
-    overlap_stress = ifelse(overlap_stress==1, "Same", "Different")
+    lv_log = log(lv)
   )
@@
 trials_coded %>% 
+  ggplot(aes(group, lv, fill = group, colour = group)) +
+  geom_violin() +
+  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
+  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
+  labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme_custom() +
+  theme(
+    legend.position = "none",
+    axis.title.x = element_blank(),
+    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+  ) +
+
+  trials_coded %>% 
   select(trial_id, group, ends_with("_ratio")) %>% 
@@
   ggplot(aes(group, ratio, fill = group, colour = group)) +
-  facet_wrap(~type) +
+  facet_wrap(~type, ncol = 1) +
   geom_violin() +
@@
     legend.position = "none",
+    axis.title.y = element_blank(),
+    axis.title.x = element_blank(),
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  )
+  ) +
+
+  plot_layout(ncol = 2)
 
@@
 We modelled the probability of an average participant providing a correct response, conditional to a set of predictors (aka.
-fixed effects):
-
-* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database.
-* Vowel similarity (`vowel_ratio`, 0-1): edit distance between the sequence of vowels of the presented word  and the target word. Computed using custom code.
-* Consonant similarity (`consonant_ratio`, 0-1): edit distance between the sequence of consonants of the presented word and the target word. Computed using custom code.
-* Shared phonological onset (`onset`, 0 or 1): do the presented word and the target word share phonological onset? Manually coded.
-
-We also added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model.
+fixed effects) and their interactions:
+
+* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database. Standardised before entering the regression model.
+* Phonological similarity (`lv`, 0-1): Inverse Levenshtein distance between the IPA transcription of the presented word and its correct translation in the target language, calculated using the `stringdist` function of the `stringdist` package [@loo2014stringdist].
+Standardised before entering the regression model.
+* Shared phonological onset (`onset`, Same/Different): Whether the presented and the target word share phonological onset, as judged by experimenters, and sum coded as $Different = -0.5$ and $Same = +0.5$.
+
+We added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model.
 We specified random intercepts slopes for all fixed effects within the `participant` grouping variable.
@@
 
-The model was implemented using the following formula:
-
-```
-~ pthn + consonant_ratio + vowel_ratio + onset + (1 + pthn + consonant_ratio + vowel_ratio + onset | participant)
+The model was implemented using the following formula (`fit_3`):
+
+```
+~ pthn*lv*onset + (1 + pthn*lv*onset | participant)
 ```
@@
 
-* fit_2: `~ pthn + consonant_ratio + vowel_ratio + (1 + pthn + consonant_ratio + vowel_ratio | participant)`
-* fit_3: `~ pthn + consonant_ratio + (1 + pthn + consonant_ratio | participant)`
-* fit_4: `~ pthn + (1 + pthn | participant)`
-* fit_5: `~ 1 + (1  | participant)`
+* fit_0: `~ 1 + (1  | participant)`
+* fit_1: `~ pthn + (1 + pthn | participant)`
+* fit_2: `~ pthn*lv + (1 + pthn*lv | participant)`
 
@@
   vector[N] Z_1_7;
+  vector[N] Z_1_8;
   int<lower=1> NC_1;  // number of group-level correlations
-// data for group-level effects of ID 2
-int<lower=1> N_2;  // number of grouping levels
-int<lower=1> M_2;  // number of coefficients per level
-int<lower=1> J_2[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_2_1;
   int prior_only;  // should the likelihood be ignored?
@@
   real Intercept;  // temporary intercept for centered predictors
+  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
   matrix[M_1, N_1] z_1;  // standardized group-level effects
   cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-vector<lower=0>[M_2] sd_2;  // group-level standard deviations
-vector[N_2] z_2[M_2];  // standardized group-level effects
 }
 transformed parameters {
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
   matrix[N_1, M_1] r_1;  // actual group-level effects
@@
   vector[N_1] r_1_7;
-vector[N_2] r_2_1;  // actual group-level effects
-sd_1 = rep_vector(1, rows(sd_1));
+  vector[N_1] r_1_8;
   // compute actual group-level effects
@@
   r_1_7 = r_1[, 7];
-r_2_1 = (sd_2[1] * (z_2[1]));
+  r_1_8 = r_1[, 8];
 }
@@
       // add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n];
+      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_1_8[J_1[n]] * Z_1_8[n];
     }
@@
   // priors including constants
-target += normal_lpdf(b | 0, 1);
-target += normal_lpdf(Intercept | 0, 1);
+  target += normal_lpdf(b | 0, 3);
+  target += normal_lpdf(Intercept | 0, 3);
+  target += cauchy_lpdf(sd_1 | 0, 3)
+    - 8 * cauchy_lccdf(0 | 0, 3);
   target += std_normal_lpdf(to_vector(z_1));
   target += lkj_corr_cholesky_lpdf(L_1 | 5);
-target += cauchy_lpdf(sd_2 | 0, 3)
-- 1 * cauchy_lccdf(0 | 0, 3);
-target += std_normal_lpdf(z_2[1]);
 }
@@
   mutate_at(
-    vars(consonant_ratio, vowel_ratio, pthn, frequency),
+    vars(lv, consonant_ratio, vowel_ratio, pthn, frequency),
     function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
@@
     footnote = "Transformed using the inverse logit to get the average probability of correct response",
-    locations = cells_body(columns = "term", rows = term=="(Intercept)")
+    locations = cells_body(columns = "term", rows = term=="Intercept")
   ) %>% 
@@
     footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("pthn", "consonant_ratio", "vowel_ratio", "onset1"))
+    locations = cells_body(columns = "term", rows = term %in% c("pthn", "lv", "onset1", "pthn:lv", "pthn:onset1", "lv:onset1", "pthn:lv:onset1"))
   ) %>% 
@@
 
-```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+The intercept has been transformed using the inverse logit to get the average probability of correct response.
+The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
 post <- gather_draws(fit_responses, `b_.*`, regex = TRUE) %>% 
@@
     legend.position = "none",
-    axis.text = element_text(size = 7),
+    plot.caption.position = "plot",
+    plot.caption = element_text(hjust = 0),
+    axis.text = element_text(size= 7),
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
@@
 ```{r model_comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
 map(loo$loos, "estimates") %>% 
@@
   mutate(formula = c(
-    "... + onset + (0 + onset | participant)",
-    "... + vowel_ratio + (0 + vowel_ratio | participant)",
-    "... + consonant_ratio + (0 + consonant_ratio | participant)",
-    "... + pthn + (0 + pthn | participant)",
-    "~ 1 + (1  | participant)"
+    "...+onset+pthn:onset+lv:onset+pthn:lv:onset+(0+onset+pthn:onset+lv:onset+pthn:lv:onset|participant)",
+    "...+lv+pthn:lv+(0+lv+pthn:lv|participant)",
+    "...+pthn+(0+pthn|participant)",
+    "~1+(1 |participant)"
   )) %>% 
@@
 ggplot(post_re, aes(.value, sort(as.factor(participant)), xmin = .lower, xmax = .upper, colour = group)) +
-  facet_grid(group~.param, scales = "free_y") +
+  facet_grid(group~.param, scales = "free") +
   geom_errorbarh(height = 0.5, alpha = 0.75) +
@@
 nd <- expand_grid(
-  pthn = seq(
-    min(responses_coded$pthn, na.rm = TRUE),
-    max(responses_coded$pthn, na.rm = TRUE),
-    by = 0.1
-  ),
+  pthn = seq(min(responses_coded$pthn, na.rm = TRUE), max(responses_coded$pthn, na.rm = TRUE), by = 0.1),
   onset = unique(responses_coded$onset),
-  overlap_stress = unique(responses_coded$overlap_stress),
-  vowel_ratio = c(-1, 1),
-  consonant_ratio = c(-1, 1)
-) %>% 
-  mutate(frequency = 0)
+  lv = c(-1, 1)
+)
 
 m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
-  mutate_at(vars(ends_with("_ratio")), function(x) paste0(x, " SD")) %>% 
-  mutate(
-    overlap_stress = paste0("Stress: ", overlap_stress),
-    onset = paste0("Onset: ", onset)
-  )
-
-ggplot(m, aes(pthn, .value, 
-              colour = interaction(consonant_ratio, vowel_ratio, sep = "/"),
-              fill = interaction(consonant_ratio, vowel_ratio, sep = "/"))) +
+  mutate_at(vars(lv), function(x) paste0(x, " SD")) %>% 
+  mutate(onset = paste0("Onset: ", onset))
+
+ggplot(m, aes(pthn, .value, colour = lv, fill = lv)) +
   facet_wrap(~onset) +
@@
   labs(x = "Phonological neighbourhood density\n(higher frequency neighbours)",
-       y = "P(Correct)", colour = "Consonant/Vowel similarity",
-       fill = "Consonant similarity/Vowel similarity") +
+       y = "P(Correct)", colour = "Phonological similarity",
+       fill = "Phonological similarity") +
   guides(colour = guide_legend(ncol = 2)) +
@@
 
-##### By consonant similarity
+##### By phonological similarity similarity
 
@@
   onset = unique(responses_coded$onset),
-  overlap_stress = unique(responses_coded$overlap_stress),
-  vowel_ratio = c(-1, 1),
-  consonant_ratio = seq(
-    min(responses_coded$consonant_ratio, na.rm = TRUE),
-    max(responses_coded$consonant_ratio, na.rm = TRUE),
+  lv = seq(
+    min(responses_coded$lv, na.rm = TRUE),
+    max(responses_coded$lv, na.rm = TRUE),
     by = 0.1
   )
-) %>% 
-  mutate(frequency = 0)
+)
 
 m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
-  mutate_at(vars(vowel_ratio, pthn), function(x) paste0(x, " SD")) %>% 
-  mutate(
-    overlap_stress = paste0("Stress: ", overlap_stress),
-    onset = paste0("Onset: ", onset)
-  )
-
-ggplot(m, aes(consonant_ratio, .value, 
-              colour = interaction(pthn, vowel_ratio, sep = "/"),
-              fill = interaction(pthn, vowel_ratio, sep = "/"))) +
+  mutate_at(vars(pthn), function(x) paste0(x, " SD")) %>% 
+  mutate(onset = paste0("Onset: ", onset))
+
+ggplot(m, aes(lv, .value, colour = pthn, fill = pthn)) +
   facet_wrap(~onset) +
@@
   labs(x = "Consonant similarity (SD)",
-       y = "P(Correct)", colour = "PTHN/Vowel similarity",
+       y = "P(Correct)", colour = "PTHN",
        fill = "PTHN/Vowel similarity") +
@@
 
-```
-
-##### By vowel similarity
-
-```{r emmeans_vowel, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-nd <- expand_grid(
-  pthn = c(-1, 1),
-  onset = unique(responses_coded$onset),
-  overlap_stress = unique(responses_coded$overlap_stress),
-  vowel_ratio = seq(
-    min(responses_coded$vowel_ratio, na.rm = TRUE),
-    max(responses_coded$vowel_ratio, na.rm = TRUE),
-    by = 0.1
-  ),
-  consonant_ratio = c(-1, 1),
-) %>% 
-  mutate(frequency = 0)
-
-m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
-  mutate_at(vars(consonant_ratio, pthn), function(x) paste0(x, " SD")) %>% 
-  mutate(
-    overlap_stress = paste0("Stress: ", overlap_stress),
-    onset = paste0("Onset: ", onset)
-  )
-
-ggplot(m, aes(vowel_ratio, .value, 
-              colour = interaction(pthn, consonant_ratio, sep = "/"),
-              fill = interaction(pthn, consonant_ratio, sep = "/"))) +
-  facet_wrap(~onset) +
-  # stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
-  stat_summary(fun = mean, geom = "line", size = 1) +
-  labs(x = "Vowel similarity (SD)",
-       y = "P(Correct)", colour = "PTHN/Consonant similarity",
-       fill = "PTHN/Consonant similarity") +
-  guides(colour = guide_legend(ncol = 2)) +
-  theme_custom() +
-  scale_y_continuous(labels = scales::percent) +
-  theme(
-    legend.position = "top",
-    panel.grid = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "responses_vowel.png"))
 ```

### 2021-05-25 R/analysis.R

@@ -22,7 +22,7 @@ responses <- readRDS(here("Data", "responses.rds")) %>%
     mutate(frequency_zipf = relative_to_zipf(frequency)) %>% 
     # center predictors
     mutate_at(
-        vars(consonant_ratio, vowel_ratio, pthn, frequency),
+        vars(lv, consonant_ratio, vowel_ratio, pthn, frequency),
         function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
     # impute missing data
     mice(m = 5, print = FALSE, method = "pmm") %>% 
@@ -36,15 +36,10 @@ contrasts(responses$onset) <- c(-0.5, 0.5)
 
 # fit models ----
 # formula
-f <- bf(
-    correct ~ pthn + consonant_ratio + vowel_ratio + onset +
-        (1 + pthn + consonant_ratio + vowel_ratio + onset | participant),
-    family = bernoulli(link = "logit")
-)
-f_2 <- bf(correct ~ pthn + consonant_ratio + vowel_ratio + (1 + pthn + consonant_ratio + vowel_ratio | participant), family = bernoulli(link = "logit"))
-f_3 <- bf(correct ~ pthn + consonant_ratio + (1 + pthn + consonant_ratio | participant), family = bernoulli(link = "logit"))
-f_4 <- bf(correct ~ pthn + (1 + pthn | participant), family = bernoulli(link = "logit"))
-f_5 <- bf(correct ~ 1 + (1 | participant), family = bernoulli(link = "logit"))
+f_0 <- bf(correct ~ 1 + (1 | participant), family = bernoulli(link = "logit"))
+f_1 <- bf(correct ~ pthn + (1 + pthn | participant), family = bernoulli(link = "logit"))
+f_2 <- bf(correct ~ pthn*lv + (1 + pthn*lv | participant), family = bernoulli(link = "logit"))
+f_3 <- bf(correct ~ pthn*lv*onset + (1 + pthn*lv*onset | participant), family = bernoulli(link = "logit"))
 
 # prior
 priors <- c(
@@ -53,27 +48,27 @@ priors <- c(
     prior(cauchy(0, 3), class= "sd", group = "participant"),
     prior(lkj(5), class = "cor")
 )
+
 # fit models
-fit <- brm(f, data = responses, prior = priors, backend = "cmdstanr",
-           file = here("Results", "fit_responses.rds"))
[…]

### 2021-05-25 R/processing.R

@@ -5,14 +5,13 @@
 # load packages
 library(tidyverse) # for manipulating variables
 library(readxl) # for importing Excel spreadsheets
-library(googlesheets4) # for importing Google spreadsheets
 library(janitor) # for cleaning variable names
 library(lubridate) # for working with dates
 library(data.table) # for importing data
-library(stringdist) # for calculating Levenshtein distance
+library(here) # for locating files
 
 # create/load functions
-source("R/utils.R") # helper functions
+source(here("R", "utils.R")) # helper functions
 
 # set parameters
 set.seed(888) # for reproducibility
@@ -20,45 +19,17 @@ practice_trials <- c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)
 spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
 
 # import trial data ----
-# import Clearpond data
-clearpond_english <- read.csv("Data/clearpond/clearpond_english.csv") %>% clean_names() %>% as_tibble() %>% select(-c(x, s_pthn))
-clearpond_spanish <- read.csv("Data/clearpond/clearpond_spanish.csv")[, -1] %>% clean_names() %>% as_tibble()
-clearpond <- bind_rows(
-    list(
-        `ENG-CAT` = clearpond_english,
-        `ENG-SPA` = clearpond_english,
-        `SPA-CAT` = clearpond_spanish),
-    .id = "group"
-) %>% 
-    rename(word2 = word, frequency = freq_per_million) %>% 
-    mutate(word2 = str_to_lower(word2)) %>% 
-    select(word2, group, pthn, frequency)
 
 # import trials
-trace <- read_xlsx("Stimuli/trace.xlsx") %>% 
-    list("ENG-SPA" = ., "ENG-CAT" = .) %>% 
-    bind_rows(.id = "group") %>% 
-    rename(word2 = ort_eng, trace = trace_eng)
-
-trials <- c("ENG-SPA" = "English-Spanish",
-            "ENG-CAT" = "English-Catalan",
-            "SPA-CAT" = "Spanish-Catalan") %>% 
-    map(function(x) range_read(ss = "1pZE58wHYGFxaIC90hSr-202Qv9go6-VZLixl_lF1QKY", sheet = x)) %>% 
-    bind_rows(.id = "group") %>% 
-    clean_names() %>% 
[…]

### 2021-05-26 Rmd/report.Rmd
Drop onset from model and fix y-axis in interacion plots.

@@
     toc: true
-    toc_depth: 2
+    toc_depth: 3
     toc_float: true
@@
 fit_jtrace <- readRDS(here("Results", "fit_jtrace.rds"))
-fit_responses <- readRDS(here("Results", "fit_responses_3.rds"))
+fit_responses <- readRDS(here("Results", "fit_responses_2.rds"))
 loo <- readRDS(here("Results", "loo.rds"))
@@
 
-#### Vowel and consonant similarity
+#### Phonological similarity
 
@@
 
+#### Phonological similarity and onset
+
+```{r trials_similarity_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+trials_coded %>% 
+  select(trial_id, group, lv, onset) %>% 
+  ggplot(aes(lv, fill = onset, colour = onset)) +
+  facet_wrap(~group, ncol = 2) +
+  geom_histogram() +
+  labs(x = "Phonological Leveshtein distance\n(corrected for string length)", y = "# trials",
+       fill = "Phonological onset", colour = "Phonological onset") +
+  theme_custom() +
+  theme(
+    legend.position = c(0.75, 0.25),
+    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+  )
+```
+
 #### {-}
@@
 
-* fit_0: `~ 1 + (1  | participant)`
-* fit_1: `~ pthn + (1 + pthn | participant)`
-* fit_2: `~ pthn*lv + (1 + pthn*lv | participant)`
+* fit_0: `~ 1 + frequency + (1  | participant)`
+* fit_1: `~ 1 + frequency + pthn + (1 + pthn | participant)`
+* fit_2: `~ 1 + frequency + pthn*lv + (1 + pthn*lv | participant)`
 
@@
 vector[N] Z_1_4;
-  vector[N] Z_1_5;
-  vector[N] Z_1_6;
-  vector[N] Z_1_7;
-  vector[N] Z_1_8;
 int<lower=1> NC_1;  // number of group-level correlations
@@
 vector[N_1] r_1_4;
-  vector[N_1] r_1_5;
-  vector[N_1] r_1_6;
-  vector[N_1] r_1_7;
-  vector[N_1] r_1_8;
 // compute actual group-level effects
@@
 r_1_4 = r_1[, 4];
-  r_1_5 = r_1[, 5];
-  r_1_6 = r_1[, 6];
-  r_1_7 = r_1[, 7];
-  r_1_8 = r_1[, 8];
 }
@@
 // add more terms to the linear predictor
-      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_1_8[J_1[n]] * Z_1_8[n];
+mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n];
 }
@@
 target += cauchy_lpdf(sd_1 | 0, 3)
-    - 8 * cauchy_lccdf(0 | 0, 3);
+- 4 * cauchy_lccdf(0 | 0, 3);
 target += std_normal_lpdf(to_vector(z_1));
@@
 
-ggplot(post, aes(.value, fill = .variable)) +
-  facet_wrap(~.variable, scales = "free") +
+ggplot(post, aes(.value, .variable, fill = .variable)) +
   geom_vline(xintercept = 0) +
@@
     plot.caption = element_text(hjust = 0),
-    axis.text = element_text(size= 7),
+    # axis.text = element_text(size = 7),
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
@@
   mutate(formula = c(
-    "...+onset+pthn:onset+lv:onset+pthn:lv:onset+(0+onset+pthn:onset+lv:onset+pthn:lv:onset|participant)",
-    "...+lv+pthn:lv+(0+lv+pthn:lv|participant)",
-    "...+pthn+(0+pthn|participant)",
-    "~1+(1 |participant)"
+    "~ 1 + frequency + pthn*lv + (1 + pthn*lv | participant)",
+    "~ 1 + frequency + pthn + (1 + pthn | participant)",
+    "~ 1 + frequency + (1 | participant)"
   )) %>% 
@@
 
-#### Marginal effects {.tabset .tab-pills}
+#### Marginal effects
 
-##### By PTHN
+```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=6}
 
-```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+# by pthn
 nd <- expand_grid(
   pthn = seq(min(responses_coded$pthn, na.rm = TRUE), max(responses_coded$pthn, na.rm = TRUE), by = 0.1),
-  onset = unique(responses_coded$onset),
+  frequency = 0,
   lv = c(-1, 1)
@@
 m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
-  mutate_at(vars(lv), function(x) paste0(x, " SD")) %>% 
-  mutate(onset = paste0("Onset: ", onset))
+  mutate_at(vars(lv), function(x) paste0(x, " SD"))
 
 ggplot(m, aes(pthn, .value, colour = lv, fill = lv)) +
-  facet_wrap(~onset) +
-  # stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
+  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
   stat_summary(fun = mean, geom = "line", size = 1) +
-  labs(x = "Phonological neighbourhood density\n(higher frequency neighbours)",
-       y = "P(Correct)", colour = "Phonological similarity",
-       fill = "Phonological similarity") +
+  labs(x = "PTHN", y = "P(Correct)", colour = "Phon. similarity", fill = "Phon. similarity") +
+  scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
   guides(colour = guide_legend(ncol = 2)) +
   theme_custom() +
-  scale_y_continuous(labels = scales::percent) +
   theme(
@@
     panel.grid = element_line(colour = "grey", linetype = "dotted")
-  ) +
+  ) 
 ggsave(here("Figures", "responses_pthn.png"))
@@
 
-##### By phonological similarity similarity
-
-```{r emmeans_consonant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r emmeans_lv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=6}
+# by lv
 nd <- expand_grid(
   pthn = c(-1, 1),
-  onset = unique(responses_coded$onset),
+  frequency = 0,
   lv = seq(
@@
 m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
-  mutate_at(vars(pthn), function(x) paste0(x, " SD")) %>% 
-  mutate(onset = paste0("Onset: ", onset))
+  mutate_at(vars(pthn), function(x) paste0(x, " SD"))
 
 ggplot(m, aes(lv, .value, colour = pthn, fill = pthn)) +
-  facet_wrap(~onset) +
-  # stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
+  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
   stat_summary(fun = mean, geom = "line", size = 1) +
-  labs(x = "Consonant similarity (SD)",
-       y = "P(Correct)", colour = "PTHN",
-       fill = "PTHN/Vowel similarity") +
+  labs(x = "Phon. similarity", y = "P(Correct)", colour = "PTHN", fill = "PTHN") +
+  scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
   guides(colour = guide_legend(ncol = 2)) +
   theme_custom() +
-  scale_y_continuous(labels = scales::percent) +
   theme(
@@
   ) +
-  ggsave(here("Figures", "responses_consonant.png"))
+  ggsave(here("Figures", "responses_lv.png"))
 
 ```
-
-##### {-}
 

### 2021-05-28 Rmd/compare_jtrace_translation.Rmd
Add analysis code
Added code to run model predictions for jTRACE output, and code to integrate word stats (lv, freq, pthn)

@@
 library(ggplot2)
+library(janitor) # for cleaning colnames
+library(stringdist)
+library(brms) # for fitting Bayesian models
+library(tidybayes) # for processing Bayesian models
+library(here)
 
+# set parameters
+set.seed(888) # for reproducibility
+options(mc.cores = 4, chains = 4, iter = 500, seed = 888)
+source(here("R", "utils.R"))
 ```
@@
   rownames(temp) <- 1:nrow(temp)
-  names(temp)[1] <- "output" # (rename col) output item
-  names(temp)[2] <- "activation" # (rename col) max activation value
+  names(temp)[1] <- "trace_eng" # (rename col) output item
+  names(temp)[2] <- "trace_activation" # (rename col) max activation value
 
   temp <- temp %>%
-    filter(!output == "cycle") %>% # remove row corresponding to cycle number
+    filter(!trace_eng == "cycle") %>% # remove row corresponding to cycle number
     mutate(input = fnames[i]) %>% # specify input word
-    arrange(desc(activation)) %>%
-    filter(!activation <= 0) # remove low activation (maybe even apply a higher threshold? TBD)
+    arrange(desc(trace_activation)) %>%
+    filter(!trace_activation <= 0) # remove low activation (maybe even apply a higher threshold? TBD)
 
@@
 
+# LEXICON / OUTPUT
 # xlsx file with 1 column for TRACE transcriptions and 1 column for orthography (English words) 
-df_transcription <- read_excel("trace_transcription.xlsx") %>%
+df_lexicon <- read_excel("trace_transcription.xlsx", sheet = "lexicon") %>%
   select(trace_eng, ort_eng) %>%
@@
 # Match the jTRACE output transcriptions with their orthography
-trace_output <- left_join(trace_output, df_transcription, by = c("output" = "trace_eng")) %>%
+trace_output_phon <- left_join(trace_output, df_lexicon, by = c("trace_eng")) %>%
   rename(output_word = ort_eng)
 
+# INPUT -------------------------------------
+# jTRACE transcriptions of input
+df_trial_phon <- read_excel("trace_transcription.xlsx", sheet = "trials") %>%
+  select(ort_presented, trace_presented, language) %>%
+  mutate(language = tolower(language)) %>%
+  unique()
+
+trace_output_phon <- left_join(trace_output_phon, df_trial_phon, by = c("input" = "ort_presented", "language"))
+
 ```
+
+```{r load clearpond}
+
+df_clearpond <- read_excel("trace_transcription.xlsx", sheet = "clearpond") %>%
+  select(Word, Freq_per_million, PTHN) %>%
+  clean_names() %>%
+  rename(frequency = freq_per_million) %>%
+  unique() %>%
+  mutate(word = tolower(word))
+
+trace_output_stats <- left_join(trace_output_phon, df_clearpond, by = c("output_word" = "word")) %>%
+  mutate(n_char = ifelse(nchar(trace_eng) > nchar(trace_presented), nchar(trace_eng), nchar(trace_presented)),
+    lv = stringsim(trace_eng, trace_presented, method = "lv"))
+
+```
+
+#### Model predictions for jTRACE output
+
+```{r run models jtrace}
+
+responses <- trace_output_stats %>% 
+    # transform relative frequency to Zipf score
+    mutate(frequency_zipf = relative_to_zipf(frequency)) %>% 
+    # center predictors
+    mutate_at(
+        vars(lv, pthn, frequency),
+        function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
+    complete() %>% 
+    as_tibble() %>% 
+    drop_na(lv, pthn, frequency_zipf)
+
+# fit models ----
+# formula
+f_0 <- bf(trace_activation ~ 1 + frequency, family = bernoulli(link = "logit"))
+f_1 <- bf(trace_activation ~ 1+ frequency + pthn, family = bernoulli(link = "logit"))
+f_2 <- bf(trace_activation ~ 1 + frequency + pthn*lv, family = bernoulli(link = "logit"))
+
+# prior
+priors <- c(
+    prior(normal(0, 3), class = "Intercept"),
+    prior(normal(0, 3), clas = "b"),
+    prior(cauchy(0, 3), class= "sd"),
+    prior(lkj(5), class = "cor")
+)
+
+# fit models
+fit_0 <- brm(f_0, data = responses, prior = priors[c(1, 3),], backend = "cmdstanr",
+             file = here("Results", "fit_responses_jtrace_0.rds"))
+fit_1 <- brm(f_1, data = responses, prior = priors, backend = "cmdstanr",
+             file = here("Results", "fit_responses_jtrace_1.rds"))
+fit_2 <- brm(f_2, data = responses, prior = priors, backend = "cmdstanr",
+             file = here("Results", "fit_responses_jtrace_2.rds"))
+
+# model comparison
+# model fit is the best
+loos <- list(fit_0 = loo(fit_0), fit_1 = loo(fit_1), fit_2 = loo(fit_2))
+loos_comp <- loo_compare(loos$fit_0, loos$fit_1, loos$fit_2)
+
+```
+
+## Comparing against Translation Task
 
@@
 
-```{r plot, warning = FALSE}
+```{r match data, warning = FALSE}
 
@@
 
-# If "output" is <NA>, the word is not in the currently-used jTRACE lexicon
+# If "trace_activation" is <NA>, the word is not in the currently-used jTRACE lexicon
 df_match %>%
@@
   slice(1) %>%
-  select(input, output_word, pct_participant, output, activation) %>%
+  select(input, output_word, pct_participant, trace_activation) %>%
   arrange(desc(pct_participant))
@@
 # if no data for a given output word is available in trace output, it means that activation is outside of Top 100 words and/or below 0
-df_match$activation[is.na(df_match$activation)] <- 0
+df_match$trace_activation[is.na(df_match$trace_activation)] <- 0
 
@@
 
+```
+
+### Plot relationship
+
+```{r plot, warning = FALSE}
+
 # plot
-ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
+ggplot(df_match_trace, aes(x = trace_activation, y = pct_participant)) +
   geom_smooth() +
@@
 # correlation test
-cor.test(df_match_trace$activation, df_match_trace$pct_participant)
+cor.test(df_match_trace$trace_activation, df_match_trace$pct_participant)
 
 # plot by word
-ggplot(df_match_trace, aes(x = activation, y = pct_participant)) +
+ggplot(df_match_trace, aes(x = trace_activation, y = pct_participant)) +
   geom_smooth() +

### 2021-07-01 Rmd/00_outline.Rmd
Rename files, include job, add index.html

---
title: "Outline"
date: "Updated: `r format(Sys.Date(), '%d/%m%/%Y')`"
output:
    pdf_document:
        latex_engine: xelatex
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The impact of cognateness in bilingual lexical acess

### Comprehension

### Production

### Translation

* [@dijkstra2010how]

### Phonological neighbourhoods and parallel activation

### Limitations of current measures of cognateness

* Levenshtein distance
* Coltheart N

## The present study

### 2021-07-01 Rmd/02_methods.Rmd
Rename files, include job, add index.html

--- 2020-08-06 Manuscript/te_manuscript.Rmd
@@
 ---
-title             : "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
-shorttitle        : "Phoneme-weighted Levenshtein Distance"
-
-author: 
-  - name          : "First Author"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
-    email         : "my@email.com"
-  - name          : "Second Author"
-    affiliation   : "1,2"
-
-authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
-
-abstract: |
-  Many studies in bilingualism have explored how form similarity between translation equivalents (i.e.
-cognate- ness) impacts language processing.
-Cognates range from being identical to sharing only a few key features.
-Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity.
-In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value.
-A widely used metric for word similarity is Levenshtein distance [@levenshtein1966binary].
-This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other.
-However, a challenge when defining phonological similarity is that some phonemes (e.g.
-[k] and [g]) are perceptually more similar than others (e.g.
-[k] and [o]).
-The standard Lev- enshtein distance calculation treats each change equally.
-Thus it often overestimates phonological distance between words.
-We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g.
-vowel or consonant insertion) and position in the word.
-To estimate the weights, we are col- lecting behavioural data in an auditory translation elicitation task.
-Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words.
-The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent.
-Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation.
-The higher the probability, the higher the perceptual similarity.
-Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "keywords"
-wordcount         : "X"
-
-bibliography      : ["r-references.bib"]
-
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  papaja::apa6_pdf:
-    latex_engine: "xelatex"
+title: "Methods"
+date: "Updated: `r format(Sys.Date(), '%d/%m%/%Y')`"
+output:
+    pdf_document:
+        latex_engine: xelatex
+bibliography: "references.bib"
 ---
 
-```{r setup, include = FALSE}
-library(papaja)
-```
+```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
 
-```{r analysis-preferences}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(
+	out.width = "100%",
+	message = FALSE,
+	warning = FALSE
+)
+
 ```
@@
 
-library(dplyr)
-library(tidyr)
-library(tibble)
-library(stringr)
-library(ggplot2)
+library(tidyverse)
+library(papaja)
+library(gt)
+library(knitr)
+library(lubridate)
 library(here)
-library(data.table)
-library(tidybayes)
-library(gt)
-library(lubridate)
 
-participants <- fread(here("Data", "01_participants.csv"), na.strings = "") %>% as_tibble() %>% mutate(date = as_date(date))
-dat <- fread(here("Data", "01_processed_coded.csv"), na.strings = "") %>% as_tibble()
-posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()
-predictions <- fread(here("Results", "posterior_predictions.csv"), na.strings = "") %>% as_tibble()
-fit <- readRDS(here("Results", "fit3_phon.rds"))
-comparison <- readRDS(here("Results", "model_comparison.rds"))
-item_effects <- fread(here("Results", "item_effects.csv"), na.strings = "") %>% as_tibble()
-group_effects <- fread(here("Results", "group_effects.csv"), na.strings = "") %>% as_tibble()
-corrs <- fread(here("Results", "corrs.csv"), na.strings = "") %>% as_tibble()
+participants <- readRDS(here("Data", "participants.rds"))
+stimuli <- readRDS(here("Data", "stimuli.rds"))
+responses <- readRDS(here("Data", "responses.rds"))
 
@@
 
-# Methods
-
-## Algorithm
-
-## Translation Elicitation task
-
-### Participants
+## Participants
 
 Data collection took place from `r format(min(participants$date), "%dth %B, %Y")` to  `r format(max(participants$date), "%dth %B, %Y")`.
-We collected data from `r nrow(participants)` participants ($M_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(sd(participants$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
+We collected data from `r nrow(participants)` participants ($M_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
 `r filter(participants, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(participants, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(participants, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(participants, country=="Spain", sex =="Female") %>% nrow()` female).
@@
 We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION?
-(Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No)".
-However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`).
+(Yes/No)" in English, and as *¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
+(Sí/No)*.
+However, the proportion of Spanish participants that reported visual impairment was implausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
+This is possibly due to these participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`).
 
-### Procedure
+## Procedure
 
@@
 Third, participants completed the experimental task.
-Before the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
 They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
 English participants were randomly assigned to the list of Catalan or Spanish trials.
-Spanish participants completed the list of Catalan trials.
-
-### Design
+Participants in the Catalan list were presented with `r count(stimuli, group)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, group)$n[2]` trials.
 
 Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio offset.
+After one second, the audio started playing while the dot remained being displayed until the audio ended.
 Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
@@
 We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Typos were allowed, however.
-Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
+Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
+Participants contributed a total of `r distinct(responses, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(responses, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(responses, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
 The task took approximately 15 minutes to be completed.
 
-### Stimuli
+```{r design, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+include_graphics(here("Figures", "design.png"))
+
+```
+
+## Stimuli
+
+Participants listened to one audio file in each trial.
+This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@boersma2018].
+
+Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT [@boada2013] and SUBTLEX-ESP [@cuetos2011], respectively, and are expressed as Zipf scores [@van-heuven2012; @zipf1949].The average lexical frequency (counts per million) of the words was `r printnum(mean(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`).
+The average lexical frequency of the Catalan words was `r printnum(mean(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`), and the average lexical frequency of the Spanish words was `r printnum(mean(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`).
 
@@
 To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using Paretho-Smoothed Importance Sampling (PSIS-LOO, @vehtari2016).
+We compared models using leave-one-out cross-validation (LOO, @vehtari2016).
 More information about the models and model comparison can be found in Appendix X.
@@
 All analyses were performed in R environment [@rcoreteam2019].
-We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] to impute missing data, and the `brms` R package to [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
-
-# Results
-
-## Posterior draws
-
-```{r posterior_draws}
-
-posterior %>%
-    ggplot(aes(.value, .variable)) +
-    geom_vline(xintercept = 0, linetype = "dashed") +
-    stat_slab(fill = "black") + 
-    stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.15)) +
-    labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
-         title = "Posterior distribution of coefficients of fixed effects") +
-    scale_colour_brewer(palette = "Blues") +
-    theme_classic() +
-    theme(panel.grid.major.y = element_line(colour = "grey"),
-          legend.position = "top",
-          legend.key = element_rect(fill = "transparent")) +
-    ggsave(here("Figures", "coefs.png"))
-
-```
-
-## Posterior predictions
-
-```{r posterior_predictions}
-
-predictions %>%
-    ggplot(aes(similarity_phon, .value, colour = same_onset, fill = same_onset)) +
-    facet_grid(consecutive_longest~pthn)+
-    stat_summary(fun = mean, geom = "line", size = 1) +
-    geom_vline(xintercept = 0, colour= "grey") +
-    geom_hline(yintercept = 0.5, colour= "grey") +
-   labs(x = "Phonological similarity", y = "Probability of correct translation",
-         fill = "Credible interval", colour = "Onset",
-         title = "Effect of phonological similarity on probability of correct translation",
-         subtitle = "Model predictions are displayed for different quartiles of PTHN") +
-    scale_fill_brewer(palette = "Set1") +
-    scale_colour_brewer(palette = "Set1") +
-    theme_classic() +
-    theme(legend.position = "top") +
-    ggsave(here("Figures", "predictions.png"), height = 4)
-
-```
-
-## Random effects
-
-### Item effects
-
-```{r item_effects}
-
-item_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  ggplot(aes(estimate, fill = param)) +
-  facet_wrap(~param, scales = "free") +
-  geom_histogram(bins = 20) +
-  geom_rug(alpha = 0.1) +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  labs(x = "Estimate", y = "Density", fill = "Parameter") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_classic() +
-  theme(legend.position = "none",
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
-
-```
-
-### Group effects
-
-```{r goup_effects_table}
-
-group_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(hdi = paste0("[", lower, ", ", upper, "]")) %>% 
-  select(param, group, error, hdi) %>% 
-  group_by(param) %>% 
-  gt() %>% 
-  cols_label(group = "Group", error = "SE", hdi = "95% HDI")
-
-```
-
-```{r group_effects}
-
-group_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  ggplot(aes(group, estimate)) +
-  facet_wrap(~param) +
-  geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = lower, ymax = upper, colour =  group), width = 0, size = 5) +
-  geom_errorbar(aes(ymin = estimate-error, ymax = estimate+error), width = 0, size = 1, alpha = 0.5) +
-  geom_point(size = 3) +
-  labs(x = "Group", y = "Estimate", fill = "Group") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_classic() +
-  theme(legend.position = "none",
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
-
-```
-### Correlations
-
-```{r correlations}
-
-corrs %>% 
-  filter(str_detect(variable, "trial_id")) %>%
-  mutate(variable = str_remove(variable, "cor_group:trial_id__"),
-         variable = str_replace(variable, "__" , " - ")) %>% 
-  ggplot(aes(value, variable, xmin = lower, xmax = upper)) +
-  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
-  geom_errorbarh(show.legend = FALSE, height = 0) +
-  geom_point(size = 2, show.legend = FALSE) +
-  labs(x = "Estimate", y = "Correlation") +
-  theme_classic() +
-  theme(axis.title.y = element_blank())
-
-```
-
-# Discussion
-
-\newpage
-
-# References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup
+We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].

### 2021-07-01 Rmd/report.Rmd
Rename files, include job, add index.html

@@
     toc: true
-    toc_depth: 3
+    toc_depth: 4
     toc_float: true
+bibliography: 
 ---
@@
 # import data
-trials <- readRDS(here("Data", "trials.rds"))
+stimuli <- readRDS(here("Data", "stimuli.rds"))
 participants <- readRDS(here("Data", "participants.rds"))
@@
 
-```{r participants_trials, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r participants_stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
 
-Participants in the ENG-SPA listened to `r nrow(trials[trials$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(trials[trials$group=="ENG-CAT",])` trials.
+Participants in the ENG-SPA listened to `r nrow(stimuli[stimuli$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials.
 Five trials in each condition were practice trials and were excluded from analyses.
 
-```{r trials_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
-trials_coded <- trials %>% 
+```{r stimuli_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
+stimuli_coded <- stimuli %>% 
   mutate(
@@
 
-```{r trials_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials_coded %>% 
+```{r stimuli_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
   ggplot(aes(zipf, fill = test_language)) +
@@
 
-```{r trials_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials_coded %>% 
+```{r stimuli_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
   count(test_language, pthn) %>% 
@@
 
-```{r trials_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials_coded %>% 
+```{r stimuli_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
   ggplot(aes(group, lv, fill = group, colour = group)) +
@@
 
-  trials_coded %>% 
+  stimuli_coded %>% 
   select(trial_id, group, ends_with("_ratio")) %>% 
@@
 
-```{r trials_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials_coded %>% 
+```{r stimuli_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
   select(trial_id, group, onset) %>% 
@@
 
-```{r trials_similarity_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-trials_coded %>% 
+```{r stimuli_similarity_onset, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
   select(trial_id, group, lv, onset) %>% 
@@
 
-m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
   mutate_at(vars(lv), function(x) paste0(x, " SD"))
@@
   stat_summary(fun = mean, geom = "line", size = 1) +
-  labs(x = "PTHN", y = "P(Correct)", colour = "Phon. similarity", fill = "Phon. similarity") +
+  labs(x = "Phon. neigh. dens. (SD)", y = "P(Correct)",
+       colour = "Phon. similarity", fill = "Phon. similarity") +
   scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
@@
 
-m <- add_fitted_draws(nd, fit_responses, n = 20, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
   mutate_at(vars(pthn), function(x) paste0(x, " SD"))
@@
   stat_summary(fun = mean, geom = "line", size = 1) +
-  labs(x = "Phon. similarity", y = "P(Correct)", colour = "PTHN", fill = "PTHN") +
+  labs(x = "Phon. similarity (SD)", y = "P(Correct)", colour = "Phon. neigh. dens.",
+       fill = "Phon. neigh. dens.") +
   scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
@@
   ) +
-  ggsave(here("Figures", "responses_lv.png"))
+  ggsave(here("Figures", "responses_lv.png"), height = 4, width = 4.5)
 
@@
     footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "consonant_ratio", "vowel_ratio", "onset1", "overlap_stress1") | str_detect(term, "activation"))
+    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "lv", "pthn:lv") | str_detect(term, "activation"))
   ) %>% 
@@
   facet_wrap(~.variable, scales = "free") +
-  geom_vline(xintercept = 0, colour = "red") +
+  geom_vline(xintercept = 0, colour = "black") +
   stat_halfeye() +
@@
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-    axis.text = element_text(size = 7)
-  )
-```
-
-### Random coefficients
-
-```{r jtrace_post_re, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-re <- ranef(fit_jtrace$fit)$`group:input`
-re <- rbind(Intercept = re[,,1], activation = re[,,2]) %>% 
-  as.data.frame() %>% 
-  rownames_to_column("term") %>% 
-  mutate(
-    term = str_replace(str_remove(term, ".1"), "\\.", "\\-"),
-    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
-    Q2.5 = ifelse(term=="Intercept", inv_logit_scaled(Q2.5), Q2.5/4),
-    Q97.5 = ifelse(term=="Intercept", inv_logit_scaled(Q97.5), Q97.5/4),
-    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
-  ) %>% 
-  separate(term, c("group", "input"), sep = "_")
-re$.param <- rep(c("Intercept", "activation"), each = 9)
-
-ggplot(re, aes(Estimate, input, colour = input)) +
-  facet_wrap(~.param) +
-  geom_vline(xintercept = 0, colour = "red") +
-  geom_errorbarh(aes(xmin = Estimate-Est.Error, xmax = Estimate+Est.Error), height = 0, size = 5, alpha = 0.5) +
-  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), height = 0, size = 1) +
-  geom_point() +
-  labs(x = "Estimate", y = "Input", colour = "Input") +
-  theme_custom() +
-  theme(
-    legend.position = "top",
-    legend.direction = "horizontal",
-    axis.title.x = element_blank(),
-    legend.title = element_blank(),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-    axis.text.y = element_blank(),
-    axis.ticks = element_blank()
+    axis.text = element_text(size = 9)
   )
@@
 
-trials_trace <- trials %>% 
+stimuli_trace <- stimuli %>% 
   as_tibble() %>% 
@@
   # add experimental trials
-  left_join(trials_trace) %>% 
+  left_join(stimuli_trace) %>% 
   mutate(correct = response==phon_trace) %>% 
@@
 ggplot(d_jtrace, aes(response_reordered, activation, fill = correct)) +
-  facet_wrap(input_label~source, scales = "free_y", ncol = 2) +
+  facet_wrap(input_label~source, scales = "free_y", ncol = 3) +
   geom_col(colour = "white") +

### 2021-07-01 Rmd/te_manuscript.Rmd
Rename files, include job, add index.html

@@
 ---
-title             : "A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity"
-shorttitle        : "Phoneme-weighted Levenshtein Distance"
+title             : "The role of cognateness in non-native spoken word recognition"
+shorttitle        : "Cognateness and non-native word recognition"
 
 author: 
-  - name          : "First Author"
+  - name          : "Gonzalo Garcia-Castro"
     affiliation   : "1"
     corresponding : yes    # Define only one corresponding author
-    address       : "Postal address"
-    email         : "my@email.com"
-  - name          : "Second Author"
-    affiliation   : "1,2"
+    address       : "Ramon Trial Fargas, 25-27, 08005 Barcelona"
+    email         : "gonzalo.garciadecastro@upf.edu"
+  - name          : "Serene Siow"
+    affiliation   : "2"
+  - name          : "Nuria Sebastian-Galles"
+    affiliation   : "1"
+  - name          : "Kim Plunkett"
+    affiliation   : "2"
 
 authornote: |
-  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
-
-  Enter author note here.
+  GG and SS contributed equally and share first authorship.
 
 abstract: |
-  Many studies in bilingualism have explored how form similarity between translation equivalents (i.e.
-cognate- ness) impacts language processing.
-Cognates range from being identical to sharing only a few key features.
-Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity.
-In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value.
-A widely used metric for word similarity is Levenshtein distance [@levenshtein1966binary].
-This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other.
-However, a challenge when defining phonological similarity is that some phonemes (e.g.
-[k] and [g]) are perceptually more similar than others (e.g.
-[k] and [o]).
-The standard Lev- enshtein distance calculation treats each change equally.
-Thus it often overestimates phonological distance between words.
-We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g.
-vowel or consonant insertion) and position in the word.
-To estimate the weights, we are col- lecting behavioural data in an auditory translation elicitation task.
-Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words.
-The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent.
-Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation.
-The higher the probability, the higher the perceptual similarity.
-Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.
+  There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
+For example, bilinguals produce and translated cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
+Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
+These measures partially ignore potential sources of variability during lexical access like vowel vs.
+consonant overlap, overlap in stressed syllables, onset overlap or the distance in features between both translations.
+In this study, we explored the impact of some of these variables on non-native word translation task: Spanish and English participants listened to non-native words (Catalan or Spanish) and were prompted to type their translation in their native language.
+Critically, participants where unfamiliar with the testing language, ensuring that they were not able to translate words based on previous knowledge on their meaning, leaving phonological information as the only cue participants were able to exploit to translate words to their native language correctly.
+We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.
 
@@
 
-keywords          : "keywords"
+keywords          : "cognate, word recognition, translation, non-native, spoken word recognition"
 wordcount         : "X"
 
-bibliography      : ["r-references.bib"]
+bibliography      : ["r-references.bib", "references.bib"]
 
@@
 
-```{r analysis-preferences}
+```{r analysis-preferences, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 # Seed for random number generation
@@
 
-participants <- readRDS(here("Results", "participants.rds"))
-stimuli <- readRDS(here("Results", "stimuli.rds"))
-accuracy <- readRDS(here("Results", "accuracy.rds"))
-fit <- readRDS(here("Results", "fit3_phon.rds"))
-comparison <- readRDS(here("Results", "model_comparison.rds"))
+participants <- readRDS(here("Data", "participants.rds"))
+stimuli <- readRDS(here("Data", "stimuli.rds"))
+responses <- readRDS(here("Data", "responses.rds"))
 
@@
 English participants were randomly assigned to the list of Catalan or Spanish trials.
-Participants in the Catalan list were presented with `r count(stimuli, language)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, language)$n[2]` trials.
+Participants in the Catalan list were presented with `r count(stimuli, group)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, group)$n[2]` trials.
 
@@
 Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
+Participants contributed a total of `r distinct(responses, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(responses, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(responses, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
-Audio duration, intensity and pitch were extracted using the `PraatR` R package [@albin2014].
 

### 2021-07-01 R/00_stimuli.R

@@ -1,0 +1,75 @@
+# trial information
+
+# set up ----
+
+# load packages
+library(tidyverse)
+library(readxl) # to import  Excel spreadsheets
+library(janitor) # to clean column names
+library(stringdist) # to compute Levenshtein distance
+library(here) # to locate files
+
+# load/create functions
+source(here("R", "utils.R"))
+
+# set parameters
+individual_plots <- FALSE
+
+# clearpond ----
+# relative frequencies and phonological neighbours
+clearpond <- map(
+  list(
+    `ENG-CAT` = here("Data", "clearpond", "clearpond_english.csv"),
+    `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"),
+    `SPA-CAT` = here("Data", "clearpond", "clearpond_spanish.csv")
+  ),
+  function(x){
+    read.csv(x) %>% 
+      clean_names() %>% 
+      select("word", "pho_word", "freq_per_million", "pthn")
+  }
+) %>% 
+  bind_rows(.id = "group") %>% 
+  as_tibble() %>%
+  distinct(word, group, .keep_all = TRUE) %>% 
+  rename(word2 = word, phon_clearpond = pho_word, frequency = freq_per_million)
+
+# trace ----
+# jTRACE transcriptions
+trace <- read_xlsx(here("Stimuli", "trace.xlsx")) %>% 
+  distinct(ort_eng, .keep_all = TRUE) %>% 
+  rename(word2 = ort_eng, phon_trace = trace_eng)
+
+# levenshtein distance ----
+lv <- map(
+  c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
+  function(x) read_xlsx(here("Stimuli", "trials.xlsx"), sheet = x)
+) %>% 
+  bind_rows(.id = "group") %>% 
+  clean_names() %>% 
+  select(trial_id, group, ipa1, ipa2) %>% 
+  mutate(
+    n_char = ifelse(nchar(ipa1) > nchar(ipa2), nchar(ipa1), nchar(ipa2)),
+    lv = stringsim(ipa1, ipa2, method = "lv")
+  )
+
+
+# trials ----
+stimuli <- map(
+  c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
[…]

### 2021-07-01 R/03_trace.R

@@ -1,0 +1,344 @@
+# jtrace analyses
+
+# set up ----
+library(tidyverse)
+library(readxl) # for importing Excel spreadsheets
+library(patchwork) # for arranging plots
+library(stringdist) # for Levenshtein distance
+library(brms) # for fitting Bayesian models
+library(tidybayes) # for manipulating MCMC draws
+library(tidytext) # for reordering levels in ggplot2
+library(janitor) # for cleaning column names
+library(here) # for locating files
+
+# set parameters
+set.seed(888) # for reproducibility
+source(here("R", "utils.R")) # custom functions
+
+#### import data ----
+# import trials
+trials <- readRDS(here("Data", "stimuli.rds")) %>% 
+    mutate(
+        test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"),
+        zipf = log10(frequency)+3,
+        pthn_log = log(pthn),
+        overlap_stress = ifelse(overlap_stress==1, "Same", "Different")
+    )
+# import participant responses
+responses <- readRDS(here("Data", "responses.rds"))
+
+# by Serene (some fixes by Gonzalo) ----
+
+# TRACE simulations ----
+
+# import TRACE simulations
+filedir <- here("Data", "jtrace")
+## 1 file per input word, with 1 output word per column
+## files should be named with input word orthography and language (e.g. "gato_spanish.xlsx")
+## files should be xlsx (csv doesn't deal well with special characters like ^ and @)
+## folder should not contain any other files
+
+# Note: it's effortful to have to convert to xlsx (jTRACE doesn't allow direct xlsx export), but it's the most reliable method to preserve special characters I've found so far. Would welcome ideas
+
+# Export jTRACE output with Top 100 items and 100 cycles (arbitrary value)
+# Max ad-hoc activation
+
+fnames <- dir(filedir) # extract list of file names in directory
+trace_output <- data.frame() # create empty data frame
+# for loop to extract highest activations for each input word
+for (i in 1:length(fnames)){
+    filename <- paste(filedir, fnames[i], sep = "/") # get file name
+    temp <- read_excel(filename) # read file
+    
+    temp <- apply(temp,2,max) # get highest value in each column
+    temp <- as.data.frame(temp)
[…]

### 2021-07-01 R/processing.R

@@ -1,188 +1,0 @@
-# preprocessing
-
-# set up ----
-
-# load packages
-library(tidyverse) # for manipulating variables
-library(readxl) # for importing Excel spreadsheets
-library(janitor) # for cleaning variable names
-library(lubridate) # for working with dates
-library(data.table) # for importing data
-library(here) # for locating files
-
-# create/load functions
-source(here("R", "utils.R")) # helper functions
-
-# set parameters
-set.seed(888) # for reproducibility
-practice_trials <- c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)
-spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
-
-# import trial data ----
-
-# import trials
-trials <- readRDS(here("Data", "trials.rds"))
-
-# process data ----
-# participant files
-filenames <- list.files(here("Data", "Raw")) %>% 
-    str_extract(".*?\\_") %>% 
-    str_remove("_")
-
-processed <- list.files(here("Data", "Raw"), full.names = TRUE) %>% 
-    map(fread, na.string = "") %>%  # import participant files
-    map(mutate, participant = as.character(participant)) %>% 
-    set_names(filenames) %>% # label each dataset with the participant's ID
-    bind_rows(.id = "filename") %>% # merge datasets
-    as_tibble() %>%
-    clean_names() %>% 
-    # select relevant variables and rename if necessary
-    select(
-        participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
-        matches("language_l|language_s|language_c|demo_|setup_"),
-        -matches("_rt|_time"),
-        key_press_time
-    ) %>% 
-    rename(demo_impairment = demo_language_key_keys) %>% 
-    rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
-    # clean text input by participants (because of typos of need to translate) and redefine location    
-    group_by(participant) %>% 
-    mutate_at(
-        vars(
-            starts_with("demo_"),
-            starts_with("l"),
-            starts_with("setup_"),
-            age, city, setup_location, setup_noise
-        ),
-        first_non_na) %>% 
-    mutate(date = max(date, na.rm = TRUE)) %>% 
-    ungroup() %>% 
[…]

### 2021-07-01 Rmd/report.Rmd
Add new README.md

@@
 output:
+  md_document:
+    variant: markdown_github
   html_document:
@@
     toc_float: true
-bibliography: 
+bibliography: "references.bib"
 ---
@@
 fit_jtrace <- readRDS(here("Results", "fit_jtrace.rds"))
-fit_responses <- readRDS(here("Results", "fit_responses_2.rds"))
+fit_responses <- readRDS(here("Results", "fit_responses_4.rds"))
 loo <- readRDS(here("Results", "loo.rds"))
@@
   mutate(formula = c(
-    "~ 1 + frequency + pthn*lv + (1 + pthn*lv | participant)",
+    "~ 1 + frequency + (1 | participant)",
     "~ 1 + frequency + pthn + (1 + pthn | participant)",
-    "~ 1 + frequency + (1 | participant)"
+    "~ 1 + frequency + pthn*vowel_ratio + (1 + pthn*vowel_ratio | participant)",
+    "~ 1 + frequency + pthn*consonant_ratio + (1 + pthn*consonant_ratio  | participant)",
+    "~ 1 + frequency + pthn*(consonant_ratio + consonant_ratio) + (1 + pthn*(consonant_ratio + consonant_ratio) | participant)"
   )) %>% 
@@
 nd <- expand_grid(
-  pthn = seq(min(responses_coded$pthn, na.rm = TRUE), max(responses_coded$pthn, na.rm = TRUE), by = 0.1),
+  pthn = seq(min(responses_coded$pthn, na.rm = TRUE),
+             max(responses_coded$pthn, na.rm = TRUE), by = 0.1),
   frequency = 0,
-  lv = c(-1, 1)
+  consonant_ratio = c(-1, 1),
+  vowel_ratio = c(-1, 1)
 )
@@
 m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
-  mutate_at(vars(lv), function(x) paste0(x, " SD"))
+  mutate_at(vars(consonant_ratio, vowel_ratio), function(x) paste0(x, " SD"))
 
-ggplot(m, aes(pthn, .value, colour = lv, fill = lv)) +
+ggplot(m, aes(pthn, .value, colour = interaction(consonant_ratio, vowel_ratio, sep = "/"), fill = interaction(consonant_ratio, vowel_ratio, sep = "/"))) +
   stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
@@
   labs(x = "Phon. neigh. dens. (SD)", y = "P(Correct)",
-       colour = "Phon. similarity", fill = "Phon. similarity") +
+       colour = "Consonant/Vowel similarity", fill = "Consonant/Vowel similarity") +
   scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
@@
 
-```{r emmeans_lv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=6}
-# by lv
+```{r emmeans_consonant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=6}
+# by consonant_ratio
 nd <- expand_grid(
@@
   frequency = 0,
-  lv = seq(
-    min(responses_coded$lv, na.rm = TRUE),
-    max(responses_coded$lv, na.rm = TRUE),
+  consonant_ratio = seq(
+    min(responses_coded$consonant_ratio, na.rm = TRUE),
+    max(responses_coded$consonant_ratio, na.rm = TRUE),
     by = 0.1
-  )
+  ),
+  vowel_ratio = c(-1, 1)
 )
@@
 m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
-  mutate_at(vars(pthn), function(x) paste0(x, " SD"))
+  mutate_at(vars(pthn, vowel_ratio), function(x) paste0(x, " SD"))
 
-ggplot(m, aes(lv, .value, colour = pthn, fill = pthn)) +
+ggplot(m, aes(consonant_ratio, .value, colour = interaction(vowel_ratio, pthn, sep = "/"), fill = interaction(vowel_ratio, pthn, sep = "/"))) +
   stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
   stat_summary(fun = mean, geom = "line", size = 1) +
-  labs(x = "Phon. similarity (SD)", y = "P(Correct)", colour = "Phon. neigh. dens.",
-       fill = "Phon. neigh. dens.") +
+  labs(x = "Consonant similarity", y = "P(Correct)", colour = "PTHN/Vowel similarity",
+       fill = "PTHN/Vowel similarity") +
   scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
@@
   ) +
-  ggsave(here("Figures", "responses_lv.png"), height = 4, width = 4.5)
+  ggsave(here("Figures", "responses_consonant.png"), height = 4, width = 4.5)
+
+```
+
+```{r emmeans_vowel, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=6}
+# by vowel_ratio
+nd <- expand_grid(
+  pthn = c(-1, 1),
+  frequency = 0,
+  vowel_ratio = seq(
+    min(responses_coded$vowel_ratio, na.rm = TRUE),
+    max(responses_coded$vowel_ratio, na.rm = TRUE),
+    by = 0.1
+  ),
+  consonant_ratio = c(-1, 1)
+)
+
+m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
+  mutate_at(vars(pthn, consonant_ratio), function(x) paste0(x, " SD"))
+
+ggplot(m, aes(vowel_ratio, .value, colour = interaction(consonant_ratio, pthn, sep = "/"), fill = interaction(consonant_ratio, pthn, sep = "/"))) +
+  stat_lineribbon(.width = 0.95, alpha = 0.25, size = 0) +
+  stat_summary(fun = mean, geom = "line", size = 1) +
+  labs(x = "Vowel similarity", y = "P(Correct)", colour = "PTHN/Consonant similarity",
+       fill = "PTHN/Consonant similarity") +
+  scale_y_continuous(labels = scales::percent, limits= c(0, 1)) +
+  guides(colour = guide_legend(ncol = 2)) +
+  theme_custom() +
+  theme(
+    legend.position = "top",
+    panel.grid = element_line(colour = "grey", linetype = "dotted")
+  ) +
+  ggsave(here("Figures", "responses_vowel.png"), height = 4, width = 4.5)
 

### 2021-09-29 Rmd/Old/02_methods.Rmd
Remove big files

--- 2021-07-01 Rmd/02_methods.Rmd
@@
 ```
-
-## Participants
-
-Data collection took place from `r format(min(participants$date), "%dth %B, %Y")` to  `r format(max(participants$date), "%dth %B, %Y")`.
-We collected data from `r nrow(participants)` participants ($M_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
-`r filter(participants, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(participants, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(participants, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(participants, country=="Spain", sex =="Female") %>% nrow()` female).
-Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
-Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants were asked to complete the experiment in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION?
-(Yes/No)" in English, and as *¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No)*.
-However, the proportion of Spanish participants that reported visual impairment was implausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
-This is possibly due to these participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`).
-
-## Procedure
-
-The experiment was implemented online using Psychopy/Pavlovia [@peirce2019].
-Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-First, participants were informed about the aims of the study and gave informed consent for participating.
-Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
-Third, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
-They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the list of Catalan or Spanish trials.
-Participants in the Catalan list were presented with `r count(stimuli, group)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, group)$n[2]` trials.
-
-Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio ended.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
-Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
-Participants were allowed to correct their answer.
-Then, participants pressed the RETURN key to start and new trial.
-We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Participants contributed a total of `r distinct(responses, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(responses, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(responses, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
-The task took approximately 15 minutes to be completed.
-
-```{r design, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-include_graphics(here("Figures", "design.png"))
-
-```
-
-## Stimuli
-
-Participants listened to one audio file in each trial.
-This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@boersma2018].
-
-Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT [@boada2013] and SUBTLEX-ESP [@cuetos2011], respectively, and are expressed as Zipf scores [@van-heuven2012; @zipf1949].The average lexical frequency (counts per million) of the words was `r printnum(mean(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`).
-The average lexical frequency of the Catalan words was `r printnum(mean(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`), and the average lexical frequency of the Spanish words was `r printnum(mean(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`).
-
-## Data analysis
-
-We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT).
-We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
-We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance.
-We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.
-Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020].
-Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013].
-To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (LOO, @vehtari2016).
-More information about the models and model comparison can be found in Appendix X.
-
-All analyses were performed in R environment [@rcoreteam2019].
-We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].

### 2021-09-29 Rmd/report.Rmd
Remove big files

@@
 
-```{r setup, include=FALSE}
+```{r, include=FALSE}
 knitr::opts_chunk$set(
@@
 
-# set params
-source(here("R", "utils.R"))
+# import data
+tar_load_globals()
+tar_load(stimuli)
+tar_load(participants)
+tar_load(responses)
+tar_load(model_fits)
+tar_load(model_loos)
 
-# import data
-stimuli <- readRDS(here("Data", "stimuli.rds"))
-participants <- readRDS(here("Data", "participants.rds"))
-responses <- readRDS(here("Data", "responses.rds"))
-jtrace <- readRDS(here("Data", "jtrace.rds"))
-
-# import models
-fit_jtrace <- readRDS(here("Results", "fit_jtrace.rds"))
-fit_responses <- readRDS(here("Results", "fit_responses_4.rds"))
-loo <- readRDS(here("Results", "loo.rds"))
+theme_set(theme_github())
 
@@
 
-# TRACE simulations
-
-## TRACE vs. human data {.tabset .tab-pill}
-
-### TRACE activation vs. responses
-
-```{r jtrace_all_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-ggplot(jtrace, aes(x = activation, y = pct_participant)) +
-  geom_smooth(colour = wes_palettes$Darjeeling1[1],
-              fill = wes_palettes$Darjeeling1[1],
-              alpha = 0.5) +
-  geom_point(shape = 1, stroke = 1, alpha = 0.5) +
-  labs(
-    x = "jTRACE activation value for the word",
-    y = "Participants who responded with the word"
-  ) +
-  scale_y_continuous(labels = scales::percent) +
-  theme_custom() +
-  theme(
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "trace_average.png"))
-
-```
-
-### By items
-
-```{r jtrace_word_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-# plot by word
-jtrace %>% 
-  mutate(input_language = paste0(input, " (", language, ")")) %>% 
-  ggplot(aes(x = activation, y = pct_participant)) +
-  facet_wrap(~input_language) +
-  geom_smooth(colour = wes_palettes$Darjeeling1[1],
-              fill = wes_palettes$Darjeeling1[1],
-              alpha = 0.5) +
-  geom_point(shape = 1, stroke = 1, alpha = 0.5) +
-  labs(
-    x = "jTRACE activation value for the word",
-    y = "% participants who responded with the word",
-    title = "jTRACE activation against % participants who gave word as response"
-  ) +
-  theme_custom() +
-  theme(
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  )
-
-```
-
-### {-}
-
-## Regression model {.tabset .tab-pill}
-
-### Fixed coefficients
-
-```{r jtrace_fixed_table, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-summary(fit_jtrace$fit)$fixed %>% 
-  as.data.frame() %>%
-  rownames_to_column("term") %>% 
-  mutate(
-    Estimate = ifelse(term=="Intercept", inv_logit_scaled(Estimate), Estimate/4),
-    `l-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`l-95% CI`), `l-95% CI`/4),
-    `u-95% CI` = ifelse(term=="Intercept", inv_logit_scaled(`u-95% CI`), `u-95% CI`/4),
-    Est.Error = ifelse(term=="Intercept", inv_logit_scaled(Est.Error), Est.Error/4)
-  ) %>% 
-  gt() %>% 
-  fmt_number(2:6, decimals = 2) %>% 
-  fmt_number(7:8, decimals = 0) %>% 
-  cols_merge(vars("l-95% CI", "u-95% CI"), pattern = "[{1}, {2}]") %>% 
-  cols_label(
-    term = md("**Predictor**"),
-    Estimate = md("**Mean**"),
-    Est.Error = md("**SEM**"),
-    "l-95% CI" = md("**95\\% CrI**"),
-    Rhat = md("**Rhat**"),
-    Bulk_ESS = md("**Bulk ESS**"),
-    Tail_ESS = md("**Tail ESS**")
-  ) %>% 
-  tab_footnote(
-    footnote = "Transformed using the inverse logit to get the average probability of correct response",
-    locations = cells_body(columns = "term", rows = term=="Intercept")
-  ) %>% 
-  tab_footnote(
-    footnote = "Transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable.",
-    locations = cells_body(columns = "term", rows = term %in% c("pthn", "frequency", "lv", "pthn:lv") | str_detect(term, "activation"))
-  ) %>% 
-  tab_footnote(
-    footnote = "ESS: Effective sample size",
-    locations = cells_column_labels(columns = c("Bulk_ESS", "Tail_ESS"))
-  )
-```
-
-```{r jtrace_fixed_plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%"}
-fit_jtrace$fixed %>%
-  mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4)) %>% 
-  ggplot(aes(.value, fill = .variable)) +
-  facet_wrap(~.variable, scales = "free") +
-  geom_vline(xintercept = 0, colour = "black") +
-  stat_halfeye() +
-  labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
-  theme_custom() +
-  theme(
-    legend.position = "none",
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-    axis.text = element_text(size = 9)
-  )
-```
-
-### {-}
-
-## Candidate pool
-
-```{r jtrace_candidates, echo=FALSE, fig.height=30, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-stimuli_trace <- stimuli %>% 
-  as_tibble() %>% 
-  select(group, input = word1, target = word2, phon_trace)
-
-# process TRACE data ----
-trace <- map(
-  list.files(here("Data", "jtrace"), full.names = TRUE),
-  function(x){
-    read_xlsx(x) %>% 
-      pivot_longer(-cycle, names_to = "response", values_to = "activation")
-  }
-) %>%
-  set_names(list.files(here("Data", "jtrace"))) %>% 
-  bind_rows(.id = "target") %>% 
-  separate(target, c("input", "language"), sep = "_") %>% 
-  mutate(language = str_remove(language, ".xlsx"),
-         language = str_to_sentence(language))
-
-# get top candidates
-trace_top <- trace %>% 
-  group_by(input, language, response) %>% 
-  summarise(
-    activation = max(activation, na.rm = TRUE),
-    .groups = "drop"
-  ) %>% 
-  group_by(input, language) %>% 
-  slice_max(activation, n = 10) %>% 
-  ungroup() %>% 
-  mutate(group = ifelse(language=="Catalan", "ENG-CAT", "ENG-SPA")) %>% 
-  select(-language) %>% 
-  # add experimental trials
-  left_join(stimuli_trace) %>% 
-  mutate(correct = response==phon_trace) %>% 
-  select(group, input, response, activation, correct)
-
-# experimental results ----
-responses_jtrace <- responses %>% 
-  filter(valid_response, group %in% c("ENG-CAT", "ENG-SPA")) %>% 
-  rename(response = input_text, input = word) %>% 
-  count(group, input, response, correct) %>% 
-  group_by(group, input) %>%
-  mutate(total = sum(n), activation = n/total) %>% 
-  ungroup() %>% 
-  group_by(group, input) %>% 
-  slice_max(order_by = activation, n = 10) %>% 
-  ungroup() %>% 
-  select(group, input, response, correct, activation) %>% 
-  filter(input %in% trace_top$input)
-
-# merge everything
-d_jtrace <- list(TRACE = trace_top, Participants = responses_jtrace) %>% 
-  bind_rows(.id = "source") %>% 
-  mutate(
-    # to reorder variable within panels
-    input_label = paste0(input, " (", group, ")"),
-    response_reordered = reorder_within(response, activation, input_label)
-  )
-
-ggplot(d_jtrace, aes(response_reordered, activation, fill = correct)) +
-  facet_wrap(input_label~source, scales = "free_y", ncol = 3) +
-  geom_col(colour = "white") +
-  labs(x = "Top 10 candidates", y = "Max. TRACE activation") +
-  coord_flip() +
-  scale_x_reordered() +
-  # scale_y_continuous(limits = c(0, 1)) +
-  theme_bw() +
-  theme(
-    legend.position = "none"
-  )
-
-```
-
 ## {-}
 
+# {-}
+
 # References

### 2021-09-29 _targets.R

@@ -1,0 +1,74 @@
+library(targets)
+library(tarchetypes)
+
+source("R/utils.R")
+source("R/00_stimuli.R")
+source("R/01_responses.R")
+source("R/02_models.R")
+
+# set parameters
+tar_option_set(
+    packages = c(
+        "dplyr", "tidyr", "stringr", "ggplot2", "tibble", "forcats", "multilex", "keyring",
+        "readxl", "janitor", "mice", "here", "lubridate", "purrr", "scales", "stringdist",
+        "brms", "tidybayes", "gt", "patchwork", "wesanderson", "papaja", "knitr", "data.table"
+    )
+)
+
+list(
+    
+    # items
+    tar_target(clearpond_path, list(`ENG-CAT` = here("Data", "clearpond", "clearpond_english.csv"), `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"), `SPA-CAT` = here("Data", "clearpond", "clearpond_spanish.csv")),),
+    tar_target(clearpond, get_clearpond(clearpond_path)),
+    tar_target(trace_path, here("Stimuli", "trace.xlsx")),
+    tar_target(trace, get_trace(trace_path)),
+    tar_target(stimuli_path, here("Stimuli", "trials.xlsx")),
+    tar_target(levenshtein, get_levenshtein(stimuli_path = stimuli_path)),
+    tar_target(stimuli, get_stimuli(stimuli_path = stimuli_path, clearpond = clearpond, trace = trace, levenshtein = levenshtein)),
+    tar_target(practice_trials, c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)),
+    
+    # responses
+    tar_target(responses_path, list.files(here("Data", "Raw"), full.names = TRUE)),
+    tar_target(responses_processed, get_responses_processed(responses_path = responses_path, stimuli = stimuli, practice_trials = practice_trials)),
+    tar_target(responses_clean, get_responses_clean(responses_processed = responses_processed, stimuli = stimuli)),
+    tar_target(responses_coded_path, here("Data", "02_coded.xlsx")),
+    tar_target(responses_coded, read_xlsx(responses_coded_path)),
+    tar_target(participants, get_participants(responses_processed = responses_processed,responses_coded = responses_coded)),
+    tar_target(responses, get_responses(responses_coded = responses_coded, practice_trials = practice_trials, participants = participants, stimuli = stimuli)),
+    
+    # models
+    tar_target(
[…]

### 2021-09-29 Rmd/manuscript.Rmd
Remove big files
Render report

--- 2021-07-01 Rmd/te_manuscript.Rmd
@@
 
-participants <- readRDS(here("Data", "participants.rds"))
-stimuli <- readRDS(here("Data", "stimuli.rds"))
-responses <- readRDS(here("Data", "responses.rds"))
+tar_load(participants)
+tar_load(stimuli)
+tar_load(responses)
+tar_load(model_fits)
 
@@
 
-## Algorithm
-
-## Translation Elicitation task
-
-### Participants
+## Participants
 
@@
 (Sí/No)".
-However, the proportion of Spanish participants that reported visual impairment was unplausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
+However, the proportion of Spanish participants that reported visual impairment was implausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
 This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`).
 
-### Procedure
+## Procedure
 
@@
 Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-First, participants were informed about the aims of the study and gave informed consent for participating.
-Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+Participants were informed about the aims of the study and gave informed consent for participating.
+Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
 Third, participants completed the experimental task.
@@
 
-### Stimuli
+## Stimuli
 
@@
 We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
-We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance.
+
 We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.

### 2021-09-29 Rmd/report.Rmd
Render report

@@
 )
-options(knitr.kable.NA = '-',
+options(
+  knitr.kable.NA = '-',
+  knitr.duplicate.label = "allow",
   ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
@@
 ```{r coefs_fixed_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-summary(fit_responses)$fixed %>% 
+summary(model_fits$fit_4)$fixed %>% 
   as.data.frame() %>%
@@
 The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
-post <- gather_draws(fit_responses, `b_.*`, regex = TRUE) %>% 
+post <- gather_draws(model_fits$fit_4, `b_.*`, regex = TRUE) %>% 
   mutate(.value = ifelse(.variable=="b_Intercept", inv_logit_scaled(.value), .value/4))
@@
     panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "responses_fixed.png"))
+  )
 ```
@@
 ```{r model_comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-map(loo$loos, "estimates") %>% 
-  map(function(x){
-    as.data.frame(x) %>% 
-      rownames_to_column("ic")
-  }
-  ) %>% 
-  bind_rows(.id = "model") %>% 
-  clean_names() %>% 
-  filter(ic %in% c("elpd_loo", "looic")) %>% 
-  pivot_wider(names_from = ic, values_from = c("estimate", "se")) %>% 
-  left_join(rownames_to_column(as.data.frame(loo$comparison, "model"))) %>% 
-  mutate(formula = c(
-    "~ 1 + frequency + (1 | participant)",
-    "~ 1 + frequency + pthn + (1 + pthn | participant)",
-    "~ 1 + frequency + pthn*vowel_ratio + (1 + pthn*vowel_ratio | participant)",
-    "~ 1 + frequency + pthn*consonant_ratio + (1 + pthn*consonant_ratio  | participant)",
-    "~ 1 + frequency + pthn*(consonant_ratio + consonant_ratio) + (1 + pthn*(consonant_ratio + consonant_ratio) | participant)"
-  )) %>% 
+model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
   arrange(desc(model)) %>% 
-  select(formula, estimate_elpd_loo, se_elpd_loo, estimate_looic, se_looic, elpd_diff, se_diff) %>%
-  rename(model = formula) %>% 
-  gt() %>% 
-  fmt_number(2:7) %>% 
-  cols_merge(c("estimate_elpd_loo", "se_elpd_loo"), pattern = "{1} ({2})") %>% 
-  cols_merge(c("estimate_looic", "se_looic"), pattern = "{1} ({2})") %>% 
-  cols_merge(c("elpd_diff", "se_diff"), pattern = "{1} ({2})") %>% 
+  select(model, matches("elpd_loo"), matches("diff")) %>%
+  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  fmt_number(2:5) %>% 
   cols_label(
     model = md("**Model**"),
-    estimate_elpd_loo = md("**ELPD<sub>LOO</sub> (SE)**"),
-    estimate_looic = md("**LOO-IC (SE)**"),
-    elpd_diff = md("**ELPD<sub>diff</sub> (SE)**")
-
+    elpd_loo = md("**ELPD**"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("**diff**"),
+    se_diff = md("***SE* diff**")
   )
@@
 ```{r post_re, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-post_re <- gather_draws(fit_responses, r_participant[participant, .param]) %>% 
+post_re <- gather_draws(model_fits$fit_4, r_participant[participant, .param]) %>% 
   mean_qi() %>% 
@@
 
-m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, model_fits$fit_4, n = 50, re_formula = NA) %>% 
   mutate_at(vars(consonant_ratio, vowel_ratio), function(x) paste0(x, " SD"))
@@
   )
-ggsave(here("Figures", "responses_pthn.png"))
 
@@
 
-m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, model_fits$fit_4, n = 50, re_formula = NA) %>% 
   mutate_at(vars(pthn, vowel_ratio), function(x) paste0(x, " SD"))
@@
     panel.grid = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "responses_consonant.png"), height = 4, width = 4.5)
+  )
 
@@
 
-m <- add_fitted_draws(nd, fit_responses, n = 50, re_formula = NA) %>% 
+m <- add_fitted_draws(nd, model_fits$fit_4, n = 50, re_formula = NA) %>% 
   mutate_at(vars(pthn, consonant_ratio), function(x) paste0(x, " SD"))
@@
     panel.grid = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "responses_vowel.png"), height = 4, width = 4.5)
+  )
 
@@
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "accuracy_item.png"))
+  )
 
@@
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-  ) +
-  ggsave(here("Figures", "accuracy_participant.png"))
+  )
 

### 2021-09-29 R/02_models.R

@@ -59,14 +59,15 @@ get_model_fits <- function(
         file = here("Results", "fit_responses_4.rds")
     )
     
-    fits <- list(fit_0 = fit_0, fit_1 = fit_1, fit_1 = fit_2, fit_1 = fit_3, fit_1 = fit_4)
+    fits <- list(fit_0 = fit_0, fit_1 = fit_1, fit_2 = fit_2, fit_3 = fit_3, fit_4 = fit_4)
     
     return(fits)
 }
 
+
 # compre models
 get_model_loos <- function(fits){
-    loos <- map(fits, loo)
-    loos_comp <- loo_compare(loos)
-    return(loos_comp)
+    loo <- loo_compare(map(fits, loo))
 }
+
+

### 2021-09-29 _targets.R

@@ -64,10 +64,10 @@ list(
     tar_target(
         model_loos,
         get_model_loos(model_fits)
-    )
+    ),
     
     # render report.Rmd
-    # tar_render(report, "Rmd/report.Rmd")
+    tar_render(report, "Rmd/report.Rmd")
 )
 
 

### 2021-09-29 Rmd/report.Rmd
Fix gt tables and ggplot2 theme in report.md

@@
   labs(x = "Group", y = "# participants", fill = "Valid?") +
-  theme_custom() +
   theme(
@@
   scale_y_continuous(limits = c(0.8, 1), labels = scales::percent) +
-  theme_custom() +
   theme(
@@
   plot_annotation() &
-  theme_custom() +
   theme(
@@
   scale_x_reordered() +
-  theme_custom() +
   theme(
@@
   coord_flip() +
-  theme_custom() +
   theme(
@@
   labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
-  theme_custom() +
   theme(
@@
   labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
-  theme_custom() +
   theme(
@@
   scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  theme_custom() +
   theme(
@@
   labs(x = "Group", y = "# trials", fill = "Phonological onset", colour = "Phonological onset") +
-  theme_custom() +
   theme(
@@
        fill = "Phonological onset", colour = "Phonological onset") +
-  theme_custom() +
   theme(
@@
     columns = 2:4
-  )
+  ) %>% 
+  as_raw_html()
 ```
@@
   labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
-  theme_custom() +
   theme(
@@
     se_diff = md("***SE* diff**")
-  )
+  ) %>% 
+  as_raw_html()
 
@@
   labs(x = "Participant-level adjustment", y = "Participant", colour = "Group") +
-  theme_custom() +
   theme(
@@
   guides(colour = guide_legend(ncol = 2)) +
-  theme_custom() +
   theme(

### 2021-09-29 Rmd/report.Rmd
Fix figures

@@
   cache.extra = knitr::rand_seed,
-  out.width = "100%",
+  out.width = "80%",
   results = "asis",
+  dpi = 300,
   dev.args = list(png = list(type = "cairo"))
@@
 
-#### Stan code
-
-Stan code generated by `brms`:
-
-```{stan stan_model, echo=TRUE, output.var="model"}
-// generated with brms 2.15.0
-functions {
-/* compute correlated group-level effects
-* Args: 
-*   z: matrix of unscaled group-level effects
-*   SD: vector of standard deviation parameters
-*   L: cholesky factor correlation matrix
-* Returns: 
-*   matrix of scaled group-level effects
-*/ 
-matrix scale_r_cor(matrix z, vector SD, matrix L) {
-// r is stored in another dimension order than z
-return transpose(diag_pre_multiply(SD, L) * z);
-}
-}
-data {
-int<lower=1> N;  // total number of observations
-int Y[N];  // response variable
-int<lower=1> K;  // number of population-level effects
-matrix[N, K] X;  // population-level design matrix
-// data for group-level effects of ID 1
-int<lower=1> N_1;  // number of grouping levels
-int<lower=1> M_1;  // number of coefficients per level
-int<lower=1> J_1[N];  // grouping indicator per observation
-// group-level predictor values
-vector[N] Z_1_1;
-vector[N] Z_1_2;
-vector[N] Z_1_3;
-vector[N] Z_1_4;
-int<lower=1> NC_1;  // number of group-level correlations
-int prior_only;  // should the likelihood be ignored?
-}
-transformed data {
-int Kc = K - 1;
-matrix[N, Kc] Xc;  // centered version of X without an intercept
-vector[Kc] means_X;  // column means of X before centering
-for (i in 2:K) {
-means_X[i - 1] = mean(X[, i]);
-Xc[, i - 1] = X[, i] - means_X[i - 1];
-}
-}
-parameters {
-vector[Kc] b;  // population-level effects
-real Intercept;  // temporary intercept for centered predictors
-vector<lower=0>[M_1] sd_1;  // group-level standard deviations
-matrix[M_1, N_1] z_1;  // standardized group-level effects
-cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
-}
-transformed parameters {
-matrix[N_1, M_1] r_1;  // actual group-level effects
-// using vectors speeds up indexing in loops
-vector[N_1] r_1_1;
-vector[N_1] r_1_2;
-vector[N_1] r_1_3;
-vector[N_1] r_1_4;
-// compute actual group-level effects
-r_1 = scale_r_cor(z_1, sd_1, L_1);
-r_1_1 = r_1[, 1];
-r_1_2 = r_1[, 2];
-r_1_3 = r_1[, 3];
-r_1_4 = r_1[, 4];
-}
-model {
-// likelihood including constants
-if (!prior_only) {
-// initialize linear predictor term
-vector[N] mu = Intercept + rep_vector(0.0, N);
-for (n in 1:N) {
-// add more terms to the linear predictor
-mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n];
-}
-target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
-}
-// priors including constants
-target += normal_lpdf(b | 0, 3);
-target += normal_lpdf(Intercept | 0, 3);
-target += cauchy_lpdf(sd_1 | 0, 3)
-- 4 * cauchy_lccdf(0 | 0, 3);
-target += std_normal_lpdf(to_vector(z_1));
-target += lkj_corr_cholesky_lpdf(L_1 | 5);
-}
-generated quantities {
-// actual population-level intercept
-real b_Intercept = Intercept - dot_product(means_X, b);
-// compute group-level correlations
-corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
-vector<lower=-1,upper=1>[NC_1] cor_1;
-// extract upper diagonal of correlation matrix
-for (k in 1:M_1) {
-for (j in 1:(k - 1)) {
-cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
-}
-}
-}
-```
-
 #### {-}
@@
 
-```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 50% and 95% CrI.
+```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 95% CrI.
 The intercept has been transformed using the inverse logit to get the average probability of correct response.
@@
 
-ggplot(post, aes(.value, .variable, fill = .variable)) +
+ggplot(post, aes(.value, .variable)) +
   geom_vline(xintercept = 0) +
-  stat_halfeye() +
+  stat_halfeye(point_colour = "white", interval_colour = "white", slab_fill = "grey80", .width = 0.95) +
   labs(x = "Value", y = "Posterior probability density", fill = "Variable") +
@@
 
+# Appendix
+
+## Appendix 1: Stan code
+
+#### Stan code
+
+Stan code generated by `brms`:
+
+```{stan stan_model, echo=TRUE, output.var="model"}
+// generated with brms 2.15.0
+functions {
+/* compute correlated group-level effects
+* Args: 
+*   z: matrix of unscaled group-level effects
+*   SD: vector of standard deviation parameters
+*   L: cholesky factor correlation matrix
+* Returns: 
+*   matrix of scaled group-level effects
+*/ 
+matrix scale_r_cor(matrix z, vector SD, matrix L) {
+// r is stored in another dimension order than z
+return transpose(diag_pre_multiply(SD, L) * z);
+}
+}
+data {
+int<lower=1> N;  // total number of observations
+int Y[N];  // response variable
+int<lower=1> K;  // number of population-level effects
+matrix[N, K] X;  // population-level design matrix
+// data for group-level effects of ID 1
+int<lower=1> N_1;  // number of grouping levels
+int<lower=1> M_1;  // number of coefficients per level
+int<lower=1> J_1[N];  // grouping indicator per observation
+// group-level predictor values
+vector[N] Z_1_1;
+vector[N] Z_1_2;
+vector[N] Z_1_3;
+vector[N] Z_1_4;
+int<lower=1> NC_1;  // number of group-level correlations
+int prior_only;  // should the likelihood be ignored?
+}
+transformed data {
+int Kc = K - 1;
+matrix[N, Kc] Xc;  // centered version of X without an intercept
+vector[Kc] means_X;  // column means of X before centering
+for (i in 2:K) {
+means_X[i - 1] = mean(X[, i]);
+Xc[, i - 1] = X[, i] - means_X[i - 1];
+}
+}
+parameters {
+vector[Kc] b;  // population-level effects
+real Intercept;  // temporary intercept for centered predictors
+vector<lower=0>[M_1] sd_1;  // group-level standard deviations
+matrix[M_1, N_1] z_1;  // standardized group-level effects
+cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
+}
+transformed parameters {
+matrix[N_1, M_1] r_1;  // actual group-level effects
+// using vectors speeds up indexing in loops
+vector[N_1] r_1_1;
+vector[N_1] r_1_2;
+vector[N_1] r_1_3;
+vector[N_1] r_1_4;
+// compute actual group-level effects
+r_1 = scale_r_cor(z_1, sd_1, L_1);
+r_1_1 = r_1[, 1];
+r_1_2 = r_1[, 2];
+r_1_3 = r_1[, 3];
+r_1_4 = r_1[, 4];
+}
+model {
+// likelihood including constants
+if (!prior_only) {
+// initialize linear predictor term
+vector[N] mu = Intercept + rep_vector(0.0, N);
+for (n in 1:N) {
+// add more terms to the linear predictor
+mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n];
+}
+target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
+}
+// priors including constants
+target += normal_lpdf(b | 0, 3);
+target += normal_lpdf(Intercept | 0, 3);
+target += cauchy_lpdf(sd_1 | 0, 3)
+- 4 * cauchy_lccdf(0 | 0, 3);
+target += std_normal_lpdf(to_vector(z_1));
+target += lkj_corr_cholesky_lpdf(L_1 | 5);
+}
+generated quantities {
+// actual population-level intercept
+real b_Intercept = Intercept - dot_product(means_X, b);
+// compute group-level correlations
+corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
+vector<lower=-1,upper=1>[NC_1] cor_1;
+// extract upper diagonal of correlation matrix
+for (k in 1:M_1) {
+for (j in 1:(k - 1)) {
+cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
+}
+}
+}
+```
+
 # References

### 2021-09-29 Rmd/manuscript.Rmd
Render manuscript on targets

@@
 output            :
-  papaja::apa6_pdf:
-    latex_engine: "xelatex"
+  word_document:
+    reference_docx: "manuscript_template.docx"
 ---
@@
 tar_load(model_fits)
+tar_load(model_loos)
 
@@
 The average duration of the audios was `r printnum(mean(stimuli$duration), digits = 2)` (*SD* = `r printnum(sd(stimuli$duration), digits = 2)`, *Min* = `r printnum(min(stimuli$duration), digits = 2)`, *Max* = `r printnum(max(stimuli$duration), digits = 2)`).
-The average duration of the Catalan audios was `r printnum(mean(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Catalan",]$duration), digits = 2)`), and the average duration of the Spanish audios was `r printnum(mean(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Spanish",]$duration), digits = 2)`).
-The average lexical frequency of the words was `r printnum(mean(stimuli$frequency_zipf), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf), digits = 2)`).
-The average lexical frequency of the Catalan audios was `r printnum(mean(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Catalan",]$frequency_zipf), digits = 2)`), and the average lexical frequency of the Spanish audios was `r printnum(mean(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$language=="Spanish",]$frequency_zipf), digits = 2)`).
+The average lexical frequency of the words was `r printnum(mean(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`).
+The average lexical frequency of the Catalan audios was `r printnum(mean(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`), and the average lexical frequency of the Spanish audios was `r printnum(mean(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`).
 Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT [@boada2013] and SUBTLEX-ESP [@cuetos2011], respectively, and are expressed as Zipf scores [@van-heuven2012; @zipf1949].
@@
 
-```{r posterior_draws}
-
-posterior %>%
-  ggplot(aes(.value, .variable)) +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  stat_slab(fill = "black") + 
-  stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.15)) +
-  labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
-       title = "Posterior distribution of coefficients of fixed effects") +
-  scale_colour_brewer(palette = "Blues") +
-  theme_classic() +
-  theme(panel.grid.major.y = element_line(colour = "grey"),
-        legend.position = "top",
-        legend.key = element_rect(fill = "transparent")) +
-  ggsave(here("Figures", "coefs.png"))
-
-```
-
 ## Posterior predictions
-
-```{r posterior_predictions}
-
-predictions %>%
-  ggplot(aes(similarity_phon, .value, colour = same_onset, fill = same_onset)) +
-  facet_grid(consecutive_longest~pthn)+
-  stat_summary(fun = mean, geom = "line", size = 1) +
-  geom_vline(xintercept = 0, colour= "grey") +
-  geom_hline(yintercept = 0.5, colour= "grey") +
-  labs(x = "Phonological similarity", y = "Probability of correct translation",
-       fill = "Credible interval", colour = "Onset",
-       title = "Effect of phonological similarity on probability of correct translation",
-       subtitle = "Model predictions are displayed for different quartiles of PTHN") +
-  scale_fill_brewer(palette = "Set1") +
-  scale_colour_brewer(palette = "Set1") +
-  theme_classic() +
-  theme(legend.position = "top") +
-  ggsave(here("Figures", "predictions.png"), height = 4)
-
-```
 
@@
 
-```{r item_effects}
+### Group effects
 
-item_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  ggplot(aes(estimate, fill = param)) +
-  facet_wrap(~param, scales = "free") +
-  geom_histogram(bins = 20) +
-  geom_rug(alpha = 0.1) +
-  geom_vline(xintercept = 0, linetype = "dashed") +
-  labs(x = "Estimate", y = "Density", fill = "Parameter") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_classic() +
-  theme(legend.position = "none",
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
-
+```{r group_effects_table}
 ```
 
-### Group effects
-
-```{r goup_effects_table}
-
-group_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  mutate_if(is.numeric, round, 2) %>% 
-  mutate(hdi = paste0("[", lower, ", ", upper, "]")) %>% 
-  select(param, group, error, hdi) %>% 
-  group_by(param) %>% 
-  gt() %>% 
-  cols_label(group = "Group", error = "SE", hdi = "95% HDI")
-
-```
-
-```{r group_effects}
-
-group_effects %>%
-  pivot_wider(names_from = estimate, values_from = value) %>% 
-  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
-                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
-  ggplot(aes(group, estimate)) +
-  facet_wrap(~param) +
-  geom_hline(yintercept = 0) +
-  geom_errorbar(aes(ymin = lower, ymax = upper, colour =  group), width = 0, size = 5) +
-  geom_errorbar(aes(ymin = estimate-error, ymax = estimate+error), width = 0, size = 1, alpha = 0.5) +
-  geom_point(size = 3) +
-  labs(x = "Group", y = "Estimate", fill = "Group") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_classic() +
-  theme(legend.position = "none",
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
-        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))
-
-```
 ### Correlations
-
-```{r correlations}
-
-corrs %>% 
-  filter(str_detect(variable, "trial_id")) %>%
-  mutate(variable = str_remove(variable, "cor_group:trial_id__"),
-         variable = str_replace(variable, "__" , " - ")) %>% 
-  ggplot(aes(value, variable, xmin = lower, xmax = upper)) +
-  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
-  geom_errorbarh(show.legend = FALSE, height = 0) +
-  geom_point(size = 2, show.legend = FALSE) +
-  labs(x = "Estimate", y = "Correlation") +
-  theme_classic() +
-  theme(axis.title.y = element_blank())
-
-```
 

### 2021-09-29 Rmd/manuscript.tex
Render manuscript on targets

--- 2020-07-30 Manuscript/te_manuscript.tex
@@
   man]{apa6}
+\usepackage{amsmath,amssymb}
 \usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-\usepackage{ifxetex,ifluatex}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
@@
 \hypersetup{
-  pdftitle={A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity},
-  pdfauthor={First Author1 \& Second Author1,2},
+  pdftitle={The role of cognateness in non-native spoken word recognition},
+  pdfauthor={Gonzalo Garcia-Castro1, Serene Siow2, Nuria Sebastian-Galles1, \& Kim Plunkett2},
   pdflang={en-EN},
-  pdfkeywords={keywords},
+  pdfkeywords={cognate, word recognition, translation, non-native, spoken word recognition},
   hidelinks,
@@
 \urlstyle{same} % disable monospaced font for URLs
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 % \newenvironment{ltable}
-%   {\begin{landscape}\begin{center}\begin{threeparttable}}
-%   {\end{threeparttable}\end{center}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}
+%   {\begin{landscape}\centering\begin{threeparttable}}
+%   {\end{threeparttable}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
 
@@
 \makeatother
-\shorttitle{Phoneme-weighted Levenshtein Distance}
-\keywords{keywords\newline\indent Word count: X}
+\shorttitle{Cognateness and non-native word recognition}
+\keywords{cognate, word recognition, translation, non-native, spoken word recognition\newline\indent Word count: X}
 \DeclareDelayedFloatFlavor{ThreePartTable}{table}
@@
 \usepackage{csquotes}
-\ifxetex
+\ifXeTeX
   % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
@@
 \else
-  \usepackage[shorthands=off,main=english]{babel}
+  \usepackage[main=english]{babel}
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
 \fi
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-\title{A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity}
-\author{First Author\textsuperscript{1} \& Second Author\textsuperscript{1,2}}
+\title{The role of cognateness in non-native spoken word recognition}
+\author{Gonzalo Garcia-Castro\textsuperscript{1}, Serene Siow\textsuperscript{2}, Nuria Sebastian-Galles\textsuperscript{1}, \& Kim Plunkett\textsuperscript{2}}
 \date{}
@@
 
-Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.
+GG and SS contributed equally and share first authorship.
 
-Enter author note here.
-
-Correspondence concerning this article should be addressed to First Author, Postal address. E-mail: \href{mailto:my@email.com}{\nolinkurl{my@email.com}}
+Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trial Fargas, 25-27, 08005 Barcelona.
+E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}
 
@@
 \abstract{
-Many studies in bilingualism have explored how form similarity between translation equivalents (i.e.~cognate- ness) impacts language processing.
-Cognates range from being identical to sharing only a few key features.
-Although there exists some objective measures of orthographic similarity between word-forms, there is still no representative equivalent for phonological similarity.
-In infant bilingual research, a measure that reflects perceptual similarity of spoken words as presented in infant-directed speech would be of great value.
-A widely used metric for word similarity is Levenshtein distance ({\textbf{???}}).
-This measure calculates word-form similarity by computing the smallest total number of substitutions, insertions and deletions needed to change one word to the other.
-However, a challenge when defining phonological similarity is that some phonemes (e.g.~{[}k{]} and {[}g{]}) are perceptually more similar than others (e.g.~{[}k{]} and {[}o{]}).
-The standard Lev- enshtein distance calculation treats each change equally.
-Thus it often overestimates phonological distance between words.
-We are developing an adaptation of Levenshtein distance for phonological representations, where edit operations are weighted according to the degree of phonemic feature changes involved, type of change (e.g.~vowel or consonant insertion) and position in the word.
-To estimate the weights, we are col- lecting behavioural data in an auditory translation elicitation task.
-Monolingual adult participants with no exposure to the target language will be asked to guess and produce the translations of auditorily-presented words.
-The only way for participants to guess the correct translation is by mapping the presented unknown phonological form onto the phonology of its known translational equivalent.
-Cross-linguistic phonological similarity will be operationalised as the probability that participants correctly guess the corresponding trans- lation.
-The higher the probability, the higher the perceptual similarity.
-Our weighted similarity metric will provide a useful tool for identifying phonological similarity between words in infant-directed speech.
+There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
+For example, bilinguals produce and translated cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
+Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
+These measures partially ignore potential sources of variability during lexical access like vowel vs.~consonant overlap, overlap in stressed syllables, onset overlap or the distance in features between both translations.
+In this study, we explored the impact of some of these variables on non-native word translation task: Spanish and English participants listened to non-native words (Catalan or Spanish) and were prompted to type their translation in their native language.
+Critically, participants where unfamiliar with the testing language, ensuring that they were not able to translate words based on previous knowledge on their meaning, leaving phonological information as the only cue participants were able to exploit to translate words to their native language correctly.
+We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.
 }
@@
 
-\hypertarget{algorithm}{%
-\subsection{Algorithm}\label{algorithm}}
+\hypertarget{participants}{%
+\subsection{Participants}\label{participants}}
 
-\hypertarget{translation-elicitation-task}{%
-\subsection{Translation Elicitation task}\label{translation-elicitation-task}}
-
-\hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
-
-Data collection took place from 04th junio, 2020 to 28th junio, 2020.
-We collected data from 107 participants (\(M_{age}\) = 21.79, \(SD_{age}\) = 2.41, \(Range_{age}\) = 18-33).
-74 participants were British English native speakers living in United Kingdom (49 female), and 74 participants were Spanish native speakers living in Spain (28 female).
+Data collection took place from 04th June, 2020 to 28th June, 2020.
+We collected data from 105 participants (\(M_{age}\) = 21.79, \(SD_{age}\) = 2.43, \(Range_{age}\) = 18-33).
+73 participants were British English native speakers living in United Kingdom (47 female), and 73 participants were Spanish native speakers living in Spain (27 female).
 Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
@@
 Participants were asked to complete the experiment in a quiet place with good internet connection.
-We exluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were diagnosed with a language (\emph{n} = 2)\footnote{We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as \enquote{Do you have normal or corrected-to-normal VISION?
-(Yes/No)} in English, and as \enquote{¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No)}.
-However, the proportion of Spanish participants that reported visual impairment was unplausibly large (\emph{n} = 6, 18.18\%).
+We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were diagnosed with a language (\emph{n} = 2)\footnote{We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as ``Do you have normal or corrected-to-normal VISION?
+(Yes/No)'' in English, and as ``¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
+(Sí/No).'' However, the proportion of Spanish participants that reported visual impairment was implausibly large (\emph{n} = 6, 18.75\%).
 This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.}, or c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
@@
 \hypertarget{procedure}{%
-\subsubsection{Procedure}\label{procedure}}
+\subsection{Procedure}\label{procedure}}
 
-The experiment was implemented online using Psychopy/Pavlovia ({\textbf{???}}).
+The experiment was implemented online using Psychopy/Pavlovia (\textbf{peirce2019?}).
 Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-First, participants were informed about the aims of the study and gave informed consent for participating.
-Second, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+Participants were informed about the aims of the study and gave informed consent for participating.
+Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
 Third, participants completed the experimental task.
-Before the task, participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
 They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
 English participants were randomly assigned to the list of Catalan or Spanish trials.
-Spanish participants completed the list of Catalan trials.
+Participants in the Catalan list were presented with 86 trials, and participants in the Spanish list were presented with 103 trials.
 
-\hypertarget{design}{%
-\subsubsection{Design}\label{design}}
-
-Each trial started with a yellow fixation dot presented during one second on the center of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio offset.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a \enquote{\textgreater{}} symbol.
+Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
+After one second, the audio started playing while the dot remained being displayed until the audio ended.
+Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ``\textgreater{}'' symbol.
 Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
@@
 We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Typos were allowed, however.
-Participants contributed a total of 9536 valid trials (5901 in Catalan, 3635 in Spanish).
+Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
+Participants contributed a total of 8077 valid trials (5235 in Catalan, 2842 in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 \hypertarget{stimuli}{%
-\subsubsection{Stimuli}\label{stimuli}}
+\subsection{Stimuli}\label{stimuli}}
+
+Participants listened to one audio file in each trial.
+This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infancia of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat (\textbf{boersma2018?}).
+The average duration of the audios was NA (\emph{SD} = NA, \emph{Min} = \(\infty\), \emph{Max} = \(-\infty\)).
+The average lexical frequency of the words was 4.29 (\emph{SD} = 0.59, \emph{Min} = 2.87, \emph{Max} = 6.14).
+The average lexical frequency of the Catalan audios was 4.31 seconds (\emph{SD} = 0.59, \emph{Min} = 2.87, \emph{Max} = 5.68), and the average lexical frequency of the Spanish audios was 4.32 seconds (\emph{SD} = 0.56, \emph{Min} = 2.87, \emph{Max} = 5.68).
+Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT (\textbf{boada2013?}) and SUBTLEX-ESP (\textbf{cuetos2011?}), respectively, and are expressed as Zipf scores (\textbf{van-heuven2012?}; \textbf{zipf1949?}).
 
@@
 We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
-We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance.
-We fit several Bayesian linear models ({\textbf{???}}), each including one of the aforementioned variables as predictor, and the proportion of correct reponses as output.
-Continuous predictors were standadised and categorical predictors were sum-coded ({\textbf{???}}).
-Missing predictor scores were imputed before fitting the models using multiple imputation ({\textbf{???}}).
+
+We fit several Bayesian linear models (\textbf{sorensen2015?}), each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.
+Continuous predictors were standardised and categorical predictors were sum-coded (\textbf{schad2020?}).
+Missing predictor scores were imputed before fitting the models using multiple imputation (\textbf{vanbuuren2013?}).
 To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using Paretho-Smoothed Importance Sampling (PSIS-LOO, ({\textbf{???}})).
+We compared models using leave-one-out cross-validation (LOO, (\textbf{vehtari2016?})).
 More information about the models and model comparison can be found in Appendix X.
 
-All analyses were performed in R environment ({\textbf{???}}).
-We used the \texttt{tidyverse} family of R packages to process data and to generate figures, the \texttt{mice} R package ({\textbf{???}}) to impute missing data, and the \texttt{brms} R package to ({\textbf{???}}) to fit and compare models, which uses the Stan probabilistic programming language language ({\textbf{???}}).
+All analyses were performed in R environment (\textbf{rcoreteam2019?}).
+We used the \texttt{tidyverse} family of R packages to process data and to generate figures, the \texttt{mice} R package (\textbf{vanbuuren2013?}) for multiple imputation of missing data, and the \texttt{brms} R package (\textbf{burkner2017?}) to fit and compare models, which uses the Stan probabilistic programming language language (\textbf{carpenter2016?}).
 
@@
 
-\includegraphics{te_manuscript_files/figure-latex/posterior_draws-1.pdf}
-
 \hypertarget{posterior-predictions}{%
 \subsection{Posterior predictions}\label{posterior-predictions}}
-
-\includegraphics{te_manuscript_files/figure-latex/posterior_predictions-1.pdf}
 
@@
 
-\includegraphics{te_manuscript_files/figure-latex/item_effects-1.pdf}
-
 \hypertarget{group-effects}{%
@@
 
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{lrl}
-\toprule
-Group & SE & 95\% HDI \\ 
-\midrule
-\multicolumn{1}{l}{Intercept} \\ 
-\midrule
-ENG-CAT & 1.15 & [-4.27, 0.09] \\ 
-ENG-SPA & 1.08 & [-4.05, 0.15] \\ 
-SPA-CAT & 0.92 & [-2.87, 0.84] \\ 
-\midrule
-\multicolumn{1}{l}{Different onset vs. Same onset} \\ 
-\midrule
-ENG-CAT & 1.73 & [-1.05, 5.49] \\ 
-ENG-SPA & 1.62 & [-1.42, 5.09] \\ 
-SPA-CAT & 1.49 & [-1.3, 4.56] \\ 
-\midrule
-\multicolumn{1}{l}{Same onset: Vowel vs. Consonant} \\ 
-\midrule
-ENG-CAT & 2.58 & [-4.11, 6.94] \\ 
-ENG-SPA & 2.59 & [-3.61, 7.29] \\ 
-SPA-CAT & 1.98 & [-3.59, 4.64] \\ 
-\midrule
-\multicolumn{1}{l}{Consecutive longest} \\ 
-\midrule
-ENG-CAT & 1.09 & [-2.54, 2.18] \\ 
-ENG-SPA & 1.13 & [-3.27, 1.59] \\ 
-SPA-CAT & 1.11 & [-1.73, 2.94] \\ 
-\midrule
-\multicolumn{1}{l}{Phonological similarity} \\ 
-\midrule
-ENG-CAT & 1.05 & [-2.45, 1.96] \\ 
-ENG-SPA & 1.04 & [-2.71, 1.69] \\ 
-SPA-CAT & 1.10 & [-1.62, 2.87] \\ 
-\bottomrule
-\end{longtable}
-
-\includegraphics{te_manuscript_files/figure-latex/group_effects-1.pdf}
-\#\#\# Correlations
-
-\includegraphics{te_manuscript_files/figure-latex/correlations-1.pdf}
+\hypertarget{correlations}{%
+\subsubsection{Correlations}\label{correlations}}
 
@@
 
+\begin{verbatim}
+## Warning in utils::citation(x[pkg], auto = if (no_citations[pkg]) TRUE else
+## NULL): no date field in DESCRIPTION file of package 'multilex'
+\end{verbatim}
+
+\begin{verbatim}
+## Warning in utils::citation(x[pkg], auto = if (no_citations[pkg]) TRUE else
+## NULL): could not determine year for 'multilex' from package DESCRIPTION file
+\end{verbatim}
+
 \begingroup
@@
 \hypertarget{refs}{}
+\begin{CSLReferences}{0}{0}
+\end{CSLReferences}
 

### 2021-09-29 _targets.R

@@ -67,8 +67,12 @@ list(
     ),
     
     # render report.Rmd
-    tar_render(report, "Rmd/report.Rmd")
-)
+    tar_render(report, "Rmd/report.Rmd"),
+
+    # render manuscript.Rmd
+    tar_render(manuscript, "Rmd/manuscript.Rmd")
+    
+    )
 
 
 

### 2021-11-25 Rmd/manuscript.Rmd
Starts manuscript

@@
 ---
-title             : "The role of cognateness in non-native spoken word recognition"
-shorttitle        : "Cognateness and non-native word recognition"
-
+title: "The role of cognateness in non-native spoken word recognition"
 author:
-  - name          : "Gonzalo Garcia-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trial Fargas, 25-27, 08005 Barcelona"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Serene Siow"
-    affiliation   : "2"
-  - name          : "Nuria Sebastian-Galles"
-    affiliation   : "1"
-  - name          : "Kim Plunkett"
-    affiliation   : "2"
-
-authornote: |
-  GG and SS contributed equally and share first authorship.
-
+- name: Gonzalo Garcia-Castro
+  affiliation: '1'
+  corresponding: yes
+  address: Ramon Trial Fargas, 25-27, 08005 Barcelona
+  email: gonzalo.garciadecastro@upf.edu
+- name: Serene Siow
+  affiliation: '2'
+- name: Nuria Sebastian-Galles
+  affiliation: '1'
+- name: Kim Plunkett
+  affiliation: '2'
+shorttitle: Cognateness and non-native word recognition
+output:
+  word_document:
+    reference_docx: manuscript_template.docx
 abstract: |
@@
 We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.
-
   <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "cognate, word recognition, translation, non-native, spoken word recognition"
-wordcount         : "X"
-
-bibliography      : ["r-references.bib", "references.bib"]
-
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  word_document:
-    reference_docx: "manuscript_template.docx"
+keywords: cognate, word recognition, translation, non-native, spoken word recognition
+wordcount: X
+bibliography:
+- r-references.bib
+- references.bib
+floatsintext: no
+figurelist: no
+tablelist: no
+footnotelist: no
+linenumbers: yes
+mask: no
+draft: no
+documentclass: apa6
+classoption: man
+authornote: |
+  GG and SS contributed equally and share first authorship.
 ---
 
-```{r setup, include = FALSE}
-library(papaja)
-```
+```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
 
-```{r analysis-preferences, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-# Seed for random number generation
 set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  cache = FALSE, 
+  cache.vars = FALSE,
+  cache.rebuild = FALSE,
+  cache.lazy = FALSE,
+  results = "asis", 
+  dpi = 300
+)
+
+options(
+  knitr.kable.NA = '-',
+  knitr.duplicate.label = "allow",
+  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
+  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
+  ggplot2.continuous.fill = wesanderson::wes_palettes$Darjeeling1,
+  ggplot2.continuous.colour = wesanderson::wes_palettes$Darjeeling1
+)
+
 ```
@@
 
-library(tidyverse)
-library(data.table)
-library(tidybayes)
-library(gt)
-library(lubridate)
-library(here)
-
+tar_load_globals()
+tar_load(stimuli)
 tar_load(participants)
-tar_load(stimuli)
 tar_load(responses)
@@
 
+theme_set(theme_ggdist())
+
 ```
+
+# Introduction
+
+## Non-native language comprehension is difficult, but cognateness helps
+
+Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech [@de1982relationship] or the loss of segmental information [@warren1970perceptual].
+But even in the most ideal of the situations, the processes engaged in word recognition do not occur without uncertainty.
+The speech input activates multiple candidate lexical entries based on their similarity with the signal, and for word recognition to take place, the lexical system must select one of the candidates.
+There is extensive literature about how the number of candidates, their frequency, and their phonological and semantic similarity with the target word impacts the dynamics of word recognition.
+However, less is known about how these factors affect the recognition of words in a non-native language.
+
+Listening to speech in a non-native language is more costly than doing so in the native language, even if one is proficient in the non-native language.
+The source of this increased cognitive effort is likely to originate at multiple levels.
+For instance, some sounds in the speech signal do not correspond to any phoneme in the native language, segmenting the speech signal relies on previous familiarity with word forms or with the statistical regularities of the language, or word order might be different in both languages (e.g., subject-object-verb vs.
+subject-verb-object.).
+The listener, however, is rarely completely naïve to the language they are listening to.
+All languages share, to some extent, similarities, frequently due to their typological closeness, which can be exploited by listeners of a non-native language.
+
+Lexical similarity, for example, provides useful information to sequential language learners.
+Romance language like Spanish, Italian, French or Catalan, share many similarities at the lexical level in the form of cognates [@schepens2012distributions].
+Cognates are cross-language synonyms whose form (e.g., phonology, orthography, signature) is similar.
+The cause of this similarity is frequently attributed to a shared etymological origin.
+This is the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively)^[Some form-similar cross-language synonyms are technically not cognates.
+For example, *sun* and *sol* (in Spanish), share their phonological onset, but their etymology points to different origins.
+We will use the word *cognateness* to include all form-similar cross-language synonyms for simplicity.
+It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
+There is evidence that cognates are learnt more easily in L2 than non-cognates.
+@de2000hard presented 40 Dutch natives 60 pairs of translation equivalents.
+In each learning trial, two words were presented in a screen side-by-side: one in Dutch and one pseudoword.
+Pseudowords were generated in such way that were easily pronounceable and were phonotactically legal in Dutch.
+Word pairs varied in their form-similarity (number of shared letters, 40-70%).
+Participants were tested twice in the same task with one week of difference.
+Participants' performance was better for cognates across both testing sessions, suggesting that cognates were learnt and retained more easily than non-cognates.
+Converging evidence was provided by @lotto1998effects in Dutch natives learning Italian words: cognates were learnt more easily than non-cognates.
+
+Cognates are not only learnt faster than non-cognates, but also translated faster.
+Early accounts of bilingual lexical access described two possible routes for translating a word from L1 to L2 or *vice versa*: (1) via direct links between L1-L2 links or (2) through word-concept links.
+Both translation routes are available during translation [@potter1984lexical; @kroll1988lexical].
+These models assumed that the strength of the word-concept associations in both L1 and L2 is a function of their frequency of use, in order to account for the influence of  both the lexical frequency of the presented word and the target word in translation tasks on participant's performance.
+Under the assumption that bilinguals use L1 more often, @kroll1994category proposed the asymmetric hypothesis, predicting stronger word-concept associations between L1 representations compared to L2 representations.
+This would make L2-to-L1 translation more reliant to the direct translation route (L1-L2 association) than L1-to-L2 translation.
+This prediction has three consequences: (1) L2-to-L1 translation is expected to be faster than L1-to-L2 translation (as it does not rely on conceptual mediation during translation), and (2) since word-to-word connections (like the L1-L2 one) occurs at the lexical level (and not at the conceptual level), it should be sensitive to the form-similarity between both words, and therefore to cognateness, and (3) since low-proficiency bilinguals rely more on the direct route during translation, and this route is sensitive to cognateness, their performance during backward translation tasks should not only be better than during forward translation, but also enhanced by the cognate status of L1 and L2 representations.
+
+To test these predictions, @degroot1994forward and @de1992determinants asked 52 Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+In each trial, participants were presented visually with a word in English or Dutch, and were asked to speak out loud its translation in the other language as soon as possible.
+The authors reported three main findings: (1) reaction times and accuracy were roughly equivalent in across both forward and backward conditions (although slightly faster in the former), (2) semantic variables, such as imageability, were positively associated with participants' performance more strongly in the forward translation conditional, although the effect size was small, (3) cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward translation than during forward translation, and (4) when translating cognates participants' performance across both conditions was less sensitive to semantic variables than when translating non-cognates.
+These results prompted the authors to reject a hard version @kroll1994category's asymmetric hypothesis, and suggest that both conceptually mediated and direct translation routes are active during translation, but backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness than forward translation.
+When the authors tested a group of bilinguals with higher proficiency in English, their results pointed in the same direction.
+
+The idealised lexicon considered by the models describe above only considers word-word connections between translation equivalents.
+More recent studies have highlighted the role of the rich network of connections that given word establishes with other phonologically or conceptually related words.
+@collins1975spreading introduced this notion into their theory of semantic processing, and later @luce1998recognizing operationalised this dimension in their Neighbourhood Activation Model (NAM).
+In this model, lexical selection is mediated not only by the lexical frequency of the target word, and the number of activated candidates, but also by the frequency of the candidates.
+@luce1998recognizing designed a lexical decision task in which English native participants listened to a word across several conditions of noise-to-signal ratio.
+The authors then used a computerised lexicon to calculate the number of phonological neighbours around each of the presented words, based on phonetic similarity matrices, and the average lexical frequency of such neighbourhoods.
+Participants answered more faster and more accurately to high frequency words.
+Words from high density neighbourhoods were responded to more accurately, but more slowly.
+The average lexical frequency of the neighbourhood was associated to a decrease in accuracy.
+Low frequency words were responded to more accurately in low density neighbours than in high density neighbours.
+The authors concluded that, although both word lexical frequency and neighbourhood density can, in principle, facilitate spoken word recognition, this effect was modulated by the frequency of the neighbourhood: when phonological neighbours are more frequent, lexical selection is hindered [@goldinger1989priming; @luce1990similarity].
+
+The role of neighbourhood lexical frequency on bilingual spoken word recognition is unclear.
+Across several modalities, bilinguals' recognition of words in a language is affected by form-similarity and frequency of words in the other language.
+The parallel activation hypothesis accounts for this phenomenon, suggesting that lexical access is language-non selective.
+There is strong evidence supporting this claim in both comprehension, production, and translation, and across modalities.
+Much of this evidence stems from the fact that bilinguals' production times are faster when naming cognates (i.e., words whose translation in the other language is form-similar) than non-cognates.
+Both phonological and semantic links between translations underlie facilitatory effect of cognateness.
+The activation of form-similar words in the non-target language that do not share meaning with the target word, however, interferes with lexical selection.
+Disentangling the effect of semantic and phonological links during spoken word recognition is challenging when participants are bilingual.
+
+This has important implications to non-native word recognition: is word recognition facilitated by cognateness via translation?
+The answer is not straightforward.
+Disentangling the effect of form- and conceptual similarity in a bilingual sample is daunting, given that second language learners are able to exploit ...
+
+In the present study, we used a fully monolingual sample to disentangle the effect of neighbourhood frequency and phonological similarity.
+We asked participants to translate word in a language they had no significant experience with.
+This made sure that participants were unaware of any semantic relationship between the (non-native) words they heard and their translation, and could only rely on their phonological similarity to guess their translation.
+We tested the role of the amount of phonological similarity between the presented and the target words, the lexical frequency of the target word, and the average lexical frequency of the target word's phonological neighbourhood.
 
@@
 
-Data collection took place from `r format(min(participants$date), "%dth %B, %Y")` to  `r format(max(participants$date), "%dth %B, %Y")`.
-We collected data from `r nrow(participants)` participants ($M_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(sd(participants$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
-`r filter(participants, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(participants, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(participants, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(participants, country=="Spain", sex =="Female") %>% nrow()` female).
+Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`.
+We collected data from `r length(unique(participants$participant))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, *Min* = `r printnum(min(participants$age, na.rm = TRUE))`, *Max* = `r printnum(max(participants$age, na.rm = TRUE))`).
+`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA") & participants$sex=="Female"]))`), and `r length(unique(participants$participant[participants$group %in% c("SPA-CAT")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant[participants$group %in% c("SPA-CAT") & participants$sex=="Female"]))` female).
 Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
@@
 Participants were asked to complete the experiment in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION?
-(Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No)".
-However, the proportion of Spanish participants that reported visual impairment was implausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`).
+We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`) , or c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
 
@@
 
-The experiment was implemented online using Psychopy/Pavlovia [@peirce2019].
+The experiment was implemented online using Psychopy/Pavlovia [@peirce2019psychopy2].
 Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
 Participants were informed about the aims of the study and gave informed consent for participating.
-Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+This experiment was  Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
 Third, participants completed the experimental task.
@@
 English participants were randomly assigned to the list of Catalan or Spanish trials.
-Participants in the Catalan list were presented with `r count(stimuli, group)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, group)$n[2]` trials.
+Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
 
-Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
+Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
 After one second, the audio started playing while the dot remained being displayed until the audio ended.
@@
 Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Participants contributed a total of `r distinct(responses, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(responses, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(responses, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
+Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infancia of Universitat Pompeu Fabra (Barcelona, Spain).
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
 These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@boersma2018].
-The average duration of the audios was `r printnum(mean(stimuli$duration), digits = 2)` (*SD* = `r printnum(sd(stimuli$duration), digits = 2)`, *Min* = `r printnum(min(stimuli$duration), digits = 2)`, *Max* = `r printnum(max(stimuli$duration), digits = 2)`).
-The average lexical frequency of the words was `r printnum(mean(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`).
-The average lexical frequency of the Catalan audios was `r printnum(mean(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`), and the average lexical frequency of the Spanish audios was `r printnum(mean(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`).
-Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT [@boada2013] and SUBTLEX-ESP [@cuetos2011], respectively, and are expressed as Zipf scores [@van-heuven2012; @zipf1949].
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
+The average duration of the audios was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, *Min* = `r printnum(min(stimuli$duration, na.rm = TRUE))`, *Max* = `r printnum(max(stimuli$duration, na.rm = TRUE))`).
+The average duration of the Catalan audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, *Min* = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, *Max* = `r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, *Min* = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, *Max* = `r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
+Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+
+```{r stimuli_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+stimuli %>% 
+  select(group, trial_id, frequency, frequency_zipf, pthn, consonant_ratio, vowel_ratio) %>% 
+  group_by(group) %>% 
+  summarise_at(vars(frequency:vowel_ratio), list(mean = ~mean(., na.rm = TRUE),  sd = ~sd(., na.rm = TRUE))) %>% 
+  gt(rowname_col = "group") %>% 
+  fmt_number(frequency_mean:vowel_ratio_sd) %>% 
+  tab_spanner(md("**Frequency (per million)**"), c("frequency_mean", "frequency_sd")) %>% 
+  tab_spanner(md("**Frequency (Zipf score)**"), matches("frequency_zipf")) %>% 
+  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
+  tab_spanner(md("**Consonant similarity**"), matches("consonant")) %>% 
+  tab_spanner(md("**Vowel similarity**"), matches("vowel")) %>% 
+  cols_label(
+    frequency_mean = md("Mean"),
+    frequency_sd = md("*SD*"),
+    frequency_zipf_mean = md("Mean"),
+    frequency_zipf_sd = md("*SD*"),
+    pthn_mean = md("Mean"),
+    pthn_sd = md("*SD*"),
+    consonant_ratio_mean = md("Mean"),
+    consonant_ratio_sd = md("*SD*"),
+    vowel_ratio_mean = md("Mean"),
+    vowel_ratio_sd = md("*SD*")
+  )
+
+```
 
@@
 
-We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT).
-We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
+We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We first fitted a base model (Model 0) that only included the lexical frequency (`frequency`) of the target word as a fixed effect, with a random intercept per participant.
+Second, we extended the model to include `pthn` as a fixed effect, with a random slope by participant (Model 1).
+Third, we added the fixed effect `consonant_ratio` and the `pthn:consonant_ratio`, and random slopes for both effects by participant (Model 2).
+Third, we added the fixed effect `vowel_ratio` and the `pthn:vowel_ratio`, and random slopes for both effects by participant (Model 3).
+Finally, we fit a model that included the two-way interactions `pthn:consonant_ratio` and `pthn:vowel_ratio`, an their random slopes by participant (Model 4).
+Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020capitalize].
+Model 4 can be formally expressed as:
 
-We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.
-Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020].
-Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013].
+$$
+\begin{align*}
+
+&\textbf{Likelihood}  \\
+y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
+
+&\textbf{Parameters}  \\
+
+logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
+& \beta_{1[p]} ~ Frequency_{i} ~ + \\
+& \beta_{2[p]} ~ PTHN_i ~ + \\
+& \beta_{3[p]} ~ Similarity_i ~ + \\
+& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
+
+\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
+\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
+&& \text{[participant-level coefficients]} \\ \\
+
+&\textbf{Prior}  \\
+
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 3) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ Cauchy(0, 3) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(5) && \text{[correlation between participant-level coefficients]} \\
+
+\end{align*}
+$$
+
 To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (LOO, @vehtari2016).
-More information about the models and model comparison can be found in Appendix X.
-
-All analyses were performed in R environment [@rcoreteam2019].
-We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
+We compared models using leave-one-out cross-validation (*LOO*) [@vehtari2017practical].
+More information about the models and model comparison can be found in Appendix 2.
+All analyses were performed in R environment [@rcore2019r].
+We used the tidyverse family of R packages to process data and to generate figures, and the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
 
@@
 
-## Posterior draws
+All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
+Model 4, which included all main effects and the two-way interactions between PTHN and vowel similarity, and PTHN and consonant similarity, showed the best performance (see Table 2).
 
-## Posterior predictions
+```{r model_values, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-## Random effects
+coefs <- fixef(model_fits$fit_3) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("variable") %>% 
+  clean_names() %>% 
+  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
+  group_split(variable) %>% 
+  set_names(make_clean_names(map(., "variable"))) %>% 
+  map(select, -variable) %>% 
+  map(unlist)
 
-### Item effects
-
-### Group effects
-
-```{r group_effects_table}
 ```
 
-### Correlations
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=4}
+
+loos_table <- model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
+  arrange(desc(elpd_loo)) %>% 
+  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
+  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  fmt_number(2:7) %>% 
+  cols_label(
+    model = md("**Model**"),
+    looic = md("***LOO<sub>IC</sub>***"),
+    se_looic = md("***SE*<sub>IC</sub>**"),
+    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("***LOO<sub>diff</sub>***"),
+    se_diff = md("***SE<sub>diff</sub>***")
+  )
+
+gtsave(loos_table, here("Figures", "loos.png"))
+
+```
+
+We now report the mean of the posterior distribution of each coefficient in Model 3, along with its associated measures of uncertainty.
+For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
+Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
+This way, the coefficients are expressed as increases/decreases in probability of correct translation [@gelman2020regression].
+
+```{r posterior_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 3.
+A) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of each coefficient.
+Credible intervals (CrI), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95%, 80%, and 50%, given the data we have collected.
+Points represent the mean of the distribution.
+The intercept has been transformed using the inverse logit to get the average probability of correct response.
+The rest of the coefficients have been transformed using the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in each predictor variable.
+B) Participant-level coefficient variability.
+Our model estimated participant-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (SD).
+C) Correlation between participant-level effects.
+Our model allowed participant-level coefficients to co-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each correlation.
+Coefficients are represented in the X- and Y-axis in the same order as indicated in the Y-axis of panels A and C.", fig.height=5, fig.width=6}
+
+post <- gather_draws(model_fits$fit_3, `b_.*`, `sd_.*`, regex = TRUE)
+
+# fixed effects
+str_repl <- c(
+  "Intercept" = "Intercept",
+  "frequency_zipf" = "Frequency (+1 SD)",
+  "pthn:lv" = "PTHN \u00d7 Levenshtein", 
+  "lv" = "Levenshtein (+1 SD)",
+  "pthn" = "PTHN (+1 SD)"
+)
+
+post_fix <- post %>% 
+  filter(str_detect(.variable, "b_")) %>% 
+  mutate(.variable_name = str_remove(.variable, "b_") %>% 
+           str_replace_all(str_repl) %>% 
+           factor(levels = c("Intercept", "Frequency (+1 SD)", "PTHN (+1 SD)", "Levenshtein (+1 SD)", "PTHN \u00d7 Levenshtein")),
+         .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)) %>% 
+  arrange(.variable) %>% 
+  ggplot(aes(.value, fct_rev(.variable_name))) +
+  geom_vline(xintercept = 0) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = scales::percent_format())))) +
+  stat_pointinterval(.width = 0, point_size = 1) +
+  # geom_text(
+  #   data = filter(rhats, str_detect(.variable, "b_")), 
+  #   aes(x = -0.1, y = .variable_name, label = paste0("Rhat = ", printnum(.value, digits = 3))),
+  #   size = 3, position = position_nudge(y = 0.1)
+  # ) +
+  # geom_text(
+  #   data = filter(n_effs, str_detect(.variable, "b_")), 
+  #   aes(x = -0.1, y = .variable_name, label = paste0("N eff. ratio = ", printnum(.value, digits = 2))),
+  #   size = 3, position = position_nudge(y = 0.25)
+  # ) +
+  scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
+  scale_x_continuous(labels = percent) +
+  labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
+  theme(
+    legend.position = c(1, 0.1),
+    legend.background = element_rect(fill = NA),
+    legend.justification = "right",
+    legend.direction = "horizontal",
+    plot.caption.position = "plot",
+    # axis.text = element_text(size = 7),
+    axis.title.y = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black", vjust = 0),
+    panel.grid.major.y = element_line(colour = "grey", size = 0.25)
+  )
+
+post_sd <- post %>% 
+  filter(str_detect(.variable, "sd_"),) %>% 
+  mutate(
+    .variable_name = str_remove(.variable, "sd_participant__") %>% 
+      str_replace_all(str_repl) %>% 
+      factor(levels = c("Intercept", "Frequency (+1 SD)", "PTHN (+1 SD)", "Levenshtein (+1 SD)", "PTHN \u00d7 Levenshtein"))
+  ) %>%   
+  ggplot(aes(.value, fct_rev(.variable_name))) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = scales::percent_format())))) +
+  stat_pointinterval(.width = 0, point_size = 1) +
+  #   geom_text(
+  #   data = filter(rhats, str_detect(.variable, "b_")), 
+  #   aes(x = -0.1, y = .variable_name, label = paste0("Rhat = ", printnum(.value, digits = 3))),
+  #   size = 3, position = position_nudge(y = 0.1)
+  # ) +
+  # geom_text(
+  #   data = filter(n_effs, str_detect(.variable, "b_")), 
+  #   aes(x = -0.1, y = .variable_name, label = paste0("N eff. ratio = ", printnum(.value, digits = 2))),
+  #   size = 3, position = position_nudge(y = 0.25)
+  # ) +
+  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
+  scale_x_continuous(labels = percent) +
+  labs(x = "SDs in P(Correct)", y = "Posterior probability density", fill = "CrI") +
+  theme(
+    legend.position = c(1, 0.1),
+    legend.background = element_rect(fill = NA),
+    legend.justification = "right",
+    legend.direction = "horizontal",
+    plot.caption.position = "plot",
+    plot.caption = element_text(hjust = 0),
+    # axis.text = element_text(size = 7),
+    panel.grid.major.y = element_line(colour = "grey", size = 0.25),
+    axis.title.y = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black")
+  )
+
+corr_mat <- cov2cor(vcov(model_fits$fit_3)) 
+corr_mat[lower.tri(corr_mat)] <- NA
+
+post_cors <- corr_mat %>% 
+  as.data.frame() %>% 
+  rownames_to_column("term1") %>% 
+  pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
+  drop_na(.value) %>% 
+  mutate(term1 = factor(term1, levels = names(str_repl), ordered = TRUE),
+         term2 = factor(term2, levels = names(str_repl), ordered = TRUE)) %>% 
+  mutate_at(vars(term1, term2), str_replace_all, str_repl) %>% 
+  mutate(
+    .value_label = case_when(
+      term1==term2 ~ NA_character_,
+      TRUE ~ printnum(.value, gt1 = FALSE)
+    ),
+    .value = case_when(
+      term1==term2 ~ NA_real_,
+      TRUE ~ .value
+    )
+  ) %>% 
+  ggplot(aes(fct_inorder(term1), fct_rev(fct_inorder(term2)), fill = .value)) +
+  geom_tile(na.rm = TRUE) +
+  geom_text(aes(label = .value_label), size = 3, na.rm = TRUE) +
+  labs(x = "Term 1", y = "Term 2", fill = "Posterior correlation") +
+  scale_fill_gradient(low = "#FFC20A", high = "white", na.value = "white") +
+  coord_equal() +
+  theme(
+    legend.position = "none",
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    plot.background = element_rect(fill = NA)
+  )
+
+post_fix / (post_sd + post_cors + plot_layout(widths = c(0.6, 0.4))) +
+  plot_layout(guides = "keep") & 
+  plot_annotation(tag_levels = "A") &
+  theme(
+    title = element_text(size = 8, face = "bold"),
+    plot.title.position = "plot", 
+    plot.title = element_text(hjust = 0.5)
+  )
+
+```
+
+Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (*SE* = `r percent(coefs$intercept["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations.
+Every standard deviation increment in the translation's lexical frequency (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))`) increased the probability of a correct responses in `r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)` (*SE* = `r percent(coefs$frequency_zipf["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$frequency_zipf["q2_5"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf["q97_5"], accuracy = 0.01)`]).
+The number of the translation's more frequent phonological neighbours, on the other hand, decreased the probability of a correct responses in `r percent(coefs$pthn["estimate"], accuracy = 0.01)` (*SE* = `r percent(coefs$pthn["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$pthn["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn["q97_5"], accuracy = 0.01)`]) for every increase in 1 *SD* (`r printnum(sd(stimuli$pthn, na.rm = TRUE))`).
+The effect of phonological similarity was conditional to the phonological density of the translation.
+Phonological similarity barely increased the probability of a correct responses (`r percent(coefs$lv["estimate"], accuracy = 0.01)`, *SE* = `r percent(coefs$lv["est_error"])`, 95% *CrI* = [`r percent(coefs$lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv["q97_5"])`]) by itself.
+However, for every *SD* increase in PTHN, phonological similarity increased in `r percent(coefs$pthn_lv["estimate"])`, (*SE* = `r percent(coefs$pthn_lv["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$pthn_lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_lv["q97_5"], accuracy = 0.01)`]) such probability.
+
+```{r marginal_effects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=5, fig.height=5, fig.cap="Expected mean posterior predictions.
+The X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 200 observations from our model: 100 simulations for words with low PTHN (-1 SD) and 100 simulations for high PTHN (+1 SD) words.
+For each simulation, we drew a single sample from the posterior distribution of each coefficient.
+Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high PTHN words.
+We also plotted the mean of the high PTHN (black solid line) and low PTHN (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
+The dispersion of the lines indicates the uncertainty of our predictions."}
+
+# by vowel_ratio
+nd <- expand.grid(
+  pthn = c(-1, 1),
+  frequency_zipf = 0,
+  lv = seq(
+    min(responses$lv, na.rm = TRUE),
+    max(responses$lv, na.rm = TRUE),
+    by = 0.1
+  )
+)
+
+m <- add_epred_draws(nd, model_fits$fit_3, ndraws = 100, re_formula = NA) %>% 
+  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
+  mutate(
+    pthn = case_when(
+      pthn=="-1 SD" ~ "Low PTHN",
+      pthn=="0 SD" ~ "Mean PTHN",
+      pthn=="1 SD" ~ "High PTHN"
+    ),
+    frequency_zipf = case_when(
+      frequency_zipf=="-1 SD" ~ "Low frequency",
+      frequency_zipf=="0 SD" ~ "Mean frequency",
+      frequency_zipf=="1 SD" ~ "High frequency"
+    )
+  )
+
+nd_obs <- model_fits$fit_3$data %>% 
+  as_tibble() %>% 
+  mutate(pthn = cut_interval(pthn, 3, labels = c("Low PTHN", "Mean PTHN", "High PTHN")))
+
+ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
+  # geom_jitter(data = filter(nd_obs, correct), aes(y = 1), alpha = 0.1,
+  #             size = 0.25, height = 0.05, width = 0.1, show.legend = FALSE) +
+  # geom_jitter(data = filter(nd_obs, !correct), aes(y = 0), alpha = 0.25, 
+  #             size = 0.25, height = 0.05, width = 0.1, show.legend = FALSE) +
+  geom_hline(yintercept = 0.5, colour = "grey") +
+  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 1) +
+  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = FALSE) +
+  labs(x = "Levenshtein (phonological similarity)", y = "P(Correct)", colour = "PTHN", fill = "PTHN", linetype = "PTHN") +
+  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
+  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 0.6), breaks = seq(0, 1, 0.1)) +
+  guides(colour = guide_legend(ncol = 2)) +
+  theme(
+    legend.position = "top",
+    legend.title = element_blank(),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+    panel.grid = element_line(colour = "grey")
+  )
+
+```
 
@@
 # References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
 
@@
 \endgroup
+
+\newpage
+
+# Appendix
+
+```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
+Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.
+Dots and whiskers represent the average accuracy and 95% confidence interval of each word.
+Accuracy is plotted separately for each group.", fig.height=10, fig.width=9}
+
+responses_item <- responses %>%
+  group_by(trial_id, group, word) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  left_join(select(stimuli, group, trial_id, word2, ipa1, ipa2)) %>% 
+  mutate(
+    word = paste0(word, " (", ipa1, ") / ", word2, " (", ipa2, ")"),
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    word_label = reorder_within(word, by = correct_prop, within = group)
+  )
+
+ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5, colour = "black") +
+  geom_errorbar(width = 0) +
+  geom_point(size = 0.5) +
+  labs(y = "Proportion of correct responses", x = "Word", colour = "Group") +
+  coord_flip() +
+  scale_color_manual(values = c("#1E88E5", "#ff2976", "#FFC20A")) +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  scale_x_discrete(labels = function(x) str_remove(x, "___ENG-CAT|___ENG-SPA|___SPA-CAT")) +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_text(size = 6, colour = "black"),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 7, colour = "black"),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  )
+
+```

### 2021-11-25 Rmd/manuscript.tex
Starts manuscript

@@
 \documentclass[
-  english,
   man]{apa6}
@@
   pdftitle={The role of cognateness in non-native spoken word recognition},
-  pdfauthor={Gonzalo Garcia-Castro1, Serene Siow2, Nuria Sebastian-Galles1, \& Kim Plunkett2},
-  pdflang={en-EN},
-  pdfkeywords={cognate, word recognition, translation, non-native, spoken word recognition},
+  pdfkeywords={cognate, word recognition, translation, non-native,
+spoken word recognition},
   hidelinks,
@@
 \setcounter{secnumdepth}{-\maxdimen} % remove section numbering
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-% Manuscript styling
-\usepackage{upgreek}
-\captionsetup{font=singlespacing,justification=justified}
-
-% Table formatting
+\usepackage{amsmath}
+\usepackage{booktabs}
+\usepackage{caption}
 \usepackage{longtable}
-\usepackage{lscape}
-% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
-\usepackage{multirow}		% Table styling
-\usepackage{tabularx}		% Control Column width
-\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
-\usepackage{threeparttablex}            % Lets threeparttable work with longtable
-
-% Create new environments so endfloat can handle them
-% \newenvironment{ltable}
-%   {\begin{landscape}\centering\begin{threeparttable}}
-%   {\end{threeparttable}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
-
-% Enables adjusting longtable caption width to table width
-% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
-\makeatletter
-\newcommand\LastLTentrywidth{1em}
-\newlength\longtablewidth
-\setlength{\longtablewidth}{1in}
-\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
-
-% \setlength{\parindent}{0.5in}
-% \setlength{\parskip}{0pt plus 0pt minus 0pt}
-
-% \usepackage{etoolbox}
-\makeatletter
-\patchcmd{\HyOrg@maketitle}
-  {\section{\normalfont\normalsize\abstractname}}
-  {\section*{\normalfont\normalsize\abstractname}}
-  {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
-  {\section{\protect\normalfont{\@title}}}
-  {\section*{\protect\normalfont{\@title}}}
-  {}{\typeout{Failed to patch title.}}
-\makeatother
-\shorttitle{Cognateness and non-native word recognition}
-\keywords{cognate, word recognition, translation, non-native, spoken word recognition\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
-\usepackage{lineno}
-
-\linenumbers
-\usepackage{csquotes}
-\ifXeTeX
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
 \ifLuaTeX
@@
 \title{The role of cognateness in non-native spoken word recognition}
-\author{Gonzalo Garcia-Castro\textsuperscript{1}, Serene Siow\textsuperscript{2}, Nuria Sebastian-Galles\textsuperscript{1}, \& Kim Plunkett\textsuperscript{2}}
+\author{true \and true \and true \and true}
 \date{}
-
-\authornote{
-
-GG and SS contributed equally and share first authorship.
-
-Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trial Fargas, 25-27, 08005 Barcelona.
-E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}
-
-}
-
-\affiliation{\phantom{0}}
-
-\abstract{
-There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
-For example, bilinguals produce and translated cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
-Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
-These measures partially ignore potential sources of variability during lexical access like vowel vs.~consonant overlap, overlap in stressed syllables, onset overlap or the distance in features between both translations.
-In this study, we explored the impact of some of these variables on non-native word translation task: Spanish and English participants listened to non-native words (Catalan or Spanish) and were prompted to type their translation in their native language.
-Critically, participants where unfamiliar with the testing language, ensuring that they were not able to translate words based on previous knowledge on their meaning, leaving phonological information as the only cue participants were able to exploit to translate words to their native language correctly.
-We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.
-}
 
@@
 \maketitle
+\begin{abstract}
+There is compelling evidence that bilinguals access their lexicon in a
+language non-selective way. For example, bilinguals produce and
+translated cognates (words whose translation in the other language is
+form-similar) faster than non-cognates, suggesting that the phonology of
+both languages interact during word production and comprehension (Costa
+et al., 2000; Christoffels et al., 2006). Previous literature on this
+effect has often relied on measures of overall form-similarity to
+categorise words into cognates andnon-cognates, such as the Levenshtein
+distance. These measures partially ignore potential sources of
+variability during lexical access like vowel vs.~consonant overlap,
+overlap in stressed syllables, onset overlap or the distance in features
+between both translations. In this study, we explored the impact of some
+of these variables on non-native word translation task: Spanish and
+English participants listened to non-native words (Catalan or Spanish)
+and were prompted to type their translation in their native language.
+Critically, participants where unfamiliar with the testing language,
+ensuring that they were not able to translate words based on previous
+knowledge on their meaning, leaving phonological information as the only
+cue participants were able to exploit to translate words to their native
+language correctly. We analysed the probability of correct translations,
+adjusting for the amount of vowel overlap, consonant overlap, overlap at
+stress position, overlap at onset, and distance in features between
+replaced phonemes, including the average frequency of phonological
+neighbours of the target words are a covariate.
+\end{abstract}
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+\hypertarget{non-native-language-comprehension-is-difficult-but-cognateness-helps}{%
+\subsection{Non-native language comprehension is difficult, but
+cognateness
+helps}\label{non-native-language-comprehension-is-difficult-but-cognateness-helps}}
+
+Humans are able to recognise spoken words without much effort, even in
+adverse conditions such as compressed speech (De Haan 1982) or the loss
+of segmental information (Warren 1970). But even in the most ideal of
+the situations, the processes engaged in word recognition do not occur
+without uncertainty. The speech input activates multiple candidate
+lexical entries based on their similarity with the signal, and for word
+recognition to take place, the lexical system must select one of the
+candidates. There is extensive literature about how the number of
+candidates, their frequency, and their phonological and semantic
+similarity with the target word impacts the dynamics of word
+recognition. However, less is known about how these factors affect the
+recognition of words in a non-native language.
+
+Listening to speech in a non-native language is more costly than doing
+so in the native language, even if one is proficient in the non-native
+language. The source of this increased cognitive effort is likely to
+originate at multiple levels. For instance, some sounds in the speech
+signal do not correspond to any phoneme in the native language,
+segmenting the speech signal relies on previous familiarity with word
+forms or with the statistical regularities of the language, or word
+order might be different in both languages (e.g., subject-object-verb
+vs.~subject-verb-object.). The listener, however, is rarely completely
+naïve to the language they are listening to. All languages share, to
+some extent, similarities, frequently due to their typological
+closeness, which can be exploited by listeners of a non-native language.
+
+Lexical similarity, for example, provides useful information to
+sequential language learners. Romance language like Spanish, Italian,
+French or Catalan, share many similarities at the lexical level in the
+form of cognates (Schepens, Dijkstra, and Grootjen 2012). Cognates are
+cross-language synonyms whose form (e.g., phonology, orthography,
+signature) is similar. The cause of this similarity is frequently
+attributed to a shared etymological origin. This is the case of
+\emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan,
+respectively)\footnote{Some form-similar cross-language synonyms are
+  technically not cognates. For example, \emph{sun} and \emph{sol} (in
+  Spanish), share their phonological onset, but their etymology points
+  to different origins. We will use the word \emph{cognateness} to
+  include all form-similar cross-language synonyms for simplicity. It is
+  highly implausible that etymology plays a direct role on language
+  perception if it is not via form-similarity, since it is not necessary
+  for participants in psycholinguistic experiments to be aware of the
+  etymology of the words they encounter in the tasks to be subject to
+  the effect of form-similarity.}. There is evidence that cognates are
+learnt more easily in L2 than non-cognates. De Groot and Keijzer (2000)
+presented 40 Dutch natives 60 pairs of translation equivalents. In each
+learning trial, two words were presented in a screen side-by-side: one
+in Dutch and one pseudoword. Pseudowords were generated in such way that
+were easily pronounceable and were phonotactically legal in Dutch. Word
+pairs varied in their form-similarity (number of shared letters,
+40-70\%). Participants were tested twice in the same task with one week
+of difference. Participants' performance was better for cognates across
+both testing sessions, suggesting that cognates were learnt and retained
+more easily than non-cognates. Converging evidence was provided by Lotto
+and De Groot (1998) in Dutch natives learning Italian words: cognates
+were learnt more easily than non-cognates.
+
+Cognates are not only learnt faster than non-cognates, but also
+translated faster. Early accounts of bilingual lexical access described
+two possible routes for translating a word from L1 to L2 or \emph{vice
+versa}: (1) via direct links between L1-L2 links or (2) through
+word-concept links. Both translation routes are available during
+translation (Potter et al. 1984; Kroll, Curley, and others 1988). These
+models assumed that the strength of the word-concept associations in
+both L1 and L2 is a function of their frequency of use, in order to
+account for the influence of both the lexical frequency of the presented
+word and the target word in translation tasks on participant's
+performance. Under the assumption that bilinguals use L1 more often,
+Kroll and Stewart (1994) proposed the asymmetric hypothesis, predicting
+stronger word-concept associations between L1 representations compared
+to L2 representations. This would make L2-to-L1 translation more reliant
+to the direct translation route (L1-L2 association) than L1-to-L2
+translation. This prediction has three consequences: (1) L2-to-L1
+translation is expected to be faster than L1-to-L2 translation (as it
+does not rely on conceptual mediation during translation), and (2) since
+word-to-word connections (like the L1-L2 one) occurs at the lexical
+level (and not at the conceptual level), it should be sensitive to the
+form-similarity between both words, and therefore to cognateness, and
+(3) since low-proficiency bilinguals rely more on the direct route
+during translation, and this route is sensitive to cognateness, their
+performance during backward translation tasks should not only be better
+than during forward translation, but also enhanced by the cognate status
+of L1 and L2 representations.
+
+To test these predictions, Degroot, Dannenburg, and Vanhell (1994) and
+Groot (1992) asked 52 Dutch natives with high (yet non-native) English
+proficiency to translate words from either Dutch to English (forward
+translation, L1 to L2) or from English to Dutch (backward translation,
+L2 to L1). In each trial, participants were presented visually with a
+word in English or Dutch, and were asked to speak out loud its
+translation in the other language as soon as possible. The authors
+reported three main findings: (1) reaction times and accuracy were
+roughly equivalent in across both forward and backward conditions
+(although slightly faster in the former), (2) semantic variables, such
+as imageability, were positively associated with participants'
+performance more strongly in the forward translation conditional,
+although the effect size was small, (3) cognates were translated equally
+fast in both conditions, but non-cognates were translated faster during
+backward translation than during forward translation, and (4) when
+translating cognates participants' performance across both conditions
+was less sensitive to semantic variables than when translating
+non-cognates. These results prompted the authors to reject a hard
+version Kroll and Stewart (1994)'s asymmetric hypothesis, and suggest
+that both conceptually mediated and direct translation routes are active
+during translation, but backward translation relies more strongly on
+direct links between L1 and L2 representations, which makes it more
+sensitive to cognateness than forward translation. When the authors
+tested a group of bilinguals with higher proficiency in English, their
+results pointed in the same direction.
+
+The idealised lexicon considered by the models describe above only
+considers word-word connections between translation equivalents. More
+recent studies have highlighted the role of the rich network of
+connections that given word establishes with other phonologically or
+conceptually related words. Collins and Loftus (1975) introduced this
+notion into their theory of semantic processing, and later Luce and
+Pisoni (1998) operationalised this dimension in their Neighbourhood
+Activation Model (NAM). In this model, lexical selection is mediated not
+only by the lexical frequency of the target word, and the number of
+activated candidates, but also by the frequency of the candidates. Luce
+and Pisoni (1998) designed a lexical decision task in which English
+native participants listened to a word across several conditions of
+noise-to-signal ratio. The authors then used a computerised lexicon to
+calculate the number of phonological neighbours around each of the
+presented words, based on phonetic similarity matrices, and the average
+lexical frequency of such neighbourhoods. Participants answered more
+faster and more accurately to high frequency words. Words from high
+density neighbourhoods were responded to more accurately, but more
+slowly. The average lexical frequency of the neighbourhood was
+associated to a decrease in accuracy. Low frequency words were responded
+to more accurately in low density neighbours than in high density
+neighbours. The authors concluded that, although both word lexical
+frequency and neighbourhood density can, in principle, facilitate spoken
+word recognition, this effect was modulated by the frequency of the
+neighbourhood: when phonological neighbours are more frequent, lexical
+selection is hindered (Goldinger, Luce, and Pisoni 1989; Luce, Pisoni,
+and Goldinger 1990).
+
+The role of neighbourhood lexical frequency on bilingual spoken word
+recognition is unclear. Across several modalities, bilinguals'
+recognition of words in a language is affected by form-similarity and
+frequency of words in the other language. The parallel activation
+hypothesis accounts for this phenomenon, suggesting that lexical access
+is language-non selective. There is strong evidence supporting this
+claim in both comprehension, production, and translation, and across
+modalities. Much of this evidence stems from the fact that bilinguals'
+production times are faster when naming cognates (i.e., words whose
+translation in the other language is form-similar) than non-cognates.
+Both phonological and semantic links between translations underlie
+facilitatory effect of cognateness. The activation of form-similar words
+in the non-target language that do not share meaning with the target
+word, however, interferes with lexical selection. Disentangling the
+effect of semantic and phonological links during spoken word recognition
+is challenging when participants are bilingual.
+
+This has important implications to non-native word recognition: is word
+recognition facilitated by cognateness via translation? The answer is
+not straightforward. Disentangling the effect of form- and conceptual
+similarity in a bilingual sample is daunting, given that second language
+learners are able to exploit \ldots{}
+
+In the present study, we used a fully monolingual sample to disentangle
+the effect of neighbourhood frequency and phonological similarity. We
+asked participants to translate word in a language they had no
+significant experience with. This made sure that participants were
+unaware of any semantic relationship between the (non-native) words they
+heard and their translation, and could only rely on their phonological
+similarity to guess their translation. We tested the role of the amount
+of phonological similarity between the presented and the target words,
+the lexical frequency of the target word, and the average lexical
+frequency of the target word's phonological neighbourhood.
 
@@
 
-Data collection took place from 04th June, 2020 to 28th June, 2020.
-We collected data from 105 participants (\(M_{age}\) = 21.79, \(SD_{age}\) = 2.43, \(Range_{age}\) = 18-33).
-73 participants were British English native speakers living in United Kingdom (47 female), and 73 participants were Spanish native speakers living in Spain (27 female).
-Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
-Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants were asked to complete the experiment in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were diagnosed with a language (\emph{n} = 2)\footnote{We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as ``Do you have normal or corrected-to-normal VISION?
-(Yes/No)'' in English, and as ``¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No).'' However, the proportion of Spanish participants that reported visual impairment was implausibly large (\emph{n} = 6, 18.75\%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.}, or c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
+Data collection took place from June 04th, 2020 to June 28th, 2020. We
+collected data from 104 participants (\emph{Mean} = 21.79 years,
+\emph{SD} = 2.43, \emph{Min} = 18, \emph{Max} = 33). 72 participants
+were British English native speakers living in United Kingdom (46), and
+32 participants were Spanish native speakers living in Spain (27
+female). Participants in UK were recruited via Prolific (5£
+compensation) and SONA (compensation in academic credits). Participants
+in Spain were contacted via announcements in Faculties, and were
+compensated 5€ or an Amazon voucher for the same value. Participants
+were asked to complete the experiment in a quiet place with good
+internet connection. We excluded data from participants that a)
+self-rated their oral and/or written skills in a second or third
+language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were
+diagnosed with a language (\emph{n} = 2) , or c) did not contribute more
+than 80\% of valid trials (\emph{n} = 9).
 
@@
 
-The experiment was implemented online using Psychopy/Pavlovia (\textbf{peirce2019?}).
-Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-Participants were informed about the aims of the study and gave informed consent for participating.
-Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
-Third, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
-They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the list of Catalan or Spanish trials.
-Participants in the Catalan list were presented with 86 trials, and participants in the Spanish list were presented with 103 trials.
+The experiment was implemented online using Psychopy/Pavlovia (Peirce et
+al. 2019). Participants accessed the study from a link provided by
+Prolific or SONA and completed the experiment from a browser (Chrome or
+Mozilla). Participants were informed about the aims of the study and
+gave informed consent for participating. This experiment was Then,
+participants answered a series of questions about their demographic
+status, their language background, and the set up they were using for
+completing the study. Third, participants completed the experimental
+task. Participants were informed that they would listen to a series of
+pre-recorded words in Catalan or Spanish (English participants) or only
+Catalan (Spanish participants). They were instructed to listen to each
+word, guess its meaning in English (English participants) or Spanish
+(Spanish participants), and type their answer as soon as possible.
+English participants were randomly assigned to the list of Catalan or
+Spanish trials. Participants in the Catalan list were presented with 86
+trials, and participants in the Spanish list were presented with 103
+trials.
 
-Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio ended.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ``\textgreater{}'' symbol.
-Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
-Participants were allowed to correct their answer.
-Then, participants pressed the RETURN key to start and new trial.
-We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Participants contributed a total of 8077 valid trials (5235 in Catalan, 2842 in Spanish).
-The task took approximately 15 minutes to be completed.
+Each trial started with a yellow fixation point presented during one
+second on the centre of the screen over a black background. After one
+second, the audio started playing while the dot remained being displayed
+until the audio ended. Upon the offset of the fixation point and audio,
+participants were prompted to write their answer by a ``\textgreater{}''
+symbol. Typed letters were displayed in the screen in real time to
+provide visual feed-back to participants. Participants were allowed to
+correct their answer. Then, participants pressed the RETURN key to start
+and new trial. We excluded trials where participants did not type an
+existing word in the correspondent language, or did not type anything at
+all. Trials where the response was mistyped by only one character were
+counted as correct, as long as the respond did not correspond to a
+distinct word. Participants contributed a total of 8077 valid trials
+(5235 in Catalan, 2842 in Spanish). The task took approximately 15
+minutes to be completed.
 
@@
 
-Participants listened to one audio file in each trial.
-This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infancia of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat (\textbf{boersma2018?}).
-The average duration of the audios was NA (\emph{SD} = NA, \emph{Min} = \(\infty\), \emph{Max} = \(-\infty\)).
-The average lexical frequency of the words was 4.29 (\emph{SD} = 0.59, \emph{Min} = 2.87, \emph{Max} = 6.14).
-The average lexical frequency of the Catalan audios was 4.31 seconds (\emph{SD} = 0.59, \emph{Min} = 2.87, \emph{Max} = 5.68), and the average lexical frequency of the Spanish audios was 4.32 seconds (\emph{SD} = 0.56, \emph{Min} = 2.87, \emph{Max} = 5.68).
-Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT (\textbf{boada2013?}) and SUBTLEX-ESP (\textbf{cuetos2011?}), respectively, and are expressed as Zipf scores (\textbf{van-heuven2012?}; \textbf{zipf1949?}).
+Participants listened to one audio file in each trial. This audio file
+corresponded to a word in Catalan (for Spanish speakers, and for English
+participants allocated in the Catalan condition) or Spanish (for English
+speakers allocated in the Spanish condition). The audio files were the
+same ones used in child experiments conducted in the Laboratori de
+Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female
+bilingual from the Metropolitan Area of Barcelona in a child-directed
+manner. Catalan and Spanish words were recorded at 44,100 Hz in separate
+files in the same session, and then de-noised using Audacity and
+normalised at peak intensity using Praat (Broersma and Weenink 2021).
+The average duration of the audios was 1.20 (\emph{SD} = 0.18,
+\emph{Min} = 0.78, \emph{Max} = 1.58). The average duration of the
+Catalan audios was 1.23 seconds (\emph{SD} = 0.19, \emph{Min} = 0.80,
+\emph{Max} = 1.58), and the average duration of the Spanish audios was
+1.16 seconds (\emph{SD} = 0.15, \emph{Min} = 0.78, \emph{Max} = 1.53).
+Table 1 summarises the lexical frequency, phonological neighbourhood
+density and phonological overlap of the words included in the Catalan
+and the Spanish lists.
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lrrrrrrrrrr}
+\toprule
+ & \multicolumn{2}{c}{\textbf{Frequency (per million)}} & \multicolumn{2}{c}{\textbf{Frequency (Zipf score)}} & \multicolumn{2}{c}{\emph{\textbf{PTHN}}} & \multicolumn{2}{c}{\textbf{Consonant similarity}} & \multicolumn{2}{c}{\textbf{Vowel similarity}} \\
+ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9} \cmidrule(lr){10-11}
+ & Mean & \emph{SD} & Mean & \emph{SD} & Mean & \emph{SD} & Mean & \emph{SD} & Mean & \emph{SD} \\ 
+\midrule
+ENG-CAT & $46.80$ & $77.23$ & $4.31$ & $0.59$ & $3.26$ & $3.74$ & $0.34$ & $0.17$ & $0.17$ & $0.11$ \\ 
+ENG-SPA & $43.91$ & $67.53$ & $4.32$ & $0.56$ & $3.52$ & $4.10$ & $0.34$ & $0.15$ & $0.14$ & $0.10$ \\ 
+SPA-CAT & $53.32$ & $156.19$ & $4.23$ & $0.62$ & $1.62$ & $2.62$ & $0.42$ & $0.15$ & $0.27$ & $0.12$ \\ 
+ \bottomrule
+\end{longtable}
 
@@
 
-We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT).
-We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
+We modelled the probability of participants guessing the correct
+translation of each input word using a generalised multilevel Bayesian
+regression model with a Bernoulli logit link distribution. We first
+fitted a base model (Model 0) that only included the lexical frequency
+(\texttt{frequency}) of the target word as a fixed effect, with a random
+intercept per participant. Second, we extended the model to include
+\texttt{pthn} as a fixed effect, with a random slope by participant
+(Model 1). Third, we added the fixed effect \texttt{consonant\_ratio}
+and the \texttt{pthn:consonant\_ratio}, and random slopes for both
+effects by participant (Model 2). Third, we added the fixed effect
+\texttt{vowel\_ratio} and the \texttt{pthn:vowel\_ratio}, and random
+slopes for both effects by participant (Model 3). Finally, we fit a
+model that included the two-way interactions
+\texttt{pthn:consonant\_ratio} and \texttt{pthn:vowel\_ratio}, an their
+random slopes by participant (Model 4). Continuous predictors were
+standardised and categorical predictors were sum-coded (Schad et al.
+2020). Model 4 can be formally expressed as:
 
-We fit several Bayesian linear models (\textbf{sorensen2015?}), each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.
-Continuous predictors were standardised and categorical predictors were sum-coded (\textbf{schad2020?}).
-Missing predictor scores were imputed before fitting the models using multiple imputation (\textbf{vanbuuren2013?}).
-To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (LOO, (\textbf{vehtari2016?})).
-More information about the models and model comparison can be found in Appendix X.
+\$\$ \begin{align*}
 
-All analyses were performed in R environment (\textbf{rcoreteam2019?}).
-We used the \texttt{tidyverse} family of R packages to process data and to generate figures, the \texttt{mice} R package (\textbf{vanbuuren2013?}) for multiple imputation of missing data, and the \texttt{brms} R package (\textbf{burkner2017?}) to fit and compare models, which uses the Stan probabilistic programming language language (\textbf{carpenter2016?}).
+&\textbf{Likelihood}  \\
+y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
+
+&\textbf{Parameters}  \\
+
+logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
+& \beta_{1[p]} ~ Frequency_{i} ~ + \\
+& \beta_{2[p]} ~ PTHN_i ~ + \\
+& \beta_{3[p]} ~ Similarity_i ~ + \\
+& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
+
+\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
+\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
+&& \text{[participant-level coefficients]} \\ \\
+
+&\textbf{Prior}  \\
+
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 3) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ Cauchy(0, 3) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(5) && \text{[correlation between participant-level coefficients]} \\
+
+\end{align*} \$\$
+
+To test and account for cross-group differences, we included a random
+intercept for each group. We compared models using leave-one-out
+cross-validation (\emph{LOO}) (Vehtari, Gelman, and Gabry 2017). More
+information about the models and model comparison can be found in
+Appendix 2. All analyses were performed in R environment (RCore 2019).
+We used the tidyverse family of R packages to process data and to
+generate figures, and the brms R package (Bürkner 2017) using the
+cmdstanr backend to the Stan probabilistic language (Carpenter et al.
+2017) to estimate and compare the models (see Appendix 1 for mode
+details on the models).
 
@@
 
-\hypertarget{posterior-draws}{%
-\subsection{Posterior draws}\label{posterior-draws}}
+All models showed good out-of-sample predictive validity, as suggested
+by the fact that the expected log-probability density was many times
+larger than its associated standard error. Model 4, which included all
+main effects and the two-way interactions between PTHN and vowel
+similarity, and PTHN and consonant similarity, showed the best
+performance (see Table 2).
 
-\hypertarget{posterior-predictions}{%
-\subsection{Posterior predictions}\label{posterior-predictions}}
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lrrrrrr}
+\toprule
+ & \emph{\textbf{LOOELPD}} & \emph{\textbf{SE}} & \emph{\textbf{LOOIC}} & \textbf{\emph{SE}IC} & \emph{\textbf{LOOdiff}} & \emph{\textbf{SEdiff}} \\ 
+\midrule
+Model 3 & $-4,276.17$ & $45.33$ & $8,552.34$ & $90.66$ & - & - \\ 
+Model 2 & $-4,293.17$ & $45.33$ & $8,586.34$ & $90.65$ & $-17.00$ & $6.40$ \\ 
+Model 1 & $-4,308.60$ & $45.28$ & $8,617.19$ & $90.56$ & $-32.43$ & $9.07$ \\ 
+Model 0 & $-4,319.55$ & $45.11$ & $8,639.10$ & $90.21$ & $-43.38$ & $10.34$ \\ 
+ \bottomrule
+\end{longtable}
 
-\hypertarget{random-effects}{%
-\subsection{Random effects}\label{random-effects}}
+We now report the mean of the posterior distribution of each coefficient
+in Model 3, along with its associated measures of uncertainty. For
+interpretability, we transformed the estimates of the intercept using
+the inverse logit function so that the values are expressed in
+probability of correct response instead of log-odds, and we transformed
+the coefficients of the rest of predictors divided by four. Dividing a
+coefficient expressed in log-odds by four returns an approximate of the
+derivative of the logistic function indicating the maximum steepness of
+the logistic curve. This way, the coefficients are expressed as
+increases/decreases in probability of correct translation (Gelman, Hill,
+and Vehtari 2020).
 
-\hypertarget{item-effects}{%
-\subsubsection{Item effects}\label{item-effects}}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/posterior_fix-1.pdf}
+\caption{Estimated posterior distributions of coefficients in Model 3.
+A) Population-level effects. Distributions indicate the estimated
+posterior likelihood density of each coefficient. Credible intervals
+(CrI), represented with increasingly lighter segmentents in the
+distribution indicate the range of values that contain the true value
+with 95\%, 80\%, and 50\%, given the data we have collected. Points
+represent the mean of the distribution. The intercept has been
+transformed using the inverse logit to get the average probability of
+correct response. The rest of the coefficients have been transformed
+using the divide-by-four rule to get the maximum change in probability
+of correct response associated with a unit increase in each predictor
+variable. B) Participant-level coefficient variability. Our model
+estimated participant-level coefficients to account for the dependency
+between responses from the same participant. Distributions in this panel
+indicate the estimated variability across coefficients from different
+participants, expressed as standard deviations (SD). C) Correlation
+between participant-level effects. Our model allowed participant-level
+coefficients to co-vary. This panel represents the Pearson correlations
+between each pair of coefficients, expressed as the mean of the
+posterior distribution of each correlation. Coefficients are represented
+in the X- and Y-axis in the same order as indicated in the Y-axis of
+panels A and C.}
+\end{figure}
 
-\hypertarget{group-effects}{%
-\subsubsection{Group effects}\label{group-effects}}
+Overall, participants were 25.03\%, (\emph{SE} = 56.22\%, 95\%
+\emph{CrI} = {[}56.22\%, 56.22\%{]}) likely to produce correct
+translations. Every standard deviation increment in the translation's
+lexical frequency (\emph{SD} = 0.59) increased the probability of a
+correct responses in 3.90\% (\emph{SE} = 0.98\%, 95\% \emph{CrI} =
+{[}0.98\%, 0.98\%{]}). The number of the translation's more frequent
+phonological neighbours, on the other hand, decreased the probability of
+a correct responses in -4.60\% (\emph{SE} = -1.15\%, 95\% \emph{CrI} =
+{[}-1.15\%, -1.15\%{]}) for every increase in 1 \emph{SD} (3.65). The
+effect of phonological similarity was conditional to the phonological
+density of the translation. Phonological similarity barely increased the
+probability of a correct responses (3.05\%, \emph{SE} = 1\%, 95\%
+\emph{CrI} = {[}0.76\%, 1\%{]}) by itself. However, for every \emph{SD}
+increase in PTHN, phonological similarity increased in -2\%, (\emph{SE}
+= -0.62\%, 95\% \emph{CrI} = {[}-0.62\%, -0.62\%{]}) such probability.
 
-\hypertarget{correlations}{%
-\subsubsection{Correlations}\label{correlations}}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/marginal_effects-1.pdf}
+\caption{Expected mean posterior predictions.}
+\end{figure}
+
+\includegraphics{manuscript_files/figure-latex/accuracy-1.pdf}
 
@@
 
-\begin{verbatim}
-## Warning in utils::citation(x[pkg], auto = if (no_citations[pkg]) TRUE else
-## NULL): no date field in DESCRIPTION file of package 'multilex'
-\end{verbatim}
-
-\begin{verbatim}
-## Warning in utils::citation(x[pkg], auto = if (no_citations[pkg]) TRUE else
-## NULL): could not determine year for 'multilex' from package DESCRIPTION file
-\end{verbatim}
-
 \begingroup
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-broersma2021praat}{}}%
+Broersma, Paul, and David Weenink. 2021. {``Praat: Doing Phonetics by
+Computer {[}Computer Program{]}.''} \url{http://www.praat.org/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
+Bürkner, Paul-Christian. 2017. {``Brms: An r Package for Bayesian
+Multilevel Models Using Stan.''} \emph{Journal of Statistical Software}
+80 (1): 1--28.
+
+\leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
+Carpenter, Bob, Andrew Gelman, Matthew D Hoffman, Daniel Lee, Ben
+Goodrich, Michael Betancourt, Marcus A Brubaker, Jiqiang Guo, Peter Li,
+and Allen Riddell. 2017. {``Stan: A Probabilistic Programming
+Language.''} \emph{Grantee Submission} 76 (1): 1--32.
+
+\leavevmode\vadjust pre{\hypertarget{ref-collins1975spreading}{}}%
+Collins, Allan M, and Elizabeth F Loftus. 1975. {``A
+Spreading-Activation Theory of Semantic Processing.''}
+\emph{Psychological Review} 82 (6): 407.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de2000hard}{}}%
+De Groot, Annette MB, and Rineke Keijzer. 2000. {``What Is Hard to Learn
+Is Easy to Forget: The Roles of Word Concreteness, Cognate Status, and
+Word Frequency in Foreign-Language Vocabulary Learning and
+Forgetting.''} \emph{Language Learning} 50 (1): 1--56.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de1982relationship}{}}%
+De Haan, Henry J. 1982. {``The Relationship of Estimated
+Comprehensibility to the Rate of Connected Speech.''} \emph{Perception
+\& Psychophysics} 32 (1): 27--31.
+
+\leavevmode\vadjust pre{\hypertarget{ref-degroot1994forward}{}}%
+Degroot, Annette MB, Lucia Dannenburg, and Janet G Vanhell. 1994.
+{``Forward and Backward Word Translation by Bilinguals.''} \emph{Journal
+of Memory and Language} 33 (5): 600--629.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
+Gelman, Andrew, Jennifer Hill, and Aki Vehtari. 2020. \emph{Regression
+and Other Stories}. Cambridge University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-goldinger1989priming}{}}%
+Goldinger, Stephen D, Paul A Luce, and David B Pisoni. 1989. {``Priming
+Lexical Neighbors of Spoken Words: Effects of Competition and
+Inhibition.''} \emph{Journal of Memory and Language} 28 (5): 501--18.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de1992determinants}{}}%
+Groot, Annette M de. 1992. {``Determinants of Word Translation.''}
+\emph{Journal of Experimental Psychology: Learning, Memory, and
+Cognition} 18 (5): 1001.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroll1988lexical}{}}%
+Kroll, Judith F, Janet Curley, and others. 1988. {``Lexical Memory in
+Novice Bilinguals: The Role of Concepts in Retrieving Second Language
+Words.''} \emph{Practical Aspects of Memory} 2 (389-395): 8.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroll1994category}{}}%
+Kroll, Judith F, and Erika Stewart. 1994. {``Category Interference in
+Translation and Picture Naming: Evidence for Asymmetric Connections
+Between Bilingual Memory Representations.''} \emph{Journal of Memory and
+Language} 33 (2): 149--74.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lotto1998effects}{}}%
+Lotto, Lorella, and Annette MB De Groot. 1998. {``Effects of Learning
+Method and Word Type on Acquiring Vocabulary in an Unfamiliar
+Language.''} \emph{Language Learning} 48 (1): 31--69.
+
+\leavevmode\vadjust pre{\hypertarget{ref-luce1998recognizing}{}}%
+Luce, Paul A, and David B Pisoni. 1998. {``Recognizing Spoken Words: The
+Neighborhood Activation Model.''} \emph{Ear and Hearing} 19 (1): 1.
+
+\leavevmode\vadjust pre{\hypertarget{ref-luce1990similarity}{}}%
+Luce, Paul A, David B Pisoni, and Steven D Goldinger. 1990.
+{``Similarity Neighborhoods of Spoken Words.''}
+
+\leavevmode\vadjust pre{\hypertarget{ref-peirce2019psychopy2}{}}%
+Peirce, Jonathan, Jeremy R Gray, Sol Simpson, Michael MacAskill, Richard
+Höchenberger, Hiroyuki Sogo, Erik Kastman, and Jonas Kristoffer
+Lindeløv. 2019. {``PsychoPy2: Experiments in Behavior Made Easy.''}
+\emph{Behavior Research Methods} 51 (1): 195--203.
+
+\leavevmode\vadjust pre{\hypertarget{ref-potter1984lexical}{}}%
+Potter, Mary C, Kwok-Fai So, Barbara Von Eckardt, and Laurie B Feldman.
+1984. {``Lexical and Conceptual Representation in Beginning and
+Proficient Bilinguals.''} \emph{Journal of Verbal Learning and Verbal
+Behavior} 23 (1): 23--38.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rcore2019r}{}}%
+RCore, TEAM. 2019. {``R: A Language and Environment for Statistical
+Computing. R Foundation for Statistical Computing, Austria.''}
+
+\leavevmode\vadjust pre{\hypertarget{ref-schad2020capitalize}{}}%
+Schad, Daniel J, Shravan Vasishth, Sven Hohenstein, and Reinhold Kliegl.
+2020. {``How to Capitalize on a Priori Contrasts in Linear (Mixed)
+Models: A Tutorial.''} \emph{Journal of Memory and Language} 110:
+104038.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schepens2012distributions}{}}%
+Schepens, Job, Ton Dijkstra, and Franc Grootjen. 2012. {``Distributions
+of Cognates in Europe as Based on Levenshtein Distance.''}
+\emph{Bilingualism: Language and Cognition} 15 (1): 157--66.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vehtari2017practical}{}}%
+Vehtari, Aki, Andrew Gelman, and Jonah Gabry. 2017. {``Practical
+Bayesian Model Evaluation Using Leave-One-Out Cross-Validation and
+WAIC.''} \emph{Statistics and Computing} 27 (5): 1413--32.
+
+\leavevmode\vadjust pre{\hypertarget{ref-warren1970perceptual}{}}%
+Warren, Richard M. 1970. {``Perceptual Restoration of Missing Speech
+Sounds.''} \emph{Science} 167 (3917): 392--93.
+
 \end{CSLReferences}

### 2021-11-25 Rmd/manuscript_appendix.Rmd
Starts manuscript

--- 2021-09-29 Rmd/manuscript.Rmd
@@
 ---
-title             : "The role of cognateness in non-native spoken word recognition"
-shorttitle        : "Cognateness and non-native word recognition"
-
-author: 
-  - name          : "Gonzalo Garcia-Castro"
-    affiliation   : "1"
-    corresponding : yes    # Define only one corresponding author
-    address       : "Ramon Trial Fargas, 25-27, 08005 Barcelona"
-    email         : "gonzalo.garciadecastro@upf.edu"
-  - name          : "Serene Siow"
-    affiliation   : "2"
-  - name          : "Nuria Sebastian-Galles"
-    affiliation   : "1"
-  - name          : "Kim Plunkett"
-    affiliation   : "2"
-
-authornote: |
-  GG and SS contributed equally and share first authorship.
-
-abstract: |
-  There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
-For example, bilinguals produce and translated cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
-Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
-These measures partially ignore potential sources of variability during lexical access like vowel vs.
-consonant overlap, overlap in stressed syllables, onset overlap or the distance in features between both translations.
-In this study, we explored the impact of some of these variables on non-native word translation task: Spanish and English participants listened to non-native words (Catalan or Spanish) and were prompted to type their translation in their native language.
-Critically, participants where unfamiliar with the testing language, ensuring that they were not able to translate words based on previous knowledge on their meaning, leaving phonological information as the only cue participants were able to exploit to translate words to their native language correctly.
-We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords          : "cognate, word recognition, translation, non-native, spoken word recognition"
-wordcount         : "X"
-
-bibliography      : ["r-references.bib", "references.bib"]
-
-floatsintext      : no
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-linenumbers       : yes
-mask              : no
-draft             : no
-
-documentclass     : "apa6"
-classoption       : "man"
-output            :
-  word_document:
-    reference_docx: "manuscript_template.docx"
 ---
 
-```{r setup, include = FALSE}
-library(papaja)
-```
+```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
+Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.
+Dots and whiskers represent the average accuracy and 95% confidence interval of each word.
+Accuracy is plotted separately for each group."}
 
-```{r analysis-preferences, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-# Seed for random number generation
-set.seed(42)
-knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
-```
+responses_item <- responses %>%
+  group_by(trial_id, group, word) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  left_join(select(stimuli, group, trial_id, word2, ipa1, ipa2)) %>% 
+  mutate(
+    word = paste0(word, " (", ipa1, ") / ", word2, " (", ipa2, ")"),
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    word_label = reorder_within(word, by = correct_prop, within = group)
+  )
 
-```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
-library(tidyverse)
-library(data.table)
-library(tidybayes)
-library(gt)
-library(lubridate)
-library(here)
-
-tar_load(participants)
-tar_load(stimuli)
-tar_load(responses)
-tar_load(model_fits)
-tar_load(model_loos)
+ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5, colour = "black") +
+  geom_errorbar(width = 0) +
+  geom_point(size = 0.5) +
+  labs(y = "Proportion of correct responses", x = "Word", colour = "Group") +
+  coord_flip() +
+  scale_color_manual(values = c("#1E88E5", "#ff2976", "#FFC20A")) +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  scale_x_discrete(labels = function(x) str_remove(x, "___ENG-CAT|___ENG-SPA|___SPA-CAT")) +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_text(size = 5, colour = "black"),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9, colour = "black"),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  )
 
 ```
-
-# Methods
-
-## Participants
-
-Data collection took place from `r format(min(participants$date), "%dth %B, %Y")` to  `r format(max(participants$date), "%dth %B, %Y")`.
-We collected data from `r nrow(participants)` participants ($M_{age}$ = `r printnum(mean(participants$age, na.rm = TRUE))`, $SD_{age}$ = `r printnum(sd(participants$age, na.rm = TRUE))`, $Range_{age}$ = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
-`r filter(participants, country=="UK") %>% nrow()` participants were British English native speakers living in United Kingdom (`r filter(participants, country=="UK", sex =="Female") %>% nrow()` female), and `r filter(participants, country=="UK") %>% nrow()` participants were Spanish native speakers living in Spain (`r filter(participants, country=="Spain", sex =="Female") %>% nrow()` female).
-Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
-Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants were asked to complete the experiment in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r filter(participants, l2oral>4 | l2written>4) %>% nrow()`), b) were diagnosed with a language (*n* = `r filter(participants, impairment) %>% nrow()`)^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as "Do you have normal or corrected-to-normal VISION?
-(Yes/No)" in English, and as "¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No)".
-However, the proportion of Spanish participants that reported visual impairment was implausibly large (*n* = `r nrow(filter(participants, country %in% "Spain", vision))`, `r printnum((1-nrow(filter(participants, country %in% "Spain", !vision))/nrow(filter(participants, country=="Spain")))*100)`%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.], or c) did not contribute more than 80% of valid trials (*n* = `r sum(participants$invalid_participant_trials, na.rm = TRUE)`).
-
-## Procedure
-
-The experiment was implemented online using Psychopy/Pavlovia [@peirce2019].
-Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-Participants were informed about the aims of the study and gave informed consent for participating.
-Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
-Third, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
-They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the list of Catalan or Spanish trials.
-Participants in the Catalan list were presented with `r count(stimuli, group)$n[1]` trials, and participants in the Spanish list were presented with `r count(stimuli, group)$n[2]` trials.
-
-Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio ended.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
-Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
-Participants were allowed to correct their answer.
-Then, participants pressed the RETURN key to start and new trial.
-We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Participants contributed a total of `r distinct(responses, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(responses, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(responses, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish).
-The task took approximately 15 minutes to be completed.
-
-## Stimuli
-
-Participants listened to one audio file in each trial.
-This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infancia of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@boersma2018].
-The average duration of the audios was `r printnum(mean(stimuli$duration), digits = 2)` (*SD* = `r printnum(sd(stimuli$duration), digits = 2)`, *Min* = `r printnum(min(stimuli$duration), digits = 2)`, *Max* = `r printnum(max(stimuli$duration), digits = 2)`).
-The average lexical frequency of the words was `r printnum(mean(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)` (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli$frequency_zipf, na.rm = TRUE), digits = 2)`).
-The average lexical frequency of the Catalan audios was `r printnum(mean(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-CAT",]$frequency_zipf, na.rm = TRUE), digits = 2)`), and the average lexical frequency of the Spanish audios was `r printnum(mean(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)` seconds (*SD* = `r printnum(sd(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Min* = `r printnum(min(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`, *Max* = `r printnum(max(stimuli[stimuli$group=="ENG-SPA",]$frequency_zipf, na.rm = TRUE), digits = 2)`).
-Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT [@boada2013] and SUBTLEX-ESP [@cuetos2011], respectively, and are expressed as Zipf scores [@van-heuven2012; @zipf1949].
-
-## Data analysis
-
-We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT).
-We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
-
-We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.
-Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020].
-Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013].
-To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (LOO, @vehtari2016).
-More information about the models and model comparison can be found in Appendix X.
-
-All analyses were performed in R environment [@rcoreteam2019].
-We used the `tidyverse` family of R packages to process data and to generate figures, the `mice` R package [@vanbuuren2013] for multiple imputation of missing data, and the `brms` R package [@burkner2017] to fit and compare models, which uses the Stan probabilistic programming language language [@carpenter2016].
-
-# Results
-
-## Posterior draws
-
-## Posterior predictions
-
-## Random effects
-
-### Item effects
-
-### Group effects
-
-```{r group_effects_table}
-```
-
-### Correlations
-
-# Discussion
-
-\newpage
-
-# References
-```{r create_r-references}
-r_refs(file = "r-references.bib")
-```
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup

### 2021-11-25 Rmd/report.Rmd
Starts manuscript

@@
 library(patchwork)
+library(tidytext)
 library(janitor)

### 2021-11-25 R/02_models.R

@@ -20,7 +20,8 @@ get_model_fits <- function(
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
         save_model = here("Stan", "fit_responses_0.stan"),
-        file = here("Results", "fit_responses_0.rds")
+        file = here("Results", "fit_responses_0.rds"), 
+        control = list(adapt_delta = 0.95)
     )
     fit_1 <- brm(
         model_formulas$f_1,
@@ -29,7 +30,8 @@ get_model_fits <- function(
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
         save_model = here("Stan", "fit_responses_1.stan"),
-        file = here("Results", "fit_responses_1.rds")
+        file = here("Results", "fit_responses_1.rds"), 
+        control = list(adapt_delta = 0.95)
     )
     fit_2 <- brm(
         model_formulas$f_2,
@@ -38,7 +40,8 @@ get_model_fits <- function(
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
         save_model = here("Stan", "fit_responses_2.stan"),
-        file = here("Results", "fit_responses_2.rds")
+        file = here("Results", "fit_responses_2.rds"), 
+        control = list(adapt_delta = 0.95)
     )
     fit_3 <- brm(
         model_formulas$f_3,
@@ -47,27 +50,28 @@ get_model_fits <- function(
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
         save_model = here("Stan", "fit_responses_3.stan"),
-        file = here("Results", "fit_responses_3.rds")
-    )
-    fit_4 <- brm(
-        model_formulas$f_4,
-        data = responses,
-        prior = model_prior,
-        save_pars = save_pars(all = TRUE),
-        backend = "cmdstanr",
-        save_model = here("Stan", "fit_responses_4.stan"),
-        file = here("Results", "fit_responses_4.rds")
+        file = here("Results", "fit_responses_3.rds"), 
+        control = list(adapt_delta = 0.95)
     )
     
-    fits <- list(fit_0 = fit_0, fit_1 = fit_1, fit_2 = fit_2, fit_3 = fit_3, fit_4 = fit_4)
+    
+    fits <- list(fit_0 = fit_0, fit_1 = fit_1, fit_2 = fit_2, fit_3 = fit_3)
     
     return(fits)
 }
 
 
-# compre models
+# compare models
[…]

### 2021-11-25 R/compare_methods.R

@@ -1,0 +1,45 @@
+tar_load(responses)
+
+library(stringdist)
+
+threshold <- 0.6
+
+d <- responses %>% 
+    mutate(
+        overlap_lv = stringsim(input_text, word1, method = "lv"),
+        overlap_osa = stringsim(input_text, word1, method = "osa"),
+        overlap_hamming = stringsim(input_text, word1, method = "hamming"),
+        overlap_jw = stringsim(input_text, word1, method = "jw"),
+        overlap_cosine = stringsim(input_text, word1, method = "cosine")
+    ) %>% 
+    mutate(
+        is_correct_exact = input_text==word2,
+        is_correct_manual = correct,
+        is_correct_lv = overlap_lv >= threshold,
+        is_correct_osa = overlap_lv >= threshold,
+        is_correct_hamming = overlap_lv >= threshold,
+        is_correct_cosine = overlap_lv >= threshold
+    ) %>% 
+    select(group, trial_id, word, participant, matches("is_correct")) %>% 
+    pivot_longer(c(matches("is_correct"), -is_correct_manual), names_to = "method", values_to = "is_correct") %>% 
+    mutate(method = str_to_sentence(str_remove(method, "is_correct_"))) %>% 
+    group_by(group, word, method) %>% 
+    summarise(
+        is_correct_manual = sum(is_correct_manual),
+        is_correct = sum(is_correct),
+        n = n(),
+        prop_correct_manual = is_correct_manual/n,
+        prop_correct = is_correct/n,
+        .groups = "drop"
+    )
+    
+ggplot(d, aes(prop_correct_manual, prop_correct, colour = group)) +
+    facet_wrap(~method) +
+    # geom_rug(sides = "tr") +
+    geom_point(size = 0.75, alpha = 0.5) +
+    labs(x = "% correct (manual classification)",
+         y = "% correct (automatic classification) \n Similarity threshold: 60%") +
+    scale_x_continuous(labels = percent) +
+    scale_y_continuous(labels = percent) +
+    coord_equal() +
+    theme_ggdist()

### 2021-11-25 R/utils.R

@@ -60,142 +60,6 @@ theme_custom <- function(){
     )
 }
 
-# extract formants
-extract_formants <- function(
-    file,
-    time_step = 0.001,
-    n = 2,
-    max_freq = 5500,
-    window_length = 0.001,
-    pre = 50,
-    include_frames = TRUE,
-    include_time = TRUE,
-    dec = 3,
-    include_intensity = TRUE,
-    dec_intensity = 5,
-    include_n_formats = FALSE,
-    dec_freq = 3,
-    include_bandwidths = FALSE) {
-    require(PraatR)
-    require(rlang)
-    require(purrr)
-    require(stringr)
-    require(dplyr)
-    require(data.table)
-    require(janitor)
-    
-    file_paths <- list.files({{ file }}, pattern = ".wav", full.names = TRUE)
-    file_names <- list.files({{ file }}, pattern = ".wav", full.names = FALSE)
-    
-    pb <- progress_estimated(length(file_paths)*2)
-    arguments1 <- list({{ time_step }}, {{ n }}, {{ max_freq }}, {{ window_length }}, {{ pre }})
-    arguments2 <- list({{ include_frames }}, {{ include_time }}, {{ dec }}, {{ include_intensity }}, {{ dec_intensity }}, {{ include_n_formats }}, {{ dec_freq }}, {{ include_bandwidths }}) 
-    
-    dir <- tempdir()
-    dir_formants <- paste(dir, file_names, sep = "/") %>% str_replace(., ".wav", ".Formant")
-    dir_formants_table <- paste(dir, file_names, sep = "/")  %>% str_replace(., ".wav", ".txt")
-    
-    
-    fun1 <- function(x, y) {
-        pb$tick()$print()
-        praat("To Formant (burg)...",
-              arguments = arguments1,
-              input = x,
-              output = y,
-              overwrite = TRUE)
-    }
-    
-    fun2 <- function(x, y){
-        pb$tick()$print()
-        praat("Down to Table...",
-              arguments = arguments2,
-              input = x,
-              output = y,
-              filetype = "tab-separated",
-              overwrite = TRUE)
-    }
-    
-    invisible(
-        map2(.x = as.list(file_paths),
-             .y = as.list(dir_formants),
-             .f = ~fun1(.x, .y))
-    )
-    
-    invisible(
-        map2(.x = as.list(dir_formants),
-             .y = as.list(dir_formants_table),
[…]

### 2021-11-25 _targets.R

@@ -11,20 +11,21 @@ tar_option_set(
     packages = c(
         "dplyr", "tidyr", "stringr", "ggplot2", "tibble", "forcats", "multilex", "keyring",
         "readxl", "janitor", "mice", "here", "lubridate", "purrr", "scales", "stringdist",
-        "brms", "tidybayes", "gt", "patchwork", "wesanderson", "papaja", "knitr", "data.table"
+        "brms", "tidybayes", "gt", "patchwork", "wesanderson", "papaja", "knitr", "data.table",
+        "audio", "tidytext"
     )
 )
 
 list(
     
-    # items
+    # stimuli
     tar_target(clearpond_path, list(`ENG-CAT` = here("Data", "clearpond", "clearpond_english.csv"), `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"), `SPA-CAT` = here("Data", "clearpond", "clearpond_spanish.csv")),),
     tar_target(clearpond, get_clearpond(clearpond_path)),
-    tar_target(trace_path, here("Stimuli", "trace.xlsx")),
-    tar_target(trace, get_trace(trace_path)),
     tar_target(stimuli_path, here("Stimuli", "trials.xlsx")),
     tar_target(levenshtein, get_levenshtein(stimuli_path = stimuli_path)),
-    tar_target(stimuli, get_stimuli(stimuli_path = stimuli_path, clearpond = clearpond, trace = trace, levenshtein = levenshtein)),
+    tar_target(audios_path, here("Stimuli/Sounds")),
+    tar_target(durations, get_duration(stimuli_path = stimuli_path, audios_path = audios_path)),
+    tar_target(stimuli, get_stimuli(stimuli_path = stimuli_path, clearpond = clearpond, levenshtein = levenshtein, durations = durations)),
     tar_target(practice_trials, c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)),
     
     # responses
@@ -40,11 +41,23 @@ list(
     tar_target(
         model_formulas,
         formulas <- list(
-            f_0 = bf(correct ~ 1 + frequency + (1 | participant), family = bernoulli(link = "logit")),
-            f_1 = bf(correct ~ 1+ frequency + pthn + (1 + pthn | participant), family = bernoulli(link = "logit")),
-            f_2 = bf(correct ~ 1 + frequency + pthn*consonant_ratio + (1 + pthn*consonant_ratio | participant), family = bernoulli(link = "logit")),
[…]

### 2021-11-26 Rmd/manuscript.Rmd
knit manuscript to PDF

@@
 title: "The role of cognateness in non-native spoken word recognition"
+
+shorttitle: Cognateness and non-native word recognition
+
 author:
@@
   email: gonzalo.garciadecastro@upf.edu
+  role: 
+    - Conceptualization
+    - Data analysis
+    - Writing - Original Draft Preparation
+    - Writing - Review & Editing
 - name: Serene Siow
   affiliation: '2'
+  role:
+    - Conceptualization
+    - Writing - Review & Editing
 - name: Nuria Sebastian-Galles
   affiliation: '1'
+  role:
+    - Conceptualization
+    - Writing - Review & Editing
 - name: Kim Plunkett
   affiliation: '2'
-shorttitle: Cognateness and non-native word recognition
-output:
-  word_document:
-    reference_docx: manuscript_template.docx
+  role:
+    - Conceptualization
+    - Writing - Review & Editing
+affiliation:
+  - id            : "1"
+    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
+  - id            : "2"
+    institution   : "Department of Experimental Psychology, University of Oxford"
+
+authornote: |
+  GG and SS contributed equally and share first authorship.
+
 abstract: |
   There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
-For example, bilinguals produce and translated cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
+For example, bilinguals produce and translate cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
 Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
@@
   <!-- https://tinyurl.com/ybremelq -->
+
 keywords: cognate, word recognition, translation, non-native, spoken word recognition
 wordcount: X
+
 bibliography:
@@
   - references.bib
-floatsintext: no
+floatsintext: yes
 figurelist: no
@@
 classoption: man
-authornote: |
-  GG and SS contributed equally and share first authorship.
+output:
+  papaja::apa6_pdf:
+    keep_tex: FALSE
 ---
@@
   warning = FALSE,
-  cache = FALSE, 
-  cache.vars = FALSE,
-  cache.rebuild = FALSE,
-  cache.lazy = FALSE,
   results = "asis", 
-  dpi = 300
+  dpi = 500
 )
@@
 tar_load(model_loos)
+tar_load(posterior_draws_fixed)
+tar_load(posterior_epreds_fixed)
+tar_load(posterior_epreds_random)
 
@@
 But even in the most ideal of the situations, the processes engaged in word recognition do not occur without uncertainty.
-The speech input activates multiple candidate lexical entries based on their similarity with the signal, and for word recognition to take place, the lexical system must select one of the candidates.
+The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
 There is extensive literature about how the number of candidates, their frequency, and their phonological and semantic similarity with the target word impacts the dynamics of word recognition.
-However, less is known about how these factors affect the recognition of words in a non-native language.
+But less is known about how these factors affect the recognition of words in a non-native language.
 
-Listening to speech in a non-native language is more costly than doing so in the native language, even if one is proficient in the non-native language.
+Listening to speech in a non-native language is more costly than doing so in the native language, even if one is proficient in such language.
 The source of this increased cognitive effort is likely to originate at multiple levels.
-For instance, some sounds in the speech signal do not correspond to any phoneme in the native language, segmenting the speech signal relies on previous familiarity with word forms or with the statistical regularities of the language, or word order might be different in both languages (e.g., subject-object-verb vs.
+Some examples: (1) some sounds in the speech signal do not correspond to any phoneme in the native language, (2) segmenting the speech signal relies on previous familiarity with word forms or with the statistical regularities of the language: (3) word order might be different in both languages (e.g., subject-object-verb vs.
 subject-verb-object.).
 The listener, however, is rarely completely naïve to the language they are listening to.
-All languages share, to some extent, similarities, frequently due to their typological closeness, which can be exploited by listeners of a non-native language.
+All languages share, to some extent, similarities, frequently due to their typological closeness.
+These similarities which can be exploited by non-native listeners.
 
-Lexical similarity, for example, provides useful information to sequential language learners.
-Romance language like Spanish, Italian, French or Catalan, share many similarities at the lexical level in the form of cognates [@schepens2012distributions].
+One of such commonalities occurs at the lexical level, in the for of cognateness.
 Cognates are cross-language synonyms whose form (e.g., phonology, orthography, signature) is similar.
 The cause of this similarity is frequently attributed to a shared etymological origin.
-This is the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively)^[Some form-similar cross-language synonyms are technically not cognates.
+Romance languages like Spanish, Italian, French or Catalan, share many cognates [@schepens2012distributions].
+For example, this is the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively)^[Some form-similar cross-language synonyms are technically not cognates.
 For example, *sun* and *sol* (in Spanish), share their phonological onset, but their etymology points to different origins.
@@
 It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
+Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access their lexicon in a language non-selective way: during word production and comprehension, both lexical representations are activated, and the form similarity between both impacts participants' performance in naming or lexical decision tasks.
+For instance, in their seminal study, @costa2000cognate asked Spanish-Catalan bilinguals to name pictures in Spanish.
+Unbeknownst to participants, half of the pictures' associated labels were cognates in Spanish and Catalan (e.g., *puerta*-*porta*) while the other half were non-cognates (e.g., *mesa*-*taula*).
+Surprisingly, participants named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
+This suggested that bilinguals activated picture's word representations in both languages, and that the phonological overlap between both labels facilitated the naming process.
+
 There is evidence that cognates are learnt more easily in L2 than non-cognates.
-@de2000hard presented 40 Dutch natives 60 pairs of translation equivalents.
+For instance, @de2000hard presented 40 Dutch natives 60 pairs of translation equivalents.
 In each learning trial, two words were presented in a screen side-by-side: one in Dutch and one pseudoword.
@@
 
-Cognates are not only learnt faster than non-cognates, but also translated faster.
-Early accounts of bilingual lexical access described two possible routes for translating a word from L1 to L2 or *vice versa*: (1) via direct links between L1-L2 links or (2) through word-concept links.
-Both translation routes are available during translation [@potter1984lexical; @kroll1988lexical].
-These models assumed that the strength of the word-concept associations in both L1 and L2 is a function of their frequency of use, in order to account for the influence of  both the lexical frequency of the presented word and the target word in translation tasks on participant's performance.
-Under the assumption that bilinguals use L1 more often, @kroll1994category proposed the asymmetric hypothesis, predicting stronger word-concept associations between L1 representations compared to L2 representations.
-This would make L2-to-L1 translation more reliant to the direct translation route (L1-L2 association) than L1-to-L2 translation.
-This prediction has three consequences: (1) L2-to-L1 translation is expected to be faster than L1-to-L2 translation (as it does not rely on conceptual mediation during translation), and (2) since word-to-word connections (like the L1-L2 one) occurs at the lexical level (and not at the conceptual level), it should be sensitive to the form-similarity between both words, and therefore to cognateness, and (3) since low-proficiency bilinguals rely more on the direct route during translation, and this route is sensitive to cognateness, their performance during backward translation tasks should not only be better than during forward translation, but also enhanced by the cognate status of L1 and L2 representations.
+Cognates are not only learnt faster than non-cognates, but translated faster too.
+The mechanisms behind this effect, though, are still unclear.
+Early accounts of bilingual lexical access suggested that low-proficiency learners first established links between newly learnt words in L2 and their meanings through their L1 translation equivalents [@potter1984lexical].
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
+The Revised Hierarchical Model [RHM; @kroll1994category] captured this assumption and predicted that translating words from L1 to L2 (forward translation) should take longer than translating from L2 to L1 (backward translation).
+The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, while forward translation would rely more strongly on the mediation between the concept and the two word forms.
+One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words iduring backward translation.
 
@@
 In each trial, participants were presented visually with a word in English or Dutch, and were asked to speak out loud its translation in the other language as soon as possible.
-The authors reported three main findings: (1) reaction times and accuracy were roughly equivalent in across both forward and backward conditions (although slightly faster in the former), (2) semantic variables, such as imageability, were positively associated with participants' performance more strongly in the forward translation conditional, although the effect size was small, (3) cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward translation than during forward translation, and (4) when translating cognates participants' performance across both conditions was less sensitive to semantic variables than when translating non-cognates.
-These results prompted the authors to reject a hard version @kroll1994category's asymmetric hypothesis, and suggest that both conceptually mediated and direct translation routes are active during translation, but backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness than forward translation.
+The authors reported three main findings: (1) translation times and accuracy were roughly equivalent in across both forward and backward conditions (although slightly faster in the former), (2) semantic variables, such as imageability, were positively associated with participants' performance more strongly in the forward translation conditional, although the effect size was small, (3) cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward translation than during forward translation, and (4) when translating cognates participants' performance across both conditions was less sensitive to semantic variables than when translating non-cognates.
+These results prompted the authors to reject a hard version @kroll1994category's account, and suggested that both conceptually mediated and direct translation routes are active during translation, but backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness than forward translation.
 When the authors tested a group of bilinguals with higher proficiency in English, their results pointed in the same direction.
+Subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language], contrary to the predictions of the RHM.
 
-The idealised lexicon considered by the models describe above only considers word-word connections between translation equivalents.
+The idealised lexicon considered by the models described above only considers word-word connections between translation equivalents.
 More recent studies have highlighted the role of the rich network of connections that given word establishes with other phonologically or conceptually related words.
-@collins1975spreading introduced this notion into their theory of semantic processing, and later @luce1998recognizing operationalised this dimension in their Neighbourhood Activation Model (NAM).
+This notion was first introduced in monolingual research by @collins1975spreading in their theory of semantic processing, and later @luce1998recognizing formalised this dimension in their Neighbourhood Activation Model (NAM).
 In this model, lexical selection is mediated not only by the lexical frequency of the target word, and the number of activated candidates, but also by the frequency of the candidates.
@@
 
-The role of neighbourhood lexical frequency on bilingual spoken word recognition is unclear.
+The fact that phonological neighbourhoods and their structure impacts lexical processing poses an important question for bilingualism research.
+Are word representations in one language part of phonological or orthographic neighbours in the other language?
+@van1998orthographic addressed this issue in an experimental series and proposed the Bilingual Interactive Activation (BIA, later revised and relabelled as BIA+).
+In line with connectionist accounts of lexical processing this model advocated for lexical representations to establish both excitatory and inhibitory connections with other representations from the same of the other language [@mcclelland1981interactive], forming an integrated lexicon for both languages.
+@van1998orthographic tested Dutch-English bilinguals in a progressive demasking task (in each trial a words in presented in increasingly longer time windows interrupted by a checkerboard until the participant presses a button to type the word in a keyboard) or a lexical decision task (words a visually presented and the participant must decide as quickly as possible whether its a word in the target language or a non-word).
+The authors manipulated by size of the orthographic neighbourhood size of the words presented in its language (target language) or in the other language (non-target language).
+Interestingly, the authors reported participants' performance to be affected by the neighbourhood size of the presented word in the non-target language across both tasks.
+Specifically, participants took longer to respond to words with high neighbourhood density in the non-target language, even when participants completed the task in the target language exclusively (e.g., were presented with Dutch words exclusively).
+Monolinguals, on the other hand, were only affected by neighbourhood size in the target language.
+This provided evidence of cross-language interference between the word representation in the non-target language and the those in the target language.
+
+More recently, @dijkstra2019multilink developed a model, Multilink, that integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
+This model
+
+These studies tested the role of neighbourhood lexical frequency in samples of fairly proficient (i.e., balanced) bilinguals, and addressed this issue mostly at the orthographic level.
+Although these models contemplate the role of phonology in their accounts, the fact the it plays a secondary role in them limits their theoretical generalisability.
+One of the main reasons why a substantial part of the literature has avoided phonology as their modality of interest stems from the methodological level: working with phonological transcriptions is computationally challenging, and perhaps more critically, there is no clear way of measuring the similarity/dissimilarity between phonological forms of words within or across languages.
+Any possible way of measuring the overlap between two phonological forms at the surface level (e.g., between phonological transcriptions of the words) is very likely to neglect important sources of perceived similarity/dissimilarity.
+For example, one could measure
+
+spoken word recognition is unclear.
 Across several modalities, bilinguals' recognition of words in a language is affected by form-similarity and frequency of words in the other language.
@@
 Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`.
-We collected data from `r length(unique(participants$participant))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, *Min* = `r printnum(min(participants$age, na.rm = TRUE))`, *Max* = `r printnum(max(participants$age, na.rm = TRUE))`).
-`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA") & participants$sex=="Female"]))`), and `r length(unique(participants$participant[participants$group %in% c("SPA-CAT")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant[participants$group %in% c("SPA-CAT") & participants$sex=="Female"]))` female).
+We collected data from `r length(unique(participants$participant))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
+`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant[participants$group %in% c("SPA-CAT")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant[participants$group %in% c("SPA-CAT") & participants$sex=="Female"]))` female).
 Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
 Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants were asked to complete the experiment in a quiet place with good internet connection.
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the local ethical committee (XXXXXXXXXXXX).
+Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
 We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`) , or c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
@@
 The experiment was implemented online using Psychopy/Pavlovia [@peirce2019psychopy2].
-Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-Participants were informed about the aims of the study and gave informed consent for participating.
-This experiment was  Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
-Third, participants completed the experimental task.
+Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla).
+After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+Then, participants completed the experimental task.
 Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
@@
 Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
+
+```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%"}
+
+include_graphics(here("Figures", "design.png"))
+
+```
 
@@
 
+```{r stimulilengths, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+lengths <- stimuli %>% 
+  select(trial_id, group, ipa1, ipa2, word1, word2) %>% 
+  mutate_at(vars(ipa1:word2), nchar) %>% 
+  group_by(group) %>% 
+  summarise_at(vars(ipa1, ipa2, word1, word2), list(mean = mean, sd = sd, min = min, max = max)) %>%
+  split(.$group) %>% 
+  map(select, -group) %>% 
+  set_names(make_clean_names(names(.))) %>% 
+  map(unlist)
+```
+
+We arranged two lists of words: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
+Words in the Catalan list (listened to by participants) were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
+Their translations to English (typed by participants in the keyboard) were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), and their translations to Spanish were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
+Words in the Spanish list were `r printnum(lengths$eng_spa["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_spa["ipa1_sd"])`, Range = `r printnum(lengths$eng_spa["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_spa["ipa1_max"], digits = 0)`).
+Their translations to English were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
+
+We extracted the lexical frequencies from SUBTLEX-UK for English words [@van2014subtlex], SUBTLEX-ESP for Spanish words [@cuetos2011subtlex], and SUBTLEX-CAT for Catalan words [@boada2020subtlex].
+We extracted or transformed scores as/into Zipf scores [@van2014subtlex] to so correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses [@van2014subtlex].
+
+We retrieved PTHN scores from the CLEARPOND database [@marian2012clearpond].
+PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
+CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target word in only one addition, deletion, or substitution.
+PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
+
+Finally, we measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist].
+This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
+First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical.
+This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
+The result is a score that ranges from 0 to 1, where 0 indicates no similarity between the two strings and one indicates that both strings are identical.
+We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
+Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+
 Participants listened to one audio file in each trial.
@@
 Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the audios was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, *Min* = `r printnum(min(stimuli$duration, na.rm = TRUE))`, *Max* = `r printnum(max(stimuli$duration, na.rm = TRUE))`).
-The average duration of the Catalan audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, *Min* = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, *Max* = `r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, *Min* = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, *Max* = `r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
-Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+The average duration of the audios was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
+The average duration of the Catalan audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
 
-```{r stimuli_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-stimuli %>% 
-  select(group, trial_id, frequency, frequency_zipf, pthn, consonant_ratio, vowel_ratio) %>% 
+stimuli_table <- stimuli %>% 
+  select(group, trial_id, frequency, frequency_zipf, pthn, lv) %>% 
   group_by(group) %>% 
-  summarise_at(vars(frequency:vowel_ratio), list(mean = ~mean(., na.rm = TRUE),  sd = ~sd(., na.rm = TRUE))) %>% 
+  summarise_at(
+    vars(frequency:lv), 
+    list(
+      mean = ~mean(., na.rm = TRUE),  
+      sd = ~sd(., na.rm = TRUE)
+    )
+  ) %>%
   gt(rowname_col = "group") %>% 
-  fmt_number(frequency_mean:vowel_ratio_sd) %>% 
-  tab_spanner(md("**Frequency (per million)**"), c("frequency_mean", "frequency_sd")) %>% 
-  tab_spanner(md("**Frequency (Zipf score)**"), matches("frequency_zipf")) %>% 
+  fmt_number(frequency_mean:lv_sd) %>% 
+  tab_spanner(md("**Freq./million**"), c("frequency_mean", "frequency_sd")) %>% 
+  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
   tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
-  tab_spanner(md("**Consonant similarity**"), matches("consonant")) %>% 
-  tab_spanner(md("**Vowel similarity**"), matches("vowel")) %>% 
+  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
   cols_label(
@@
     pthn_sd = md("*SD*"),
-    consonant_ratio_mean = md("Mean"),
-    consonant_ratio_sd = md("*SD*"),
-    vowel_ratio_mean = md("Mean"),
-    vowel_ratio_sd = md("*SD*")
-  )
+    lv_mean = md("Mean"),
+    lv_sd = md("*SD*")
+  ) %>% 
+  as_latex()
+
+stimuli_table
 
@@
 Finally, we fit a model that included the two-way interactions `pthn:consonant_ratio` and `pthn:vowel_ratio`, an their random slopes by participant (Model 4).
-Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020capitalize].
+All predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
 Model 4 can be formally expressed as:
-
-$$
-\begin{align*}
-
-&\textbf{Likelihood}  \\
-y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
-
-&\textbf{Parameters}  \\
-
-logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
-& \beta_{1[p]} ~ Frequency_{i} ~ + \\
-& \beta_{2[p]} ~ PTHN_i ~ + \\
-& \beta_{3[p]} ~ Similarity_i ~ + \\
-& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
-
-\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
-\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
-&& \text{[participant-level coefficients]} \\ \\
-
-&\textbf{Prior}  \\
-
-\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 3) && \text{[participant-level coefficients]} \\
-\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ Cauchy(0, 3) && \text{[SD for population and participant]} \\
-\rho_{p}, ~ \rho_{w} \sim& ~LKJ(5) && \text{[correlation between participant-level coefficients]} \\
-
-\end{align*}
-$$
 
@@
 
-```{r model_values, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r modelvalues, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
@@
     model = md("**Model**"),
-    looic = md("***LOO<sub>IC</sub>***"),
+    looic = md("***LOO_{IC}***"),
     se_looic = md("***SE*<sub>IC</sub>**"),
@@
     se_diff = md("***SE<sub>diff</sub>***")
-  )
+  ) %>% 
+  as_latex()
 
-gtsave(loos_table, here("Figures", "loos.png"))
+loos_table
 
@@
 
-```{r posterior_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 3.
-A) Population-level effects.
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 3.
+A\\) Population-level effects.
 Distributions indicate the estimated posterior likelihood density of each coefficient.
-Credible intervals (CrI), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95%, 80%, and 50%, given the data we have collected.
-Points represent the mean of the distribution.
-The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The rest of the coefficients have been transformed using the divide-by-four rule to get the maximum change in probability of correct response associated with a unit increase in each predictor variable.
-B) Participant-level coefficient variability.
-Our model estimated participant-level coefficients to account for the dependency between responses from the same participant.
-Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (SD).
-C) Correlation between participant-level effects.
-Our model allowed participant-level coefficients to co-vary.
+Credible intervals \\(CrI\\), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Dots represent the mean of the distribution.
+B\\) Participant\\-level coefficient variability.
+Our model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations \\(SD\\).
+C\\) Correlation between participant\\-level effects.
+Our model allowed participant-level coefficients to co\\-vary.
 This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each correlation.
-Coefficients are represented in the X- and Y-axis in the same order as indicated in the Y-axis of panels A and C.", fig.height=5, fig.width=6}
+Coefficients are represented in the X\\-axis and Y\\-axis in the same order as indicated in the Y\\-axis of panels A and C.", fig.height=5, fig.width=6}
 
-post <- gather_draws(model_fits$fit_3, `b_.*`, `sd_.*`, regex = TRUE)
-
-# fixed effects
 str_repl <- c(
@@
 
-post_fix <- post %>% 
+# fixed effects
+post_fix <- posterior_draws_fixed %>% 
   filter(str_detect(.variable, "b_")) %>% 
-  mutate(.variable_name = str_remove(.variable, "b_") %>% 
+  mutate(
+    .variable_name = str_remove(.variable, "b_") %>% 
       str_replace_all(str_repl) %>% 
       factor(levels = c("Intercept", "Frequency (+1 SD)", "PTHN (+1 SD)", "Levenshtein (+1 SD)", "PTHN \u00d7 Levenshtein")),
-         .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)) %>% 
+    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+  ) %>% 
   arrange(.variable) %>% 
@@
   geom_vline(xintercept = 0) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = scales::percent_format())))) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
   stat_pointinterval(.width = 0, point_size = 1) +
-  # geom_text(
-  #   data = filter(rhats, str_detect(.variable, "b_")), 
-  #   aes(x = -0.1, y = .variable_name, label = paste0("Rhat = ", printnum(.value, digits = 3))),
-  #   size = 3, position = position_nudge(y = 0.1)
-  # ) +
-  # geom_text(
-  #   data = filter(n_effs, str_detect(.variable, "b_")), 
-  #   aes(x = -0.1, y = .variable_name, label = paste0("N eff. ratio = ", printnum(.value, digits = 2))),
-  #   size = 3, position = position_nudge(y = 0.25)
-  # ) +
   scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
@@
 
-post_sd <- post %>% 
-  filter(str_detect(.variable, "sd_"),) %>% 
+post_sd <- posterior_draws_fixed %>% 
+  filter(str_detect(.variable, "sd_")) %>% 
   mutate(
@@
   ggplot(aes(.value, fct_rev(.variable_name))) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = scales::percent_format())))) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
   stat_pointinterval(.width = 0, point_size = 1) +
-  #   geom_text(
-  #   data = filter(rhats, str_detect(.variable, "b_")), 
-  #   aes(x = -0.1, y = .variable_name, label = paste0("Rhat = ", printnum(.value, digits = 3))),
-  #   size = 3, position = position_nudge(y = 0.1)
-  # ) +
-  # geom_text(
-  #   data = filter(n_effs, str_detect(.variable, "b_")), 
-  #   aes(x = -0.1, y = .variable_name, label = paste0("N eff. ratio = ", printnum(.value, digits = 2))),
-  #   size = 3, position = position_nudge(y = 0.25)
-  # ) +
   scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
@@
 
-```{r marginal_effects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=5, fig.height=5, fig.cap="Expected mean posterior predictions.
+```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=3, fig.cap="Expected mean posterior predictions.
+A) Population-level expected mean posterior predictions.
 The X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
 We simulated 200 observations from our model: 100 simulations for words with low PTHN (-1 SD) and 100 simulations for high PTHN (+1 SD) words.
+We did this across the range of values of the Levenshtein scores.
 For each simulation, we drew a single sample from the posterior distribution of each coefficient.
@@
 We also plotted the mean of the high PTHN (black solid line) and low PTHN (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
-The dispersion of the lines indicates the uncertainty of our predictions."}
+The dispersion of the lines indicates the uncertainty of our predictions.
+B) Participant-level expected mean posterior predictions.
+We conducted the same procedure for each individual participant, simulating 100 observations for high-PTHN and 100 for low-PTHN from our model acrosss the range of Levenshtein scores.
+We averaged the resulting predictions for each participant and then plotted each participant's predictions in separate panels to show how the effect of PTHN and phonological similarity changed at the individual level."}
 
 # by vowel_ratio
-nd <- expand.grid(
-  pthn = c(-1, 1),
-  frequency_zipf = 0,
-  lv = seq(
-    min(responses$lv, na.rm = TRUE),
-    max(responses$lv, na.rm = TRUE),
-    by = 0.1
-  )
-)
-
-m <- add_epred_draws(nd, model_fits$fit_3, ndraws = 100, re_formula = NA) %>% 
+m <- posterior_epreds_fixed %>% 
   mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
@@
 
-nd_obs <- model_fits$fit_3$data %>% 
-  as_tibble() %>% 
-  mutate(pthn = cut_interval(pthn, 3, labels = c("Low PTHN", "Mean PTHN", "High PTHN")))
+m_re <- posterior_epreds_random %>% 
+  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
+  mutate(
+    pthn = case_when(
+      pthn=="-1 SD" ~ "Low PTHN",
+      pthn=="0 SD" ~ "Mean PTHN",
+      pthn=="1 SD" ~ "High PTHN"
+    ),
+    frequency_zipf = case_when(
+      frequency_zipf=="-1 SD" ~ "Low frequency",
+      frequency_zipf=="0 SD" ~ "Mean frequency",
+      frequency_zipf=="1 SD" ~ "High frequency"
+    )
+  ) %>% 
+  group_by(participant, lv, pthn) %>% 
+  summarise(.epred = mean(.epred, na.rm = TRUE), .groups = "drop")
 
-ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
+plot_m <- ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
   # geom_jitter(data = filter(nd_obs, correct), aes(y = 1), alpha = 0.1,
@@
   geom_hline(yintercept = 0.5, colour = "grey") +
-  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 1) +
-  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = FALSE) +
-  labs(x = "Levenshtein (phonological similarity)", y = "P(Correct)", colour = "PTHN", fill = "PTHN", linetype = "PTHN") +
+  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.25, show.legend = FALSE) +
+  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.5, show.legend = FALSE) +
+  labs(x = "Levenshtein (phonological similarity)", y = "P(Correct)", 
+       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
+       title = "Population-level predictions") +
   scale_color_manual(values = c("#1E88E5", "#ff2976")) +
-  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 0.6), breaks = seq(0, 1, 0.1)) +
+  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
   guides(colour = guide_legend(ncol = 2)) +
@@
     panel.grid = element_line(colour = "grey")
+  )
+
+plot_m_re <- ggplot(m_re, aes(lv, .epred, colour = pthn, fill = pthn)) +
+  facet_wrap(~participant) +
+  geom_line(aes(group = interaction(pthn, participant)), size = 0.35) +
+  labs(x = "Levenshtein (phonological similarity)", y = "P(Correct)", 
+       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
+       title = "Participant-level predictions") +
+  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
+  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
+  guides(colour = guide_legend(ncol = 2)) +
+  theme(
+    legend.position = "none",
+    axis.text = element_blank(),
+    panel.grid = element_blank(),
+    strip.text = element_blank(),
+    strip.background = element_blank(),
+    axis.ticks = element_blank(),
+    axis.line = element_line(colour = "grey"),
+    panel.spacing = unit(0.1, "lines")
+  )
+
+plot_m + plot_m_re +
+  plot_layout(guides = "collect") &
+  plot_annotation(tag_levels = "A") &
+  theme(
+    legend.position = "top",
+    legend.title = element_blank(),
+    title = element_text(size = 8, face = "bold"),
+    plot.title.position = "plot", 
+    plot.title = element_text(hjust = 0.5)
   )
@@
 ```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
-Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.
-Dots and whiskers represent the average accuracy and 95% confidence interval of each word.
+Words presented to participants in the English\\-Spanish or in the English\\-Catalan and Spanish\\-Catalan are listed in the Y\\-axis, ordered from higher to lower average translation accuracy, depicted in the X\\-axis.
+Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word.
 Accuracy is plotted separately for each group.", fig.height=10, fig.width=9}

### 2021-11-26 R/02_models.R

@@ -19,6 +19,7 @@ get_model_fits <- function(
         prior = model_prior[c(1, 3),],
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
+        seed = 888,
         save_model = here("Stan", "fit_responses_0.stan"),
         file = here("Results", "fit_responses_0.rds"), 
         control = list(adapt_delta = 0.95)
@@ -29,6 +30,7 @@ get_model_fits <- function(
         prior = model_prior,
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
+        seed = 888,
         save_model = here("Stan", "fit_responses_1.stan"),
         file = here("Results", "fit_responses_1.rds"), 
         control = list(adapt_delta = 0.95)
@@ -39,6 +41,7 @@ get_model_fits <- function(
         prior = model_prior,
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
+        seed = 888,
         save_model = here("Stan", "fit_responses_2.stan"),
         file = here("Results", "fit_responses_2.rds"), 
         control = list(adapt_delta = 0.95)
@@ -49,6 +52,8 @@ get_model_fits <- function(
         prior = model_prior,
         save_pars = save_pars(all = TRUE),
         backend = "cmdstanr",
+        sample_prior = "yes",
+        seed = 888,
         save_model = here("Stan", "fit_responses_3.stan"),
         file = here("Results", "fit_responses_3.rds"), 
         control = list(adapt_delta = 0.95)
@@ -71,7 +76,44 @@ get_model_loos <- function(fits){
         saveRDS(loo, path)
     }
     return(loo)
-    
+}
+
+# get model posterior draws (fixed effects)
+get_model_draws_fixed <- function(fit){
+    post <- gather_draws(fit, `b_.*`, `sd_.*`, regex = TRUE)
+    return(post)
+}
+
+
+# get model posterior expected predictions (fixed effects)
+get_model_epreds_fixed <- function(fit, ndraws = 100){
+    nd <- expand.grid(
+        pthn = c(-1, 1),
+        frequency_zipf = 0,
+        lv = seq(
+            min(fit$data$lv, na.rm = TRUE),
+            max(fit$data$lv, na.rm = TRUE),
+            by = 0.1
+        )
+    )
+    m <- add_epred_draws(nd, fit, ndraws = 100, re_formula = NA)
+    return(m)
+}
+
[…]

### 2021-11-26 R/utils.R

@@ -1,23 +1,4 @@
-# utils
-
-# dark GitHub theme
-theme_github <- function(){
-    theme_dark() +
-        theme(
-            text = element_text(colour = "white", size = 12),
-            axis.text = element_text(colour = "white", size = 8),
-            legend.title = element_text(colour = "white", size = 10),
-            legend.text = element_text(colour = "white", size = 10),
-            legend.background = element_rect(fill = "#0D1117"),
-            strip.text = element_text(size = 10),
-            legend.key = element_rect(fill = "#0D1117", colour = "#0D1117"),
-            strip.background = element_rect(fill = "#161B22"),
-            plot.background = element_rect(fill = "#0D1117"),
-            panel.background = element_rect(fill = "#0D1117"),
-            panel.border = element_rect(fill = "transparent", colour = "#0D1117"),
-            panel.grid = element_blank()
-        )
-}
+#### utils: helper functions ---------------------------------------------------
 
 # evaluate if x is NOT an element of y
 "%!in%" <- function(x, y) !(x %in% y)
@@ -60,6 +41,142 @@ theme_custom <- function(){
     )
 }
 
+# extract formants
+extract_formants <- function(
+    file,
+    time_step = 0.001,
+    n = 2,
+    max_freq = 5500,
+    window_length = 0.001,
+    pre = 50,
+    include_frames = TRUE,
+    include_time = TRUE,
+    dec = 3,
+    include_intensity = TRUE,
+    dec_intensity = 5,
+    include_n_formats = FALSE,
+    dec_freq = 3,
+    include_bandwidths = FALSE) {
+    require(PraatR)
+    require(rlang)
+    require(purrr)
+    require(stringr)
+    require(dplyr)
+    require(data.table)
+    require(janitor)
+    
+    file_paths <- list.files({{ file }}, pattern = ".wav", full.names = TRUE)
+    file_names <- list.files({{ file }}, pattern = ".wav", full.names = FALSE)
+    
+    pb <- progress_estimated(length(file_paths)*2)
+    arguments1 <- list({{ time_step }}, {{ n }}, {{ max_freq }}, {{ window_length }}, {{ pre }})
+    arguments2 <- list({{ include_frames }}, {{ include_time }}, {{ dec }}, {{ include_intensity }}, {{ dec_intensity }}, {{ include_n_formats }}, {{ dec_freq }}, {{ include_bandwidths }}) 
+    
[…]

### 2021-11-26 Rmd/manuscript.Rmd
Update manuscript

@@
 
-options(
-  knitr.kable.NA = '-',
-  knitr.duplicate.label = "allow",
-  ggplot2.discrete.fill = wesanderson::wes_palettes$Darjeeling1,
-  ggplot2.discrete.colour = wesanderson::wes_palettes$Darjeeling1,
-  ggplot2.continuous.fill = wesanderson::wes_palettes$Darjeeling1,
-  ggplot2.continuous.colour = wesanderson::wes_palettes$Darjeeling1
-)
+options(knitr.kable.NA = '-', knitr.duplicate.label = "allow")
 
@@
 
+# import objects
 tar_load_globals()
@@
 
-theme_set(theme_ggdist())
+# set custom ggplot2 theme
+theme_set(
+  theme(
+    title = element_text(size = 8, face = "bold"),
+
+    legend.background = element_rect(fill = NA, colour = NA),
+    legend.key = element_rect(fill = "white", colour = NA),
+    legend.justification = "right",
+    legend.direction = "horizontal",
+
+    axis.line = element_line(colour = "black", size = 0.5),
+    axis.ticks = element_line(colour = "black", size = 0.5),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+
+    plot.caption.position = "plot",
+    plot.title = element_text(hjust = 0.5),
+
+    panel.grid = element_blank(),
+    panel.background = element_rect(fill = "white", colour = NA)
+  )
+)
 
@@
 
-## Non-native language comprehension is difficult, but cognateness helps
+## Native speech processing is effortless
 
@@
 But less is known about how these factors affect the recognition of words in a non-native language.
+
+## Non-antive speech processing is costly
 
@@
 
+## Cognateness aids non-native speech processing
+
 One of such commonalities occurs at the lexical level, in the for of cognateness.
@@
 It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
-Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access their lexicon in a language non-selective way: during word production and comprehension, both lexical representations are activated, and the form similarity between both impacts participants' performance in naming or lexical decision tasks.
+
+## The cognate advantage in comprehension, production, and learning
+
+Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access their lexicon in a language non-selective way: during word production and comprehension, both lexical representations are activated, and the form similarity between both impacts participants' performance in naming or lexical decision tasks [e.g., @costa2000cognate; @thierry2007brain].
 For instance, in their seminal study, @costa2000cognate asked Spanish-Catalan bilinguals to name pictures in Spanish.
@@
 
-There is evidence that cognates are learnt more easily in L2 than non-cognates.
+There is also evidence that cognates are learnt more easily in L2 than non-cognates.
 For instance, @de2000hard presented 40 Dutch natives 60 pairs of translation equivalents.
@@
 
+### The cognate advantage in translation: the RMH model
+
 Cognates are not only learnt faster than non-cognates, but translated faster too.
@@
 The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, while forward translation would rely more strongly on the mediation between the concept and the two word forms.
-One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words iduring backward translation.
+One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
 
@@
 Subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language], contrary to the predictions of the RHM.
+
+## The impact of neighbourhoods lexical processing: the NAM and the BIA/BIA+ models
 
@@
 
-More recently, @dijkstra2019multilink developed a model, Multilink, that integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
-This model
+## Is the lexical route necessary for translation?: The Multilink model
 
-These studies tested the role of neighbourhood lexical frequency in samples of fairly proficient (i.e., balanced) bilinguals, and addressed this issue mostly at the orthographic level.
-Although these models contemplate the role of phonology in their accounts, the fact the it plays a secondary role in them limits their theoretical generalisability.
-One of the main reasons why a substantial part of the literature has avoided phonology as their modality of interest stems from the methodological level: working with phonological transcriptions is computationally challenging, and perhaps more critically, there is no clear way of measuring the similarity/dissimilarity between phonological forms of words within or across languages.
-Any possible way of measuring the overlap between two phonological forms at the surface level (e.g., between phonological transcriptions of the words) is very likely to neglect important sources of perceived similarity/dissimilarity.
-For example, one could measure
+More recently, @dijkstra2019multilink implemented a localist-connectionist model, Multilink, that integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
+This model assumes (1) an integrated lexicon in which word representations from the two languages establish connections as words from the same language do; (2) during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1; (3) contrary to the RHM model's assumption that translation equivalents are linked by excitatory connections (e.g., translation can be done via word-word connections), @dijkstra2019multilink suggest that such connections would spread activation though irrelevant words in the target language and therefore initially assumed that translation equivalents are exclusively connect throught their shared concept; (4) the strength of the association between L2 representations and their concept is a function of the language user's proficiceny in L2.
+Multilink considers the particular case of Dutch-English bilinguals to generated simulations and test them against experimental data.
+The authors simulated data from Multilink to mirror @christoffels2006memory translation elicitation task, and reported a remarkable correlation between Multilink's simulations and participant's data, with forward translation being faster than backward translation, contrary to the RHM model predictions.
+More relevant to the present study is the fact that the model also generated data suggesting that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+Overall, evidence from this model provides evidence for an integrated bilingual lexicon and for a facilitatory effect of cross-language similarity during translation.
+One of the main conclusions of @dijkstra2019multilink study is that word-word connections are not necessary for translation, as suggested by @kroll1994category: Multilink did not include such connections and still fitted experimental data.
+However, as they note, both their simulations and the data collected by @christoffels2006memory corresponded to high-proficiency bilinguals.
+In the case of the model simulations, vocabulary sizes were balanced in both languages and word-concept connections were equally strong for each member of the translation pairs in both languages.
+This scenario is actually contemplated by the @kroll1994category's RHM model.
+What @dijkstra2019multilink did not test, is the fit of Mutilink's predictions for low-proficiency, unbalanced bilinguals, whose vocabulary sizes are smaller or their word-concept connections are weaker in L2 than in L1.
+It is under this set of assumptions that the RHM model predicts participants to rely strongly on word-word connections to be able to translation words, especially from L2 to L1.
+This prediction, however, is untested in the Multilink model.
 
-spoken word recognition is unclear.
-Across several modalities, bilinguals' recognition of words in a language is affected by form-similarity and frequency of words in the other language.
-The parallel activation hypothesis accounts for this phenomenon, suggesting that lexical access is language-non selective.
-There is strong evidence supporting this claim in both comprehension, production, and translation, and across modalities.
-Much of this evidence stems from the fact that bilinguals' production times are faster when naming cognates (i.e., words whose translation in the other language is form-similar) than non-cognates.
-Both phonological and semantic links between translations underlie facilitatory effect of cognateness.
-The activation of form-similar words in the non-target language that do not share meaning with the target word, however, interferes with lexical selection.
-Disentangling the effect of semantic and phonological links during spoken word recognition is challenging when participants are bilingual.
+## The present study
 
-This has important implications to non-native word recognition: is word recognition facilitated by cognateness via translation?
-The answer is not straightforward.
-Disentangling the effect of form- and conceptual similarity in a bilingual sample is daunting, given that second language learners are able to exploit ...
-
-In the present study, we used a fully monolingual sample to disentangle the effect of neighbourhood frequency and phonological similarity.
-We asked participants to translate word in a language they had no significant experience with.
-This made sure that participants were unaware of any semantic relationship between the (non-native) words they heard and their translation, and could only rely on their phonological similarity to guess their translation.
-We tested the role of the amount of phonological similarity between the presented and the target words, the lexical frequency of the target word, and the average lexical frequency of the target word's phonological neighbourhood.
+In this study, we explored the plausibility of the lexical route in backward translation by testing monolingual participants.
+Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which which participants' vocabulary size in L2 is null, and word-concept connections are absent (i.e., set to zero in formal terms).
+In this scenario, participants can only rely on the similarity between the L2 form (presented as stimulus) and the L1 form (the target word).
+In other words, these participants can only use the lexical route to suceeed in the translation task.
+If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would sugges that the lexical route is in place for low-proficiency bilinguals.
+If participants are not able to translation words from L2 to L1 regardless of their form-similarity, this would suggest that low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
 
@@
 This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
-First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical.
+First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
 This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
@@
   scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
-  scale_x_continuous(labels = percent) +
+  scale_x_continuous(labels = function(x) percent(round(x, 1))) +
   labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
@@
     legend.position = c(1, 0.1),
-    legend.background = element_rect(fill = NA),
-    legend.justification = "right",
-    legend.direction = "horizontal",
-    plot.caption.position = "plot",
-    # axis.text = element_text(size = 7),
     axis.title.y = element_blank(),
@@
     axis.text.y = element_text(colour = "black", vjust = 0),
-    panel.grid.major.y = element_line(colour = "grey", size = 0.25)
+    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
   )
 
+# SD of random effects
 post_sd <- posterior_draws_fixed %>% 
@@
     legend.position = c(1, 0.1),
-    legend.background = element_rect(fill = NA),
-    legend.justification = "right",
-    legend.direction = "horizontal",
-    plot.caption.position = "plot",
-    plot.caption = element_text(hjust = 0),
-    # axis.text = element_text(size = 7),
-    panel.grid.major.y = element_line(colour = "grey", size = 0.25),
     axis.title.y = element_blank(),
     axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black")
+    axis.text.y = element_text(colour = "black", vjust = 0),
+    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
   )
 
+# Pearson correlations between random effects
 corr_mat <- cov2cor(vcov(model_fits$fit_3)) 
@@
     axis.text.x = element_blank(),
-    axis.text.y = element_blank(),
-    plot.background = element_rect(fill = NA)
+    axis.text.y = element_blank()
   )
@@
   plot_layout(guides = "keep") & 
-  plot_annotation(tag_levels = "A") &
-  theme(
-    title = element_text(size = 8, face = "bold"),
-    plot.title.position = "plot", 
-    plot.title = element_text(hjust = 0.5)
-  )
+  plot_annotation(tag_levels = "A")
 
@@
   scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
-  guides(colour = guide_legend(ncol = 2)) +
-  theme(
-    legend.position = "top",
-    legend.title = element_blank(),
-    axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black"),
-    panel.grid = element_line(colour = "grey")
-  )
+  guides(colour = guide_legend(ncol = 2))
 
@@
     legend.position = "none",
-    axis.text = element_blank(),
-    panel.grid = element_blank(),
+    axis.text.x = element_blank(),
+    axis.text.y = element_blank(),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks = element_blank(),
     strip.text = element_blank(),
     strip.background = element_blank(),
-    axis.ticks = element_blank(),
-    axis.line = element_line(colour = "grey"),
     panel.spacing = unit(0.1, "lines")
@@
     legend.title = element_blank(),
-    title = element_text(size = 8, face = "bold"),
-    plot.title.position = "plot", 
-    plot.title = element_text(hjust = 0.5)
   )
@@
     legend.position = "none",
-    axis.text.y = element_text(size = 6, colour = "black"),
+    axis.text.y = element_text(size = 5, colour = "black", hjust = 1),
     axis.ticks = element_blank(),
     axis.text.x = element_text(size = 7, colour = "black"),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
   )

### 2021-11-26 Rmd/manuscript.tex
Update manuscript

--- 2021-09-29 Rmd/manuscript.tex
@@
   english,
-  man]{apa6}
+  man,floatsintext]{apa6}
 \usepackage{amsmath,amssymb}
@@
 \keywords{cognate, word recognition, translation, non-native, spoken word recognition\newline\indent Word count: X}
-\DeclareDelayedFloatFlavor{ThreePartTable}{table}
-\DeclareDelayedFloatFlavor{lltable}{table}
-\DeclareDelayedFloatFlavor*{longtable}{table}
-\makeatletter
-\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
-\makeatother
 \usepackage{lineno}
@@
 
+The authors made the following contributions.
+Gonzalo Garcia-Castro: Conceptualization, Data analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Serene Siow: Conceptualization, Writing - Review \& Editing; Nuria Sebastian-Galles: Conceptualization, Writing - Review \& Editing; Kim Plunkett: Conceptualization, Writing - Review \& Editing.
+
 Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trial Fargas, 25-27, 08005 Barcelona.
@@
 
-\affiliation{\phantom{0}}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra\\\textsuperscript{2} Department of Experimental Psychology, University of Oxford}
 
@@
 There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
-For example, bilinguals produce and translated cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
+For example, bilinguals produce and translate cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
 Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
@@
 
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+\hypertarget{native-speech-processing-is-effortless}{%
+\subsection{Native speech processing is effortless}\label{native-speech-processing-is-effortless}}
+
+Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech (De Haan, 1982) or the loss of segmental information (Warren, 1970).
+But even in the most ideal of the situations, the processes engaged in word recognition do not occur without uncertainty.
+The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
+There is extensive literature about how the number of candidates, their frequency, and their phonological and semantic similarity with the target word impacts the dynamics of word recognition.
+But less is known about how these factors affect the recognition of words in a non-native language.
+
+\hypertarget{non-antive-speech-processing-is-costly}{%
+\subsection{Non-antive speech processing is costly}\label{non-antive-speech-processing-is-costly}}
+
+Listening to speech in a non-native language is more costly than doing so in the native language, even if one is proficient in such language.
+The source of this increased cognitive effort is likely to originate at multiple levels.
+Some examples: (1) some sounds in the speech signal do not correspond to any phoneme in the native language, (2) segmenting the speech signal relies on previous familiarity with word forms or with the statistical regularities of the language: (3) word order might be different in both languages (e.g., subject-object-verb vs.~subject-verb-object.).
+The listener, however, is rarely completely naïve to the language they are listening to.
+All languages share, to some extent, similarities, frequently due to their typological closeness.
+These similarities which can be exploited by non-native listeners.
+
+\hypertarget{cognateness-aids-non-native-speech-processing}{%
+\subsection{Cognateness aids non-native speech processing}\label{cognateness-aids-non-native-speech-processing}}
+
+One of such commonalities occurs at the lexical level, in the for of cognateness.
+Cognates are cross-language synonyms whose form (e.g., phonology, orthography, signature) is similar.
+The cause of this similarity is frequently attributed to a shared etymological origin.
+Romance languages like Spanish, Italian, French or Catalan, share many cognates (Schepens, Dijkstra, \& Grootjen, 2012).
+For example, this is the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively)\footnote{Some form-similar cross-language synonyms are technically not cognates.
+For example, \emph{sun} and \emph{sol} (in Spanish), share their phonological onset, but their etymology points to different origins.
+We will use the word \emph{cognateness} to include all form-similar cross-language synonyms for simplicity.
+It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.}.
+
+\hypertarget{the-cognate-advantage-in-comprehension-production-and-learning}{%
+\subsection{The cognate advantage in comprehension, production, and learning}\label{the-cognate-advantage-in-comprehension-production-and-learning}}
+
+Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access their lexicon in a language non-selective way: during word production and comprehension, both lexical representations are activated, and the form similarity between both impacts participants' performance in naming or lexical decision tasks (e.g., Costa, Caramazza, \& Sebastian-Galles, 2000; Thierry \& Wu, 2007).
+For instance, in their seminal study, Costa et al.
+(2000) asked Spanish-Catalan bilinguals to name pictures in Spanish.
+Unbeknownst to participants, half of the pictures' associated labels were cognates in Spanish and Catalan (e.g., \emph{puerta}-\emph{porta}) while the other half were non-cognates (e.g., \emph{mesa}-\emph{taula}).
+Surprisingly, participants named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
+This suggested that bilinguals activated picture's word representations in both languages, and that the phonological overlap between both labels facilitated the naming process.
+
+There is also evidence that cognates are learnt more easily in L2 than non-cognates.
+For instance, De Groot and Keijzer (2000) presented 40 Dutch natives 60 pairs of translation equivalents.
+In each learning trial, two words were presented in a screen side-by-side: one in Dutch and one pseudoword.
+Pseudowords were generated in such way that were easily pronounceable and were phonotactically legal in Dutch.
+Word pairs varied in their form-similarity (number of shared letters, 40-70\%).
+Participants were tested twice in the same task with one week of difference.
+Participants' performance was better for cognates across both testing sessions, suggesting that cognates were learnt and retained more easily than non-cognates.
+Converging evidence was provided by Lotto and De Groot (1998) in Dutch natives learning Italian words: cognates were learnt more easily than non-cognates.
+
+\hypertarget{the-cognate-advantage-in-translation-the-rmh-model}{%
+\subsubsection{The cognate advantage in translation: the RMH model}\label{the-cognate-advantage-in-translation-the-rmh-model}}
+
+Cognates are not only learnt faster than non-cognates, but translated faster too.
+The mechanisms behind this effect, though, are still unclear.
+Early accounts of bilingual lexical access suggested that low-proficiency learners first established links between newly learnt words in L2 and their meanings through their L1 translation equivalents (Potter, So, Von Eckardt, \& Feldman, 1984).
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
+The Revised Hierarchical Model {[}RHM; Kroll and Stewart (1994){]} captured this assumption and predicted that translating words from L1 to L2 (forward translation) should take longer than translating from L2 to L1 (backward translation).
+The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, while forward translation would rely more strongly on the mediation between the concept and the two word forms.
+One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
+
+To test these predictions, Degroot, Dannenburg, and Vanhell (1994) and Groot (1992) asked 52 Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+In each trial, participants were presented visually with a word in English or Dutch, and were asked to speak out loud its translation in the other language as soon as possible.
+The authors reported three main findings: (1) translation times and accuracy were roughly equivalent in across both forward and backward conditions (although slightly faster in the former), (2) semantic variables, such as imageability, were positively associated with participants' performance more strongly in the forward translation conditional, although the effect size was small, (3) cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward translation than during forward translation, and (4) when translating cognates participants' performance across both conditions was less sensitive to semantic variables than when translating non-cognates.
+These results prompted the authors to reject a hard version Kroll and Stewart (1994)'s account, and suggested that both conceptually mediated and direct translation routes are active during translation, but backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness than forward translation.
+When the authors tested a group of bilinguals with higher proficiency in English, their results pointed in the same direction.
+Subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels, De Groot, \& Kroll, 2006; Christoffels, Ganushchak, \& Koester, 2013), contrary to the predictions of the RHM.
+
+\hypertarget{the-impact-of-neighbourhoods-lexical-processing-the-nam-and-the-biabia-models}{%
+\subsection{The impact of neighbourhoods lexical processing: the NAM and the BIA/BIA+ models}\label{the-impact-of-neighbourhoods-lexical-processing-the-nam-and-the-biabia-models}}
+
+The idealised lexicon considered by the models described above only considers word-word connections between translation equivalents.
+More recent studies have highlighted the role of the rich network of connections that given word establishes with other phonologically or conceptually related words.
+This notion was first introduced in monolingual research by Collins and Loftus (1975) in their theory of semantic processing, and later Luce and Pisoni (1998) formalised this dimension in their Neighbourhood Activation Model (NAM).
+In this model, lexical selection is mediated not only by the lexical frequency of the target word, and the number of activated candidates, but also by the frequency of the candidates.
+Luce and Pisoni (1998) designed a lexical decision task in which English native participants listened to a word across several conditions of noise-to-signal ratio.
+The authors then used a computerised lexicon to calculate the number of phonological neighbours around each of the presented words, based on phonetic similarity matrices, and the average lexical frequency of such neighbourhoods.
+Participants answered more faster and more accurately to high frequency words.
+Words from high density neighbourhoods were responded to more accurately, but more slowly.
+The average lexical frequency of the neighbourhood was associated to a decrease in accuracy.
+Low frequency words were responded to more accurately in low density neighbours than in high density neighbours.
+The authors concluded that, although both word lexical frequency and neighbourhood density can, in principle, facilitate spoken word recognition, this effect was modulated by the frequency of the neighbourhood: when phonological neighbours are more frequent, lexical selection is hindered (Goldinger, Luce, \& Pisoni, 1989; Luce, Pisoni, \& Goldinger, 1990).
+
+The fact that phonological neighbourhoods and their structure impacts lexical processing poses an important question for bilingualism research.
+Are word representations in one language part of phonological or orthographic neighbours in the other language?
+Van Heuven, Dijkstra, and Grainger (1998) addressed this issue in an experimental series and proposed the Bilingual Interactive Activation (BIA, later revised and relabelled as BIA+).
+In line with connectionist accounts of lexical processing this model advocated for lexical representations to establish both excitatory and inhibitory connections with other representations from the same of the other language (McClelland \& Rumelhart, 1981), forming an integrated lexicon for both languages.
+Van Heuven et al.
+(1998) tested Dutch-English bilinguals in a progressive demasking task (in each trial a words in presented in increasingly longer time windows interrupted by a checkerboard until the participant presses a button to type the word in a keyboard) or a lexical decision task (words a visually presented and the participant must decide as quickly as possible whether its a word in the target language or a non-word).
+The authors manipulated by size of the orthographic neighbourhood size of the words presented in its language (target language) or in the other language (non-target language).
+Interestingly, the authors reported participants' performance to be affected by the neighbourhood size of the presented word in the non-target language across both tasks.
+Specifically, participants took longer to respond to words with high neighbourhood density in the non-target language, even when participants completed the task in the target language exclusively (e.g., were presented with Dutch words exclusively).
+Monolinguals, on the other hand, were only affected by neighbourhood size in the target language.
+This provided evidence of cross-language interference between the word representation in the non-target language and the those in the target language.
+
+\hypertarget{is-the-lexical-route-necessary-for-translation-the-multilink-model}{%
+\subsection{Is the lexical route necessary for translation?: The Multilink model}\label{is-the-lexical-route-necessary-for-translation-the-multilink-model}}
+
+More recently, Dijkstra et al.
+(2019) implemented a localist-connectionist model, Multilink, that integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
+This model assumes (1) an integrated lexicon in which word representations from the two languages establish connections as words from the same language do; (2) during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1; (3) contrary to the RHM model's assumption that translation equivalents are linked by excitatory connections (e.g., translation can be done via word-word connections), Dijkstra et al.
+(2019) suggest that such connections would spread activation though irrelevant words in the target language and therefore initially assumed that translation equivalents are exclusively connect throught their shared concept; (4) the strength of the association between L2 representations and their concept is a function of the language user's proficiceny in L2.
+Multilink considers the particular case of Dutch-English bilinguals to generated simulations and test them against experimental data.
+The authors simulated data from Multilink to mirror Christoffels et al.
+(2006) translation elicitation task, and reported a remarkable correlation between Multilink's simulations and participant's data, with forward translation being faster than backward translation, contrary to the RHM model predictions.
+More relevant to the present study is the fact that the model also generated data suggesting that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+Overall, evidence from this model provides evidence for an integrated bilingual lexicon and for a facilitatory effect of cross-language similarity during translation.
+One of the main conclusions of Dijkstra et al.
+(2019) study is that word-word connections are not necessary for translation, as suggested by Kroll and Stewart (1994): Multilink did not include such connections and still fitted experimental data.
+However, as they note, both their simulations and the data collected by Christoffels et al.
+(2006) corresponded to high-proficiency bilinguals.
+In the case of the model simulations, vocabulary sizes were balanced in both languages and word-concept connections were equally strong for each member of the translation pairs in both languages.
+This scenario is actually contemplated by the Kroll and Stewart (1994)'s RHM model.
+What Dijkstra et al.
+(2019) did not test, is the fit of Mutilink's predictions for low-proficiency, unbalanced bilinguals, whose vocabulary sizes are smaller or their word-concept connections are weaker in L2 than in L1.
+It is under this set of assumptions that the RHM model predicts participants to rely strongly on word-word connections to be able to translation words, especially from L2 to L1.
+This prediction, however, is untested in the Multilink model.
+
+\hypertarget{the-present-study}{%
+\subsection{The present study}\label{the-present-study}}
+
+In this study, we explored the plausibility of the lexical route in backward translation by testing monolingual participants.
+Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which which participants' vocabulary size in L2 is null, and word-concept connections are absent (i.e., set to zero in formal terms).
+In this scenario, participants can only rely on the similarity between the L2 form (presented as stimulus) and the L1 form (the target word).
+In other words, these participants can only use the lexical route to suceeed in the translation task.
+If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would sugges that the lexical route is in place for low-proficiency bilinguals.
+If participants are not able to translation words from L2 to L1 regardless of their form-similarity, this would suggest that low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
+
 \hypertarget{methods}{%
@@
 
-Data collection took place from 04th June, 2020 to 28th June, 2020.
-We collected data from 105 participants (\(M_{age}\) = 21.79, \(SD_{age}\) = 2.43, \(Range_{age}\) = 18-33).
-73 participants were British English native speakers living in United Kingdom (47 female), and 73 participants were Spanish native speakers living in Spain (27 female).
+Data collection took place from June 04th, 2020 to June 28th, 2020.
+We collected data from 104 participants (\emph{Mean} = 21.79 years, \emph{SD} = 2.43, Range = 18-33).
+72 participants were British English native speakers living in United Kingdom (46 female), and 32 participants were Spanish native speakers living in Spain (27 female).
 Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
 Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants were asked to complete the experiment in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were diagnosed with a language (\emph{n} = 2)\footnote{We originally planned to exclude participants that reported any visual impairment that glasses would not correct This item was phrased as ``Do you have normal or corrected-to-normal VISION?
-(Yes/No)'' in English, and as ``¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
-(Sí/No).'' However, the proportion of Spanish participants that reported visual impairment was implausibly large (\emph{n} = 6, 18.75\%).
-This is possibly due to these participants using glasses dayly and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision.}, or c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the local ethical committee (XXXXXXXXXXXX).
+Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
+We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were diagnosed with a language (\emph{n} = 2) , or c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
 
@@
 
-The experiment was implemented online using Psychopy/Pavlovia (\textbf{peirce2019?}).
-Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from a browser (Chrome or Mozilla).
-Participants were informed about the aims of the study and gave informed consent for participating.
-Then, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
-Third, participants completed the experimental task.
+The experiment was implemented online using Psychopy/Pavlovia (Peirce et al., 2019).
+Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla).
+After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+Then, participants completed the experimental task.
 Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
@@
 
-Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background.
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{C:/Users/u155880/Documents/translation-elicitation/Figures/design} \caption{ }\label{fig:procedurefigure}
+\end{figure}
+
+Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
 After one second, the audio started playing while the dot remained being displayed until the audio ended.
@@
 
+We arranged two lists of words: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
+Words in the Catalan list (listened to by participants) were 5.01 phonemes long on average (\emph{SD} = 1.48, Range = 2-8).
+Their translations to English (typed by participants in the keyboard) were 5.14 characters long on average (\emph{SD} = 1.57, Range = 3-9), and their translations to Spanish were 5.44 characters long on average (\emph{SD} = 1.56, Range = 3-9).
+Words in the Spanish list were 5.50 phonemes long on average (\emph{SD} = 1.46, Range = 3-9).
+Their translations to English were 5.27 characters long on average (\emph{SD} = 1.75, Range = 3-12).
+
+We extracted the lexical frequencies from SUBTLEX-UK for English words (Van Heuven, Mandera, Keuleers, \& Brysbaert, 2014), SUBTLEX-ESP for Spanish words (Cuetos, Glez-Nosti, Barbon, \& Brysbaert, 2011), and SUBTLEX-CAT for Catalan words (Boada, Guasch, Haro, Demestre, \& Ferré, 2020).
+We extracted or transformed scores as/into Zipf scores (Van Heuven et al., 2014) to so correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses (Van Heuven et al., 2014).
+
+We retrieved PTHN scores from the CLEARPOND database (Marian, Bartolotti, Chabal, \& Shook, 2012).
+PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
+CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, \url{http://espeak.sourceforge.net/}) differs from that of the target word in only one addition, deletion, or substitution.
+PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
+
+Finally, we measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the \texttt{stringsim} function of the stringdist R package (van der Loo, 2014).
+This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
+First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical (Levenshtein \& others, 1966).
+This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
+The result is a score that ranges from 0 to 1, where 0 indicates no similarity between the two strings and one indicates that both strings are identical.
+We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
+Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+
 Participants listened to one audio file in each trial.
 This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infancia of Universitat Pompeu Fabra (Barcelona, Spain).
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
 These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat (\textbf{boersma2018?}).
-The average duration of the audios was NA (\emph{SD} = NA, \emph{Min} = \(\infty\), \emph{Max} = \(-\infty\)).
-The average lexical frequency of the words was 4.29 (\emph{SD} = 0.59, \emph{Min} = 2.87, \emph{Max} = 6.14).
-The average lexical frequency of the Catalan audios was 4.31 seconds (\emph{SD} = 0.59, \emph{Min} = 2.87, \emph{Max} = 5.68), and the average lexical frequency of the Spanish audios was 4.32 seconds (\emph{SD} = 0.56, \emph{Min} = 2.87, \emph{Max} = 5.68).
-Lexical frequencies in Catalan and Spanish were extracted from SUBTLEX-CAT (\textbf{boada2013?}) and SUBTLEX-ESP (\textbf{cuetos2011?}), respectively, and are expressed as Zipf scores (\textbf{van-heuven2012?}; \textbf{zipf1949?}).
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat (Broersma \& Weenink, 2021).
+The average duration of the audios was 1.20 (\emph{SD} = 0.18, Range = 0.78-1.58).
+The average duration of the Catalan audios was 1.23 seconds (\emph{SD} = 0.19, Range = 0.80-1.58), and the average duration of the Spanish audios was 1.16 seconds (\emph{SD} = 0.15, Range = 0.78-1.53).
+
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lrrrrrrrr}
+\toprule
+ & \multicolumn{2}{c}{\textbf{Freq./million}} & \multicolumn{2}{c}{\textbf{Freq. (Zipf)}} & \multicolumn{2}{c}{\emph{\textbf{PTHN}}} & \multicolumn{2}{c}{\textbf{Levenshtein}} \\ 
+ \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
+ & Mean & \emph{SD} & Mean & \emph{SD} & Mean & \emph{SD} & Mean & \emph{SD} \\ 
+\midrule
+ENG-CAT & $46.80$ & $77.23$ & $4.31$ & $0.59$ & $3.26$ & $3.74$ & $0.16$ & $0.19$ \\ 
+ENG-SPA & $43.91$ & $67.53$ & $4.32$ & $0.56$ & $3.52$ & $4.10$ & $0.13$ & $0.17$ \\ 
+SPA-CAT & $53.32$ & $156.19$ & $4.23$ & $0.62$ & $1.62$ & $2.62$ & $0.37$ & $0.25$ \\ 
+ \bottomrule
+\end{longtable}
 
@@
 
-We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT).
-We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group.
+We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We first fitted a base model (Model 0) that only included the lexical frequency (\texttt{frequency}) of the target word as a fixed effect, with a random intercept per participant.
+Second, we extended the model to include \texttt{pthn} as a fixed effect, with a random slope by participant (Model 1).
+Third, we added the fixed effect \texttt{consonant\_ratio} and the \texttt{pthn:consonant\_ratio}, and random slopes for both effects by participant (Model 2).
+Third, we added the fixed effect \texttt{vowel\_ratio} and the \texttt{pthn:vowel\_ratio}, and random slopes for both effects by participant (Model 3).
+Finally, we fit a model that included the two-way interactions \texttt{pthn:consonant\_ratio} and \texttt{pthn:vowel\_ratio}, an their random slopes by participant (Model 4).
+All predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
+Model 4 can be formally expressed as:
 
-We fit several Bayesian linear models (\textbf{sorensen2015?}), each including one of the aforementioned variables as predictor, and the proportion of correct responses as output.
-Continuous predictors were standardised and categorical predictors were sum-coded (\textbf{schad2020?}).
-Missing predictor scores were imputed before fitting the models using multiple imputation (\textbf{vanbuuren2013?}).
 To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (LOO, (\textbf{vehtari2016?})).
-More information about the models and model comparison can be found in Appendix X.
-
-All analyses were performed in R environment (\textbf{rcoreteam2019?}).
-We used the \texttt{tidyverse} family of R packages to process data and to generate figures, the \texttt{mice} R package (\textbf{vanbuuren2013?}) for multiple imputation of missing data, and the \texttt{brms} R package (\textbf{burkner2017?}) to fit and compare models, which uses the Stan probabilistic programming language language (\textbf{carpenter2016?}).
+We compared models using leave-one-out cross-validation (\emph{LOO}) (Vehtari, Gelman, \& Gabry, 2017).
+More information about the models and model comparison can be found in Appendix 2.
+All analyses were performed in R environment (RCore, 2019).
+We used the tidyverse family of R packages to process data and to generate figures, and the brms R package (Bürkner, 2017) using the cmdstanr backend to the Stan probabilistic language (Carpenter et al., 2017) to estimate and compare the models (see Appendix 1 for mode details on the models).
 
@@
 
-\hypertarget{posterior-draws}{%
-\subsection{Posterior draws}\label{posterior-draws}}
+All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
+Model 4, which included all main effects and the two-way interactions between PTHN and vowel similarity, and PTHN and consonant similarity, showed the best performance (see Table 2).
 
-\hypertarget{posterior-predictions}{%
-\subsection{Posterior predictions}\label{posterior-predictions}}
+\captionsetup[table]{labelformat=empty,skip=1pt}
+\begin{longtable}{lrrrrrr}
+\toprule
+ & \emph{\textbf{LOOELPD}} & \emph{\textbf{SE}} & \emph{\textbf{LOO\_\{IC\}}} & \textbf{\emph{SE}IC} & \emph{\textbf{LOOdiff}} & \emph{\textbf{SEdiff}} \\ 
+\midrule
+Model 3 & $-4,277.50$ & $44.29$ & $8,555.00$ & $88.58$ & - & - \\ 
+Model 2 & $-4,295.13$ & $44.44$ & $8,590.26$ & $88.89$ & $-17.63$ & $6.08$ \\ 
+Model 1 & $-4,310.21$ & $44.50$ & $8,620.43$ & $89.00$ & $-32.72$ & $8.28$ \\ 
+Model 0 & $-4,321.68$ & $44.31$ & $8,643.37$ & $88.62$ & $-44.19$ & $9.44$ \\ 
+ \bottomrule
+\end{longtable}
 
-\hypertarget{random-effects}{%
-\subsection{Random effects}\label{random-effects}}
+We now report the mean of the posterior distribution of each coefficient in Model 3, along with its associated measures of uncertainty.
+For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
+Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
+This way, the coefficients are expressed as increases/decreases in probability of correct translation (Gelman, Hill, \& Vehtari, 2020).
 
-\hypertarget{item-effects}{%
-\subsubsection{Item effects}\label{item-effects}}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/posteriorfix-1.pdf}
+\caption{\label{fig:posteriorfix}Estimated posterior distributions of coefficients in Model 3.
+A) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of each coefficient.
+Credible intervals \(CrI\), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\%, 80\%, and 50\% probability.
+Dots represent the mean of the distribution.
+B) Participant-level coefficient variability.
+Our model estimated participant-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations \(SD\).
+C) Correlation between participant-level effects.
+Our model allowed participant-level coefficients to co-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each correlation.
+Coefficients are represented in the X-axis and Y-axis in the same order as indicated in the Y-axis of panels A and C.}
+\end{figure}
 
-\hypertarget{group-effects}{%
-\subsubsection{Group effects}\label{group-effects}}
+Overall, participants were 35.45\%, (\emph{SE} = 52.06\%, 95\% \emph{CrI} = {[}32.09\%, 39.47\%{]}) likely to produce correct translations.
+Every standard deviation increment in the translation's lexical frequency (\emph{SD} = 0.59) increased the probability of a correct responses in 3.53\% (\emph{SE} = 0.67\%, 95\% \emph{CrI} = {[}2.21\%, 4.85\%{]}).
+The number of the translation's more frequent phonological neighbours, on the other hand, decreased the probability of a correct responses in -3.90\% (\emph{SE} = 0.76\%, 95\% \emph{CrI} = {[}-5.42\%, -2.43\%{]}) for every increase in 1 \emph{SD} (3.65).
+The effect of phonological similarity was conditional to the phonological density of the translation.
+Phonological similarity barely increased the probability of a correct responses (0.78\%, \emph{SE} = 1\%, 95\% \emph{CrI} = {[}-1.02\%, 3\%{]}) by itself.
+However, for every \emph{SD} increase in PTHN, phonological similarity increased in -2\%, (\emph{SE} = 0.80\%, 95\% \emph{CrI} = {[}-3.78\%, -0.68\%{]}) such probability.
 
-\hypertarget{correlations}{%
-\subsubsection{Correlations}\label{correlations}}
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/marginaleffects-1.pdf}
+\caption{\label{fig:marginaleffects}Expected mean posterior predictions.
+A) Population-level expected mean posterior predictions.
+The X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 200 observations from our model: 100 simulations for words with low PTHN (-1 SD) and 100 simulations for high PTHN (+1 SD) words.
+We did this across the range of values of the Levenshtein scores.
+For each simulation, we drew a single sample from the posterior distribution of each coefficient.
+Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high PTHN words.
+We also plotted the mean of the high PTHN (black solid line) and low PTHN (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
+The dispersion of the lines indicates the uncertainty of our predictions.
+B) Participant-level expected mean posterior predictions.
+We conducted the same procedure for each individual participant, simulating 100 observations for high-PTHN and 100 for low-PTHN from our model acrosss the range of Levenshtein scores.
+We averaged the resulting predictions for each participant and then plotted each participant's predictions in separate panels to show how the effect of PTHN and phonological similarity changed at the individual level.}
+\end{figure}
 
@@
 \hypertarget{refs}{}
-\begin{CSLReferences}{0}{0}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-boada2020subtlex}{}}%
+Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P.
+(2020).
+SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for catalan.
+\emph{Behavior Research Methods}, \emph{52}(1), 360--375.
+
+\leavevmode\vadjust pre{\hypertarget{ref-broersma2021praat}{}}%
+Broersma, P., \& Weenink, D. (2021). Praat: Doing phonetics by computer {[}computer program{]} (Version Version 6.1.54). Retrieved from \url{http://www.praat.org/}
+
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
+Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+
+\leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
+Carpenter, B., Gelman, A., Hoffman, M.
+D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A.
+(2017).
+Stan: A probabilistic programming language.
+\emph{Grantee Submission}, \emph{76}(1), 1--32.
+
+\leavevmode\vadjust pre{\hypertarget{ref-christoffels2006memory}{}}%
+Christoffels, I.
+K., De Groot, A.
+M., \& Kroll, J.
+F.
+(2006).
+Memory and language skills in simultaneous interpreters: The role of expertise and language proficiency.
+\emph{Journal of Memory and Language}, \emph{54}(3), 324--345.
+
+\leavevmode\vadjust pre{\hypertarget{ref-christoffels2013language}{}}%
+Christoffels, I. K., Ganushchak, L., \& Koester, D. (2013). Language conflict in translation: An ERP study of translation production. \emph{Journal of Cognitive Psychology}, \emph{25}(5), 646--664.
+
+\leavevmode\vadjust pre{\hypertarget{ref-collins1975spreading}{}}%
+Collins, A. M., \& Loftus, E. F. (1975). A spreading-activation theory of semantic processing. \emph{Psychological Review}, \emph{82}(6), 407.
+
+\leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
+Costa, A., Caramazza, A., \& Sebastian-Galles, N.
+(2000).
+The cognate facilitation effect: Implications for models of lexical access.
+\emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{26}(5), 1283.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cuetos2011subtlex}{}}%
+Cuetos, F., Glez-Nosti, M., Barbon, A., \& Brysbaert, M.
+(2011).
+SUBTLEX-ESP: Frecuencias de las palabras espanolas basadas en los subtitulos de las peliculas.
+\emph{Psicol{ó}gica}, \emph{32}(2), 133--144.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de2000hard}{}}%
+De Groot, A.
+M., \& Keijzer, R.
+(2000).
+What is hard to learn is easy to forget: The roles of word concreteness, cognate status, and word frequency in foreign-language vocabulary learning and forgetting.
+\emph{Language Learning}, \emph{50}(1), 1--56.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de1982relationship}{}}%
+De Haan, H. J. (1982). The relationship of estimated comprehensibility to the rate of connected speech. \emph{Perception \& Psychophysics}, \emph{32}(1), 27--31.
+
+\leavevmode\vadjust pre{\hypertarget{ref-degroot1994forward}{}}%
+Degroot, A. M., Dannenburg, L., \& Vanhell, J. G. (1994). Forward and backward word translation by bilinguals. \emph{Journal of Memory and Language}, \emph{33}(5), 600--629.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dijkstra2019multilink}{}}%
+Dijkstra, T., Wahl, A., Buytenhuijs, F., Van Halem, N., Al-Jibouri, Z., De Korte, M., \& Rekké, S.
+(2019).
+Multilink: A computational model for bilingual word recognition and word translation.
+\emph{Bilingualism: Language and Cognition}, \emph{22}(4), 657--679.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
+Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-goldinger1989priming}{}}%
+Goldinger, S. D., Luce, P. A., \& Pisoni, D. B. (1989). Priming lexical neighbors of spoken words: Effects of competition and inhibition. \emph{Journal of Memory and Language}, \emph{28}(5), 501--518.
+
+\leavevmode\vadjust pre{\hypertarget{ref-de1992determinants}{}}%
+Groot, A. M. de. (1992). Determinants of word translation. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{18}(5), 1001.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kroll1994category}{}}%
+Kroll, J.
+F., \& Stewart, E.
+(1994).
+Category interference in translation and picture naming: Evidence for asymmetric connections between bilingual memory representations.
+\emph{Journal of Memory and Language}, \emph{33}(2), 149--174.
+
+\leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
+Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. In \emph{Soviet physics doklady} (Vol. 10, pp. 707--710). Soviet Union.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lotto1998effects}{}}%
+Lotto, L., \& De Groot, A. M. (1998). Effects of learning method and word type on acquiring vocabulary in an unfamiliar language. \emph{Language Learning}, \emph{48}(1), 31--69.
+
+\leavevmode\vadjust pre{\hypertarget{ref-luce1998recognizing}{}}%
+Luce, P. A., \& Pisoni, D. B. (1998). Recognizing spoken words: The neighborhood activation model. \emph{Ear and Hearing}, \emph{19}(1), 1.
+
+\leavevmode\vadjust pre{\hypertarget{ref-luce1990similarity}{}}%
+Luce, P. A., Pisoni, D. B., \& Goldinger, S. D. (1990). Similarity neighborhoods of spoken words.
+
+\leavevmode\vadjust pre{\hypertarget{ref-marian2012clearpond}{}}%
+Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012). CLEARPOND: Cross-linguistic easy-access resource for phonological and orthographic neighborhood densities.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mcclelland1981interactive}{}}%
+McClelland, J. L., \& Rumelhart, D. E. (1981). An interactive activation model of context effects in letter perception: I. An account of basic findings. \emph{Psychological Review}, \emph{88}(5), 375.
+
+\leavevmode\vadjust pre{\hypertarget{ref-peirce2019psychopy2}{}}%
+Peirce, J., Gray, J.
+R., Simpson, S., MacAskill, M., Höchenberger, R., Sogo, H., \ldots{} Lindeløv, J.
+K.
+(2019).
+PsychoPy2: Experiments in behavior made easy.
+\emph{Behavior Research Methods}, \emph{51}(1), 195--203.
+
+\leavevmode\vadjust pre{\hypertarget{ref-potter1984lexical}{}}%
+Potter, M.
+C., So, K.-F., Von Eckardt, B., \& Feldman, L.
+B.
+(1984).
+Lexical and conceptual representation in beginning and proficient bilinguals.
+\emph{Journal of Verbal Learning and Verbal Behavior}, \emph{23}(1), 23--38.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rcore2019r}{}}%
+RCore, T. (2019). R: A language and environment for statistical computing. R foundation for statistical computing, austria.
+
+\leavevmode\vadjust pre{\hypertarget{ref-schepens2012distributions}{}}%
+Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of cognates in europe as based on levenshtein distance. \emph{Bilingualism: Language and Cognition}, \emph{15}(1), 157--166.
+
+\leavevmode\vadjust pre{\hypertarget{ref-thierry2007brain}{}}%
+Thierry, G., \& Wu, Y.
+J.
+(2007).
+Brain potentials reveal unconscious translation during foreign-language comprehension.
+\emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535.
+
+\leavevmode\vadjust pre{\hypertarget{ref-van2014stringdist}{}}%
+van der Loo, M. P. J. (2014). The stringdist package for approximate string matching. \emph{The {R} {J}ournal}, \emph{6}, 111--122. Retrieved from \url{https://CRAN.R-project.org/package=stringdist}
+
+\leavevmode\vadjust pre{\hypertarget{ref-van1998orthographic}{}}%
+Van Heuven, W. J., Dijkstra, T., \& Grainger, J. (1998). Orthographic neighborhood effects in bilingual word recognition. \emph{Journal of Memory and Language}, \emph{39}(3), 458--483.
+
+\leavevmode\vadjust pre{\hypertarget{ref-van2014subtlex}{}}%
+Van Heuven, W.
+J., Mandera, P., Keuleers, E., \& Brysbaert, M.
+(2014).
+SUBTLEX-UK: A new and improved word frequency database for british english.
+\emph{Quarterly Journal of Experimental Psychology}, \emph{67}(6), 1176--1190.
+
+\leavevmode\vadjust pre{\hypertarget{ref-vehtari2017practical}{}}%
+Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model evaluation using leave-one-out cross-validation and WAIC. \emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
+
+\leavevmode\vadjust pre{\hypertarget{ref-warren1970perceptual}{}}%
+Warren, R. M. (1970). Perceptual restoration of missing speech sounds. \emph{Science}, \emph{167}(3917), 392--393.
+
 \end{CSLReferences}
@@
 
+\newpage
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_files/figure-latex/accuracy-1.pdf}
+\caption{\label{fig:accuracy}Proportion of correct translation by item.
+Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.
+Dots and whiskers represent the average accuracy and 95\% confidence interval of each word.
+Accuracy is plotted separately for each group.}
+\end{figure}
+
 \end{document}

### 2021-12-02 Rmd/manuscript.Rmd
Add comments by Chiara Santolin and APA 7th style

@@
 draft: no
-documentclass: apa6
+csl: "apa7.csl"
+documentclass: apa7
 classoption: man
 output:
-  papaja::apa6_pdf:
-    keep_tex: FALSE
+  word_document:
+    reference_docx: "manuscript_template.docx"
 ---
@@
 
-## Native speech processing is effortless
+Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech [@de1982relationship] or the loss of segmental information [@warren1970perceptual].
+But even in the most ideal of the situations, the cognitive processes engaged in word recognition do not occur without uncertainty.
+The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
+There is extensive literature about the influence on word recognition of the number of lexical candidates, their frequency, and their  phonological and semantic similarity with target words.
+Less is known about how all these factors affect the dynamics of word recognition in non-native languages.
 
-Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech [@de1982relationship] or the loss of segmental information [@warren1970perceptual].
-But even in the most ideal of the situations, the processes engaged in word recognition do not occur without uncertainty.
-The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
-There is extensive literature about how the number of candidates, their frequency, and their phonological and semantic similarity with the target word impacts the dynamics of word recognition.
-But less is known about how these factors affect the recognition of words in a non-native language.
+Listening to speech in a non-native language is more costly than doing so in native languages, regardless of the level of proficiency [REF].
+The source of this increased cognitive effort is likely to originate at multiple levels.
+For example, some sounds in the speech signal do not correspond to any phoneme in the native language, making the process of matching sounds to phonemes complicated.
+This is the case of a Spanish native listening to words containing the guttural /ʁ/ in French.
+Language learners, however, are rarely completely naïve to the language they are listening to.
+All languages share, to some extent, similarities, frequently due to their typological closeness.
+These similarities which can be exploited by non-native listeners to process the speech signal.
+In the present research we investigate the extent to which adult listeners rely on phonological similarity between cross-language synonyms during word translation.
 
-## Non-antive speech processing is costly
-
-Listening to speech in a non-native language is more costly than doing so in the native language, even if one is proficient in such language.
-The source of this increased cognitive effort is likely to originate at multiple levels.
-Some examples: (1) some sounds in the speech signal do not correspond to any phoneme in the native language, (2) segmenting the speech signal relies on previous familiarity with word forms or with the statistical regularities of the language: (3) word order might be different in both languages (e.g., subject-object-verb vs.
-subject-verb-object.).
-The listener, however, is rarely completely naïve to the language they are listening to.
-All languages share, to some extent, similarities, frequently due to their typological closeness.
-These similarities which can be exploited by non-native listeners.
-
-## Cognateness aids non-native speech processing
-
-One of such commonalities occurs at the lexical level, in the for of cognateness.
 Cognates are cross-language synonyms whose form (e.g., phonology, orthography, signature) is similar.
-The cause of this similarity is frequently attributed to a shared etymological origin.
-Romance languages like Spanish, Italian, French or Catalan, share many cognates [@schepens2012distributions].
-For example, this is the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively)^[Some form-similar cross-language synonyms are technically not cognates.
+The cause of such similarity is frequently attributed to a shared etymological origin.
+Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively)^[Some form-similar cross-language synonyms are technically not cognates.
 For example, *sun* and *sol* (in Spanish), share their phonological onset, but their etymology points to different origins.
@@
 
-## The cognate advantage in comprehension, production, and learning
+Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+During word production and comprehension, lexical representations of both languages are activated, and the form similarity between both impacts bilinguals' performance in naming or lexical decision tasks [e.g., @costa2000cognate; @thierry2007brain].
+In a seminal study, @costa2000cognate asked Spanish-Catalan adult bilinguals to name pictures in Spanish.
+Unbeknownst to participants, half of the labels associated to pictures were cognates in Spanish and Catalan (e.g., *puerta*-*porta*) while the other half were non-cognates (e.g., *mesa*-*taula*, meaning *table* in Spanish and Catalan).
+Surprisingly, Spanish-Catalan bilinguals named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
+This suggests that bilinguals activate the picture's word representations in both languages, and the phonological overlap between both labels facilitated the naming process.
 
-Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access their lexicon in a language non-selective way: during word production and comprehension, both lexical representations are activated, and the form similarity between both impacts participants' performance in naming or lexical decision tasks [e.g., @costa2000cognate; @thierry2007brain].
-For instance, in their seminal study, @costa2000cognate asked Spanish-Catalan bilinguals to name pictures in Spanish.
-Unbeknownst to participants, half of the pictures' associated labels were cognates in Spanish and Catalan (e.g., *puerta*-*porta*) while the other half were non-cognates (e.g., *mesa*-*taula*).
-Surprisingly, participants named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
-This suggested that bilinguals activated picture's word representations in both languages, and that the phonological overlap between both labels facilitated the naming process.
-
-There is also evidence that cognates are learnt more easily in L2 than non-cognates.
-For instance, @de2000hard presented 40 Dutch natives 60 pairs of translation equivalents.
-In each learning trial, two words were presented in a screen side-by-side: one in Dutch and one pseudoword.
-Pseudowords were generated in such way that were easily pronounceable and were phonotactically legal in Dutch.
-Word pairs varied in their form-similarity (number of shared letters, 40-70%).
-Participants were tested twice in the same task with one week of difference.
-Participants' performance was better for cognates across both testing sessions, suggesting that cognates were learnt and retained more easily than non-cognates.
-Converging evidence was provided by @lotto1998effects in Dutch natives learning Italian words: cognates were learnt more easily than non-cognates.
-
-### The cognate advantage in translation: the RMH model
+Additionally, cognates seem to be learnt more easily in L2 than non-cognates.
+@de2000hard presented Dutch natives pairs of Dutch words and pseudowords.
+Participants' performance learnt and retained better pairs with high orthographic similarity than pairs with low orthographic similarity [see also @lotto1998effects for Dutch natives learning Italian words].
 
 Cognates are not only learnt faster than non-cognates, but translated faster too.
-The mechanisms behind this effect, though, are still unclear.
-Early accounts of bilingual lexical access suggested that low-proficiency learners first established links between newly learnt words in L2 and their meanings through their L1 translation equivalents [@potter1984lexical].
+Early accounts of bilingual lexical access suggest that low-proficiency learners first established links between newly acquired words in L2 and their meanings through their L1 translation equivalents [@potter1984lexical].
 As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
@@
 
-To test these predictions, @degroot1994forward and @de1992determinants asked 52 Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
-In each trial, participants were presented visually with a word in English or Dutch, and were asked to speak out loud its translation in the other language as soon as possible.
-The authors reported three main findings: (1) translation times and accuracy were roughly equivalent in across both forward and backward conditions (although slightly faster in the former), (2) semantic variables, such as imageability, were positively associated with participants' performance more strongly in the forward translation conditional, although the effect size was small, (3) cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward translation than during forward translation, and (4) when translating cognates participants' performance across both conditions was less sensitive to semantic variables than when translating non-cognates.
-These results prompted the authors to reject a hard version @kroll1994category's account, and suggested that both conceptually mediated and direct translation routes are active during translation, but backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness than forward translation.
-When the authors tested a group of bilinguals with higher proficiency in English, their results pointed in the same direction.
-Subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language], contrary to the predictions of the RHM.
+To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+Participants were presented with written words in English or Dutch, and asked to speak out loud their translation in the other language as fast as possible.
+Results who that translation times and accuracy were roughly equivalent in across both forward and backward conditions, and that semantic variables (e.g., concreteness), were positively associated with participants' performance, especially in the forward translation condition.
+Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected  by semantic variables than non-cognates.
 
-## The impact of neighbourhoods lexical processing: the NAM and the BIA/BIA+ models
+These findings suggest that both conceptually mediated and direct translation routes are active during translation, although backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness (than forward translation).
+Interestingly, higher-proficiency bilinguals tested in the same task showed similar results.
+However, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language], contrary to the predictions of the RHM.
+Further models build upon this model, introducing a wider range of connections in their theorised lexica.
 
-The idealised lexicon considered by the models described above only considers word-word connections between translation equivalents.
-More recent studies have highlighted the role of the rich network of connections that given word establishes with other phonologically or conceptually related words.
-This notion was first introduced in monolingual research by @collins1975spreading in their theory of semantic processing, and later @luce1998recognizing formalised this dimension in their Neighbourhood Activation Model (NAM).
-In this model, lexical selection is mediated not only by the lexical frequency of the target word, and the number of activated candidates, but also by the frequency of the candidates.
-@luce1998recognizing designed a lexical decision task in which English native participants listened to a word across several conditions of noise-to-signal ratio.
-The authors then used a computerised lexicon to calculate the number of phonological neighbours around each of the presented words, based on phonetic similarity matrices, and the average lexical frequency of such neighbourhoods.
-Participants answered more faster and more accurately to high frequency words.
-Words from high density neighbourhoods were responded to more accurately, but more slowly.
-The average lexical frequency of the neighbourhood was associated to a decrease in accuracy.
-Low frequency words were responded to more accurately in low density neighbours than in high density neighbours.
-The authors concluded that, although both word lexical frequency and neighbourhood density can, in principle, facilitate spoken word recognition, this effect was modulated by the frequency of the neighbourhood: when phonological neighbours are more frequent, lexical selection is hindered [@goldinger1989priming; @luce1990similarity].
+Since @kroll1994category proposed their RHM model, subsequent studies brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+The Neighbourhood Activation Model [NAM; @luce1998recognizing] suggests that lexical selection is mediated not only by the lexical frequency of the target word (and the number of activated candidates), but also by the frequency of the candidates.
 
-The fact that phonological neighbourhoods and their structure impacts lexical processing poses an important question for bilingualism research.
-Are word representations in one language part of phonological or orthographic neighbours in the other language?
-@van1998orthographic addressed this issue in an experimental series and proposed the Bilingual Interactive Activation (BIA, later revised and relabelled as BIA+).
-In line with connectionist accounts of lexical processing this model advocated for lexical representations to establish both excitatory and inhibitory connections with other representations from the same of the other language [@mcclelland1981interactive], forming an integrated lexicon for both languages.
-@van1998orthographic tested Dutch-English bilinguals in a progressive demasking task (in each trial a words in presented in increasingly longer time windows interrupted by a checkerboard until the participant presses a button to type the word in a keyboard) or a lexical decision task (words a visually presented and the participant must decide as quickly as possible whether its a word in the target language or a non-word).
-The authors manipulated by size of the orthographic neighbourhood size of the words presented in its language (target language) or in the other language (non-target language).
-Interestingly, the authors reported participants' performance to be affected by the neighbourhood size of the presented word in the non-target language across both tasks.
-Specifically, participants took longer to respond to words with high neighbourhood density in the non-target language, even when participants completed the task in the target language exclusively (e.g., were presented with Dutch words exclusively).
-Monolinguals, on the other hand, were only affected by neighbourhood size in the target language.
-This provided evidence of cross-language interference between the word representation in the non-target language and the those in the target language.
+Subsequent behavioural experiments [e.g., @luce1998recognizing] show that English adult participants were tested in a lexical decision task in which words were presented across several conditions of noise-to-signal ratio.
+A computerised lexicon calculated number and average lexical frequency of phonological neighbours around each of the presented words.
+Participants answered more faster and more accurately to high frequency words, and words from high density neighbourhoods were also responded detected more accurately, but slowly.
+Therefore, average lexical frequency of  neighbourhoods was associated to a decrease in accuracy.
+Low frequency words were in fact detected more accurately in low density than in high density neighbours.
+These findings suggested that both word lexical frequency and neighbourhood density can facilitate spoken word recognition, but this effect is modulated by the frequency of the neighbourhood; when phonological neighbours are more frequent, lexical selection is hindered [@goldinger1989priming; @luce1990similarity].
 
-## Is the lexical route necessary for translation?: The Multilink model
+The influence of phonological neighbourhoods and their structure on lexical processing poses rise important question in the field of bilingualism research.
+Do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+The Bilingual Interactive Activation [BIA+; @van1998orthographic] proposes that lexical representations establish both excitatory and inhibitory connections with other representations from the same or the other language, forming an integrated lexicon for both languages [@mcclelland1981interactive].
+Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
+@van1998orthographic presented bilingual participants with words in only one language (target language) in a lexical decision task.
+When the translation of the presented word in the non-target language was part of a large neighbourhood, reaction times were slower.
+This provides evidence that the presented words activated orthographic neighbourhoods in the non-target language, which slowed down word recognition.
 
-More recently, @dijkstra2019multilink implemented a localist-connectionist model, Multilink, that integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
-This model assumes (1) an integrated lexicon in which word representations from the two languages establish connections as words from the same language do; (2) during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1; (3) contrary to the RHM model's assumption that translation equivalents are linked by excitatory connections (e.g., translation can be done via word-word connections), @dijkstra2019multilink suggest that such connections would spread activation though irrelevant words in the target language and therefore initially assumed that translation equivalents are exclusively connect throught their shared concept; (4) the strength of the association between L2 representations and their concept is a function of the language user's proficiceny in L2.
-Multilink considers the particular case of Dutch-English bilinguals to generated simulations and test them against experimental data.
-The authors simulated data from Multilink to mirror @christoffels2006memory translation elicitation task, and reported a remarkable correlation between Multilink's simulations and participant's data, with forward translation being faster than backward translation, contrary to the RHM model predictions.
-More relevant to the present study is the fact that the model also generated data suggesting that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
-Overall, evidence from this model provides evidence for an integrated bilingual lexicon and for a facilitatory effect of cross-language similarity during translation.
-One of the main conclusions of @dijkstra2019multilink study is that word-word connections are not necessary for translation, as suggested by @kroll1994category: Multilink did not include such connections and still fitted experimental data.
-However, as they note, both their simulations and the data collected by @christoffels2006memory corresponded to high-proficiency bilinguals.
-In the case of the model simulations, vocabulary sizes were balanced in both languages and word-concept connections were equally strong for each member of the translation pairs in both languages.
-This scenario is actually contemplated by the @kroll1994category's RHM model.
-What @dijkstra2019multilink did not test, is the fit of Mutilink's predictions for low-proficiency, unbalanced bilinguals, whose vocabulary sizes are smaller or their word-concept connections are weaker in L2 than in L1.
-It is under this set of assumptions that the RHM model predicts participants to rely strongly on word-word connections to be able to translation words, especially from L2 to L1.
-This prediction, however, is untested in the Multilink model.
+A more recent model of bilingual lexical processing is Multilink @dijkstra2019multilink, which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
+This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
+Such model also assumes that during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1 and it proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model (which suggests that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
+In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
+A remarkable correlation between Multilink simulations of @christoffels2006memory translation elicitation task and behavioural data with Dutch-English bilinguals has been reported, showing forward translation being faster than backward translation, contrary to the predictions made by the RHM model.
+Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+Overall, results from Multilink from this model are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
 
-## The present study
+The fact that Multilink fitted the experimental data from @christoffels2006memory with remarkable success, even in absence of word-word connections between translation equivalents might put into question the necessity of such connections during translation, in favour of a simpler model in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+This claim, however, should be taken.
+It should be taken into account, however, that @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
+Additionally, data generated from Multilink model assumed similar vocabulary sizes across languages and, perhaps more critically in this regard, that word-concept connections were equally strong for L1 and L2 representations.
+These conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
+It is low-proficiency bilinguals whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation.
+This prediction remains untested my Multilink.
 
-In this study, we explored the plausibility of the lexical route in backward translation by testing monolingual participants.
-Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which which participants' vocabulary size in L2 is null, and word-concept connections are absent (i.e., set to zero in formal terms).
-In this scenario, participants can only rely on the similarity between the L2 form (presented as stimulus) and the L1 form (the target word).
-In other words, these participants can only use the lexical route to suceeed in the translation task.
-If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would sugges that the lexical route is in place for low-proficiency bilinguals.
+The role of this study was to investigate the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
+Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
+In this case, monolinguals can only use the lexical route to suceeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
+If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would suggest that the lexical route is in place for low-proficiency bilinguals.
 If participants are not able to translation words from L2 to L1 regardless of their form-similarity, this would suggest that low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
@@
 Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
-
-```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%"}
-
-include_graphics(here("Figures", "design.png"))
-
-```
-
 Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
@@
 The task took approximately 15 minutes to be completed.
+
+```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%", fig.cap="Schematic representation of a trial in the experimental task."}
+
+include_graphics(here("Figures", "design.png"))
+
+```
 
@@
 
-Finally, we measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist].
+We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist].
 This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
@@
 
-```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Stimuli data.
+Mean and standard deviation (SD) of lexical frequencies od target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words."}
 
@@
 
-```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=4}
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=4, fig.cap="Model comparison.
+Outcomes of the leave-one-out cross-validation procedure.
+Each row indicates the values of the expected log-predicted density (ELPD), leave-one-out information creterion (LOO-IC), and the difference in ELPD (LOO-diff) between each model and the inmediately simpler model.
+Each value is accompanied by a standard error (SE) indicating the uncertainty of its estimate."}
 

### 2021-12-02 Rmd/manuscript.tex
Add comments by Chiara Santolin and APA 7th style

@@
 
-\hypertarget{native-speech-processing-is-effortless}{%
-\subsection{Native speech processing is effortless}\label{native-speech-processing-is-effortless}}
+Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech (De Haan, 1982) or the loss of segmental information (Warren, 1970).
+But even in the most ideal of the situations, the cognitive processes engaged in word recognition do not occur without uncertainty.
+The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
+There is extensive literature about the influence on word recognition of the number of lexical candidates, their frequency, and their phonological and semantic similarity with target words.
+Less is known about how all these factors affect the dynamics of word recognition in non-native languages.
 
-Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech (De Haan, 1982) or the loss of segmental information (Warren, 1970).
-But even in the most ideal of the situations, the processes engaged in word recognition do not occur without uncertainty.
-The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
-There is extensive literature about how the number of candidates, their frequency, and their phonological and semantic similarity with the target word impacts the dynamics of word recognition.
-But less is known about how these factors affect the recognition of words in a non-native language.
+Listening to speech in a non-native language is more costly than doing so in native languages, regardless of the level of proficiency {[}REF{]}.
+The source of this increased cognitive effort is likely to originate at multiple levels.
+For example, some sounds in the speech signal do not correspond to any phoneme in the native language, making the process of matching sounds to phonemes complicated.
+This is the case of a Spanish native listening to words containing the guttural /ʁ/ in French.
+Language learners, however, are rarely completely naïve to the language they are listening to.
+All languages share, to some extent, similarities, frequently due to their typological closeness.
+These similarities which can be exploited by non-native listeners to process the speech signal.
+In the present research we investigate the extent to which adult listeners rely on phonological similarity between cross-language synonyms during word translation.
 
-\hypertarget{non-antive-speech-processing-is-costly}{%
-\subsection{Non-antive speech processing is costly}\label{non-antive-speech-processing-is-costly}}
-
-Listening to speech in a non-native language is more costly than doing so in the native language, even if one is proficient in such language.
-The source of this increased cognitive effort is likely to originate at multiple levels.
-Some examples: (1) some sounds in the speech signal do not correspond to any phoneme in the native language, (2) segmenting the speech signal relies on previous familiarity with word forms or with the statistical regularities of the language: (3) word order might be different in both languages (e.g., subject-object-verb vs.~subject-verb-object.).
-The listener, however, is rarely completely naïve to the language they are listening to.
-All languages share, to some extent, similarities, frequently due to their typological closeness.
-These similarities which can be exploited by non-native listeners.
-
-\hypertarget{cognateness-aids-non-native-speech-processing}{%
-\subsection{Cognateness aids non-native speech processing}\label{cognateness-aids-non-native-speech-processing}}
-
-One of such commonalities occurs at the lexical level, in the for of cognateness.
 Cognates are cross-language synonyms whose form (e.g., phonology, orthography, signature) is similar.
-The cause of this similarity is frequently attributed to a shared etymological origin.
-Romance languages like Spanish, Italian, French or Catalan, share many cognates (Schepens, Dijkstra, \& Grootjen, 2012).
-For example, this is the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively)\footnote{Some form-similar cross-language synonyms are technically not cognates.
+The cause of such similarity is frequently attributed to a shared etymological origin.
+Romance languages such as Spanish and Catalan share many cognates (Schepens, Dijkstra, \& Grootjen, 2012), as in the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively)\footnote{Some form-similar cross-language synonyms are technically not cognates.
 For example, \emph{sun} and \emph{sol} (in Spanish), share their phonological onset, but their etymology points to different origins.
@@
 
-\hypertarget{the-cognate-advantage-in-comprehension-production-and-learning}{%
-\subsection{The cognate advantage in comprehension, production, and learning}\label{the-cognate-advantage-in-comprehension-production-and-learning}}
+Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+During word production and comprehension, lexical representations of both languages are activated, and the form similarity between both impacts bilinguals' performance in naming or lexical decision tasks (e.g., Costa, Caramazza, \& Sebastian-Galles, 2000; Thierry \& Wu, 2007).
+In a seminal study, Costa et al.
+(2000) asked Spanish-Catalan adult bilinguals to name pictures in Spanish.
+Unbeknownst to participants, half of the labels associated to pictures were cognates in Spanish and Catalan (e.g., \emph{puerta}-\emph{porta}) while the other half were non-cognates (e.g., \emph{mesa}-\emph{taula}, meaning \emph{table} in Spanish and Catalan).
+Surprisingly, Spanish-Catalan bilinguals named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
+This suggests that bilinguals activate the picture's word representations in both languages, and the phonological overlap between both labels facilitated the naming process.
 
-Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access their lexicon in a language non-selective way: during word production and comprehension, both lexical representations are activated, and the form similarity between both impacts participants' performance in naming or lexical decision tasks (e.g., Costa, Caramazza, \& Sebastian-Galles, 2000; Thierry \& Wu, 2007).
-For instance, in their seminal study, Costa et al.
-(2000) asked Spanish-Catalan bilinguals to name pictures in Spanish.
-Unbeknownst to participants, half of the pictures' associated labels were cognates in Spanish and Catalan (e.g., \emph{puerta}-\emph{porta}) while the other half were non-cognates (e.g., \emph{mesa}-\emph{taula}).
-Surprisingly, participants named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
-This suggested that bilinguals activated picture's word representations in both languages, and that the phonological overlap between both labels facilitated the naming process.
-
-There is also evidence that cognates are learnt more easily in L2 than non-cognates.
-For instance, De Groot and Keijzer (2000) presented 40 Dutch natives 60 pairs of translation equivalents.
-In each learning trial, two words were presented in a screen side-by-side: one in Dutch and one pseudoword.
-Pseudowords were generated in such way that were easily pronounceable and were phonotactically legal in Dutch.
-Word pairs varied in their form-similarity (number of shared letters, 40-70\%).
-Participants were tested twice in the same task with one week of difference.
-Participants' performance was better for cognates across both testing sessions, suggesting that cognates were learnt and retained more easily than non-cognates.
-Converging evidence was provided by Lotto and De Groot (1998) in Dutch natives learning Italian words: cognates were learnt more easily than non-cognates.
-
-\hypertarget{the-cognate-advantage-in-translation-the-rmh-model}{%
-\subsubsection{The cognate advantage in translation: the RMH model}\label{the-cognate-advantage-in-translation-the-rmh-model}}
+Additionally, cognates seem to be learnt more easily in L2 than non-cognates.
+De Groot and Keijzer (2000) presented Dutch natives pairs of Dutch words and pseudowords.
+Participants' performance learnt and retained better pairs with high orthographic similarity than pairs with low orthographic similarity (see also Lotto \& De Groot, 1998 for Dutch natives learning Italian words).
 
 Cognates are not only learnt faster than non-cognates, but translated faster too.
-The mechanisms behind this effect, though, are still unclear.
-Early accounts of bilingual lexical access suggested that low-proficiency learners first established links between newly learnt words in L2 and their meanings through their L1 translation equivalents (Potter, So, Von Eckardt, \& Feldman, 1984).
+Early accounts of bilingual lexical access suggest that low-proficiency learners first established links between newly acquired words in L2 and their meanings through their L1 translation equivalents (Potter, So, Von Eckardt, \& Feldman, 1984).
 As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
@@
 
-To test these predictions, Degroot, Dannenburg, and Vanhell (1994) and Groot (1992) asked 52 Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
-In each trial, participants were presented visually with a word in English or Dutch, and were asked to speak out loud its translation in the other language as soon as possible.
-The authors reported three main findings: (1) translation times and accuracy were roughly equivalent in across both forward and backward conditions (although slightly faster in the former), (2) semantic variables, such as imageability, were positively associated with participants' performance more strongly in the forward translation conditional, although the effect size was small, (3) cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward translation than during forward translation, and (4) when translating cognates participants' performance across both conditions was less sensitive to semantic variables than when translating non-cognates.
-These results prompted the authors to reject a hard version Kroll and Stewart (1994)'s account, and suggested that both conceptually mediated and direct translation routes are active during translation, but backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness than forward translation.
-When the authors tested a group of bilinguals with higher proficiency in English, their results pointed in the same direction.
-Subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels, De Groot, \& Kroll, 2006; Christoffels, Ganushchak, \& Koester, 2013), contrary to the predictions of the RHM.
+To test these predictions, Degroot, Dannenburg, and Vanhell (1994) and Groot (1992) asked Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+Participants were presented with written words in English or Dutch, and asked to speak out loud their translation in the other language as fast as possible.
+Results who that translation times and accuracy were roughly equivalent in across both forward and backward conditions, and that semantic variables (e.g., concreteness), were positively associated with participants' performance, especially in the forward translation condition.
+Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected by semantic variables than non-cognates.
 
-\hypertarget{the-impact-of-neighbourhoods-lexical-processing-the-nam-and-the-biabia-models}{%
-\subsection{The impact of neighbourhoods lexical processing: the NAM and the BIA/BIA+ models}\label{the-impact-of-neighbourhoods-lexical-processing-the-nam-and-the-biabia-models}}
+These findings suggest that both conceptually mediated and direct translation routes are active during translation, although backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness (than forward translation).
+Interestingly, higher-proficiency bilinguals tested in the same task showed similar results.
+However, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels, De Groot, \& Kroll, 2006; Christoffels, Ganushchak, \& Koester, 2013), contrary to the predictions of the RHM.
+Further models build upon this model, introducing a wider range of connections in their theorised lexica.
 
-The idealised lexicon considered by the models described above only considers word-word connections between translation equivalents.
-More recent studies have highlighted the role of the rich network of connections that given word establishes with other phonologically or conceptually related words.
-This notion was first introduced in monolingual research by Collins and Loftus (1975) in their theory of semantic processing, and later Luce and Pisoni (1998) formalised this dimension in their Neighbourhood Activation Model (NAM).
-In this model, lexical selection is mediated not only by the lexical frequency of the target word, and the number of activated candidates, but also by the frequency of the candidates.
-Luce and Pisoni (1998) designed a lexical decision task in which English native participants listened to a word across several conditions of noise-to-signal ratio.
-The authors then used a computerised lexicon to calculate the number of phonological neighbours around each of the presented words, based on phonetic similarity matrices, and the average lexical frequency of such neighbourhoods.
-Participants answered more faster and more accurately to high frequency words.
-Words from high density neighbourhoods were responded to more accurately, but more slowly.
-The average lexical frequency of the neighbourhood was associated to a decrease in accuracy.
-Low frequency words were responded to more accurately in low density neighbours than in high density neighbours.
-The authors concluded that, although both word lexical frequency and neighbourhood density can, in principle, facilitate spoken word recognition, this effect was modulated by the frequency of the neighbourhood: when phonological neighbours are more frequent, lexical selection is hindered (Goldinger, Luce, \& Pisoni, 1989; Luce, Pisoni, \& Goldinger, 1990).
+Since Kroll and Stewart (1994) proposed their RHM model, subsequent studies brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+The Neighbourhood Activation Model {[}NAM; Luce and Pisoni (1998){]} suggests that lexical selection is mediated not only by the lexical frequency of the target word (and the number of activated candidates), but also by the frequency of the candidates.
 
-The fact that phonological neighbourhoods and their structure impacts lexical processing poses an important question for bilingualism research.
-Are word representations in one language part of phonological or orthographic neighbours in the other language?
-Van Heuven, Dijkstra, and Grainger (1998) addressed this issue in an experimental series and proposed the Bilingual Interactive Activation (BIA, later revised and relabelled as BIA+).
-In line with connectionist accounts of lexical processing this model advocated for lexical representations to establish both excitatory and inhibitory connections with other representations from the same of the other language (McClelland \& Rumelhart, 1981), forming an integrated lexicon for both languages.
+Subsequent behavioural experiments (e.g., Luce \& Pisoni, 1998) show that English adult participants were tested in a lexical decision task in which words were presented across several conditions of noise-to-signal ratio.
+A computerised lexicon calculated number and average lexical frequency of phonological neighbours around each of the presented words.
+Participants answered more faster and more accurately to high frequency words, and words from high density neighbourhoods were also responded detected more accurately, but slowly.
+Therefore, average lexical frequency of neighbourhoods was associated to a decrease in accuracy.
+Low frequency words were in fact detected more accurately in low density than in high density neighbours.
+These findings suggested that both word lexical frequency and neighbourhood density can facilitate spoken word recognition, but this effect is modulated by the frequency of the neighbourhood; when phonological neighbours are more frequent, lexical selection is hindered (Goldinger, Luce, \& Pisoni, 1989; Luce, Pisoni, \& Goldinger, 1990).
+
+The influence of phonological neighbourhoods and their structure on lexical processing poses rise important question in the field of bilingualism research.
+Do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+The Bilingual Interactive Activation {[}BIA+; Van Heuven, Dijkstra, and Grainger (1998){]} proposes that lexical representations establish both excitatory and inhibitory connections with other representations from the same or the other language, forming an integrated lexicon for both languages (McClelland \& Rumelhart, 1981).
+Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
 Van Heuven et al.
-(1998) tested Dutch-English bilinguals in a progressive demasking task (in each trial a words in presented in increasingly longer time windows interrupted by a checkerboard until the participant presses a button to type the word in a keyboard) or a lexical decision task (words a visually presented and the participant must decide as quickly as possible whether its a word in the target language or a non-word).
-The authors manipulated by size of the orthographic neighbourhood size of the words presented in its language (target language) or in the other language (non-target language).
-Interestingly, the authors reported participants' performance to be affected by the neighbourhood size of the presented word in the non-target language across both tasks.
-Specifically, participants took longer to respond to words with high neighbourhood density in the non-target language, even when participants completed the task in the target language exclusively (e.g., were presented with Dutch words exclusively).
-Monolinguals, on the other hand, were only affected by neighbourhood size in the target language.
-This provided evidence of cross-language interference between the word representation in the non-target language and the those in the target language.
+(1998) presented bilingual participants with words in only one language (target language) in a lexical decision task.
+When the translation of the presented word in the non-target language was part of a large neighbourhood, reaction times were slower.
+This provides evidence that the presented words activated orthographic neighbourhoods in the non-target language, which slowed down word recognition.
 
-\hypertarget{is-the-lexical-route-necessary-for-translation-the-multilink-model}{%
-\subsection{Is the lexical route necessary for translation?: The Multilink model}\label{is-the-lexical-route-necessary-for-translation-the-multilink-model}}
+A more recent model of bilingual lexical processing is Multilink Dijkstra et al.
+(2019), which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
+This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
+Such model also assumes that during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1 and it proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model (which suggests that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
+In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
+A remarkable correlation between Multilink simulations of Christoffels et al.
+(2006) translation elicitation task and behavioural data with Dutch-English bilinguals has been reported, showing forward translation being faster than backward translation, contrary to the predictions made by the RHM model.
+Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+Overall, results from Multilink from this model are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
 
-More recently, Dijkstra et al.
-(2019) implemented a localist-connectionist model, Multilink, that integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
-This model assumes (1) an integrated lexicon in which word representations from the two languages establish connections as words from the same language do; (2) during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1; (3) contrary to the RHM model's assumption that translation equivalents are linked by excitatory connections (e.g., translation can be done via word-word connections), Dijkstra et al.
-(2019) suggest that such connections would spread activation though irrelevant words in the target language and therefore initially assumed that translation equivalents are exclusively connect throught their shared concept; (4) the strength of the association between L2 representations and their concept is a function of the language user's proficiceny in L2.
-Multilink considers the particular case of Dutch-English bilinguals to generated simulations and test them against experimental data.
-The authors simulated data from Multilink to mirror Christoffels et al.
-(2006) translation elicitation task, and reported a remarkable correlation between Multilink's simulations and participant's data, with forward translation being faster than backward translation, contrary to the RHM model predictions.
-More relevant to the present study is the fact that the model also generated data suggesting that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
-Overall, evidence from this model provides evidence for an integrated bilingual lexicon and for a facilitatory effect of cross-language similarity during translation.
-One of the main conclusions of Dijkstra et al.
-(2019) study is that word-word connections are not necessary for translation, as suggested by Kroll and Stewart (1994): Multilink did not include such connections and still fitted experimental data.
-However, as they note, both their simulations and the data collected by Christoffels et al.
-(2006) corresponded to high-proficiency bilinguals.
-In the case of the model simulations, vocabulary sizes were balanced in both languages and word-concept connections were equally strong for each member of the translation pairs in both languages.
-This scenario is actually contemplated by the Kroll and Stewart (1994)'s RHM model.
-What Dijkstra et al.
-(2019) did not test, is the fit of Mutilink's predictions for low-proficiency, unbalanced bilinguals, whose vocabulary sizes are smaller or their word-concept connections are weaker in L2 than in L1.
-It is under this set of assumptions that the RHM model predicts participants to rely strongly on word-word connections to be able to translation words, especially from L2 to L1.
-This prediction, however, is untested in the Multilink model.
+The fact that Multilink fitted the experimental data from Christoffels et al.
+(2006) with remarkable success, even in absence of word-word connections between translation equivalents might put into question the necessity of such connections during translation, in favour of a simpler model in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+This claim, however, should be taken.
+It should be taken into account, however, that Christoffels et al.
+(2006) only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
+Additionally, data generated from Multilink model assumed similar vocabulary sizes across languages and, perhaps more critically in this regard, that word-concept connections were equally strong for L1 and L2 representations.
+These conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
+It is low-proficiency bilinguals whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation.
+This prediction remains untested my Multilink.
 
-\hypertarget{the-present-study}{%
-\subsection{The present study}\label{the-present-study}}
-
-In this study, we explored the plausibility of the lexical route in backward translation by testing monolingual participants.
-Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which which participants' vocabulary size in L2 is null, and word-concept connections are absent (i.e., set to zero in formal terms).
-In this scenario, participants can only rely on the similarity between the L2 form (presented as stimulus) and the L1 form (the target word).
-In other words, these participants can only use the lexical route to suceeed in the translation task.
-If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would sugges that the lexical route is in place for low-proficiency bilinguals.
+The role of this study was to investigate the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
+Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
+In this case, monolinguals can only use the lexical route to suceeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
+If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would suggest that the lexical route is in place for low-proficiency bilinguals.
 If participants are not able to translation words from L2 to L1 regardless of their form-similarity, this would suggest that low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
@@
 Participants in the Catalan list were presented with 86 trials, and participants in the Spanish list were presented with 103 trials.
-
-\begin{figure}
-\includegraphics[width=0.8\linewidth]{C:/Users/u155880/Documents/translation-elicitation/Figures/design} \caption{ }\label{fig:procedurefigure}
-\end{figure}
-
 Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
@@
 The task took approximately 15 minutes to be completed.
+
+\begin{figure}
+\includegraphics[width=0.8\linewidth]{C:/Users/u155880/Documents/translation-elicitation/Figures/design} \caption{Schematic representation of a trial in the experimental task.}\label{fig:procedurefigure}
+\end{figure}
 
@@
 
-Finally, we measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the \texttt{stringsim} function of the stringdist R package (van der Loo, 2014).
+We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the \texttt{stringsim} function of the stringdist R package (van der Loo, 2014).
 This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-collins1975spreading}{}}%
-Collins, A. M., \& Loftus, E. F. (1975). A spreading-activation theory of semantic processing. \emph{Psychological Review}, \emph{82}(6), 407.
-
 \leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%

### 2021-12-03 Rmd/manuscript.Rmd
Add manuscript plots with diagnostics

@@
 classoption: man
+appendix: "manuscript_appendix.Rmd"
 output:
@@
 
-Listening to speech in a non-native language is more costly than doing so in native languages, regardless of the level of proficiency [REF].
+Listening to speech in a non-native language is more costly than doing so in native languages, regardless of the level of proficiency [@lecumberri2010non].
 The source of this increased cognitive effort is likely to originate at multiple levels.
@@
 This suggests that bilinguals activate the picture's word representations in both languages, and the phonological overlap between both labels facilitated the naming process.
-
 Additionally, cognates seem to be learnt more easily in L2 than non-cognates.
@@
 As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
-The Revised Hierarchical Model [RHM; @kroll1994category] captured this assumption and predicted that translating words from L1 to L2 (forward translation) should take longer than translating from L2 to L1 (backward translation).
+The Revised Hierarchical Model [RHM,   @kroll1994category] captured this assumption and predicted that translating words from L1 to L2 (forward translation) should take longer than translating from L2 to L1 (backward translation).
 The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, while forward translation would rely more strongly on the mediation between the concept and the two word forms.
@@
 Since @kroll1994category proposed their RHM model, subsequent studies brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-The Neighbourhood Activation Model [NAM; @luce1998recognizing] suggests that lexical selection is mediated not only by the lexical frequency of the target word (and the number of activated candidates), but also by the frequency of the candidates.
+The Neighbourhood Activation Model [NAM, @luce1998recognizing] suggests that lexical selection is mediated not only by the lexical frequency of the target word (and the number of activated candidates), but also by the frequency of the candidates.
 

### 2021-12-03 Rmd/manuscript_appendix.Rmd
Add manuscript plots with diagnostics

@@
 
-```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
+```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+tar_load_globals()
+tar_load(stimuli)
+tar_load(participants)
+tar_load(responses)
+tar_load(model_fits)
+
+bayesplot_theme_set(
+  new = theme_custom() + 
+    theme(
+      axis.text.x = element_text(size = 8), 
+      axis.text.y = element_text(size = 8),
+      strip.text = element_text(size = 10),
+      legend.key = element_rect(colour = NA, fill = "white")
+    )
+)
+
+```
+
+# Appendix 1: Stan code
+
+```{r appendix-model-code, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+model_fits$fit_3$model
+color_scheme_set(c("#1A85FF", "#9ccaff", "#1A85FF",  "#1A85FF", "#1A85FF", "#1A85FF"))
+
+```
+
+# Appendix 2: MCMC diagnostics
+
+## Pairs plot
+
+```{r appendix-mcmc-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+posterior_cp <- as.array(model_fits$fit_3)
+
+plot_pairs <- mcmc_pairs(
+  posterior_cp, pars = c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv"),
+  off_diag_args = list(size = 0.75),
+  alpha = 1, size = 0
+)
+
+plot_pairs
+```
+
+## Trace plots
+
+```{r appendix-trace-plot}
+mcmc_trace(model_fits$fit_3, pars = c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv"), n_warmup = 0) +
+  annotate(geom = "rect", xmin = -Inf, xmax = 500, ymin = -Inf, ymax = Inf, alpha = 0.5, fill = "grey")
+
+```
+
+## Gelman-Rubin convergence criterion (R-hat)
+
+```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+rhats <- rhat(model_fits$fit_3)
+ratios <- neff_ratio(model_fits$fit_3)
+
+mcmc_rhat(rhats) + 
+  mcmc_neff(ratios) +
+  plot_layout(guides = "keep", ncol = 1) &
+  theme(
+    legend.position = "right"
+  )
+
+```
+
+# Appendix 3: by item accuracy
+
+```{r appendix-accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
 Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.

### 2021-12-03 _targets_packages.R

@@ -1,5 +1,6 @@
 # Generated by targets::tar_renv(): do not edit by hand
 library(audio)
+library(bayesplot)
 library(brms)
 library(bs4Dash)
 library(clustermq)
@@ -32,5 +33,6 @@ library(stringr)
 library(tibble)
 library(tidybayes)
 library(tidyr)
+library(tidytext)
 library(visNetwork)
 library(wesanderson)

### 2021-12-04 Rmd/manuscript.Rmd
knit to Word

@@
 classoption: man
-appendix: "manuscript_appendix.Rmd"
-output:
-  word_document:
-    reference_docx: "manuscript_template.docx"
+output: papaja::apa6_docx
 ---
@@
 
-options(knitr.kable.NA = '-', knitr.duplicate.label = "allow")
+options(
+  knitr.kable.NA = '-', 
+  knitr.duplicate.label = "allow",
+  gt.html_tag_check = FALSE
+  )
 
@@
 Model 4 can be formally expressed as:
+
+$$
+\begin{align*}
+
+&\textbf{Likelihood}  \\
+y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
+
+&\textbf{Parameters}  \\
+
+logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
+& \beta_{1[p]} ~ Frequency_{i} ~ + \\
+& \beta_{2[p]} ~ PTHN_i ~ + \\
+& \beta_{3[p]} ~ Similarity_i ~ + \\
+& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
+
+\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
+\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
+&& \text{[participant-level coefficients]} \\ \\
+
+&\textbf{Prior}  \\
+
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 3) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ \text{half}- Cauchy(0, 3) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(5) && \text{[correlation between participant-level coefficients]} \\
+
+\end{align*}
+$$
 
@@
 
-# Appendix
+# Appendix 1: MCMC diagnostics
 
-```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
-Words presented to participants in the English\\-Spanish or in the English\\-Catalan and Spanish\\-Catalan are listed in the Y\\-axis, ordered from higher to lower average translation accuracy, depicted in the X\\-axis.
-Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word.
-Accuracy is plotted separately for each group.", fig.height=10, fig.width=9}
+```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+suppressWarnings({
+  bayesplot_theme_set(
+    new = theme_custom() + 
+      theme(
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8),
+        strip.text = element_text(size = 10),
+        legend.key = element_rect(colour = NA, fill = "white")
+      )
+  )
+})
+
+color_scheme_set("brightblue")
+pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")
+
+```
+
+## Pairs plot
+
+```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+posterior_cp <- as.array(model_fits$fit_3)
+np_cp <- nuts_params(model_fits$fit_3)
+
+plot_pairs <- mcmc_pairs(
+  posterior_cp, pars = pars,
+  off_diag_args = list(size = 0.75)
+)
+
+plot_pairs
+```
+
+## Trace and Rank plots
+
+```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+(mcmc_trace(model_fits$fit_3, pars) + legend_none()) /
+  mcmc_rank_overlay(model_fits$fit_3, pars = pars) +
+  plot_layout(guides = "collect") &
+  theme(
+    legend.position = "top"
+  )
+```
+
+## Gelman-Rubin convergence criterion (R-hat)
+
+```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+rhats <- rhat(model_fits$fit_3)
+ratios <- neff_ratio(model_fits$fit_3)
+
+mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
+  mcmc_neff_hist(ratios, binwidth = 0.1) +
+  plot_layout(guides = "keep", ncol = 1) &
+  theme(
+    legend.position = "top"
+  )
+```
+
+# Appendix 2: by item accuracy
+
+```{r appendix-accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
+Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.
+Dots and whiskers represent the average accuracy and 95% confidence interval of each word.
+Accuracy is plotted separately for each group."}
 
@@
     legend.position = "none",
-    axis.text.y = element_text(size = 5, colour = "black", hjust = 1),
+    axis.text.y = element_text(size = 5, colour = "black"),
     axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 7, colour = "black"),
+    axis.text.x = element_text(size = 9, colour = "black"),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
   )

### 2021-12-04 _targets.R

@@ -93,11 +93,10 @@ list(
     
     # render report.Rmd
     # tar_render(report, "Rmd/report.Rmd"),
-
+    
     # render manuscript.Rmd
     tar_render(manuscript, "Rmd/manuscript.Rmd")
-    
-    )
+)
 
 
 

### 2021-12-09 Rmd/manuscript.Rmd
Add paragraph

@@
 
-Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech [@de1982relationship] or the loss of segmental information [@warren1970perceptual].
-But even in the most ideal of the situations, the cognitive processes engaged in word recognition do not occur without uncertainty.
-The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
-There is extensive literature about the influence on word recognition of the number of lexical candidates, their frequency, and their  phonological and semantic similarity with target words.
-Less is known about how all these factors affect the dynamics of word recognition in non-native languages.
+Even for highly proficient bilinguals, listening to non-native speech is more costly than listening to native speech [@lecumberri2010non].
+One of the sources of this increased difficulty stems from the potential mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal may not overlap perfectly with any phonemic category in the listener's native language.
+For example, a Spanish native listening to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
+The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
+This mismatch at the phonological level increases the uncertainty under which lexical selection takes place during word recognition.
+Yet, non-native listeners--even those with low proficiency--are rarely completely naïve to the language they are listening to.
+All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness, and these similarities can be exploited by non-native listeners to process the speech signal.
+In the present research we investigate the extent to which adult listeners rely on this similarity during lexical access, and we do so capitalising on the role of *cognates*.
 
-Listening to speech in a non-native language is more costly than doing so in native languages, regardless of the level of proficiency [@lecumberri2010non].
-The source of this increased cognitive effort is likely to originate at multiple levels.
-For example, some sounds in the speech signal do not correspond to any phoneme in the native language, making the process of matching sounds to phonemes complicated.
-This is the case of a Spanish native listening to words containing the guttural /ʁ/ in French.
-Language learners, however, are rarely completely naïve to the language they are listening to.
-All languages share, to some extent, similarities, frequently due to their typological closeness.
-These similarities which can be exploited by non-native listeners to process the speech signal.
-In the present research we investigate the extent to which adult listeners rely on phonological similarity between cross-language synonyms during word translation.
-
-Cognates are cross-language synonyms whose form (e.g., phonology, orthography, signature) is similar.
-The cause of such similarity is frequently attributed to a shared etymological origin.
+Cognates are cross-language synonyms whose form (e.g., phonology, orthography, sign) is similar.
+The source of such similarity is frequently attributed to a shared etymological origin.
 Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively)^[Some form-similar cross-language synonyms are technically not cognates.
@@
 It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
+Cognates play a pivotal role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+For example, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan) [@costa2000cognate].
+This cognate advantage is not restricted to production, but rather extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does], and word translation [@christoffels2006memory].
 
-Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-During word production and comprehension, lexical representations of both languages are activated, and the form similarity between both impacts bilinguals' performance in naming or lexical decision tasks [e.g., @costa2000cognate; @thierry2007brain].
-In a seminal study, @costa2000cognate asked Spanish-Catalan adult bilinguals to name pictures in Spanish.
-Unbeknownst to participants, half of the labels associated to pictures were cognates in Spanish and Catalan (e.g., *puerta*-*porta*) while the other half were non-cognates (e.g., *mesa*-*taula*, meaning *table* in Spanish and Catalan).
-Surprisingly, Spanish-Catalan bilinguals named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
-This suggests that bilinguals activate the picture's word representations in both languages, and the phonological overlap between both labels facilitated the naming process.
-Additionally, cognates seem to be learnt more easily in L2 than non-cognates.
-@de2000hard presented Dutch natives pairs of Dutch words and pseudowords.
-Participants' performance learnt and retained better pairs with high orthographic similarity than pairs with low orthographic similarity [see also @lotto1998effects for Dutch natives learning Italian words].
-
-Cognates are not only learnt faster than non-cognates, but translated faster too.
-Early accounts of bilingual lexical access suggest that low-proficiency learners first established links between newly acquired words in L2 and their meanings through their L1 translation equivalents [@potter1984lexical].
+Early theoretical accounts of this phenomenon pointed that the impact of cognateness on lexical processing is dependent on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for converging evidence].
 As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
-The Revised Hierarchical Model [RHM,   @kroll1994category] captured this assumption and predicted that translating words from L1 to L2 (forward translation) should take longer than translating from L2 to L1 (backward translation).
-The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, while forward translation would rely more strongly on the mediation between the concept and the two word forms.
+The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
+The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
 One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
@@
 Participants were presented with written words in English or Dutch, and asked to speak out loud their translation in the other language as fast as possible.
-Results who that translation times and accuracy were roughly equivalent in across both forward and backward conditions, and that semantic variables (e.g., concreteness), were positively associated with participants' performance, especially in the forward translation condition.
+Results showed that translation times and accuracy were roughly equivalent  in forward and backward translation, and that semantic variables (e.g., concreteness) were positively associated with participants' performance, especially in the forward translation condition.
 Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected  by semantic variables than non-cognates.
-
-These findings suggest that both conceptually mediated and direct translation routes are active during translation, although backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness (than forward translation).
+These findings suggest that both conceptually-mediated and direct translation routes are active during translation, although backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness (than forward translation).
 Interestingly, higher-proficiency bilinguals tested in the same task showed similar results.
 However, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language], contrary to the predictions of the RHM.
-Further models build upon this model, introducing a wider range of connections in their theorised lexica.
+Further models built upon this model, introducing a wider range of connections in their theorised lexica.
 
-Since @kroll1994category proposed their RHM model, subsequent studies brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-The Neighbourhood Activation Model [NAM, @luce1998recognizing] suggests that lexical selection is mediated not only by the lexical frequency of the target word (and the number of activated candidates), but also by the frequency of the candidates.
-
-Subsequent behavioural experiments [e.g., @luce1998recognizing] show that English adult participants were tested in a lexical decision task in which words were presented across several conditions of noise-to-signal ratio.
-A computerised lexicon calculated number and average lexical frequency of phonological neighbours around each of the presented words.
-Participants answered more faster and more accurately to high frequency words, and words from high density neighbourhoods were also responded detected more accurately, but slowly.
-Therefore, average lexical frequency of  neighbourhoods was associated to a decrease in accuracy.
-Low frequency words were in fact detected more accurately in low density than in high density neighbours.
-These findings suggested that both word lexical frequency and neighbourhood density can facilitate spoken word recognition, but this effect is modulated by the frequency of the neighbourhood; when phonological neighbours are more frequent, lexical selection is hindered [@goldinger1989priming; @luce1990similarity].
-
-The influence of phonological neighbourhoods and their structure on lexical processing poses rise important question in the field of bilingualism research.
-Do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation [BIA+; @van1998orthographic] proposes that lexical representations establish both excitatory and inhibitory connections with other representations from the same or the other language, forming an integrated lexicon for both languages [@mcclelland1981interactive].
+Since the RHM model was proposed, later studies brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted the fact that lexical selection is affected by the number of the phonological neighbours of the selected word.
+Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when such neighbours; lexical frequency is higher [@luce1998recognizing].
+This finding poses an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+The Bilingual Interactive Activation [BIA+, @van1998orthographic] proposes that lexical representations establish both excitatory and inhibitory connections with other representations from the same or the other language, forming an integrated lexicon for both languages [@mcclelland1981interactive].
 Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
@@
 Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
-In this case, monolinguals can only use the lexical route to suceeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
+In this case, monolinguals can only use the lexical route to succeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
 If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would suggest that the lexical route is in place for low-proficiency bilinguals.

### 2021-12-16 Rmd/appendix.Rmd
Fix error when joining stimuli info and participant responses

---

---

# Appendix 1: MCMC diagnostics

```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

suppressWarnings({
  bayesplot_theme_set(
    new = theme_custom() + 
      theme(
        axis.text.x = element_text(size = 8), 
        axis.text.y = element_text(size = 8),
        strip.text = element_text(size = 10),
        legend.key = element_rect(colour = NA, fill = "white")
      )
  )
})

color_scheme_set("brightblue")
pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")
pars_labels <- c(
  "Intercept",
  "Frequency (+1 SD)",
  "PTHN \u00d7 Levenshtein", 
  "Levenshtein (+1 SD)",
  "PTHN (+1 SD)"
)
```

## Pairs plot

```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
posterior_cp <- as.array(model_fits$fit_4)
np_cp <- nuts_params(model_fits$fit_4)

plot_pairs <- mcmc_pairs(
  posterior_cp, pars = pars,
  off_diag_args = list(size = 0.75),
  diag_args = list(facet_args = list(labeller = label_value(labels = pars_labels)))
)

plot_pairs
```

## Trace and Rank plots

```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
(mcmc_trace(model_fits$fit_4, pars) + legend_none()) /
  mcmc_rank_overlay(model_fits$fit_4, pars = pars) +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "top"
  )
```

## Gelman-Rubin convergence criterion (R-hat)

```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
rhats <- rhat(model_fits$fit_4)
ratios <- neff_ratio(model_fits$fit_4)

mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
  mcmc_neff_hist(ratios, binwidth = 0.1) +
  plot_layout(guides = "keep", ncol = 1) &
  theme(
    legend.position = "top"
  )
```

# Appendix 3: participant-level posterior predictions

```{r acc_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
# by participant
m_re_participant <- posterior_epreds_random$participants %>% 
  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
  mutate(
    pthn = case_when(
      pthn=="-1 SD" ~ "Low PTHN",
      pthn=="0 SD" ~ "Mean PTHN",
      pthn=="1 SD" ~ "High PTHN"
    ),
    frequency_zipf = case_when(
      frequency_zipf=="-1 SD" ~ "Low frequency",
      frequency_zipf=="0 SD" ~ "Mean frequency",
      frequency_zipf=="1 SD" ~ "High frequency"
    )
  ) %>% 
  group_by(participant, lv, pthn) %>% 
  summarise(.epred = mean(.epred, na.rm = TRUE), .groups = "drop") %>% 
  mutate(participant = as.numeric(participant))

ggplot(m_re_participant, aes(lv, .epred, colour = pthn, fill = pthn)) +
  facet_wrap(~participant) +
  geom_line(aes(group = interaction(pthn, participant)), size = 0.35) +
  geom_text(
    data = distinct(m_re_participant, participant), 
    aes(x = -1, y = 1, label = participant),
    size = 1.5, vjust = 1, hjust = 0, inherit.aes = FALSE, 
    position = position_nudge(x = 0.2),
    show.legend = FALSE
  ) +
  labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
       title = "Participant-level predictions") +
  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
  guides(colour = guide_legend(ncol = 2)) +
  theme(
    legend.position = "none",
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks = element_blank(),
    strip.text = element_blank(),
    strip.background = element_blank(),
    panel.spacing = unit(0.1, "lines")
  )

```

# Appendix 2: accuracy by item

## English-Spanish group

```{r appendix-accuracy-engspa, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.height = 7, fig.width = 4}

acc_engspa <- responses %>%
  filter(group == "ENG-SPA") %>% 
  group_by(word) %>%
  summarise(
    correct_sum = sum(correct, na.rm = TRUE),
    correct_n = sum(!is.na(correct), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  left_join(select(filter(stimuli, group=="ENG-SPA"), word = word1, word2, ipa1, ipa2)) %>% 
  drop_na(word2) %>% 
  rowwise() %>% 
  mutate(
    word = paste0(word, " /", ipa1, "/"),
    word2 = paste0(word2, " /", ipa2, "/"),
    correct_prop = prop_adj(correct_sum, correct_n),
    correct_se = prop_adj_se(correct_sum, correct_n),
    correct_ci_lower = correct_prop + qnorm((1-.95)/2)*correct_se,
    correct_ci_upper = correct_prop + qnorm(1-(1-.95)/2)*correct_se,
    correct_ci_lower = ifelse(correct_ci_lower < 0, 0, correct_ci_lower),
    correct_ci_upper = ifelse(correct_ci_upper > 1, 1, correct_ci_upper)
  ) %>% 
  ungroup() %>% 
  arrange(-correct_prop) %>% 
  select(word, word2, correct_prop, correct_se, correct_ci_lower, correct_ci_upper)

if ((nrow(acc_engspa) %% 2) > 0) {
  acc_engspa <- add_row(
    acc_engspa,
    data.frame(
      word = NA, word2 = NA, correct_prop = NA, correct_se = NA, 
      correct_ci_lower = NA, correct_ci_upper = NA
    )
  )
}

acc_engspa <- mutate(
  acc_engspa, 
  row = 1:nrow(acc_engspa), 
  col = ifelse(row <= nrow(acc_engspa)/2, 1, 2)
)

tab_engspa <- cbind(
  rename_all(acc_engspa[acc_engspa$col==1,], ~paste0(., "_1")), 
  rename_all(acc_engspa[acc_engspa$col==2,], ~paste0(., "_2"))
) %>% 
  select(-c(col_1, col_2, row_1, row_2)) %>% 
  as_tibble() %>% 
  gt() %>% 
  tab_header(
    title = "Table A.1",
    subtitle = "Proportion of correct translation by item in the English-Spanish group"
  ) %>% 
  tab_source_note("Words presented to participants in the English-Spanish group are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis. Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word. Accuracy is plotted separately for each group.") %>% 
  tab_spanner("(1)", ends_with("_1")) %>% 
  tab_spanner("(2)", ends_with("_2")) %>% 
  cols_merge(5:6, pattern = "[{1}, {2}]") %>% 
  cols_merge(11:12, pattern = "[{1}, {2}]") %>%

  fmt_percent(3:6, decimals = 0) %>% 
  fmt_percent(9:12, decimals = 0)  %>%

  cols_align("left") %>% 
  cols_label(
    word_1 = "Word",
    word2_1 = "Translation",
    correct_prop_1 = "Accuracy",
    correct_se_1 = md("*SE*"),
    correct_ci_lower_1 = md("*CI*<sub>95%</sub>"),
    word_2 = "Word",
    word2_2 = "Translation",
    correct_prop_2 = "Accuracy",
    correct_se_2 = md("*SE*"),
    correct_ci_lower_2 = md("*CI*<sub>95%</sub>")
  ) %>% 
  tab_style(
    style = cell_borders(sides = c("left", "right"), color = "white"),
    locations = list(
      cells_body(), cells_row_groups(),
      cells_stub(), 
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "small"), cell_borders(sides = "all", color = "white")),
    locations = cells_stub()
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "small", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = c(1:12))
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:6)
  ) %>% 
  tab_style(
    cell_text(weight = "bold"), cells_body(columns = c(1, 7))
  ) %>% 
  tab_style(
    cell_text(size = "small"), 
    cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_style(cell_text(size = "small"), cells_body(columns = everything(), rows = everything())) %>%   tab_options(data_row.padding = px(1)) %>% 
  gtsave(here("Figures", "accuracy_engspa.png"))

include_graphics(here("Figures", "accuracy_engspa.png"))
```

## English-Catalan group

```{r appendix-accuracy-engcat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.height = 7, fig.width = 4}

acc_engcat <- responses %>%
  filter(group == "ENG-CAT") %>% 
  group_by(word) %>%
  summarise(
    correct_sum = sum(correct, na.rm = TRUE),
    correct_n = sum(!is.na(correct), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  left_join(select(filter(stimuli, group=="ENG-CAT"), word = word1, word2, ipa1, ipa2)) %>% 
  drop_na(word2) %>% 
  rowwise() %>% 
  mutate(
    word = paste0(word, " /", ipa1, "/"),
    word2 = paste0(word2, " /", ipa2, "/"),
    correct_prop = prop_adj(correct_sum, correct_n),
    correct_se = prop_adj_se(correct_sum, correct_n),
    correct_ci_lower = correct_prop + qnorm((1-.95)/2)*correct_se,
    correct_ci_upper = correct_prop + qnorm(1-(1-.95)/2)*correct_se,
    correct_ci_lower = ifelse(correct_ci_lower < 0, 0, correct_ci_lower),
    correct_ci_upper = ifelse(correct_ci_upper > 1, 1, correct_ci_upper)
  ) %>% 
  ungroup() %>% 
  arrange(-correct_prop) %>% 
  select(word, word2, correct_prop, correct_se, correct_ci_lower, correct_ci_upper)

if ((nrow(acc_engcat) %% 2) > 0) {
  acc_engcat <- add_row(
    acc_engcat,
    data.frame(
      word = NA, word2 = NA, correct_prop = NA, correct_se = NA, 
      correct_ci_lower = NA, correct_ci_upper = NA
    )
  )
}

acc_engcat <- mutate(acc_engcat, row = 1:nrow(acc_engcat), col = ifelse(row <= nrow(acc_engcat)/2, 1, 2))

tab_engcat <- cbind(
  rename_all(acc_engcat[acc_engcat$col==1,], ~paste0(., "_1")), 
  rename_all(acc_engcat[acc_engcat$col==2,], ~paste0(., "_2"))
) %>% 
  select(-c(col_1, col_2, row_1, row_2)) %>% 
  as_tibble() %>% 
  gt() %>% 
  tab_header(
    title = "Table A.2",
    subtitle = "Proportion of correct translation by item in the English-Catalan group"
  ) %>% 
  tab_source_note("Words presented to participants in the English-Catalan group are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis. Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word. Accuracy is plotted separately for each group.") %>% 
  tab_spanner("(1)", ends_with("_1")) %>% 
  tab_spanner("(2)", ends_with("_2")) %>% 
  cols_merge(5:6, pattern = "[{1}, {2}]") %>% 
  cols_merge(11:12, pattern = "[{1}, {2}]") %>%

  fmt_percent(3:6, decimals = 0) %>% 
  fmt_percent(9:12, decimals = 0)  %>%

  cols_align("left") %>% 
  cols_label(
    word_1 = "Word",
    word2_1 = "Translation",
    correct_prop_1 = "Accuracy",
    correct_se_1 = md("*SE*"),
    correct_ci_lower_1 = md("*CI*<sub>95%</sub>"),
    word_2 = "Word",
    word2_2 = "Translation",
    correct_prop_2 = "Accuracy",
    correct_se_2 = md("*SE*"),
    correct_ci_lower_2 = md("*CI*<sub>95%</sub>")
  ) %>% 
  tab_style(
    style = cell_borders(sides = c("left", "right"), color = "white"),
    locations = list(
      cells_body(), cells_row_groups(),
      cells_stub(), 
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "small"), cell_borders(sides = "all", color = "white")),
    locations = cells_stub()
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "small", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = c(1:12))
  ) %>% 
  tab_style(
    cell_text(weight = "bold"), cells_body(columns = c(1, 7))
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:6)
  ) %>% 
  tab_style(
    cell_text(size = "small"), 
    cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_style(cell_text(size = "small"), cells_body(columns = everything(), rows = everything())) %>%   tab_options(data_row.padding = px(1)) %>% 
  gtsave(here("Figures", "accuracy_engcat.png"))

include_graphics(here("Figures", "accuracy_engcat.png"))

```

## Spanish-Catalan group

```{r appendix-accuracy-spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.height = 7, fig.width = 4}

acc_spacat <- responses %>%
  filter(group == "SPA-CAT") %>% 
  group_by(word) %>%
  summarise(
    correct_sum = sum(correct, na.rm = TRUE),
    correct_n = sum(!is.na(correct), na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  left_join(select(filter(stimuli, group=="SPA-CAT"), word = word1, word2, ipa1, ipa2)) %>% 
  drop_na(word2) %>% 
  rowwise() %>% 
  mutate(
    word = paste0(word, " /", ipa1, "/"),
    word2 = paste0(word2, " /", ipa2, "/"),
    correct_prop = prop_adj(correct_sum, correct_n),
    correct_se = prop_adj_se(correct_sum, correct_n),
    correct_ci_lower = correct_prop + qnorm((1-.95)/2)*correct_se,
    correct_ci_upper = correct_prop + qnorm(1-(1-.95)/2)*correct_se,
    correct_ci_lower = ifelse(correct_ci_lower < 0, 0, correct_ci_lower),
    correct_ci_upper = ifelse(correct_ci_upper > 1, 1, correct_ci_upper)
  ) %>% 
  ungroup() %>% 
  arrange(-correct_prop) %>% 
  select(word, word2, correct_prop, correct_se, correct_ci_lower, correct_ci_upper)

if ((nrow(acc_spacat) %% 2) > 0) {
  acc_spacat <- add_row(
    acc_spacat,
    data.frame(
      word = NA, word2 = NA, correct_prop = NA, correct_se = NA, 
      correct_ci_lower = NA, correct_ci_upper = NA
    )
  )
}

acc_spacat <- mutate(acc_spacat, row = 1:nrow(acc_spacat), col = ifelse(row <= nrow(acc_spacat)/2, 1, 2))

tab_spacat <- cbind(
  rename_all(acc_spacat[acc_spacat$col==1,], ~paste0(., "_1")), 
  rename_all(acc_spacat[acc_spacat$col==2,], ~paste0(., "_2"))
) %>% 
  select(-c(col_1, col_2, row_1, row_2)) %>% 
  as_tibble() %>% 
  gt() %>% 
  tab_header(
    title = "Table A.3",
    subtitle = "Proportion of correct translation by item in the Spanish-Catalan group"
  ) %>% 
  tab_source_note("Words presented to participants in the Spanish-Catalan group are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis. Dots and whiskers represent the average accuracy and 95\\% confidence interval of each word. Accuracy is plotted separately for each group.") %>% 
  tab_spanner("(1)", ends_with("_1")) %>% 
  tab_spanner("(2)", ends_with("_2")) %>% 
  cols_merge(5:6, pattern = "[{1}, {2}]") %>% 
  cols_merge(11:12, pattern = "[{1}, {2}]") %>%

  fmt_percent(3:6, decimals = 0) %>% 
  fmt_percent(9:12, decimals = 0)  %>%

  cols_align("left") %>% 
  cols_label(
    word_1 = "Word",
    word2_1 = "Translation",
    correct_prop_1 = "Accuracy",
    correct_se_1 = md("*SE*"),
    correct_ci_lower_1 = md("*CI*<sub>95%</sub>"),
    word_2 = "Word",
    word2_2 = "Translation",
    correct_prop_2 = "Accuracy",
    correct_se_2 = md("*SE*"),
    correct_ci_lower_2 = md("*CI*<sub>95%</sub>")
  ) %>% 
  tab_style(
    style = cell_borders(sides = c("left", "right"), color = "white"),
    locations = list(
      cells_body(), cells_row_groups(),
      cells_stub(), 
      cells_column_labels(), cells_stubhead()
    )
  ) %>% 
  tab_style(
    style = list(cell_text(align = "left", size = "small"), cell_borders(sides = "all", color = "white")),
    locations = cells_stub()
  ) %>% 
  tab_style(
    style = cell_text(align = "center", size = "small", weight = "normal", style = "normal"),
    locations = list(cells_column_spanners(), cells_column_labels())
  ) %>% 
  tab_style(
    style = cell_borders(sides = "all", color = "white"),
    locations = cells_body(columns = c(1:12))
  ) %>% 
  tab_style(
    style = cell_borders(sides = "top", color = "white"),
    locations = cells_body(columns = 1:6)
  ) %>% 
  tab_style(
    cell_text(weight = "bold"), cells_body(columns = c(1, 7))
  ) %>% 
  tab_style(
    cell_text(size = "small"), 
    cells_body(columns = everything(), rows = everything())
  ) %>%
  tab_options(data_row.padding = px(1)) %>% 
  gtsave(here("Figures", "accuracy_spacat.png"))

include_graphics(here("Figures", "accuracy_spacat.png"))

```

### 2021-12-16 Rmd/manuscript.Rmd
Fix error when joining stimuli info and participant responses

@@
   email: gonzalo.garciadecastro@upf.edu
-  role: 
-    - Conceptualization
-    - Data analysis
-    - Writing - Original Draft Preparation
-    - Writing - Review & Editing
 - name: Serene Siow
   affiliation: '2'
-  role:
-    - Conceptualization
-    - Writing - Review & Editing
 - name: Nuria Sebastian-Galles
   affiliation: '1'
-  role:
-    - Conceptualization
-    - Writing - Review & Editing
 - name: Kim Plunkett
   affiliation: '2'
-  role:
-    - Conceptualization
-    - Writing - Review & Editing
 affiliation:
@@
 authornote: |
-  GG and SS contributed equally and share first authorship.
 
 abstract: |
-  There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
-For example, bilinguals produce and translate cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
-Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
-These measures partially ignore potential sources of variability during lexical access like vowel vs.
-consonant overlap, overlap in stressed syllables, onset overlap or the distance in features between both translations.
-In this study, we explored the impact of some of these variables on non-native word translation task: Spanish and English participants listened to non-native words (Catalan or Spanish) and were prompted to type their translation in their native language.
-Critically, participants where unfamiliar with the testing language, ensuring that they were not able to translate words based on previous knowledge on their meaning, leaving phonological information as the only cue participants were able to exploit to translate words to their native language correctly.
-We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.
+  When listening to non-native speech, words whose translation in the native language is phonologically similar are easier to recognise.
+Previous studies suggest that low-proficiency bilinguals benefit more from this facilitation effect than high-proficiency bilinguals, but this claim has remained relatively unaddressed.
+In this study, we tested English and Spanish monolinguals in a translation elicitation task, in which they had to guess the translation of a list of words in an unfamiliar language (Spanish or Catalan).
+Presented words varied in their phonological similarity to their correct translation.
+This allowed us to test the impact of such phonological similarity on word recognition in a sample of listeners with very low proficiency in a non-native language.
+Translation accuracy increased substantially the likelihood of correct translations, but this effect was mediated by the size of the phonological neighbourhood of the correct translation.
+Participants benefitted from phonological similarity only when the correct translation was part of a sparse phonological neighbourhood.
+When the correct translation was part of a larger phonological neighbourhood, translation accuracy remained unaffected.
+Overall, these results provide evidence that litening to non-native words activated lexical representations in the native language, even in monolingual participants, resulting in competition for selection in the native language during translation.
+
   <!-- https://tinyurl.com/ybremelq -->
 
-keywords: cognate, word recognition, translation, non-native, spoken word recognition
-wordcount: X
+keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian
+wordcount: 4553
 
@@
 draft: no
+numbersections: yes
 csl: "apa7.csl"
@@
   results = "asis", 
-  dpi = 500
+  dpi = 1000
 )
@@
 
-Even for highly proficient bilinguals, listening to non-native speech is more costly than listening to native speech [@lecumberri2010non].
-One of the sources of this increased difficulty stems from the potential mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal may not overlap perfectly with any phonemic category in the listener's native language.
+Listening to non-native speech is more costly than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
+One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
 For example, a Spanish native listening to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
 The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
-This mismatch at the phonological level increases the uncertainty under which lexical selection takes place during word recognition.
-Yet, non-native listeners--even those with low proficiency--are rarely completely naïve to the language they are listening to.
-All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness, and these similarities can be exploited by non-native listeners to process the speech signal.
-In the present research we investigate the extent to which adult listeners rely on this similarity during lexical access, and we do so capitalising on the role of *cognates*.
+These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical].
+Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
+In the present study we investigate the extent to which adult listeners rely on the phonological similarity between the native and the non-native languages during word recognition, capitalising on the role of *cognates*.
 
-Cognates are cross-language synonyms whose form (e.g., phonology, orthography, sign) is similar.
-The source of such similarity is frequently attributed to a shared etymological origin.
-Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively)^[Some form-similar cross-language synonyms are technically not cognates.
+All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the populations of speakers of each language.
+Cognates embody a particular case of such commonalities.
+Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
 For example, *sun* and *sol* (in Spanish), share their phonological onset, but their etymology points to different origins.
@@
 It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
+For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
 Cognates play a pivotal role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-For example, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan) [@costa2000cognate].
+For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
 This cognate advantage is not restricted to production, but rather extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does], and word translation [@christoffels2006memory].
 
-Early theoretical accounts of this phenomenon pointed that the impact of cognateness on lexical processing is dependent on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for converging evidence].
-As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
+Early theoretical accounts of this phenomenon pointed to the fact that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), and not to their shared concept (conceptual route).
+This word-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection).
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations, therefore to cognateness.
 The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
-The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
+The rationale behind this prediction was that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
 One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
 
-To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
-Participants were presented with written words in English or Dutch, and asked to speak out loud their translation in the other language as fast as possible.
-Results showed that translation times and accuracy were roughly equivalent  in forward and backward translation, and that semantic variables (e.g., concreteness) were positively associated with participants' performance, especially in the forward translation condition.
+To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
+Results showed that translation times and accuracy were roughly equivalent in forward and backward translation.
 Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected  by semantic variables than non-cognates.
-These findings suggest that both conceptually-mediated and direct translation routes are active during translation, although backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness (than forward translation).
-Interestingly, higher-proficiency bilinguals tested in the same task showed similar results.
-However, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language], contrary to the predictions of the RHM.
-Further models built upon this model, introducing a wider range of connections in their theorised lexica.
+Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
+Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
+Although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
+These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
 
-Since the RHM model was proposed, later studies brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted the fact that lexical selection is affected by the number of the phonological neighbours of the selected word.
-Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when such neighbours; lexical frequency is higher [@luce1998recognizing].
-This finding poses an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation [BIA+, @van1998orthographic] proposes that lexical representations establish both excitatory and inhibitory connections with other representations from the same or the other language, forming an integrated lexicon for both languages [@mcclelland1981interactive].
+Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted the fact that lexical selection is affected by the number of phonological neighbours of the selected word.
+Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when such neighbours' lexical frequency is higher [@luce1998recognizing].
+This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+The Bilingual Interactive Activation [BIA+, @van1998orthographic] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
 Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
-@van1998orthographic presented bilingual participants with words in only one language (target language) in a lexical decision task.
-When the translation of the presented word in the non-target language was part of a large neighbourhood, reaction times were slower.
-This provides evidence that the presented words activated orthographic neighbourhoods in the non-target language, which slowed down word recognition.
+When bilingual participants are presented with words in only one language (target language) in a lexical decision task, translating them to the non-target language took longer when they were part of large  neighbourhoods [@van1998orthographic].
+This suggests that words presented in the target language activated orthographic neighbourhoods in the non-target language, which competed for selection with the target word during recognition.
 
-A more recent model of bilingual lexical processing is Multilink @dijkstra2019multilink, which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
+A more recent model of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
 This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
-Such model also assumes that during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1 and it proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model (which suggests that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
+Such model also assumes that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
+It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
 In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
-A remarkable correlation between Multilink simulations of @christoffels2006memory translation elicitation task and behavioural data with Dutch-English bilinguals has been reported, showing forward translation being faster than backward translation, contrary to the predictions made by the RHM model.
+A remarkable correlation between Multilink simulations of a translation elicitation task and behavioural data on Dutch-English bilinguals [@christoffels2006memory] has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
 Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
-Overall, results from Multilink from this model are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
+Overall, results from Multilink are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
 
-The fact that Multilink fitted the experimental data from @christoffels2006memory with remarkable success, even in absence of word-word connections between translation equivalents might put into question the necessity of such connections during translation, in favour of a simpler model in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
-This claim, however, should be taken.
-It should be taken into account, however, that @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
-Additionally, data generated from Multilink model assumed similar vocabulary sizes across languages and, perhaps more critically in this regard, that word-concept connections were equally strong for L1 and L2 representations.
-These conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation.
-This prediction remains untested my Multilink.
+The fact that Multilink successfully fitted the experimental data from @christoffels2006memory, despite the absence of any word-word connections between translation equivalents, is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated, assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
+The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
+It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation, and this prediction remains untested my Multilink.
 
-The role of this study was to investigate the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
+In the present study we investigated the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
 Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
 In this case, monolinguals can only use the lexical route to succeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
-If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would suggest that the lexical route is in place for low-proficiency bilinguals.
-If participants are not able to translation words from L2 to L1 regardless of their form-similarity, this would suggest that low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
+We hypothesised that if participants are able to translate words in L2, without knowledge of the language it belongs to, and based on their form-similarity with the target word, then the lexical route should be in place for low-proficiency bilinguals.
+In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then should low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
+
+Following @luce1998recognizing we also hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+If this is the case, words whose translation in the native language are part of larger phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target word.
 
 # Methods
+
+All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/) and a GitHub repository [https://github.com/bilingual-project/translation-elicitation.git](https://github.com/bilingual-project/translation-elicitation.git), along with additional notes.
 
@@
 lengths <- stimuli %>% 
-  select(trial_id, group, ipa1, ipa2, word1, word2) %>% 
+  select(group, ipa1, ipa2, word1, word2) %>% 
   mutate_at(vars(ipa1:word2), nchar) %>% 
@@
 
-We retrieved PTHN scores from the CLEARPOND database [@marian2012clearpond].
+We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) from the CLEARPOND database [@marian2012clearpond].
 PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
@@
 
+```{r stimuli-correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=4, fig.cap="Marginal distribution and pairwise correlation of lexical frequency, PTHN, and Levenshtein similarity.
+Statistics are indicated for the three experimental conditions (ENG-SPA, ENG-CAT, and SPA-CAT) with different colours.
+The upper triangle of the plot matrix shows the Pearson correlation between each pair of variables (asterisks indicate statistical significance at .05, .01, and .001 significance thresholds).
+The diagonal of the plot matrix indicates the marginal distribution of eacch variable.
+The lower triangle of the plot matrix representes graphically the association between each pair of variables, together with a linear regression line."}
+my_bin <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_density(..., size = 1, alpha = 0.25) +
+    geom_rug(...)
+}
+
+my_smooth <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_point(..., shape = 1, stroke = 1, alpha = 0.25, size = 0.5) +
+    geom_smooth(..., method = "lm", formula = y ~ x, alpha = 0.25, size = 1)
+}
+
+my_cor <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    ggally_cor(..., size = 7, title = "r", )
+}
+
+stimuli %>% 
+  drop_na(frequency_zipf, pthn, lv) %>% 
+  ggpairs(
+    mapping = aes(colour = group, fill = group),
+    columns = c("frequency_zipf", "pthn", "lv"),
+    columnLabels = c("Frequency (Zipf)", "PTHN", "Levenshtein"),
+    lower = list(continuous = my_smooth),
+    diag = list(continuous = my_bin),
+    upper = list("my_cor")
+  ) +
+  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
+  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE)
+
+```
+
 Participants listened to one audio file in each trial.
@@
 The average duration of the audios was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
-The average duration of the Catalan audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
+The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
 
-```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Stimuli data.
-Mean and standard deviation (SD) of lexical frequencies od target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words."}
+```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-stimuli_table <- stimuli %>% 
-  select(group, trial_id, frequency, frequency_zipf, pthn, lv) %>% 
+tb_stimuli <- stimuli %>% 
+  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
   group_by(group) %>% 
@@
   gt(rowname_col = "group") %>% 
-  fmt_number(frequency_mean:lv_sd) %>% 
-  tab_spanner(md("**Freq./million**"), c("frequency_mean", "frequency_sd")) %>% 
+  tab_header(
+    title = md("Table 1"),
+    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
+  ) %>% 
+  fmt_number(frequency_mean:pthn_sd) %>%
+  fmt_percent(matches("lv")) %>% 
+  tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
   tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
@@
   ) %>% 
-  as_latex()
+  summary_rows(
+    columns = c(matches("mean"), -matches("lv")),
+    fns = list(
+      Mean = "mean",
+      SD = "sd"
+    )
+  ) %>% 
+  summary_rows(
+    columns = matches("lv"),
+    fns = list(
+      Mean = "mean",
+      SD = "sd"
+    ), formatter = fmt_percent
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium", style = "italic"),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_stub_grand_summary()
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = c("left", "right"), color = "white"),
+    locations = list(
+      cells_body(), cells_row_groups(), cells_column_spanners(), 
+      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
+      cells_column_labels(), cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
+    locations = cells_stub()
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+    locations = list(cells_column_spanners(), cells_column_labels())
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "all", color = "white"),
+    locations = cells_body(columns = 1:9)
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "top", color = "white"),
+    locations = cells_body(columns = 1:9)
+  ) %>% 
+  gtsave(here("Figures", "stimuli.png"))
 
-stimuli_table
+include_graphics(here("Figures", "stimuli.png"), auto_pdf = FALSE)
 
@@
 We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We first fitted a base model (Model 0) that only included the lexical frequency (`frequency`) of the target word as a fixed effect, with a random intercept per participant.
-Second, we extended the model to include `pthn` as a fixed effect, with a random slope by participant (Model 1).
-Third, we added the fixed effect `consonant_ratio` and the `pthn:consonant_ratio`, and random slopes for both effects by participant (Model 2).
-Third, we added the fixed effect `vowel_ratio` and the `pthn:vowel_ratio`, and random slopes for both effects by participant (Model 3).
-Finally, we fit a model that included the two-way interactions `pthn:consonant_ratio` and `pthn:vowel_ratio`, an their random slopes by participant (Model 4).
-All predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
-Model 4 can be formally expressed as:
+We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0).
+Second, we added the fixed effect of lexical frequency of the target word ($Frequency$).
+We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the $Levenshtein \times Group$ interaction).
+We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.5) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize].
+Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random].
+We will later refer to these models as Model 0 to Model 6.
+For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
 
 $$
-\begin{align*}
+\begin{align}
 
@@
 
-\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 3) && \text{[participant-level coefficients]} \\
-\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ \text{half}- Cauchy(0, 3) && \text{[SD for population and participant]} \\
-\rho_{p}, ~ \rho_{w} \sim& ~LKJ(5) && \text{[correlation between participant-level coefficients]} \\
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
 
-\end{align*}
+\end{align}
 $$
@@
 All analyses were performed in R environment [@rcore2019r].
-We used the tidyverse family of R packages to process data and to generate figures, and the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
+We used the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
 
@@
 All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
-Model 4, which included all main effects and the two-way interactions between PTHN and vowel similarity, and PTHN and consonant similarity, showed the best performance (see Table 2).
+Model 6, which included all main effects and the two-way interactions between $PTHN$ and $Levenshtein$, and between $Group$ and $Levenshtein$ and $PTHN$ and consonant similarity, showed the best performance.
 
-```{r modelvalues, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
 
-coefs <- fixef(model_fits$fit_3) %>% 
+tab_loos <- model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
+  arrange(desc(elpd_loo)) %>% 
+  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
+  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  tab_header(
+    title = md("Table 2"),
+    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Outcomes of the leave-one-out cross-validation procedure.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
+  ) %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  fmt_number(2:7) %>% 
+  cols_label(
+    model = md("**Model**"),
+    looic = md("***LOO<sub>IC</sub>***"),
+    se_looic = md("***SE*<sub>IC</sub>**"),
+    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("***LOO<sub>diff</sub>***"),
+    se_diff = md("***SE<sub>diff</sub>***")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium", style = "italic"),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = list(cell_borders(sides = c("left", "right"), color = "white"),
+                 cell_text(align = "center")),
+    locations = list(
+      cells_body(), cells_row_groups(), cells_column_spanners(), 
+      cells_stub(), cells_source_notes(),
+      cells_column_labels(), cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(cell_text(align = "left", size = "medium"), 
+                 cell_borders(sides = "all", color = "white")),
+    locations = list(cells_stub(), cells_source_notes())
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+    locations = list(cells_column_spanners(), cells_column_labels())
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "all", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "top", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  gtsave(here("Figures", "loos.png"))
+
+include_graphics(here("Figures", "loos.png"))
+
+```
+
+We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
+For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
+Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
+This way, the coefficients are expressed as increments or decrements in probability of correct translation [@gelman2020regression].
+
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
+A\\) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Black dots represent the mean of each distribution.
+B\\) Participant\\-level coefficient variability.
+The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
+C\\) Correlation between participant\\-level effects.
+The model allowed participant-level coefficients to co\\-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}
+
+coefs <- fixef(model_fits$fit_6) %>% 
   as.data.frame() %>% 
@@
 
-```
-
-```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=4, fig.cap="Model comparison.
-Outcomes of the leave-one-out cross-validation procedure.
-Each row indicates the values of the expected log-predicted density (ELPD), leave-one-out information creterion (LOO-IC), and the difference in ELPD (LOO-diff) between each model and the inmediately simpler model.
-Each value is accompanied by a standard error (SE) indicating the uncertainty of its estimate."}
-
-loos_table <- model_loos %>% 
-  as.data.frame() %>% 
-  rownames_to_column("model") %>% 
-  arrange(desc(elpd_loo)) %>% 
-  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
-  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
-  mutate_all(~ifelse(. == 0, NA, .)) %>% 
-  gt(rowname_col = "model") %>% 
-  fmt_missing(everything(), missing_text = "-") %>% 
-  fmt_number(2:7) %>% 
-  cols_label(
-    model = md("**Model**"),
-    looic = md("***LOO_{IC}***"),
-    se_looic = md("***SE*<sub>IC</sub>**"),
-    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
-    se_elpd_loo = md("***SE***"),
-    elpd_diff = md("***LOO<sub>diff</sub>***"),
-    se_diff = md("***SE<sub>diff</sub>***")
-  ) %>% 
-  as_latex()
-
-loos_table
-
-```
-
-We now report the mean of the posterior distribution of each coefficient in Model 3, along with its associated measures of uncertainty.
-For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
-Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increases/decreases in probability of correct translation [@gelman2020regression].
-
-```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 3.
-A\\) Population-level effects.
-Distributions indicate the estimated posterior likelihood density of each coefficient.
-Credible intervals \\(CrI\\), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
-Dots represent the mean of the distribution.
-B\\) Participant\\-level coefficient variability.
-Our model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
-Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations \\(SD\\).
-C\\) Correlation between participant\\-level effects.
-Our model allowed participant-level coefficients to co\\-vary.
-This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each correlation.
-Coefficients are represented in the X\\-axis and Y\\-axis in the same order as indicated in the Y\\-axis of panels A and C.", fig.height=5, fig.width=6}
-
 str_repl <- c(
   "Intercept" = "Intercept",
-  "frequency_zipf" = "Frequency (+1 SD)",
+  "frequency_zipf" = "Frequency\n[+1 SD]",
   "pthn:lv" = "PTHN \u00d7 Levenshtein", 
-  "lv" = "Levenshtein (+1 SD)",
-  "pthn" = "PTHN (+1 SD)"
+  "lv:group1" = "Levenshtein \u00d7 Group 1",
+  "lv:group2" = "Levenshtein \u00d7 Group 2",
+  "lv" = "Levenshtein\n[+1 SD]",
+  "pthn" = "PTHN\n[+1 SD]",
+  "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
+  "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
 )
@@
       str_replace_all(str_repl) %>%  
-      factor(levels = c("Intercept", "Frequency (+1 SD)", "PTHN (+1 SD)", "Levenshtein (+1 SD)", "PTHN \u00d7 Levenshtein")),
+      factor(
+        levels = c(
+          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
+          "Group 1\n[ENG-* vs. SPA-CAT]",
+          "Group 2\n[ENG-CAT vs. ENG-SPA]",
+          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+        )
+      ),
     .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
@@
   theme(
-    legend.position = c(1, 0.1),
+    legend.position = c(1, 0.25),
+    legend.direction = "vertical",
+    legend.background = element_rect(fill = "grey95"),
+    legend.key = element_rect(fill = "grey95", colour = "grey95"), 
+    legend.key.height = unit(0.1, "cm"),
     axis.title.y = element_blank(),
+    legend.text = element_text(size = 7),
     axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black", vjust = 0),
+    axis.text.y = element_text(colour = "black", hjust = 1),
+    axis.ticks.y = element_blank(),
     panel.grid.major.y = element_line(colour = "grey", size = 0.5)
@@
 # SD of random effects
-post_sd <- posterior_draws_fixed %>% 
+post_sd_data <- posterior_draws_fixed %>% 
   filter(str_detect(.variable, "sd_")) %>% 
   mutate(
-    .variable_name = str_remove(.variable, "sd_participant__") %>% 
+    .variable_name = str_remove(.variable, "sd_participant__|sd_word__") %>% 
       str_replace_all(str_repl) %>% 
-      factor(levels = c("Intercept", "Frequency (+1 SD)", "PTHN (+1 SD)", "Levenshtein (+1 SD)", "PTHN \u00d7 Levenshtein"))
-  ) %>%   
-  ggplot(aes(.value, fct_rev(.variable_name))) +
+      factor(
+        levels = c(
+          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
+          "Group 1\n[ENG-* vs. SPA-CAT]",
+          "Group 2\n[ENG-CAT vs. ENG-SPA]",
+          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+        )
+      ),
+    group = case_when(
+      str_detect(.variable, "participant") ~ "Participants",
+      str_detect(.variable, "word") ~ "Words"
+    ),
+    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+  )
+
+post_sd <- ggplot(filter(post_sd_data, group=="Participants"), aes(.value, fct_rev(.variable_name))) +
   stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
@@
   scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
-  scale_x_continuous(labels = percent) +
-  labs(x = "SDs in P(Correct)", y = "Posterior probability density", fill = "CrI") +
+  scale_x_continuous(labels = percent, limits = c(0, 1)) +
+  labs(x = "SDs in P(Correct)", title = "By participant", fill = "CrI") +
+  theme(axis.text.x = element_blank()) +
+
+  ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
+  stat_pointinterval(.width = 0, point_size = 1) +
+  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
+  scale_x_continuous(labels = percent, limits = c(0, 1)) +
+  labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +
+
+  plot_layout(ncol = 1, heights = c(0.85, 0.15), guides = "collect") &
   theme(
-    legend.position = c(1, 0.1),
+    legend.position = "none",
+    axis.text.y = element_text(hjust = 1),
+    axis.title.x = element_blank(),
+    legend.direction = "vertical",
     axis.title.y = element_blank(),
-    axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black", vjust = 0),
+    legend.background = element_rect(fill = "grey95"),
+    legend.key.height = unit(0.1, "cm"),
+    legend.key = element_rect(fill = "grey95", colour = "grey95"),
+    legend.text = element_text(size = 7),
+    axis.ticks.y = element_blank(),
     panel.grid.major.y = element_line(colour = "grey", size = 0.5)
@@
 # Pearson correlations between random effects
-corr_mat <- cov2cor(vcov(model_fits$fit_3)) 
+corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)
+
+corr_mat <- corr_mat %>% 
+  as_tibble() %>% 
+  select(matches("Estimate")) %>% 
+  rename_all(str_remove, "Estimate.") %>% 
+  mutate(term1 = corr_names)
+
 corr_mat[lower.tri(corr_mat)] <- NA
 
-post_cors <- corr_mat %>% 
+corr_data <- corr_mat %>% 
   as.data.frame() %>% 
-  rownames_to_column("term1") %>% 
   pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
-  drop_na(.value) %>% 
+  drop_na(.value)
+
+post_cors <- corr_data %>% 
   mutate(term1 = factor(term1, levels = names(str_repl), ordered = TRUE),
@@
   mutate(
-    .value_label = case_when(
-      term1==term2 ~ NA_character_,
-      TRUE ~ printnum(.value, gt1 = FALSE)
-    ),
-    .value = case_when(
-      term1==term2 ~ NA_real_,
-      TRUE ~ .value
-    )
+    .value_label = case_when(term1==term2 ~ NA_character_, TRUE ~ printnum(.value, gt1 = FALSE)),
+    .value = case_when(term1==term2 ~ NA_real_, TRUE ~ .value)
   ) %>% 
+  filter(term1 != "PTHN × Levenshtein", term2 != "Intercept") %>% 
   ggplot(aes(fct_inorder(term1), fct_rev(fct_inorder(term2)), fill = .value)) +
@@
   geom_text(aes(label = .value_label), size = 3, na.rm = TRUE) +
-  labs(x = "Term 1", y = "Term 2", fill = "Posterior correlation") +
-  scale_fill_gradient(low = "#FFC20A", high = "white", na.value = "white") +
+  labs(x = "Term 1", y = "Term 2", fill = "Posterior\ncorrelation") +
+  scale_fill_gradient2(
+    low = "#04bf78", mid = "white", high = "#FFC20A", 
+    na.value = "white", limits = c(-0.4, 0.4)
+  ) +
+  scale_x_discrete(position = "top") +
   coord_equal() +
   theme(
-    legend.position = "none",
+    legend.position = c(1, 1),
+    legend.direction = "horizontal",
+    legend.key.height = unit(0.1, "cm"),
+    legend.key.width = unit(0.3, "cm"),
+    legend.justification = c(1, 1),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 5),
     axis.title.x = element_blank(),
     axis.title.y = element_blank(),
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank()
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 6, angle = 90, hjust = 0, vjust = 1),
+    axis.text.y = element_text(size = 6, vjust = 1, hjust = 1)
   )
 
-post_fix / (post_sd + post_cors + plot_layout(widths = c(0.6, 0.4))) +
-  plot_layout(guides = "keep") & 
+post_fix / ((post_sd | post_cors) + plot_layout(widths = c(0.6, 0.4))) +
+  plot_layout() +
   plot_annotation(tag_levels = "A")
@@
 
-Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (*SE* = `r percent(coefs$intercept["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations.
-Every standard deviation increment in the translation's lexical frequency (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))`) increased the probability of a correct responses in `r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)` (*SE* = `r percent(coefs$frequency_zipf["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$frequency_zipf["q2_5"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf["q97_5"], accuracy = 0.01)`]).
-The number of the translation's more frequent phonological neighbours, on the other hand, decreased the probability of a correct responses in `r percent(coefs$pthn["estimate"], accuracy = 0.01)` (*SE* = `r percent(coefs$pthn["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$pthn["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn["q97_5"], accuracy = 0.01)`]) for every increase in 1 *SD* (`r printnum(sd(stimuli$pthn, na.rm = TRUE))`).
-The effect of phonological similarity was conditional to the phonological density of the translation.
-Phonological similarity barely increased the probability of a correct responses (`r percent(coefs$lv["estimate"], accuracy = 0.01)`, *SE* = `r percent(coefs$lv["est_error"])`, 95% *CrI* = [`r percent(coefs$lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv["q97_5"])`]) by itself.
-However, for every *SD* increase in PTHN, phonological similarity increased in `r percent(coefs$pthn_lv["estimate"])`, (*SE* = `r percent(coefs$pthn_lv["est_error"], accuracy = 0.01)`, 95% *CrI* = [`r percent(coefs$pthn_lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_lv["q97_5"], accuracy = 0.01)`]) such probability.
+```{r bayesfactors}
+h_frequency <- abs(hypothesis(model_fits$fit_6, hypothesis = "frequency_zipf = 0")$hypothesis$Evid.Ratio)
+h_pthn <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn = 0")$hypothesis$Evid.Ratio)
+h_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv = 0")$hypothesis$Evid.Ratio)
+h_pthn_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn:lv = 0")$hypothesis$Evid.Ratio)
+h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
+h_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
+h_lv_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group1 = 0")$hypothesis$Evid.Ratio)
+h_lv_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group2 = 0")$hypothesis$Evid.Ratio)
 
-```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=3, fig.cap="Expected mean posterior predictions.
-A) Population-level expected mean posterior predictions.
-The X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 200 observations from our model: 100 simulations for words with low PTHN (-1 SD) and 100 simulations for high PTHN (+1 SD) words.
+icc_conditional <- icc(model_fits$fit_6, by_group = TRUE)$ICC[2]
+
+sds <- gather_draws(model_fits$fit_6, `sd_.*`, regex = TRUE) %>% 
+  mean_qi() %>% 
+  mutate(
+    .value = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.value), .value/4),
+    .lower = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.lower), .lower/4),
+    .upper = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.upper), .upper/4)
+  ) %>% 
+  split(.$.variable)
+
+names(sds) <- make_clean_names(names(sds))
+
+word_example <- ranef(model_fits$fit_6)$word[,,"Intercept"] %>% 
+  as.data.frame() %>% 
+  rownames_to_column("word") %>% 
+  arrange(-Estimate) %>% 
+  clean_names() %>% 
+  mutate_if(is.numeric, inv_logit_scaled) %>% 
+  slice(1, n()) %>% 
+  left_join(distinct(stimuli, word1, ipa1), by = c("word" = "word1")) %>% 
+  split(.$word)
+
+participant_ranef_range <- sds %>% 
+  bind_rows() %>% 
+  filter(!grepl("Intercept|word|group1", .variable)) %>% 
+  summarise(.value = range(.value)) %>% 
+  unlist()
+
+cors_range <- gather_draws(model_fits$fit_6, `cor_.*`, regex = TRUE) %>% 
+  median_qi() %>% 
+  summarise(.value = range(.value))
+
+corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)
+
+corr_data <- corr_mat %>% 
+  as_tibble() %>% 
+  clean_names() %>% 
+  mutate(term1 = corr_names) %>% 
+  relocate(term1) %>% 
+  split(.$term1)
+
+names(corr_data) <- make_clean_names(names(corr_data))
+
+h_cor_intercept_frequency <- hypothesis(model_fits$fit_6, "participant__Intercept__frequency_zipf = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_frequency_zipf_lv_group1 <- hypothesis(model_fits$fit_6, "participant__frequency_zipf__lv:group1 = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_intercept_pthn <- hypothesis(model_fits$fit_6, "participant__Intercept__pthn = 0", class = "cor")$hypothesis$Evid.Ratio
+```
+
+Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations.
+The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
+The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
+Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
+This difference increased in `r percent(coefs$pthn_lv["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
+
+Spanish natives translating Catalan words were `r percent(coefs$group1["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
+This difference increased in `r percent(coefs$lv_group1["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+English natives translating Catalan words were `r percent(abs(coefs$group2["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
+For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_group2["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+
+```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
+A) Population-level predictions: the X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 200 observations from our model: 100 simulations for words with low *PTHN* (-1 *SD*) and 100 simulations for high *PTHN* (+1 *SD*) words.
 We did this across the range of values of the Levenshtein scores.
 For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high PTHN words.
-We also plotted the mean of the high PTHN (black solid line) and low PTHN (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
+Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high *PTHN* words.
+We also plotted the mean of the high *PTHN* (black solid line) and low *PTHN* (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
 The dispersion of the lines indicates the uncertainty of our predictions.
-B) Participant-level expected mean posterior predictions.
-We conducted the same procedure for each individual participant, simulating 100 observations for high-PTHN and 100 for low-PTHN from our model acrosss the range of Levenshtein scores.
-We averaged the resulting predictions for each participant and then plotted each participant's predictions in separate panels to show how the effect of PTHN and phonological similarity changed at the individual level."}
+We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
 
-# by vowel_ratio
 m <- posterior_epreds_fixed %>% 
@@
 
-m_re <- posterior_epreds_random %>% 
-  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
-  mutate(
-    pthn = case_when(
-      pthn=="-1 SD" ~ "Low PTHN",
-      pthn=="0 SD" ~ "Mean PTHN",
-      pthn=="1 SD" ~ "High PTHN"
-    ),
-    frequency_zipf = case_when(
-      frequency_zipf=="-1 SD" ~ "Low frequency",
-      frequency_zipf=="0 SD" ~ "Mean frequency",
-      frequency_zipf=="1 SD" ~ "High frequency"
-    )
-  ) %>% 
-  group_by(participant, lv, pthn) %>% 
-  summarise(.epred = mean(.epred, na.rm = TRUE), .groups = "drop")
-
-plot_m <- ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
-  # geom_jitter(data = filter(nd_obs, correct), aes(y = 1), alpha = 0.1,
-  #             size = 0.25, height = 0.05, width = 0.1, show.legend = FALSE) +
-  # geom_jitter(data = filter(nd_obs, !correct), aes(y = 0), alpha = 0.25, 
-  #             size = 0.25, height = 0.05, width = 0.1, show.legend = FALSE) +
+ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
+  facet_wrap(~group) +
   geom_hline(yintercept = 0.5, colour = "grey") +
-  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.25, show.legend = FALSE) +
-  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.5, show.legend = FALSE) +
-  labs(x = "Levenshtein (phonological similarity)", y = "P(Correct)", 
+  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
+  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
+  annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
+  annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
+  annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
+  annotate("text", x = 2.15, y = 0.95, size = 2.75, label = "More similar") +
+  labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
        colour = "PTHN", fill = "PTHN", linetype = "PTHN",
@@
   scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
-  guides(colour = guide_legend(ncol = 2))
-
-plot_m_re <- ggplot(m_re, aes(lv, .epred, colour = pthn, fill = pthn)) +
-  facet_wrap(~participant) +
-  geom_line(aes(group = interaction(pthn, participant)), size = 0.35) +
-  labs(x = "Levenshtein (phonological similarity)", y = "P(Correct)", 
-       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
-       title = "Participant-level predictions") +
-  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
-  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.1)) +
-  guides(colour = guide_legend(ncol = 2)) +
-  theme(
-    legend.position = "none",
-    axis.text.x = element_blank(),
-    axis.text.y = element_blank(),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank(),
-    axis.ticks = element_blank(),
-    strip.text = element_blank(),
-    strip.background = element_blank(),
-    panel.spacing = unit(0.1, "lines")
-  )
-
-plot_m + plot_m_re +
-  plot_layout(guides = "collect") &
-  plot_annotation(tag_levels = "A") &
   theme(
     legend.position = "top",
-    legend.title = element_blank(),
+    plot.title = element_blank()
   )
@@
 ```
+
+The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data].
+Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_Intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_Intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
+Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants.
+Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
+The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]).
+This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
+
+Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions.
+The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r `round(h_cor_intercept_frequency, 2)`).
+The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf$estimate_lv_group1, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf$q2_5_lv_group1, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf$q97_5_lv_group1, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
+The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).
+The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
 
@@
 \newpage
-
-# Appendix 1: MCMC diagnostics
-
-```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-suppressWarnings({
-  bayesplot_theme_set(
-    new = theme_custom() + 
-      theme(
-        axis.text.x = element_text(size = 8), 
-        axis.text.y = element_text(size = 8),
-        strip.text = element_text(size = 10),
-        legend.key = element_rect(colour = NA, fill = "white")
-      )
-  )
-})
-
-color_scheme_set("brightblue")
-pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")
-
-```
-
-## Pairs plot
-
-```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-posterior_cp <- as.array(model_fits$fit_3)
-np_cp <- nuts_params(model_fits$fit_3)
-
-plot_pairs <- mcmc_pairs(
-  posterior_cp, pars = pars,
-  off_diag_args = list(size = 0.75)
-)
-
-plot_pairs
-```
-
-## Trace and Rank plots
-
-```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-(mcmc_trace(model_fits$fit_3, pars) + legend_none()) /
-  mcmc_rank_overlay(model_fits$fit_3, pars = pars) +
-  plot_layout(guides = "collect") &
-  theme(
-    legend.position = "top"
-  )
-```
-
-## Gelman-Rubin convergence criterion (R-hat)
-
-```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-rhats <- rhat(model_fits$fit_3)
-ratios <- neff_ratio(model_fits$fit_3)
-
-mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
-  mcmc_neff_hist(ratios, binwidth = 0.1) +
-  plot_layout(guides = "keep", ncol = 1) &
-  theme(
-    legend.position = "top"
-  )
-```
-
-# Appendix 2: by item accuracy
-
-```{r appendix-accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
-Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.
-Dots and whiskers represent the average accuracy and 95% confidence interval of each word.
-Accuracy is plotted separately for each group."}
-
-responses_item <- responses %>%
-  group_by(trial_id, group, word) %>%
-  summarise(
-    correct_sum = sum(correct, na.rm = TRUE),
-    correct_n = sum(!is.na(correct), na.rm = TRUE),
-    .groups = "drop"
-  ) %>% 
-  left_join(select(stimuli, group, trial_id, word2, ipa1, ipa2)) %>% 
-  mutate(
-    word = paste0(word, " (", ipa1, ") / ", word2, " (", ipa2, ")"),
-    correct_prop = prop_adj(correct_sum, correct_n),
-    correct_se = prop_adj_se(correct_sum, correct_n),
-    word_label = reorder_within(word, by = correct_prop, within = group)
-  )
-
-ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
-  facet_wrap(~group, scales = "free_y") +
-  geom_hline(yintercept = 0.5, colour = "black") +
-  geom_errorbar(width = 0) +
-  geom_point(size = 0.5) +
-  labs(y = "Proportion of correct responses", x = "Word", colour = "Group") +
-  coord_flip() +
-  scale_color_manual(values = c("#1E88E5", "#ff2976", "#FFC20A")) +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
-  scale_x_discrete(labels = function(x) str_remove(x, "___ENG-CAT|___ENG-SPA|___SPA-CAT")) +
-  theme(
-    legend.position = "none",
-    axis.text.y = element_text(size = 5, colour = "black"),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 9, colour = "black"),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-  )
-
-```

### 2021-12-16 R/compare_methods.R

@@ -1,45 +1,0 @@
-tar_load(responses)
-
-library(stringdist)
-
-threshold <- 0.6
-
-d <- responses %>% 
-    mutate(
-        overlap_lv = stringsim(input_text, word1, method = "lv"),
-        overlap_osa = stringsim(input_text, word1, method = "osa"),
-        overlap_hamming = stringsim(input_text, word1, method = "hamming"),
-        overlap_jw = stringsim(input_text, word1, method = "jw"),
-        overlap_cosine = stringsim(input_text, word1, method = "cosine")
-    ) %>% 
-    mutate(
-        is_correct_exact = input_text==word2,
-        is_correct_manual = correct,
-        is_correct_lv = overlap_lv >= threshold,
-        is_correct_osa = overlap_lv >= threshold,
-        is_correct_hamming = overlap_lv >= threshold,
-        is_correct_cosine = overlap_lv >= threshold
-    ) %>% 
-    select(group, trial_id, word, participant, matches("is_correct")) %>% 
-    pivot_longer(c(matches("is_correct"), -is_correct_manual), names_to = "method", values_to = "is_correct") %>% 
-    mutate(method = str_to_sentence(str_remove(method, "is_correct_"))) %>% 
-    group_by(group, word, method) %>% 
-    summarise(
-        is_correct_manual = sum(is_correct_manual),
-        is_correct = sum(is_correct),
-        n = n(),
-        prop_correct_manual = is_correct_manual/n,
-        prop_correct = is_correct/n,
-        .groups = "drop"
-    )
-    
-ggplot(d, aes(prop_correct_manual, prop_correct, colour = group)) +
-    facet_wrap(~method) +
-    # geom_rug(sides = "tr") +
-    geom_point(size = 0.75, alpha = 0.5) +
-    labs(x = "% correct (manual classification)",
-         y = "% correct (automatic classification) \n Similarity threshold: 60%") +
-    scale_x_continuous(labels = percent) +
-    scale_y_continuous(labels = percent) +
-    coord_equal() +
-    theme_ggdist()

### 2021-12-16 R/utils.R

@@ -41,6 +41,25 @@ theme_custom <- function(){
     )
 }
 
+# remove special characters
+replace_non_ascii <- function(x){
+    str_replace_all(
+        x,
+        c(
+            "á" = "a",
+            "é" = "e",
+            "í" = "i",
+            "ó" = "o",
+            "ú" = "u",
+            "à" = "a",
+            "è" = "e",
+            "ò" = "o",
+            "ñ" = "n",
+            "ç" = "c"
+        )
+        )
+}
+
 # extract formants
 extract_formants <- function(
     file,

### 2021-12-16 _targets_packages.R

@@ -8,6 +8,7 @@ library(data.table)
 library(dplyr)
 library(forcats)
 library(future)
+library(GGally)
 library(ggplot2)
 library(gt)
 library(here)

### 2021-12-20 docs/00_stimuli.Rmd
Reorganise repo

---
title: "Stimuli"
output:
  md_document:
    variant: markdown_github
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "80%",
  results = "asis",
  dpi = 300,
  dev.args = list(png = list(type = "cairo"))
)

options(
  knitr.kable.NA = '-',
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.continuous.fill = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  ),
  ggplot2.continuous.colour = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  )
)
```

```{r prepare}
# import data
tar_load_globals()
tar_load(stimuli)

theme_set(theme_custom())
```

# Overview

Participants in the ENG-SPA listened to `r nrow(stimuli[stimuli$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials. Five trials in each condition were practice trials and were excluded from analyses.

```{r stimuli_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
stimuli_coded <- mutate(stimuli, test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"))
```

# Lexical frequency

```{r stimuli_frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  distinct(test_language, word1, .keep_all = TRUE) %>% 
  left_join(count(., test_language, name = "n_total")) %>% 
  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
  ggplot(aes(frequency_zipf, fill = test_language)) +
  facet_wrap(~test_language, ncol = 1) +
  geom_histogram(bins = 25, colour = "white") +
  labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language")  +
  scale_y_continuous(breaks = seq(0, 12, 2)) +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

```

# PTHN

```{r stimuli_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  left_join(count(., test_language, name = "n_total")) %>% 
  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
  count(test_language, pthn) %>% 
  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
  facet_wrap(~test_language, ncol = 1) +
  geom_col(position = position_dodge()) +
  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
  theme(
    legend.position = "none",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

```

# Phonological similarity

```{r stimuli_similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
  left_join(count(., group, name = "n_total")) %>% 
  mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
  ggplot(aes(group, lv, fill = group, colour = group)) +
  geom_violin() +
  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
  labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme(
    legend.position = "none",
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

```

### 2021-12-20 docs/02_data-analysis.Rmd
Reorganise repo

---
title: "Data analysis"
output:
  md_document:
    variant: markdown_github
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "80%",
  results = "asis",
  dpi = 300,
  dev.args = list(png = list(type = "cairo"))
)
options(
  knitr.kable.NA = '-',
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.continuous.fill = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  ),
  ggplot2.continuous.colour = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  )
)
```

```{r prepare}
# import data
tar_load_globals()

theme_set(theme_custom())
```

# Model details

We fitted a Bayesian multilevel regression model using the R package `brms` [@burkner2017brms], with correct responses (`correct`, 0 is incorrect response, 1 if correct response) as the response variable with a Bernoulli distribution and a logit link function. We modelled the probability of an average participant providing a correct response, conditional to a set of predictors (aka. fixed effects) and their interactions:

* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database. Standardised before entering the regression model.
* Phonological similarity (`lv`, 0-1): Inverse Levenshtein distance between the IPA transcription of the presented word and its correct translation in the target language, calculated using the `stringdist` function of the `stringdist` package [@loo2014stringdist]. Standardised before entering the regression model.
* Shared phonological onset (`onset`, Same/Different): Whether the presented and the target word share phonological onset, as judged by experimenters, and sum coded as $Different = -0.5$ and $Same = +0.5$.

We added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model. We specified random intercepts slopes for all fixed effects within the `participant` grouping variable.

We performed multiple imputation via predictive mean matching using the `mice` function of the `mice` package [@van2011mice] to impute missing values in the response variable or the predictors.

The model was implemented using the following formula (`fit_3`):

$$
\begin{align}

&\textbf{Likelihood}  \\
y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\

&\textbf{Parameters}  \\

logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
& \beta_{1[p]} ~ Frequency_{i} ~ + \\
& \beta_{2[p]} ~ PTHN_i ~ + \\
& \beta_{3[p]} ~ Similarity_i ~ + \\
& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\

\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
&& \text{[participant-level coefficients]} \\ \\

&\textbf{Prior}  \\

\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\

\end{align}
$$

The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 iterations each (1,000 warm-ups iterations). The output of this model is the approximated posterior distribution of all the parameters estimated, therefore indicating the probability of each combination of values of the parameters, conditional to the observed data.

We compared this (extended model) to other models dropping one predictor at a time using Leave-one-out cross-validation (LOO-CV), using the `loo` and `loo_compare` functions of the `brms` package. We compared our extended model against the following models:

* fit_0: `~ 1 + (1 | participant) + (1 | word)`
* fit_1: `~ 1 + frequency_zipf + (1 + frequency_zipf | participant) + (1 | word)`
* fit_2: `~ 1 + frequency_zipf + pthn + (1 + frequency_zipf + pthn | participant) + (1 | word)`
* fit_3: `~ 1 + frequency_zipf + pthn + lv + (1 + frequency_zipf + pthn + lv | participant) + (1 | word)`
* fit_4: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_5: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_6: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + group:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`

# Model implementation

## Stan code

```stan stan_model
// generated with brms 2.16.1
functions {
 /* compute correlated group-level effects
  * Args: 
  *   z: matrix of unscaled group-level effects
  *   SD: vector of standard deviation parameters
  *   L: cholesky factor correlation matrix
  * Returns: 
  *   matrix of scaled group-level effects
  */ 
  matrix scale_r_cor(matrix z, vector SD, matrix L) {
    // r is stored in another dimension order than z
    return transpose(diag_pre_multiply(SD, L) * z);
  }
}
data {
  int<lower=1> N;  // total number of observations
  int Y[N];  // response variable
  int<lower=1> K;  // number of population-level effects
  matrix[N, K] X;  // population-level design matrix
  // data for group-level effects of ID 1
  int<lower=1> N_1;  // number of grouping levels
  int<lower=1> M_1;  // number of coefficients per level
  int<lower=1> J_1[N];  // grouping indicator per observation
  // group-level predictor values
  vector[N] Z_1_1;
  vector[N] Z_1_2;
  vector[N] Z_1_3;
  vector[N] Z_1_4;
  vector[N] Z_1_5;
  vector[N] Z_1_6;
  vector[N] Z_1_7;
  int<lower=1> NC_1;  // number of group-level correlations
  // data for group-level effects of ID 2
  int<lower=1> N_2;  // number of grouping levels
  int<lower=1> M_2;  // number of coefficients per level
  int<lower=1> J_2[N];  // grouping indicator per observation
  // group-level predictor values
  vector[N] Z_2_1;
  int prior_only;  // should the likelihood be ignored?
}
transformed data {
  int Kc = K - 1;
  matrix[N, Kc] Xc;  // centered version of X without an intercept
  vector[Kc] means_X;  // column means of X before centering
  for (i in 2:K) {
    means_X[i - 1] = mean(X[, i]);
    Xc[, i - 1] = X[, i] - means_X[i - 1];
  }
}
parameters {
  vector[Kc] b;  // population-level effects
  real Intercept;  // temporary intercept for centered predictors
  vector<lower=0>[M_1] sd_1;  // group-level standard deviations
  matrix[M_1, N_1] z_1;  // standardized group-level effects
  cholesky_factor_corr[M_1] L_1;  // cholesky factor of correlation matrix
  vector<lower=0>[M_2] sd_2;  // group-level standard deviations
  vector[N_2] z_2[M_2];  // standardized group-level effects
}
transformed parameters {
  matrix[N_1, M_1] r_1;  // actual group-level effects
  // using vectors speeds up indexing in loops
  vector[N_1] r_1_1;
  vector[N_1] r_1_2;
  vector[N_1] r_1_3;
  vector[N_1] r_1_4;
  vector[N_1] r_1_5;
  vector[N_1] r_1_6;
  vector[N_1] r_1_7;
  vector[N_2] r_2_1;  // actual group-level effects
  // compute actual group-level effects
  r_1 = scale_r_cor(z_1, sd_1, L_1);
  r_1_1 = r_1[, 1];
  r_1_2 = r_1[, 2];
  r_1_3 = r_1[, 3];
  r_1_4 = r_1[, 4];
  r_1_5 = r_1[, 5];
  r_1_6 = r_1[, 6];
  r_1_7 = r_1[, 7];
  r_2_1 = (sd_2[1] * (z_2[1]));
}
model {
  // likelihood including constants
  if (!prior_only) {
    // initialize linear predictor term
    vector[N] mu = Intercept + rep_vector(0.0, N);
    for (n in 1:N) {
      // add more terms to the linear predictor
      mu[n] += r_1_1[J_1[n]] * Z_1_1[n] + r_1_2[J_1[n]] * Z_1_2[n] + r_1_3[J_1[n]] * Z_1_3[n] + r_1_4[J_1[n]] * Z_1_4[n] + r_1_5[J_1[n]] * Z_1_5[n] + r_1_6[J_1[n]] * Z_1_6[n] + r_1_7[J_1[n]] * Z_1_7[n] + r_2_1[J_2[n]] * Z_2_1[n];
    }
    target += bernoulli_logit_glm_lpmf(Y | Xc, mu, b);
  }
  // priors including constants
  target += normal_lpdf(b | 0, 0.1);
  target += normal_lpdf(Intercept | 0, 0.1);
  target += cauchy_lpdf(sd_1 | 0, 0.1)
    - 7 * cauchy_lccdf(0 | 0, 0.1);
  target += std_normal_lpdf(to_vector(z_1));
  target += lkj_corr_cholesky_lpdf(L_1 | 8);
  target += cauchy_lpdf(sd_2 | 0, 0.1)
    - 1 * cauchy_lccdf(0 | 0, 0.1);
  target += std_normal_lpdf(z_2[1]);
}
generated quantities {
  // actual population-level intercept
  real b_Intercept = Intercept - dot_product(means_X, b);
  // compute group-level correlations
  corr_matrix[M_1] Cor_1 = multiply_lower_tri_self_transpose(L_1);
  vector<lower=-1,upper=1>[NC_1] cor_1;
  // additionally sample draws from priors
  real prior_b = normal_rng(0,0.1);
  real prior_Intercept = normal_rng(0,0.1);
  real prior_sd_1 = cauchy_rng(0,0.1);
  real prior_cor_1 = lkj_corr_rng(M_1,8)[1, 2];
  real prior_sd_2 = cauchy_rng(0,0.1);
  // extract upper diagonal of correlation matrix
  for (k in 1:M_1) {
    for (j in 1:(k - 1)) {
      cor_1[choose(k - 1, 2) + j] = Cor_1[j, k];
    }
  }
  // use rejection sampling for truncated priors
  while (prior_sd_1 < 0) {
    prior_sd_1 = cauchy_rng(0,0.1);
  }
  while (prior_sd_2 < 0) {
    prior_sd_2 = cauchy_rng(0,0.1);
  }
}
```

### 2021-12-20 manuscript/manuscript.Rmd
Reorganise repo

--- 2021-12-16 Rmd/manuscript.Rmd
@@
 # set custom ggplot2 theme
-theme_set(
-  theme(
-    title = element_text(size = 8, face = "bold"),
-
-    legend.background = element_rect(fill = NA, colour = NA),
-    legend.key = element_rect(fill = "white", colour = NA),
-    legend.justification = "right",
-    legend.direction = "horizontal",
-
-    axis.line = element_line(colour = "black", size = 0.5),
-    axis.ticks = element_line(colour = "black", size = 0.5),
-    axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black"),
-
-    plot.caption.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-
-    panel.grid = element_blank(),
-    panel.background = element_rect(fill = "white", colour = NA)
-  )
-)
+theme_set(theme_custom()) # in utlis.R
 
@@
 
-include_graphics(here("Figures", "design.png"))
+include_graphics(here("img", "design.png"))
 
@@
   ) %>% 
-  gtsave(here("Figures", "stimuli.png"))
+  gtsave(here("img", "stimuli.png"))
 
-include_graphics(here("Figures", "stimuli.png"), auto_pdf = FALSE)
+include_graphics(here("img", "stimuli.png"), auto_pdf = FALSE)
 
@@
 
-include_graphics(here("Figures", "loos.png"))
+include_graphics(here("img", "loos.png"))
 
@@
 
-```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 6.
 A\\) Population-level effects.
@@
   "pthn:lv" = "PTHN \u00d7 Levenshtein", 
-  "lv:group1" = "Levenshtein \u00d7 Group 1",
-  "lv:group2" = "Levenshtein \u00d7 Group 2",
+  "lv:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
   "lv" = "Levenshtein\n[+1 SD]",
   "pthn" = "PTHN\n[+1 SD]",
-  "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-  "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
 )
@@
 h_pthn_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn:lv = 0")$hypothesis$Evid.Ratio)
-h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
-h_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
-h_lv_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group1 = 0")$hypothesis$Evid.Ratio)
-h_lv_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group2 = 0")$hypothesis$Evid.Ratio)
+h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
+h_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)
+h_lv_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
+h_lv_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)
 
@@
 
-h_cor_intercept_frequency <- hypothesis(model_fits$fit_6, "participant__Intercept__frequency_zipf = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_frequency_zipf_lv_group1 <- hypothesis(model_fits$fit_6, "participant__frequency_zipf__lv:group1 = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_intercept_pthn <- hypothesis(model_fits$fit_6, "participant__Intercept__pthn = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_intercept_frequency <- hypothesis(
+  model_fits$fit_6, 
+  "participant__Intercept__frequency_zipf = 0", 
+  class = "cor"
+  )$hypothesis$Evid.Ratio
+
+h_cor_frequency_zipf_lv_group1 <- hypothesis(
+  model_fits$fit_6, 
+  "participant__frequency_zipf__lv:groupENGMCATvs.ENGMSPA  = 0", 
+  class = "cor"
+  )$hypothesis$Evid.Ratio
+
+h_cor_intercept_pthn <- hypothesis(
+  model_fits$fit_6,
+  "participant__Intercept__pthn = 0", 
+  class = "cor"
+  )$hypothesis$Evid.Ratio
 ```
@@
 
-Spanish natives translating Catalan words were `r percent(coefs$group1["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
-This difference increased in `r percent(coefs$lv_group1["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-English natives translating Catalan words were `r percent(abs(coefs$group2["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
-For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_group2["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+Spanish natives translating Catalan words were `r percent(coefs$group_en_gvs_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
+This difference increased in `r percent(coefs$lv_group_en_gvs_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+English natives translating Catalan words were `r percent(abs(coefs$group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
+For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
 
@@
   geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
-  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
-  annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
-  annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
+  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", 
+               colour = "black", size = 0.75, show.legend = TRUE) +
+  annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, 
+           arrow = arrow(length = unit(0.1, "cm")), size = .5) +
+  annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, 
+           arrow = arrow(length = unit(0.1, "cm")), size = .5) +
   annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
@@
 Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
-The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]).
+The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group_en_gvs_spa$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group_en_gvs_spa$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group_en_gvs_spa$.upper, accuracy = 0.01)`]).
 This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.

### 2021-12-20 manuscript/manuscript_appendix.Rmd
Reorganise repo

--- 2021-11-25 Rmd/manuscript_appendix.Rmd
@@
 
-```{r accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
+```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+tar_load_globals()
+tar_load(stimuli)
+tar_load(participants)
+tar_load(responses)
+tar_load(model_fits)
+
+suppressWarnings({
+  bayesplot_theme_set(
+    new = theme_custom() + 
+      theme(
+        axis.text.x = element_text(size = 8), 
+        axis.text.y = element_text(size = 8),
+        strip.text = element_text(size = 10),
+        legend.key = element_rect(colour = NA, fill = "white")
+      )
+  )
+})
+
+color_scheme_set("brightblue")
+pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")
+
+```
+
+# Appendix 1: Stan code
+
+```{r appendix-code, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+model_fits$fit_3$model
+```
+
+# Appendix 2: MCMC diagnostics
+
+## Pairs plot
+
+```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+posterior_cp <- as.array(model_fits$fit_3)
+np_cp <- nuts_params(model_fits$fit_3)
+
+plot_pairs <- mcmc_pairs(
+  posterior_cp, pars = pars,
+  off_diag_args = list(size = 0.75)
+)
+
+plot_pairs
+```
+
+## Trace plots
+
+```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+(mcmc_trace(model_fits$fit_3, pars) + legend_none()) /
+  mcmc_rank_overlay(model_fits$fit_3, pars = pars) +
+  plot_layout(guides = "collect") &
+  theme(
+    legend.position = "top"
+  )
+```
+
+## Gelman-Rubin convergence criterion (R-hat)
+
+```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+rhats <- rhat(model_fits$fit_3)
+ratios <- neff_ratio(model_fits$fit_3)
+
+mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
+  mcmc_neff_hist(ratios, binwidth = 0.1) +
+  plot_layout(guides = "keep", ncol = 1) &
+  theme(
+    legend.position = "right"
+  )
+```
+
+# Appendix 3: by item accuracy
+
+```{r appendix-accuracy, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Proportion of correct translation by item.
 Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.

### 2021-12-20 R/00_stimuli.R

@@ -1,71 +1,0 @@
-# relative frequencies and phonological neighbours
-get_clearpond <- function(clearpond_path){
-    clearpond <- clearpond_path %>% 
-        map(
-            function(x){
-                read.csv(x) %>% 
-                    clean_names() %>% 
-                    select("word", "pho_word", "freq_per_million", "pthn")
-            }
-        ) %>% 
-        bind_rows(.id = "group") %>% 
-        as_tibble() %>%
-        distinct(word, group, .keep_all = TRUE) %>% 
-        rename(word2 = word, phon_clearpond = pho_word, frequency = freq_per_million)
-    
-    return(clearpond)
-}
-
-get_levenshtein <- function(stimuli_path){
-
-    levenshtein <- map(
-        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
-        function(x) read_xlsx(stimuli_path, sheet = x)
-    ) %>% 
-        bind_rows(.id = "group") %>% 
-        clean_names() %>% 
-        select(group, word1, word2, ipa1, ipa2) %>% 
-        mutate(
-            n_char = ifelse(nchar(ipa1) > nchar(ipa2), nchar(ipa1), nchar(ipa2)), # number of characters
-            lv = stringsim(ipa1, ipa2, method = "lv") # Levenshtein similarity (proportion)
-        )
-    
-    return(levenshtein)
-}
-
-
-# get audio duration
-get_duration <- function(stimuli_path, audios_path){
-    stimuli <- map(
-        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
-        ~read_xlsx(stimuli_path, sheet = .)) %>% 
-        bind_rows(.id = "group") %>% 
-        select(word1, group, file) %>% 
-        mutate(file = here("Stimuli", "Sounds", file))
-    
-    audios <- map(stimuli$file, load.wave) # load audio
-    lengths <- map_dbl(audios, length)/2 # get number of samples (only one channel)
-    sampling_rate <- unique(map_dbl(audios, ~attr(., "rate"))) # get sampling rates
-    duration <- lengths/sampling_rate # get duration (in seconds)
-    return(duration)
-}
-
-get_stimuli <- function(stimuli_path, clearpond, levenshtein, durations){
-    
-    stimuli <- map(
[…]

### 2021-12-20 R/01_responses.R

@@ -1,209 +1,0 @@
-# preprocess responses
-get_responses_processed <- function(
-    responses_path,
-    stimuli,
-    practice_trials,
-    seed = 888
-){
-    # set parameters
-    set.seed(seed) # for reproducibility
-    spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
-    
-    # list CSV files in Data/Raw/ folder
-    filenames <- list.files(here("Data", "Raw")) %>% 
-        str_extract(".*?\\_") %>% 
-        str_remove("_")
-    
-    # participant files
-    processed <- responses_path %>% 
-        map(fread, na.string = "") %>%  # import participant files
-        map(mutate, participant = as.character(participant)) %>% 
-        set_names(filenames) %>% # label each dataset with the participant's ID
-        bind_rows(.id = "filename") %>% # merge datasets
-        as_tibble() %>%
-        clean_names() %>% 
-        # select relevant variables and rename if necessary
-        select(
-            participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
-            matches("language_l|language_s|language_c|demo_|setup_"),
-            -matches("_rt|_time"),
-            key_press_time
-        ) %>% 
-        rename(demo_impairment = demo_language_key_keys) %>% 
-        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
-        # clean text input by participants (because of typos of need to translate) and redefine location    
-        group_by(participant) %>% 
-        # for each participant, get first row of each demo variable
-        mutate_at(
-            vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), age, city, setup_location, setup_noise),
-            first_non_na # defined in R/utils.R
-        ) %>% 
-        mutate(date = max(date, na.rm = TRUE)) %>% 
-        ungroup() %>% 
-        mutate_at(
-            vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), 
-            clean_input_text # defiend in R/utils.R
-        ) %>% 
-        mutate(
-            city = str_to_sentence(city), # capitalise first letter of city names
-            country = ifelse(city %in% spanish_cities, "Spain", "UK"), # group cities into countries
-            date = as_datetime(ymd_hms(date)), # format dates
-            demo_sex = case_when(
[…]

### 2021-12-20 R/utils.R

@@ -1,175 +1,0 @@
-#### utils: helper functions ---------------------------------------------------
-
-# evaluate if x is NOT an element of y
-"%!in%" <- function(x, y) !(x %in% y)
-
-# find first/last non-missing value in vector
-first_non_na <- function(x) ifelse(is.logical(first(x[!is.na(x)])), NA, first(x[!is.na(x)]))
-
-last_non_na <- function(x) ifelse(is.logical(last(x[!is.na(x)])), NA, last(x[!is.na(x)]))
-
-add_big_mark <- function(x) format(x, big.mark = ",", scientific = FALSE)
-
-
-# text input responses are messy: we need to homogenise them
-clean_input_text <- function(x) {
-    {{ x }} %>% 
-        str_to_sentence() %>%
-        str_replace("Lshift|space|minus", "") %>%
-        as.character() %>% 
-        as_tibble(x = .) %>% 
-        mutate(. , value = case_when( 
-            nchar(value) > 12 ~ "Several",
-            value %in% c("Francés", "Frances", "Basic french") ~ "French",
-            value %in% c("Italiano", "Italino") ~ "Italian",
-            value %in% c("Alemán") ~ "German",
-            value %in% c("NrLondon") ~ "London",
-            value %in% c("Rnorthampton") ~ "London",
-            value %in% c("Newcastleminusuponminustyne") ~ "Newcastle upon Tyne",
-            value %in% c("Newcastlespaceuk") ~ "Newcastle",
-            value %in% c("Miltonkeynes") ~ "Milton Keynes",
-            value %in% c("Irthlingboroug") ~ "Irthlingborough",
-            value %in% c("spalding", "berkshire", "bedfordshire") ~ str_to_sentence(value),
-            value %in% c("No", "Cuál", "Space") ~ NA_character_,
-            TRUE ~ value)) %>% 
-        pull(value)
-}
-
-# ggplot custom theme
-theme_custom <- function(){
-    theme(
-        panel.background = element_rect(fill = "transparent"),
-        panel.grid = element_blank(),
-        #panel.grid = element_line(colour = "grey", linetype = "dotted"),
-        panel.border = element_rect(fill = "transparent", colour = "black"),
-        text = element_text(colour = "black", size = 15),
-        axis.text = element_text(colour = "black")
-    )
-}
-
[…]

### 2021-12-20 scripts/R/00_stimuli.R

@@ -1,0 +1,73 @@
+# relative frequencies and phonological neighbours
+get_clearpond <- function(clearpond_path){
+    clearpond <- clearpond_path %>% 
+        map(
+            function(x){
+                read.csv(x) %>% 
+                    clean_names() %>% 
+                    select("word", "pho_word", "freq_per_million", "pthn")
+            }
+        ) %>% 
+        bind_rows(.id = "group") %>% 
+        as_tibble() %>%
+        distinct(word, group, .keep_all = TRUE) %>% 
+        rename(word2 = word, phon_clearpond = pho_word, frequency = freq_per_million)
+    
+    return(clearpond)
+}
+
+get_levenshtein <- function(stimuli_path){
+
+    levenshtein <- map(
+        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
+        ~read_xlsx(stimuli_path, sheet = .)
+    ) %>% 
+        bind_rows(.id = "group") %>% 
+        clean_names() %>% 
+        select(group, word1, word2, ipa1, ipa2) %>% 
+        mutate(
+            n_char = ifelse(nchar(ipa1) > nchar(ipa2), nchar(ipa1), nchar(ipa2)), # number of characters
+            lv = stringsim(ipa1, ipa2, method = "lv") # Levenshtein similarity (proportion)
+        )
+    
+    return(levenshtein)
+}
+
+
+# get audio duration
+get_duration <- function(stimuli_path, audios_path){
+    stimuli <- map(
+        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
+        ~read_xlsx(stimuli_path, sheet = .)) %>% 
+        bind_rows(.id = "group") %>% 
+        select(word1, group, file) %>% 
+        mutate(file = here("stimuli", file))
+    
+    audios <- map(stimuli$file, load.wave) # load audio
+    lengths <- map_dbl(audios, length)/2 # get number of samples (only one channel)
+    sampling_rate <- unique(map_dbl(audios, ~attr(., "rate"))) # get sampling rates
+    duration <- lengths/sampling_rate # get duration (in seconds)
+    return(duration)
+}
+
+get_stimuli <- function(stimuli_path, clearpond, levenshtein, durations, save = TRUE){
+    
+    stimuli <- map(
[…]

### 2021-12-20 scripts/R/01_responses.R

@@ -1,0 +1,207 @@
+# preprocess responses
+get_responses_processed <- function(
+    responses_path,
+    stimuli,
+    practice_trials,
+    seed = 888
+){
+    # set parameters
+    set.seed(seed) # for reproducibility
+    spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
+    
+    # list CSV files in Data/Raw/ folder
+    filenames <- list.files(here("data", "raw")) %>% 
+        str_extract(".*?\\_") %>% 
+        str_remove("_")
+    
+    # participant files
+    processed <- responses_path %>% 
+        map(fread, na.string = "") %>%  # import participant files
+        map(mutate, participant = as.character(participant)) %>% 
+        set_names(filenames) %>% # label each dataset with the participant's ID
+        bind_rows(.id = "filename") %>% # merge datasets
+        as_tibble() %>%
+        clean_names() %>% 
+        # select relevant variables and rename if necessary
+        select(
+            participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
+            matches("language_l|language_s|language_c|demo_|setup_"),
+            -matches("_rt|_time"),
+            key_press_time
+        ) %>% 
+        rename(demo_impairment = demo_language_key_keys) %>% 
+        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
+        # clean text input by participants (because of typos of need to translate) and redefine location    
+        group_by(participant) %>% 
+        # for each participant, get first row of each demo variable
+        mutate_at(
+            vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), age, city, setup_location, setup_noise),
+            first_non_na # defined in R/utils.R
+        ) %>% 
+        mutate(date = max(date, na.rm = TRUE)) %>% 
+        ungroup() %>% 
+        mutate_at(
+            vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), 
+            clean_input_text # defiend in R/utils.R
+        ) %>% 
+        mutate( 
+            city = str_to_sentence(city), # capitalise first letter of city names
+            city = ifelse(participant == "55lvve0vwnrihv0rihqipejv1nfnr", "Málaga", city),
+            country = ifelse(city %in% spanish_cities, "Spain", "UK"), # group cities into countries
[…]

### 2021-12-20 scripts/R/02_models.R

@@ -1,0 +1,214 @@
+# models
+
+get_model_fits <- function(
+    responses,
+    model_formulas, # this is a list of formulas (see _targets.R)
+    model_prior, 
+    iter = 4000,
+    cores = 4,
+    chains = 4,
+    seed = 888
+){
+    # set parameters
+    options(mc.cores = cores, chains = chains, iter = iter, seed = seed)
+    
+    # fit models
+    fit_0 <- brm(
+        model_formulas$f_0,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior[c(1, 3),],
+        save_pars = save_pars(all = TRUE), # for computing LOO
+        backend = "cmdstanr", # faster compilation times
+        seed = seed, # for reproducibility
+        save_model = here("scripts", "stan", "fit_responses_0.stan"),
+        file = here("results", "fit_responses_0.rds"), 
+        control = list(adapt_delta = 0.95) # for better chain convergence
+    )
+    fit_1 <- brm(
+        model_formulas$f_1,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior,
+        save_pars = save_pars(all = TRUE), # for computing LOO
+        backend = "cmdstanr", # faster compilation times
+        seed = seed, # for reproducibility
+        save_model = here("scripts", "stan", "fit_responses_1.stan"),
+        file = here("results", "fit_responses_1.rds"), 
+        control = list(adapt_delta = 0.95) # for better chain convergence
+    )
+    fit_2 <- brm(
+        model_formulas$f_2,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior,
+        save_pars = save_pars(all = TRUE), # for computing LOO
+        backend = "cmdstanr", # faster compilation times
+        seed = seed, # for reproducibility
+        save_model = here("scripts", "stan", "fit_responses_2.stan"),
+        file = here("results", "fit_responses_2.rds"), 
+        control = list(adapt_delta = 0.95) # for better chain convergence
+    )
+    fit_3 <- brm(
+        model_formulas$f_3,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior,
+        save_pars = save_pars(all = TRUE), # for computing LOO
+        backend = "cmdstanr", # faster compilation times
+        seed = seed, # for reproducibility
+        save_model = here("scripts", "stan", "fit_responses_3.stan"),
[…]

### 2021-12-20 scripts/R/utils.R

@@ -1,0 +1,188 @@
+#### utils: helper functions ---------------------------------------------------
+
+# evaluate if x is NOT an element of y
+"%!in%" <- function(x, y) !(x %in% y)
+
+# find first/last non-missing value in vector
+first_non_na <- function(x) ifelse(is.logical(first(x[!is.na(x)])), NA, first(x[!is.na(x)]))
+
+last_non_na <- function(x) ifelse(is.logical(last(x[!is.na(x)])), NA, last(x[!is.na(x)]))
+
+add_big_mark <- function(x) format(x, big.mark = ",", scientific = FALSE)
+
+
+# text input responses are messy: we need to homogenise them
+clean_input_text <- function(x) {
+    {{ x }} %>% 
+        str_to_sentence() %>%
+        str_replace("Lshift|space|minus", "") %>%
+        as.character() %>% 
+        as_tibble(x = .) %>% 
+        mutate(. , value = case_when( 
+            nchar(value) > 12 ~ "Several",
+            value %in% c("Francés", "Frances", "Basic french") ~ "French",
+            value %in% c("Italiano", "Italino") ~ "Italian",
+            value %in% c("Alemán") ~ "German",
+            value %in% c("NrLondon") ~ "London",
+            value %in% c("Rnorthampton") ~ "London",
+            value %in% c("Newcastleminusuponminustyne") ~ "Newcastle upon Tyne",
+            value %in% c("Newcastlespaceuk") ~ "Newcastle",
+            value %in% c("Mã\u0081laga") ~ "Málaga",
+            value %in% c("Miltonkeynes") ~ "Milton Keynes",
+            value %in% c("Irthlingboroug") ~ "Irthlingborough",
+            value %in% c("spalding", "berkshire", "bedfordshire") ~ str_to_sentence(value),
+            value %in% c("No", "Cuál", "Space") ~ NA_character_,
+            TRUE ~ value)) %>% 
+        pull(value)
+}
+
+# ggplot custom theme
+theme_custom <- function(){
+    theme(
+        title = element_text(size = 8, face = "bold"),
+        
+        legend.background = element_rect(fill = NA, colour = NA),
+        legend.key = element_rect(fill = "white", colour = NA),
+        legend.justification = "right",
+        legend.direction = "horizontal",
+        
+        axis.line = element_line(colour = "black", size = 0.5),
[…]

### 2021-12-22 manuscript/manuscript.Rmd
Fix participant group

@@
 Outcomes of the leave-one-out cross-validation procedure.
-Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information criterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
 Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
@@
   ) %>% 
-  gtsave(here("Figures", "loos.png"))
+  gtsave(here("img", "loos.png"))
 
@@
 The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data].
-Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_Intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_Intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
+Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
 Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants.

### 2021-12-22 scripts/R/01_responses.R

@@ -24,7 +24,8 @@ get_responses_processed <- function(
         clean_names() %>% 
         # select relevant variables and rename if necessary
         select(
-            participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
+            participant, test_language, trial_id, word, soundfile, key_pressed,
+            input_text, key_press_time, error, age, date, city,
             matches("language_l|language_s|language_c|demo_|setup_"),
             -matches("_rt|_time"),
             key_press_time
@@ -35,7 +36,8 @@ get_responses_processed <- function(
         group_by(participant) %>% 
         # for each participant, get first row of each demo variable
         mutate_at(
-            vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), age, city, setup_location, setup_noise),
+            vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), 
+                 age, city, setup_location, setup_noise),
             first_non_na # defined in R/utils.R
         ) %>% 
         mutate(date = max(date, na.rm = TRUE)) %>% 

### 2022-01-14 docs/Old/00_outline.Rmd
update figures

---
title: "Outline"
date: "Updated: `r format(Sys.Date(), '%d/%m%/%Y')`"
output:
    pdf_document:
        latex_engine: xelatex
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The cognate facilitation effect

Previous studies have observed a cognate facilitation effect in bilinguals: comprehension, production, and translation of cognates (form similar translation equivalents) is faster than that of non-cognates.

## Measuring form similarity is non-trivial

There are multiple ways of measuring the similarity between two words across multiple modalities (e.g., orthographic, phonological, signed), but all fail, to different extent, to capture the listeners' subjective experience when assessing the similarity between two presented words.

This is partly due to the fact that perceived similarity is a function of variables not considered in the objective measure, or that human listeners weight such variables differently than the objective measure.

For example, one could count the number of characters/letters shared by two words, and use the result as a measure of form similarity between both. However, this measure would quite frequently likely underestimate the similarity perceived by participants (),

one could measure the orthographic similarity between two written word forms using the Levenshtein distance, a classic metric of edit distance that counts the number of additions, deletions, and substitutions one of the words has to go through to be identical to the other word.

* Although this effect has been reported across different modalities (e.g., orthographic, auditory, signed), th
* What mechanisms underlie this effect, and how the lexical representation of cognates and non-cognates might differ is still a hotly debated topic: the phonology

* Comprehension of speech in an unfamiliar language is not necessarily impossible. Words that share a large phonological similarity with their translation in the native language are easier to translate.

## The cognate facilitation effect

### Comprehension

### Production

### Translation

### Phonological neighbourhoods

### Can non-native word-forms

### Limitations of current measures of cognateness

* Levenshtein distance
* Coltheart N

## The present study

### 2022-01-14 docs/report.Rmd
update figures

--- 2021-11-25 Rmd/report.Rmd
@@
 
-* Aged 18 to 25 years
+* Aged 18 to 26 years
 * Did not report being fluent in Catalan, Spanish, or Italian
@@
 stimuli_coded %>% 
-  count(test_language, pthn) %>% 
-  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
-  geom_col(position = position_dodge()) +
-  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
+  count(group, pthn) %>% 
+  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = group)) +
+  facet_wrap(~group, ncol = 1) +
+  geom_col() +
+  labs(x = "number of phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
   theme(

### 2022-01-14 manuscript/manuscript.Rmd
update figures

--- 2021-12-16 Rmd/manuscript.Rmd
@@
 abstract: |
-  When listening to non-native speech, words whose translation in the native language is phonologically similar are easier to recognise.
-Previous studies suggest that low-proficiency bilinguals benefit more from this facilitation effect than high-proficiency bilinguals, but this claim has remained relatively unaddressed.
-In this study, we tested English and Spanish monolinguals in a translation elicitation task, in which they had to guess the translation of a list of words in an unfamiliar language (Spanish or Catalan).
-Presented words varied in their phonological similarity to their correct translation.
-This allowed us to test the impact of such phonological similarity on word recognition in a sample of listeners with very low proficiency in a non-native language.
-Translation accuracy increased substantially the likelihood of correct translations, but this effect was mediated by the size of the phonological neighbourhood of the correct translation.
-Participants benefitted from phonological similarity only when the correct translation was part of a sparse phonological neighbourhood.
-When the correct translation was part of a larger phonological neighbourhood, translation accuracy remained unaffected.
-Overall, these results provide evidence that litening to non-native words activated lexical representations in the native language, even in monolingual participants, resulting in competition for selection in the native language during translation.
 
@@
 Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
+Trials in which participants took longer than 10 seconds to respond were also filtered out.
 Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish).
@@
 
+```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.cap="Distribution and summary statistics of stimuli properties.
+A\\) Lexical frequency of target words, expressed in counts per million and as Zipf scores.
+B\\) Number of higher-frequency phonological neighbourhoods of target words (PTHN).
+C\\) Normalised Levenshtein phonological similarity between the presented and the target words, expressed in percentages."}
+
+medians <- stimuli %>% 
+  group_by(group) %>% 
+  summarise(
+    pthn_median = median(pthn),
+    pthn_sd = sd(pthn),
+    frequency_zipf_median = median(frequency_zipf),
+    frequency_zipf_sd = sd(frequency_zipf),
+    lv_median = median(lv),
+    lv_sd = sd(lv)
+  )
+
+stimuli %>% 
+  count(group, pthn) %>% 
+  ggplot(aes(pthn, n, fill = group)) + 
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0, xmax = pthn_median+pthn_sd, 
+      ymin = -Inf, ymax = Inf, fill = group
+    ), inherit.aes = FALSE, alpha = 0.15
+  ) +   
+  geom_col(colour = "white") + 
+  geom_vline(
+    data = medians, aes(xintercept = pthn_median, group = group), 
+    colour = "black", size = 0.5, linetype = "dashed"
+  ) + 
+  geom_text(data = medians, aes(label = "Median ± SD", x = pthn_median+10, y = 40), size = 3) +
+  geom_curve(
+    data = medians, aes(x = pthn_median+10, xend = pthn_median+0.2, y = 35, yend = 35),
+    arrow = arrow(length = unit(0.2, "cm")), curvature = -0.25, colour = "grey30") +
+  labs(x = "PTHN", y = "Number of trials", title = "Higher-frequency neighbours") +
+
+  ggplot(stimuli, aes(frequency_zipf, fill = group)) + 
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd, 
+      ymin = -Inf, ymax = Inf, fill = group
+    ), inherit.aes = FALSE, alpha = 0.15
+  ) +   
+  geom_histogram(bins = 20, colour = "white") +
+  geom_vline(
+    data = medians, aes(xintercept = frequency_zipf_median, group = group), 
+    colour = "black", size = 0.5, linetype = "dashed"
+  ) + 
+  labs(x = "Lexical frequency (Zipf)", y = "Number of trials", title = "Frequency") +
+
+  ggplot(stimuli, aes(lv, fill = group)) +
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0, xmax = lv_median+lv_sd, 
+      ymin = -Inf, ymax = Inf, fill = group
+    ), inherit.aes = FALSE, alpha = 0.15
+  ) +   
+  geom_histogram(bins = 20, colour = "white") +
+  geom_vline(
+    data = medians, aes(xintercept = lv_median, group = group), 
+    colour = "black", size = 0.5, linetype = "dashed"
+  ) + 
+  scale_x_continuous(labels = percent) +
+  labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity") +
+
+  plot_layout(guides = "collect") &
+  plot_annotation(tag_levels = "A") &
+  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
+  theme(legend.position = "none")
+```
+
 We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist].
@@
 This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is a score that ranges from 0 to 1, where 0 indicates no similarity between the two strings and one indicates that both strings are identical.
+The result is percentage score indicating no similarity between the two strings if 0%, and a perfect match between two strings if 100%.
 We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
-Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
-
-```{r stimuli-correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=4, fig.cap="Marginal distribution and pairwise correlation of lexical frequency, PTHN, and Levenshtein similarity.
-Statistics are indicated for the three experimental conditions (ENG-SPA, ENG-CAT, and SPA-CAT) with different colours.
-The upper triangle of the plot matrix shows the Pearson correlation between each pair of variables (asterisks indicate statistical significance at .05, .01, and .001 significance thresholds).
-The diagonal of the plot matrix indicates the marginal distribution of eacch variable.
-The lower triangle of the plot matrix representes graphically the association between each pair of variables, together with a linear regression line."}
-my_bin <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_density(..., size = 1, alpha = 0.25) +
-    geom_rug(...)
-}
-
-my_smooth <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_point(..., shape = 1, stroke = 1, alpha = 0.25, size = 0.5) +
-    geom_smooth(..., method = "lm", formula = y ~ x, alpha = 0.25, size = 1)
-}
-
-my_cor <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    ggally_cor(..., size = 7, title = "r", )
-}
-
-stimuli %>% 
-  drop_na(frequency_zipf, pthn, lv) %>% 
-  ggpairs(
-    mapping = aes(colour = group, fill = group),
-    columns = c("frequency_zipf", "pthn", "lv"),
-    columnLabels = c("Frequency (Zipf)", "PTHN", "Levenshtein"),
-    lower = list(continuous = my_smooth),
-    diag = list(continuous = my_bin),
-    upper = list("my_cor")
-  ) +
-  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
-  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE)
-
-```
+Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
 
@@
 
-```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tb_stimuli <- stimuli %>% 
-  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
-  group_by(group) %>% 
-  summarise_at(
-    vars(frequency:lv), 
-    list(
-      mean = ~mean(., na.rm = TRUE),  
-      sd = ~sd(., na.rm = TRUE)
-    )
-  ) %>%
-  gt(rowname_col = "group") %>% 
-  tab_header(
-    title = md("Table 1"),
-    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
-  ) %>% 
-  tab_source_note(
-    md("*Note*.
-Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
-  ) %>% 
-  fmt_number(frequency_mean:pthn_sd) %>%
-  fmt_percent(matches("lv")) %>% 
-  tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
-  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
-  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
-  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
-  cols_label(
-    frequency_mean = md("Mean"),
-    frequency_sd = md("*SD*"),
-    frequency_zipf_mean = md("Mean"),
-    frequency_zipf_sd = md("*SD*"),
-    pthn_mean = md("Mean"),
-    pthn_sd = md("*SD*"),
-    lv_mean = md("Mean"),
-    lv_sd = md("*SD*")
-  ) %>% 
-  summary_rows(
-    columns = c(matches("mean"), -matches("lv")),
-    fns = list(
-      Mean = "mean",
-      SD = "sd"
-    )
-  ) %>% 
-  summary_rows(
-    columns = matches("lv"),
-    fns = list(
-      Mean = "mean",
-      SD = "sd"
-    ), formatter = fmt_percent
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", weight = "bold", size = "medium"),
-    locations = cells_title(groups = "title")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium", style = "italic"),
-    locations = cells_title(groups = "subtitle")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium"),
-    locations = cells_source_notes()
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium"),
-    locations = cells_stub_grand_summary()
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = c("left", "right"), color = "white"),
-    locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
-    )
-  ) %>% 
-  tab_style(
-    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
-    locations = cells_stub()
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "all", color = "white"),
-    locations = cells_body(columns = 1:9)
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "top", color = "white"),
-    locations = cells_body(columns = 1:9)
-  ) %>% 
-  gtsave(here("Figures", "stimuli.png"))
-
-include_graphics(here("Figures", "stimuli.png"), auto_pdf = FALSE)
-
-```
-
 ## Data analysis
@@
 We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the $Levenshtein \times Group$ interaction).
-We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.5) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize].
+We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.25) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize].
 Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random].
@@
   ) %>% 
-  gtsave(here("Figures", "loos.png"))
+  gtsave(here("img", "loos.png"))
 
-include_graphics(here("Figures", "loos.png"))
+include_graphics(here("img", "loos.png"))
 
@@
   stat_pointinterval(.width = 0, point_size = 1) +
-  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
   scale_x_continuous(labels = percent, limits = c(0, 1)) +

### 2022-01-14 R/00_stimuli.R

@@ -1,76 +1,0 @@
-# relative frequencies and phonological neighbours
-get_clearpond <- function(clearpond_path){
-    clearpond <- clearpond_path %>% 
-        map(
-            function(x){
-                read.csv(x) %>% 
-                    clean_names() %>% 
-                    select("word", "pho_word", "freq_per_million", "pthn")
-            }
-        ) %>% 
-        bind_rows(.id = "group") %>% 
-        as_tibble() %>%
-        distinct(word, group, .keep_all = TRUE) %>% 
-        rename(word2 = word, phon_clearpond = pho_word, frequency = freq_per_million)
-    return(clearpond)
-}
-
-get_levenshtein <- function(stimuli_path){
-
-    levenshtein <- map(
-        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
-        function(x) read_xlsx(stimuli_path, sheet = x)
-    ) %>% 
-        bind_rows(.id = "group") %>% 
-        clean_names() %>% 
-        select(group, word1, word2, ipa1, ipa2) %>% 
-        mutate(
-            n_char = ifelse(nchar(ipa1) > nchar(ipa2), nchar(ipa1), nchar(ipa2)),
-            lv = stringsim(ipa1, ipa2, method = "lv")
-        )
-    
-    return(levenshtein)
-}
-
-get_duration <- function(
-    stimuli_path,
-    audios_path
-){
-    stimuli <- map(
-        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
-        ~read_xlsx(stimuli_path, sheet = .)) %>% 
-        bind_rows(.id = "group") %>% 
-        select(word1, group, file) %>% 
-        mutate(file = here("Stimuli", "Sounds", file))
-    
-    audios <- map(stimuli$file, load.wave) 
-    lengths <- map_dbl(audios, length)/2
-    sampling_rate <- unique(map_dbl(audios, ~attr(., "rate")))
-    duration <- lengths/sampling_rate
-    return(duration)
-}
-
-get_stimuli <- function(
-    stimuli_path,
-    clearpond,
-    levenshtein,
-    durations
-){
-    
-    stimuli <- map(
-        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
[…]

### 2022-01-14 R/01_responses.R

@@ -1,199 +1,0 @@
-# preprocess responses
-get_responses_processed <- function(
-    responses_path,
-    stimuli,
-    practice_trials,
-    seed = 888
-){
-    # set parameters
-    set.seed(seed) # for reproducibility
-    spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
-    
-    filenames <- list.files(here("Data", "Raw")) %>% 
-        str_extract(".*?\\_") %>% 
-        str_remove("_")
-    
-    # participant files
-    processed <- responses_path %>% 
-        map(fread, na.string = "") %>%  # import participant files
-        map(mutate, participant = as.character(participant)) %>% 
-        set_names(filenames) %>% # label each dataset with the participant's ID
-        bind_rows(.id = "filename") %>% # merge datasets
-        as_tibble() %>%
-        clean_names() %>% 
-        # select relevant variables and rename if necessary
-        select(
-            participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
-            matches("language_l|language_s|language_c|demo_|setup_"),
-            -matches("_rt|_time"),
-            key_press_time
-        ) %>% 
-        rename(demo_impairment = demo_language_key_keys) %>% 
-        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
-        # clean text input by participants (because of typos of need to translate) and redefine location    
-        group_by(participant) %>% 
-        mutate_at(
-            vars(
-                starts_with("demo_"),
-                starts_with("l"),
-                starts_with("setup_"),
-                age, city, setup_location, setup_noise
-            ),
-            first_non_na) %>% 
-        mutate(date = max(date, na.rm = TRUE)) %>% 
-        ungroup() %>% 
-        mutate_at(vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), clean_input_text) %>% 
-        mutate(
-            city = str_to_sentence(city),
-            country = ifelse(city %in% spanish_cities, "Spain", "UK"),
-            date = as_datetime(ymd_hms(date)),
-            demo_sex = case_when(
-                country %in% "UK" & demo_sex %in% "M" ~ "Male",
-                country %in% "UK" & demo_sex %in% "F" ~ "Female",
-                country %in% "Spain" & demo_sex %in% "M" ~ "Female",
[…]

### 2022-01-14 _targets.R

@@ -1,10 +1,10 @@
 library(targets)
 library(tarchetypes)
 
-source("R/utils.R")
-source("R/00_stimuli.R")
-source("R/01_responses.R")
-source("R/02_models.R")
+source("scripts/R/utils.R")
+source("scripts/R/00_stimuli.R")
+source("scripts/R/01_responses.R")
+source("scripts/R/02_models.R")
 
 # set parameters
 tar_option_set(
@@ -19,20 +19,20 @@ tar_option_set(
 list(
     
     # stimuli
-    tar_target(clearpond_path, list(`ENG-CAT` = here("Data", "clearpond", "clearpond_english.csv"), `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"), `SPA-CAT` = here("Data", "clearpond", "clearpond_spanish.csv")),),
+    tar_target(clearpond_path, list(`ENG-CAT` = here("data", "clearpond", "clearpond_english.csv"), `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"), `SPA-CAT` = here("Data", "clearpond", "clearpond_spanish.csv")),),
     tar_target(clearpond, get_clearpond(clearpond_path)),
-    tar_target(stimuli_path, here("Stimuli", "trials.xlsx")),
+    tar_target(stimuli_path, here("stimuli", "trials.xlsx")),
     tar_target(levenshtein, get_levenshtein(stimuli_path = stimuli_path)),
-    tar_target(audios_path, here("Stimuli/Sounds")),
+    tar_target(audios_path, here("stimuli", "sounds")),
     tar_target(durations, get_duration(stimuli_path = stimuli_path, audios_path = audios_path)),
     tar_target(stimuli, get_stimuli(stimuli_path = stimuli_path, clearpond = clearpond, levenshtein = levenshtein, durations = durations)),
     tar_target(practice_trials, c(109, 147, 159, 167, 179, 1, 26, 70, 86, 96)),
     
     # responses
-    tar_target(responses_path, list.files(here("Data", "Raw"), full.names = TRUE)),
+    tar_target(responses_path, list.files(here("data", "raw"), full.names = TRUE)),
     tar_target(responses_processed, get_responses_processed(responses_path = responses_path, stimuli = stimuli, practice_trials = practice_trials)),
     tar_target(responses_clean, get_responses_clean(responses_processed = responses_processed, stimuli = stimuli)),
-    tar_target(responses_coded_path, here("Data", "02_coded.xlsx")),
[…]

### 2022-01-14 scripts/R/00_stimuli.R

@@ -1,0 +1,86 @@
+# relative frequencies and phonological neighbours
+get_clearpond <- function(clearpond_path){
+    clearpond <- clearpond_path %>% 
+        map(
+            function(x){
+                read.csv(x) %>% 
+                    clean_names() %>% 
+                    select("word", "pho_word", "freq_per_million", "pthn")
+            }
+        ) %>% 
+        bind_rows(.id = "group") %>% 
+        as_tibble() %>%
+        distinct(word, group, .keep_all = TRUE) %>% 
+        rename(word2 = word, phon_clearpond = pho_word, frequency = freq_per_million)
+    return(clearpond)
+}
+
+# get Leveshtein similarity scores
+get_levenshtein <- function(stimuli_path){
+    
+    levenshtein <- map(
+        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
+        function(x) read_xlsx(stimuli_path, sheet = x)
+    ) %>% 
+        bind_rows(.id = "group") %>% 
+        clean_names() %>% 
+        select(group, word1, word2, ipa1, ipa2) %>% 
+        mutate(
+            n_char = ifelse(nchar(ipa1) > nchar(ipa2), nchar(ipa1), nchar(ipa2)),
+            lv = stringsim(ipa1, ipa2, method = "lv")
+        )
+    
+    return(levenshtein)
+}
+
+# get audio duration
+get_duration <- function(
+    stimuli_path,
+    audios_path
+){
+    stimuli <- map(
+        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
+        ~read_xlsx(stimuli_path, sheet = .)) %>% 
+        bind_rows(.id = "group") %>% 
+        select(word1, group, file) %>% 
+        mutate(file = here("stimuli", "sounds", file))
+    
+    audios <- map(stimuli$file, load.wave) 
+    lengths <- map_dbl(audios, length)/2
+    sampling_rate <- unique(map_dbl(audios, ~attr(., "rate")))
+    duration <- lengths/sampling_rate
+    return(duration)
+}
+
+# process stimuli and add information
+get_stimuli <- function(
+    stimuli_path,
+    clearpond,
+    levenshtein,
+    durations
+){
+    
+    stimuli <- map(
[…]

### 2022-01-14 scripts/R/02_models.R

@@ -1,0 +1,197 @@
+# models
+
+get_model_fits <- function(
+    responses,
+    model_formulas,
+    model_prior,
+    iter = 500,
+    cores = 4,
+    chains = 4,
+    seed = 888
+){
+    # set parameters
+    options(mc.cores = cores, chains = chains, iter = iter, seed = seed)
+    
+    # fit models
+    fit_0 <- brm(
+        model_formulas$f_0,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior[c(1, 3),],
+        save_pars = save_pars(all = TRUE),
+        backend = "cmdstanr",
+        seed = 888,
+        save_model = here("scripts", "stan", "fit_responses_0.stan"),
+        file = here("results", "fit_responses_0.rds"), 
+        control = list(adapt_delta = 0.95)
+    )
+    fit_1 <- brm(
+        model_formulas$f_1,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior,
+        save_pars = save_pars(all = TRUE),
+        backend = "cmdstanr",
+        seed = 888,
+        save_model = here("scripts", "stan", "fit_responses_1.stan"),
+        file = here("results", "fit_responses_1.rds"), 
+        control = list(adapt_delta = 0.95)
+    )
+    fit_2 <- brm(
+        model_formulas$f_2,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior,
+        save_pars = save_pars(all = TRUE),
+        backend = "cmdstanr",
+        seed = 888,
+        save_model = here("scripts", "stan", "fit_responses_2.stan"),
+        file = here("results", "fit_responses_2.rds"), 
+        control = list(adapt_delta = 0.95)
+    )
+    fit_3 <- brm(
+        model_formulas$f_3,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior,
+        save_pars = save_pars(all = TRUE),
+        backend = "cmdstanr",
+        sample_prior = "yes",
+        seed = 888,
+        save_model = here("scripts", "stan", "fit_responses_3.stan"),
+        file = here("results", "fit_responses_3.rds"), 
+        control = list(adapt_delta = 0.95)
+    )
+    fit_4 <- brm(
+        model_formulas$f_4,
+        data = responses,
+        family = bernoulli("logit"),
+        prior = model_prior,
+        save_pars = save_pars(all = TRUE),
[…]

### 2022-01-14 scripts/R/utils.R

@@ -1,0 +1,169 @@
+#### utils: helper functions ---------------------------------------------------
+
+# evaluate if x is NOT an element of y
+"%!in%" <- function(x, y) !(x %in% y)
+
+# find first/last non-missing value in vector
+first_non_na <- function(x) ifelse(is.logical(first(x[!is.na(x)])), NA, first(x[!is.na(x)]))
+last_non_na <- function(x) ifelse(is.logical(last(x[!is.na(x)])), NA, last(x[!is.na(x)]))
+add_big_mark <- function(x) format(x, big.mark = ",", scientific = FALSE)
+clean_input_text <- function(x) {
+    {{ x }} %>% 
+        str_to_sentence() %>%
+        str_replace("Lshift|space|minus", "") %>%
+        as.character() %>% 
+        as_tibble(x = .) %>% 
+        mutate(. , value = case_when( 
+            nchar(value) > 12 ~ "Several",
+            value %in% c("Francés", "Frances", "Basic french") ~ "French",
+            value %in% c("Italiano", "Italino") ~ "Italian",
+            value %in% c("Alemán") ~ "German",
+            value %in% c("NrLondon") ~ "London",
+            value %in% c("Rnorthampton") ~ "London",
+            value %in% c("Newcastleminusuponminustyne") ~ "Newcastle upon Tyne",
+            value %in% c("Newcastlespaceuk") ~ "Newcastle",
+            value %in% c("Miltonkeynes") ~ "Milton Keynes",
+            value %in% c("Irthlingboroug") ~ "Irthlingborough",
+            value %in% c("spalding", "berkshire", "bedfordshire") ~ str_to_sentence(value),
+            value %in% c("No", "Cuál", "Space") ~ NA_character_,
+            TRUE ~ value)) %>% 
+        pull(value)
+}
+# ggplot custom theme
+theme_custom <- function(){
+    theme(
+        panel.background = element_rect(fill = "transparent"),
+        panel.grid = element_blank(),
+        #panel.grid = element_line(colour = "grey", linetype = "dotted"),
+        panel.border = element_rect(fill = "transparent", colour = "black"),
+        text = element_text(colour = "black", size = 15),
+        axis.text = element_text(colour = "black")
+    )
+}
+
+# remove special characters
+replace_non_ascii <- function(x){
[…]

### 2022-01-14 docs/00_stimuli.Rmd

@@
 
+```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+tb_stimuli <- stimuli %>% 
+  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
+  group_by(group) %>% 
+  summarise_at(
+    vars(frequency:lv), 
+    list(
+      mean = ~mean(., na.rm = TRUE),  
+      sd = ~sd(., na.rm = TRUE)
+    )
+  ) %>%
+  gt(rowname_col = "group") %>% 
+  tab_header(
+    title = md("Table 1"),
+    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
+  ) %>% 
+  fmt_number(frequency_mean:pthn_sd) %>%
+  fmt_percent(matches("lv")) %>% 
+  tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
+  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
+  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
+  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
+  cols_label(
+    frequency_mean = md("Mean"),
+    frequency_sd = md("*SD*"),
+    frequency_zipf_mean = md("Mean"),
+    frequency_zipf_sd = md("*SD*"),
+    pthn_mean = md("Mean"),
+    pthn_sd = md("*SD*"),
+    lv_mean = md("Mean"),
+    lv_sd = md("*SD*")
+  ) %>% 
+  summary_rows(
+    columns = c(matches("mean"), -matches("lv")),
+    fns = list(
+      Mean = "mean",
+      SD = "sd"
+    )
+  ) %>% 
+  summary_rows(
+    columns = matches("lv"),
+    fns = list(
+      Mean = "mean",
+      SD = "sd"
+    ), formatter = fmt_percent
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium", style = "italic"),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_stub_grand_summary()
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = c("left", "right"), color = "white"),
+    locations = list(
+      cells_body(), cells_row_groups(), cells_column_spanners(), 
+      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
+      cells_column_labels(), cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
+    locations = cells_stub()
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+    locations = list(cells_column_spanners(), cells_column_labels())
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "all", color = "white"),
+    locations = cells_body(columns = 1:9)
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "top", color = "white"),
+    locations = cells_body(columns = 1:9)
+  ) %>% 
+  gtsave(here("img", "stimuli.png"))
+
+include_graphics(here("img", "stimuli.png"), auto_pdf = FALSE)
+
+```
+
 # Lexical frequency
@@
 ```
+
+# Correlations
+
+```{r stimuli-correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=4, fig.cap="Marginal distribution and pairwise correlation of lexical frequency, PTHN, and Levenshtein similarity.
+Statistics are indicated for the three experimental conditions (ENG-SPA, ENG-CAT, and SPA-CAT) with different colours.
+The upper triangle of the plot matrix shows the Pearson correlation between each pair of variables (asterisks indicate statistical significance at .05, .01, and .001 significance thresholds).
+The diagonal of the plot matrix indicates the marginal distribution of eacch variable.
+The lower triangle of the plot matrix representes graphically the association between each pair of variables, together with a linear regression line."}
+my_bin <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_density(..., size = 1, alpha = 0.25) +
+    geom_rug(...)
+}
+
+my_smooth <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_point(..., shape = 1, stroke = 1, alpha = 0.25, size = 0.5) +
+    geom_smooth(..., method = "lm", formula = y ~ x, alpha = 0.25, size = 1)
+}
+
+my_cor <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    ggally_cor(..., size = 7, title = "r", )
+}
+
+stimuli %>% 
+  drop_na(frequency_zipf, pthn, lv) %>% 
+  ggpairs(
+    mapping = aes(colour = group, fill = group),
+    columns = c("frequency_zipf", "pthn", "lv"),
+    columnLabels = c("Frequency (Zipf)", "PTHN", "Levenshtein"),
+    lower = list(continuous = my_smooth),
+    diag = list(continuous = my_bin),
+    upper = list("my_cor")
+  ) +
+  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
+  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE)
+
+```

### 2022-01-14 docs/01_participants.Rmd

---
title: "Participants"
output:
  md_document:
    variant: markdown_github
  html_document:
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true
bibliography: "references.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = TRUE,
  warning = TRUE, 
  cache.extra = knitr::rand_seed,
  out.width = "80%",
  results = "asis",
  dpi = 300,
  dev.args = list(png = list(type = "cairo"))
)

options(
  knitr.kable.NA = '-',
  knitr.duplicate.label = "allow",
  ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
  ggplot2.continuous.fill = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  ),
  ggplot2.continuous.colour = scale_fill_gradient2(
    low = "#04bf78", mid = "white", high = "#FFC20A", 
    na.value = "white", limits = c(-0.4, 0.4)
  )
)
```

```{r prepare}
# import data
tar_load_globals()
tar_load(participants)
tar_load(responses)

theme_set(theme_custom())
```

```{r participants_coded}
participants_coded <- participants %>% 
  mutate(valid_participant = ifelse(valid_participant, "Valid", "Not valid"))
```

We collected data from `r format(as_date(min(participants$date)), "%B %dth %Y")` to `r format(as_date(max(participants$date)), "%B %dth %Y")`. We tested `r length(unique(participants$participant))` participants. `r length(unique(participants$participant[participants$group=="ENG-SPA"]))` were English natives tested in Spanish, `r length(unique(participants$participant[participants$group=="ENG-CAT"]))` were English natives tested in Catalan, and `r length(unique(participants$participant[participants$group=="SPA-CAT"]))` were Spanish natives tested in Catalan. `r sum(!participants$valid_participant)` failed to meet all the inclusion criteria and were excluded from further analyses. The final sample comprised data from `r sum(!participants$valid_participant)` participants, `r sum(participants$valid_participant[participants$group=="ENG-SPA"])` in the ENG-SPA group, `r sum(participants$valid_participant[participants$group=="ENG-CAT"])` in the ENG-CAT group, and `r sum(participants$valid_participant[participants$group=="SPA-CAT"])` in the SPA-CAT group.

Participants were included if^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct. This item was phrased as Do you have normal or corrected-to-normal VISION? (Yes/No) in English, and as ¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan? (Sí/No). Surprisingly, the proportion of Spanish participants that reported visual impairment was implausibly large (n = 6, 18.18%). This is possibly due to some participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision]:

* Aged 18 to 25 years
* Did not report being fluent in Catalan, Spanish, or Italian
* Provided at least 80% of the presented trials
* Did not report motor, auditory or visual (other than wearing glasses) problems

Trials were considered valid if:

* Participant did not leave a blank response
* Participant took longer than 10 seconds to respond

# Summary

```{r participants_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_coded %>% 
  count(group, valid_participant) %>% 
  mutate(valid_participant = factor(valid_participant)) %>% 
  ggplot(aes(group, n, fill = valid_participant, order = n)) +
  geom_col() +
  labs(x = "Group", y = "# participants", fill = "Valid?") +
  theme(
    legend.title = element_blank(),
    legend.position = "top",
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
  )

```

# Valid trials

```{r participants_stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_coded %>% 
  filter(valid_participant=="Valid") %>% 
  select(participant, group, n, n_valid) %>%
  mutate(valid_prop = n_valid/n) %>% 
  ggplot(aes(valid_prop, fill = group, colour = group)) +
  facet_wrap(~group, ncol = 1) +
  geom_histogram() +
  geom_rug() +
  labs(x = "Valid trials (%)", y = "N participants", fill = "Group", colour = "Group") +
  scale_x_continuous(limits = c(0, 1), labels = percent) +
  theme(
    legend.title = element_blank(),
    legend.position = "none",
    axis.title.x = element_blank(),
    panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
  )
```

# Demographic information

```{r participants_demo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants_valid <- filter(participants_coded, valid_participant=="Valid")

p_sex <- participants_valid %>% 
  left_join(count(participants_valid, group, name = "n_total")) %>% 
  mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
  ggplot(aes(group, fill = sex)) +
  geom_bar(position = position_fill()) +
  labs(x = "Group", y = "# participants", fill = "Sex") +
  scale_y_continuous(labels = percent) +
  theme(
    legend.position = "top",
    axis.title.x = element_blank()
  )

p_age <- participants_valid %>% 
  left_join(count(participants_valid, group, name = "n_total")) %>% 
  mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
  ggplot(aes(as.factor(age), fill = group, colour = group)) +
  facet_wrap(~group, ncol = 1) +
  geom_bar() +
  labs(x = "Age (years)", y = "% participants", fill = "Group", colour = "Group") +
  theme(
    legend.position = "none"
  )

(p_sex + p_age) &
  plot_layout() &
  plot_annotation(tag_levels = "A") &
  theme(
    legend.title = element_blank(),
    panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
  )

```

# Second language

```{r participants_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_valid %>% 
  count(group, l2) %>% 
  mutate(
    l2 = case_when(
      l2=="Alemã\u0081n" ~ "German",
      l2=="Francã\u2030s" ~ "French",
      l2=="Cuã\u0081l" ~ "Missing",
      TRUE ~ l2
    ),
  ) %>% 
  left_join(count(participants_valid, group, name = "n_total")) %>% 
  mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
  ggplot(aes(group, n, fill = l2)) +
  facet_wrap(~group, ncol = 1, scales = "free_y") +
  geom_col(position = position_fill()) +
  labs(x = "L2", y = "# participants", fill = "L2") +
  scale_y_continuous(labels = percent) +
  coord_flip() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    strip.text = element_blank(),
    panel.grid = element_blank(),
    axis.text.x = element_text(angle = 90),
    axis.title.y = element_blank()
  )

```

# Language profile proficiency

```{r participants_proficiency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_valid %>% 
  select(participant, group, catalan_oral, spanish_oral, catalan_written, spanish_written) %>% 
  mutate_at(vars(catalan_oral, spanish_oral, catalan_written, spanish_written), replace_na, 0) %>% 
  pivot_longer(c(catalan_oral, spanish_oral, catalan_written, spanish_written),
               names_to = "type", values_to = "score") %>% 
  separate(type, c("language", "modality"), sep = "_") %>% 
  mutate_at(vars(language, modality), str_to_sentence) %>%
  left_join(count(participants_valid, group, name = "n_total")) %>% 
  mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
  ggplot(aes(as.integer(score), fill = language)) +
  facet_grid(modality~group) +
  geom_bar(position = position_dodge()) +
  labs(y = "# participants", x = "Self-reported proficiency (0-5)", fill = "Language") +
  theme(
    legend.position = "top",
    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
  )

```

### 2022-01-14 docs/03_results.Rmd

--- 2021-12-16 Rmd/manuscript.Rmd
@@
 ---
-title: "The role of cognateness in non-native spoken word recognition"
-
-shorttitle: Cognateness and non-native word recognition
-
-author:
-- name: Gonzalo Garcia-Castro
-  affiliation: '1'
-  corresponding: yes
-  address: Ramon Trial Fargas, 25-27, 08005 Barcelona
-  email: gonzalo.garciadecastro@upf.edu
-- name: Serene Siow
-  affiliation: '2'
-- name: Nuria Sebastian-Galles
-  affiliation: '1'
-- name: Kim Plunkett
-  affiliation: '2'
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
-  - id            : "2"
-    institution   : "Department of Experimental Psychology, University of Oxford"
-
-authornote: |
-
-abstract: |
-  When listening to non-native speech, words whose translation in the native language is phonologically similar are easier to recognise.
-Previous studies suggest that low-proficiency bilinguals benefit more from this facilitation effect than high-proficiency bilinguals, but this claim has remained relatively unaddressed.
-In this study, we tested English and Spanish monolinguals in a translation elicitation task, in which they had to guess the translation of a list of words in an unfamiliar language (Spanish or Catalan).
-Presented words varied in their phonological similarity to their correct translation.
-This allowed us to test the impact of such phonological similarity on word recognition in a sample of listeners with very low proficiency in a non-native language.
-Translation accuracy increased substantially the likelihood of correct translations, but this effect was mediated by the size of the phonological neighbourhood of the correct translation.
-Participants benefitted from phonological similarity only when the correct translation was part of a sparse phonological neighbourhood.
-When the correct translation was part of a larger phonological neighbourhood, translation accuracy remained unaffected.
-Overall, these results provide evidence that litening to non-native words activated lexical representations in the native language, even in monolingual participants, resulting in competition for selection in the native language during translation.
-
-  <!-- https://tinyurl.com/ybremelq -->
-
-keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian
-wordcount: 4553
-
-bibliography:
-  - r-references.bib
-  - references.bib
-floatsintext: yes
-figurelist: no
-tablelist: no
-footnotelist: no
-linenumbers: yes
-mask: no
-draft: no
-numbersections: yes
-csl: "apa7.csl"
-documentclass: apa7
-classoption: man
-output: papaja::apa6_docx
+title: "Results"
+output:
+  md_document:
+    variant: markdown_github
+  html_document:
+    df_print: paged
+    toc: true
+    toc_depth: 4
+    toc_float: true
+bibliography: "references.bib"
 ---
 
-```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
-set.seed(42)
+```{r setup, include=FALSE}
 knitr::opts_chunk$set(
-  message = FALSE,
-  warning = FALSE,
+  echo = FALSE,
+  message = TRUE,
+  warning = TRUE, 
+  cache.extra = knitr::rand_seed,
+  out.width = "80%",
   results = "asis",
-  dpi = 1000
+  dpi = 300,
+  dev.args = list(png = list(type = "cairo"))
 )
@@
   knitr.duplicate.label = "allow",
-  gt.html_tag_check = FALSE
+  ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
+  ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
+  ggplot2.continuous.fill = scale_fill_gradient2(
+    low = "#04bf78", mid = "white", high = "#FFC20A", 
+    na.value = "white", limits = c(-0.4, 0.4)
+  ),
+  ggplot2.continuous.colour = scale_fill_gradient2(
+    low = "#04bf78", mid = "white", high = "#FFC20A", 
+    na.value = "white", limits = c(-0.4, 0.4)
   )
-
+)
 ```
 
-```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
-# import objects
+```{r prepare}
+# import data
 tar_load_globals()
-tar_load(stimuli)
 tar_load(participants)
@@
 tar_load(posterior_draws_fixed)
+tar_load(posterior_draws_random)
 tar_load(posterior_epreds_fixed)
-tar_load(posterior_epreds_random)
 
-# set custom ggplot2 theme
-theme_set(
-  theme(
-    title = element_text(size = 8, face = "bold"),
+theme_set(theme_custom())
+```
 
-    legend.background = element_rect(fill = NA, colour = NA),
-    legend.key = element_rect(fill = "white", colour = NA),
-    legend.justification = "right",
-    legend.direction = "horizontal",
+# Model outputs {.tabset .tab-pill}
 
-    axis.line = element_line(colour = "black", size = 0.5),
-    axis.ticks = element_line(colour = "black", size = 0.5),
-    axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black"),
+```{r responses_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+responses_coded <- responses %>% 
+  # typos are considered correct responses
+  mutate_at(vars(group), as.factor) %>% 
+  # center predictors
+  mutate_at(
+    vars(lv, pthn, frequency),
+    function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
+  # impute missing data
+  mice(m = 5, print = FALSE, method = "pmm") %>% 
+  complete() %>% 
+  as_tibble() %>% 
+  arrange(group, word)
+```
 
-    plot.caption.position = "plot",
-    plot.title = element_text(hjust = 0.5),
+## Fixed coefficients
 
-    panel.grid = element_blank(),
-    panel.background = element_rect(fill = "white", colour = NA)
+```{r coefs_fixed_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+coefs <- fixef(model_fits$fit_6) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("variable") %>% 
+  clean_names() %>% 
+  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4))
+
+str_repl <- c(
+  "Intercept" = "Intercept",
+  "frequency_zipf" = "Frequency\n[+1 SD]",
+  "pthn:lv" = "PTHN \u00d7 Levenshtein", 
+  "lv:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv" = "Levenshtein\n[+1 SD]",
+  "pthn" = "PTHN\n[+1 SD]",
+  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+)
+
+coefs %>% 
+  mutate(
+    variable = str_remove(variable, "b_") %>% 
+      str_replace_all(str_repl) %>%  
+      factor(
+        levels = c(
+          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
+          "Group 1\n[ENG-* vs. SPA-CAT]",
+          "Group 2\n[ENG-CAT vs. ENG-SPA]",
+          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
         )
       )
-
-```
-
-# Introduction
-
-Listening to non-native speech is more costly than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
-One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, a Spanish native listening to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
-The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
-These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical].
-Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
-In the present study we investigate the extent to which adult listeners rely on the phonological similarity between the native and the non-native languages during word recognition, capitalising on the role of *cognates*.
-
-All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the populations of speakers of each language.
-Cognates embody a particular case of such commonalities.
-Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
-For example, *sun* and *sol* (in Spanish), share their phonological onset, but their etymology points to different origins.
-We will use the word *cognateness* to include all form-similar cross-language synonyms for simplicity.
-It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
-For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Cognates play a pivotal role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
-This cognate advantage is not restricted to production, but rather extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does], and word translation [@christoffels2006memory].
-
-Early theoretical accounts of this phenomenon pointed to the fact that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
-The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), and not to their shared concept (conceptual route).
-This word-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection).
-As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations, therefore to cognateness.
-The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
-The rationale behind this prediction was that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
-One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
-
-To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
-Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
-Results showed that translation times and accuracy were roughly equivalent in forward and backward translation.
-Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected  by semantic variables than non-cognates.
-Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
-Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
-Although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
-These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
-
-Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted the fact that lexical selection is affected by the number of phonological neighbours of the selected word.
-Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when such neighbours' lexical frequency is higher [@luce1998recognizing].
-This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation [BIA+, @van1998orthographic] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
-Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
-When bilingual participants are presented with words in only one language (target language) in a lexical decision task, translating them to the non-target language took longer when they were part of large  neighbourhoods [@van1998orthographic].
-This suggests that words presented in the target language activated orthographic neighbourhoods in the non-target language, which competed for selection with the target word during recognition.
-
-A more recent model of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
-This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
-Such model also assumes that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
-It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
-In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
-A remarkable correlation between Multilink simulations of a translation elicitation task and behavioural data on Dutch-English bilinguals [@christoffels2006memory] has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
-Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
-Overall, results from Multilink are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
-
-The fact that Multilink successfully fitted the experimental data from @christoffels2006memory, despite the absence of any word-word connections between translation equivalents, is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
-It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated, assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
-The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation, and this prediction remains untested my Multilink.
-
-In the present study we investigated the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
-Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
-In this case, monolinguals can only use the lexical route to succeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
-We hypothesised that if participants are able to translate words in L2, without knowledge of the language it belongs to, and based on their form-similarity with the target word, then the lexical route should be in place for low-proficiency bilinguals.
-In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then should low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
-
-Following @luce1998recognizing we also hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
-If this is the case, words whose translation in the native language are part of larger phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target word.
-
-# Methods
-
-All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/) and a GitHub repository [https://github.com/bilingual-project/translation-elicitation.git](https://github.com/bilingual-project/translation-elicitation.git), along with additional notes.
-
-## Participants
-
-Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`.
-We collected data from `r length(unique(participants$participant))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
-`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant[participants$group %in% c("SPA-CAT")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant[participants$group %in% c("SPA-CAT") & participants$sex=="Female"]))` female).
-Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
-Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the local ethical committee (XXXXXXXXXXXX).
-Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`) , or c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
-
-## Procedure
-
-The experiment was implemented online using Psychopy/Pavlovia [@peirce2019psychopy2].
-Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla).
-After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
-Then, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
-They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the list of Catalan or Spanish trials.
-Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
-Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio ended.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
-Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
-Participants were allowed to correct their answer.
-Then, participants pressed the RETURN key to start and new trial.
-We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish).
-The task took approximately 15 minutes to be completed.
-
-```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%", fig.cap="Schematic representation of a trial in the experimental task."}
-
-include_graphics(here("Figures", "design.png"))
-
-```
-
-## Stimuli
-
-```{r stimulilengths, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-lengths <- stimuli %>% 
-  select(group, ipa1, ipa2, word1, word2) %>% 
-  mutate_at(vars(ipa1:word2), nchar) %>% 
-  group_by(group) %>% 
-  summarise_at(vars(ipa1, ipa2, word1, word2), list(mean = mean, sd = sd, min = min, max = max)) %>%
-  split(.$group) %>% 
-  map(select, -group) %>% 
-  set_names(make_clean_names(names(.))) %>% 
-  map(unlist)
-```
-
-We arranged two lists of words: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
-Words in the Catalan list (listened to by participants) were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
-Their translations to English (typed by participants in the keyboard) were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), and their translations to Spanish were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
-Words in the Spanish list were `r printnum(lengths$eng_spa["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_spa["ipa1_sd"])`, Range = `r printnum(lengths$eng_spa["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_spa["ipa1_max"], digits = 0)`).
-Their translations to English were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
-
-We extracted the lexical frequencies from SUBTLEX-UK for English words [@van2014subtlex], SUBTLEX-ESP for Spanish words [@cuetos2011subtlex], and SUBTLEX-CAT for Catalan words [@boada2020subtlex].
-We extracted or transformed scores as/into Zipf scores [@van2014subtlex] to so correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses [@van2014subtlex].
-
-We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) from the CLEARPOND database [@marian2012clearpond].
-PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
-CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target word in only one addition, deletion, or substitution.
-PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
-
-We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist].
-This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
-First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
-This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is a score that ranges from 0 to 1, where 0 indicates no similarity between the two strings and one indicates that both strings are identical.
-We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
-Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
-
-```{r stimuli-correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=4, fig.cap="Marginal distribution and pairwise correlation of lexical frequency, PTHN, and Levenshtein similarity.
-Statistics are indicated for the three experimental conditions (ENG-SPA, ENG-CAT, and SPA-CAT) with different colours.
-The upper triangle of the plot matrix shows the Pearson correlation between each pair of variables (asterisks indicate statistical significance at .05, .01, and .001 significance thresholds).
-The diagonal of the plot matrix indicates the marginal distribution of eacch variable.
-The lower triangle of the plot matrix representes graphically the association between each pair of variables, together with a linear regression line."}
-my_bin <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_density(..., size = 1, alpha = 0.25) +
-    geom_rug(...)
-}
-
-my_smooth <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_point(..., shape = 1, stroke = 1, alpha = 0.25, size = 0.5) +
-    geom_smooth(..., method = "lm", formula = y ~ x, alpha = 0.25, size = 1)
-}
-
-my_cor <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    ggally_cor(..., size = 7, title = "r", )
-}
-
-stimuli %>% 
-  drop_na(frequency_zipf, pthn, lv) %>% 
-  ggpairs(
-    mapping = aes(colour = group, fill = group),
-    columns = c("frequency_zipf", "pthn", "lv"),
-    columnLabels = c("Frequency (Zipf)", "PTHN", "Levenshtein"),
-    lower = list(continuous = my_smooth),
-    diag = list(continuous = my_bin),
-    upper = list("my_cor")
-  ) +
-  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
-  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE)
-
-```
-
-Participants listened to one audio file in each trial.
-This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the audios was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
-The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
-
-```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tb_stimuli <- stimuli %>% 
-  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
-  group_by(group) %>% 
-  summarise_at(
-    vars(frequency:lv), 
-    list(
-      mean = ~mean(., na.rm = TRUE),  
-      sd = ~sd(., na.rm = TRUE)
-    )
   ) %>% 
-  gt(rowname_col = "group") %>% 
-  tab_header(
-    title = md("Table 1"),
-    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
+  select(-est_error) %>% 
+  gt() %>% 
+  fmt_percent(2:4) %>% 
+  cols_merge(vars("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
+  cols_label(
+    variable = md("**Predictor**"),
+    estimate = md("**Mean**"),
+    q2_5 = md("**95\\% CrI**"),
   ) %>% 
-  tab_source_note(
-    md("*Note*.
-Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
-  ) %>% 
-  fmt_number(frequency_mean:pthn_sd) %>%
-  fmt_percent(matches("lv")) %>% 
-  tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
-  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
-  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
-  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
-  cols_label(
-    frequency_mean = md("Mean"),
-    frequency_sd = md("*SD*"),
-    frequency_zipf_mean = md("Mean"),
-    frequency_zipf_sd = md("*SD*"),
-    pthn_mean = md("Mean"),
-    pthn_sd = md("*SD*"),
-    lv_mean = md("Mean"),
-    lv_sd = md("*SD*")
-  ) %>% 
-  summary_rows(
-    columns = c(matches("mean"), -matches("lv")),
-    fns = list(
-      Mean = "mean",
-      SD = "sd"
-    )
-  ) %>% 
-  summary_rows(
-    columns = matches("lv"),
-    fns = list(
-      Mean = "mean",
-      SD = "sd"
-    ), formatter = fmt_percent
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", weight = "bold", size = "medium"),
-    locations = cells_title(groups = "title")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium", style = "italic"),
-    locations = cells_title(groups = "subtitle")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium"),
-    locations = cells_source_notes()
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium"),
-    locations = cells_stub_grand_summary()
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = c("left", "right"), color = "white"),
-    locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
-    )
-  ) %>% 
-  tab_style(
-    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
-    locations = cells_stub()
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "all", color = "white"),
-    locations = cells_body(columns = 1:9)
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "top", color = "white"),
-    locations = cells_body(columns = 1:9)
-  ) %>% 
-  gtsave(here("Figures", "stimuli.png"))
-
-include_graphics(here("Figures", "stimuli.png"), auto_pdf = FALSE)
-
-```
-
-## Data analysis
-
-We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0).
-Second, we added the fixed effect of lexical frequency of the target word ($Frequency$).
-We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the $Levenshtein \times Group$ interaction).
-We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.5) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize].
-Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random].
-We will later refer to these models as Model 0 to Model 6.
-For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
-
-$$
-\begin{align}
-
-&\textbf{Likelihood}  \\
-y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
-
-&\textbf{Parameters}  \\
-
-logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
-& \beta_{1[p]} ~ Frequency_{i} ~ + \\
-& \beta_{2[p]} ~ PTHN_i ~ + \\
-& \beta_{3[p]} ~ Similarity_i ~ + \\
-& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
-
-\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
-\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
-&& \text{[participant-level coefficients]} \\ \\
-
-&\textbf{Prior}  \\
-
-\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
-\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
-\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
-
-\end{align}
-$$
-
-To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (*LOO*) [@vehtari2017practical].
-More information about the models and model comparison can be found in Appendix 2.
-All analyses were performed in R environment [@rcore2019r].
-We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
-We used the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
-
-# Results
-
-All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
-Model 6, which included all main effects and the two-way interactions between $PTHN$ and $Levenshtein$, and between $Group$ and $Levenshtein$ and $PTHN$ and consonant similarity, showed the best performance.
-
-```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tab_loos <- model_loos %>% 
-  as.data.frame() %>% 
-  rownames_to_column("model") %>% 
-  arrange(desc(elpd_loo)) %>% 
-  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
-  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
-  mutate_all(~ifelse(. == 0, NA, .)) %>% 
-  gt(rowname_col = "model") %>% 
-  tab_header(
-    title = md("Table 2"),
-    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
-  ) %>% 
-  tab_source_note(
-    md("*Note*.
-Outcomes of the leave-one-out cross-validation procedure.
-Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
-Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
-  ) %>% 
-  fmt_missing(everything(), missing_text = "-") %>% 
-  fmt_number(2:7) %>% 
-  cols_label(
-    model = md("**Model**"),
-    looic = md("***LOO<sub>IC</sub>***"),
-    se_looic = md("***SE*<sub>IC</sub>**"),
-    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
-    se_elpd_loo = md("***SE***"),
-    elpd_diff = md("***LOO<sub>diff</sub>***"),
-    se_diff = md("***SE<sub>diff</sub>***")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", weight = "bold", size = "medium"),
-    locations = cells_title(groups = "title")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium", style = "italic"),
-    locations = cells_title(groups = "subtitle")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium"),
-    locations = cells_source_notes()
-  ) %>% 
-  tab_style(
-    style = list(cell_borders(sides = c("left", "right"), color = "white"),
-                 cell_text(align = "center")),
-    locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
-    )
-  ) %>% 
-  tab_style(
-    style = list(cell_text(align = "left", size = "medium"), 
-                 cell_borders(sides = "all", color = "white")),
-    locations = list(cells_stub(), cells_source_notes())
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "all", color = "white"),
-    locations = cells_body(columns = 1:7)
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "top", color = "white"),
-    locations = cells_body(columns = 1:7)
-  ) %>% 
-  gtsave(here("Figures", "loos.png"))
-
-include_graphics(here("Figures", "loos.png"))
-
-```
-
-We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
-For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
-Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increments or decrements in probability of correct translation [@gelman2020regression].
-
-```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
+  tab_source_note("Estimated posterior distributions of coefficients in Model 6.
 A\\) Population-level effects.
@@
 Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
-Black dots represent the mean of each distribution.
-B\\) Participant\\-level coefficient variability.
-The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
-Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
-C\\) Correlation between participant\\-level effects.
-The model allowed participant-level coefficients to co\\-vary.
-This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}
+Black dots represent the mean of each distribution.") %>%
+  cols_align(align = "left") %>% 
+  as_raw_html()
+```
+
+```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 95% CrI.
+The intercept has been transformed using the inverse logit to get the average probability of correct response.
+The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
 
@@
 
-```{r bayesfactors}
-h_frequency <- abs(hypothesis(model_fits$fit_6, hypothesis = "frequency_zipf = 0")$hypothesis$Evid.Ratio)
-h_pthn <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn = 0")$hypothesis$Evid.Ratio)
-h_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv = 0")$hypothesis$Evid.Ratio)
-h_pthn_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn:lv = 0")$hypothesis$Evid.Ratio)
-h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
-h_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
-h_lv_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group1 = 0")$hypothesis$Evid.Ratio)
-h_lv_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group2 = 0")$hypothesis$Evid.Ratio)
+## Model comparison
 
-icc_conditional <- icc(model_fits$fit_6, by_group = TRUE)$ICC[2]
+Leave-one-out cross-validation.
+The more negative is `elpd_diff` (and the larger its magnitude compared to its corresponding standard error), the better it first the data compared to the null model (`~ 1 + (1 | participant)`).
+The model that fits the data the best is the extended model.
 
-sds <- gather_draws(model_fits$fit_6, `sd_.*`, regex = TRUE) %>% 
-  mean_qi() %>% 
-  mutate(
-    .value = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.value), .value/4),
-    .lower = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.lower), .lower/4),
-    .upper = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.upper), .upper/4)
+```{r model_comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+tab_loos <- model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
+  arrange(desc(elpd_loo)) %>% 
+  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
+  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  tab_header(
+    title = md("Table 2"),
+    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
   ) %>% 
-  split(.$.variable)
+  tab_source_note(
+    md("*Note*.
+Outcomes of the leave-one-out cross-validation procedure.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
+  ) %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  fmt_number(2:7) %>% 
+  cols_label(
+    model = md("**Model**"),
+    looic = md("***LOO<sub>IC</sub>***"),
+    se_looic = md("***SE*<sub>IC</sub>**"),
+    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("***LOO<sub>diff</sub>***"),
+    se_diff = md("***SE<sub>diff</sub>***")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium", style = "italic"),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = list(cell_borders(sides = c("left", "right"), color = "white"),
+                 cell_text(align = "center")),
+    locations = list(
+      cells_body(), cells_row_groups(), cells_column_spanners(), 
+      cells_stub(), cells_source_notes(),
+      cells_column_labels(), cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(cell_text(align = "left", size = "medium"), 
+                 cell_borders(sides = "all", color = "white")),
+    locations = list(cells_stub(), cells_source_notes())
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+    locations = list(cells_column_spanners(), cells_column_labels())
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "all", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "top", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  as_raw_html()
 
-names(sds) <- make_clean_names(names(sds))
-
-word_example <- ranef(model_fits$fit_6)$word[,,"Intercept"] %>% 
-  as.data.frame() %>% 
-  rownames_to_column("word") %>% 
-  arrange(-Estimate) %>% 
-  clean_names() %>% 
-  mutate_if(is.numeric, inv_logit_scaled) %>% 
-  slice(1, n()) %>% 
-  left_join(distinct(stimuli, word1, ipa1), by = c("word" = "word1")) %>% 
-  split(.$word)
-
-participant_ranef_range <- sds %>% 
-  bind_rows() %>% 
-  filter(!grepl("Intercept|word|group1", .variable)) %>% 
-  summarise(.value = range(.value)) %>% 
-  unlist()
-
-cors_range <- gather_draws(model_fits$fit_6, `cor_.*`, regex = TRUE) %>% 
-  median_qi() %>% 
-  summarise(.value = range(.value))
-
-corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
-corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)
-
-corr_data <- corr_mat %>% 
-  as_tibble() %>% 
-  clean_names() %>% 
-  mutate(term1 = corr_names) %>% 
-  relocate(term1) %>% 
-  split(.$term1)
-
-names(corr_data) <- make_clean_names(names(corr_data))
-
-h_cor_intercept_frequency <- hypothesis(model_fits$fit_6, "participant__Intercept__frequency_zipf = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_frequency_zipf_lv_group1 <- hypothesis(model_fits$fit_6, "participant__frequency_zipf__lv:group1 = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_intercept_pthn <- hypothesis(model_fits$fit_6, "participant__Intercept__pthn = 0", class = "cor")$hypothesis$Evid.Ratio
 ```
 
-Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations.
-The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
-The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
-Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
-This difference increased in `r percent(coefs$pthn_lv["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
+## Random effects
 
-Spanish natives translating Catalan words were `r percent(coefs$group1["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
-This difference increased in `r percent(coefs$lv_group1["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-English natives translating Catalan words were `r percent(abs(coefs$group2["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
-For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_group2["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+```{r post_re_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+post_re <- posterior_draws_random$participants   %>% 
+  mean_qi() %>% 
+  ungroup() %>% 
+  left_join(select(participants, participant, group)) %>% 
+  arrange(group, participant, .param) %>% 
+  mutate(.param_name = str_replace_all(.param, str_repl) %>%  
+           factor(
+             levels = c(
+               "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
+               "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
+               "Group 1\n[ENG-* vs. SPA-CAT]",
+               "Group 2\n[ENG-CAT vs. ENG-SPA]",
+               "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+             )
+           ))
 
-```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
-A) Population-level predictions: the X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 200 observations from our model: 100 simulations for words with low *PTHN* (-1 *SD*) and 100 simulations for high *PTHN* (+1 *SD*) words.
-We did this across the range of values of the Levenshtein scores.
-For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high *PTHN* words.
-We also plotted the mean of the high *PTHN* (black solid line) and low *PTHN* (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
-The dispersion of the lines indicates the uncertainty of our predictions.
-We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
+ggplot(post_re, 
+       aes(.value, sort(as.factor(participant)), 
+           xmin = .lower, xmax = .upper, colour = group)) +
+  facet_grid(group~.param_name, scales = "free") +
+  geom_errorbarh(height = 0, alpha = 0.75) +
+  geom_point(size = 0.1, shape = 4) +
+  geom_vline(xintercept = 0) +
+  labs(x = "Participant-level adjustment", y = "Participant", colour = "Group",
+       title = "By participant") +
+  theme(
+    legend.position = "none",
+    legend.title = element_blank(),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    axis.text.x = element_text(size = 9),
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank()
+  )
+```
+
+```{r post_re_word, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+post_re <- posterior_draws_random$words   %>% 
+  mean_qi() %>% 
+  ungroup() %>% 
+  arrange(word, .param) %>% 
+  mutate(.param_name = str_replace_all(.param, str_repl) %>%  
+           factor(
+             levels = c(
+               "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
+               "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
+               "Group 1\n[ENG-* vs. SPA-CAT]",
+               "Group 2\n[ENG-CAT vs. ENG-SPA]",
+               "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+             )
+           ))
+
+ggplot(post_re, 
+       aes(.value, reorder(as.factor(word), -.value), 
+           xmin = .lower, xmax = .upper, colour = "#1A85FF")) +
+  geom_errorbarh(height = 0, alpha = 0.75) +
+  geom_point(size = 0.1, shape = 4) +
+  geom_vline(xintercept = 0) +
+  labs(x = "Participant-level adjustment", y = "Participant", colour = "Group",
+       title = "By participant") +
+  theme(
+    legend.position = "none",
+    legend.title = element_blank(),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    axis.text.x = element_text(size = 9),
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank()
+  )
+```
+
+## Marginal effects
+
+```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=6}
 
@@
 
-The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data].
-Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_Intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_Intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
-Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants.
-Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
-The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]).
-This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
+# Accuracy
 
-Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions.
-The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r `round(h_cor_intercept_frequency, 2)`).
-The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf$estimate_lv_group1, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf$q2_5_lv_group1, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf$q97_5_lv_group1, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
-The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).
-The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
+## Accuracy by item
 
-# Discussion
+```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_item <- responses %>%
+  group_by(group, word) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    word_label = reorder_within(word, by = correct_prop, within = group)
+  )
 
-\newpage
+ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5, colour = "black") +
+  geom_errorbar() +
+  geom_point(size = 0.5) +
+  labs(y = "Correct responses", x = "Word", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  )
 
-# References
+```
 
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
+## Accuracy by participant
 
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup
+```{r responses_participant, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+responses_participant <- responses %>%
+  group_by(group, participant) %>%
+  summarise(
+    correct_sum = sum(correct, na.rm = TRUE),
+    correct_n = sum(!is.na(correct), na.rm = TRUE),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    correct_prop = prop_adj(correct_sum, correct_n),
+    correct_se = prop_adj_se(correct_sum, correct_n),
+    participant_label = reorder_within(participant, by = correct_prop, within = group)
+  )
 
-\newpage
+ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_hline(yintercept = 0.5) +
+  geom_errorbar(width = 0.5) +
+  geom_point(size = 1) +
+  labs(y = "Correct responses", x = "Participant", colour = "Group") +
+  coord_flip() +
+  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  theme(
+    legend.position = "none",
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 9),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
+  )
+
+```

### 2022-01-14 manuscript/manuscript.Rmd

--- 2021-12-22 manuscript/manuscript.Rmd
@@
+<<<<<<< HEAD:manuscript/manuscript.Rmd
 ---
@@
 # set custom ggplot2 theme
-theme_set(theme_custom()) # in utlis.R
+theme_set(
+  theme(
+    title = element_text(size = 8, face = "bold"),
+
+    legend.background = element_rect(fill = NA, colour = NA),
+    legend.key = element_rect(fill = "white", colour = NA),
+    legend.justification = "right",
+    legend.direction = "horizontal",
+
+    axis.line = element_line(colour = "black", size = 0.5),
+    axis.ticks = element_line(colour = "black", size = 0.5),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+
+    plot.caption.position = "plot",
+    plot.title = element_text(hjust = 0.5),
+
+    panel.grid = element_blank(),
+    panel.background = element_rect(fill = "white", colour = NA)
+  )
+)
 
@@
 
-```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-tb_stimuli <- stimuli %>% 
-  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
-  group_by(group) %>% 
-  summarise_at(
-    vars(frequency:lv), 
-    list(
-      mean = ~mean(., na.rm = TRUE),  
-      sd = ~sd(., na.rm = TRUE)
-    )
-  ) %>%
-  gt(rowname_col = "group") %>% 
-  tab_header(
-    title = md("Table 1"),
-    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
-  ) %>% 
-  tab_source_note(
-    md("*Note*.
-Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
-  ) %>% 
-  fmt_number(frequency_mean:pthn_sd) %>%
-  fmt_percent(matches("lv")) %>% 
-  tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
-  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
-  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
-  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
-  cols_label(
-    frequency_mean = md("Mean"),
-    frequency_sd = md("*SD*"),
-    frequency_zipf_mean = md("Mean"),
-    frequency_zipf_sd = md("*SD*"),
-    pthn_mean = md("Mean"),
-    pthn_sd = md("*SD*"),
-    lv_mean = md("Mean"),
-    lv_sd = md("*SD*")
-  ) %>% 
-  summary_rows(
-    columns = c(matches("mean"), -matches("lv")),
-    fns = list(
-      Mean = "mean",
-      SD = "sd"
-    )
-  ) %>% 
-  summary_rows(
-    columns = matches("lv"),
-    fns = list(
-      Mean = "mean",
-      SD = "sd"
-    ), formatter = fmt_percent
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", weight = "bold", size = "medium"),
-    locations = cells_title(groups = "title")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium", style = "italic"),
-    locations = cells_title(groups = "subtitle")
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium"),
-    locations = cells_source_notes()
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "left", size = "medium"),
-    locations = cells_stub_grand_summary()
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = c("left", "right"), color = "white"),
-    locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
-    )
-  ) %>% 
-  tab_style(
-    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
-    locations = cells_stub()
-  ) %>% 
-  tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "all", color = "white"),
-    locations = cells_body(columns = 1:9)
-  ) %>% 
-  tab_style(
-    style = cell_borders(sides = "top", color = "white"),
-    locations = cells_body(columns = 1:9)
-  ) %>% 
-  gtsave(here("img", "stimuli.png"))
-
-include_graphics(here("img", "stimuli.png"), auto_pdf = FALSE)
-
-```
-
 ## Data analysis
@@
 Outcomes of the leave-one-out cross-validation procedure.
-Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information criterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
 Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
@@
 
-```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 6.
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
 A\\) Population-level effects.
@@
   "pthn:lv" = "PTHN \u00d7 Levenshtein", 
-  "lv:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
-  "lv:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv:group1" = "Levenshtein \u00d7 Group 1",
+  "lv:group2" = "Levenshtein \u00d7 Group 2",
   "lv" = "Levenshtein\n[+1 SD]",
   "pthn" = "PTHN\n[+1 SD]",
-  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
+  "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
 )
@@
 h_pthn_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn:lv = 0")$hypothesis$Evid.Ratio)
+h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
+h_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
+h_lv_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group1 = 0")$hypothesis$Evid.Ratio)
+h_lv_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group2 = 0")$hypothesis$Evid.Ratio)
+
+icc_conditional <- icc(model_fits$fit_6, by_group = TRUE)$ICC[2]
+
+sds <- gather_draws(model_fits$fit_6, `sd_.*`, regex = TRUE) %>% 
+  mean_qi() %>% 
+  mutate(
+    .value = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.value), .value/4),
+    .lower = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.lower), .lower/4),
+    .upper = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.upper), .upper/4)
+  ) %>% 
+  split(.$.variable)
+
+names(sds) <- make_clean_names(names(sds))
+
+word_example <- ranef(model_fits$fit_6)$word[,,"Intercept"] %>% 
+  as.data.frame() %>% 
+  rownames_to_column("word") %>% 
+  arrange(-Estimate) %>% 
+  clean_names() %>% 
+  mutate_if(is.numeric, inv_logit_scaled) %>% 
+  slice(1, n()) %>% 
+  left_join(distinct(stimuli, word1, ipa1), by = c("word" = "word1")) %>% 
+  split(.$word)
+
+participant_ranef_range <- sds %>% 
+  bind_rows() %>% 
+  filter(!grepl("Intercept|word|group1", .variable)) %>% 
+  summarise(.value = range(.value)) %>% 
+  unlist()
+
+cors_range <- gather_draws(model_fits$fit_6, `cor_.*`, regex = TRUE) %>% 
+  median_qi() %>% 
+  summarise(.value = range(.value))
+
+corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)
+
+corr_data <- corr_mat %>% 
+  as_tibble() %>% 
+  clean_names() %>% 
+  mutate(term1 = corr_names) %>% 
+  relocate(term1) %>% 
+  split(.$term1)
+
+names(corr_data) <- make_clean_names(names(corr_data))
+
+h_cor_intercept_frequency <- hypothesis(model_fits$fit_6, "participant__Intercept__frequency_zipf = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_frequency_zipf_lv_group1 <- hypothesis(model_fits$fit_6, "participant__frequency_zipf__lv:group1 = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_intercept_pthn <- hypothesis(model_fits$fit_6, "participant__Intercept__pthn = 0", class = "cor")$hypothesis$Evid.Ratio
+```
+
+Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations.
+The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
+The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
+Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
+This difference increased in `r percent(coefs$pthn_lv["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
+
+Spanish natives translating Catalan words were `r percent(coefs$group1["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
+This difference increased in `r percent(coefs$lv_group1["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+English natives translating Catalan words were `r percent(abs(coefs$group2["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
+For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_group2["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+
+```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
+A) Population-level predictions: the X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 200 observations from our model: 100 simulations for words with low *PTHN* (-1 *SD*) and 100 simulations for high *PTHN* (+1 *SD*) words.
+We did this across the range of values of the Levenshtein scores.
+For each simulation, we drew a single sample from the posterior distribution of each coefficient.
+Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high *PTHN* words.
+We also plotted the mean of the high *PTHN* (black solid line) and low *PTHN* (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
+The dispersion of the lines indicates the uncertainty of our predictions.
+We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
+
+m <- posterior_epreds_fixed %>% 
+  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
+  mutate(
+    pthn = case_when(
+      pthn=="-1 SD" ~ "Low PTHN",
+      pthn=="0 SD" ~ "Mean PTHN",
+      pthn=="1 SD" ~ "High PTHN"
+    ),
+    frequency_zipf = case_when(
+      frequency_zipf=="-1 SD" ~ "Low frequency",
+      frequency_zipf=="0 SD" ~ "Mean frequency",
+      frequency_zipf=="1 SD" ~ "High frequency"
+    )
+  )
+
+ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
+  facet_wrap(~group) +
+  geom_hline(yintercept = 0.5, colour = "grey") +
+  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
+  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
+  annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
+  annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
+  annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
+  annotate("text", x = 2.15, y = 0.95, size = 2.75, label = "More similar") +
+  labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
+       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
+       title = "Population-level predictions") +
+  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
+  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  theme(
+    legend.position = "top",
+    plot.title = element_blank()
+  )
+
+```
+
+The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data].
+Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
+Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants.
+Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
+The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]).
+This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
+
+Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions.
+The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r `round(h_cor_intercept_frequency, 2)`).
+The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf$estimate_lv_group1, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf$q2_5_lv_group1, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf$q97_5_lv_group1, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
+The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).
+The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
+
+# Discussion
+
+\newpage
+
+# References
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id="refs" custom-style="Bibliography"></div>
+\endgroup
+
+\newpage
+
+=======
+---
+title: "The role of cognateness in non-native spoken word recognition"
+
+shorttitle: Cognateness and non-native word recognition
+
+author:
+- name: Gonzalo Garcia-Castro
+  affiliation: '1'
+  corresponding: yes
+  address: Ramon Trial Fargas, 25-27, 08005 Barcelona
+  email: gonzalo.garciadecastro@upf.edu
+- name: Serene Siow
+  affiliation: '2'
+- name: Nuria Sebastian-Galles
+  affiliation: '1'
+- name: Kim Plunkett
+  affiliation: '2'
+affiliation:
+  - id            : "1"
+    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
+  - id            : "2"
+    institution   : "Department of Experimental Psychology, University of Oxford"
+
+authornote: |
+
+abstract: |
+  When listening to non-native speech, words whose translation in the native language is phonologically similar are easier to recognise.
+Previous studies suggest that low-proficiency bilinguals benefit more from this facilitation effect than high-proficiency bilinguals, but this claim has remained relatively unaddressed.
+In this study, we tested English and Spanish monolinguals in a translation elicitation task, in which they had to guess the translation of a list of words in an unfamiliar language (Spanish or Catalan).
+Presented words varied in their phonological similarity to their correct translation.
+This allowed us to test the impact of such phonological similarity on word recognition in a sample of listeners with very low proficiency in a non-native language.
+Translation accuracy increased substantially the likelihood of correct translations, but this effect was mediated by the size of the phonological neighbourhood of the correct translation.
+Participants benefitted from phonological similarity only when the correct translation was part of a sparse phonological neighbourhood.
+When the correct translation was part of a larger phonological neighbourhood, translation accuracy remained unaffected.
+Overall, these results provide evidence that litening to non-native words activated lexical representations in the native language, even in monolingual participants, resulting in competition for selection in the native language during translation.
+
+  <!-- https://tinyurl.com/ybremelq -->
+
+keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian
+wordcount: 4553
+
+bibliography:
+  - r-references.bib
+  - references.bib
+floatsintext: yes
+figurelist: no
+tablelist: no
+footnotelist: no
+linenumbers: yes
+mask: no
+draft: no
+numbersections: yes
+csl: "apa7.csl"
+documentclass: apa7
+classoption: man
+output: papaja::apa6_docx
+---
+
+```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+
+set.seed(42)
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  results = "asis", 
+  dpi = 1000
+)
+
+options(
+  knitr.kable.NA = '-', 
+  knitr.duplicate.label = "allow",
+  gt.html_tag_check = FALSE
+)
+
+```
+
+```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+
+# import objects
+tar_load_globals()
+tar_load(stimuli)
+tar_load(participants)
+tar_load(responses)
+tar_load(model_fits)
+tar_load(model_loos)
+tar_load(posterior_draws_fixed)
+tar_load(posterior_epreds_fixed)
+tar_load(posterior_epreds_random)
+
+# set custom ggplot2 theme
+theme_set(theme_custom()) # in utlis.R
+
+```
+
+# Introduction
+
+Listening to non-native speech is more costly than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
+One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
+For example, a Spanish native listening to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
+The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
+These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical].
+Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
+In the present study we investigate the extent to which adult listeners rely on the phonological similarity between the native and the non-native languages during word recognition, capitalising on the role of *cognates*.
+
+All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the populations of speakers of each language.
+Cognates embody a particular case of such commonalities.
+Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
+For example, *sun* and *sol* (in Spanish), share their phonological onset, but their etymology points to different origins.
+We will use the word *cognateness* to include all form-similar cross-language synonyms for simplicity.
+It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
+For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+Cognates play a pivotal role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
+This cognate advantage is not restricted to production, but rather extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does], and word translation [@christoffels2006memory].
+
+Early theoretical accounts of this phenomenon pointed to the fact that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), and not to their shared concept (conceptual route).
+This word-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection).
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations, therefore to cognateness.
+The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
+The rationale behind this prediction was that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
+One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
+
+To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
+Results showed that translation times and accuracy were roughly equivalent in forward and backward translation.
+Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected  by semantic variables than non-cognates.
+Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
+Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
+Although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
+These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
+
+Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted the fact that lexical selection is affected by the number of phonological neighbours of the selected word.
+Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when such neighbours' lexical frequency is higher [@luce1998recognizing].
+This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+The Bilingual Interactive Activation [BIA+, @van1998orthographic] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
+Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
+When bilingual participants are presented with words in only one language (target language) in a lexical decision task, translating them to the non-target language took longer when they were part of large  neighbourhoods [@van1998orthographic].
+This suggests that words presented in the target language activated orthographic neighbourhoods in the non-target language, which competed for selection with the target word during recognition.
+
+A more recent model of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
+This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
+Such model also assumes that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
+It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
+In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
+A remarkable correlation between Multilink simulations of a translation elicitation task and behavioural data on Dutch-English bilinguals [@christoffels2006memory] has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
+Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+Overall, results from Multilink are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
+
+The fact that Multilink successfully fitted the experimental data from @christoffels2006memory, despite the absence of any word-word connections between translation equivalents, is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated, assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
+The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
+It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation, and this prediction remains untested my Multilink.
+
+In the present study we investigated the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
+Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
+In this case, monolinguals can only use the lexical route to succeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
+We hypothesised that if participants are able to translate words in L2, without knowledge of the language it belongs to, and based on their form-similarity with the target word, then the lexical route should be in place for low-proficiency bilinguals.
+In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then should low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
+
+Following @luce1998recognizing we also hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+If this is the case, words whose translation in the native language are part of larger phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target word.
+
+# Methods
+
+All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/) and a GitHub repository [https://github.com/bilingual-project/translation-elicitation.git](https://github.com/bilingual-project/translation-elicitation.git), along with additional notes.
+
+## Participants
+
+Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`.
+We collected data from `r length(unique(participants$participant))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
+`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant[participants$group %in% c("SPA-CAT")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant[participants$group %in% c("SPA-CAT") & participants$sex=="Female"]))` female).
+Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
+Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the local ethical committee (XXXXXXXXXXXX).
+Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
+We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`) , or c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
+
+## Procedure
+
+The experiment was implemented online using Psychopy/Pavlovia [@peirce2019psychopy2].
+Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla).
+After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+Then, participants completed the experimental task.
+Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
+English participants were randomly assigned to the list of Catalan or Spanish trials.
+Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
+Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
+After one second, the audio started playing while the dot remained being displayed until the audio ended.
+Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
+Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
+Participants were allowed to correct their answer.
+Then, participants pressed the RETURN key to start and new trial.
+We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
+Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
+Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish).
+The task took approximately 15 minutes to be completed.
+
+```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%", fig.cap="Schematic representation of a trial in the experimental task."}
+
+include_graphics(here("img", "design.png"))
+
+```
+
+## Stimuli
+
+```{r stimulilengths, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+lengths <- stimuli %>% 
+  select(group, ipa1, ipa2, word1, word2) %>% 
+  mutate_at(vars(ipa1:word2), nchar) %>% 
+  group_by(group) %>% 
+  summarise_at(vars(ipa1, ipa2, word1, word2), list(mean = mean, sd = sd, min = min, max = max)) %>%
+  split(.$group) %>% 
+  map(select, -group) %>% 
+  set_names(make_clean_names(names(.))) %>% 
+  map(unlist)
+```
+
+We arranged two lists of words: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
+Words in the Catalan list (listened to by participants) were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
+Their translations to English (typed by participants in the keyboard) were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), and their translations to Spanish were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
+Words in the Spanish list were `r printnum(lengths$eng_spa["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_spa["ipa1_sd"])`, Range = `r printnum(lengths$eng_spa["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_spa["ipa1_max"], digits = 0)`).
+Their translations to English were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
+
+We extracted the lexical frequencies from SUBTLEX-UK for English words [@van2014subtlex], SUBTLEX-ESP for Spanish words [@cuetos2011subtlex], and SUBTLEX-CAT for Catalan words [@boada2020subtlex].
+We extracted or transformed scores as/into Zipf scores [@van2014subtlex] to so correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses [@van2014subtlex].
+
+We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) from the CLEARPOND database [@marian2012clearpond].
+PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
+CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target word in only one addition, deletion, or substitution.
+PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
+
+We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist].
+This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
+First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
+This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
+The result is a score that ranges from 0 to 1, where 0 indicates no similarity between the two strings and one indicates that both strings are identical.
+We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
+Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+
+```{r stimuli-correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=4, fig.cap="Marginal distribution and pairwise correlation of lexical frequency, PTHN, and Levenshtein similarity.
+Statistics are indicated for the three experimental conditions (ENG-SPA, ENG-CAT, and SPA-CAT) with different colours.
+The upper triangle of the plot matrix shows the Pearson correlation between each pair of variables (asterisks indicate statistical significance at .05, .01, and .001 significance thresholds).
+The diagonal of the plot matrix indicates the marginal distribution of eacch variable.
+The lower triangle of the plot matrix representes graphically the association between each pair of variables, together with a linear regression line."}
+my_bin <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_density(..., size = 1, alpha = 0.25) +
+    geom_rug(...)
+}
+
+my_smooth <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    geom_point(..., shape = 1, stroke = 1, alpha = 0.25, size = 0.5) +
+    geom_smooth(..., method = "lm", formula = y ~ x, alpha = 0.25, size = 1)
+}
+
+my_cor <- function(data, mapping, ...) {
+  ggplot(data = data, mapping = mapping) +
+    ggally_cor(..., size = 7, title = "r", )
+}
+
+stimuli %>% 
+  drop_na(frequency_zipf, pthn, lv) %>% 
+  ggpairs(
+    mapping = aes(colour = group, fill = group),
+    columns = c("frequency_zipf", "pthn", "lv"),
+    columnLabels = c("Frequency (Zipf)", "PTHN", "Levenshtein"),
+    lower = list(continuous = my_smooth),
+    diag = list(continuous = my_bin),
+    upper = list("my_cor")
+  ) +
+  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
+  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE)
+
+```
+
+Participants listened to one audio file in each trial.
+This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
+The average duration of the audios was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
+The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
+
+```{r stimulitable, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+tb_stimuli <- stimuli %>% 
+  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
+  group_by(group) %>% 
+  summarise_at(
+    vars(frequency:lv), 
+    list(
+      mean = ~mean(., na.rm = TRUE),  
+      sd = ~sd(., na.rm = TRUE)
+    )
+  ) %>%
+  gt(rowname_col = "group") %>% 
+  tab_header(
+    title = md("Table 1"),
+    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
+  ) %>% 
+  fmt_number(frequency_mean:pthn_sd) %>%
+  fmt_percent(matches("lv")) %>% 
+  tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
+  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
+  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
+  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
+  cols_label(
+    frequency_mean = md("Mean"),
+    frequency_sd = md("*SD*"),
+    frequency_zipf_mean = md("Mean"),
+    frequency_zipf_sd = md("*SD*"),
+    pthn_mean = md("Mean"),
+    pthn_sd = md("*SD*"),
+    lv_mean = md("Mean"),
+    lv_sd = md("*SD*")
+  ) %>% 
+  summary_rows(
+    columns = c(matches("mean"), -matches("lv")),
+    fns = list(
+      Mean = "mean",
+      SD = "sd"
+    )
+  ) %>% 
+  summary_rows(
+    columns = matches("lv"),
+    fns = list(
+      Mean = "mean",
+      SD = "sd"
+    ), formatter = fmt_percent
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium", style = "italic"),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_stub_grand_summary()
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = c("left", "right"), color = "white"),
+    locations = list(
+      cells_body(), cells_row_groups(), cells_column_spanners(), 
+      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
+      cells_column_labels(), cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
+    locations = cells_stub()
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+    locations = list(cells_column_spanners(), cells_column_labels())
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "all", color = "white"),
+    locations = cells_body(columns = 1:9)
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "top", color = "white"),
+    locations = cells_body(columns = 1:9)
+  ) %>% 
+  gtsave(here("img", "stimuli.png"))
+
+include_graphics(here("img", "stimuli.png"), auto_pdf = FALSE)
+
+```
+
+## Data analysis
+
+We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0).
+Second, we added the fixed effect of lexical frequency of the target word ($Frequency$).
+We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the $Levenshtein \times Group$ interaction).
+We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.5) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize].
+Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random].
+We will later refer to these models as Model 0 to Model 6.
+For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
+
+$$
+\begin{align}
+
+&\textbf{Likelihood}  \\
+y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
+
+&\textbf{Parameters}  \\
+
+logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
+& \beta_{1[p]} ~ Frequency_{i} ~ + \\
+& \beta_{2[p]} ~ PTHN_i ~ + \\
+& \beta_{3[p]} ~ Similarity_i ~ + \\
+& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
+
+\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
+\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
+&& \text{[participant-level coefficients]} \\ \\
+
+&\textbf{Prior}  \\
+
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
+
+\end{align}
+$$
+
+To test and account for cross-group differences, we included a random intercept for each group.
+We compared models using leave-one-out cross-validation (*LOO*) [@vehtari2017practical].
+More information about the models and model comparison can be found in Appendix 2.
+All analyses were performed in R environment [@rcore2019r].
+We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
+We used the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+
+# Results
+
+All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
+Model 6, which included all main effects and the two-way interactions between $PTHN$ and $Levenshtein$, and between $Group$ and $Levenshtein$ and $PTHN$ and consonant similarity, showed the best performance.
+
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+tab_loos <- model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
+  arrange(desc(elpd_loo)) %>% 
+  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
+  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  tab_header(
+    title = md("Table 2"),
+    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Outcomes of the leave-one-out cross-validation procedure.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information criterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
+  ) %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  fmt_number(2:7) %>% 
+  cols_label(
+    model = md("**Model**"),
+    looic = md("***LOO<sub>IC</sub>***"),
+    se_looic = md("***SE*<sub>IC</sub>**"),
+    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("***LOO<sub>diff</sub>***"),
+    se_diff = md("***SE<sub>diff</sub>***")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium", style = "italic"),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = list(cell_borders(sides = c("left", "right"), color = "white"),
+                 cell_text(align = "center")),
+    locations = list(
+      cells_body(), cells_row_groups(), cells_column_spanners(), 
+      cells_stub(), cells_source_notes(),
+      cells_column_labels(), cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(cell_text(align = "left", size = "medium"), 
+                 cell_borders(sides = "all", color = "white")),
+    locations = list(cells_stub(), cells_source_notes())
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+    locations = list(cells_column_spanners(), cells_column_labels())
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "all", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "top", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  gtsave(here("img", "loos.png"))
+
+include_graphics(here("img", "loos.png"))
+
+```
+
+We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
+For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
+Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
+This way, the coefficients are expressed as increments or decrements in probability of correct translation [@gelman2020regression].
+
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 6.
+A\\) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Black dots represent the mean of each distribution.
+B\\) Participant\\-level coefficient variability.
+The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
+C\\) Correlation between participant\\-level effects.
+The model allowed participant-level coefficients to co\\-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}
+
+coefs <- fixef(model_fits$fit_6) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("variable") %>% 
+  clean_names() %>% 
+  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
+  group_split(variable) %>% 
+  set_names(make_clean_names(map(., "variable"))) %>% 
+  map(select, -variable) %>% 
+  map(unlist)
+
+str_repl <- c(
+  "Intercept" = "Intercept",
+  "frequency_zipf" = "Frequency\n[+1 SD]",
+  "pthn:lv" = "PTHN \u00d7 Levenshtein", 
+  "lv:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv" = "Levenshtein\n[+1 SD]",
+  "pthn" = "PTHN\n[+1 SD]",
+  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+)
+
+# fixed effects
+post_fix <- posterior_draws_fixed %>% 
+  filter(str_detect(.variable, "b_")) %>% 
+  mutate(
+    .variable_name = str_remove(.variable, "b_") %>% 
+      str_replace_all(str_repl) %>%  
+      factor(
+        levels = c(
+          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
+          "Group 1\n[ENG-* vs. SPA-CAT]",
+          "Group 2\n[ENG-CAT vs. ENG-SPA]",
+          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+        )
+      ),
+    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+  ) %>% 
+  arrange(.variable) %>% 
+  ggplot(aes(.value, fct_rev(.variable_name))) +
+  geom_vline(xintercept = 0) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
+  stat_pointinterval(.width = 0, point_size = 1) +
+  scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
+  scale_x_continuous(labels = function(x) percent(round(x, 1))) +
+  labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
+  theme(
+    legend.position = c(1, 0.25),
+    legend.direction = "vertical",
+    legend.background = element_rect(fill = "grey95"),
+    legend.key = element_rect(fill = "grey95", colour = "grey95"), 
+    legend.key.height = unit(0.1, "cm"),
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 7),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black", hjust = 1),
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
+  )
+
+# SD of random effects
+post_sd_data <- posterior_draws_fixed %>% 
+  filter(str_detect(.variable, "sd_")) %>% 
+  mutate(
+    .variable_name = str_remove(.variable, "sd_participant__|sd_word__") %>% 
+      str_replace_all(str_repl) %>% 
+      factor(
+        levels = c(
+          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
+          "Group 1\n[ENG-* vs. SPA-CAT]",
+          "Group 2\n[ENG-CAT vs. ENG-SPA]",
+          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+        )
+      ),
+    group = case_when(
+      str_detect(.variable, "participant") ~ "Participants",
+      str_detect(.variable, "word") ~ "Words"
+    ),
+    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+  )
+
+post_sd <- ggplot(filter(post_sd_data, group=="Participants"), aes(.value, fct_rev(.variable_name))) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
+  stat_pointinterval(.width = 0, point_size = 1) +
+  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
+  scale_x_continuous(labels = percent, limits = c(0, 1)) +
+  labs(x = "SDs in P(Correct)", title = "By participant", fill = "CrI") +
+  theme(axis.text.x = element_blank()) +
+
+  ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
+  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
+  stat_pointinterval(.width = 0, point_size = 1) +
+  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
+  scale_x_continuous(labels = percent, limits = c(0, 1)) +
+  labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +
+
+  plot_layout(ncol = 1, heights = c(0.85, 0.15), guides = "collect") &
+  theme(
+    legend.position = "none",
+    axis.text.y = element_text(hjust = 1),
+    axis.title.x = element_blank(),
+    legend.direction = "vertical",
+    axis.title.y = element_blank(),
+    legend.background = element_rect(fill = "grey95"),
+    legend.key.height = unit(0.1, "cm"),
+    legend.key = element_rect(fill = "grey95", colour = "grey95"),
+    legend.text = element_text(size = 7),
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
+  )
+
+# Pearson correlations between random effects
+corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)
+
+corr_mat <- corr_mat %>% 
+  as_tibble() %>% 
+  select(matches("Estimate")) %>% 
+  rename_all(str_remove, "Estimate.") %>% 
+  mutate(term1 = corr_names)
+
+corr_mat[lower.tri(corr_mat)] <- NA
+
+corr_data <- corr_mat %>% 
+  as.data.frame() %>% 
+  pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
+  drop_na(.value)
+
+post_cors <- corr_data %>% 
+  mutate(term1 = factor(term1, levels = names(str_repl), ordered = TRUE),
+         term2 = factor(term2, levels = names(str_repl), ordered = TRUE)) %>% 
+  mutate_at(vars(term1, term2), str_replace_all, str_repl) %>% 
+  mutate(
+    .value_label = case_when(term1==term2 ~ NA_character_, TRUE ~ printnum(.value, gt1 = FALSE)),
+    .value = case_when(term1==term2 ~ NA_real_, TRUE ~ .value)
+  ) %>% 
+  filter(term1 != "PTHN × Levenshtein", term2 != "Intercept") %>% 
+  ggplot(aes(fct_inorder(term1), fct_rev(fct_inorder(term2)), fill = .value)) +
+  geom_tile(na.rm = TRUE) +
+  geom_text(aes(label = .value_label), size = 3, na.rm = TRUE) +
+  labs(x = "Term 1", y = "Term 2", fill = "Posterior\ncorrelation") +
+  scale_fill_gradient2(
+    low = "#04bf78", mid = "white", high = "#FFC20A", 
+    na.value = "white", limits = c(-0.4, 0.4)
+  ) +
+  scale_x_discrete(position = "top") +
+  coord_equal() +
+  theme(
+    legend.position = c(1, 1),
+    legend.direction = "horizontal",
+    legend.key.height = unit(0.1, "cm"),
+    legend.key.width = unit(0.3, "cm"),
+    legend.justification = c(1, 1),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 5),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 6, angle = 90, hjust = 0, vjust = 1),
+    axis.text.y = element_text(size = 6, vjust = 1, hjust = 1)
+  )
+
+post_fix / ((post_sd | post_cors) + plot_layout(widths = c(0.6, 0.4))) +
+  plot_layout() +
+  plot_annotation(tag_levels = "A")
+
+```
+
+```{r bayesfactors}
+h_frequency <- abs(hypothesis(model_fits$fit_6, hypothesis = "frequency_zipf = 0")$hypothesis$Evid.Ratio)
+h_pthn <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn = 0")$hypothesis$Evid.Ratio)
+h_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv = 0")$hypothesis$Evid.Ratio)
+h_pthn_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn:lv = 0")$hypothesis$Evid.Ratio)
 h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
@@
 \newpage
+
+>>>>>>> c92465ffee5b01a15c56bc645bd58fd50fe89122:Rmd/manuscript.Rmd

### 2022-01-14 _targets.R
Merge versions and run targets

@@ -18,11 +18,7 @@ tar_option_set(
 
 list(
     # stimuli
-<<<<<<< HEAD
     tar_target(clearpond_path, list(`ENG-CAT` = here("data", "clearpond", "clearpond_english.csv"), `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"), `SPA-CAT` = here("Data", "clearpond", "clearpond_spanish.csv")),),
-=======
-    tar_target(clearpond_path, list(`ENG-CAT` = here("data", "clearpond", "clearpond_english.csv"), `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"), `SPA-CAT` = here("data", "clearpond", "clearpond_spanish.csv")),),
->>>>>>> c92465ffee5b01a15c56bc645bd58fd50fe89122
     tar_target(clearpond, get_clearpond(clearpond_path)),
     tar_target(stimuli_path, here("stimuli", "trials.xlsx")),
     tar_target(levenshtein, get_levenshtein(stimuli_path = stimuli_path)),
@@ -35,11 +31,7 @@ list(
     tar_target(responses_path, list.files(here("data", "raw"), full.names = TRUE)),
     tar_target(responses_processed, get_responses_processed(responses_path = responses_path, stimuli = stimuli, practice_trials = practice_trials)),
     tar_target(responses_clean, get_responses_clean(responses_processed = responses_processed, stimuli = stimuli)),
-<<<<<<< HEAD
     tar_target(responses_coded_path, here("data", "processed", "02_coded.xlsx")),
-=======
-    tar_target(responses_coded_path, here("data", "responses-processed.xlsx")),
->>>>>>> c92465ffee5b01a15c56bc645bd58fd50fe89122
     tar_target(responses_coded, read_xlsx(responses_coded_path)),
     tar_target(participants, get_participants(responses_processed = responses_processed,responses_coded = responses_coded)),
     tar_target(responses, get_responses(responses_coded = responses_coded, practice_trials = practice_trials, participants = participants, stimuli = stimuli)),

### 2022-01-14 scripts/R/01_responses.R

@@ -1,4 +1,3 @@
-<<<<<<< HEAD:scripts/R/01_responses.R
 # preprocess responses
 get_responses_processed <- function(
     responses_path,
@@ -228,214 +227,3 @@ get_responses <- function(
     return(responses)
 }
 
-=======
-# preprocess responses
-get_responses_processed <- function(
-    responses_path,
-    stimuli,
-    practice_trials,
-    seed = 888
-){
-    # set parameters
-    set.seed(seed) # for reproducibility
-    spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
-    
-    # list CSV files in Data/Raw/ folder
-    filenames <- list.files(here("data", "raw")) %>% 
-        str_extract(".*?\\_") %>% 
-        str_remove("_")
-    
-    # participant files
-    processed <- responses_path %>% 
-        map(fread, na.string = "") %>%  # import participant files
-        map(mutate, participant = as.character(participant)) %>% 
-        set_names(filenames) %>% # label each dataset with the participant's ID
-        bind_rows(.id = "filename") %>% # merge datasets
-        as_tibble() %>%
-        clean_names() %>% 
-        # select relevant variables and rename if necessary
-        select(
-            participant, test_language, trial_id, word, soundfile, key_pressed,
-            input_text, key_press_time, error, age, date, city,
-            matches("language_l|language_s|language_c|demo_|setup_"),
-            -matches("_rt|_time"),
-            key_press_time
-        ) %>% 
-        rename(demo_impairment = demo_language_key_keys) %>% 
-        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
-        # clean text input by participants (because of typos of need to translate) and redefine location    
-        group_by(participant) %>% 
-        # for each participant, get first row of each demo variable
-        mutate_at(
-            vars(starts_with("demo_"), starts_with("l"), starts_with("setup_"), 
-                 age, city, setup_location, setup_noise),
-            first_non_na # defined in R/utils.R
-        ) %>% 
-        mutate(date = max(date, na.rm = TRUE)) %>% 
-        ungroup() %>% 
-        mutate_at(
-            vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), 
-            clean_input_text # defiend in R/utils.R
-        ) %>% 
-        mutate( 
[…]

### 2022-01-14 scripts/R/02_models.R

@@ -195,3 +195,20 @@ get_model_epreds_random <- function(fit, ndraws = 100, group = c("participant", 
 }
 
 
+get_model_draws_random <- function(fit, ndraws = 100, group = c("participant", "word")){
+    
+    draws_participant <- NULL
+    if ("participant" %in% group){
+        draws_participant <- gather_draws(fit, r_participant[participant, .param], ndraws = ndraws)
+    }
+    
+    draws_word <- NULL
+    if ("word" %in% group){
+        draws_word <- gather_draws(fit, r_word[word, .param], ndraws = ndraws)
+    }
+    
+    draws <- list(participants = draws_participant, words = draws_word)
+    
+    return(draws)
+}
+

### 2022-01-14 docs/03_results.Rmd
Add traceplots

@@
 tar_load_globals()
-tar_load(participants)
-tar_load(responses)
-tar_load(model_fits)
-tar_load(model_loos)
-tar_load(posterior_draws_fixed)
-tar_load(posterior_draws_random)
-tar_load(posterior_epreds_fixed)
+tar_load_all()
 
-theme_set(theme_custom())
+theme_set(
+  theme(
+    title = element_text(size = 8, face = "bold"),
+
+    legend.background = element_rect(fill = NA, colour = NA),
+    legend.key = element_rect(fill = "white", colour = NA),
+    legend.justification = "right",
+    legend.direction = "horizontal",
+
+    axis.line = element_line(colour = "black", size = 0.5),
+    axis.ticks = element_line(colour = "black", size = 0.5),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+
+    plot.caption.position = "plot",
+    plot.title = element_text(hjust = 0.5),
+
+    panel.grid = element_blank(),
+    panel.background = element_rect(fill = "white", colour = NA)
+  )
+)
+```
+
+# Model comparison
+
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+tab_loos <- model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
+  arrange(desc(elpd_loo)) %>% 
+  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
+  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  tab_header(
+    title = md("Table 2"),
+    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Outcomes of the leave-one-out cross-validation procedure.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
+  ) %>% 
+  fmt_missing(everything(), missing_text = "-") %>% 
+  fmt_number(2:7) %>% 
+  cols_label(
+    model = md("**Model**"),
+    looic = md("***LOO<sub>IC</sub>***"),
+    se_looic = md("***SE*<sub>IC</sub>**"),
+    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("***LOO<sub>diff</sub>***"),
+    se_diff = md("***SE<sub>diff</sub>***")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium", style = "italic"),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "left", size = "medium"),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = list(cell_borders(sides = c("left", "right"), color = "white"),
+                 cell_text(align = "center")),
+    locations = list(
+      cells_body(), cells_row_groups(), cells_column_spanners(), 
+      cells_stub(), cells_source_notes(),
+      cells_column_labels(), cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(cell_text(align = "left", size = "medium"), 
+                 cell_borders(sides = "all", color = "white")),
+    locations = list(cells_stub(), cells_source_notes())
+  ) %>% 
+  tab_style(
+    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+    locations = list(cells_column_spanners(), cells_column_labels())
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "all", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  tab_style(
+    style = cell_borders(sides = "top", color = "white"),
+    locations = cells_body(columns = 1:7)
+  ) %>% 
+  gtsave(here("img", "loos.png"))
+
+include_graphics(here("img", "loos.png"))
+
 ```
@@
   fmt_percent(2:4) %>% 
-  cols_merge(vars("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
+  cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
   cols_label(
@@
 
-```{r post_fix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Points and error bars in indicate the posterior means, and 95% CrI.
-The intercept has been transformed using the inverse logit to get the average probability of correct response.
-The resto of the coefficients has been transformed using the divide-by-four- rule to get the maximum change in probability of correct response, associated with a unit increase in this variable."}
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
+A\\) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Black dots represent the mean of each distribution.
+B\\) Participant\\-level coefficient variability.
+The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
+C\\) Correlation between participant\\-level effects.
+The model allowed participant-level coefficients to co\\-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}
 
@@
        title = "By participant") +
+  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
   theme(
@@
 ```{r post_re_word, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+fixed_intercept <- fixef(model_fits$fit_6)[1,c(1, 3, 4)] %>% 
+  inv_logit_scaled()
+
 post_re <- posterior_draws_random$words   %>% 
-  mean_qi() %>% 
   ungroup() %>% 
   arrange(word, .param) %>% 
-  mutate(.param_name = str_replace_all(.param, str_repl) %>%  
+  mutate(
+    .param_name = str_replace_all(.param, str_repl) %>%  
       factor(
@@
         )
-           ))
+      ),
+    .value = inv_logit_scaled(.value)
+  )
 
-ggplot(post_re, 
-       aes(.value, reorder(as.factor(word), -.value), 
-           xmin = .lower, xmax = .upper, colour = "#1A85FF")) +
-  geom_errorbarh(height = 0, alpha = 0.75) +
-  geom_point(size = 0.1, shape = 4) +
-  geom_vline(xintercept = 0) +
-  labs(x = "Participant-level adjustment", y = "Participant", colour = "Group",
-       title = "By participant") +
+ggplot(post_re, aes(.value, reorder(as.factor(word), .value))) +
+  geom_vline(xintercept = fixed_intercept[1], colour = "grey", size = 1) +
+
+  stat_interval(
+    size = 1.5,
+    .width = c(0.5, 0.67, 0.95)
+  ) +
+  annotate(
+    geom = "text",
+    x = 0.7,
+    y = 100,
+    size = 3,
+    hjust = 0.5,
+    label = "Item accuracy \n(95%, 67%, and 50% CrI \nof item-level intercepts)"
+  ) +
+  annotate(
+    geom = "rect",
+    alpha = 0.15,
+    xmin = fixed_intercept[2],
+    xmax = fixed_intercept[3],
+    ymin = -Inf,
+    ymax = Inf
+  ) +
+  annotate(
+    x = 0.7,
+    xend = 0.6,
+    y = 108,
+    yend = 130,
+    geom = "curve",
+    curvature = 0.25,
+    arrow = arrow(length = unit(0.2, "cm"))
+  ) +
+  annotate(
+    geom = "text",
+    x = 0.8,
+    y = 10,
+    size = 3,
+    hjust = 0.5,
+    label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
+  ) +
+  annotate(
+    x = 0.7,
+    xend = fixed_intercept[1]+0.05,
+    y = 17,
+    yend = 30,
+    geom = "curve",
+    curvature = 0.2,
+    arrow = arrow(length = unit(0.2, "cm"))
+  ) +
+  labs(
+    x = "Posterior probability of correct response",
+    y = "Item",
+    colour = "Credible Interval (CrI)",
+    caption = "Credible intervals cover the range of values of the sampling space that \ncontain the true value with a given certainty."
+  ) +
+  scale_color_manual(
+    values = c("#d2e5fc", "#9ccaff", "#1A85FF"),
+    na.translate = FALSE, 
+    labels = paste0(c(95, 67, 50), "%")
+  ) +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
   theme(
-    legend.position = "none",
-    legend.title = element_blank(),
+    legend.position = c(0.1, 0.9),
+    legend.direction = "vertical",
+    legend.justification = "left",
+    legend.key = element_rect(fill = NA, colour = NA),
     panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
@@
     axis.text.y = element_blank(),
-    axis.ticks = element_blank()
+    axis.ticks = element_blank(),
+    axis.title = element_text(size = 10),
+    plot.caption.position = "plot",
+    plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm")))
   )
+
+ggsave(here("img", "accuracy.png"), height = 9, width = 5, dpi = 500)
 ```
@@
 
-```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=6}
+```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=6}
 
@@
   scale_color_manual(values = c("#1E88E5", "#ff2976")) +
-  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+  scale_y_continuous(
+    labels = function(x) percent(round(x, 2)),
+    limits = c(0, 1),
+    breaks = seq(0, 1, 0.1)
+  ) +
   theme(
@@
   coord_flip() +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
+  scale_y_continuous(limits = c(0, 1), labels = percent) +
   theme(
@@
   coord_flip() +
+  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
   scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
@@
 ```
+
+# Model diagnostics
+
+```{r traceplots, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+posterior_draws_fixed %>% 
+  mutate(
+    .chain = as.factor(.chain),
+    .variable_name = str_replace_all(.variable, str_repl) %>% 
+      str_remove("b_")
+    ) %>% 
+  ggplot(aes(.iteration, .value, colour = .chain)) +
+  facet_wrap(~.variable_name, scales = "free_y") +
+  geom_line() +
+  labs(x = "Iteration", y = "Value", color = "Chain") +
+  scale_colour_manual(values = c("#1A85FF", "#ff2976", "#FFC20A", "04bf78"), na.translate = FALSE) +
+  theme(legend.position = "top")
+
+```

### 2022-01-28 manuscript/manuscript.Rmd
Merge versions and run targets
New manuscript version

--- 2022-01-14 manuscript/manuscript.Rmd
@@
   corresponding: yes
-  address: Ramon Trial Fargas, 25-27, 08005 Barcelona
+  address: Ramon Trias Fargas, 25-27, 08005 Barcelona
   email: gonzalo.garciadecastro@upf.edu
@@
 tar_load_globals()
-tar_load(stimuli)
-tar_load(participants)
-tar_load(responses)
-tar_load(model_fits)
-tar_load(model_loos)
-tar_load(posterior_draws_fixed)
-tar_load(posterior_epreds_fixed)
-tar_load(posterior_epreds_random)
+tar_load_all()
 
@@
 
+[//]: # (NSG: I think that a nice way to start the paper (or to finish it) is by citing one of those web pages that say that native English speakers already know tons of words in Spanish, French, German...
+because they are very similar (aka cognates).
+But this may work for writing, because for speech the situation is very different.
+This may be at the beginning or at the end of the paper.
+Given that most reviewers/ editors are likely to be native English-speakers, this may convince them that the paper is interesting and that they want to see it published.)
+
 All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the populations of speakers of each language.
@@
 For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
-This cognate advantage is not restricted to production, but rather extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does], and word translation [@christoffels2006memory].
+This cognate advantage is not restricted to production, but rather extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
 
@@
 This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation [BIA+, @van1998orthographic] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
+The Bilingual Interactive Activation (BIA/BIA+)  [@van1998orthographic; @dijkstra2002architecture] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
 Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
@@
 This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
-Such model also assumes that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
+It also makes the assumption that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
 It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
@@
 The fact that Multilink successfully fitted the experimental data from @christoffels2006memory, despite the absence of any word-word connections between translation equivalents, is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
-It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated, assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
+It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
 The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
@@
 
-In the present study we investigated the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
-Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
-In this case, monolinguals can only use the lexical route to succeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
+Most of the studies referenced so far rely heavily on production and comprehension of printed words: in these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
+This is a limitation regarding the generalisation of their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
+Consider the words *radio* (English) and *radio* (Catalan).
+Both words are evidently cognates, as their orthographical word form is identical.
+However, their phonological forms (/ˈɹeɪdiˌəʊ/-/ˈraðjo/) are not identical.
+Although many English listeners might understand the Spanish word-form, even with no previous knowledge of
+Spanish, it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
+For the same reason, measuring the phonological similarity between two word forms is substantially more challenging than measuring their orthographic similarity.
+
+Most studies use symbolic representations of the word-forms of interest for compute their similarity in either modality: character strings in the case of orthographic word-forms (*radio*-*radio*), and phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA notation) in the case of phonological word forms, which also consists on a string of characters [see @floccia2018introduction for example].
+Either way, the similarity between the pair of word-forms is frequently measured by applying some algorithm that computes the edit distance between both.
+The Leveshtein distance is one of the most widely known: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string [@levenshtein1966binary], and has been successfully used to estimate the similarity of cross-language word-pairs [@gooskens2004perceptive; @schepens2012distributions].
+The Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio")`: both strings are identical.
+But the distance between the phonological transcriptions ɹeɪdiəʊ and raðjo is `r stringdist::stringdist("ɹeɪdiəʊ", "raðjo")`, which is the number of editions that separates both character strings.
+There is yet an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ are symbolic representations that do not capture the all of the features that are present in the acoustic signal when either word is spoken, and therefore any measure of similarity that uses such transcriptions as input will overestimate their actual similarity.
+Taking into account that measuring phonological similarity is considerably more complex than measuyring orthographic complexitym it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
+In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be overestimating the availability of such information in the auditory modality.
+
+In the present study we test listener's reliance on phonological similarity when translating words from a non-native language, in order to investigate the plausibility of the lexical route during backward translation.
+We tested monolingual adults in a translation elicitation task: they were presented with spoken words in a non-native language, and later asked to translate them to their native language.
+The choice of monolinguals as out population of interest is motivated by the fact that monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and therefore word-concept connections are, in principle, absent.
+As a consequence, monolinguals should only be able use the lexical route to succeed in the translation task, which they can only do by exploiting the cognateness of the words presented and their translation in the native language.
 We hypothesised that if participants are able to translate words in L2, without knowledge of the language it belongs to, and based on their form-similarity with the target word, then the lexical route should be in place for low-proficiency bilinguals.
-In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then should low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
+In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then low-proficiency bilinguals should rely entirely on word-concept connections for backward translation.
 
-Following @luce1998recognizing we also hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+We tested three groups of participants in the translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
+Catalan and Spanish are both Roman languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from a typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
+
+Following @luce1998recognizing we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
 If this is the case, words whose translation in the native language are part of larger phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target word.
+For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button* (whose lexical frequency is higher).
+If the phonological neighbourhood density of the target words affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
 
@@
 
-All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/) and a GitHub repository [https://github.com/bilingual-project/translation-elicitation.git](https://github.com/bilingual-project/translation-elicitation.git), along with additional notes.
+All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
 
@@
 We extracted the lexical frequencies from SUBTLEX-UK for English words [@van2014subtlex], SUBTLEX-ESP for Spanish words [@cuetos2011subtlex], and SUBTLEX-CAT for Catalan words [@boada2020subtlex].
-We extracted or transformed scores as/into Zipf scores [@van2014subtlex] to so correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses [@van2014subtlex].
+We extracted or transformed scores as/into Zipf scores [@van2014subtlex] to  correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses [@van2014subtlex].
+
+```{r pthn_example, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+pthn_example <- clearpond_path %>% 
+  map(read.csv) %>% 
+  bind_rows(.id = "group") %>% 
+  clean_names() %>% 
+  filter(word %in% c("book", "sock"), group=="ENG-CAT") %>% 
+  select(word, pthn)
+
+```
 
 We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) from the CLEARPOND database [@marian2012clearpond].
-PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
+PTHN scores indicate the number of phonological neighbours of the target word (the correct translation in the native language of the presented word) with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
 CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target word in only one addition, deletion, or substitution.
+For instance, the PTHN score of *book* is `r pthn_example$pthn[1]`, while that of *sock* is `r pthn_example$pthn[2]`.
 PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
@@
 We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim("teɪbəl", "mesa"))` similarity, while the *train* (trEIn)-*guitarra* (tɾEn) translation pair had a `r percent(stringsim("trEIn", "tɾEn"))` similarity.
 Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.

### 2022-01-28 manuscript/manuscript.tex
New manuscript version

--- 2021-12-02 Rmd/manuscript.tex
@@
   pdflang={en-EN},
-  pdfkeywords={cognate, word recognition, translation, non-native, spoken word recognition},
+  pdfkeywords={cognate, word recognition, translation, non-native, spoken word recognition, bayesian},
   hidelinks,
@@
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 \shorttitle{Cognateness and non-native word recognition}
-\keywords{cognate, word recognition, translation, non-native, spoken word recognition\newline\indent Word count: X}
+\keywords{cognate, word recognition, translation, non-native, spoken word recognition, bayesian\newline\indent Word count: 4553}
 \usepackage{lineno}
@@
 
-GG and SS contributed equally and share first authorship.
-
-The authors made the following contributions.
-Gonzalo Garcia-Castro: Conceptualization, Data analysis, Writing - Original Draft Preparation, Writing - Review \& Editing; Serene Siow: Conceptualization, Writing - Review \& Editing; Nuria Sebastian-Galles: Conceptualization, Writing - Review \& Editing; Kim Plunkett: Conceptualization, Writing - Review \& Editing.
-
-Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trial Fargas, 25-27, 08005 Barcelona.
+Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona.
 E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}
@@
 \affiliation{\vspace{0.5cm}\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra\\\textsuperscript{2} Department of Experimental Psychology, University of Oxford}
-
-\abstract{
-There is compelling evidence that bilinguals access their lexicon in a language non-selective way.
-For example, bilinguals produce and translate cognates (words whose translation in the other language is form-similar) faster than non-cognates, suggesting that the phonology of both languages interact during word production and comprehension (Costa et al., 2000; Christoffels et al., 2006).
-Previous literature on this effect has often relied on measures of overall form-similarity to categorise words into cognates andnon-cognates, such as the Levenshtein distance.
-These measures partially ignore potential sources of variability during lexical access like vowel vs.~consonant overlap, overlap in stressed syllables, onset overlap or the distance in features between both translations.
-In this study, we explored the impact of some of these variables on non-native word translation task: Spanish and English participants listened to non-native words (Catalan or Spanish) and were prompted to type their translation in their native language.
-Critically, participants where unfamiliar with the testing language, ensuring that they were not able to translate words based on previous knowledge on their meaning, leaving phonological information as the only cue participants were able to exploit to translate words to their native language correctly.
-We analysed the probability of correct translations, adjusting for the amount of vowel overlap, consonant overlap, overlap at stress position, overlap at onset, and distance in features between replaced phonemes, including the average frequency of phonological neighbours of the target words are a covariate.
-}
 
@@
 
-Humans are able to recognise spoken words without much effort, even in adverse conditions such as compressed speech (De Haan, 1982) or the loss of segmental information (Warren, 1970).
-But even in the most ideal of the situations, the cognitive processes engaged in word recognition do not occur without uncertainty.
-The speech input activates multiple candidate lexical representations based on their similarity with the signal, and for word recognition to take place, one of the candidates must be selected.
-There is extensive literature about the influence on word recognition of the number of lexical candidates, their frequency, and their phonological and semantic similarity with target words.
-Less is known about how all these factors affect the dynamics of word recognition in non-native languages.
+Listening to non-native speech is more costly than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations (Lecumberri, Cooke, \& Cutler, 2010; Takata \& Nábělek, 1990).
+One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
+For example, a Spanish native listening to French may encounter the word /pɔʁt/ (\emph{porte}), translation of \emph{door} in French.
+The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
+These discrepancies can nonetheless be perceived as allophonic variations of native phonemes (Best, McRoberts, \& Goodell, 2001), and the non-native speech signal engages lexical processing mechanisms (Weber \& Cutler, 2004).
+Although this mismatch has a noticeable toll on comprehension (Cutler, Weber, Smits, \& Cooper, 2004), the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
+In the present study we investigate the extent to which adult listeners rely on the phonological similarity between the native and the non-native languages during word recognition, capitalising on the role of \emph{cognates}.
 
-Listening to speech in a non-native language is more costly than doing so in native languages, regardless of the level of proficiency {[}REF{]}.
-The source of this increased cognitive effort is likely to originate at multiple levels.
-For example, some sounds in the speech signal do not correspond to any phoneme in the native language, making the process of matching sounds to phonemes complicated.
-This is the case of a Spanish native listening to words containing the guttural /ʁ/ in French.
-Language learners, however, are rarely completely naïve to the language they are listening to.
-All languages share, to some extent, similarities, frequently due to their typological closeness.
-These similarities which can be exploited by non-native listeners to process the speech signal.
-In the present research we investigate the extent to which adult listeners rely on phonological similarity between cross-language synonyms during word translation.
-
-Cognates are cross-language synonyms whose form (e.g., phonology, orthography, signature) is similar.
-The cause of such similarity is frequently attributed to a shared etymological origin.
-Romance languages such as Spanish and Catalan share many cognates (Schepens, Dijkstra, \& Grootjen, 2012), as in the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively)\footnote{Some form-similar cross-language synonyms are technically not cognates.
+All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the populations of speakers of each language.
+Cognates embody a particular case of such commonalities.
+Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin\footnote{Some form-similar cross-language synonyms are technically not cognates.
 For example, \emph{sun} and \emph{sol} (in Spanish), share their phonological onset, but their etymology points to different origins.
@@
 It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.}.
+For example, Romance languages such as Spanish and Catalan share many cognates (Schepens, Dijkstra, \& Grootjen, 2012), as in the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively).
+Cognates play a pivotal role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., \emph{puerta}-\emph{porta}), compared to when their labels are non-cognates (\emph{mesa}-\emph{taula}, \emph{table} in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval (Costa, Caramazza, \& Sebastian-Galles, 2000).
+This cognate advantage is not restricted to production, but rather extends to word recognition (Midgley, Holcomb, \& Grainger, 2011; Thierry \& Wu, 2007), word learning (De Groot \& Keijzer, 2000; Lotto \& De Groot, 1998; Valente, Ferré, Soares, Rato, \& Comesaña, 2018), and word translation (Christoffels, De Groot, \& Kroll, 2006).
 
-Cognates play a major role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-During word production and comprehension, lexical representations of both languages are activated, and the form similarity between both impacts bilinguals' performance in naming or lexical decision tasks (e.g., Costa, Caramazza, \& Sebastian-Galles, 2000; Thierry \& Wu, 2007).
-In a seminal study, Costa et al.
-(2000) asked Spanish-Catalan adult bilinguals to name pictures in Spanish.
-Unbeknownst to participants, half of the labels associated to pictures were cognates in Spanish and Catalan (e.g., \emph{puerta}-\emph{porta}) while the other half were non-cognates (e.g., \emph{mesa}-\emph{taula}, meaning \emph{table} in Spanish and Catalan).
-Surprisingly, Spanish-Catalan bilinguals named cognate pictures faster than non-cognate pictures, while Spanish monolinguals did not show this effect.
-This suggests that bilinguals activate the picture's word representations in both languages, and the phonological overlap between both labels facilitated the naming process.
-
-Additionally, cognates seem to be learnt more easily in L2 than non-cognates.
-De Groot and Keijzer (2000) presented Dutch natives pairs of Dutch words and pseudowords.
-Participants' performance learnt and retained better pairs with high orthographic similarity than pairs with low orthographic similarity (see also Lotto \& De Groot, 1998 for Dutch natives learning Italian words).
-
-Cognates are not only learnt faster than non-cognates, but translated faster too.
-Early accounts of bilingual lexical access suggest that low-proficiency learners first established links between newly acquired words in L2 and their meanings through their L1 translation equivalents (Potter, So, Von Eckardt, \& Feldman, 1984).
-As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on the mediation of L1 representations.
-The Revised Hierarchical Model {[}RHM; Kroll and Stewart (1994){]} captured this assumption and predicted that translating words from L1 to L2 (forward translation) should take longer than translating from L2 to L1 (backward translation).
-The rationale behind this prediction is that backward translation relies more strongly on direct word-word links between L1 and L2 representations, while forward translation would rely more strongly on the mediation between the concept and the two word forms.
+Early theoretical accounts of this phenomenon pointed to the fact that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) (Potter, So, Von Eckardt, \& Feldman, 1984; see Chen \& Leung, 1989 for similar results).
+The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), and not to their shared concept (conceptual route).
+This word-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection).
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations, therefore to cognateness.
+The Revised Hierarchical Model (RHM, Kroll \& Stewart, 1994) captured this hypothesis and predicted that translating words from L1 to L2 (\emph{forward} translation) should take longer than translating from L2 to L1 (\emph{backward} translation).
+The rationale behind this prediction was that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
 One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
 
-To test these predictions, Degroot, Dannenburg, and Vanhell (1994) and Groot (1992) asked Dutch natives with high (yet non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
-Participants were presented with written words in English or Dutch, and asked to speak out loud their translation in the other language as fast as possible.
-Results who that translation times and accuracy were roughly equivalent in across both forward and backward conditions, and that semantic variables (e.g., concreteness), were positively associated with participants' performance, especially in the forward translation condition.
+To test these predictions, Annette MB de Groot, Dannenburg, and Vanhell (1994) and Annette M.
+de Groot (1992) asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
+Results showed that translation times and accuracy were roughly equivalent in forward and backward translation.
 Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected by semantic variables than non-cognates.
+Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
+Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
+Although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels et al., 2006; Christoffels, Ganushchak, \& Koester, 2013).
+These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
 
-These findings suggest that both conceptually mediated and direct translation routes are active during translation, although backward translation relies more strongly on direct links between L1 and L2 representations, which makes it more sensitive to cognateness (than forward translation).
-Interestingly, higher-proficiency bilinguals tested in the same task showed similar results.
-However, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels, De Groot, \& Kroll, 2006; Christoffels, Ganushchak, \& Koester, 2013), contrary to the predictions of the RHM.
-Further models build upon this model, introducing a wider range of connections in their theorised lexica.
+Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+For instance, the Neighbourhood Activation Model (NAM, Luce \& Pisoni, 1998) highlighted the fact that lexical selection is affected by the number of phonological neighbours of the selected word.
+Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods (Goldinger, Luce, \& Pisoni, 1989; Luce, Pisoni, \& Goldinger, 1990), especially when such neighbours' lexical frequency is higher (Luce \& Pisoni, 1998).
+This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+The Bilingual Interactive Activation (Dijkstra \& Van Heuven, 2002; BIA/BIA+, Van Heuven, Dijkstra, \& Grainger, 1998) addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing (McClelland \& Rumelhart, 1981).
+Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
+When bilingual participants are presented with words in only one language (target language) in a lexical decision task, translating them to the non-target language took longer when they were part of large neighbourhoods (Van Heuven et al., 1998).
+This suggests that words presented in the target language activated orthographic neighbourhoods in the non-target language, which competed for selection with the target word during recognition.
 
-Since Kroll and Stewart (1994) proposed their RHM model, subsequent studies brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-The Neighbourhood Activation Model {[}NAM; Luce and Pisoni (1998){]} suggests that lexical selection is mediated not only by the lexical frequency of the target word (and the number of activated candidates), but also by the frequency of the candidates.
+A more recent model of bilingual lexical processing is Multilink (Dijkstra et al., 2019), which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
+This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
+Such model also assumes that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
+It also proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
+In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
+A remarkable correlation between Multilink simulations of a translation elicitation task and behavioural data on Dutch-English bilinguals (Christoffels et al., 2006) has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
+Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+Overall, results from Multilink are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
 
-Subsequent behavioural experiments (e.g., Luce \& Pisoni, 1998) show that English adult participants were tested in a lexical decision task in which words were presented across several conditions of noise-to-signal ratio.
-A computerised lexicon calculated number and average lexical frequency of phonological neighbours around each of the presented words.
-Participants answered more faster and more accurately to high frequency words, and words from high density neighbourhoods were also responded detected more accurately, but slowly.
-Therefore, average lexical frequency of neighbourhoods was associated to a decrease in accuracy.
-Low frequency words were in fact detected more accurately in low density than in high density neighbours.
-These findings suggested that both word lexical frequency and neighbourhood density can facilitate spoken word recognition, but this effect is modulated by the frequency of the neighbourhood; when phonological neighbours are more frequent, lexical selection is hindered (Goldinger, Luce, \& Pisoni, 1989; Luce, Pisoni, \& Goldinger, 1990).
+The fact that Multilink successfully fitted the experimental data from Christoffels et al.
+(2006), despite the absence of any word-word connections between translation equivalents, is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+It should be taken into account, however, that Christoffels et al.
+(2006) only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated, assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
+The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
+It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation, and this prediction remains untested my Multilink.
 
-The influence of phonological neighbourhoods and their structure on lexical processing poses rise important question in the field of bilingualism research.
-Do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation {[}BIA+; Van Heuven, Dijkstra, and Grainger (1998){]} proposes that lexical representations establish both excitatory and inhibitory connections with other representations from the same or the other language, forming an integrated lexicon for both languages (McClelland \& Rumelhart, 1981).
-Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
-Van Heuven et al.
-(1998) presented bilingual participants with words in only one language (target language) in a lexical decision task.
-When the translation of the presented word in the non-target language was part of a large neighbourhood, reaction times were slower.
-This provides evidence that the presented words activated orthographic neighbourhoods in the non-target language, which slowed down word recognition.
+Most of the studies referenced so far rely heavily on production and comprehension of printed words: in these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
+This is a limitation regarding the generalisation of their findings to the auditory modality, especially when cognateness or more generally any measure of similarity between word-forms, is involved.
+Consider the words \emph{radio} (English) and \emph{radio} (Catalan).
+Both words are evidently cognates, as their orthographical word form is identical.
+However, their phonological form is not identical: /ˈɹeɪdiˌəʊ/ and /ˈraðjo/.
+Although many English listeners might understand the Spanish word-form, even with no previous knowledge of
+Spanish, it is likely that they will find the translation process more challenging when presented with the auditory word-form than when presented with the visual word-form.
+For the same reason, measuring the phonological similarity between two word forms is substantially more challenging than measuring their orthographic similarity.
+Most studies use symbolic representations of the word-forms of interest for compute their similarity in either modality: character strings in the case of orthographic word-forms (\emph{radio}-\emph{radio}), and phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA transcription) in the case of phonological word forms, which also consists on a string of characters.
+Either way, the similarity between the pair of word-forms is frequently measured by applying some algorithm that computes the edit distance between both.
+The Leveshtein distance is one of the most widely known: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string (Levenshtein \& others, 1966), and has been successfully used to estimate the similarity of cross-language word-pairs (Gooskens \& Heeringa, 2004; Schepens et al., 2012).
+The Levenshtein distance between the orthographic word-forms is 0: both strings are identical.
+But the distance between ɹeɪdiəʊ and raðjo is 35, which is the number of editions that separates both character strings.
+There is yet an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ are symbolic representations that do not capture the all of the features that are present in the acoustic signal when either word is spoken, and therefore any measure of similarity that uses such transcriptions as input will overestimate their actual similarity.
+Taken this into consideration, it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
+In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be overestimating the availability of such information in the auditory modality.
 
-A more recent model of bilingual lexical processing is Multilink Dijkstra et al.
-(2019), which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and more relevant to the aims of the present study, translation.
-This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
-Such model also assumes that during backward translation, the presented word in L2 can activate form-similar words in L1 that compete with the target word in L1 and it proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model (which suggests that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
-In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
-A remarkable correlation between Multilink simulations of Christoffels et al.
-(2006) translation elicitation task and behavioural data with Dutch-English bilinguals has been reported, showing forward translation being faster than backward translation, contrary to the predictions made by the RHM model.
-Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
-Overall, results from Multilink from this model are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
+In the present study we test listener's reliance on phonological similarity when translating words from a non-native language, in order to investigate the plausibility of the lexical route during backward translation.
+We tested monolingual adults in a translation elicitation task: they were presented with spoken words in a non-native language, and later asked to translate them to their native language.
+Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
+In this case, monolinguals can only use the lexical route to succeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
+We hypothesised that if participants are able to translate words in L2, without knowledge of the language it belongs to, and based on their form-similarity with the target word, then the lexical route should be in place for low-proficiency bilinguals.
+In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then low-proficiency bilinguals should rely entirely on word-concept connections for backward translation.
 
-The fact that Multilink fitted the experimental data from Christoffels et al.
-(2006) with remarkable success, even in absence of word-word connections between translation equivalents might put into question the necessity of such connections during translation, in favour of a simpler model in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
-This claim, however, should be taken.
-It should be taken into account, however, that Christoffels et al.
-(2006) only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
-Additionally, data generated from Multilink model assumed similar vocabulary sizes across languages and, perhaps more critically in this regard, that word-concept connections were equally strong for L1 and L2 representations.
-These conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation.
-This prediction remains untested my Multilink.
+Following Luce and Pisoni (1998) we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+If this is the case, words whose translation in the native language are part of larger phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target word.
 
-The role of this study was to investigate the plausibility of the lexical route as a mechanism exploited by monolingual adults during backward translation.
-Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
-In this case, monolinguals can only use the lexical route to suceeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
-If participants are able to translate words in L2, even with no knowledge of the language it belongs to, based on their form-similarity with the target word, this would suggest that the lexical route is in place for low-proficiency bilinguals.
-If participants are not able to translation words from L2 to L1 regardless of their form-similarity, this would suggest that low-proficiency bilinguals rely entirely on word-concept connections for backward translation.
+We tested three groups of participants in the translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
+Catalan and Spanish are both Roman languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from a typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
 
@@
 \section{Methods}\label{methods}}
+
+All materials, data, and code used in this study are hosted in the Open Science Framework \url{https://osf.io/9fjxm/} and a GitHub repository \url{https://github.com/bilingual-project/translation-elicitation.git}, along with additional notes.
 
@@
 Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Participants contributed a total of 8077 valid trials (5235 in Catalan, 2842 in Spanish).
+Trials in which participants took longer than 10 seconds to respond were also filtered out.
+Participants contributed a total of 7239 valid trials (4809 in Catalan, 2430 in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{C:/Users/u155880/Documents/translation-elicitation/Figures/design} \caption{Schematic representation of a trial in the experimental task.}\label{fig:procedurefigure}
+\includegraphics[width=0.8\linewidth]{C:/Users/u155880/Documents/translation-elicitation/img/design} \caption{Schematic representation of a trial in the experimental task.}\label{fig:procedurefigure}
 \end{figure}
@@
 
-We retrieved PTHN scores from the CLEARPOND database (Marian, Bartolotti, Chabal, \& Shook, 2012).
+We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) from the CLEARPOND database (Marian, Bartolotti, Chabal, \& Shook, 2012).
 PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
@@
 PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
+
+\begin{figure}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stimuliplot-1} \caption{Distribution and summary statistics of stimuli properties.
+A\) Lexical frequency of target words, expressed in counts per million and as Zipf scores.
+B\) Number of higher-frequency phonological neighbourhoods of target words (PTHN).
+C\) Normalised Levenshtein phonological similarity between the presented and the target words, expressed in percentages.}\label{fig:stimuliplot}
+\end{figure}
 
@@
 This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is a score that ranges from 0 to 1, where 0 indicates no similarity between the two strings and one indicates that both strings are identical.
+The result is percentage score indicating no similarity between the two strings if 0\%, and a perfect match between two strings if 100\%.
 We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
-Table 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
 
@@
 The average duration of the audios was 1.20 (\emph{SD} = 0.18, Range = 0.78-1.58).
-The average duration of the Catalan audios was 1.23 seconds (\emph{SD} = 0.19, Range = 0.80-1.58), and the average duration of the Spanish audios was 1.16 seconds (\emph{SD} = 0.15, Range = 0.78-1.53).
-
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{lrrrrrrrr}
-\toprule
- & \multicolumn{2}{c}{\textbf{Freq./million}} & \multicolumn{2}{c}{\textbf{Freq. (Zipf)}} & \multicolumn{2}{c}{\emph{\textbf{PTHN}}} & \multicolumn{2}{c}{\textbf{Levenshtein}} \\ 
- \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7} \cmidrule(lr){8-9}
- & Mean & \emph{SD} & Mean & \emph{SD} & Mean & \emph{SD} & Mean & \emph{SD} \\ 
-\midrule
-ENG-CAT & $46.80$ & $77.23$ & $4.31$ & $0.59$ & $3.26$ & $3.74$ & $0.16$ & $0.19$ \\ 
-ENG-SPA & $43.91$ & $67.53$ & $4.32$ & $0.56$ & $3.52$ & $4.10$ & $0.13$ & $0.17$ \\ 
-SPA-CAT & $53.32$ & $156.19$ & $4.23$ & $0.62$ & $1.62$ & $2.62$ & $0.37$ & $0.25$ \\ 
- \bottomrule
-\end{longtable}
+The average duration of the Catalan audio files was 1.23 seconds (\emph{SD} = 0.19, Range = 0.80-1.58), and the average duration of the Spanish audios was 1.16 seconds (\emph{SD} = 0.15, Range = 0.78-1.53).
 
@@
 We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We first fitted a base model (Model 0) that only included the lexical frequency (\texttt{frequency}) of the target word as a fixed effect, with a random intercept per participant.
-Second, we extended the model to include \texttt{pthn} as a fixed effect, with a random slope by participant (Model 1).
-Third, we added the fixed effect \texttt{consonant\_ratio} and the \texttt{pthn:consonant\_ratio}, and random slopes for both effects by participant (Model 2).
-Third, we added the fixed effect \texttt{vowel\_ratio} and the \texttt{pthn:vowel\_ratio}, and random slopes for both effects by participant (Model 3).
-Finally, we fit a model that included the two-way interactions \texttt{pthn:consonant\_ratio} and \texttt{pthn:vowel\_ratio}, an their random slopes by participant (Model 4).
-All predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
-Model 4 can be formally expressed as:
+We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant (Model 0).
+Second, we added the fixed effect of lexical frequency of the target word (\(Frequency\)).
+We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood (\(PTHN\)), the main effect of Levenshtein similarity (\(Levenshtein\)), the \(PTHN \times Levenshtein\) interaction, the group of participants (\(Group\), with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the \(Levenshtein \times Group\) interaction).
+We specified two sum-coded \emph{a priori} contrasts for the \(Group\) predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.25) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) (Schad, Vasishth, Hohenstein, \& Kliegl, 2020).
+Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) (Barr, Levy, Scheepers, \& Tily, 2013).
+We will later refer to these models as Model 0 to Model 6.
+For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
+
+\$\$
+\begin{align}
+
+&\textbf{Likelihood}  \\
+y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
+
+&\textbf{Parameters}  \\
+
+logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
+& \beta_{1[p]} ~ Frequency_{i} ~ + \\
+& \beta_{2[p]} ~ PTHN_i ~ + \\
+& \beta_{3[p]} ~ Similarity_i ~ + \\
+& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
+
+\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
+\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
+&& \text{[participant-level coefficients]} \\ \\
+
+&\textbf{Prior}  \\
+
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
+
+\end{align}
+\$\$
 
@@
 All analyses were performed in R environment (RCore, 2019).
-We used the tidyverse family of R packages to process data and to generate figures, and the brms R package (Bürkner, 2017) using the cmdstanr backend to the Stan probabilistic language (Carpenter et al., 2017) to estimate and compare the models (see Appendix 1 for mode details on the models).
+We used the tidyverse family of R packages (Wickham et al., 2019) to process data and to generate figures.
+We used the brms R package (Bürkner, 2017) using the cmdstanr backend to the Stan probabilistic language (Carpenter et al., 2017) to estimate and compare the models (see Appendix 1 for mode details on the models).
 
@@
 All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
-Model 4, which included all main effects and the two-way interactions between PTHN and vowel similarity, and PTHN and consonant similarity, showed the best performance (see Table 2).
+Model 6, which included all main effects and the two-way interactions between \(PTHN\) and \(Levenshtein\), and between \(Group\) and \(Levenshtein\) and \(PTHN\) and consonant similarity, showed the best performance.
 
-\captionsetup[table]{labelformat=empty,skip=1pt}
-\begin{longtable}{lrrrrrr}
-\toprule
- & \emph{\textbf{LOOELPD}} & \emph{\textbf{SE}} & \emph{\textbf{LOO\_\{IC\}}} & \textbf{\emph{SE}IC} & \emph{\textbf{LOOdiff}} & \emph{\textbf{SEdiff}} \\ 
-\midrule
-Model 3 & $-4,277.50$ & $44.29$ & $8,555.00$ & $88.58$ & - & - \\ 
-Model 2 & $-4,295.13$ & $44.44$ & $8,590.26$ & $88.89$ & $-17.63$ & $6.08$ \\ 
-Model 1 & $-4,310.21$ & $44.50$ & $8,620.43$ & $89.00$ & $-32.72$ & $8.28$ \\ 
-Model 0 & $-4,321.68$ & $44.31$ & $8,643.37$ & $88.62$ & $-44.19$ & $9.44$ \\ 
- \bottomrule
-\end{longtable}
+\includegraphics{C:/Users/u155880/Documents/translation-elicitation/img/loos.png}
 
-We now report the mean of the posterior distribution of each coefficient in Model 3, along with its associated measures of uncertainty.
+We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
 For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
 Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increases/decreases in probability of correct translation (Gelman, Hill, \& Vehtari, 2020).
+This way, the coefficients are expressed as increments or decrements in probability of correct translation (Gelman, Hill, \& Vehtari, 2020).
 
@@
 \includegraphics{manuscript_files/figure-latex/posteriorfix-1.pdf}
-\caption{\label{fig:posteriorfix}Estimated posterior distributions of coefficients in Model 3.
+\caption{\label{fig:posteriorfix}Estimated posterior distributions of coefficients in Model 4.
 A) Population-level effects.
-Distributions indicate the estimated posterior likelihood density of each coefficient.
-Credible intervals \(CrI\), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\%, 80\%, and 50\% probability.
-Dots represent the mean of the distribution.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (\emph{CrI}), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\%, 80\%, and 50\% probability.
+Black dots represent the mean of each distribution.
 B) Participant-level coefficient variability.
-Our model estimated participant-level coefficients to account for the dependency between responses from the same participant.
-Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations \(SD\).
+The model estimated participant-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (\emph{SD}).
 C) Correlation between participant-level effects.
-Our model allowed participant-level coefficients to co-vary.
-This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each correlation.
-Coefficients are represented in the X-axis and Y-axis in the same order as indicated in the Y-axis of panels A and C.}
+The model allowed participant-level coefficients to co-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each estimated correlation.}
 \end{figure}
 
-Overall, participants were 35.45\%, (\emph{SE} = 52.06\%, 95\% \emph{CrI} = {[}32.09\%, 39.47\%{]}) likely to produce correct translations.
-Every standard deviation increment in the translation's lexical frequency (\emph{SD} = 0.59) increased the probability of a correct responses in 3.53\% (\emph{SE} = 0.67\%, 95\% \emph{CrI} = {[}2.21\%, 4.85\%{]}).
-The number of the translation's more frequent phonological neighbours, on the other hand, decreased the probability of a correct responses in -3.90\% (\emph{SE} = 0.76\%, 95\% \emph{CrI} = {[}-5.42\%, -2.43\%{]}) for every increase in 1 \emph{SD} (3.65).
-The effect of phonological similarity was conditional to the phonological density of the translation.
-Phonological similarity barely increased the probability of a correct responses (0.78\%, \emph{SE} = 1\%, 95\% \emph{CrI} = {[}-1.02\%, 3\%{]}) by itself.
-However, for every \emph{SD} increase in PTHN, phonological similarity increased in -2\%, (\emph{SE} = 0.80\%, 95\% \emph{CrI} = {[}-3.78\%, -0.68\%{]}) such probability.
+Overall, participants were 42.65\%, (95\% \emph{CrI} = {[}37.89\%, 47.63\%{]}) likely to produce correct translations.
+The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (\emph{SD} = 0.57 Zipf points) decreased the probability of correct translation in -0.63\% (95\% \emph{CrI} = {[}-0.63\%, -3.72\%{]}, \(BF_{\beta\neq< 0}\) = 1.53).
+The number of phonological neighbours of the correct translation with higher lexical frequency (\emph{PTHN}) slightly decreased the probability of a correct responses in -2.91\% (95\% \emph{CrI} = {[}-5.90\%, 0.07\%{]}, \(BF_{\beta \neq 0}\) = 0.32) for every increase standard deviation increment (1 \emph{SD} = 3.94 neighbours).
+Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every \emph{SD} increment in Levenshtein similarity (1 \emph{SD} = 23\% similarity), translation accuracy increased in 12.18\% (95\% \emph{CrI} = {[}7.63\%, 16.64\%{]}, \(BF_{\beta \neq 0}\) = \(>3.09 \times 10^{18}\)).
+This difference increased in 0.23\% for every \emph{SD} increment in \emph{PTHN} (95\% \emph{CrI} = {[}-2.79\%, 3.30\%{]}, \(BF_{\beta \neq 0}\) = 1.68)
+
+Spanish natives translating Catalan words were 3.67\% more accurate than English natives translating Spanish and Catalan words (95\% \emph{CrI} = {[}-1.18\%, 8.64\%{]}, \(BF_{\beta \neq 0}\) = 0.38).
+This difference increased in 3.38\% for every one \emph{SD} increment in Levenshtein similarity (95\% \emph{CrI} = {[}-1.64\%, 8.34\%{]}, \(BF_{\beta \neq 0}\) = 0.47).
+English natives translating Catalan words were 0.87\% more accurate than those translating Spanish words (95\% \emph{CrI} = {[}-5.39\%, 3.63\%{]}, \(BF_{\beta \neq 0}\) = 0.38).
+For every one \emph{SD} increment in Levenshtein similarity, the effect increased in 3.60\% (95\% \emph{CrI} = {[}-7.71\%, 0.48\%{]}, \(BF_{\beta \neq 0}\) = 0.47).
 
@@
 \includegraphics{manuscript_files/figure-latex/marginaleffects-1.pdf}
-\caption{\label{fig:marginaleffects}Expected mean posterior predictions.
-A) Population-level expected mean posterior predictions.
-The X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 200 observations from our model: 100 simulations for words with low PTHN (-1 SD) and 100 simulations for high PTHN (+1 SD) words.
+\caption{\label{fig:marginaleffects}Expected mean posterior predictions from Model 4.
+A) Population-level predictions: the X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 200 observations from our model: 100 simulations for words with low \emph{PTHN} (-1 \emph{SD}) and 100 simulations for high \emph{PTHN} (+1 \emph{SD}) words.
 We did this across the range of values of the Levenshtein scores.
 For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high PTHN words.
-We also plotted the mean of the high PTHN (black solid line) and low PTHN (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
+Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high \emph{PTHN} words.
+We also plotted the mean of the high \emph{PTHN} (black solid line) and low \emph{PTHN} (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
 The dispersion of the lines indicates the uncertainty of our predictions.
-B) Participant-level expected mean posterior predictions.
-We conducted the same procedure for each individual participant, simulating 100 observations for high-PTHN and 100 for low-PTHN from our model acrosss the range of Levenshtein scores.
-We averaged the resulting predictions for each participant and then plotted each participant's predictions in separate panels to show how the effect of PTHN and phonological similarity changed at the individual level.}
+We computed these posterior predictions for each group of participants, and plot tem in separate panels.}
 \end{figure}
+
+The conditional intra-class correlation (\emph{ICC}) for items was 69.64\%, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items (Gelman \& Hill, 2006).
+Item-level intercepts showed remarkable variability (\emph{SD} = 96.92\%, 95\% CrI = {[}94.75\%, 98.40\%{]}), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish \emph{tomate}-\emph{tomato}, /tomate/-/təmɑːtoʊ/, 99.73\%, 95\% CrI = {[}96.72\%, 99.99\%{]}), while others were almost never translated correctly (e.g., Catalan \emph{colom}-\emph{pigeon}, /kulom/-/pɪdʒɪn/, 0.08\%, 95\% CrI = {[}0.00\%, 0.81\%{]}).
+Participant-level intercepts also showed high variability (\emph{SD} = 74.72\%, 95\% CrI = {[}70.39\%, 79.47\%{]}), suggesting that average translation accuracy also varied significantly across participants.
+Finally, participant-level estimated slopes showed less variability: \emph{SD} of \(Frequency\), \(PTHN\), \(Levenshtein\), \(PTHN \times Levenshtein\), and \(Levenshtein \times Group 2\) ranged between 2.32\%, and 4.30\%.
+The participant-level estimated slopes of the \(Leveshtein \times Group 1\) interaction are an exception, though (\emph{SD} = 41.18\%, 95\% CrI = {[}24.79\%, 57.80\%{]}).
+This indicates that the impact of phonological similarity (\(Levenshtein\)) on the difference in accuracy between English and Spanish natives differed substantially across participants.
+
+Finally, participant-level effects showed weak Pearson correlations overall (\textless10\%), with three exceptions.
+The estimated correlation between the intercepts and the slopes of \(Frequency\) was 29.30\% (95\% CrI = {[}-14.99\%, 63.21\%{]}, \(BF_{Cor \neq 0}\) = \texttt{r}round(h\_cor\_intercept\_frequency, 2)`).
+The estimated correlation between the slopes of \(Frequency\) and \(Levenshtein \times Group 1\) was 21.17\% (95\% CrI = {[}-19.73\%, 56.32\%{]}, \(BF_{Cor \neq 0}\) = 0.60).
+The estimated correlation between the intercepts and the slopes of \(PTHN\) was 10.88\% (95\% CrI = {[}-28.52\%, 45.58\%{]}, \(BF_{Cor \neq 0}\) = 0.96).
+The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-barr2013random}{}}%
+Barr, D.
+J., Levy, R., Scheepers, C., \& Tily, H.
+J.
+(2013).
+Random effects structure for confirmatory hypothesis testing: Keep it maximal.
+\emph{Journal of Memory and Language}, \emph{68}(3), 255--278.
+
+\leavevmode\vadjust pre{\hypertarget{ref-best2001discrimination}{}}%
+Best, C.
+T., McRoberts, G.
+W., \& Goodell, E.
+(2001).
+Discrimination of non-native consonant contrasts varying in perceptual assimilation to the listener's native phonological system.
+\emph{The Journal of the Acoustical Society of America}, \emph{109}(2), 775--794.
+
 \leavevmode\vadjust pre{\hypertarget{ref-boada2020subtlex}{}}%
@@
 \emph{Grantee Submission}, \emph{76}(1), 1--32.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chen1989patterns}{}}%
+Chen, H.-C., \& Leung, Y.-S. (1989). Patterns of lexical processing in a nonnative language. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{15}(2), 316.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cutler2004patterns}{}}%
+Cutler, A., Weber, A., Smits, R., \& Cooper, N.
+(2004).
+Patterns of english phoneme confusions by native and non-native listeners.
+\emph{The Journal of the Acoustical Society of America}, \emph{116}(6), 3668--3678.
+
 \leavevmode\vadjust pre{\hypertarget{ref-de2000hard}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-de1982relationship}{}}%
-De Haan, H. J. (1982). The relationship of estimated comprehensibility to the rate of connected speech. \emph{Perception \& Psychophysics}, \emph{32}(1), 27--31.
-
-\leavevmode\vadjust pre{\hypertarget{ref-degroot1994forward}{}}%
-Degroot, A. M., Dannenburg, L., \& Vanhell, J. G. (1994). Forward and backward word translation by bilinguals. \emph{Journal of Memory and Language}, \emph{33}(5), 600--629.
+\leavevmode\vadjust pre{\hypertarget{ref-dijkstra2002architecture}{}}%
+Dijkstra, T., \& Van Heuven, W.
+J.
+(2002).
+The architecture of the bilingual word recognition system: From identification to decision.
+\emph{Bilingualism: Language and Cognition}, \emph{5}(3), 175--197.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-gelman2006data}{}}%
+Gelman, A., \& Hill, J. (2006). \emph{Data analysis using regression and multilevel/hierarchical models}. Cambridge university press.
+
 \leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-gooskens2004perceptive}{}}%
+Gooskens, C., \& Heeringa, W. (2004). Perceptive evaluation of levenshtein dialect distance measurements using norwegian dialect data. \emph{Language Variation and Change}, \emph{16}(3), 189--207.
+
 \leavevmode\vadjust pre{\hypertarget{ref-de1992determinants}{}}%
-Groot, A. M. de. (1992). Determinants of word translation. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{18}(5), 1001.
+Groot, Annette M. de. (1992). Determinants of word translation. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{18}(5), 1001.
+
+\leavevmode\vadjust pre{\hypertarget{ref-degroot1994forward}{}}%
+Groot, Annette MB de, Dannenburg, L., \& Vanhell, J. G. (1994). Forward and backward word translation by bilinguals. \emph{Journal of Memory and Language}, \emph{33}(5), 600--629.
 
@@
 \emph{Journal of Memory and Language}, \emph{33}(2), 149--174.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lecumberri2010non}{}}%
+Lecumberri, M. L. G., Cooke, M., \& Cutler, A. (2010). Non-native speech perception in adverse conditions: A review. \emph{Speech Communication}, \emph{52}(11-12), 864--886.
 
@@
 McClelland, J. L., \& Rumelhart, D. E. (1981). An interactive activation model of context effects in letter perception: I. An account of basic findings. \emph{Psychological Review}, \emph{88}(5), 375.
+
+\leavevmode\vadjust pre{\hypertarget{ref-midgley2011effects}{}}%
+Midgley, K.
+J., Holcomb, P.
+J., \& Grainger, J.
+(2011).
+Effects of cognate status on word comprehension in second language learners: An ERP investigation.
+\emph{Journal of Cognitive Neuroscience}, \emph{23}(7), 1634--1647.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-schad2020capitalize}{}}%
+Schad, D.
+J., Vasishth, S., Hohenstein, S., \& Kliegl, R.
+(2020).
+How to capitalize on a priori contrasts in linear (mixed) models: A tutorial.
+\emph{Journal of Memory and Language}, \emph{110}, 104038.
+
 \leavevmode\vadjust pre{\hypertarget{ref-schepens2012distributions}{}}%
 Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of cognates in europe as based on levenshtein distance. \emph{Bilingualism: Language and Cognition}, \emph{15}(1), 157--166.
+
+\leavevmode\vadjust pre{\hypertarget{ref-takata1990english}{}}%
+Takata, Y., \& Nábělek, A.
+K.
+(1990).
+English consonant recognition in noise and in reverberation by japanese and american listeners.
+\emph{The Journal of the Acoustical Society of America}, \emph{88}(2), 663--666.
 
@@
 \emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535.
+
+\leavevmode\vadjust pre{\hypertarget{ref-valente2018does}{}}%
+Valente, D., Ferré, P., Soares, A., Rato, A., \& Comesaña, M.
+(2018).
+Does phonological overlap of cognate words modulate cognate acquisition and processing in developing and skilled readers?
+\emph{Language Acquisition}, \emph{25}(4), 438--453.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-warren1970perceptual}{}}%
-Warren, R. M. (1970). Perceptual restoration of missing speech sounds. \emph{Science}, \emph{167}(3917), 392--393.
+\leavevmode\vadjust pre{\hypertarget{ref-weber2004lexical}{}}%
+Weber, A., \& Cutler, A. (2004). Lexical competition in non-native spoken-word recognition. \emph{Journal of Memory and Language}, \emph{50}(1), 1--25.
+
+\leavevmode\vadjust pre{\hypertarget{ref-wickham2019tidyverse}{}}%
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
+D., François, R., \ldots{} Yutani, H.
+(2019).
+Welcome to the {tidyverse}.
+\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
+\url{https://doi.org/10.21105/joss.01686}
 
@@
 
-\hypertarget{appendix}{%
-\section{Appendix}\label{appendix}}
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/accuracy-1.pdf}
-\caption{\label{fig:accuracy}Proportion of correct translation by item.
-Words presented to participants in the English-Spanish or in the English-Catalan and Spanish-Catalan are listed in the Y-axis, ordered from higher to lower average translation accuracy, depicted in the X-axis.
-Dots and whiskers represent the average accuracy and 95\% confidence interval of each word.
-Accuracy is plotted separately for each group.}
-\end{figure}
-
 \end{document}

### 2022-02-14 manuscript/manuscript.Rmd
Implement NS's second round of comments

@@
 Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
-For example, *sun* and *sol* (in Spanish), share their phonological onset, but their etymology points to different origins.
-We will use the word *cognateness* to include all form-similar cross-language synonyms for simplicity.
+For example, *sun* and *sol* (in Spanish) share phonological onset but their etymology points to different origins.
+The complimentary case is equally problematic for the definition of cognateness in this context: translation equivalents that share etymological origin may hardly be recognised as cognates by the average listener.
+Such is the case of *cheese* and *queso* (both from Latin), but phonologically dissimilar.
+We will use the term *cognateness* to include all form-similar cross-language synonyms for simplicity.
 It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
@@
 This is a limitation regarding the generalisation of their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
-Consider the words *radio* (English) and *radio* (Catalan).
+Consider the words *radio* (English) and *radio* (Spanish).
 Both words are evidently cognates, as their orthographical word form is identical.
 However, their phonological forms (/ˈɹeɪdiˌəʊ/-/ˈraðjo/) are not identical.
-Although many English listeners might understand the Spanish word-form, even with no previous knowledge of
-Spanish, it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
+Although many English listeners might understand the Spanish word-form, even with no previous knowledge of Spanish, it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
 For the same reason, measuring the phonological similarity between two word forms is substantially more challenging than measuring their orthographic similarity.
@@
 There is yet an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ are symbolic representations that do not capture the all of the features that are present in the acoustic signal when either word is spoken, and therefore any measure of similarity that uses such transcriptions as input will overestimate their actual similarity.
-Taking into account that measuring phonological similarity is considerably more complex than measuyring orthographic complexitym it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
+Taking into account that measuring phonological similarity is considerably more complex than measuring orthographic complexity it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
 In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be overestimating the availability of such information in the auditory modality.
@@
 In the present study we test listener's reliance on phonological similarity when translating words from a non-native language, in order to investigate the plausibility of the lexical route during backward translation.
-We tested monolingual adults in a translation elicitation task: they were presented with spoken words in a non-native language, and later asked to translate them to their native language.
-The choice of monolinguals as out population of interest is motivated by the fact that monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and therefore word-concept connections are, in principle, absent.
-As a consequence, monolinguals should only be able use the lexical route to succeed in the translation task, which they can only do by exploiting the cognateness of the words presented and their translation in the native language.
-We hypothesised that if participants are able to translate words in L2, without knowledge of the language it belongs to, and based on their form-similarity with the target word, then the lexical route should be in place for low-proficiency bilinguals.
-In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then low-proficiency bilinguals should rely entirely on word-concept connections for backward translation.
+We tested monolingual adults in a translation elicitation task: they were presented with spoken words in an unfamiliar, non-native language, and were asked to translate them to their native language.
+The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which vocabulary size in L2 is null, and therefore word-concept connections are, in principle, absent.
+As a consequence, these participants should only be able use the lexical route to succeed in the translation of words in an unfamiliar language.
+We predicted that, if the information provided by the phonological form of the unfamiliar words is sufficient for translation, participants' performance should increase as the translation pairs are phonologically more similar.
 
@@
 Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the local ethical committee (XXXXXXXXXXXX).
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`) , or c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
+We excluded data from participants that (a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), (b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`), or (c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
 
@@
 Then, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+Participants were informed that they would listen to a series of pre-recorded words in either Catalan or Spanish (English participants) or Catalan (Spanish participants).
 They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the list of Catalan or Spanish trials.
+English participants were randomly assigned to the Catalan or Spanish lists.
 Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
@@
 We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
+Trials where the response was mistyped by only one character were counted as correct, as long as the response did not correspond to a distinct word.
 Trials in which participants took longer than 10 seconds to respond were also filtered out.
@@
   group_by(group) %>% 
-  summarise_at(vars(ipa1, ipa2, word1, word2), list(mean = mean, sd = sd, min = min, max = max)) %>%
+  summarise_at(
+    vars(ipa1, ipa2, word1, word2), 
+    list(mean = mean, sd = sd, min = min, max = max)
+  ) %>%
   split(.$group) %>% 
@@
 
-We arranged two lists of words: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
+### Input stimuli (Catalan/Spanish)
+
+We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
 Words in the Catalan list (listened to by participants) were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
-Their translations to English (typed by participants in the keyboard) were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), and their translations to Spanish were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
 Words in the Spanish list were `r printnum(lengths$eng_spa["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_spa["ipa1_sd"])`, Range = `r printnum(lengths$eng_spa["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_spa["ipa1_max"], digits = 0)`).
-Their translations to English were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
+Participants listened to one audio file in each trial.
+This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
+The average duration of the audio files was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
+The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
 
-We extracted the lexical frequencies from SUBTLEX-UK for English words [@van2014subtlex], SUBTLEX-ESP for Spanish words [@cuetos2011subtlex], and SUBTLEX-CAT for Catalan words [@boada2020subtlex].
+### Translations (English/Spanish)
+
+Catalan translations to English (typed by participants in the keyboard) were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), and their translations to Spanish were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
+Spanish translations to English were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
+
+#### Translation lexical frequency
+
+We extracted the lexical frequencies from SUBTLEX-UK for English translations [@van2014subtlex], and from SUBTLEX-ESP for Spanish translations [@cuetos2011subtlex].
 We extracted or transformed scores as/into Zipf scores [@van2014subtlex] to  correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses [@van2014subtlex].
@@
 
-We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) from the CLEARPOND database [@marian2012clearpond].
+#### Translation phonological neighbourhood density (PTHN)
+
+We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) of the English and Spanish translations  from the CLEARPOND database [@marian2012clearpond].
 PTHN scores indicate the number of phonological neighbours of the target word (the correct translation in the native language of the presented word) with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
@@
 
-```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.cap="Distribution and summary statistics of stimuli properties.
-A\\) Lexical frequency of target words, expressed in counts per million and as Zipf scores.
-B\\) Number of higher-frequency phonological neighbourhoods of target words (PTHN).
-C\\) Normalised Levenshtein phonological similarity between the presented and the target words, expressed in percentages."}
+```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of stimuli translations, plotted in different panels for English translations of Catalan stimuli (cat-ENG), English translations of Spanish stimuli (spa-ENG), and Spanish translation of Catalan stimuli (cat-SPA).
+A\\) Lexical frequency of stimuli translations, expressed in counts per million and as Zipf scores.
+B\\) Number of higher-frequency phonological neighbourhoods (PTHN) of stimuli translations.
+C\\) Normalised Levenshtein phonological similarity between the presented stimuli and their translations, expressed in percentages.
+Lexical frequencies and PTHN scores of the stimuli (presented auditorily in Catalan or Spanish) are not described, as participants lack any lexical representation of such words."}
 
@@
     lv_sd = sd(lv)
-  )
+  ) %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  ))
 
+# frequency
 stimuli %>% 
-  count(group, pthn) %>% 
-  ggplot(aes(pthn, n, fill = group)) + 
-  facet_wrap(~group, ncol = 1) + 
-  geom_rect(
-    data = medians, 
-    aes(
-      xmin = 0, xmax = pthn_median+pthn_sd, 
-      ymin = -Inf, ymax = Inf, fill = group
-    ), inherit.aes = FALSE, alpha = 0.15
-  ) +   
-  geom_col(colour = "white") + 
-  geom_vline(
-    data = medians, aes(xintercept = pthn_median, group = group), 
-    colour = "black", size = 0.5, linetype = "dashed"
-  ) + 
-  geom_text(data = medians, aes(label = "Median ± SD", x = pthn_median+10, y = 40), size = 3) +
-  geom_curve(
-    data = medians, aes(x = pthn_median+10, xend = pthn_median+0.2, y = 35, yend = 35),
-    arrow = arrow(length = unit(0.2, "cm")), curvature = -0.25, colour = "grey30") +
-  labs(x = "PTHN", y = "Number of trials", title = "Higher-frequency neighbours") +
-
-  ggplot(stimuli, aes(frequency_zipf, fill = group)) + 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  )) %>% 
+  ggplot(aes(frequency_zipf, fill = group)) + 
   facet_wrap(~group, ncol = 1) + 
@@
   geom_vline(
-    data = medians, aes(xintercept = frequency_zipf_median, group = group), 
+    data = medians, 
+    aes(xintercept = frequency_zipf_median, group = group), 
+    colour = "black",
+    size = 0.5,
+    linetype = "dashed"
+  ) + 
+  geom_text(
+    data = medians, 
+    aes(
+      label = "Median ± SD", 
+      x = frequency_zipf_median+1.25,
+      y = 13
+    ), 
+    size = 3
+  ) +
+  geom_curve(
+    data = medians, 
+    aes(
+      x = frequency_zipf_median+0.85, 
+      xend = frequency_zipf_median+0.1,
+      y = 12, 
+      yend = 11
+    ),
+    arrow = arrow(length = unit(0.2, "cm")),
+    curvature = -0.25,
+    colour = "grey30"
+  ) +
+  labs(
+    x = "Lexical frequency (Zipf)", 
+    y = "Number of trials", 
+    title = "Frequency\n(stimuli translations)\n"
+  ) +
+
+  # PTHN
+  stimuli %>% 
+  count(group, pthn) %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  )) %>% 
+  ggplot(aes(pthn, n, fill = group)) + 
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0, 
+      xmax = pthn_median+pthn_sd, 
+      ymin = -Inf, 
+      ymax = Inf,
+      fill = group
+    ), inherit.aes = FALSE, 
+    alpha = 0.15
+  ) +   
+  geom_col(colour = "white") + 
+  geom_vline(
+    data = medians, aes(xintercept = pthn_median, group = group), 
     colour = "black", size = 0.5, linetype = "dashed"
   ) + 
-  labs(x = "Lexical frequency (Zipf)", y = "Number of trials", title = "Frequency") +
+  labs(
+    x = "PTHN", 
+    y = "Number of trials", 
+    title = "Higher-frequency neighbours\n(stimuli translations)\n"
+  ) +
 
-  ggplot(stimuli, aes(lv, fill = group)) +
+  # Levenshtein
+  stimuli %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  )) %>% 
+  ggplot(aes(lv, fill = group)) +
   facet_wrap(~group, ncol = 1) + 
@@
   scale_x_continuous(labels = percent) +
-  labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity") +
+  labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity\n(stimuli and stimuli translations)\n") +
 
@@
 ```
+
+### Cross-language phonological overlap (Levenshtein)
 
@@
 
-Participants listened to one audio file in each trial.
-This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the audios was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
-The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audios was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
-
 ## Data analysis
@@
 Second, we added the fixed effect of lexical frequency of the target word ($Frequency$).
-We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the $Levenshtein \times Group$ interaction).
-We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.25) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) [@schad2020capitalize].
+We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels cat-ENG, spa-ENG, and cat-SPA), and finally, the $Levenshtein \times Group$ interaction).
+We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) [@schad2020capitalize].
 Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random].

### 2022-02-18 docs/00_stimuli.Rmd
Render README as target

@@
 
-```{r stimuli-correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=6, fig.height=4, fig.cap="Marginal distribution and pairwise correlation of lexical frequency, PTHN, and Levenshtein similarity.
-Statistics are indicated for the three experimental conditions (ENG-SPA, ENG-CAT, and SPA-CAT) with different colours.
-The upper triangle of the plot matrix shows the Pearson correlation between each pair of variables (asterisks indicate statistical significance at .05, .01, and .001 significance thresholds).
-The diagonal of the plot matrix indicates the marginal distribution of eacch variable.
-The lower triangle of the plot matrix representes graphically the association between each pair of variables, together with a linear regression line."}
-my_bin <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_density(..., size = 1, alpha = 0.25) +
-    geom_rug(...)
-}
+```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of stimuli translations, plotted in different panels for English translations of Catalan stimuli (cat-ENG), English translations of Spanish stimuli (spa-ENG), and Spanish translation of Catalan stimuli (cat-SPA).
+A\\) Lexical frequency of stimuli translations, expressed in counts per million and as Zipf scores.
+B\\) Number of higher-frequency phonological neighbourhoods (PTHN) of stimuli translations.
+C\\) Normalised Levenshtein phonological similarity between the presented stimuli and their translations, expressed in percentages.
+Lexical frequencies and PTHN scores of the stimuli (presented auditorily in Catalan or Spanish) are not described, as participants lack any lexical representation of such words."}
 
-my_smooth <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    geom_point(..., shape = 1, stroke = 1, alpha = 0.25, size = 0.5) +
-    geom_smooth(..., method = "lm", formula = y ~ x, alpha = 0.25, size = 1)
-}
+medians <- stimuli %>% 
+  group_by(group) %>% 
+  summarise(
+    pthn_median = median(pthn),
+    pthn_sd = sd(pthn),
+    frequency_zipf_median = median(frequency_zipf),
+    frequency_zipf_sd = sd(frequency_zipf),
+    lv_median = median(lv),
+    lv_sd = sd(lv)
+  ) %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  ))
 
-my_cor <- function(data, mapping, ...) {
-  ggplot(data = data, mapping = mapping) +
-    ggally_cor(..., size = 7, title = "r", )
-}
+# frequency
+stimuli %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  )) %>% 
+  ggplot(aes(frequency_zipf, fill = group)) + 
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd, 
+      ymin = -Inf, ymax = Inf, fill = group
+    ), inherit.aes = FALSE, alpha = 0.15
+  ) +   
+  geom_histogram(bins = 20, colour = "white") +
+  geom_vline(
+    data = medians, 
+    aes(xintercept = frequency_zipf_median, group = group), 
+    colour = "black",
+    size = 0.5,
+    linetype = "dashed"
+  ) + 
+  geom_text(
+    data = medians, 
+    aes(
+      label = "Median ± SD", 
+      x = frequency_zipf_median+1.25,
+      y = 13
+    ), 
+    size = 3
+  ) +
+  geom_curve(
+    data = medians, 
+    aes(
+      x = frequency_zipf_median+0.85, 
+      xend = frequency_zipf_median+0.1,
+      y = 12, 
+      yend = 11
+    ),
+    arrow = arrow(length = unit(0.2, "cm")),
+    curvature = -0.25,
+    colour = "grey30"
+  ) +
+  labs(
+    x = "Lexical frequency (Zipf)", 
+    y = "Number of trials", 
+    title = "Frequency\n(stimuli translations)\n"
+  ) +
 
+  # PTHN
   stimuli %>% 
-  drop_na(frequency_zipf, pthn, lv) %>% 
-  ggpairs(
-    mapping = aes(colour = group, fill = group),
-    columns = c("frequency_zipf", "pthn", "lv"),
-    columnLabels = c("Frequency (Zipf)", "PTHN", "Levenshtein"),
-    lower = list(continuous = my_smooth),
-    diag = list(continuous = my_bin),
-    upper = list("my_cor")
+  count(group, pthn) %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  )) %>% 
+  ggplot(aes(pthn, n, fill = group)) + 
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0, 
+      xmax = pthn_median+pthn_sd, 
+      ymin = -Inf, 
+      ymax = Inf,
+      fill = group
+    ), inherit.aes = FALSE, 
+    alpha = 0.15
   ) +   
-  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
-  scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE)
+  geom_col(colour = "white") + 
+  geom_vline(
+    data = medians, aes(xintercept = pthn_median, group = group), 
+    colour = "black", size = 0.5, linetype = "dashed"
+  ) + 
+  labs(
+    x = "PTHN", 
+    y = "Number of trials", 
+    title = "Higher-frequency neighbours\n(stimuli translations)\n"
+  ) +
+
+  # Levenshtein
+  stimuli %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  )) %>% 
+  ggplot(aes(lv, fill = group)) +
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0, xmax = lv_median+lv_sd, 
+      ymin = -Inf, ymax = Inf, fill = group
+    ), inherit.aes = FALSE, alpha = 0.15
+  ) +   
+  geom_histogram(bins = 20, colour = "white") +
+  geom_vline(
+    data = medians, aes(xintercept = lv_median, group = group), 
+    colour = "black", size = 0.5, linetype = "dashed"
+  ) + 
+  scale_x_continuous(labels = percent) +
+  labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity\n(stimuli and stimuli translations)\n") +
+
+  plot_layout(guides = "collect") &
+  plot_annotation(tag_levels = "A") &
+  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
+  theme(legend.position = "none")
 

### 2022-02-18 manuscript/manuscript.Rmd
Render README as target

@@
 abstract: |
+  Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar.
 
@@
 
-coefs <- fixef(model_fits$fit_6) %>% 
+coefs <- fixef(fit_6) %>% 
   as.data.frame() %>% 
@@
   "Intercept" = "Intercept",
-  "frequency_zipf" = "Frequency\n[+1 SD]",
-  "pthn:lv" = "PTHN \u00d7 Levenshtein", 
-  "lv:group1" = "Levenshtein \u00d7 Group 1",
-  "lv:group2" = "Levenshtein \u00d7 Group 2",
-  "lv" = "Levenshtein\n[+1 SD]",
-  "pthn" = "PTHN\n[+1 SD]",
-  "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-  "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "frequency_zipf_std" = "Frequency\n[+1 SD]",
+  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
+  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv_std" = "Levenshtein\n[+1 SD]",
+  "pthn_std" = "PTHN\n[+1 SD]",
+  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
 )
@@
 # Pearson correlations between random effects
-corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
-corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)
+corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
 
@@
 ```{r bayesfactors}
-h_frequency <- abs(hypothesis(model_fits$fit_6, hypothesis = "frequency_zipf = 0")$hypothesis$Evid.Ratio)
-h_pthn <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn = 0")$hypothesis$Evid.Ratio)
-h_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv = 0")$hypothesis$Evid.Ratio)
-h_pthn_lv <- abs(hypothesis(model_fits$fit_6, hypothesis = "pthn:lv = 0")$hypothesis$Evid.Ratio)
-h_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
-h_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "group1 = 0")$hypothesis$Evid.Ratio)
-h_lv_group1 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group1 = 0")$hypothesis$Evid.Ratio)
-h_lv_group2 <- abs(hypothesis(model_fits$fit_6, hypothesis = "lv:group2 = 0")$hypothesis$Evid.Ratio)
+h_frequency <- abs(hypothesis(fit_6, hypothesis = "frequency_zipf_std = 0")$hypothesis$Evid.Ratio)
+h_pthn <- abs(hypothesis(fit_6, hypothesis = "pthn_std = 0")$hypothesis$Evid.Ratio)
+h_lv <- abs(hypothesis(fit_6, hypothesis = "lv_std = 0")$hypothesis$Evid.Ratio)
+h_pthn_lv <- abs(hypothesis(fit_6, hypothesis = "pthn_std:lv_std = 0")$hypothesis$Evid.Ratio)
+h_group1 <- abs(hypothesis(fit_6, hypothesis = "groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
+h_group2 <- abs(hypothesis(fit_6, hypothesis = "groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)
+h_lv_group1 <- abs(hypothesis(fit_6, hypothesis = "lv_std:groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
+h_lv_group2 <- abs(hypothesis(fit_6, hypothesis = "lv_std:groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)
 
-icc_conditional <- icc(model_fits$fit_6, by_group = TRUE)$ICC[2]
+icc_conditional <- icc(fit_6, by_group = TRUE)$ICC[2]
 
-sds <- gather_draws(model_fits$fit_6, `sd_.*`, regex = TRUE) %>% 
+sds <- gather_draws(fit_6, `sd_.*`, regex = TRUE) %>% 
   mean_qi() %>% 
@@
 
-word_example <- ranef(model_fits$fit_6)$word[,,"Intercept"] %>% 
+word_example <- ranef(fit_6)$word[,,"Intercept"] %>% 
   as.data.frame() %>% 
@@
 
-cors_range <- gather_draws(model_fits$fit_6, `cor_.*`, regex = TRUE) %>% 
+cors_range <- gather_draws(fit_6, `cor_.*`, regex = TRUE) %>% 
   median_qi() %>% 
@@
 
-corr_mat <- VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor 
-corr_names <- rownames(VarCorr(model_fits$fit_6, summary = TRUE)$participant$cor)
+corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
 
@@
 
-h_cor_intercept_frequency <- hypothesis(model_fits$fit_6, "participant__Intercept__frequency_zipf = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_frequency_zipf_lv_group1 <- hypothesis(model_fits$fit_6, "participant__frequency_zipf__lv:group1 = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_intercept_pthn <- hypothesis(model_fits$fit_6, "participant__Intercept__pthn = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_intercept_frequency <- hypothesis(fit_6, "participant__Intercept__frequency_zipf_std = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_frequency_zipf_lv_group1 <- hypothesis(fit_6, "participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_intercept_pthn <- hypothesis(fit_6, "participant__Intercept__pthn_std = 0", class = "cor")$hypothesis$Evid.Ratio
 ```
@@
 Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations.
-The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
-The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
-Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
-This difference increased in `r percent(coefs$pthn_lv["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_lv["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_lv["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
+The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf_std["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
+The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
+Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
+This difference increased in `r percent(coefs$pthn_std_lv_std["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_std_lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std_lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
 
-Spanish natives translating Catalan words were `r percent(coefs$group1["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
-This difference increased in `r percent(coefs$lv_group1["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_group1["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group1["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-English natives translating Catalan words were `r percent(abs(coefs$group2["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
-For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_group2["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_group2["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_group2["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+Spanish natives translating Catalan words were `r percent(coefs$group_en_gvs_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
+This difference increased in `r percent(coefs$lv_std_group_en_gvs_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+English natives translating Catalan words were `r percent(abs(coefs$group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
+For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_std_group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
 
@@
 m <- posterior_epreds_fixed %>% 
-  mutate_at(vars(pthn, frequency_zipf), function(x) paste0(x, " SD")) %>% 
+  mutate_at(vars(pthn_std, frequency_zipf_std), function(x) paste0(x, " SD")) %>% 
   mutate(
-    pthn = case_when(
-      pthn=="-1 SD" ~ "Low PTHN",
-      pthn=="0 SD" ~ "Mean PTHN",
-      pthn=="1 SD" ~ "High PTHN"
+    pthn_std = case_when(
+      pthn_std=="-1 SD" ~ "Low PTHN",
+      pthn_std=="0 SD" ~ "Mean PTHN",
+      pthn_std=="1 SD" ~ "High PTHN"
     ),
-    frequency_zipf = case_when(
-      frequency_zipf=="-1 SD" ~ "Low frequency",
-      frequency_zipf=="0 SD" ~ "Mean frequency",
-      frequency_zipf=="1 SD" ~ "High frequency"
+    frequency_zipf_std = case_when(
+      frequency_zipf_std=="-1 SD" ~ "Low frequency",
+      frequency_zipf_std=="0 SD" ~ "Mean frequency",
+      frequency_zipf_std=="1 SD" ~ "High frequency"
     )
@@
 
-ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
+ggplot(m, aes(lv_std, .epred, colour = pthn_std, fill = pthn_std)) +
   facet_wrap(~group) +
   geom_hline(yintercept = 0.5, colour = "grey") +
-  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
-  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
+  geom_line(aes(group = interaction(pthn_std, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
+  stat_summary(aes(linetype = pthn_std), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
   annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
@@
 Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
-The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_group1$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_group1$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_group1$.upper, accuracy = 0.01)`]).
+The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_std_group_en_gvs_spaE$.upper, accuracy = 0.01)`]).
 This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
@@
 Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions.
-The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r `round(h_cor_intercept_frequency, 2)`).
-The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf$estimate_lv_group1, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf$q2_5_lv_group1, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf$q97_5_lv_group1, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
-The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).
+The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_frequency, 2)`).
+The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf_std$estimate_lv_std_group_en_gvs_spa, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf_std$q2_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf_std$q97_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
+The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).
 The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.

### 2022-02-18 R/01_responses.R

@@ -208,11 +208,12 @@ get_responses <- function(
         arrange(group, trial_id) %>%  
         # typos are considered correct responses
         # transform relative frequency to Zipf score
-        mutate(frequency_zipf = relative_to_zipf(frequency)) %>% 
-        # center predictors
-        mutate_at(
-            vars(lv, pthn, frequency_zipf),
-            function(x) scale(x, center = TRUE, scale = TRUE)[,1]
+        mutate(
+            frequency_zipf = 3+log10(frequency),
+            # center predictors
+            frequency_zipf_std = scale(frequency_zipf)[,1],
+            pthn_std = scale(pthn)[,1],
+            lv_std = scale(lv)[,1]
         ) %>% 
         arrange(trial_id, group) %>% 
         mutate_at(vars(group), as.factor) %>% 

### 2022-02-18 R/utils.R

@@ -4,7 +4,7 @@
 # find first/last non-missing value in vector
 first_non_na <- function(x) ifelse(is.logical(first(x[!is.na(x)])), NA, first(x[!is.na(x)]))
 last_non_na <- function(x) ifelse(is.logical(last(x[!is.na(x)])), NA, last(x[!is.na(x)]))
-add_big_mark <- function(x) format(x, big.mark = ",", scientific = FALSE)
+# add_big_mark <- function(x) format(x, big.mark = ",", scientific = FALSE)
 clean_input_text <- function(x) {
     {{ x }} %>% 
         str_to_sentence() %>%
@@ -27,7 +27,8 @@ clean_input_text <- function(x) {
             TRUE ~ value)) %>% 
         pull(value)
 }
-# ggplot custom theme
+
+#  custom theme
 theme_custom <- function(){
     theme(
         panel.background = element_rect(fill = "transparent"),
@@ -55,80 +56,64 @@ replace_non_ascii <- function(x){
             "ñ" = "n",
             "ç" = "c"
         )
-        )
+    )
 }
 
 
 # download clearpond
-import_clearpond <- function(language = c("english", "dutch", "french", "spanish", "german")){
-    require(tidyverse)
-    require(janitor)
-    urls <- tribble(
-        ~lang, ~url, ~data, ~header,
-        "english", "https://clearpond.northwestern.edu/englishCPdatabase2.zip", "englishCPdatabase2.txt", "clearpondHeaders_EN.txt",
-        "dutch", "https://clearpond.northwestern.edu/dutchCPdatabase2.zip", "dutchCPdatabase2.txt", "clearpondHeaders_NL.txt",
-        "french", "https://clearpond.northwestern.edu/frenchCPdatabase2.zip", "frenchCPdatabase2.txt", "clearpondHeaders_FR.txt",
-        "german", "https://clearpond.northwestern.edu/germanCPdatabase2.zip", "germanCPdatabase2.txt", "clearpondHeaders_DE.txt",
-        "spanish", "https://clearpond.northwestern.edu/spanishCPdatabase2.zip", "spanishCPdatabase2.txt", "clearpondHeaders_SP.txt"
-    ) %>% 
-        filter(lang %in% language)
-    
-    dir <- tempdir()
-    files <- replicate(tempfile(), n = length(urls$lang))
-    d <- pmap(
-        .l = list(url = as.list(urls$url), file = as.list(files), data = as.list(urls$data), header = as.list(urls$header)),
[…]

### 2022-02-18 _targets_packages.R

@@ -8,7 +8,6 @@ library(data.table)
 library(dplyr)
 library(forcats)
 library(future)
-library(GGally)
 library(ggplot2)
 library(gt)
 library(here)
@@ -21,6 +20,7 @@ library(mice)
 library(multilex)
 library(papaja)
 library(patchwork)
+library(performance)
 library(pingr)
 library(purrr)
 library(readxl)

### 2022-02-18 README.Rmd
Render README as target

---
output: github_document
clean: true
---

# TranslationElicitation

More info in the Open Science Framework repository.

## How to make this repository work on your local machine

### 1. Downloading the repository

You can either download the project as a .zip file by clocking on the green button that says "Code", or if you feel comfortable with Git, you can clone the repository from your console running:

```bash
git clone https://github.com/gongcastro/translation-elicitation.git
```

### 2. Setting up R packages with renv

We used the R package renv to keep track of the dependencies of the project. renv allows to install and update the R packages necessary to run the project in a project-specific environment. Packages installed using renv in this project will only be installed or updated in the context of this project. The R packages you had installed for other projects will remain unaffected.

1) Open this project in a new RStudio session (open the project folder and click on the `translation-elicitation.Rproj` file). 
2) In the R console, run the following lines of code:

```r
# install.packages("renv") # in case yopu need to install renv
renv::activate() # this will create the renv infrastructure if missing
renv::restore() # this will install/update/move the necessary R pakcages into this project
```

### 3. Running the code with targets

Restart your R session. The code in the .Rprofile file should run at start, and define the function `make()`. You should be able to run `make()` in your console. This will run all lines of code in the right order, and will generate all objects and outputs (e.g., data frames) used in the project, and render the manuscript.

Here are some details about the code, in case you want to dive into it. We use the targets package to keep track of what code should be run, and in what order. The `_targets.R` script contains the main commands in the workflow. Each command is indicated inside of a call to the `tar_target()` function, and is called a *target*. For readability, we have defined most of the functions used inside the targets on separate scripts, hosted in the R/ folder of the project. Take a look at them if you want to see what exactly what happens inside those function.

You can import the objects generated by the targets using the `tar_load()` function. First, load all necessary R packages running `tar_load_globals()`. Then, if you have previously run `make()`, you can run `tar_load(responses)` to get the main data set or `tar_load(model_fits)` to inspect the models, for instance. You don't have to assign the call to the `tar_load()` function to a variable: the object will be directly save as a variable in your workspace.

```{r targets, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

targets::tar_visnetwork()

```

## How is this repository organised?

This repository is organised as follows:

* **data**: raw, processed, and coded data from the behavioural task
* **img**: figures resulting from the scripts
* **docs**: lab notes
* **manuscript**: Rmarkdown files and outputs of the manuscript (generated using the {papaja} R package)
* **results**: outputs of the models.
* **scripts**: R code used to process and analyse the data, and Stan code used to fit the Bayesian models (called via the R scripts).
* **stimuli**: Stimuli and trial lists used in the behavioural task.

The file `.gitignore` indicates Git what files/subfolders must *not* be kept track of (for privacy and storage limit, e.g., audios).

The experimental set up and code (PsychoPy/Pavlovia) are stored in the following GitLab repositories:

* Version in Catalan for **Spanish speakers**: https://gitlab.pavlovia.org/gongcastro/translationelicitationspa
* Version in Spanish or Catalan for **English speakers**: https://gitlab.pavlovia.org/SiowSerene/translationelicitation_eng

### 2022-02-18 docs/03_results.Rmd
Render README as target

@@
   mutate_at(
-    vars(lv, pthn, frequency),
+    vars(lv_std, pthn_std, frequency_zipf_std),
     function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
@@
   "Intercept" = "Intercept",
-  "frequency_zipf" = "Frequency\n[+1 SD]",
-  "pthn:lv" = "PTHN \u00d7 Levenshtein", 
-  "lv:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
-  "lv:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
-  "lv" = "Levenshtein\n[+1 SD]",
-  "pthn" = "PTHN\n[+1 SD]",
+  "frequency_zipf_std" = "Frequency\n[+1 SD]",
+  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
+  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv_std" = "Levenshtein\n[+1 SD]",
+  "pthn_std" = "PTHN\n[+1 SD]",
   "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
@@
   "Intercept" = "Intercept",
-  "frequency_zipf" = "Frequency\n[+1 SD]",
-  "pthn:lv" = "PTHN \u00d7 Levenshtein", 
-  "lv:group1" = "Levenshtein \u00d7 Group 1",
-  "lv:group2" = "Levenshtein \u00d7 Group 2",
-  "lv" = "Levenshtein\n[+1 SD]",
-  "pthn" = "PTHN\n[+1 SD]",
+  "frequency_zipf_std" = "Frequency\n[+1 SD]",
+  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
+  "lv_std:group1" = "Levenshtein \u00d7 Group 1",
+  "lv_std:group2" = "Levenshtein \u00d7 Group 2",
+  "lv_std" = "Levenshtein\n[+1 SD]",
+  "pthn_std" = "PTHN\n[+1 SD]",
   "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
@@
 ```{r post_re_word, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-fixed_intercept <- fixef(model_fits$fit_6)[1,c(1, 3, 4)] %>% 
+fixed_intercept <- fixef(fit_6)[1,c(1, 3, 4)] %>% 
   inv_logit_scaled()
@@
 
-ggplot(m, aes(lv, .epred, colour = pthn, fill = pthn)) +
+ggplot(m, aes(lv_std, .epred, colour = pthn_std, fill = pthn_std)) +
   facet_wrap(~group) +
   geom_hline(yintercept = 0.5, colour = "grey") +
-  geom_line(aes(group = interaction(pthn, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
-  stat_summary(aes(linetype = pthn), fun = mean, geom = "line", 
+  geom_line(aes(group = interaction(pthn_std, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
+  stat_summary(aes(linetype = pthn_std), fun = mean, geom = "line", 
                colour = "black", size = 0.75, show.legend = TRUE) +

### 2022-02-18 R/00_stimuli.R

@@ -1,0 +1,84 @@
+# relative frequencies and phonological neighbours
+get_clearpond <- function(clearpond_path){
+    clearpond <- clearpond_path %>% 
+        map(
+            function(x){
+                read.csv(x) %>% 
+                    clean_names() %>% 
+                    select("word", "pho_word", "freq_per_million", "pthn")
+            }
+        ) %>% 
+        bind_rows(.id = "group") %>% 
+        as_tibble() %>%
+        distinct(word, group, .keep_all = TRUE) %>% 
+        rename(word2 = word, phon_clearpond = pho_word, frequency = freq_per_million)
+    return(clearpond)
+}
+
+# get Leveshtein similarity scores
+get_levenshtein <- function(stimuli_path){
+    
+    levenshtein <- map(
+        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
+        function(x) read_xlsx(stimuli_path, sheet = x)
+    ) %>% 
+        bind_rows(.id = "group") %>% 
+        clean_names() %>% 
+        select(group, word1, word2, ipa1, ipa2) %>% 
+        mutate(
+            n_char = ifelse(nchar(ipa1) > nchar(ipa2), nchar(ipa1), nchar(ipa2)),
+            lv = stringsim(ipa1, ipa2, method = "lv")
+        )
+    
+    return(levenshtein)
+}
+
+# get audio duration
+get_duration <- function(
+    stimuli_path,
+    audios_path
+){
+    stimuli <- map(
+        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
+        ~read_xlsx(stimuli_path, sheet = .)) %>% 
+        bind_rows(.id = "group") %>% 
+        select(word1, group, file) %>% 
+        mutate(file = here("stimuli", "sounds", file))
+    
+    audios <- map(stimuli$file, load.wave) 
+    lengths <- map_dbl(audios, length)/2
+    sampling_rate <- unique(map_dbl(audios, ~attr(., "rate")))
+    duration <- lengths/sampling_rate
+    return(duration)
+}
+
+# process stimuli and add information
+get_stimuli <- function(
+    stimuli_path,
+    clearpond,
+    levenshtein,
+    durations
+){
+    
[…]

### 2022-02-18 R/01_responses.R

@@ -1,0 +1,230 @@
+# preprocess responses
+get_responses_processed <- function(
+    responses_path,
+    stimuli,
+    practice_trials,
+    seed = 888
+){
+    # set parameters
+    set.seed(seed) # for reproducibility
+    spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
+    
+    filenames <- list.files(here("data", "raw")) %>% 
+        str_extract(".*?\\_") %>% 
+        str_remove("_")
+    
+    # participant files
+    processed <- responses_path %>% 
+        map(fread, na.string = "") %>%  # import participant files
+        map(mutate, participant = as.character(participant)) %>% 
+        set_names(filenames) %>% # label each dataset with the participant's ID
+        bind_rows(.id = "filename") %>% # merge datasets
+        as_tibble() %>%
+        clean_names() %>% 
+        # select relevant variables and rename if necessary
+        select(
+            participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
+            matches("language_l|language_s|language_c|demo_|setup_"),
+            -matches("_rt|_time"),
+            key_press_time
+        ) %>% 
+        rename(demo_impairment = demo_language_key_keys) %>% 
+        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
+        # clean text input by participants (because of typos of need to translate) and redefine location    
+        group_by(participant) %>% 
+        mutate_at(
+            vars(
+                starts_with("demo_"),
+                starts_with("l"),
+                starts_with("setup_"),
+                age, city, setup_location, setup_noise
+            ),
+            first_non_na) %>% 
+        mutate(date = max(date, na.rm = TRUE)) %>% 
+        ungroup() %>% 
+        mutate_at(vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), clean_input_text) %>% 
+        mutate(
+            city = str_to_sentence(city),
+            country = ifelse(city %in% spanish_cities, "Spain", "UK"),
+            date = as_datetime(ymd_hms(date)),
+            demo_sex = case_when(
+                country %in% "UK" & demo_sex %in% "M" ~ "Male",
+                country %in% "UK" & demo_sex %in% "F" ~ "Female",
+                country %in% "Spain" & demo_sex %in% "M" ~ "Female",
+                country %in% "Spain" & demo_sex %in% "H" ~ "Male"
+            ),
[…]

### 2022-02-18 _targets.R

@@ -1,92 +1,343 @@
 library(targets)
 library(tarchetypes)
 
-source("scripts/R/utils.R")
-source("scripts/R/00_stimuli.R")
-source("scripts/R/01_responses.R")
-source("scripts/R/02_models.R")
+# load functions
+source("R/utils.R")
+source("R/00_stimuli.R")
+source("R/01_responses.R")
+source("R/02_models.R")
+
+# set number of cores to use with brms
+options(mc.cores = 1)
 
 # set parameters
 tar_option_set(
+    # manifest dependencies (via tar_renv)
     packages = c(
-        "dplyr", "tidyr", "stringr", "ggplot2", "tibble", "forcats", "multilex", "keyring",
-        "readxl", "janitor", "mice", "here", "lubridate", "purrr", "scales", "stringdist",
-        "brms", "tidybayes", "gt", "patchwork", "wesanderson", "papaja", "knitr", "data.table",
-        "audio", "tidytext", "bayesplot", "GGally", "performance"
+        "dplyr", 
+        "tidyr", 
+        "stringr",
+        "ggplot2",
+        "tibble", 
+        "forcats",
+        "multilex",
+        "keyring",
+        "readxl",
+        "janitor", 
+        "mice", 
+        "here", 
+        "lubridate", 
+        "purrr", 
+        "scales", 
+        "stringdist",
+        "brms", 
+        "tidybayes",
+        "gt", 
+        "patchwork", 
+        "wesanderson", 
+        "papaja", 
+        "knitr", 
+        "data.table",
+        "audio",
+        "tidytext",
+        "bayesplot",
+        "performance"
     )
 )
 
 list(
-    # stimuli
-    tar_target(clearpond_path, list(`ENG-CAT` = here("data", "clearpond", "clearpond_english.csv"), `ENG-SPA` = here("Data", "clearpond", "clearpond_english.csv"), `SPA-CAT` = here("Data", "clearpond", "clearpond_spanish.csv")),),
-    tar_target(clearpond, get_clearpond(clearpond_path)),
-    tar_target(stimuli_path, here("stimuli", "trials.xlsx")),
-    tar_target(levenshtein, get_levenshtein(stimuli_path = stimuli_path)),
-    tar_target(audios_path, here("stimuli", "sounds")),
-    tar_target(durations, get_duration(stimuli_path = stimuli_path, audios_path = audios_path)),
[…]

### 2022-02-18 scripts/R/01_responses.R

@@ -1,229 +1,0 @@
-# preprocess responses
-get_responses_processed <- function(
-    responses_path,
-    stimuli,
-    practice_trials,
-    seed = 888
-){
-    # set parameters
-    set.seed(seed) # for reproducibility
-    spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
-    
-    filenames <- list.files(here("data", "raw")) %>% 
-        str_extract(".*?\\_") %>% 
-        str_remove("_")
-    
-    # participant files
-    processed <- responses_path %>% 
-        map(fread, na.string = "") %>%  # import participant files
-        map(mutate, participant = as.character(participant)) %>% 
-        set_names(filenames) %>% # label each dataset with the participant's ID
-        bind_rows(.id = "filename") %>% # merge datasets
-        as_tibble() %>%
-        clean_names() %>% 
-        # select relevant variables and rename if necessary
-        select(
-            participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
-            matches("language_l|language_s|language_c|demo_|setup_"),
-            -matches("_rt|_time"),
-            key_press_time
-        ) %>% 
-        rename(demo_impairment = demo_language_key_keys) %>% 
-        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
-        # clean text input by participants (because of typos of need to translate) and redefine location    
-        group_by(participant) %>% 
-        mutate_at(
-            vars(
-                starts_with("demo_"),
-                starts_with("l"),
-                starts_with("setup_"),
-                age, city, setup_location, setup_noise
-            ),
-            first_non_na) %>% 
-        mutate(date = max(date, na.rm = TRUE)) %>% 
-        ungroup() %>% 
-        mutate_at(vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), clean_input_text) %>% 
-        mutate(
-            city = str_to_sentence(city),
-            country = ifelse(city %in% spanish_cities, "Spain", "UK"),
-            date = as_datetime(ymd_hms(date)),
-            demo_sex = case_when(
-                country %in% "UK" & demo_sex %in% "M" ~ "Male",
-                country %in% "UK" & demo_sex %in% "F" ~ "Female",
-                country %in% "Spain" & demo_sex %in% "M" ~ "Female",
-                country %in% "Spain" & demo_sex %in% "H" ~ "Male"
-            ),
[…]

### 2022-02-18 scripts/R/02_models.R

@@ -1,214 +1,0 @@
-# models
-
-get_model_fits <- function(
-    responses,
-    model_formulas,
-    model_prior,
-    iter = 500,
-    cores = 4,
-    chains = 4,
-    seed = 888
-){
-    # set parameters
-    options(mc.cores = cores, chains = chains, iter = iter, seed = seed)
-    
-    # fit models
-    fit_0 <- brm(
-        model_formulas$f_0,
-        data = responses,
-        family = bernoulli("logit"),
-        prior = model_prior[c(1, 3),],
-        save_pars = save_pars(all = TRUE),
-        backend = "cmdstanr",
-        seed = 888,
-        save_model = here("scripts", "stan", "fit_responses_0.stan"),
-        file = here("results", "fit_responses_0.rds"), 
-        control = list(adapt_delta = 0.95)
-    )
-    fit_1 <- brm(
-        model_formulas$f_1,
-        data = responses,
-        family = bernoulli("logit"),
-        prior = model_prior,
-        save_pars = save_pars(all = TRUE),
-        backend = "cmdstanr",
-        seed = 888,
-        save_model = here("scripts", "stan", "fit_responses_1.stan"),
-        file = here("results", "fit_responses_1.rds"), 
-        control = list(adapt_delta = 0.95)
-    )
-    fit_2 <- brm(
-        model_formulas$f_2,
-        data = responses,
-        family = bernoulli("logit"),
-        prior = model_prior,
-        save_pars = save_pars(all = TRUE),
-        backend = "cmdstanr",
-        seed = 888,
-        save_model = here("scripts", "stan", "fit_responses_2.stan"),
-        file = here("results", "fit_responses_2.rds"), 
-        control = list(adapt_delta = 0.95)
-    )
-    fit_3 <- brm(
-        model_formulas$f_3,
-        data = responses,
-        family = bernoulli("logit"),
-        prior = model_prior,
-        save_pars = save_pars(all = TRUE),
-        backend = "cmdstanr",
-        sample_prior = "yes",
-        seed = 888,
-        save_model = here("scripts", "stan", "fit_responses_3.stan"),
-        file = here("results", "fit_responses_3.rds"), 
-        control = list(adapt_delta = 0.95)
-    )
-    fit_4 <- brm(
-        model_formulas$f_4,
-        data = responses,
-        family = bernoulli("logit"),
-        prior = model_prior,
-        save_pars = save_pars(all = TRUE),
[…]

### 2022-02-18 scripts/R/utils.R

@@ -1,167 +1,0 @@
-# evaluate if x is NOT an element of y
-"%!in%" <- function(x, y) !(x %in% y)
-
-# find first/last non-missing value in vector
-first_non_na <- function(x) ifelse(is.logical(first(x[!is.na(x)])), NA, first(x[!is.na(x)]))
-last_non_na <- function(x) ifelse(is.logical(last(x[!is.na(x)])), NA, last(x[!is.na(x)]))
-add_big_mark <- function(x) format(x, big.mark = ",", scientific = FALSE)
-clean_input_text <- function(x) {
-    {{ x }} %>% 
-        str_to_sentence() %>%
-        str_replace("Lshift|space|minus", "") %>%
-        as.character() %>% 
-        as_tibble(x = .) %>% 
-        mutate(. , value = case_when( 
-            nchar(value) > 12 ~ "Several",
-            value %in% c("Francés", "Frances", "Basic french") ~ "French",
-            value %in% c("Italiano", "Italino") ~ "Italian",
-            value %in% c("Alemán") ~ "German",
-            value %in% c("NrLondon") ~ "London",
-            value %in% c("Rnorthampton") ~ "London",
-            value %in% c("Newcastleminusuponminustyne") ~ "Newcastle upon Tyne",
-            value %in% c("Newcastlespaceuk") ~ "Newcastle",
-            value %in% c("Miltonkeynes") ~ "Milton Keynes",
-            value %in% c("Irthlingboroug") ~ "Irthlingborough",
-            value %in% c("spalding", "berkshire", "bedfordshire") ~ str_to_sentence(value),
-            value %in% c("No", "Cuál", "Space") ~ NA_character_,
-            TRUE ~ value)) %>% 
-        pull(value)
-}
-# ggplot custom theme
-theme_custom <- function(){
-    theme(
-        panel.background = element_rect(fill = "transparent"),
-        panel.grid = element_blank(),
-        #panel.grid = element_line(colour = "grey", linetype = "dotted"),
-        panel.border = element_rect(fill = "transparent", colour = "black"),
-        text = element_text(colour = "black", size = 15),
-        axis.text = element_text(colour = "black")
-    )
-}
-
-# remove special characters
-replace_non_ascii <- function(x){
-    str_replace_all(
-        x,
-        c(
-            "á" = "a",
[…]

### 2022-02-22 _targets.R
Add abstract for talk at Psycholinguistics Coffee

@@ -330,13 +330,10 @@ list(
         )
     ),
     
-    # # render docs ----
-    tar_render(readme, "README.Rmd"),
-    
-    tar_render(docs_stimuli, "docs/00_stimuli.Rmd"),
-    tar_render(docs_participants, "docs/01_participants.Rmd"),
-    tar_render(docs_analysis, "docs/02_data-analysis.Rmd"),
-    tar_render(docs_results, "docs/03_results.Rmd"),
+    # render docs ----
+    # tar_render(readme, "README.Rmd"),
+
+    tar_render(docs, "docs/index.Rmd"),
 
     tar_render(manuscript, "manuscript/manuscript.Rmd")
 )

### 2022-02-22 communications/2022-03-16_psycholinguistics-coffee/abstract.Rmd
Add abstract for talk at Psycholinguistics Coffee

---
title: "The role of cognateness during word recognition in a non-native language"
author: "Gonzalo Garcia-Castro, Serene Siow, Núria Sebastian-Galles, and Kim Plunkett"
date: "16/03/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar. Phonological similarity plays a more important role in low-proficiency bilinguals, compared to high-proficiency bilinguals, and points to the latter as relying more strongly on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route). Disentangling bilinguals' reliance on the conceptual and the lexical route during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals. In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language. Given their lack of previous knowledge on the unfamiliar language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task. This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route. Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation. This effect was larger in Spanish participants translating Catalan words than in English participants translating Catalan or Spanish. These results point to cross-language word-to-word links as providing rich information during word translation.

### 2022-02-22 docs/index.Rmd
Add abstract for talk at Psycholinguistics Coffee

---
title: "Translation Elicitation: Report"
date: "Last updated: `r format(Sys.Date(), '%d/%m/%Y')`"
output:
  html_document:
    theme: paper
    highlight: pygments
    toc: true
    toc_depth: 2
    toc_float:
        collapsed: false
    number_sections: true
    keep_md: true
bibliography: "references.bib"
csl: "apa7.csl"
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
    message = FALSE,
    warning = FALSE,
    cache.extra = knitr::rand_seed,
    dev.args = list(png = list(type = "cairo")),
    dpi = 300,
    include = TRUE,
    out.width = "80%",
    results = "asis",
    echo = TRUE
)

options(
    knitr.kable.NA = '-',
    htmltools.preserve.raw = FALSE,
    knitr.duplicate.label = "allow",
    ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
    ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
    ggplot2.continuous.fill = scale_fill_gradient2(
        low = "#04bf78", mid = "white", high = "#FFC20A", 
        na.value = "white", limits = c(-0.4, 0.4)
    ),
    ggplot2.continuous.colour = scale_fill_gradient2(
        low = "#04bf78", mid = "white", high = "#FFC20A", 
        na.value = "white", limits = c(-0.4, 0.4)
    )
)
```

```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
# import data
tar_load_globals()
tar_load_all()

theme_set(
    theme(
        title = element_text(size = 8, face = "bold"),

        legend.background = element_rect(fill = NA, colour = NA),
        legend.key = element_rect(fill = "white", colour = NA),
        legend.justification = "right",
        legend.direction = "horizontal",

        axis.line = element_line(colour = "black", size = 0.5),
        axis.ticks = element_line(colour = "black", size = 0.5),
        axis.text.x = element_text(colour = "black"),
        axis.text.y = element_text(colour = "black"),

        plot.caption.position = "plot",
        plot.title = element_text(hjust = 0.5),

        panel.grid = element_blank(),
        panel.background = element_rect(fill = "white", colour = NA)
    )
)

```

# Stimuli

## Summary

Participants in the ENG-SPA listened to `r nrow(stimuli[stimuli$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials. Five trials in each condition were practice trials and were excluded from analyses.

```{r stimuli-coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}

stimuli_coded <- stimuli %>% 
    mutate(test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"))
```

```{r stimuli-table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli %>% 
    select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
    group_by(group) %>% 
    summarise_at(
        vars(frequency:lv), 
        list(
            mean = ~mean(., na.rm = TRUE),  
            sd = ~sd(., na.rm = TRUE)
        )
    ) %>%
    gt(rowname_col = "group") %>% 
    tab_header(
        title = md("Table 1"),
        subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
    ) %>% 
    tab_source_note(
        md("*Note*. Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
    ) %>% 
    fmt_number(frequency_mean:pthn_sd) %>%
    fmt_percent(matches("lv")) %>% 
    tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
    tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
    tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
    tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
    cols_label(
        frequency_mean = md("Mean"),
        frequency_sd = md("*SD*"),
        frequency_zipf_mean = md("Mean"),
        frequency_zipf_sd = md("*SD*"),
        pthn_mean = md("Mean"),
        pthn_sd = md("*SD*"),
        lv_mean = md("Mean"),
        lv_sd = md("*SD*")
    ) %>% 
    summary_rows(
        columns = c(matches("mean"), -matches("lv")),
        fns = list(
            Mean = "mean",
            SD = "sd"
        )
    ) %>% 
    summary_rows(
        columns = matches("lv"),
        fns = list(
            Mean = "mean",
            SD = "sd"
        ), formatter = fmt_percent
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_stub_grand_summary()
    ) %>% 
    tab_style(
        style = cell_borders(sides = c("left", "right"), color = "white"),
        locations = list(
            cells_body(), cells_row_groups(), cells_column_spanners(), 
            cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
            cells_column_labels(), cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
        locations = cells_stub()
    ) %>% 
    tab_style(
        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
        locations = list(cells_column_spanners(), cells_column_labels())
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:9)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:9)
    )

```

## Lexical frequency

```{r stimuli-frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
    distinct(test_language, word1, .keep_all = TRUE) %>% 
    left_join(count(., test_language, name = "n_total")) %>% 
    mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(frequency_zipf, fill = test_language)) +
    facet_wrap(~test_language, ncol = 1) +
    geom_histogram(bins = 25, colour = "white") +
    labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language")  +
    scale_y_continuous(breaks = seq(0, 12, 2)) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

## PTHN

```{r stimuli-pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
    left_join(count(., test_language, name = "n_total")) %>% 
    mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
    count(test_language, pthn) %>% 
    ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
    facet_wrap(~test_language, ncol = 1) +
    geom_col(position = position_dodge()) +
    labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

## Levenshtein distance

```{r stimuli-similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
stimuli_coded %>% 
    left_join(count(., group, name = "n_total")) %>% 
    mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(group, lv, fill = group, colour = group)) +
    geom_violin() +
    geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
    geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
    labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
    theme(
        legend.position = "none",
        axis.title.x = element_blank(),
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

## Distributions

```{r stimuli-plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of stimuli translations, plotted in different panels for English translations of Catalan stimuli (cat-ENG), English translations of Spanish stimuli (spa-ENG), and Spanish translation of Catalan stimuli (cat-SPA). A\\) Lexical frequency of stimuli translations, expressed in counts per million and as Zipf scores. B\\) Number of higher-frequency phonological neighbourhoods (PTHN) of stimuli translations. C\\) Normalised Levenshtein phonological similarity between the presented stimuli and their translations, expressed in percentages. Lexical frequencies and PTHN scores of the stimuli (presented auditorily in Catalan or Spanish) are not described, as participants lack any lexical representation of such words."}

medians <- stimuli %>% 
    group_by(group) %>% 
    summarise(
        pthn_median = median(pthn),
        pthn_sd = sd(pthn),
        frequency_zipf_median = median(frequency_zipf),
        frequency_zipf_sd = sd(frequency_zipf),
        lv_median = median(lv),
        lv_sd = sd(lv)
    ) %>% 
    mutate(group = case_when(
        group=="ENG-CAT" ~ "cat-ENG",
        group=="ENG-SPA" ~ "spa-ENG",
        group=="SPA-CAT" ~ "cat-SPA",
    ))

# frequency
stimuli %>% 
    mutate(group = case_when(
        group=="ENG-CAT" ~ "cat-ENG",
        group=="ENG-SPA" ~ "spa-ENG",
        group=="SPA-CAT" ~ "cat-SPA",
    )) %>% 
    ggplot(aes(frequency_zipf, fill = group)) + 
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(
            xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd, 
            ymin = -Inf, ymax = Inf, fill = group
        ), inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_histogram(bins = 20, colour = "white") +
    geom_vline(
        data = medians, 
        aes(xintercept = frequency_zipf_median, group = group), 
        colour = "black",
        size = 0.5,
        linetype = "dashed"
    ) + 
    geom_text(
        data = medians, 
        aes(
            label = "Median ± SD", 
            x = frequency_zipf_median+1.25,
            y = 13
        ), 
        size = 3
    ) +
    geom_curve(
        data = medians, 
        aes(
            x = frequency_zipf_median+0.85, 
            xend = frequency_zipf_median+0.1,
            y = 12, 
            yend = 11
        ),
        arrow = arrow(length = unit(0.2, "cm")),
        curvature = -0.25,
        colour = "grey30"
    ) +
    labs(
        x = "Lexical frequency (Zipf)", 
        y = "Number of trials", 
        title = "Frequency\n(stimuli translations)\n"
    ) +

    # PTHN
    stimuli %>% 
    count(group, pthn) %>% 
    mutate(group = case_when(
        group=="ENG-CAT" ~ "cat-ENG",
        group=="ENG-SPA" ~ "spa-ENG",
        group=="SPA-CAT" ~ "cat-SPA",
    )) %>% 
    ggplot(aes(pthn, n, fill = group)) + 
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(
            xmin = 0, 
            xmax = pthn_median+pthn_sd, 
            ymin = -Inf, 
            ymax = Inf,
            fill = group
        ), inherit.aes = FALSE, 
        alpha = 0.15
    ) +   
    geom_col(colour = "white") + 
    geom_vline(
        data = medians, aes(xintercept = pthn_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed"
    ) + 
    labs(
        x = "PTHN", 
        y = "Number of trials", 
        title = "Higher-frequency neighbours\n(stimuli translations)\n"
    ) +

    # Levenshtein
    stimuli %>% 
    mutate(group = case_when(
        group=="ENG-CAT" ~ "cat-ENG",
        group=="ENG-SPA" ~ "spa-ENG",
        group=="SPA-CAT" ~ "cat-SPA",
    )) %>% 
    ggplot(aes(lv, fill = group)) +
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(
            xmin = 0, xmax = lv_median+lv_sd, 
            ymin = -Inf, ymax = Inf, fill = group
        ), inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_histogram(bins = 20, colour = "white") +
    geom_vline(
        data = medians, aes(xintercept = lv_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed"
    ) + 
    scale_x_continuous(labels = percent) +
    labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity\n(stimuli and stimuli translations)\n") +

    plot_layout(guides = "collect") &
    plot_annotation(tag_levels = "A") &
    scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
    theme(legend.position = "none")

```

# Participants

```{r participants_coded, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants_coded <- participants %>% 
    mutate(valid_participant = ifelse(valid_participant, "Valid", "Not valid"))
```

We collected data from `r format(as_date(min(participants$date)), "%B %dth %Y")` to `r format(as_date(max(participants$date)), "%B %dth %Y")`. We tested `r length(unique(participants$participant))` participants. `r length(unique(participants$participant[participants$group=="ENG-SPA"]))` were English natives tested in Spanish, `r length(unique(participants$participant[participants$group=="ENG-CAT"]))` were English natives tested in Catalan, and `r length(unique(participants$participant[participants$group=="SPA-CAT"]))` were Spanish natives tested in Catalan. `r sum(!participants$valid_participant)` failed to meet all the inclusion criteria and were excluded from further analyses. The final sample comprised data from `r sum(!participants$valid_participant)` participants, `r sum(participants$valid_participant[participants$group=="ENG-SPA"])` in the ENG-SPA group, `r sum(participants$valid_participant[participants$group=="ENG-CAT"])` in the ENG-CAT group, and `r sum(participants$valid_participant[participants$group=="SPA-CAT"])` in the SPA-CAT group.

Participants were included if^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct. This item was phrased as Do you have normal or corrected-to-normal VISION? (Yes/No) in English, and as ¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan? (Sí/No). Surprisingly, the proportion of Spanish participants that reported visual impairment was implausibly large (n = 6, 18.18%). This is possibly due to some participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision]:

* Aged 18 to 25 years
* Did not report being fluent in Catalan, Spanish, or Italian
* Provided at least 80% of the presented trials
* Did not report motor, auditory or visual (other than wearing glasses) problems

Trials were considered valid if:

* Participant did not leave a blank response
* Participant took longer than 10 seconds to respond

## Summary

```{r participants_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_coded %>% 
    count(group, valid_participant) %>% 
    mutate(valid_participant = factor(valid_participant)) %>% 
    ggplot(aes(group, n, fill = valid_participant, order = n)) +
    geom_col() +
    labs(x = "Group", y = "# participants", fill = "Valid?") +
    theme(
        legend.title = element_blank(),
        legend.position = "top",
        axis.title.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )

```

## Valid trials

```{r participants_stimuli, echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_coded %>% 
    filter(valid_participant=="Valid") %>% 
    select(participant, group, n, n_valid) %>%
    mutate(valid_prop = n_valid/n) %>% 
    ggplot(aes(valid_prop, fill = group, colour = group)) +
    facet_wrap(~group, ncol = 1) +
    geom_histogram() +
    geom_rug() +
    labs(x = "Valid trials (%)", y = "N participants", fill = "Group", colour = "Group") +
    scale_x_continuous(limits = c(0, 1), labels = percent) +
    theme(
        legend.title = element_blank(),
        legend.position = "none",
        axis.title.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )
```

## Demographic information

```{r participants_demo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants_valid <- filter(participants_coded, valid_participant=="Valid")

p_sex <- participants_valid %>% 
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
    ggplot(aes(group, fill = sex)) +
    geom_bar(position = position_fill()) +
    labs(x = "Group", y = "# participants", fill = "Sex") +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        axis.title.x = element_blank()
    )

p_age <- participants_valid %>% 
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(as.factor(age), fill = group, colour = group)) +
    facet_wrap(~group, ncol = 1) +
    geom_bar() +
    labs(x = "Age (years)", y = "% participants", fill = "Group", colour = "Group") +
    theme(
        legend.position = "none"
    )

(p_sex + p_age) &
    plot_layout() &
    plot_annotation(tag_levels = "A") &
    theme(
        legend.title = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )

```

## Second language

```{r participants_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_valid %>% 
    count(group, l2) %>% 
    mutate(
        l2 = case_when(
            l2=="Alemã\u0081n" ~ "German",
            l2=="Francã\u2030s" ~ "French",
            l2=="Cuã\u0081l" ~ "Missing",
            TRUE ~ l2
        ),
    ) %>% 
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
    ggplot(aes(group, n, fill = l2)) +
    facet_wrap(~group, ncol = 1, scales = "free_y") +
    geom_col(position = position_fill()) +
    labs(x = "L2", y = "# participants", fill = "L2") +
    scale_y_continuous(labels = percent) +
    coord_flip() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        strip.text = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(angle = 90),
        axis.title.y = element_blank()
    )

```

## Language profile proficiency

```{r participants_proficiency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants_valid %>% 
    select(participant, group, catalan_oral, spanish_oral, catalan_written, spanish_written) %>% 
    mutate_at(vars(catalan_oral, spanish_oral, catalan_written, spanish_written), replace_na, 0) %>% 
    pivot_longer(c(catalan_oral, spanish_oral, catalan_written, spanish_written),
                 names_to = "type", values_to = "score") %>% 
    separate(type, c("language", "modality"), sep = "_") %>% 
    mutate_at(vars(language, modality), str_to_sentence) %>%
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(as.integer(score), fill = language)) +
    facet_grid(modality~group) +
    geom_bar(position = position_dodge()) +
    labs(y = "# participants", x = "Self-reported proficiency (0-5)", fill = "Language") +
    theme(
        legend.position = "top",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

# Data analysis

We fitted a Bayesian multilevel regression model using the R package `brms` [@burkner2017brms], with correct responses (`correct`, 0 is incorrect response, 1 if correct response) as the response variable with a Bernoulli distribution and a logit link function. We modelled the probability of an average participant providing a correct response, conditional to a set of predictors (aka. fixed effects) and their interactions:

* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database. Standardised before entering the regression model.
* Phonological similarity (`lv`, 0-1): Inverse Levenshtein distance between the IPA transcription of the presented word and its correct translation in the target language, calculated using the `stringdist` function of the `stringdist` package [@loo2014stringdist]. Standardised before entering the regression model.
* Shared phonological onset (`onset`, Same/Different): Whether the presented and the target word share phonological onset, as judged by experimenters, and sum coded as $Different = -0.5$ and $Same = +0.5$.

We added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model. We specified random intercepts slopes for all fixed effects within the `participant` grouping variable.

We performed multiple imputation via predictive mean matching using the `mice` function of the `mice` package [@van2011mice] to impute missing values in the response variable or the predictors.

The model was implemented using the following formula (`fit_3`):

$$\begin{align}

&\textbf{Likelihood}  \\
y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\

&\textbf{Parameters}  \\

logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
& \beta_{1[p]} ~ Frequency_{i} ~ + \\
& \beta_{2[p]} ~ PTHN_i ~ + \\
& \beta_{3[p]} ~ Similarity_i ~ + \\
& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\

\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
&& \text{[participant-level coefficients]} \\ \\

&\textbf{Prior}  \\

\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\

\end{align}$$

The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 iterations each (1,000 warm-ups iterations). The output of this model is the approximated posterior distribution of all the parameters estimated, therefore indicating the probability of each combination of values of the parameters, conditional to the observed data.

We compared this (extended model) to other models dropping one predictor at a time using Leave-one-out cross-validation (LOO-CV), using the `loo` and `loo_compare` functions of the `brms` package. We compared our extended model against the following models:

* fit_0: `~ 1 + (1 | participant) + (1 | word)`
* fit_1: `~ 1 + frequency_zipf + (1 + frequency_zipf | participant) + (1 | word)`
* fit_2: `~ 1 + frequency_zipf + pthn + (1 + frequency_zipf + pthn | participant) + (1 | word)`
* fit_3: `~ 1 + frequency_zipf + pthn + lv + (1 + frequency_zipf + pthn + lv | participant) + (1 | word)`
* fit_4: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_5: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_6: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + group:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`

# Results

```{r responses_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
responses_coded <- responses %>% 
    # typos are considered correct responses
    mutate_at(vars(group), as.factor) %>% 
    # center predictors
    mutate_at(
        vars(lv_std, pthn_std, frequency_zipf_std),
        function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
    # impute missing data
    mice(m = 5, print = FALSE, method = "pmm") %>% 
    complete() %>% 
    as_tibble() %>% 
    arrange(group, word)
```

## Model comparison

Leave-one-out cross-validation. The more negative is `elpd_diff` (and the larger its magnitude compared to its corresponding standard error), the better it first the data compared to the null model (`~ 1 + (1 | participant)`). The model that fits the data the best is the extended model.

```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

model_loos %>% 
    as.data.frame() %>% 
    rownames_to_column("model") %>% 
    arrange(desc(elpd_loo)) %>% 
    select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
    mutate(model = str_replace(model, "fit_", "Model ")) %>% 
    mutate_all(~ifelse(. == 0, NA, .)) %>% 
    gt(rowname_col = "model") %>% 
    tab_header(
        title = md("Table 2"),
        subtitle = md("Model performance and comparison using leave-one-out cross-validation")
    ) %>% 
    tab_source_note(
        md("*Note*. Outcomes of the leave-one-out cross-validation procedure. Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model. Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
    ) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_number(2:7) %>% 
    cols_label(
        model = md("**Model**"),
        looic = md("***LOO<sub>IC</sub>***"),
        se_looic = md("***SE*<sub>IC</sub>**"),
        elpd_loo = md("***LOO<sub>ELPD</sub>***"),
        se_elpd_loo = md("***SE***"),
        elpd_diff = md("***LOO<sub>diff</sub>***"),
        se_diff = md("***SE<sub>diff</sub>***")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = list(cell_borders(sides = c("left", "right"), color = "white"),
                     cell_text(align = "center")),
        locations = list(
            cells_body(), cells_row_groups(), cells_column_spanners(), 
            cells_stub(), cells_source_notes(),
            cells_column_labels(), cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(cell_text(align = "left", size = "medium"), 
                     cell_borders(sides = "all", color = "white")),
        locations = list(cells_stub(), cells_source_notes())
    ) %>% 
    tab_style(
        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
        locations = list(cells_column_spanners(), cells_column_labels())
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:7)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:7)
    ) %>% 
    as_raw_html()

```

### Fixed coefficients

```{r coefs_fixed_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

coefs <- fixef(fit_6) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4))

str_repl <- c(
    "Intercept" = "Intercept",
    "frequency_zipf_std" = "Frequency\n[+1 SD]",
    "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
    "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
    "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
    "lv_std" = "Levenshtein\n[+1 SD]",
    "pthn_std" = "PTHN\n[+1 SD]",
    "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
    "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
)

coefs %>% 
    mutate(
        variable = str_remove(variable, "b_") %>% 
            str_replace_all(str_repl) %>%  
            factor(
                levels = c(
                    "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
                    "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
                    "Group 1\n[ENG-* vs. SPA-CAT]",
                    "Group 2\n[ENG-CAT vs. ENG-SPA]",
                    "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
                )
            )
    ) %>% 
    select(-est_error) %>% 
    gt() %>% 
    fmt_percent(2:4) %>% 
    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
    cols_label(
        variable = md("**Predictor**"),
        estimate = md("**Mean**"),
        q2_5 = md("**95\\% CrI**"),
    ) %>% 
    tab_source_note("Estimated posterior distributions of coefficients in Model 6. A\\) Population-level effects. Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects. Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability. Black dots represent the mean of each distribution.") %>% 
    cols_align(align = "left") %>% 
    as_raw_html()
```

```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4. A\\) Population-level effects. Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects. Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability. Black dots represent the mean of each distribution. B\\) Participant\\-level coefficient variability. The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant. Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*). C\\) Correlation between participant\\-level effects. The model allowed participant-level coefficients to co\\-vary. This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}

coefs <- fixef(fit_6) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
    group_split(variable) %>% 
    set_names(make_clean_names(map(., "variable"))) %>% 
    map(select, -variable) %>% 
    map(unlist)

str_repl <- c(
    "Intercept" = "Intercept",
    "frequency_zipf_std" = "Frequency\n[+1 SD]",
    "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
    "lv_std:group1" = "Levenshtein \u00d7 Group 1",
    "lv_std:group2" = "Levenshtein \u00d7 Group 2",
    "lv_std" = "Levenshtein\n[+1 SD]",
    "pthn_std" = "PTHN\n[+1 SD]",
    "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
    "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
)

# fixed effects
post_fix <- posterior_draws_fixed %>% 
    filter(str_detect(.variable, "b_")) %>% 
    mutate(
        .variable_name = str_remove(.variable, "b_") %>% 
            str_replace_all(str_repl) %>%  
            factor(
                levels = c(
                    "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
                    "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
                    "Group 1\n[ENG-* vs. SPA-CAT]",
                    "Group 2\n[ENG-CAT vs. ENG-SPA]",
                    "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
                )
            ),
        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
    ) %>% 
    arrange(.variable) %>% 
    ggplot(aes(.value, fct_rev(.variable_name))) +
    geom_vline(xintercept = 0) +
    stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
    stat_pointinterval(.width = 0, point_size = 1) +
    scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
    scale_x_continuous(labels = function(x) percent(round(x, 1))) +
    labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
    theme(
        legend.position = c(1, 0.25),
        legend.direction = "vertical",
        legend.background = element_rect(fill = "grey95"),
        legend.key = element_rect(fill = "grey95", colour = "grey95"), 
        legend.key.height = unit(0.1, "cm"),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        axis.text.x = element_text(colour = "black"),
        axis.text.y = element_text(colour = "black", hjust = 1),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_line(colour = "grey", size = 0.5)
    )

# SD of random effects
post_sd_data <- posterior_draws_fixed %>% 
    filter(str_detect(.variable, "sd_")) %>% 
    mutate(
        .variable_name = str_remove(.variable, "sd_participant__|sd_word__") %>% 
            str_replace_all(str_repl) %>% 
            factor(
                levels = c(
                    "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
                    "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
                    "Group 1\n[ENG-* vs. SPA-CAT]",
                    "Group 2\n[ENG-CAT vs. ENG-SPA]",
                    "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
                )
            ),
        group = case_when(
            str_detect(.variable, "participant") ~ "Participants",
            str_detect(.variable, "word") ~ "Words"
        ),
        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
    )

post_sd <- ggplot(filter(post_sd_data, group=="Participants"), aes(.value, fct_rev(.variable_name))) +
    stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
    stat_pointinterval(.width = 0, point_size = 1) +
    scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
    scale_x_continuous(labels = percent, limits = c(0, 1)) +
    labs(x = "SDs in P(Correct)", title = "By participant", fill = "CrI") +
    theme(axis.text.x = element_blank()) +

    ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
    stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
    stat_pointinterval(.width = 0, point_size = 1) +
    scale_x_continuous(labels = percent, limits = c(0, 1)) +
    labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +

    plot_layout(ncol = 1, heights = c(0.85, 0.15), guides = "collect") &
    theme(
        legend.position = "none",
        axis.text.y = element_text(hjust = 1),
        axis.title.x = element_blank(),
        legend.direction = "vertical",
        axis.title.y = element_blank(),
        legend.background = element_rect(fill = "grey95"),
        legend.key.height = unit(0.1, "cm"),
        legend.key = element_rect(fill = "grey95", colour = "grey95"),
        legend.text = element_text(size = 7),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_line(colour = "grey", size = 0.5)
    )

# Pearson correlations between random effects
corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)

corr_mat <- corr_mat %>% 
    as_tibble() %>% 
    select(matches("Estimate")) %>% 
    rename_all(str_remove, "Estimate.") %>% 
    mutate(term1 = corr_names)

corr_mat[lower.tri(corr_mat)] <- NA

corr_data <- corr_mat %>% 
    as.data.frame() %>% 
    pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
    drop_na(.value)

post_cors <- corr_data %>% 
    mutate(term1 = factor(term1, levels = names(str_repl), ordered = TRUE),
           term2 = factor(term2, levels = names(str_repl), ordered = TRUE)) %>% 
    mutate_at(vars(term1, term2), str_replace_all, str_repl) %>% 
    mutate(
        .value_label = case_when(term1==term2 ~ NA_character_, TRUE ~ printnum(.value, gt1 = FALSE)),
        .value = case_when(term1==term2 ~ NA_real_, TRUE ~ .value)
    ) %>% 
    filter(term1 != "PTHN × Levenshtein", term2 != "Intercept") %>% 
    ggplot(aes(fct_inorder(term1), fct_rev(fct_inorder(term2)), fill = .value)) +
    geom_tile(na.rm = TRUE) +
    geom_text(aes(label = .value_label), size = 3, na.rm = TRUE) +
    labs(x = "Term 1", y = "Term 2", fill = "Posterior\ncorrelation") +
    scale_fill_gradient2(
        low = "#04bf78", mid = "white", high = "#FFC20A", 
        na.value = "white", limits = c(-0.4, 0.4)
    ) +
    scale_x_discrete(position = "top") +
    coord_equal() +
    theme(
        legend.position = c(1, 1),
        legend.direction = "horizontal",
        legend.key.height = unit(0.1, "cm"),
        legend.key.width = unit(0.3, "cm"),
        legend.justification = c(1, 1),
        legend.title = element_blank(),
        legend.text = element_text(size = 5),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 6, angle = 90, hjust = 0, vjust = 1),
        axis.text.y = element_text(size = 6, vjust = 1, hjust = 1)
    )

post_fix / ((post_sd | post_cors) + plot_layout(widths = c(0.6, 0.4))) +
    plot_layout() +
    plot_annotation(tag_levels = "A")

```

## Random effects

```{r post_re_participant, echo=FALSE, fig.height=6, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}
post_re <- posterior_draws_random$participants %>% 
    mean_qi() %>% 
    ungroup() %>% 
    left_join(select(participants, participant, group)) %>% 
    arrange(group, participant, .param) %>% 
    mutate(.param_name = str_replace_all(.param, str_repl) %>%  
               factor(
                   levels = c(
                       "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
                       "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
                       "Group 1\n[ENG-* vs. SPA-CAT]",
                       "Group 2\n[ENG-CAT vs. ENG-SPA]",
                       "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
                   )
               ))

ggplot(post_re, 
       aes(.value, sort(as.factor(participant)), 
           xmin = .lower, xmax = .upper, colour = group)) +
    facet_grid(group~.param_name, scales = "free") +
    geom_vline(xintercept = 0, colour = "grey") +
    geom_errorbarh(height = 0, alpha = 0.75) +
    geom_point(size = 0.1, shape = 4) +
    labs(
        x = "Participant-level adjustment",
        y = "Participant", 
        colour = "Group",
        title = "By participant"
    ) +
    scale_color_manual(
        values = c("#1A85FF", "#ff2976", "#FFC20A"),
        na.translate = FALSE
    ) &
    theme(
        legend.position = "none",
        legend.title = element_blank(),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_blank(),
        axis.ticks = element_blank()
    )
```

```{r post_re_word, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
fixed_intercept <- fixef(fit_6)[1,c(1, 3, 4)] %>% 
    inv_logit_scaled()

post_re <- posterior_draws_random$words   %>% 
    ungroup() %>% 
    arrange(word, .param) %>% 
    mutate(
        .param_name = str_replace_all(.param, str_repl) %>%  
            factor(
                levels = c(
                    "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
                    "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
                    "Group 1\n[ENG-* vs. SPA-CAT]",
                    "Group 2\n[ENG-CAT vs. ENG-SPA]",
                    "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
                )
            ),
        .value = inv_logit_scaled(.value)
    )

ggplot(post_re, aes(.value, reorder(as.factor(word), .value))) +
    geom_vline(xintercept = fixed_intercept[1], colour = "grey", size = 1) +

    stat_interval(
        size = 1.5,
        .width = c(0.5, 0.67, 0.95)
    ) +
    annotate(
        geom = "text",
        x = 0.7,
        y = 100,
        size = 3,
        hjust = 0.5,
        label = "Item accuracy \n(95%, 67%, and 50% CrI \nof item-level intercepts)"
    ) +
    annotate(
        geom = "rect",
        alpha = 0.15,
        xmin = fixed_intercept[2],
        xmax = fixed_intercept[3],
        ymin = -Inf,
        ymax = Inf
    ) +
    annotate(
        x = 0.7,
        xend = 0.6,
        y = 108,
        yend = 130,
        geom = "curve",
        curvature = 0.25,
        arrow = arrow(length = unit(0.2, "cm"))
    ) +
    annotate(
        geom = "text",
        x = 0.8,
        y = 10,
        size = 3,
        hjust = 0.5,
        label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
    ) +
    annotate(
        x = 0.7,
        xend = fixed_intercept[1]+0.05,
        y = 17,
        yend = 30,
        geom = "curve",
        curvature = 0.2,
        arrow = arrow(length = unit(0.2, "cm"))
    ) +
    labs(
        x = "Posterior probability of correct response",
        y = "Item",
        colour = "Credible Interval (CrI)",
        caption = "Credible intervals cover the range of values of the sampling space that \ncontain the true value with a given certainty."
    ) +
    scale_color_manual(
        values = c("#d2e5fc", "#9ccaff", "#1A85FF"),
        na.translate = FALSE, 
        labels = paste0(c(95, 67, 50), "%")
    ) +
    scale_x_continuous(limits = c(0, 1), labels = percent) +
    theme(
        legend.position = c(0.1, 0.9),
        legend.direction = "vertical",
        legend.justification = "left",
        legend.key = element_rect(fill = NA, colour = NA),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 10),
        plot.caption.position = "plot",
        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm")))
    )

ggsave(here("img", "accuracy.png"), height = 9, width = 5, dpi = 500)
```

## Marginal effects

```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=6}

m <- posterior_epreds_fixed %>% 
    mutate_at(vars(pthn_std, frequency_zipf_std), function(x) paste0(x, " SD")) %>% 
    mutate(
        pthn_std = case_when(
            pthn_std=="-1 SD" ~ "Low PTHN",
            pthn_std=="0 SD" ~ "Mean PTHN",
            pthn_std=="1 SD" ~ "High PTHN"
        ),
        frequency_zipf_std = case_when(
            frequency_zipf_std=="-1 SD" ~ "Low frequency",
            frequency_zipf_std=="0 SD" ~ "Mean frequency",
            frequency_zipf_std=="1 SD" ~ "High frequency"
        )
    )

ggplot(m, aes(lv_std, .epred, colour = pthn_std, fill = pthn_std)) +
    facet_wrap(~group) +
    geom_hline(yintercept = 0.5, colour = "grey") +
    geom_line(aes(group = interaction(pthn_std, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
    stat_summary(aes(linetype = pthn_std), fun = mean, geom = "line", 
                 colour = "black", size = 0.75, show.legend = TRUE) +
    annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, 
             arrow = arrow(length = unit(0.1, "cm")), size = .5) +
    annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, 
             arrow = arrow(length = unit(0.1, "cm")), size = .5) +
    annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
    annotate("text", x = 2.15, y = 0.95, size = 2.75, label = "More similar") +
    labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
         colour = "PTHN", fill = "PTHN", linetype = "PTHN",
         title = "Population-level predictions") +
    scale_color_manual(values = c("#1E88E5", "#ff2976")) +
    scale_y_continuous(
        labels = function(x) percent(round(x, 2)),
        limits = c(0, 1),
        breaks = seq(0, 1, 0.1)
    ) +
    theme(
        legend.position = "top",
        plot.title = element_blank()
    )

```

## Accuracy

### Accuracy by item

```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
responses_item <- responses %>%
    group_by(group, word) %>%
    summarise(
        correct_sum = sum(correct, na.rm = TRUE),
        correct_n = sum(!is.na(correct), na.rm = TRUE),
        .groups = "drop"
    ) %>% 
    mutate(
        correct_prop = prop_adj(correct_sum, correct_n),
        correct_se = prop_adj_se(correct_sum, correct_n),
        word_label = reorder_within(word, by = correct_prop, within = group)
    )

ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
    facet_wrap(~group, scales = "free_y") +
    geom_hline(yintercept = 0.5, colour = "black") +
    geom_errorbar() +
    geom_point(size = 0.5) +
    labs(y = "Correct responses", x = "Word", colour = "Group") +
    coord_flip() +
    scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
    scale_y_continuous(limits = c(0, 1), labels = percent) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 9),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
    )

```

### Accuracy by participant

```{r responses_participant, echo=FALSE, fig.height=7, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}
responses_participant <- responses %>%
    group_by(group, participant) %>%
    summarise(
        correct_sum = sum(correct, na.rm = TRUE),
        correct_n = sum(!is.na(correct), na.rm = TRUE),
        .groups = "drop"
    ) %>% 
    mutate(
        correct_prop = prop_adj(correct_sum, correct_n),
        correct_se = prop_adj_se(correct_sum, correct_n),
        participant_label = reorder_within(participant, by = correct_prop, within = group)
    )

ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
    facet_wrap(~group, scales = "free_y") +
    geom_hline(yintercept = 0.5) +
    geom_errorbar(width = 0.5) +
    geom_point(size = 1) +
    labs(
        y = "Correct responses",
        x = "Participant",
        colour = "Group"
    ) +
    coord_flip() +
    scale_color_manual(
        values = c("#1A85FF", "#ff2976", "#FFC20A"),
        na.translate = FALSE
    ) +
    scale_y_continuous(
        limits = c(0, 1), 
        labels = scales::percent
    ) +
    theme(
        legend.position = "none",
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        axis.text.x = element_text(size = 9),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
    )

```

## Model diagnostics

```{r traceplots, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

posterior_draws_fixed %>% 
    mutate(
        .chain = as.factor(.chain),
        .variable_name = str_replace_all(.variable, str_repl) %>% 
            str_remove("b_")
    ) %>% 
    ggplot(aes(.iteration, .value, colour = .chain)) +
    facet_wrap(~.variable_name, scales = "free_y") +
    geom_line() +
    labs(x = "Iteration", y = "Value", color = "Chain") +
    scale_colour_manual(values = c("#1A85FF", "#ff2976", "#FFC20A", "04bf78"), na.translate = FALSE) +
    theme(legend.position = "top")

```

### 2022-02-22 manuscript/manuscript.Rmd
Add abstract for talk at Psycholinguistics Coffee

@@
   Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar.
+Phonological similarity plays a more important role in low-proficiency bilinguals, compared to high-proficiency bilinguals, and points to the latter as relying more strongly on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
+Disentangling bilinguals' reliance on the conceptual and the lexical route during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals.
+In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
+Given their lack of previous knowledge on the unfamiliar language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
+This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route.
+Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation.
+This effect was larger in Spanish participants translating Catalan words than in English participants translating Catalan or Spanish.
+These results point to cross-language word-to-word links as providing rich information during word translation.
 

### 2022-03-10 R/01_responses.R
Minor refactoring

@@ -174,6 +174,9 @@ get_participants <- function(
                 # L2 is not a blocked one
                 l2 %!in% blocked_languages
         )
+    
+    saveRDS(participants, "results/participants.rds")
+    
     return(participants)
 }
 
@@ -224,6 +227,8 @@ get_responses <- function(
         "ENG vs. SPA" = c(-0.25, -0.25, 0.5), 
         "ENG-CAT vs. ENG-SPA" = c(0.5, -0.5, 0)
     )
+    
+    saveRDS(responses, "results/responses.rds")
     
     return(responses)
 }

### 2022-03-16 manuscript/manuscript.Rmd
Minor refactoring
WIP

@@
 We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
-For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim("teɪbəl", "mesa"))` similarity, while the *train* (trEIn)-*guitarra* (tɾEn) translation pair had a `r percent(stringsim("trEIn", "tɾEn"))` similarity.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim("teɪbəl", "mesa"))` similarity, while the *train* (trEIn)-*tren* (tɾEn) translation pair had a `r percent(stringsim("trEIn", "tɾEn"))` similarity.
 Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
@@
 
-$$
-\begin{align}
+$$\begin{align}
 
@@
 
-\end{align}
-$$
+\end{align}$$
 

### 2022-03-17 docs/index.Rmd
Add suggestions from @siowserene and refactor some code

@@
     list(
-            mean = ~mean(., na.rm = TRUE),  
+      median = ~median(., na.rm = TRUE),  
       sd = ~sd(., na.rm = TRUE)
@@
     md("*Note*.
-Mean and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
-    ) %>% 
-    fmt_number(frequency_mean:pthn_sd) %>%
+Median and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
+  ) %>% 
+  fmt_number(frequency_median:pthn_sd) %>%
   fmt_percent(matches("lv")) %>% 
-    tab_spanner(md("**Freq. (million)**"), c("frequency_mean", "frequency_sd")) %>% 
+  tab_spanner(md("**Freq. (million)**"), c("frequency_median", "frequency_sd")) %>% 
   tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
@@
   cols_label(
-        frequency_mean = md("Mean"),
+    frequency_median = md("Median"),
     frequency_sd = md("*SD*"),
-        frequency_zipf_mean = md("Mean"),
+    frequency_zipf_median = md("Median"),
     frequency_zipf_sd = md("*SD*"),
-        pthn_mean = md("Mean"),
+    pthn_median = md("Median"),
     pthn_sd = md("*SD*"),
-        lv_mean = md("Mean"),
+    lv_median = md("Median"),
     lv_sd = md("*SD*")
@@
   summary_rows(
-        columns = c(matches("mean"), -matches("lv")),
+    columns = c(matches("median"), -matches("lv")),
     fns = list(
-            Mean = "mean",
+      Median = "median",
       SD = "sd"
@@
     fns = list(
-            Mean = "mean",
+      Median = "median",
       SD = "sd"
@@
 participants_valid %>% 
-    count(group, l2) %>% 
+  count(group, l2, .drop = FALSE) %>% 
   mutate(
@@
 
-ggplot(post_re, aes(.value, reorder(as.factor(word), .value))) +
-    geom_vline(xintercept = fixed_intercept[1], colour = "grey", size = 1) +
+ggplot(
+  post_re,
+  aes(
+    .value,
+    reorder(as.factor(word), .value)
+  )
+) +
+  geom_vline(
+    xintercept = fixed_intercept[1],
+    colour = "grey", 
+    size = 1
+  ) +
 
@@
     axis.text.x = element_text(size = 7),
-        axis.text.y = element_blank(),
+    axis.text.y = element_text(size = 5, hjust = 1),
     axis.ticks = element_blank(),
@@
 
+```{r post_re_word_empirical, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+means <- responses %>% 
+  group_by(group) %>% 
+  summarise(
+    successes = sum(correct, na.rm = TRUE),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    prop = prop_adj(successes, n),
+    prop_se = prop_adj_se(successes, n)
+  )
+
+responses %>% 
+  mutate(
+    stim = paste0(
+      word, " (", ipa1, ") / ", 
+      word2, " (", ipa2, ") / ",
+      percent(lv, accuracy = 1)
+    )
+  ) %>% 
+  group_by(group, stim, lv) %>% 
+  summarise(
+    successes = sum(correct, na.rm = TRUE),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    prop = prop_adj(successes, n),
+    prop_se = prop_adj_se(successes, n)
+  ) %>% 
+  ggplot() +
+  aes(
+    x = prop,
+    y = reorder(as.factor(stim), lv),
+    xmin = prop-prop_se,
+    xmax = prop+prop_se,
+    label = percent(lv),
+    colour = group
+  ) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_vline(
+    data = means,
+    aes(xintercept = prop),
+    size = 1,
+    colour = "grey"
+  ) +
+  geom_errorbar(width = 0.1) +
+  geom_point() +
+  annotate(
+    geom = "text",
+    x = 0.7,
+    y = 10,
+    size = 3,
+    hjust = 0.5,
+    label = "Item accuracy \n(95% CI)"
+  ) +
+  annotate(
+    x = 0.7,
+    xend = 0.15,
+    y = 13,
+    yend = 20,
+    geom = "curve",
+    curvature = 0.25,
+    arrow = arrow(length = unit(0.2, "cm"))
+  ) +
+  # annotate(
+  #   geom = "text",
+  #   x = 0.8,
+  #   y = 10,
+  #   size = 3,
+  #   hjust = 0.5,
+  #   label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
+  # ) +
+  # annotate(
+  #   x = 0.7,
+  #   xend = fixed_intercept[1]+0.05,
+#   y = 17,
+#   yend = 30,
+#   geom = "curve",
+#   curvature = 0.2,
+#   arrow = arrow(length = unit(0.2, "cm"))
+# ) +
+labs(
+  x = "Probability of correct response",
+  y = "Presented word / Correct translation / Levenshtein similarity"
+) +
+  scale_colour_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    limits = c(0, 1), 
+    labels = percent
+  ) +
+  theme(
+    legend.position = "none",
+    legend.direction = "vertical",
+    legend.justification = "left",
+    legend.key = element_rect(fill = NA, colour = NA),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    axis.text.x = element_text(size = 7),
+    axis.text.y = element_text(size = 5, hjust = 1),
+    axis.ticks = element_blank(),
+    axis.title = element_text(size = 10),
+    plot.caption.position = "plot",
+    plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm")))
+  )
+
+ggsave(here("img", "accuracy_empirical.png"), height = 9, width = 15, dpi = 500)
+```
+
 ## Marginal effects
 
-```{r emmeans_pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=6}
-
-m <- posterior_epreds_fixed %>% 
-    mutate_at(vars(pthn_std, frequency_zipf_std), function(x) paste0(x, " SD")) %>% 
+```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
+A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein).
+We did this across the range of values of the PTHN scores.
+For each simulation, we drew a single sample from the posterior distribution of each coefficient.
+Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
+Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
+The dispersion of the lines indicates the uncertainty of our predictions.
+We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
+
+nd <- expand.grid(
+	lv_std = c(
+		min(fit_6$data$lv_std),
+		0,
+		max(fit_6$data$lv_std)
+	),
+	frequency_zipf_std = 0,
+	pthn_std = seq(
+		min(fit_6$data$pthn_std, na.rm = TRUE),
+		max(fit_6$data$pthn_std, na.rm = TRUE),
+		by = 0.1
+	),
+	group = unique(fit_6$data$group)
+)
+
+m <- add_epred_draws(
+	nd, 
+	fit_6, 
+	ndraws = 100, 
+	re_formula = NA
+) %>% 
 	mutate(
-        pthn_std = case_when(
-            pthn_std=="-1 SD" ~ "Low PTHN",
-            pthn_std=="0 SD" ~ "Mean PTHN",
-            pthn_std=="1 SD" ~ "High PTHN"
+		lv_std = factor(
+			lv_std,
+			levels = c(
+				min(fit_6$data$lv_std),
+				0,
+				max(fit_6$data$lv_std)
 			),
-        frequency_zipf_std = case_when(
-            frequency_zipf_std=="-1 SD" ~ "Low frequency",
-            frequency_zipf_std=="0 SD" ~ "Mean frequency",
-            frequency_zipf_std=="1 SD" ~ "High frequency"
-        )
-    )
-
-ggplot(m, aes(lv_std, .epred, colour = pthn_std, fill = pthn_std)) +
+			labels = c(
+				paste0("-1 SD (", percent(min(responses$lv, na.rm = TRUE)), ")"),
+				paste0("Mean (", percent(mean(responses$lv, na.rm = TRUE)), ")"),
+				paste0("+1 SD (", percent(max(responses$lv, na.rm = TRUE)), ")")
+			)
+		) 
+	)
+
+ggplot(m) +
+	aes(
+		pthn_std, 
+		.epred,
+		colour = lv_std,
+		fill = lv_std
+	) +
 	facet_wrap(~group) +
-    geom_hline(yintercept = 0.5, colour = "grey") +
-    geom_line(aes(group = interaction(pthn_std, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
-    stat_summary(aes(linetype = pthn_std), fun = mean, geom = "line", 
-                 colour = "black", size = 0.75, show.legend = TRUE) +
-    annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, 
-             arrow = arrow(length = unit(0.1, "cm")), size = .5) +
-    annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, 
-             arrow = arrow(length = unit(0.1, "cm")), size = .5) +
-    annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
-    annotate("text", x = 2.15, y = 0.95, size = 2.75, label = "More similar") +
-    labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
-         colour = "PTHN", fill = "PTHN", linetype = "PTHN",
-         title = "Population-level predictions") +
-    scale_color_manual(values = c("#1E88E5", "#ff2976")) +
+	geom_hline(
+		yintercept = 0.5, 
+		colour = "grey"
+	) +
+	geom_line(
+		aes(group = interaction(lv_std, .draw)), 
+		size = 0.65,
+		alpha = 0.5,
+		show.legend = TRUE
+	) +
+	stat_summary(
+		aes(linetype = lv_std), 
+		fun = mean, 
+		geom = "line",
+		colour = "black", 
+		size = 0.75,
+		show.legend = TRUE
+	) +
+	annotate(
+		"segment", 
+		x = -0.5, 
+		xend = -0.9, 
+		y = 0.85, 
+		yend = 0.85, 
+		arrow = arrow(length = unit(0.1, "cm")), 
+		size = .5
+	) +
+	annotate(
+		"segment", 
+		x = 2.25, 
+		xend = 2.65, 
+		y = 0.85, 
+		yend = 0.85, 
+		arrow = arrow(length = unit(0.1, "cm")),
+		size = .5
+	) +
+	annotate(
+		"text", 
+		x = -0.5, 
+		y = 0.95, 
+		size = 2.75,
+		label = "Less similar"
+	) +
+	annotate(
+		"text", 
+		x = 2.15, 
+		y = 0.95, 
+		size = 2.75,
+		label = "More similar"
+	) +
+	labs(
+		x = paste0("PTHN (+1 SD, ", round(sd(responses$pthn), 1), " higher-frequency neighbours)"),
+		y = "P(Correct)", 
+		colour = "Levenshtein similarity",
+		fill = "Levenshtein similarity",
+		linetype = "Levenshtein similarity",
+		title = "Population-level posterior predictions"
+	) +
 	scale_y_continuous(
@@
 ```
-
-## Accuracy
-
-### Accuracy by item
-
-```{r accuracy_item, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses_item <- responses %>%
-    group_by(group, word) %>%
-    summarise(
-        correct_sum = sum(correct, na.rm = TRUE),
-        correct_n = sum(!is.na(correct), na.rm = TRUE),
-        .groups = "drop"
-    ) %>% 
-    mutate(
-        correct_prop = prop_adj(correct_sum, correct_n),
-        correct_se = prop_adj_se(correct_sum, correct_n),
-        word_label = reorder_within(word, by = correct_prop, within = group)
-    )
-
-ggplot(responses_item, aes(reorder(word_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
-    facet_wrap(~group, scales = "free_y") +
-    geom_hline(yintercept = 0.5, colour = "black") +
-    geom_errorbar() +
-    geom_point(size = 0.5) +
-    labs(y = "Correct responses", x = "Word", colour = "Group") +
-    coord_flip() +
-    scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) +
-    scale_y_continuous(limits = c(0, 1), labels = percent) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text.x = element_text(size = 9),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-    )
-
-```
-
-### Accuracy by participant
-
-```{r responses_participant, echo=FALSE, fig.height=7, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}
-responses_participant <- responses %>%
-    group_by(group, participant) %>%
-    summarise(
-        correct_sum = sum(correct, na.rm = TRUE),
-        correct_n = sum(!is.na(correct), na.rm = TRUE),
-        .groups = "drop"
-    ) %>% 
-    mutate(
-        correct_prop = prop_adj(correct_sum, correct_n),
-        correct_se = prop_adj_se(correct_sum, correct_n),
-        participant_label = reorder_within(participant, by = correct_prop, within = group)
-    )
-
-ggplot(responses_participant, aes(reorder(participant_label, correct_prop), correct_prop, ymin = correct_prop-correct_se, ymax = correct_prop+correct_se, colour = group)) +
-    facet_wrap(~group, scales = "free_y") +
-    geom_hline(yintercept = 0.5) +
-    geom_errorbar(width = 0.5) +
-    geom_point(size = 1) +
-    labs(
-        y = "Correct responses",
-        x = "Participant",
-        colour = "Group"
-    ) +
-    coord_flip() +
-    scale_color_manual(
-        values = c("#1A85FF", "#ff2976", "#FFC20A"),
-        na.translate = FALSE
-    ) +
-    scale_y_continuous(
-        limits = c(0, 1), 
-        labels = scales::percent
-    ) +
-    theme(
-        legend.position = "none",
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank(),
-        axis.text.x = element_text(size = 9),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
-    )
-
-```
-
-## Model diagnostics
-
-```{r traceplots, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-posterior_draws_fixed %>% 
-    mutate(
-        .chain = as.factor(.chain),
-        .variable_name = str_replace_all(.variable, str_repl) %>% 
-            str_remove("b_")
-    ) %>% 
-    ggplot(aes(.iteration, .value, colour = .chain)) +
-    facet_wrap(~.variable_name, scales = "free_y") +
-    geom_line() +
-    labs(x = "Iteration", y = "Value", color = "Chain") +
-    scale_colour_manual(values = c("#1A85FF", "#ff2976", "#FFC20A", "04bf78"), na.translate = FALSE) +
-    theme(legend.position = "top")
-
-```

### 2022-03-17 manuscript/manuscript.Rmd
Add suggestions from @siowserene and refactor some code

@@
 
-Listening to non-native speech is more costly than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
+Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
 One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
@@
 Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
-In the present study we investigate the extent to which adult listeners rely on the phonological similarity between the native and the non-native languages during word recognition, capitalising on the role of *cognates*.
+In the present study we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
 
@@
 
-All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the populations of speakers of each language.
+All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration or social contact).
 Cognates embody a particular case of such commonalities.
@@
 For example, *sun* and *sol* (in Spanish) share phonological onset but their etymology points to different origins.
-The complimentary case is equally problematic for the definition of cognateness in this context: translation equivalents that share etymological origin may hardly be recognised as cognates by the average listener.
-Such is the case of *cheese* and *queso* (both from Latin), but phonologically dissimilar.
-We will use the term *cognateness* to include all form-similar cross-language synonyms for simplicity.
-It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
+The complimentary case is equally problematic for the definition of cognateness in this context: cross-language synonyms that share etymological origin may not be necessarily be recognised as cognates by the average listener.
+Such is the case of *cheese* and *queso* (Latin root), but are phonologically dissimilar.
+for simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
+We do not expect that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
 For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Cognates play a pivotal role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+Cognates play a pivotal role in many models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
 For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
-This cognate advantage is not restricted to production, but rather extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
-
-Early theoretical accounts of this phenomenon pointed to the fact that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
-The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), and not to their shared concept (conceptual route).
-This word-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection).
-As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations, therefore to cognateness.
+This cognate advantage is not restricted to production, instead also extending to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
+
+Early theoretical accounts of this phenomenon proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), instead of directly to their shared concept (conceptual route).
+This word-to-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection.
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
+cognates subsequently exert less impact on lexical processing in proficient speakers.
 The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
-The rationale behind this prediction was that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
-One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
+The rationale behind this prediction was that backward translation relies more strongly on direct word-to-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
+Backward translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
 
@@
 Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
-Although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
+However, although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
 These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
@@
 Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted the fact that lexical selection is affected by the number of phonological neighbours of the selected word.
-Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when such neighbours' lexical frequency is higher [@luce1998recognizing].
+For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
+Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when these neighbours' lexical frequency is higher [@luce1998recognizing].
 This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation (BIA/BIA+)  [@van1998orthographic; @dijkstra2002architecture] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
-Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
-When bilingual participants are presented with words in only one language (target language) in a lexical decision task, translating them to the non-target language took longer when they were part of large  neighbourhoods [@van1998orthographic].
-This suggests that words presented in the target language activated orthographic neighbourhoods in the non-target language, which competed for selection with the target word during recognition.
+The Bilingual Interactive Activation (BIA/BIA+)  [@van1998orthographic; @dijkstra2002architecture] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with each other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
+Evidence supporting such an integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word's translation.
+When bilingual participants are presented with words in one language in a lexical decision task, translating them to the other language took longer when they were part of large  neighbourhoods [@van1998orthographic].
+This suggests that presented words activated orthographic neighbourhoods in the other language, which competed for selection with the target word during recognition.
 
 A more recent model of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
-This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
+This model assumes an integrated lexicon in which word representations from two languages establish connections in a way that is similar to connections between words of a single language.
 It also makes the assumption that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
-It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
-In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
-A remarkable correlation between Multilink simulations of a translation elicitation task and behavioural data on Dutch-English bilinguals [@christoffels2006memory] has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
+It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-to-word connections).
+In addition, this model suggests that the strength of the association between L2 representations and their concept depends on language users' proficiency in L2.
+A strong correlation between Multilink simulations of a translation elicitation task and behavioural data from Dutch-English bilinguals [@christoffels2006memory] has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
 Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
@@
 
-The fact that Multilink successfully fitted the experimental data from @christoffels2006memory, despite the absence of any word-word connections between translation equivalents, is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
-It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
+The fact that Multilink successfully fitted the experimental data from @christoffels2006memory in the absence of word-to-word connections between translation equivalents is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
+The model specification under which the quantitative predictions were generated assumes similar vocabulary sizes across languages, and also that the word-to-concept connections of L1 and L2 representations have same strength.
 The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation, and this prediction remains untested my Multilink.
+It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-to-word connections during translation, especially during backward translation, and this prediction remains untested by Multilink.
 
 Most of the studies referenced so far rely heavily on production and comprehension of printed words: in these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
-This is a limitation regarding the generalisation of their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
+There are limitations in generalising their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
 Consider the words *radio* (English) and *radio* (Spanish).
-Both words are evidently cognates, as their orthographical word form is identical.
-However, their phonological forms (/ˈɹeɪdiˌəʊ/-/ˈraðjo/) are not identical.
-Although many English listeners might understand the Spanish word-form, even with no previous knowledge of Spanish, it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
-For the same reason, measuring the phonological similarity between two word forms is substantially more challenging than measuring their orthographic similarity.
-
-Most studies use symbolic representations of the word-forms of interest for compute their similarity in either modality: character strings in the case of orthographic word-forms (*radio*-*radio*), and phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA notation) in the case of phonological word forms, which also consists on a string of characters [see @floccia2018introduction for example].
-Either way, the similarity between the pair of word-forms is frequently measured by applying some algorithm that computes the edit distance between both.
-The Leveshtein distance is one of the most widely known: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string [@levenshtein1966binary], and has been successfully used to estimate the similarity of cross-language word-pairs [@gooskens2004perceptive; @schepens2012distributions].
-The Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio")`: both strings are identical.
-But the distance between the phonological transcriptions ɹeɪdiəʊ and raðjo is `r stringdist::stringdist("ɹeɪdiəʊ", "raðjo")`, which is the number of editions that separates both character strings.
-There is yet an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ are symbolic representations that do not capture the all of the features that are present in the acoustic signal when either word is spoken, and therefore any measure of similarity that uses such transcriptions as input will overestimate their actual similarity.
+Both words are clearly cognates, as their orthographical word form is identical.
+However, their phonological forms (English /ˈɹeɪ.di.əʊ/-- Spanish /ˈra.ðjo/) are not identical.
+Although many English listeners might recognise the Spanish word-form--even with no previous knowledge of Spanish--it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
+
+Measuring the phonological similarity between two word forms is also substantially more challenging than measuring their orthographic similarity.
+Many studies use symbolic representations to compute the simiarity between word-forms of interest in either modality.
+Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (*radio*-*radio*); phonological word-forms are represented by their  phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA notation) [see @floccia2018introduction for example].
+The similarity between each pair of word-forms is operationalised numerically by applying some algorithm that computes the edit distance between their symbolic representation.
+The Leveshtein distance is one of the most widely known algorithms: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string [@levenshtein1966binary], and has been successfully used to estimate the similarity of cross-language word-pairs [@gooskens2004perceptive; @schepens2012distributions].
+Using our previous examples, the Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio")`: both strings are identical.
+But the distance between the phonological transcriptions ɹeɪdiəʊ and raðjo is `r stringdist::stringdist("ɹeɪdiəʊ", "raðjo")`, which is the number of edits that separates both character strings.
+There is also an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.
+whether to include elements of aspiration, nasalisation, etc.) can affect how accurate these symbolic representation capture all of the features that are present in the acoustic signal when either word is spoken.
+Therefore, any measure of similarity that uses phonological transcriptions as input may overestimate or underestimate their actual similarity.
 Taking into account that measuring phonological similarity is considerably more complex than measuring orthographic complexity it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
-In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be overestimating the availability of such information in the auditory modality.
-
-In the present study we test listener's reliance on phonological similarity when translating words from a non-native language, in order to investigate the plausibility of the lexical route during backward translation.
-We tested monolingual adults in a translation elicitation task: they were presented with spoken words in an unfamiliar, non-native language, and were asked to translate them to their native language.
-The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which vocabulary size in L2 is null, and therefore word-concept connections are, in principle, absent.
-As a consequence, these participants should only be able use the lexical route to succeed in the translation of words in an unfamiliar language.
-We predicted that, if the information provided by the phonological form of the unfamiliar words is sufficient for translation, participants' performance should increase as the translation pairs are phonologically more similar.
+
+In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be misrepresenting the availability of such information in the auditory modality.
+In the present study we tested listener's reliance on phonological similarity when translating words from an unfamiliar language, in order to investigate the plausibility of word-to-word during backward translation.
+We tested English and Spanish native speakers in a translation elicitation task: they were presented with spoken words in an unfamiliar, non-native language, and were asked to translate them to their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as **presented words**, and the correct translation for the presented words as **target translations**.
+The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and therefore word-to-concept connections are, in principle, absent.
+As a consequence, these participants should only be able use word-to-word connections to succeed in the translation of words in an unfamiliar language.
+We predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
 
 We tested three groups of participants in the translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
-Catalan and Spanish are both Roman languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from a typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
-
-Following @luce1998recognizing we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
-If this is the case, words whose translation in the native language are part of larger phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target word.
-For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button* (whose lexical frequency is higher).
-If the phonological neighbourhood density of the target words affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
+Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
+
+Following @luce1998recognizing's NAM model, we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+If this is the case,  has more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target translation.
+For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button* (because the lexical frequency of *bottom* is higher than that of *button*).
+If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
 
@@
 Participants were allowed to correct their answer.
-Then, participants pressed the RETURN key to start and new trial.
-We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Trials where the response was mistyped by only one character were counted as correct, as long as the response did not correspond to a distinct word.
-Trials in which participants took longer than 10 seconds to respond were also filtered out.
+Then, participants pressed the RETURN/ENTER key to confirm their answer and start and new trial.
+We excluded trials where the participant's answer was not a valid word in the target language or no answer was given.
+Trials in which the response was mistyped by only one character were counted as correct (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct word.
+Trials in which participants took longer than 10 seconds to respond were also excluded.
 Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish).
@@
 We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
-Words in the Catalan list (listened to by participants) were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
+Words in the Catalan list were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
 Words in the Spanish list were `r printnum(lengths$eng_spa["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_spa["ipa1_sd"])`, Range = `r printnum(lengths$eng_spa["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_spa["ipa1_max"], digits = 0)`).
-Participants listened to one audio file in each trial.
-This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
+Participants listened to one audio file in each trial, each containing a single word presented in isolation.
 The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
@@
 
-### Translations (English/Spanish)
-
-Catalan translations to English (typed by participants in the keyboard) were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), and their translations to Spanish were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
-Spanish translations to English were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
-
-#### Translation lexical frequency
-
-We extracted the lexical frequencies from SUBTLEX-UK for English translations [@van2014subtlex], and from SUBTLEX-ESP for Spanish translations [@cuetos2011subtlex].
-We extracted or transformed scores as/into Zipf scores [@van2014subtlex] to  correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses [@van2014subtlex].
+### Participant responses (English/Spanish)
+
+Responses given by English participants for Catalan presented words were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), while their translations to Spani responses given by Spanish participants for Catalan words were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
+Responses tgiven by English participants for Spanish words were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
+
+#### Target translation lexical frequency
+
+We extracted the lexical frequencies for English target translations from SUBTLEX-UK[@van2014subtlex], and frequencies for Spanish from SUBTLEX-ESP [@cuetos2011subtlex].
+We extracted scores as Zipf scores where available or transformed raw scores to Zipf scores to  correct their logarithmic distribution, limiting their range roughly between 0 and 7, allowing an easier interpretation of further analyses [@van2014subtlex].
 
@@
 
-#### Translation phonological neighbourhood density (PTHN)
-
-We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) of the English and Spanish translations  from the CLEARPOND database [@marian2012clearpond].
-PTHN scores indicate the number of phonological neighbours of the target word (the correct translation in the native language of the presented word) with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
-CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target word in only one addition, deletion, or substitution.
-For instance, the PTHN score of *book* is `r pthn_example$pthn[1]`, while that of *sock* is `r pthn_example$pthn[2]`.
+#### Target translatiuon phonological neighbourhood density (PTHN)
+
+We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) of the English and Spanish target translations from the CLEARPOND database [@marian2012clearpond].
+PTHN scores indicate the number of phonological neighbours with higher lexical frequency than the target translation, as indicated by its frequency score in the corresponding SUBTLEX database.
+CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target translation in only one addition, deletion, or substitution.
 PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
 
-```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of stimuli translations, plotted in different panels for English translations of Catalan stimuli (cat-ENG), English translations of Spanish stimuli (spa-ENG), and Spanish translation of Catalan stimuli (cat-SPA).
-A\\) Lexical frequency of stimuli translations, expressed in counts per million and as Zipf scores.
-B\\) Number of higher-frequency phonological neighbourhoods (PTHN) of stimuli translations.
-C\\) Normalised Levenshtein phonological similarity between the presented stimuli and their translations, expressed in percentages.
-Lexical frequencies and PTHN scores of the stimuli (presented auditorily in Catalan or Spanish) are not described, as participants lack any lexical representation of such words."}
-
-medians <- stimuli %>% 
+```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA).
+Columns indicate: A\\) Lexical frequency of target translations, expressed as Zipf scores.
+B\\) PTHN of target translations, and C\\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages.
+Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words."}
+
+d <- stimuli %>% 
+  mutate(
+    group = case_when(
+      group=="ENG-CAT" ~ "cat-ENG",
+      group=="ENG-SPA" ~ "spa-ENG",
+      group=="SPA-CAT" ~ "cat-SPA"
+    ),
+    group = factor(
+      group,
+      levels = c("cat-ENG", "spa-ENG", "cat-SPA")
+    )
+  )
+
+medians <- d %>% 
   group_by(group) %>% 
@@
     lv_sd = sd(lv)
-  ) %>% 
-  mutate(group = case_when(
-    group=="ENG-CAT" ~ "cat-ENG",
-    group=="ENG-SPA" ~ "spa-ENG",
-    group=="SPA-CAT" ~ "cat-SPA",
-  ))
+  )
 
 # frequency
-stimuli %>% 
-  mutate(group = case_when(
-    group=="ENG-CAT" ~ "cat-ENG",
-    group=="ENG-SPA" ~ "spa-ENG",
-    group=="SPA-CAT" ~ "cat-SPA",
-  )) %>% 
-  ggplot(aes(frequency_zipf, fill = group)) + 
+d %>% 
+  ggplot() +
+  aes(
+    frequency_zipf, 
+    fill = group) +
   facet_wrap(~group, ncol = 1) + 
@@
     aes(
-      xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd, 
-      ymin = -Inf, ymax = Inf, fill = group
-    ), inherit.aes = FALSE, alpha = 0.15
-  ) +   
-  geom_histogram(bins = 20, colour = "white") +
+      xmin = frequency_zipf_median-frequency_zipf_sd, 
+      xmax = frequency_zipf_median+frequency_zipf_sd, 
+      ymin = -Inf, 
+      ymax = Inf, 
+      fill = group
+    ), 
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20,
+    colour = "white"
+    ) +
   geom_vline(
     data = medians, 
-    aes(xintercept = frequency_zipf_median, group = group), 
+    aes(
+      xintercept = frequency_zipf_median,
+      group = group
+      ), 
     colour = "black",
@@
   # PTHN
-  stimuli %>% 
-  count(group, pthn) %>% 
-  mutate(group = case_when(
-    group=="ENG-CAT" ~ "cat-ENG",
-    group=="ENG-SPA" ~ "spa-ENG",
-    group=="SPA-CAT" ~ "cat-SPA",
-  )) %>% 
-  ggplot(aes(pthn, n, fill = group)) + 
+  d %>% 
+  count(
+    group, 
+    pthn
+    ) %>% 
+  ggplot() + 
+  aes(
+    pthn,
+    n, 
+    fill = group
+    ) +
   facet_wrap(~group, ncol = 1) + 
@@
   geom_vline(
-    data = medians, aes(xintercept = pthn_median, group = group), 
-    colour = "black", size = 0.5, linetype = "dashed"
+    data = medians, 
+    aes(
+      xintercept = pthn_median, 
+      group = group
+      ), 
+    colour = "black", 
+    size = 0.5, 
+    linetype = "dashed"
   ) + 
@@
   # Levenshtein
-  stimuli %>% 
-  mutate(group = case_when(
-    group=="ENG-CAT" ~ "cat-ENG",
-    group=="ENG-SPA" ~ "spa-ENG",
-    group=="SPA-CAT" ~ "cat-SPA",
-  )) %>% 
-  ggplot(aes(lv, fill = group)) +
+  d %>% 
+  ggplot() +
+  aes(
+    lv, 
+    fill = group
+    ) +
   facet_wrap(~group, ncol = 1) + 
@@
     aes(
-      xmin = 0, xmax = lv_median+lv_sd, 
-      ymin = -Inf, ymax = Inf, fill = group
-    ), inherit.aes = FALSE, alpha = 0.15
-  ) +   
-  geom_histogram(bins = 20, colour = "white") +
+      xmin = 0,
+      xmax = lv_median+lv_sd, 
+      ymin = -Inf,
+      ymax = Inf, 
+      fill = group
+    ),
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20, 
+    colour = "white"
+    ) +
   geom_vline(
-    data = medians, aes(xintercept = lv_median, group = group), 
-    colour = "black", size = 0.5, linetype = "dashed"
+    data = medians, 
+    aes(
+      xintercept = lv_median,
+        group = group
+      ), 
+    colour = "black",
+    size = 0.5, 
+    linetype = "dashed"
   ) + 
   scale_x_continuous(labels = percent) +
-  labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity\n(stimuli and stimuli translations)\n") +
+  labs(
+    x = "Levenshtein", 
+    y = "Number of trials", 
+       title = "Phonological similarity\n(stimuli and stimuli translations)\n"
+    ) +
 
@@
   plot_annotation(tag_levels = "A") &
-  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
-  theme(legend.position = "none")
+  scale_fill_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A"),
+    na.translate = FALSE) &
+  theme(
+    legend.position = "none",
+    title = element_text(size = 10),
+    axis.title = element_text(size = 10)
+  )
 ```
 
-### Cross-language phonological overlap (Levenshtein)
-
-We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the `stringsim` function of the stringdist R package [@van2014stringdist].
+### Cross-language phonological similarity (Levenshtein)
+
+We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their IPA transcriptions using the `stringsim` function of the stringdist R package [@van2014stringdist].
 This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
-First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
+First, it computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
 This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is percentage score indicating no similarity between the two strings if 0%, and a perfect match between two strings if 100%.
-We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
+The result is a percentage score, with 0% indicating no similarity between the two strings, and 100% indicating a perfect match between the two strings.
+We computed this similarity measure (refred simply as Levenshtein, from now on) for every translation pair in our stimuli lists.
 For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim("teɪbəl", "mesa"))` similarity, while the *train* (trEIn)-*tren* (tɾEn) translation pair had a `r percent(stringsim("trEIn", "tɾEn"))` similarity.
@@
 
-We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We modelled the probability of participants guessing the correct target translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
 We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0).
-Second, we added the fixed effect of lexical frequency of the target word ($Frequency$).
+Secondly, we added the fixed effect of lexical frequency of the target translation ($Frequency$).
 We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels cat-ENG, spa-ENG, and cat-SPA), and finally, the $Levenshtein \times Group$ interaction).
 We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) [@schad2020capitalize].
-Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) [@barr2013random].
+Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) [@barr2013random].
 We will later refer to these models as Model 0 to Model 6.
@@
   arrange(desc(elpd_loo)) %>% 
-  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
-  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  select(
+    model,
+    matches("elpd_loo"),
+    matches("looic"), 
+    matches("diff")) %>%
+  mutate(
+    model = str_replace(
+      model,
+      "fit_", 
+      "Model "
+    )
+  ) %>% 
   mutate_all(~ifelse(. == 0, NA, .)) %>% 
@@
   ) %>% 
-  fmt_missing(everything(), missing_text = "-") %>% 
+  fmt_missing(
+    everything(),
+    missing_text = "-"
+  ) %>% 
   fmt_number(2:7) %>% 
@@
   tab_style(
-    style = cell_text(align = "left", weight = "bold", size = "medium"),
-    locations = cells_title(groups = "title")
+    style = cell_text(
+      align = "left",
+      weight = "bold", 
+      size = "medium"
+    ),
+    locations = cells_title(
+      groups = "title"
+    )
   ) %>% 
   tab_style(
-    style = cell_text(align = "left", size = "medium", style = "italic"),
-    locations = cells_title(groups = "subtitle")
+    style = cell_text(
+      align = "left", 
+      size = "medium", 
+      style = "italic"
+    ),
+    locations = cells_title(
+      groups = "subtitle"
+    )
   ) %>% 
   tab_style(
-    style = cell_text(align = "left", size = "medium"),
+    style = cell_text(
+      align = "left", 
+      size = "medium"
+    ),
     locations = cells_source_notes()
@@
   tab_style(
-    style = list(cell_borders(sides = c("left", "right"), color = "white"),
-                 cell_text(align = "center")),
+    style = list(
+      cell_borders(
+        sides = c("left", "right"),
+        color = "white"
+      ),
+      cell_text(
+        align = "center"
+      )
+    ),
     locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
+      cells_body(), 
+      cells_row_groups(), 
+      cells_column_spanners(), 
+      cells_stub(), 
+      cells_source_notes(),
+      cells_column_labels(), 
+      cells_stubhead()
     )
@@
   tab_style(
-    style = list(cell_text(align = "left", size = "medium"), 
-                 cell_borders(sides = "all", color = "white")),
-    locations = list(cells_stub(), cells_source_notes())
+    style = list(
+      cell_text(
+        align = "left",
+        size = "medium"
+      ), 
+      cell_borders(
+        sides = "all", 
+        color = "white"
+      )
+    ),
+    locations = list(
+      cells_stub(), 
+      cells_source_notes()
+    )
   ) %>% 
   tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
+    style = cell_text(
+      align = "center",
+      size = "medium", 
+      weight = "normal",
+      style = "normal"
+    ),
+    locations = list(
+      cells_column_spanners(), 
+      cells_column_labels()
+    )
   ) %>% 
   tab_style(
-    style = cell_borders(sides = "all", color = "white"),
-    locations = cells_body(columns = 1:7)
+    style = cell_borders(
+      sides = "all",
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
   ) %>% 
   tab_style(
-    style = cell_borders(sides = "top", color = "white"),
-    locations = cells_body(columns = 1:7)
+    style = cell_borders(
+      sides = "top", 
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
   ) %>% 
@@
 Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increments or decrements in probability of correct translation [@gelman2020regression].
+This way, the coefficients are expressed as increments or decrements in the probability of correct translation [@gelman2020regression].
 
@@
   clean_names() %>% 
-  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
+  mutate_at(
+    vars(estimate:q97_5),
+    ~ifelse(
+      variable=="Intercept", 
+      inv_logit_scaled(.), 
+      ./4
+    )
+  ) %>% 
   group_split(variable) %>% 
@@
   "pthn_std" = "PTHN\n[+1 SD]",
-  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "groupENGvs.SPA" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
         levels = c(
-          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-          "Group 1\n[ENG-* vs. SPA-CAT]",
-          "Group 2\n[ENG-CAT vs. ENG-SPA]",
-          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]",
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1", 
+          "Levenshtein \u00d7 Group 2"
         )
@@
   arrange(.variable) %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
+  ggplot() +
+  aes(
+    .value,
+    fct_rev(.variable_name)
+  ) +
   geom_vline(xintercept = 0) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
-  scale_x_continuous(labels = function(x) percent(round(x, 1))) +
-  labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95),
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#1A85FF", "#9ccaff", "#d2e5fc"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = function(x) percent(round(x, 1))
+  ) +
+  labs(
+    x = "P(Correct)", 
+    y = "Posterior probability density",
+    fill = "CrI"
+  ) +
   theme(
@@
     legend.direction = "vertical",
-    legend.background = element_rect(fill = "grey95"),
-    legend.key = element_rect(fill = "grey95", colour = "grey95"), 
+    legend.background = element_rect(
+      fill = "grey95"
+    ),
+    legend.key = element_rect(
+      fill = "grey95", 
+      colour = "grey95"), 
     legend.key.height = unit(0.1, "cm"),
     axis.title.y = element_blank(),
-    legend.text = element_text(size = 7),
-    axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black", hjust = 1),
+    legend.text = element_text(
+      size = 7
+    ),
+    axis.text.x = element_text(
+      colour = "black"
+    ),
+    axis.text.y = element_text(
+      colour = "black",
+      hjust = 1
+    ),
     axis.ticks.y = element_blank(),
-    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      size = 0.5
+    )
   )
@@
   mutate(
-    .variable_name = str_remove(.variable, "sd_participant__|sd_word__") %>% 
+    .variable_name = str_remove(
+      .variable,
+      "sd_participant__|sd_word__"
+    ) %>% 
       str_replace_all(str_repl) %>% 
@@
         levels = c(
-          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-          "Group 1\n[ENG-* vs. SPA-CAT]",
-          "Group 2\n[ENG-CAT vs. ENG-SPA]",
-          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1",
+          "Levenshtein \u00d7 Group 2"
         )
@@
     ),
-    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
-  )
-
-post_sd <- ggplot(filter(post_sd_data, group=="Participants"), aes(.value, fct_rev(.variable_name))) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, limits = c(0, 1)) +
-  labs(x = "SDs in P(Correct)", title = "By participant", fill = "CrI") +
-  theme(axis.text.x = element_blank()) +
-
-  ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_x_continuous(labels = percent, limits = c(0, 1)) +
-  labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +
-
-  plot_layout(ncol = 1, heights = c(0.85, 0.15), guides = "collect") &
+    .value = ifelse(
+      str_detect(.variable, "Intercept"), 
+      inv_logit_scaled(.value), 
+      .value/4
+    )
+  )
+
+post_sd <- post_sd_data %>% 
+  filter(group=="Participants") %>% 
+  ggplot() +
+  aes(.value, fct_rev(.variable_name)) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95),
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#ff2976", "#f2a7c2", "#fcd9e6"), 
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By participant", 
+    fill = "CrI"
+  ) +
+  theme(
+    axis.text.x = element_blank()
+  ) +
+
+  post_sd_data %>% 
+  filter(group=="Words") %>% 
+  ggplot() +
+  aes(
+    .value, 
+    fct_rev(.variable_name)
+  ) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95), 
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By item",
+    fill = "CrI"
+  ) +
+  plot_layout(
+    ncol = 1,
+    heights = c(0.85, 0.15), 
+    guides = "collect"
+  ) &
   theme(
@@
   as.data.frame() %>% 
-  pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
+  pivot_longer(
+    -term1, 
+    names_to = "term2",
+    values_to = ".value"
+  ) %>% 
   drop_na(.value)
@@
 post_cors <- corr_data %>% 
-  mutate(term1 = factor(term1, levels = names(str_repl), ordered = TRUE),
-         term2 = factor(term2, levels = names(str_repl), ordered = TRUE)) %>% 
-  mutate_at(vars(term1, term2), str_replace_all, str_repl) %>% 
   mutate(
-    .value_label = case_when(term1==term2 ~ NA_character_, TRUE ~ printnum(.value, gt1 = FALSE)),
-    .value = case_when(term1==term2 ~ NA_real_, TRUE ~ .value)
-  ) %>% 
-  filter(term1 != "PTHN × Levenshtein", term2 != "Intercept") %>% 
-  ggplot(aes(fct_inorder(term1), fct_rev(fct_inorder(term2)), fill = .value)) +
+    term1 = factor(
+      term1, 
+      levels = names(str_repl),
+      ordered = TRUE
+    ),
+    term2 = factor(
+      term2, 
+      levels = names(str_repl),
+      ordered = TRUE
+    )
+  ) %>% 
+  mutate_at(
+    vars(term1, term2),
+    str_replace_all, 
+    str_repl
+  ) %>% 
+  mutate(
+    .value_label = case_when(
+      term1==term2 ~ NA_character_, 
+      TRUE ~ printnum(.value, gt1 = FALSE)
+    ),
+    .value = case_when(
+      term1==term2 ~ NA_real_, 
+      TRUE ~ .value
+    )
+  ) %>% 
+  filter(
+    term1 != "PTHN × Levenshtein", 
+    term2 != "Intercept"
+  ) %>% 
+  ggplot() +
+  aes(
+    fct_inorder(term1),
+    fct_rev(fct_inorder(term2)), 
+    fill = .value
+  ) +
   geom_tile(na.rm = TRUE) +
-  geom_text(aes(label = .value_label), size = 3, na.rm = TRUE) +
-  labs(x = "Term 1", y = "Term 2", fill = "Posterior\ncorrelation") +
+  geom_text(
+    aes(label = .value_label), 
+    size = 3, 
+    na.rm = TRUE) +
+  labs(
+    x = "Term 1", 
+    y = "Term 2",
+    fill = "Posterior\ncorrelation"
+  ) +
   scale_fill_gradient2(
-    low = "#04bf78", mid = "white", high = "#FFC20A", 
-    na.value = "white", limits = c(-0.4, 0.4)
+    low = "#04bf78", 
+    mid = "white",
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
   ) +
@@
     axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 6, angle = 90, hjust = 0, vjust = 1),
-    axis.text.y = element_text(size = 6, vjust = 1, hjust = 1)
-  )
-
-post_fix / ((post_sd | post_cors) + plot_layout(widths = c(0.6, 0.4))) +
+    axis.text.x = element_text(
+      size = 6,
+      angle = 90,
+      hjust = 0, 
+      vjust = 1),
+    axis.text.y = element_text(
+      size = 6,
+      vjust = 1,
+      hjust = 1
+    )
+  )
+
+post_fix / (
+  (post_sd | post_cors) +
+    plot_layout(widths = c(0.6, 0.4))
+) +
   plot_layout() +
@@
 ```{r bayesfactors}
-h_frequency <- abs(hypothesis(fit_6, hypothesis = "frequency_zipf_std = 0")$hypothesis$Evid.Ratio)
-h_pthn <- abs(hypothesis(fit_6, hypothesis = "pthn_std = 0")$hypothesis$Evid.Ratio)
-h_lv <- abs(hypothesis(fit_6, hypothesis = "lv_std = 0")$hypothesis$Evid.Ratio)
-h_pthn_lv <- abs(hypothesis(fit_6, hypothesis = "pthn_std:lv_std = 0")$hypothesis$Evid.Ratio)
-h_group1 <- abs(hypothesis(fit_6, hypothesis = "groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
-h_group2 <- abs(hypothesis(fit_6, hypothesis = "groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)
-h_lv_group1 <- abs(hypothesis(fit_6, hypothesis = "lv_std:groupENGvs.SPA = 0")$hypothesis$Evid.Ratio)
-h_lv_group2 <- abs(hypothesis(fit_6, hypothesis = "lv_std:groupENGMCATvs.ENGMSPA = 0")$hypothesis$Evid.Ratio)
+
+get_bf <- function(hyp, abs = TRUE, ...){
+  bf <- hypothesis(fit_6, hypothesis = hyp, ...)$hypothesis$Evid.Ratio
+  if (abs) bf <- abs(bf)
+  return(bf)
+}
+
+h_frequency <- get_bf("frequency_zipf_std = 0")
+h_pthn <- get_bf("pthn_std = 0")
+h_lv <- get_bf("pthn_std = 0")
+
+h_pthn_lv <- get_bf("pthn_std:lv_std = 0")
+h_group1 <- get_bf("groupENGvs.SPA = 0")
+h_group2 <- get_bf("groupENGMCATvs.ENGMSPA = 0")
+h_lv_group1 <- get_bf("lv_std:groupENGvs.SPA = 0")
+h_lv_group2 <- get_bf("lv_std:groupENGMCATvs.ENGMSPA = 0")
 
@@
   mutate(
-    .value = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.value), .value/4),
-    .lower = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.lower), .lower/4),
-    .upper = ifelse(grepl("Intercept", .variable), inv_logit_scaled(.upper), .upper/4)
+    .value = ifelse(
+      grepl("Intercept", .variable), 
+      inv_logit_scaled(.value), 
+      .value/4
+    ),
+    .lower = ifelse(
+      grepl("Intercept", .variable),
+      inv_logit_scaled(.lower),
+      .lower/4
+    ),
+    .upper = ifelse(
+      grepl("Intercept", .variable),
+      inv_logit_scaled(.upper), 
+      .upper/4
+    )
   ) %>% 
@@
   slice(1, n()) %>% 
-  left_join(distinct(stimuli, word1, ipa1), by = c("word" = "word1")) %>% 
+  left_join(
+    distinct(
+      stimuli,
+      word1, 
+      ipa1), 
+    by = c("word" = "word1")) %>% 
   split(.$word)
@@
 
-cors_range <- gather_draws(fit_6, `cor_.*`, regex = TRUE) %>% 
+cors_range <- gather_draws(
+  fit_6, 
+  `cor_.*`, 
+  regex = TRUE
+) %>% 
   median_qi() %>% 
@@
 
-h_cor_intercept_frequency <- hypothesis(fit_6, "participant__Intercept__frequency_zipf_std = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_frequency_zipf_lv_group1 <- hypothesis(fit_6, "participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", class = "cor")$hypothesis$Evid.Ratio
-h_cor_intercept_pthn <- hypothesis(fit_6, "participant__Intercept__pthn_std = 0", class = "cor")$hypothesis$Evid.Ratio
+h_cor_intercept_frequency <- get_bf("participant__Intercept__frequency_zipf_std = 0", class = "cor")
+h_cor_frequency_zipf_lv_group1 <- get_bf("participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", class = "cor")
+h_cor_intercept_pthn <- get_bf("participant__Intercept__pthn_std = 0", class = "cor")
 ```
 
-Overall, participants were `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likely to produce correct translations.
-The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf_std["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
-The number of phonological neighbours of the correct translation with higher lexical frequency (*PTHN*) slightly decreased the probability of a correct responses in `r percent(coefs$pthn_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
-Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
-This difference increased in `r percent(coefs$pthn_std_lv_std["estimate"], accuracy = 0.01)` for every *SD* increment in *PTHN* (95% *CrI* = [`r percent(coefs$pthn_std_lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std_lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
+Overall, participants had `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likelihood to produce correct translations.
+The lexical frequency of the target translation had little impact on accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf_std["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
+The number of phonological neighbours with higher lexical frequency than the target translation (*PTHN*) slightly decreased the probability of a correct responses by `r percent(coefs$pthn_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
+Finally, phonological similarity between the presented word and target translation had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
+The interaction effect between $PTHN \times Levenshtein$ was small: a +1 SD increment in Levenshtein increased the effect of *PTHN* by `r percent(coefs$pthn_std_lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std_lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std_lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
 
 Spanish natives translating Catalan words were `r percent(coefs$group_en_gvs_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
-This difference increased in `r percent(coefs$lv_std_group_en_gvs_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in `r percent(coefs$lv_std_group_en_gvs_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
 English natives translating Catalan words were `r percent(abs(coefs$group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
-For every one *SD* increment in Levenshtein similarity, the effect increased in `r percent(abs(coefs$lv_std_group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+For every one *SD* increment in Levenshtein similarity, the effect of presented language increased in `r percent(abs(coefs$lv_std_group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
 
 ```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
-A) Population-level predictions: the X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 200 observations from our model: 100 simulations for words with low *PTHN* (-1 *SD*) and 100 simulations for high *PTHN* (+1 *SD*) words.
-We did this across the range of values of the Levenshtein scores.
+A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein).
+We did this across the range of values of the PTHN scores.
 For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high *PTHN* words.
-We also plotted the mean of the high *PTHN* (black solid line) and low *PTHN* (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
+Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
+Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
 The dispersion of the lines indicates the uncertainty of our predictions.
@@
 
-m <- posterior_epreds_fixed %>% 
-  mutate_at(vars(pthn_std, frequency_zipf_std), function(x) paste0(x, " SD")) %>% 
+nd <- expand.grid(
+	lv_std = c(
+		min(fit_6$data$lv_std),
+		0,
+		max(fit_6$data$lv_std)
+	),
+	frequency_zipf_std = 0,
+	pthn_std = seq(
+		min(fit_6$data$pthn_std, na.rm = TRUE),
+		max(fit_6$data$pthn_std, na.rm = TRUE),
+		by = 0.1
+	),
+	group = unique(fit_6$data$group)
+)
+
+m <- add_epred_draws(
+	nd, 
+	fit_6, 
+	ndraws = 100, 
+	re_formula = NA
+) %>% 
 	mutate(
-    pthn_std = case_when(
-      pthn_std=="-1 SD" ~ "Low PTHN",
-      pthn_std=="0 SD" ~ "Mean PTHN",
-      pthn_std=="1 SD" ~ "High PTHN"
-    ),
-    frequency_zipf_std = case_when(
-      frequency_zipf_std=="-1 SD" ~ "Low frequency",
-      frequency_zipf_std=="0 SD" ~ "Mean frequency",
-      frequency_zipf_std=="1 SD" ~ "High frequency"
-    )
-  )
-
-ggplot(m, aes(lv_std, .epred, colour = pthn_std, fill = pthn_std)) +
+		lv_std = factor(
+			lv_std,
+			levels = c(
+				min(fit_6$data$lv_std),
+				0,
+				max(fit_6$data$lv_std)
+			),
+			labels = c(
+				paste0("Min (", percent(min(responses$lv, na.rm = TRUE)), ")"),
+				paste0("Mean (", percent(mean(responses$lv, na.rm = TRUE)), ")"),
+				paste0("Max (", percent(max(responses$lv, na.rm = TRUE)), ")")
+			)
+		) 
+	)
+
+ggplot(m) +
+	aes(
+		pthn_std, 
+		.epred,
+		colour = lv_std,
+		fill = lv_std
+	) +
 	facet_wrap(~group) +
-  geom_hline(yintercept = 0.5, colour = "grey") +
-  geom_line(aes(group = interaction(pthn_std, .draw)), alpha = 0.25, size = 0.5, show.legend = TRUE) +
-  stat_summary(aes(linetype = pthn_std), fun = mean, geom = "line", colour = "black", size = 0.75, show.legend = TRUE) +
-  annotate("segment", x = -0.5, xend = -0.9, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
-  annotate("segment", x = 2.25, xend = 2.65, y = 0.85, yend = 0.85, arrow = arrow(length = unit(0.1, "cm")), size = .5) +
-  annotate("text", x = -0.5, y = 0.95, size = 2.75, label = "Less similar") +
-  annotate("text", x = 2.15, y = 0.95, size = 2.75, label = "More similar") +
-  labs(x = "Phonological similarity\n(Levenshtein, SD)", y = "P(Correct)", 
-       colour = "PTHN", fill = "PTHN", linetype = "PTHN",
-       title = "Population-level predictions") +
-  scale_color_manual(values = c("#1E88E5", "#ff2976")) +
-  scale_y_continuous(labels = function(x) percent(round(x, 2)), limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
+	geom_hline(
+		yintercept = 0.5, 
+		colour = "grey"
+	) +
+	geom_line(
+		aes(group = interaction(lv_std, .draw)), 
+		size = 0.65,
+		alpha = 0.5,
+		show.legend = TRUE
+	) +
+	stat_summary(
+		aes(linetype = lv_std), 
+		fun = mean, 
+		geom = "line",
+		colour = "black", 
+		size = 0.75,
+		show.legend = TRUE
+	) +
+	labs(
+		x = paste0("PTHN (+1 SD, ", round(sd(responses$pthn), 1), " higher-frequency neighbours)"),
+		y = "P(Correct)", 
+		colour = "Levenshtein similarity",
+		fill = "Levenshtein similarity",
+		linetype = "Levenshtein similarity",
+		title = "Population-level posterior predictions"
+	) +
+  scale_color_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A")
+  ) +
+	scale_y_continuous(
+		labels = function(x) percent(round(x, 2)),
+		limits = c(0, 1),
+		breaks = seq(0, 1, 0.1)
+	) +
 	theme(
@@
 
-The conditional intra-class correlation (*ICC*) for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items [@gelman2006data].
-Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
+The conditional intra-class correlation [*ICC*; @gelman2006data] for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items.
+Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were very likely to be correctly translated (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
 Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants.

### 2022-03-17 R/02_models.R
Add suggestions from @siowserene and refactor some code

@@ -43,18 +43,18 @@ get_model_draws_fixed <- function(fit){
 
 
 # get model posterior expected predictions (fixed effects)
-get_model_epreds_fixed <- function(fit, ndraws = 100){
+get_model_epreds_fixed <- function(fit, ndraws = 50){
     nd <- expand.grid(
-        pthn_std = c(-1, 1),
+        lv_std = c(-1, 0, 1),
         frequency_zipf_std = 0,
-        lv_std = seq(
-            min(fit$data$lv_std, na.rm = TRUE),
-            max(fit$data$lv_std, na.rm = TRUE),
+        pthn_std = seq(
+            min(fit$data$pthn_std, na.rm = TRUE),
+            max(fit$data$pthn_std, na.rm = TRUE),
             by = 0.1
         ),
         group = unique(fit$data$group)
     )
-    m <- add_epred_draws(nd, fit, ndraws = 100, re_formula = NA)
+    m <- add_epred_draws(nd, fit, ndraws = 50, re_formula = NA)
     return(m)
 }
 
@@ -66,11 +66,11 @@ get_model_epreds_random <- function(fit, ndraws = 100, group = c("participant", 
     if ("participant" %in% group){
         nd_re_participants <- expand.grid(
             participant = unique(fit$data$participant),
-            pthn_std = c(-1, 1),
+            lv_std = c(-1, 0, 1),
             frequency_zipf_std = 0,
-            lv_std = seq(
-                min(fit$data$lv_std, na.rm = TRUE),
-                max(fit$data$lv_std, na.rm = TRUE),
+            pthn_std = seq(
+                min(fit$data$pthn_std, na.rm = TRUE),
+                max(fit$data$pthn_std, na.rm = TRUE),
                 by = 0.1
             ),
             group = NA

### 2022-03-18 manuscript/manuscript.Rmd

@@
   theme(
-    title = element_text(size = 8, face = "bold"),
-
-    legend.background = element_rect(fill = NA, colour = NA),
-    legend.key = element_rect(fill = "white", colour = NA),
+    title = element_text(
+      size = 8, 
+      face = "bold"
+    ),
+    legend.background = element_rect(
+      fill = NA,
+      colour = NA
+    ),
+    legend.key = element_rect(
+      fill = "white", 
+      colour = NA),
     legend.justification = "right",
     legend.direction = "horizontal",
-
-    axis.line = element_line(colour = "black", size = 0.5),
-    axis.ticks = element_line(colour = "black", size = 0.5),
-    axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black"),
-
+    axis.line = element_line(
+      colour = "black",
+      size = 0.5
+    ),
+    axis.ticks = element_line(
+      colour = "black", 
+      size = 0.5
+    ),
+    axis.text.x = element_text(
+      colour = "black"
+    ),
+    axis.text.y = element_text(
+      colour = "black"
+    ),
     plot.caption.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-
+    plot.title = element_text(
+      hjust = 0.5
+    ),
     panel.grid = element_blank(),
-    panel.background = element_rect(fill = "white", colour = NA)
+    panel.background = element_rect(
+      fill = "white", 
+      colour = NA
+    )
   )
@@
 lengths <- stimuli %>% 
-  select(group, ipa1, ipa2, word1, word2) %>% 
-  mutate_at(vars(ipa1:word2), nchar) %>% 
+  select(
+    group, 
+    ipa1,
+    ipa2, 
+    word1, 
+    word2
+  ) %>% 
+  mutate_at(
+    vars(ipa1:word2),
+    nchar
+  ) %>% 
   group_by(group) %>% 
   summarise_at(
-    vars(ipa1, ipa2, word1, word2), 
-    list(mean = mean, sd = sd, min = min, max = max)
+    vars(
+      ipa1, 
+      ipa2, 
+      word1, 
+      word2
+    ), 
+    list(
+      mean = mean, 
+      sd = sd,
+      min = min, 
+      max = max
+    )
   ) %>%
@@
   clean_names() %>% 
-  filter(word %in% c("book", "sock"), group=="ENG-CAT") %>% 
+  filter(
+    word %in% c("book", "sock"),
+    group=="ENG-CAT"
+  ) %>% 
   select(word, pthn)
@@
 
-#### Target translatiuon phonological neighbourhood density (PTHN)
+#### Target translation phonological neighbourhood density (PTHN)
 
@@
       group,
-      levels = c("cat-ENG", "spa-ENG", "cat-SPA")
+      levels = c(
+        "cat-ENG",
+        "spa-ENG", 
+        "cat-SPA"
+      )
     )
@@
     frequency_zipf, 
-    fill = group) +
-  facet_wrap(~group, ncol = 1) + 
+    fill = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) + 
   geom_rect(
@@
   ) +
-  facet_wrap(~group, ncol = 1) + 
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) + 
   geom_rect(
@@
   mutate_all(~ifelse(. == 0, NA, .)) %>% 
-  gt(rowname_col = "model") %>% 
+  gt(
+    rowname_col = "model", 
+    id = "model-comparison"
+  ) %>% 
   tab_header(
@@
       ),
-    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+    .value = ifelse(
+      str_detect(.variable, "Intercept"),
+      inv_logit_scaled(.value), 
+      .value/4
+      )
   ) %>% 
@@
 
-```{r bayesfactors}
-
+```{r bayesfactors, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+
+# wrapper to extract Bayes Factors
 get_bf <- function(hyp, abs = TRUE, ...){
@@
 h_lv <- get_bf("pthn_std = 0")
-
 h_pthn_lv <- get_bf("pthn_std:lv_std = 0")
@@
 
-sds <- gather_draws(fit_6, `sd_.*`, regex = TRUE) %>% 
+sds <- gather_draws(
+  fit_6,
+  `sd_.*`, 
+  regex = TRUE
+  ) %>% 
   mean_qi() %>% 
@@
 
-h_cor_intercept_frequency <- get_bf("participant__Intercept__frequency_zipf_std = 0", class = "cor")
-h_cor_frequency_zipf_lv_group1 <- get_bf("participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", class = "cor")
-h_cor_intercept_pthn <- get_bf("participant__Intercept__pthn_std = 0", class = "cor")
+h_cor_intercept_frequency <- get_bf(
+  "participant__Intercept__frequency_zipf_std = 0", 
+  class = "cor"
+  )
+h_cor_frequency_zipf_lv_group1 <- get_bf(
+  "participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", 
+  class = "cor"
+  )
+h_cor_intercept_pthn <- get_bf(
+  "participant__Intercept__pthn_std = 0",
+  class = "cor"
+  )
 ```

### 2022-03-21 docs/index.Rmd
Minor refactoring

@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
-  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv_std:group1" = "Levenshtein \u00d7 Group 1",
+  "lv_std:group2" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
   "pthn_std" = "PTHN\n[+1 SD]",
-  "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-  "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "group1" =  "Group 1\n[*-ENG vs. cat-SPA]",
+  "group2" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
   "pthn_std" = "PTHN\n[+1 SD]",
-  "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-  "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "group1" =  "Group 1\n[*-ENG vs. cat-SPA]",
+  "group2" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
       factor(
-        levels = c(
-          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-          "Group 1\n[ENG-* vs. SPA-CAT]",
-          "Group 2\n[ENG-CAT vs. ENG-SPA]",
-          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+        levels =c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1",
+          "Levenshtein \u00d7 Group 2"
         )
@@
              levels = c(
-               "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-               "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-               "Group 1\n[ENG-* vs. SPA-CAT]",
-               "Group 2\n[ENG-CAT vs. ENG-SPA]",
-               "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
-             )
-           ))
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1",
+          "Levenshtein \u00d7 Group 2"
+        )
+           )
+        )
 

### 2022-03-21 manuscript/manuscript.Rmd
Minor refactoring

--- 2022-03-17 manuscript/manuscript.Rmd
@@
 classoption: man
-output: papaja::apa6_docx
+output:
+  papaja::apa6_docx: default
+  papaja::apa6_pdf: default
 ---
@@
 We computed this similarity measure (refred simply as Levenshtein, from now on) for every translation pair in our stimuli lists.
-For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim("teɪbəl", "mesa"))` similarity, while the *train* (trEIn)-*tren* (tɾEn) translation pair had a `r percent(stringsim("trEIn", "tɾEn"))` similarity.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a 17% similarity, while the *train* (trEIn)-*tren* (tɾEn) translation pair had a 60% similarity.
 Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
@@
 A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein).
-We did this across the range of values of the PTHN scores.
+We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean similarity (20.51% Levenshtein), and 100 simulations for translations with the maximum similarity observed in the dataset (83.33% Levenshtein).
+These simulations were obtained for the range of values of the PTHN scores.
 For each simulation, we drew a single sample from the posterior distribution of each coefficient.

### 2022-03-21 manuscript/manuscript.tex
Minor refactoring

@@
   english,
-  man,floatsintext]{apa6}
+  man,floatsintext]{apa7}
 \usepackage{amsmath,amssymb}
@@
 
+\abstract{
+Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar.
+Phonological similarity plays a more important role in low-proficiency bilinguals, compared to high-proficiency bilinguals, and points to the latter as relying more strongly on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
+Disentangling bilinguals' reliance on the conceptual and the lexical route during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals.
+In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
+Given their lack of previous knowledge on the unfamiliar language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
+This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route.
+Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation.
+This effect was larger in Spanish participants translating Catalan words than in English participants translating Catalan or Spanish.
+These results point to cross-language word-to-word links as providing rich information during word translation.
+}
+
 \begin{document}
@@
 
-Listening to non-native speech is more costly than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations (Lecumberri, Cooke, \& Cutler, 2010; Takata \& Nábělek, 1990).
+Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations (Lecumberri et al., 2010; Takata \& Nábělek, 1990).
 One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, a Spanish native listening to French may encounter the word /pɔʁt/ (\emph{porte}), translation of \emph{door} in French.
-The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
-These discrepancies can nonetheless be perceived as allophonic variations of native phonemes (Best, McRoberts, \& Goodell, 2001), and the non-native speech signal engages lexical processing mechanisms (Weber \& Cutler, 2004).
-Although this mismatch has a noticeable toll on comprehension (Cutler, Weber, Smits, \& Cooper, 2004), the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
-In the present study we investigate the extent to which adult listeners rely on the phonological similarity between the native and the non-native languages during word recognition, capitalising on the role of \emph{cognates}.
+For example, a Spanish native listening to French may encounter the word \ipatext{p<U+0254><U+0281>t} (\emph{porte}), translation of \emph{door} in French.
+The voiced uvular fricative consonant \ipatext{<U+0281>} and the open-mid back rounded vowel \ipatext{<U+0254>} do not exist in Spanish.
+These discrepancies can nonetheless be perceived as allophonic variations of native phonemes (Best et al., 2001), and the non-native speech signal engages lexical processing mechanisms (Weber \& Cutler, 2004).
+Although this mismatch has a noticeable toll on comprehension (Cutler et al., 2004), the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
+In the present study we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of \emph{cognates}.
 
-All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the populations of speakers of each language.
+All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration or social contact).
 Cognates embody a particular case of such commonalities.
 Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin\footnote{Some form-similar cross-language synonyms are technically not cognates.
-For example, \emph{sun} and \emph{sol} (in Spanish), share their phonological onset, but their etymology points to different origins.
-We will use the word \emph{cognateness} to include all form-similar cross-language synonyms for simplicity.
-It is highly implausible that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.}.
-For example, Romance languages such as Spanish and Catalan share many cognates (Schepens, Dijkstra, \& Grootjen, 2012), as in the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively).
-Cognates play a pivotal role in virtually all models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., \emph{puerta}-\emph{porta}), compared to when their labels are non-cognates (\emph{mesa}-\emph{taula}, \emph{table} in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval (Costa, Caramazza, \& Sebastian-Galles, 2000).
-This cognate advantage is not restricted to production, but rather extends to word recognition (Midgley, Holcomb, \& Grainger, 2011; Thierry \& Wu, 2007), word learning (De Groot \& Keijzer, 2000; Lotto \& De Groot, 1998; Valente, Ferré, Soares, Rato, \& Comesaña, 2018), and word translation (Christoffels, De Groot, \& Kroll, 2006).
+For example, \emph{sun} and \emph{sol} (in Spanish) share phonological onset but their etymology points to different origins.
+The complimentary case is equally problematic for the definition of cognateness in this context: cross-language synonyms that share etymological origin may not be necessarily be recognised as cognates by the average listener.
+Such is the case of \emph{cheese} and \emph{queso} (Latin root), but are phonologically dissimilar.
+for simplicity, we will use the term \emph{cognateness} in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
+We do not expect that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.}.
+For example, Romance languages such as Spanish and Catalan share many cognates (Schepens et al., 2012), as in the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively).
+Cognates play a pivotal role in many models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., \emph{puerta}-\emph{porta}), compared to when their labels are non-cognates (\emph{mesa}-\emph{taula}, \emph{table} in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval (Costa et al., 2000).
+This cognate advantage is not restricted to production, instead also extending to word recognition (Midgley et al., 2011; Thierry \& Wu, 2007), word learning (De Groot \& Keijzer, 2000; Elias \& Degani, 2022; Lotto \& De Groot, 1998; Valente et al., 2018), and word translation (Christoffels et al., 2006).
 
-Early theoretical accounts of this phenomenon pointed to the fact that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) (Potter, So, Von Eckardt, \& Feldman, 1984; see Chen \& Leung, 1989 for similar results).
-The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), and not to their shared concept (conceptual route).
-This word-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection).
-As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations, therefore to cognateness.
+Early theoretical accounts of this phenomenon proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) (Potter et al., 1984; see Chen \& Leung, 1989 for similar results).
+The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), instead of directly to their shared concept (conceptual route).
+This word-to-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection.
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
+cognates subsequently exert less impact on lexical processing in proficient speakers.
 The Revised Hierarchical Model (RHM, Kroll \& Stewart, 1994) captured this hypothesis and predicted that translating words from L1 to L2 (\emph{forward} translation) should take longer than translating from L2 to L1 (\emph{backward} translation).
-The rationale behind this prediction was that backward translation relies more strongly on direct word-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
-One of the consequences of this prediction is that backward translation should be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
+The rationale behind this prediction was that backward translation relies more strongly on direct word-to-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
+Backward translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
 
-To test these predictions, Annette MB de Groot, Dannenburg, and Vanhell (1994) and Annette M.
-de Groot (1992) asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+To test these predictions, Groot et al.
+(1994) and Groot (1992) asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
 Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
@@
 Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
-Although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels et al., 2006; Christoffels, Ganushchak, \& Koester, 2013).
+However, although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels et al., 2006, 2013).
 These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
@@
 Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-For instance, the Neighbourhood Activation Model (NAM, Luce \& Pisoni, 1998) highlighted the fact that lexical selection is affected by the number of phonological neighbours of the selected word.
-Words surrounded by a larger number of phonological neighbours (only differing in one phoneme from the target words) are responded more slowly and less accurately than words from sparser phonological neighbourhoods (Goldinger, Luce, \& Pisoni, 1989; Luce, Pisoni, \& Goldinger, 1990), especially when such neighbours' lexical frequency is higher (Luce \& Pisoni, 1998).
+For instance, the Neighbourhood Activation Model (NAM, Luce \& Pisoni, 1998) highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
+Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods (Goldinger et al., 1989; Luce et al., 1990), especially when these neighbours' lexical frequency is higher (Luce \& Pisoni, 1998).
 This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation (Dijkstra \& Van Heuven, 2002; BIA/BIA+, Van Heuven, Dijkstra, \& Grainger, 1998) addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing (McClelland \& Rumelhart, 1981).
-Evidence supporting such integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the translation of the presented word.
-When bilingual participants are presented with words in only one language (target language) in a lexical decision task, translating them to the non-target language took longer when they were part of large neighbourhoods (Van Heuven et al., 1998).
-This suggests that words presented in the target language activated orthographic neighbourhoods in the non-target language, which competed for selection with the target word during recognition.
+The Bilingual Interactive Activation (BIA/BIA+) (Dijkstra \& Van Heuven, 2002; Van Heuven et al., 1998) addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with each other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing (McClelland \& Rumelhart, 1981).
+Evidence supporting such an integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word's translation.
+When bilingual participants are presented with words in one language in a lexical decision task, translating them to the other language took longer when they were part of large neighbourhoods (Van Heuven et al., 1998).
+This suggests that presented words activated orthographic neighbourhoods in the other language, which competed for selection with the target word during recognition.
 
 A more recent model of bilingual lexical processing is Multilink (Dijkstra et al., 2019), which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
-This model assumes an integrated lexicon in which word representations from the two languages establish connections in a way that is similar to what words of a single language do.
-Such model also assumes that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
-It also proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-word connections).
-In addition, this model suggests that the strength of the association between L2 representations and their concept is a function of language users' proficiency in L2.
-A remarkable correlation between Multilink simulations of a translation elicitation task and behavioural data on Dutch-English bilinguals (Christoffels et al., 2006) has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
+This model assumes an integrated lexicon in which word representations from two languages establish connections in a way that is similar to connections between words of a single language.
+It also makes the assumption that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
+It also proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-to-word connections).
+In addition, this model suggests that the strength of the association between L2 representations and their concept depends on language users' proficiency in L2.
+A strong correlation between Multilink simulations of a translation elicitation task and behavioural data from Dutch-English bilinguals (Christoffels et al., 2006) has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
 Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
@@
 The fact that Multilink successfully fitted the experimental data from Christoffels et al.
-(2006), despite the absence of any word-word connections between translation equivalents, is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+(2006) in the absence of word-to-word connections between translation equivalents is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
 It should be taken into account, however, that Christoffels et al.
-(2006) only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2), and that the model specification under which the quantitative predictions were generated, assumes similar vocabulary sizes across languages and same strength of word-concept connections of L1 and L2 representations.
+(2006) only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
+The model specification under which the quantitative predictions were generated assumes similar vocabulary sizes across languages, and also that the word-to-concept connections of L1 and L2 representations have same strength.
 The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-word connections during translation, especially during backward translation, and this prediction remains untested my Multilink.
+It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-to-word connections during translation, especially during backward translation, and this prediction remains untested by Multilink.
 
 Most of the studies referenced so far rely heavily on production and comprehension of printed words: in these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
-This is a limitation regarding the generalisation of their findings to the auditory modality, especially when cognateness or more generally any measure of similarity between word-forms, is involved.
-Consider the words \emph{radio} (English) and \emph{radio} (Catalan).
-Both words are evidently cognates, as their orthographical word form is identical.
-However, their phonological form is not identical: /ˈɹeɪdiˌəʊ/ and /ˈraðjo/.
-Although many English listeners might understand the Spanish word-form, even with no previous knowledge of
-Spanish, it is likely that they will find the translation process more challenging when presented with the auditory word-form than when presented with the visual word-form.
-For the same reason, measuring the phonological similarity between two word forms is substantially more challenging than measuring their orthographic similarity.
-Most studies use symbolic representations of the word-forms of interest for compute their similarity in either modality: character strings in the case of orthographic word-forms (\emph{radio}-\emph{radio}), and phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA transcription) in the case of phonological word forms, which also consists on a string of characters.
-Either way, the similarity between the pair of word-forms is frequently measured by applying some algorithm that computes the edit distance between both.
-The Leveshtein distance is one of the most widely known: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string (Levenshtein \& others, 1966), and has been successfully used to estimate the similarity of cross-language word-pairs (Gooskens \& Heeringa, 2004; Schepens et al., 2012).
-The Levenshtein distance between the orthographic word-forms is 0: both strings are identical.
-But the distance between ɹeɪdiəʊ and raðjo is 35, which is the number of editions that separates both character strings.
-There is yet an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ are symbolic representations that do not capture the all of the features that are present in the acoustic signal when either word is spoken, and therefore any measure of similarity that uses such transcriptions as input will overestimate their actual similarity.
-Taken this into consideration, it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
-In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be overestimating the availability of such information in the auditory modality.
+There are limitations in generalising their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
+Consider the words \emph{radio} (English) and \emph{radio} (Spanish).
+Both words are clearly cognates, as their orthographical word form is identical.
+However, their phonological forms (English /ˈɹeɪ.di.əʊ/-- Spanish /ˈra.ðjo/) are not identical.
+Although many English listeners might recognise the Spanish word-form--even with no previous knowledge of Spanish--it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
 
-In the present study we test listener's reliance on phonological similarity when translating words from a non-native language, in order to investigate the plausibility of the lexical route during backward translation.
-We tested monolingual adults in a translation elicitation task: they were presented with spoken words in a non-native language, and later asked to translate them to their native language.
-Monolinguals can be considered a particular (extreme) case of unbalanced bilingualism, in which their vocabulary size in L2 is null, and word-concept connections are absent.
-In this case, monolinguals can only use the lexical route to succeed in the translation task, and they can only do so by exploiting the cognateness of the words presented and their translation in the native language.
-We hypothesised that if participants are able to translate words in L2, without knowledge of the language it belongs to, and based on their form-similarity with the target word, then the lexical route should be in place for low-proficiency bilinguals.
-In contrast, if participants are not able to translate words from L2 to L1 regardless of their form-similarity, then low-proficiency bilinguals should rely entirely on word-concept connections for backward translation.
+Measuring the phonological similarity between two word forms is also substantially more challenging than measuring their orthographic similarity.
+Many studies use symbolic representations to compute the simiarity between word-forms of interest in either modality.
+Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (\emph{radio}-\emph{radio}); phonological word-forms are represented by their phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA notation) (see Floccia et al., 2018 for example).
+The similarity between each pair of word-forms is operationalised numerically by applying some algorithm that computes the edit distance between their symbolic representation.
+The Leveshtein distance is one of the most widely known algorithms: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string (Levenshtein \& others, 1966), and has been successfully used to estimate the similarity of cross-language word-pairs (Gooskens \& Heeringa, 2004; Schepens et al., 2012).
+Using our previous examples, the Levenshtein distance between the orthographic word-forms \emph{radio} (English) and \emph{radio} (Spanish) is 0: both strings are identical.
+But the distance between the phonological transcriptions and \ipatext{'ra<U+0329>ðjo} is 35, which is the number of edits that separates both character strings.
+There is also an additional layer of complexity: the phonological transcriptions \ipatext{'<U+0279>e<U+026A>di<U+0329><U+0259><U+028A>} and \ipatext{'ra<U+0329>ðjo} can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.~whether to include elements of aspiration, nasalisation, etc.) can affect how accurate these symbolic representation capture all of the features that are present in the acoustic signal when either word is spoken.
+Therefore, any measure of similarity that uses phonological transcriptions as input may overestimate or underestimate their actual similarity.
+Taking into account that measuring phonological similarity is considerably more complex than measuring orthographic complexity it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
 
-Following Luce and Pisoni (1998) we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
-If this is the case, words whose translation in the native language are part of larger phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target word.
+In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be misrepresenting the availability of such information in the auditory modality.
+In the present study we tested listener's reliance on phonological similarity when translating words from an unfamiliar language, in order to investigate the plausibility of word-to-word during backward translation.
+We tested English and Spanish native speakers in a translation elicitation task: they were presented with spoken words in an unfamiliar, non-native language, and were asked to translate them to their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as \textbf{presented words}, and the correct translation for the presented words as \textbf{target translations}.
+The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and therefore word-to-concept connections are, in principle, absent.
+As a consequence, these participants should only be able use word-to-word connections to succeed in the translation of words in an unfamiliar language.
+We predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
 
 We tested three groups of participants in the translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
-Catalan and Spanish are both Roman languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from a typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
+Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
+
+Following Luce and Pisoni (1998)`s NAM model, we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+If this is the case, has more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target translation.
+For instance, one could expect English participants to incorrectly translate the Spanish word \emph{botón} as \emph{bottom} instead of as its correct translation \emph{button} (because the lexical frequency of \emph{bottom} is higher than that of \emph{button}).
+If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
 
@@
 
-All materials, data, and code used in this study are hosted in the Open Science Framework \url{https://osf.io/9fjxm/} and a GitHub repository \url{https://github.com/bilingual-project/translation-elicitation.git}, along with additional notes.
+All materials, data, and code used in this study are hosted in the Open Science Framework \url{https://osf.io/9fjxm/} and a GitHub repository \url{https://github.com/gongcastro/translation-elicitation.git}, along with additional notes.
 
@@
 Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the local ethical committee (XXXXXXXXXXXX).
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-We excluded data from participants that a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), b) were diagnosed with a language (\emph{n} = 2) , or c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
+We excluded data from participants that (a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), (b) were diagnosed with a language (\emph{n} = 2), or (c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
 
@@
 Then, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in Catalan or Spanish (English participants) or only Catalan (Spanish participants).
+Participants were informed that they would listen to a series of pre-recorded words in either Catalan or Spanish (English participants) or Catalan (Spanish participants).
 They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the list of Catalan or Spanish trials.
+English participants were randomly assigned to the Catalan or Spanish lists.
 Participants in the Catalan list were presented with 86 trials, and participants in the Spanish list were presented with 103 trials.
@@
 Participants were allowed to correct their answer.
-Then, participants pressed the RETURN key to start and new trial.
-We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all.
-Trials where the response was mistyped by only one character were counted as correct, as long as the respond did not correspond to a distinct word.
-Trials in which participants took longer than 10 seconds to respond were also filtered out.
+Then, participants pressed the RETURN/ENTER key to confirm their answer and start and new trial.
+We excluded trials where the participant's answer was not a valid word in the target language or no answer was given.
+Trials in which the response was mistyped by only one character were counted as correct (e.g., ``pengiun'' instead of ``penguin''), as long as the response did not correspond to a distinct word.
+Trials in which participants took longer than 10 seconds to respond were also excluded.
 Participants contributed a total of 7239 valid trials (4809 in Catalan, 2430 in Spanish).
@@
 
-We arranged two lists of words: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
-Words in the Catalan list (listened to by participants) were 5.01 phonemes long on average (\emph{SD} = 1.48, Range = 2-8).
-Their translations to English (typed by participants in the keyboard) were 5.14 characters long on average (\emph{SD} = 1.57, Range = 3-9), and their translations to Spanish were 5.44 characters long on average (\emph{SD} = 1.56, Range = 3-9).
+\hypertarget{input-stimuli-catalanspanish}{%
+\subsubsection{Input stimuli (Catalan/Spanish)}\label{input-stimuli-catalanspanish}}
+
+We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
+Words in the Catalan list were 5.01 phonemes long on average (\emph{SD} = 1.48, Range = 2-8).
 Words in the Spanish list were 5.50 phonemes long on average (\emph{SD} = 1.46, Range = 3-9).
-Their translations to English were 5.27 characters long on average (\emph{SD} = 1.75, Range = 3-12).
+Participants listened to one audio file in each trial, each containing a single word presented in isolation.
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat (Broersma \& Weenink, 2021).
+The average duration of the audio files was 1.20 (\emph{SD} = 0.18, Range = 0.78-1.58).
+The average duration of the Catalan audio files was 1.23 seconds (\emph{SD} = 0.19, Range = 0.80-1.58), and the average duration of the Spanish audio files was 1.16 seconds (\emph{SD} = 0.15, Range = 0.78-1.53).
 
-We extracted the lexical frequencies from SUBTLEX-UK for English words (Van Heuven, Mandera, Keuleers, \& Brysbaert, 2014), SUBTLEX-ESP for Spanish words (Cuetos, Glez-Nosti, Barbon, \& Brysbaert, 2011), and SUBTLEX-CAT for Catalan words (Boada, Guasch, Haro, Demestre, \& Ferré, 2020).
-We extracted or transformed scores as/into Zipf scores (Van Heuven et al., 2014) to so correct their logarithmic distribution, limiting their range roughly between 0 to 7, allowing an easier interpretation of further analyses (Van Heuven et al., 2014).
+\hypertarget{participant-responses-englishspanish}{%
+\subsubsection{Participant responses (English/Spanish)}\label{participant-responses-englishspanish}}
 
-We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) from the CLEARPOND database (Marian, Bartolotti, Chabal, \& Shook, 2012).
-PTHN scores indicate the number of phonological neighbours of the target word with higher lexical frequency, as indicated by its score in the corresponding SUBTLEX database.
-CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, \url{http://espeak.sourceforge.net/}) differs from that of the target word in only one addition, deletion, or substitution.
+Responses given by English participants for Catalan presented words were 5.14 characters long on average (\emph{SD} = 1.57, Range = 3-9), while their translations to Spani responses given by Spanish participants for Catalan words were 5.44 characters long on average (\emph{SD} = 1.56, Range = 3-9).
+Responses tgiven by English participants for Spanish words were 5.27 characters long on average (\emph{SD} = 1.75, Range = 3-12).
+
+\hypertarget{target-translation-lexical-frequency}{%
+\paragraph{Target translation lexical frequency}\label{target-translation-lexical-frequency}}
+
+We extracted the lexical frequencies for English target translations from SUBTLEX-UK(Van Heuven et al., 2014), and frequencies for Spanish from SUBTLEX-ESP (Cuetos et al., 2011).
+We extracted scores as Zipf scores where available or transformed raw scores to Zipf scores to correct their logarithmic distribution, limiting their range roughly between 0 and 7, allowing an easier interpretation of further analyses (Van Heuven et al., 2014).
+
+\hypertarget{target-translation-phonological-neighbourhood-density-pthn}{%
+\paragraph{Target translation phonological neighbourhood density (PTHN)}\label{target-translation-phonological-neighbourhood-density-pthn}}
+
+We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) of the English and Spanish target translations from the CLEARPOND database (Marian et al., 2012).
+PTHN scores indicate the number of phonological neighbours with higher lexical frequency than the target translation, as indicated by its frequency score in the corresponding SUBTLEX database.
+CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, \url{http://espeak.sourceforge.net/}) differs from that of the target translation in only one addition, deletion, or substitution.
 PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
@@
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stimuliplot-1} \caption{Distribution and summary statistics of stimuli properties.
-A\) Lexical frequency of target words, expressed in counts per million and as Zipf scores.
-B\) Number of higher-frequency phonological neighbourhoods of target words (PTHN).
-C\) Normalised Levenshtein phonological similarity between the presented and the target words, expressed in percentages.}\label{fig:stimuliplot}
+\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stimuliplot-1} \caption{Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA).
+Columns indicate: A\) Lexical frequency of target translations, expressed as Zipf scores.
+B\) PTHN of target translations, and C\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages.
+Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words.}\label{fig:stimuliplot}
 \end{figure}
 
-We measured the phonological similarity between translation pairs by computing the Levenshtein similarity between their IPA translations using the \texttt{stringsim} function of the stringdist R package (van der Loo, 2014).
+\hypertarget{cross-language-phonological-similarity-levenshtein}{%
+\subsubsection{Cross-language phonological similarity (Levenshtein)}\label{cross-language-phonological-similarity-levenshtein}}
+
+We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their IPA transcriptions using the \texttt{stringsim} function of the stringdist R package (van der Loo, 2014).
 This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
-First, it computes the edit distance between two character strings (in this case, phoneme symbols) by counting the number of additions, deletions, and substitutions necessary to make both strings identical (Levenshtein \& others, 1966).
+First, it computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical (Levenshtein \& others, 1966).
 This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is percentage score indicating no similarity between the two strings if 0\%, and a perfect match between two strings if 100\%.
-We computed this similarity measure (Levenshtein, from now on) for every translation pair in our stimuli lists.
+The result is a percentage score, with 0\% indicating no similarity between the two strings, and 100\% indicating a perfect match between the two strings.
+We computed this similarity measure (refred simply as Levenshtein, from now on) for every translation pair in our stimuli lists.
+For example, the \emph{table} (\ipatext{'te<U+026A><U+0329>b<U+0259>l})-\emph{mesa} (mesa) translation pair had a 17\% similarity, while the \emph{train} (trEIn)-\emph{tren} (\ipatext{t<U+027E>en}) translation pair had a 60\% similarity.
 Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
-
-Participants listened to one audio file in each trial.
-This audio file corresponded to a word in Catalan (for Spanish speakers, and for English participants allocated in the Catalan condition) or Spanish (for English speakers allocated in the Spanish condition).
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat (Broersma \& Weenink, 2021).
-The average duration of the audios was 1.20 (\emph{SD} = 0.18, Range = 0.78-1.58).
-The average duration of the Catalan audio files was 1.23 seconds (\emph{SD} = 0.19, Range = 0.80-1.58), and the average duration of the Spanish audios was 1.16 seconds (\emph{SD} = 0.15, Range = 0.78-1.53).
 
@@
 
-We modelled the probability of participants guessing the correct translation of each input word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We modelled the probability of participants guessing the correct target translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
 We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant (Model 0).
-Second, we added the fixed effect of lexical frequency of the target word (\(Frequency\)).
-We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood (\(PTHN\)), the main effect of Levenshtein similarity (\(Levenshtein\)), the \(PTHN \times Levenshtein\) interaction, the group of participants (\(Group\), with levels ENG-CAT, ENG-SPA, and SPA-CAT), and finally, the \(Levenshtein \times Group\) interaction).
-We specified two sum-coded \emph{a priori} contrasts for the \(Group\) predictor: one comparing English natives (ENG-CAT = -0.25, ENG-SPA = -0.25) against Spanish natives (SPA-CAT = +0.5), and one comparing English natives translating Spanish words (ENG-SPA = -0.5) against English natives translating Catalan words (ENG-CAT = +0.5) (Schad, Vasishth, Hohenstein, \& Kliegl, 2020).
-Each model included random intercepts by item (to account for the correlation of responses from different participants to the same word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different words) (Barr, Levy, Scheepers, \& Tily, 2013).
+Secondly, we added the fixed effect of lexical frequency of the target translation (\(Frequency\)).
+We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood (\(PTHN\)), the main effect of Levenshtein similarity (\(Levenshtein\)), the \(PTHN \times Levenshtein\) interaction, the group of participants (\(Group\), with levels cat-ENG, spa-ENG, and cat-SPA), and finally, the \(Levenshtein \times Group\) interaction).
+We specified two sum-coded \emph{a priori} contrasts for the \(Group\) predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) (Schad et al., 2020).
+Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) (Barr et al., 2013).
 We will later refer to these models as Model 0 to Model 6.
@@
 
-\$\$
-\begin{align}
+\$\$\begin{align}
 
@@
 
-\end{align}
-\$\$
+\end{align}\$\$
 
 To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (\emph{LOO}) (Vehtari, Gelman, \& Gabry, 2017).
+We compared models using leave-one-out cross-validation (\emph{LOO}) (Vehtari et al., 2017).
 More information about the models and model comparison can be found in Appendix 2.
@@
 Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increments or decrements in probability of correct translation (Gelman, Hill, \& Vehtari, 2020).
+This way, the coefficients are expressed as increments or decrements in the probability of correct translation (Gelman et al., 2020).
 
@@
 
-Overall, participants were 42.65\%, (95\% \emph{CrI} = {[}37.89\%, 47.63\%{]}) likely to produce correct translations.
-The lexical frequency of the correct translation had little impact in accuracy: every standard deviation increment (\emph{SD} = 0.57 Zipf points) decreased the probability of correct translation in -0.63\% (95\% \emph{CrI} = {[}-0.63\%, -3.72\%{]}, \(BF_{\beta\neq< 0}\) = 1.53).
-The number of phonological neighbours of the correct translation with higher lexical frequency (\emph{PTHN}) slightly decreased the probability of a correct responses in -2.91\% (95\% \emph{CrI} = {[}-5.90\%, 0.07\%{]}, \(BF_{\beta \neq 0}\) = 0.32) for every increase standard deviation increment (1 \emph{SD} = 3.94 neighbours).
-Finally, phonological similarity between translations had a strong effect on participants' accuracy: for every \emph{SD} increment in Levenshtein similarity (1 \emph{SD} = 23\% similarity), translation accuracy increased in 12.18\% (95\% \emph{CrI} = {[}7.63\%, 16.64\%{]}, \(BF_{\beta \neq 0}\) = \(>3.09 \times 10^{18}\)).
-This difference increased in 0.23\% for every \emph{SD} increment in \emph{PTHN} (95\% \emph{CrI} = {[}-2.79\%, 3.30\%{]}, \(BF_{\beta \neq 0}\) = 1.68)
+Overall, participants had 43.08\%, (95\% \emph{CrI} = {[}38.20\%, 48.00\%{]}) likelihood to produce correct translations.
+The lexical frequency of the target translation had little impact on accuracy: every standard deviation increment (\emph{SD} = 0.57 Zipf points) decreased the probability of correct translation in 1.14\% (95\% \emph{CrI} = {[}1.14\%, -2.40\%{]}, \(BF_{\beta\neq< 0}\) = 1.20).
+The number of phonological neighbours with higher lexical frequency than the target translation (\emph{PTHN}) slightly decreased the probability of a correct responses by -1.70\% (95\% \emph{CrI} = {[}-5.26\%, 1.98\%{]}, \(BF_{\beta \neq 0}\) = 0.91) for every increase standard deviation increment (1 \emph{SD} = 3.94 neighbours).
+Finally, phonological similarity between the presented word and target translation had a strong effect on participants' accuracy: for every \emph{SD} increment in Levenshtein similarity (1 \emph{SD} = 23\% similarity), translation accuracy increased in 12.30\% (95\% \emph{CrI} = {[}7.48\%, 17.39\%{]}, \(BF_{\beta \neq 0}\) = \(>3.09 \times 10^{18}\)).
+The interaction effect between \(PTHN \times Levenshtein\) was small: a +1 SD increment in Levenshtein increased the effect of \emph{PTHN} by -0.11\% (95\% \emph{CrI} = {[}-3.72\%, 3.54\%{]}, \(BF_{\beta \neq 0}\) = 1.39)
 
-Spanish natives translating Catalan words were 3.67\% more accurate than English natives translating Spanish and Catalan words (95\% \emph{CrI} = {[}-1.18\%, 8.64\%{]}, \(BF_{\beta \neq 0}\) = 0.38).
-This difference increased in 3.38\% for every one \emph{SD} increment in Levenshtein similarity (95\% \emph{CrI} = {[}-1.64\%, 8.34\%{]}, \(BF_{\beta \neq 0}\) = 0.47).
-English natives translating Catalan words were 0.87\% more accurate than those translating Spanish words (95\% \emph{CrI} = {[}-5.39\%, 3.63\%{]}, \(BF_{\beta \neq 0}\) = 0.38).
-For every one \emph{SD} increment in Levenshtein similarity, the effect increased in 3.60\% (95\% \emph{CrI} = {[}-7.71\%, 0.48\%{]}, \(BF_{\beta \neq 0}\) = 0.47).
+Spanish natives translating Catalan words were 3.74\% more accurate than English natives translating Spanish and Catalan words (95\% \emph{CrI} = {[}-1.09\%, 8.64\%{]}, \(BF_{\beta \neq 0}\) = 0.35).
+The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in 3.42\% for every one \emph{SD} increment in Levenshtein similarity (95\% \emph{CrI} = {[}-1.52\%, 8.45\%{]}, \(BF_{\beta \neq 0}\) = 0.41).
+English natives translating Catalan words were 0.60\% more accurate than those translating Spanish words (95\% \emph{CrI} = {[}-3.98\%, 5.17\%{]}, \(BF_{\beta \neq 0}\) = 1.00).
+For every one \emph{SD} increment in Levenshtein similarity, the effect of presented language increased in 3.08\% (95\% \emph{CrI} = {[}-1.24\%, 7.43\%{]}, \(BF_{\beta \neq 0}\) = 0.41).
 
@@
 \caption{\label{fig:marginaleffects}Expected mean posterior predictions from Model 4.
-A) Population-level predictions: the X-axis and the Y-axis represent the Levenshtein distance (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 200 observations from our model: 100 simulations for words with low \emph{PTHN} (-1 \emph{SD}) and 100 simulations for high \emph{PTHN} (+1 \emph{SD}) words.
-We did this across the range of values of the Levenshtein scores.
+A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0\% Levenshtein), 100 simulations for translations with mean similarity (20.51\% Levenshtein), and 100 simulations for translations with the maximum similarity observed in the dataset (83.33\% Levenshtein).
+These simulations were obtained for the range of values of the PTHN scores.
 For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink in the case of low PTHN words, and blue in the case of high \emph{PTHN} words.
-We also plotted the mean of the high \emph{PTHN} (black solid line) and low \emph{PTHN} (black dashed line) simulations to indicate the expected mean value of the posterior predictions of the model.
+Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
+Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
 The dispersion of the lines indicates the uncertainty of our predictions.
@@
 
-The conditional intra-class correlation (\emph{ICC}) for items was 69.64\%, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items (Gelman \& Hill, 2006).
-Item-level intercepts showed remarkable variability (\emph{SD} = 96.92\%, 95\% CrI = {[}94.75\%, 98.40\%{]}), indicating that participants' accuracy differed substantially across words: some words were correctly translated nearly always (e.g., Spanish \emph{tomate}-\emph{tomato}, /tomate/-/təmɑːtoʊ/, 99.73\%, 95\% CrI = {[}96.72\%, 99.99\%{]}), while others were almost never translated correctly (e.g., Catalan \emph{colom}-\emph{pigeon}, /kulom/-/pɪdʒɪn/, 0.08\%, 95\% CrI = {[}0.00\%, 0.81\%{]}).
-Participant-level intercepts also showed high variability (\emph{SD} = 74.72\%, 95\% CrI = {[}70.39\%, 79.47\%{]}), suggesting that average translation accuracy also varied significantly across participants.
-Finally, participant-level estimated slopes showed less variability: \emph{SD} of \(Frequency\), \(PTHN\), \(Levenshtein\), \(PTHN \times Levenshtein\), and \(Levenshtein \times Group 2\) ranged between 2.32\%, and 4.30\%.
-The participant-level estimated slopes of the \(Leveshtein \times Group 1\) interaction are an exception, though (\emph{SD} = 41.18\%, 95\% CrI = {[}24.79\%, 57.80\%{]}).
+The conditional intra-class correlation {[}\emph{ICC}; Gelman and Hill (2006){]} for items was 71.49\%, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items.
+Item-level intercepts showed remarkable variability (\emph{SD} = 96.91\%, 95\% CrI = {[}94.66\%, 98.36\%{]}), indicating that participants' accuracy differed substantially across words: some words were very likely to be correctly translated (e.g., Spanish \emph{tomate}-\emph{tomato}, /tomate/-/təmɑːtoʊ/, 99.73\%, 95\% CrI = {[}96.40\%, 99.99\%{]}), while others were almost never translated correctly (e.g., Catalan \emph{colom}-\emph{pigeon}, /kulom/-/pɪdʒɪn/, 0.08\%, 95\% CrI = {[}0.00\%, 1.00\%{]}).
+Participant-level intercepts also showed high variability (\emph{SD} = 75.73\%, 95\% CrI = {[}71.33\%, 80.33\%{]}), suggesting that average translation accuracy also varied significantly across participants.
+Finally, participant-level estimated slopes showed less variability: \emph{SD} of \(Frequency\), \(PTHN\), \(Levenshtein\), \(PTHN \times Levenshtein\), and \(Levenshtein \times Group 2\) ranged between 2.56\%, and 37.19\%.
+The participant-level estimated slopes of the \(Leveshtein \times Group 1\) interaction are an exception, though (\emph{SD} = 37.19\%, 95\% CrI = {[}14.92\%, {]}).
 This indicates that the impact of phonological similarity (\(Levenshtein\)) on the difference in accuracy between English and Spanish natives differed substantially across participants.
@@
 Finally, participant-level effects showed weak Pearson correlations overall (\textless10\%), with three exceptions.
-The estimated correlation between the intercepts and the slopes of \(Frequency\) was 29.30\% (95\% CrI = {[}-14.99\%, 63.21\%{]}, \(BF_{Cor \neq 0}\) = \texttt{r}round(h\_cor\_intercept\_frequency, 2)`).
-The estimated correlation between the slopes of \(Frequency\) and \(Levenshtein \times Group 1\) was 21.17\% (95\% CrI = {[}-19.73\%, 56.32\%{]}, \(BF_{Cor \neq 0}\) = 0.60).
-The estimated correlation between the intercepts and the slopes of \(PTHN\) was 10.88\% (95\% CrI = {[}-28.52\%, 45.58\%{]}, \(BF_{Cor \neq 0}\) = 0.96).
+The estimated correlation between the intercepts and the slopes of \(Frequency\) was 35.81\% (95\% CrI = {[}-2.19\%, 65.22\%{]}, \(BF_{Cor \neq 0}\) = 0.15).
+The estimated correlation between the slopes of \(Frequency\) and \(Levenshtein \times Group 1\) was 26.07\% (95\% CrI = {[}-13.12\%, 57.14\%{]}, \(BF_{Cor \neq 0}\) = 0.37).
+The estimated correlation between the intercepts and the slopes of \(PTHN\) was 1.81\% (95\% CrI = {[}-39.01\%, 38.35\%{]}, \(BF_{Cor \neq 0}\) = 1).
 The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-boada2020subtlex}{}}%
-Boada, R., Guasch, M., Haro, J., Demestre, J., \& Ferré, P.
-(2020).
-SUBTLEX-CAT: Subtitle word frequencies and contextual diversity for catalan.
-\emph{Behavior Research Methods}, \emph{52}(1), 360--375.
-
 \leavevmode\vadjust pre{\hypertarget{ref-broersma2021praat}{}}%
-Broersma, P., \& Weenink, D. (2021). Praat: Doing phonetics by computer {[}computer program{]} (Version Version 6.1.54). Retrieved from \url{http://www.praat.org/}
+Broersma, P., \& Weenink, D. (2021). \emph{Praat: Doing phonetics by computer {[}computer program{]}} (Version 6.1.54). \url{http://www.praat.org/}
 
@@
 Carpenter, B., Gelman, A., Hoffman, M.
-D., Lee, D., Goodrich, B., Betancourt, M., \ldots{} Riddell, A.
+D., Lee, D., Goodrich, B., Betancourt, M., Brubaker, M.
+A., Guo, J., Li, P., \& Riddell, A.
 (2017).
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-elias2022cross}{}}%
+Elias, M., \& Degani, T.
+(2022).
+Cross-language interactions during novel word learning: The contribution of form similarity and participant characteristics.
+\emph{Bilingualism: Language and Cognition}, 1--18.
+
+\leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
+Floccia, C., Sambrook, T.
+D., Delle Luche, C., Kwok, R., Goslin, J., White, L., Cattani, A., Sullivan, E., Abbot-Smith, K., Krott, A., \& others.
+(2018).
+I: INTRODUCTION.
+\emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29.
+
 \leavevmode\vadjust pre{\hypertarget{ref-gelman2006data}{}}%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-de1992determinants}{}}%
-Groot, Annette M. de. (1992). Determinants of word translation. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{18}(5), 1001.
+Groot, A. M. de. (1992). Determinants of word translation. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{18}(5), 1001.
 
 \leavevmode\vadjust pre{\hypertarget{ref-degroot1994forward}{}}%
-Groot, Annette MB de, Dannenburg, L., \& Vanhell, J. G. (1994). Forward and backward word translation by bilinguals. \emph{Journal of Memory and Language}, \emph{33}(5), 600--629.
+Groot, A. M. de, Dannenburg, L., \& Vanhell, J. G. (1994). Forward and backward word translation by bilinguals. \emph{Journal of Memory and Language}, \emph{33}(5), 600--629.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
-Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. In \emph{Soviet physics doklady} (Vol. 10, pp. 707--710). Soviet Union.
+Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. \emph{Soviet Physics Doklady}, \emph{10}, 707--710.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-luce1990similarity}{}}%
-Luce, P. A., Pisoni, D. B., \& Goldinger, S. D. (1990). Similarity neighborhoods of spoken words.
+Luce, P. A., Pisoni, D. B., \& Goldinger, S. D. (1990). \emph{Similarity neighborhoods of spoken words.}
 
 \leavevmode\vadjust pre{\hypertarget{ref-marian2012clearpond}{}}%
-Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012). CLEARPOND: Cross-linguistic easy-access resource for phonological and orthographic neighborhood densities.
+Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012). \emph{CLEARPOND: Cross-linguistic easy-access resource for phonological and orthographic neighborhood densities}.
 
@@
 Peirce, J., Gray, J.
-R., Simpson, S., MacAskill, M., Höchenberger, R., Sogo, H., \ldots{} Lindeløv, J.
+R., Simpson, S., MacAskill, M., Höchenberger, R., Sogo, H., Kastman, E., \& Lindeløv, J.
 K.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-rcore2019r}{}}%
-RCore, T. (2019). R: A language and environment for statistical computing. R foundation for statistical computing, austria.
+RCore, T. (2019). \emph{R: A language and environment for statistical computing. R foundation for statistical computing, austria}.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-van2014stringdist}{}}%
-van der Loo, M. P. J. (2014). The stringdist package for approximate string matching. \emph{The {R} {J}ournal}, \emph{6}, 111--122. Retrieved from \url{https://CRAN.R-project.org/package=stringdist}
+van der Loo, M. P. J. (2014). The stringdist package for approximate string matching. \emph{The {R} {J}ournal}, \emph{6}, 111--122. \url{https://CRAN.R-project.org/package=stringdist}
 
@@
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
-D., François, R., \ldots{} Yutani, H.
+D., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T.
+L., Miller, E., Bache, S.
+M., Müller, K., Ooms, J., Robinson, D., Seidel, D.
+P., Spinu, V., \ldots{} Yutani, H.
 (2019).

### 2022-03-21 R/00_stimuli.R

@@ -5,29 +5,60 @@ get_clearpond <- function(clearpond_path){
             function(x){
                 read.csv(x) %>% 
                     clean_names() %>% 
-                    select("word", "pho_word", "freq_per_million", "pthn")
+                    select(
+                        "word",
+                        "pho_word", 
+                        "freq_per_million", 
+                        "pthn"
+                    )
             }
         ) %>% 
         bind_rows(.id = "group") %>% 
         as_tibble() %>%
-        distinct(word, group, .keep_all = TRUE) %>% 
-        rename(word2 = word, phon_clearpond = pho_word, frequency = freq_per_million)
+        distinct(
+            word, 
+            group, 
+            .keep_all = TRUE
+        ) %>% 
+        rename(
+            word2 = word,
+            phon_clearpond = pho_word,
+            frequency = freq_per_million
+        )
     return(clearpond)
 }
 
 # get Leveshtein similarity scores
 get_levenshtein <- function(stimuli_path){
     
-    levenshtein <- map(
-        c("ENG-SPA" = "English-Spanish", "ENG-CAT" = "English-Catalan", "SPA-CAT" = "Spanish-Catalan"),
-        function(x) read_xlsx(stimuli_path, sheet = x)
+    levenshtein <- c(
+        "ENG-SPA" = "English-Spanish", 
+        "ENG-CAT" = "English-Catalan",
+        "SPA-CAT" = "Spanish-Catalan"
     ) %>% 
+        map(
+            function(x) read_xlsx(stimuli_path, sheet = x)
+        ) %>% 
         bind_rows(.id = "group") %>% 
         clean_names() %>% 
-        select(group, word1, word2, ipa1, ipa2) %>% 
+        select(
+            group, 
+            word1, 
+            word2, 
+            ipa1,
+            ipa2
+        ) %>% 
         mutate(
-            n_char = ifelse(nchar(ipa1) > nchar(ipa2), nchar(ipa1), nchar(ipa2)),
-            lv = stringsim(ipa1, ipa2, method = "lv")
+            n_char = ifelse(
+                nchar(ipa1) > nchar(ipa2),
+                nchar(ipa1),
+                nchar(ipa2)
+            ),
+            lv = stringsim(
+                ipa1, 
+                ipa2, 
+                method = "lv"
+            )
         )
     
     return(levenshtein)
[…]

### 2022-03-21 R/utils.R

@@ -61,49 +61,101 @@ replace_non_ascii <- function(x){
 
 
 # download clearpond
-# import_clearpond <- function(language = c("english", "dutch", "french", "spanish", "german")){
-#     require(tidyverse)
-#     require(janitor)
-#     urls <- tribble(
-#         ~lang, ~url, ~data, ~header,
-#         "english", "https://clearpond.northwestern.edu/englishCPdatabase2.zip", "englishCPdatabase2.txt", "clearpondHeaders_EN.txt",
-#         "dutch", "https://clearpond.northwestern.edu/dutchCPdatabase2.zip", "dutchCPdatabase2.txt", "clearpondHeaders_NL.txt",
-#         "french", "https://clearpond.northwestern.edu/frenchCPdatabase2.zip", "frenchCPdatabase2.txt", "clearpondHeaders_FR.txt",
-#         "german", "https://clearpond.northwestern.edu/germanCPdatabase2.zip", "germanCPdatabase2.txt", "clearpondHeaders_DE.txt",
-#         "spanish", "https://clearpond.northwestern.edu/spanishCPdatabase2.zip", "spanishCPdatabase2.txt", "clearpondHeaders_SP.txt"
-#     ) %>% 
-#         filter(lang %in% language)
-#     
-#     dir <- tempdir()
-#     files <- replicate(tempfile(), n = length(urls$lang))
-#     d <- pmap(
-#         .l = list(url = as.list(urls$url), file = as.list(files), data = as.list(urls$data), header = as.list(urls$header)),
-#         .f = function(url = .l[[1]], file = .l[[2]], data = .l[[3]], header = .l[[4]]) {
-#             download.file(url, destfile = file)
-#             unzip(zipfile = file, exdir = dir)
-#             headers <- c("word", read.delim(paste0(dir, .Platform$file.sep, header))[,1])
-#             d <- read.delim(paste0(dir, .Platform$file.sep, data)) %>% 
-#                 `colnames<-`(., headers) %>% 
-#                 as_tibble() %>% 
-#                 mutate_at(vars(ends_with("W")), ~str_split(., pattern = ";"))
-#             return(d)
-#         }
-#     ) %>% 
-#         set_names(language) %>% 
-#         bind_rows(.id = "language") %>% 
-#         clean_names()
-#     return(d)
-# }
+import_clearpond <- function(
+    language = c(
+        "english", 
+        "dutch", 
+        "french",
+        "spanish",
[…]

### 2022-03-28 _targets.R
Add group as a word-level slope

@@ -156,22 +156,19 @@ list(
                     (1 | participant) + 
                     (1 | word)
             ),
-            
             f_1 = bf(
                 correct ~ 1 + 
                     frequency_zipf_std +
-                    (1 + frequency_zipf_std | participant) +
+                    (1 + frequency_zipf_std | participant) + 
                     (1 | word)
             ),
-            
             f_2 = bf(
                 correct ~ 1 + 
-                    frequency_zipf_std + 
+                    frequency_zipf_std +
                     pthn_std + 
                     (1 + frequency_zipf_std + pthn_std | participant) +
                     (1 | word)
             ),
-            
             f_3 = bf(
                 correct ~ 1 + 
                     frequency_zipf_std +
@@ -180,7 +177,6 @@ list(
                     (1 + frequency_zipf_std + pthn_std + lv_std | participant) 
                 + (1 | word)
             ),
-            
             f_4 = bf(
                 correct ~ 1 + 
                     frequency_zipf_std +
@@ -190,28 +186,26 @@ list(
                     (1 + frequency_zipf_std + pthn_std*lv_std | participant) + 
                     (1 | word)
             ),
-            
             f_5 = bf(
                 correct ~ 1 + 
-                    frequency_zipf_std +  
+                    frequency_zipf_std +
                     pthn_std + 
                     lv_std + 
                     lv_std:pthn_std + 
                     group +
                     (1 + frequency_zipf_std + pthn_std + lv_std + lv_std:pthn_std | participant) + 
-                    (1 | word)
+                    (1 + group | word)
             ),
-            
             f_6 = bf(
                 correct ~ 1 +
-                    frequency_zipf_std + 
+                    frequency_zipf_std +
                     pthn_std + 
                     lv_std + 
                     lv_std:pthn_std + 
                     group + 
                     group:lv_std +
                     (1 + frequency_zipf_std + pthn_std + lv_std + lv_std:pthn_std + group:lv_std | participant) + 
-                    (1 | word))
+                    (1 + group | word))
         )
     ),
     # model prior
@@ -219,7 +213,7 @@ list(
         model_prior,
         c(
             prior(normal(0, 0.1), class = "Intercept"),
-            prior(normal(0, 0.1), clas = "b"),
+            prior(normal(0, 0.1), class = "b"),
[…]

### 2022-03-28 docs/index.Rmd
Add group as a word-level slope

@@
   ggplot2.continuous.fill = scale_fill_gradient2(
-    low = "#04bf78", mid = "white", high = "#FFC20A", 
-    na.value = "white", limits = c(-0.4, 0.4)
+    low = "#04bf78",
+    mid = "white",
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
   ),
   ggplot2.continuous.colour = scale_fill_gradient2(
-    low = "#04bf78", mid = "white", high = "#FFC20A", 
-    na.value = "white", limits = c(-0.4, 0.4)
+    low = "#04bf78",
+    mid = "white",
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
   )
@@
 stimuli %>% 
-  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
+  select(
+    group,
+    word1,
+    frequency, 
+    frequency_zipf,
+    pthn, 
+    lv
+  ) %>% 
   group_by(group) %>% 
@@
   fmt_percent(matches("lv")) %>% 
-  tab_spanner(md("**Freq. (million)**"), c("frequency_median", "frequency_sd")) %>% 
-  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
-  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
-  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
+  tab_spanner(
+    md("**Freq. (million)**"), 
+    c("frequency_median", "frequency_sd")
+  ) %>% 
+  tab_spanner(
+    md("**Freq. (Zipf)**"),
+    matches("frequency_zipf")) %>% 
+  tab_spanner(
+    md("***PTHN***"), 
+    matches("pthn")) %>% 
+  tab_spanner(
+    md("**Levenshtein**"),
+    matches("lv")) %>% 
   cols_label(
@@
   summary_rows(
-    columns = c(matches("median"), -matches("lv")),
+    columns = c(
+      matches("median"), 
+      -matches("lv")
+    ),
     fns = list(
@@
   tab_style(
-    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    style = cell_text(
+      align = "left",
+      weight = "bold", 
+      size = "medium"
+    ),
     locations = cells_title(groups = "title")
@@
   tab_style(
-    style = cell_text(align = "left", size = "medium", style = "italic"),
+    style = cell_text(
+      align = "left", 
+      size = "medium",
+      style = "italic"
+    ),
     locations = cells_title(groups = "subtitle")
@@
   tab_style(
-    style = cell_text(align = "left", size = "medium"),
+    style = cell_text(
+      align = "left",
+      size = "medium"
+    ),
     locations = cells_source_notes()
@@
   tab_style(
-    style = cell_text(align = "left", size = "medium"),
+    style = cell_text(
+      align = "left",
+      size = "medium"
+    ),
     locations = cells_stub_grand_summary()
@@
   tab_style(
-    style = cell_borders(sides = c("left", "right"), color = "white"),
+    style = cell_borders(
+      sides = c("left", "right"),
+      color = "white"
+    ),
     locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
+      cells_body(), 
+      cells_row_groups(), 
+      cells_column_spanners(), 
+      cells_stub(), 
+      cells_stub_grand_summary(),
+      cells_source_notes(),
+      cells_column_labels(), 
+      cells_stubhead()
     )
@@
   tab_style(
-    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
+    style = list(
+      cell_text(
+        align = "left", 
+        size = "medium"
+      ), 
+      cell_borders(
+        sides = "all", 
+        color = "white"
+      )
+    ),
     locations = cells_stub()
@@
   tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
+    style = cell_text(
+      align = "center",
+      size = "medium", 
+      weight = "normal", 
+      style = "normal"
+    ),
+    locations = list(
+      cells_column_spanners(),
+      cells_column_labels()
+    )
   ) %>% 
   tab_style(
-    style = cell_borders(sides = "all", color = "white"),
+    style = cell_borders(
+      sides = "all",
+      color = "white"
+    ),
     locations = cells_body(columns = 1:9)
@@
   tab_style(
-    style = cell_borders(sides = "top", color = "white"),
+    style = cell_borders(
+      sides = "top", 
+      color = "white"
+    ),
     locations = cells_body(columns = 1:9)
@@
 stimuli_coded %>% 
-  distinct(test_language, word1, .keep_all = TRUE) %>% 
-  left_join(count(., test_language, name = "n_total")) %>% 
-  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
-  ggplot(aes(frequency_zipf, fill = test_language)) +
-  facet_wrap(~test_language, ncol = 1) +
-  geom_histogram(bins = 25, colour = "white") +
-  labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language")  +
-  scale_y_continuous(breaks = seq(0, 12, 2)) +
+  distinct(
+    test_language,
+    word1, 
+    .keep_all = TRUE
+  ) %>% 
+  left_join(
+    count(
+      ., 
+      test_language,
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    test_language = paste0(
+      test_language,  
+      " (N = ", n_total, 
+      ")"
+    )
+  ) %>% 
+  ggplot() +
+  aes(
+    frequency_zipf,
+    fill = test_language
+  ) +
+  facet_wrap(
+    ~test_language,
+    ncol = 1
+  ) +
+  geom_histogram(
+    bins = 25, 
+    colour = "white"
+  ) +
+  labs(
+    x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database",
+    y = "# trials",
+    fill = "Test language"
+  )  +
+  scale_y_continuous(
+    breaks = seq(0, 12, 2)
+  ) +
   theme(
     legend.position = "none",
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    )
   )
@@
 stimuli_coded %>% 
-  left_join(count(., test_language, name = "n_total")) %>% 
-  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
-  count(test_language, pthn) %>% 
-  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
-  facet_wrap(~test_language, ncol = 1) +
+  left_join(
+    count(
+      ., 
+      test_language, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    test_language = paste0(
+      test_language, 
+      " (N = ", 
+      n_total, 
+      ")"
+    )
+  ) %>% 
+  count(
+    test_language, 
+    pthn
+  ) %>% 
+  ggplot() +
+  aes(
+    reorder(as.factor(pthn), -n), 
+    n,
+    fill = test_language
+  ) +
+facet_wrap(
+  ~test_language,
+  ncol = 1
+) +
   geom_col(position = position_dodge()) +
-  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
+  labs(
+    x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database",
+    y = "# trials",
+    fill = "Test language"
+  ) +
   theme(
     legend.position = "none",
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    )
   )
@@
 stimuli_coded %>% 
-  left_join(count(., group, name = "n_total")) %>% 
-  mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
-  ggplot(aes(group, lv, fill = group, colour = group)) +
+  left_join(
+    count(
+      ., 
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(
+      group,  
+      " (N = ",
+      n_total,
+      ")"
+    )
+  ) %>% 
+  ggplot() +
+  aes(
+    group,
+    lv,
+    fill = group, 
+    colour = group
+  ) +
   geom_violin() +
-  geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
-  geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
-  labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
-  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  geom_point(
+    shape = 1, 
+    stroke = 1, 
+    position = position_jitter(width = 0.1),
+    colour = "black",
+    alpha = 0.25
+  ) +
+  geom_boxplot(
+    colour = "black",
+    fill = "white", 
+    width = 0.1, 
+    size = 0.75
+  ) +
+  labs(
+    x = "Group", 
+    y = "Phonological Leveshtein distance\n(corrected for string length)",
+    fill = "Test language"
+  ) +
+  scale_y_continuous(
+    limits = c(0, 1),
+    labels = scales::percent
+  ) +
   theme(
@@
     axis.title.x = element_blank(),
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    )
   )
@@
     aes(
-      xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd, 
-      ymin = -Inf, ymax = Inf, fill = group
-    ), inherit.aes = FALSE, alpha = 0.15
-  ) +   
-  geom_histogram(bins = 20, colour = "white") +
+      xmin = frequency_zipf_median-frequency_zipf_sd,
+      xmax = frequency_zipf_median+frequency_zipf_sd, 
+      ymin = -Inf,
+      ymax = Inf, 
+      fill = group
+    ),
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20, 
+    colour = "white"
+  ) +
   geom_vline(
     data = medians, 
-    aes(xintercept = frequency_zipf_median, group = group), 
+    aes(
+      xintercept = frequency_zipf_median,
+      group = group
+    ), 
     colour = "black",
@@
   stimuli %>% 
-  count(group, pthn) %>% 
-  mutate(group = case_when(
+  count(
+    group, 
+    pthn
+  ) %>% 
+  mutate(
+    group = case_when(
       group=="ENG-CAT" ~ "cat-ENG",
@@
     )) %>% 
-  ggplot(aes(pthn, n, fill = group)) + 
-  facet_wrap(~group, ncol = 1) + 
+  ggplot() + 
+  aes(
+    pthn, 
+    n,
+    fill = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) + 
   geom_rect(
@@
       fill = group
-    ), inherit.aes = FALSE, 
+    ),
+    inherit.aes = FALSE, 
     alpha = 0.15
@@
   geom_vline(
-    data = medians, aes(xintercept = pthn_median, group = group), 
-    colour = "black", size = 0.5, linetype = "dashed"
+    data = medians, 
+    aes(
+      xintercept = pthn_median, 
+      group = group
+    ), 
+    colour = "black", 
+    size = 0.5,
+    linetype = "dashed"
   ) + 
@@
   stimuli %>% 
-  mutate(group = case_when(
+  mutate(
+    group = case_when(
       group=="ENG-CAT" ~ "cat-ENG",
@@
     )) %>% 
-  ggplot(aes(lv, fill = group)) +
-  facet_wrap(~group, ncol = 1) + 
+  ggplot() +
+  aes(
+    lv, 
+    fill = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) + 
   geom_rect(
@@
     aes(
-      xmin = 0, xmax = lv_median+lv_sd, 
-      ymin = -Inf, ymax = Inf, fill = group
-    ), inherit.aes = FALSE, alpha = 0.15
-  ) +   
-  geom_histogram(bins = 20, colour = "white") +
+      xmin = 0,
+      xmax = lv_median+lv_sd, 
+      ymin = -Inf,
+      ymax = Inf, 
+      fill = group
+    ), 
+    inherit.aes = FALSE, 
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20,
+    colour = "white"
+  ) +
   geom_vline(
-    data = medians, aes(xintercept = lv_median, group = group), 
-    colour = "black", size = 0.5, linetype = "dashed"
+    data = medians, 
+    aes(
+      xintercept = lv_median,
+      group = group
+    ), 
+    colour = "black", 
+    size = 0.5, 
+    linetype = "dashed"
   ) + 
   scale_x_continuous(labels = percent) +
-  labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity\n(stimuli and stimuli translations)\n") +
-
-  plot_layout(guides = "collect") &
-  plot_annotation(tag_levels = "A") &
-  scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
-  theme(legend.position = "none")
+  labs(
+    x = "Levenshtein", 
+    y = "Number of trials",
+    title = "Phonological similarity\n(stimuli and stimuli translations)\n"
+  ) +
+
+  plot_layout(
+    guides = "collect"
+  ) &
+  plot_annotation(
+    tag_levels = "A"
+  ) &
+  scale_fill_manual(
+    values = c(
+      "#1A85FF",
+      "#ff2976",
+      "#FFC20A"
+    ),
+    na.translate = FALSE
+  ) &
+  theme(
+    legend.position = "none"
+  )
 
@@
 participants_coded <- participants %>% 
-  mutate(valid_participant = ifelse(valid_participant, "Valid", "Not valid"))
+  mutate(
+    valid_participant = ifelse(
+      valid_participant, 
+      "Valid", 
+      "Not valid"
+    )
+  )
 ```
@@
 participants_coded %>% 
-  count(group, valid_participant) %>% 
-  mutate(valid_participant = factor(valid_participant)) %>% 
-  ggplot(aes(group, n, fill = valid_participant, order = n)) +
+  count(
+    group,
+    valid_participant
+  ) %>% 
+  mutate(
+    valid_participant = factor(valid_participant)
+  ) %>% 
+  ggplot() +
+  aes(
+    group, 
+    n, 
+    fill = valid_participant, 
+    order = n
+  ) +
   geom_col() +
-  labs(x = "Group", y = "# participants", fill = "Valid?") +
+  labs(
+    x = "Group",
+    y = "# participants", 
+    fill = "Valid?"
+  ) +
   theme(
@@
     axis.title.x = element_blank(),
-    panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
+    panel.grid.major.y = element_line(
+      linetype = "dotted", 
+      colour = "grey"
+    )
   )
@@
 participants_coded %>% 
-  filter(valid_participant=="Valid") %>% 
-  select(participant, group, n, n_valid) %>%
-  mutate(valid_prop = n_valid/n) %>% 
-  ggplot(aes(valid_prop, fill = group, colour = group)) +
-  facet_wrap(~group, ncol = 1) +
+  filter(
+    valid_participant=="Valid"
+  ) %>% 
+  select(
+    participant, 
+    group, 
+    n, 
+    n_valid
+  ) %>%
+  mutate(
+    valid_prop = n_valid/n
+  ) %>% 
+  ggplot() +
+  aes(
+    valid_prop,
+    fill = group,
+    colour = group
+  ) +
+  facet_wrap(
+    ~group, 
+    ncol = 1
+  ) +
   geom_histogram() +
   geom_rug() +
-  labs(x = "Valid trials (%)", y = "N participants", fill = "Group", colour = "Group") +
-  scale_x_continuous(limits = c(0, 1), labels = percent) +
+  labs(
+    x = "Valid trials (%)", 
+    y = "N participants",
+    fill = "Group", 
+    colour = "Group"
+  ) +
+  scale_x_continuous(
+    limits = c(0, 1), 
+    labels = percent
+  ) +
   theme(
@@
 ```{r participants_demo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-participants_valid <- filter(participants_coded, valid_participant=="Valid")
+participants_valid <- participants_coded %>% 
+  filter(valid_participant=="Valid")
 
 p_sex <- participants_valid %>% 
-  left_join(count(participants_valid, group, name = "n_total")) %>% 
-  mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
-  ggplot(aes(group, fill = sex)) +
+  left_join(
+    count(
+      participants_valid, 
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  "\n(N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    group, 
+    fill = sex
+  ) +
   geom_bar(position = position_fill()) +
-  labs(x = "Group", y = "# participants", fill = "Sex") +
+  labs(
+    x = "Group",
+    y = "# participants",
+    fill = "Sex"
+  ) +
   scale_y_continuous(labels = percent) +
@@
 p_age <- participants_valid %>% 
-  left_join(count(participants_valid, group, name = "n_total")) %>% 
-  mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
-  ggplot(aes(as.factor(age), fill = group, colour = group)) +
-  facet_wrap(~group, ncol = 1) +
+  left_join(
+    count(
+      participants_valid, 
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  " (N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    as.factor(age), 
+    fill = group, 
+    colour = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) +
   geom_bar() +
-  labs(x = "Age (years)", y = "% participants", fill = "Group", colour = "Group") +
+  labs(
+    x = "Age (years)",
+    y = "% participants", 
+    fill = "Group", 
+    colour = "Group"
+  ) +
   theme(
@@
     legend.title = element_blank(),
-    panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
+    panel.grid.major.y = element_line(
+      linetype = "dotted", 
+      colour = "grey"
+    )
   )
@@
 participants_valid %>% 
-  count(group, l2, .drop = FALSE) %>% 
+  count(
+    group,
+    l2, 
+    .drop = FALSE
+  ) %>% 
   mutate(
@@
   ) %>% 
-  left_join(count(participants_valid, group, name = "n_total")) %>% 
-  mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
-  ggplot(aes(group, n, fill = l2)) +
-  facet_wrap(~group, ncol = 1, scales = "free_y") +
+  left_join(
+    count(
+      participants_valid,
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  "\n(N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    group,
+    n, 
+    fill = l2
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1, 
+    scales = "free_y"
+  ) +
   geom_col(position = position_fill()) +
-  labs(x = "L2", y = "# participants", fill = "L2") +
-  scale_y_continuous(labels = percent) +
+  labs(
+    x = "L2",
+    y = "# participants",
+    fill = "L2"
+  ) +
+  scale_y_continuous(
+    labels = percent
+  ) +
   coord_flip() +
@@
 participants_valid %>% 
-  select(participant, group, catalan_oral, spanish_oral, catalan_written, spanish_written) %>% 
-  mutate_at(vars(catalan_oral, spanish_oral, catalan_written, spanish_written), replace_na, 0) %>% 
-  pivot_longer(c(catalan_oral, spanish_oral, catalan_written, spanish_written),
-               names_to = "type", values_to = "score") %>% 
-  separate(type, c("language", "modality"), sep = "_") %>% 
-  mutate_at(vars(language, modality), str_to_sentence) %>%
-  left_join(count(participants_valid, group, name = "n_total")) %>% 
-  mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
-  ggplot(aes(as.integer(score), fill = language)) +
+  select(
+    participant,
+    group, 
+    catalan_oral,
+    spanish_oral, 
+    catalan_written,
+    spanish_written
+  ) %>% 
+  mutate_at(
+    vars(
+      catalan_oral,
+      spanish_oral,
+      catalan_written, 
+      spanish_written
+    ),
+    replace_na, 
+    0
+  ) %>% 
+  pivot_longer(
+    c(
+      catalan_oral, 
+      spanish_oral,
+      catalan_written,
+      spanish_written
+    ),
+    names_to = "type", 
+    values_to = "score"
+  ) %>% 
+  separate(
+    type, 
+    c("language", "modality"),
+    sep = "_"
+  ) %>% 
+  mutate_at(
+    vars(language, modality), 
+    str_to_sentence
+  ) %>%
+  left_join(
+    count(
+      participants_valid,
+      group,
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  " (N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    as.integer(score), 
+    fill = language
+  ) +
   facet_grid(modality~group) +
-  geom_bar(position = position_dodge()) +
-  labs(y = "# participants", x = "Self-reported proficiency (0-5)", fill = "Language") +
+  geom_bar(
+    position = position_dodge()
+  ) +
+  labs(
+    y = "# participants",
+    x = "Self-reported proficiency (0-5)",
+    fill = "Language"
+  ) +
   theme(
     legend.position = "top",
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+    panel.grid.major.y = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    )
   )
@@
   # typos are considered correct responses
-  mutate_at(vars(group), as.factor) %>% 
+  mutate_at(
+    vars(group), 
+    as.factor
+  ) %>% 
   # center predictors
   mutate_at(
-    vars(lv_std, pthn_std, frequency_zipf_std),
-    function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
+    vars(
+      lv_std, 
+      pthn_std,
+      frequency_zipf_std
+    ),
+    function(x) {
+      scale(
+        x, 
+        center = TRUE, 
+        scale = TRUE
+      )[,1]
+    }
+  ) %>% 
   # impute missing data
-  mice(m = 5, print = FALSE, method = "pmm") %>% 
+  mice(
+    m = 5, 
+    print = FALSE,
+    method = "pmm"
+  ) %>% 
   complete() %>% 
   as_tibble() %>% 
-  arrange(group, word)
+  arrange(
+    group,
+    word
+  )
 ```
@@
   arrange(desc(elpd_loo)) %>% 
-  select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
-  mutate(model = str_replace(model, "fit_", "Model ")) %>% 
+  select(
+    model,
+    matches("elpd_loo"),
+    matches("looic"),
+    matches("diff")
+  ) %>%
+  mutate(
+    model = str_replace(
+      model,
+      "fit_", 
+      "Model "
+    )
+  ) %>% 
   mutate_all(~ifelse(. == 0, NA, .)) %>% 
@@
   tab_style(
-    style = cell_text(align = "left", weight = "bold", size = "medium"),
+    style = cell_text(
+      align = "left",
+      weight = "bold",
+      size = "medium"
+    ),
     locations = cells_title(groups = "title")
@@
   tab_style(
-    style = cell_text(align = "left", size = "medium", style = "italic"),
+    style = cell_text(
+      align = "left",
+      size = "medium", 
+      style = "italic"
+    ),
     locations = cells_title(groups = "subtitle")
@@
   tab_style(
-    style = list(cell_borders(sides = c("left", "right"), color = "white"),
+    style = list(
+      cell_borders(
+        sides = c("left", "right"),
+        color = "white"
+      ),
       cell_text(align = "center")),
     locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
+      cells_body(), 
+      cells_row_groups(),
+      cells_column_spanners(), 
+      cells_stub(), 
+      cells_source_notes(),
+      cells_column_labels(),
+      cells_stubhead()
     )
@@
   tab_style(
-    style = list(cell_text(align = "left", size = "medium"), 
-                 cell_borders(sides = "all", color = "white")),
-    locations = list(cells_stub(), cells_source_notes())
+    style = list(
+      cell_text(
+        align = "left", 
+        size = "medium"
+      ), 
+      cell_borders(
+        sides = "all", 
+        color = "white"
+      )
+    ),
+    locations = list(
+      cells_stub(), 
+      cells_source_notes()
+    )
   ) %>% 
   tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
+    style = cell_text(
+      align = "center",
+      size = "medium", 
+      weight = "normal", 
+      style = "normal"
+    ),
+    locations = list(
+      cells_column_spanners(),
+      cells_column_labels()
+    )
   ) %>% 
   tab_style(
-    style = cell_borders(sides = "all", color = "white"),
-    locations = cells_body(columns = 1:7)
+    style = cell_borders(
+      sides = "all", 
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
   ) %>% 
   tab_style(
-    style = cell_borders(sides = "top", color = "white"),
-    locations = cells_body(columns = 1:7)
+    style = cell_borders(
+      sides = "top", 
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
   ) %>% 
@@
   clean_names() %>% 
-  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4))
+  mutate_at(
+    vars(estimate:q97_5), 
+    ~ifelse(
+      variable=="Intercept",
+      inv_logit_scaled(.),
+      ./4
+    )
+  )
 
@@
   fmt_percent(2:4) %>% 
-  cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
+  cols_merge(
+    c("q2_5", "q97_5"), 
+    pattern = "[{1}, {2}]"
+  ) %>% 
   cols_label(
@@
   ) %>% 
-  tab_source_note("Estimated posterior distributions of coefficients in Model 6.
+  tab_source_note(
+    "Estimated posterior distributions of coefficients in Model 6.
 A\\) Population-level effects.
@@
 Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
-Black dots represent the mean of each distribution.") %>%
+Black dots represent the mean of each distribution."
+  ) %>% 
   cols_align(align = "left") %>% 
@@
   clean_names() %>% 
-  mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
+  mutate_at(
+    vars(estimate:q97_5), 
+    ~ifelse(
+      variable=="Intercept", 
+      inv_logit_scaled(.), 
+      ./4
+    )
+  ) %>% 
   group_split(variable) %>% 
@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:group1" = "Levenshtein \u00d7 Group 1",
-  "lv_std:group2" = "Levenshtein \u00d7 Group 2",
+  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
   "pthn_std" = "PTHN\n[+1 SD]",
-  "group1" =  "Group 1\n[*-ENG vs. cat-SPA]",
-  "group2" = "Group 2\n[cat-ENG vs. spa-ENG]"
+  "groupENGvs.SPA" =  "Group 1\n[*-ENG vs. cat-SPA]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
       factor(
-        levels =c(
+        levels = c(
           "Intercept", 
@@
           "PTHN \u00d7 Levenshtein",
-          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 1\n[*-ENG vs. cat-SPA]",
           "Group 2\n[cat-ENG vs. spa-ENG]",
@@
       ),
-    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+    .value = ifelse(
+      str_detect(.variable, "Intercept"), 
+      inv_logit_scaled(.value),
+      .value/4
+    )
   ) %>% 
   arrange(.variable) %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
+  ggplot() +
+  aes(
+    .value, 
+    fct_rev(.variable_name)
+  ) +
   geom_vline(xintercept = 0) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
-  scale_x_continuous(labels = function(x) percent(round(x, 1))) +
-  labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
+  stat_slab(
+    aes(fill = stat(
+      cut_cdf_qi(
+        cdf, 
+        .width = c(.5, .8, .95),
+        labels = percent_format()
+      )
+    )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0,
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#1A85FF", "#9ccaff", "#d2e5fc"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = function(x) percent(round(x, 1))
+  ) +
+  labs(
+    x = "P(Correct)",
+    y = "Posterior probability density", 
+    fill = "CrI"
+  ) +
   theme(
@@
 post_sd_data <- posterior_draws_fixed %>% 
-  filter(str_detect(.variable, "sd_")) %>% 
-  mutate(
-    .variable_name = str_remove(.variable, "sd_participant__|sd_word__") %>% 
+  filter(
+    str_detect(.variable, "sd_")
+  ) %>% 
+  mutate(
+    .variable_name = str_remove(
+      .variable, 
+      "sd_participant__|sd_word__"
+    ) %>% 
       str_replace_all(str_repl) %>% 
@@
     ),
-    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
-  )
-
-post_sd <- ggplot(filter(post_sd_data, group=="Participants"), aes(.value, fct_rev(.variable_name))) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, limits = c(0, 1)) +
-  labs(x = "SDs in P(Correct)", title = "By participant", fill = "CrI") +
+    .value = ifelse(
+      str_detect(.variable, "Intercept"),
+      inv_logit_scaled(.value),
+      .value/4
+    )
+  )
+
+post_sd <- filter(post_sd_data, group=="Participants") %>% 
+  ggplot() +
+  aes(
+    .value,
+    fct_rev(.variable_name)
+  ) +
+  stat_slab(
+    aes(fill = stat(
+      cut_cdf_qi(
+        cdf, 
+        .width = c(.5, .8, .95),
+        labels = percent_format()
+      )
+    )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0,
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#ff2976", "#f2a7c2", "#fcd9e6"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = percent,
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By participant", 
+    fill = "CrI"
+  ) +
   theme(axis.text.x = element_blank()) +
 
-  ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_x_continuous(labels = percent, limits = c(0, 1)) +
-  labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +
-
-  plot_layout(ncol = 1, heights = c(0.85, 0.15), guides = "collect") &
+  post_sd_data %>% 
+  filter(group=="Words") %>% 
+  ggplot() +
+  aes(
+    .value, 
+    fct_rev(.variable_name)
+  ) +
+  stat_slab(
+    aes(fill = stat(
+      cut_cdf_qi(
+        cdf, 
+        .width = c(.5, .8, .95),
+        labels = percent_format()
+      )
+    )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By item", 
+    fill = "CrI"
+  ) +
+
+  plot_layout(
+    ncol = 1, 
+    heights = c(0.85, 0.15), 
+    guides = "collect"
+  ) &
   theme(
@@
     axis.ticks.y = element_blank(),
-    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
+    panel.grid.major.y = element_line(
+      colour = "grey",
+      size = 0.5
+      )
   )
@@
   as.data.frame() %>% 
-  pivot_longer(-term1, names_to = "term2", values_to = ".value") %>% 
+  pivot_longer(
+    -term1, 
+    names_to = "term2", 
+    values_to = ".value"
+    ) %>% 
   drop_na(.value)
@@
 post_cors <- corr_data %>% 
-  mutate(term1 = factor(term1, levels = names(str_repl), ordered = TRUE),
-         term2 = factor(term2, levels = names(str_repl), ordered = TRUE)) %>% 
-  mutate_at(vars(term1, term2), str_replace_all, str_repl) %>% 
-  mutate(
-    .value_label = case_when(term1==term2 ~ NA_character_, TRUE ~ printnum(.value, gt1 = FALSE)),
-    .value = case_when(term1==term2 ~ NA_real_, TRUE ~ .value)
-  ) %>% 
-  filter(term1 != "PTHN × Levenshtein", term2 != "Intercept") %>% 
-  ggplot(aes(fct_inorder(term1), fct_rev(fct_inorder(term2)), fill = .value)) +
+  mutate(
+    term1 = factor(
+      term1,
+      levels = names(str_repl), 
+      ordered = TRUE
+    ),
+    term2 = factor(
+      term2,
+      levels = names(str_repl),
+      ordered = TRUE
+    )
+  ) %>% 
+  mutate_at(
+    vars(term1, term2), 
+    str_replace_all, 
+    str_repl
+  ) %>% 
+  mutate(
+    .value_label = case_when(
+      term1==term2 ~ NA_character_,
+      TRUE ~ printnum(.value, gt1 = FALSE)
+    ),
+    .value = case_when(
+      term1==term2 ~ NA_real_,
+      TRUE ~ .value
+    )
+  ) %>% 
+  filter(
+    term1 != "PTHN × Levenshtein", 
+    term2 != "Intercept"
+  ) %>% 
+  ggplot() +
+  aes(
+    fct_inorder(term1), 
+    fct_rev(fct_inorder(term2)),
+    fill = .value
+  )
 geom_tile(na.rm = TRUE) +
-  geom_text(aes(label = .value_label), size = 3, na.rm = TRUE) +
-  labs(x = "Term 1", y = "Term 2", fill = "Posterior\ncorrelation") +
+  geom_text(
+    aes(label = .value_label), 
+    size = 3, 
+    na.rm = TRUE
+  ) +
+  labs(
+    x = "Term 1", 
+    y = "Term 2", 
+    fill = "Posterior\ncorrelation"
+  ) +
   scale_fill_gradient2(
-    low = "#04bf78", mid = "white", high = "#FFC20A", 
-    na.value = "white", limits = c(-0.4, 0.4)
+    low = "#04bf78", 
+    mid = "white", 
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
   ) +
@@
 
-post_fix / ((post_sd | post_cors) + plot_layout(widths = c(0.6, 0.4))) +
+post_fix / 
+  (
+    (post_sd | post_cors) +
+      plot_layout(widths = c(0.6, 0.4))
+  ) +
   plot_layout() +
@@
   ungroup() %>% 
-  left_join(select(participants, participant, group)) %>% 
-  arrange(group, participant, .param) %>% 
-  mutate(.param_name = str_replace_all(.param, str_repl) %>%  
+  left_join(
+    select(
+      participants,
+      participant,
+      group
+    )
+  ) %>% 
+  arrange(
+    group, 
+    participant, 
+    .param
+  ) %>% 
+  mutate(
+    .param_name = str_replace_all(
+      .param, 
+      str_repl
+    ) %>%  
       factor(
@@
 
-ggplot(post_re, 
-       aes(.value, sort(as.factor(participant)), 
-           xmin = .lower, xmax = .upper, colour = group)) +
-  facet_grid(group~.param_name, scales = "free") +
-  geom_vline(xintercept = 0, colour = "grey") +
-  geom_errorbarh(height = 0, alpha = 0.75) +
-  geom_point(size = 0.1, shape = 4) +
+ggplot(
+  post_re, 
+  aes(
+    .value, 
+    sort(as.factor(participant)), 
+    xmin = .lower,
+    xmax = .upper,
+    colour = group
+  )
+) +
+  facet_grid(
+    group~.param_name,
+    scales = "free"
+  ) +
+  geom_vline(
+    xintercept = 0, 
+    colour = "grey"
+  ) +
+  geom_errorbarh(
+    height = 0,
+    alpha = 0.75
+  ) +
+  geom_point(
+    size = 0.1,
+    shape = 4
+  ) +
   labs(
@@
     legend.title = element_blank(),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    panel.grid.major.x = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    ),
     axis.text.x = element_text(size = 7),
@@
   ungroup() %>% 
-  arrange(word, .param) %>% 
-  mutate(
-    .param_name = str_replace_all(.param, str_repl) %>%  
+  arrange(
+    word, 
+    .param
+  ) %>% 
+  mutate(
+    .param_name = str_replace_all(
+      .param, 
+      str_repl
+    ) %>%  
       factor(
@@
     .value,
-    reorder(as.factor(word), .value)
+    reorder(
+      as.factor(word), 
+      .value
+    )
   )
@@
     legend.justification = "left",
-    legend.key = element_rect(fill = NA, colour = NA),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    legend.key = element_rect(
+      fill = NA, 
+      colour = NA
+    ),
+    panel.grid.major.x = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    ),
     axis.text.x = element_text(size = 7),
@@
     plot.caption.position = "plot",
-    plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm")))
-  )
-
-ggsave(here("img", "accuracy.png"), height = 9, width = 5, dpi = 500)
+    plot.caption = element_text(
+      size = 9, 
+      hjust = 0, 
+      face = "plain",
+      margin = margin(t = unit(3, "cm"))
+    )
+  )
+
+ggsave(
+  here("img", "accuracy.png"), 
+  height = 9,
+  width = 5, 
+  dpi = 500
+)
 ```
@@
   )
+
+posterior_accuracy <-
 
@@
     legend.key = element_rect(fill = NA, colour = NA),
-    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    panel.grid.major.x = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    ),
     axis.text.x = element_text(size = 7),
@@
     plot.caption.position = "plot",
-    plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm")))
-  )
-
-ggsave(here("img", "accuracy_empirical.png"), height = 9, width = 15, dpi = 500)
+    plot.caption = element_text(
+      size = 9,
+      hjust = 0, 
+      face = "plain", 
+      margin = margin(t = unit(3, "cm")))
+  )
+
+ggsave(
+  here("img", "accuracy_empirical.png"), 
+  height = 9,
+  width = 15,
+  dpi = 500
+)
 ```
@@
   labs(
-		x = paste0("PTHN (+1 SD, ", round(sd(responses$pthn), 1), " higher-frequency neighbours)"),
+    x = paste0(
+      "PTHN (+1 SD, ", 
+      round(sd(responses$pthn), 1),
+      " higher-frequency neighbours)"
+    ),
     y = "P(Correct)", 

### 2022-04-27 docs/index.Rmd
Make changes to manuscript

--- 2022-03-17 docs/index.Rmd
@@
   ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
-  ggplot2.continuous.fill = scale_fill_gradient2(
-    low = "#04bf78", mid = "white", high = "#FFC20A", 
-    na.value = "white", limits = c(-0.4, 0.4)
-  ),
-  ggplot2.continuous.colour = scale_fill_gradient2(
-    low = "#04bf78", mid = "white", high = "#FFC20A", 
-    na.value = "white", limits = c(-0.4, 0.4)
-  )
-)
+  ggplot2.continuous.fill = ggplot2::scale_fill_gradient2(
+    low = "#04bf78", 
+    mid = "white", 
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
+  ),
+  ggplot2.continuous.colour = ggplot2::scale_fill_gradient2(
+    low = "#04bf78", 
+    mid = "white",
+    high = "#FFC20A", 
+    na.value = "white",
+    limits = c(-0.4, 0.4)
+  )
+)
+
 ```
@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:group1" = "Levenshtein \u00d7 Group 1",
-  "lv_std:group2" = "Levenshtein \u00d7 Group 2",
+  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
   "pthn_std" = "PTHN\n[+1 SD]",
-  "group1" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-  "group2" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "groupENGvs.SPA" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[cat-ENG vs. spa-ENG]"
+)
+
+post_draws <-  gather_draws(
+  fit_6, 
+  `b_.*`, 
+  `sd_.*`, 
+  regex = TRUE
 )
@@
 # fixed effects
-post_fix <- posterior_draws_fixed %>% 
+post_fix <- post_draws %>% 
   filter(str_detect(.variable, "b_")) %>% 
@@
         levels = c(
-          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-          "Group 1\n[ENG-* vs. SPA-CAT]",
-          "Group 2\n[ENG-CAT vs. ENG-SPA]",
-          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
-        )
-      ),
-    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]",
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1", 
+          "Levenshtein \u00d7 Group 2"
+        )
+      ),
+    .value = ifelse(
+      str_detect(.variable, "Intercept"),
+      inv_logit_scaled(.value), 
+      .value/4
+    )
   ) %>% 
   arrange(.variable) %>% 
-  ggplot(aes(.value, fct_rev(.variable_name))) +
+  ggplot() +
+  aes(
+    .value,
+    fct_rev(.variable_name)
+  ) +
   geom_vline(xintercept = 0) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
-  scale_x_continuous(labels = function(x) percent(round(x, 1))) +
-  labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95),
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#1A85FF", "#9ccaff", "#d2e5fc"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = function(x) percent(round(x, 1))
+  ) +
+  labs(
+    x = "P(Correct)", 
+    y = "Posterior probability density",
+    fill = "CrI"
+  ) +
   theme(
@@
     legend.direction = "vertical",
-    legend.background = element_rect(fill = "grey95"),
-    legend.key = element_rect(fill = "grey95", colour = "grey95"), 
+    legend.background = element_rect(
+      fill = "grey95"
+    ),
+    legend.key = element_rect(
+      fill = "grey95", 
+      colour = "grey95"
+    ), 
     legend.key.height = unit(0.1, "cm"),
     axis.title.y = element_blank(),
-    legend.text = element_text(size = 7),
-    axis.text.x = element_text(colour = "black"),
-    axis.text.y = element_text(colour = "black", hjust = 1),
+    legend.text = element_text(
+      size = 7
+    ),
+    axis.text.x = element_text(
+      colour = "black"
+    ),
+    axis.text.y = element_text(
+      colour = "black",
+      hjust = 1
+    ),
     axis.ticks.y = element_blank(),
-    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      size = 0.5
+    )
   )
@@
 # SD of random effects
-post_sd_data <- posterior_draws_fixed %>% 
-  filter(str_detect(.variable, "sd_")) %>% 
+post_sd_data <- post_draws 
+filter(str_detect(.variable, "sd_")) 
 mutate(
-    .variable_name = str_remove(.variable, "sd_participant__|sd_word__") %>% 
+  group = case_when(
+    str_detect(.variable, "participant") ~ "Participants",
+    str_detect(.variable, "word") ~ "Words"
+  ),    
+  .variable_name = str_remove(
+    .variable,
+    "sd_participant__|sd_word__"
+  ) )
 str_replace_all(str_repl) %>% 
@@
     levels = c(
-          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-          "Group 1\n[ENG-* vs. SPA-CAT]",
-          "Group 2\n[ENG-CAT vs. ENG-SPA]",
-          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
-        )
-      ),
-    group = case_when(
-      str_detect(.variable, "participant") ~ "Participants",
-      str_detect(.variable, "word") ~ "Words"
-    ),
-    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
-  )
-
-post_sd <- ggplot(filter(post_sd_data, group=="Participants"), aes(.value, fct_rev(.variable_name))) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_fill_manual(values = c("#ff2976", "#f2a7c2", "#fcd9e6"), na.translate = FALSE) +
-  scale_x_continuous(labels = percent, limits = c(0, 1)) +
-  labs(x = "SDs in P(Correct)", title = "By participant", fill = "CrI") +
-  theme(axis.text.x = element_blank()) +
-
-  ggplot(filter(post_sd_data, group=="Words"), aes(.value, fct_rev(.variable_name))) +
-  stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-  stat_pointinterval(.width = 0, point_size = 1) +
-  scale_x_continuous(labels = percent, limits = c(0, 1)) +
-  labs(x = "SDs in P(Correct)", title = "By item", fill = "CrI") +
-
-  plot_layout(ncol = 1, heights = c(0.85, 0.15), guides = "collect") &
+      "Intercept", 
+      "Frequency\n[+1 SD]",
+      "PTHN\n[+1 SD]", 
+      "Levenshtein\n[+1 SD]", 
+      "PTHN \u00d7 Levenshtein",
+      "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+      "Group 2\n[cat-ENG vs. spa-ENG]",
+      "Levenshtein \u00d7 Group 1",
+      "Levenshtein \u00d7 Group 2"
+    )
+  ),
+.value = ifelse(
+  str_detect(.variable, "Intercept"), 
+  inv_logit_scaled(.value), 
+  .value/4
+)
+)
+
+post_sd <- post_sd_data %>% 
+  ggplot() +
+  aes(.value, fct_rev(.variable_name)) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95),
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#ff2976", "#f2a7c2", "#fcd9e6"), 
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By participant", 
+    fill = "CrI"
+  ) +
+  theme(
+    axis.text.x = element_blank()
+  ) +
+
+  post_sd_data %>% 
+  filter(group=="Words") %>% 
+  ggplot() +
+  aes(
+    .value, 
+    fct_rev(.variable_name)
+  ) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95), 
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By item",
+    fill = "CrI"
+  ) +
+  scale_fill_manual(
+    values = c("#ff2976", "#f2a7c2", "#fcd9e6"), 
+    na.translate = FALSE
+  ) +
+  plot_layout(
+    ncol = 1,
+    heights = c(0.85, 0.15), 
+    guides = "collect"
+  ) &
   theme(
@@
 ```{r post_re_participant, echo=FALSE, fig.height=6, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}
-post_re <- posterior_draws_random$participants %>% 
+
+nd_re_participants <- expand.grid(
+  participant = unique(fit$data$participant),
+  lv_std = c(-1, 0, 1),
+  frequency_zipf_std = 0,
+  pthn_std = seq(
+    min(fit$data$pthn_std, na.rm = TRUE),
+    max(fit$data$pthn_std, na.rm = TRUE),
+    by = 0.1
+  ),
+  group = NA
+)
+
+post_re <- add_epred_draws(
+  nd_re_participants, 
+  fit, 
+  re_formula = ~ (1 + frequency_zipf_std + pthn_std*lv_std | participant),
+  ndraws = 50
+)  %>% 
   mean_qi() %>% 
@@
 
-post_re <- posterior_draws_random$words   %>% 
-  ungroup() %>% 
-  arrange(word, .param) %>% 
-  mutate(
-    .param_name = str_replace_all(.param, str_repl) %>%  
-      factor(
-        levels = c(
-          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-          "Group 1\n[ENG-* vs. SPA-CAT]",
-          "Group 2\n[ENG-CAT vs. ENG-SPA]",
-          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
-        )
-      ),
-    .value = inv_logit_scaled(.value)
-  )
-
-ggplot(
-  post_re,
+nd_re_words <- expand.grid(
+  word = unique(fit_6$data$word),
+  pthn_std = c(-1, 1),
+  frequency_zipf_std = 0,
+  lv_std = seq(
+    min(fit_6$data$lv_std, na.rm = TRUE), 
+    max(fit_6$data$lv_std, na.rm = TRUE
+    ), by = 0.1
+  ),
+  group = NA
+)
+
+post_re <- add_epred_draws(
+  nd_re_words, 
+  fit_6, 
+  re_formula = ~ (1| word),
+  ndraws = 50
+) %>% 
+  ungroup()
+
+post_re %>% 
+  ggplot() +
   aes(
-    .value,
-    reorder(as.factor(word), .value)
-  )
+    .epred,
+    reorder(as.factor(word), .epred)
   ) +
@@
   ) +
-
   stat_interval(
@@
 
-responses %>% 
+word_accuracy <- responses %>% 
   mutate(
@@
     prop_se = prop_adj_se(successes, n)
-  ) %>% 
+  )
+
+word_accuracy %>% 
   ggplot() +
@@
 ```
+
+# Additional analyses
+
+## Previous familiarity
+
+```{r familiarity-frequency-distribution, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=10, fig.width=10}
+
+familiarity_problematic <- here("data", "familiarity", "familiarity.xlsx") %>% read_xlsx() %>% 
+  rename(
+    word = presented,
+    word_2 = translation,
+    test_language = language,
+    problem = note
+  ) %>% 
+  mutate(
+    problem = ifelse(
+      str_detect(problem, "check coding"), 
+      "Check coding",
+      str_to_sentence(problem)
+    ),
+    group = ifelse(test_language=="Catalan", "ENG-CAT", "ENG-SPA")
+  ) %>% 
+  select(word, group, problem)
+
+familiarity <- stimuli %>% 
+  filter(group != "SPA-CAT") %>% 
+  mutate(
+    stim = paste0(
+      word1, " (", ipa1, ")", " / ",
+      word2, " (", ipa2, ")", " / ",
+      percent(lv, accuracy = 1)
+    ),
+    test_language = case_when(
+      group=="ENG-CAT" ~ "Catalan",
+      group=="ENG-SPA" ~ "Spanish",
+      group=="SPA-CAT" ~ "Catalan"
+    ),
+    word = replace_non_ascii(word1)
+  ) %>% 
+  left_join(familiarity_problematic) %>% 
+  select(
+    group, 
+    word = word,
+    stim,
+    problem
+  ) %>% 
+  left_join(filter(subtlex, test_language=="English")) %>% 
+  replace_na(
+    list(
+      frequency_count = 1,
+      frequency_zipf = 0.555,
+      problem = "Normal"
+    )
+  ) %>% 
+  left_join(select(word_accuracy, group, stim, prop, prop_se)) %>% 
+  group_by(group) %>% 
+  mutate(
+    word_reorder = reorder_within(word, frequency_zipf, group)
+  ) %>% 
+  ungroup()
+
+familiarity %>% 
+  ggplot() +
+  aes(
+    x = frequency_zipf,
+    y = word_reorder,
+    fill = problem,
+    colour = problem,
+    label = word
+  ) +
+  facet_wrap(~group, scales = "free_y") +
+  geom_text(size = 3) +
+  labs(
+    x = "Frequency (Zipf) of presented word in English",
+    y = "Word",
+    colour = "Problem"
+  ) +
+  scale_y_reordered() +
+  theme(
+    legend.position = "top",
+    panel.border = element_rect(fill = NA, colour = "grey"),
+    panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    legend.title = element_blank(),
+    axis.text.y = element_blank(),
+    axis.ticks.y = element_blank(),
+    axis.title.y = element_blank()
+  )
+
+ggsave(here("img", "problematic_frequency-distribution.png"), height = 10)
+
+```
+
+```{r previous-familiarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.height=3.5, fig.width=8}
+
+familiarity %>% 
+  ggplot() +
+  aes(
+    x = frequency_zipf,
+    y = prop,
+    colour = problem,
+    fill = problem,
+    label = word
+  ) +
+  facet_wrap(~group) +
+  geom_text(
+    size = 3
+  ) +
+  # geom_point(
+  #   shape = 1,
+  #   size = 1,
+  #   alpha = 0.5,
+  #   stroke = 1
+  # ) +
+  geom_smooth(
+    method = "lm",
+    colour = "black",
+    fill = "grey"
+  ) +
+  labs(
+    x = "Frequency (Zipf) of presented word in English",
+    y = "Accuracy",
+    colour = "Group",
+    fill = "Group"
+  ) +
+  scale_x_continuous(
+    limits = c(2, 6)
+  ) +
+  scale_y_continuous(
+    limits = c(0, 1),
+    labels = percent
+  ) +
+  theme(
+    legend.position = "none"
+  )
+
+ggsave(here("img", "problematic_frequency-accuracy.png"), height = 4, width = 10)
+
+```

### 2022-04-27 manuscript/manuscript.Rmd
Make changes to manuscript

--- 2022-03-18 manuscript/manuscript.Rmd
@@
 Phonological similarity plays a more important role in low-proficiency bilinguals, compared to high-proficiency bilinguals, and points to the latter as relying more strongly on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
-Disentangling bilinguals' reliance on the conceptual and the lexical route during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals.
+Disentangling bilinguals' reliance on the conceptual and lexical routes during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals.
 In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
@@
 One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, a Spanish native listening to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
+For example, imagine a Spanish native, with no previous familiarity with any other language, listens for the first time to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
 The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
-These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical].
-Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
+These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /'pwer.ta/ (*puerta*).
+The recognition of this French, unfamiliar word relies almost entirely on its similarity with its Spanish translation.
+Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
+This time, phonological similarity is of little help: /dɔː/ and /'pwer.ta/ share very few phonemes.
+It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations (which they lack, they are unfamiliar to English).
+
+Thus, cross-language phonological similarity seems, at face value, an important source of information for unfamiliar listeners during auditory word recognition, even translations are not phonologically identical.
+Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
 In the present study we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
@@
 
-All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration or social contact).
+Many languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration or social contact).
 Cognates embody a particular case of such commonalities.
@@
 Such is the case of *cheese* and *queso* (Latin root), but are phonologically dissimilar.
-for simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
+For simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
 We do not expect that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
@@
 For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
-This cognate advantage is not restricted to production, instead also extending to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
-
-Early theoretical accounts of this phenomenon proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+This cognate advantage is not restricted to production, but also extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
+
+Early theoretical accounts of this the 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
 The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), instead of directly to their shared concept (conceptual route).
-This word-to-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection.
+This word-to-word connectivity should be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection.
 As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
-cognates subsequently exert less impact on lexical processing in proficient speakers.
+Cognates subsequently exert less impact on lexical processing in proficient speakers.
 The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
-The rationale behind this prediction was that backward translation relies more strongly on direct word-to-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
+The rationale behind this prediction was that backward translation relies more strongly on word-to-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
 Backward translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
@@
 Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
-Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
+Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on word-to-word links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
 However, although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
@@
 
-Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+[//]: # (KP: I find this para quite complicated to follow. Would it be possible to distill it to a simpler story that is directly relevant for the manuscript.)
+
+Since the RHM model was proposed, later studies on monolingual populations have focused on the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
 For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
@@
 
+post_draws <-  gather_draws(
+  fit_6, 
+  `b_.*`, 
+  `sd_.*`, 
+  regex = TRUE
+)
+
 # fixed effects
-post_fix <- posterior_draws_fixed %>% 
+post_fix <- post_draws %>% 
   filter(str_detect(.variable, "b_")) %>% 
@@
 # SD of random effects
-post_sd_data <- posterior_draws_fixed %>% 
+post_sd_data <- post_draws %>% 
   filter(str_detect(.variable, "sd_")) %>% 
@@
 
-post_sd <- post_sd_data %>% 
+post_sd <- post_draws %>% 
   filter(group=="Participants") %>% 

### 2022-04-27 R/00_stimuli.R

@@ -1,3 +1,94 @@
+# get SUBTLEX lexical frequencies ----
+get_subtlex <- function(){
+    
+    paths <- list(
+        English = here("data", "subtlex", "SUBTLEX-UK.txt"),
+        Spanish = here("data", "subtlex", "SUBTLEX-ESP.txt"),
+        Catalan = here("data", "subtlex", "SUBTLEX-CAT.txt")
+    )
+    
+    # import SUBTLEX-UK (English)
+    subtlex_eng <- paths$English %>% 
+        fread(
+            verbose = FALSE,
+            showProgress = TRUE
+        ) %>%
+        as_tibble() %>% 
+        clean_names() %>% 
+        rename(
+            word = spelling,
+            frequency_count = freq_count,
+            frequency_zipf = log_freq_zipf
+        ) %>% 
+        select(word, frequency_count, frequency_zipf)
+    
+    
+    # import SUBTLEX-ESP (Spanish)
+    subtlex_spa <- paths$Spanish %>%
+        fread(
+            verbose = FALSE,
+            showProgress = FALSE,
+            select = 1:4
+        ) %>% 
+        clean_names() %>%
+        mutate(frequency_zipf = log10(freq_per_million)+3) %>% 
+        rename(frequency_count = freq_count) %>% 
+        select(
+            word, 
+            frequency_count, 
+            frequency_zipf
+        ) %>% 
+        drop_na() 
+    
+    
+    # import SUBTLEX-CAT (Catalan)
+    subtlex_cat <- paths$Catalan %>%
+        fread(
+            verbose = FALSE,
+            showProgress = FALSE, dec = "."
+        ) %>% 
+        clean_names() %>% 
+        rename(
+            word = words, 
+            frequency_count = abs_wf, 
+            frequency_zipf = zipf
+        ) %>% 
+        mutate(frequency_count = as.numeric(str_remove_all(frequency_count, ","))) %>% 
+        select(word, frequency_count, frequency_zipf)
+    
+    # merge datasets
+    frequency <- list(
+        English = subtlex_eng,
+        Spanish = subtlex_spa,
+        Catalan = subtlex_cat
+    ) %>% 
+        bind_rows(.id = "test_language") %>% 
+        select(word, test_language, frequency_count, frequency_zipf) 
+    
+}
+
+
+# import subtlex
+import_subtlex <- function() {
[…]

### 2022-04-27 docs/index.Rmd

@@
         "sd_participant__|sd_word__"
-  ) )
+    ),
     str_replace_all(str_repl) %>% 

### 2022-04-27 manuscript/manuscript.Rmd

@@
+<<<<<<< HEAD
 ---
@@
 \newpage
+=======
+---
+title: "The role of cognateness in non-native spoken word recognition"
+
+shorttitle: Cognateness and non-native word recognition
+
+author:
+- name: Gonzalo Garcia-Castro
+  affiliation: '1'
+  corresponding: yes
+  address: Ramon Trias Fargas, 25-27, 08005 Barcelona
+  email: gonzalo.garciadecastro@upf.edu
+- name: Serene Siow
+  affiliation: '2'
+- name: Nuria Sebastian-Galles
+  affiliation: '1'
+- name: Kim Plunkett
+  affiliation: '2'
+affiliation:
+  - id            : "1"
+    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
+  - id            : "2"
+    institution   : "Department of Experimental Psychology, University of Oxford"
+
+authornote: |
+
+abstract: |
+  Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar.
+Phonological similarity plays a more important role in low-proficiency bilinguals, compared to high-proficiency bilinguals, and points to the latter as relying more strongly on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
+Disentangling bilinguals' reliance on the conceptual and the lexical route during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals.
+In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
+Given their lack of previous knowledge on the unfamiliar language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
+This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route.
+Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation.
+This effect was larger in Spanish participants translating Catalan words than in English participants translating Catalan or Spanish.
+These results point to cross-language word-to-word links as providing rich information during word translation.
+
+  <!-- https://tinyurl.com/ybremelq -->
+
+keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian
+wordcount: 4553
+
+bibliography:
+  - r-references.bib
+  - references.bib
+floatsintext: yes
+figurelist: no
+tablelist: no
+footnotelist: no
+linenumbers: yes
+mask: no
+draft: no
+numbersections: yes
+csl: "apa7.csl"
+documentclass: apa7
+classoption: man
+output:
+  papaja::apa6_docx: default
+  papaja::apa6_pdf: default
+---
+
+```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+
+set.seed(42)
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  results = "asis", 
+  dpi = 1000
+)
+
+options(
+  knitr.kable.NA = '-', 
+  knitr.duplicate.label = "allow",
+  gt.html_tag_check = FALSE
+)
+
+```
+
+```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+
+# import objects
+tar_load_globals()
+tar_load_all()
+
+# set custom ggplot2 theme
+theme_set(
+  theme(
+    title = element_text(size = 8, face = "bold"),
+
+    legend.background = element_rect(fill = NA, colour = NA),
+    legend.key = element_rect(fill = "white", colour = NA),
+    legend.justification = "right",
+    legend.direction = "horizontal",
+
+    axis.line = element_line(colour = "black", size = 0.5),
+    axis.ticks = element_line(colour = "black", size = 0.5),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+
+    plot.caption.position = "plot",
+    plot.title = element_text(hjust = 0.5),
+
+    panel.grid = element_blank(),
+    panel.background = element_rect(fill = "white", colour = NA)
+  )
+)
+
+```
+
+# Introduction
+
+Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
+One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
+For example, a Spanish native listening to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
+The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
+These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical].
+Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
+In the present study we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
+
+[//]: # (NSG: I think that a nice way to start the paper (or to finish it) is by citing one of those web pages that say that native English speakers already know tons of words in Spanish, French, German...
+because they are very similar (aka cognates).
+But this may work for writing, because for speech the situation is very different.
+This may be at the beginning or at the end of the paper.
+Given that most reviewers/ editors are likely to be native English-speakers, this may convince them that the paper is interesting and that they want to see it published.)
+
+All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration or social contact).
+Cognates embody a particular case of such commonalities.
+Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
+For example, *sun* and *sol* (in Spanish) share phonological onset but their etymology points to different origins.
+The complimentary case is equally problematic for the definition of cognateness in this context: cross-language synonyms that share etymological origin may not be necessarily be recognised as cognates by the average listener.
+Such is the case of *cheese* and *queso* (Latin root), but are phonologically dissimilar.
+for simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
+We do not expect that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
+For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+Cognates play a pivotal role in many models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
+For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
+This cognate advantage is not restricted to production, instead also extending to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
+
+Early theoretical accounts of this phenomenon proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), instead of directly to their shared concept (conceptual route).
+This word-to-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection.
+As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
+cognates subsequently exert less impact on lexical processing in proficient speakers.
+The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
+The rationale behind this prediction was that backward translation relies more strongly on direct word-to-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
+Backward translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
+
+To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
+Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
+Results showed that translation times and accuracy were roughly equivalent in forward and backward translation.
+Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected  by semantic variables than non-cognates.
+Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
+Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
+However, although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
+These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
+
+Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
+Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when these neighbours' lexical frequency is higher [@luce1998recognizing].
+This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+The Bilingual Interactive Activation (BIA/BIA+)  [@van1998orthographic; @dijkstra2002architecture] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with each other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
+Evidence supporting such an integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word's translation.
+When bilingual participants are presented with words in one language in a lexical decision task, translating them to the other language took longer when they were part of large  neighbourhoods [@van1998orthographic].
+This suggests that presented words activated orthographic neighbourhoods in the other language, which competed for selection with the target word during recognition.
+
+A more recent model of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
+This model assumes an integrated lexicon in which word representations from two languages establish connections in a way that is similar to connections between words of a single language.
+It also makes the assumption that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
+It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-to-word connections).
+In addition, this model suggests that the strength of the association between L2 representations and their concept depends on language users' proficiency in L2.
+A strong correlation between Multilink simulations of a translation elicitation task and behavioural data from Dutch-English bilinguals [@christoffels2006memory] has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
+Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+Overall, results from Multilink are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
+
+The fact that Multilink successfully fitted the experimental data from @christoffels2006memory in the absence of word-to-word connections between translation equivalents is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
+The model specification under which the quantitative predictions were generated assumes similar vocabulary sizes across languages, and also that the word-to-concept connections of L1 and L2 representations have same strength.
+The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
+It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-to-word connections during translation, especially during backward translation, and this prediction remains untested by Multilink.
+
+Most of the studies referenced so far rely heavily on production and comprehension of printed words: in these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
+There are limitations in generalising their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
+Consider the words *radio* (English) and *radio* (Spanish).
+Both words are clearly cognates, as their orthographical word form is identical.
+However, their phonological forms (English /ˈɹeɪ.di.əʊ/-- Spanish /ˈra.ðjo/) are not identical.
+Although many English listeners might recognise the Spanish word-form--even with no previous knowledge of Spanish--it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
+
+Measuring the phonological similarity between two word forms is also substantially more challenging than measuring their orthographic similarity.
+Many studies use symbolic representations to compute the simiarity between word-forms of interest in either modality.
+Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (*radio*-*radio*); phonological word-forms are represented by their  phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA notation) [see @floccia2018introduction for example].
+The similarity between each pair of word-forms is operationalised numerically by applying some algorithm that computes the edit distance between their symbolic representation.
+The Leveshtein distance is one of the most widely known algorithms: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string [@levenshtein1966binary], and has been successfully used to estimate the similarity of cross-language word-pairs [@gooskens2004perceptive; @schepens2012distributions].
+Using our previous examples, the Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio")`: both strings are identical.
+But the distance between the phonological transcriptions ɹeɪdiəʊ and raðjo is `r stringdist::stringdist("ɹeɪdiəʊ", "raðjo")`, which is the number of edits that separates both character strings.
+There is also an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.
+whether to include elements of aspiration, nasalisation, etc.) can affect how accurate these symbolic representation capture all of the features that are present in the acoustic signal when either word is spoken.
+Therefore, any measure of similarity that uses phonological transcriptions as input may overestimate or underestimate their actual similarity.
+Taking into account that measuring phonological similarity is considerably more complex than measuring orthographic complexity it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
+
+In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be misrepresenting the availability of such information in the auditory modality.
+In the present study we tested listener's reliance on phonological similarity when translating words from an unfamiliar language, in order to investigate the plausibility of word-to-word during backward translation.
+We tested English and Spanish native speakers in a translation elicitation task: they were presented with spoken words in an unfamiliar, non-native language, and were asked to translate them to their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as **presented words**, and the correct translation for the presented words as **target translations**.
+The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and therefore word-to-concept connections are, in principle, absent.
+As a consequence, these participants should only be able use word-to-word connections to succeed in the translation of words in an unfamiliar language.
+We predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
+
+We tested three groups of participants in the translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
+Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
+
+Following @luce1998recognizing's NAM model, we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+If this is the case,  has more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target translation.
+For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button* (because the lexical frequency of *bottom* is higher than that of *button*).
+If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
+
+# Methods
+
+All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
+
+## Participants
+
+Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`.
+We collected data from `r length(unique(participants$participant))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
+`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant[participants$group %in% c("SPA-CAT")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant[participants$group %in% c("SPA-CAT") & participants$sex=="Female"]))` female).
+Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
+Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
+Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
+We excluded data from participants that (a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), (b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`), or (c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
+
+## Procedure
+
+The experiment was implemented online using Psychopy/Pavlovia [@peirce2019psychopy2].
+Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla).
+After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
+Then, participants completed the experimental task.
+Participants were informed that they would listen to a series of pre-recorded words in either Catalan or Spanish (English participants) or Catalan (Spanish participants).
+They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
+English participants were randomly assigned to the Catalan or Spanish lists.
+Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
+Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
+After one second, the audio started playing while the dot remained being displayed until the audio ended.
+Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
+Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
+Participants were allowed to correct their answer.
+Then, participants pressed the RETURN/ENTER key to confirm their answer and start and new trial.
+We excluded trials where the participant's answer was not a valid word in the target language or no answer was given.
+Trials in which the response was mistyped by only one character were counted as correct (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct word.
+Trials in which participants took longer than 10 seconds to respond were also excluded.
+Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish).
+The task took approximately 15 minutes to be completed.
+
+```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%", fig.cap="Schematic representation of a trial in the experimental task."}
+
+include_graphics(here("img", "design.png"))
+
+```
+
+## Stimuli
+
+```{r stimulilengths, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+lengths <- stimuli %>% 
+  select(group, ipa1, ipa2, word1, word2) %>% 
+  mutate_at(vars(ipa1:word2), nchar) %>% 
+  group_by(group) %>% 
+  summarise_at(
+    vars(ipa1, ipa2, word1, word2), 
+    list(mean = mean, sd = sd, min = min, max = max)
+  ) %>%
+  split(.$group) %>% 
+  map(select, -group) %>% 
+  set_names(make_clean_names(names(.))) %>% 
+  map(unlist)
+```
+
+### Input stimuli (Catalan/Spanish)
+
+We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
+Words in the Catalan list were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
+Words in the Spanish list were `r printnum(lengths$eng_spa["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_spa["ipa1_sd"])`, Range = `r printnum(lengths$eng_spa["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_spa["ipa1_max"], digits = 0)`).
+Participants listened to one audio file in each trial, each containing a single word presented in isolation.
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
+The average duration of the audio files was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
+The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
+
+### Participant responses (English/Spanish)
+
+Responses given by English participants for Catalan presented words were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), while their translations to Spani responses given by Spanish participants for Catalan words were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
+Responses tgiven by English participants for Spanish words were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
+
+#### Target translation lexical frequency
+
+We extracted the lexical frequencies for English target translations from SUBTLEX-UK[@van2014subtlex], and frequencies for Spanish from SUBTLEX-ESP [@cuetos2011subtlex].
+We extracted scores as Zipf scores where available or transformed raw scores to Zipf scores to  correct their logarithmic distribution, limiting their range roughly between 0 and 7, allowing an easier interpretation of further analyses [@van2014subtlex].
+
+```{r pthn_example, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+pthn_example <- clearpond_path %>% 
+  map(read.csv) %>% 
+  bind_rows(.id = "group") %>% 
+  clean_names() %>% 
+  filter(word %in% c("book", "sock"), group=="ENG-CAT") %>% 
+  select(word, pthn)
+
+```
+
+#### Target translatiuon phonological neighbourhood density (PTHN)
+
+We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) of the English and Spanish target translations from the CLEARPOND database [@marian2012clearpond].
+PTHN scores indicate the number of phonological neighbours with higher lexical frequency than the target translation, as indicated by its frequency score in the corresponding SUBTLEX database.
+CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target translation in only one addition, deletion, or substitution.
+PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
+
+```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA).
+Columns indicate: A\\) Lexical frequency of target translations, expressed as Zipf scores.
+B\\) PTHN of target translations, and C\\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages.
+Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words."}
+
+d <- stimuli %>% 
+  mutate(
+    group = case_when(
+      group=="ENG-CAT" ~ "cat-ENG",
+      group=="ENG-SPA" ~ "spa-ENG",
+      group=="SPA-CAT" ~ "cat-SPA"
+    ),
+    group = factor(
+      group,
+      levels = c("cat-ENG", "spa-ENG", "cat-SPA")
+    )
+  )
+
+medians <- d %>% 
+  group_by(group) %>% 
+  summarise(
+    pthn_median = median(pthn),
+    pthn_sd = sd(pthn),
+    frequency_zipf_median = median(frequency_zipf),
+    frequency_zipf_sd = sd(frequency_zipf),
+    lv_median = median(lv),
+    lv_sd = sd(lv)
+  )
+
+# frequency
+d %>% 
+  ggplot() +
+  aes(
+    frequency_zipf, 
+    fill = group) +
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = frequency_zipf_median-frequency_zipf_sd, 
+      xmax = frequency_zipf_median+frequency_zipf_sd, 
+      ymin = -Inf, 
+      ymax = Inf, 
+      fill = group
+    ), 
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20,
+    colour = "white"
+    ) +
+  geom_vline(
+    data = medians, 
+    aes(
+      xintercept = frequency_zipf_median,
+      group = group
+      ), 
+    colour = "black",
+    size = 0.5,
+    linetype = "dashed"
+  ) + 
+  geom_text(
+    data = medians, 
+    aes(
+      label = "Median ± SD", 
+      x = frequency_zipf_median+1.25,
+      y = 13
+    ), 
+    size = 3
+  ) +
+  geom_curve(
+    data = medians, 
+    aes(
+      x = frequency_zipf_median+0.85, 
+      xend = frequency_zipf_median+0.1,
+      y = 12, 
+      yend = 11
+    ),
+    arrow = arrow(length = unit(0.2, "cm")),
+    curvature = -0.25,
+    colour = "grey30"
+  ) +
+  labs(
+    x = "Lexical frequency (Zipf)", 
+    y = "Number of trials", 
+    title = "Frequency\n(stimuli translations)\n"
+  ) +
+
+  # PTHN
+  d %>% 
+  count(
+    group, 
+    pthn
+    ) %>% 
+  ggplot() + 
+  aes(
+    pthn,
+    n, 
+    fill = group
+    ) +
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0, 
+      xmax = pthn_median+pthn_sd, 
+      ymin = -Inf, 
+      ymax = Inf,
+      fill = group
+    ), inherit.aes = FALSE, 
+    alpha = 0.15
+  ) +   
+  geom_col(colour = "white") + 
+  geom_vline(
+    data = medians, 
+    aes(
+      xintercept = pthn_median, 
+      group = group
+      ), 
+    colour = "black", 
+    size = 0.5, 
+    linetype = "dashed"
+  ) + 
+  labs(
+    x = "PTHN", 
+    y = "Number of trials", 
+    title = "Higher-frequency neighbours\n(stimuli translations)\n"
+  ) +
+
+  # Levenshtein
+  d %>% 
+  ggplot() +
+  aes(
+    lv, 
+    fill = group
+    ) +
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0,
+      xmax = lv_median+lv_sd, 
+      ymin = -Inf,
+      ymax = Inf, 
+      fill = group
+    ),
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20, 
+    colour = "white"
+    ) +
+  geom_vline(
+    data = medians, 
+    aes(
+      xintercept = lv_median,
+        group = group
+      ), 
+    colour = "black",
+    size = 0.5, 
+    linetype = "dashed"
+  ) + 
+  scale_x_continuous(labels = percent) +
+  labs(
+    x = "Levenshtein", 
+    y = "Number of trials", 
+       title = "Phonological similarity\n(stimuli and stimuli translations)\n"
+    ) +
+
+  plot_layout(guides = "collect") &
+  plot_annotation(tag_levels = "A") &
+  scale_fill_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A"),
+    na.translate = FALSE) &
+  theme(
+    legend.position = "none",
+    title = element_text(size = 10),
+    axis.title = element_text(size = 10)
+  )
+```
+
+### Cross-language phonological similarity (Levenshtein)
+
+We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their IPA transcriptions using the `stringsim` function of the stringdist R package [@van2014stringdist].
+This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
+First, it computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
+This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
+The result is a percentage score, with 0% indicating no similarity between the two strings, and 100% indicating a perfect match between the two strings.
+We computed this similarity measure (refred simply as Levenshtein, from now on) for every translation pair in our stimuli lists.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a 17% similarity, while the *train* (trEIn)-*tren* (tɾEn) translation pair had a 60% similarity.
+Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+
+## Data analysis
+
+We modelled the probability of participants guessing the correct target translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0).
+Secondly, we added the fixed effect of lexical frequency of the target translation ($Frequency$).
+We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels cat-ENG, spa-ENG, and cat-SPA), and finally, the $Levenshtein \times Group$ interaction).
+We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) [@schad2020capitalize].
+Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) [@barr2013random].
+We will later refer to these models as Model 0 to Model 6.
+For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
+
+$$\begin{align}
+
+&\textbf{Likelihood}  \\
+y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
+
+&\textbf{Parameters}  \\
+
+logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
+& \beta_{1[p]} ~ Frequency_{i} ~ + \\
+& \beta_{2[p]} ~ PTHN_i ~ + \\
+& \beta_{3[p]} ~ Similarity_i ~ + \\
+& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
+
+\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
+\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
+&& \text{[participant-level coefficients]} \\ \\
+
+&\textbf{Prior}  \\
+
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
+
+\end{align}$$
+
+To test and account for cross-group differences, we included a random intercept for each group.
+We compared models using leave-one-out cross-validation (*LOO*) [@vehtari2017practical].
+More information about the models and model comparison can be found in Appendix 2.
+All analyses were performed in R environment [@rcore2019r].
+We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
+We used the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+
+# Results
+
+All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
+Model 6, which included all main effects and the two-way interactions between $PTHN$ and $Levenshtein$, and between $Group$ and $Levenshtein$ and $PTHN$ and consonant similarity, showed the best performance.
+
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+tab_loos <- model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
+  arrange(desc(elpd_loo)) %>% 
+  select(
+    model,
+    matches("elpd_loo"),
+    matches("looic"), 
+    matches("diff")) %>%
+  mutate(
+    model = str_replace(
+      model,
+      "fit_", 
+      "Model "
+    )
+  ) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  tab_header(
+    title = md("Table 2"),
+    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Outcomes of the leave-one-out cross-validation procedure.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
+  ) %>% 
+  fmt_missing(
+    everything(),
+    missing_text = "-"
+  ) %>% 
+  fmt_number(2:7) %>% 
+  cols_label(
+    model = md("**Model**"),
+    looic = md("***LOO<sub>IC</sub>***"),
+    se_looic = md("***SE*<sub>IC</sub>**"),
+    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("***LOO<sub>diff</sub>***"),
+    se_diff = md("***SE<sub>diff</sub>***")
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left",
+      weight = "bold", 
+      size = "medium"
+    ),
+    locations = cells_title(
+      groups = "title"
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left", 
+      size = "medium", 
+      style = "italic"
+    ),
+    locations = cells_title(
+      groups = "subtitle"
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left", 
+      size = "medium"
+    ),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = list(
+      cell_borders(
+        sides = c("left", "right"),
+        color = "white"
+      ),
+      cell_text(
+        align = "center"
+      )
+    ),
+    locations = list(
+      cells_body(), 
+      cells_row_groups(), 
+      cells_column_spanners(), 
+      cells_stub(), 
+      cells_source_notes(),
+      cells_column_labels(), 
+      cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(
+      cell_text(
+        align = "left",
+        size = "medium"
+      ), 
+      cell_borders(
+        sides = "all", 
+        color = "white"
+      )
+    ),
+    locations = list(
+      cells_stub(), 
+      cells_source_notes()
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "center",
+      size = "medium", 
+      weight = "normal",
+      style = "normal"
+    ),
+    locations = list(
+      cells_column_spanners(), 
+      cells_column_labels()
+    )
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "all",
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "top", 
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
+  ) %>% 
+  gtsave(here("img", "loos.png"))
+
+include_graphics(here("img", "loos.png"))
+
+```
+
+We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
+For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
+Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
+This way, the coefficients are expressed as increments or decrements in the probability of correct translation [@gelman2020regression].
+
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
+A\\) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Black dots represent the mean of each distribution.
+B\\) Participant\\-level coefficient variability.
+The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
+C\\) Correlation between participant\\-level effects.
+The model allowed participant-level coefficients to co\\-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}
+
+coefs <- fixef(fit_6) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("variable") %>% 
+  clean_names() %>% 
+  mutate_at(
+    vars(estimate:q97_5),
+    ~ifelse(
+      variable=="Intercept", 
+      inv_logit_scaled(.), 
+      ./4
+    )
+  ) %>% 
+  group_split(variable) %>% 
+  set_names(make_clean_names(map(., "variable"))) %>% 
+  map(select, -variable) %>% 
+  map(unlist)
+
+str_repl <- c(
+  "Intercept" = "Intercept",
+  "frequency_zipf_std" = "Frequency\n[+1 SD]",
+  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
+  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv_std" = "Levenshtein\n[+1 SD]",
+  "pthn_std" = "PTHN\n[+1 SD]",
+  "groupENGvs.SPA" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+  "groupENGMCATvs.ENGMSPA" = "Group 2\n[cat-ENG vs. spa-ENG]"
+)
+
+# fixed effects
+post_fix <- posterior_draws_fixed %>% 
+  filter(str_detect(.variable, "b_")) %>% 
+  mutate(
+    .variable_name = str_remove(.variable, "b_") %>% 
+      str_replace_all(str_repl) %>%  
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]",
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1", 
+          "Levenshtein \u00d7 Group 2"
+        )
+      ),
+    .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+  ) %>% 
+  arrange(.variable) %>% 
+  ggplot() +
+  aes(
+    .value,
+    fct_rev(.variable_name)
+  ) +
+  geom_vline(xintercept = 0) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95),
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#1A85FF", "#9ccaff", "#d2e5fc"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = function(x) percent(round(x, 1))
+  ) +
+  labs(
+    x = "P(Correct)", 
+    y = "Posterior probability density",
+    fill = "CrI"
+  ) +
+  theme(
+    legend.position = c(1, 0.25),
+    legend.direction = "vertical",
+    legend.background = element_rect(
+      fill = "grey95"
+    ),
+    legend.key = element_rect(
+      fill = "grey95", 
+      colour = "grey95"), 
+    legend.key.height = unit(0.1, "cm"),
+    axis.title.y = element_blank(),
+    legend.text = element_text(
+      size = 7
+    ),
+    axis.text.x = element_text(
+      colour = "black"
+    ),
+    axis.text.y = element_text(
+      colour = "black",
+      hjust = 1
+    ),
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      size = 0.5
+    )
+  )
+
+# SD of random effects
+post_sd_data <- posterior_draws_fixed %>% 
+  filter(str_detect(.variable, "sd_")) %>% 
+  mutate(
+    .variable_name = str_remove(
+      .variable,
+      "sd_participant__|sd_word__"
+    ) %>% 
+      str_replace_all(str_repl) %>% 
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1",
+          "Levenshtein \u00d7 Group 2"
+        )
+      ),
+    group = case_when(
+      str_detect(.variable, "participant") ~ "Participants",
+      str_detect(.variable, "word") ~ "Words"
+    ),
+    .value = ifelse(
+      str_detect(.variable, "Intercept"), 
+      inv_logit_scaled(.value), 
+      .value/4
+    )
+  )
+
+post_sd <- post_sd_data %>% 
+  filter(group=="Participants") %>% 
+  ggplot() +
+  aes(.value, fct_rev(.variable_name)) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95),
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#ff2976", "#f2a7c2", "#fcd9e6"), 
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By participant", 
+    fill = "CrI"
+  ) +
+  theme(
+    axis.text.x = element_blank()
+  ) +
+
+  post_sd_data %>% 
+  filter(group=="Words") %>% 
+  ggplot() +
+  aes(
+    .value, 
+    fct_rev(.variable_name)
+  ) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95), 
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By item",
+    fill = "CrI"
+  ) +
+  plot_layout(
+    ncol = 1,
+    heights = c(0.85, 0.15), 
+    guides = "collect"
+  ) &
+  theme(
+    legend.position = "none",
+    axis.text.y = element_text(hjust = 1),
+    axis.title.x = element_blank(),
+    legend.direction = "vertical",
+    axis.title.y = element_blank(),
+    legend.background = element_rect(fill = "grey95"),
+    legend.key.height = unit(0.1, "cm"),
+    legend.key = element_rect(fill = "grey95", colour = "grey95"),
+    legend.text = element_text(size = 7),
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
+  )
+
+# Pearson correlations between random effects
+corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
+
+corr_mat <- corr_mat %>% 
+  as_tibble() %>% 
+  select(matches("Estimate")) %>% 
+  rename_all(str_remove, "Estimate.") %>% 
+  mutate(term1 = corr_names)
+
+corr_mat[lower.tri(corr_mat)] <- NA
+
+corr_data <- corr_mat %>% 
+  as.data.frame() %>% 
+  pivot_longer(
+    -term1, 
+    names_to = "term2",
+    values_to = ".value"
+  ) %>% 
+  drop_na(.value)
+
+post_cors <- corr_data %>% 
+  mutate(
+    term1 = factor(
+      term1, 
+      levels = names(str_repl),
+      ordered = TRUE
+    ),
+    term2 = factor(
+      term2, 
+      levels = names(str_repl),
+      ordered = TRUE
+    )
+  ) %>% 
+  mutate_at(
+    vars(term1, term2),
+    str_replace_all, 
+    str_repl
+  ) %>% 
+  mutate(
+    .value_label = case_when(
+      term1==term2 ~ NA_character_, 
+      TRUE ~ printnum(.value, gt1 = FALSE)
+    ),
+    .value = case_when(
+      term1==term2 ~ NA_real_, 
+      TRUE ~ .value
+    )
+  ) %>% 
+  filter(
+    term1 != "PTHN × Levenshtein", 
+    term2 != "Intercept"
+  ) %>% 
+  ggplot() +
+  aes(
+    fct_inorder(term1),
+    fct_rev(fct_inorder(term2)), 
+    fill = .value
+  ) +
+  geom_tile(na.rm = TRUE) +
+  geom_text(
+    aes(label = .value_label), 
+    size = 3, 
+    na.rm = TRUE) +
+  labs(
+    x = "Term 1", 
+    y = "Term 2",
+    fill = "Posterior\ncorrelation"
+  ) +
+  scale_fill_gradient2(
+    low = "#04bf78", 
+    mid = "white",
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
+  ) +
+  scale_x_discrete(position = "top") +
+  coord_equal() +
+  theme(
+    legend.position = c(1, 1),
+    legend.direction = "horizontal",
+    legend.key.height = unit(0.1, "cm"),
+    legend.key.width = unit(0.3, "cm"),
+    legend.justification = c(1, 1),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 5),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(
+      size = 6,
+      angle = 90,
+      hjust = 0, 
+      vjust = 1),
+    axis.text.y = element_text(
+      size = 6,
+      vjust = 1,
+      hjust = 1
+    )
+  )
+
+post_fix / (
+  (post_sd | post_cors) +
+    plot_layout(widths = c(0.6, 0.4))
+) +
+  plot_layout() +
+  plot_annotation(tag_levels = "A")
+
+```
+
+```{r bayesfactors}
+
+get_bf <- function(hyp, abs = TRUE, ...){
+  bf <- hypothesis(fit_6, hypothesis = hyp, ...)$hypothesis$Evid.Ratio
+  if (abs) bf <- abs(bf)
+  return(bf)
+}
+
+h_frequency <- get_bf("frequency_zipf_std = 0")
+h_pthn <- get_bf("pthn_std = 0")
+h_lv <- get_bf("pthn_std = 0")
+
+h_pthn_lv <- get_bf("pthn_std:lv_std = 0")
+h_group1 <- get_bf("groupENGvs.SPA = 0")
+h_group2 <- get_bf("groupENGMCATvs.ENGMSPA = 0")
+h_lv_group1 <- get_bf("lv_std:groupENGvs.SPA = 0")
+h_lv_group2 <- get_bf("lv_std:groupENGMCATvs.ENGMSPA = 0")
+
+icc_conditional <- icc(fit_6, by_group = TRUE)$ICC[2]
+
+sds <- gather_draws(fit_6, `sd_.*`, regex = TRUE) %>% 
+  mean_qi() %>% 
+  mutate(
+    .value = ifelse(
+      grepl("Intercept", .variable), 
+      inv_logit_scaled(.value), 
+      .value/4
+    ),
+    .lower = ifelse(
+      grepl("Intercept", .variable),
+      inv_logit_scaled(.lower),
+      .lower/4
+    ),
+    .upper = ifelse(
+      grepl("Intercept", .variable),
+      inv_logit_scaled(.upper), 
+      .upper/4
+    )
+  ) %>% 
+  split(.$.variable)
+
+names(sds) <- make_clean_names(names(sds))
+
+word_example <- ranef(fit_6)$word[,,"Intercept"] %>% 
+  as.data.frame() %>% 
+  rownames_to_column("word") %>% 
+  arrange(-Estimate) %>% 
+  clean_names() %>% 
+  mutate_if(is.numeric, inv_logit_scaled) %>% 
+  slice(1, n()) %>% 
+  left_join(
+    distinct(
+      stimuli,
+      word1, 
+      ipa1), 
+    by = c("word" = "word1")) %>% 
+  split(.$word)
+
+participant_ranef_range <- sds %>% 
+  bind_rows() %>% 
+  filter(!grepl("Intercept|word|group1", .variable)) %>% 
+  summarise(.value = range(.value)) %>% 
+  unlist()
+
+cors_range <- gather_draws(
+  fit_6, 
+  `cor_.*`, 
+  regex = TRUE
+) %>% 
+  median_qi() %>% 
+  summarise(.value = range(.value))
+
+corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
+
+corr_data <- corr_mat %>% 
+  as_tibble() %>% 
+  clean_names() %>% 
+  mutate(term1 = corr_names) %>% 
+  relocate(term1) %>% 
+  split(.$term1)
+
+names(corr_data) <- make_clean_names(names(corr_data))
+
+h_cor_intercept_frequency <- get_bf("participant__Intercept__frequency_zipf_std = 0", class = "cor")
+h_cor_frequency_zipf_lv_group1 <- get_bf("participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", class = "cor")
+h_cor_intercept_pthn <- get_bf("participant__Intercept__pthn_std = 0", class = "cor")
+```
+
+Overall, participants had `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likelihood to produce correct translations.
+The lexical frequency of the target translation had little impact on accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf_std["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
+The number of phonological neighbours with higher lexical frequency than the target translation (*PTHN*) slightly decreased the probability of a correct responses by `r percent(coefs$pthn_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
+Finally, phonological similarity between the presented word and target translation had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
+The interaction effect between $PTHN \times Levenshtein$ was small: a +1 SD increment in Levenshtein increased the effect of *PTHN* by `r percent(coefs$pthn_std_lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std_lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std_lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
+
+Spanish natives translating Catalan words were `r percent(coefs$group_en_gvs_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
+The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in `r percent(coefs$lv_std_group_en_gvs_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+English natives translating Catalan words were `r percent(abs(coefs$group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
+For every one *SD* increment in Levenshtein similarity, the effect of presented language increased in `r percent(abs(coefs$lv_std_group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+
+```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
+A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean similarity (20.51% Levenshtein), and 100 simulations for translations with the maximum similarity observed in the dataset (83.33% Levenshtein).
+These simulations were obtained for the range of values of the PTHN scores.
+For each simulation, we drew a single sample from the posterior distribution of each coefficient.
+Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
+Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
+The dispersion of the lines indicates the uncertainty of our predictions.
+We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
+
+nd <- expand.grid(
+	lv_std = c(
+		min(fit_6$data$lv_std),
+		0,
+		max(fit_6$data$lv_std)
+	),
+	frequency_zipf_std = 0,
+	pthn_std = seq(
+		min(fit_6$data$pthn_std, na.rm = TRUE),
+		max(fit_6$data$pthn_std, na.rm = TRUE),
+		by = 0.1
+	),
+	group = unique(fit_6$data$group)
+)
+
+m <- add_epred_draws(
+	nd, 
+	fit_6, 
+	ndraws = 100, 
+	re_formula = NA
+) %>% 
+	mutate(
+		lv_std = factor(
+			lv_std,
+			levels = c(
+				min(fit_6$data$lv_std),
+				0,
+				max(fit_6$data$lv_std)
+			),
+			labels = c(
+				paste0("Min (", percent(min(responses$lv, na.rm = TRUE)), ")"),
+				paste0("Mean (", percent(mean(responses$lv, na.rm = TRUE)), ")"),
+				paste0("Max (", percent(max(responses$lv, na.rm = TRUE)), ")")
+			)
+		) 
+	)
+
+ggplot(m) +
+	aes(
+		pthn_std, 
+		.epred,
+		colour = lv_std,
+		fill = lv_std
+	) +
+	facet_wrap(~group) +
+	geom_hline(
+		yintercept = 0.5, 
+		colour = "grey"
+	) +
+	geom_line(
+		aes(group = interaction(lv_std, .draw)), 
+		size = 0.65,
+		alpha = 0.5,
+		show.legend = TRUE
+	) +
+	stat_summary(
+		aes(linetype = lv_std), 
+		fun = mean, 
+		geom = "line",
+		colour = "black", 
+		size = 0.75,
+		show.legend = TRUE
+	) +
+	labs(
+		x = paste0("PTHN (+1 SD, ", round(sd(responses$pthn), 1), " higher-frequency neighbours)"),
+		y = "P(Correct)", 
+		colour = "Levenshtein similarity",
+		fill = "Levenshtein similarity",
+		linetype = "Levenshtein similarity",
+		title = "Population-level posterior predictions"
+	) +
+  scale_color_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A")
+  ) +
+	scale_y_continuous(
+		labels = function(x) percent(round(x, 2)),
+		limits = c(0, 1),
+		breaks = seq(0, 1, 0.1)
+	) +
+	theme(
+		legend.position = "top",
+		plot.title = element_blank()
+	)
+
+```
+
+The conditional intra-class correlation [*ICC*; @gelman2006data] for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items.
+Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were very likely to be correctly translated (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
+Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants.
+Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
+The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_std_group_en_gvs_spaE$.upper, accuracy = 0.01)`]).
+This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
+
+Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions.
+The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_frequency, 2)`).
+The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf_std$estimate_lv_std_group_en_gvs_spa, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf_std$q2_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf_std$q97_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
+The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).
+The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
+
+# Discussion
+
+\newpage
+
+# References
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+\setlength{\leftskip}{0.5in}
+
+<div id="refs" custom-style="Bibliography"></div>
+\endgroup
+
+\newpage
+
+>>>>>>> 849feac7cbd8456a1cac154c57c0b33b987d4387

### 2022-04-27 docs/index.Rmd
No idea what changed

--- 2022-03-28 docs/index.Rmd
@@
       ., 
-      test_language, 
+      group, 
       name = "n_total"
@@
     test_language = paste0(
-      test_language, 
+      group, 
       " (N = ", 
@@
   count(
-    test_language, 
+    group, 
     pthn
@@
   aes(
-    reorder(as.factor(pthn), -n), 
+    as.factor(pthn), 
     n,
-    fill = test_language
+    fill = group
   ) +
   facet_wrap(
-  ~test_language,
+    ~group,
     ncol = 1
@@
 
-* fit_0: `~ 1 + (1 | participant) + (1 | word)`
-* fit_1: `~ 1 + frequency_zipf + (1 + frequency_zipf | participant) + (1 | word)`
-* fit_2: `~ 1 + frequency_zipf + pthn + (1 + frequency_zipf + pthn | participant) + (1 | word)`
-* fit_3: `~ 1 + frequency_zipf + pthn + lv + (1 + frequency_zipf + pthn + lv | participant) + (1 | word)`
-* fit_4: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
-* fit_5: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
-* fit_6: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + group:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
+* fit_0: `~ 1 + (1 | participant_id) + (1 | translation_id)`
+* fit_1: `~ 1 + frequency_zipf + (1 + frequency_zipf | participant_id) + (1 | translation_id)`
+* fit_2: `~ 1 + frequency_zipf + pthn + (1 + frequency_zipf + pthn | participant_id) + (1 | translation_id)`
+* fit_3: `~ 1 + frequency_zipf + pthn + lv + (1 + frequency_zipf + pthn + lv | participant_id) + (1 | translation_id)`
+* fit_4: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant_id) + (1 | translation_id)`
 
@@
 
-coefs <- fixef(fit_6) %>% 
+coefs <- fixef(fit_4) %>% 
   as.data.frame() %>% 
@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:group1" = "Levenshtein \u00d7 Group 1",
-  "lv_std:group2" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
-  "pthn_std" = "PTHN\n[+1 SD]",
-  "group1" =  "Group 1\n[*-ENG vs. cat-SPA]",
-  "group2" = "Group 2\n[cat-ENG vs. spa-ENG]"
+  "pthn_std" = "PTHN\n[+1 SD]"
 )
@@
           "Levenshtein\n[+1 SD]", 
-          "PTHN \u00d7 Levenshtein",
-          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-          "Group 2\n[cat-ENG vs. spa-ENG]",
-          "Levenshtein \u00d7 Group 1",
-          "Levenshtein \u00d7 Group 2"
+          "PTHN \u00d7 Levenshtein"
         )
@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
-  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
-  "pthn_std" = "PTHN\n[+1 SD]",
-  "groupENGvs.SPA" =  "Group 1\n[*-ENG vs. cat-SPA]",
-  "groupENGMCATvs.ENGMSPA" = "Group 2\n[cat-ENG vs. spa-ENG]"
+  "pthn_std" = "PTHN\n[+1 SD]"
 )
@@
       .variable, 
-      "sd_participant__|sd_word__"
+      "sd_participant_id__|sd_translation_id__"
     ) %>% 
@@
       str_detect(.variable, "participant") ~ "Participants",
-      str_detect(.variable, "word") ~ "Words"
+      str_detect(.variable, "translation") ~ "Words"
     ),
@@
 
-post_sd <- filter(post_sd_data, group=="Participants") %>% 
+post_sd <- post_sd_data %>% 
+  filter(group=="Participants") %>% 
   ggplot() +
@@
   stat_slab(
-    aes(fill = stat(
+    aes(
+      fill = stat(
         cut_cdf_qi(
@@
   ) +
-
   plot_layout(
@@
 # Pearson correlations between random effects
-corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
-corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
+corr_mat <- VarCorr(fit_4, summary = TRUE)$participant_id$cor 
+corr_names <- rownames(VarCorr(fit_4, summary = TRUE)$participant_id$cor)
 
@@
     fill = .value
-  )
+  ) +
   geom_tile(na.rm = TRUE) +
@@
     legend.key.height = unit(0.1, "cm"),
-    legend.key.width = unit(0.3, "cm"),
+    legend.key.width = unit(0.4, "cm"),
     legend.justification = c(1, 1),
     legend.title = element_blank(),
-    legend.text = element_text(size = 5),
+    legend.text = element_text(size = 7),
     axis.title.x = element_blank(),
@@
     axis.ticks = element_blank(),
-    axis.text.x = element_text(size = 6, angle = 90, hjust = 0, vjust = 1),
-    axis.text.y = element_text(size = 6, vjust = 1, hjust = 1)
+    axis.text.x = element_text(size = 8, angle = 90, hjust = 0, vjust = 1),
+    axis.text.y = element_text(size = 8, vjust = 1, hjust = 1)
   )
@@
       participants,
-      participant,
+      participant_id,
       group
@@
     group, 
-    participant, 
+    participant_id, 
     .param
@@
     .value, 
-    sort(as.factor(participant)), 
+    sort(as.factor(participant_id)), 
     xmin = .lower,
@@
 ```{r post_re_word, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
-fixed_intercept <- fixef(fit_6)[1,c(1, 3, 4)] %>% 
+fixed_intercept <- fixef(fit_4)[1,c(1, 3, 4)] %>% 
   inv_logit_scaled()
+
+empirical_accuracy <- responses %>% 
+  group_by(group, translation_id, lv) %>% 
+  summarise(
+    successes = sum(correct, na.rm = TRUE),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    prop = prop_adj(successes, n),
+    prop_se = prop_adj_se(successes, n)
+  )
 
@@
   arrange(
-    word, 
+    translation_id, 
     .param
@@
     .value = inv_logit_scaled(.value)
-  )
-
+  ) %>% 
+  left_join(
+    empirical_accuracy
+  )
+
+post_re %>% 
+  left_join(
+    select(
+      stimuli,
+      group,
+      translation_id,
+      translation,
+      lv
+    ) %>% 
+      mutate(translation = paste0(translation, " (", percent(lv), ")"))
+  ) %>% 
   ggplot(
-  post_re,
-  aes(
-    .value,
-    reorder(
-      as.factor(word), 
+    aes(
+      x = .value,
+      y = reorder(
+        as.factor(translation), 
         .value
@@
     )
+  ) +
+  facet_wrap(
+    ~group,
+    scales = "free_y"
   ) +
@@
   ) +
-  annotate(
-    geom = "text",
-    x = 0.7,
-    y = 100,
-    size = 3,
-    hjust = 0.5,
-    label = "Item accuracy \n(95%, 67%, and 50% CrI \nof item-level intercepts)"
-  ) +
-  annotate(
-    geom = "rect",
-    alpha = 0.15,
-    xmin = fixed_intercept[2],
-    xmax = fixed_intercept[3],
-    ymin = -Inf,
-    ymax = Inf
-  ) +
-  annotate(
-    x = 0.7,
-    xend = 0.6,
-    y = 108,
-    yend = 130,
-    geom = "curve",
-    curvature = 0.25,
-    arrow = arrow(length = unit(0.2, "cm"))
-  ) +
-  annotate(
-    geom = "text",
-    x = 0.8,
-    y = 10,
-    size = 3,
-    hjust = 0.5,
-    label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
-  ) +
-  annotate(
-    x = 0.7,
-    xend = fixed_intercept[1]+0.05,
-    y = 17,
-    yend = 30,
-    geom = "curve",
-    curvature = 0.2,
-    arrow = arrow(length = unit(0.2, "cm"))
+  geom_point(
+    aes(
+      x = prop
+    ),
+    size = 0.5,
+    shape = 3
+  ) +
+  geom_errorbar(
+    aes(
+      x = prop,
+      xmin = prop-prop_se,
+      xmax = prop+prop_se
+    ),
+    width = 0,
+    size = 0.5
   ) +
   labs(
-    x = "Posterior probability of correct response",
-    y = "Item",
-    colour = "Credible Interval (CrI)",
-    caption = "Credible intervals cover the range of values of the sampling space that \ncontain the true value with a given certainty."
+    x = "Probability of correct response",
+    y = "Translation",
+    colour = "Model prediction (credible interval)",
+    caption = "Credible intervals cover the range of values of the sampling space that contain the true value with a given certainty."
   ) +
@@
   theme(
-    legend.position = c(0.1, 0.9),
-    legend.direction = "vertical",
-    legend.justification = "left",
+    legend.position = "top",
     legend.key = element_rect(
@@
   height = 9,
-  width = 5, 
-  dpi = 500
-)
-```
-
-```{r post_re_word_empirical, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-means <- responses %>% 
-  group_by(group) %>% 
+  width = 14,
+  dpi = 1000
+)
+```
+
+```{r post_re_word_lv, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+fixed_intercept <- fixef(fit_4)[1,c(1, 3, 4)] %>% 
+  inv_logit_scaled()
+
+empirical_accuracy <- responses %>% 
+  group_by(group, translation_id, lv) %>% 
   summarise(
@@
 
-posterior_accuracy <-
-
-  responses %>% 
-  mutate(
-    stim = paste0(
-      word, " (", ipa1, ") / ", 
-      word2, " (", ipa2, ") / ",
-      percent(lv, accuracy = 1)
-    )
-  ) %>% 
-  group_by(group, stim, lv) %>% 
-  summarise(
-    successes = sum(correct, na.rm = TRUE),
-    n = n(),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    prop = prop_adj(successes, n),
-    prop_se = prop_adj_se(successes, n)
-  ) %>% 
-  ggplot() +
+post_re <- posterior_draws_random$words %>% 
+  ungroup() %>% 
+  arrange(
+    translation_id, 
+    .param
+  ) %>% 
+  mutate(
+    .param_name = str_replace_all(
+      .param, 
+      str_repl
+    ) %>%  
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein"
+        )
+      ),
+    .value = inv_logit_scaled(.value)
+  ) %>% 
+  left_join(
+    empirical_accuracy
+  )
+
+post_re %>% 
+  left_join(
+    select(
+      stimuli,
+      group,
+      translation_id,
+      translation,
+      lv
+    ) %>% 
+      mutate(translation = paste0(translation, " (", percent(lv), ")"))
+  ) %>% 
+  ggplot(
+    aes(
+      x = .value,
+      y = reorder(
+        as.factor(translation), 
+        lv
+      )
+    )
+  ) +
+  facet_wrap(
+    ~group,
+    scales = "free_y"
+  ) +
+  geom_vline(
+    xintercept = fixed_intercept[1],
+    colour = "grey", 
+    size = 1
+  ) +
+  stat_interval(
+    size = 1.5,
+    .width = c(0.5, 0.67, 0.95)
+  ) +
+  geom_point(
+    aes(
+      x = prop
+    ),
+    size = 0.5,
+    shape = 4
+  ) +
+  geom_errorbar(
     aes(
       x = prop,
-    y = reorder(as.factor(stim), lv),
       xmin = prop-prop_se,
-    xmax = prop+prop_se,
-    label = percent(lv),
-    colour = group
-  ) +
-  facet_wrap(~group, scales = "free_y") +
-  geom_vline(
-    data = means,
-    aes(xintercept = prop),
-    size = 1,
-    colour = "grey"
-  ) +
-  geom_errorbar(width = 0.1) +
-  geom_point() +
-  annotate(
-    geom = "text",
-    x = 0.7,
-    y = 10,
-    size = 3,
-    hjust = 0.5,
-    label = "Item accuracy \n(95% CI)"
-  ) +
-  annotate(
-    x = 0.7,
-    xend = 0.15,
-    y = 13,
-    yend = 20,
-    geom = "curve",
-    curvature = 0.25,
-    arrow = arrow(length = unit(0.2, "cm"))
-  ) +
-  # annotate(
-  #   geom = "text",
-  #   x = 0.8,
-  #   y = 10,
-  #   size = 3,
-  #   hjust = 0.5,
-  #   label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
-  # ) +
-  # annotate(
-  #   x = 0.7,
-  #   xend = fixed_intercept[1]+0.05,
-#   y = 17,
-#   yend = 30,
-#   geom = "curve",
-#   curvature = 0.2,
-#   arrow = arrow(length = unit(0.2, "cm"))
-# ) +
+      xmax = prop+prop_se
+    ),
+    width = 0,
+    size = 0.5
+  ) +
   labs(
-  x = "Probability of correct response",
-  y = "Presented word / Correct translation / Levenshtein similarity"
-) +
-  scale_colour_manual(
-    values = c("#1A85FF", "#ff2976", "#FFC20A"),
-    na.translate = FALSE
-  ) +
-  scale_x_continuous(
-    limits = c(0, 1), 
-    labels = percent
-  ) +
+    x = "Posterior probability of correct response",
+    y = "Translation (Levenshtein similarity)",
+    colour = "Model prediction (credible interval)",
+    caption = "Credible intervals cover the range of values of the sampling space that contain the true value with a given certainty."
+  ) +
+  scale_color_manual(
+    values = c("#d2e5fc", "#9ccaff", "#1A85FF"),
+    na.translate = FALSE, 
+    labels = paste0(c(95, 67, 50), "%")
+  ) +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
   theme(
-    legend.position = "none",
-    legend.direction = "vertical",
-    legend.justification = "left",
-    legend.key = element_rect(fill = NA, colour = NA),
+    legend.position = "top",
+    legend.key = element_rect(
+      fill = NA, 
+      colour = NA
+    ),
     panel.grid.major.x = element_line(
@@
       face = "plain",
-      margin = margin(t = unit(3, "cm")))
+      margin = margin(t = unit(3, "cm"))
+    )
   )
@@
 ggsave(
-  here("img", "accuracy_empirical.png"), 
+  here("img", "accuracy_lv.png"), 
   height = 9,
-  width = 15,
-  dpi = 500
+  width = 16,
+  dpi = 1000
 )
@@
   lv_std = c(
-    min(fit_6$data$lv_std),
+    min(fit_4$data$lv_std),
     0,
-    max(fit_6$data$lv_std)
+    max(fit_4$data$lv_std)
   ),
@@
   pthn_std = seq(
-    min(fit_6$data$pthn_std, na.rm = TRUE),
-    max(fit_6$data$pthn_std, na.rm = TRUE),
+    min(fit_4$data$pthn_std, na.rm = TRUE),
+    max(fit_4$data$pthn_std, na.rm = TRUE),
     by = 0.1
-  ),
-  group = unique(fit_6$data$group)
+  )
 )
@@
   nd, 
-  fit_6, 
+  fit_4, 
   ndraws = 100, 
@@
       levels = c(
-        min(fit_6$data$lv_std),
+        min(fit_4$data$lv_std),
         0,
-        max(fit_6$data$lv_std)
+        max(fit_4$data$lv_std)
       ),
@@
   ) +
-  facet_wrap(~group) +
   geom_hline(

### 2022-04-27 R/02_models.R

@@ -25,25 +25,42 @@ get_model_fit <- function(name, formula, prior, ...){
 
 # leave-one-out cross-validation
 get_model_loos <- function(fits){
+    
     path <- here("results", "model_loos.rds")
+    
     if (file.exists(path)){
         loo <- readRDS(here(path))
+        
     } else {
         loo <- loo_compare(map(fits, loo))
         saveRDS(loo, path)
     }
+    
     return(loo)
 }
 
 # get model posterior draws (fixed effects)
-get_model_draws_fixed <- function(fit){
-    post <- gather_draws(fit, `b_.*`, `sd_.*`, regex = TRUE)
+get_model_draws_fixed <- function(
+    fit
+){
+    
+    post <- gather_draws(
+        fit, 
+        `b_.*`,
+        `sd_.*`,
+        regex = TRUE
+    )
+    
     return(post)
 }
 
 
 # get model posterior expected predictions (fixed effects)
-get_model_epreds_fixed <- function(fit, ndraws = 50){
+get_model_epreds_fixed <- function(
+    fit, 
+    ndraws = 100
+){
+    
     nd <- expand.grid(
         lv_std = c(-1, 0, 1),
         frequency_zipf_std = 0,
@@ -51,67 +68,124 @@ get_model_epreds_fixed <- function(fit, ndraws = 50){
             min(fit$data$pthn_std, na.rm = TRUE),
             max(fit$data$pthn_std, na.rm = TRUE),
             by = 0.1
-        ),
-        group = unique(fit$data$group)
+        )
     )
-    m <- add_epred_draws(nd, fit, ndraws = 50, re_formula = NA)
+    
+    m <- add_epred_draws(
+        nd, 
+        fit, 
+        ndraws = 50,
+        re_formula = NA
+    )
+    
     return(m)
 }
 
 # get model posterior expected predictions (random effects)
-get_model_epreds_random <- function(fit, ndraws = 100, group = c("participant", "word")){
+get_model_epreds_random <- function(
+    fit, 
+    ndraws = 100,
+    group = c(
+        "participant",
+        "word"
+    )
+){
     
     # by participant
     m_re_participants <- NULL
+    
     if ("participant" %in% group){
+        
         nd_re_participants <- expand.grid(
-            participant = unique(fit$data$participant),
+            participant_id = unique(fit$data$participant_id),
             lv_std = c(-1, 0, 1),
             frequency_zipf_std = 0,
             pthn_std = seq(
                 min(fit$data$pthn_std, na.rm = TRUE),
[…]

### 2022-04-27 R/01_responses.R
No idea what changed

@@ -23,13 +23,28 @@ get_responses_processed <- function(
         clean_names() %>% 
         # select relevant variables and rename if necessary
         select(
-            participant, test_language, trial_id, word, soundfile, key_pressed, input_text, key_press_time, error, age, date, city,
+            participant, 
+            test_language, 
+            trial_id, 
+            word, 
+            soundfile, 
+            key_pressed, 
+            input_text, 
+            key_press_time, 
+            error,
+            age, 
+            date, 
+            city,
             matches("language_l|language_s|language_c|demo_|setup_"),
             -matches("_rt|_time"),
             key_press_time
         ) %>% 
         rename(demo_impairment = demo_language_key_keys) %>% 
-        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
+        rename_all(
+            gsub,
+            pattern = "_key_keys|key_keys",
+            replacement = ""
+        ) %>% 
         # clean text input by participants (because of typos of need to translate) and redefine location    
         group_by(participant) %>% 
         mutate_at(
@@ -37,15 +52,31 @@ get_responses_processed <- function(
                 starts_with("demo_"),
                 starts_with("l"),
                 starts_with("setup_"),
-                age, city, setup_location, setup_noise
+                age, 
+                city,
+                setup_location, 
+                
             ),
-            first_non_na) %>% 
+            first_non_na
+        ) %>% 
         mutate(date = max(date, na.rm = TRUE)) %>% 
         ungroup() %>% 
-        mutate_at(vars(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), clean_input_text) %>% 
+        mutate_at(
+            vars(
+                starts_with("language_"),
+                city, 
+                starts_with("demo_"),
+                starts_with("l")
+            ),
+            clean_input_text
+        ) %>% 
         mutate(
             city = str_to_sentence(city),
-            country = ifelse(city %in% spanish_cities, "Spain", "UK"),
+            country = ifelse(
+                city %in% spanish_cities,
+                "Spain", 
+                "UK"
+            ),
             date = as_datetime(ymd_hms(date)),
             demo_sex = case_when(
                 country %in% "UK" & demo_sex %in% "M" ~ "Male",
@@ -69,11 +100,23 @@ get_responses_processed <- function(
                 country %in% "UK" & test_language %in% "Spanish" ~ "ENG-SPA",
                 TRUE ~ "SPA-CAT"
             ),
[…]

### 2022-04-27 docs/index.Rmd

--- 2022-04-27 docs/index.Rmd
@@
+<<<<<<< HEAD
+---
+title: "Translation Elicitation: Report"
+date: "Last updated: `r format(Sys.Date(), '%d/%m/%Y')`"
+output:
+  html_document:
+    theme: paper
+    highlight: pygments
+    toc: true
+    toc_depth: 2
+    toc_float:
+        collapsed: false
+    number_sections: true
+    keep_md: true
+bibliography: "references.bib"
+csl: "apa7.csl"
+---
+
+```{r setup, include=FALSE}
+
+knitr::opts_chunk$set(
+  message = FALSE,
+  warning = FALSE,
+  cache.extra = knitr::rand_seed,
+  dev.args = list(png = list(type = "cairo")),
+  dpi = 300,
+  include = TRUE,
+  out.width = "80%",
+  results = "asis",
+  echo = TRUE
+)
+
+options(
+  knitr.kable.NA = '-',
+  htmltools.preserve.raw = FALSE,
+  knitr.duplicate.label = "allow",
+  ggplot2.discrete.fill = c("#1A85FF", "#ff2976", "#FFC20A"),
+  ggplot2.discrete.colour = c("#1A85FF", "#ff2976", "#FFC20A"),
+  ggplot2.continuous.fill = scale_fill_gradient2(
+    low = "#04bf78",
+    mid = "white",
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
+  ),
+  ggplot2.continuous.colour = scale_fill_gradient2(
+    low = "#04bf78",
+    mid = "white",
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
+  )
+)
+```
+
+```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+# import data
+tar_load_globals()
+tar_load_all()
+
+theme_set(
+  theme(
+    title = element_text(size = 8, face = "bold"),
+
+    legend.background = element_rect(fill = NA, colour = NA),
+    legend.key = element_rect(fill = "white", colour = NA),
+    legend.justification = "right",
+    legend.direction = "horizontal",
+
+    axis.line = element_line(colour = "black", size = 0.5),
+    axis.ticks = element_line(colour = "black", size = 0.5),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black"),
+
+    plot.caption.position = "plot",
+    plot.title = element_text(hjust = 0.5),
+
+    panel.grid = element_blank(),
+    panel.background = element_rect(fill = "white", colour = NA)
+  )
+)
+
+```
+
+# Stimuli
+
+## Summary
+
+Participants in the ENG-SPA listened to `r nrow(stimuli[stimuli$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials.
+Five trials in each condition were practice trials and were excluded from analyses.
+
+```{r stimuli-coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
+
+stimuli_coded <- stimuli %>% 
+  mutate(test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"))
+```
+
+```{r stimuli-table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli %>% 
+  select(
+    group,
+    word1,
+    frequency, 
+    frequency_zipf,
+    pthn, 
+    lv
+  ) %>% 
+  group_by(group) %>% 
+  summarise_at(
+    vars(frequency:lv), 
+    list(
+      median = ~median(., na.rm = TRUE),  
+      sd = ~sd(., na.rm = TRUE)
+    )
+  ) %>%
+  gt(rowname_col = "group") %>% 
+  tab_header(
+    title = md("Table 1"),
+    subtitle = md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Median and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
+  ) %>% 
+  fmt_number(frequency_median:pthn_sd) %>%
+  fmt_percent(matches("lv")) %>% 
+  tab_spanner(
+    md("**Freq. (million)**"), 
+    c("frequency_median", "frequency_sd")
+  ) %>% 
+  tab_spanner(
+    md("**Freq. (Zipf)**"),
+    matches("frequency_zipf")) %>% 
+  tab_spanner(
+    md("***PTHN***"), 
+    matches("pthn")) %>% 
+  tab_spanner(
+    md("**Levenshtein**"),
+    matches("lv")) %>% 
+  cols_label(
+    frequency_median = md("Median"),
+    frequency_sd = md("*SD*"),
+    frequency_zipf_median = md("Median"),
+    frequency_zipf_sd = md("*SD*"),
+    pthn_median = md("Median"),
+    pthn_sd = md("*SD*"),
+    lv_median = md("Median"),
+    lv_sd = md("*SD*")
+  ) %>% 
+  summary_rows(
+    columns = c(
+      matches("median"), 
+      -matches("lv")
+    ),
+    fns = list(
+      Median = "median",
+      SD = "sd"
+    )
+  ) %>% 
+  summary_rows(
+    columns = matches("lv"),
+    fns = list(
+      Median = "median",
+      SD = "sd"
+    ), formatter = fmt_percent
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left",
+      weight = "bold", 
+      size = "medium"
+    ),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left", 
+      size = "medium",
+      style = "italic"
+    ),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left",
+      size = "medium"
+    ),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left",
+      size = "medium"
+    ),
+    locations = cells_stub_grand_summary()
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = c("left", "right"),
+      color = "white"
+    ),
+    locations = list(
+      cells_body(), 
+      cells_row_groups(), 
+      cells_column_spanners(), 
+      cells_stub(), 
+      cells_stub_grand_summary(),
+      cells_source_notes(),
+      cells_column_labels(), 
+      cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(
+      cell_text(
+        align = "left", 
+        size = "medium"
+      ), 
+      cell_borders(
+        sides = "all", 
+        color = "white"
+      )
+    ),
+    locations = cells_stub()
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "center",
+      size = "medium", 
+      weight = "normal", 
+      style = "normal"
+    ),
+    locations = list(
+      cells_column_spanners(),
+      cells_column_labels()
+    )
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "all",
+      color = "white"
+    ),
+    locations = cells_body(columns = 1:9)
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "top", 
+      color = "white"
+    ),
+    locations = cells_body(columns = 1:9)
+  )
+
+```
+
+## Lexical frequency
+
+```{r stimuli-frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
+  distinct(
+    test_language,
+    word1, 
+    .keep_all = TRUE
+  ) %>% 
+  left_join(
+    count(
+      ., 
+      test_language,
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    test_language = paste0(
+      test_language,  
+      " (N = ", n_total, 
+      ")"
+    )
+  ) %>% 
+  ggplot() +
+  aes(
+    frequency_zipf,
+    fill = test_language
+  ) +
+  facet_wrap(
+    ~test_language,
+    ncol = 1
+  ) +
+  geom_histogram(
+    bins = 25, 
+    colour = "white"
+  ) +
+  labs(
+    x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database",
+    y = "# trials",
+    fill = "Test language"
+  )  +
+  scale_y_continuous(
+    breaks = seq(0, 12, 2)
+  ) +
+  theme(
+    legend.position = "none",
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    )
+  )
+
+```
+
+## PTHN
+
+```{r stimuli-pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
+  left_join(
+    count(
+      ., 
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    test_language = paste0(
+      group, 
+      " (N = ", 
+      n_total, 
+      ")"
+    )
+  ) %>% 
+  count(
+    group, 
+    pthn
+  ) %>% 
+  ggplot() +
+  aes(
+    as.factor(pthn), 
+    n,
+    fill = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) +
+  geom_col(position = position_dodge()) +
+  labs(
+    x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database",
+    y = "# trials",
+    fill = "Test language"
+  ) +
+  theme(
+    legend.position = "none",
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    )
+  )
+
+```
+
+## Levenshtein distance
+
+```{r stimuli-similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli_coded %>% 
+  left_join(
+    count(
+      ., 
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(
+      group,  
+      " (N = ",
+      n_total,
+      ")"
+    )
+  ) %>% 
+  ggplot() +
+  aes(
+    group,
+    lv,
+    fill = group, 
+    colour = group
+  ) +
+  geom_violin() +
+  geom_point(
+    shape = 1, 
+    stroke = 1, 
+    position = position_jitter(width = 0.1),
+    colour = "black",
+    alpha = 0.25
+  ) +
+  geom_boxplot(
+    colour = "black",
+    fill = "white", 
+    width = 0.1, 
+    size = 0.75
+  ) +
+  labs(
+    x = "Group", 
+    y = "Phonological Leveshtein distance\n(corrected for string length)",
+    fill = "Test language"
+  ) +
+  scale_y_continuous(
+    limits = c(0, 1),
+    labels = scales::percent
+  ) +
+  theme(
+    legend.position = "none",
+    axis.title.x = element_blank(),
+    panel.grid.major.y = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    )
+  )
+
+```
+
+## Distributions
+
+```{r stimuli-plot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of stimuli translations, plotted in different panels for English translations of Catalan stimuli (cat-ENG), English translations of Spanish stimuli (spa-ENG), and Spanish translation of Catalan stimuli (cat-SPA).
+A\\) Lexical frequency of stimuli translations, expressed in counts per million and as Zipf scores.
+B\\) Number of higher-frequency phonological neighbourhoods (PTHN) of stimuli translations.
+C\\) Normalised Levenshtein phonological similarity between the presented stimuli and their translations, expressed in percentages.
+Lexical frequencies and PTHN scores of the stimuli (presented auditorily in Catalan or Spanish) are not described, as participants lack any lexical representation of such words."}
+
+medians <- stimuli %>% 
+  group_by(group) %>% 
+  summarise(
+    pthn_median = median(pthn),
+    pthn_sd = sd(pthn),
+    frequency_zipf_median = median(frequency_zipf),
+    frequency_zipf_sd = sd(frequency_zipf),
+    lv_median = median(lv),
+    lv_sd = sd(lv)
+  ) %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  ))
+
+# frequency
+stimuli %>% 
+  mutate(group = case_when(
+    group=="ENG-CAT" ~ "cat-ENG",
+    group=="ENG-SPA" ~ "spa-ENG",
+    group=="SPA-CAT" ~ "cat-SPA",
+  )) %>% 
+  ggplot(aes(frequency_zipf, fill = group)) + 
+  facet_wrap(~group, ncol = 1) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = frequency_zipf_median-frequency_zipf_sd,
+      xmax = frequency_zipf_median+frequency_zipf_sd, 
+      ymin = -Inf,
+      ymax = Inf, 
+      fill = group
+    ),
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20, 
+    colour = "white"
+  ) +
+  geom_vline(
+    data = medians, 
+    aes(
+      xintercept = frequency_zipf_median,
+      group = group
+    ), 
+    colour = "black",
+    size = 0.5,
+    linetype = "dashed"
+  ) + 
+  geom_text(
+    data = medians, 
+    aes(
+      label = "Median ± SD", 
+      x = frequency_zipf_median+1.25,
+      y = 13
+    ), 
+    size = 3
+  ) +
+  geom_curve(
+    data = medians, 
+    aes(
+      x = frequency_zipf_median+0.85, 
+      xend = frequency_zipf_median+0.1,
+      y = 12, 
+      yend = 11
+    ),
+    arrow = arrow(length = unit(0.2, "cm")),
+    curvature = -0.25,
+    colour = "grey30"
+  ) +
+  labs(
+    x = "Lexical frequency (Zipf)", 
+    y = "Number of trials", 
+    title = "Frequency\n(stimuli translations)\n"
+  ) +
+
+  # PTHN
+  stimuli %>% 
+  count(
+    group, 
+    pthn
+  ) %>% 
+  mutate(
+    group = case_when(
+      group=="ENG-CAT" ~ "cat-ENG",
+      group=="ENG-SPA" ~ "spa-ENG",
+      group=="SPA-CAT" ~ "cat-SPA",
+    )) %>% 
+  ggplot() + 
+  aes(
+    pthn, 
+    n,
+    fill = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0, 
+      xmax = pthn_median+pthn_sd, 
+      ymin = -Inf, 
+      ymax = Inf,
+      fill = group
+    ),
+    inherit.aes = FALSE, 
+    alpha = 0.15
+  ) +   
+  geom_col(colour = "white") + 
+  geom_vline(
+    data = medians, 
+    aes(
+      xintercept = pthn_median, 
+      group = group
+    ), 
+    colour = "black", 
+    size = 0.5,
+    linetype = "dashed"
+  ) + 
+  labs(
+    x = "PTHN", 
+    y = "Number of trials", 
+    title = "Higher-frequency neighbours\n(stimuli translations)\n"
+  ) +
+
+  # Levenshtein
+  stimuli %>% 
+  mutate(
+    group = case_when(
+      group=="ENG-CAT" ~ "cat-ENG",
+      group=="ENG-SPA" ~ "spa-ENG",
+      group=="SPA-CAT" ~ "cat-SPA",
+    )) %>% 
+  ggplot() +
+  aes(
+    lv, 
+    fill = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) + 
+  geom_rect(
+    data = medians, 
+    aes(
+      xmin = 0,
+      xmax = lv_median+lv_sd, 
+      ymin = -Inf,
+      ymax = Inf, 
+      fill = group
+    ), 
+    inherit.aes = FALSE, 
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20,
+    colour = "white"
+  ) +
+  geom_vline(
+    data = medians, 
+    aes(
+      xintercept = lv_median,
+      group = group
+    ), 
+    colour = "black", 
+    size = 0.5, 
+    linetype = "dashed"
+  ) + 
+  scale_x_continuous(labels = percent) +
+  labs(
+    x = "Levenshtein", 
+    y = "Number of trials",
+    title = "Phonological similarity\n(stimuli and stimuli translations)\n"
+  ) +
+
+  plot_layout(
+    guides = "collect"
+  ) &
+  plot_annotation(
+    tag_levels = "A"
+  ) &
+  scale_fill_manual(
+    values = c(
+      "#1A85FF",
+      "#ff2976",
+      "#FFC20A"
+    ),
+    na.translate = FALSE
+  ) &
+  theme(
+    legend.position = "none"
+  )
+
+```
+
+# Participants
+
+```{r participants_coded, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+participants_coded <- participants %>% 
+  mutate(
+    valid_participant = ifelse(
+      valid_participant, 
+      "Valid", 
+      "Not valid"
+    )
+  )
+```
+
+We collected data from `r format(as_date(min(participants$date)), "%B %dth %Y")` to `r format(as_date(max(participants$date)), "%B %dth %Y")`.
+We tested `r length(unique(participants$participant))` participants.
+`r length(unique(participants$participant[participants$group=="ENG-SPA"]))` were English natives tested in Spanish, `r length(unique(participants$participant[participants$group=="ENG-CAT"]))` were English natives tested in Catalan, and `r length(unique(participants$participant[participants$group=="SPA-CAT"]))` were Spanish natives tested in Catalan.
+`r sum(!participants$valid_participant)` failed to meet all the inclusion criteria and were excluded from further analyses.
+The final sample comprised data from `r sum(!participants$valid_participant)` participants, `r sum(participants$valid_participant[participants$group=="ENG-SPA"])` in the ENG-SPA group, `r sum(participants$valid_participant[participants$group=="ENG-CAT"])` in the ENG-CAT group, and `r sum(participants$valid_participant[participants$group=="SPA-CAT"])` in the SPA-CAT group.
+
+Participants were included if^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct.
+This item was phrased as Do you have normal or corrected-to-normal VISION?
+(Yes/No) in English, and as ¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan?
+(Sí/No).
+Surprisingly, the proportion of Spanish participants that reported visual impairment was implausibly large (n = 6, 18.18%).
+This is possibly due to some participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision]:
+
+* Aged 18 to 25 years
+* Did not report being fluent in Catalan, Spanish, or Italian
+* Provided at least 80% of the presented trials
+* Did not report motor, auditory or visual (other than wearing glasses) problems
+
+Trials were considered valid if:
+
+* Participant did not leave a blank response
+* Participant took longer than 10 seconds to respond
+
+## Summary
+
+```{r participants_summary, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+participants_coded %>% 
+  count(
+    group,
+    valid_participant
+  ) %>% 
+  mutate(
+    valid_participant = factor(valid_participant)
+  ) %>% 
+  ggplot() +
+  aes(
+    group, 
+    n, 
+    fill = valid_participant, 
+    order = n
+  ) +
+  geom_col() +
+  labs(
+    x = "Group",
+    y = "# participants", 
+    fill = "Valid?"
+  ) +
+  theme(
+    legend.title = element_blank(),
+    legend.position = "top",
+    axis.title.x = element_blank(),
+    panel.grid.major.y = element_line(
+      linetype = "dotted", 
+      colour = "grey"
+    )
+  )
+
+```
+
+## Valid trials
+
+```{r participants_stimuli, echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+participants_coded %>% 
+  filter(
+    valid_participant=="Valid"
+  ) %>% 
+  select(
+    participant, 
+    group, 
+    n, 
+    n_valid
+  ) %>%
+  mutate(
+    valid_prop = n_valid/n
+  ) %>% 
+  ggplot() +
+  aes(
+    valid_prop,
+    fill = group,
+    colour = group
+  ) +
+  facet_wrap(
+    ~group, 
+    ncol = 1
+  ) +
+  geom_histogram() +
+  geom_rug() +
+  labs(
+    x = "Valid trials (%)", 
+    y = "N participants",
+    fill = "Group", 
+    colour = "Group"
+  ) +
+  scale_x_continuous(
+    limits = c(0, 1), 
+    labels = percent
+  ) +
+  theme(
+    legend.title = element_blank(),
+    legend.position = "none",
+    axis.title.x = element_blank(),
+    panel.grid.major.y = element_line(
+      linetype = "dotted", 
+      colour = "grey"
+    )
+  )
+```
+
+## Demographic information
+
+```{r participants_demo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+participants_valid <- participants_coded %>% 
+  filter(valid_participant=="Valid")
+
+p_sex <- participants_valid %>% 
+  left_join(
+    count(
+      participants_valid, 
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  "\n(N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    group, 
+    fill = sex
+  ) +
+  geom_bar(position = position_fill()) +
+  labs(
+    x = "Group",
+    y = "# participants",
+    fill = "Sex"
+  ) +
+  scale_y_continuous(labels = percent) +
+  theme(
+    legend.position = "top",
+    axis.title.x = element_blank()
+  )
+
+p_age <- participants_valid %>% 
+  left_join(
+    count(
+      participants_valid, 
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  " (N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    as.factor(age), 
+    fill = group, 
+    colour = group
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1
+  ) +
+  geom_bar() +
+  labs(
+    x = "Age (years)",
+    y = "% participants", 
+    fill = "Group", 
+    colour = "Group"
+  ) +
+  theme(
+    legend.position = "none"
+  )
+
+(p_sex + p_age) &
+  plot_layout() &
+  plot_annotation(tag_levels = "A") &
+  theme(
+    legend.title = element_blank(),
+    panel.grid.major.y = element_line(
+      linetype = "dotted", 
+      colour = "grey"
+    )
+  )
+
+```
+
+## Second language
+
+```{r participants_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+participants_valid %>% 
+  count(
+    group,
+    l2, 
+    .drop = FALSE
+  ) %>% 
+  mutate(
+    l2 = case_when(
+      l2=="Alemã\u0081n" ~ "German",
+      l2=="Francã\u2030s" ~ "French",
+      l2=="Cuã\u0081l" ~ "Missing",
+      TRUE ~ l2
+    ),
+  ) %>% 
+  left_join(
+    count(
+      participants_valid,
+      group, 
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  "\n(N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    group,
+    n, 
+    fill = l2
+  ) +
+  facet_wrap(
+    ~group,
+    ncol = 1, 
+    scales = "free_y"
+  ) +
+  geom_col(position = position_fill()) +
+  labs(
+    x = "L2",
+    y = "# participants",
+    fill = "L2"
+  ) +
+  scale_y_continuous(
+    labels = percent
+  ) +
+  coord_flip() +
+  theme(
+    legend.position = "top",
+    legend.title = element_blank(),
+    strip.text = element_blank(),
+    panel.grid = element_blank(),
+    axis.text.x = element_text(angle = 90),
+    axis.title.y = element_blank()
+  )
+
+```
+
+## Language profile proficiency
+
+```{r participants_proficiency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+participants_valid %>% 
+  select(
+    participant,
+    group, 
+    catalan_oral,
+    spanish_oral, 
+    catalan_written,
+    spanish_written
+  ) %>% 
+  mutate_at(
+    vars(
+      catalan_oral,
+      spanish_oral,
+      catalan_written, 
+      spanish_written
+    ),
+    replace_na, 
+    0
+  ) %>% 
+  pivot_longer(
+    c(
+      catalan_oral, 
+      spanish_oral,
+      catalan_written,
+      spanish_written
+    ),
+    names_to = "type", 
+    values_to = "score"
+  ) %>% 
+  separate(
+    type, 
+    c("language", "modality"),
+    sep = "_"
+  ) %>% 
+  mutate_at(
+    vars(language, modality), 
+    str_to_sentence
+  ) %>%
+  left_join(
+    count(
+      participants_valid,
+      group,
+      name = "n_total"
+    )
+  ) %>% 
+  mutate(
+    group = paste0(group,  " (N = ", n_total, ")")
+  ) %>% 
+  ggplot() +
+  aes(
+    as.integer(score), 
+    fill = language
+  ) +
+  facet_grid(modality~group) +
+  geom_bar(
+    position = position_dodge()
+  ) +
+  labs(
+    y = "# participants",
+    x = "Self-reported proficiency (0-5)",
+    fill = "Language"
+  ) +
+  theme(
+    legend.position = "top",
+    panel.grid.major.y = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    )
+  )
+
+```
+
+# Data analysis
+
+We fitted a Bayesian multilevel regression model using the R package `brms` [@burkner2017brms], with correct responses (`correct`, 0 is incorrect response, 1 if correct response) as the response variable with a Bernoulli distribution and a logit link function.
+We modelled the probability of an average participant providing a correct response, conditional to a set of predictors (aka.
+fixed effects) and their interactions:
+
+* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database. Standardised before entering the regression model.
+* Phonological similarity (`lv`, 0-1): Inverse Levenshtein distance between the IPA transcription of the presented word and its correct translation in the target language, calculated using the `stringdist` function of the `stringdist` package [@loo2014stringdist].
+Standardised before entering the regression model.
+* Shared phonological onset (`onset`, Same/Different): Whether the presented and the target word share phonological onset, as judged by experimenters, and sum coded as $Different = -0.5$ and $Same = +0.5$.
+
+We added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model.
+We specified random intercepts slopes for all fixed effects within the `participant` grouping variable.
+
+We performed multiple imputation via predictive mean matching using the `mice` function of the `mice` package [@van2011mice] to impute missing values in the response variable or the predictors.
+
+The model was implemented using the following formula (`fit_3`):
+
+$$\begin{align}
+
+&\textbf{Likelihood}  \\
+y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
+
+&\textbf{Parameters}  \\
+
+logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
+& \beta_{1[p]} ~ Frequency_{i} ~ + \\
+& \beta_{2[p]} ~ PTHN_i ~ + \\
+& \beta_{3[p]} ~ Similarity_i ~ + \\
+& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
+
+\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
+\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
+&& \text{[participant-level coefficients]} \\ \\
+
+&\textbf{Prior}  \\
+
+\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
+\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
+\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
+
+\end{align}$$
+
+The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 iterations each (1,000 warm-ups iterations).
+The output of this model is the approximated posterior distribution of all the parameters estimated, therefore indicating the probability of each combination of values of the parameters, conditional to the observed data.
+
+We compared this (extended model) to other models dropping one predictor at a time using Leave-one-out cross-validation (LOO-CV), using the `loo` and `loo_compare` functions of the `brms` package.
+We compared our extended model against the following models:
+
+* fit_0: `~ 1 + (1 | participant_id) + (1 | translation_id)`
+* fit_1: `~ 1 + frequency_zipf + (1 + frequency_zipf | participant_id) + (1 | translation_id)`
+* fit_2: `~ 1 + frequency_zipf + pthn + (1 + frequency_zipf + pthn | participant_id) + (1 | translation_id)`
+* fit_3: `~ 1 + frequency_zipf + pthn + lv + (1 + frequency_zipf + pthn + lv | participant_id) + (1 | translation_id)`
+* fit_4: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant_id) + (1 | translation_id)`
+
+# Results
+
+```{r responses_coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+responses_coded <- responses %>% 
+  # typos are considered correct responses
+  mutate_at(
+    vars(group), 
+    as.factor
+  ) %>% 
+  # center predictors
+  mutate_at(
+    vars(
+      lv_std, 
+      pthn_std,
+      frequency_zipf_std
+    ),
+    function(x) {
+      scale(
+        x, 
+        center = TRUE, 
+        scale = TRUE
+      )[,1]
+    }
+  ) %>% 
+  # impute missing data
+  mice(
+    m = 5, 
+    print = FALSE,
+    method = "pmm"
+  ) %>% 
+  complete() %>% 
+  as_tibble() %>% 
+  arrange(
+    group,
+    word
+  )
+```
+
+## Model comparison
+
+Leave-one-out cross-validation.
+The more negative is `elpd_diff` (and the larger its magnitude compared to its corresponding standard error), the better it first the data compared to the null model (`~ 1 + (1 | participant)`).
+The model that fits the data the best is the extended model.
+
+```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+model_loos %>% 
+  as.data.frame() %>% 
+  rownames_to_column("model") %>% 
+  arrange(desc(elpd_loo)) %>% 
+  select(
+    model,
+    matches("elpd_loo"),
+    matches("looic"),
+    matches("diff")
+  ) %>%
+  mutate(
+    model = str_replace(
+      model,
+      "fit_", 
+      "Model "
+    )
+  ) %>% 
+  mutate_all(~ifelse(. == 0, NA, .)) %>% 
+  gt(rowname_col = "model") %>% 
+  tab_header(
+    title = md("Table 2"),
+    subtitle = md("Model performance and comparison using leave-one-out cross-validation")
+  ) %>% 
+  tab_source_note(
+    md("*Note*.
+Outcomes of the leave-one-out cross-validation procedure.
+Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
+Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
+  ) %>% 
+  fmt_missing(
+    everything(),
+    missing_text = "-"
+  ) %>% 
+  fmt_number(2:7) %>% 
+  cols_label(
+    model = md("**Model**"),
+    looic = md("***LOO<sub>IC</sub>***"),
+    se_looic = md("***SE*<sub>IC</sub>**"),
+    elpd_loo = md("***LOO<sub>ELPD</sub>***"),
+    se_elpd_loo = md("***SE***"),
+    elpd_diff = md("***LOO<sub>diff</sub>***"),
+    se_diff = md("***SE<sub>diff</sub>***")
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left",
+      weight = "bold",
+      size = "medium"
+    ),
+    locations = cells_title(groups = "title")
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left",
+      size = "medium", 
+      style = "italic"
+    ),
+    locations = cells_title(groups = "subtitle")
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "left",
+      size = "medium"
+    ),
+    locations = cells_source_notes()
+  ) %>% 
+  tab_style(
+    style = list(
+      cell_borders(
+        sides = c("left", "right"),
+        color = "white"
+      ),
+      cell_text(align = "center")),
+    locations = list(
+      cells_body(), 
+      cells_row_groups(),
+      cells_column_spanners(), 
+      cells_stub(), 
+      cells_source_notes(),
+      cells_column_labels(),
+      cells_stubhead()
+    )
+  ) %>% 
+  tab_style(
+    style = list(
+      cell_text(
+        align = "left", 
+        size = "medium"
+      ), 
+      cell_borders(
+        sides = "all", 
+        color = "white"
+      )
+    ),
+    locations = list(
+      cells_stub(), 
+      cells_source_notes()
+    )
+  ) %>% 
+  tab_style(
+    style = cell_text(
+      align = "center",
+      size = "medium", 
+      weight = "normal", 
+      style = "normal"
+    ),
+    locations = list(
+      cells_column_spanners(),
+      cells_column_labels()
+    )
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "all", 
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
+  ) %>% 
+  tab_style(
+    style = cell_borders(
+      sides = "top", 
+      color = "white"
+    ),
+    locations = cells_body(
+      columns = 1:7
+    )
+  ) %>% 
+  as_raw_html()
+
+```
+
+### Fixed coefficients
+
+```{r coefs_fixed_table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+coefs <- fixef(fit_4) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("variable") %>% 
+  clean_names() %>% 
+  mutate_at(
+    vars(estimate:q97_5), 
+    ~ifelse(
+      variable=="Intercept",
+      inv_logit_scaled(.),
+      ./4
+    )
+  )
+
+str_repl <- c(
+  "Intercept" = "Intercept",
+  "frequency_zipf_std" = "Frequency\n[+1 SD]",
+  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
+  "lv_std" = "Levenshtein\n[+1 SD]",
+  "pthn_std" = "PTHN\n[+1 SD]"
+)
+
+coefs %>% 
+  mutate(
+    variable = str_remove(variable, "b_") %>% 
+      str_replace_all(str_repl) %>%  
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein"
+        )
+      )
+  ) %>% 
+  select(-est_error) %>% 
+  gt() %>% 
+  fmt_percent(2:4) %>% 
+  cols_merge(
+    c("q2_5", "q97_5"), 
+    pattern = "[{1}, {2}]"
+  ) %>% 
+  cols_label(
+    variable = md("**Predictor**"),
+    estimate = md("**Mean**"),
+    q2_5 = md("**95\\% CrI**"),
+  ) %>% 
+  tab_source_note(
+    "Estimated posterior distributions of coefficients in Model 6.
+A\\) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Black dots represent the mean of each distribution."
+  ) %>% 
+  cols_align(align = "left") %>% 
+  as_raw_html()
+```
+
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
+A\\) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Black dots represent the mean of each distribution.
+B\\) Participant\\-level coefficient variability.
+The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
+C\\) Correlation between participant\\-level effects.
+The model allowed participant-level coefficients to co\\-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}
+
+coefs <- fixef(fit_4) %>% 
+  as.data.frame() %>% 
+  rownames_to_column("variable") %>% 
+  clean_names() %>% 
+  mutate_at(
+    vars(estimate:q97_5), 
+    ~ifelse(
+      variable=="Intercept", 
+      inv_logit_scaled(.), 
+      ./4
+    )
+  ) %>% 
+  group_split(variable) %>% 
+  set_names(make_clean_names(map(., "variable"))) %>% 
+  map(select, -variable) %>% 
+  map(unlist)
+
+str_repl <- c(
+  "Intercept" = "Intercept",
+  "frequency_zipf_std" = "Frequency\n[+1 SD]",
+  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
+  "lv_std" = "Levenshtein\n[+1 SD]",
+  "pthn_std" = "PTHN\n[+1 SD]"
+)
+
+# fixed effects
+post_fix <- posterior_draws_fixed %>% 
+  filter(str_detect(.variable, "b_")) %>% 
+  mutate(
+    .variable_name = str_remove(.variable, "b_") %>% 
+      str_replace_all(str_repl) %>%  
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein",
+          "Group 1\n[*-ENG vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
+          "Levenshtein \u00d7 Group 1",
+          "Levenshtein \u00d7 Group 2"
+        )
+      ),
+    .value = ifelse(
+      str_detect(.variable, "Intercept"), 
+      inv_logit_scaled(.value),
+      .value/4
+    )
+  ) %>% 
+  arrange(.variable) %>% 
+  ggplot() +
+  aes(
+    .value, 
+    fct_rev(.variable_name)
+  ) +
+  geom_vline(xintercept = 0) +
+  stat_slab(
+    aes(fill = stat(
+      cut_cdf_qi(
+        cdf, 
+        .width = c(.5, .8, .95),
+        labels = percent_format()
+      )
+    )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0,
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#1A85FF", "#9ccaff", "#d2e5fc"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = function(x) percent(round(x, 1))
+  ) +
+  labs(
+    x = "P(Correct)",
+    y = "Posterior probability density", 
+    fill = "CrI"
+  ) +
+  theme(
+    legend.position = c(1, 0.25),
+    legend.direction = "vertical",
+    legend.background = element_rect(fill = "grey95"),
+    legend.key = element_rect(fill = "grey95", colour = "grey95"), 
+    legend.key.height = unit(0.1, "cm"),
+    axis.title.y = element_blank(),
+    legend.text = element_text(size = 7),
+    axis.text.x = element_text(colour = "black"),
+    axis.text.y = element_text(colour = "black", hjust = 1),
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
+  )
+
+# SD of random effects
+post_sd_data <- posterior_draws_fixed %>% 
+  filter(
+    str_detect(.variable, "sd_")
+  ) %>% 
+  mutate(
+    .variable_name = str_remove(
+      .variable, 
+      "sd_participant_id__|sd_translation_id__"
+    ) %>% 
+      str_replace_all(str_repl) %>% 
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein"
+        )
+      ),
+    group = case_when(
+      str_detect(.variable, "participant") ~ "Participants",
+      str_detect(.variable, "translation") ~ "Words"
+    ),
+    .value = ifelse(
+      str_detect(.variable, "Intercept"),
+      inv_logit_scaled(.value),
+      .value/4
+    )
+  )
+
+post_sd <- post_sd_data %>% 
+  filter(group=="Participants") %>% 
+  ggplot() +
+  aes(
+    .value,
+    fct_rev(.variable_name)
+  ) +
+  stat_slab(
+    aes(fill = stat(
+      cut_cdf_qi(
+        cdf, 
+        .width = c(.5, .8, .95),
+        labels = percent_format()
+      )
+    )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0,
+    point_size = 1
+  ) +
+  scale_fill_manual(
+    values = c("#ff2976", "#f2a7c2", "#fcd9e6"),
+    na.translate = FALSE
+  ) +
+  scale_x_continuous(
+    labels = percent,
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By participant", 
+    fill = "CrI"
+  ) +
+  theme(axis.text.x = element_blank()) +
+
+  post_sd_data %>% 
+  filter(group=="Words") %>% 
+  ggplot() +
+  aes(
+    .value, 
+    fct_rev(.variable_name)
+  ) +
+  stat_slab(
+    aes(
+      fill = stat(
+        cut_cdf_qi(
+          cdf, 
+          .width = c(.5, .8, .95),
+          labels = percent_format()
+        )
+      )
+    )
+  ) +
+  stat_pointinterval(
+    .width = 0, 
+    point_size = 1
+  ) +
+  scale_x_continuous(
+    labels = percent, 
+    limits = c(0, 1)
+  ) +
+  labs(
+    x = "SDs in P(Correct)",
+    title = "By item", 
+    fill = "CrI"
+  ) +
+  plot_layout(
+    ncol = 1, 
+    heights = c(0.85, 0.15), 
+    guides = "collect"
+  ) &
+  theme(
+    legend.position = "none",
+    axis.text.y = element_text(hjust = 1),
+    axis.title.x = element_blank(),
+    legend.direction = "vertical",
+    axis.title.y = element_blank(),
+    legend.background = element_rect(fill = "grey95"),
+    legend.key.height = unit(0.1, "cm"),
+    legend.key = element_rect(fill = "grey95", colour = "grey95"),
+    legend.text = element_text(size = 7),
+    axis.ticks.y = element_blank(),
+    panel.grid.major.y = element_line(
+      colour = "grey",
+      size = 0.5
+    )
+  )
+
+# Pearson correlations between random effects
+corr_mat <- VarCorr(fit_4, summary = TRUE)$participant_id$cor 
+corr_names <- rownames(VarCorr(fit_4, summary = TRUE)$participant_id$cor)
+
+corr_mat <- corr_mat %>% 
+  as_tibble() %>% 
+  select(matches("Estimate")) %>% 
+  rename_all(str_remove, "Estimate.") %>% 
+  mutate(term1 = corr_names)
+
+corr_mat[lower.tri(corr_mat)] <- NA
+
+corr_data <- corr_mat %>% 
+  as.data.frame() %>% 
+  pivot_longer(
+    -term1, 
+    names_to = "term2", 
+    values_to = ".value"
+  ) %>% 
+  drop_na(.value)
+
+post_cors <- corr_data %>% 
+  mutate(
+    term1 = factor(
+      term1,
+      levels = names(str_repl), 
+      ordered = TRUE
+    ),
+    term2 = factor(
+      term2,
+      levels = names(str_repl),
+      ordered = TRUE
+    )
+  ) %>% 
+  mutate_at(
+    vars(term1, term2), 
+    str_replace_all, 
+    str_repl
+  ) %>% 
+  mutate(
+    .value_label = case_when(
+      term1==term2 ~ NA_character_,
+      TRUE ~ printnum(.value, gt1 = FALSE)
+    ),
+    .value = case_when(
+      term1==term2 ~ NA_real_,
+      TRUE ~ .value
+    )
+  ) %>% 
+  filter(
+    term1 != "PTHN × Levenshtein", 
+    term2 != "Intercept"
+  ) %>% 
+  ggplot() +
+  aes(
+    fct_inorder(term1), 
+    fct_rev(fct_inorder(term2)),
+    fill = .value
+  ) +
+  geom_tile(na.rm = TRUE) +
+  geom_text(
+    aes(label = .value_label), 
+    size = 3, 
+    na.rm = TRUE
+  ) +
+  labs(
+    x = "Term 1", 
+    y = "Term 2", 
+    fill = "Posterior\ncorrelation"
+  ) +
+  scale_fill_gradient2(
+    low = "#04bf78", 
+    mid = "white", 
+    high = "#FFC20A", 
+    na.value = "white", 
+    limits = c(-0.4, 0.4)
+  ) +
+  scale_x_discrete(position = "top") +
+  coord_equal() +
+  theme(
+    legend.position = c(1, 1),
+    legend.direction = "horizontal",
+    legend.key.height = unit(0.1, "cm"),
+    legend.key.width = unit(0.4, "cm"),
+    legend.justification = c(1, 1),
+    legend.title = element_blank(),
+    legend.text = element_text(size = 7),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank(),
+    axis.ticks = element_blank(),
+    axis.text.x = element_text(size = 8, angle = 90, hjust = 0, vjust = 1),
+    axis.text.y = element_text(size = 8, vjust = 1, hjust = 1)
+  )
+
+post_fix / 
+  (
+    (post_sd | post_cors) +
+      plot_layout(widths = c(0.6, 0.4))
+  ) +
+  plot_layout() +
+  plot_annotation(tag_levels = "A")
+
+```
+
+## Random effects
+
+```{r post_re_participant, echo=FALSE, fig.height=6, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}
+post_re <- posterior_draws_random$participants %>% 
+  mean_qi() %>% 
+  ungroup() %>% 
+  left_join(
+    select(
+      participants,
+      participant_id,
+      group
+    )
+  ) %>% 
+  arrange(
+    group, 
+    participant_id, 
+    .param
+  ) %>% 
+  mutate(
+    .param_name = str_replace_all(
+      .param, 
+      str_repl
+    ) %>%  
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein"
+        )
+      )
+  )
+
+ggplot(
+  post_re, 
+  aes(
+    .value, 
+    sort(as.factor(participant_id)), 
+    xmin = .lower,
+    xmax = .upper,
+    colour = group
+  )
+) +
+  facet_grid(
+    group~.param_name,
+    scales = "free"
+  ) +
+  geom_vline(
+    xintercept = 0, 
+    colour = "grey"
+  ) +
+  geom_errorbarh(
+    height = 0,
+    alpha = 0.75
+  ) +
+  geom_point(
+    size = 0.1,
+    shape = 4
+  ) +
+  labs(
+    x = "Participant-level adjustment",
+    y = "Participant", 
+    colour = "Group",
+    title = "By participant"
+  ) +
+  scale_color_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A"),
+    na.translate = FALSE
+  ) &
+  theme(
+    legend.position = "none",
+    legend.title = element_blank(),
+    panel.grid.major.x = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    ),
+    axis.text.x = element_text(size = 7),
+    axis.text.y = element_blank(),
+    axis.ticks = element_blank()
+  )
+```
+
+```{r post_re_word, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
+fixed_intercept <- fixef(fit_4)[1,c(1, 3, 4)] %>% 
+  inv_logit_scaled()
+
+empirical_accuracy <- responses %>% 
+  group_by(group, translation_id, lv) %>% 
+  summarise(
+    successes = sum(correct, na.rm = TRUE),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    prop = prop_adj(successes, n),
+    prop_se = prop_adj_se(successes, n)
+  )
+
+post_re <- posterior_draws_random$words %>% 
+  ungroup() %>% 
+  arrange(
+    translation_id, 
+    .param
+  ) %>% 
+  mutate(
+    .param_name = str_replace_all(
+      .param, 
+      str_repl
+    ) %>%  
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein"
+        )
+      ),
+    .value = inv_logit_scaled(.value)
+  ) %>% 
+  left_join(
+    empirical_accuracy
+  )
+
+post_re %>% 
+  left_join(
+    select(
+      stimuli,
+      group,
+      translation_id,
+      translation,
+      lv
+    ) %>% 
+      mutate(translation = paste0(translation, " (", percent(lv), ")"))
+  ) %>% 
+  ggplot(
+    aes(
+      x = .value,
+      y = reorder(
+        as.factor(translation), 
+        .value
+      )
+    )
+  ) +
+  facet_wrap(
+    ~group,
+    scales = "free_y"
+  ) +
+  geom_vline(
+    xintercept = fixed_intercept[1],
+    colour = "grey", 
+    size = 1
+  ) +
+  stat_interval(
+    size = 1.5,
+    .width = c(0.5, 0.67, 0.95)
+  ) +
+  geom_point(
+    aes(
+      x = prop
+    ),
+    size = 0.5,
+    shape = 3
+  ) +
+  geom_errorbar(
+    aes(
+      x = prop,
+      xmin = prop-prop_se,
+      xmax = prop+prop_se
+    ),
+    width = 0,
+    size = 0.5
+  ) +
+  labs(
+    x = "Probability of correct response",
+    y = "Translation",
+    colour = "Model prediction (credible interval)",
+    caption = "Credible intervals cover the range of values of the sampling space that contain the true value with a given certainty."
+  ) +
+  scale_color_manual(
+    values = c("#d2e5fc", "#9ccaff", "#1A85FF"),
+    na.translate = FALSE, 
+    labels = paste0(c(95, 67, 50), "%")
+  ) +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
+  theme(
+    legend.position = "top",
+    legend.key = element_rect(
+      fill = NA, 
+      colour = NA
+    ),
+    panel.grid.major.x = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    ),
+    axis.text.x = element_text(size = 7),
+    axis.text.y = element_text(size = 5, hjust = 1),
+    axis.ticks = element_blank(),
+    axis.title = element_text(size = 10),
+    plot.caption.position = "plot",
+    plot.caption = element_text(
+      size = 9, 
+      hjust = 0, 
+      face = "plain",
+      margin = margin(t = unit(3, "cm"))
+    )
+  )
+
+ggsave(
+  here("img", "accuracy.png"), 
+  height = 9,
+  width = 14,
+  dpi = 1000
+)
+```
+
+```{r post_re_word_lv, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
+
+fixed_intercept <- fixef(fit_4)[1,c(1, 3, 4)] %>% 
+  inv_logit_scaled()
+
+empirical_accuracy <- responses %>% 
+  group_by(group, translation_id, lv) %>% 
+  summarise(
+    successes = sum(correct, na.rm = TRUE),
+    n = n(),
+    .groups = "drop"
+  ) %>% 
+  mutate(
+    prop = prop_adj(successes, n),
+    prop_se = prop_adj_se(successes, n)
+  )
+
+post_re <- posterior_draws_random$words %>% 
+  ungroup() %>% 
+  arrange(
+    translation_id, 
+    .param
+  ) %>% 
+  mutate(
+    .param_name = str_replace_all(
+      .param, 
+      str_repl
+    ) %>%  
+      factor(
+        levels = c(
+          "Intercept", 
+          "Frequency\n[+1 SD]",
+          "PTHN\n[+1 SD]", 
+          "Levenshtein\n[+1 SD]", 
+          "PTHN \u00d7 Levenshtein"
+        )
+      ),
+    .value = inv_logit_scaled(.value)
+  ) %>% 
+  left_join(
+    empirical_accuracy
+  )
+
+post_re %>% 
+  left_join(
+    select(
+      stimuli,
+      group,
+      translation_id,
+      translation,
+      lv
+    ) %>% 
+      mutate(translation = paste0(translation, " (", percent(lv), ")"))
+  ) %>% 
+  ggplot(
+    aes(
+      x = .value,
+      y = reorder(
+        as.factor(translation), 
+        lv
+      )
+    )
+  ) +
+  facet_wrap(
+    ~group,
+    scales = "free_y"
+  ) +
+  geom_vline(
+    xintercept = fixed_intercept[1],
+    colour = "grey", 
+    size = 1
+  ) +
+  stat_interval(
+    size = 1.5,
+    .width = c(0.5, 0.67, 0.95)
+  ) +
+  geom_point(
+    aes(
+      x = prop
+    ),
+    size = 0.5,
+    shape = 4
+  ) +
+  geom_errorbar(
+    aes(
+      x = prop,
+      xmin = prop-prop_se,
+      xmax = prop+prop_se
+    ),
+    width = 0,
+    size = 0.5
+  ) +
+  labs(
+    x = "Posterior probability of correct response",
+    y = "Translation (Levenshtein similarity)",
+    colour = "Model prediction (credible interval)",
+    caption = "Credible intervals cover the range of values of the sampling space that contain the true value with a given certainty."
+  ) +
+  scale_color_manual(
+    values = c("#d2e5fc", "#9ccaff", "#1A85FF"),
+    na.translate = FALSE, 
+    labels = paste0(c(95, 67, 50), "%")
+  ) +
+  scale_x_continuous(limits = c(0, 1), labels = percent) +
+  theme(
+    legend.position = "top",
+    legend.key = element_rect(
+      fill = NA, 
+      colour = NA
+    ),
+    panel.grid.major.x = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    ),
+    axis.text.x = element_text(size = 7),
+    axis.text.y = element_text(size = 5, hjust = 1),
+    axis.ticks = element_blank(),
+    axis.title = element_text(size = 10),
+    plot.caption.position = "plot",
+    plot.caption = element_text(
+      size = 9, 
+      hjust = 0, 
+      face = "plain",
+      margin = margin(t = unit(3, "cm"))
+    )
+  )
+
+ggsave(
+  here("img", "accuracy_lv.png"), 
+  height = 9,
+  width = 16,
+  dpi = 1000
+)
+```
+
+## Marginal effects
+
+```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
+A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein).
+We did this across the range of values of the PTHN scores.
+For each simulation, we drew a single sample from the posterior distribution of each coefficient.
+Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
+Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
+The dispersion of the lines indicates the uncertainty of our predictions.
+We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
+
+nd <- expand.grid(
+  lv_std = c(
+    min(fit_4$data$lv_std),
+    0,
+    max(fit_4$data$lv_std)
+  ),
+  frequency_zipf_std = 0,
+  pthn_std = seq(
+    min(fit_4$data$pthn_std, na.rm = TRUE),
+    max(fit_4$data$pthn_std, na.rm = TRUE),
+    by = 0.1
+  )
+)
+
+m <- add_epred_draws(
+  nd, 
+  fit_4, 
+  ndraws = 100, 
+  re_formula = NA
+) %>% 
+  mutate(
+    lv_std = factor(
+      lv_std,
+      levels = c(
+        min(fit_4$data$lv_std),
+        0,
+        max(fit_4$data$lv_std)
+      ),
+      labels = c(
+        paste0("-1 SD (", percent(min(responses$lv, na.rm = TRUE)), ")"),
+        paste0("Mean (", percent(mean(responses$lv, na.rm = TRUE)), ")"),
+        paste0("+1 SD (", percent(max(responses$lv, na.rm = TRUE)), ")")
+      )
+    ) 
+  )
+
+ggplot(m) +
+  aes(
+    pthn_std, 
+    .epred,
+    colour = lv_std,
+    fill = lv_std
+  ) +
+  geom_hline(
+    yintercept = 0.5, 
+    colour = "grey"
+  ) +
+  geom_line(
+    aes(group = interaction(lv_std, .draw)), 
+    size = 0.65,
+    alpha = 0.5,
+    show.legend = TRUE
+  ) +
+  stat_summary(
+    aes(linetype = lv_std), 
+    fun = mean, 
+    geom = "line",
+    colour = "black", 
+    size = 0.75,
+    show.legend = TRUE
+  ) +
+  annotate(
+    "segment", 
+    x = -0.5, 
+    xend = -0.9, 
+    y = 0.85, 
+    yend = 0.85, 
+    arrow = arrow(length = unit(0.1, "cm")), 
+    size = .5
+  ) +
+  annotate(
+    "segment", 
+    x = 2.25, 
+    xend = 2.65, 
+    y = 0.85, 
+    yend = 0.85, 
+    arrow = arrow(length = unit(0.1, "cm")),
+    size = .5
+  ) +
+  annotate(
+    "text", 
+    x = -0.5, 
+    y = 0.95, 
+    size = 2.75,
+    label = "Less similar"
+  ) +
+  annotate(
+    "text", 
+    x = 2.15, 
+    y = 0.95, 
+    size = 2.75,
+    label = "More similar"
+  ) +
+  labs(
+    x = paste0(
+      "PTHN (+1 SD, ", 
+      round(sd(responses$pthn), 1),
+      " higher-frequency neighbours)"
+    ),
+    y = "P(Correct)", 
+    colour = "Levenshtein similarity",
+    fill = "Levenshtein similarity",
+    linetype = "Levenshtein similarity",
+    title = "Population-level posterior predictions"
+  ) +
+  scale_y_continuous(
+    labels = function(x) percent(round(x, 2)),
+    limits = c(0, 1),
+    breaks = seq(0, 1, 0.1)
+  ) +
+  theme(
+    legend.position = "top",
+    plot.title = element_blank()
+  )
+
+```
+
+=======
 ---
@@
 ```
+
+>>>>>>> eb2ad2e01316fbb4878ef163afbb685d3a91ad49

### 2022-04-27 docs/index.Rmd
Restore workflow

--- 2022-04-27 docs/index.Rmd
@@
 
-Participants in the ENG-SPA listened to `r nrow(stimuli[stimuli$group=="ENG-SPA",])` Spanish words, and participants in the ENG-CAT and the SPA-CAT groups listened to `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials.
+Participants in the spa-ENG listened to `r nrow(stimuli[stimuli$group=="spa-ENG",])` Spanish words, and participants in the cat-ENG and the cat-SPA groups listened to `r nrow(stimuli[stimuli$group=="cat-ENG",])` trials.
 Five trials in each condition were practice trials and were excluded from analyses.
@@
 stimuli_coded <- stimuli %>% 
-    mutate(test_language = ifelse(group=="ENG-SPA", "Spanish", "Catalan"))
+  mutate(test_language = ifelse(group=="spa-ENG", "Spanish", "Catalan"))
 ```
@@
     locations = cells_body(columns = 1:9)
-    )
+  ) %>% 
+  as_raw_html()
 
@@
 stimuli_coded %>% 
-    distinct(test_language, word1, .keep_all = TRUE) %>% 
+  distinct(
+    test_language,
+    word1,
+    .keep_all = TRUE
+  ) %>% 
   left_join(count(., test_language, name = "n_total")) %>% 
   mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
-    ggplot(aes(frequency_zipf, fill = test_language)) +
+  ggplot(
+    aes(
+      frequency_zipf, 
+      fill = test_language
+    )
+  ) +
   facet_wrap(~test_language, ncol = 1) +
-    geom_histogram(bins = 25, colour = "white") +
-    labs(x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language")  +
+  geom_histogram(
+    bins = 25,
+    colour = "white"
+  ) +
+  labs(
+    x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", 
+    y = "# trials", 
+    fill = "Test language"
+  )  +
   scale_y_continuous(breaks = seq(0, 12, 2)) +
@@
   mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
-    ggplot(aes(group, lv, fill = group, colour = group)) +
+  ggplot() +
+  aes(
+    x = group, 
+    y = lv,
+    fill = group,
+    colour = group
+  ) +
   geom_violin() +
-    geom_point(shape = 1, stroke = 1, position = position_jitter(width = 0.1), colour = "black", alpha = 0.25) +
-    geom_boxplot(colour = "black", fill = "white", width = 0.1, size = 0.75) +
-    labs(x = "Group", y = "Phonological Leveshtein distance\n(corrected for string length)", fill = "Test language") +
-    scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
+  geom_point(
+    shape = 1,
+    stroke = 1, 
+    position = position_jitter(width = 0.1), 
+    colour = "black", 
+    alpha = 0.25
+  ) +
+  geom_boxplot(
+    colour = "black", 
+    fill = "white",
+    width = 0.1, 
+    size = 0.75
+  ) +
+  labs(
+    x = "Group", 
+    y = "Phonological Leveshtein distance\n(corrected for string length)",
+    fill = "Test language"
+  ) +
+  scale_y_continuous(
+    labels = scales::percent
+  ) +
   theme(
@@
     axis.title.x = element_blank(),
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+    panel.grid.major.y = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    )
   )
@@
     lv_sd = sd(lv)
-    ) %>% 
-    mutate(group = case_when(
-        group=="ENG-CAT" ~ "cat-ENG",
-        group=="ENG-SPA" ~ "spa-ENG",
-        group=="SPA-CAT" ~ "cat-SPA",
-    ))
+  )
 
@@
 stimuli %>% 
-    mutate(group = case_when(
-        group=="ENG-CAT" ~ "cat-ENG",
-        group=="ENG-SPA" ~ "spa-ENG",
-        group=="SPA-CAT" ~ "cat-SPA",
-    )) %>% 
-    ggplot(aes(frequency_zipf, fill = group)) + 
+  ggplot() + 
+  aes(
+    x = frequency_zipf,
+    fill = group
+  ) + 
   facet_wrap(~group, ncol = 1) + 
@@
     aes(
-            xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd, 
-            ymin = -Inf, ymax = Inf, fill = group
-        ), inherit.aes = FALSE, alpha = 0.15
-    ) +   
-    geom_histogram(bins = 20, colour = "white") +
+      xmin = frequency_zipf_median-frequency_zipf_sd,
+      xmax = frequency_zipf_median+frequency_zipf_sd, 
+      ymin = -Inf,
+      ymax = Inf, 
+      fill = group
+    ),
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20,
+    colour = "white"
+  ) +
   geom_vline(
     data = medians, 
-        aes(xintercept = frequency_zipf_median, group = group), 
+    aes(
+      xintercept = frequency_zipf_median,
+      group = group
+    ), 
     colour = "black",
@@
   count(group, pthn) %>% 
-    mutate(group = case_when(
-        group=="ENG-CAT" ~ "cat-ENG",
-        group=="ENG-SPA" ~ "spa-ENG",
-        group=="SPA-CAT" ~ "cat-SPA",
-    )) %>% 
-    ggplot(aes(pthn, n, fill = group)) + 
+  ggplot(
+    aes(
+      x = pthn, 
+      y = n,
+      fill = group
+    )
+  ) + 
   facet_wrap(~group, ncol = 1) + 
@@
       fill = group
-        ), inherit.aes = FALSE, 
+    ), 
+    inherit.aes = FALSE, 
     alpha = 0.15
@@
   geom_vline(
-        data = medians, aes(xintercept = pthn_median, group = group), 
-        colour = "black", size = 0.5, linetype = "dashed"
+    data = medians, 
+    aes(
+      xintercept = pthn_median,
+      group = group
+    ), 
+    colour = "black", 
+    size = 0.5,
+    linetype = "dashed"
   ) + 
@@
   stimuli %>% 
-    mutate(group = case_when(
-        group=="ENG-CAT" ~ "cat-ENG",
-        group=="ENG-SPA" ~ "spa-ENG",
-        group=="SPA-CAT" ~ "cat-SPA",
-    )) %>% 
-    ggplot(aes(lv, fill = group)) +
+  ggplot() +
+  aes(lv, fill = group) +
   facet_wrap(~group, ncol = 1) + 
@@
     aes(
-            xmin = 0, xmax = lv_median+lv_sd, 
-            ymin = -Inf, ymax = Inf, fill = group
-        ), inherit.aes = FALSE, alpha = 0.15
-    ) +   
-    geom_histogram(bins = 20, colour = "white") +
+      xmin = 0,
+      xmax = lv_median+lv_sd, 
+      ymin = -Inf,
+      ymax = Inf, fill = group
+    ),
+    inherit.aes = FALSE,
+    alpha = 0.15
+  ) +   
+  geom_histogram(
+    bins = 20,
+    colour = "white"
+  ) +
   geom_vline(
-        data = medians, aes(xintercept = lv_median, group = group), 
-        colour = "black", size = 0.5, linetype = "dashed"
+    data = medians,
+    aes(
+      xintercept = lv_median, 
+      group = group
+    ), 
+    colour = "black",
+    size = 0.5,
+    linetype = "dashed"
   ) + 
   scale_x_continuous(labels = percent) +
-    labs(x = "Levenshtein", y = "Number of trials", title = "Phonological similarity\n(stimuli and stimuli translations)\n") +
+  labs(
+    x = "Levenshtein",
+    y = "Number of trials",
+    title = "Phonological similarity\n(stimuli and stimuli translations)\n"
+  ) +
 
@@
   plot_annotation(tag_levels = "A") &
-    scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
+  scale_fill_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A"), 
+    na.translate = FALSE
+  ) &
   theme(legend.position = "none")
@@
 We collected data from `r format(as_date(min(participants$date)), "%B %dth %Y")` to `r format(as_date(max(participants$date)), "%B %dth %Y")`.
-We tested `r length(unique(participants$participant))` participants.
-`r length(unique(participants$participant[participants$group=="ENG-SPA"]))` were English natives tested in Spanish, `r length(unique(participants$participant[participants$group=="ENG-CAT"]))` were English natives tested in Catalan, and `r length(unique(participants$participant[participants$group=="SPA-CAT"]))` were Spanish natives tested in Catalan.
+We tested `r length(unique(participants$participant_id))` participants.
+`r length(unique(participants$participant_id[participants$group=="spa-ENG"]))` were English natives tested in Spanish, `r length(unique(participants$participant_id[participants$group=="cat-ENG"]))` were English natives tested in Catalan, and `r length(unique(participants$participant_id[participants$group=="cat-SPA"]))` were Spanish natives tested in Catalan.
 `r sum(!participants$valid_participant)` failed to meet all the inclusion criteria and were excluded from further analyses.
-The final sample comprised data from `r sum(!participants$valid_participant)` participants, `r sum(participants$valid_participant[participants$group=="ENG-SPA"])` in the ENG-SPA group, `r sum(participants$valid_participant[participants$group=="ENG-CAT"])` in the ENG-CAT group, and `r sum(participants$valid_participant[participants$group=="SPA-CAT"])` in the SPA-CAT group.
+The final sample comprised data from `r sum(!participants$valid_participant)` participants, `r sum(participants$valid_participant[participants$group=="spa-ENG"])` in the spa-ENG group, `r sum(participants$valid_participant[participants$group=="cat-ENG"])` in the cat-ENG group, and `r sum(participants$valid_participant[participants$group=="cat-SPA"])` in the cat-SPA group.
 
@@
 participants_coded %>% 
-    count(group, valid_participant) %>% 
+  count(
+    group, 
+    valid_participant
+  ) %>% 
   mutate(valid_participant = factor(valid_participant)) %>% 
-    ggplot(aes(group, n, fill = valid_participant, order = n)) +
+  ggplot() +
+  aes(
+    group,
+    n, 
+    fill = valid_participant,
+    order = n
+  ) + 
   geom_col() +
-    labs(x = "Group", y = "# participants", fill = "Valid?") +
+  labs(
+    x = "Group", 
+    y = "# participants", 
+    fill = "Valid?"
+  ) +
   theme(
@@
     axis.title.x = element_blank(),
-        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
+    panel.grid.major.y = element_line(
+      linetype = "dotted", 
+      colour = "grey"
+    )
   )
@@
   filter(valid_participant=="Valid") %>% 
-    select(participant, group, n, n_valid) %>%
+  select(
+    participant_id, 
+    group, 
+    n,
+    n_valid
+  ) %>%
   mutate(valid_prop = n_valid/n) %>% 
-    ggplot(aes(valid_prop, fill = group, colour = group)) +
+  ggplot() +
+  aes(
+    x = valid_prop, 
+    fill = group,
+    colour = group
+  ) + 
   facet_wrap(~group, ncol = 1) +
@@
   geom_rug() +
-    labs(x = "Valid trials (%)", y = "N participants", fill = "Group", colour = "Group") +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
+  labs(
+    x = "Valid trials (%)",
+    y = "N participants",
+    fill = "Group",
+    colour = "Group"
+  ) +
+  scale_x_continuous(labels = percent) +
   theme(
@@
 ```{r participants_demo, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-participants_valid <- filter(participants_coded, valid_participant=="Valid")
+participants_valid <- filter(
+  participants_coded,
+  valid_participant=="Valid"
+)
 
@@
   mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
-    ggplot(aes(group, fill = sex)) +
+  ggplot() +
+  aes(
+    x = group, 
+    fill = sex
+  ) +
   geom_bar(position = position_fill()) +
-    labs(x = "Group", y = "# participants", fill = "Sex") +
+  labs(
+    x = "Group", 
+    y = "# participants", 
+    fill = "Sex"
+  ) +
   scale_y_continuous(labels = percent) +
@@
   mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
-    ggplot(aes(as.factor(age), fill = group, colour = group)) +
+  ggplot(
+    aes(
+      as.factor(age),
+      fill = group,
+      colour = group
+    )
+  ) +
   facet_wrap(~group, ncol = 1) +
-    geom_bar() +
-    labs(x = "Age (years)", y = "% participants", fill = "Group", colour = "Group") +
+  geom_bar(colour = "white") +
+  labs(
+    x = "Age (years)", 
+    y = "% participants", 
+    fill = "Group", 
+    colour = "Group"
+  ) +
   theme(
@@
     legend.title = element_blank(),
-        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
+    panel.grid.major.y = element_line(
+      linetype = "dotted", 
+      colour = "grey"
+    )
   )
@@
   ) %>% 
-    left_join(count(participants_valid, group, name = "n_total")) %>% 
+  left_join(
+    count(
+      participants_valid, 
+      group, 
+      name = "n_total"
+      )
+    ) %>% 
   mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
-    ggplot(aes(group, n, fill = l2)) +
-    facet_wrap(~group, ncol = 1, scales = "free_y") +
+  ggplot() +
+  aes(
+    group,
+    n,
+    fill = l2
+  ) + 
+  facet_wrap(
+    ~group, 
+    ncol = 1,
+    scales = "free_y"
+  ) +
   geom_col(position = position_fill()) +
-    labs(x = "L2", y = "# participants", fill = "L2") +
+  labs(
+    x = "L2", 
+    y = "# participants", 
+    fill = "L2"
+  ) +
   scale_y_continuous(labels = percent) +
@@
 participants_valid %>% 
-    select(participant, group, catalan_oral, spanish_oral, catalan_written, spanish_written) %>% 
-    mutate_at(vars(catalan_oral, spanish_oral, catalan_written, spanish_written), replace_na, 0) %>% 
-    pivot_longer(c(catalan_oral, spanish_oral, catalan_written, spanish_written),
-                 names_to = "type", values_to = "score") %>% 
-    separate(type, c("language", "modality"), sep = "_") %>% 
-    mutate_at(vars(language, modality), str_to_sentence) %>%
-    left_join(count(participants_valid, group, name = "n_total")) %>% 
+  select(
+    participant_id, 
+    group,
+    catalan_oral, 
+    spanish_oral, 
+    catalan_written,
+    spanish_written
+  ) %>% 
+  mutate_at(
+    vars(
+      catalan_oral,
+      spanish_oral,
+      catalan_written,
+      spanish_written
+    ), 
+    replace_na, 
+    0
+  ) %>% 
+  pivot_longer(
+    c(
+      catalan_oral,
+      spanish_oral, 
+      catalan_written, 
+      spanish_written
+    ),
+    names_to = "type",
+    values_to = "score"
+  ) %>% 
+  separate(
+    type, 
+    c("language", "modality"), 
+    sep = "_"
+  ) %>% 
+  mutate_at(
+    vars(language, modality), 
+    str_to_sentence
+  ) %>%
+  left_join(
+    count(
+      participants_valid,
+      group, 
+      name = "n_total"
+      )
+    ) %>% 
   mutate(group = paste0(group, " (N = ", n_total, ")")) %>% 
-    ggplot(aes(as.integer(score), fill = language)) +
+  ggplot(
+    aes(
+      as.integer(score), 
+      fill = language)
+  ) +
   facet_grid(modality~group) +
-    geom_bar(position = position_dodge()) +
-    labs(y = "# participants", x = "Self-reported proficiency (0-5)", fill = "Language") +
+  geom_bar(
+    position = position_dodge(),
+    colour = "white"
+  ) +
+  labs(
+    y = "# participants", 
+    x = "Self-reported proficiency (0-5)",
+    fill = "Language"
+  ) +
   theme(
     legend.position = "top",
-        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+    panel.grid.major.y = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    )
   )
@@
   # impute missing data
-    mice(m = 5, print = FALSE, method = "pmm") %>% 
-    complete() %>% 
   as_tibble() %>% 
@@
   arrange(desc(elpd_loo)) %>% 
-    select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
+  select(
+    model,
+    matches("elpd_loo"),
+    matches("looic"),
+    matches("diff")
+  ) %>%
   mutate(model = str_replace(model, "fit_", "Model ")) %>% 
@@
   tab_style(
-        style = list(cell_borders(sides = c("left", "right"), color = "white"),
-                     cell_text(align = "center")),
+    style = list(
+      cell_borders(
+        sides = c("left", "right"), 
+        color = "white"
+      ),
+      cell_text(
+        align = "center"
+      )
+    ),
     locations = list(
-            cells_body(), cells_row_groups(), cells_column_spanners(), 
-            cells_stub(), cells_source_notes(),
-            cells_column_labels(), cells_stubhead()
+      cells_body(), 
+      cells_row_groups(), 
+      cells_column_spanners(), 
+      cells_stub(),
+      cells_source_notes(),
+      cells_column_labels(),
+      cells_stubhead()
     )
@@
   tab_style(
-        style = list(cell_text(align = "left", size = "medium"), 
-                     cell_borders(sides = "all", color = "white")),
-        locations = list(cells_stub(), cells_source_notes())
+    style = list(
+      cell_text(align = "left", size = "medium"), 
+      cell_borders(sides = "all", color = "white")
+    ),
+    locations = list(
+      cells_stub(), 
+      cells_source_notes()
+    )
   ) %>% 
   tab_style(
-        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-        locations = list(cells_column_spanners(), cells_column_labels())
+    style = cell_text(
+      align = "center", 
+      size = "medium",
+      weight = "normal",
+      style = "normal"
+    ),
+    locations = list(
+      cells_column_spanners(),
+      cells_column_labels()
+    )
   ) %>% 
@@
   clean_names() %>% 
-    mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4))
+  mutate_at(
+    vars(estimate:q97_5), 
+    ~ifelse(
+      variable=="Intercept",
+      inv_logit_scaled(.),
+      ./4
+    )
+  )
 
@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-    "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
-    "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv_std:groupeng_spa" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupcateng_spaeng" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
   "pthn_std" = "PTHN\n[+1 SD]",
-    "groupENGvs.SPA" =  "Group 1\n[ENG-* vs. SPA-CAT]",
-    "groupENGMCATvs.ENGMSPA" = "Group 2\n[ENG-CAT vs. ENG-SPA]"
+  "groupeng_spa" =  "Group 1\n[ENG-* vs. cat-SPA]",
+  "groupcateng_spaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
           "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-                    "Group 1\n[ENG-* vs. SPA-CAT]",
-                    "Group 2\n[ENG-CAT vs. ENG-SPA]",
+          "Group 1\n[ENG-* vs. cat-SPA]",
+          "Group 2\n[cat-ENG vs. spa-ENG]",
           "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
@@
   fmt_percent(2:4) %>% 
-    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
+  cols_merge(
+    c("q2_5", "q97_5"), 
+    pattern = "[{1}, {2}]"
+  ) %>% 
   cols_label(
@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-    "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
-    "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv_std:groupeng_spa" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupcateng_spaeng" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
   "pthn_std" = "PTHN\n[+1 SD]",
-    "groupENGvs.SPA" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-    "groupENGMCATvs.ENGMSPA" = "Group 2\n[cat-ENG vs. spa-ENG]"
+  "groupeng_spa" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+  "groupcateng_spaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
 # SD of random effects
-post_sd_data <- post_draws 
-filter(str_detect(.variable, "sd_")) 
+post_sd_data <- post_draws %>% 
+  filter(str_detect(.variable, "sd_")) %>% 
   mutate(
@@
       "sd_participant__|sd_word__"
-    ),
+    ) %>% 
       str_replace_all(str_repl) %>% 
@@
 
-## Random effects
-
-```{r post_re_participant, echo=FALSE, fig.height=6, fig.width=9, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-nd_re_participants <- expand.grid(
-    participant = unique(fit$data$participant),
-    lv_std = c(-1, 0, 1),
-    frequency_zipf_std = 0,
-    pthn_std = seq(
-        min(fit$data$pthn_std, na.rm = TRUE),
-        max(fit$data$pthn_std, na.rm = TRUE),
-        by = 0.1
-    ),
-    group = NA
-)
-
-post_re <- add_epred_draws(
-    nd_re_participants, 
-    fit, 
-    re_formula = ~ (1 + frequency_zipf_std + pthn_std*lv_std | participant),
-    ndraws = 50
-)  %>% 
-    mean_qi() %>% 
-    ungroup() %>% 
-    left_join(select(participants, participant, group)) %>% 
-    arrange(group, participant, .param) %>% 
-    mutate(.param_name = str_replace_all(.param, str_repl) %>%  
-               factor(
-                   levels = c(
-                       "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-                       "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-                       "Group 1\n[ENG-* vs. SPA-CAT]",
-                       "Group 2\n[ENG-CAT vs. ENG-SPA]",
-                       "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
-                   )
-               ))
-
-ggplot(post_re, 
-       aes(.value, sort(as.factor(participant)), 
-           xmin = .lower, xmax = .upper, colour = group)) +
-    facet_grid(group~.param_name, scales = "free") +
-    geom_vline(xintercept = 0, colour = "grey") +
-    geom_errorbarh(height = 0, alpha = 0.75) +
-    geom_point(size = 0.1, shape = 4) +
-    labs(
-        x = "Participant-level adjustment",
-        y = "Participant", 
-        colour = "Group",
-        title = "By participant"
-    ) +
-    scale_color_manual(
-        values = c("#1A85FF", "#ff2976", "#FFC20A"),
-        na.translate = FALSE
-    ) &
-    theme(
-        legend.position = "none",
-        legend.title = element_blank(),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_blank(),
-        axis.ticks = element_blank()
-    )
-```
-
 ```{r post_re_word, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
-fixed_intercept <- fixef(fit_6)[1,c(1, 3, 4)] %>% 
+fixed_intercept <- fixef(fit_6)[1 ,c(1, 3, 4)] %>% 
   inv_logit_scaled()
@@
 nd_re_words <- expand.grid(
-    word = unique(fit_6$data$word),
+  translation_id = unique(fit_6$data$translation_id),
   pthn_std = c(-1, 1),
@@
   fit_6, 
-    re_formula = ~ (1| word),
+  re_formula = ~ (1 | translation_id),
   ndraws = 50
 ) %>% 
-    ungroup()
+  ungroup() %>% 
+  select(-group) %>% 
+  left_join(select(stimuli, group, translation, translation_id))
 
@@
     .epred,
-        reorder(as.factor(word), .epred)
-    ) +
+    reorder(as.factor(translation), .epred)
+  ) +
+  facet_wrap(~group, scales = "free_y") +
   geom_vline(
@@
   ) +
-    annotate(
-        geom = "text",
-        x = 0.7,
-        y = 100,
-        size = 3,
-        hjust = 0.5,
-        label = "Item accuracy \n(95%, 67%, and 50% CrI \nof item-level intercepts)"
-    ) +
-    annotate(
-        geom = "rect",
-        alpha = 0.15,
-        xmin = fixed_intercept[2],
-        xmax = fixed_intercept[3],
-        ymin = -Inf,
-        ymax = Inf
-    ) +
-    annotate(
-        x = 0.7,
-        xend = 0.6,
-        y = 108,
-        yend = 130,
-        geom = "curve",
-        curvature = 0.25,
-        arrow = arrow(length = unit(0.2, "cm"))
-    ) +
-    annotate(
-        geom = "text",
-        x = 0.8,
-        y = 10,
-        size = 3,
-        hjust = 0.5,
-        label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
-    ) +
-    annotate(
-        x = 0.7,
-        xend = fixed_intercept[1]+0.05,
-        y = 17,
-        yend = 30,
-        geom = "curve",
-        curvature = 0.2,
-        arrow = arrow(length = unit(0.2, "cm"))
-    ) +
+  # annotate(
+  #   geom = "text",
+  #   x = 0.7,
+  #   y = 100,
+  #   size = 3,
+  #   hjust = 0.5,
+  #   label = "Item accuracy \n(95%, 67%, and 50% CrI \nof item-level intercepts)"
+  # ) +
+  # annotate(
+  #   geom = "rect",
+  #   alpha = 0.15,
+#   xmin = fixed_intercept[2],
+#   xmax = fixed_intercept[3],
+#   ymin = -Inf,
+#   ymax = Inf
+# ) +
+# annotate(
+#   x = 0.7,
+#   xend = 0.6,
+#   y = 108,
+#   yend = 130,
+#   geom = "curve",
+#   curvature = 0.25,
+#   arrow = arrow(length = unit(0.2, "cm"))
+# ) +
+# annotate(
+#   geom = "text",
+#   x = 0.8,
+#   y = 10,
+#   size = 3,
+#   hjust = 0.5,
+#   label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
+# ) +
+# annotate(
+#   x = 0.7,
+#   xend = fixed_intercept[1]+0.05,
+#   y = 17,
+#   yend = 30,
+#   geom = "curve",
+#   curvature = 0.2,
+#   arrow = arrow(length = unit(0.2, "cm"))
+# ) +
 labs(
@@
   ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
+  scale_x_continuous(
+    limits = c(0, 1), 
+    labels = percent
+  ) +
   theme(
@@
     legend.key = element_rect(fill = NA, colour = NA),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    panel.grid.major.x = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    ),
     axis.text.x = element_text(size = 7),
@@
     plot.caption.position = "plot",
-        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm")))
+    plot.caption = element_text(
+      size = 9,
+      hjust = 0, 
+      face = "plain", 
+      margin = margin(
+        t = unit(3, "cm")
+      )
+    )
   )
@@
     legend.justification = "left",
-        legend.key = element_rect(fill = NA, colour = NA),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    legend.key = element_rect(
+      fill = NA, 
+      colour = NA
+    ),
+    panel.grid.major.x = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    ),
     axis.text.x = element_text(size = 7),
@@
     plot.caption.position = "plot",
-        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm")))
+    plot.caption = element_text(
+      size = 9, 
+      hjust = 0,
+      face = "plain", 
+      margin = margin(t = unit(3, "cm"))
+    )
   )
@@
   ) +
+  scale_colour_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A"),
+    na.translate = FALSE
+  ) +
   scale_y_continuous(
@@
     ),
-        group = ifelse(test_language=="Catalan", "ENG-CAT", "ENG-SPA")
+    group = ifelse(test_language=="Catalan", "cat-ENG", "spa-ENG")
   ) %>% 
@@
 familiarity <- stimuli %>% 
-    filter(group != "SPA-CAT") %>% 
+  filter(group != "cat-SPA") %>%
   mutate(
@@
     test_language = case_when(
-            group=="ENG-CAT" ~ "Catalan",
-            group=="ENG-SPA" ~ "Spanish",
-            group=="SPA-CAT" ~ "Catalan"
+      group=="cat-ENG" ~ "Catalan",
+      group=="spa-ENG" ~ "Spanish",
+      group=="cat-SPA" ~ "Catalan"
     ),
@@
   ) %>% 
-    left_join(select(word_accuracy, group, stim, prop, prop_se)) %>% 
+  left_join(
+    select(
+      word_accuracy, 
+      group,
+      stim,
+      prop,
+      prop_se
+    )
+  ) %>% 
   group_by(group) %>% 
   mutate(
-        word_reorder = reorder_within(word, frequency_zipf, group)
+    word_reorder = reorder_within(
+      word, 
+      frequency_zipf, 
+      group
+    )
   ) %>% 
@@
   ) +
+  scale_colour_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A", "black"),
+    na.translate = FALSE
+  ) +
   scale_y_reordered() +
@@
     panel.border = element_rect(fill = NA, colour = "grey"),
-        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+    panel.grid.major.x = element_line(
+      colour = "grey", 
+      linetype = "dotted"
+    ),
     legend.title = element_blank(),
@@
 familiarity %>% 
+  filter(group != "cat-SPA") %>% 
   ggplot() +
@@
   ) +
-    scale_x_continuous(
-        limits = c(2, 6)
-    ) +
-    scale_y_continuous(
-        limits = c(0, 1),
-        labels = percent
-    ) +
+  scale_colour_manual(
+    values = c("#1A85FF", "#ff2976", "#FFC20A", "black"),
+    na.translate = FALSE
+  ) +
+  # scale_x_continuous(
+  #   limits = c(2, 6)
+  # ) +
+  # scale_y_continuous(
+  #   limits = c(0, 1),
+  #   labels = percent
+  # ) +
   theme(
@@
 
-ggsave(here("img", "problematic_frequency-accuracy.png"), height = 4, width = 10)
-
-```
+ggsave(
+  here("img", "problematic_frequency-accuracy.png"), 
+  height = 4, 
+  width = 10
+)
+
+```

### 2022-04-27 manuscript/manuscript.Rmd
Restore workflow

--- 2022-04-27 manuscript/manuscript.Rmd
@@
 One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, imagine a Spanish native, with no previous familiarity with any other language, listens for the first time to French may encounter the word /pɔʁt/ (*porte*), translation of *door* in French.
-The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish.
+For example, imagine a Spanish native, with no previous familiarity with any other language, listens for the first time to French may encounter the word p`r knitr::asis_output("\U0254")``r knitr::asis_output("\U0281")`t/ (*porte*), translation of *door* in French.
+The voiced uvular fricative consonant /`r knitr::asis_output("\U0281")`/ and the open-mid back rounded vowel /`r knitr::asis_output("\U0254")`/ do not exist in Spanish.
 These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /'pwer.ta/ (*puerta*).
 The recognition of this French, unfamiliar word relies almost entirely on its similarity with its Spanish translation.
-Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
-This time, phonological similarity is of little help: /dɔː/ and /'pwer.ta/ share very few phonemes.
+Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /d`r knitr::asis_output("\U0254")`ː/ (*door*).
+This time, phonological similarity is of little help: /d`r knitr::asis_output("\U0254")`ː/ and /'pwer.ta/ share very few phonemes.
 It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations (which they lack, they are unfamiliar to English).
@@
 Many studies use symbolic representations to compute the simiarity between word-forms of interest in either modality.
-Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (*radio*-*radio*); phonological word-forms are represented by their  phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA notation) [see @floccia2018introduction for example].
+Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (*radio*-*radio*); phonological word-forms are represented by their  phonological transcriptions (`r knitr::asis_output("\U0279")`eɪdiəʊ-raðjo, in IPA notation) [see @floccia2018introduction for example].
 The similarity between each pair of word-forms is operationalised numerically by applying some algorithm that computes the edit distance between their symbolic representation.
 The Leveshtein distance is one of the most widely known algorithms: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string [@levenshtein1966binary], and has been successfully used to estimate the similarity of cross-language word-pairs [@gooskens2004perceptive; @schepens2012distributions].
-Using our previous examples, the Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio")`: both strings are identical.
-But the distance between the phonological transcriptions ɹeɪdiəʊ and raðjo is `r stringdist::stringdist("ɹeɪdiəʊ", "raðjo")`, which is the number of edits that separates both character strings.
-There is also an additional layer of complexity: the phonological transcriptions /ˈɹeɪdiˌəʊ/ and /ˈraðjo/ can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.
+Using our previous examples, the Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio", useBytes = TRUE)`: both strings are identical.
+But the distance between the phonological transcriptions `r knitr::asis_output("\U0279")`eɪdiə`r knitr::asis_output("\U028A")` and ra`r knitr::asis_output("\U00F0")`jo is 7, which is the number of edits that separates both character strings.
+There is also an additional layer of complexity: the phonological transcriptions /ˈ`r knitr::asis_output("\U0279")`eɪdiˌə`r knitr::asis_output("\U028A")`/ and /ˈra`r knitr::asis_output("\U00F0")`jo/ can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.
 whether to include elements of aspiration, nasalisation, etc.) can affect how accurate these symbolic representation capture all of the features that are present in the acoustic signal when either word is spoken.
@@
 Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`.
-We collected data from `r length(unique(participants$participant))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
-`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant[participants$group %in% c("ENG-CAT", "ENG-SPA") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant[participants$group %in% c("SPA-CAT")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant[participants$group %in% c("SPA-CAT") & participants$sex=="Female"]))` female).
+We collected data from `r length(unique(participants$participant_id))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
+`r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant_id[participants$group %in% c("cat-SPA")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant_id[participants$group %in% c("cat-SPA") & participants$sex=="Female"]))` female).
 Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
@@
 English participants were randomly assigned to the Catalan or Spanish lists.
-Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="ENG-CAT",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="ENG-SPA",])` trials.
+Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="cat-ENG",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="spa-ENG",])` trials.
 Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
@@
 Trials in which participants took longer than 10 seconds to respond were also excluded.
-Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("ENG-CAT", "SPA-CAT"),])` in Catalan, `r nrow(responses[responses$group %in% c("ENG-SPA"),])` in Spanish).
+Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("cat-ENG", "cat-SPA"),])` in Catalan, `r nrow(responses[responses$group %in% c("spa-ENG"),])` in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
-Words in the Catalan list were `r printnum(lengths$eng_cat["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_cat["ipa1_sd"])`, Range = `r printnum(lengths$eng_cat["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_cat["ipa1_max"], digits = 0)`).
-Words in the Spanish list were `r printnum(lengths$eng_spa["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$eng_spa["ipa1_sd"])`, Range = `r printnum(lengths$eng_spa["ipa1_min"], digits = 0)`-`r printnum(lengths$eng_spa["ipa1_max"], digits = 0)`).
+Words in the Catalan list were `r printnum(lengths$cat_eng["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$cat_eng["ipa1_sd"])`, Range = `r printnum(lengths$cat_eng["ipa1_min"], digits = 0)`-`r printnum(lengths$cat_eng["ipa1_max"], digits = 0)`).
+Words in the Spanish list were `r printnum(lengths$spa_eng["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$spa_eng["ipa1_sd"])`, Range = `r printnum(lengths$spa_eng["ipa1_min"], digits = 0)`-`r printnum(lengths$spa_eng["ipa1_max"], digits = 0)`).
 Participants listened to one audio file in each trial, each containing a single word presented in isolation.
@@
 The average duration of the audio files was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
-The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-CAT")], na.rm = TRUE))`), and the average duration of the Spanish audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("ENG-SPA")], na.rm = TRUE))`).
+The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`), and the average duration of the Spanish audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`).
 
@@
 
-Responses given by English participants for Catalan presented words were `r printnum(lengths$eng_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`), while their translations to Spani responses given by Spanish participants for Catalan words were `r printnum(lengths$spa_cat["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_cat["word2_sd"])`, Range = `r printnum(lengths$eng_cat["word2_min"], digits = 0)`-`r printnum(lengths$eng_cat["word2_max"], digits = 0)`).
-Responses tgiven by English participants for Spanish words were `r printnum(lengths$eng_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$eng_spa["word2_sd"])`, Range = `r printnum(lengths$eng_spa["word2_min"], digits = 0)`-`r printnum(lengths$eng_spa["word2_max"], digits = 0)`).
+Responses given by English participants for Catalan presented words were `r printnum(lengths$cat_eng["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_eng["word2_sd"])`, Range = `r printnum(lengths$cat_eng["word2_min"], digits = 0)`-`r printnum(lengths$cat_eng["word2_max"], digits = 0)`), while their translations to Spanish responses given by Spanish participants for Catalan words were `r printnum(lengths$cat_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_spa["word2_sd"])`, Range = `r printnum(lengths$cat_spa["word2_min"], digits = 0)`-`r printnum(lengths$cat_spa["word2_max"], digits = 0)`).
+Responses given by English participants for Spanish words were `r printnum(lengths$spa_eng["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_eng["word2_sd"])`, Range = `r printnum(lengths$spa_eng["word2_min"], digits = 0)`-`r printnum(lengths$spa_eng["word2_max"], digits = 0)`).
 
@@
     word %in% c("book", "sock"),
-    group=="ENG-CAT"
+    group=="cat-ENG"
   ) %>% 
@@
   mutate(
-    group = case_when(
-      group=="ENG-CAT" ~ "cat-ENG",
-      group=="ENG-SPA" ~ "spa-ENG",
-      group=="SPA-CAT" ~ "cat-SPA"
-    ),
     group = factor(
@@
 We computed this similarity measure (refred simply as Levenshtein, from now on) for every translation pair in our stimuli lists.
-For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim("teɪbəl", "mesa"))` similarity, while the *train* (trEIn)-*tren* (tɾEn) translation pair had a `r percent(stringsim("trEIn", "tɾEn"))` similarity.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
 Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
@@
   "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:groupENGvs.SPA" = "Levenshtein \u00d7 Group 1",
-  "lv_std:groupENGMCATvs.ENGMSPA" = "Levenshtein \u00d7 Group 2",
+  "lv_std:groupeng_spa" = "Levenshtein \u00d7 Group 1",
+  "lv_std:groupcateng_spaeng" = "Levenshtein \u00d7 Group 2",
   "lv_std" = "Levenshtein\n[+1 SD]",
   "pthn_std" = "PTHN\n[+1 SD]",
-  "groupENGvs.SPA" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-  "groupENGMCATvs.ENGMSPA" = "Group 2\n[cat-ENG vs. spa-ENG]"
+  "groupeng_spa" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
+  "groupcatengspaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
 
-post_sd <- post_draws %>% 
+post_sd <- post_sd_data %>% 
   filter(group=="Participants") %>% 
   ggplot() +
-  aes(.value, fct_rev(.variable_name)) +
+  aes(
+    x = .value, 
+    y = fct_rev(.variable_name)
+  ) +
   stat_slab(
@@
 
-# Pearson correlations between random effects
+# pearson correlations between random effects
 corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
@@
 h_pthn_lv <- get_bf("pthn_std:lv_std = 0")
-h_group1 <- get_bf("groupENGvs.SPA = 0")
-h_group2 <- get_bf("groupENGMCATvs.ENGMSPA = 0")
-h_lv_group1 <- get_bf("lv_std:groupENGvs.SPA = 0")
-h_lv_group2 <- get_bf("lv_std:groupENGMCATvs.ENGMSPA = 0")
+h_group1 <- get_bf("groupeng_spa = 0")
+h_group2 <- get_bf("groupcateng_spaeng = 0")
+h_lv_group1 <- get_bf("lv_std:groupeng_spa = 0")
+h_lv_group2 <- get_bf("lv_std:groupcateng_spaeng = 0")
 
@@
 
-word_example <- ranef(fit_6)$word[,,"Intercept"] %>% 
+word_example <- ranef(fit_6)$translation_id[,,"Intercept"] %>% 
   as.data.frame() %>% 
@@
 
-corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
-corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
+corr_mat <- VarCorr(fit_6, summary = TRUE)$participant_id$cor 
+corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant_id$cor)
 
@@
 h_cor_intercept_frequency <- get_bf(
-  "participant__Intercept__frequency_zipf_std = 0", 
+  "participant_id__Intercept__frequency_zipf_std = 0", 
   class = "cor"
@@
 h_cor_frequency_zipf_lv_group1 <- get_bf(
-  "participant__frequency_zipf_std__lv_std:groupENGvs.SPA = 0", 
+  "participant_id__frequency_zipf_std__lv_std:groupeng_spa = 0", 
   class = "cor"
@@
 h_cor_intercept_pthn <- get_bf(
-  "participant__Intercept__pthn_std = 0",
+  "participant_id__Intercept__pthn_std = 0",
   class = "cor"
@@
 
-Spanish natives translating Catalan words were `r percent(coefs$group_en_gvs_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
-The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in `r percent(coefs$lv_std_group_en_gvs_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_group_en_gvs_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_en_gvs_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-English natives translating Catalan words were `r percent(abs(coefs$group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
-For every one *SD* increment in Levenshtein similarity, the effect of presented language increased in `r percent(abs(coefs$lv_std_group_engmca_tvs_engmspa["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_group_engmca_tvs_engmspa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_group_engmca_tvs_engmspa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+Spanish natives translating Catalan words were `r percent(coefs$groupeng_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$groupeng_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$groupeng_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
+The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in `r percent(coefs$lv_std_groupeng_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_groupeng_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_groupeng_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
+English natives translating Catalan words were `r percent(abs(coefs$groupcateng_spaeng["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$groupcateng_spaeng["q2_5"], accuracy = 0.01)`, `r percent(coefs$groupcateng_spaeng["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
+For every one *SD* increment in Levenshtein similarity, the effect of presented language increased in `r percent(abs(coefs$lv_std_groupcateng_spaeng["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_groupcateng_spaeng["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_groupcateng_spaeng["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
 
@@
 Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
-The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_std_group_en_gvs_spa$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_std_group_en_gvs_spaE$.upper, accuracy = 0.01)`]).
+The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_std_group_eng_spa$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_std_group_eng_spa$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_std_group_eng_spa$.upper, accuracy = 0.01)`]).
 This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
@@
 The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_frequency, 2)`).
-The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf_std$estimate_lv_std_group_en_gvs_spa, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf_std$q2_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf_std$q97_5_lv_std_group_en_gvs_spa, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
+The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf_std$estimate_lv_std_group_eng_spa, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf_std$q2_5_lv_std_group_eng_spa, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf_std$q97_5_lv_std_group_eng_spa, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
 The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).

### 2022-04-27 R/00_stimuli.R

@@ -1,158 +1,3 @@
-<<<<<<< HEAD
-# relative frequencies and phonological neighbours
-get_clearpond <- function(clearpond_path){
-    clearpond <- clearpond_path %>% 
-        map(
-            function(x){
-                read.csv(x) %>% 
-                    clean_names() %>% 
-                    select(
-                        "word",
-                        "pho_word", 
-                        "freq_per_million", 
-                        "pthn"
-                    )
-            }
-        ) %>% 
-        bind_rows(.id = "group") %>% 
-        as_tibble() %>%
-        distinct(
-            word, 
-            group, 
-            .keep_all = TRUE
-        ) %>% 
-        rename(
-            word2 = word,
-            phon_clearpond = pho_word,
-            frequency = freq_per_million
-        )
-    return(clearpond)
-}
-
-# get Leveshtein similarity scores
-get_levenshtein <- function(stimuli_path){
-    
-    levenshtein <- c(
-        "ENG-SPA" = "English-Spanish", 
-        "ENG-CAT" = "English-Catalan",
-        "SPA-CAT" = "Spanish-Catalan"
-    ) %>% 
-        map(
-            function(x) read_xlsx(stimuli_path, sheet = x)
-        ) %>% 
-        bind_rows(.id = "group") %>% 
-        clean_names() %>% 
-        select(
-            group, 
-            word1, 
-            word2, 
-            ipa1,
-            ipa2
-        ) %>% 
-        mutate(
-            n_char = ifelse(
-                nchar(ipa1) > nchar(ipa2),
-                nchar(ipa1),
-                nchar(ipa2)
-            ),
-            lv = stringsim(
-                ipa1, 
-                ipa2, 
-                method = "lv"
-            )
-        )
-    
-    return(levenshtein)
-}
-
-# get audio duration
-get_duration <- function(
-    stimuli_path,
-    audios_path
-){
-    stimuli <- c(
-        "ENG-SPA" = "English-Spanish",
-        "ENG-CAT" = "English-Catalan", 
-        "SPA-CAT" = "Spanish-Catalan"
-    ) %>% 
-        map(
-            ~read_xlsx(
-                stimuli_path, 
-                sheet = .
-            )
-        ) %>% 
-        bind_rows(.id = "group") %>% 
-        select(
-            word1, 
-            group, 
-            file
[…]

### 2022-04-27 _targets.R

@@ -9,8 +9,7 @@ source("R/02_models.R", encoding = "UTF-8")
 
 # set number of cores to use with brms
 options(
-    mc.cores = 1,
-    encoding = "UTF-8"
+    mc.cores = 1
 )
 
 # set parameters
@@ -62,9 +61,9 @@ list(
     tar_target(
         clearpond_path, 
         lst(
-            `ENG-CAT` = here("data", "clearpond", "clearpond_english.csv"), 
-            `ENG-SPA` = here("data", "clearpond", "clearpond_english.csv"), 
-            `SPA-CAT` = here("data", "clearpond", "clearpond_spanish.csv")
+            "cat-ENG" = here("data", "clearpond", "clearpond_english.csv"), 
+            "spa-ENG" = here("data", "clearpond", "clearpond_english.csv"), 
+            "cat-SPAT" = here("data", "clearpond", "clearpond_spanish.csv")
         )
     ),
     tar_target(
@@ -193,13 +192,9 @@ list(
                     pthn_std + 
                     lv_std + 
                     pthn_std:lv_std + 
-<<<<<<< HEAD
                     (1 + frequency_zipf_std + pthn_std*lv_std | participant_id) + 
                     (1 | translation_id)
-            )
-=======
-                    (1 + frequency_zipf_std + pthn_std*lv_std | participant) + 
-                    (1 | word)
+                
             ),
             f_5 = bf(
                 correct ~ 1 + 
@@ -208,8 +203,8 @@ list(
                     lv_std + 
                     lv_std:pthn_std + 
                     group +
-                    (1 + frequency_zipf_std + pthn_std + lv_std + lv_std:pthn_std | participant) + 
-                    (1 + group | word)
+                    (1 + frequency_zipf_std + pthn_std + lv_std + lv_std:pthn_std | participant_id) + 
+                    (1 + group | translation_id)
             ),
             f_6 = bf(
                 correct ~ 1 +
@@ -219,9 +214,8 @@ list(
                     lv_std:pthn_std + 
                     group + 
                     group:lv_std +
-                    (1 + frequency_zipf_std + pthn_std + lv_std + lv_std:pthn_std + group +group:lv_std | participant) + 
-                    (1 | word))
->>>>>>> eb2ad2e01316fbb4878ef163afbb685d3a91ad49
+                    (1 + frequency_zipf_std + pthn_std + lv_std + lv_std:pthn_std + group + group:lv_std | participant_id) + 
[…]

### 2022-04-28 manuscript/manuscript.Rmd
Implement suggestions by KP

@@
 One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, imagine a Spanish native, with no previous familiarity with any other language, listens for the first time to French may encounter the word p`r knitr::asis_output("\U0254")``r knitr::asis_output("\U0281")`t/ (*porte*), translation of *door* in French.
-The voiced uvular fricative consonant /`r knitr::asis_output("\U0281")`/ and the open-mid back rounded vowel /`r knitr::asis_output("\U0254")`/ do not exist in Spanish.
-These discrepancies can nonetheless be perceived as allophonic variations of native phonemes [@best2001discrimination], and the non-native speech signal engages lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /'pwer.ta/ (*puerta*).
+For example, imagine a Spanish native, with no previous familiarity with any other language, who listens for the first time to French may encounter the word /`r knitr::asis_output("\U0070\U0254\U0281\U0074")` (*porte*), translation of *door* in French.
+The voiced uvular fricative consonant /`r knitr::asis_output("\U0281")`/ and the open-mid back rounded vowel /`r knitr::asis_output("\U0254")`/ do not exist in Spanish as phonemic categories.
+These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /r/ and /o/ [@best2001discrimination], and the non-native speech signal may engage lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ (*puerta*).
 The recognition of this French, unfamiliar word relies almost entirely on its similarity with its Spanish translation.
-Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /d`r knitr::asis_output("\U0254")`ː/ (*door*).
-This time, phonological similarity is of little help: /d`r knitr::asis_output("\U0254")`ː/ and /'pwer.ta/ share very few phonemes.
-It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations (which they lack, they are unfamiliar to English).
-
-Thus, cross-language phonological similarity seems, at face value, an important source of information for unfamiliar listeners during auditory word recognition, even translations are not phonologically identical.
+Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /`r knitr::asis_output("\U0064\U0254\U02D0")`/ (*door*).
+This time, phonological similarity is of little help: /`r knitr::asis_output("\U0064\U0254\U02D0")`/ and /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ share very few phonemes.
+It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations (which they lack, as they are unfamiliar to English).
+
+Cross-language phonological similarity seems, at face value, an important source of information for unfamiliar listeners during auditory word recognition, even if translations are not phonologically identical.
 Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
-In the present study we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
-
-[//]: # (NSG: I think that a nice way to start the paper (or to finish it) is by citing one of those web pages that say that native English speakers already know tons of words in Spanish, French, German...
-because they are very similar (aka cognates).
-But this may work for writing, because for speech the situation is very different.
-This may be at the beginning or at the end of the paper.
-Given that most reviewers/ editors are likely to be native English-speakers, this may convince them that the paper is interesting and that they want to see it published.)
-
-Many languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration or social contact).
+In the present study, we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
+
+Many languages share, to some extent, similarities at the lexical level.
+This is frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
 Cognates embody a particular case of such commonalities.
-Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
+Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level) commonly due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
 For example, *sun* and *sol* (in Spanish) share phonological onset but their etymology points to different origins.
-The complimentary case is equally problematic for the definition of cognateness in this context: cross-language synonyms that share etymological origin may not be necessarily be recognised as cognates by the average listener.
-Such is the case of *cheese* and *queso* (Latin root), but are phonologically dissimilar.
+The complimentary case is equally problematic when defining cognateness: cross-language synonyms that share etymological origin may not be necessarily be recognised as cognates by the average listener.
+Such is the case of *cheese* and *queso* which share Latin root but are phonologically dissimilar.
 For simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
@@
 
-Early theoretical accounts of this the 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+Early theoretical accounts of this 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
 The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), instead of directly to their shared concept (conceptual route).
@@
 As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
-Cognates subsequently exert less impact on lexical processing in proficient speakers.
-The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 (*forward* translation) should take longer than translating from L2 to L1 (*backward* translation).
-The rationale behind this prediction was that backward translation relies more strongly on word-to-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
-Backward translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
-
-To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
-Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
-Results showed that translation times and accuracy were roughly equivalent in forward and backward translation.
-Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected  by semantic variables than non-cognates.
-Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
-Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on word-to-word links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
-However, although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation [@christoffels2006memory; @christoffels2013language].
-These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
-
-[//]: # (KP: I find this para quite complicated to follow. Would it be possible to distill it to a simpler story that is directly relevant for the manuscript.)
+Cognates subsequently exert less impact on lexical processing in proficient speakers [see @andras2022cognate for recent experimental evidence for this claim].
+The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 should take longer than translating from L2 to L1.
+The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, whereas L2-to-L1 translation relies more strongly on the mediation between the concept and the two word forms.
+L1-to-L2 translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during L1-to-L2 translation.
+
+To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate Dutch words to English (L1-to-L2) or English words to Dutch (L2-to-L1).
+Although participants' performance in both conditions was roughly equivalent for cognates, non-cognates were translated more slowly from Dutch to English than from English to Dutch, suggesting that L1-to-L2 translation is more sensitive to form-similarity than L2-to-L1 translation.
+Conversely, participants' performance was more sensitive to semantic variables (e.g., concreteness) when translating English words to Dutch, compared to when translating Dutch words to English, suggesting that L2-to-L1 translation was more sensitive to the shared conceptual properties of the translation pair.
+
+These findings support a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations than L2-to-L1 translation, and therefore is more sensitive to cognateness.
+Further evidence for the presence of differences in reliance on the lexical and the conceptual routes between L1-to-L2 and L2-to-L1 translation was provided by @phillips2006erp, who used eletrophysiological recordings of bilinguals translating words from L1 to L2 or *vice versa* after semantic or phonological priming.
+The authors reported stronger N400 responses (associated with semantic priming) in L1-to-L2 translation than in L2-to-L2 translation, and stronger phonological mismatch negativity responses (associated with phonological priming) during L2-to-L1 translation, compared to L1-to-L2 translation.
+
+There is mixed evidence, however, about whether these differences are dependent on participants' proficiency in L2.
+While @degroot1994forward found an equivalent pattern of results in high- and low-proficiency bilinguals, other studies showed different performances in both groups during L2-to-L1 and L1-to-L2 translation [@christoffels2006memory; @christoffels2013language].
+These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backing a more complex interaction between the L1 and L2 lexica during word recognition.
 
@@
 
-A more recent model of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
-This model assumes an integrated lexicon in which word representations from two languages establish connections in a way that is similar to connections between words of a single language.
-It also makes the assumption that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
-It also  proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-to-word connections).
-In addition, this model suggests that the strength of the association between L2 representations and their concept depends on language users' proficiency in L2.
-A strong correlation between Multilink simulations of a translation elicitation task and behavioural data from Dutch-English bilinguals [@christoffels2006memory] has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
-Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
-Overall, results from Multilink are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
-
-The fact that Multilink successfully fitted the experimental data from @christoffels2006memory in the absence of word-to-word connections between translation equivalents is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
+A more recent mmake(odel of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
+This model assumes an integrated lexicon, in which word representations from the two languages are activated by the input in a language non-selective way.
+It also proposes that translation equivalents are exclusively connected through their shared concept, to which they both send activation.
+The model does not implement excitatory or inhibitory connections between word representations, within or across languages.
+The cognate effect under this model arises from the fact that the input activates form-similar word in both L1 and L2 word-forms, so when such word-forms send their activation to a shared semantic node (i.e., cognates) such semantic node receives more excitation than when only receiving activation from the word representation in one language (i.e., non-cognates).
+This specification crucially differs from that of the RHM model, which suggests includes excitatory connections between the L1 and L2 word-forms at the lexical level, allowing translation to be achieved without the mediation of semantic nodes.
+In addition, Multilink suggests that the strength of the association between L2 representations and their concept depends on language users' proficiency in L2.
+A simulation series by @dijkstra2019multilink shows that data generated from Multilink are parallel to those observed in translation elicitation task and behavioural data from Dutch-English bilinguals [@christoffels2006memory].
+Overall, this  model provides support to the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+
+The fact that Multilink successfully fitted the experimental data from @christoffels2006memory in the absence of word-to-word connections between translation equivalents is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which L1-to-L1 translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
 It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
@@
 The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-to-word connections during translation, especially during backward translation, and this prediction remains untested by Multilink.
-
-Most of the studies referenced so far rely heavily on production and comprehension of printed words: in these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
+It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-to-word connections during translation, especially during backward translation.
+This prediction remains untested by Multilink.
+
+One caveat concerning the generalisability of these conclusions is that most of the studies referenced so far rely heavily on production and comprehension of printed words.
+In these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
 There are limitations in generalising their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
@@
 Both words are clearly cognates, as their orthographical word form is identical.
-However, their phonological forms (English /ˈɹeɪ.di.əʊ/-- Spanish /ˈra.ðjo/) are not identical.
+However, their phonological forms (English /`r knitr::asis_output("\U02C8\U0279\U0065\U026A\U002E\U0064\U0069\U0259\U028A")`/-- Spanish /`r knitr::asis_output("\U02C8\U0072\U0061\U002E\U00F0\U006A\U006F")`/) are not identical.
 Although many English listeners might recognise the Spanish word-form--even with no previous knowledge of Spanish--it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
@@
 Measuring the phonological similarity between two word forms is also substantially more challenging than measuring their orthographic similarity.
-Many studies use symbolic representations to compute the simiarity between word-forms of interest in either modality.
-Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (*radio*-*radio*); phonological word-forms are represented by their  phonological transcriptions (`r knitr::asis_output("\U0279")`eɪdiəʊ-raðjo, in IPA notation) [see @floccia2018introduction for example].
+Many studies use symbolic representations to compute the similarity between word-forms of interest in either modality.
+Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (*radio*-*radio*); phonological word-forms are represented by their  phonological transcriptions (/`r knitr::asis_output("\U02C8\U0279\U0065\U026A\U002E\U0064\U0069\U0259\U028A")`/-/`r knitr::asis_output("\U02C8\U0072\U0061\U002E\U00F0\U006A\U006F")`/, in IPA notation) [see @floccia2018introduction for example].
 The similarity between each pair of word-forms is operationalised numerically by applying some algorithm that computes the edit distance between their symbolic representation.
@@
 Using our previous examples, the Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio", useBytes = TRUE)`: both strings are identical.
-But the distance between the phonological transcriptions `r knitr::asis_output("\U0279")`eɪdiə`r knitr::asis_output("\U028A")` and ra`r knitr::asis_output("\U00F0")`jo is 7, which is the number of edits that separates both character strings.
-There is also an additional layer of complexity: the phonological transcriptions /ˈ`r knitr::asis_output("\U0279")`eɪdiˌə`r knitr::asis_output("\U028A")`/ and /ˈra`r knitr::asis_output("\U00F0")`jo/ can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.
+But the distance between the phonological transcriptions `r knitr::asis_output("\U0279\U0065\U026A\U0064\U0069\U0259\U028A")` and `r knitr::asis_output("\U0072\U0061\U00F0\U006A\U006F")` is 7, which is the number of edits that separates both character strings.
+There is also an additional layer of complexity: the phonological transcriptions `r knitr::asis_output("\U0279\U0065\U026A\U0064\U0069\U0259\U028A")` and `r knitr::asis_output("\U0072\U0061\U00F0\U006A\U006F")` can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.
 whether to include elements of aspiration, nasalisation, etc.) can affect how accurate these symbolic representation capture all of the features that are present in the acoustic signal when either word is spoken.
@@
 Taking into account that measuring phonological similarity is considerably more complex than measuring orthographic complexity it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
-
-In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be misrepresenting the availability of such information in the auditory modality.
-In the present study we tested listener's reliance on phonological similarity when translating words from an unfamiliar language, in order to investigate the plausibility of word-to-word during backward translation.
-We tested English and Spanish native speakers in a translation elicitation task: they were presented with spoken words in an unfamiliar, non-native language, and were asked to translate them to their native language.
+In summary, it is likely that previous investigations on bilinguals' reliance on form-similarity during translation might be misrepresenting the availability of such information in the auditory modality.
+
+In the present study we tested listeners' reliance on phonological similarity when translating words from an unfamiliar language, in order to investigate the plausibility of word-to-word connections during L1-to-L2 translation.
+We tested English and Spanish native speakers in a translation elicitation task.
+Participants
+were presented with spoken words in an unfamiliar (non-native language) and were asked to translate them to their native language.
 We will henceforth refer to the auditory-presented words heard by participants on each trial as **presented words**, and the correct translation for the presented words as **target translations**.
@@
 
-We tested three groups of participants in the translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
+We tested three groups of participants in this translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
 Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
@@
 
-```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of coefficients in Model 4.
-A\\) Population-level effects.
-Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of population-level regression coefficients in Model 4.
+Distributions indicate the estimated posterior likelihood density of regression coefficients.
 Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
-Black dots represent the mean of each distribution.
-B\\) Participant\\-level coefficient variability.
-The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
-Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
-C\\) Correlation between participant\\-level effects.
-The model allowed participant-level coefficients to co\\-vary.
-This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation.", fig.height=7, fig.width=6.5}
+Black dots represent the mean of each distribution.", fig.height=4.5, fig.width=6.5}
 
@@
   "groupeng_spa" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-  "groupcatengspaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
+  "groupcateng_spaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
 )
@@
 # fixed effects
-post_fix <- post_draws %>% 
+post_draws %>% 
   filter(str_detect(.variable, "b_")) %>% 
@@
 
-# SD of random effects
-post_sd_data <- post_draws %>% 
-  filter(str_detect(.variable, "sd_")) %>% 
-  mutate(
-    .variable_name = str_remove(
-      .variable,
-      "sd_participant__|sd_word__"
-    ) %>% 
-      str_replace_all(str_repl) %>% 
-      factor(
-        levels = c(
-          "Intercept", 
-          "Frequency\n[+1 SD]",
-          "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", 
-          "PTHN \u00d7 Levenshtein",
-          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-          "Group 2\n[cat-ENG vs. spa-ENG]",
-          "Levenshtein \u00d7 Group 1",
-          "Levenshtein \u00d7 Group 2"
-        )
-      ),
-    group = case_when(
-      str_detect(.variable, "participant") ~ "Participants",
-      str_detect(.variable, "word") ~ "Words"
-    ),
-    .value = ifelse(
-      str_detect(.variable, "Intercept"), 
-      inv_logit_scaled(.value), 
-      .value/4
-    )
-  )
-
-post_sd <- post_sd_data %>% 
-  filter(group=="Participants") %>% 
-  ggplot() +
-  aes(
-    x = .value, 
-    y = fct_rev(.variable_name)
-  ) +
-  stat_slab(
-    aes(
-      fill = stat(
-        cut_cdf_qi(
-          cdf, 
-          .width = c(.5, .8, .95),
-          labels = percent_format()
-        )
-      )
-    )
-  ) +
-  stat_pointinterval(
-    .width = 0, 
-    point_size = 1
-  ) +
-  scale_fill_manual(
-    values = c("#ff2976", "#f2a7c2", "#fcd9e6"), 
-    na.translate = FALSE
-  ) +
-  scale_x_continuous(
-    labels = percent, 
-    limits = c(0, 1)
-  ) +
-  labs(
-    x = "SDs in P(Correct)",
-    title = "By participant", 
-    fill = "CrI"
-  ) +
-  theme(
-    axis.text.x = element_blank()
-  ) +
-
-  post_sd_data %>% 
-  filter(group=="Words") %>% 
-  ggplot() +
-  aes(
-    .value, 
-    fct_rev(.variable_name)
-  ) +
-  stat_slab(
-    aes(
-      fill = stat(
-        cut_cdf_qi(
-          cdf, 
-          .width = c(.5, .8, .95), 
-          labels = percent_format()
-        )
-      )
-    )
-  ) +
-  stat_pointinterval(
-    .width = 0, 
-    point_size = 1
-  ) +
-  scale_x_continuous(
-    labels = percent, 
-    limits = c(0, 1)
-  ) +
-  labs(
-    x = "SDs in P(Correct)",
-    title = "By item",
-    fill = "CrI"
-  ) +
-  plot_layout(
-    ncol = 1,
-    heights = c(0.85, 0.15), 
-    guides = "collect"
-  ) &
-  theme(
-    legend.position = "none",
-    axis.text.y = element_text(hjust = 1),
-    axis.title.x = element_blank(),
-    legend.direction = "vertical",
-    axis.title.y = element_blank(),
-    legend.background = element_rect(fill = "grey95"),
-    legend.key.height = unit(0.1, "cm"),
-    legend.key = element_rect(fill = "grey95", colour = "grey95"),
-    legend.text = element_text(size = 7),
-    axis.ticks.y = element_blank(),
-    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
-  )
-
-# pearson correlations between random effects
-corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
-corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
-
-corr_mat <- corr_mat %>% 
-  as_tibble() %>% 
-  select(matches("Estimate")) %>% 
-  rename_all(str_remove, "Estimate.") %>% 
-  mutate(term1 = corr_names)
-
-corr_mat[lower.tri(corr_mat)] <- NA
-
-corr_data <- corr_mat %>% 
-  as.data.frame() %>% 
-  pivot_longer(
-    -term1, 
-    names_to = "term2",
-    values_to = ".value"
-  ) %>% 
-  drop_na(.value)
-
-post_cors <- corr_data %>% 
-  mutate(
-    term1 = factor(
-      term1, 
-      levels = names(str_repl),
-      ordered = TRUE
-    ),
-    term2 = factor(
-      term2, 
-      levels = names(str_repl),
-      ordered = TRUE
-    )
-  ) %>% 
-  mutate_at(
-    vars(term1, term2),
-    str_replace_all, 
-    str_repl
-  ) %>% 
-  mutate(
-    .value_label = case_when(
-      term1==term2 ~ NA_character_, 
-      TRUE ~ printnum(.value, gt1 = FALSE)
-    ),
-    .value = case_when(
-      term1==term2 ~ NA_real_, 
-      TRUE ~ .value
-    )
-  ) %>% 
-  filter(
-    term1 != "PTHN × Levenshtein", 
-    term2 != "Intercept"
-  ) %>% 
-  ggplot() +
-  aes(
-    fct_inorder(term1),
-    fct_rev(fct_inorder(term2)), 
-    fill = .value
-  ) +
-  geom_tile(na.rm = TRUE) +
-  geom_text(
-    aes(label = .value_label), 
-    size = 3, 
-    na.rm = TRUE) +
-  labs(
-    x = "Term 1", 
-    y = "Term 2",
-    fill = "Posterior\ncorrelation"
-  ) +
-  scale_fill_gradient2(
-    low = "#04bf78", 
-    mid = "white",
-    high = "#FFC20A", 
-    na.value = "white", 
-    limits = c(-0.4, 0.4)
-  ) +
-  scale_x_discrete(position = "top") +
-  coord_equal() +
-  theme(
-    legend.position = c(1, 1),
-    legend.direction = "horizontal",
-    legend.key.height = unit(0.1, "cm"),
-    legend.key.width = unit(0.3, "cm"),
-    legend.justification = c(1, 1),
-    legend.title = element_blank(),
-    legend.text = element_text(size = 5),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank(),
-    axis.ticks = element_blank(),
-    axis.text.x = element_text(
-      size = 6,
-      angle = 90,
-      hjust = 0, 
-      vjust = 1),
-    axis.text.y = element_text(
-      size = 6,
-      vjust = 1,
-      hjust = 1
-    )
-  )
-
-post_fix / (
-  (post_sd | post_cors) +
-    plot_layout(widths = c(0.6, 0.4))
-) +
-  plot_layout() +
-  plot_annotation(tag_levels = "A")
-
 ```
@@
 The dispersion of the lines indicates the uncertainty of our predictions.
-We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
+We computed these posterior predictions for each group of participants, and plotted them in separate panels."}
 
@@
   aes(
-    pthn_std, 
-    .epred,
+    x = pthn_std, 
+    y = .epred,
     colour = lv_std,
@@
   labs(
-    x = paste0("PTHN (+1 SD, ", round(sd(responses$pthn), 1), " higher-frequency neighbours)"),
+    x = paste0(
+      "PTHN (+1 SD, ", 
+      round(sd(responses$pthn), 1), 
+      " higher-frequency neighbours)"
+    ),
     y = "P(Correct)", 
@@
   ) +
+  scale_linetype_manual(
+    values = c("solid", "dashed", "dotdash")
+  ) +
   scale_y_continuous(
@@
 
-The conditional intra-class correlation [*ICC*; @gelman2006data] for items was `r percent(icc_conditional, accuracy = 0.01)`, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items.
-Item-level intercepts showed remarkable variability (*SD* = `r percent(sds$sd_word_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_word_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_word_intercept$.upper, accuracy = 0.01)`]), indicating that participants' accuracy differed substantially across words: some words were very likely to be correctly translated (e.g., Spanish *tomate*-*tomato*, /tomate/-/təmɑːtoʊ/, `r percent(word_example$tomate$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$tomate$q2_5, accuracy = 0.01)`, `r percent(word_example$tomate$q97_5, accuracy = 0.01)`]), while others were almost never translated correctly (e.g., Catalan *colom*-*pigeon*, /kulom/-/pɪdʒɪn/, `r percent(word_example$colom$estimate, accuracy = 0.01)`, 95% CrI = [`r percent(word_example$colom$q2_5, accuracy = 0.01)`, `r percent(word_example$colom$q97_5, accuracy = 0.01)`]).
-Participant-level intercepts also showed high variability (*SD* = `r percent(sds$sd_participant_intercept$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_intercept$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_intercept$.upper, accuracy = 0.01)`]), suggesting that average translation accuracy also varied significantly across participants.
-Finally, participant-level estimated slopes showed less variability: *SD* of $Frequency$, $PTHN$, $Levenshtein$, $PTHN \times Levenshtein$, and $Levenshtein \times Group 2$ ranged between `r percent(participant_ranef_range[1], accuracy = 0.01)`, and `r percent(participant_ranef_range[2], accuracy = 0.01)`.
-The participant-level estimated slopes of the $Leveshtein \times Group 1$ interaction are an exception, though (*SD* = `r percent(sds$sd_participant_lv_std_group_eng_spa$.value, accuracy = 0.01)`, 95% CrI = [`r percent(sds$sd_participant_lv_std_group_eng_spa$.lower, accuracy = 0.01)`, `r percent(sds$sd_participant_lv_std_group_eng_spa$.upper, accuracy = 0.01)`]).
-This indicates that the impact of phonological similarity ($Levenshtein$) on the difference in accuracy between English and Spanish natives differed substantially across participants.
-
-Finally, participant-level effects showed weak Pearson correlations overall (<10%), with three exceptions.
-The estimated correlation between the intercepts and the slopes of $Frequency$ was `r percent(corr_data$intercept$estimate_frequency_zipf_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_frequency_zipf_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_frequency_zipf_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_frequency, 2)`).
-The estimated correlation between the slopes of $Frequency$ and $Levenshtein \times Group 1$ was `r percent(corr_data$frequency_zipf_std$estimate_lv_std_group_eng_spa, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$frequency_zipf_std$q2_5_lv_std_group_eng_spa, accuracy = 0.01)`, `r percent(corr_data$frequency_zipf_std$q97_5_lv_std_group_eng_spa, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_frequency_zipf_lv_group1, 2)`).
-The estimated correlation between the intercepts and the slopes of $PTHN$ was `r percent(corr_data$intercept$estimate_pthn_std, accuracy = 0.01)` (95% CrI = [`r percent(corr_data$intercept$q2_5_pthn_std, accuracy = 0.01)`, `r percent(corr_data$intercept$q97_5_pthn_std, accuracy = 0.01)`], $BF_{Cor \neq 0}$ = `r round(h_cor_intercept_pthn, 2)`).
-The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
-
 # Discussion

### 2022-04-28 R/utils.R

@@ -75,39 +75,58 @@ replace_non_ascii <- function(x){
 
 
 # download clearpond
-# import_clearpond <- function(language = c("english", "dutch", "french", "spanish", "german")){
-#     require(tidyverse)
-#     require(janitor)
-#     urls <- tribble(
-#         ~lang, ~url, ~data, ~header,
-#         "english", "https://clearpond.northwestern.edu/englishCPdatabase2.zip", "englishCPdatabase2.txt", "clearpondHeaders_EN.txt",
-#         "dutch", "https://clearpond.northwestern.edu/dutchCPdatabase2.zip", "dutchCPdatabase2.txt", "clearpondHeaders_NL.txt",
-#         "french", "https://clearpond.northwestern.edu/frenchCPdatabase2.zip", "frenchCPdatabase2.txt", "clearpondHeaders_FR.txt",
-#         "german", "https://clearpond.northwestern.edu/germanCPdatabase2.zip", "germanCPdatabase2.txt", "clearpondHeaders_DE.txt",
-#         "spanish", "https://clearpond.northwestern.edu/spanishCPdatabase2.zip", "spanishCPdatabase2.txt", "clearpondHeaders_SP.txt"
-#     ) %>% 
-#         filter(lang %in% language)
-#     
-#     dir <- tempdir()
-#     files <- replicate(tempfile(), n = length(urls$lang))
-#     d <- pmap(
-#         .l = list(url = as.list(urls$url), file = as.list(files), data = as.list(urls$data), header = as.list(urls$header)),
-#         .f = function(url = .l[[1]], file = .l[[2]], data = .l[[3]], header = .l[[4]]) {
-#             download.file(url, destfile = file)
-#             unzip(zipfile = file, exdir = dir)
-#             headers <- c("word", read.delim(paste0(dir, .Platform$file.sep, header))[,1])
-#             d <- read.delim(paste0(dir, .Platform$file.sep, data)) %>% 
-#                 `colnames<-`(., headers) %>% 
-#                 as_tibble() %>% 
-#                 mutate_at(vars(ends_with("W")), ~str_split(., pattern = ";"))
-#             return(d)
-#         }
-#     ) %>% 
-#         set_names(language) %>% 
-#         bind_rows(.id = "language") %>% 
-#         clean_names()
-#     return(d)
-# }
+import_clearpond <- function(
+    language = c(
+        "english", 
+        "dutch", 
+        "french", 
+        "spanish", 
+        "german"
+    )
+){
[…]

### 2022-10-28 R/additional_analyses.qmd
Add questionnaire data and analyses, and change reports to Quarto

---
title: "Additional analyses"
format: html
---

## Additional analyses

### Previous familiarity

```{r}
#| label: familiarity-frequency
#| fig-height: 10
#| fig-width: 10
familiarity_problematic <- here("data", "familiarity", "familiarity.xlsx") %>% read_xlsx() %>% 
    rename(
        word = presented,
        word_2 = translation,
        test_language = language,
        problem = note
    ) %>% 
    mutate(
        problem = ifelse(
            str_detect(problem, "check coding"), 
            "Check coding",
            str_to_sentence(problem)
        ),
        group = ifelse(test_language=="Catalan", "cat-ENG", "spa-ENG")
    ) %>% 
    select(word, group, problem)

familiarity <- stimuli %>% 
    filter(group != "cat-SPA") %>%
    mutate(
        stim = paste0(
            word_1, " /", ipa_flat_1, "/", " - ",
            word_2, " /", ipa_flat_2, "/", " - ",
            percent(lv, accuracy = 1)
        ),
        test_language = case_when(
            group=="cat-ENG" ~ "Catalan",
            group=="spa-ENG" ~ "Spanish",
            group=="cat-SPA" ~ "Catalan"
        ),
        word = replace_non_ascii(word_1)
    ) %>% 
    left_join(familiarity_problematic) %>% 
    select(group, word = word, stim, problem) %>% 
    left_join(filter(subtlex, test_language=="English")) %>% 
    replace_na(list(frequency_count = 1, frequency_zipf = 0.555, problem = "Normal")) %>% 
    left_join(select(word_accuracy, group, stim, prop,prop_se)) %>% 
    group_by(group) %>% 
    mutate(word_reorder = reorder_within(word, frequency_zipf, group)
    ) %>% 
    ungroup()

familiarity %>% 
    ggplot() +
    aes(x = frequency_zipf, y = word_reorder, fill = problem, colour = problem, label = word) +
    facet_wrap(~group, scales = "free_y") +
    geom_text(size = 3) +
    labs(
        x = "Frequency (Zipf) of presented word in English",
        y = "Word",
        colour = "Problem"
    ) +
    scale_colour_manual(values = my_colours, na.translate = FALSE) +
    scale_y_reordered() +
    theme(
        legend.position = "top",
        panel.border = element_rect(fill = NA, colour = "grey"),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        legend.title = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank()
    )

ggsave(here("img", "problematic_frequency-distribution.png"), height = 10)

```

```{r}
#| label: previous-familiarity
#| fig-height: 3.5
#| fig-width: 8
familiarity %>% 
    filter(group != "cat-SPA") %>% 
    ggplot() +
    aes(x = frequency_zipf, y = prop, colour = problem, fill = problem, label = word) +
    facet_wrap(~group) +
    geom_text(size = 3) +
    # geom_point(
    #   shape = 1,
    #   size = 1,
    #   alpha = 0.5,
    #   stroke = 1
    # ) +
    geom_smooth(method = "lm", colour = "black", fill = "grey") +
    labs(
        x = "Frequency (Zipf) of presented word in English",
        y = "Accuracy",
        colour = "Group",
        fill = "Group"
    ) +
    scale_colour_manual(values = my_colours, na.translate = FALSE) +
    # scale_x_continuous(
    #   limits = c(2, 6)
    # ) +
    # scale_y_continuous(
    #   limits = c(0, 1),
    #   labels = percent
    # ) +
    theme(
        legend.position = "none"
    )

ggsave(here("img", "problematic_frequency-accuracy.png"), height = 4, width = 10)

```

### 2022-10-28 index.qmd
Add questionnaire data and analyses, and change reports to Quarto

---
title: "Translation Elicitation: laboratory notes"
date: "Last updated: `r format(Sys.Date(), '%d/%m/%Y')`"
format:
    html:
        toc: true
        number-sections: true
        bibliography: "references.bib"
        csl: "apa7.csl"
        warnings: false
        messages: false
        fig-dpi: 600
        out-width: "80%"
        code-fold: true
---

```{r}
#| label: setup
#| include: false
my_colours <- c("#1A85FF", "#ff2976", "#FFC20A")

options(
    knitr.kable.NA = '-',
    htmltools.preserve.raw = FALSE,
    knitr.duplicate.label = "allow",
    ggplot2.discrete.fill = my_colours,
    ggplot2.discrete.colour = my_colours,
    ggplot2.continuous.fill = ggplot2::scale_fill_gradient2(
        low = "#04bf78", 
        mid = "white", 
        high = "#FFC20A", 
        na.value = "white", 
        limits = c(-0.4, 0.4)
    ),
    ggplot2.continuous.colour = ggplot2::scale_fill_gradient2(
        low = "#04bf78", 
        mid = "white",
        high = my_colours[3], 
        na.value = "white",
        limits = c(-0.4, 0.4)
    )
)

# load targets
tar_load_all()

# set custom ggplot theme
theme_set(theme_custom())
```

# Study 1

## Stimuli

::: {.panel-tabset}

### Summary

Participants in the spa-ENG listened to `r nrow(stimuli[stimuli$group=="spa-ENG",])` Spanish words, and participants in the cat-ENG and the cat-SPA groups listened to `r nrow(stimuli[stimuli$group=="cat-ENG",])` trials. Five trials in each condition were practice trials and were excluded from analyses.

```{r}
#| label: tbl-stimuli
tab_subtitle <- md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")

tab_note <- md("*Note*. Median and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")

stimuli %>% 
    select(group, word_1, freq_2, freq_zipf_2, nd, lv) %>% 
    group_by(group) %>% 
    summarise(
        across(
            freq_2:lv, 
            list(
                median = ~median(., na.rm = TRUE),
                sd = ~sd(., na.rm = TRUE)
            )
        )
    ) %>%
    gt(rowname_col = "group") %>% 
    tab_header(title = md("Table 1"), subtitle = tab_subtitle) %>% 
    tab_source_note(tab_note) %>% 
    fmt_number(freq_2_median:nd_sd) %>%
    fmt_percent(matches("lv|global")) %>% 
    tab_spanner(
        md("**Freq. (million)**"),
        c("freq_2_median", "freq_2_sd", "freq_zipf_2_median", "freq_zipf_2_sd")
    ) %>% 
    tab_spanner(md("**Freq. (Zipf)**"), matches("freq_zipf_2")) %>% 
    tab_spanner(md("**Neighbours**"), matches("nd")) %>% 
    tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
    cols_label(
        freq_2_median = md("Median"),
        freq_2_sd = md("*SD*"),
        freq_zipf_2_median = md("Median"),
        freq_zipf_2_sd = md("*SD*"),
        nd_median = md("Median"),
        nd_sd = md("*SD*"),
        lv_median = md("Median"),
        lv_sd = md("*SD*")
    ) %>% 
    summary_rows(
        columns = c(matches("median"), -matches("lv")),
        fns = list(Median = "median", SD = "sd")
    ) %>% 
    summary_rows(
        columns = matches("lv"),
        fns = list(Median = "median", SD = "sd"), 
        formatter = fmt_percent
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_stub_grand_summary()
    ) %>% 
    tab_style(
        style = cell_borders(sides = c("left", "right"), color = "white"),
        locations = list(
            cells_body(),
            cells_row_groups(),
            cells_column_spanners(), 
            cells_stub(),
            cells_stub_grand_summary(), 
            cells_source_notes(),
            cells_column_labels(),
            cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(
            cell_text(align = "left", size = "medium"),
            cell_borders(sides = "all", color = "white")
        ),
        locations = cells_stub()
    ) %>% 
    tab_style(
        style = cell_text(
            align = "center", 
            size = "medium",
            weight = "normal",
            style = "normal"
        ),
        locations = list(
            cells_column_spanners(),
            cells_column_labels()
        )
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:9)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:9)
    )
```

### cat-ENG

```{r}
#| label: tbl-stimuli-cateng
stimuli %>% 
    filter(group=="cat-ENG") %>% 
    mutate(
        word_1 = paste0(word_1, " (", ipa_flat_1, ")"),
        word_2 = paste0(word_2, " (", ipa_flat_2, ")"),
        # nd_list = paste0(nd_list, collapse = ", ")
    ) %>% 
    select(word_1, word_2, duration, freq_zipf_2, lv, lv_dist, nd, nd_list) %>% 
    gt() %>% 
    fmt_number(where(is.numeric)) %>% 
    fmt_integer(c(nd, lv_dist)) %>%
    fmt_percent(lv) %>% 
    fmt_missing(everything(), everything(), "---") %>% 
    tab_spanner("Cross-language phonological neighbours", nd:nd_list) %>% 
    tab_spanner("Phonological similarity", lv:lv_dist) %>% 
    cols_label(
        word_1 = "Presented word (Catalan)",
        word_2 = "Correct translation (English)",
        freq_zipf_2 = str_wrap("Translation lexical frequency (Zipf)", 12),
        lv = "Levenshtein similarity",
        lv_dist = "Levenshtein distance",
        duration = "Audio duration (s)",
        nd = "N",
        nd_list = "Neighbours"
    ) %>% 
    tab_style(
        cell_text(align = "left", weight = "bold"),
        cells_column_spanners()
    ) %>% 
    tab_style(
        cell_text(style = "italic", align = "left"),
        list(cells_column_labels())
    ) %>% 
    tab_style(
        cell_text(align = "left"),
        cells_body()
    ) %>% 
    tab_style(
        cell_text(size = "x-small"),
        cells_body(columns = nd_list, rows = everything())
    )
```

### spa-ENG

```{r}
#| label: tbl-stimuli-spaeng
stimuli %>% 
    filter(group=="spa-ENG") %>% 
    mutate(
        word_1 = paste0(word_1, " (", ipa_flat_1, ")"),
        word_2 = paste0(word_2, " (", ipa_flat_2, ")"),
        # nd_list = paste0(nd_list, collapse = ", ")
    ) %>% 
    select(word_1, word_2, duration, freq_zipf_2, lv, lv_dist, nd, nd_list) %>% 
    gt() %>% 
    fmt_number(where(is.numeric)) %>% 
    fmt_integer(c(nd, lv_dist)) %>%
    fmt_percent(lv) %>% 
    fmt_missing(everything(), everything(), "---") %>% 
    tab_spanner("Cross-language phonological neighbours", nd:nd_list) %>% 
    tab_spanner("Phonological similarity", lv:lv_dist) %>% 
    cols_label(
        word_1 = "Presented word (Spanish)",
        word_2 = "Correct translation (English)",
        freq_zipf_2 = str_wrap("Translation lexical frequency (Zipf)", 12),
        lv = "Levenshtein similarity",
        lv_dist = "Levenshtein distance",
        duration = "Audio duration (s)",
        nd = "N",
        nd_list = "Neighbours"
    ) %>% 
    tab_style(
        cell_text(align = "left", weight = "bold"),
        cells_column_spanners()
    ) %>% 
    tab_style(
        cell_text(style = "italic", align = "left"),
        list(cells_column_labels())
    ) %>% 
    tab_style(
        cell_text(align = "left"),
        cells_body()
    ) %>% 
    tab_style(
        cell_text(size = "x-small"),
        cells_body(columns = nd_list, rows = everything())
    )
```

### cat-SPA

```{r}
#| label: tbl-stimuli-catspa
stimuli %>% 
    filter(group=="cat-SPA") %>% 
    mutate(
        word_1 = paste0(word_1, " (", ipa_flat_1, ")"),
        word_2 = paste0(word_2, " (", ipa_flat_2, ")"),
        # nd_list = paste0(nd_list, collapse = ", ")
    ) %>% 
    select(word_1, word_2, duration, freq_zipf_2, lv, lv_dist, nd, nd_list) %>% 
    gt() %>% 
    fmt_number(where(is.numeric)) %>% 
    fmt_integer(c(nd, lv_dist)) %>%
    fmt_percent(lv) %>% 
    fmt_missing(everything(), everything(), "---") %>% 
    tab_spanner("Cross-language phonological neighbours", nd:nd_list) %>% 
    tab_spanner("Phonological similarity", lv:lv_dist) %>% 
    cols_label(
        word_1 = "Presented word (Catalan)",
        word_2 = "Correct translation (Spanish)",
        freq_zipf_2 = str_wrap("Translation lexical frequency (Zipf)", 12),
        lv = "Levenshtein similarity",
        lv_dist = "Levenshtein distance",
        duration = "Audio duration (s)",
        nd = "N",
        nd_list = "Neighbours"
    ) %>% 
    tab_style(
        cell_text(align = "left", weight = "bold"),
        cells_column_spanners()
    ) %>% 
    tab_style(
        cell_text(style = "italic", align = "left"),
        list(cells_column_labels())
    ) %>% 
    tab_style(
        cell_text(align = "left"),
        cells_body()
    ) %>% 
    tab_style(
        cell_text(size = "x-small"),
        cells_body(columns = nd_list, rows = everything())
    )
```
:::

### Lexical frequency

```{r}
#| label: fig-stimuli-freq
stimuli %>% 
    distinct(group, word_1, .keep_all = TRUE) %>% 
    left_join(count(., group, name = "n_total")) %>% 
    mutate(test_language = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(freq_zipf_2, fill = group)) +
    facet_wrap(~group, ncol = 1) +
    geom_histogram(bins = 25, colour = "white") +
    labs(
        x = "Lexical frequency (Zipf score)\nExtracted from CLEARPOND database", 
        y = "# trials", 
        fill = "Group"
    )  +
    scale_y_continuous(breaks = seq(0, 12, 2)) +
    theme(
        legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

### PTHN

```{r}
#| label: fig-stimuli-pthn
#| fig-height: 10
stimuli %>% 
    left_join(count(., group, name = "n_total")) %>% 
    mutate(translation = paste0(word_1, "/", word_2)) %>% 
    ggplot(aes(x = nd, y = reorder(translation, nd), colour = group, fill = group, label = translation)) +
    facet_wrap(~group, scales = "free_y") +
    geom_text(size = 3, hjust = 0) +
    labs(
        x = "Group", 
        y = "# phonological neighbours", 
    ) +
    scale_x_continuous(breaks = seq(0, 30, 5)) +
    theme(
        legend.position = "none",
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank()
    )
```

### Levenshtein distance

```{r}
#| label: fig-stimuli-lv
#| fig-width: 9
#| fig-height: 12
stimuli %>% 
    left_join(count(., group, name = "n_total")) %>% 
    mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot() +
    aes(x = lv, y = reorder_within(translation, lv, group), fill = group, colour = group) +
    facet_wrap(~group, scales = "free_y") +
    geom_text(aes(label = percent(lv, accuracy = 0.1)), size = 3, hjust = 0) +
    labs(
        x = "Group", 
        y = "Phonological Leveshtein distance\n(corrected for string length)",
        fill = "Test language"
    ) +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    theme(
        legend.position = "none",
        axis.title.y = element_blank(),
        axis.text.y = element_text(size = 7, hjust = 1),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
    ) 
ggsave(here("img", "lv.png"), height = 12)
```

### Distributions

```{r}
#| label: fig-prepare
#| out-width: "100%"
#| fig-width: 7
#| fig-height: 5.5
#| fig-cap: "Distribution and summary statistics of stimuli translations, plotted in different panels for English translations of Catalan stimuli (cat-ENG), English translations of Spanish stimuli (spa-ENG), and Spanish translation of Catalan stimuli (cat-SPA). A\\) Lexical frequency of stimuli translations, expressed in counts per million and as Zipf scores. B\\) Number of higher-frequency phonological neighbourhoods (PTHN) of stimuli translations. C\\) Normalised Levenshtein phonological similarity between the presented stimuli and their translations, expressed in percentages. Lexical frequencies and PTHN scores of the stimuli (presented auditorily in Catalan or Spanish) are not described, as participants lack any lexical representation of such words."
medians <- stimuli %>% 
    select(group, word_1, freq_2, freq_zipf_2, nd, lv) %>% 
    group_by(group) %>% 
    summarise(across(freq_2:lv,list(median = ~median(., na.rm = TRUE),  sd = ~sd(., na.rm = TRUE))))

# frequency
stimuli %>% 
    ggplot() + 
    aes(x = freq_zipf_2, fill = group) + 
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(
            xmin = freq_zipf_2_median-freq_zipf_2_sd,
            xmax = freq_zipf_2_median+freq_zipf_2_sd, 
            ymin = -Inf,
            ymax = Inf, 
            fill = group
        ),
        inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_histogram(bins = 20, colour = "white") +
    geom_vline(
        data = medians, 
        aes(xintercept = freq_zipf_2_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed"
    ) + 
    geom_text(
        data = medians, 
        aes(label = "Median ± SD", x = freq_zipf_2_median+1.25, y = 13), 
        size = 3
    ) +
    geom_curve(
        data = medians, 
        aes(x = freq_zipf_2_median+0.85,  xend = freq_zipf_2_median+0.1,
            y = 12, yend = 11),
        arrow = arrow(length = unit(0.2, "cm")),
        curvature = -0.25, colour = "grey30"
    ) +
    labs(
        x = "Lexical frequency (Zipf)", 
        y = "Number of trials", 
        title = "Frequency\n(stimuli translations)\n"
    ) +

    # PTHN 1
    stimuli %>% 
    count(group, nd) %>% 
    ggplot(aes(x = nd, y = n, fill = group)) + 
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(xmin = 0, xmax = nd_median+nd_sd, ymin = -Inf, ymax = Inf, fill = group), 
        inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_col(colour = "white") + 
    geom_vline(
        data = medians, 
        aes(xintercept = nd_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed") + 
    labs(
        x = "PTHN across languages", 
        y = "Number of trials", 
        title = "Higher-frequency neighbours\n(stimuli translations)\n"
    ) +

    # Levenshtein
    stimuli %>% 
    ggplot(aes(lv, fill = group)) +
    facet_wrap(~group, ncol = 1) + 
    geom_rect(
        data = medians, 
        aes(xmin = 0, xmax = lv_median+lv_sd, ymin = -Inf, ymax = Inf, fill = group),
        inherit.aes = FALSE, alpha = 0.15
    ) +   
    geom_histogram(bins = 20, colour = "white") +
    geom_vline(
        data = medians,
        aes(xintercept = lv_median, group = group), 
        colour = "black", size = 0.5, linetype = "dashed"
    ) + 
    scale_x_continuous(labels = percent) +
    labs(
        x = "Levenshtein",
        y = "Number of trials",
        title = "Phonological similarity\n(stimuli and stimuli translations)\n"
    ) +

    plot_layout(guides = "collect", nrow = 1) &
    plot_annotation(tag_levels = "A") &
    scale_fill_manual(values = my_colours, na.translate = FALSE) &
    theme(legend.position = "none")

```

## Participants

```{r}
#| label: participants-coded
#| include: false
participants_coded <- participants %>% 
    mutate(valid_participant = ifelse(valid_participant, "Valid", "Not valid"))
```

We collected data from `r format(as_date(min(participants$date)), "%B %dth %Y")` to `r format(as_date(max(participants$date)), "%B %dth %Y")`. We tested `r length(unique(participants$participant_id))` participants. `r length(unique(participants$participant_id[participants$group=="spa-ENG"]))` were English natives tested in Spanish, `r length(unique(participants$participant_id[participants$group=="cat-ENG"]))` were English natives tested in Catalan, and `r length(unique(participants$participant_id[participants$group=="cat-SPA"]))` were Spanish natives tested in Catalan. `r sum(!participants$valid_participant)` failed to meet all the inclusion criteria and were excluded from further analyses. The final sample comprised data from `r sum(!participants$valid_participant)` participants, `r sum(participants$valid_participant[participants$group=="spa-ENG"])` in the spa-ENG group, `r sum(participants$valid_participant[participants$group=="cat-ENG"])` in the cat-ENG group, and `r sum(participants$valid_participant[participants$group=="cat-SPA"])` in the cat-SPA group.

Participants were included if^[We originally planned to exclude participants that reported any visual impairment that glasses would not correct. This item was phrased as Do you have normal or corrected-to-normal VISION? (Yes/No) in English, and as ¿Tienes problemas de VISIÓN que unas gafas o lentes de contacto NO corrijan? (Sí/No). Surprisingly, the proportion of Spanish participants that reported visual impairment was implausibly large (n = 6, 18.18%). This is possibly due to some participants using glasses daily and not having read the item until the end, where it is indicated that the use of glasses is considered as normal vision]:

* Aged 18 to 25 years
* Did not report being fluent in Catalan, Spanish, or Italian
* Provided at least 80% of the presented trials
* Did not report motor, auditory or visual (other than wearing glasses) problems

Trials were considered valid if:

* Participant did not leave a blank response
* Participant took longer than 10 seconds to respond

### Summary

```{r}
#| label: fig-participants-summary
participants_coded %>% 
    count(group, valid_participant) %>% 
    mutate(valid_participant = factor(valid_participant)) %>% 
    ggplot() +
    aes(group, n, fill = valid_participant, order = n) + 
    geom_col() +
    labs(
        x = "Group", 
        y = "# participants", 
        fill = "Valid?"
    ) +
    theme(
        legend.title = element_blank(),
        legend.position = "top",
        axis.title.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )

```

### Valid trials

```{r}
#| label: fig-participants-trials
participants_coded %>% 
    filter(valid_participant=="Valid") %>% 
    select(participant_id, group, n, n_valid) %>%
    mutate(valid_prop = n_valid/n) %>% 
    ggplot() +
    aes(x = valid_prop, fill = group, colour = group) + 
    facet_wrap(~group) +
    geom_histogram() +
    labs(
        x = "Valid trials (%)",
        y = "N participants",
        fill = "Group",
        colour = "Group"
    ) +
    scale_x_continuous(labels = percent) +
    theme(
        legend.title = element_blank(),
        legend.position = "none",
        axis.title.x = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )
```

### Demographic information

```{r}
#| label: fig-participants-demo
participants_valid <- filter(participants_coded, valid_participant=="Valid")

p_sex <- participants_valid %>% 
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
    ggplot() +
    aes(x = group, fill = sex) +
    geom_bar(position = position_fill()) +
    labs(
        x = "Group", 
        y = "# participants", 
        fill = "Sex"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "top",
        axis.title.x = element_blank()
    )

p_age <- participants_valid %>% 
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
    ggplot(aes(as.factor(age), fill = group, colour = group)) +
    facet_wrap(~group, ncol = 1) +
    geom_bar(colour = "white") +
    labs(
        x = "Age (years)", 
        y = "% participants", 
        fill = "Group", 
        colour = "Group"
    ) +
    theme(legend.position = "none")

(p_sex + p_age) &
    plot_layout() &
    plot_annotation(tag_levels = "A") &
    theme(
        legend.title = element_blank(),
        panel.grid.major.y = element_line(linetype = "dotted", colour = "grey")
    )

```

### Second language

```{r}
#| label: fig-participants-l2
participants_valid %>% 
    count(group, l2, .drop = FALSE) %>% 
    mutate(
        l2 = case_when(
            l2=="Alemã\u0081n" ~ "German",
            l2=="Francã\u2030s" ~ "French",
            l2=="Cuã\u0081l" ~ "Missing",
            TRUE ~ l2
        ),
    ) %>% 
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group,  "\n(N = ", n_total, ")")) %>% 
    ggplot() +
    aes(group, n, fill = l2) + 
    facet_wrap(~group, ncol = 1, scales = "free_y") +
    geom_col(position = position_fill()) +
    labs(
        x = "L2", 
        y = "# participants", 
        fill = "L2"
    ) +
    scale_y_continuous(labels = percent) +
    coord_flip() +
    theme(
        legend.position = "top",
        legend.title = element_blank(),
        strip.text = element_blank(),
        panel.grid = element_blank(),
        axis.text.x = element_text(),
        axis.title.y = element_blank()
    )

```

### Language profile proficiency

```{r}
#| label: fig-participants-proficiency
participants_valid %>% 
    select(participant_id, group, catalan_oral, spanish_oral, catalan_written, spanish_written) %>% 
    mutate(across(c(catalan_oral, spanish_oral, catalan_written, spanish_written), replace_na, 0)) %>% 
    pivot_longer(
        c(catalan_oral, spanish_oral, catalan_written, spanish_written),
        names_to = "type",
        values_to = "score"
    ) %>% 
    separate(type, c("language", "modality"), sep = "_") %>% 
    mutate(across(c(language, modality), str_to_sentence)) %>%
    left_join(count(participants_valid, group, name = "n_total")) %>% 
    mutate(group = paste0(group, " (N = ", n_total, ")")) %>% 
    ggplot(aes(as.integer(score), fill = language)) +
    facet_grid(modality~group) +
    geom_bar(position = position_dodge(), colour = "white") +
    labs(
        y = "# participants", 
        x = "Self-reported proficiency (0-5)",
        fill = "Language"
    ) +
    theme(
        legend.position = "top",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
    )

```

## Data analysis

We fitted a Bayesian multilevel regression model using the R package `brms` [@burkner2017brms], with correct responses (`correct`, 0 is incorrect response, 1 if correct response) as the response variable with a Bernoulli distribution and a logit link function. We modelled the probability of an average participant providing a correct response, conditional to a set of predictors (aka. fixed effects) and their interactions:

* PTHN (`pthn`, 0-Inf): number of phonological neighbours with higher frequency of the target word. Extracted from the CLEARPOND database. Standardised before entering the regression model.
* Phonological similarity (`lv`, 0-1): Inverse Levenshtein distance between the IPA transcription of the presented word and its correct translation in the target language, calculated using the `stringdist` function of the `stringdist` package [@loo2014stringdist]. Standardised before entering the regression model.
* Shared phonological onset (`onset`, Same/Different): Whether the presented and the target word share phonological onset, as judged by experimenters, and sum coded as $Different = -0.5$ and $Same = +0.5$.

We added participants (`participants`) and as a grouping variable (aka., random effects), therefore adding participant-level adjustments to the model. We specified random intercepts slopes for all fixed effects within the `participant` grouping variable.

We performed multiple imputation via predictive mean matching using the `mice` function of the `mice` package [@van2011mice] to impute missing values in the response variable or the predictors.

The model was implemented using the following formula (`fit_3`):

$$\begin{align}

&\textbf{Likelihood}  \\
y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\

&\textbf{Parameters}  \\

logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
& \beta_{1[p]} ~ Frequency_{i} ~ + \\
& \beta_{2[p]} ~ PTHN_i ~ + \\
& \beta_{3[p]} ~ Similarity_i ~ + \\
& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\

\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
&& \text{[participant-level coefficients]} \\ \\

&\textbf{Prior}  \\

\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\

\end{align}$$

The estimation of the model was performed using Bayesian inference via Hamiltonian Monte-Carlo in Stan [@carpenter2017stan], with 4 chains, 2,000 iterations each (1,000 warm-ups iterations). The output of this model is the approximated posterior distribution of all the parameters estimated, therefore indicating the probability of each combination of values of the parameters, conditional to the observed data.

We compared this (extended model) to other models dropping one predictor at a time using Leave-one-out cross-validation (LOO-CV), using the `loo` and `loo_compare` functions of the `brms` package. We compared our extended model against the following models:

* fit_0: `~ 1 + (1 | participant) + (1 | word)`
* fit_1: `~ 1 + frequency_zipf + (1 + frequency_zipf | participant) + (1 | word)`
* fit_2: `~ 1 + frequency_zipf + pthn + (1 + frequency_zipf + pthn | participant) + (1 | word)`
* fit_3: `~ 1 + frequency_zipf + pthn + lv + (1 + frequency_zipf + pthn + lv | participant) + (1 | word)`
* fit_4: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_5: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`
* fit_6: `~ 1 + frequency_zipf + pthn + lv + pthn:lv + group + group:lv + (1 + frequency_zipf + pthn + pthn:lv + lv | participant) + (1 | word)`

## Results

### Raw data

### Model outputs

```{r}
#| label: responses-coded
responses_coded <- responses %>% 
    # typos are considered correct responses
    mutate(
        group = as.factor(group), 
        # center predictors
        across(c(lv_std, nd_std, freq_zipf_2_std), ~scale(., center = TRUE, scale = TRUE)[,1])
    ) %>% 
    # impute missing data
    as_tibble() %>% 
    arrange(group, word)
```

#### Model comparison

Leave-one-out cross-validation. The more negative is `elpd_diff` (and the larger its magnitude compared to its corresponding standard error), the better it first the data compared to the null model (`~ 1 + (1 | participant)`). The model that fits the data the best is the extended model.

```{r}
#| label: tbl-model-comparison
tab_subtitle <- md("Model performance and comparison using leave-one-out cross-validation")

tab_caption  <- md("*Note*. Outcomes of the leave-one-out cross-validation procedure. Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model. Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")

model_loos %>% 
    as.data.frame() %>% 
    rownames_to_column("model") %>% 
    arrange(desc(elpd_loo)) %>% 
    select(model, matches("elpd_loo"), matches("looic"), matches("diff") ) %>%
    mutate(
        model = str_replace(model, "fit_", "Model "),
        across(everything(), ~na_if(., 0))
    ) %>% 
    gt(rowname_col = "model") %>% 
    tab_header(title = md("Table 2"), subtitle = tab_subtitle) %>% 
    tab_source_note(tab_caption) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_number(2:7) %>% 
    cols_label(
        model = md("**Model**"),
        looic = md("***LOO<sub>IC</sub>***"),
        se_looic = md("***SE*<sub>IC</sub>**"),
        elpd_loo = md("***LOO<sub>ELPD</sub>***"),
        se_elpd_loo = md("***SE***"),
        elpd_diff = md("***LOO<sub>diff</sub>***"),
        se_diff = md("***SE<sub>diff</sub>***")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = list(
            cell_borders(sides = c("left", "right"), color = "white"),
            cell_text(align = "center")
        ),
        locations = list(
            cells_body(), 
            cells_row_groups(), 
            cells_column_spanners(), 
            cells_stub(),
            cells_source_notes(),
            cells_column_labels(),
            cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(
            cell_text(align = "left", size = "medium"), 
            cell_borders(sides = "all", color = "white")
        ),
        locations = list(
            cells_stub(), 
            cells_source_notes()
        )
    ) %>% 
    tab_style(
        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
        locations = list(
            cells_column_spanners(),
            cells_column_labels()
        )
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:7)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:7)
    )
```

#### Regression coefficients

```{r}
#| label: tbl-model-coefs
coefs <- fixef(fit_4) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)))

str_repl <- c(
    "Intercept" = "Intercept",
    "freq_zipf_2_std" = "Frequency [+1 SD]",
    "nd_std" = "Neighbours [+1 SD]",
    "lv_std" = "Levenshtein [+1 SD]",
    "nd_std:lv_std" = "Neigbours \u00d7 Cognateness"
)

tab_caption <- md("Estimated posterior distributions of coefficients in Model 6. A\\) Population-level effects. Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects. Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability. Black dots represent the mean of each distribution.")

coefs %>% 
    mutate(
        variable = variable %>% 
            str_remove("b_") %>% 
            factor(levels = names(str_repl), labels = str_repl)
    ) %>% 
    select(-est_error) %>% 
    gt() %>% 
    fmt_percent(2:4) %>% 
    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
    cols_label(
        variable = md("**Predictor**"),
        estimate = md("**Mean**"),
        q2_5 = md("**95\\% CrI**"),
    ) %>% 
    tab_source_note(tab_caption) %>% 
    cols_align(align = "left")
```

```{r}
#| label: fig-coefs-fixed
#| fig-height: 7
#| fig-width: 6.5
#| fig-cap: "Estimated posterior distributions of coefficients in Model 4. A\\) Population-level effects. Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects. Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability. Black dots represent the mean of each distribution. B\\) Participant\\-level coefficient variability. The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant. Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*). C\\) Correlation between participant\\-level effects. The model allowed participant-level coefficients to co\\-vary. This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation."
coefs <- fixef(fit_4) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4))) %>% 
    group_split(variable) %>% 
    set_names(make_clean_names(map(., "variable"))) %>% 
    map(select, -variable) %>% 
    map(unlist)

post_draws <-  gather_draws(fit_4, `b_.*`, `sd_.*`, regex = TRUE)

# fixed effects
post_draws %>% 
    filter(str_detect(.variable, "b_")) %>% 
    mutate(
        .variable_name = .variable %>% 
            str_remove("b_") %>% 
            factor(levels = names(str_repl), labels = str_repl),
        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
    ) %>% 
    arrange(.variable) %>% 
    ggplot() +
    aes(.value, fct_rev(.variable_name)) +
    geom_vline(xintercept = 0) +
    stat_slab(fill = my_colours[1], colour = my_colours[1]) +
    stat_pointinterval() +
    scale_fill_manual(values = my_colours, na.translate = FALSE) +
    scale_x_continuous(labels = function(x) percent(round(x, 1))) +
    labs(
        x = "P(Correct)", 
        y = "Posterior probability density",
        fill = "CrI"
    ) +
    theme(
        legend.position = c(1, 0.25),
        legend.direction = "vertical",
        legend.background = element_rect(fill = "grey95"),
        legend.key = element_rect(fill = "grey95", colour = "grey95"), 
        legend.key.height = unit(0.1, "cm"),
        axis.title.y = element_blank(),
        legend.text = element_text(size = 7),
        axis.text.x = element_text(colour = "black"),
        axis.text.y = element_text(colour = "black", hjust = 1),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_line(colour = "grey", size = 0.5)
    ) 
```

```{r}
#| label: coefs-word
empirical_means <- responses %>% 
    group_by(group) %>% 
    summarise(
        successes = sum(correct, na.rm = TRUE),
        n = n(),
        .groups = "drop"
    ) %>% 
    mutate(
        prop = prop_adj(successes, n),
        prop_se = prop_adj_se(successes, n)
    )

empirical_accuracy <- responses %>% 
    mutate(translation = paste0(word_1, " /", ipa_flat_1, "/ - ",  word_2, " /", ipa_flat_2, "/ - ", percent(lv, accuracy = 1))) %>% 
    group_by(group, translation, lv) %>% 
    summarise(
        successes = sum(correct, na.rm = TRUE),
        n = n(),
        .groups = "drop"
    ) %>% 
    rowwise() %>% 
    mutate(
        prop = prop_adj(successes, n),
        prop_se = prop_adj_se(successes, n), 
    )

fixed_intercept <- fixef(fit_4)[1 ,c(1, 3, 4)] %>% 
    inv_logit_scaled()

nd_re_words <- expand.grid(
    translation_id = unique(fit_4$data$translation_id),
    nd_std = 0,
    freq_zipf_2_std = 0,
    lv_std = 0,
    group = NA
)

post_re <- add_epred_draws(nd_re_words, fit_4, ndraws = 50,
                           re_formula = ~ (1 | translation_id)) %>% 
    group_by(translation_id, nd_std, freq_zipf_2_std, lv_std, group) %>% 
    summarise(
        .mean = mean_qi(.epred)[,1], 
        .lower = mean_qi(.epred)[,2],
        .upper = mean_qi(.epred)[,3],
        .groups = "drop"
    ) %>% 
    ungroup() %>% 
    select(-c(group)) %>% 
    left_join(select(stimuli, group, translation, translation_id, lv)) %>% 
    mutate(translation = paste0(translation, " - ", percent(lv, accuracy = 1))) %>% 
    left_join(select(empirical_accuracy, translation, n, prop, prop_se))

post_re %>% 
    ggplot() +
    aes(x = .mean, y = reorder_within(translation, prop, group)) +
    facet_wrap(~group, scales = "free_y") +
    geom_vline(data = empirical_means, aes(xintercept = prop), size = 0.75, colour = "grey") +
    geom_errorbar(aes(xmin = .lower, xmax = .upper), size = 0.5, colour = "grey", width = 0.5) + 
    geom_point(aes(x = .mean), size = 0.5, colour = "grey") + 
    geom_point(aes(x = prop, colour = group), size = 1, shape = 15) +
    labs(
        x = "Percentage of correct responses",
        y = "Item",
        title = "Observed translation accuracy",
        subtitle = "Dots indicate proportions of correct responses, error bars indicate +/-1 SE",
        colour = "Credible Interval (CrI)",
        caption = "Percentages next to items indicate the Levenshtein distance between both translations."
    ) +
    scale_x_continuous(limits = c(0, 1), labels = percent) +
    scale_y_reordered() +
    theme(
        legend.position = "none",
        legend.key = element_rect(fill = NA, colour = NA),
        axis.text.x = element_text(size = 7),
        axis.text.y = element_text(size = 5, hjust = 1),
        axis.ticks = element_blank(),
        axis.title = element_text(size = 10),
        axis.title.y = element_blank(),
        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
        plot.caption.position = "plot",
        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
        plot.title.position = "plot",
        plot.title = element_text(hjust = 0, size = 13),
        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
    )

ggsave(here("img", "accuracy.png"), height = 9, width = 10, dpi = 1000)
```

#### Marginal effects

```{r}
#| label: fig-marginal-effects
#| fig-width: 7
#| fig-height: 3.5
#| fig-cap: "Expected mean posterior predictions from Model 4. A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively. We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein). We did this across the range of values of the PTHN scores. For each simulation, we drew a single sample from the posterior distribution of each coefficient. Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations. Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition. The dispersion of the lines indicates the uncertainty of our predictions. We computed these posterior predictions for each group of participants, and plot tem in separate panels."
rescale_values <- function(x, vct) (x*sd(vct))+mean(vct)
scale_values <- function(x, vct) (x-mean(vct))/sd(vct)

nd <- expand.grid(
    lv_std = seq(
        min(fit_4$data$lv_std, na.rm = TRUE),
        max(fit_4$data$lv_std, na.rm = TRUE),
        by = 0.1
    ),
    freq_zipf_2_std = 0,
    nd_std = scale_values(c(0, 3, 6, 10, 15, 25), fit_4$data$nd)
)

m <- add_epred_draws(nd, fit_4, re_formula = NA, ndraws = 50) %>% 
    mutate(nd_lab = rescale_values(nd_std, fit_4$data$nd) %>% 
               paste0(" neighbours") %>% 
               factor(levels = paste0(c(0, 3, 6, 10, 15, 25), " neighbours"))
    )

ggplot(m, aes(x = lv_std, y = .epred)) +
    facet_wrap(~nd_lab, nrow = 1) +
    geom_line(aes(group = interaction(nd_std, .draw)),
              colour = "grey", size = 0.75, alpha = 0.25) +
    geom_hline(yintercept = 0.5, colour = "grey", size = 1) +
    stat_summary(fun = mean, geom = "line", size = 1, colour = my_colours[2]) +
    labs(
        x = "Phonological similarity (Levenshtein)",
        y = "P(Correct)",
        title = "Population-level posterior predictions"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        axis.text.x = element_text(size = 10),
        plot.title = element_blank()
    )
```

# Study 2

## Participants

### Language condition

```{r}
#| label: participants-language
#| warning: false
# participants language condition
questionnaire_participants %>% 
    count(language) %>% 
    ggplot(aes(language, n, colour = language, fill = language, label = n)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(y = n+2), position = position_dodge(width = 1), size = 4, show.legend = FALSE, hjust = 0.5) +
    labs(x = "Educational attainment", y = "# participants", colour = "Language", fill = "Language") +
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    theme_custom() +
    theme(legend.position = "none")

```

## Age

```{r}
#| label: participants-age
#| warning: false
questionnaire_participants %>% 
    count(age, language) %>% 
    ggplot(aes(age, n, colour = language, fill = language, label = n)) +
    facet_wrap(~language) +
    geom_col() +
    geom_text(aes(y = n+0.5), position = position_dodge(width = 1), size = 4, show.legend = FALSE, hjust = 0.5) +
    labs(x = "Age (years)", y = "# participants", colour = "Language", fill = "Language") +
    scale_y_continuous(breaks = 0:10) +
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    theme_custom() +
    theme(legend.position = "top")
```

## Gender

```{r}
#| label: participants-gender
#| warning: false
questionnaire_participants %>% 
    count(language, sex) %>% 
    ggplot(aes(sex, n, colour = language, fill = language, label = n)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(y = n+1), position = position_dodge(width = 1), size = 4, show.legend = FALSE, hjust = 0.5) +
    labs(x = "Gender", y = "# participants", colour = "Language", fill = "Language") +
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    theme_custom() +
    theme(legend.position = "top")
```

### Educational attainment

```{r}
#| label: participants-education
#| warning: false
questionnaire_participants %>% 
    count(language, education) %>% 
    mutate(education = str_wrap(education, width = 23)) %>% 
    ggplot(aes(education, n, colour = language, fill = language, label = n)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(y = n+1), position = position_dodge(width = 1), size = 4, show.legend = FALSE, hjust = 0.5) +
    labs(x = "Educational attainment", y = "# participants", colour = "Language", fill = "Language") +
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    theme_custom() 
```

### Language profile

```{r}
#| label: participants-profile
#| warning: false
questionnaire_participants %>% 
    pivot_longer(c(l_1, l_2, l_3), names_to = "language_number", values_to = "language_response") %>% 
    count(language, language_number, language_response) %>% 
    ggplot(aes(reorder(language_response, -n), n, colour = language, fill = language, label = n)) +
    facet_wrap(~language_number) +
    geom_col(position = position_dodge()) +
    geom_text(aes(y = n+1), position = position_dodge(width = 1), size = 4, show.legend = FALSE, hjust = 0.5) +
    labs(x = "Familiar languages", y = "# participants", colour = "Language", fill = "Language") +
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    theme_custom() +
    theme(
        legend.position = "top",
        axis.text.x = element_text(angle = 30, hjust = 1)
    )
```

### Number of trials

```{r}
#| label: participants-trials
#| warning: false
questionnaire_participants %>% 
    count(language, num_trials) %>% 
    mutate(num_trials = as.factor(num_trials)) %>% 
    ggplot(aes(num_trials, n, colour = language, fill = language, label = n)) +
    geom_col(position = position_dodge()) +
    geom_text(aes(y = n+1), position = position_dodge(width = 1), size = 4, show.legend = FALSE, hjust = 0.5) +
    labs(x = "Completed trials", y = "# participants", colour = "Language", fill = "Language") +
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    theme_custom() +
    theme(
        legend.position = "top"
    )
```

## Results

### Raw data

#### The relationship between confidence and knowledge

```{r}
#| label: fig-confidence-knowledge
#| warning: false
#| fig-height: 6

# plot confidence ratings by knowledge
questionnaire_responses %>% 
    left_join(select(questionnaire_data_clean, participant_id, word_1, knowledge, confidence)) %>% 
    count(knowledge, confidence, group) %>% 
    replace_na(list(n = 0)) %>% 
    ggplot(aes(confidence, n, fill = knowledge, colour = knowledge, label = format(n, big.mark = ","))) +
    facet_wrap(~group) +
    geom_col(position = position_dodge(width = 1), show.legend = FALSE) +
    geom_text(aes(y = n+70), position = position_dodge(width = 1), size = 3, show.legend = FALSE, hjust = 0.5) +
    scale_y_continuous(labels = ~format(., big.mark = ",")) +
    theme(
        axis.title.x = element_blank()
    ) +

    questionnaire_responses %>%  
    left_join(select(questionnaire_data_clean, participant_id, word_1, knowledge, confidence)) %>% 
    count(knowledge, confidence, group) %>%

    ggplot(aes(confidence, n, fill = knowledge, colour = knowledge, label = n)) +
    facet_wrap(~group) +
    geom_col(position = position_fill()) +
    scale_y_continuous(labels = percent) +

    plot_layout(nrow = 2) &
    labs(x = "Confidence rating", y = "Responses", colour = "Knows the word", fill = "Knows the word") &
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) &
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) &
    scale_x_continuous(breaks = 0:7) &
    theme_custom() &
    theme(legend.position = "top")
```

#### Participant-level accuracy

```{r}
#| label: results-participant-accuracy
#| warning: false
#| fig-height: 7
# plot participant accuracy
questionnaire_responses %>% 
    group_by(participant_id, group) %>% 
    summarise(
        sum_correct = sum(correct),
        num_trials = n(),
        .groups = "drop"
    ) %>% 
    mutate(
        prop = sum_correct/num_trials,
        prop_se = prop_adj_se(sum_correct, num_trials)
    ) %>% 
    ggplot(aes(prop, reorder(participant_id, prop), colour = group)) +
    facet_wrap(~group, scales = "free_y") + 
    geom_errorbar(aes(xmin = prop-prop_se, xmax = prop+prop_se), width = 0.5) +
    geom_point() +
    labs(x = "Proportion of correct translations", y = "Participant", colour = "Language") +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    scale_x_continuous(labels = percent) +
    theme_custom() +
    theme(
        legend.position = "top",
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.text.y = element_text(size = 7, hjust = 1),
        axis.ticks.y = element_blank()
    )
```

#### Word-level accuracy

```{r}
#| label: results-word-accuracy
#| warning: false
#| fig-height: 10
# plot word accuracy
questionnaire_responses %>% 
    group_by(word_1, group) %>% 
    summarise(
        sum_correct = sum(correct),
        num_trials = n(),
        .groups = "drop"
    ) %>% 
    mutate(
        prop = sum_correct/num_trials,
        prop_se = prop_adj_se(sum_correct, num_trials)
    ) %>% 
    ggplot(aes(prop, reorder_within(word_1, prop, group), colour = group)) +
    facet_wrap(~group, scales = "free_y") +
    geom_errorbar(aes(xmin = prop-prop_se, xmax = prop+prop_se), width = 0.5) +
    geom_point() +
    labs(x = "Proportion of correct translations", y = "Word", colour = "Language") +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    scale_x_continuous(labels = percent) +
    scale_y_reordered() +
    theme_custom() +
    theme(
        legend.position = "top",
        axis.text.y = element_text(size = 7, hjust = 1),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.ticks.y = element_blank()
    )
```

#### Accuracy by confidence and knowledge

```{r}
#| label: results-accuracy-confidence
#| warning: false
# plot response confidence
questionnaire_responses %>%
    mutate(correct = as.integer(correct)) %>% 
    ggplot(aes(confidence, correct, colour = group, fill = group)) +
    facet_grid(~knowledge, labeller = label_both) +
    geom_point(
        position = position_jitterdodge(jitter.width = 0.15, jitter.height = 0.1, dodge.width = 0.3),
        alpha = 0.25, size = 0.5
    ) +
    geom_smooth(method = "glm", method.args = list(family = "binomial")) +
    labs(x = "Confidence", y = "P(correct)", colour = "Language", fill = "Language") +
    scale_colour_manual(values = c("#1A85FF", "#ff2976")) +
    scale_fill_manual(values = c("#1A85FF", "#ff2976")) +
    scale_x_continuous(breaks = 0:7) +
    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.25)) +
    theme_custom() +
    theme(
        legend.position = "top"
    )
```

### Model outputs

#### Model comparisons

Leave-one-out cross-validation. The more negative is `elpd_diff` (and the larger its magnitude compared to its corresponding standard error), the better it first the data compared to the null model (`~ 1 + (1 | participant)`). The model that fits the data the best is the extended model.

```{r}
#| label: tbl-model-comparison-2
tab_subtitle <- md("Model performance and comparison using leave-one-out cross-validation")

tab_caption  <- md("*Note*. Outcomes of the leave-one-out cross-validation procedure. Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model. Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")

questionnaire_model_loos %>% 
    as.data.frame() %>% 
    rownames_to_column("model") %>% 
    arrange(desc(elpd_loo)) %>% 
    select(model, matches("elpd_loo"), matches("looic"), matches("diff") ) %>%
    mutate(
        model = str_replace(model, "questionnaire_fit_", "Model "),
        across(everything(), ~na_if(., 0))
    ) %>% 
    gt(rowname_col = "model") %>% 
    tab_header(title = md("Table 2"), subtitle = tab_subtitle) %>% 
    tab_source_note(tab_caption) %>% 
    fmt_missing(everything(), missing_text = "-") %>% 
    fmt_number(2:7) %>% 
    cols_label(
        model = md("**Model**"),
        looic = md("***LOO<sub>IC</sub>***"),
        se_looic = md("***SE*<sub>IC</sub>**"),
        elpd_loo = md("***LOO<sub>ELPD</sub>***"),
        se_elpd_loo = md("***SE***"),
        elpd_diff = md("***LOO<sub>diff</sub>***"),
        se_diff = md("***SE<sub>diff</sub>***")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", weight = "bold", size = "medium"),
        locations = cells_title(groups = "title")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium", style = "italic"),
        locations = cells_title(groups = "subtitle")
    ) %>% 
    tab_style(
        style = cell_text(align = "left", size = "medium"),
        locations = cells_source_notes()
    ) %>% 
    tab_style(
        style = list(
            cell_borders(sides = c("left", "right"), color = "white"),
            cell_text(align = "center")
        ),
        locations = list(
            cells_body(), 
            cells_row_groups(), 
            cells_column_spanners(), 
            cells_stub(),
            cells_source_notes(),
            cells_column_labels(),
            cells_stubhead()
        )
    ) %>% 
    tab_style(
        style = list(
            cell_text(align = "left", size = "medium"), 
            cell_borders(sides = "all", color = "white")
        ),
        locations = list(
            cells_stub(), 
            cells_source_notes()
        )
    ) %>% 
    tab_style(
        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
        locations = list(
            cells_column_spanners(),
            cells_column_labels()
        )
    ) %>% 
    tab_style(
        style = cell_borders(sides = "all", color = "white"),
        locations = cells_body(columns = 1:7)
    ) %>% 
    tab_style(
        style = cell_borders(sides = "top", color = "white"),
        locations = cells_body(columns = 1:7)
    )
```

#### Regression coefficients

```{r}
#| label: tbl-questionnaire-model-coefs
#| warning: false
coefs <- fixef(questionnaire_fit_4) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)))

tab_caption <- md("Estimated posterior distributions of coefficients in Model 6. A\\) Population-level effects. Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects. Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability. Black dots represent the mean of each distribution.")

coefs %>% 
    mutate(
        variable = variable %>% 
            str_remove("b_") %>% 
            factor(levels = names(str_repl), labels = str_repl)
    ) %>% 
    select(-est_error) %>% 
    gt() %>% 
    fmt_percent(2:4) %>% 
    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
    cols_label(
        variable = md("**Predictor**"),
        estimate = md("**Mean**"),
        q2_5 = md("**95\\% CrI**"),
    ) %>% 
    tab_source_note(tab_caption) %>% 
    cols_align(align = "left")
```

```{r}
#| label: fig-questionnaire-coefs-fixed
#| fig-height: 7
#| fig-width: 6.5
#| fig-cap: "Estimated posterior distributions of coefficients in Model 4. A\\) Population-level effects. Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects. Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability. Black dots represent the mean of each distribution. B\\) Participant\\-level coefficient variability. The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant. Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*). C\\) Correlation between participant\\-level effects. The model allowed participant-level coefficients to co\\-vary. This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation."
coefs <- fixef(questionnaire_fit_4) %>% 
    as.data.frame() %>% 
    rownames_to_column("variable") %>% 
    clean_names() %>% 
    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4))) %>% 
    group_split(variable) %>% 
    set_names(make_clean_names(map(., "variable"))) %>% 
    map(select, -variable) %>% 
    map(unlist)

post_draws <-  gather_draws(questionnaire_fit_4, `b_.*`, `sd_.*`, regex = TRUE)

# fixed effects
post_draws %>% 
    filter(str_detect(.variable, "b_")) %>% 
    mutate(
        .variable_name = .variable %>% 
            str_remove("b_") %>% 
            factor(levels = names(str_repl), labels = str_repl),
        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
    ) %>% 
    arrange(.variable) %>% 
    ggplot() +
    aes(.value, fct_rev(.variable_name)) +
    geom_vline(xintercept = 0, size = 1, colour = "grey") +
    stat_slab(fill = my_colours[1], colour = my_colours[1]) +
    stat_pointinterval() +
    labs(
        x = "P(Correct)", 
        y = "Posterior probability density",
        fill = "CrI"
    ) +
    scale_x_continuous(labels = function(x) percent(round(x, 1)), breaks = seq(-0.1, 0.5, 0.1)) +
    theme(
        axis.title.y = element_blank(),
        axis.text = element_text(colour = "black", hjust = 1, size = 12),
        axis.line = element_line(size = 0.75),
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.ticks.y = element_blank(),
        panel.background = element_blank(),
        panel.border = element_blank()

    ) 
```

#### Marginal effects

```{r}
#| label: questionnaire-marginal-effects
#| fig-width: 7
#| fig-height: 3.5
#| fig-cap: "Expected mean posterior predictions from Model 4. A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively. We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein). We did this across the range of values of the PTHN scores. For each simulation, we drew a single sample from the posterior distribution of each coefficient. Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations. Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition. The dispersion of the lines indicates the uncertainty of our predictions. We computed these posterior predictions for each group of participants, and plot tem in separate panels."
questionnaire_nd <- expand.grid(
    lv_std = seq(
        min(questionnaire_fit_4$data$lv_std, na.rm = TRUE),
        max(questionnaire_fit_4$data$lv_std, na.rm = TRUE),
        by = 0.1
    ),
    freq_zipf_2_std = 0,
    nd_std = scale_values(c(0, 3, 6, 10, 15, 25), questionnaire_fit_4$data$nd)
)

questionnaire_m <- add_epred_draws(nd, questionnaire_fit_4, re_formula = NA, ndraws = 50) %>% 
    mutate(nd_lab = rescale_values(nd_std, questionnaire_fit_4$data$nd) %>% 
               paste0(" neighbours") %>% 
               factor(levels = paste0(c(0, 3, 6, 10, 15, 25), " neighbours"))
    )

ggplot(questionnaire_m, aes(x = lv_std, y = .epred)) +
    facet_wrap(~nd_lab, nrow = 1) +
    geom_line(aes(group = interaction(nd_std, .draw)),
              colour = "grey", size = 0.75, alpha = 0.25) +
    geom_hline(yintercept = 0.5, colour = "grey", size = 1) +
    stat_summary(fun = mean, geom = "line", size = 1, colour = my_colours[2]) +
    labs(
        x = "Phonological similarity (Levenshtein)",
        y = "P(Correct)",
        title = "Population-level posterior predictions"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        axis.text.x = element_text(size = 10),
        plot.title = element_blank()
    )
```

# Joint analyses

```{r}
#| label: fig-joint-marginal-effects
#| fig-width: 9
#| fig-height: 4
#| fig-cap: "Expected mean posterior predictions from Model 4. A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively. We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein). We did this across the range of values of the PTHN scores. For each simulation, we drew a single sample from the posterior distribution of each coefficient. Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations. Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition. The dispersion of the lines indicates the uncertainty of our predictions. We computed these posterior predictions for each group of participants, and plot tem in separate panels."

bind_rows(list("Study 1" = m, "Study 2" = questionnaire_m), .id = "study") %>% 
    ggplot(aes(x = lv_std, y = .epred, colour = study)) +
    facet_wrap(~nd_lab, nrow = 1) +
    geom_line(aes(group = interaction(nd_std, .draw, study)),
              size = 0.75, alpha = 0.05) +
    geom_hline(yintercept = 0.5, size = 1) +
    stat_summary(fun = mean, geom = "line", size = 1) +
    labs(
        x = "Phonological similarity (Levenshtein)",
        y = "P(Correct)",
        title = "Population-level posterior predictions"
    ) +
    scale_y_continuous(labels = percent) +
    theme(
        legend.position = "none",
        axis.text.x = element_text(size = 10),
        plot.title = element_blank()
    )
```

### 2022-10-28 manuscript.qmd
Add questionnaire data and analyses, and change reports to Quarto

--- 2022-04-28 manuscript/manuscript.Rmd
@@
 
-shorttitle: Cognateness and non-native word recognition
-
 author:
   - name: Gonzalo Garcia-Castro
-  affiliation: '1'
-  corresponding: yes
-  address: Ramon Trias Fargas, 25-27, 08005 Barcelona
-  email: gonzalo.garciadecastro@upf.edu
+    affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
   - name: Serene Siow
-  affiliation: '2'
+    affiliation: Department of Experimental Psychology, University of Oxford
   - name: Nuria Sebastian-Galles
-  affiliation: '1'
+    affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
   - name: Kim Plunkett
-  affiliation: '2'
-affiliation:
-  - id            : "1"
-    institution   : "Center for Brain and Cognition, Universitat Pompeu Fabra"
-  - id            : "2"
-    institution   : "Department of Experimental Psychology, University of Oxford"
-
-authornote: |
+    affiliation: Department of Experimental Psychology, University of Oxford
 
@@
 
-  <!-- https://tinyurl.com/ybremelq -->
-
 keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian
-wordcount: 4553
 
@@
   - references.bib
-floatsintext: yes
-figurelist: no
-tablelist: no
-footnotelist: no
-linenumbers: yes
-mask: no
-draft: no
-numbersections: yes
+toc: false
+number-sections: true
 csl: "apa7.csl"
-documentclass: apa7
-classoption: man
-output: papaja::apa6_docx
+format:
+  pdf:
+    message: false
+    warning: false
+    results: "asis" 
+    fig-dpi: 1000
 ---
 
-```{r setup, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
+```{r}
+#| label: setup
+#| message: false
+#| warning: false
+#| include: false
 
 set.seed(42)
-knitr::opts_chunk$set(
-  message = FALSE,
-  warning = FALSE,
-  results = "asis", 
-  dpi = 1000
-)
 
@@
 
-```{r prepare, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
+```{r}
+#| label: prepare
+#| message: false
+#| warning: false
 # import objects
@@
     theme(
-    title = element_text(
-      size = 8, 
-      face = "bold"
-    ),
-    legend.background = element_rect(
-      fill = NA,
-      colour = NA
-    ),
-    legend.key = element_rect(
-      fill = "white", 
-      colour = NA),
+        title = element_text(size = 8, face = "bold"),
+        legend.background = element_rect(fill = NA, colour = NA),
+        legend.key = element_rect(fill = "white", colour = NA),
         legend.justification = "right",
         legend.direction = "horizontal",
-    axis.line = element_line(
-      colour = "black",
-      size = 0.5
-    ),
-    axis.ticks = element_line(
-      colour = "black", 
-      size = 0.5
-    ),
-    axis.text.x = element_text(
-      colour = "black"
-    ),
-    axis.text.y = element_text(
-      colour = "black"
-    ),
+        axis.line = element_line(colour = "black", size = 0.5),
+        axis.ticks = element_line(colour = "black", size = 0.5),
+        axis.text.x = element_text(colour = "black"),
+        axis.text.y = element_text(colour = "black"),
         plot.caption.position = "plot",
-    plot.title = element_text(
-      hjust = 0.5
-    ),
+        plot.title = element_text(hjust = 0.5),
         panel.grid = element_blank(),
-    panel.background = element_rect(
-      fill = "white", 
-      colour = NA
-    )
+        panel.background = element_rect(fill = "white", colour = NA)
     )
@@
 Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I) and by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
@@
 
-```{r procedurefigure, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="80%", fig.cap="Schematic representation of a trial in the experimental task."}
-
+```{r}
+#| label: procedure-figure
+#| echo: false
+#| message: false
+#| warning: false
+#| out-width: "80%"
+#| fig-cap: "Schematic representation of a trial in the experimental task."
 include_graphics(here("img", "design.png"))
-
 ```
@@
 
-```{r stimulilengths, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+```{r}
+#| label: stimuli-lengths
+#| message: false
+#| warning: false
 lengths <- stimuli %>% 
-  select(
-    group, 
-    ipa1,
-    ipa2, 
-    word1, 
-    word2
-  ) %>% 
-  mutate_at(
-    vars(ipa1:word2),
-    nchar
-  ) %>% 
+    select(group, ipa_flat_1, ipa_flat_2, word_1, word_2) %>% 
+    mutate_at(vars(ipa_flat_1:word_2), nchar) %>% 
     group_by(group) %>% 
     summarise_at(
-    vars(
-      ipa1, 
-      ipa2, 
-      word1, 
-      word2
-    ), 
-    list(
-      mean = mean, 
-      sd = sd,
-      min = min, 
-      max = max
-    )
+        vars(ipa_flat_1, ipa_flat_2, word_1, word_2), 
+        list(mean = mean, sd = sd, min = min, max = max)
     ) %>%
@@
 We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
-Words in the Catalan list were `r printnum(lengths$cat_eng["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$cat_eng["ipa1_sd"])`, Range = `r printnum(lengths$cat_eng["ipa1_min"], digits = 0)`-`r printnum(lengths$cat_eng["ipa1_max"], digits = 0)`).
-Words in the Spanish list were `r printnum(lengths$spa_eng["ipa1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$spa_eng["ipa1_sd"])`, Range = `r printnum(lengths$spa_eng["ipa1_min"], digits = 0)`-`r printnum(lengths$spa_eng["ipa1_max"], digits = 0)`).
+Words in the Catalan list were `r printnum(lengths$cat_eng["ipa_flat_1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$cat_eng["ipa_flat_1_sd"])`, Range = `r printnum(lengths$cat_eng["ipa_flat_1_min"], digits = 0)`-`r printnum(lengths$cat_eng["ipa_flat_1_max"], digits = 0)`).
+Words in the Spanish list were `r printnum(lengths$spa_eng["ipa_flat_1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$spa_eng["ipa_flat_1_sd"])`, Range = `r printnum(lengths$spa_eng["ipa_flat_1_min"], digits = 0)`-`r printnum(lengths$spa_eng["ipa_flat_1_max"], digits = 0)`).
 Participants listened to one audio file in each trial, each containing a single word presented in isolation.
@@
 
-Responses given by English participants for Catalan presented words were `r printnum(lengths$cat_eng["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_eng["word2_sd"])`, Range = `r printnum(lengths$cat_eng["word2_min"], digits = 0)`-`r printnum(lengths$cat_eng["word2_max"], digits = 0)`), while their translations to Spanish responses given by Spanish participants for Catalan words were `r printnum(lengths$cat_spa["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_spa["word2_sd"])`, Range = `r printnum(lengths$cat_spa["word2_min"], digits = 0)`-`r printnum(lengths$cat_spa["word2_max"], digits = 0)`).
-Responses given by English participants for Spanish words were `r printnum(lengths$spa_eng["word2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_eng["word2_sd"])`, Range = `r printnum(lengths$spa_eng["word2_min"], digits = 0)`-`r printnum(lengths$spa_eng["word2_max"], digits = 0)`).
+Responses given by English participants for Catalan presented words were `r printnum(lengths$cat_eng["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_eng["word_2_sd"])`, Range = `r printnum(lengths$cat_eng["word_2_min"], digits = 0)`-`r printnum(lengths$cat_eng["word_2_max"], digits = 0)`), while their translations to Spanish responses given by Spanish participants for Catalan words were `r printnum(lengths$cat_spa["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_spa["word_2_sd"])`, Range = `r printnum(lengths$cat_spa["word_2_min"], digits = 0)`-`r printnum(lengths$cat_spa["word_2_max"], digits = 0)`).
+Responses given by English participants for Spanish words were `r printnum(lengths$spa_eng["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_eng["word_2_sd"])`, Range = `r printnum(lengths$spa_eng["word_2_min"], digits = 0)`-`r printnum(lengths$spa_eng["word_2_max"], digits = 0)`).
 
@@
 
-```{r pthn_example, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+```{r}
+#| label: pthn-example
+#| message: false
+#| warning: false
 pthn_example <- clearpond_path %>% 
@@
     clean_names() %>% 
-  filter(
-    word %in% c("book", "sock"),
-    group=="cat-ENG"
-  ) %>% 
+    filter(word %in% c("book", "sock"), group=="cat-ENG") %>% 
     select(word, pthn)
@@
 
-```{r stimuliplot, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, out.width="100%", fig.width=7, fig.height=5.5, fig.cap="Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA).
+```{r}
+#| label: stimuli-plot
+#| message: false
+#| warning: false
+#| echo: false
+#| out-width: "100%"
+#| fig-width: 7
+#| fig-height: 5.5
+#| fig-cap: "Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA).
 Columns indicate: A\\) Lexical frequency of target translations, expressed as Zipf scores.
 B\\) PTHN of target translations, and C\\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages.
-Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words."}
-
-d <- stimuli %>% 
-  mutate(
-    group = factor(
-      group,
-      levels = c(
-        "cat-ENG",
-        "spa-ENG", 
-        "cat-SPA"
-      )
-    )
-  )
+Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words."
+d <-  mutate(stimuli, group = factor(group, levels = c("cat-ENG", "spa-ENG", "cat-SPA")))
 
@@
 d %>% 
-  ggplot() +
-  aes(
-    frequency_zipf, 
-    fill = group
-  ) +
-  facet_wrap(
-    ~group,
-    ncol = 1
-  ) + 
+    ggplot(aes(frequency_zipf, fill = group)) +
+    facet_wrap(~group, ncol = 1) + 
     geom_rect(
@@
         aes(
-      xmin = frequency_zipf_median-frequency_zipf_sd, 
-      xmax = frequency_zipf_median+frequency_zipf_sd, 
-      ymin = -Inf, 
-      ymax = Inf, 
-      fill = group
+            xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd,
+            ymin = -Inf, ymax = Inf, fill = group
         ), 
-    inherit.aes = FALSE,
-    alpha = 0.15
+        inherit.aes = FALSE, alpha = 0.15
     ) +   
@@
         data = medians, 
-    aes(
-      xintercept = frequency_zipf_median,
-      group = group
-    ), 
-    colour = "black",
-    size = 0.5,
-    linetype = "dashed"
+        aes(xintercept = frequency_zipf_median, group = group), 
+        colour = "black", size = 0.5, linetype = "dashed"
     ) + 
@@
         data = medians, 
-    aes(
-      label = "Median ± SD", 
-      x = frequency_zipf_median+1.25,
-      y = 13
-    ), 
+        aes(label = "Median ± SD", x = frequency_zipf_median+1.25, y = 13), 
         size = 3
@@
         data = medians, 
-    aes(
-      x = frequency_zipf_median+0.85, 
-      xend = frequency_zipf_median+0.1,
-      y = 12, 
-      yend = 11
-    ),
-    arrow = arrow(length = unit(0.2, "cm")),
-    curvature = -0.25,
-    colour = "grey30"
+        aes(x = frequency_zipf_median+0.85, xend = frequency_zipf_median+0.1, y = 12, yend = 11),
+        arrow = arrow(length = unit(0.2, "cm")), curvature = -0.25, colour = "grey30"
     ) +
@@
     d %>% 
-  count(
-    group, 
-    pthn
-  ) %>% 
-  ggplot() + 
-  aes(
-    pthn,
-    n, 
-    fill = group
-  ) +
-  facet_wrap(
-    ~group,
-    ncol = 1
-  ) + 
+    count(group, pthn) %>% 
+    ggplot(aes(pthn, n, fill = group)) + 
+    facet_wrap(~group, ncol = 1) + 
     geom_rect(
         data = medians, 
-    aes(
-      xmin = 0, 
-      xmax = pthn_median+pthn_sd, 
-      ymin = -Inf, 
-      ymax = Inf,
-      fill = group
-    ), inherit.aes = FALSE, 
-    alpha = 0.15
+        aes(xmin = 0, xmax = pthn_median+pthn_sd, ymin = -Inf, ymax = Inf, fill = group),
+        inherit.aes = FALSE, alpha = 0.15
     ) +   
@@
         data = medians, 
-    aes(
-      xintercept = pthn_median, 
-      group = group
-    ), 
-    colour = "black", 
-    size = 0.5, 
-    linetype = "dashed"
+        aes(xintercept = pthn_median, group = group), 
+        colour = "black", size = 0.5, linetype = "dashed"
     ) + 
@@
     d %>% 
-  ggplot() +
-  aes(
-    lv, 
-    fill = group
-  ) +
-  facet_wrap(
-    ~group,
-    ncol = 1
-  ) + 
+    ggplot(aes(lv, fill = group)) +
+    facet_wrap(~group, ncol = 1) + 
     geom_rect(
         data = medians, 
-    aes(
-      xmin = 0,
-      xmax = lv_median+lv_sd, 
-      ymin = -Inf,
-      ymax = Inf, 
-      fill = group
-    ),
-    inherit.aes = FALSE,
-    alpha = 0.15
+        aes(xmin = 0, xmax = lv_median+lv_sd, ymin = -Inf, ymax = Inf, fill = group),
+        inherit.aes = FALSE, alpha = 0.15
     ) +   
-  geom_histogram(
-    bins = 20, 
-    colour = "white"
-  ) +
+    geom_histogram(bins = 20, colour = "white") +
     geom_vline(
         data = medians, 
-    aes(
-      xintercept = lv_median,
-      group = group
-    ), 
-    colour = "black",
-    size = 0.5, 
-    linetype = "dashed"
+        aes(xintercept = lv_median, group = group), 
+        colour = "black", size = 0.5, linetype = "dashed"
     ) + 
@@
     ) +
-
     plot_layout(guides = "collect") &
     plot_annotation(tag_levels = "A") &
-  scale_fill_manual(
-    values = c("#1A85FF", "#ff2976", "#FFC20A"),
-    na.translate = FALSE) &
+    scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
     theme(
@@
 
-```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
+```{r}
+#| label: model-comparison
+#| message: false
+#| warning: false
 tab_loos <- model_loos %>% 
@@
     arrange(desc(elpd_loo)) %>% 
-  select(
-    model,
-    matches("elpd_loo"),
-    matches("looic"), 
-    matches("diff")) %>%
-  mutate(
-    model = str_replace(
-      model,
-      "fit_", 
-      "Model "
-    )
-  ) %>% 
+    select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
+    mutate(model = str_replace(model, "fit_", "Model ")) %>% 
     mutate_all(~ifelse(. == 0, NA, .)) %>% 
-  gt(
-    rowname_col = "model", 
-    id = "model-comparison"
-  ) %>% 
+    gt(rowname_col = "model", id = "model-comparison") %>% 
     tab_header(
@@
     ) %>% 
-  fmt_missing(
-    everything(),
-    missing_text = "-"
-  ) %>% 
+    fmt_missing(everything(), missing_text = "-") %>% 
     fmt_number(2:7) %>% 
@@
     tab_style(
-    style = cell_text(
-      align = "left",
-      weight = "bold", 
-      size = "medium"
-    ),
-    locations = cells_title(
-      groups = "title"
-    )
+        style = cell_text(align = "left", weight = "bold", size = "medium"),
+        locations = cells_title(groups = "title")
     ) %>% 
     tab_style(
-    style = cell_text(
-      align = "left", 
-      size = "medium", 
-      style = "italic"
-    ),
-    locations = cells_title(
-      groups = "subtitle"
-    )
+        style = cell_text(align = "left", size = "medium", style = "italic"),
+        locations = cells_title(groups = "subtitle")
     ) %>% 
     tab_style(
-    style = cell_text(
-      align = "left", 
-      size = "medium"
-    ),
+        style = cell_text(align = "left", size = "medium"),
         locations = cells_source_notes()
@@
         style = list(
-      cell_borders(
-        sides = c("left", "right"),
-        color = "white"
-      ),
-      cell_text(
-        align = "center"
-      )
+            cell_borders(sides = c("left", "right"), color = "white"),
+            cell_text(align = "center")
         ),
@@
         style = list(
-      cell_text(
-        align = "left",
-        size = "medium"
+            cell_text(align = "left", size = "medium"), 
+            cell_borders(sides = "all", color = "white")
         ),
-      cell_borders(
-        sides = "all", 
-        color = "white"
-      )
-    ),
-    locations = list(
-      cells_stub(), 
-      cells_source_notes()
-    )
+        locations = list(cells_stub(), cells_source_notes())
     ) %>% 
     tab_style(
-    style = cell_text(
-      align = "center",
-      size = "medium", 
-      weight = "normal",
-      style = "normal"
-    ),
-    locations = list(
-      cells_column_spanners(), 
-      cells_column_labels()
-    )
+        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
+        locations = list(cells_column_spanners(), cells_column_labels())
     ) %>% 
     tab_style(
-    style = cell_borders(
-      sides = "all",
-      color = "white"
-    ),
-    locations = cells_body(
-      columns = 1:7
-    )
+        style = cell_borders(sides = "all", color = "white"),
+        locations = cells_body(columns = 1:7)
     ) %>% 
     tab_style(
-    style = cell_borders(
-      sides = "top", 
-      color = "white"
-    ),
-    locations = cells_body(
-      columns = 1:7
-    )
+        style = cell_borders(sides = "top", color = "white"),
+        locations = cells_body(columns = 1:7)
     ) %>% 
@@
     clean_names() %>% 
-  mutate_at(
-    vars(estimate:q97_5),
-    ~ifelse(
-      variable=="Intercept", 
-      inv_logit_scaled(.), 
-      ./4
-    )
-  ) %>% 
+    mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
     group_split(variable) %>% 
@@
             ),
-    .value = ifelse(
-      str_detect(.variable, "Intercept"),
-      inv_logit_scaled(.value), 
-      .value/4
-    )
+        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
     ) %>% 
     arrange(.variable) %>% 
-  ggplot() +
-  aes(
-    .value,
-    fct_rev(.variable_name)
-  ) +
+    ggplot(aes(.value, fct_rev(.variable_name))) +
     geom_vline(xintercept = 0) +
-  stat_slab(
-    aes(
-      fill = stat(
-        cut_cdf_qi(
-          cdf, 
-          .width = c(.5, .8, .95),
-          labels = percent_format()
-        )
-      )
-    )
-  ) +
-  stat_pointinterval(
-    .width = 0, 
-    point_size = 1
-  ) +
-  scale_fill_manual(
-    values = c("#1A85FF", "#9ccaff", "#d2e5fc"),
-    na.translate = FALSE
-  ) +
-  scale_x_continuous(
-    labels = function(x) percent(round(x, 1))
-  ) +
-  labs(
-    x = "P(Correct)", 
-    y = "Posterior probability density",
-    fill = "CrI"
-  ) +
+    stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
+    stat_pointinterval(.width = 0, point_size = 1) +
+    scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
+    scale_x_continuous(labels = function(x) percent(round(x, 1))) +
+    labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
     theme(
@@
         legend.direction = "vertical",
-    legend.background = element_rect(
-      fill = "grey95"
-    ),
-    legend.key = element_rect(
-      fill = "grey95", 
-      colour = "grey95"), 
+        legend.background = element_rect(fill = "grey95"),
+        legend.key = element_rect(fill = "grey95", colour = "grey95"), 
         legend.key.height = unit(0.1, "cm"),
         axis.title.y = element_blank(),
-    legend.text = element_text(
-      size = 7
-    ),
-    axis.text.x = element_text(
-      colour = "black"
-    ),
-    axis.text.y = element_text(
-      colour = "black",
-      hjust = 1
-    ),
+        legend.text = element_text(size = 7),
+        axis.text.x = element_text(colour = "black"),
+        axis.text.y = element_text(colour = "black", hjust = 1),
         axis.ticks.y = element_blank(),
@@
 
-```{r bayesfactors, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
-
+```{r}
+#| label: bayes-factors
+#| message: false
+#| warning: false
 # wrapper to extract Bayes Factors
@@
     left_join(
-    distinct(
-      stimuli,
-      word1, 
-      ipa1), 
-    by = c("word" = "word1")) %>% 
+        distinct(stimuli, word1, ipa1), 
+        by = c("word" = "word1")
+    ) %>% 
     split(.$word)
@@
 
-cors_range <- gather_draws(
-  fit_6, 
-  `cor_.*`, 
-  regex = TRUE
-) %>% 
+cors_range <- gather_draws(fit_6, `cor_.*`, regex = TRUE) %>% 
     median_qi() %>% 
@@
 
-h_cor_intercept_frequency <- get_bf(
-  "participant_id__Intercept__frequency_zipf_std = 0", 
-  class = "cor"
-)
-h_cor_frequency_zipf_lv_group1 <- get_bf(
-  "participant_id__frequency_zipf_std__lv_std:groupeng_spa = 0", 
-  class = "cor"
-)
-h_cor_intercept_pthn <- get_bf(
-  "participant_id__Intercept__pthn_std = 0",
-  class = "cor"
-)
+h_cor_intercept_frequency <- get_bf("participant_id__Intercept__frequency_zipf_std = 0", class = "cor")
+h_cor_frequency_zipf_lv_group1 <- get_bf("participant_id__frequency_zipf_std__lv_std:groupeng_spa = 0", class = "cor")
+h_cor_intercept_pthn <- get_bf("participant_id__Intercept__pthn_std = 0", class = "cor")
 ```
@@
 
-```{r marginaleffects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
+```{r}
+#| label: marginal-effects
+#| message: false
+#| warning: false
+#| fig-width: 7
+#| fig-height: 3.5
+#| fig-cap: "Expected mean posterior predictions from Model 4.
 A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
@@
 The dispersion of the lines indicates the uncertainty of our predictions.
-We computed these posterior predictions for each group of participants, and plotted them in separate panels."}
-
+We computed these posterior predictions for each group of participants, and plotted them in separate panels."
 nd <- expand.grid(
-  lv_std = c(
-    min(fit_6$data$lv_std),
-    0,
-    max(fit_6$data$lv_std)
-  ),
+    lv_std = c(min(fit_6$data$lv_std), 0, max(fit_6$data$lv_std)),
     frequency_zipf_std = 0,
@@
 
-m <- add_epred_draws(
-  nd, 
-  fit_6, 
-  ndraws = 100, 
-  re_formula = NA
-) %>% 
+m <- add_epred_draws(nd, fit_6, ndraws = 100, re_formula = NA) %>% 
     mutate(
@@
 ggplot(m) +
-  aes(
-    x = pthn_std, 
-    y = .epred,
-    colour = lv_std,
-    fill = lv_std
-  ) +
+    aes(x = pthn_std, y = .epred, colour = lv_std, fill = lv_std) +
     facet_wrap(~group) +
-  geom_hline(
-    yintercept = 0.5, 
-    colour = "grey"
-  ) +
-  geom_line(
-    aes(group = interaction(lv_std, .draw)), 
-    size = 0.65,
-    alpha = 0.5,
-    show.legend = TRUE
-  ) +
+    geom_hline(yintercept = 0.5, colour = "grey") +
+    geom_line(aes(group = interaction(lv_std, .draw)), size = 0.65, alpha = 0.5, show.legend = TRUE) +
     stat_summary(
@@
     ) +
-  scale_color_manual(
-    values = c("#1A85FF", "#ff2976", "#FFC20A")
-  ) +
-  scale_linetype_manual(
-    values = c("solid", "dashed", "dotdash")
-  ) +
+    scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A")) +
+    scale_linetype_manual(values = c("solid", "dashed", "dotdash")) +
     scale_y_continuous(
@@
     )
-
 ```
@@
 
-\newpage
-
 # References
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-<div id="refs" custom-style="Bibliography"></div>
-\endgroup
-
-\newpage

### 2022-10-28 R/utils.R

@@ -58,7 +58,7 @@ theme_custom <- function(){
 replace_non_ascii <- function(x){
     str_replace_all(
         x,
-        c(
+        enc2utf8(c(
             "á" = "a",
             "é" = "e",
             "í" = "i",
@@ -68,22 +68,15 @@ replace_non_ascii <- function(x){
             "è" = "e",
             "ò" = "o",
             "ñ" = "n",
-            "ç" = "c"
-        )
+            "ç" = "c",
+            "ü" = "u"
+        ))
     )
 }
 
 
 # download clearpond
-import_clearpond <- function(
-    language = c(
-        "english", 
-        "dutch", 
-        "french", 
-        "spanish", 
-        "german"
-    )
-){
+import_clearpond <- function(language = c("english", "dutch", "french", "spanish", "german")){
     urls <- tribble(
         ~lang, ~url, ~data, ~header,
         "english", "https://clearpond.northwestern.edu/englishCPdatabase2.zip", "englishCPdatabase2.txt", "clearpondHeaders_EN.txt",
@@ -104,10 +97,10 @@ import_clearpond <- function(
             header = as.list(urls$header)
         ),
         .f = function(
-            url = .l[[1]],
-            file = .l[[2]],
-            data = .l[[3]], 
-            header = .l[[4]]
+        url = .l[[1]],
+        file = .l[[2]],
+        data = .l[[3]], 
+        header = .l[[4]]
         ) {
             download.file(url, destfile = file)
             unzip(zipfile = file, exdir = dir)
@@ -191,3 +184,45 @@ prop_adj_ci <- function(x, n, .width = 0.95) {
     ci[2] <- ifelse(ci[2]>1, 1, ci[2]) # truncate at 1
     return(ci)
 }
+
+
+
+# get neighbours based on Levenshtein distance
+find_neighbours <- function(x, corpus, neighbour_threshold = 1) {
+    
+    x <- str_remove_all(x, "[:punct:]")
+    
+    neighbours <- stringdistmatrix(x, corpus) %>% 
+        as_tibble() %>% 
+        set_names(corpus) %>% 
+        mutate(row_name = x) %>% 
+        pivot_longer(-row_name, names_to = "corpus", values_to = "lv") %>% 
+        mutate(is_neighbour = lv <= neighbour_threshold) %>% 
+        filter(row_name!=corpus) %>% 
+        group_by(row_name) %>% 
+        summarise(
[…]

### 2022-10-28 _targets.R

@@ -6,6 +6,8 @@ source("R/utils.R", encoding = "UTF-8")
 source("R/00_stimuli.R", encoding = "UTF-8")
 source("R/01_responses.R", encoding = "UTF-8")
 source("R/02_models.R", encoding = "UTF-8")
+source("R/03_questionnaire.R", encoding = "UTF-8")
+
 
 # set number of cores to use with brms
 options(
@@ -16,9 +18,11 @@ options(
 tar_option_set(
     # manifest dependencies (via tar_renv)
     packages = c(
+        "arrow",
         "dplyr", 
         "tidyr", 
         "stringr",
+        "readr",
         "ggplot2",
         "tibble", 
         "forcats",
@@ -44,119 +48,57 @@ tar_option_set(
         "tidytext",
         "bayesplot",
         "performance",
-        "httr"
+        "httr",
+        "ggrepel"
     ), 
 )
+
+conflicted::conflict_prefer("filter", "dplyr")
+conflicted::conflict_prefer("first", "dplyr")
+conflicted::conflict_prefer("last", "dplyr")
+conflicted::conflict_prefer("between", "dplyr")
 
 list(
     # stimuli ----
     
-    # get SUBTLEX data
-    tar_target(
-        subtlex,
-        get_subtlex()
-    ),
-    
-    # get data on lexical frequency and phonological neighbourhood density
-    tar_target(
-        clearpond_path, 
-        lst(
-            "cat-ENG" = here("data", "clearpond", "clearpond_english.csv"), 
-            "spa-ENG" = here("data", "clearpond", "clearpond_english.csv"), 
-            "cat-SPAT" = here("data", "clearpond", "clearpond_spanish.csv")
-        )
-    ),
-    tar_target(
-        clearpond, 
-        get_clearpond(clearpond_path)
-    ),
-    tar_target(
-        stimuli_path, 
-        "stimuli/trials.xlsx", 
-        format = "file"
-    ),
+    tar_target(stimuli_path, "stimuli/trials.xlsx", format = "file"),
     
     # get data on cross-language Levenshtein distance
-    tar_target(
-        levenshtein, 
-        get_levenshtein(stimuli_path = stimuli_path)
-    ),
+    tar_target(levenshtein, get_levenshtein(stimuli_path = stimuli_path)),
     
     # get audio durations
-    tar_target(
-        audios_path, 
-        "stimuli/sounds"
-    ),
-    tar_target(
-        durations, 
-        get_duration(
[…]

### 2022-10-28 test.R

@@ -1,22 +1,0 @@
-
-
-# stimuli_transformed <- stimuli %>% 
-#     mutate(
-#         word = word %>% 
-#             gsub("á", "a", .) %>% 
-#             gsub("r", "0", .)
-#     ) 
-
-library(dplyr)
-
-data.frame(
-    word = c("árbol", "cómo")
-) %>% 
-    mutate(
-        word_replaced = gsub("á", "a", word) 
-    ) %>% 
-    print()
-
-# process responses
-
-

### 2022-10-28 docs/index.Rmd
Add questionnaire data and analyses, and change reports to Quarto

@@
 # import data
-tar_load_globals()
 tar_load_all()
@@
 
-```{r stimuli-coded, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE,  include=FALSE}
-
-stimuli_coded <- stimuli %>% 
-  mutate(test_language = ifelse(group=="spa-ENG", "Spanish", "Catalan"))
-```
-
 ```{r stimuli-table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+
 stimuli %>% 
-  select(group, word1, frequency, frequency_zipf, pthn, lv) %>% 
+    select(group, word_1, freq_2, freq_zipf_2, nd, global_similarity, lv) %>% 
     group_by(group) %>% 
     summarise_at(
-    vars(frequency:lv), 
+        vars(freq_2:lv), 
         list(
@@
     ) %>% 
-  fmt_number(frequency_median:pthn_sd) %>%
-  fmt_percent(matches("lv")) %>% 
-  tab_spanner(md("**Freq. (million)**"), c("frequency_median", "frequency_sd")) %>% 
-  tab_spanner(md("**Freq. (Zipf)**"), matches("frequency_zipf")) %>% 
-  tab_spanner(md("***PTHN***"), matches("pthn")) %>% 
-  tab_spanner(md("**Levenshtein**"), matches("lv")) %>% 
+    fmt_number(freq_2_median:nd_sd) %>%
+    fmt_percent(matches("lv|global")) %>% 
+    tab_spanner(
+        md("**Freq. (million)**"),
+        c("freq_2_median", "freq_2_sd", "freq_zipf_2_median", "freq_zipf_2_sd")
+    ) %>% 
+    tab_spanner(
+        md("**Freq. (Zipf)**"), 
+        matches("freq_zipf_2")
+    ) %>% 
+    tab_spanner(
+        md("***PTHN* (presented)**"),
+        matches("nd")
+    ) %>% 
+    tab_spanner(
+        md("**Glob. similarity**"),
+        matches("global")
+    ) %>% 
+    tab_spanner(
+        md("**Levenshtein**"),
+        matches("lv")
+    ) %>% 
     cols_label(
-    frequency_median = md("Median"),
-    frequency_sd = md("*SD*"),
-    frequency_zipf_median = md("Median"),
-    frequency_zipf_sd = md("*SD*"),
-    pthn_median = md("Median"),
-    pthn_sd = md("*SD*"),
+        freq_2_median = md("Median"),
+        freq_2_sd = md("*SD*"),
+        freq_zipf_2_median = md("Median"),
+        freq_zipf_2_sd = md("*SD*"),
+        nd_median = md("Median"),
+        nd_sd = md("*SD*"),
+        global_similarity_median = md("*Mean*"),
+        global_similarity_sd = md("*SD*"),
         lv_median = md("Median"),
@@
             SD = "sd"
-    ), formatter = fmt_percent
+        ), 
+        formatter = fmt_percent
     ) %>% 
@@
         locations = list(
-      cells_body(), cells_row_groups(), cells_column_spanners(), 
-      cells_stub(), cells_stub_grand_summary(), cells_source_notes(),
-      cells_column_labels(), cells_stubhead()
+            cells_body(),
+            cells_row_groups(),
+            cells_column_spanners(), 
+            cells_stub(),
+            cells_stub_grand_summary(), 
+            cells_source_notes(),
+            cells_column_labels(),
+            cells_stubhead()
         )
@@
     tab_style(
-    style = list(cell_text(align = "left", size = "medium"), cell_borders(sides = "all", color = "white")),
+        style = list(
+            cell_text(align = "left", size = "medium"),
+            cell_borders(sides = "all", color = "white")
+        ),
         locations = cells_stub()
@@
     tab_style(
-    style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-    locations = list(cells_column_spanners(), cells_column_labels())
+        style = cell_text(
+            align = "center", 
+            size = "medium",
+            weight = "normal",
+            style = "normal"
+        ),
+        locations = list(
+            cells_column_spanners(),
+            cells_column_labels()
+        )
     ) %>% 
@@
         style = cell_borders(sides = "all", color = "white"),
-    locations = cells_body(columns = 1:9)
+        locations = cells_body(columns = 1:11)
     ) %>% 
@@
         style = cell_borders(sides = "top", color = "white"),
-    locations = cells_body(columns = 1:9)
+        locations = cells_body(columns = 1:11)
     ) %>% 
@@
 ```{r stimuli-frequency, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-stimuli_coded %>% 
+stimuli %>% 
     distinct(
-    test_language,
-    word1,
+        group,
+        word_1,
         .keep_all = TRUE
     ) %>% 
-  left_join(count(., test_language, name = "n_total")) %>% 
-  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
+    left_join(count(., group, name = "n_total")) %>% 
+    mutate(test_language = paste0(group,  " (N = ", n_total, ")")) %>% 
     ggplot(
         aes(
-      frequency_zipf, 
-      fill = test_language
-    )
-  ) +
-  facet_wrap(~test_language, ncol = 1) +
+            freq_zipf_2, 
+            fill = group
+        )
+    ) +
+    facet_wrap(
+        ~group, 
+        ncol = 1
+    ) +
     geom_histogram(
@@
         y = "# trials", 
-    fill = "Test language"
+        fill = "Group"
     )  +
@@
 ```{r stimuli-pthn, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-stimuli_coded %>% 
-  left_join(count(., test_language, name = "n_total")) %>% 
-  mutate(test_language = paste0(test_language,  " (N = ", n_total, ")")) %>% 
-  count(test_language, pthn) %>% 
-  ggplot(aes(reorder(as.factor(pthn), -n), n, fill = test_language)) +
-  facet_wrap(~test_language, ncol = 1) +
-  geom_col(position = position_dodge()) +
-  labs(x = "# Phonological neighbours with higher lexical frequency\nExtracted from CLEARPOND database", y = "# trials", fill = "Test language") +
+stimuli %>% 
+    left_join(count(., group, name = "n_total")) %>% 
+    mutate(translation = paste0(word_1, "/", word_2)) %>% 
+    ggplot() +
+    facet_wrap(~group, scales = "free_y") +
+    aes(
+        x = nd, 
+        y = reorder(translation, nd),
+        colour = group,
+        fill = group,
+        label = translation
+    ) + 
+    geom_text(
+        size = 3,
+        hjust = 0
+    ) +
+    labs(
+        x = "Group", 
+        y = "# phonological neighbours", 
+    ) +
+    scale_x_continuous(breaks = seq(0, 30, 5)) +
     theme(
         legend.position = "none",
-    panel.grid.major.y = element_line(colour = "grey", linetype = "dotted")
+        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
+        axis.ticks.y = element_blank(),
+        axis.text.y = element_blank()
     )
@@
 ```{r stimuli-similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-stimuli_coded %>% 
+stimuli %>% 
     left_join(count(., group, name = "n_total")) %>% 
@@
     aes(
-    x = group, 
-    y = lv,
+        x = lv,
+        y = reorder_within(translation, lv, group),
         fill = group,
-    colour = group
-  ) +
-  geom_violin() +
-  geom_point(
-    shape = 1,
-    stroke = 1, 
-    position = position_jitter(width = 0.1), 
-    colour = "black", 
-    alpha = 0.25
-  ) +
-  geom_boxplot(
-    colour = "black", 
-    fill = "white",
-    width = 0.1, 
-    size = 0.75
+        colour = group,
+        label = percent(lv, accuracy = 0.1)
+    ) +
+    facet_wrap(~group, scales = "free_y") +
+    geom_text(
+        size = 3,
+        hjust = 0
     ) +
@@
     ) +
-  scale_y_continuous(
-    labels = scales::percent
-  ) +
+    scale_x_continuous(
+        labels = percent
+    ) +
+    scale_y_reordered() +
     theme(
         legend.position = "none",
-    axis.title.x = element_blank(),
-    panel.grid.major.y = element_line(
+        axis.title.y = element_blank(),
+        axis.text.y = element_text(size = 7, hjust = 1),
+        panel.grid.major.x = element_line(
             colour = "grey",
@@
     ) 
-
+ggsave(here("img", "lv.png"), height = 12)
+
+```
+
+## Global similarity
+
+```{r global-similarity, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
+stimuli %>% 
+    ggplot() +
+    aes(
+        x = global_similarity, 
+        y = reorder_within(translation, global_similarity, group),
+        label = percent(global_similarity, accuracy = 0.1),
+        fill = group,
+        colour = group,
+    ) + 
+    facet_wrap(~group, scales = "free_y") + 
+    geom_text(
+        size = 2,
+        hjust = 0
+    ) +
+    labs(
+        x = "Global similarity (mean)",
+        y = "Translation"
+    ) +
+    scale_y_reordered() +
+    scale_x_continuous(
+        limits = c(0, 0.2),
+        labels = percent
+    ) +
+    theme(
+        legend.position = "none",
+        axis.text.y = element_text(size = 7, hjust = 1),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(
+            colour = "grey",
+            linetype = "dotted"
+        )
+    )
+
+ggsave(here("img", "global_similarity.png"), height = 12)
 ```
@@
 medians <- stimuli %>% 
+    select(group, word_1, freq_2, freq_zipf_2, nd, global_similarity, lv) %>% 
     group_by(group) %>% 
-  summarise(
-    pthn_median = median(pthn),
-    pthn_sd = sd(pthn),
-    frequency_zipf_median = median(frequency_zipf),
-    frequency_zipf_sd = sd(frequency_zipf),
-    lv_median = median(lv),
-    lv_sd = sd(lv)
+    summarise_at(
+        vars(freq_2:lv), 
+        list(
+            median = ~median(., na.rm = TRUE),  
+            sd = ~sd(., na.rm = TRUE)
+        )
     )
@@
     aes(
-    x = frequency_zipf,
+        x = freq_zipf_2,
         fill = group
@@
         aes(
-      xmin = frequency_zipf_median-frequency_zipf_sd,
-      xmax = frequency_zipf_median+frequency_zipf_sd, 
+            xmin = freq_zipf_2_median-freq_zipf_2_sd,
+            xmax = freq_zipf_2_median+freq_zipf_2_sd, 
             ymin = -Inf,
@@
         aes(
-      xintercept = frequency_zipf_median,
+            xintercept = freq_zipf_2_median,
             group = group
@@
             label = "Median ± SD", 
-      x = frequency_zipf_median+1.25,
+            x = freq_zipf_2_median+1.25,
             y = 13
@@
         aes(
-      x = frequency_zipf_median+0.85, 
-      xend = frequency_zipf_median+0.1,
+            x = freq_zipf_2_median+0.85, 
+            xend = freq_zipf_2_median+0.1,
             y = 12, 
@@
 
-  # PTHN
+    # PTHN 1
     stimuli %>% 
-  count(group, pthn) %>% 
+    count(group, nd) %>% 
     ggplot(
         aes(
-      x = pthn, 
+            x = nd, 
             y = n,
@@
             xmin = 0, 
-      xmax = pthn_median+pthn_sd, 
+            xmax = nd_median+nd_sd, 
             ymin = -Inf, 
@@
         aes(
-      xintercept = pthn_median,
+            xintercept = nd_median,
             group = group
@@
     labs(
-    x = "PTHN", 
+        x = "PTHN across languages", 
         y = "Number of trials", 
         title = "Higher-frequency neighbours\n(stimuli translations)\n"
+    ) +
+
+    # global similarity
+    stimuli %>% 
+    ggplot(
+        aes(
+            x = global_similarity, 
+            fill = group
+        )
+    ) + 
+    facet_wrap(~group, ncol = 1) + 
+    geom_rect(
+        data = medians, 
+        aes(
+            xmin = 0, 
+            xmax = global_similarity_median+global_similarity_sd, 
+            ymin = -Inf, 
+            ymax = Inf,
+            fill = group
+        ), 
+        inherit.aes = FALSE, 
+        alpha = 0.15
+    ) +   
+    geom_histogram(colour = "white") + 
+    geom_vline(
+        data = medians, 
+        aes(
+            xintercept = global_similarity_median,
+            group = group
+        ), 
+        colour = "black", 
+        size = 0.5,
+        linetype = "dashed"
+    ) + 
+    labs(
+        x = "Global similarity across languages", 
+        y = "Number of trials", 
+        title = "Mean cross-language similarity\n(stimuli translations)\n"
     ) +
@@
 
-  plot_layout(guides = "collect") &
+    plot_layout(
+        guides = "collect",
+        nrow = 1
+    ) &
     plot_annotation(tag_levels = "A") &
@@
     ) + 
-  facet_wrap(~group, ncol = 1) +
+    facet_wrap(~group) +
     geom_histogram() +
-  geom_rug() +
     labs(
@@
         panel.grid = element_blank(),
-    axis.text.x = element_text(angle = 90),
+        axis.text.x = element_text(),
         axis.title.y = element_blank()
@@
     mutate_at(
-    vars(lv_std, pthn_std, frequency_zipf_std),
+        vars(lv_std, nd_std, freq_zipf_2_std),
         function(x) scale(x, center = TRUE, scale = TRUE)[,1]) %>% 
@@
 
-model_loos %>% 
+model_loos_gs %>% 
     as.data.frame() %>% 
@@
 
-coefs <- fixef(fit_6) %>% 
+coefs <- fixef(fit_4_gs) %>% 
     as.data.frame() %>% 
@@
     "Intercept" = "Intercept",
-  "frequency_zipf_std" = "Frequency\n[+1 SD]",
-  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:groupeng_spa" = "Levenshtein \u00d7 Group 1",
-  "lv_std:groupcateng_spaeng" = "Levenshtein \u00d7 Group 2",
-  "lv_std" = "Levenshtein\n[+1 SD]",
-  "pthn_std" = "PTHN\n[+1 SD]",
-  "groupeng_spa" =  "Group 1\n[ENG-* vs. cat-SPA]",
-  "groupcateng_spaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
+    "freq_zipf_2_std" = "Frequency\n[+1 SD]",
+    "global_similarity_std:lv_std" = "Glob. similarity \u00d7 Levenshtein", 
+    "global_similarity_std" = "Glob. similarity \n[+1 SD]",
+    "lv_std" = "Levenshtein\n[+1 SD]"
 )
@@
                 levels = c(
-          "Intercept", "Frequency\n[+1 SD]", "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", "PTHN \u00d7 Levenshtein",
-          "Group 1\n[ENG-* vs. cat-SPA]",
-          "Group 2\n[cat-ENG vs. spa-ENG]",
-          "Levenshtein \u00d7 Group 1", "Levenshtein \u00d7 Group 2"
+                    "Intercept", 
+                    "Frequency\n[+1 SD]", 
+                    "Glob. similarity \n[+1 SD]", 
+                    "Levenshtein\n[+1 SD]",
+                    "Glob. similarity \u00d7 Levenshtein"
                 )
@@
     "Intercept" = "Intercept",
-  "frequency_zipf_std" = "Frequency\n[+1 SD]",
-  "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-  "lv_std:groupeng_spa" = "Levenshtein \u00d7 Group 1",
-  "lv_std:groupcateng_spaeng" = "Levenshtein \u00d7 Group 2",
+    "freq_zipf_2_std" = "Frequency\n[+1 SD]",
+    "global_similarity_std:lv_std" = "Glob. similarity \u00d7 Levenshtein", 
     "lv_std" = "Levenshtein\n[+1 SD]",
-  "pthn_std" = "PTHN\n[+1 SD]",
-  "groupeng_spa" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-  "groupcateng_spaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
+    "global_similarity_std" = "Glob. similarity \n[+1 SD]"
 )
@@
 post_draws <-  gather_draws(
-  fit_6, 
+    fit_4_gs, 
     `b_.*`, 
@@
                     "Frequency\n[+1 SD]",
-          "PTHN\n[+1 SD]", 
+                    "Glob. similarity \n[+1 SD]", 
                     "Levenshtein\n[+1 SD]",
-          "PTHN \u00d7 Levenshtein",
+                    "Glob. similarity \u00d7 Levenshtein",
                     "Group 1\n[cat/spa-ENG vs. cat-SPA]",
@@
 
-# SD of random effects
-post_sd_data <- post_draws %>% 
-  filter(str_detect(.variable, "sd_")) %>% 
-  mutate(
-    group = case_when(
-      str_detect(.variable, "participant") ~ "Participants",
-      str_detect(.variable, "word") ~ "Words"
-    ),    
-    .variable_name = str_remove(
-      .variable,
-      "sd_participant__|sd_word__"
-    ) %>% 
-      str_replace_all(str_repl) %>% 
-      factor(
-        levels = c(
-          "Intercept", 
-          "Frequency\n[+1 SD]",
-          "PTHN\n[+1 SD]", 
-          "Levenshtein\n[+1 SD]", 
-          "PTHN \u00d7 Levenshtein",
-          "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-          "Group 2\n[cat-ENG vs. spa-ENG]",
-          "Levenshtein \u00d7 Group 1",
-          "Levenshtein \u00d7 Group 2"
-        )
-      ),
-    .value = ifelse(
-      str_detect(.variable, "Intercept"), 
-      inv_logit_scaled(.value), 
-      .value/4
-    )
-  )
-
-post_sd <- post_sd_data %>% 
-  ggplot() +
-  aes(.value, fct_rev(.variable_name)) +
-  stat_slab(
-    aes(
-      fill = stat(
-        cut_cdf_qi(
-          cdf, 
-          .width = c(.5, .8, .95),
-          labels = percent_format()
-        )
-      )
-    )
-  ) +
-  stat_pointinterval(
-    .width = 0, 
-    point_size = 1
-  ) +
-  scale_fill_manual(
-    values = c("#ff2976", "#f2a7c2", "#fcd9e6"), 
-    na.translate = FALSE
-  ) +
-  scale_x_continuous(
-    labels = percent, 
-    limits = c(0, 1)
-  ) +
-  labs(
-    x = "SDs in P(Correct)",
-    title = "By participant", 
-    fill = "CrI"
-  ) +
-  theme(
-    axis.text.x = element_blank()
-  ) +
-
-  post_sd_data %>% 
-  filter(group=="Words") %>% 
-  ggplot() +
-  aes(
-    .value, 
-    fct_rev(.variable_name)
-  ) +
-  stat_slab(
-    aes(
-      fill = stat(
-        cut_cdf_qi(
-          cdf, 
-          .width = c(.5, .8, .95), 
-          labels = percent_format()
-        )
-      )
-    )
-  ) +
-  stat_pointinterval(
-    .width = 0, 
-    point_size = 1
-  ) +
-  scale_x_continuous(
-    labels = percent, 
-    limits = c(0, 1)
-  ) +
-  labs(
-    x = "SDs in P(Correct)",
-    title = "By item",
-    fill = "CrI"
-  ) +
-  scale_fill_manual(
-    values = c("#ff2976", "#f2a7c2", "#fcd9e6"), 
-    na.translate = FALSE
-  ) +
-  plot_layout(
-    ncol = 1,
-    heights = c(0.85, 0.15), 
-    guides = "collect"
-  ) &
-  theme(
-    legend.position = "none",
-    axis.text.y = element_text(hjust = 1),
-    axis.title.x = element_blank(),
-    legend.direction = "vertical",
-    axis.title.y = element_blank(),
-    legend.background = element_rect(fill = "grey95"),
-    legend.key.height = unit(0.1, "cm"),
-    legend.key = element_rect(fill = "grey95", colour = "grey95"),
-    legend.text = element_text(size = 7),
-    axis.ticks.y = element_blank(),
-    panel.grid.major.y = element_line(colour = "grey", size = 0.5)
-  )
-
 # Pearson correlations between random effects
-corr_mat <- VarCorr(fit_6, summary = TRUE)$participant$cor 
-corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant$cor)
+corr_mat <- VarCorr(fit_4_gs, summary = TRUE)$participant$cor 
+corr_names <- rownames(VarCorr(fit_4_gs, summary = TRUE)$participant$cor)
 
@@
         na.value = "white", 
-    limits = c(-0.4, 0.4)
+        limits = c(-0.1, 0.1)
     ) +
@@
 
-post_fix / (
-  (post_sd | post_cors) +
-    plot_layout(widths = c(0.6, 0.4))
-) +
-  plot_layout() +
+(post_fix | post_cors) +
+    plot_layout(widths = c(0.6, 0.4)) +
     plot_annotation(tag_levels = "A")
@@
 ```{r post_re_word, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
-fixed_intercept <- fixef(fit_6)[1 ,c(1, 3, 4)] %>% 
+empirical_means <- responses %>% 
+    group_by(group) %>% 
+    summarise(
+        successes = sum(correct, na.rm = TRUE),
+        n = n(),
+        .groups = "drop"
+    ) %>% 
+    mutate(
+        prop = prop_adj(successes, n),
+        prop_se = prop_adj_se(successes, n)
+    )
+
+empirical_accuracy <- responses %>% 
+    mutate(
+        translation = paste0(
+            word_1, " /", ipa_flat_1, "/ - ", 
+            word_2, " /", ipa_flat_2, "/ - ",
+            percent(lv, accuracy = 1)
+        )
+    ) %>% 
+    group_by(group, translation, lv) %>% 
+    summarise(
+        successes = sum(correct, na.rm = TRUE),
+        n = n(),
+        .groups = "drop"
+    ) %>% 
+    mutate(
+        prop = successes/n,
+        prop_se = sqrt((prop*(1-prop))/n)
+    )
+
+fixed_intercept <- fixef(fit_4_gs)[1 ,c(1, 3, 4)] %>% 
     inv_logit_scaled()
@@
 nd_re_words <- expand.grid(
-  translation_id = unique(fit_6$data$translation_id),
-  pthn_std = c(-1, 1),
-  frequency_zipf_std = 0,
+    translation_id = unique(fit_4_gs$data$translation_id),
+    global_similarity_std = c(-1, 1),
+    freq_zipf_2_std = 0,
     lv_std = seq(
-    min(fit_6$data$lv_std, na.rm = TRUE), 
-    max(fit_6$data$lv_std, na.rm = TRUE
+        min(fit_4_gs$data$lv_std, na.rm = TRUE), 
+        max(fit_4_gs$data$lv_std, na.rm = TRUE
         ), by = 0.1
@@
     nd_re_words, 
-  fit_6, 
+    fit_4_gs, 
     re_formula = ~ (1 | translation_id),
@@
     select(-group) %>% 
-  left_join(select(stimuli, group, translation, translation_id))
+    left_join(select(stimuli, group, translation, translation_id, lv)) %>% 
+    mutate(translation = paste0(translation, " - ", percent(lv, accuracy = 1))) %>% 
+    left_join(select(empirical_accuracy, translation, n, prop, prop_se))
 
@@
     aes(
-    .epred,
-    reorder(as.factor(translation), .epred)
+        x = .epred,
+        y = reorder_within(translation, lv, group),
+        colour = group
     ) +
     facet_wrap(~group, scales = "free_y") +
+    # geom_vline(
+    #     xintercept = fixed_intercept[1],
+    #     colour = "#1A85FF", 
+    #     size = 0.5,
+    #     linetype = "dotted"
+    # ) +
     geom_vline(
-    xintercept = fixed_intercept[1],
-    colour = "grey", 
+        data = empirical_means,
+        aes(xintercept = prop),
+        size = 0.75
+    ) +
+    # stat_interval(
+    #     size = 1.5,
+    #     .width = c(0.5, 0.67, 0.95),
+    #     aes(y = reorder_within(translation, prop, group))
+    # ) +
+    geom_errorbar(
+        data = empirical_accuracy,
+        aes(
+            x = prop,
+            xmin = prop-prop_se,
+            xmax = prop+prop_se,
+            y = reorder_within(translation, prop, group)
+        ),
+        width = 0,
+        size = 0.5
+    ) +
+    geom_point(
+        data = empirical_accuracy,
+        aes(
+            x = prop,
+            y = reorder_within(translation, prop, group)
+        ),
         size = 1
     ) +
-  stat_interval(
-    size = 1.5,
-    .width = c(0.5, 0.67, 0.95)
-  ) +
-  # annotate(
-  #   geom = "text",
-  #   x = 0.7,
-  #   y = 100,
-  #   size = 3,
-  #   hjust = 0.5,
-  #   label = "Item accuracy \n(95%, 67%, and 50% CrI \nof item-level intercepts)"
+    labs(
+        x = "Percentage of correct responses",
+        y = "Item",
+        title = "Observed translation accuracy",
+        subtitle = "Dots indicate proportions of correct responses, error bars indicate +/-1 SE",
+        colour = "Credible Interval (CrI)",
+        caption = "Percentages next to items indicate the Levenshtein distance between both translations."
+    ) +
+    # scale_color_manual(
+    #     values = c("#d2e5fc", "#9ccaff", "#1A85FF"),
+    #     na.translate = FALSE, 
+    #     labels = paste0(c(95, 67, 50), "%")
     # ) +
-  # annotate(
-  #   geom = "rect",
-  #   alpha = 0.15,
-#   xmin = fixed_intercept[2],
-#   xmax = fixed_intercept[3],
-#   ymin = -Inf,
-#   ymax = Inf
-# ) +
-# annotate(
-#   x = 0.7,
-#   xend = 0.6,
-#   y = 108,
-#   yend = 130,
-#   geom = "curve",
-#   curvature = 0.25,
-#   arrow = arrow(length = unit(0.2, "cm"))
-# ) +
-# annotate(
-#   geom = "text",
-#   x = 0.8,
-#   y = 10,
-#   size = 3,
-#   hjust = 0.5,
-#   label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
-# ) +
-# annotate(
-#   x = 0.7,
-#   xend = fixed_intercept[1]+0.05,
-#   y = 17,
-#   yend = 30,
-#   geom = "curve",
-#   curvature = 0.2,
-#   arrow = arrow(length = unit(0.2, "cm"))
-# ) +
-labs(
-  x = "Posterior probability of correct response",
-  y = "Item",
-  colour = "Credible Interval (CrI)",
-  caption = "Credible intervals cover the range of values of the sampling space that \ncontain the true value with a given certainty."
-) +
-  scale_color_manual(
-    values = c("#d2e5fc", "#9ccaff", "#1A85FF"),
-    na.translate = FALSE, 
-    labels = paste0(c(95, 67, 50), "%")
-  ) +
     scale_x_continuous(
@@
     ) +
+    scale_y_reordered() +
     theme(
-    legend.position = c(0.1, 0.9),
-    legend.direction = "vertical",
-    legend.justification = "left",
+        legend.position = "none",
         legend.key = element_rect(fill = NA, colour = NA),
-    panel.grid.major.x = element_line(
-      colour = "grey", 
-      linetype = "dotted"
-    ),
         axis.text.x = element_text(size = 7),
@@
         axis.title = element_text(size = 10),
+        axis.title.y = element_blank(),
+        panel.grid.major = element_line(
+            size = 0.15,
+            linetype = "dotted",
+            colour = "grey"
+        ),
         plot.caption.position = "plot",
@@
             )
-    )
-  )
-
-ggsave(here("img", "accuracy.png"), height = 9, width = 5, dpi = 500)
-```
-
-```{r post_re_word_empirical, echo=FALSE, fig.height=9, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-means <- responses %>% 
-  group_by(group) %>% 
-  summarise(
-    successes = sum(correct, na.rm = TRUE),
-    n = n(),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    prop = prop_adj(successes, n),
-    prop_se = prop_adj_se(successes, n)
-  )
-
-word_accuracy <- responses %>% 
-  mutate(
-    stim = paste0(
-      word, " (", ipa1, ") / ", 
-      word2, " (", ipa2, ") / ",
-      percent(lv, accuracy = 1)
-    )
-  ) %>% 
-  group_by(group, stim, lv) %>% 
-  summarise(
-    successes = sum(correct, na.rm = TRUE),
-    n = n(),
-    .groups = "drop"
-  ) %>% 
-  mutate(
-    prop = prop_adj(successes, n),
-    prop_se = prop_adj_se(successes, n)
-  )
-
-word_accuracy %>% 
-  ggplot() +
-  aes(
-    x = prop,
-    y = reorder(as.factor(stim), lv),
-    xmin = prop-prop_se,
-    xmax = prop+prop_se,
-    label = percent(lv),
-    colour = group
-  ) +
-  facet_wrap(~group, scales = "free_y") +
-  geom_vline(
-    data = means,
-    aes(xintercept = prop),
-    size = 1,
-    colour = "grey"
-  ) +
-  geom_errorbar(width = 0.1) +
-  geom_point() +
-  annotate(
-    geom = "text",
-    x = 0.7,
-    y = 10,
-    size = 3,
-    hjust = 0.5,
-    label = "Item accuracy \n(95% CI)"
-  ) +
-  annotate(
-    x = 0.7,
-    xend = 0.15,
-    y = 13,
-    yend = 20,
-    geom = "curve",
-    curvature = 0.25,
-    arrow = arrow(length = unit(0.2, "cm"))
-  ) +
-  # annotate(
-  #   geom = "text",
-  #   x = 0.8,
-  #   y = 10,
-  #   size = 3,
-  #   hjust = 0.5,
-  #   label = "Global posterior accuracy\n(Mean ± 95% CrI of the \npopulation-level intercept)"
-  # ) +
-  # annotate(
-  #   x = 0.7,
-  #   xend = fixed_intercept[1]+0.05,
-#   y = 17,
-#   yend = 30,
-#   geom = "curve",
-#   curvature = 0.2,
-#   arrow = arrow(length = unit(0.2, "cm"))
-# ) +
-labs(
-  x = "Probability of correct response",
-  y = "Presented word / Correct translation / Levenshtein similarity"
-) +
-  scale_colour_manual(
-    values = c("#1A85FF", "#ff2976", "#FFC20A"),
-    na.translate = FALSE
-  ) +
-  scale_x_continuous(
-    limits = c(0, 1), 
-    labels = percent
-  ) +
-  theme(
-    legend.position = "none",
-    legend.direction = "vertical",
-    legend.justification = "left",
-    legend.key = element_rect(
-      fill = NA, 
-      colour = NA
-    ),
-    panel.grid.major.x = element_line(
-      colour = "grey",
-      linetype = "dotted"
-    ),
-    axis.text.x = element_text(size = 7),
-    axis.text.y = element_text(size = 5, hjust = 1),
-    axis.ticks = element_blank(),
-    axis.title = element_text(size = 10),
-    plot.caption.position = "plot",
-    plot.caption = element_text(
-      size = 9, 
-      hjust = 0,
-      face = "plain", 
-      margin = margin(t = unit(3, "cm"))
-    )
-  )
-
-ggsave(here("img", "accuracy_empirical.png"), height = 9, width = 15, dpi = 500)
+        ),
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0, size = 13),
+        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
+    )
+
+ggsave(here("img", "accuracy.png"), height = 9, width = 10, dpi = 1000)
 ```
@@
 ## Marginal effects
+
+### Global similarity
 
@@
     lv_std = c(
-    min(fit_6$data$lv_std),
+        min(fit_6_gs$data$lv_std),
         0,
-    max(fit_6$data$lv_std)
-  ),
-  frequency_zipf_std = 0,
-  pthn_std = seq(
-    min(fit_6$data$pthn_std, na.rm = TRUE),
-    max(fit_6$data$pthn_std, na.rm = TRUE),
+        max(fit_6_gs$data$lv_std)
+    ),
+    freq_zipf_2_std = 0,
+    global_similarity_std = seq(
+        min(fit_6_gs$data$global_similarity_std, na.rm = TRUE),
+        max(fit_6_gs$data$global_similarity_std, na.rm = TRUE),
         by = 0.1
     ),
-  group = unique(fit_6$data$group)
+    group = unique(fit_6_gs$data$group)
 )
@@
     nd, 
-  fit_6, 
-  ndraws = 100, 
+    fit_6_gs, 
+    # ndraws = 100, 
     re_formula = NA
@@
             levels = c(
-        min(fit_6$data$lv_std),
+                min(fit_6_gs$data$lv_std),
                 0,
-        max(fit_6$data$lv_std)
+                max(fit_6_gs$data$lv_std)
             ),
@@
     aes(
-    pthn_std, 
-    .epred,
-    colour = lv_std,
-    fill = lv_std
-  ) +
-  facet_wrap(~group) +
+        x = global_similarity_std, 
+        y = .epred,
+        colour = group,
+        fill = group
+    ) +
+    # facet_wrap(~group) +
     geom_hline(
         yintercept = 0.5, 
-    colour = "grey"
-  ) +
-  geom_line(
-    aes(group = interaction(lv_std, .draw)), 
-    size = 0.65,
-    alpha = 0.5,
-    show.legend = TRUE
-  ) +
+        colour = "grey",
+        size = 1
+    ) +
+    # geom_line(
+    #     aes(group = interaction(lv_std, .draw)), 
+    #     size = 0.65,
+    #     alpha = 0.5,
+    #     show.legend = TRUE
+    # # ) +
+    # stat_lineribbon(
+    #     alpha = 0.5,
+    #     .width = 0.95,
+    #     colour = NA
+    # ) +
 stat_summary(
@@
     geom = "line",
-    colour = "black", 
     size = 0.75,
@@
     labs(
-    x = paste0("PTHN (+1 SD, ", round(sd(responses$pthn), 1), " higher-frequency neighbours)"),
+        x = paste0("Neighbourhood density (+1 SD, ", round(sd(responses$nd), 1), " higher-frequency neighbours)"),
+        y = "P(Correct)", 
+        colour = "Levenshtein similarity",
+        fill = "Levenshtein similarity",
+        linetype = "Levenshtein similarity",
+        title = "Population-level posterior predictions"
+    ) +
+    scale_colour_manual(
+        values = c("#1A85FF", "#ff2976", "#FFC20A"),
+        na.translate = FALSE
+    ) +
+    scale_y_continuous(
+        labels = function(x) percent(round(x, 2)),
+        limits = c(0, 1),
+        breaks = seq(0, 1, 0.1)
+    ) +
+    theme(
+        legend.position = "top",
+        plot.title = element_blank()
+    )
+
+```
+
+### Levenshtein distance
+
+```{r marginaleffects-lv, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.width=7, fig.height=3.5, fig.cap="Expected mean posterior predictions from Model 4.
+A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
+We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein).
+We did this across the range of values of the PTHN scores.
+For each simulation, we drew a single sample from the posterior distribution of each coefficient.
+Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
+Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
+The dispersion of the lines indicates the uncertainty of our predictions.
+We computed these posterior predictions for each group of participants, and plot tem in separate panels."}
+
+nd <- expand.grid(
+    global_similarity_std = c(-1, 0, 1),
+    freq_zipf_2_std = 0,
+    lv_std = seq(
+        min(fit_6_gs$data$lv_std, na.rm = TRUE),
+        max(fit_6_gs$data$lv_std, na.rm = TRUE),
+        by = 0.1
+    ),
+    group = unique(fit_6_gs$data$group)
+)
+
+m <- add_epred_draws(
+    nd, 
+    fit_6_gs, 
+    # ndraws = 100, 
+    re_formula = NA
+) %>% 
+    mutate(
+        global_similarity_std = factor(
+            global_similarity_std,
+            levels = c(-1, 0, 1),
+            labels = c("-1 SD", "Mean", "+1 SD")
+        )
+    )
+
+ggplot(m) +
+    aes(
+        x = lv_std, 
+        y = .epred,
+        colour = group,
+        fill = group
+    ) +
+    facet_wrap(~group) +
+    geom_hline(
+        yintercept = 0.5, 
+        colour = "grey",
+        size = 1
+    ) +
+    # geom_line(
+    #     aes(group = interaction(lv_std, .draw)), 
+    #     size = 0.65,
+    #     alpha = 0.5,
+    #     show.legend = TRUE
+    # # ) +
+    # stat_lineribbon(
+    #     alpha = 0.5,
+    #     .width = 0.95,
+    #     colour = NA
+    # ) +
+stat_summary(
+    aes(linetype = global_similarity_std), 
+    fun = mean, 
+    geom = "line",
+    size = 0.75,
+    show.legend = TRUE
+) +
+    annotate(
+        "segment", 
+        x = -0.5, 
+        xend = -0.9, 
+        y = 0.85, 
+        yend = 0.85, 
+        arrow = arrow(length = unit(0.1, "cm")), 
+        size = .5
+    ) +
+    annotate(
+        "segment", 
+        x = 2.25, 
+        xend = 2.65, 
+        y = 0.85, 
+        yend = 0.85, 
+        arrow = arrow(length = unit(0.1, "cm")),
+        size = .5
+    ) +
+    annotate(
+        "text", 
+        x = -0.5, 
+        y = 0.95, 
+        size = 2.75,
+        label = "Less similar"
+    ) +
+    annotate(
+        "text", 
+        x = 2.15, 
+        y = 0.95, 
+        size = 2.75,
+        label = "More similar"
+    ) +
+    labs(
+        x = paste0("Neighbourhood density (+1 SD, ", round(sd(responses$nd), 1), " higher-frequency neighbours)"),
         y = "P(Correct)", 
@@
         stim = paste0(
-      word1, " (", ipa1, ")", " / ",
-      word2, " (", ipa2, ")", " / ",
+            word_1, " /", ipa_flat_1, "/", " - ",
+            word_2, " /", ipa_flat_2, "/", " - ",
             percent(lv, accuracy = 1)
@@
         ),
-    word = replace_non_ascii(word1)
+        word = replace_non_ascii(word_1)
     ) %>% 

### 2022-10-28 R/00_stimuli.R

@@ -1,162 +1,115 @@
-# get SUBTLEX lexical frequencies ----
-get_subtlex <- function(){
+
+# generate PCT corpora from Clearpond 
+make_pct_corpus <- function(){
     
-    paths <- list(
-        English = here("data", "subtlex", "SUBTLEX-UK.txt"),
-        Spanish = here("data", "subtlex", "SUBTLEX-ESP.txt"),
-        Catalan = here("data", "subtlex", "SUBTLEX-CAT.txt")
+    # load headers
+    corpus_spa_headers <- readLines(
+        here("data", "clearpond", "spanishCPdatabase2", "clearpondHeaders_SP.txt"),
+        warn = FALSE
     )
     
-    # import SUBTLEX-UK (English)
-    subtlex_eng <- paths$English %>% 
-        fread(
-            verbose = FALSE,
-            showProgress = TRUE
-        ) %>%
-        as_tibble() %>% 
-        clean_names() %>% 
-        rename(
-            word = spelling,
-            frequency_count = freq_count,
-            frequency_zipf = log_freq_zipf
-        ) %>% 
-        select(word, frequency_count, frequency_zipf)
+    corpus_eng_headers <- readLines(
+        here("data", "clearpond", "englishCPdatabase2", "clearpondHeaders_EN.txt"),
+        warn = FALSE
+    )
     
+    # load corpora
+    suppressWarnings({
+        corpus_spa <- here("data", "clearpond", "spanishCPdatabase2", "spanishCPdatabase2.txt") %>% 
+            read_tsv(col_names = corpus_spa_headers, show_col_types = FALSE) %>% 
+            clean_names() %>% 
+            select(Spelling = word, Transcription = phono, Frequency = frequency) %>% 
+            write_csv(
+                here("data", "pct", "pct-corpus_spa.csv")
+            )
+    })
     
-    # import SUBTLEX-ESP (Spanish)
-    subtlex_spa <- paths$Spanish %>%
-        fread(
-            verbose = FALSE,
-            showProgress = FALSE,
-            select = 1:4
-        ) %>% 
-        clean_names() %>%
-        mutate(frequency_zipf = log10(freq_per_million)+3) %>% 
-        rename(frequency_count = freq_count) %>% 
-        select(
-            word, 
-            frequency_count, 
-            frequency_zipf
-        ) %>% 
-        drop_na() 
+    message("Spanish corpus generated")
     
+    suppressWarnings({
+        
[…]

### 2022-10-28 _targets_packages.R

@@ -1,4 +1,5 @@
 # Generated by targets::tar_renv(): do not edit by hand
+library(arrow)
 library(audio)
 library(bayesplot)
 library(brms)
@@ -9,8 +10,10 @@ library(dplyr)
 library(forcats)
 library(future)
 library(ggplot2)
+library(ggrepel)
 library(gt)
 library(here)
+library(httr)
 library(janitor)
 library(keyring)
 library(knitr)
@@ -23,6 +26,7 @@ library(patchwork)
 library(performance)
 library(pingr)
 library(purrr)
+library(readr)
 library(readxl)
 library(rstudioapi)
 library(scales)

### 2022-11-14 index.qmd
Update lab notes

@@
         code-fold: true
+        code-tools: true
+        code-line-numbers: true
+        code-link: true
 ---
@@
 #| include: false
+set.seed(888)
+
 my_colours <- c("#1A85FF", "#ff2976", "#FFC20A")
@@
 #| label: tbl-stimuli
-tab_subtitle <- md("Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words.")
-
-tab_note <- md("*Note*.
-Median and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words.")
-
+#| tbl-cap: "Lexical frequency, phonological neigbourhood size, and levenshtein distance of the stimuli words."
+#| tbl-subcap: "Median and standard deviation (*SD*) of lexical frequencies of target words (expressed in counts per million and as Zipf scores), density of higher-frequency phonological neighbourhoods of target words, and normalised Levenshtein phonological similarity between the presented and the target words."
 stimuli %>% 
@@
     gt(rowname_col = "group") %>% 
-    tab_header(title = md("Table 1"), subtitle = tab_subtitle) %>% 
-    tab_source_note(tab_note) %>% 
     fmt_number(freq_2_median:nd_sd) %>%
@@
     left_join(count(., group, name = "n_total")) %>% 
-    mutate(translation = paste0(word_1, "/", word_2)) %>% 
+    mutate(
+        translation = paste0(word_1, "/", word_2),
+        group = paste0(group,  " (N = ", n_total, ")")
+    ) %>% 
     ggplot(aes(x = nd, y = reorder(translation, nd), colour = group, fill = group, label = translation)) +
     facet_wrap(~group, scales = "free_y") +
-    geom_text(size = 3, hjust = 0) +
-    labs(
-        x = "Group", 
-        y = "# phonological neighbours", 
+    geom_segment(aes(x = 0, xend = nd, y = reorder(translation, nd), yend = reorder(translation, nd))) +
+    geom_point(size = 1) +
+    labs(
+        x = "# phonological neighbours", 
+        y = "Word"
     ) +
@@
         axis.ticks.y = element_blank(),
-        axis.text.y = element_blank()
+        axis.title.y = element_blank(),
+        axis.text.y = element_text(size = 7, hjust = 1)
     )
@@
     left_join(count(., group, name = "n_total")) %>% 
-    mutate(group = paste0(group,  " (N = ", n_total, ")")) %>% 
+    mutate(
+        group = paste0(group,  " (N = ", n_total, ")"),
+        translation = paste(translation, "(", percent(lv, accuracy = 0.01), ")"),
+        translation_ord = reorder_within(translation, lv, group)
+    ) %>% 
     ggplot() +
-    aes(x = lv, y = reorder_within(translation, lv, group), fill = group, colour = group) +
+    aes(x = lv, y = translation_ord, fill = group, colour = group) +
     facet_wrap(~group, scales = "free_y") +
-    geom_text(aes(label = percent(lv, accuracy = 0.1)), size = 3, hjust = 0) +
+    geom_segment(aes(x = 0, xend = lv, y = translation_ord, yend = translation_ord)) +
+    geom_point(size = 1) +
     labs(
@@
         axis.title.y = element_blank(),
-        axis.text.y = element_text(size = 7, hjust = 1),
+        axis.text.y = element_text(size = 3, hjust = 1),
         panel.grid.major.x = element_line(colour = "grey", linetype = "dotted")
@@
         y = "Number of trials", 
-        title = "Higher-frequency neighbours\n(stimuli translations)\n"
+        title = "Cross-language neighbours"
     ) +
@@
     )
+```
+
+```{r}
+#| label: model-rocs
+model_rocs %>% 
+    bind_rows(.id = "model") %>% 
+    ggplot(aes(1-specificity, sensitivity, colour = model)) +
+    facet_wrap(~model) +
+    geom_line(aes(group = .draw), alpha = 0.1) +
+    labs(x = "1 - Specificity", y = "Sensitivity", colour = "Model") + 
+    scale_y_continuous(labels = percent) +
+    scale_x_continuous(labels = percent) +
+    theme_custom()
+
 ```
@@
     clean_names() %>% 
-    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4))) %>% 
+    mutate(across(estimate:q97_5, ~logit_to_prob(., variable))) %>%
     group_split(variable) %>% 
@@
     mutate(
-        .variable_name = .variable %>% 
-            str_remove("b_") %>% 
-            factor(levels = names(str_repl), labels = str_repl),
-        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
+        .variable_name = factor(
+            gsub("b_", "", .variable),
+            levels = names(str_repl),
+            labels = str_repl
+        ),
+        .value = logit_to_prob(.value, .variable)
     ) %>% 
@@
     ) 
-```
-
-```{r}
-#| label: coefs-word
-empirical_means <- responses %>% 
-    group_by(group) %>% 
-    summarise(
-        successes = sum(correct, na.rm = TRUE),
-        n = n(),
-        .groups = "drop"
-    ) %>% 
-    mutate(
-        prop = prop_adj(successes, n),
-        prop_se = prop_adj_se(successes, n)
-    )
-
-empirical_accuracy <- responses %>% 
-    mutate(translation = paste0(word_1, " /", ipa_flat_1, "/ - ",  word_2, " /", ipa_flat_2, "/ - ", percent(lv, accuracy = 1))) %>% 
-    group_by(group, translation, lv) %>% 
-    summarise(
-        successes = sum(correct, na.rm = TRUE),
-        n = n(),
-        .groups = "drop"
-    ) %>% 
-    rowwise() %>% 
-    mutate(
-        prop = prop_adj(successes, n),
-        prop_se = prop_adj_se(successes, n), 
-    )
-
-fixed_intercept <- fixef(fit_4)[1 ,c(1, 3, 4)] %>% 
-    inv_logit_scaled()
-
-nd_re_words <- expand.grid(
-    translation_id = unique(fit_4$data$translation_id),
-    nd_std = 0,
-    freq_zipf_2_std = 0,
-    lv_std = 0,
-    group = NA
-)
-
-post_re <- add_epred_draws(nd_re_words, fit_4, ndraws = 50,
-                           re_formula = ~ (1 | translation_id)) %>% 
-    group_by(translation_id, nd_std, freq_zipf_2_std, lv_std, group) %>% 
-    summarise(
-        .mean = mean_qi(.epred)[,1], 
-        .lower = mean_qi(.epred)[,2],
-        .upper = mean_qi(.epred)[,3],
-        .groups = "drop"
-    ) %>% 
-    ungroup() %>% 
-    select(-c(group)) %>% 
-    left_join(select(stimuli, group, translation, translation_id, lv)) %>% 
-    mutate(translation = paste0(translation, " - ", percent(lv, accuracy = 1))) %>% 
-    left_join(select(empirical_accuracy, translation, n, prop, prop_se))
-
-post_re %>% 
-    ggplot() +
-    aes(x = .mean, y = reorder_within(translation, prop, group)) +
-    facet_wrap(~group, scales = "free_y") +
-    geom_vline(data = empirical_means, aes(xintercept = prop), size = 0.75, colour = "grey") +
-    geom_errorbar(aes(xmin = .lower, xmax = .upper), size = 0.5, colour = "grey", width = 0.5) + 
-    geom_point(aes(x = .mean), size = 0.5, colour = "grey") + 
-    geom_point(aes(x = prop, colour = group), size = 1, shape = 15) +
-    labs(
-        x = "Percentage of correct responses",
-        y = "Item",
-        title = "Observed translation accuracy",
-        subtitle = "Dots indicate proportions of correct responses, error bars indicate +/-1 SE",
-        colour = "Credible Interval (CrI)",
-        caption = "Percentages next to items indicate the Levenshtein distance between both translations."
-    ) +
-    scale_x_continuous(limits = c(0, 1), labels = percent) +
-    scale_y_reordered() +
-    theme(
-        legend.position = "none",
-        legend.key = element_rect(fill = NA, colour = NA),
-        axis.text.x = element_text(size = 7),
-        axis.text.y = element_text(size = 5, hjust = 1),
-        axis.ticks = element_blank(),
-        axis.title = element_text(size = 10),
-        axis.title.y = element_blank(),
-        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
-        plot.caption.position = "plot",
-        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
-        plot.title.position = "plot",
-        plot.title = element_text(hjust = 0, size = 13),
-        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
-    )
-
-ggsave(here("img", "accuracy.png"), height = 9, width = 10, dpi = 1000)
 ```
@@
     freq_zipf_2_std = 0,
-    nd_std = scale_values(c(0, 3, 6, 10, 15, 25), fit_4$data$nd)
+    nd_std = scale_values(c(0, 3, 6), fit_4$data$nd)
 )
@@
 m <- add_epred_draws(nd, fit_4, re_formula = NA, ndraws = 50) %>% 
-    mutate(nd_lab = rescale_values(nd_std, fit_4$data$nd) %>% 
+    mutate(
+        nd_lab = rescale_values(nd_std, fit_4$data$nd) %>% 
             paste0(" neighbours") %>% 
-               factor(levels = paste0(c(0, 3, 6, 10, 15, 25), " neighbours"))
+            factor(levels = paste0(c(0, 3, 6), " neighbours"))
     )
@@
     geom_line(aes(group = interaction(nd_std, .draw)),
-              colour = "grey", size = 0.75, alpha = 0.25) +
+              colour = my_colours[1], size = 0.75, alpha = 0.1) +
     geom_hline(yintercept = 0.5, colour = "grey", size = 1) +
-    stat_summary(fun = mean, geom = "line", size = 1, colour = my_colours[2]) +
+    stat_summary(fun = mean, geom = "line", size = 1, colour = my_colours[1]) +
     labs(
@@
 #| warning: false
-coefs <- fixef(questionnaire_fit_4) %>% 
+questionnaire_coefs <- fixef(questionnaire_fit_4) %>% 
     as.data.frame() %>% 
@@
     clean_names() %>% 
-    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)))
+    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", plogis(.), ./4)))
 
@@
 
-coefs %>% 
+questionnaire_coefs %>% 
     mutate(
@@
 
-post_draws <-  gather_draws(questionnaire_fit_4, `b_.*`, `sd_.*`, regex = TRUE)
+questionnaire_post_draws <- gather_draws(questionnaire_fit_4, `b_.*`, `sd_.*`, regex = TRUE)
 
 # fixed effects
-post_draws %>% 
+questionnaire_post_draws %>% 
     filter(str_detect(.variable, "b_")) %>% 
@@
     freq_zipf_2_std = 0,
-    nd_std = scale_values(c(0, 3, 6, 10, 15, 25), questionnaire_fit_4$data$nd)
-)
-
-questionnaire_m <- add_epred_draws(nd, questionnaire_fit_4, re_formula = NA, ndraws = 50) %>% 
-    mutate(nd_lab = rescale_values(nd_std, questionnaire_fit_4$data$nd) %>% 
+    nd_std = scale_values(c(0, 3, 6), questionnaire_fit_4$data$nd)
+)
+
+questionnaire_m <- add_epred_draws(
+    questionnaire_nd,
+    questionnaire_fit_4, 
+    re_formula = NA,
+    ndraws = 50
+) %>% 
+    mutate(
+        nd_lab = rescale_values(nd_std, questionnaire_fit_4$data$nd) %>% 
+            round(2) %>% 
             paste0(" neighbours") %>% 
-               factor(levels = paste0(c(0, 3, 6, 10, 15, 25), " neighbours"))
+            factor(levels = paste0(c(0, 3, 6), " neighbours"))
     )
@@
     geom_line(aes(group = interaction(nd_std, .draw)),
-              colour = "grey", size = 0.75, alpha = 0.25) +
+              colour = my_colours[2], size = 0.75, alpha = 0.1) +
     geom_hline(yintercept = 0.5, colour = "grey", size = 1) +
@@
 # Joint analyses
+
+## Regression coefficients
+
+```{r}
+#| label: fig-joint-coefs
+#| fig-height: 5
+#| fig-width: 8.5
+#| fig-cap: "Estimated posterior distributions of coefficients in Model 4.
+A\\) Population-level effects.
+Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
+Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
+Black dots represent the mean of each distribution.
+B\\) Participant\\-level coefficient variability.
+The model estimated participant\\-level coefficients to account for the dependency between responses from the same participant.
+Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (*SD*).
+C\\) Correlation between participant\\-level effects.
+The model allowed participant-level coefficients to co\\-vary.
+This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each  estimated correlation."
+# fixed effects
+list("Study 1" = post_draws, "Study 2" = questionnaire_post_draws) %>% 
+    bind_rows(.id = "study") %>% 
+    filter(str_detect(.variable, "b_")) %>% 
+    mutate(
+        .variable_name = factor(
+            gsub("b_", "", .variable),
+            levels = names(str_repl),
+            labels = str_repl
+        ),
+        .value = logit_to_prob(.value, .variable)
+    ) %>% 
+    arrange(.variable) %>% 
+    ggplot(aes(.value, fct_rev(.variable_name), fill = study, colour = study)) +
+    stat_slab(size = 1.25, position = position_nudge(y = 0.15), fill = NA, scale = 0.5) +
+    stat_pointinterval(position = position_dodge(width = 0.25), .width = 0.95, size = 1.35) +
+    scale_fill_manual(values = my_colours, na.translate = FALSE) +
+    scale_x_continuous(labels = ~percent(., accuracy = 1), breaks = seq(-1, 1, 0.1)) +
+    labs(
+        x = "P(Correct)", 
+        y = "Posterior probability density",
+        fill = "Study",
+        colour = "Study"
+    ) +
+    theme(
+        legend.position = "top",
+        legend.title = element_blank(),
+        axis.title.y = element_blank(),
+        axis.text = element_text(colour = "black", hjust = 1, size = 12),
+        axis.ticks.y = element_blank(),
+        panel.grid.major.x = element_line(colour = "grey", size = 0.5, linetype = "dotted")
+    )
+
+```
+
+## Marginal effects
+
+### Population-level marginal effects
 
@@
 We computed these posterior predictions for each group of participants, and plot tem in separate panels."
-
+#| warnings: false
+#| messages: false
 bind_rows(list("Study 1" = m, "Study 2" = questionnaire_m), .id = "study") %>% 
@@
     geom_line(aes(group = interaction(nd_std, .draw, study)),
-              size = 0.75, alpha = 0.05) +
+              size = 0.75, alpha = 0.1) +
     geom_hline(yintercept = 0.5, size = 1) +
@@
         legend.position = "none",
-        axis.text.x = element_text(size = 10),
         plot.title = element_blank()
@@
 ```
+
+### Group-level marginal effects
+
+#### Word-level marginal effects
+
+::: {.panel-tabset}
+
+##### Catalan to English (cat-ENG)
+
+```{r}
+#| label: fig-coefs-word-cateng
+#| fig-cap: "Word-level model posterior predicted and observed mean accuracy.
+Intervals represent the 95% posterior credible interval of the accuracy each item was responded to in Study 1 (blue) and Study 2 (red).
+dots indicate the observed proportion of participants that translated the word correctly."
+#| fig-height: 13
+#| fig-width: 5
+#| warnings: false
+#| messages: false
+empirical_means <- list("Study 1" = responses, "Study 2" = questionnaire_responses) %>% 
+    bind_rows(.id = "study") %>% 
+    group_by(study, group) %>% 
+    summarise(successes = sum(correct, na.rm = TRUE), n = n(), .groups = "drop") %>% 
+    mutate(
+        prop = prop_adj(successes, n),
+        prop_se = prop_adj_se(successes, n)
+    )
+
+empirical_accuracy <- list("Study 1" = responses, "Study 2" = questionnaire_responses) %>% 
+    bind_rows(.id = "study") %>% 
+    mutate(translation = paste0(translation, " (", percent(lv, accuracy = 1), ")")) %>% 
+    group_by(study, group, translation, lv) %>% 
+    summarise(prop = mean(correct), .groups = "drop")
+
+nd_re_words <- expand.grid(
+    translation_id = unique(fit_4$data$translation_id),
+    nd_std = 0,
+    freq_zipf_2_std = 0,
+    lv_std = 0,
+    group = NA
+)
+
+questionnaire_nd_re_words <- expand.grid(
+    translation_id = unique(questionnaire_fit_4$data$translation_id),
+    nd_std = 0,
+    freq_zipf_2_std = 0,
+    lv_std = 0,
+    group = NA
+)
+
+post_re <- add_epred_draws(nd_re_words, fit_4, ndraws = 50, re_formula = ~ (1 | translation_id))
+questionnaire_post_re <- add_epred_draws(questionnaire_nd_re_words, questionnaire_fit_4, ndraws = 50, re_formula = ~ (1 | translation_id))
+
+post_re_join <- list("Study 1" = post_re, "Study 2" = questionnaire_post_re) %>% 
+    bind_rows(.id = "study") %>% 
+    group_by(study, translation_id, nd_std, freq_zipf_2_std, lv_std, group) %>% 
+    summarise(
+        .mean = mean_qi(.epred)[,1], 
+        .lower = mean_qi(.epred)[,2],
+        .upper = mean_qi(.epred)[,3],
+        .groups = "drop"
+    ) %>% 
+    ungroup() %>% 
+    select(-c(group)) %>% 
+    left_join(select(stimuli, group, translation, translation_id, lv)) %>% 
+    mutate(
+        translation = paste0(translation, " (", percent(lv, accuracy = 1), ")"),
+        translation_ord = reorder_within(translation, lv, group),
+        group = factor(group, levels = c("cat-ENG", "spa-ENG", "cat-SPA"), ordered = TRUE)
+    ) %>% 
+    left_join(select(empirical_accuracy, study, group, translation, prop))
+
+post_re_join %>% 
+    filter(group=="cat-ENG") %>% 
+    ggplot(aes(x = .mean, y = translation_ord, colour = study, fill = study)) +
+    geom_errorbar(aes(xmin = .lower, xmax = .upper),
+                  size = 0.5, width = 0,
+                  position = position_dodge(width = 0.5)) + 
+    geom_point(aes(x = prop), size = 2, shape = 18,
+               position = position_dodge(width = 0.5)) +
+    geom_smooth(aes(group = study), orientation = "y", alpha = 0.1, size = 0.5,
+                method = "glm", method.args = list(family = "binomial")) +
+    labs(
+        x = "Percentage of correct responses",
+        y = "Item",
+        colour = "Study",
+        fill = "Study"
+    ) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    scale_y_reordered() +
+    theme(
+        legend.position = "top",
+        legend.key = element_rect(fill = NA, colour = NA),
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 5, hjust = 1),
+        axis.ticks = element_blank(),
+        axis.title = element_text(size = 10),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
+        plot.caption.position = "plot",
+        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0, size = 13),
+        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
+    )
+```
+
+##### Spanish to English (spa-ENG)
+
+```{r}
+#| label: fig-coefs-word-spaeng
+#| fig-cap: "Word-level model posterior predicted and observed mean accuracy.
+Intervals represent the 95% posterior credible interval of the accuracy each item was responded to in Study 1 (blue) and Study 2 (red).
+dots indicate the observed proportion of participants that translated the word correctly."
+#| fig-height: 13
+#| fig-width: 5
+#| warnings: false
+#| messages: false
+post_re_join %>% 
+    filter(group=="spa-ENG") %>% 
+    ggplot(aes(x = .mean, y = translation_ord, colour = study, fill = study)) +
+    geom_errorbar(aes(xmin = .lower, xmax = .upper),
+                  size = 0.75, width = 0,
+                  position = position_dodge(width = 0.5)) + 
+    geom_point(aes(x = prop), size = 1.25, shape = 18,
+               position = position_dodge(width = 0.5)) +
+    geom_smooth(aes(group = study), orientation = "y", alpha = 0.1, size = 0.5,
+                method = "glm", method.args = list(family = "binomial")) +
+     labs(
+        x = "Percentage of correct responses",
+        y = "Item",
+        colour = "Study",
+        fill = "Study"
+    ) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    scale_y_reordered() +
+    theme(
+        legend.position = "top",
+        legend.key = element_rect(fill = NA, colour = NA),
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 5, hjust = 1),
+        axis.ticks = element_blank(),
+        axis.title = element_text(size = 10),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
+        plot.caption.position = "plot",
+        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0, size = 13),
+        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
+    )
+```
+
+##### Catalan to Spanish (cat-SPA)
+
+```{r}
+#| label: fig-coefs-word-catspa
+#| fig-cap: "Word-level model posterior predicted and observed mean accuracy.
+Intervals represent the 95% posterior credible interval of the accuracy each item was responded to in Study 1 (blue) and Study 2 (red).
+dots indicate the observed proportion of participants that translated the word correctly."
+#| fig-height: 13
+#| fig-width: 5
+#| warnings: false
+#| messages: false
+post_re_join %>% 
+    filter(group=="spa-ENG") %>% 
+    ggplot(aes(x = .mean, y = translation_ord, colour = study, fill = study)) +
+    geom_errorbar(aes(xmin = .lower, xmax = .upper),
+                  size = 0.75, width = 0,
+                  position = position_dodge(width = 0.5)) + 
+    geom_point(aes(x = prop), size = 1.25, shape = 18,
+               position = position_dodge(width = 0.5)) +
+    geom_smooth(aes(group = study), orientation = "y", alpha = 0.1, size = 0.5,
+                method = "glm", method.args = list(family = binomial)) +
+    labs(
+        x = "Percentage of correct responses",
+        y = "Item",
+        colour = "Study",
+        fill = "Study"
+    ) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    scale_y_reordered() +
+    theme(
+        legend.position = "top",
+        legend.key = element_rect(fill = NA, colour = NA),
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 5, hjust = 1),
+        axis.ticks = element_blank(),
+        axis.title = element_text(size = 10),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
+        plot.caption.position = "plot",
+        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0, size = 13),
+        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
+    )
+```
+
+:::
+
+#### Participant-level marginal effects
+
+::: {.panel-tabset}
+
+##### Catalan to English (cat-ENG)
+
+```{r}
+#| label: fig-coefs-participant-cateng
+#| fig-cap: "Participant-level model posterior predicted and observed mean accuracy.
+Intervals represent the 95% posterior credible interval of the accuracy of each participant from Study 1 (blue) and Study 2 (red).
+Dots indicate the observed proportion of words participants translated the correctly."
+#| fig-height: 8
+#| fig-width: 5
+#| warnings: false
+#| messages: false
+empirical_means <- list("Study 1" = responses, "Study 2" = questionnaire_responses) %>% 
+    bind_rows(.id = "study") %>% 
+    group_by(study, group) %>% 
+    summarise(successes = sum(correct, na.rm = TRUE), n = n(), .groups = "drop") %>% 
+    mutate(
+        prop = prop_adj(successes, n),
+        prop_se = prop_adj_se(successes, n)
+    )
+
+empirical_accuracy <- list("Study 1" = responses, "Study 2" = questionnaire_responses) %>% 
+    bind_rows(.id = "study") %>% 
+    mutate(translation = paste0(translation, " (", percent(lv, accuracy = 1), ")")) %>% 
+    group_by(study, group, participant_id) %>% 
+    summarise(prop = mean(correct), .groups = "drop")
+
+nd_re_participants <- expand.grid(
+    participant_id = unique(fit_4$data$participant_id),
+    nd_std = 0,
+    freq_zipf_2_std = 0,
+    lv_std = 0
+)
+
+questionnaire_nd_re_participants <- expand.grid(
+    participant_id = unique(questionnaire_fit_4$data$participant_id),
+    nd_std = 0,
+    freq_zipf_2_std = 0,
+    lv_std = 0
+)
+
+participants_joint <- list(
+    "Study 1" = select(participants, participant_id, group),
+    "Study 2" = select(questionnaire_participants, participant_id, group)
+) %>% 
+    bind_rows(.id = "study")
+
+post_re <- add_epred_draws(
+    nd_re_participants, 
+    fit_4, 
+    ndraws = 50,
+    re_formula = ~ (1 | participant_id)
+)
+
+questionnaire_post_re <- add_epred_draws(
+    questionnaire_nd_re_participants, 
+    questionnaire_fit_4,
+    ndraws = 50, 
+    re_formula = ~ (1 | participant_id)
+)
+
+post_re_join <- list("Study 1" = post_re, "Study 2" = questionnaire_post_re) %>% 
+    bind_rows(.id = "study") %>% 
+    group_by(study, participant_id) %>% 
+    summarise(
+        .mean = mean_qi(.epred)[,1], 
+        .lower = mean_qi(.epred)[,2],
+        .upper = mean_qi(.epred)[,3],
+        .groups = "drop"
+    ) %>% 
+    ungroup() %>% 
+    left_join(select(participants_joint, participant_id, group)) %>% 
+    mutate(group = factor(group, levels = c("cat-ENG", "spa-ENG", "cat-SPA"), ordered = TRUE)) %>% 
+    left_join(select(empirical_accuracy, study, group, participant_id, prop))
+
+post_re_join %>% 
+    mutate(participant_id = str_trunc(participant_id, 10)) %>% 
+    filter(group=="cat-ENG") %>% 
+    ggplot(aes(x = .mean, y = participant_id, colour = study, fill = study)) +
+    facet_wrap(~study, scales = "free_y") +
+    geom_errorbar(aes(xmin = .lower, xmax = .upper),
+                  size = 1, width = 0,
+                  position = position_dodge(width = 0.5)) + 
+    geom_point(aes(x = prop), size = 2, shape = 18,
+               position = position_dodge(width = 0.5)) +
+    labs(
+        x = "Percentage of correct responses",
+        y = "Participant",
+        colour = "Study",
+        fill = "Study"
+    ) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    scale_y_reordered() +
+    theme(
+        legend.position = "none",
+        legend.key = element_rect(fill = NA, colour = NA),
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 5, hjust = 1),
+        axis.ticks = element_blank(),
+        axis.title = element_text(size = 10),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
+        plot.caption.position = "plot",
+        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0, size = 13),
+        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
+    )
+```
+
+##### Spanish to English (spa-ENG)
+
+```{r}
+#| label: fig-coefs-participant-spaeng
+#| fig-cap: "Participant-level model posterior predicted and observed mean accuracy.
+Intervals represent the 95% posterior credible interval of the accuracy of each participant from Study 1 (blue) and Study 2 (red).
+Dots indicate the observed proportion of words participants translated the correctly."
+#| fig-height: 8
+#| fig-width: 5
+#| warnings: false
+#| messages: false
+post_re_join %>% 
+    filter(group=="spa-ENG") %>% 
+    ggplot(aes(x = .mean, y = participant_id, colour = study, fill = study)) +
+    facet_wrap(~study, scales = "free_y") +
+    geom_errorbar(aes(xmin = .lower, xmax = .upper),
+                  size = 1, width = 0,
+                  position = position_dodge(width = 0.5)) + 
+    geom_point(aes(x = prop), size = 2, shape = 18,
+               position = position_dodge(width = 0.5)) +
+    labs(
+        x = "Percentage of correct responses",
+        y = "Participant",
+        colour = "Study",
+        fill = "Study"
+    ) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    scale_y_reordered() +
+    theme(
+        legend.position = "none",
+        legend.key = element_rect(fill = NA, colour = NA),
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 5, hjust = 1),
+        axis.ticks = element_blank(),
+        axis.title = element_text(size = 10),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
+        plot.caption.position = "plot",
+        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0, size = 13),
+        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
+    )
+```
+
+##### Catalan to Spanish (cat-SPA)
+
+```{r}
+#| label: fig-coefs-participant-catspa
+#| fig-cap: "Participant-level model posterior predicted and observed mean accuracy.
+Intervals represent the 95% posterior credible interval of the accuracy of each participant from Study 1 (blue) and Study 2 (red).
+Dots indicate the observed proportion of words participants translated the correctly."
+#| fig-height: 8
+#| fig-width: 5
+#| warnings: false
+#| messages: false
+post_re_join %>% 
+    filter(group=="cat-SPA") %>% 
+    ggplot(aes(x = .mean, y = participant_id, colour = study, fill = study)) +
+    facet_wrap(~study, scales = "free_y") +
+    geom_errorbar(aes(xmin = .lower, xmax = .upper),
+                  size = 1, width = 0,
+                  position = position_dodge(width = 0.5)) + 
+    geom_point(aes(x = prop), size = 2, shape = 18,
+               position = position_dodge(width = 0.5)) +
+    labs(
+        x = "Percentage of correct responses",
+        y = "Participant",
+        colour = "Study",
+        fill = "Study"
+    ) +
+    scale_x_continuous(limits = c(0, 1), labels = percent) +
+    scale_y_reordered() +
+    theme(
+        legend.position = "none",
+        legend.key = element_rect(fill = NA, colour = NA),
+        axis.text.x = element_text(size = 7),
+        axis.text.y = element_text(size = 5, hjust = 1),
+        axis.ticks = element_blank(),
+        axis.title = element_text(size = 10),
+        axis.title.y = element_blank(),
+        panel.grid.major.x = element_line(size = 0.15, linetype = "dotted", colour = "grey"),
+        plot.caption.position = "plot",
+        plot.caption = element_text(size = 9, hjust = 0, face = "plain", margin = margin(t = unit(3, "cm"))),
+        plot.title.position = "plot",
+        plot.title = element_text(hjust = 0, size = 13),
+        plot.subtitle = element_text(hjust = 0, face = "plain", size = 13)
+    )
+```
+
+:::

### 2022-11-21 index.qmd
Add questionnaire analyses and restructure the targets script.

@@
     gt(rowname_col = "group") %>% 
+    fmt_integer(n) %>% 
     fmt_number(freq_2_median:nd_sd) %>%
@@
 ```{r}
-#| label: model-rocs
+#| label: fig-model-rocs
 model_rocs %>% 
@@
 #| label: tbl-model-coefs
-coefs <- fixef(fit_4) %>% 
-    as.data.frame() %>% 
-    rownames_to_column("variable") %>% 
+coefs <- model_parameters$fit_5 %>%
+    as_tibble() %>% 
     clean_names() %>% 
-    mutate(across(estimate:q97_5, ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)))
+    mutate(
+        parameter = str_remove(parameter, "b_"),
+        across(median:ci_high, ~ifelse(parameter=="(Intercept)", plogis(.), ./4))
+    ) %>% 
+    select(-ci)
 
@@
     "lv_std" = "Levenshtein [+1 SD]",
+    "groupeng_spa" = "Group [ENG vs. SPA]",
+    "groupeng_spa" = "Group [spa/cat-ENG vs. cat-SPA]",
+    "groupcateng_spaeng" = "Group [cat-ENG vs. spa-ENG]",
     "nd_std:lv_std" = "Neigbours \u00d7 Cognateness"
@@
     mutate(
-        variable = variable %>% 
-            str_remove("b_") %>% 
-            factor(levels = names(str_repl), labels = str_repl)
-    ) %>% 
-    select(-est_error) %>% 
+        parameter = factor(parameter, levels = names(str_repl), labels = str_repl),
+        across(starts_with("ci_"), ~percent(., accuracy = 0.01)),
+        ci = paste("[", ci_low, ", ", ci_high, "]")
+    ) %>% 
+    select(-starts_with("ci_"), -c(rhat, ess)) %>% 
+    relocate(parameter, median, ci, log_bf) %>% 
     gt() %>% 
-    fmt_percent(2:4) %>% 
-    cols_merge(c("q2_5", "q97_5"), pattern = "[{1}, {2}]") %>% 
+    fmt_percent(median) %>% 
+    fmt_number(log_bf) %>% 
     cols_label(
-        variable = md("**Predictor**"),
-        estimate = md("**Mean**"),
-        q2_5 = md("**95\\% CrI**"),
+        parameter = "Predictor",
+        median = "Median",
+        ci = "95% CrI",
+        log_bf = "Log(Bayes Factor)"
     ) %>% 
     tab_source_note(tab_caption) %>% 
-    cols_align(align = "left")
+    cols_align(align = "right")
 ```
@@
     )
+```
+
+```{r}
+#| label: fig-questionnaire-model-rocs
+questionnaire_model_rocs %>% 
+    bind_rows(.id = "model") %>% 
+    ggplot(aes(1-specificity, sensitivity, colour = model)) +
+    facet_wrap(~model) +
+    geom_line(aes(group = .draw), alpha = 0.1) +
+    labs(x = "1 - Specificity", y = "Sensitivity", colour = "Model") + 
+    scale_y_continuous(labels = percent) +
+    scale_x_continuous(labels = percent) +
+    theme_custom()
+
 ```
@@
 :::
+
+## Confidence ratings and probability of correct responses
+
+```{r}
+#| label: fig-confidence-correct 
+questionnaire_responses %>%
+    mutate(lv = cut(lv, breaks = seq(0, 1, 0.1), include.lowest = TRUE)) %>% 
+    group_by(word_1, group, lv) %>% 
+    summarise(
+        n = n(),
+        confidence_mean = mean(confidence),
+        knowledge_prop = mean(knowledge),
+        correct = sum(correct),
+        .groups = "drop"
+    ) %>% 
+    mutate(correct_prop = prop_adj(correct, n)) %>% 
+    arrange(correct_prop) %>% 
+    ggplot(aes(confidence_mean, correct_prop, label = word_1)) +
+    facet_grid(group~lv) +
+    geom_point(alpha = 0.5) +
+    geom_text_repel() +
+    labs(x = "Mean confidence score", y = "P(Correct)") +
+    scale_y_continuous(labels = percent, breaks = seq(0, 1, 0.25), limits = c(0, 1))
+
+```
+
+```{r}
+#| label: tbl-confidence-correct 
+a <- questionnaire_responses %>%
+    group_by(translation, group, lv) %>% 
+    summarise(
+        n = n(),
+        confidence_mean = mean(confidence),
+        knowledge_prop = mean(knowledge),
+        correct_prop = mean(correct),
+        .groups = "drop"
+    ) %>% 
+    arrange(group, lv) %>% 
+    select(-n) %>% 
+    gt(groupname_col = "group", rowname_col = "translation") %>% 
+    tab_header(title = "Word-level accuracy scores",
+               subtitle = "Observed proportion of correct translation of every word used in the questionnaire (only English participants translation Catalan or Spanish), with information about their Levenshtein similarity, and mean confidence of the translation and prior knowledge with the preented word reported by participants.") %>%
+    fmt_percent(c(lv, ends_with("prop"))) %>% 
+    fmt_number(confidence_mean) %>% 
+    cols_label(
+        translation = "Translation (ordered by Levenshtein similarity)",
+        lv = "Levenshtein similarity",
+        confidence_mean = "Mean confidence (0-5)",
+        knowledge_prop = "Knows the word (%)",
+        correct_prop = "Accuracy (%)"
+    ) %>% 
+    tab_style(cell_text(weight = "bold"), cells_column_labels()) %>% 
+    tab_style(list(cell_text(style = "italic", weight = "bold"), cell_fill("grey")), cells_row_groups()) %>% 
+    tab_style(cell_text(align = "left"), cells_title(groups = "subtitle")) %>% 
+    data_color(
+        columns = correct_prop,
+        colors = col_numeric(
+            palette = rev(c("green", "yellow", "orange", "red")),
+            domain = c(0, 1)
+        )
+    ) %>% 
+    tab_footnote(
+        footnote = "Presented word, correct translation, and their correspondent IPA phonological transcription",
+        locations = cells_column_labels(columns = translation)
+    ) %>% 
+    tab_footnote(
+        footnote = "Levenshtein distance between the IPA transcriptions of the presented word and its correct translation, divided by the number of characters of the longest transcription, and substracted from 1 to make it a similarity score.",
+        locations = cells_column_labels(columns = lv)
+    ) %>% 
+    tab_footnote(
+        footnote = "Mean of the reported confidence of participants' translations, in a scale from 0 to 5.",
+        locations = cells_column_labels(columns = confidence_mean)
+    ) %>% 
+    tab_footnote(
+        footnote = "Proportion of participants that reported knowing the translation of the word in Catalan or Spanish.",
+        locations = cells_column_labels(columns = knowledge_prop)
+    ) %>% 
+    tab_footnote(
+        footnote = "Proportion of participants that translated the presented word to English correctly, colour-coded with warm colours assiged to higher accuracy scores, and colder colours assigned to lower accuracy scores.",
+        locations = cells_column_labels(columns = correct_prop)
+    ) %>% 
+    tab_stubhead(label = "Translation")
+
+gtsave(a, "accuracy-confidence.html")
+
+```

### 2022-11-21 R/01_experiment.R

@@ -1,0 +1,239 @@
+# import data ------------------------------------------------------------------
+get_exp_raw <- function(){
+    # set parameters
+    spanish_cities <- c("Lorca", "Albacete", "Cieza", "Cartagena", "Murcia", "España", "Málaga", "Oviedo", "Santander", "Granada")
+    
+    responses_path <- list.files("data/experiment/raw/", full.names = TRUE)
+    
+    filenames <- list.files("data/experiment/raw") %>% 
+        str_extract(".*?\\_") %>% 
+        str_remove("_")
+    
+    # participant files
+    exp_raw <- responses_path %>% 
+        map(fread, na.string = "") %>%  # import participant files
+        map(mutate, participant = str_trunc(participant, width = 10, side = "right")) %>% 
+        set_names(filenames) %>% # label each dataset with the participant's ID
+        bind_rows(.id = "filename") %>% # merge datasets
+        as_tibble() %>%
+        clean_names() %>% 
+        rename(participant_id = participant) %>% 
+        # select relevant variables and rename if necessary
+        select(
+            participant_id, test_language, trial_id, word, soundfile, key_pressed, 
+            response = input_text, key_press_time, error, age, date, city,
+            matches("language_l|language_s|language_c|demo_|setup_"),
+            -matches("_rt|_time"), key_press_time
+        ) %>% 
+        rename(demo_impairment = demo_language_key_keys) %>% 
+        rename_all(gsub, pattern = "_key_keys|key_keys", replacement = "") %>% 
+        # clean text input by participants (because of typos of need to translate) and redefine location    
+        group_by(participant_id) %>% 
+        mutate(
+            across(c(starts_with("demo_"), starts_with("l"), starts_with("setup_"), age, city, setup_location, setup_noise), first_non_na),
+            date = max(date, na.rm = TRUE)
+        ) %>% 
+        ungroup() %>% 
+        mutate(
+            across(c(starts_with("language_"), city, starts_with("demo_"), starts_with("l")), clean_input_text),
+            city = str_to_sentence(city),
+            country = ifelse(city %in% spanish_cities, "Spain", "UK"),
+            date = as_datetime(ymd_hms(date)),
+            demo_sex = case_when(
+                country %in% "UK" & demo_sex %in% "M" ~ "Male",
+                country %in% "UK" & demo_sex %in% "F" ~ "Female",
+                country %in% "Spain" & demo_sex %in% "M" ~ "Female",
+                country %in% "Spain" & demo_sex %in% "H" ~ "Male"
+            ),
[…]

### 2022-11-21 R/02_models.R

@@ -1,72 +1,0 @@
-# models
-
-get_model_fit <- function(name, formula, prior, ...){
-    
-    stan_path <- here("stan", paste0(name, ".stan"))
-    results_path <- here("results", paste0(name, ".rds"))
-    
-    brm(
-        formula = formula,
-        family = bernoulli("logit"),
-        prior = prior,
-        save_pars = save_pars(all = TRUE),
-        backend = "cmdstanr",
-        iter = 1000,
-        cores = 4,
-        chains = 4,
-        seed = 888,
-        control = list(adapt_delta = 0.95),
-        save_model = stan_path,
-        file = results_path,
-        sample_prior = "yes",
-        ...
-    )
-}
-
-# leave-one-out cross-validation
-get_model_loos <- function(fits){
-    path <- here("results", "model_loos.rds")
-    if (file.exists(path)){
-        loo <- readRDS(here(path))
-    } else {
-        loo <- loo_compare(map(fits, loo))
-        saveRDS(loo, path)
-    }
-    return(loo)
-}
-
-# get ROC curve of Bayesian model
-get_roc_curve <- function(newdata, object, ...) {
-    
-    # enquote response variable and get brmsfit family
-    resp_var <- formula(object)[["formula"]][[2]]
-    resp_var <- enquo(resp_var)
-    model_fam <- object[["family"]][["family"]]
-    
-    # object must be a brmsfit object with a supported family
-    supported <- c("bernoulli", "binomial", "categorical", "cumulative", "sratio", "cratio", "acat")
-    stopifnot(is.brmsfit(object))
-    if (!(model_fam %in% supported)) stop(paste0("model family must be one of: ", paste0(supported, collapse = ", ")))
-    
-    if (model_fam %in% c("binomial", "bernoulli")) {
-        roc_values <- add_epred_draws(newdata, object, ndraws = 50) %>% 
-            ungroup() %>% 
-            mutate(!!resp_var := as.factor(!!resp_var)) %>% 
-            # generate a ROC curve for each posterior draw
-            split(.$.draw) %>%
-            map_dfr(~roc_curve(., truth = !!resp_var, .epred, event_level = "second"), .id = ".draw") 
-        
-    } else {
-        cat_symbols <- syms(as.character(unique(get_y(object))))
-        roc_values <- add_epred_draws(newdata, object, ...) %>% 
-            ungroup() %>% 
-            mutate(!!resp_var := as.factor(!!resp_var)) %>%
[…]

### 2022-11-21 R/utils.R
Add questionnaire analyses and restructure the targets script.

@@ -1,3 +1,41 @@
+# run all targets as an RStudio job
+make <- function(){
+    job::job(
+        title = "Translation Elicitation", {{ 
+            targets::tar_make() 
+        }}
+    )
+}
+
+# load all built targets (and packages)
+tar_load_all <- function(){
+    invisible({
+        suppressMessages({
+            tar_load_globals()
+        })
+        tars <- tar_objects()
+        message("Loading targets: ", paste0(tars, collapse = ", "))
+        lapply(tars, tar_load_raw, envir = .GlobalEnv)
+    })
+}
+
+unmake <- function(keep_models = TRUE) {
+    path <- list.files("results", full.names = TRUE)
+    tar_destroy("all", ask = FALSE)
+    
+    if (!keep_models){
+        if (length(path) > 1) {
+            for (i in 1:length(path)){
+                if (file.exists(path[i])) {
+                    file.remove(path[i])
+                }
+            }
+        }
+    }
+    message("Removed project outputs!")
+    
+}
+
 # evaluate if x is NOT an element of y
 "%!in%" <- function(x, y) !(x %in% y)
 

### 2022-11-21 _targets.R

@@ -4,14 +4,15 @@ library(tarchetypes)
 # load functions
 source("R/utils.R", encoding = "UTF-8")
 source("R/00_stimuli.R", encoding = "UTF-8")
-source("R/01_responses.R", encoding = "UTF-8")
-source("R/02_models.R", encoding = "UTF-8")
-source("R/03_questionnaire.R", encoding = "UTF-8")
-
+source("R/01_experiment.R", encoding = "UTF-8")
+source("R/02_questionnaire.R", encoding = "UTF-8")
+source("R/03_merged.R", encoding = "UTF-8")
+source("R/04_models.R", encoding = "UTF-8")
 
 # set number of cores to use with brms
 options(
-    mc.cores = 1
+    mc.cores = 4,
+    brms.backend = "cmdstanr"
 )
 
 # set parameters
@@ -26,8 +27,6 @@ tar_option_set(
         "ggplot2",
         "tibble", 
         "forcats",
-        "multilex",
-        "keyring",
         "readxl",
         "janitor", 
         "mice", 
@@ -40,18 +39,16 @@ tar_option_set(
         "tidybayes",
         "gt", 
         "patchwork", 
-        "wesanderson", 
         "papaja", 
         "knitr", 
         "data.table",
         "audio",
         "tidytext",
         "bayesplot",
-        "performance",
+        "parameters",
         "httr",
-        "ggrepel",
-        "yardstick"
-    ), 
+        "ggrepel"
+    )
 )
 
 conflicted::conflict_prefer("filter", "dplyr")
@@ -60,7 +57,7 @@ conflicted::conflict_prefer("last", "dplyr")
 conflicted::conflict_prefer("between", "dplyr")
 
 list(
-    # stimuli ----
+    # stimuli ------------------------------------------------------------------
     
     tar_target(stimuli_path, "stimuli/trials.xlsx", format = "file"),
     
@@ -70,7 +67,7 @@ list(
             "corona", # problematic, given COVID-19
             "moneda", # 2 correct responses: money, coin
             "lengua", # 2 correct responses: language, tongue
-            "porc"   # 2 correct responses: pork, pig
+            "porc"    # 2 correct responses: pork, pig
         )
     ),
     
@@ -96,158 +93,75 @@ list(
         )
     ),
     
-    # responses ----
-    tar_target(responses_path, list.files("data/raw", full.names = TRUE)),
+    # experiment responses -----------------------------------------------------
[…]

### 2022-11-21 _targets_packages.R

@@ -22,6 +22,7 @@ library(markdown)
 library(mice)
 library(multilex)
 library(papaja)
+library(parameters)
 library(patchwork)
 library(performance)
 library(pingr)
@@ -41,3 +42,4 @@ library(tidyr)
 library(tidytext)
 library(visNetwork)
 library(wesanderson)
+library(yardstick)

### 2024-04-02 docs/index.qmd
Make report in Quarto

---
title: "Translation Elicitation"
subtitle: "Lab notes"
format: html
---

```{r}
#| label: setup
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggdist)

tar_load(stimuli)
```

## Stimuli

### Lexical frequency

```{r}
stimuli |>
    ggplot(aes(group, freq_2, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    labs(x = "Average cross-linguistic phonological similarity",
         y = "Lexical frequency (log10)") +
    scale_y_continuous(transform = "log10")
```

### Cross-linguistic phonological similarity (cognateness)

```{r}
stimuli |>
    ggplot(aes(lv, group = NA)) +
    facet_wrap(~group) +
    geom_histogram() +
    labs(x = "Average cross-linguistic phonological similarity",
         y = "Lexical frequency (log10)")
```

### Cross-linguistic neighbourhood density

```{r}
#| label: fig-clnd
#| fig-cap: "Cross-language neighbourhood density"
stimuli |>
    mutate(total = n(), .by = group) |> 
    count(group, neigh_n, total) |> 
    mutate(n = n/total) |> 
    ggplot(aes(neigh_n, n)) +
    facet_wrap(~group) +
    geom_col(colour = "white") +
    labs(x = "Number of cross-language phonological neighbours",
         y = "Proportion of items")
```

### Cross-linguistic neighbourhood similarity

```{r}
#| label: fig-clnd
#| fig-cap: "Cross-language neighbourhood similarity"
stimuli |>
    ggplot(aes(group, neigh_sim, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    labs(x = "Group",
         y = "Average cross-linguistic phonological similarity")
```

### Audio duration

```{r}
#| label: fig-clnd
#| fig-cap: "Cross-language neighbourhood similarity"
stimuli |>
    ggplot(aes(group, duration, group = NA)) +
    geom_dots(aes(order = group),
              layout = "swarm", side = "both") +
    labs(x = "Audio duration (s)",
         y = "Average cross-linguistic phonological similarity")
```

### 2024-05-02 manuscript/manuscript.qmd
Make Quarto manuscript

--- 2022-10-28 manuscript.qmd
@@
 
-author:
-  - name: Gonzalo Garcia-Castro
-    affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
-  - name: Serene Siow
-    affiliation: Department of Experimental Psychology, University of Oxford
-  - name: Nuria Sebastian-Galles
-    affiliation: Center for Brain and Cognition, Universitat Pompeu Fabra
-  - name: Kim Plunkett
-    affiliation: Department of Experimental Psychology, University of Oxford
-
 abstract: |
-  Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar.
-Phonological similarity plays a more important role in low-proficiency bilinguals, compared to high-proficiency bilinguals, and points to the latter as relying more strongly on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
-Disentangling bilinguals' reliance on the conceptual and lexical routes during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals.
+    Understanding spoken words in a non-native language is easier when the words are phonologically similar to their translations into the listener's native language are phonologically similar.
+Phonological similarity plays a more important role in the language performance of low-proficiency bilinguals, compared to that of high-proficiency bilinguals., and This suggests that low-proficiency bilinguals may points to the latter as relying more strongly on cross-language word-to-word links (lexical route), while high-proficiency bilinguals rely more on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
+Disentangling bilinguals’ reliance on the conceptual and lexical and conceptual routes during translation is challenging, given that both sources of information are present during spoken word comprehensionto some extent in both high- and low-proficiency bilinguals.
 In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
-Given their lack of previous knowledge on the unfamiliar language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
+Given their lack of previous knowledge ion the unfamiliar tested language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
 This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route.
 Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation.
-This effect was larger in Spanish participants translating Catalan words than in English participants translating Catalan or Spanish.
 These results point to cross-language word-to-word links as providing rich information during word translation.
-
-keywords: cognate, word recognition, translation, non-native, spoken word recognition, bayesian
-
-bibliography:
-  - r-references.bib
-  - references.bib
-toc: false
-number-sections: true
-csl: "apa7.csl"
-format:
-  pdf:
-    message: false
-    warning: false
-    results: "asis" 
-    fig-dpi: 1000
 ---
@@
 
-set.seed(42)
+library(targets)
+library(dplyr)
+library(tidyr)
+library(ggplot2)
+library(gt)
+library(gtExtras)
+library(brms)
+library(tidybayes)
 
-options(
-    knitr.kable.NA = '-', 
-    knitr.duplicate.label = "allow",
-    gt.html_tag_check = FALSE
-)
+set.seed(1234)
 
-```
+# import objects
+tar_config_set(store = here::here("_targets"),
+                        script = here::here("_targets.R"))
+tar_load_globals()
+tar_load(participants)
+tar_load(exp_participants)
+tar_load(quest_participants)
+tar_load(stimuli)
 
-```{r}
-#| label: prepare
-#| message: false
-#| warning: false
-# import objects
-tar_load_globals()
-tar_load_all()
-
-# set custom ggplot2 theme
-theme_set(
-    theme(
-        title = element_text(size = 8, face = "bold"),
-        legend.background = element_rect(fill = NA, colour = NA),
-        legend.key = element_rect(fill = "white", colour = NA),
-        legend.justification = "right",
-        legend.direction = "horizontal",
-        axis.line = element_line(colour = "black", size = 0.5),
-        axis.ticks = element_line(colour = "black", size = 0.5),
-        axis.text.x = element_text(colour = "black"),
-        axis.text.y = element_text(colour = "black"),
-        plot.caption.position = "plot",
-        plot.title = element_text(hjust = 0.5),
-        panel.grid = element_blank(),
-        panel.background = element_rect(fill = "white", colour = NA)
-    )
-)
+theme_set(theme_ggdist())
 
@@
 Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
-One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, imagine a Spanish native, with no previous familiarity with any other language, who listens for the first time to French may encounter the word /`r knitr::asis_output("\U0070\U0254\U0281\U0074")` (*porte*), translation of *door* in French.
+However, there are some sources of information which may make non-native speech comprehension easier.
+One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener’s native language.
+For example, imagine a Spanish native, with no previous familiarity with any other language, who listens for the first time to French may encounter the word /`r knitr::asis_output("\U0070\U0254\U0281\U0074")`/ (*porte*), translation of door in French.
 The voiced uvular fricative consonant /`r knitr::asis_output("\U0281")`/ and the open-mid back rounded vowel /`r knitr::asis_output("\U0254")`/ do not exist in Spanish as phonemic categories.
@@
 Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /`r knitr::asis_output("\U0064\U0254\U02D0")`/ (*door*).
-This time, phonological similarity is of little help: /`r knitr::asis_output("\U0064\U0254\U02D0")`/ and /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ share very few phonemes.
+This time, phonological similarity is of little help: /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ and /`r knitr::asis_output("\U0064\U0254\U02D0")`/ share very few phonemes.
 It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations (which they lack, as they are unfamiliar to English).
@@
 Cross-language phonological similarity seems, at face value, an important source of information for unfamiliar listeners during auditory word recognition, even if translations are not phonologically identical.
-Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
-In the present study, we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
+Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners–even low-proficiency ones–are rarely completely naïve to the language they are listening to.
+In the present study, we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates.*
 
@@
 Cognates embody a particular case of such commonalities.
-Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level) commonly due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
+Cognates are defined as cross-language synonyms which have similar forms whose form is similar (at the phonological level, orthographic level, signed level), commonly due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
 For example, *sun* and *sol* (in Spanish) share phonological onset but their etymology points to different origins.
@@
 Such is the case of *cheese* and *queso* which share Latin root but are phonologically dissimilar.
-For simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
+For simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfill a specified threshold of form similarity, regardless of etymology.
 We do not expect that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
-For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (door in Spanish and Catalan, respectively).
 Cognates play a pivotal role in many models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, *table* in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
+For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, [table]), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
 This cognate advantage is not restricted to production, but also extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
 
-Early theoretical accounts of this 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
-The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), instead of directly to their shared concept (conceptual route).
-This word-to-word connectivity should be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection.
+## Lexical and conceptual routes for translation
+
+Early theoretical accounts of this 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals’ proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+The reasoning behind this claim was the assumption that second-language learners start acquiring words in their L2 by associating them to their translation in their L1 (lexical route), instead of directly to their shared concept (conceptual route).
+This word-to-word connectivity should be sensitive to form-similarity (i.e., cognateness): the more similar the two word- forms are, the stronger the connection.
 As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
@@
 The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 should take longer than translating from L2 to L1.
-The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, whereas L2-to-L1 translation relies more strongly on the mediation between the concept and the two word forms.
+The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, whereas L2-to-L1 translation relies more strongly on the mediation between the concept and the two word- forms.
 L1-to-L2 translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during L1-to-L2 translation.
@@
 To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate Dutch words to English (L1-to-L2) or English words to Dutch (L2-to-L1).
-Although participants' performance in both conditions was roughly equivalent for cognates, non-cognates were translated more slowly from Dutch to English than from English to Dutch, suggesting that L1-to-L2 translation is more sensitive to form-similarity than L2-to-L1 translation.
-Conversely, participants' performance was more sensitive to semantic variables (e.g., concreteness) when translating English words to Dutch, compared to when translating Dutch words to English, suggesting that L2-to-L1 translation was more sensitive to the shared conceptual properties of the translation pair.
+Although participants’ performance in both conditions was roughly equivalent for cognates, non-cognates were translated more slowly from Dutch to English than from English to Dutch, suggesting that L1-to-L2 translation is more sensitive to form-similarity than L2-to-L1 translation.
+Conversely, participants’ performance was more sensitive to semantic variables (e.g., concreteness) when translating English words to Dutch, compared to when translating Dutch words to English, suggesting that L2-to-L1 translation was more sensitive to the shared conceptual properties of the translation pair.
 
-These findings support a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations than L2-to-L1 translation, and therefore is more sensitive to cognateness.
-Further evidence for the presence of differences in reliance on the lexical and the conceptual routes between L1-to-L2 and L2-to-L1 translation was provided by @phillips2006erp, who used eletrophysiological recordings of bilinguals translating words from L1 to L2 or *vice versa* after semantic or phonological priming.
-The authors reported stronger N400 responses (associated with semantic priming) in L1-to-L2 translation than in L2-to-L2 translation, and stronger phonological mismatch negativity responses (associated with phonological priming) during L2-to-L1 translation, compared to L1-to-L2 translation.
+These findings support a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations than L2-to-L1 translation does, and therefore is more sensitive to cognateness.
+Further evidence for the presence of this difference between L1-to-L2 and L2-to-L1 translations on the degree of reliance on the lexical and or the conceptual routes between L1-to-L2 and L2-to-L1 translation was provided by @phillips2006erp, who used eletrophysiological recordings of bilinguals translating words from L1 to L2 or *vice versa* after semantic or phonological priming.
+The authors reported stronger N400 responses (associated with semantic priming) in L1-to-L2 translation than in L2-to-L12 translation, and stronger phonological mismatch negativity responses (associated with phonological priming) during L2-to-L1 translation, compared to L1-to-L2 translation.
 
-There is mixed evidence, however, about whether these differences are dependent on participants' proficiency in L2.
-While @degroot1994forward found an equivalent pattern of results in high- and low-proficiency bilinguals, other studies showed different performances in both groups during L2-to-L1 and L1-to-L2 translation [@christoffels2006memory; @christoffels2013language].
-These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backing a more complex interaction between the L1 and L2 lexica during word recognition.
+## Competition from lexical neighbours
 
-Since the RHM model was proposed, later studies on monolingual populations have focused on the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
+Since the RHM model was proposed, later studies on monolingual populations have focused how on the role of the network of connections that words establish networks of connections with each other at the form-level (phonological or orthographic) or at the conceptual level.
 For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
-Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when these neighbours' lexical frequency is higher [@luce1998recognizing].
-This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation (BIA/BIA+)  [@van1998orthographic; @dijkstra2002architecture] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with each other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
-Evidence supporting such an integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word's translation.
+Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when these neighbours’ lexical frequency is higher [@luce1998recognizing].
+
+The existence of phonological links between words is supported with experimental evidence from the visual world paradigm and phonological priming.
+@meyer2007activation showed that target pictures were named faster when presented together with a phonologically-related distractor, suggesting that the activation of shared phonological information allowed the target to be activated and produced faster.
+There are also findings suggesting that facilitatory or inhibitory effects on target recognition depend on the degree of similarity the competitor shares with the target [@hamburger1996phonological; @dufour2003lexical].
+When there is only mismatch of one phoneme between prime and target, inhibitory effects are exerted on target recognition [@dufour2003lexical].
+This inhibition was proposed to be due to strong lexical activation of the prime which results in lexical interference, as opposed to simple phonological activation from less similar primes that triggers facilitation.
+
+These findings of phonological links between words in the monolingual lexicon pose an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
+@weber2004lexical observed that bilinguals displayed lexical competition effects from words in their native language when performing a word-referent matching task in their L2.
+This highlights that lexical competition for bilinguals during language comprehension may not only come from the tested language but also the native language, despite the native language not being actively required for the task.
+Other studies have similarly found interference when spoken target words in one language had phonological overlap with a word in the participant’s other language, even when the other language was not required by the task [@spivey1999cross; @desroches2022dynamics].
+Such patterns have even been found with young children less than 4 years old [@von2012language].
+
+The Bilingual Interactive Activation (BIA/BIA+) [@van1998orthographic; @dijkstra2002architecture] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with each other, resulting in an integrated lexicon for both languages.
+This is, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
+Evidence supporting such an integrated lexicon is provided by the fact that bilinguals’ performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word’s translation.
 When bilingual participants are presented with words in one language in a lexical decision task, translating them to the other language took longer when they were part of large neighbourhoods [@van1998orthographic].
@@
 
-A more recent mmake(odel of bilingual lexical processing is Multilink [@dijkstra2019multilink], which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
-This model assumes an integrated lexicon, in which word representations from the two languages are activated by the input in a language non-selective way.
-It also proposes that translation equivalents are exclusively connected through their shared concept, to which they both send activation.
-The model does not implement excitatory or inhibitory connections between word representations, within or across languages.
-The cognate effect under this model arises from the fact that the input activates form-similar word in both L1 and L2 word-forms, so when such word-forms send their activation to a shared semantic node (i.e., cognates) such semantic node receives more excitation than when only receiving activation from the word representation in one language (i.e., non-cognates).
-This specification crucially differs from that of the RHM model, which suggests includes excitatory connections between the L1 and L2 word-forms at the lexical level, allowing translation to be achieved without the mediation of semantic nodes.
-In addition, Multilink suggests that the strength of the association between L2 representations and their concept depends on language users' proficiency in L2.
-A simulation series by @dijkstra2019multilink shows that data generated from Multilink are parallel to those observed in translation elicitation task and behavioural data from Dutch-English bilinguals [@christoffels2006memory].
-Overall, this  model provides support to the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
+## The present study
 
-The fact that Multilink successfully fitted the experimental data from @christoffels2006memory in the absence of word-to-word connections between translation equivalents is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which L1-to-L1 translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
-It should be taken into account, however, that  @christoffels2006memory only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
-The model specification under which the quantitative predictions were generated assumes similar vocabulary sizes across languages, and also that the word-to-concept connections of L1 and L2 representations have same strength.
-The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-to-word connections during translation, especially during backward translation.
-This prediction remains untested by Multilink.
+In the present study, we tested listeners’ reliance on phonological similarity when translating words from an unfamiliar language to their native language, in order to investigate the plausibility of word-to-word connections during L1-to-L2 translation.
+We tested English native speakers and Spanish native speakers in a translation elicitation task.
+This paradigm has been used by bilingualism researchers to identify easily-recognisable cognates for experimental studies.
+Using this paradigm, we aim to delve deeper into the factors affecting foreign word recognition by monolingual participants with no prior exposure to the language.
 
-One caveat concerning the generalisability of these conclusions is that most of the studies referenced so far rely heavily on production and comprehension of printed words.
-In these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
-There are limitations in generalising their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
-Consider the words *radio* (English) and *radio* (Spanish).
-Both words are clearly cognates, as their orthographical word form is identical.
-However, their phonological forms (English /`r knitr::asis_output("\U02C8\U0279\U0065\U026A\U002E\U0064\U0069\U0259\U028A")`/-- Spanish /`r knitr::asis_output("\U02C8\U0072\U0061\U002E\U00F0\U006A\U006F")`/) are not identical.
-Although many English listeners might recognise the Spanish word-form--even with no previous knowledge of Spanish--it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
-
-Measuring the phonological similarity between two word forms is also substantially more challenging than measuring their orthographic similarity.
-Many studies use symbolic representations to compute the similarity between word-forms of interest in either modality.
-Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (*radio*-*radio*); phonological word-forms are represented by their  phonological transcriptions (/`r knitr::asis_output("\U02C8\U0279\U0065\U026A\U002E\U0064\U0069\U0259\U028A")`/-/`r knitr::asis_output("\U02C8\U0072\U0061\U002E\U00F0\U006A\U006F")`/, in IPA notation) [see @floccia2018introduction for example].
-The similarity between each pair of word-forms is operationalised numerically by applying some algorithm that computes the edit distance between their symbolic representation.
-The Leveshtein distance is one of the most widely known algorithms: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string [@levenshtein1966binary], and has been successfully used to estimate the similarity of cross-language word-pairs [@gooskens2004perceptive; @schepens2012distributions].
-Using our previous examples, the Levenshtein distance between the orthographic word-forms *radio* (English) and *radio* (Spanish) is `r stringdist::stringdist("radio", "radio", useBytes = TRUE)`: both strings are identical.
-But the distance between the phonological transcriptions `r knitr::asis_output("\U0279\U0065\U026A\U0064\U0069\U0259\U028A")` and `r knitr::asis_output("\U0072\U0061\U00F0\U006A\U006F")` is 7, which is the number of edits that separates both character strings.
-There is also an additional layer of complexity: the phonological transcriptions `r knitr::asis_output("\U0279\U0065\U026A\U0064\U0069\U0259\U028A")` and `r knitr::asis_output("\U0072\U0061\U00F0\U006A\U006F")` can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.
-whether to include elements of aspiration, nasalisation, etc.) can affect how accurate these symbolic representation capture all of the features that are present in the acoustic signal when either word is spoken.
-Therefore, any measure of similarity that uses phonological transcriptions as input may overestimate or underestimate their actual similarity.
-Taking into account that measuring phonological similarity is considerably more complex than measuring orthographic complexity it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
-In summary, it is likely that previous investigations on bilinguals' reliance on form-similarity during translation might be misrepresenting the availability of such information in the auditory modality.
-
-In the present study we tested listeners' reliance on phonological similarity when translating words from an unfamiliar language, in order to investigate the plausibility of word-to-word connections during L1-to-L2 translation.
-We tested English and Spanish native speakers in a translation elicitation task.
-Participants
-were presented with spoken words in an unfamiliar (non-native language) and were asked to translate them to their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as **presented words**, and the correct translation for the presented words as **target translations**.
+Participants were presented with spoken words in an unfamiliar (non-native language) and were asked to translate them to their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
 The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and therefore word-to-concept connections are, in principle, absent.
@@
 We predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
+
+Following @luce1998recognizing's NAM model, we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
+If this is the case, non-native words that have more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency and are phonologically closer to the presented word than the target translation.
+For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as bottom instead of as its correct translation *button* (because the lexical frequency of bottom is higher than that of button and also phonological closer).
+This combined constraint led us to calculate a specific form of phonological neighbours where a neighbour is counted only if it is higher frequency and has stronger phonological similarity than the target translation as judged by Levenshtein distance.
+If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
 
@@
 
-# Methods
+# Experiment 1
+
+## Methods
 
@@
 
-## Participants
+### Participants
 
-Data collection took place from `r format(min(participants$date, na.rm = TRUE), "%B %dth, %Y")` to `r format(max(participants$date, na.rm = TRUE), "%B %dth, %Y")`.
-We collected data from `r length(unique(participants$participant_id))` participants (*Mean* = `r printnum(mean(participants$age, na.rm = TRUE))` years, *SD* = `r printnum(sd(participants$age, na.rm = TRUE))`, Range = `r printnum(min(participants$age, na.rm = TRUE))`-`r printnum(max(participants$age, na.rm = TRUE))`).
-`r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG")]))` participants were British English native speakers living in United Kingdom (`r length(unique(participants$participant_id[participants$group %in% c("cat-ENG", "spa-ENG") & participants$sex=="Female"]))` female), and `r length(unique(participants$participant_id[participants$group %in% c("cat-SPA")]))` participants were Spanish native speakers living in Spain (`r length(unique(participants$participant_id[participants$group %in% c("cat-SPA") & participants$sex=="Female"]))` female).
-Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
-Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I) and by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+```{r}
+#| label: participants-numbers-1
+participants_1 <- exp_participants |> 
+    filter(group!="cat-SPA")
+
+collection_dates <- participants_1 |> 
+    pull(date) |> 
+    range() |> 
+    format("%B %dth, %Y")
+
+n_participants <- n_distinct(participants_1)
+n_participants_group <- table(participants_1$group)
+participants_age <- participants_1 |> 
+    summarise(across(age, lst(mean, sd, min, max))) 
+n_female <- table(participants_1$gender)
+```
+
+Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
+We collected data from `r n_participants` British English native participants living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+
+```{r}
+#| label: participants
+tbl_data <- participants |> 
+    mutate(exp = case_when(
+        source=="Experiment" & group!= "cat-SPA" ~ "Experiment 1",
+        source=="Experiment" & group== "cat-SPA" ~ "Experiment 2",
+        source=="Questionnaire" ~ "Experiment 3",
+    )) |> 
+    summarise(n = n(),
+              n_excluded = sum(!valid_participant),
+              across(age, lst(mean, sd, min, max)),
+              knows_l2 = sum(l_2 != "None"),
+              l2_lst = lst(unique(l_2)[unique(l_2)!="None"]),
+              .by = c(group, exp))
+
+tbl_data |> 
+    gt(groupname_col = "exp", rowname_col = "group") |> 
+    fmt_number(matches("age")) |> 
+    fmt_integer(matches("min|max")) |> 
+    cols_merge_range(age_min, age_max) |> 
+    tab_spanner("Age", matches("age")) |> 
+    tab_spanner("Sample size", starts_with("n")) |> 
+    tab_spanner("L2", matches("l2")) |> 
+    cols_label(group = "Group",
+               exp = "Experiment",
+               n = "N",
+               n_excluded = "Excluded",
+               age_mean = "M",
+               age_sd = "SD",
+               age_min = "Range",
+               knows_l2 = "N knows L2?",
+               l2_lst = "Which L2?") |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels()) |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(align = "left"),
+              list(cells_body(), cells_column_labels()))
+```
+
+```{r}
+#| label: participants-excluded-1
+excluded_lang <- participants_1 |> 
+    filter(l_2_writ_prod > 3 |
+               l_2_oral_comp > 3 |
+               cat_oral_comp > 3 | 
+               cat_writ_prod > 3 |
+               spa_oral_comp > 3 |
+               spa_writ_prod > 3) |> 
+    nrow()
+
+excluded_impairment <- participants_1 |> 
+    filter(has_language_problems) |> 
+    nrow()
+
+excluded_ntrials <- participants_1 |> 
+    filter(invalid_participant_trials) |> 
+    nrow()
+```
+Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-We excluded data from participants that (a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (*n* = `r participants %>% filter(l2written > 4 | l2oral > 4) %>% nrow()`), (b) were diagnosed with a language (*n* = `r participants %>% filter(impairment) %>% nrow()`), or (c) did not contribute more than 80% of valid trials (*n* = `r participants %>% filter(invalid_participant_trials) %>% nrow()`).
+We excluded data from participants that (a) self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), (b) were diagnosed with a language (*n* = `r excluded_impairment`), or (c) did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
 
-## Procedure
+### Stimuli
+
+```{r}
+#| label: stimuli-lengths
+lengths <- stimuli |>  
+    select(group, ipa_flat_1, ipa_flat_2, word_1, word_2) |>  
+    mutate(across(ipa_flat_1:word_2, nchar)) |> 
+    summarise(across(c(ipa_flat_1, ipa_flat_2, word_1, word_2), 
+                     lst(mean, sd, min, max)),
+              .by = group)
+lengths <- split(lengths, lengths$group)
+names(lengths) <- janitor::make_clean_names(names(lengths))
+
+durations <- stimuli |> 
+    summarise(across(duration, lst(mean, sd, min, max)),
+              .by = group) 
+durations <- split(durations, durations$group)
+names(durations) <- janitor::make_clean_names(names(durations))
+
+```
+
+#### Presented stimuli
+
+We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
+Words in the Catalan list were `r round(lengths$cat_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_flat_1_sd, 2)`, *Range* = `r round(lengths$cat_eng$ipa_flat_1_min, 2)`-`r round(lengths$cat_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations (which participants had to type), were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, *Range* = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`.
+Words in the Spanish list were `r round(lengths$spa_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_flat_1_sd, 2)`, *Range* = `r round(lengths$spa_eng$ipa_flat_1_min, 2)`-`r round(lengths$spa_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations (which participants had to type), were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, *Range* = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`.
+
+Participants listened to one audio file in each trial, each containing a single word presented in isolation.
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
+The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, *Range* = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
+The average duration of the Catalan audio files was `r round(durations$spa_eng$duration_mean, 2)` (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, *Range* = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
+
+```{r}
+stim_exc_freq <- stimuli |> 
+    mutate(is_na = is.na(freq_zipf_2)) |> 
+    count(group, is_na) |> 
+    filter(is_na)
+stim_exc_freq <- split(stim_exc_freq, stim_exc_freq$group)
+names(stim_exc_freq) <- janitor::make_clean_names(names(stim_exc_freq))
+```
+
+### Procedure
 
@@
 Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="cat-ENG",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="spa-ENG",])` trials.
+
 Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
@@
 Trials in which participants took longer than 10 seconds to respond were also excluded.
-Participants contributed a total of `r nrow(responses)` valid trials (`r nrow(responses[responses$group %in% c("cat-ENG", "cat-SPA"),])` in Catalan, `r nrow(responses[responses$group %in% c("spa-ENG"),])` in Spanish).
+Participants contributed a total of `r nrow(exp_responses)` valid trials (`r nrow(exp_responses[exp_responses$group %in% c("cat-ENG", "cat-SPA"),])` in Catalan, `r nrow(exp_responses[exp_responses$group %in% c("spa-ENG"),])` in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 
-## Stimuli
+### Data analysis
 
 ```{r}
-#| label: stimuli-lengths
-#| message: false
-#| warning: false
-lengths <- stimuli %>% 
-    select(group, ipa_flat_1, ipa_flat_2, word_1, word_2) %>% 
-    mutate_at(vars(ipa_flat_1:word_2), nchar) %>% 
-    group_by(group) %>% 
-    summarise_at(
-        vars(ipa_flat_1, ipa_flat_2, word_1, word_2), 
-        list(mean = mean, sd = sd, min = min, max = max)
-    ) %>%
-    split(.$group) %>% 
-    map(select, -group) %>% 
-    set_names(make_clean_names(names(.))) %>% 
-    map(unlist)
+lengths <- dataset_1 |> 
+    mutate(len = nchar(response)) |> 
+    summarise(across(len,
+                     lst(mean, median, sd, min, max)),
+              .by = c(group)) 
+lengths <- split(lengths, lengths$group)
+
+names(lengths) <- janitor::make_clean_names(names(lengths))
 ```
 
-### Input stimuli (Catalan/Spanish)
-
-We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
-Words in the Catalan list were `r printnum(lengths$cat_eng["ipa_flat_1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$cat_eng["ipa_flat_1_sd"])`, Range = `r printnum(lengths$cat_eng["ipa_flat_1_min"], digits = 0)`-`r printnum(lengths$cat_eng["ipa_flat_1_max"], digits = 0)`).
-Words in the Spanish list were `r printnum(lengths$spa_eng["ipa_flat_1_mean"])` phonemes long on average (*SD* = `r printnum(lengths$spa_eng["ipa_flat_1_sd"])`, Range = `r printnum(lengths$spa_eng["ipa_flat_1_min"], digits = 0)`-`r printnum(lengths$spa_eng["ipa_flat_1_max"], digits = 0)`).
-Participants listened to one audio file in each trial, each containing a single word presented in isolation.
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the audio files was `r printnum(mean(stimuli$duration, na.rm = TRUE))` (*SD* = `r printnum(sd(stimuli$duration, na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration, na.rm = TRUE))`-`r printnum(max(stimuli$duration, na.rm = TRUE))`).
-The average duration of the Catalan audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("cat-ENG")], na.rm = TRUE))`), and the average duration of the Spanish audio files was `r printnum(mean(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))` seconds (*SD* = `r printnum(sd(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`, Range = `r printnum(min(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`-`r printnum(max(stimuli$duration[stimuli$group %in% c("spa-ENG")], na.rm = TRUE))`).
-
-### Participant responses (English/Spanish)
-
-Responses given by English participants for Catalan presented words were `r printnum(lengths$cat_eng["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_eng["word_2_sd"])`, Range = `r printnum(lengths$cat_eng["word_2_min"], digits = 0)`-`r printnum(lengths$cat_eng["word_2_max"], digits = 0)`), while their translations to Spanish responses given by Spanish participants for Catalan words were `r printnum(lengths$cat_spa["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$cat_spa["word_2_sd"])`, Range = `r printnum(lengths$cat_spa["word_2_min"], digits = 0)`-`r printnum(lengths$cat_spa["word_2_max"], digits = 0)`).
-Responses given by English participants for Spanish words were `r printnum(lengths$spa_eng["word_2_mean"])` characters long on average (*SD* = `r printnum(lengths$spa_eng["word_2_sd"])`, Range = `r printnum(lengths$spa_eng["word_2_min"], digits = 0)`-`r printnum(lengths$spa_eng["word_2_max"], digits = 0)`).
-
-#### Target translation lexical frequency
-
-We extracted the lexical frequencies for English target translations from SUBTLEX-UK[@van2014subtlex], and frequencies for Spanish from SUBTLEX-ESP [@cuetos2011subtlex].
-We extracted scores as Zipf scores where available or transformed raw scores to Zipf scores to  correct their logarithmic distribution, limiting their range roughly between 0 and 7, allowing an easier interpretation of further analyses [@van2014subtlex].
+Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, Range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, Range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
 
 ```{r}
-#| label: pthn-example
-#| message: false
-#| warning: false
-pthn_example <- clearpond_path %>% 
-    map(read.csv) %>% 
-    bind_rows(.id = "group") %>% 
-    clean_names() %>% 
-    filter(word %in% c("book", "sock"), group=="cat-ENG") %>% 
-    select(word, pthn)
+tbl_data <- list("Experiment 1" = dataset_1,
+                 "Experiment 2" = dataset_2,
+                 "Experiment 3" = dataset_3) |> 
+    bind_rows(.id = "exp") |> 
+    count(exp, group, response_type) |> 
+    pivot_wider(names_from = response_type,
+                values_from = n,
+                names_repair = janitor::make_clean_names) |>
+    relocate(exp, group, correct, typo, wrong, false_friend) |> 
+    mutate(n_correct = rowSums(cbind(correct, typo)),
+           n_incorrect = rowSums(cbind(wrong, false_friend)),
+           n_total = rowSums(cbind(n_correct, n_incorrect)),
+           across(correct:false_friend, 
+                  \(x) x / n_total,
+                  .names = "{col}_prop")) |> 
+    select(exp, group, correct, typo, wrong, false_friend,
+           matches("prop"))
+
+tbl_data |> 
+    gt(rowname_col = "group", groupname_col = "exp") |> 
+    fmt_missing(everything(), everything(), "--") |> 
+    tab_spanner("Correct responses", 
+                c(correct, typo)) |> 
+    tab_spanner("Incorrect responses", 
+                c(wrong, false_friend)) |> 
+    fmt_integer(c(correct, typo, wrong, false_friend)) |> 
+    fmt_percent(is.double) |> 
+    cols_merge_n_pct(correct, col_pct = correct_prop) |> 
+    cols_merge_n_pct(typo, col_pct = typo_prop) |> 
+    cols_merge_n_pct(wrong, col_pct = wrong_prop) |> 
+    cols_merge_n_pct(false_friend, col_pct = false_friend_prop) |>
+
+    cols_label(correct = "Correct",
+               typo = "Typo",
+               wrong = "Wrong translation",
+               false_friend = "False friend") |> 
+    tab_style(cell_text(weight = "bold"), 
+              cells_column_spanners()) |> 
+    tab_style(cell_text(style = "italic"), 
+              cells_column_labels())
 
@@
 
-#### Target translation phonological neighbourhood density (PTHN)
+#### Predictors
 
-We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) of the English and Spanish target translations from the CLEARPOND database [@marian2012clearpond].
-PTHN scores indicate the number of phonological neighbours with higher lexical frequency than the target translation, as indicated by its frequency score in the corresponding SUBTLEX database.
-CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differs from that of the target translation in only one addition, deletion, or substitution.
-PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
-
-```{r}
-#| label: stimuli-plot
-#| message: false
-#| warning: false
-#| echo: false
-#| out-width: "100%"
-#| fig-width: 7
-#| fig-height: 5.5
-#| fig-cap: "Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA).
-Columns indicate: A\\) Lexical frequency of target translations, expressed as Zipf scores.
-B\\) PTHN of target translations, and C\\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages.
-Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words."
-d <-  mutate(stimuli, group = factor(group, levels = c("cat-ENG", "spa-ENG", "cat-SPA")))
-
-medians <- d %>% 
-    group_by(group) %>% 
-    summarise(
-        pthn_median = median(pthn),
-        pthn_sd = sd(pthn),
-        frequency_zipf_median = median(frequency_zipf),
-        frequency_zipf_sd = sd(frequency_zipf),
-        lv_median = median(lv),
-        lv_sd = sd(lv)
-    )
-
-# frequency
-d %>% 
-    ggplot(aes(frequency_zipf, fill = group)) +
-    facet_wrap(~group, ncol = 1) + 
-    geom_rect(
-        data = medians, 
-        aes(
-            xmin = frequency_zipf_median-frequency_zipf_sd, xmax = frequency_zipf_median+frequency_zipf_sd,
-            ymin = -Inf, ymax = Inf, fill = group
-        ), 
-        inherit.aes = FALSE, alpha = 0.15
-    ) +   
-    geom_histogram(
-        bins = 20,
-        colour = "white"
-    ) +
-    geom_vline(
-        data = medians, 
-        aes(xintercept = frequency_zipf_median, group = group), 
-        colour = "black", size = 0.5, linetype = "dashed"
-    ) + 
-    geom_text(
-        data = medians, 
-        aes(label = "Median ± SD", x = frequency_zipf_median+1.25, y = 13), 
-        size = 3
-    ) +
-    geom_curve(
-        data = medians, 
-        aes(x = frequency_zipf_median+0.85, xend = frequency_zipf_median+0.1, y = 12, yend = 11),
-        arrow = arrow(length = unit(0.2, "cm")), curvature = -0.25, colour = "grey30"
-    ) +
-    labs(
-        x = "Lexical frequency (Zipf)", 
-        y = "Number of trials", 
-        title = "Frequency\n(stimuli translations)\n"
-    ) +
-
-    # PTHN
-    d %>% 
-    count(group, pthn) %>% 
-    ggplot(aes(pthn, n, fill = group)) + 
-    facet_wrap(~group, ncol = 1) + 
-    geom_rect(
-        data = medians, 
-        aes(xmin = 0, xmax = pthn_median+pthn_sd, ymin = -Inf, ymax = Inf, fill = group),
-        inherit.aes = FALSE, alpha = 0.15
-    ) +   
-    geom_col(colour = "white") + 
-    geom_vline(
-        data = medians, 
-        aes(xintercept = pthn_median, group = group), 
-        colour = "black", size = 0.5, linetype = "dashed"
-    ) + 
-    labs(
-        x = "PTHN", 
-        y = "Number of trials", 
-        title = "Higher-frequency neighbours\n(stimuli translations)\n"
-    ) +
-
-    # Levenshtein
-    d %>% 
-    ggplot(aes(lv, fill = group)) +
-    facet_wrap(~group, ncol = 1) + 
-    geom_rect(
-        data = medians, 
-        aes(xmin = 0, xmax = lv_median+lv_sd, ymin = -Inf, ymax = Inf, fill = group),
-        inherit.aes = FALSE, alpha = 0.15
-    ) +   
-    geom_histogram(bins = 20, colour = "white") +
-    geom_vline(
-        data = medians, 
-        aes(xintercept = lv_median, group = group), 
-        colour = "black", size = 0.5, linetype = "dashed"
-    ) + 
-    scale_x_continuous(labels = percent) +
-    labs(
-        x = "Levenshtein", 
-        y = "Number of trials", 
-        title = "Phonological similarity\n(stimuli and stimuli translations)\n"
-    ) +
-    plot_layout(guides = "collect") &
-    plot_annotation(tag_levels = "A") &
-    scale_fill_manual(values = c("#1A85FF", "#ff2976", "#FFC20A"), na.translate = FALSE) &
-    theme(
-        legend.position = "none",
-        title = element_text(size = 10),
-        axis.title = element_text(size = 10)
-    )
-```
-
-### Cross-language phonological similarity (Levenshtein)
-
-We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their IPA transcriptions using the `stringsim` function of the stringdist R package [@van2014stringdist].
+We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their X-SAMPA transcriptions using the `stringsim` function of the `stringdist` R package [@van2014stringdist].
 This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
@@
 The result is a percentage score, with 0% indicating no similarity between the two strings, and 100% indicating a perfect match between the two strings.
-We computed this similarity measure (refred simply as Levenshtein, from now on) for every translation pair in our stimuli lists.
+We computed this similarity measure (referred simply as Levenshtein, from here on) for every translation pair in our stimuli lists.
 For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
@@
 
-## Data analysis
+The lexical frequencies in this database were extracted from SUBTLEX-UK [@van2014subtlex].
+We transformed lexical frequency scores to Zipf scores [@van2014subtlex].
+Word in the stimuli list without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
+For each presented Catalan or Spanish word presented in each list, we calculated its number of higher-frequency cross-language phonological neighbours (CLPN).
+This measure was calculated by counting the number of higher-frequency English words whose phonological word-form (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differed by one phoneme from that of the presented Catalan or Spanish word.
+Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond].
+
+#### Modelling approach
 
@@
 
+#### Statistical inference
+
 To test and account for cross-group differences, we included a random intercept for each group.
@@
 
-# Results
+## Results
+
+```{r}
+#| label: tbl-dataset
+sem <- function(x) mean(x) / (sqrt(length(x)))
+
+tbl_data <- list("Experiment 1" = dataset_1,
+                 "Experiment 2" = dataset_2,
+                 "Experiment 3" = dataset_3) |> 
+    bind_rows(.id = "exp") |> 
+    add_count(exp, group, participant_id, name = "trials") |> 
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(group, exp, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              across(n_trials, lst(mean, sum, sd, min, max)),
+              n_participants = n_distinct(participant_id),
+              .by = c(group, exp)) |> 
+    relocate(n_participants, matches("n_trials"))
+
+tbl_data |> 
+    gt(rowname_col = "group",
+       groupname_col = "exp") |> 
+    fmt_number(is.numeric, decimals = 2) |> 
+    gtExtras::gt_plt_dist(correct_list, type = "density") |> 
+    fmt_integer(c(matches("sum|min|max"), n_participants)) |> 
+    fmt_number(matches("correct"), scale_by = 100) |> 
+    tab_spanner("Valid trials", matches("n_trials")) |> 
+    tab_spanner("Accuracy (%)", matches("correct")) |> 
+    cols_merge_range(n_trials_min, n_trials_max) |> 
+    cols_merge_range(correct_min, correct_max) |> 
+    cols_label(n_trials_sum = "N trials",
+               n_participants = "N",
+               n_trials_mean = "Mean",
+               n_trials_sd = "SD",
+               n_trials_min = "Range",
+               correct_mean = "Mean",
+               correct_sd = "SD",
+               correct_sem = "SE",
+               correct_min = "Range",
+               correct_list = "") |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels()) |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(align = "left"),
+              list(cells_body(), cells_column_labels()))
+```
 

### 2024-05-02 docs/index.qmd
Update report

@@
 library(ggplot2)
-library(ggdist)
+library(gt)
+library(gtExtras)
+library(patchwork)
+library(gghighlight)
+library(tidybayes)
+
+targets::tar_config_set(
+    store = here::here("_targets"),
+    script = here::here("_targets.R")
+)
 
 tar_load(stimuli)
+tar_load(exp_participants)
+tar_load(exp_responses)
+tar_load(quest_participants)
+tar_load(quest_responses)
+tar_load(dataset_1)
+tar_load(dataset_2)
+tar_load(dataset_3)
+tar_load(fit_1)
+
+theme_set(theme_ggdist() + 
+              theme(panel.grid.major.y = element_line(
+                  linetype = "dotted",
+                  colour = "grey")))
+
+# proportion adjusted from boundary values (Gelman, Hill & Vehtari, 2020)
+prop_adj <- function(x, n){
+    e <- (x+2)/(n+4)
+    return(e)
+}
+
+# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+prop_adj_se <- function(x, n) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    return(se)
+}
+
+# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+prop_adj_ci <- function(x, n, .width = 0.95) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    ci <-  e + qnorm(c((1-.width)/2, (1-(1-.width)/2)))*se
+    ci[1] <- ifelse(ci[1]<0, 0, ci[1]) # truncate at 0
+    ci[2] <- ifelse(ci[2]>1, 1, ci[2]) # truncate at 1
+    return(ci)
+}
 ```
 
-## Stimuli
+# Stimuli
 
-### Lexical frequency
+## Audio duration
 
 ```{r}
-stimuli |>
-    ggplot(aes(group, freq_2, group = NA)) +
-    geom_dots(aes(order = group),
-              layout = "swarm", side = "both") +
-    labs(x = "Average cross-linguistic phonological similarity",
-         y = "Lexical frequency (log10)") +
-    scale_y_continuous(transform = "log10")
-```
-
-### Cross-linguistic phonological similarity (cognateness)
-
-```{r}
-stimuli |>
-    ggplot(aes(lv, group = NA)) +
-    facet_wrap(~group) +
-    geom_histogram() +
-    labs(x = "Average cross-linguistic phonological similarity",
-         y = "Lexical frequency (log10)")
-```
-
-### Cross-linguistic neighbourhood density
-
-```{r}
-#| label: fig-clnd
-#| fig-cap: "Cross-language neighbourhood density"
-stimuli |>
-    mutate(total = n(), .by = group) |> 
-    count(group, neigh_n, total) |> 
-    mutate(n = n/total) |> 
-    ggplot(aes(neigh_n, n)) +
-    facet_wrap(~group) +
-    geom_col(colour = "white") +
-    labs(x = "Number of cross-language phonological neighbours",
-         y = "Proportion of items")
-```
-
-### Cross-linguistic neighbourhood similarity
-
-```{r}
-#| label: fig-clnd
-#| fig-cap: "Cross-language neighbourhood similarity"
-stimuli |>
-    ggplot(aes(group, neigh_sim, group = NA)) +
-    geom_dots(aes(order = group),
-              layout = "swarm", side = "both") +
-    labs(x = "Group",
-         y = "Average cross-linguistic phonological similarity")
-```
-
-### Audio duration
-
-```{r}
-#| label: fig-clnd
+#| label: fig-duration
 #| fig-cap: "Cross-language neighbourhood similarity"
@@
               layout = "swarm", side = "both") +
+    stat_summary(fun.data = mean_se,
+                 geom = "errorbar",
+                 colour = "red",
+                 linewidth = 1,
+                 width = 0.1) +
+    stat_summary(fun = mean,
+                 geom = "point",
+                 colour = "red",
+                 size = 2) +
     labs(x = "Audio duration (s)",
@@
 ```
+
+### Lexical frequency
+
+Lexical frequency in the TARGET language (English or Spanish).
+
+```{r}
+#| label: frequency-2
+missing <- stimuli[is.na(stimuli$freq_zipf_2), "ipa_flat_1"]
+```
+
+Missing observations: `r paste0(missing, collapse = ", ")`.
+
+```{r}
+stimuli |>
+    drop_na(freq_zipf_2) |> 
+    ggplot(aes(group, freq_zipf_2, group = NA)) +
+    geom_dots(aes(order = group),
+              layout = "swarm", side = "both",
+              alpha = 0.5) +
+    stat_summary(fun.data = mean_se, color = "red",
+                 geom = "errorbar",
+                 width = 0.5,
+                 linewidth = 1) +
+    labs(x = "Average cross-linguistic phonological similarity",
+         y = "Lexical frequency (Zipf)")
+```
+
+## Phonological similarity between translation pairs (cognateness)
+
+```{r}
+stimuli |> 
+    mutate(t_label = paste0(ipa_flat_1, "-", ipa_flat_2)) |> 
+    ggplot(aes(group, lv)) +
+    facet_wrap(~group, scales = "free_x") +
+    stat_interval(size = 8) +
+    # geom_text(data = stimuli |>
+    #               mutate(t_label = paste0(ipa_flat_1, "-", ipa_flat_2)) |>
+    #               top_n(10, lv),
+    #           aes(label = t_label),
+    #           position = position_nudge(x = 0.1),
+    #           hjust = 0)
+    # geom_point() +
+    ggrepel::geom_text_repel(aes(label = t_label),
+                             seed = 1234, size = 3,
+                             direction = "y",
+                             box.padding = unit(0.2, "lines"),
+                             max.overlaps = 13,
+                             alpha = 0.55,
+                             segment.size = 0,
+                             segment.curvature = 1/4) +
+    labs(x = "Group",
+         y = "Cognateness\n(Levenshtein similarity)",
+         colour = "Proportion of translations") +
+    scale_y_continuous(labels = scales::percent) +
+    scale_color_brewer(palette = "Reds") +
+    theme(legend.position = "top",
+          axis.text.x = element_blank())
+```
+
+## Cross-linguistic neighbourhood density
+
+```{r}
+#| label: tbl-clnd
+#| fig-cap: "Cross-language neighbourhood density"
+stimuli |> 
+    arrange(group, desc(neigh_n)) |> 
+    group_by(group) |> 
+    top_n(n = 10, neigh_n) |> 
+    select(group, translation, avg_sim, neigh_n, neigh_lst) |> 
+    gt(rowname_col = "translation",
+       groupname_col = "group") |> 
+    fmt_percent(avg_sim) |> 
+    tab_style(cell_text(align = "left"), cells_body()) |> 
+    cols_label(neigh_n = "N neighbours",
+               avg_sim = "Average cross-language similarity",
+               neigh_lst = "List of neighbours")
+```
+
+## Cross-linguistic neighbourhood density (higher frequency than translation)
+
+```{r}
+#| label: tbl-clnd
+#| fig-cap: "Cross-language neighbourhood density (higher frequency than translation)"
+stimuli |> 
+    arrange(group, desc(neigh_n_h)) |> 
+    group_by(group) |> 
+    top_n(n = 10, neigh_n_h) |> 
+    select(group, translation, avg_sim_h, neigh_n_h, neigh_lst_h) |> 
+    gt(rowname_col = "translation",
+       groupname_col = "group") |> 
+    fmt_percent(avg_sim_h) |> 
+    tab_style(cell_text(align = "left"), cells_body()) |> 
+    cols_label(neigh_n_h = "N neighbours",
+               avg_sim_h = "Average cross-language similarity",
+               neigh_lst_h = "List of neighbours")
+```
+
+## Cross-linguistic neighbourhood similarity
+
+```{r}
+#| label: fig-clns
+#| fig-cap: "Cross-language neighbourhood similarity"
+stimuli |>
+    ggplot(aes(group, avg_sim, group = NA)) +
+    geom_dots(aes(order = group),
+              layout = "swarm", side = "both") +
+    stat_summary(fun.data = mean_se,
+                 geom = "errorbar",
+                 colour = "red",
+                 linewidth = 3/4,
+                 width = 0.1) +
+    stat_summary(fun = mean,
+                 geom = "point",
+                 colour = "red",
+                 size = 2) +
+    labs(x = "Group",
+         y = "Average cross-linguistic phonological similarity")
+```
+
+## Cognateness by average cross-linguistic similarity
+
+```{r}
+#| label: fig-cognateness-avg-sim
+stimuli |> 
+    ggplot(aes(lv, avg_sim)) +
+    geom_rug(colour = "grey") +
+    geom_point(alpha = 0.5,
+               shape = 1,
+               size = 2,
+               stroke = 1) +
+    geom_smooth(method = "lm",
+                colour = "red",
+                fill = "red") + 
+    labs(x = "Cognateness (Levenshtein distance)",
+         y = "Average cross-language similarity")
+```
+
+# Participants
+
+```{r}
+#| label: participants
+tbl_data <- list(Experiment = exp_participants,
+                 Questionnaire = quest_participants) |> 
+    bind_rows(.id = "source") |> 
+    summarise(n = n(),
+              n_excluded = sum(!valid_participant),
+              across(age, lst(mean, sd, min, max)),
+              knows_l2 = sum(l_2 != "None"),
+              l2_lst = lst(unique(l_2)[unique(l_2)!="None"]),
+              .by = c(group, source))
+
+tbl_data |> 
+    gt(groupname_col = "source", rowname_col = "group") |> 
+    fmt_number(matches("age")) |> 
+    fmt_integer(matches("min|max")) |> 
+    cols_merge_range(age_min, age_max) |> 
+    tab_spanner("Age", matches("age")) |> 
+    tab_spanner("Sample size", starts_with("n")) |> 
+    tab_spanner("L2", matches("l2")) |> 
+    cols_label(group = "Group",
+               source = "Source",
+               n = "N",
+               n_excluded = "Excluded",
+               age_mean = "M",
+               age_sd = "SD",
+               age_min = "Range",
+               knows_l2 = "N knows L2?",
+               l2_lst = "Which L2?") |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels()) |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(align = "left"),
+              list(cells_body(), cells_column_labels()))
+```
+
+## Target language proficiency
+
+```{r}
+tbl_data <- list(Experiment = exp_participants,
+                 Questionnaire = quest_participants) |> 
+    bind_rows(.id = "source") |> 
+    summarise(across(matches("writ|oral"),
+                     lst(mean, sd, min, max)),
+              .by = c(group, source)) |> 
+    select(-matches("l_2"))
+
+tbl_data |> 
+    gt(rowname_col = "group", groupname_col = "source") |> 
+    fmt_number(is.numeric) |> 
+    fmt_integer(matches("min|max")) |> 
+    fmt_missing(rows = everything(), columns = everything(),
+                missing_text = "--") |> 
+    cols_merge_range(cat_oral_comp_min, cat_oral_comp_max) |> 
+    cols_merge_range(spa_oral_comp_min, spa_oral_comp_max) |> 
+    cols_merge_range(cat_writ_prod_min, cat_writ_prod_max) |> 
+    cols_merge_range(spa_writ_prod_min, spa_writ_prod_max) |> 
+    tab_spanner("Catalan", matches("cat_")) |> 
+    tab_spanner("Spanish", matches(("spa_"))) |> 
+    tab_spanner("Oral comprehension", matches("oral")) |> 
+    tab_spanner("Written production", matches("writ")) |> 
+    cols_label(cat_oral_comp_mean = "M",
+               spa_oral_comp_mean = "M",
+               cat_writ_prod_mean = "M",
+               spa_writ_prod_mean = "M",
+               cat_oral_comp_sd = "SD",
+               spa_oral_comp_sd = "SD",
+               cat_writ_prod_sd = "SD",
+               spa_writ_prod_sd = "SD",
+               cat_oral_comp_min = "Range",
+               spa_oral_comp_min = "Range",
+               cat_writ_prod_min = "Range",
+               spa_writ_prod_min = "Range") |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels()) |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(align = "left"),
+              list(cells_body(), cells_column_labels()))
+```
+
+# Results
+
+```{r}
+#| label: tbl-dataset
+sem <- function(x) mean(x) / (sqrt(length(x)))
+
+dataset_2 |> 
+    add_count(source, group, participant_id, name = "trials") |> 
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(group, source, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              across(n_trials, lst(mean, sum, sd, min, max)),
+              n_participants = n_distinct(participant_id),
+              .by = c(group, source)) |> 
+    relocate(n_participants, matches("n_trials")) |> 
+    gt(rowname_col = "group",
+       groupname_col = "source") |> 
+    fmt_number(is.numeric, decimals = 2) |> 
+    gtExtras::gt_plt_dist(correct_list, type = "density") |> 
+    fmt_integer(c(matches("sum|min|max"), n_participants)) |> 
+    fmt_number(matches("correct"), scale_by = 100) |> 
+    tab_spanner("Valid trials", matches("n_trials")) |> 
+    tab_spanner("Accuracy (%)", matches("correct")) |> 
+    cols_merge_range(n_trials_min, n_trials_max) |> 
+    cols_merge_range(correct_min, correct_max) |> 
+    cols_label(n_trials_sum = "N trials",
+               n_participants = "N",
+               n_trials_mean = "Mean",
+               n_trials_sd = "SD",
+               n_trials_min = "Range",
+               correct_mean = "Mean",
+               correct_sd = "SD",
+               correct_sem = "SE",
+               correct_min = "Range",
+               correct_list = "") |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels()) |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(align = "left"),
+              list(cells_body(), cells_column_labels()))
+```
+
+## Questionnaire results
+
+```{r}
+quest_responses |> 
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(group, confidence, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              .by = c(group, confidence)) |> 
+    ggplot(aes(confidence, correct_mean,
+               colour = group,
+               ymin = correct_mean - correct_sem,
+               ymax = correct_mean + correct_sem)) +
+    geom_errorbar(width = 0.5,
+                  linewidth = 3/4,
+                  position = position_dodge(width = 0.5)) +
+    geom_point(size = 2,
+               position = position_dodge(width = 0.5)) +
+    scale_y_continuous(limits = c(0, 1)) +
+    labs(x = "Reported confidence (1-5)",
+         y = "Accuracy",
+         colour = "Group") +
+
+    quest_responses |> 
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(group, knowledge, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              .by = c(group, knowledge)) |> 
+    mutate(knowledge = if_else(knowledge, "Yes", "No")) |> 
+    ggplot(aes(knowledge, correct_mean,
+               colour = group,
+               ymin = correct_mean - correct_sem,
+               ymax = correct_mean + correct_sem)) +
+    geom_errorbar(width = 0.25,
+                  linewidth = 3/4,
+                  position = position_dodge(width = 0.5)) +
+    geom_point(size = 2,
+               position = position_dodge(width = 0.5)) +
+    scale_y_continuous(limits = c(0, 1)) +
+    labs(x = "Reported word knowledge",
+         y = "Accuracy",
+         colour = "Group") +
+    theme(axis.text.y = element_blank(),
+          axis.title.y = element_blank()) +
+
+    quest_responses |> 
+    mutate(knowledge = if_else(knowledge, "Yes", "No")) |> 
+    summarise(across(confidence, lst(mean, sem)),
+              .by = c(group, knowledge)) |> 
+    ggplot(aes(knowledge, confidence_mean,
+               colour = group,
+               ymin = confidence_mean - confidence_sem,
+               ymax = confidence_mean + confidence_sem)) +
+    geom_errorbar(width = 0.25,
+                  linewidth = 3/4,
+                  position = position_dodge(width = 0.5)) +
+    geom_point(size = 2,
+               position = position_dodge(width = 0.5)) +
+    labs(x = "Reported word knowledge",
+         y = "Resported confidence (1-5)",
+         colour = "Group") +
+    scale_y_continuous(labels = scales::percent) +
+    plot_layout(nrow = 1,
+                guides = "collect") &
+    theme(legend.position = "top")
+```
+
+```{r}
+#| label: fig-results-lv
+#| fig-cap: "Accuracy by cognateness."
+dataset_2 |> 
+    mutate(lv_std = cut_interval(lv_std, 10, labels = FALSE),
+           group = factor(group, levels = c("cat-ENG",
+                                            "spa-ENG",
+                                            "cat-SPA"))) |>
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(group, lv_std, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              .by = c(group, lv_std)) |> 
+    ggplot(aes(lv_std, correct_mean, colour = group,
+               ymin = correct_mean-correct_sem,
+               ymax = correct_mean+correct_sem)) +
+    facet_wrap(~group) +
+    geom_errorbar(width = 0.25,
+                  linewidth = 3/4,
+                  position = position_dodge(width = 0.25)) +
+    geom_point(size = 2,
+                  position = position_dodge(width = 0.25)) +
+    labs(x = "SD Cognateness (Levenshtein distance)",
+         y = "Accuracy",
+         colour = "Group",
+         fill = "Group") +
+    scale_y_continuous(labels = scales::percent) +
+    theme(legend.position = "none")
+```
+
+```{r}
+#| label: fig-results-avg_sim
+#| fig-cap: "Accuracy by average cross-language similarity"
+dataset_2 |> 
+    mutate(avg_sim_std = cut_number(avg_sim_std, 10, labels = FALSE)) |>
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(group, avg_sim_std, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              .by = c(group, avg_sim_std)) |> 
+    ggplot(aes(avg_sim_std, correct_mean, colour = group,
+               ymin = correct_mean-correct_sem,
+               ymax = correct_mean+correct_sem)) +
+    facet_wrap(~group) +
+    geom_errorbar(width = 0.25,
+                  linewidth = 3/4) +
+    geom_point(size = 2) +
+    labs(x = "SD Average cross-language similarity",
+         y = "Accuracy",
+         colour = "Group",
+         fill = "Group") +
+    scale_y_continuous(labels = scales::percent) +
+    theme(legend.position = "none")
+```
+
+```{r}
+#| label: fig-results-avg_sim
+#| fig-cap: "Accuracy by average cross-language similarity"
+dataset_2 |> 
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(group, neigh_n, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              .by = c(group, neigh_n)) |> 
+    ggplot(aes(neigh_n, correct_mean, colour = group,
+               ymin = correct_mean-correct_sem,
+               ymax = correct_mean+correct_sem)) +
+    geom_point(size = 2) +
+    geom_smooth(aes(group = group),
+                method = "glm", 
+                method.args = list(family = "binomial"), 
+                se = FALSE) +
+    labs(x = "Cognateness (Levenshtein distance)",
+         y = "Accuracy",
+         colour = "Group",
+         fill = "Group")
+```
+## Model outputs
+
+### Analysis 1
+
+```{r}
+nd <- expand_grid(lv_std = seq(-3, 3, 0.1),
+                  avg_sim_h_std = 0,
+                  freq_zipf_2_std = 0)
+
+epreds <- add_epred_draws(nd, fit_0, re_formula = NA) |> 
+    # mean_qi() |> 
+    mutate(avg_sim_h_std = as.factor(avg_sim_h_std))
+
+dataset_1 |> 
+    summarise(n_trials = n(),
+              correct = sum(correct),
+              .by = c(lv_std, participant_id)) |> 
+    mutate(correct = prop_adj(correct, n_trials)) |> 
+    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+              .by = c(lv_std)) |> 
+    ggplot(aes(lv_std, .epred)) +
+    stat_lineribbon(data = epreds) +
+    geom_point(aes(y = correct_mean)) +
+    scale_fill_brewer(palette = "Reds")
+```
+
+# Appendix
+
+## Appendix 1: List of trials
+
+### Catalan-English (cat-ENG)
+
+```{r}
+#| label: tbl-app-stim-cateng
+#| tbl-cap: "Catalan-English stimuli"
+tar_load(trial_list)
+
+trial_list_lst <- trial_list |> 
+    split(trial_list$group) |>
+    janitor::clean_names() |> 
+    map(select, matches("word_|ipa|sampa"))
+
+trial_list_lst$cat_eng |> 
+    gt() |> 
+    tab_spanner("Presented word (1)", ends_with("_1")) |> 
+    tab_spanner("Presented word (2)", ends_with("_2")) |> 
+    cols_label(word_1 = "Orthography",
+               ipa_flat_1 = "IPA",
+               sampa_1 = "X-SAMPA",
+               word_2 = "Orthography",
+               ipa_flat_2 = "IPA",
+               sampa_2 = "X-SAMPA")
+```
+
+### Spanish-English (spa-ENG)
+
+```{r}
+#| label: tbl-app-stim-cateng
+#| tbl-cap: "Spanish-English stimuli"
+trial_list_lst$spa_eng |> 
+    gt() |> 
+    tab_spanner("Presented word (1)", ends_with("_1")) |> 
+    tab_spanner("Presented word (2)", ends_with("_2")) |> 
+    cols_label(word_1 = "Orthography",
+               ipa_flat_1 = "IPA",
+               sampa_1 = "X-SAMPA",
+               word_2 = "Orthography",
+               ipa_flat_2 = "IPA",
+               sampa_2 = "X-SAMPA")
+```
+
+### Catalan-Spanish (cat-SPA)
+
+```{r}
+#| label: tbl-app-stim-catspa
+#| tbl-cap: "Catalan-Spanish stimuli"
+trial_list_lst$cat_spa |> 
+    gt() |> 
+    tab_spanner("Presented word (1)", ends_with("_1")) |> 
+    tab_spanner("Presented word (2)", ends_with("_2")) |> 
+    cols_label(word_1 = "Orthography",
+               ipa_flat_1 = "IPA",
+               sampa_1 = "X-SAMPA",
+               word_2 = "Orthography",
+               ipa_flat_2 = "IPA",
+               sampa_2 = "X-SAMPA")
+```

### 2024-05-03 manuscript/manuscript.qmd
First results

@@
 library(ggplot2)
+library(patchwork)
 library(gt)
@@
 tar_load(stimuli)
+tar_load(epreds)
 
-theme_set(theme_ggdist())
+theme_set(theme_ggdist() + theme(panel.grid.major.y = element_line(colour = "grey",
+                                                                   linetype = "dotted")))
 
@@
 Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
-However, there are some sources of information which may make non-native speech comprehension easier.
+However, tfile.path are some sources of information which may make non-native speech comprehension easier.
 One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener’s native language.
@@
 The voiced uvular fricative consonant /`r knitr::asis_output("\U0281")`/ and the open-mid back rounded vowel /`r knitr::asis_output("\U0254")`/ do not exist in Spanish as phonemic categories.
-These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /r/ and /o/ [@best2001discrimination], and the non-native speech signal may engage lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ (*puerta*).
+These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /r/ and /o/ [@best2001discrimination], and the non-native speech signal may engage lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and tfile.pathfore the recognition of its translation to Spanish /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ (*puerta*).
 The recognition of this French, unfamiliar word relies almost entirely on its similarity with its Spanish translation.
@@
 The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 should take longer than translating from L2 to L1.
-The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, whereas L2-to-L1 translation relies more strongly on the mediation between the concept and the two word- forms.
-L1-to-L2 translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during L1-to-L2 translation.
+The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, wfile.pathas L2-to-L1 translation relies more strongly on the mediation between the concept and the two word- forms.
+L1-to-L2 translation should tfile.pathfore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during L1-to-L2 translation.
 
@@
 
-These findings support a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations than L2-to-L1 translation does, and therefore is more sensitive to cognateness.
+These findings support a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but wfile.path L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations than L2-to-L1 translation does, and tfile.pathfore is more sensitive to cognateness.
 Further evidence for the presence of this difference between L1-to-L2 and L2-to-L1 translations on the degree of reliance on the lexical and or the conceptual routes between L1-to-L2 and L2-to-L1 translation was provided by @phillips2006erp, who used eletrophysiological recordings of bilinguals translating words from L1 to L2 or *vice versa* after semantic or phonological priming.
@@
 @meyer2007activation showed that target pictures were named faster when presented together with a phonologically-related distractor, suggesting that the activation of shared phonological information allowed the target to be activated and produced faster.
-There are also findings suggesting that facilitatory or inhibitory effects on target recognition depend on the degree of similarity the competitor shares with the target [@hamburger1996phonological; @dufour2003lexical].
-When there is only mismatch of one phoneme between prime and target, inhibitory effects are exerted on target recognition [@dufour2003lexical].
+Tfile.path are also findings suggesting that facilitatory or inhibitory effects on target recognition depend on the degree of similarity the competitor shares with the target [@hamburger1996phonological; @dufour2003lexical].
+When tfile.path is only mismatch of one phoneme between prime and target, inhibitory effects are exerted on target recognition [@dufour2003lexical].
 This inhibition was proposed to be due to strong lexical activation of the prime which results in lexical interference, as opposed to simple phonological activation from less similar primes that triggers facilitation.
@@
 We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
-The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and therefore word-to-concept connections are, in principle, absent.
+The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and tfile.pathfore word-to-concept connections are, in principle, absent.
 As a consequence, these participants should only be able use word-to-word connections to succeed in the translation of words in an unfamiliar language.
@@
 For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as bottom instead of as its correct translation *button* (because the lexical frequency of bottom is higher than that of button and also phonological closer).
-This combined constraint led us to calculate a specific form of phonological neighbours where a neighbour is counted only if it is higher frequency and has stronger phonological similarity than the target translation as judged by Levenshtein distance.
+This combined constraint led us to calculate a specific form of phonological neighbours wfile.path a neighbour is counted only if it is higher frequency and has stronger phonological similarity than the target translation as judged by Levenshtein distance.
 If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
@@
 We tested three groups of participants in this translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
-Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
+Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, wfile.pathas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, tfile.pathfore allowing us to explore potential cross-language differences in participant's performance.
 
@@
     summarise(n = n(),
-              n_excluded = sum(!valid_participant),
+              n_excluded = sum(!valid_status=="Valid"),
               across(age, lst(mean, sd, min, max)),
-              knows_l2 = sum(l_2 != "None"),
-              l2_lst = lst(unique(l_2)[unique(l_2)!="None"]),
+              knows_l2 = sum(l2 != "None"),
+              l2_lst = lst(unique(l2)[unique(l2)!="None"]),
               .by = c(group, exp))
@@
 excluded_lang <- participants_1 |> 
-    filter(l_2_writ_prod > 3 |
-               l_2_oral_comp > 3 |
+    filter(l2_writ_prod > 3 |
+               l2_oral_comp > 3 |
                cat_oral_comp > 3 | 
@@
 excluded_impairment <- participants_1 |> 
-    filter(has_language_problems) |> 
+    filter(valid_status=="Language impairment") |> 
     nrow()
@@
 excluded_ntrials <- participants_1 |> 
-    filter(invalid_participant_trials) |> 
+    filter(valid_status=="Insufficient trials") |> 
     nrow()
@@
 Then, participants pressed the RETURN/ENTER key to confirm their answer and start and new trial.
-We excluded trials where the participant's answer was not a valid word in the target language or no answer was given.
+We excluded trials wfile.path the participant's answer was not a valid word in the target language or no answer was given.
 Trials in which the response was mistyped by only one character were counted as correct (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct word.
@@
 #| fig-cap: "Schematic representation of a trial in the experimental task."
-include_graphics(here("img", "design.png"))
+knitr::include_graphics(here::here("img", "design.png"))
 ```
@@
 The result is a percentage score, with 0% indicating no similarity between the two strings, and 100% indicating a perfect match between the two strings.
-We computed this similarity measure (referred simply as Levenshtein, from here on) for every translation pair in our stimuli lists.
+We computed this similarity measure (referred simply as Levenshtein, from file.path on) for every translation pair in our stimuli lists.
 For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
@@
 We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) [@schad2020capitalize].
-Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) [@barr2013random].
+Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, wfile.path our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) [@barr2013random].
 We will later refer to these models as Model 0 to Model 6.
@@
 We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
-We used the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+We used the brms R package [@burkner2017brms] using the `cmdstanr` backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
 
-## Results
+### Results
 
@@
 
-All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
-Model 6, which included all main effects and the two-way interactions between $PTHN$ and $Levenshtein$, and between $Group$ and $Levenshtein$ and $PTHN$ and consonant similarity, showed the best performance.
-
-```{r}
-#| label: model-comparison
-#| message: false
-#| warning: false
-tab_loos <- model_loos %>% 
-    as.data.frame() %>% 
-    rownames_to_column("model") %>% 
-    arrange(desc(elpd_loo)) %>% 
-    select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
-    mutate(model = str_replace(model, "fit_", "Model ")) %>% 
-    mutate_all(~ifelse(. == 0, NA, .)) %>% 
-    gt(rowname_col = "model", id = "model-comparison") %>% 
-    tab_header(
-        title = md("Table 2"),
-        subtitle = md("Model performance and comparison using leave-one-out cross-validation")
-    ) %>% 
-    tab_source_note(
-        md("*Note*.
-Outcomes of the leave-one-out cross-validation procedure.
-Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
-Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
-    ) %>% 
-    fmt_missing(everything(), missing_text = "-") %>% 
-    fmt_number(2:7) %>% 
-    cols_label(
-        model = md("**Model**"),
-        looic = md("***LOO<sub>IC</sub>***"),
-        se_looic = md("***SE*<sub>IC</sub>**"),
-        elpd_loo = md("***LOO<sub>ELPD</sub>***"),
-        se_elpd_loo = md("***SE***"),
-        elpd_diff = md("***LOO<sub>diff</sub>***"),
-        se_diff = md("***SE<sub>diff</sub>***")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", weight = "bold", size = "medium"),
-        locations = cells_title(groups = "title")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", size = "medium", style = "italic"),
-        locations = cells_title(groups = "subtitle")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", size = "medium"),
-        locations = cells_source_notes()
-    ) %>% 
-    tab_style(
-        style = list(
-            cell_borders(sides = c("left", "right"), color = "white"),
-            cell_text(align = "center")
-        ),
-        locations = list(
-            cells_body(), 
-            cells_row_groups(), 
-            cells_column_spanners(), 
-            cells_stub(), 
-            cells_source_notes(),
-            cells_column_labels(), 
-            cells_stubhead()
-        )
-    ) %>% 
-    tab_style(
-        style = list(
-            cell_text(align = "left", size = "medium"), 
-            cell_borders(sides = "all", color = "white")
-        ),
-        locations = list(cells_stub(), cells_source_notes())
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-        locations = list(cells_column_spanners(), cells_column_labels())
-    ) %>% 
-    tab_style(
-        style = cell_borders(sides = "all", color = "white"),
-        locations = cells_body(columns = 1:7)
-    ) %>% 
-    tab_style(
-        style = cell_borders(sides = "top", color = "white"),
-        locations = cells_body(columns = 1:7)
-    ) %>% 
-    gtsave(here("img", "loos.png"))
-
-include_graphics(here("img", "loos.png"))
-
+```{r exp_1_epreds}
+#| label: fig-epreds-1
+#| fig-height: 3.5
+#| fig-width: 10
+get_epreds(exp_1_m0, dataset_1) |> 
+    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+           neigh_n_h = factor(neigh_n_h,
+                              levels = paste0(c(0, 2, 4, 8, 16),
+                                              " neighbours"),
+                              ordered = TRUE)) |> 
+    ggplot(aes(lv, .epred)) +
+    facet_wrap(~neigh_n_h, nrow = 1) +
+    stat_lineribbon(aes(fill_ramp = after_stat(level)),
+                    linewidth = 1/2,
+                    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
+    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+         y = "p(Correct)",
+         colour = "Cross-language neighbourhood density",
+         fill = "Cross-language neighbourhood density",
+         linetype = "Cross-language neighbourhood density",
+         fill_ramp = "CrI (%)") +
+    scale_x_continuous(labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent) +
+    scale_fill_brewer(palette = "Reds") +
+    theme(legend.position = "none",
+          legend.box = "vertical") 
 ```
@@
 
-# Discussion
+### Discussion
+
+## Experiment 2
+
+### Methods
+
+### Results
+
+```{r exp_2_epreds}
+#| label: fig-epreds-2
+#| fig-height: 3.5
+#| fig-width: 10
+get_epreds(exp_2_m0, dataset_2) |> 
+    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+           neigh_n_h = factor(neigh_n_h,
+                              levels = paste0(c(0 , 1, 2, 4, 8, 16),
+                                              " neighbours"),
+                              ordered = TRUE)) |>
+    ggplot(aes(lv, .epred)) +
+    facet_wrap(~neigh_n_h, nrow = 1) +
+    stat_lineribbon(aes(fill_ramp = after_stat(level)),
+                    linewidth = 1/2,
+                    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
+    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+         y = "P(Correct)",
+         colour = "Cross-language neighbourhood density",
+         fill = "Cross-language neighbourhood density",
+         linetype = "Cross-language neighbourhood density",
+         fill_ramp = "CrI (%)") +
+    scale_x_continuous(labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent) +
+    scale_fill_brewer(palette = "Reds") +
+    theme(legend.position = "none",
+          legend.box = "vertical") 
+```
+
+### Discussion
+
+## Experiment 3
+
+```{r exp_3_epreds}
+#| label: fig-epreds-3
+#| fig-height: 3.5
+#| fig-width: 10
+get_epreds(exp_3_m0, dataset_3) |> 
+    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+           neigh_n_h = factor(neigh_n_h,
+                              levels = paste0(c(0 , 1, 2, 4, 8, 16),
+                                              " neighbours"),
+                              ordered = TRUE)) |> 
+    ggplot(aes(lv, .epred)) +
+    facet_wrap(~neigh_n_h, nrow = 1) +
+    stat_lineribbon(aes(fill_ramp = after_stat(level)),
+                    linewidth = 1/2,
+                    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
+    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+         y = "p(Correct)",
+         colour = "Cross-language neighbourhood density",
+         fill = "Cross-language neighbourhood density",
+         linetype = "Cross-language neighbourhood density",
+         fill_ramp = "CrI (%)") +
+    scale_x_continuous(labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent) +
+    scale_fill_brewer(palette = "Reds") +
+    theme(legend.position = "none",
+          legend.box = "vertical") 
+```
+
+### Methods
+
+### Results
+
+### Discussion
 

### 2024-05-03 R/models.R

@@ -1,29 +1,27 @@
-# models
-
+#' Fit Bayesian regression model in brms 
+#' 
 get_model_fit <- function(name, formula, prior, ...){
     
-    stan_path <- here("stan", paste0(name, ".stan"))
-    results_path <- here("results", paste0(name, ".rds"))
+    results_path <- file.path("results", paste0(name, ".rds"))
     
-    brm(
-        formula = formula,
-        family = bernoulli("logit"),
-        prior = prior,
-        save_pars = save_pars(all = TRUE),
-        backend = "cmdstanr",
-        iter = 1000,
-        cores = 4,
-        chains = 4,
-        seed = 888,
-        control = list(adapt_delta = 0.95),
-        save_model = stan_path,
-        file = results_path,
-        sample_prior = "yes",
-        ...
-    )
+    out <- brm(formula = formula,
+               family = bernoulli("logit"),
+               save_pars = save_pars(all = TRUE),
+               iter = 1000,
+               cores = 4,
+               chains = 4,
+               seed = 1234,
+               silent = 1,
+               control = list(adapt_delta = 0.95),
+               file = results_path,
+               file_refit = "on_change",
+               ...)
+    
+    return(out)
 }
 
-# leave-one-out cross-validation
+#' Leave-one-out cross-validation
+#' 
 get_model_loos <- function(fits){
     path <- here("results", "model_loos.rds")
     if (file.exists(path)){
@@ -35,40 +33,23 @@ get_model_loos <- function(fits){
     return(loo)
 }
 
-# get ROC curve of Bayesian model
-get_roc_curve <- function(newdata, object, ...) {
+
+#' Get expected posterior predictions
+#' 
+get_epreds <- function(model, data, n = 100, 
+                       lv = seq(0, 1, length.out = 100), 
+                       neigh_n_h = c(0, 2, 4, 8, 16),
+                       ...) {
     
-    # enquote response variable and get brmsfit family
-    resp_var <- formula(object)[["formula"]][[2]]
-    resp_var <- enquo(resp_var)
-    model_fam <- object[["family"]][["family"]]
+    knowledge <- unique(model$data$knowledge)
+    confidence <- unique(model$data$confidence)
+    group <- unique(model$data$group)
+    nd <- expand_grid(neigh_n_h, lv, knowledge, confidence, group)
+    epreds <- tidybayes::add_epred_draws(nd, model, re_formula = NA, ...)
     
[…]

### 2024-05-03 R/stimuli.R

@@ -1,3 +1,38 @@
+#' Import CLEARPOND database
+#'
+import_clearpond <- function(cp_c_path_eng, 
+                             cp_c_path_spa,
+                             cp_h_path_eng,
+                             cp_h_path_spa) {
+    suppressWarnings({
+        header <- readLines(file.path(cp_h_path_eng), warn = FALSE)
+        eng <- read_tsv(file.path(cp_c_path_eng), 
+                        col_names = header,
+                        progress = FALSE, 
+                        show_col_types = FALSE) |> 
+            janitor::clean_names() 
+        
+        header <- readLines(file.path(cp_h_path_spa), warn = FALSE)
+        spa <- read_tsv(file.path(cp_c_path_spa), 
+                        col_names = header,
+                        progress = FALSE,
+                        show_col_types = FALSE) |> 
+            janitor::clean_names() 
+        
+        out <- bind_rows(lst("English" = eng,
+                             "Spanish" = spa),
+                         .id = "language") |>  
+            rename(sampa = phono,
+                   freq = frequency) |> 
+            mutate(sampa = str_remove_all(sampa, "\\."),
+                   freq_zipf = log10(1e6 * freq/n()) + 3,
+                   .by = c(language)) |> 
+            select(language, word, sampa, freq_zipf)
+    })
+    
+    return(out)
+}
+
 #' Import list of trials and process
 #' 
 get_trial_list <- function(trial_list_path) {

### 2024-05-03 _targets.R

@@ -1,18 +1,20 @@
 suppressWarnings({
-    library(tarchetypes)
-    library(dplyr)
-    library(tidyr)
-    library(stringr)
-    library(ggplot2)
-    library(ggdist)
-    library(beeswarm)
-    library(readxl)
-    library(brms)
-    library(cmdstanr)
-    library(stringdist)
-    library(cli)
-    library(readr)
-    library(purrr)
+    suppressPackageStartupMessages({
+        library(tarchetypes)
+        library(dplyr)
+        library(tidyr)
+        library(stringr)
+        library(ggplot2)
+        library(ggdist)
+        library(beeswarm)
+        library(readxl)
+        library(brms)
+        library(cmdstanr)
+        library(stringdist)
+        library(cli)
+        library(readr)
+        library(purrr)
+    })
 })
 
 # load functions
@@ -28,7 +30,6 @@ options(mc.cores = 4,
 
 list(
     # CLEARPOND database
-    tar_target(cp_download, download_clearpond()),
     tar_target(cp_c_path_eng, "data-raw/clearpond/englishCPdatabase2.txt", format = "file"),
     tar_target(cp_h_path_eng, "data-raw/clearpond/clearpondHeaders_EN.txt", format = "file"),
     tar_target(cp_c_path_spa, "data-raw/clearpond/spanishCPdatabase2.txt", format = "file"),
@@ -64,7 +65,7 @@ list(
                get_stimuli(trial_list = trial_list, 
                            levenshtein = levenshtein,
                            durations = durations,
-                           corpus = clearpond,
+                           corpus = clearpond, 
                            stimuli_exclude = stimuli_exclude)),
     
     # experiment responses -----------------------------------------------------
@@ -79,19 +80,20 @@ list(
     
     # questionnaire responses --------------------------------------------------
     tar_target(quest_raw_files_path, "data-raw/questionnaire/"),
-    tar_target(question_raw_files,
-               list.files(quest_raw_files_path, full.names = TRUE),
+    tar_target(quest_raw_files,
+               list.files(quest_raw_files_path, 
+                          full.names = TRUE, pattern = "quest"),
                format = "file"),
-    tar_target(quest_raw, get_quest_raw(question_raw_files)),
+    tar_target(quest_raw, get_quest_raw(quest_raw_files)),
     tar_target(quest_processed, get_quest_processed(quest_raw)),
     tar_target(quest_participants, get_quest_participants(quest_processed)),
[…]

### 2024-05-06 manuscript/_extensions/mikemahoney218/arxiv/partials/_authors.tex
Manuscript

$-- You can use as many custom partials as you need. Convention is to prefix name with '_'
$-- It can be useful to use such template to split some template parts in smaller pieces, which is easier to reuse.
$-- This '_custom.tex' is used on 'title.tex' as example.
$-- See other existing format in quarto-journals/ organisation.
$-- %%%% TODO %%%%%
$-- Use it if you need to insert content at this specific place of the main Pandoc's template. Otherwise, remove it.
$-- Here we are using it to format the authors part of the template.
$-- %%%%%%%%%%%%%%%
\textbf{$it.name.literal$}$if(it.orcid)$~\orcidlink{$it.orcid$}$endif$$for(it.affiliations/first)$\\$it.department$\\$it.name$\\$if(it.city)$$it.city$$if(it.postal-code)$,\ $it.postal-code$$endif$$endif$$endfor$\\$if(it.email)$\texttt{\href{mailto:$it.email$}{$it.email$}}$endif$

### 2024-05-06 manuscript/manuscript.tex
Manuscript

@@
 \PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
 %
 \documentclass[
-  english,
-  man,floatsintext]{apa7}
+]{article}
+
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else  
+    % xetex/luatex font selection
+  \setmainfont[]{Latin Modern Roman}
+  \setmathfont[]{Latin Modern Math}
 \fi
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={The role of cognateness in non-native spoken word recognition},
-  pdfauthor={Gonzalo Garcia-Castro1, Serene Siow2, Nuria Sebastian-Galles1, \& Kim Plunkett2},
-  pdflang={en-EN},
-  pdfkeywords={cognate, word recognition, translation, non-native, spoken word recognition, bayesian},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
 \usepackage{graphicx}
@@
 \makeatother
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-% Manuscript styling
-\usepackage{upgreek}
-\captionsetup{font=singlespacing,justification=justified}
-
-% Table formatting
-\usepackage{longtable}
-\usepackage{lscape}
-% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
-\usepackage{multirow}		% Table styling
-\usepackage{tabularx}		% Control Column width
-\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
-\usepackage{threeparttablex}            % Lets threeparttable work with longtable
-
-% Create new environments so endfloat can handle them
-% \newenvironment{ltable}
-%   {\begin{landscape}\centering\begin{threeparttable}}
-%   {\end{threeparttable}\end{landscape}}
-\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
-
-% Enables adjusting longtable caption width to table width
-% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
+% definitions for citeproc citations
+\NewDocumentCommand\citeproctext{}{}
+\NewDocumentCommand\citeproc{mm}{%
+  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
 \makeatletter
-\newcommand\LastLTentrywidth{1em}
-\newlength\longtablewidth
-\setlength{\longtablewidth}{1in}
-\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
-
-% \setlength{\parindent}{0.5in}
-% \setlength{\parskip}{0pt plus 0pt minus 0pt}
-
-% \usepackage{etoolbox}
-\makeatletter
-\patchcmd{\HyOrg@maketitle}
-  {\section{\normalfont\normalsize\abstractname}}
-  {\section*{\normalfont\normalsize\abstractname}}
-  {}{\typeout{Failed to patch abstract.}}
-\patchcmd{\HyOrg@maketitle}
-  {\section{\protect\normalfont{\@title}}}
-  {\section*{\protect\normalfont{\@title}}}
-  {}{\typeout{Failed to patch title.}}
+ % allow citations to break across lines
+ \let\@cite@ofmt\@firstofone
+ % avoid brackets around text for \cite:
+ \def\@biblabel#1{}
+ \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
 \makeatother
-\shorttitle{Cognateness and non-native word recognition}
-\keywords{cognate, word recognition, translation, non-native, spoken word recognition, bayesian\newline\indent Word count: 4553}
-\usepackage{lineno}
-
-\linenumbers
-\usepackage{csquotes}
-\ifXeTeX
-  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
-  \usepackage{polyglossia}
-  \setmainlanguage[]{english}
-\else
-  \usepackage[main=english]{babel}
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\fi
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \newlength{\cslhangindent}
@@
 \setlength{\csllabelwidth}{3em}
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
+\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
+ {\begin{list}{}{%
+  \setlength{\itemindent}{0pt}
+  \setlength{\leftmargin}{0pt}
+  \setlength{\parsep}{0pt}
   % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  \ifodd #1
+   \setlength{\leftmargin}{\cslhangindent}
+   \setlength{\itemindent}{-1\cslhangindent}
+  \fi
   % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
+  \setlength{\itemsep}{#2\baselineskip}}}
+ {\end{list}}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+\usepackage{booktabs}
+\usepackage{caption}
+\usepackage{longtable}
+\usepackage{colortbl}
+\usepackage{array}
+\usepackage{tipa}
+\usepackage{booktabs}
+\usepackage{arxiv}
+\usepackage{orcidlink}
+\usepackage{amsmath}
+\usepackage[T1]{fontenc}
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
 \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\ifLuaTeX
+\usepackage[bidi=basic]{babel}
+\else
+\usepackage[bidi=default]{babel}
+\fi
+\babelprovide[main,import]{english}
+\ifPDFTeX
+\else
+\babelfont{rm}[]{Latin Modern Roman}
+\fi
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\usepackage{bookmark}
+
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={The role of cognateness in non-native spoken word recognition},
+  pdfauthor={Gonzalo Garcia-Castro; Serene Siow; Kim Plunkett; Nuria Sebastian-Galles},
+  pdflang={en},
+  pdfkeywords={cognate, spoken word
+recognition, phonology, bilingualism, non-native speech
+processing, lexical access},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
+
+\usepackage{lineno}
+\linenumbers
+\newcommand{\runninghead}{A Preprint }
+\renewcommand{\runninghead}{Cognateness and non-native word
+recognition }
 \title{The role of cognateness in non-native spoken word recognition}
-\author{Gonzalo Garcia-Castro\textsuperscript{1}, Serene Siow\textsuperscript{2}, Nuria Sebastian-Galles\textsuperscript{1}, \& Kim Plunkett\textsuperscript{2}}
+\author{\textbf{Gonzalo
+Garcia-Castro}~\orcidlink{0000-0002-8553-4209}\\Center for Brain and
+Cognition\\Universitat Pompeu Fabra\\Barcelona,
+08005\\\texttt{\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}}\And\textbf{Serene
+Siow}~\orcidlink{0000-0001-6482-2191}\\Department of Experimental
+Psychology\\University of Oxford\\Oxford, OX2
+6GG\\\texttt{\href{mailto:siow.serene@gmail.com}{siow.serene@gmail.com}}\And\textbf{Kim
+Plunkett}~\orcidlink{0000-0003-0216-7480}\\Department of Experimental
+Psychology\\University of Oxford\\Oxford, OX2
+6GG\\\texttt{\href{mailto:kim.plunkett@psy.ox.ac.uk}{kim.plunkett@psy.ox.ac.uk}}\And\textbf{Nuria
+Sebastian-Galles}~\orcidlink{0000-0001-6938-2498}\\Center for Brain and
+Cognition\\Universitat Pompeu Fabra\\Barcelona,
+08005\\\texttt{\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}}
 \date{}
-
-\authornote{
-
-Correspondence concerning this article should be addressed to Gonzalo Garcia-Castro, Ramon Trias Fargas, 25-27, 08005 Barcelona.
-E-mail: \href{mailto:gonzalo.garciadecastro@upf.edu}{\nolinkurl{gonzalo.garciadecastro@upf.edu}}
-
-}
-
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Center for Brain and Cognition, Universitat Pompeu Fabra\\\textsuperscript{2} Department of Experimental Psychology, University of Oxford}
-
-\abstract{
-Understanding spoken words in a non-native language is easier when their translations to the native language are phonologically similar.
-Phonological similarity plays a more important role in low-proficiency bilinguals, compared to high-proficiency bilinguals, and points to the latter as relying more strongly on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
-Disentangling bilinguals' reliance on the conceptual and the lexical route during translation is challenging, given that both sources of information are present to some extent in both high- and low-proficiency bilinguals.
-In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
-Given their lack of previous knowledge on the unfamiliar language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
-This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route.
-Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation.
-This effect was larger in Spanish participants translating Catalan words than in English participants translating Catalan or Spanish.
-These results point to cross-language word-to-word links as providing rich information during word translation.
-}
-
 \begin{document}
 \maketitle
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations (Lecumberri et al., 2010; Takata \& Nábělek, 1990).
-One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, a Spanish native listening to French may encounter the word \ipatext{p<U+0254><U+0281>t} (\emph{porte}), translation of \emph{door} in French.
-The voiced uvular fricative consonant \ipatext{<U+0281>} and the open-mid back rounded vowel \ipatext{<U+0254>} do not exist in Spanish.
-These discrepancies can nonetheless be perceived as allophonic variations of native phonemes (Best et al., 2001), and the non-native speech signal engages lexical processing mechanisms (Weber \& Cutler, 2004).
-Although this mismatch has a noticeable toll on comprehension (Cutler et al., 2004), the fact that word recognition can take place in such circumstances illustrates that non-native listeners--even low-proficiency ones--are rarely completely naïve to the language they are listening to.
-In the present study we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of \emph{cognates}.
-
-All languages share, to some extent, similarities at the lexical level, frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration or social contact).
-Cognates embody a particular case of such commonalities.
-Cognates are defined as cross-language synonyms whose form is similar (at the phonological level, orthographic level, signed level, etc.), frequently due to their shared etymological origin\footnote{Some form-similar cross-language synonyms are technically not cognates.
-For example, \emph{sun} and \emph{sol} (in Spanish) share phonological onset but their etymology points to different origins.
-The complimentary case is equally problematic for the definition of cognateness in this context: cross-language synonyms that share etymological origin may not be necessarily be recognised as cognates by the average listener.
-Such is the case of \emph{cheese} and \emph{queso} (Latin root), but are phonologically dissimilar.
-for simplicity, we will use the term \emph{cognateness} in this paper to include all cross-language synonyms that fulfil a specified threshold of form similarity, regardless of etymology.
-We do not expect that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.}.
-For example, Romance languages such as Spanish and Catalan share many cognates (Schepens et al., 2012), as in the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and Catalan, respectively).
-Cognates play a pivotal role in many models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., \emph{puerta}-\emph{porta}), compared to when their labels are non-cognates (\emph{mesa}-\emph{taula}, \emph{table} in Spanish and Catalan), which suggests that word representations of both languages are activated during word retrieval (Costa et al., 2000).
-This cognate advantage is not restricted to production, instead also extending to word recognition (Midgley et al., 2011; Thierry \& Wu, 2007), word learning (De Groot \& Keijzer, 2000; Elias \& Degani, 2022; Lotto \& De Groot, 1998; Valente et al., 2018), and word translation (Christoffels et al., 2006).
-
-Early theoretical accounts of this phenomenon proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) (Potter et al., 1984; see Chen \& Leung, 1989 for similar results).
-The reason behind this claim was the assumption that second-language learners start acquiring words in L2 by associating them to their translation in L1 (lexical route), instead of directly to their shared concept (conceptual route).
-This word-to-word connection would be sensitive to form-similarity (i.e., cognateness): the more similar the two word forms are, the stronger the connection.
-As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
-cognates subsequently exert less impact on lexical processing in proficient speakers.
-The Revised Hierarchical Model (RHM, Kroll \& Stewart, 1994) captured this hypothesis and predicted that translating words from L1 to L2 (\emph{forward} translation) should take longer than translating from L2 to L1 (\emph{backward} translation).
-The rationale behind this prediction was that backward translation relies more strongly on direct word-to-word links between L1 and L2 representations, whereas forward translation relies more strongly on the mediation between the concept and the two word forms.
-Backward translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during backward translation.
+\begin{abstract}
+Understanding spoken words in a non-native language is easier when the
+words are phonologically similar to their translations into the
+listener's native language are phonologically similar. Phonological
+similarity plays a more important role in the language performance of
+low-proficiency bilinguals, compared to that of high-proficiency
+bilinguals., and This suggests that low-proficiency bilinguals may
+points to the latter as relying more strongly on cross-language
+word-to-word links (lexical route), while high-proficiency bilinguals
+rely more on word-concept associations (conceptual route) than on
+cross-language word-to-word links (lexical route). Disentangling
+bilinguals' reliance on the conceptual and lexical and conceptual routes
+during translation is challenging, given that both sources of
+information are present during spoken word comprehensionto some extent
+in both high- and low-proficiency bilinguals. In this study, we tested
+English and Spanish native speakers in a translation elicitation task,
+in which they listened to words in an unfamiliar language (Spanish or
+Catalan for English natives, and Catalan for Spanish natives), and then
+had to translate them to their native language. Given their lack of
+previous knowledge ion the unfamiliar tested language, phonological
+similarity between the presented words and their correct translations
+was the only cue available for participants to succeed in the task. This
+allowed us to explore the informativeness of the lexical route for word
+translation, in the absence of information from the conceptual route.
+Our results indicate that participants benefited from the phonological
+similarity between translation equivalents: the more similar the word
+they listened to and its translational equivalent, the higher the
+probability of a correct translation. These results point to
+cross-language word-to-word links as providing rich information during
+word translation.
+\end{abstract}
+{\bfseries \emph Keywords}
+\def\sep{\textbullet\ }
+cognate \sep spoken word
+recognition \sep phonology \sep bilingualism \sep non-native speech
+processing \sep 
+lexical access
+
+\section{Introduction}\label{introduction}
+
+Listening to non-native speech is costlier than listening to native
+speech, even for highly proficient bilinguals, and especially in
+acoustically adverse situations
+(\citeproc{ref-lecumberri2010non}{Lecumberri et al., 2010};
+\citeproc{ref-takata1990english}{Takata \& Nábělek, 1990}). However,
+there are some sources of information which may make non-native speech
+comprehension easier. One of the sources of this increased difficulty
+stems from the possible mismatch between the phonology of the native and
+the non-native languages: some acoustic features embedded in the
+non-native speech signal do not overlap perfectly with any phonemic
+category in the listener's native language. For example, imagine a
+Spanish native, with no previous familiarity with any other language,
+who listens for the first time to French may encounter the word /pɔʁt/
+(\emph{porte}), translation of door in French. The voiced uvular
+fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not
+exist in Spanish as phonemic categories. These discrepancies can
+nonetheless be perceived as allophonic variations the native (Spanish)
+phonemes /r/ and /o/ (\citeproc{ref-best2001discrimination}{Best et al.,
+2001}), and the non-native speech signal may engage lexical processing
+mechanisms (\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}),
+leading to the activation and selection, and therefore the recognition
+of its translation to Spanish /ˈpwer.ta/ (\emph{puerta}). The
+recognition of this French, unfamiliar word relies almost entirely on
+its similarity with its Spanish translation. Conversely, imagine now
+that the same Spanish listener hears English for the first time and
+encounters the word /dɔː/ (\emph{door}). This time, phonological
+similarity is of little help: /ˈpwer.ta/ and /dɔː/ share very few
+phonemes. It would come as a surprise if the listener was able to
+translate the English word successfully, if not by relying on previously
+learnt word-concept associations (which they lack, as they are
+unfamiliar to English).
+
+Cross-language phonological similarity seems, at face value, an
+important source of information for unfamiliar listeners during auditory
+word recognition, even if translations are not phonologically identical.
+Although this mismatch has a noticeable toll on comprehension
+(\citeproc{ref-cutler2004patterns}{Cutler et al., 2004}), the fact that
+word recognition can take place at all in such circumstances illustrates
+that non-native listeners--even low-proficiency ones--are rarely
+completely naïve to the language they are listening to. In the present
+study, we investigate the extent to which adult listeners rely on the
+phonological similarity between their native and non-native languages
+during word recognition, capitalising on the role of \emph{cognates.}
+
+Many languages share, to some extent, similarities at the lexical level.
+This is frequently due to their typological closeness and/or
+socio-historical events involving the speakers of these languages (e.g.,
+migration, social contact). Cognates embody a particular case of such
+commonalities. Cognates are defined as cross-language synonyms which
+have similar forms whose form is similar (at the phonological level,
+orthographic level, signed level), commonly due to their shared
+etymological origin\footnote{Some form-similar cross-language synonyms
+  are technically not cognates. For example, \emph{sun} and \emph{sol}
+  (in Spanish) share phonological onset but their etymology points to
+  different origins. The complimentary case is equally problematic when
+  defining cognateness: cross-language synonyms that share etymological
+  origin may not be necessarily be recognised as cognates by the average
+  listener. Such is the case of \emph{cheese} and \emph{queso} which
+  share Latin root but are phonologically dissimilar. For simplicity, we
+  will use the term \emph{cognateness} in this paper to include all
+  cross-language synonyms that fulfill a specified threshold of form
+  similarity, regardless of etymology. We do not expect that etymology
+  plays a direct role on language perception if it is not via
+  form-similarity, since it is not necessary for participants in
+  psycholinguistic experiments to be aware of the etymology of the words
+  they encounter in the tasks to be subject to the effect of
+  form-similarity.}. For example, Romance languages such as Spanish and
+Catalan share many cognates
+(\citeproc{ref-schepens2012distributions}{Schepens et al., 2012}), as in
+the case of \emph{puerta} and \emph{porta} (door in Spanish and Catalan,
+respectively). Cognates play a pivotal role in many models of bilingual
+lexical processing because they provide evidence that bilinguals access
+the lexicon in a language non-selective way. For instance, pictures are
+named faster by Spanish-Catalan bilinguals when their associated labels
+in both languages are cognates (e.g., \emph{puerta}-\emph{porta}),
+compared to when their labels are non-cognates
+(\emph{mesa}-\emph{taula}, {[}table{]}), which suggests that word
+representations of both languages are activated during word retrieval
+(\citeproc{ref-costa2000cognate}{Costa et al., 2000}). This cognate
+advantage is not restricted to production, but also extends to word
+recognition (\citeproc{ref-midgley2011effects}{Midgley et al., 2011};
+\citeproc{ref-thierry2007brain}{Thierry \& Wu, 2007}), word learning
+(\citeproc{ref-de2000hard}{De Groot \& Keijzer, 2000};
+\citeproc{ref-elias2022cross}{Elias \& Degani, 2022};
+\citeproc{ref-lotto1998effects}{Lotto \& De Groot, 1998};
+\citeproc{ref-valente2018does}{Valente et al., 2018}), and word
+translation (\citeproc{ref-christoffels2006memory}{Christoffels et al.,
+2006}).
+
+\subsection{Lexical and conceptual routes for
+translation}\label{lexical-and-conceptual-routes-for-translation}
+
+Early theoretical accounts of this `Cognate effect' proposed that the
+impact of cognateness on lexical processing depends on bilinguals'
+proficiency in their second language (L2)
+(\citeproc{ref-potter1984lexical}{Potter et al., 1984}; see
+\citeproc{ref-chen1989patterns}{Chen \& Leung, 1989} for similar
+results). The reasoning behind this claim was the assumption that
+second-language learners start acquiring words in their L2 by
+associating them to their translation in their L1 (lexical route),
+instead of directly to their shared concept (conceptual route). This
+word-to-word connectivity should be sensitive to form-similarity (i.e.,
+cognateness): the more similar the two word- forms are, the stronger the
+connection. As learners become more proficient, the connection between
+L2 representations and their meaning grows stronger, and L2 word
+processing becomes less reliant on word-word connections between L1 and
+L2 representations. Cognates subsequently exert less impact on lexical
+processing in proficient speakers (see
+\citeproc{ref-andras2022cognate}{Andras et al., 2022} for recent
+experimental evidence for this claim). The Revised Hierarchical Model
+(RHM, \citeproc{ref-kroll1994category}{Kroll \& Stewart, 1994}) captured
+this hypothesis and predicted that translating words from L1 to L2
+should take longer than translating from L2 to L1. The rationale behind
+this prediction was that L1-to-L2 translation relies more strongly on
+word-to-word links between L1 and L2 representations, where L2-to-L1
+translation relies more strongly on the mediation between the concept
+and the two word- forms. L1-to-L2 translation should therefore be more
+sensitive to the form similarity between the L1 and the L2
+representations: cognate words should be retrieved faster than
+non-cognate words during L1-to-L2 translation.
 
 To test these predictions, Groot et al.
-(1994) and Groot (1992) asked Dutch natives with high (but non-native) English proficiency to translate words from either Dutch to English (forward translation, L1 to L2) or from English to Dutch (backward translation, L2 to L1).
-Participants were presented with written words in English or Dutch, and were asked to speak out loud their translation in the other language as fast as possible.
-Results showed that translation times and accuracy were roughly equivalent in forward and backward translation.
-Cognates were translated equally fast in both conditions, but non-cognates were translated faster during backward than forward translation, and translation of cognates was less affected by semantic variables than non-cognates.
-Moreover, forward translation was more sensitive than backward translation to the effect of semantic variables (e.g., concreteness).
-Overall, these findings supported a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where backward translation relies more strongly on direct links between L1 and L2 representations, making it more sensitive to cognateness (than forward translation).
-However, although higher-proficiency bilinguals tested in the same task showed similar results, subsequent studies did not find differences in participant's performance during forward and backward translation, or even found better performances in forward translation (Christoffels et al., 2006, 2013).
-These later findings put into question the RHM's predictions regarding the relationship between proficiency and cognateness, and backed a more complex interaction between the L1 and L2 lexica during word recognition.
-
-Since the RHM model was proposed, later studies on monolingual populations brought attention to the role of the network of connections that words establish with each other at the form-level (phonological or orthographic) or at the conceptual level.
-For instance, the Neighbourhood Activation Model (NAM, Luce \& Pisoni, 1998) highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
-Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods (Goldinger et al., 1989; Luce et al., 1990), especially when these neighbours' lexical frequency is higher (Luce \& Pisoni, 1998).
-This finding posed an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-The Bilingual Interactive Activation (BIA/BIA+) (Dijkstra \& Van Heuven, 2002; Van Heuven et al., 1998) addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with each other, resulting in an integrated lexicon for both languages, in line with connectionist approaches to lexical processing (McClelland \& Rumelhart, 1981).
-Evidence supporting such an integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word's translation.
-When bilingual participants are presented with words in one language in a lexical decision task, translating them to the other language took longer when they were part of large neighbourhoods (Van Heuven et al., 1998).
-This suggests that presented words activated orthographic neighbourhoods in the other language, which competed for selection with the target word during recognition.
-
-A more recent model of bilingual lexical processing is Multilink (Dijkstra et al., 2019), which integrates and formalises previous claims and predictions on how an interactive account of bilingual lexical access impacts word recognition, production, and translation.
-This model assumes an integrated lexicon in which word representations from two languages establish connections in a way that is similar to connections between words of a single language.
-It also makes the assumption that during backward translation, the presented word in L2 activates form-similar words in L1, which in turn compete with the target word in L1.
-It also proposes that translation equivalents are exclusively connected through their shared concept, contrary to the RHM model's specifications (which suggest that translation equivalents are connected by excitatory connections at the lexical level; e.g., translation can be done via word-to-word connections).
-In addition, this model suggests that the strength of the association between L2 representations and their concept depends on language users' proficiency in L2.
-A strong correlation between Multilink simulations of a translation elicitation task and behavioural data from Dutch-English bilinguals (Christoffels et al., 2006) has been reported, showing that forward translation is faster than backward translation, contrary to the predictions made by the RHM model.
-Importantly, the model also generates data supporting the claim that cognates are translated faster than non-cognates, in line with previous literature on the cognate advantage during translation.
-Overall, results from Multilink are compatible with an integrated bilingual lexicon and with the existence of a facilitatory effect of cross-language similarity during translation.
-
-The fact that Multilink successfully fitted the experimental data from Christoffels et al.
-(2006) in the absence of word-to-word connections between translation equivalents is remarkable, and questions the necessity of such connections during translation, in favour of simpler models in which backward translation relies exclusively on the association between L2 and L1 word forms via their shared concept.
-It should be taken into account, however, that Christoffels et al.
-(2006) only tested fairly balanced bilinguals (i.e., bilinguals with high proficiency in L2).
-The model specification under which the quantitative predictions were generated assumes similar vocabulary sizes across languages, and also that the word-to-concept connections of L1 and L2 representations have same strength.
-The aforementioned conclusions should therefore be taken with caution when extended to unbalanced bilinguals.
-It is low-proficiency bilinguals for whom the RHM model predicts strong reliance on word-to-word connections during translation, especially during backward translation, and this prediction remains untested by Multilink.
-
-Most of the studies referenced so far rely heavily on production and comprehension of printed words: in these studies, stimuli are presented in a screen, and participants must read them in order to complete the experimental task at hand.
-There are limitations in generalising their findings to the auditory modality, especially when cognateness (or more generally any measure of similarity between word-forms) is involved.
-Consider the words \emph{radio} (English) and \emph{radio} (Spanish).
-Both words are clearly cognates, as their orthographical word form is identical.
-However, their phonological forms (English /ˈɹeɪ.di.əʊ/-- Spanish /ˈra.ðjo/) are not identical.
-Although many English listeners might recognise the Spanish word-form--even with no previous knowledge of Spanish--it is likely that they will find the translation process more challenging when presented with the auditory word-forms than when presented with the visual word-forms.
-
-Measuring the phonological similarity between two word forms is also substantially more challenging than measuring their orthographic similarity.
-Many studies use symbolic representations to compute the simiarity between word-forms of interest in either modality.
-Orthographic word-forms in alphabetical languages are represented by the strings of letters that make up the word (\emph{radio}-\emph{radio}); phonological word-forms are represented by their phonological transcriptions (ɹeɪdiəʊ-raðjo, in IPA notation) (see Floccia et al., 2018 for example).
-The similarity between each pair of word-forms is operationalised numerically by applying some algorithm that computes the edit distance between their symbolic representation.
-The Leveshtein distance is one of the most widely known algorithms: it computes the number of insertions, deletions or substitutions of the of the strings must go through in order to become identical to the other string (Levenshtein \& others, 1966), and has been successfully used to estimate the similarity of cross-language word-pairs (Gooskens \& Heeringa, 2004; Schepens et al., 2012).
-Using our previous examples, the Levenshtein distance between the orthographic word-forms \emph{radio} (English) and \emph{radio} (Spanish) is 0: both strings are identical.
-But the distance between the phonological transcriptions and \ipatext{'ra<U+0329>ðjo} is 35, which is the number of edits that separates both character strings.
-There is also an additional layer of complexity: the phonological transcriptions \ipatext{'<U+0279>e<U+026A>di<U+0329><U+0259><U+028A>} and \ipatext{'ra<U+0329>ðjo} can also be transcribed in different ways depending on variation in pronunciation between speakers, and decisions on the transcriptions (e.g.~whether to include elements of aspiration, nasalisation, etc.) can affect how accurate these symbolic representation capture all of the features that are present in the acoustic signal when either word is spoken.
-Therefore, any measure of similarity that uses phonological transcriptions as input may overestimate or underestimate their actual similarity.
-Taking into account that measuring phonological similarity is considerably more complex than measuring orthographic complexity it is no surprise that most literature on word recognition (or even on psycholinguistics) has been devoted to study orthography instead of phonology.
-
-In summary, it is very likely that previous investigations on bilinguals' reliance on form-similarity during translation might be misrepresenting the availability of such information in the auditory modality.
-In the present study we tested listener's reliance on phonological similarity when translating words from an unfamiliar language, in order to investigate the plausibility of word-to-word during backward translation.
-We tested English and Spanish native speakers in a translation elicitation task: they were presented with spoken words in an unfamiliar, non-native language, and were asked to translate them to their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as \textbf{presented words}, and the correct translation for the presented words as \textbf{target translations}.
-The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and therefore word-to-concept connections are, in principle, absent.
-As a consequence, these participants should only be able use word-to-word connections to succeed in the translation of words in an unfamiliar language.
-We predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
-
-We tested three groups of participants in the translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
-Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
-
-Following Luce and Pisoni (1998)`s NAM model, we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
-If this is the case, has more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target translation.
-For instance, one could expect English participants to incorrectly translate the Spanish word \emph{botón} as \emph{bottom} instead of as its correct translation \emph{button} (because the lexical frequency of \emph{bottom} is higher than that of \emph{button}).
-If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
-
-\hypertarget{methods}{%
-\section{Methods}\label{methods}}
-
-All materials, data, and code used in this study are hosted in the Open Science Framework \url{https://osf.io/9fjxm/} and a GitHub repository \url{https://github.com/gongcastro/translation-elicitation.git}, along with additional notes.
-
-\hypertarget{participants}{%
-\subsection{Participants}\label{participants}}
-
-Data collection took place from June 04th, 2020 to June 28th, 2020.
-We collected data from 104 participants (\emph{Mean} = 21.79 years, \emph{SD} = 2.43, Range = 18-33).
-72 participants were British English native speakers living in United Kingdom (46 female), and 32 participants were Spanish native speakers living in Spain (27 female).
-Participants in UK were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
-Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
-Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-We excluded data from participants that (a) self-rated their oral and/or written skills in a second or third language as higher than 4 in a 5-point scale (\emph{n} = 1), (b) were diagnosed with a language (\emph{n} = 2), or (c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
-
-\hypertarget{procedure}{%
-\subsection{Procedure}\label{procedure}}
-
-The experiment was implemented online using Psychopy/Pavlovia (Peirce et al., 2019).
-Participants accessed the study from a link provided by Prolific or SONA and completed the experiment from an internet browser (Chrome or Mozilla).
-After giving their consent for participating, participants answered a series of questions about their demographic status, their language background, and the set up they were using for completing the study.
-Then, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in either Catalan or Spanish (English participants) or Catalan (Spanish participants).
-They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the Catalan or Spanish lists.
-Participants in the Catalan list were presented with 86 trials, and participants in the Spanish list were presented with 103 trials.
-Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
-After one second, the audio started playing while the dot remained being displayed until the audio ended.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ``\textgreater{}'' symbol.
-Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
-Participants were allowed to correct their answer.
-Then, participants pressed the RETURN/ENTER key to confirm their answer and start and new trial.
-We excluded trials where the participant's answer was not a valid word in the target language or no answer was given.
-Trials in which the response was mistyped by only one character were counted as correct (e.g., ``pengiun'' instead of ``penguin''), as long as the response did not correspond to a distinct word.
-Trials in which participants took longer than 10 seconds to respond were also excluded.
-Participants contributed a total of 7239 valid trials (4809 in Catalan, 2430 in Spanish).
-The task took approximately 15 minutes to be completed.
+(\citeproc{ref-degroot1994forward}{1994}) and Groot
+(\citeproc{ref-de1992determinants}{1992}) asked Dutch natives with high
+(but non-native) English proficiency to translate Dutch words to English
+(L1-to-L2) or English words to Dutch (L2-to-L1). Although participants'
+performance in both conditions was roughly equivalent for cognates,
+non-cognates were translated more slowly from Dutch to English than from
+English to Dutch, suggesting that L1-to-L2 translation is more sensitive
+to form-similarity than L2-to-L1 translation. Conversely, participants'
+performance was more sensitive to semantic variables (e.g.,
+concreteness) when translating English words to Dutch, compared to when
+translating Dutch words to English, suggesting that L2-to-L1 translation
+was more sensitive to the shared conceptual properties of the
+translation pair.
+
+These findings support a soft version of the RHM model, in which both
+the conceptual and the lexical translation routes are active during
+translation, but where L1-to-L2 translation relies more strongly on
+word-to-word links between L1 and L2 representations than L2-to-L1
+translation does, and therefore is more sensitive to cognateness.
+Further evidence for the presence of this difference between L1-to-L2
+and L2-to-L1 translations on the degree of reliance on the lexical and
+or the conceptual routes between L1-to-L2 and L2-to-L1 translation was
+provided by Phillips et al. (\citeproc{ref-phillips2006erp}{2006}), who
+used eletrophysiological recordings of bilinguals translating words from
+L1 to L2 or \emph{vice versa} after semantic or phonological priming.
+The authors reported stronger N400 responses (associated with semantic
+priming) in L1-to-L2 translation than in L2-to-L12 translation, and
+stronger phonological mismatch negativity responses (associated with
+phonological priming) during L2-to-L1 translation, compared to L1-to-L2
+translation.
+
+\subsection{Competition from lexical
+neighbours}\label{competition-from-lexical-neighbours}
+
+Since the RHM model was proposed, later studies on monolingual
+populations have focused how on the role of the network of connections
+that words establish networks of connections with each other at the
+form-level (phonological or orthographic) or at the conceptual level.
+For instance, the Neighbourhood Activation Model (NAM,
+\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}) highlighted
+how lexical selection is affected by the number of phonological
+neighbours that the selected word has. Words surrounded by a larger
+number of phonological neighbours (words that only differ in one phoneme
+from the target word) are responded to more slowly and less accurately
+than words from sparser phonological neighbourhoods
+(\citeproc{ref-goldinger1989priming}{Goldinger et al., 1989};
+\citeproc{ref-luce1990similarity}{Luce et al., 1990}), especially when
+these neighbours' lexical frequency is higher
+(\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}).
+
+The existence of phonological links between words is supported with
+experimental evidence from the visual world paradigm and phonological
+priming. Meyer \& Damian (\citeproc{ref-meyer2007activation}{2007})
+showed that target pictures were named faster when presented together
+with a phonologically-related distractor, suggesting that the activation
+of shared phonological information allowed the target to be activated
+and produced faster. there are also findings suggesting that
+facilitatory or inhibitory effects on target recognition depend on the
+degree of similarity the competitor shares with the target
+(\citeproc{ref-dufour2003lexical}{Dufour \& Peereman, 2003};
+\citeproc{ref-hamburger1996phonological}{Hamburger \& Slowiaczek,
+1996}). When there is only mismatch of one phoneme between prime and
+target, inhibitory effects are exerted on target recognition
+(\citeproc{ref-dufour2003lexical}{Dufour \& Peereman, 2003}). This
+inhibition was proposed to be due to strong lexical activation of the
+prime which results in lexical interference, as opposed to simple
+phonological activation from less similar primes that triggers
+facilitation.
+
+These findings of phonological links between words in the monolingual
+lexicon pose an important question in the field of bilingualism
+research: do word representations in one language form part of the
+phonological or orthographic neighbours of the other language? Weber \&
+Cutler (\citeproc{ref-weber2004lexical}{2004}) observed that bilinguals
+displayed lexical competition effects from words in their native
+language when performing a word-referent matching task in their L2. This
+highlights that lexical competition for bilinguals during language
+comprehension may not only come from the tested language but also the
+native language, despite the native language not being actively required
+for the task. Other studies have similarly found interference when
+spoken target words in one language had phonological overlap with a word
+in the participant's other language, even when the other language was
+not required by the task (\citeproc{ref-desroches2022dynamics}{Desroches
+et al., 2022}; \citeproc{ref-spivey1999cross}{Spivey \& Marian, 1999}).
+Such patterns have even been found with young children less than 4 years
+old (\citeproc{ref-von2012language}{Von Holzen \& Mani, 2012}).
+
+The Bilingual Interactive Activation (BIA/BIA+)
+(\citeproc{ref-dijkstra2002architecture}{Dijkstra \& Van Heuven, 2002};
+\citeproc{ref-van1998orthographic}{Van Heuven et al., 1998}) addressed
+this issue and suggested that lexical representations from both
+languages establish both excitatory and inhibitory connections with each
+other, resulting in an integrated lexicon for both languages. This is,
+in line with connectionist approaches to lexical processing
+(\citeproc{ref-mcclelland1981interactive}{McClelland \& Rumelhart,
+1981}). Evidence supporting such an integrated lexicon is provided by
+the fact that bilinguals' performance in word recognition tasks is
+sensitive to the orthographic neighbourhood density of the presented
+word's translation. When bilingual participants are presented with words
+in one language in a lexical decision task, translating them to the
+other language took longer when they were part of large neighbourhoods
+(\citeproc{ref-van1998orthographic}{Van Heuven et al., 1998}). This
+suggests that presented words activated orthographic neighbourhoods in
+the other language, which competed for selection with the target word
+during recognition.
+
+\subsection{The present study}\label{the-present-study}
+
+In the present study, we tested listeners' reliance on phonological
+similarity when translating words from an unfamiliar language to their
+native language, in order to investigate the plausibility of
+word-to-word connections during L1-to-L2 translation. We tested English
+native speakers and Spanish native speakers in a translation elicitation
+task. This paradigm has been used by bilingualism researchers to
+identify easily-recognisable cognates for experimental studies. Using
+this paradigm, we aim to delve deeper into the factors affecting foreign
+word recognition by monolingual participants with no prior exposure to
+the language.
+
+Participants were presented with spoken words in an unfamiliar
+(non-native language) and were asked to translate them to their native
+language. We will henceforth refer to the auditory-presented words heard
+by participants on each trial as presented words, and the correct
+translation for the presented words as target translations. The
+motivation behind testing participants in an unfamiliar language was
+that these participants can be considered a particular (extreme) case of
+unbalanced bilingualism, in which they do not know any words in L2, and
+therefore word-to-concept connections are, in principle, absent. As a
+consequence, these participants should only be able use word-to-word
+connections to succeed in the translation of words in an unfamiliar
+language. We predicted that, if the information provided by the
+phonological form of the unfamiliar presented words is sufficient for
+translation, participants' performance should increase when the
+translation pairs are phonologically more similar.
+
+Following Luce \& Pisoni (\citeproc{ref-luce1998recognizing}{1998})`s
+NAM model, we further hypothesised that if participants are able to
+correctly translate words from an unfamiliar language based on their
+phonological similarity to the target word, similar sounding but
+incorrect translations (e.g., false friends) should be activated too. If
+this is the case, non-native words that have more phonological
+neighbours should be less likely to be correctly translated, especially
+if such neighbours have higher lexical frequency and are phonologically
+closer to the presented word than the target translation. For instance,
+one could expect English participants to incorrectly translate the
+Spanish word \emph{botón} as bottom instead of as its correct
+translation \emph{button} (because the lexical frequency of
+\emph{bottom} is higher than that of \emph{button} and also phonological
+closer). This combined constraint led us to calculate a specific form of
+phonological neighbours where a neighbour is counted only if it is
+higher frequency and has stronger phonological similarity than the
+target translation as judged by Levenshtein distance. If the
+phonological neighbourhood density of the target translation affects
+participants' performance negatively, this would suggest that non-native
+words trigger competition between words in the native language during
+translation.
+
+In Experiment 1, we collected data from two groups of British English
+natives. One group was presented with Catalan words, the other with
+Spanish words. We examined the extent to which participants were able to
+use the phonological similarity between the presented (Catalan or
+Spanish) word and its translation (i.e., cognateness) to provide
+accurate responses. In Experiment 2, we tested a group of (European)
+Spanish natives in the same task, who were presented with Catalan words.
+Catalan and Spanish are both Romance languages whose close typological
+distance is reflected in the fact that they share many cognates, where
+English is a Germanic language that shares considerably fewer cognates
+with Catalan and Spanish. By testing participants translating words from
+typologically close and distant languages, we expected to widen the
+range of the phonological similarity scores of the translation pairs
+involved in the experimental task, therefore allowing us to explore
+potential cross-language differences in participant's performance.
+Participants in Experiments 1 and 2 were surprisingly good at
+translation some words, which had low phonological similarity with their
+correct translation. In Experiment 3, we collected additional data from
+a new group of British English natives, who were presented with Catalan
+and Spanish words. In contrast with the procedure in Experiment 1,
+participants in Experiment 3 where asked, after providing a response in
+each trial, to report whether they had previous knowledge of the meaning
+of the presented word (e.g., some Spanish words are common knowledge,
+even for monolingual speakers of English). Finally, we present analyses
+on the joint data sets of Experiments 1 to 3.
+
+\section{Experiment 1}\label{experiment-1}
+
+\subsection{Methods}\label{methods}
+
+All materials, data, and code used in this study are hosted in the Open
+Science Framework \url{https://osf.io/9fjxm/} and a GitHub repository
+\url{https://github.com/gongcastro/translation-elicitation.git}, along
+with additional notes.
+
+\subsubsection{Participants}\label{participants}
+
+Data collection took place from June 04th, 2020 to June 25th, 2020. We
+collected data from 71 British English native participants living in
+United Kingdom (\emph{Mean} = 21.76 years, \emph{SD} = 2.15, Range =
+18-26, 46, 25 female).
+
+\captionsetup{labelsep=none}
+
+\begin{longtable}{l|rrrrrrc}
+
+\caption{\label{tbl-participants}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{2}{c}{Sample size} & \multicolumn{3}{c}{Age} & \multicolumn{2}{c}{L2} \\ 
+\cmidrule(lr){2-3} \cmidrule(lr){4-6} \cmidrule(lr){7-8}
+\multicolumn{1}{l}{} & N & Excluded & M & SD & Range & N knows L2? & Which L2? \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & 35 & 8 & $21.80$ & $2.08$ & $18$–$26$ & 1 & Russian \\ 
+cat-ENG & 36 & 4 & $21.72$ & $2.25$ & $18$–$25$ & 5 & German, Punjabi, French, Italian, Several \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & 33 & 12 & $21.85$ & $3.00$ & $18$–$33$ & 12 & Italian, French, German \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & 32 & 1 & $21.72$ & $2.59$ & $18$–$26$ & 3 & Japanese, German \\ 
+cat-ENG & 32 & 0 & $22.31$ & $2.39$ & $18$–$26$ & 2 & Cantonese, Irish \\ 
+\bottomrule
+
+\end{longtable}
+
+Participants were recruited via Prolific (5£ compensation) and SONA
+(compensation in academic credits), and gave informed consent before
+providing any data and the study was conducted in accordance with
+ethical standards of the Declaration of Helsinki and the protocol was by
+the University of Oxford Medical Sciences Inter-Divisional Research
+Ethics Committee (IDREC) (R60939/RE007). Participants were asked to
+complete the experiment using a laptop in a quiet place with good
+internet connection. We excluded data from participants that (a)
+self-rated their oral and/or written skills in Catalan and Spanish, or
+any other second language as four or higher in a five-point scale
+(\emph{n} = 2), (b) were diagnosed with a language (\emph{n} = 1), or
+(c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
+
+\subsubsection{Stimuli}\label{stimuli}
+
+We arranged two lists of input words to be presented to participants in
+the auditory modality: one in Catalan and one in Spanish. Words in the
+Catalan list were 5.02 phonemes long on average (\emph{SD} = 1.49,
+\emph{Range} = 2-8), and the orthographic form of their English
+translations (which participants had to type), were 5.12 characters long
+on average (\emph{SD} = 1.56, \emph{Range} = 3-9. Words in the Spanish
+list were 5.52 phonemes long on average (\emph{SD} = 1.47, \emph{Range}
+= 3-9), and the orthographic form of their English translations (which
+participants had to type), were 5.29 characters long on average
+(\emph{SD} = 1.77, \emph{Range} = 3-12.
+
+Participants listened to one audio file in each trial, each containing a
+single word presented in isolation. The audio files were the same ones
+used in child experiments conducted in the Laboratori de Recerca en
+Infància of Universitat Pompeu Fabra (Barcelona, Spain). These audio
+files were recorded by a proficient Catalan-Spanish female bilingual
+from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files
+in the same session, and then de-noised using Audacity and normalised at
+peak intensity using Praat (\citeproc{ref-broersma2021praat}{Broersma \&
+Weenink, 2021}). The average duration of the Catalan audio files was
+1.24 (\emph{SD} = 0.19, \emph{Range} = 0.8-1.58). The average duration
+of the Catalan audio files was 1.16 (\emph{SD} = 0.15, \emph{Range} =
+0.78-1.53).
+
+For each translation pair, we defined three predictors of interest: the
+lexical frequency of the correct translation (\emph{Frequency}), the
+phonological similarity between the presented (Catalan or Spanish) word
+and their correct English translation (\emph{Cognateness}), and
+presented word's number of cross-language phonological neighbours
+(\emph{CLPN}).
+
+We included \emph{Frequency} under the hypothesis that, keeping other
+predictors constant, participants would translate higher-frequency words
+more accurately and faster than lower-frequency words. Lexical
+frequencies of correct translations were extracted from SUBTLEX-UK
+(\citeproc{ref-van2014subtlex}{Van Heuven et al., 2014}), and
+transformed to Zipf scores (\citeproc{ref-van2014subtlex}{Van Heuven et
+al., 2014}). Words in the stimuli list without a lexical frequency value
+were excluded from data analysis (2 in the Catalan list, 5 in the
+Spanish list).
+
+We calculated \emph{Cognatenesss}, our main predictor of interest, by
+computing the Levenshtein similarity between the X-SAMPA transcriptions
+of each pair of translations using the \texttt{stringdist} R package
+(\citeproc{ref-van2014stringdist}{van der Loo, 2014}). The Levenshtein
+distance computes the edit distance between two character strings (in
+this case, phoneme transcriptions) by counting the number of additions,
+deletions, and substitutions necessary to make both strings identical
+(\citeproc{ref-levenshtein1966binary}{Levenshtein et al., 1966}). We
+divided this edit distance by the number of characters included in the
+longest X-SAMPA transcription of the translation pair. This results in a
+proportion score, in which values closer to zero correspond to lower
+Levenshtein distances between phonological transcriptions (i.e., higher
+similarity), and values closer to 1 correspond to higher Levenshtein
+distances (i.e., lower similarity). This transformation also help
+account for the fact that the Levenshtein distance tends to increase
+with the length of the transcriptions. For interpretability, we
+subtracted this proportion from one, so that values closer to one
+correspond to higher similarity between phonological transcriptions, and
+values closer to zero correspond to lower similarity between
+phonological transcriptions. For example, the \emph{table}
+(teɪbəl)-\emph{mesa} (mesa) translation pair had a 17\% similarity,
+while the \emph{train} (trEIn)-\emph{tren} (tɾEn) translation pair had a
+60\% similarity. Figure 1 summarises the lexical frequency, phonological
+neighbourhood density and phonological overlap of the words included in
+the Catalan and the Spanish lists.
+
+For each Catalan and Spanish word, we calculated the number of
+\emph{CLPN} by counting the number of English words with same or higher
+lexical frequency, and whose phonological transcription (in X-SAMPA
+format) different in up to one phoneme from that of the presented
+Catalan or Spanish word. Lexical frequencies and phonological
+transcriptions were extracted from the multilingual database CLEARPOND
+(\citeproc{ref-marian2012clearpond}{Marian et al., 2012})\footnote{{[}Phonological
+  trnascriptions in CLEARPOND were generated from eSPEAK,
+  \url{http://espeak.sourceforge.net/}{]}}.
+
+\captionsetup{labelsep=none}
+
+\begin{longtable}{l|rrrrrr}
+
+\caption{\label{tbl-stimuli}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{2}{c}{Frequency} & \multicolumn{2}{c}{Cognateness} & \multicolumn{2}{c}{CLPN} \\ 
+\cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
+\multicolumn{1}{l}{} & Mean ± SD & Range & Mean ± SD & Range & Mean ± SD & Range \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & $5.88$ ± $0.57$ & $4.43$–$7.24$ & $0.13$ ± $0.18$ & $0.00$–$0.75$ & $0.19$ ± $0.79$ & $0$–$5$ \\ 
+cat-ENG & $5.89$ ± $0.59$ & $4.43$–$7.27$ & $0.16$ ± $0.18$ & $0.00$–$0.67$ & $0.76$ ± $2.53$ & $0$–$15$ \\ 
+cat-SPA & $5.79$ ± $0.63$ & $4.48$–$7.70$ & $0.38$ ± $0.26$ & $0.00$–$1.00$ & $0.87$ ± $2.08$ & $0$–$12$ \\ 
+\bottomrule
+
+\end{longtable}
+
+\subsubsection{Procedure}\label{procedure}
+
+The experiment was implemented online using Psychopy/Pavlovia
+(\citeproc{ref-peirce2019psychopy2}{Peirce et al., 2019}). Participants
+accessed the study from a link provided by Prolific or SONA and
+completed the experiment from an internet browser (Chrome or Mozilla).
+After giving their consent for participating, participants answered a
+series of questions about their demographic status, their language
+background, and the set up they were using for completing the study.
+Then, participants completed the experimental task. Participants were
+informed that they would listen to a series of pre-recorded words in
+either Catalan or Spanish (English participants) or Catalan (Spanish
+participants). They were instructed to listen to each word, guess its
+meaning in English (English participants) or Spanish (Spanish
+participants), and type their answer as soon as possible. English
+participants were randomly assigned to the Catalan or Spanish lists.
+Participants in the Catalan list were presented with 83 trials, and
+participants in the Spanish list were presented with 99 trials.
+
+Each trial started with a yellow fixation point presented during one
+second on the centre of the screen over a black background. After one
+second, the audio started playing while the dot remained being displayed
+until the audio ended. Upon the offset of the fixation point and audio,
+participants were prompted to write their answer by a
+\texttt{\textgreater{}} symbol. Typed letters were displayed in the
+screen in real time to provide visual feed-back to participants.
+Participants were allowed to correct their answer. Then, participants
+pressed the \texttt{RETURN/ENTER} key to confirm their answer and start
+and new trial. We excluded trials where the participant's answer was not
+a valid word in the target language or no answer was given. Trials in
+which the response was mistyped by only one character were counted as
+correct (e.g., ``pengiun'' instead of ``penguin''), as long as the
+response did not correspond to a distinct word. Trials in which
+participants took longer than 10 seconds to respond were also excluded.
+Participants contributed a total of 9152 valid trials (5694 in Catalan,
+3458 in Spanish). The task took approximately 15 minutes to be
+completed.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{../img/design.png}
+
+}
+
+\caption{Schematic representation of a trial in the experimental task.}
+
+\end{figure}%
+
+\subsubsection{Data analysis}\label{data-analysis}
+
+Responses given by English participants to Catalan presented words were
+5.35 characters long on average (\emph{Median} = 5, \emph{SD} = 1.79,
+Range = 1-14), while their translations to Spanish responses were 5.57
+characters long on average (\emph{Median} = 5, \emph{SD} = 1.97, Range =
+2-21).
+
+\captionsetup{labelsep=none}
+
+\begin{longtable}{l|rrrr}
+
+\caption{\label{tbl-results}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{2}{c}{Correct responses} & \multicolumn{2}{c}{Incorrect responses} \\ 
+\cmidrule(lr){2-3} \cmidrule(lr){4-5}
+\multicolumn{1}{l}{} & Correct & Typo & Wrong translation & False friend \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $429$ ($16.47\%$) & $11$ ($0.42\%$) & $2,082$ ($79.95\%$) & $82$ ($3.15\%$) \\ 
+spa-ENG & $374$ ($14.37\%$) & $2$ ($0.08\%$) & $2,117$ ($81.36\%$) & $109$ ($4.19\%$) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $505$ ($18.41\%$) & $7$ ($0.26\%$) & $2,052$ ($74.81\%$) & $179$ ($6.53\%$) \\ 
+spa-ENG & $630$ ($20.03\%$) & $6$ ($0.19\%$) & $2,352$ ($74.79\%$) & $157$ ($4.99\%$) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & $780$ ($46.93\%$) & $20$ ($1.20\%$) & $736$ ($44.28\%$) & $126$ ($7.58\%$) \\ 
+\bottomrule
+
+\end{longtable}
+
+We modelled the probability of participants guessing the correct
+translation of each presented word using a generalised multilevel
+Bayesian regression model with a Bernoulli logit link distribution. We
+included as fixed effects the intercept, the main effects of
+\emph{Frequency}, \emph{Cognateness}, and \emph{CLPN}, and the two-way
+interaction between \emph{Cognateness} and \emph{CLPN}. We also included
+participant-level random intercepts and slopes for the the main effects
+and the interaction. \textbf{?@eq-model} shows a formal description of
+the model.
+
+To test the practical relevance of each predictor we followed
+(\citeproc{ref-krushke2018bayesian}{\textbf{krushke2018bayesian?}}). We
+first specified a region of practical equivalence (\emph{ROPE}) around
+zero. This area indicates the values of the regression coefficients that
+we considered as equivalent to zero. We then computed the 95\% posterior
+credible intervals (CrI) of the regression coefficients of interest.
+Finally, we calculated the proportion of the 95\% CrI that fell inside
+the ROPE. This proportion indicates the probability that the true value
+of the coefficient is equivalent to zero.
+
+All analyses were performed in R environment
+(\citeproc{ref-rcore2019r}{RCore, 2019}). We used the tidyverse family
+of R packages (\citeproc{ref-wickham2019tidyverse}{Wickham et al.,
+2019}) to process data and to generate figures. We used the
+\texttt{brms} R package (\citeproc{ref-burkner2017brms}{Bürkner, 2017})
+using the \texttt{cmdstanr} back-end to the Stan probabilistic language
+(\citeproc{ref-carpenter2017stan}{Carpenter et al., 2017}) to estimate
+and compare the models (see Appendix 1 for mode details on the models).
+
+\subsection{Results}\label{results}
+
+\textbf{?@tbl-results-1} shows a summary of participants' accuracy
+across Experiments 1, 2, and 3. Overall, participants responded less
+accurately to words with more CLPNs than to words with fewer CLPNs,
+regardless of the amount of phonological similarity between the
+presented word and its translation. This is indicated by the size of the
+regression coefficient of the two-way interaction between
+\emph{Cognateness} and \emph{CLPN} (\(\beta\) = -0.629, 95\% CrI =
+{[}-0.982, -0.344{]}, \emph{p}(ROPE) = 0). As anticipated, participants'
+performance benefited from an increase in \emph{Cognateness} (\(\beta\)
+= 7.11, 95\% CrI = {[}6.567, 7.635{]}, \emph{p}(ROPE) = 0), while the
+number of \emph{CLND} had the opposite effect (\(\beta\) = 0.007, 95\%
+CrI = {[}-0.107, 0.103{]}, \emph{p}(ROPE) = 0.927). A model including
+\emph{Group} as a main effect did not improve the predictive performance
+of the model (\(\Delta_{ELPD}\) = -1.33, \(SE\) = 1.14), indicating that
+participants in both groups performed equivalently.
+Figure~\ref{fig-epreds-1} illustrates the posterior of the average
+predictions of the model for words with different values of
+\emph{Cognateness} and \emph{CLPN}.
+
+\captionsetup{labelsep=none}
+
+\begin{longtable}{l|rrrrrrr}
+
+\caption{\label{tbl-dataset}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} &  & \multicolumn{3}{c}{Valid trials} & \multicolumn{3}{c}{Accuracy (\%)} \\ 
+\cmidrule(lr){3-5} \cmidrule(lr){6-8}
+\multicolumn{1}{l}{} & N & Mean ± SD & N trials & Range & Mean ± SD & SE & Range \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & $27$ & $96.37$ ± $2.88$ ± $2.88$ & $2,602$ & $87$–$98$ & $15.86$ ± $5.20$ & $3.05$ & $8.82$–$28.71$ \\ 
+cat-ENG & $32$ & $81.38$ ± $3.17$ ± $3.17$ & $2,604$ & $71$–$83$ & $18.48$ ± $4.89$ & $3.27$ & $10.84$–$32.56$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & $31$ & $101.45$ ± $1.80$ ± $1.80$ & $3,145$ & $92$–$102$ & $21.35$ ± $8.10$ & $3.83$ & $6.60$–$44.34$ \\ 
+cat-ENG & $32$ & $85.72$ ± $0.52$ ± $0.52$ & $2,743$ & $84$–$86$ & $20.06$ ± $5.37$ & $3.55$ & $10.00$–$30.00$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & $21$ & $79.14$ ± $3.14$ ± $3.14$ & $1,662$ & $72$–$82$ & $48.30$ ± $5.29$ & $10.54$ & $38.27$–$58.97$ \\ 
+\bottomrule
+
+\end{longtable}
 
 \begin{figure}
-\includegraphics[width=0.8\linewidth]{C:/Users/u155880/Documents/translation-elicitation/img/design} \caption{Schematic representation of a trial in the experimental task.}\label{fig:procedurefigure}
-\end{figure}
-
-\hypertarget{stimuli}{%
-\subsection{Stimuli}\label{stimuli}}
-
-\hypertarget{input-stimuli-catalanspanish}{%
-\subsubsection{Input stimuli (Catalan/Spanish)}\label{input-stimuli-catalanspanish}}
-
-We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan (to be presented to English and Spanish natives) and one in Spanish (to be presented to English natives only).
-Words in the Catalan list were 5.01 phonemes long on average (\emph{SD} = 1.48, Range = 2-8).
-Words in the Spanish list were 5.50 phonemes long on average (\emph{SD} = 1.46, Range = 3-9).
-Participants listened to one audio file in each trial, each containing a single word presented in isolation.
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat (Broersma \& Weenink, 2021).
-The average duration of the audio files was 1.20 (\emph{SD} = 0.18, Range = 0.78-1.58).
-The average duration of the Catalan audio files was 1.23 seconds (\emph{SD} = 0.19, Range = 0.80-1.58), and the average duration of the Spanish audio files was 1.16 seconds (\emph{SD} = 0.15, Range = 0.78-1.53).
-
-\hypertarget{participant-responses-englishspanish}{%
-\subsubsection{Participant responses (English/Spanish)}\label{participant-responses-englishspanish}}
-
-Responses given by English participants for Catalan presented words were 5.14 characters long on average (\emph{SD} = 1.57, Range = 3-9), while their translations to Spani responses given by Spanish participants for Catalan words were 5.44 characters long on average (\emph{SD} = 1.56, Range = 3-9).
-Responses tgiven by English participants for Spanish words were 5.27 characters long on average (\emph{SD} = 1.75, Range = 3-12).
-
-\hypertarget{target-translation-lexical-frequency}{%
-\paragraph{Target translation lexical frequency}\label{target-translation-lexical-frequency}}
-
-We extracted the lexical frequencies for English target translations from SUBTLEX-UK(Van Heuven et al., 2014), and frequencies for Spanish from SUBTLEX-ESP (Cuetos et al., 2011).
-We extracted scores as Zipf scores where available or transformed raw scores to Zipf scores to correct their logarithmic distribution, limiting their range roughly between 0 and 7, allowing an easier interpretation of further analyses (Van Heuven et al., 2014).
-
-\hypertarget{target-translation-phonological-neighbourhood-density-pthn}{%
-\paragraph{Target translation phonological neighbourhood density (PTHN)}\label{target-translation-phonological-neighbourhood-density-pthn}}
-
-We retrieved PTHN scores (Phonological Total Higher-frequency Neighbourhood size) of the English and Spanish target translations from the CLEARPOND database (Marian et al., 2012).
-PTHN scores indicate the number of phonological neighbours with higher lexical frequency than the target translation, as indicated by its frequency score in the corresponding SUBTLEX database.
-CLEARPOND defines a phonological neighbour as a word whose phonological transcription in International Phonetic Alphabet (IPA) format (generated from eSPEAK, \url{http://espeak.sourceforge.net/}) differs from that of the target translation in only one addition, deletion, or substitution.
-PTHN scores in CLEARPOND measure have been calculated using corpora of similar size across language, allowing reliable cross-language comparisons.
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}
+
+}
+
+\caption{\label{fig-epreds-1}}
+
+\end{figure}%
+
+\subsection{Discussion}\label{discussion}
+
+\section{Experiment 2}\label{experiment-2}
+
+\subsection{Methods}\label{methods-1}
+
+\subsection{Results}\label{results-1}
+
+\textbf{?@tbl-results-1} shows a summary of participants' accuracy
+across Experiments 1, 2, and 3. Overall, participants responded less
+accurately to words with more CLPNs than to words with fewer CLPNs,
+regardless of the amount of phonological similarity between the
+presented word and its translation. This is indicated by the size of the
+regression coefficient of the two-way interaction between
+\emph{Cognateness} and \emph{CLPN} (\(\beta\) = -0.584, 95\% CrI =
+{[}-0.852, -0.331{]}, \emph{p}(ROPE) = 0). As anticipated, participants'
+performance benefited from an increase in \emph{Cognateness} (\(\beta\)
+= 6.926, 95\% CrI = {[}6.4, 7.461{]}, \emph{p}(ROPE) = 0), while the
+number of \emph{CLND} had the opposite effect (\(\beta\) = 0, 95\% CrI =
+{[}-0.113, 0.085{]}, \emph{p}(ROPE) = 0.95). A model including
+\emph{Group} as a main effect did not improve the predictive performance
+of the model (\(\Delta_{ELPD}\) = -6.56, \(SE\) = 3.21), indicating that
+participants in both groups performed equivalently.
+Figure~\ref{fig-epreds-1} illustrates the posterior of the average
+predictions of the model for words with different values of
+\emph{Cognateness} and \emph{CLPN}.
 
 \begin{figure}
-\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stimuliplot-1} \caption{Distribution and summary statistics of target translations, plotted in blue for English translations of Catalan stimuli (cat-ENG), pink for English translations of Spanish stimuli (spa-ENG), and yellow for Spanish translation of Catalan stimuli (cat-SPA).
-Columns indicate: A\) Lexical frequency of target translations, expressed as Zipf scores.
-B\) PTHN of target translations, and C\) Normalised Levenshtein phonological similarity between the presented words and their target translations, expressed in percentages.
-Lexical frequencies and PTHN scores of the presented words are not used in our analyses, as participants lack any lexical representation of such words.}\label{fig:stimuliplot}
-\end{figure}
-
-\hypertarget{cross-language-phonological-similarity-levenshtein}{%
-\subsubsection{Cross-language phonological similarity (Levenshtein)}\label{cross-language-phonological-similarity-levenshtein}}
-
-We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their IPA transcriptions using the \texttt{stringsim} function of the stringdist R package (van der Loo, 2014).
-This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
-First, it computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical (Levenshtein \& others, 1966).
-This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is a percentage score, with 0\% indicating no similarity between the two strings, and 100\% indicating a perfect match between the two strings.
-We computed this similarity measure (refred simply as Levenshtein, from now on) for every translation pair in our stimuli lists.
-For example, the \emph{table} (\ipatext{'te<U+026A><U+0329>b<U+0259>l})-\emph{mesa} (mesa) translation pair had a 17\% similarity, while the \emph{train} (trEIn)-\emph{tren} (\ipatext{t<U+027E>en}) translation pair had a 60\% similarity.
-Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
-
-\hypertarget{data-analysis}{%
-\subsection{Data analysis}\label{data-analysis}}
-
-We modelled the probability of participants guessing the correct target translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant (Model 0).
-Secondly, we added the fixed effect of lexical frequency of the target translation (\(Frequency\)).
-We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood (\(PTHN\)), the main effect of Levenshtein similarity (\(Levenshtein\)), the \(PTHN \times Levenshtein\) interaction, the group of participants (\(Group\), with levels cat-ENG, spa-ENG, and cat-SPA), and finally, the \(Levenshtein \times Group\) interaction).
-We specified two sum-coded \emph{a priori} contrasts for the \(Group\) predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) (Schad et al., 2020).
-Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) (Barr et al., 2013).
-We will later refer to these models as Model 0 to Model 6.
-For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
-
-\$\$\begin{align}
-
-&\textbf{Likelihood}  \\
-y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
-
-&\textbf{Parameters}  \\
-
-logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
-& \beta_{1[p]} ~ Frequency_{i} ~ + \\
-& \beta_{2[p]} ~ PTHN_i ~ + \\
-& \beta_{3[p]} ~ Similarity_i ~ + \\
-& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
-
-\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
-\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
-&& \text{[participant-level coefficients]} \\ \\
-
-&\textbf{Prior}  \\
-
-\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
-\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
-\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
-
-\end{align}\$\$
-
-To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (\emph{LOO}) (Vehtari et al., 2017).
-More information about the models and model comparison can be found in Appendix 2.
-All analyses were performed in R environment (RCore, 2019).
-We used the tidyverse family of R packages (Wickham et al., 2019) to process data and to generate figures.
-We used the brms R package (Bürkner, 2017) using the cmdstanr backend to the Stan probabilistic language (Carpenter et al., 2017) to estimate and compare the models (see Appendix 1 for mode details on the models).
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
-Model 6, which included all main effects and the two-way interactions between \(PTHN\) and \(Levenshtein\), and between \(Group\) and \(Levenshtein\) and \(PTHN\) and consonant similarity, showed the best performance.
-
-\includegraphics{C:/Users/u155880/Documents/translation-elicitation/img/loos.png}
-
-We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
-For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
-Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increments or decrements in the probability of correct translation (Gelman et al., 2020).
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}
+
+}
+
+\caption{\label{fig-epreds-2}}
+
+\end{figure}%
+
+\subsection{Discussion}\label{discussion-1}
+
+\section{Experiment 3}\label{experiment-3}
+
+\subsection{Methods}\label{methods-2}
+
+Same as in Experiment 1.
+
+\subsection{Results}\label{results-2}
+
+\textbf{?@tbl-results-1} shows a summary of participants' accuracy
+across Experiments 1, 2, and 3. Overall, participants responded less
+accurately to words with more CLPNs than to words with fewer CLPNs,
+regardless of the amount of phonological similarity between the
+presented word and its translation. This is indicated by the size of the
+regression coefficient of the two-way interaction between
+\emph{Cognateness} and \emph{CLPN} (\(\beta\) = -0.409, 95\% CrI =
+{[}-0.895, 0.111{]}, \emph{p}(ROPE) = 0.081). As anticipated,
+participants' performance benefited from an increase in
+\emph{Cognateness} (\(\beta\) = 9.274, 95\% CrI = {[}8.473, 10.269{]},
+\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
+effect (\(\beta\) = -0.069, 95\% CrI = {[}-0.353, 0.192{]},
+\emph{p}(ROPE) = 0.492). Figure~\ref{fig-epreds-1} illustrates the
+posterior of the average predictions of the model for words with
+different values of \emph{Cognateness} and \emph{CLPN}.
 
 \begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/posteriorfix-1.pdf}
-\caption{\label{fig:posteriorfix}Estimated posterior distributions of coefficients in Model 4.
-A) Population-level effects.
-Distributions indicate the estimated posterior likelihood density of regression coefficients of fixed effects.
-Credible intervals (\emph{CrI}), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\%, 80\%, and 50\% probability.
-Black dots represent the mean of each distribution.
-B) Participant-level coefficient variability.
-The model estimated participant-level coefficients to account for the dependency between responses from the same participant.
-Distributions in this panel indicate the estimated variability across coefficients from different participants, expressed as standard deviations (\emph{SD}).
-C) Correlation between participant-level effects.
-The model allowed participant-level coefficients to co-vary.
-This panel represents the Pearson correlations between each pair of coefficients, expressed as the mean of the posterior distribution of each estimated correlation.}
-\end{figure}
-
-Overall, participants had 43.08\%, (95\% \emph{CrI} = {[}38.20\%, 48.00\%{]}) likelihood to produce correct translations.
-The lexical frequency of the target translation had little impact on accuracy: every standard deviation increment (\emph{SD} = 0.57 Zipf points) decreased the probability of correct translation in 1.14\% (95\% \emph{CrI} = {[}1.14\%, -2.40\%{]}, \(BF_{\beta\neq< 0}\) = 1.20).
-The number of phonological neighbours with higher lexical frequency than the target translation (\emph{PTHN}) slightly decreased the probability of a correct responses by -1.70\% (95\% \emph{CrI} = {[}-5.26\%, 1.98\%{]}, \(BF_{\beta \neq 0}\) = 0.91) for every increase standard deviation increment (1 \emph{SD} = 3.94 neighbours).
-Finally, phonological similarity between the presented word and target translation had a strong effect on participants' accuracy: for every \emph{SD} increment in Levenshtein similarity (1 \emph{SD} = 23\% similarity), translation accuracy increased in 12.30\% (95\% \emph{CrI} = {[}7.48\%, 17.39\%{]}, \(BF_{\beta \neq 0}\) = \(>3.09 \times 10^{18}\)).
-The interaction effect between \(PTHN \times Levenshtein\) was small: a +1 SD increment in Levenshtein increased the effect of \emph{PTHN} by -0.11\% (95\% \emph{CrI} = {[}-3.72\%, 3.54\%{]}, \(BF_{\beta \neq 0}\) = 1.39)
-
-Spanish natives translating Catalan words were 3.74\% more accurate than English natives translating Spanish and Catalan words (95\% \emph{CrI} = {[}-1.09\%, 8.64\%{]}, \(BF_{\beta \neq 0}\) = 0.35).
-The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in 3.42\% for every one \emph{SD} increment in Levenshtein similarity (95\% \emph{CrI} = {[}-1.52\%, 8.45\%{]}, \(BF_{\beta \neq 0}\) = 0.41).
-English natives translating Catalan words were 0.60\% more accurate than those translating Spanish words (95\% \emph{CrI} = {[}-3.98\%, 5.17\%{]}, \(BF_{\beta \neq 0}\) = 1.00).
-For every one \emph{SD} increment in Levenshtein similarity, the effect of presented language increased in 3.08\% (95\% \emph{CrI} = {[}-1.24\%, 7.43\%{]}, \(BF_{\beta \neq 0}\) = 0.41).
-
-\begin{figure}
-\centering
-\includegraphics{manuscript_files/figure-latex/marginaleffects-1.pdf}
-\caption{\label{fig:marginaleffects}Expected mean posterior predictions from Model 4.
-A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0\% Levenshtein), 100 simulations for translations with mean similarity (20.51\% Levenshtein), and 100 simulations for translations with the maximum similarity observed in the dataset (83.33\% Levenshtein).
-These simulations were obtained for the range of values of the PTHN scores.
-For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
-Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
-The dispersion of the lines indicates the uncertainty of our predictions.
-We computed these posterior predictions for each group of participants, and plot tem in separate panels.}
-\end{figure}
-
-The conditional intra-class correlation {[}\emph{ICC}; Gelman and Hill (2006){]} for items was 71.49\%, indicating that responses from different participants to the same item were, on average, more similar to each other than responses to other items.
-Item-level intercepts showed remarkable variability (\emph{SD} = 96.91\%, 95\% CrI = {[}94.66\%, 98.36\%{]}), indicating that participants' accuracy differed substantially across words: some words were very likely to be correctly translated (e.g., Spanish \emph{tomate}-\emph{tomato}, /tomate/-/təmɑːtoʊ/, 99.73\%, 95\% CrI = {[}96.40\%, 99.99\%{]}), while others were almost never translated correctly (e.g., Catalan \emph{colom}-\emph{pigeon}, /kulom/-/pɪdʒɪn/, 0.08\%, 95\% CrI = {[}0.00\%, 1.00\%{]}).
-Participant-level intercepts also showed high variability (\emph{SD} = 75.73\%, 95\% CrI = {[}71.33\%, 80.33\%{]}), suggesting that average translation accuracy also varied significantly across participants.
-Finally, participant-level estimated slopes showed less variability: \emph{SD} of \(Frequency\), \(PTHN\), \(Levenshtein\), \(PTHN \times Levenshtein\), and \(Levenshtein \times Group 2\) ranged between 2.56\%, and 37.19\%.
-The participant-level estimated slopes of the \(Leveshtein \times Group 1\) interaction are an exception, though (\emph{SD} = 37.19\%, 95\% CrI = {[}14.92\%, {]}).
-This indicates that the impact of phonological similarity (\(Levenshtein\)) on the difference in accuracy between English and Spanish natives differed substantially across participants.
-
-Finally, participant-level effects showed weak Pearson correlations overall (\textless10\%), with three exceptions.
-The estimated correlation between the intercepts and the slopes of \(Frequency\) was 35.81\% (95\% CrI = {[}-2.19\%, 65.22\%{]}, \(BF_{Cor \neq 0}\) = 0.15).
-The estimated correlation between the slopes of \(Frequency\) and \(Levenshtein \times Group 1\) was 26.07\% (95\% CrI = {[}-13.12\%, 57.14\%{]}, \(BF_{Cor \neq 0}\) = 0.37).
-The estimated correlation between the intercepts and the slopes of \(PTHN\) was 1.81\% (95\% CrI = {[}-39.01\%, 38.35\%{]}, \(BF_{Cor \neq 0}\) = 1).
-The large uncertainty associated with these estimates, however, prevents us from interpreting the size of these correlations as practically meaningful.
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-\newpage
-
-\hypertarget{references}{%
-\section{References}\label{references}}
-
-\begingroup
-\setlength{\parindent}{-0.5in}
-\setlength{\leftskip}{0.5in}
-
-\hypertarget{refs}{}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}
+
+}
+
+\caption{\label{fig-epreds-3}}
+
+\end{figure}%
+
+\subsection{Discussion}\label{discussion-2}
+
+\section{References}\label{references}
+
+\phantomsection\label{refs}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-barr2013random}{}}%
-Barr, D.
-J., Levy, R., Scheepers, C., \& Tily, H.
-J.
-(2013).
-Random effects structure for confirmatory hypothesis testing: Keep it maximal.
-\emph{Journal of Memory and Language}, \emph{68}(3), 255--278.
-
-\leavevmode\vadjust pre{\hypertarget{ref-best2001discrimination}{}}%
-Best, C.
-T., McRoberts, G.
-W., \& Goodell, E.
-(2001).
-Discrimination of non-native consonant contrasts varying in perceptual assimilation to the listener's native phonological system.
-\emph{The Journal of the Acoustical Society of America}, \emph{109}(2), 775--794.
-
-\leavevmode\vadjust pre{\hypertarget{ref-broersma2021praat}{}}%
-Broersma, P., \& Weenink, D. (2021). \emph{Praat: Doing phonetics by computer {[}computer program{]}} (Version 6.1.54). \url{http://www.praat.org/}
-
-\leavevmode\vadjust pre{\hypertarget{ref-burkner2017brms}{}}%
-Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
-
-\leavevmode\vadjust pre{\hypertarget{ref-carpenter2017stan}{}}%
-Carpenter, B., Gelman, A., Hoffman, M.
-D., Lee, D., Goodrich, B., Betancourt, M., Brubaker, M.
-A., Guo, J., Li, P., \& Riddell, A.
-(2017).
-Stan: A probabilistic programming language.
-\emph{Grantee Submission}, \emph{76}(1), 1--32.
-
-\leavevmode\vadjust pre{\hypertarget{ref-chen1989patterns}{}}%
-Chen, H.-C., \& Leung, Y.-S. (1989). Patterns of lexical processing in a nonnative language. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{15}(2), 316.
-
-\leavevmode\vadjust pre{\hypertarget{ref-christoffels2006memory}{}}%
-Christoffels, I.
-K., De Groot, A.
-M., \& Kroll, J.
-F.
-(2006).
-Memory and language skills in simultaneous interpreters: The role of expertise and language proficiency.
-\emph{Journal of Memory and Language}, \emph{54}(3), 324--345.
-
-\leavevmode\vadjust pre{\hypertarget{ref-christoffels2013language}{}}%
-Christoffels, I. K., Ganushchak, L., \& Koester, D. (2013). Language conflict in translation: An ERP study of translation production. \emph{Journal of Cognitive Psychology}, \emph{25}(5), 646--664.
-
-\leavevmode\vadjust pre{\hypertarget{ref-costa2000cognate}{}}%
-Costa, A., Caramazza, A., \& Sebastian-Galles, N.
-(2000).
-The cognate facilitation effect: Implications for models of lexical access.
-\emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{26}(5), 1283.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cuetos2011subtlex}{}}%
-Cuetos, F., Glez-Nosti, M., Barbon, A., \& Brysbaert, M.
-(2011).
-SUBTLEX-ESP: Frecuencias de las palabras espanolas basadas en los subtitulos de las peliculas.
-\emph{Psicol{ó}gica}, \emph{32}(2), 133--144.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cutler2004patterns}{}}%
-Cutler, A., Weber, A., Smits, R., \& Cooper, N.
-(2004).
-Patterns of english phoneme confusions by native and non-native listeners.
-\emph{The Journal of the Acoustical Society of America}, \emph{116}(6), 3668--3678.
-
-\leavevmode\vadjust pre{\hypertarget{ref-de2000hard}{}}%
-De Groot, A.
-M., \& Keijzer, R.
-(2000).
-What is hard to learn is easy to forget: The roles of word concreteness, cognate status, and word frequency in foreign-language vocabulary learning and forgetting.
+\bibitem[\citeproctext]{ref-andras2022cognate}
+Andras, F., Rivera, M., Bajo, T., Dussias, P. E., \& Paolieri, D.
+(2022). Cognate facilitation effect during auditory comprehension of a
+second language: A visual world eye-tracking study. \emph{International
+Journal of Bilingualism}, 13670069211033359.
+
+\bibitem[\citeproctext]{ref-best2001discrimination}
+Best, C. T., McRoberts, G. W., \& Goodell, E. (2001). Discrimination of
+non-native consonant contrasts varying in perceptual assimilation to the
+listener's native phonological system. \emph{The Journal of the
+Acoustical Society of America}, \emph{109}(2), 775--794.
+
+\bibitem[\citeproctext]{ref-broersma2021praat}
+Broersma, P., \& Weenink, D. (2021). \emph{Praat: Doing phonetics by
+computer {[}computer program{]}} (Version 6.1.54).
+\url{http://www.praat.org/}
+
+\bibitem[\citeproctext]{ref-burkner2017brms}
+Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
+using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+
+\bibitem[\citeproctext]{ref-carpenter2017stan}
+Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B.,
+Betancourt, M., Brubaker, M. A., Guo, J., Li, P., \& Riddell, A. (2017).
+Stan: A probabilistic programming language. \emph{Grantee Submission},
+\emph{76}(1), 1--32.
+
+\bibitem[\citeproctext]{ref-chen1989patterns}
+Chen, H.-C., \& Leung, Y.-S. (1989). Patterns of lexical processing in a
+nonnative language. \emph{Journal of Experimental Psychology: Learning,
+Memory, and Cognition}, \emph{15}(2), 316.
+
+\bibitem[\citeproctext]{ref-christoffels2006memory}
+Christoffels, I. K., De Groot, A. M., \& Kroll, J. F. (2006). Memory and
+language skills in simultaneous interpreters: The role of expertise and
+language proficiency. \emph{Journal of Memory and Language},
+\emph{54}(3), 324--345.
+
+\bibitem[\citeproctext]{ref-costa2000cognate}
+Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate
+facilitation effect: Implications for models of lexical access.
+\emph{Journal of Experimental Psychology: Learning, Memory, and
+Cognition}, \emph{26}(5), 1283.
+
+\bibitem[\citeproctext]{ref-cutler2004patterns}
+Cutler, A., Weber, A., Smits, R., \& Cooper, N. (2004). Patterns of
+english phoneme confusions by native and non-native listeners. \emph{The
+Journal of the Acoustical Society of America}, \emph{116}(6),
+3668--3678.
+
+\bibitem[\citeproctext]{ref-de2000hard}
+De Groot, A. M., \& Keijzer, R. (2000). What is hard to learn is easy to
+forget: The roles of word concreteness, cognate status, and word
+frequency in foreign-language vocabulary learning and forgetting.
 \emph{Language Learning}, \emph{50}(1), 1--56.
 
-\leavevmode\vadjust pre{\hypertarget{ref-dijkstra2002architecture}{}}%
-Dijkstra, T., \& Van Heuven, W.
-J.
-(2002).
-The architecture of the bilingual word recognition system: From identification to decision.
+\bibitem[\citeproctext]{ref-desroches2022dynamics}
+Desroches, A. S., Friesen, D. C., Teles, M., Korade, C. A., \& Forest,
+E. W. (2022). The dynamics of spoken word recognition in bilinguals.
+\emph{Bilingualism: Language and Cognition}, \emph{25}(4), 705--710.
+
+\bibitem[\citeproctext]{ref-dijkstra2002architecture}
+Dijkstra, T., \& Van Heuven, W. J. (2002). The architecture of the
+bilingual word recognition system: From identification to decision.
 \emph{Bilingualism: Language and Cognition}, \emph{5}(3), 175--197.
 
-\leavevmode\vadjust pre{\hypertarget{ref-dijkstra2019multilink}{}}%
-Dijkstra, T., Wahl, A., Buytenhuijs, F., Van Halem, N., Al-Jibouri, Z., De Korte, M., \& Rekké, S.
-(2019).
-Multilink: A computational model for bilingual word recognition and word translation.
-\emph{Bilingualism: Language and Cognition}, \emph{22}(4), 657--679.
-
-\leavevmode\vadjust pre{\hypertarget{ref-elias2022cross}{}}%
-Elias, M., \& Degani, T.
-(2022).
-Cross-language interactions during novel word learning: The contribution of form similarity and participant characteristics.
-\emph{Bilingualism: Language and Cognition}, 1--18.
-
-\leavevmode\vadjust pre{\hypertarget{ref-floccia2018introduction}{}}%
-Floccia, C., Sambrook, T.
-D., Delle Luche, C., Kwok, R., Goslin, J., White, L., Cattani, A., Sullivan, E., Abbot-Smith, K., Krott, A., \& others.
-(2018).
-I: INTRODUCTION.
-\emph{Monographs of the Society for Research in Child Development}, \emph{83}(1), 7--29.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gelman2006data}{}}%
-Gelman, A., \& Hill, J. (2006). \emph{Data analysis using regression and multilevel/hierarchical models}. Cambridge university press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gelman2020regression}{}}%
-Gelman, A., Hill, J., \& Vehtari, A. (2020). \emph{Regression and other stories}. Cambridge University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-goldinger1989priming}{}}%
-Goldinger, S. D., Luce, P. A., \& Pisoni, D. B. (1989). Priming lexical neighbors of spoken words: Effects of competition and inhibition. \emph{Journal of Memory and Language}, \emph{28}(5), 501--518.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gooskens2004perceptive}{}}%
-Gooskens, C., \& Heeringa, W. (2004). Perceptive evaluation of levenshtein dialect distance measurements using norwegian dialect data. \emph{Language Variation and Change}, \emph{16}(3), 189--207.
-
-\leavevmode\vadjust pre{\hypertarget{ref-de1992determinants}{}}%
-Groot, A. M. de. (1992). Determinants of word translation. \emph{Journal of Experimental Psychology: Learning, Memory, and Cognition}, \emph{18}(5), 1001.
-
-\leavevmode\vadjust pre{\hypertarget{ref-degroot1994forward}{}}%
-Groot, A. M. de, Dannenburg, L., \& Vanhell, J. G. (1994). Forward and backward word translation by bilinguals. \emph{Journal of Memory and Language}, \emph{33}(5), 600--629.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kroll1994category}{}}%
-Kroll, J.
-F., \& Stewart, E.
-(1994).
-Category interference in translation and picture naming: Evidence for asymmetric connections between bilingual memory representations.
-\emph{Journal of Memory and Language}, \emph{33}(2), 149--174.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lecumberri2010non}{}}%
-Lecumberri, M. L. G., Cooke, M., \& Cutler, A. (2010). Non-native speech perception in adverse conditions: A review. \emph{Speech Communication}, \emph{52}(11-12), 864--886.
-
-\leavevmode\vadjust pre{\hypertarget{ref-levenshtein1966binary}{}}%
-Levenshtein, V. I., \& others. (1966). Binary codes capable of correcting deletions, insertions, and reversals. \emph{Soviet Physics Doklady}, \emph{10}, 707--710.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lotto1998effects}{}}%
-Lotto, L., \& De Groot, A. M. (1998). Effects of learning method and word type on acquiring vocabulary in an unfamiliar language. \emph{Language Learning}, \emph{48}(1), 31--69.
-
-\leavevmode\vadjust pre{\hypertarget{ref-luce1998recognizing}{}}%
-Luce, P. A., \& Pisoni, D. B. (1998). Recognizing spoken words: The neighborhood activation model. \emph{Ear and Hearing}, \emph{19}(1), 1.
-
-\leavevmode\vadjust pre{\hypertarget{ref-luce1990similarity}{}}%
-Luce, P. A., Pisoni, D. B., \& Goldinger, S. D. (1990). \emph{Similarity neighborhoods of spoken words.}
-
-\leavevmode\vadjust pre{\hypertarget{ref-marian2012clearpond}{}}%
-Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012). \emph{CLEARPOND: Cross-linguistic easy-access resource for phonological and orthographic neighborhood densities}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-mcclelland1981interactive}{}}%
-McClelland, J. L., \& Rumelhart, D. E. (1981). An interactive activation model of context effects in letter perception: I. An account of basic findings. \emph{Psychological Review}, \emph{88}(5), 375.
-
-\leavevmode\vadjust pre{\hypertarget{ref-midgley2011effects}{}}%
-Midgley, K.
-J., Holcomb, P.
-J., \& Grainger, J.
-(2011).
-Effects of cognate status on word comprehension in second language learners: An ERP investigation.
-\emph{Journal of Cognitive Neuroscience}, \emph{23}(7), 1634--1647.
-
-\leavevmode\vadjust pre{\hypertarget{ref-peirce2019psychopy2}{}}%
-Peirce, J., Gray, J.
-R., Simpson, S., MacAskill, M., Höchenberger, R., Sogo, H., Kastman, E., \& Lindeløv, J.
-K.
-(2019).
-PsychoPy2: Experiments in behavior made easy.
-\emph{Behavior Research Methods}, \emph{51}(1), 195--203.
-
-\leavevmode\vadjust pre{\hypertarget{ref-potter1984lexical}{}}%
-Potter, M.
-C., So, K.-F., Von Eckardt, B., \& Feldman, L.
-B.
-(1984).
-Lexical and conceptual representation in beginning and proficient bilinguals.
-\emph{Journal of Verbal Learning and Verbal Behavior}, \emph{23}(1), 23--38.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rcore2019r}{}}%
-RCore, T. (2019). \emph{R: A language and environment for statistical computing. R foundation for statistical computing, austria}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-schad2020capitalize}{}}%
-Schad, D.
-J., Vasishth, S., Hohenstein, S., \& Kliegl, R.
-(2020).
-How to capitalize on a priori contrasts in linear (mixed) models: A tutorial.
-\emph{Journal of Memory and Language}, \emph{110}, 104038.
-
-\leavevmode\vadjust pre{\hypertarget{ref-schepens2012distributions}{}}%
-Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of cognates in europe as based on levenshtein distance. \emph{Bilingualism: Language and Cognition}, \emph{15}(1), 157--166.
-
-\leavevmode\vadjust pre{\hypertarget{ref-takata1990english}{}}%
-Takata, Y., \& Nábělek, A.
-K.
-(1990).
-English consonant recognition in noise and in reverberation by japanese and american listeners.
-\emph{The Journal of the Acoustical Society of America}, \emph{88}(2), 663--666.
-
-\leavevmode\vadjust pre{\hypertarget{ref-thierry2007brain}{}}%
-Thierry, G., \& Wu, Y.
-J.
-(2007).
-Brain potentials reveal unconscious translation during foreign-language comprehension.
-\emph{Proceedings of the National Academy of Sciences}, \emph{104}(30), 12530--12535.
-
-\leavevmode\vadjust pre{\hypertarget{ref-valente2018does}{}}%
-Valente, D., Ferré, P., Soares, A., Rato, A., \& Comesaña, M.
-(2018).
-Does phonological overlap of cognate words modulate cognate acquisition and processing in developing and skilled readers?
-\emph{Language Acquisition}, \emph{25}(4), 438--453.
-
-\leavevmode\vadjust pre{\hypertarget{ref-van2014stringdist}{}}%
-van der Loo, M. P. J. (2014). The stringdist package for approximate string matching. \emph{The {R} {J}ournal}, \emph{6}, 111--122. \url{https://CRAN.R-project.org/package=stringdist}
-
-\leavevmode\vadjust pre{\hypertarget{ref-van1998orthographic}{}}%
-Van Heuven, W. J., Dijkstra, T., \& Grainger, J. (1998). Orthographic neighborhood effects in bilingual word recognition. \emph{Journal of Memory and Language}, \emph{39}(3), 458--483.
-
-\leavevmode\vadjust pre{\hypertarget{ref-van2014subtlex}{}}%
-Van Heuven, W.
-J., Mandera, P., Keuleers, E., \& Brysbaert, M.
-(2014).
-SUBTLEX-UK: A new and improved word frequency database for british english.
-\emph{Quarterly Journal of Experimental Psychology}, \emph{67}(6), 1176--1190.
-
-\leavevmode\vadjust pre{\hypertarget{ref-vehtari2017practical}{}}%
-Vehtari, A., Gelman, A., \& Gabry, J. (2017). Practical bayesian model evaluation using leave-one-out cross-validation and WAIC. \emph{Statistics and Computing}, \emph{27}(5), 1413--1432.
-
-\leavevmode\vadjust pre{\hypertarget{ref-weber2004lexical}{}}%
-Weber, A., \& Cutler, A. (2004). Lexical competition in non-native spoken-word recognition. \emph{Journal of Memory and Language}, \emph{50}(1), 1--25.
-
-\leavevmode\vadjust pre{\hypertarget{ref-wickham2019tidyverse}{}}%
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L.
-D., François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M., Pedersen, T.
-L., Miller, E., Bache, S.
-M., Müller, K., Ooms, J., Robinson, D., Seidel, D.
-P., Spinu, V., \ldots{} Yutani, H.
-(2019).
-Welcome to the {tidyverse}.
-\emph{Journal of Open Source Software}, \emph{4}(43), 1686.
-\url{https://doi.org/10.21105/joss.01686}
+\bibitem[\citeproctext]{ref-dufour2003lexical}
+Dufour, S., \& Peereman, R. (2003). Lexical competition in phonological
+priming: Assessing the role of phonological match and mismatch lengths
+between primes and targets. \emph{Memory \& Cognition}, \emph{31},
+1271--1283.
+
+\bibitem[\citeproctext]{ref-elias2022cross}
+Elias, M., \& Degani, T. (2022). Cross-language interactions during
+novel word learning: The contribution of form similarity and participant
+characteristics. \emph{Bilingualism: Language and Cognition}, 1--18.
+
+\bibitem[\citeproctext]{ref-goldinger1989priming}
+Goldinger, S. D., Luce, P. A., \& Pisoni, D. B. (1989). Priming lexical
+neighbors of spoken words: Effects of competition and inhibition.
+\emph{Journal of Memory and Language}, \emph{28}(5), 501--518.
+
+\bibitem[\citeproctext]{ref-de1992determinants}
+Groot, A. M. de. (1992). Determinants of word translation. \emph{Journal
+of Experimental Psychology: Learning, Memory, and Cognition},
+\emph{18}(5), 1001.
+
+\bibitem[\citeproctext]{ref-degroot1994forward}
+Groot, A. M. de, Dannenburg, L., \& Vanhell, J. G. (1994). Forward and
+backward word translation by bilinguals. \emph{Journal of Memory and
+Language}, \emph{33}(5), 600--629.
+
+\bibitem[\citeproctext]{ref-hamburger1996phonological}
+Hamburger, M., \& Slowiaczek, L. M. (1996). Phonological priming
+reflects lexical competition. \emph{Psychonomic Bulletin \& Review},
+\emph{3}(4), 520--525.
+
+\bibitem[\citeproctext]{ref-kroll1994category}
+Kroll, J. F., \& Stewart, E. (1994). Category interference in
+translation and picture naming: Evidence for asymmetric connections
+between bilingual memory representations. \emph{Journal of Memory and
+Language}, \emph{33}(2), 149--174.
+
+\bibitem[\citeproctext]{ref-lecumberri2010non}
+Lecumberri, M. L. G., Cooke, M., \& Cutler, A. (2010). Non-native speech
+perception in adverse conditions: A review. \emph{Speech Communication},
+\emph{52}(11-12), 864--886.
+
+\bibitem[\citeproctext]{ref-levenshtein1966binary}
+Levenshtein, V. I. et al. (1966). Binary codes capable of correcting
+deletions, insertions, and reversals. \emph{Soviet Physics Doklady},
+\emph{10}, 707--710.
+
+\bibitem[\citeproctext]{ref-lotto1998effects}
+Lotto, L., \& De Groot, A. M. (1998). Effects of learning method and
+word type on acquiring vocabulary in an unfamiliar language.
+\emph{Language Learning}, \emph{48}(1), 31--69.
+
+\bibitem[\citeproctext]{ref-luce1998recognizing}
+Luce, P. A., \& Pisoni, D. B. (1998). Recognizing spoken words: The
+neighborhood activation model. \emph{Ear and Hearing}, \emph{19}(1), 1.
+
+\bibitem[\citeproctext]{ref-luce1990similarity}
+Luce, P. A., Pisoni, D. B., \& Goldinger, S. D. (1990). \emph{Similarity
+neighborhoods of spoken words.}
+
+\bibitem[\citeproctext]{ref-marian2012clearpond}
+Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012).
+\emph{CLEARPOND: Cross-linguistic easy-access resource for phonological
+and orthographic neighborhood densities}.
+
+\bibitem[\citeproctext]{ref-mcclelland1981interactive}
+McClelland, J. L., \& Rumelhart, D. E. (1981). An interactive activation
+model of context effects in letter perception: I. An account of basic
+findings. \emph{Psychological Review}, \emph{88}(5), 375.
+
+\bibitem[\citeproctext]{ref-meyer2007activation}
+Meyer, A. S., \& Damian, M. F. (2007). Activation of distractor names in
+the picture-picture interference paradigm. \emph{Memory \& Cognition},
+\emph{35}(3), 494--503.
+
+\bibitem[\citeproctext]{ref-midgley2011effects}
+Midgley, K. J., Holcomb, P. J., \& Grainger, J. (2011). Effects of
+cognate status on word comprehension in second language learners: An ERP
+investigation. \emph{Journal of Cognitive Neuroscience}, \emph{23}(7),
+1634--1647.
+
+\bibitem[\citeproctext]{ref-peirce2019psychopy2}
+Peirce, J., Gray, J. R., Simpson, S., MacAskill, M., Höchenberger, R.,
+Sogo, H., Kastman, E., \& Lindeløv, J. K. (2019). PsychoPy2: Experiments
+in behavior made easy. \emph{Behavior Research Methods}, \emph{51}(1),
+195--203.
+
+\bibitem[\citeproctext]{ref-phillips2006erp}
+Phillips, N. A., Klein, D., Mercier, J., \& Boysson, C. de. (2006). ERP
+measures of auditory word repetition and translation priming in
+bilinguals. \emph{Brain Research}, \emph{1125}(1), 116--131.
+
+\bibitem[\citeproctext]{ref-potter1984lexical}
+Potter, M. C., So, K.-F., Von Eckardt, B., \& Feldman, L. B. (1984).
+Lexical and conceptual representation in beginning and proficient
+bilinguals. \emph{Journal of Verbal Learning and Verbal Behavior},
+\emph{23}(1), 23--38.
+
+\bibitem[\citeproctext]{ref-rcore2019r}
+RCore, T. (2019). \emph{R: A language and environment for statistical
+computing. R foundation for statistical computing, austria}.
+
+\bibitem[\citeproctext]{ref-schepens2012distributions}
+Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of
+cognates in europe as based on levenshtein distance. \emph{Bilingualism:
+Language and Cognition}, \emph{15}(1), 157--166.
+
+\bibitem[\citeproctext]{ref-spivey1999cross}
+Spivey, M. J., \& Marian, V. (1999). Cross talk between native and
+second languages: Partial activation of an irrelevant lexicon.
+\emph{Psychological Science}, \emph{10}(3), 281--284.
+
+\bibitem[\citeproctext]{ref-takata1990english}
+Takata, Y., \& Nábělek, A. K. (1990). English consonant recognition in
+noise and in reverberation by japanese and american listeners. \emph{The
+Journal of the Acoustical Society of America}, \emph{88}(2), 663--666.
+
+\bibitem[\citeproctext]{ref-thierry2007brain}
+Thierry, G., \& Wu, Y. J. (2007). Brain potentials reveal unconscious
+translation during foreign-language comprehension. \emph{Proceedings of
+the National Academy of Sciences}, \emph{104}(30), 12530--12535.
+
+\bibitem[\citeproctext]{ref-valente2018does}
+Valente, D., Ferré, P., Soares, A., Rato, A., \& Comesaña, M. (2018).
+Does phonological overlap of cognate words modulate cognate acquisition
+and processing in developing and skilled readers? \emph{Language
+Acquisition}, \emph{25}(4), 438--453.
+
+\bibitem[\citeproctext]{ref-van2014stringdist}
+van der Loo, M. P. J. (2014). The stringdist package for approximate
+string matching. \emph{The {R} {J}ournal}, \emph{6}, 111--122.
+\url{https://CRAN.R-project.org/package=stringdist}
+
+\bibitem[\citeproctext]{ref-van1998orthographic}
+Van Heuven, W. J., Dijkstra, T., \& Grainger, J. (1998). Orthographic
+neighborhood effects in bilingual word recognition. \emph{Journal of
+Memory and Language}, \emph{39}(3), 458--483.
+
+\bibitem[\citeproctext]{ref-van2014subtlex}
+Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
+SUBTLEX-UK: A new and improved word frequency database for british
+english. \emph{Quarterly Journal of Experimental Psychology},
+\emph{67}(6), 1176--1190.
+
+\bibitem[\citeproctext]{ref-von2012language}
+Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access
+in bilingual toddlers. \emph{Journal of Experimental Child Psychology},
+\emph{113}(4), 569--586.
+
+\bibitem[\citeproctext]{ref-weber2004lexical}
+Weber, A., \& Cutler, A. (2004). Lexical competition in non-native
+spoken-word recognition. \emph{Journal of Memory and Language},
+\emph{50}(1), 1--25.
+
+\bibitem[\citeproctext]{ref-wickham2019tidyverse}
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
+Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
+Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
+Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
+\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
+
+\bibitem[\citeproctext]{ref-andras2022cognate}
+Andras, F., Rivera, M., Bajo, T., Dussias, P. E., \& Paolieri, D.
+(2022). Cognate facilitation effect during auditory comprehension of a
+second language: A visual world eye-tracking study. \emph{International
+Journal of Bilingualism}, 13670069211033359.
+
+\bibitem[\citeproctext]{ref-best2001discrimination}
+Best, C. T., McRoberts, G. W., \& Goodell, E. (2001). Discrimination of
+non-native consonant contrasts varying in perceptual assimilation to the
+listener's native phonological system. \emph{The Journal of the
+Acoustical Society of America}, \emph{109}(2), 775--794.
+
+\bibitem[\citeproctext]{ref-broersma2021praat}
+Broersma, P., \& Weenink, D. (2021). \emph{Praat: Doing phonetics by
+computer {[}computer program{]}} (Version 6.1.54).
+\url{http://www.praat.org/}
+
+\bibitem[\citeproctext]{ref-burkner2017brms}
+Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
+using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+
+\bibitem[\citeproctext]{ref-carpenter2017stan}
+Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B.,
+Betancourt, M., Brubaker, M. A., Guo, J., Li, P., \& Riddell, A. (2017).
+Stan: A probabilistic programming language. \emph{Grantee Submission},
+\emph{76}(1), 1--32.
+
+\bibitem[\citeproctext]{ref-chen1989patterns}
+Chen, H.-C., \& Leung, Y.-S. (1989). Patterns of lexical processing in a
+nonnative language. \emph{Journal of Experimental Psychology: Learning,
+Memory, and Cognition}, \emph{15}(2), 316.
+
+\bibitem[\citeproctext]{ref-christoffels2006memory}
+Christoffels, I. K., De Groot, A. M., \& Kroll, J. F. (2006). Memory and
+language skills in simultaneous interpreters: The role of expertise and
+language proficiency. \emph{Journal of Memory and Language},
+\emph{54}(3), 324--345.
+
+\bibitem[\citeproctext]{ref-costa2000cognate}
+Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate
+facilitation effect: Implications for models of lexical access.
+\emph{Journal of Experimental Psychology: Learning, Memory, and
+Cognition}, \emph{26}(5), 1283.
+
+\bibitem[\citeproctext]{ref-cutler2004patterns}
+Cutler, A., Weber, A., Smits, R., \& Cooper, N. (2004). Patterns of
+english phoneme confusions by native and non-native listeners. \emph{The
+Journal of the Acoustical Society of America}, \emph{116}(6),
+3668--3678.
+
+\bibitem[\citeproctext]{ref-de2000hard}
+De Groot, A. M., \& Keijzer, R. (2000). What is hard to learn is easy to
+forget: The roles of word concreteness, cognate status, and word
+frequency in foreign-language vocabulary learning and forgetting.
+\emph{Language Learning}, \emph{50}(1), 1--56.
+
+\bibitem[\citeproctext]{ref-desroches2022dynamics}
+Desroches, A. S., Friesen, D. C., Teles, M., Korade, C. A., \& Forest,
+E. W. (2022). The dynamics of spoken word recognition in bilinguals.
+\emph{Bilingualism: Language and Cognition}, \emph{25}(4), 705--710.
+
+\bibitem[\citeproctext]{ref-dijkstra2002architecture}
+Dijkstra, T., \& Van Heuven, W. J. (2002). The architecture of the
+bilingual word recognition system: From identification to decision.
+\emph{Bilingualism: Language and Cognition}, \emph{5}(3), 175--197.
+
+\bibitem[\citeproctext]{ref-dufour2003lexical}
+Dufour, S., \& Peereman, R. (2003). Lexical competition in phonological
+priming: Assessing the role of phonological match and mismatch lengths
+between primes and targets. \emph{Memory \& Cognition}, \emph{31},
+1271--1283.
+
+\bibitem[\citeproctext]{ref-elias2022cross}
+Elias, M., \& Degani, T. (2022). Cross-language interactions during
+novel word learning: The contribution of form similarity and participant
+characteristics. \emph{Bilingualism: Language and Cognition}, 1--18.
+
+\bibitem[\citeproctext]{ref-goldinger1989priming}
+Goldinger, S. D., Luce, P. A., \& Pisoni, D. B. (1989). Priming lexical
+neighbors of spoken words: Effects of competition and inhibition.
+\emph{Journal of Memory and Language}, \emph{28}(5), 501--518.
+
+\bibitem[\citeproctext]{ref-de1992determinants}
+Groot, A. M. de. (1992). Determinants of word translation. \emph{Journal
+of Experimental Psychology: Learning, Memory, and Cognition},
+\emph{18}(5), 1001.
+
+\bibitem[\citeproctext]{ref-degroot1994forward}
+Groot, A. M. de, Dannenburg, L., \& Vanhell, J. G. (1994). Forward and
+backward word translation by bilinguals. \emph{Journal of Memory and
+Language}, \emph{33}(5), 600--629.
+
+\bibitem[\citeproctext]{ref-hamburger1996phonological}
+Hamburger, M., \& Slowiaczek, L. M. (1996). Phonological priming
+reflects lexical competition. \emph{Psychonomic Bulletin \& Review},
+\emph{3}(4), 520--525.
+
+\bibitem[\citeproctext]{ref-kroll1994category}
+Kroll, J. F., \& Stewart, E. (1994). Category interference in
+translation and picture naming: Evidence for asymmetric connections
+between bilingual memory representations. \emph{Journal of Memory and
+Language}, \emph{33}(2), 149--174.
+
+\bibitem[\citeproctext]{ref-lecumberri2010non}
+Lecumberri, M. L. G., Cooke, M., \& Cutler, A. (2010). Non-native speech
+perception in adverse conditions: A review. \emph{Speech Communication},
+\emph{52}(11-12), 864--886.
+
+\bibitem[\citeproctext]{ref-levenshtein1966binary}
+Levenshtein, V. I. et al. (1966). Binary codes capable of correcting
+deletions, insertions, and reversals. \emph{Soviet Physics Doklady},
+\emph{10}, 707--710.
+
+\bibitem[\citeproctext]{ref-lotto1998effects}
+Lotto, L., \& De Groot, A. M. (1998). Effects of learning method and
+word type on acquiring vocabulary in an unfamiliar language.
+\emph{Language Learning}, \emph{48}(1), 31--69.
+
+\bibitem[\citeproctext]{ref-luce1998recognizing}
+Luce, P. A., \& Pisoni, D. B. (1998). Recognizing spoken words: The
+neighborhood activation model. \emph{Ear and Hearing}, \emph{19}(1), 1.
+
+\bibitem[\citeproctext]{ref-luce1990similarity}
+Luce, P. A., Pisoni, D. B., \& Goldinger, S. D. (1990). \emph{Similarity
+neighborhoods of spoken words.}
+
+\bibitem[\citeproctext]{ref-marian2012clearpond}
+Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012).
+\emph{CLEARPOND: Cross-linguistic easy-access resource for phonological
+and orthographic neighborhood densities}.
+
+\bibitem[\citeproctext]{ref-mcclelland1981interactive}
+McClelland, J. L., \& Rumelhart, D. E. (1981). An interactive activation
+model of context effects in letter perception: I. An account of basic
+findings. \emph{Psychological Review}, \emph{88}(5), 375.
+
+\bibitem[\citeproctext]{ref-meyer2007activation}
+Meyer, A. S., \& Damian, M. F. (2007). Activation of distractor names in
+the picture-picture interference paradigm. \emph{Memory \& Cognition},
+\emph{35}(3), 494--503.
+
+\bibitem[\citeproctext]{ref-midgley2011effects}
+Midgley, K. J., Holcomb, P. J., \& Grainger, J. (2011). Effects of
+cognate status on word comprehension in second language learners: An ERP
+investigation. \emph{Journal of Cognitive Neuroscience}, \emph{23}(7),
+1634--1647.
+
+\bibitem[\citeproctext]{ref-peirce2019psychopy2}
+Peirce, J., Gray, J. R., Simpson, S., MacAskill, M., Höchenberger, R.,
+Sogo, H., Kastman, E., \& Lindeløv, J. K. (2019). PsychoPy2: Experiments
+in behavior made easy. \emph{Behavior Research Methods}, \emph{51}(1),
+195--203.
+
+\bibitem[\citeproctext]{ref-phillips2006erp}
+Phillips, N. A., Klein, D., Mercier, J., \& Boysson, C. de. (2006). ERP
+measures of auditory word repetition and translation priming in
+bilinguals. \emph{Brain Research}, \emph{1125}(1), 116--131.
+
+\bibitem[\citeproctext]{ref-potter1984lexical}
+Potter, M. C., So, K.-F., Von Eckardt, B., \& Feldman, L. B. (1984).
+Lexical and conceptual representation in beginning and proficient
+bilinguals. \emph{Journal of Verbal Learning and Verbal Behavior},
+\emph{23}(1), 23--38.
+
+\bibitem[\citeproctext]{ref-rcore2019r}
+RCore, T. (2019). \emph{R: A language and environment for statistical
+computing. R foundation for statistical computing, austria}.
+
+\bibitem[\citeproctext]{ref-schepens2012distributions}
+Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of
+cognates in europe as based on levenshtein distance. \emph{Bilingualism:
+Language and Cognition}, \emph{15}(1), 157--166.
+
+\bibitem[\citeproctext]{ref-spivey1999cross}
+Spivey, M. J., \& Marian, V. (1999). Cross talk between native and
+second languages: Partial activation of an irrelevant lexicon.
+\emph{Psychological Science}, \emph{10}(3), 281--284.
+
+\bibitem[\citeproctext]{ref-takata1990english}
+Takata, Y., \& Nábělek, A. K. (1990). English consonant recognition in
+noise and in reverberation by japanese and american listeners. \emph{The
+Journal of the Acoustical Society of America}, \emph{88}(2), 663--666.
+
+\bibitem[\citeproctext]{ref-thierry2007brain}
+Thierry, G., \& Wu, Y. J. (2007). Brain potentials reveal unconscious
+translation during foreign-language comprehension. \emph{Proceedings of
+the National Academy of Sciences}, \emph{104}(30), 12530--12535.
+
+\bibitem[\citeproctext]{ref-valente2018does}
+Valente, D., Ferré, P., Soares, A., Rato, A., \& Comesaña, M. (2018).
+Does phonological overlap of cognate words modulate cognate acquisition
+and processing in developing and skilled readers? \emph{Language
+Acquisition}, \emph{25}(4), 438--453.
+
+\bibitem[\citeproctext]{ref-van2014stringdist}
+van der Loo, M. P. J. (2014). The stringdist package for approximate
+string matching. \emph{The {R} {J}ournal}, \emph{6}, 111--122.
+\url{https://CRAN.R-project.org/package=stringdist}
+
+\bibitem[\citeproctext]{ref-van1998orthographic}
+Van Heuven, W. J., Dijkstra, T., \& Grainger, J. (1998). Orthographic
+neighborhood effects in bilingual word recognition. \emph{Journal of
+Memory and Language}, \emph{39}(3), 458--483.
+
+\bibitem[\citeproctext]{ref-van2014subtlex}
+Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
+SUBTLEX-UK: A new and improved word frequency database for british
+english. \emph{Quarterly Journal of Experimental Psychology},
+\emph{67}(6), 1176--1190.
+
+\bibitem[\citeproctext]{ref-von2012language}
+Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access
+in bilingual toddlers. \emph{Journal of Experimental Child Psychology},
+\emph{113}(4), 569--586.
+
+\bibitem[\citeproctext]{ref-weber2004lexical}
+Weber, A., \& Cutler, A. (2004). Lexical competition in non-native
+spoken-word recognition. \emph{Journal of Memory and Language},
+\emph{50}(1), 1--25.
+
+\bibitem[\citeproctext]{ref-wickham2019tidyverse}
+Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
+François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
+Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
+Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
+Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
+\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
 
@@
 
-\endgroup
-
-\newpage
-
 \end{document}

### 2024-05-06 manuscript/test.qmd
Manuscript

--- 2024-05-02 manuscript/manuscript.qmd
@@
 title: "The role of cognateness in non-native spoken word recognition"
-
-abstract: |
-    Understanding spoken words in a non-native language is easier when the words are phonologically similar to their translations into the listener's native language are phonologically similar.
-Phonological similarity plays a more important role in the language performance of low-proficiency bilinguals, compared to that of high-proficiency bilinguals., and This suggests that low-proficiency bilinguals may points to the latter as relying more strongly on cross-language word-to-word links (lexical route), while high-proficiency bilinguals rely more on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
-Disentangling bilinguals’ reliance on the conceptual and lexical and conceptual routes during translation is challenging, given that both sources of information are present during spoken word comprehensionto some extent in both high- and low-proficiency bilinguals.
-In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
-Given their lack of previous knowledge ion the unfamiliar tested language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
-This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route.
-Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation.
-These results point to cross-language word-to-word links as providing rich information during word translation.
+link-citations: true
+editor: source
+lang: en
+freeze: true
+fig-dpi: 1000
+echo: false
+warning: false
+format: pdf
+csl: "_assets/apa7.csl"
+bibliography: "_assets/references.bib"
 ---
@@
 tar_load(exp_participants)
+tar_load(exp_responses)
+tar_load(quest_responses)
 tar_load(quest_participants)
 tar_load(stimuli)
+tar_load(epreds)
+tar_load(dataset_1)
+tar_load(dataset_2)
+tar_load(dataset_3)
+tar_load(exp_1_m0)
+tar_load(exp_2_m0)
+tar_load(exp_3_m0)
+tar_load(exp_1_loos)
+tar_load(exp_2_loos_1)
 
-theme_set(theme_ggdist())
+theme_set(theme_ggdist() + 
+              theme(panel.grid.major.y = element_line(colour = "grey",
+                                                      linetype = "dotted")))
 
@@
 However, there are some sources of information which may make non-native speech comprehension easier.
-One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener’s native language.
+One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
 For example, imagine a Spanish native, with no previous familiarity with any other language, who listens for the first time to French may encounter the word /`r knitr::asis_output("\U0070\U0254\U0281\U0074")`/ (*porte*), translation of door in French.
@@
 
-Early theoretical accounts of this 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals’ proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
+Early theoretical accounts of this 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
 The reasoning behind this claim was the assumption that second-language learners start acquiring words in their L2 by associating them to their translation in their L1 (lexical route), instead of directly to their shared concept (conceptual route).
@@
 The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 should take longer than translating from L2 to L1.
-The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, whereas L2-to-L1 translation relies more strongly on the mediation between the concept and the two word- forms.
+The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, where L2-to-L1 translation relies more strongly on the mediation between the concept and the two word- forms.
 L1-to-L2 translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during L1-to-L2 translation.
@@
 To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate Dutch words to English (L1-to-L2) or English words to Dutch (L2-to-L1).
-Although participants’ performance in both conditions was roughly equivalent for cognates, non-cognates were translated more slowly from Dutch to English than from English to Dutch, suggesting that L1-to-L2 translation is more sensitive to form-similarity than L2-to-L1 translation.
-Conversely, participants’ performance was more sensitive to semantic variables (e.g., concreteness) when translating English words to Dutch, compared to when translating Dutch words to English, suggesting that L2-to-L1 translation was more sensitive to the shared conceptual properties of the translation pair.
+Although participants' performance in both conditions was roughly equivalent for cognates, non-cognates were translated more slowly from Dutch to English than from English to Dutch, suggesting that L1-to-L2 translation is more sensitive to form-similarity than L2-to-L1 translation.
+Conversely, participants' performance was more sensitive to semantic variables (e.g., concreteness) when translating English words to Dutch, compared to when translating Dutch words to English, suggesting that L2-to-L1 translation was more sensitive to the shared conceptual properties of the translation pair.
 
@@
 For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
-Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when these neighbours’ lexical frequency is higher [@luce1998recognizing].
+Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when these neighbours' lexical frequency is higher [@luce1998recognizing].
 
@@
 @meyer2007activation showed that target pictures were named faster when presented together with a phonologically-related distractor, suggesting that the activation of shared phonological information allowed the target to be activated and produced faster.
-There are also findings suggesting that facilitatory or inhibitory effects on target recognition depend on the degree of similarity the competitor shares with the target [@hamburger1996phonological; @dufour2003lexical].
+there are also findings suggesting that facilitatory or inhibitory effects on target recognition depend on the degree of similarity the competitor shares with the target [@hamburger1996phonological; @dufour2003lexical].
 When there is only mismatch of one phoneme between prime and target, inhibitory effects are exerted on target recognition [@dufour2003lexical].
@@
 This highlights that lexical competition for bilinguals during language comprehension may not only come from the tested language but also the native language, despite the native language not being actively required for the task.
-Other studies have similarly found interference when spoken target words in one language had phonological overlap with a word in the participant’s other language, even when the other language was not required by the task [@spivey1999cross; @desroches2022dynamics].
+Other studies have similarly found interference when spoken target words in one language had phonological overlap with a word in the participant's other language, even when the other language was not required by the task [@spivey1999cross; @desroches2022dynamics].
 Such patterns have even been found with young children less than 4 years old [@von2012language].
@@
 This is, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
-Evidence supporting such an integrated lexicon is provided by the fact that bilinguals’ performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word’s translation.
+Evidence supporting such an integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word's translation.
 When bilingual participants are presented with words in one language in a lexical decision task, translating them to the other language took longer when they were part of large neighbourhoods [@van1998orthographic].
@@
 
-In the present study, we tested listeners’ reliance on phonological similarity when translating words from an unfamiliar language to their native language, in order to investigate the plausibility of word-to-word connections during L1-to-L2 translation.
+In the present study, we tested listeners' reliance on phonological similarity when translating words from an unfamiliar language to their native language, in order to investigate the plausibility of word-to-word connections during L1-to-L2 translation.
 We tested English native speakers and Spanish native speakers in a translation elicitation task.
@@
 If this is the case, non-native words that have more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency and are phonologically closer to the presented word than the target translation.
-For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as bottom instead of as its correct translation *button* (because the lexical frequency of bottom is higher than that of button and also phonological closer).
+For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as bottom instead of as its correct translation *button* (because the lexical frequency of *bottom* is higher than that of *button* and also phonological closer).
 This combined constraint led us to calculate a specific form of phonological neighbours where a neighbour is counted only if it is higher frequency and has stronger phonological similarity than the target translation as judged by Levenshtein distance.
-If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
+If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
 
-We tested three groups of participants in this translation elicitation task: English natives translating Spanish words, English natives translating Catalan words, and Spanish natives translating Catalan words.
-Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, whereas English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+In Experiment 1, we collected data from two groups of British English natives.
+One group was presented with Catalan words, the other with Spanish words.
+We examined the extent to which participants were able to use the phonological similarity between the presented (Catalan or Spanish) word and its translation (i.e., cognateness) to provide accurate responses.
+In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
+Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
 By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
-
-Following @luce1998recognizing's NAM model, we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
-If this is the case,  has more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency than the target translation.
-For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button* (because the lexical frequency of *bottom* is higher than that of *button*).
-If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
+Participants in Experiments 1 and 2 were surprisingly good at translation some words, which had low phonological similarity with their correct translation.
+In Experiment 3, we collected additional data from a new group of British English natives, who were presented with Catalan and Spanish words.
+In contrast with the procedure in Experiment 1, participants in Experiment 3 where asked, after providing a response in each trial, to report whether they had previous knowledge of the meaning of the presented word (e.g., some Spanish words are common knowledge, even for monolingual speakers of English).
+Finally, we present analyses on the joint data sets of Experiments 1 to 3.
 
@@
 
-```{r}
-#| label: participants
+```{r tbl-participants}
+#| label: tbl-participants
+#| eval: false
 tbl_data <- participants |> 
@@
 excluded_ntrials <- participants_1 |> 
-    filter(invalid_participant_trials) |> 
+    filter(valid_status=="Insufficient trials") |> 
     nrow()
 ```
+
 Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
@@
 names(durations) <- janitor::make_clean_names(names(durations))
-
 ```
-
-#### Presented stimuli
 
@@
 names(stim_exc_freq) <- janitor::make_clean_names(names(stim_exc_freq))
+
+stim_exc_freq_n <- sum(map_int(stim_exc_freq, "n"))
+```
+
+For each translation pair, we defined three predictors of interest: the lexical frequency of the correct translation (*Frequency*), the phonological similarity between the presented (Catalan or Spanish) word and their correct English translation (*Cognateness*), and presented word's number of cross-language phonological neighbours (*CLPN*).
+
+We included *Frequency* under the hypothesis that, keeping other predictors constant, participants would translate higher-frequency words more accurately and faster than lower-frequency words.
+Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores [@van2014subtlex].
+Words in the stimuli list without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
+
+We calculated *Cognatenesss*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
+The Levenshtein distance computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
+We divided this edit distance by the number of characters included in the longest X-SAMPA transcription of the translation pair.
+This results in a proportion score, in which values closer to zero correspond to lower Levenshtein distances between phonological transcriptions (i.e., higher similarity), and values closer to 1 correspond to higher Levenshtein distances (i.e., lower similarity).
+This transformation also help account for the fact that the Levenshtein distance tends to increase with the length of the transcriptions.
+For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+
+For each Catalan and Spanish word, we calculated the number of *CLPN* by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) different in up to one phoneme from that of the presented Catalan or Spanish word.
+Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological trnascriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
+
+```{r tbl-stimuli}
+#| label: tbl-stimuli
+#| eval: false
+stimuli |> 
+    summarise(across(c(freq_zipf_2, lv, neigh_n_h),
+                     lst(mean, median, sd, min, max), 
+                     na.rm = TRUE),
+              .by = c(group)) |> 
+    gt(rowname_col = "group") |> 
+    tab_spanner("Frequency", matches("freq")) |> 
+    tab_spanner("Cognateness", matches("lv")) |> 
+    tab_spanner("CLPN", matches("neigh")) |> 
+    fmt_number(is.numeric) |> 
+    fmt_integer(c(neigh_n_h_median,
+                  neigh_n_h_min,
+                  neigh_n_h_max)) |> 
+    cols_merge_uncert(freq_zipf_2_mean, freq_zipf_2_sd) |> 
+    cols_merge_uncert(lv_mean, lv_sd) |> 
+    cols_merge_uncert(neigh_n_h_mean, neigh_n_h_sd) |> 
+    cols_merge_range(freq_zipf_2_min, freq_zipf_2_max) |> 
+    cols_merge_range(lv_min, lv_max) |> 
+    cols_merge_range(neigh_n_h_min, neigh_n_h_max) |> 
+    cols_label(freq_zipf_2_mean = "Mean",
+               freq_zipf_2_sd = "SD",
+               freq_zipf_2_median = "Median",
+               freq_zipf_2_min = "Range",
+               lv_mean = "Mean",
+               lv_sd = "SD",
+               lv_median = "Median",
+               lv_min = "Range",
+               neigh_n_h_mean = "Mean",
+               neigh_n_h_sd = "SD",
+               neigh_n_h_median = "Median",
+               neigh_n_h_min = "Range") |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels())
+
 ```
@@
 After one second, the audio started playing while the dot remained being displayed until the audio ended.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
+Upon the offset of the fixation point and audio, participants were prompted to write their answer by a `>` symbol.
 Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
 Participants were allowed to correct their answer.
-Then, participants pressed the RETURN/ENTER key to confirm their answer and start and new trial.
+Then, participants pressed the `RETURN/ENTER` key to confirm their answer and start and new trial.
 We excluded trials where the participant's answer was not a valid word in the target language or no answer was given.
@@
 #| warning: false
+#| eval: false
 #| out-width: "80%"
 #| fig-cap: "Schematic representation of a trial in the experimental task."
-include_graphics(here("img", "design.png"))
+knitr::include_graphics(here::here("img", "design.png"))
 ```
@@
 ```{r}
+#| label: tbl-results
+#| eval: false
 tbl_data <- list("Experiment 1" = dataset_1,
@@
 
-#### Predictors
+We modelled the probability of participants guessing the correct translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We included as fixed effects the intercept, the main effects of *Frequency*, *Cognateness*, and *CLPN*, and the two-way interaction between *Cognateness* and *CLPN*.
+We also included participant-level random intercepts and slopes for the the main effects and the interaction.
+@eq-model shows a formal description of the model.
 
-We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their X-SAMPA transcriptions using the `stringsim` function of the `stringdist` R package [@van2014stringdist].
-This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
-First, it computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
-This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is a percentage score, with 0% indicating no similarity between the two strings, and 100% indicating a perfect match between the two strings.
-We computed this similarity measure (referred simply as Levenshtein, from here on) for every translation pair in our stimuli lists.
-For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
-Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+To test the practical relevance of each predictor we followed [@krushke2018bayesian].
+We first specified a region of practical equivalence (*ROPE*) around zero.
+This area indicates the values of the regression coefficients that we considered as equivalent to zero.
+We then computed the 95% posterior credible intervals (CrI) of the regression coefficients of interest.
+Finally, we calculated the proportion of the 95% CrI that fell inside the ROPE.
+This proportion indicates the probability that the true value of the coefficient is equivalent to zero.
 
-The lexical frequencies in this database were extracted from SUBTLEX-UK [@van2014subtlex].
-We transformed lexical frequency scores to Zipf scores [@van2014subtlex].
-Word in the stimuli list without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
-For each presented Catalan or Spanish word presented in each list, we calculated its number of higher-frequency cross-language phonological neighbours (CLPN).
-This measure was calculated by counting the number of higher-frequency English words whose phonological word-form (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differed by one phoneme from that of the presented Catalan or Spanish word.
-Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond].
-
-#### Modelling approach
-
-We modelled the probability of participants guessing the correct target translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0).
-Secondly, we added the fixed effect of lexical frequency of the target translation ($Frequency$).
-We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels cat-ENG, spa-ENG, and cat-SPA), and finally, the $Levenshtein \times Group$ interaction).
-We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) [@schad2020capitalize].
-Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) [@barr2013random].
-We will later refer to these models as Model 0 to Model 6.
-For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
-
-$$\begin{align}
-
-&\textbf{Likelihood}  \\
-y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
-
-&\textbf{Parameters}  \\
-
-logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
-& \beta_{1[p]} ~ Frequency_{i} ~ + \\
-& \beta_{2[p]} ~ PTHN_i ~ + \\
-& \beta_{3[p]} ~ Similarity_i ~ + \\
-& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
-
-\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
-\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
-&& \text{[participant-level coefficients]} \\ \\
-
-&\textbf{Prior}  \\
-
-\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
-\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
-\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
-
-\end{align}$$
-
-#### Statistical inference
-
-To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (*LOO*) [@vehtari2017practical].
-More information about the models and model comparison can be found in Appendix 2.
 All analyses were performed in R environment [@rcore2019r].
 We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
-We used the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
 
@@
 ```{r}
+c1 <- gather_draws(exp_1_m0, `b_.*`, `sd_.*`, regex = TRUE) |> 
+    summarise(across(.value, lst(.median = median, 
+                                 .lower = \(x) hdi(x)[1],
+                                 .upper = \(x) hdi(x)[2],
+                                 .rope = \(x) mean(between(x, -0.1, 0.1))),
+                     .names = "{fn}"))
+c1 <- split(c1, c1$.variable)
+names(c1) <- janitor::make_clean_names(names(c1)) 
+```
+
+@tbl-results-1 shows a summary of participants' accuracy across Experiments 1, 2, and 3.
+Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Cognateness* and *CLPN* ($\beta$ = `r round(c1$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_lv$.lower, 3)`, `r round(c1$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_lv$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Cognateness* ($\beta$ = `r round(c1$b_lv$.median, 3)`, 95% CrI = [`r round(c1$b_lv$.lower, 3)`, `r round(c1$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h$.lower, 3)`, `r round(c1$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h$.rope, 3)`).
+A model including *Group* as a main effect did not improve the predictive performance of the model ($\Delta_{ELPD}$ = `r round(exp_1_loos[2], 2)`, $SE$ = `r round(exp_1_loos[4], 2)`), indicating that participants in both groups performed equivalently.
+@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Cognateness* and *CLPN*.
+
+```{r tbl-dataset}
 #| label: tbl-dataset
 sem <- function(x) mean(x) / (sqrt(length(x)))
+# proportion adjusted from boundary values (Gelman, Hill & Vehtari, 2020)
+prop_adj <- function(x, n){
+    e <- (x+2)/(n+4)
+    return(e)
+}
+
+# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+prop_adj_se <- function(x, n) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    return(se)
+}
+
+# adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+prop_adj_ci <- function(x, n, .width = 0.95) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    ci <-  e + qnorm(c((1-.width)/2, (1-(1-.width)/2)))*se
+    ci[1] <- ifelse(ci[1]<0, 0, ci[1]) # truncate at 0
+    ci[2] <- ifelse(ci[2]>1, 1, ci[2]) # truncate at 1
+    return(ci)
+}
 
@@
     fmt_number(is.numeric, decimals = 2) |> 
-    gtExtras::gt_plt_dist(correct_list, type = "density") |> 
+    #gtExtras::gt_plt_dist(correct_list, type = "density") |> 
     fmt_integer(c(matches("sum|min|max"), n_participants)) |> 
@@
 ```
-
-All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
-Model 6, which included all main effects and the two-way interactions between $PTHN$ and $Levenshtein$, and between $Group$ and $Levenshtein$ and $PTHN$ and consonant similarity, showed the best performance.
-
-```{r}
-#| label: model-comparison
-#| message: false
-#| warning: false
-tab_loos <- model_loos %>% 
-    as.data.frame() %>% 
-    rownames_to_column("model") %>% 
-    arrange(desc(elpd_loo)) %>% 
-    select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
-    mutate(model = str_replace(model, "fit_", "Model ")) %>% 
-    mutate_all(~ifelse(. == 0, NA, .)) %>% 
-    gt(rowname_col = "model", id = "model-comparison") %>% 
-    tab_header(
-        title = md("Table 2"),
-        subtitle = md("Model performance and comparison using leave-one-out cross-validation")
-    ) %>% 
-    tab_source_note(
-        md("*Note*.
-Outcomes of the leave-one-out cross-validation procedure.
-Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
-Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
-    ) %>% 
-    fmt_missing(everything(), missing_text = "-") %>% 
-    fmt_number(2:7) %>% 
-    cols_label(
-        model = md("**Model**"),
-        looic = md("***LOO<sub>IC</sub>***"),
-        se_looic = md("***SE*<sub>IC</sub>**"),
-        elpd_loo = md("***LOO<sub>ELPD</sub>***"),
-        se_elpd_loo = md("***SE***"),
-        elpd_diff = md("***LOO<sub>diff</sub>***"),
-        se_diff = md("***SE<sub>diff</sub>***")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", weight = "bold", size = "medium"),
-        locations = cells_title(groups = "title")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", size = "medium", style = "italic"),
-        locations = cells_title(groups = "subtitle")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", size = "medium"),
-        locations = cells_source_notes()
-    ) %>% 
-    tab_style(
-        style = list(
-            cell_borders(sides = c("left", "right"), color = "white"),
-            cell_text(align = "center")
-        ),
-        locations = list(
-            cells_body(), 
-            cells_row_groups(), 
-            cells_column_spanners(), 
-            cells_stub(), 
-            cells_source_notes(),
-            cells_column_labels(), 
-            cells_stubhead()
-        )
-    ) %>% 
-    tab_style(
-        style = list(
-            cell_text(align = "left", size = "medium"), 
-            cell_borders(sides = "all", color = "white")
-        ),
-        locations = list(cells_stub(), cells_source_notes())
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-        locations = list(cells_column_spanners(), cells_column_labels())
-    ) %>% 
-    tab_style(
-        style = cell_borders(sides = "all", color = "white"),
-        locations = cells_body(columns = 1:7)
-    ) %>% 
-    tab_style(
-        style = cell_borders(sides = "top", color = "white"),
-        locations = cells_body(columns = 1:7)
-    ) %>% 
-    gtsave(here("img", "loos.png"))
-
-include_graphics(here("img", "loos.png"))
-
-```
-
-We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
-For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
-Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increments or decrements in the probability of correct translation [@gelman2020regression].
-
-```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of population-level regression coefficients in Model 4.
-Distributions indicate the estimated posterior likelihood density of regression coefficients.
-Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
-Black dots represent the mean of each distribution.", fig.height=4.5, fig.width=6.5}
-
-coefs <- fixef(fit_6) %>% 
-    as.data.frame() %>% 
-    rownames_to_column("variable") %>% 
-    clean_names() %>% 
-    mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
-    group_split(variable) %>% 
-    set_names(make_clean_names(map(., "variable"))) %>% 
-    map(select, -variable) %>% 
-    map(unlist)
-
-str_repl <- c(
-    "Intercept" = "Intercept",
-    "frequency_zipf_std" = "Frequency\n[+1 SD]",
-    "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-    "lv_std:groupeng_spa" = "Levenshtein \u00d7 Group 1",
-    "lv_std:groupcateng_spaeng" = "Levenshtein \u00d7 Group 2",
-    "lv_std" = "Levenshtein\n[+1 SD]",
-    "pthn_std" = "PTHN\n[+1 SD]",
-    "groupeng_spa" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-    "groupcateng_spaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
-)
-
-post_draws <- gather_draws(
-    fit_6, 
-    `b_.*`, 
-    `sd_.*`, 
-    regex = TRUE
-)
-
-# fixed effects
-post_draws %>% 
-    filter(str_detect(.variable, "b_")) %>% 
-    mutate(
-        .variable_name = str_remove(.variable, "b_") %>% 
-            str_replace_all(str_repl) %>%  
-            factor(
-                levels = c(
-                    "Intercept", 
-                    "Frequency\n[+1 SD]",
-                    "PTHN\n[+1 SD]", 
-                    "Levenshtein\n[+1 SD]",
-                    "PTHN \u00d7 Levenshtein",
-                    "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-                    "Group 2\n[cat-ENG vs. spa-ENG]",
-                    "Levenshtein \u00d7 Group 1", 
-                    "Levenshtein \u00d7 Group 2"
-                )
-            ),
-        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
-    ) %>% 
-    arrange(.variable) %>% 
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    geom_vline(xintercept = 0) +
-    stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-    stat_pointinterval(.width = 0, point_size = 1) +
-    scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
-    scale_x_continuous(labels = function(x) percent(round(x, 1))) +
-    labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
-    theme(
-        legend.position = c(1, 0.25),
-        legend.direction = "vertical",
-        legend.background = element_rect(fill = "grey95"),
-        legend.key = element_rect(fill = "grey95", colour = "grey95"), 
-        legend.key.height = unit(0.1, "cm"),
-        axis.title.y = element_blank(),
-        legend.text = element_text(size = 7),
-        axis.text.x = element_text(colour = "black"),
-        axis.text.y = element_text(colour = "black", hjust = 1),
-        axis.ticks.y = element_blank(),
-        panel.grid.major.y = element_line(
-            colour = "grey", 
-            size = 0.5
-        )
-    )
-
-```
-
-```{r}
-#| label: bayes-factors
-#| message: false
-#| warning: false
-# wrapper to extract Bayes Factors
-get_bf <- function(hyp, abs = TRUE, ...){
-    bf <- hypothesis(fit_6, hypothesis = hyp, ...)$hypothesis$Evid.Ratio
-    if (abs) bf <- abs(bf)
-    return(bf)
-}
-
-h_frequency <- get_bf("frequency_zipf_std = 0")
-h_pthn <- get_bf("pthn_std = 0")
-h_lv <- get_bf("pthn_std = 0")
-h_pthn_lv <- get_bf("pthn_std:lv_std = 0")
-h_group1 <- get_bf("groupeng_spa = 0")
-h_group2 <- get_bf("groupcateng_spaeng = 0")
-h_lv_group1 <- get_bf("lv_std:groupeng_spa = 0")
-h_lv_group2 <- get_bf("lv_std:groupcateng_spaeng = 0")
-
-icc_conditional <- icc(fit_6, by_group = TRUE)$ICC[2]
-
-sds <- gather_draws(
-    fit_6,
-    `sd_.*`, 
-    regex = TRUE
-) %>% 
-    mean_qi() %>% 
-    mutate(
-        .value = ifelse(
-            grepl("Intercept", .variable), 
-            inv_logit_scaled(.value), 
-            .value/4
-        ),
-        .lower = ifelse(
-            grepl("Intercept", .variable),
-            inv_logit_scaled(.lower),
-            .lower/4
-        ),
-        .upper = ifelse(
-            grepl("Intercept", .variable),
-            inv_logit_scaled(.upper), 
-            .upper/4
-        )
-    ) %>% 
-    split(.$.variable)
-
-names(sds) <- make_clean_names(names(sds))
-
-word_example <- ranef(fit_6)$translation_id[,,"Intercept"] %>% 
-    as.data.frame() %>% 
-    rownames_to_column("word") %>% 
-    arrange(-Estimate) %>% 
-    clean_names() %>% 
-    mutate_if(is.numeric, inv_logit_scaled) %>% 
-    slice(1, n()) %>% 
-    left_join(
-        distinct(stimuli, word1, ipa1), 
-        by = c("word" = "word1")
-    ) %>% 
-    split(.$word)
-
-participant_ranef_range <- sds %>% 
-    bind_rows() %>% 
-    filter(!grepl("Intercept|word|group1", .variable)) %>% 
-    summarise(.value = range(.value)) %>% 
-    unlist()
-
-cors_range <- gather_draws(fit_6, `cor_.*`, regex = TRUE) %>% 
-    median_qi() %>% 
-    summarise(.value = range(.value))
-
-corr_mat <- VarCorr(fit_6, summary = TRUE)$participant_id$cor 
-corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant_id$cor)
-
-corr_data <- corr_mat %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(term1 = corr_names) %>% 
-    relocate(term1) %>% 
-    split(.$term1)
-
-names(corr_data) <- make_clean_names(names(corr_data))
-
-h_cor_intercept_frequency <- get_bf("participant_id__Intercept__frequency_zipf_std = 0", class = "cor")
-h_cor_frequency_zipf_lv_group1 <- get_bf("participant_id__frequency_zipf_std__lv_std:groupeng_spa = 0", class = "cor")
-h_cor_intercept_pthn <- get_bf("participant_id__Intercept__pthn_std = 0", class = "cor")
-```
-
-Overall, participants had `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likelihood to produce correct translations.
-The lexical frequency of the target translation had little impact on accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf_std["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
-The number of phonological neighbours with higher lexical frequency than the target translation (*PTHN*) slightly decreased the probability of a correct responses by `r percent(coefs$pthn_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
-Finally, phonological similarity between the presented word and target translation had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
-The interaction effect between $PTHN \times Levenshtein$ was small: a +1 SD increment in Levenshtein increased the effect of *PTHN* by `r percent(coefs$pthn_std_lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std_lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std_lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
-
-Spanish natives translating Catalan words were `r percent(coefs$groupeng_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$groupeng_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$groupeng_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
-The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in `r percent(coefs$lv_std_groupeng_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_groupeng_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_groupeng_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-English natives translating Catalan words were `r percent(abs(coefs$groupcateng_spaeng["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$groupcateng_spaeng["q2_5"], accuracy = 0.01)`, `r percent(coefs$groupcateng_spaeng["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
-For every one *SD* increment in Levenshtein similarity, the effect of presented language increased in `r percent(abs(coefs$lv_std_groupcateng_spaeng["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_groupcateng_spaeng["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_groupcateng_spaeng["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-
-```{r}
-#| label: marginal-effects
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 3.5
-#| fig-cap: "Expected mean posterior predictions from Model 4.
-A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein).
-We did this across the range of values of the PTHN scores.
-For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
-Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
-The dispersion of the lines indicates the uncertainty of our predictions.
-We computed these posterior predictions for each group of participants, and plotted them in separate panels."
-nd <- expand.grid(
-    lv_std = c(min(fit_6$data$lv_std), 0, max(fit_6$data$lv_std)),
-    frequency_zipf_std = 0,
-    pthn_std = seq(
-        min(fit_6$data$pthn_std, na.rm = TRUE),
-        max(fit_6$data$pthn_std, na.rm = TRUE),
-        by = 0.1
-    ),
-    group = unique(fit_6$data$group)
-)
-
-m <- add_epred_draws(nd, fit_6, ndraws = 100, re_formula = NA) %>% 
-    mutate(
-        lv_std = factor(
-            lv_std,
-            levels = c(
-                min(fit_6$data$lv_std),
-                0,
-                max(fit_6$data$lv_std)
-            ),
-            labels = c(
-                paste0("Min (", percent(min(responses$lv, na.rm = TRUE)), ")"),
-                paste0("Mean (", percent(mean(responses$lv, na.rm = TRUE)), ")"),
-                paste0("Max (", percent(max(responses$lv, na.rm = TRUE)), ")")
-            )
-        ) 
-    )
-
-ggplot(m) +
-    aes(x = pthn_std, y = .epred, colour = lv_std, fill = lv_std) +
-    facet_wrap(~group) +
-    geom_hline(yintercept = 0.5, colour = "grey") +
-    geom_line(aes(group = interaction(lv_std, .draw)), size = 0.65, alpha = 0.5, show.legend = TRUE) +
-    stat_summary(
-        aes(linetype = lv_std), 
-        fun = mean, 
-        geom = "line",
-        colour = "black", 
-        size = 0.75,
-        show.legend = TRUE
-    ) +
-    labs(
-        x = paste0(
-            "PTHN (+1 SD, ", 
-            round(sd(responses$pthn), 1), 
-            " higher-frequency neighbours)"
-        ),
-        y = "P(Correct)", 
-        colour = "Levenshtein similarity",
-        fill = "Levenshtein similarity",
-        linetype = "Levenshtein similarity",
-        title = "Population-level posterior predictions"
-    ) +
-    scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A")) +
-    scale_linetype_manual(values = c("solid", "dashed", "dotdash")) +
-    scale_y_continuous(
-        labels = function(x) percent(round(x, 2)),
-        limits = c(0, 1),
-        breaks = seq(0, 1, 0.1)
-    ) +
-    theme(
-        legend.position = "top",
-        plot.title = element_blank()
-    )
-```
-
-# Discussion
-
-# References

### 2024-05-06 R/models.R

@@ -43,12 +43,12 @@ get_epreds <- function(model, data, n = 100,
     
     knowledge <- unique(model$data$knowledge)
     confidence <- unique(model$data$confidence)
+    freq_zipf_2 <- mean(data$freq_zipf_2, na.rm = TRUE)
     group <- unique(model$data$group)
-    nd <- expand_grid(neigh_n_h, lv, knowledge, confidence, group)
+    nd <- expand_grid(freq_zipf_2, neigh_n_h, lv, knowledge, confidence, group)
     epreds <- tidybayes::add_epred_draws(nd, model, re_formula = NA, ...)
     
     return(epreds)
-    
 }
 
 

### 2024-05-06 R/questionnaire.R

@@ -170,7 +170,7 @@ get_quest_responses <- function(quest_processed, quest_participants, stimuli) {
   valid_codes <- c("correct", "typo", "wrong", "false_friend")
   correct_codes <- c("correct", "typo")
   
-  responses_coded_path <- file.path("data-raw", "questionnaire", "04_responses-coded.csv")
+  responses_coded_path <- file.path("data", "questionnaire-manual-coded.csv")
   
   quest_responses <- read_csv(responses_coded_path, col_types = "cccclicc") |>  
     rename(participant_id = participant) |> 

### 2024-05-08 manuscript/manuscript.qmd
New analyses for Experiment 3

--- 2024-05-02 manuscript/manuscript.qmd
@@
 title: "The role of cognateness in non-native spoken word recognition"
-
-abstract: |
-    Understanding spoken words in a non-native language is easier when the words are phonologically similar to their translations into the listener's native language are phonologically similar.
-Phonological similarity plays a more important role in the language performance of low-proficiency bilinguals, compared to that of high-proficiency bilinguals., and This suggests that low-proficiency bilinguals may points to the latter as relying more strongly on cross-language word-to-word links (lexical route), while high-proficiency bilinguals rely more on word-concept associations (conceptual route) than on cross-language word-to-word links (lexical route).
-Disentangling bilinguals’ reliance on the conceptual and lexical and conceptual routes during translation is challenging, given that both sources of information are present during spoken word comprehensionto some extent in both high- and low-proficiency bilinguals.
-In this study, we tested English and Spanish native speakers in a translation elicitation task, in which they listened to words in an unfamiliar language (Spanish or Catalan for English natives, and Catalan for Spanish natives), and then had to translate them to their native language.
-Given their lack of previous knowledge ion the unfamiliar tested language, phonological similarity between the presented words and their correct translations was the only cue available for participants to succeed in the task.
-This allowed us to explore the informativeness of the lexical route for word translation, in the absence of information from the conceptual route.
-Our results indicate that participants benefited from the phonological similarity between translation equivalents: the more similar the word they listened to and its translational equivalent, the higher the probability of a correct translation.
-These results point to cross-language word-to-word links as providing rich information during word translation.
 ---
@@
 #| label: setup
-#| message: false
-#| warning: false
-#| include: false
-
 library(targets)
@@
 tar_load(exp_participants)
+tar_load(exp_responses)
+tar_load(quest_responses)
 tar_load(quest_participants)
 tar_load(stimuli)
+tar_load(epreds)
+tar_load(dataset_1)
+tar_load(dataset_2)
+tar_load(dataset_3)
+tar_load(exp_1_m0)
+tar_load(exp_2_m0)
+tar_load(exp_3_m1)
 
-theme_set(theme_ggdist())
+theme_set(theme_ggdist() + 
+              theme(panel.grid.major.y = element_line(colour = "grey",
+                                                      linetype = "dotted")))
 
@@
 
+Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
@@
 
-```{r}
-#| label: participants
+```{r tbl-participants}
+#| label: tbl-participants
 tbl_data <- participants |> 
@@
 
-```{r}
-#| label: participants-excluded-1
-excluded_lang <- participants_1 |> 
-    filter(l_2_writ_prod > 3 |
-               l_2_oral_comp > 3 |
-               cat_oral_comp > 3 | 
-               cat_writ_prod > 3 |
-               spa_oral_comp > 3 |
-               spa_writ_prod > 3) |> 
-    nrow()
-
-excluded_impairment <- participants_1 |> 
-    filter(has_language_problems) |> 
-    nrow()
-
-excluded_ntrials <- participants_1 |> 
-    filter(invalid_participant_trials) |> 
-    nrow()
-```
-Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
-Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-We excluded data from participants that (a) self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), (b) were diagnosed with a language (*n* = `r excluded_impairment`), or (c) did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
-
 ### Stimuli
@@
 names(stim_exc_freq) <- janitor::make_clean_names(names(stim_exc_freq))
+
+stim_exc_freq_n <- sum(map_int(stim_exc_freq, "n"))
+```
+
+For each translation pair, we defined three predictors of interest: the lexical frequency of the correct translation (*Frequency*), the phonological similarity between the presented (Catalan or Spanish) word and their correct English translation (*Cognateness*), and presented word's number of cross-language phonological neighbours (*CLPN*).
+
+We included *Frequency* under the hypothesis that, keeping other predictors constant, participants would translate higher-frequency words more accurately and faster than lower-frequency words.
+Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores [@van2014subtlex].
+Words in the stimuli list without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
+
+We calculated *Cognatenesss*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
+The Levenshtein distance computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
+We divided this edit distance by the number of characters included in the longest X-SAMPA transcription of the translation pair.
+This results in a proportion score, in which values closer to zero correspond to lower Levenshtein distances between phonological transcriptions (i.e., higher similarity), and values closer to 1 correspond to higher Levenshtein distances (i.e., lower similarity).
+This transformation also help account for the fact that the Levenshtein distance tends to increase with the length of the transcriptions.
+For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+
+For each Catalan and Spanish word, we calculated the number of *CLPN* by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) different in up to one phoneme from that of the presented Catalan or Spanish word.
+Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological trnascriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
+
+```{r tbl-stimuli}
+#| label: tbl-stimuli
+stimuli |> 
+    summarise(across(c(freq_zipf_2, lv, neigh_n_h),
+                     lst(mean, sd, min, max), 
+                     na.rm = TRUE),
+              .by = c(group)) |> 
+    gt(rowname_col = "group") |> 
+    tab_spanner("Frequency", matches("freq")) |> 
+    tab_spanner("Cognateness", matches("lv")) |> 
+    tab_spanner("CLPN", matches("neigh")) |> 
+    fmt_number(is.numeric) |> 
+    fmt_integer(c(neigh_n_h_min,
+                  neigh_n_h_max)) |> 
+    cols_merge_uncert(freq_zipf_2_mean, freq_zipf_2_sd) |> 
+    cols_merge_uncert(lv_mean, lv_sd) |> 
+    cols_merge_uncert(neigh_n_h_mean, neigh_n_h_sd) |> 
+    cols_merge_range(freq_zipf_2_min, freq_zipf_2_max) |> 
+    cols_merge_range(lv_min, lv_max) |> 
+    cols_merge_range(neigh_n_h_min, neigh_n_h_max) |> 
+    cols_label(freq_zipf_2_mean = "Mean ± SD",
+               freq_zipf_2_sd = "SD",
+               freq_zipf_2_min = "Range",
+               lv_mean = "Mean ± SD",
+               lv_sd = "SD",
+               lv_min = "Range",
+               neigh_n_h_mean = "Mean ± SD",
+               neigh_n_h_sd = "SD",
+               neigh_n_h_min = "Range") |> 
+    tab_style(cell_text(weight = "bold"),
+              cells_column_spanners()) |> 
+    tab_style(cell_text(style = "italic"),
+              cells_column_labels())
+
 ```
@@
 
-```{r}
+We modelled the probability of participants guessing the correct translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
+We included as fixed effects the intercept, the main effects of *Frequency*, *Cognateness*, and *CLPN*, and the two-way interaction between *Cognateness* and *CLPN*.
+We also included participant-level random intercepts and slopes for the the main effects and the interaction.
+Eq.
+1 shows a formal description of the model.
+
+![Eq. 1](_assets/img/model.PNG)
+
+To test the practical relevance of each predictor we followed [@kruschke2018bayesian].
+We first specified a region of practical equivalence (*ROPE*) around zero ([-0.1, +0.1], in the logit scale).
+This area indicates the values of the regression coefficients that we considered as equivalent to zero.
+We then computed the 95% posterior credible intervals (CrI) of the regression coefficients of interest.
+Finally, we calculated the proportion of the 95% CrI that fell inside the ROPE.
+This proportion indicates the probability that the true value of the coefficient is equivalent to zero.
+
+All analyses were performed in R environment [@rcore2019r].
+We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
+We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+
+## Results
+
+```{r dataset-1}
+#| label: dataset-1
+dataset <- exp_responses |> 
+    filter(group!="cat-SPA") |> 
+    group_split(group) |> 
+    set_names(c("spa-ENG", "cat-ENG")) |> 
+    janitor::clean_names()
+
+dataset <- split(exp_responses, exp_responses$group)
+names(dataset) <- janitor::make_clean_names(names(dataset))
+r_validity <- map(dataset, 
+                  \(x) janitor::clean_names(table(x$response_type)))
+n_participants <- map(dataset, \(x) n_distinct(x$participant_id))
+
+excluded_lang <- participants_1 |> 
+    filter(l2_writ_prod > 3 |
+               l2_oral_comp > 3 |
+               cat_oral_comp > 3 | 
+               cat_writ_prod > 3 |
+               spa_oral_comp > 3 |
+               spa_writ_prod > 3) |> 
+    nrow()
+
+excluded_impairment <- participants_1 |> 
+    filter(valid_status=="Language impairment") |> 
+    nrow()
+
+excluded_ntrials <- participants_1 |> 
+    filter(valid_status=="Insufficient trials") |> 
+    nrow()
+
+dataset_clean <- exp_responses |> 
+    filter(group!="cat-SPA") |> 
+    dplyr::filter(valid_participant, valid_response) |> 
+    group_split(group) |> 
+    set_names(c("spa-ENG", "cat-ENG")) |> 
+    janitor::clean_names()
+
 lengths <- dataset_1 |> 
@@
 
+We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` unique participants.
+Of those trials, `r format(nrow(dataset$cat_eng), big.mark = ",")` were provided by `r n_participants$cat_eng` unique participants who listened to Catalan words, and `r format(nrow(dataset$spa_eng), big.mark = ",")` trials were provided by `r n_participants$spa_eng` unique participants who listened to Spanish words.
+We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_eng["blank"]`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (e.g., `f`, *n* = `r r_validity$cat_eng["incomplete"]`), and in which participants added comments to the experimenter (e.g., `unsure`, (*n* = `r r_validity$cat_eng["other"]`).
+In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
+Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
 Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, Range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, Range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
 
-```{r}
+```{r tbl-results}
+#| label: tbl-results
 tbl_data <- list("Experiment 1" = dataset_1,
@@
 
-#### Predictors
+```{r}
+c1 <- gather_draws(exp_1_m0, `b_.*`, `sd_.*`, regex = TRUE) |> 
+    summarise(across(.value, lst(.median = median, 
+                                 .lower = \(x) hdi(x)[1],
+                                 .upper = \(x) hdi(x)[2],
+                                 .rope = \(x) mean(between(x, -0.1, 0.1))),
+                     .names = "{fn}"))
+c1 <- split(c1, c1$.variable)
+names(c1) <- janitor::make_clean_names(names(c1)) 
+```
 
-We measured the phonological similarity between presented words and their target translation by computing the Levenshtein similarity between their X-SAMPA transcriptions using the `stringsim` function of the `stringdist` R package [@van2014stringdist].
-This function computes the inverse of the Levenshtein distance between two character strings as a proportion.
-First, it computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
-This measure is then divided by the maximum distance (according to the length of the longest string) and then subtracted from 1.
-The result is a percentage score, with 0% indicating no similarity between the two strings, and 100% indicating a perfect match between the two strings.
-We computed this similarity measure (referred simply as Levenshtein, from here on) for every translation pair in our stimuli lists.
-For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
-Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+@tbl-results shows a summary of participants' accuracy across Experiments 1, 2, and 3.
+MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
+Participants translating Catalan words and participants translating Spanish words performed equivalently, as indicated by the regression coefficient of *Group* ($\beta$ = `r round(c1$b_group1$.median, 3)`, 95% CrI = [`r round(c1$b_group1$.lower, 3)`, `r round(c1$b_group1$.upper, 3)`], *p*(ROPE) = `r round(c1$b_group1$.rope, 3)`).
+Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Cognateness* and *CLPN* ($\beta$ = `r round(c1$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_lv$.lower, 3)`, `r round(c1$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_lv$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Cognateness* ($\beta$ = `r round(c1$b_lv$.median, 3)`, 95% CrI = [`r round(c1$b_lv$.lower, 3)`, `r round(c1$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h$.lower, 3)`, `r round(c1$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h$.rope, 3)`).
+@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Cognateness* and *CLPN*.
+@fig-coefs shows a graphic summary of the posterior distribution of the regression coefficients of interest.
 
-The lexical frequencies in this database were extracted from SUBTLEX-UK [@van2014subtlex].
-We transformed lexical frequency scores to Zipf scores [@van2014subtlex].
-Word in the stimuli list without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
-For each presented Catalan or Spanish word presented in each list, we calculated its number of higher-frequency cross-language phonological neighbours (CLPN).
-This measure was calculated by counting the number of higher-frequency English words whose phonological word-form (generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)) differed by one phoneme from that of the presented Catalan or Spanish word.
-Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond].
-
-#### Modelling approach
-
-We modelled the probability of participants guessing the correct target translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We first fitted a base model that only included the intercept as a fixed effect, and random intercepts per participant  (Model 0).
-Secondly, we added the fixed effect of lexical frequency of the target translation ($Frequency$).
-We then extended the model sequentially by adding the main effect of higher-frequency phonological neighbourhood ($PTHN$), the main effect of Levenshtein similarity ($Levenshtein$), the $PTHN \times Levenshtein$ interaction, the group of participants ($Group$, with levels cat-ENG, spa-ENG, and cat-SPA), and finally, the $Levenshtein \times Group$ interaction).
-We specified two sum-coded *a priori* contrasts for the $Group$ predictor: one comparing English natives (cat-ENG = -0.25, spa-ENG = -0.25) against Spanish natives (cat-SPA = +0.5), and one comparing English natives translating Spanish words (spa-ENG = -0.5) against English natives translating Catalan words (cat-ENG = +0.5) [@schad2020capitalize].
-Each model included random intercepts by item (to account for the correlation of responses from different participants to the same presented word), and random intercepts and slopes by participant, where our experimental designed allowed it (to account for the correlation between responses from the same participant to different presented words) [@barr2013random].
-We will later refer to these models as Model 0 to Model 6.
-For interpretability, all continuous predictor variables were standardised (transformed in standard deviations from the mean) before entering the model.
-
-$$\begin{align}
-
-&\textbf{Likelihood}  \\
-y_{i} \sim& Bernoulli(p_{i}) && \text{[probability of correct translation]} \\ \\
-
-&\textbf{Parameters}  \\
-
-logit(p_{i}) = ~ &  \beta_{0[p,w]} ~ +  && \text{[linear model]}\\
-& \beta_{1[p]} ~ Frequency_{i} ~ + \\
-& \beta_{2[p]} ~ PTHN_i ~ + \\
-& \beta_{3[p]} ~ Similarity_i ~ + \\
-& \beta_{4[p]} ~ (PTHN_i \times Similarity_i) \\ \\
-
-\beta_{0-6[p,w]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P ~\text{and  word } w ~\text{in 1, ..., } W && \text{[participant- and word-level intercepts]} \\
-\beta_{1-6[p]} \sim& ~  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p ~\text{in 1, ..., } P
-&& \text{[participant-level coefficients]} \\ \\
-
-&\textbf{Prior}  \\
-
-\mu_{\beta_{p,w}} ~ \sim& ~ \mathcal{N}(0, 0.1) && \text{[participant-level coefficients]} \\
-\sigma_{\beta_{p}}, ~ \sigma_{\beta_{w}} \sim& ~ HalfCauchy(0, 0.1) && \text{[SD for population and participant]} \\
-\rho_{p}, ~ \rho_{w} \sim& ~LKJ(8) && \text{[correlation between participant-level coefficients]} \\
-
-\end{align}$$
-
-#### Statistical inference
-
-To test and account for cross-group differences, we included a random intercept for each group.
-We compared models using leave-one-out cross-validation (*LOO*) [@vehtari2017practical].
-More information about the models and model comparison can be found in Appendix 2.
-All analyses were performed in R environment [@rcore2019r].
-We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
-We used the brms R package [@burkner2017brms] using the cmdstanr backend to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
-
-## Results
-
-```{r}
+```{r tbl-dataset}
 #| label: tbl-dataset
 sem <- function(x) mean(x) / (sqrt(length(x)))
+
+#' Proportion adjusted from boundary values (Gelman, Hill & Vehtari, 2020)
+#' 
+prop_adj <- function(x, n){
+    e <- (x+2)/(n+4)
+    return(e)
+}
+
+#' Adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+#' 
+prop_adj_se <- function(x, n) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    return(se)
+}
+
+#' Adjusted standard error of proportion (Gelman, Hill & Vehtari, 2020)
+#' 
+prop_adj_ci <- function(x, n, .width = 0.95) {
+    e <- (x+2)/(n+4)
+    se <- sqrt(e*(1-e)/(n+4))
+    ci <-  e + qnorm(c((1-.width)/2, (1-(1-.width)/2)))*se
+    ci[1] <- ifelse(ci[1]<0, 0, ci[1]) # truncate at 0
+    ci[2] <- ifelse(ci[2]>1, 1, ci[2]) # truncate at 1
+    return(ci)
+}
 
@@
     mutate(correct = prop_adj(correct, n_trials)) |> 
-    summarise(across(correct, lst(mean, sd, sem, min, max, list)),
+    summarise(across(correct, lst(mean, sd, sem, min, max)),
               across(n_trials, lst(mean, sum, sd, min, max)),
@@
               .by = c(group, exp)) |> 
-    relocate(n_participants, matches("n_trials"))
+    relocate(n_participants, matches("correct"))
 
@@
     fmt_number(is.numeric, decimals = 2) |> 
-    gtExtras::gt_plt_dist(correct_list, type = "density") |> 
+    #gtExtras::gt_plt_dist(correct_list, type = "density") |> 
     fmt_integer(c(matches("sum|min|max"), n_participants)) |> 
     fmt_number(matches("correct"), scale_by = 100) |> 
+    tab_spanner("Accuracy (%)", matches("correct")) |> 
     tab_spanner("Valid trials", matches("n_trials")) |> 
-    tab_spanner("Accuracy (%)", matches("correct")) |> 
+    cols_merge_uncert(n_trials_mean, n_trials_sd) |> 
     cols_merge_range(n_trials_min, n_trials_max) |> 
+    cols_merge_uncert(n_trials_mean, n_trials_sd) |> 
+    cols_merge_uncert(correct_mean, correct_sd) |> 
     cols_merge_range(correct_min, correct_max) |> 
@@
                n_participants = "N",
-               n_trials_mean = "Mean",
-               n_trials_sd = "SD",
+               n_trials_mean = "Mean ± SD",
                n_trials_min = "Range",
-               correct_mean = "Mean",
-               correct_sd = "SD",
+               correct_mean = "Mean ± SD",
                correct_sem = "SE",
-               correct_min = "Range",
-               correct_list = "") |> 
+               correct_min = "Range") |> 
     tab_style(cell_text(style = "italic"),
@@
 
-All models showed good out-of-sample predictive validity, as suggested by the fact that the expected log-probability density was many times larger than its associated standard error.
-Model 6, which included all main effects and the two-way interactions between $PTHN$ and $Levenshtein$, and between $Group$ and $Levenshtein$ and $PTHN$ and consonant similarity, showed the best performance.
+```{r fig-epreds-1}
+#| label: fig-epreds-1
+#| fig-height: 6
+#| fig-width: 10
+get_epreds <- function(model, data, n = 100, 
+                       lv = seq(0, 1, length.out = 100), 
+                       neigh_n_h = c(0, 2, 4, 8, 12),
+                       ...) {
+
+    knowledge <- unique(model$data$knowledge)
+    confidence <- unique(model$data$confidence)
+    freq_zipf_2 <- mean(data$freq_zipf_2, na.rm = TRUE)
+    group <- unique(model$data$group)
+    nd <- expand_grid(freq_zipf_2, neigh_n_h, lv, knowledge, confidence, group)
+    epreds <- tidybayes::add_epred_draws(nd, model, re_formula = NA, ...)
+
+    return(epreds)
+}
+
+get_epreds(exp_1_m0, dataset_1) |> 
+    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+           neigh_n_h = factor(neigh_n_h,
+                              levels = paste0(c(0, 2, 4, 8, 12),
+                                              " neighbours"),
+                              ordered = TRUE)) |> 
+    ggplot(aes(lv, .epred)) +
+    facet_grid(group~neigh_n_h) +
+    stat_lineribbon(aes(fill_ramp = after_stat(level)),
+                    linewidth = 1/2,
+                    .width = c(0.95, 0.89, 0.78, 0.67, 0.50)) +
+    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+         y = "p(Correct)",
+         colour = "Cross-language neighbourhood density",
+         fill = "Cross-language neighbourhood density",
+         linetype = "Cross-language neighbourhood density",
+         fill_ramp = "CrI (%)") +
+    scale_x_continuous(labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent) +
+    scale_fill_brewer(palette = "Reds") +
+    theme(legend.position = "none",
+          legend.box = "vertical") 
+```
+
+## Discussion
+
+In the present experiment, we investigated the extent to which cognateness (phonological similarity between translation equivalents) is sufficient for successful word translation, in the absence of semantic knowledge.
+We tested two groups of monolingual British English natives in a translation task that involved words in Catalan or Spanish, two languages participants reported having no familiarity with.
+Participants benefited strongly from cognateness, when the correct translation of the presented words in Catalan or Spanish had few English phonological neighbours with higher lexical frequency.
+This suggests that, in the absence of distractors, even naïve participants can efficiently use phonological similarity to succeed in a translation task.
+Our results suggest that word-to-word connections at the phonological level might play a strong role during L2 speech comprehension, specially in low-proficiency listeners.
+
+# Experiment 2
+
+## Methods
+
+### Participants
+
+```{r participants-numbers-2}
+#| label: participants-numbers-2
+participants_2 <- exp_participants |> 
+    filter(group=="cat-SPA")
+
+collection_dates <- participants_2 |> 
+    pull(date) |> 
+    range() |> 
+    format("%B %dth, %Y")
+
+n_participants <- n_distinct(participants_2)
+n_participants_group <- table(participants_2$group)
+participants_age <- participants_2 |> 
+    summarise(across(age, lst(mean, sd, min, max))) 
+n_female <- table(participants_2$gender)
+```
+
+Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
+Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
+We collected data from `r n_participants` Spanish native participants living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+
+### Stimuli
+
+Same list of Catalan stimuli as in Experiment 1.
+
+### Procedure
+
+Same as in Experiment 1.
+
+### Data analysis
+
+Same approach as in Experiment 1.
+
+## Results
+
+```{r dataset-2}
+#| label: dataset-2
+dataset <- exp_responses |> 
+    filter(group=="cat-SPA") |> 
+    group_split(group) |> 
+    set_names(c("cat-SPA")) |> 
+    janitor::clean_names()
+
+dataset <- split(exp_responses, exp_responses$group)
+names(dataset) <- janitor::make_clean_names(names(dataset))
+r_validity <- map(dataset, 
+                  \(x) janitor::clean_names(table(x$response_type)))
+n_participants <- map(dataset, \(x) n_distinct(x$participant_id))
+
+excluded_lang <- participants_2 |> 
+    filter(l2_writ_prod > 3 |
+               l2_oral_comp > 3 |
+               cat_oral_comp > 3 | 
+               cat_writ_prod > 3 |
+               spa_oral_comp > 3 |
+               spa_writ_prod > 3) |> 
+    nrow()
+
+excluded_impairment <- participants_1 |> 
+    filter(valid_status=="Language impairment") |> 
+    nrow()
+
+excluded_ntrials <- participants_1 |> 
+    filter(valid_status=="Insufficient trials") |> 
+    nrow()
+
+dataset_clean <- exp_responses |> 
+    filter(group=="cat-SPA") |> 
+    dplyr::filter(valid_participant, valid_response) |> 
+    group_split(group) |> 
+    set_names(c("cat-SPA")) |> 
+    janitor::clean_names()
+
+lengths <- dataset_2 |> 
+    mutate(len = nchar(response)) |> 
+    summarise(across(len,
+                     lst(mean, median, sd, min, max)),
+              .by = c(group)) 
+```
+
+We collected data for a total of `r format(nrow(dataset$cat_spa) + nrow(dataset$cat_spa), big.mark = ",")` trials completed by `r n_participants$cat_spa` unique participants.
+We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_spa["blank"]`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r r_validity$cat_spa["language"]`), in which participants did not provide a whole word (e.g., `f`, *n* = `r r_validity$cat_spa["incomplete"]`), and in which participants added comments to the experimenter (e.g., `unsure`, (*n* = `r r_validity$cat_spa["other"]`).
+In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_spa$participant_id) + n_distinct(dataset_clean$cat_spa$participant_id)` unique participants.
+Of those trials, `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Spanish words.
+Responses given by participants were `r round(lengths$len_mean, 2)` characters long on average (*Median* = `r round(lengths$len_median, 2)`, *SD* = `r round(lengths$len_sd, 2)`, Range = `r round(lengths$len_min, 2)`-`r round(lengths$len_max, 2)`).
 
 ```{r}
-#| label: model-comparison
-#| message: false
-#| warning: false
-tab_loos <- model_loos %>% 
-    as.data.frame() %>% 
-    rownames_to_column("model") %>% 
-    arrange(desc(elpd_loo)) %>% 
-    select(model, matches("elpd_loo"), matches("looic"), matches("diff")) %>%
-    mutate(model = str_replace(model, "fit_", "Model ")) %>% 
-    mutate_all(~ifelse(. == 0, NA, .)) %>% 
-    gt(rowname_col = "model", id = "model-comparison") %>% 
-    tab_header(
-        title = md("Table 2"),
-        subtitle = md("Model performance and comparison using leave-one-out cross-validation")
-    ) %>% 
-    tab_source_note(
-        md("*Note*.
-Outcomes of the leave-one-out cross-validation procedure.
-Each row indicates the values of the expected log-predicted density (*ELPD*), leave-one-out information creterion (*LOO-IC*), and the difference in *ELPD* (*LOO-diff*) between each model and the inmediately simpler model.
-Each value is accompanied by a standard error (*SE*) indicating the uncertainty of its estimate.")
-    ) %>% 
-    fmt_missing(everything(), missing_text = "-") %>% 
-    fmt_number(2:7) %>% 
-    cols_label(
-        model = md("**Model**"),
-        looic = md("***LOO<sub>IC</sub>***"),
-        se_looic = md("***SE*<sub>IC</sub>**"),
-        elpd_loo = md("***LOO<sub>ELPD</sub>***"),
-        se_elpd_loo = md("***SE***"),
-        elpd_diff = md("***LOO<sub>diff</sub>***"),
-        se_diff = md("***SE<sub>diff</sub>***")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", weight = "bold", size = "medium"),
-        locations = cells_title(groups = "title")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", size = "medium", style = "italic"),
-        locations = cells_title(groups = "subtitle")
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "left", size = "medium"),
-        locations = cells_source_notes()
-    ) %>% 
-    tab_style(
-        style = list(
-            cell_borders(sides = c("left", "right"), color = "white"),
-            cell_text(align = "center")
-        ),
-        locations = list(
-            cells_body(), 
-            cells_row_groups(), 
-            cells_column_spanners(), 
-            cells_stub(), 
-            cells_source_notes(),
-            cells_column_labels(), 
-            cells_stubhead()
-        )
-    ) %>% 
-    tab_style(
-        style = list(
-            cell_text(align = "left", size = "medium"), 
-            cell_borders(sides = "all", color = "white")
-        ),
-        locations = list(cells_stub(), cells_source_notes())
-    ) %>% 
-    tab_style(
-        style = cell_text(align = "center", size = "medium", weight = "normal", style = "normal"),
-        locations = list(cells_column_spanners(), cells_column_labels())
-    ) %>% 
-    tab_style(
-        style = cell_borders(sides = "all", color = "white"),
-        locations = cells_body(columns = 1:7)
-    ) %>% 
-    tab_style(
-        style = cell_borders(sides = "top", color = "white"),
-        locations = cells_body(columns = 1:7)
-    ) %>% 
-    gtsave(here("img", "loos.png"))
-
-include_graphics(here("img", "loos.png"))
-
+c2 <- gather_draws(exp_2_m0, `b_.*`, `sd_.*`, regex = TRUE) |> 
+    summarise(across(.value, lst(.median = median, 
+                                 .lower = \(x) hdi(x)[1],
+                                 .upper = \(x) hdi(x)[2],
+                                 .rope = \(x) mean(between(x, -0.1, 0.1))),
+                     .names = "{fn}"))
+c2 <- split(c2, c2$.variable)
+names(c2) <- janitor::make_clean_names(names(c2)) 
 ```
 
-We now report the mean of the posterior distribution of each coefficient in Model 6, along with its associated measures of uncertainty.
-For interpretability, we transformed the estimates of the intercept using the inverse logit function so that the values are expressed in probability of correct response instead of log-odds, and we transformed the coefficients of the rest of predictors divided by four.
-Dividing a coefficient expressed in log-odds by four returns an approximate of the derivative of the logistic function indicating the maximum steepness of the logistic curve.
-This way, the coefficients are expressed as increments or decrements in the probability of correct translation [@gelman2020regression].
+@tbl-results-1 shows a summary of participants' accuracy across Experiments 1, 2, and 3.
+MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
+Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Cognateness* and *CLPN* ($\beta$ = `r round(c2$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_lv$.lower, 3)`, `r round(c2$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_lv$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Cognateness* ($\beta$ = `r round(c2$b_lv$.median, 3)`, 95% CrI = [`r round(c2$b_lv$.lower, 3)`, `r round(c2$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c2$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h$.lower, 3)`, `r round(c2$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h$.rope, 3)`).
+@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Cognateness* and *CLPN*.
 
-```{r posteriorfix, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Estimated posterior distributions of population-level regression coefficients in Model 4.
-Distributions indicate the estimated posterior likelihood density of regression coefficients.
-Credible intervals (*CrI*), represented with increasingly lighter segmentents in the distribution indicate the range of values that contain the true value with 95\\%, 80\\%, and 50\\% probability.
-Black dots represent the mean of each distribution.", fig.height=4.5, fig.width=6.5}
+```{r fig-epreds-2}
+#| label: fig-epreds-2
+#| fig-height: 3.5
+#| fig-width: 10
+get_epreds(exp_2_m0, dataset_2) |> 
+    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+           neigh_n_h = factor(neigh_n_h,
+                              levels = paste0(c(0, 2, 4, 8, 12),
+                                              " neighbours"),
+                              ordered = TRUE)) |> 
+    ggplot(aes(lv, .epred)) +
+    facet_wrap(~neigh_n_h, nrow = 1) +
+    stat_lineribbon(aes(fill_ramp = after_stat(level)),
+                    linewidth = 1/2,
+                    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
+    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+         y = "p(Correct)",
+         colour = "Cross-language neighbourhood density",
+         fill = "Cross-language neighbourhood density",
+         linetype = "Cross-language neighbourhood density",
+         fill_ramp = "CrI (%)") +
+    scale_x_continuous(labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent) +
+    scale_fill_brewer(palette = "Reds") +
+    theme(legend.position = "none",
+          legend.box = "vertical") 
+```
 
-coefs <- fixef(fit_6) %>% 
-    as.data.frame() %>% 
-    rownames_to_column("variable") %>% 
-    clean_names() %>% 
-    mutate_at(vars(estimate:q97_5), ~ifelse(variable=="Intercept", inv_logit_scaled(.), ./4)) %>% 
-    group_split(variable) %>% 
-    set_names(make_clean_names(map(., "variable"))) %>% 
-    map(select, -variable) %>% 
-    map(unlist)
+## Discussion
 
-str_repl <- c(
-    "Intercept" = "Intercept",
-    "frequency_zipf_std" = "Frequency\n[+1 SD]",
-    "pthn_std:lv_std" = "PTHN \u00d7 Levenshtein", 
-    "lv_std:groupeng_spa" = "Levenshtein \u00d7 Group 1",
-    "lv_std:groupcateng_spaeng" = "Levenshtein \u00d7 Group 2",
-    "lv_std" = "Levenshtein\n[+1 SD]",
-    "pthn_std" = "PTHN\n[+1 SD]",
-    "groupeng_spa" =  "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-    "groupcateng_spaeng" = "Group 2\n[cat-ENG vs. spa-ENG]"
-)
+Experiment 2 was an extension of Experiment 1 to a population of monolinguals whose native language is typologically similar to the presented language.
+Particularly, we presented Catalan words to Spanish native speakers who were reportedly unfamiliar with Catalan.
+Our results indicate a similar pattern of results as those in Experiment 1: participants were able to provide correct translations of presented Catalan words, provided that the Catalan words shared some degree of phonological similarity with their Spanish translation, and that the number of phonological neighbours with higher lexical frequency was reduced.
+In contrasts with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 was more resilient to the interference of phonological neighbours.
+English natives in Experiment 1 provided significantly less accuracy responses when more than four phonological neighbors were present (even when translating high-similarity words), compared to when only one or neighbour were present.
+Spanish participants in Experiment 2 benefited from phonological similarity, even when eight neighbours were present.
+Spanish participants' performance declined after 8 neighbours, and was evident at 12 neighbours.
+Overall, this suggests that participants in Experiment 2, who were natives of a typologically similar language (Spanish) to the presented language (Catalan) benefited more strongly from phonological similarity than participants in Experiment 1, who were natives of typologically less similar language (English) to the presented language (Catalan, Spanish).
 
-post_draws <- gather_draws(
-    fit_6, 
-    `b_.*`, 
-    `sd_.*`, 
-    regex = TRUE
-)
+Participants from both Experiment 1 and 2 benefited strongly from phonological similarity to correctly translate words from a non-native, reportedly unfamiliar language.
+This pattern of results holds for most of the presented stimuli, but some low-similarity Catalan and Spanish words were responded to surprisingly accurately by English listeners.
+Given that participants were reportedly unfamiliar with both languages, it was expected that participants would be very unlikely to provide correct translations for words sharing little to no phonological similarity to their correct translation.
+@tbl-surprises a list of Catalan and Spanish words to which participants provided responses with $\geq$ 10 average accuracy.
 
-# fixed effects
-post_draws %>% 
-    filter(str_detect(.variable, "b_")) %>% 
-    mutate(
-        .variable_name = str_remove(.variable, "b_") %>% 
-            str_replace_all(str_repl) %>%  
-            factor(
-                levels = c(
-                    "Intercept", 
-                    "Frequency\n[+1 SD]",
-                    "PTHN\n[+1 SD]", 
-                    "Levenshtein\n[+1 SD]",
-                    "PTHN \u00d7 Levenshtein",
-                    "Group 1\n[cat/spa-ENG vs. cat-SPA]",
-                    "Group 2\n[cat-ENG vs. spa-ENG]",
-                    "Levenshtein \u00d7 Group 1", 
-                    "Levenshtein \u00d7 Group 2"
-                )
-            ),
-        .value = ifelse(str_detect(.variable, "Intercept"), inv_logit_scaled(.value), .value/4)
-    ) %>% 
-    arrange(.variable) %>% 
-    ggplot(aes(.value, fct_rev(.variable_name))) +
-    geom_vline(xintercept = 0) +
-    stat_slab(aes(fill = stat(cut_cdf_qi(cdf, .width = c(.5, .8, .95), labels = percent_format())))) +
-    stat_pointinterval(.width = 0, point_size = 1) +
-    scale_fill_manual(values = c("#1A85FF", "#9ccaff", "#d2e5fc"), na.translate = FALSE) +
-    scale_x_continuous(labels = function(x) percent(round(x, 1))) +
-    labs(x = "P(Correct)", y = "Posterior probability density", fill = "CrI") +
-    theme(
-        legend.position = c(1, 0.25),
-        legend.direction = "vertical",
-        legend.background = element_rect(fill = "grey95"),
-        legend.key = element_rect(fill = "grey95", colour = "grey95"), 
-        legend.key.height = unit(0.1, "cm"),
+```{r tbl-surprises}
+#| label: tbl-surprises
+#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen th presented word (in Catalan or Spanish) and their correct Enlgish or Spanish translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+tbl_data <- list("Experiment 1" = dataset_1,
+                 "Experiment 2" = dataset_2) |>
+    bind_rows(.id = "exp") |> 
+    summarise(n = n(),
+              correct = sum(correct),
+              .by = c(exp, group, lv, translation)) |> 
+    mutate(accuracy = prop_adj(correct, n),
+           accuracy_se = prop_adj_se(correct, n)) |>    
+    dplyr::filter(lv==0, accuracy >= 0.1) |> 
+    select(exp, translation, group, accuracy, accuracy_se) |> 
+    arrange(exp, group, desc(accuracy))
+
+tbl_data |> 
+    gt(rowname_col = "translation",
+       groupname_col = c("exp", "group")) |> 
+    fmt_number(is.numeric, scale_by = 100) |> 
+    cols_label(accuracy = "Accuracy (%)", accuracy_se = "SE")
+```
+
+It is likely that participants had prior knowledge of these words despite having reported little to no familiarity with the presented language (Catalan or Spanish).
+One possibility is that participants had previously encountered these words embedded in English linguistic input.
+Spanish words percolate English speech with relative frequency, via different sources such as popular culture, songs, TV programs, etc.
+In addition, words from languages other than Spanish, but with high similarity to the Spanish words (e.g., cognates from Italian or French) might appear in English speech as well.
+Such prior knowledge might not be specific to the low-similarity words highlighted before.
+Participants may also have had prior knowledge about higher-similarity words, which could have contributed to participants responding to such words more accurately than without such prior knowledge.
+In the case of higher-similarity words, it is more difficult to disentangle the extent to which participants' accuracy is a function of pure phonological similarity, or prior knowledge they had about the meaning of Spanish words.
+To investigate this issue, we run Experiment 3.
+
+[ADD HERE LEXICAL FREQUENCY OF CATALAN AND SPANISH WORDS IN ENGLISH, AND OF CATALAN WORDS IN SPANISH]
+
+# Experiment 3
+
+Experiment 3 is a replication of Experiment 1, in which we collected additional data about participants' prior familiarity with the presented Catalan and Spanish words, in addition to the same translation task presented to participants in Experiment 1.
+
+## Methods
+
+### Participants
+
+```{r participants-numbers-3}
+#| label: participants-numbers-3
+participants_3 <- quest_participants |> 
+    filter(group!="cat-SPA")
+
+collection_dates <- participants_3 |> 
+    pull(date) |> 
+    range() |> 
+    format("%B %dth, %Y")
+
+n_participants <- n_distinct(participants_3)
+n_participants_group <- table(participants_3$group)
+participants_age <- participants_3 |> 
+    summarise(across(age, lst(mean, sd, min, max))) 
+n_female <- table(participants_3$gender)
+```
+
+Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
+We collected data from `r n_participants` British English native participants living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+
+```{r}
+#| label: participants-excluded-3
+excluded_lang <- participants_3 |> 
+    filter(l2_writ_prod > 3 |
+               l2_oral_comp > 3 |
+               cat_oral_comp > 3 | 
+               cat_writ_prod > 3 |
+               spa_oral_comp > 3 |
+               spa_writ_prod > 3) |> 
+    nrow()
+
+excluded_impairment <- participants_3 |> 
+    filter(valid_status=="Language impairment") |> 
+    nrow()
+
+excluded_ntrials <- participants_3 |> 
+    filter(valid_status=="Insufficient trials") |> 
+    nrow()
+```
+
+Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
+We excluded data from participants that (a) self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), (b) were diagnosed with a language (*n* = `r excluded_impairment`), or (c) did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+
+### Stimuli
+
+Same as in Experiment 1.
+
+### Procedure
+
+[TO BE COMPLETED]
+
+### Data analysis
+
+Same approach as in Experiment 1.
+
+## Results
+
+```{r dataset-3}
+#| label: dataset-3
+dataset <- quest_responses |> 
+    filter(group!="cat-SPA") |> 
+    group_split(group) |> 
+    set_names(c("spa-ENG", "cat-ENG")) |> 
+    janitor::clean_names()
+
+dataset <- split(quest_responses, quest_responses$group)
+names(dataset) <- janitor::make_clean_names(names(dataset))
+r_validity <- map(dataset, 
+                  \(x) janitor::clean_names(table(x$response_type)))
+n_participants <- map(dataset, \(x) n_distinct(x$participant_id))
+
+excluded_lang <- participants_3 |> 
+    filter(l2_writ_prod > 3 |
+               l2_oral_comp > 3 |
+               cat_oral_comp > 3 | 
+               cat_writ_prod > 3 |
+               spa_oral_comp > 3 |
+               spa_writ_prod > 3) |> 
+    nrow()
+
+excluded_impairment <- participants_3 |> 
+    filter(valid_status=="Language impairment") |> 
+    nrow()
+
+excluded_ntrials <- participants_3 |> 
+    filter(valid_status=="Insufficient trials") |> 
+    nrow()
+
+dataset_clean <- quest_responses |> 
+    filter(group!="cat-SPA") |> 
+    dplyr::filter(valid_participant, valid_response) |> 
+    group_split(group) |> 
+    set_names(c("spa-ENG", "cat-ENG")) |> 
+    janitor::clean_names() 
+```
+
+We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` unique participants.
+Of those trials, `r format(nrow(dataset$cat_eng), big.mark = ",")` were provided by `r n_participants$cat_eng` unique participants who listened to Catalan words, and `r format(nrow(dataset$spa_eng), big.mark = ",")` trials were provided by `r n_participants$spa_eng` unique participants who listened to Spanish words.
+We excluded trials in which participants did not enter any text (*n* = `r dataset$cat_eng$blank`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r dataset$cat_eng$language`), in which participants did not provide a whole word (e.g., `f`, *n* = `r dataset$cat_eng$incomplete`), and in which participants added comments to the experimenter (e.g., `unsure`, (*n* = `r dataset$cat_eng$other`).
+In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
+Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
+
+```{r}
+lengths <- dataset_3 |> 
+    mutate(len = map_int(strsplit(response, ""), length)) |> 
+    summarise(across(len,
+                     lst(mean, median, sd, min, max)),
+              .by = c(group)) 
+lengths <- split(lengths, lengths$group)
+
+names(lengths) <- janitor::make_clean_names(names(lengths))
+```
+
+Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, Range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, Range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
+
+MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
+Overall, participants reported prior knowledge more often for that Spanish words with unexpectedly high accuracy (see Discussion in Experiment 2) than for words with expected accuracy (see @fig-knowledge).
+Participants reported prior knowledge of Catalan words with unexpected accuracy as often as those with expected accuracy.
+This suggests that participants in Experiment 1 may have relied, to some extent, on their prior knowledge about form-meaning mappings to correctly translate some Spanish words.
+To isolate such an effect of prior Spanish knowledge, we run the same analysis as in Experiment 1 on the newly collected translations from Experiment 3, now excluding responses to words in which participants reported prior knowledge.
+
+```{r fig-knowledge}
+#| label: fig-knowledge
+#| fig-width: 7
+#| fig-height: 5
+#| fig-cap: "Catalan/Spanish prior word knowledge as reported by English native participants in Experiment 3.
+(A) Average proportion of participants that reported prior knowledge for words with surprisingly high accuracy (no phonological similarity with the correct translation, accuracy higher than 10%), and for words with expected accuracy (low similarity-low accuracy, or high similarity-high accuracy).
+(B) Average accuracy for words with expected and unexpected accuracy."
+dataset_3 |> 
+    mutate(surprise = translation %in% tbl_data$translation) |> 
+    summarise(knowledge = mean(knowledge),
+              .by = c(translation, group, surprise)) |> 
+    select(translation, group, surprise, knowledge) |> 
+    mutate(surprise = if_else(
+        surprise,
+        "Unexpectedly high accuracy\n(no phonological similarity)",
+        "Expected accuracy")) |> 
+    ggplot(aes(knowledge, group, shape = group)) +
+    facet_wrap(~surprise, ncol = 1, switch = "y") +
+    stat_summary(fun = mean,
+                 size = 2,
+                 geom = "point") +
+    stat_summary(fun.data = mean_se,
+                 geom = "errorbar",
+                 width = 0.1) +
+    labs(x = "Prior knowledge reported",
+         y = "Group") +
+    scale_x_continuous(labels = scales::percent) +
+
+    dataset_3 |> 
+    mutate(surprise = translation %in% tbl_data$translation) |> 
+    mutate(surprise = if_else(
+        surprise,
+        "Unexpectedly high accuracy\n(no phonological similarity)",
+        "Expected accuracy")) |> 
+    summarise(correct = sum(correct),
+              n = n(),
+              .by = c(translation, group, surprise, knowledge)) |> 
+    mutate(accuracy = prop_adj(correct, n)) |> 
+    select(translation, group, accuracy, surprise, knowledge) |> 
+    ggplot(aes(accuracy, group, shape = group)) +
+    facet_wrap(~surprise, ncol = 1, switch = "y") +
+    stat_summary(fun = mean,
+                 size = 2,
+                 position = position_dodge(width = 0.75),
+                 geom = "point") +
+    stat_summary(fun.data = mean_se,
+                 geom = "errorbar",
+                 width = 0.1,
+                 position = position_dodge(width = 0.75)) +
+    labs(x = "Accuracy",
+         y = "Group") +
+    scale_x_continuous(labels = scales::percent,
+                       limits = c(0, 1)) +
+    theme(axis.title.y = element_blank(),
+          axis.text.y = element_blank(),
+          strip.text = element_blank()) +
+
+    plot_layout(nrow = 1, guides = "collect") &
+    plot_annotation(tag_levels = "A") &
+    labs(colour = "Group",
+         shape = "Group") &
+    theme(panel.border = element_rect(fill = NA,
+                                      colour = "grey",
+                                      linewidth = 1),
+          axis.line = element_blank(),
+          axis.text.y = element_blank(),
           axis.title.y = element_blank(),
-        legend.text = element_text(size = 7),
-        axis.text.x = element_text(colour = "black"),
-        axis.text.y = element_text(colour = "black", hjust = 1),
           axis.ticks.y = element_blank(),
-        panel.grid.major.y = element_line(
-            colour = "grey", 
-            size = 0.5
-        )
-    )
-
+          legend.position = "top",
+          panel.grid.major.x = element_line(colour = "grey",
+                                            linetype = "dotted"),
+          panel.grid.major.y = element_blank()) 
 ```
@@
 ```{r}
-#| label: bayes-factors
-#| message: false
-#| warning: false
-# wrapper to extract Bayes Factors
-get_bf <- function(hyp, abs = TRUE, ...){
-    bf <- hypothesis(fit_6, hypothesis = hyp, ...)$hypothesis$Evid.Ratio
-    if (abs) bf <- abs(bf)
-    return(bf)
-}
-
-h_frequency <- get_bf("frequency_zipf_std = 0")
-h_pthn <- get_bf("pthn_std = 0")
-h_lv <- get_bf("pthn_std = 0")
-h_pthn_lv <- get_bf("pthn_std:lv_std = 0")
-h_group1 <- get_bf("groupeng_spa = 0")
-h_group2 <- get_bf("groupcateng_spaeng = 0")
-h_lv_group1 <- get_bf("lv_std:groupeng_spa = 0")
-h_lv_group2 <- get_bf("lv_std:groupcateng_spaeng = 0")
-
-icc_conditional <- icc(fit_6, by_group = TRUE)$ICC[2]
-
-sds <- gather_draws(
-    fit_6,
-    `sd_.*`, 
-    regex = TRUE
-) %>% 
-    mean_qi() %>% 
-    mutate(
-        .value = ifelse(
-            grepl("Intercept", .variable), 
-            inv_logit_scaled(.value), 
-            .value/4
-        ),
-        .lower = ifelse(
-            grepl("Intercept", .variable),
-            inv_logit_scaled(.lower),
-            .lower/4
-        ),
-        .upper = ifelse(
-            grepl("Intercept", .variable),
-            inv_logit_scaled(.upper), 
-            .upper/4
-        )
-    ) %>% 
-    split(.$.variable)
-
-names(sds) <- make_clean_names(names(sds))
-
-word_example <- ranef(fit_6)$translation_id[,,"Intercept"] %>% 
-    as.data.frame() %>% 
-    rownames_to_column("word") %>% 
-    arrange(-Estimate) %>% 
-    clean_names() %>% 
-    mutate_if(is.numeric, inv_logit_scaled) %>% 
-    slice(1, n()) %>% 
-    left_join(
-        distinct(stimuli, word1, ipa1), 
-        by = c("word" = "word1")
-    ) %>% 
-    split(.$word)
-
-participant_ranef_range <- sds %>% 
-    bind_rows() %>% 
-    filter(!grepl("Intercept|word|group1", .variable)) %>% 
-    summarise(.value = range(.value)) %>% 
-    unlist()
-
-cors_range <- gather_draws(fit_6, `cor_.*`, regex = TRUE) %>% 
-    median_qi() %>% 
-    summarise(.value = range(.value))
-
-corr_mat <- VarCorr(fit_6, summary = TRUE)$participant_id$cor 
-corr_names <- rownames(VarCorr(fit_6, summary = TRUE)$participant_id$cor)
-
-corr_data <- corr_mat %>% 
-    as_tibble() %>% 
-    clean_names() %>% 
-    mutate(term1 = corr_names) %>% 
-    relocate(term1) %>% 
-    split(.$term1)
-
-names(corr_data) <- make_clean_names(names(corr_data))
-
-h_cor_intercept_frequency <- get_bf("participant_id__Intercept__frequency_zipf_std = 0", class = "cor")
-h_cor_frequency_zipf_lv_group1 <- get_bf("participant_id__frequency_zipf_std__lv_std:groupeng_spa = 0", class = "cor")
-h_cor_intercept_pthn <- get_bf("participant_id__Intercept__pthn_std = 0", class = "cor")
+c3 <- gather_draws(exp_3_m1, `b_.*`, `sd_.*`, regex = TRUE) |> 
+    summarise(across(.value, lst(.median = median, 
+                                 .lower = \(x) hdi(x)[1],
+                                 .upper = \(x) hdi(x)[2],
+                                 .rope = \(x) mean(between(x, -0.1, 0.1))),
+                     .names = "{fn}"))
+c3 <- split(c3, c3$.variable)
+names(c3) <- janitor::make_clean_names(names(c3)) 
 ```
 
-Overall, participants had `r percent(coefs$intercept["estimate"], accuracy = 0.01)`, (95% *CrI* = [`r percent(coefs$intercept["q2_5"], accuracy = 0.01)`, `r percent(coefs$intercept["q97_5"], accuracy = 0.01)`]) likelihood to produce correct translations.
-The lexical frequency of the target translation had little impact on accuracy: every standard deviation increment (*SD* = `r printnum(sd(stimuli$frequency_zipf, na.rm = TRUE))` Zipf points) decreased the probability of correct translation in `r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$frequency_zipf_std["estimate"], accuracy = 0.01)`, `r percent(coefs$frequency_zipf_std["q2_5"], accuracy = 0.01)`], $BF_{\beta\neq< 0}$ = `r printnum(h_frequency)`).
-The number of phonological neighbours with higher lexical frequency than the target translation (*PTHN*) slightly decreased the probability of a correct responses by `r percent(coefs$pthn_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn)`) for every increase standard deviation increment  (1 *SD* = `r printnum(sd(stimuli$pthn, na.rm = TRUE))` neighbours).
-Finally, phonological similarity between the presented word and target translation had a strong effect on participants' accuracy: for every *SD* increment in Levenshtein similarity (1 *SD* = `r printnum(percent(sd(stimuli$lv, na.rm = TRUE)), digits = 2)` similarity), translation accuracy increased in `r percent(coefs$lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = $>3.09 \times 10^{18}$).
-The interaction effect between $PTHN \times Levenshtein$ was small: a +1 SD increment in Levenshtein increased the effect of *PTHN* by `r percent(coefs$pthn_std_lv_std["estimate"], accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$pthn_std_lv_std["q2_5"], accuracy = 0.01)`, `r percent(coefs$pthn_std_lv_std["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_pthn_lv)`)
+Participants translating Catalan words and participants translating Spanish words, as indicated by the regression coefficient of *Group* ($\beta$ = `r round(c3$b_group1$.median, 3)`, 95% CrI = [`r round(c3$b_group1$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_group1$.rope, 3)`).
+Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Cognateness* and *CLPN* ($\beta$ = `r round(c3$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h_lv$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h_lv$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Cognateness* ($\beta$ = `r round(c3$b_lv$.median, 3)`, 95% CrI = [`r round(c3$b_lv$.lower, 3)`, `r round(c3$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c3$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h$.lower, 3)`, `r round(c3$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h$.rope, 3)`).
+@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Cognateness* and *CLPN*.
 
-Spanish natives translating Catalan words were `r percent(coefs$groupeng_spa["estimate"], accuracy = 0.01)` more accurate than English natives translating Spanish and Catalan words (95% *CrI* = [`r percent(coefs$groupeng_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$groupeng_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group1)`).
-The interaction between this contrast and Levenshtein suggests a moderate impact of phonological similarity on the difference between Spanish and English participants: this difference increased in `r percent(coefs$lv_std_groupeng_spa["estimate"], accuracy = 0.01)` for every one *SD* increment in Levenshtein similarity (95% *CrI* = [`r percent(coefs$lv_std_groupeng_spa["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_groupeng_spa["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-English natives translating Catalan words were `r percent(abs(coefs$groupcateng_spaeng["estimate"]), accuracy = 0.01)` more accurate than those translating Spanish words (95% *CrI* = [`r percent(coefs$groupcateng_spaeng["q2_5"], accuracy = 0.01)`, `r percent(coefs$groupcateng_spaeng["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_group2)`).
-For every one *SD* increment in Levenshtein similarity, the effect of presented language increased in `r percent(abs(coefs$lv_std_groupcateng_spaeng["estimate"]), accuracy = 0.01)` (95% *CrI* = [`r percent(coefs$lv_std_groupcateng_spaeng["q2_5"], accuracy = 0.01)`, `r percent(coefs$lv_std_groupcateng_spaeng["q97_5"], accuracy = 0.01)`], $BF_{\beta \neq 0}$ = `r printnum(h_lv_group1)`).
-
-```{r}
-#| label: marginal-effects
-#| message: false
-#| warning: false
-#| fig-width: 7
-#| fig-height: 3.5
-#| fig-cap: "Expected mean posterior predictions from Model 4.
-A) Population-level predictions: the X-axis and the Y-axis represent the PTHN score (in standard deviations from the mean) and the probability of correct translation, respectively.
-We simulated 300 observations from the posterior distribution of the model: 100 simulations for translations with no similarity (0% Levenshtein), 100 simulations for translations with mean accuracy (20.51% Levenshtein), and 100 simulations for translations with the maximum observed accuracy (83.33% Levenshtein).
-We did this across the range of values of the PTHN scores.
-For each simulation, we drew a single sample from the posterior distribution of each coefficient.
-Each simulation is depicted in the graph as a line: pink for minimum similarity translations, blue for hight similarity translations, and yellow for maximum similarity translations.
-Black, thick lines indicate the expected mean value of the posterior predictions of the model for each condition.
-The dispersion of the lines indicates the uncertainty of our predictions.
-We computed these posterior predictions for each group of participants, and plotted them in separate panels."
-nd <- expand.grid(
-    lv_std = c(min(fit_6$data$lv_std), 0, max(fit_6$data$lv_std)),
-    frequency_zipf_std = 0,
-    pthn_std = seq(
-        min(fit_6$data$pthn_std, na.rm = TRUE),
-        max(fit_6$data$pthn_std, na.rm = TRUE),
-        by = 0.1
-    ),
-    group = unique(fit_6$data$group)
-)
-
-m <- add_epred_draws(nd, fit_6, ndraws = 100, re_formula = NA) %>% 
-    mutate(
-        lv_std = factor(
-            lv_std,
-            levels = c(
-                min(fit_6$data$lv_std),
-                0,
-                max(fit_6$data$lv_std)
-            ),
-            labels = c(
-                paste0("Min (", percent(min(responses$lv, na.rm = TRUE)), ")"),
-                paste0("Mean (", percent(mean(responses$lv, na.rm = TRUE)), ")"),
-                paste0("Max (", percent(max(responses$lv, na.rm = TRUE)), ")")
-            )
-        ) 
-    )
-
-ggplot(m) +
-    aes(x = pthn_std, y = .epred, colour = lv_std, fill = lv_std) +
-    facet_wrap(~group) +
-    geom_hline(yintercept = 0.5, colour = "grey") +
-    geom_line(aes(group = interaction(lv_std, .draw)), size = 0.65, alpha = 0.5, show.legend = TRUE) +
-    stat_summary(
-        aes(linetype = lv_std), 
-        fun = mean, 
-        geom = "line",
-        colour = "black", 
-        size = 0.75,
-        show.legend = TRUE
-    ) +
-    labs(
-        x = paste0(
-            "PTHN (+1 SD, ", 
-            round(sd(responses$pthn), 1), 
-            " higher-frequency neighbours)"
-        ),
-        y = "P(Correct)", 
-        colour = "Levenshtein similarity",
-        fill = "Levenshtein similarity",
-        linetype = "Levenshtein similarity",
-        title = "Population-level posterior predictions"
-    ) +
-    scale_color_manual(values = c("#1A85FF", "#ff2976", "#FFC20A")) +
-    scale_linetype_manual(values = c("solid", "dashed", "dotdash")) +
-    scale_y_continuous(
-        labels = function(x) percent(round(x, 2)),
-        limits = c(0, 1),
-        breaks = seq(0, 1, 0.1)
-    ) +
-    theme(
-        legend.position = "top",
-        plot.title = element_blank()
-    )
+```{r fig-epreds-3}
+#| label: fig-epreds-3
+#| fig-height: 6
+#| fig-width: 10
+get_epreds(exp_3_m1, dataset_3) |> 
+    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+           neigh_n_h = factor(neigh_n_h,
+                              levels = paste0(c(0, 2, 4, 8, 12),
+                                              " neighbours"),
+                              ordered = TRUE)) |> 
+    ggplot(aes(lv, .epred)) +
+    facet_grid(group~neigh_n_h) +
+    stat_lineribbon(aes(fill_ramp = after_stat(level)),
+                    linewidth = 1/2,
+                    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
+    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+         y = "p(Correct)",
+         fill_ramp = "CrI (%)") +
+    scale_x_continuous(labels = scales::percent) +
+    scale_y_continuous(labels = scales::percent) +
+    scale_fill_brewer(palette = "Reds") +
+    theme(legend.position = "top",
+          legend.box = "vertical") 
 ```
 
-# Discussion
+## Discussion
+
+# Joint analyses
+
+Across Experiments 1 and 3, we found strong evidence that participants efficiently exploited phonological similarity to provide accurate translations for words in an unfamiliar language, provided that few phonological neighbours of higher lexical frequency were present.
+@fig-coefs summarizes the posterior distribution of the regression coefficients of the models in Experiments 1 to 3.
+
+```{r fig-coefs}
+#| label: fig-coefs
+#| fig-width: 9
+#| fig-height: 5
+var_labels <- c("b_Intercept" = "Intercept",
+                "b_freq_zipf_2" = "Frequency",
+                "b_lv" = "Cognateness",
+                "b_neigh_n_h" = "CLPN",
+                "b_neigh_n_h:lv" = "Cognateness × CLPN",
+                "b_group1" = "cat-ENG vs. spa-ENG")
+
+list(exp_1_m0, exp_2_m0, exp_3_m1) |> 
+    set_names(paste0("Experiment ", 1:3)) |> 
+    map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |> 
+    dplyr::filter(.variable != "b_Intercept") |> 
+    mutate(.value = .value/4,
+           .variable = factor(.variable, 
+                              levels = names(var_labels),
+                              labels = var_labels)) |> 
+    ggplot(aes(.value, reorder(exp, desc(exp)))) +
+    facet_wrap(~.variable, scales = "free_x") +
+    annotate(geom = "rect",
+             xmin = -0.1/4,
+             xmax = 0.1/4,
+             ymin = -Inf,
+             ymax = Inf,
+             fill = "grey",
+             alpha = 3/4,
+             colour = NA) +
+    geom_vline(xintercept = 0, 
+               colour = "grey",
+               linewidth = 3/4) +
+    stat_slab(aes(fill = after_stat(level)), 
+              point_interval = mode_hdi, 
+              .width = c(0.95, 0.89, 0.67, 1),
+              outline_bars = TRUE,
+              colour = "white",
+              linewidth = 0.1,
+              breaks = 20,
+              normalize = "panels") +
+    stat_spike(at = "median",
+               size = 1,
+               aes(linetype = "Median"),
+               normalize = "panels") +
+    labs(x = "Estimate", y = "Experiment",
+         colour = "CrI",
+         fill = "CrI",
+         linetype = "") +
+    scale_fill_brewer(palette = "Reds") +
+    theme(legend.position = c(1, 0),
+          legend.box = "horizontal",
+          legend.justification = c(1, 0),
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
+```
+
+# General discussion
 

### 2024-05-08 manuscript/manuscript.tex
New analyses for Experiment 3

@@
 
-Data collection took place from June 04th, 2020 to June 25th, 2020. We
-collected data from 71 British English native participants living in
-United Kingdom (\emph{Mean} = 21.76 years, \emph{SD} = 2.15, Range =
-18-26, 46, 25 female).
-
-\captionsetup{labelsep=none}
-
-\begin{longtable}{l|rrrrrrc}
-
-\caption{\label{tbl-participants}}
-
-\tabularnewline
-
-\toprule
-\multicolumn{1}{l}{} & \multicolumn{2}{c}{Sample size} & \multicolumn{3}{c}{Age} & \multicolumn{2}{c}{L2} \\ 
-\cmidrule(lr){2-3} \cmidrule(lr){4-6} \cmidrule(lr){7-8}
-\multicolumn{1}{l}{} & N & Excluded & M & SD & Range & N knows L2? & Which L2? \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & 35 & 8 & $21.80$ & $2.08$ & $18$–$26$ & 1 & Russian \\ 
-cat-ENG & 36 & 4 & $21.72$ & $2.25$ & $18$–$25$ & 5 & German, Punjabi, French, Italian, Several \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & 33 & 12 & $21.85$ & $3.00$ & $18$–$33$ & 12 & Italian, French, German \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & 32 & 1 & $21.72$ & $2.59$ & $18$–$26$ & 3 & Japanese, German \\ 
-cat-ENG & 32 & 0 & $22.31$ & $2.39$ & $18$–$26$ & 2 & Cantonese, Irish \\ 
-\bottomrule
-
-\end{longtable}
-
 Participants were recruited via Prolific (5£ compensation) and SONA
@@
 complete the experiment using a laptop in a quiet place with good
-internet connection. We excluded data from participants that (a)
-self-rated their oral and/or written skills in Catalan and Spanish, or
-any other second language as four or higher in a five-point scale
-(\emph{n} = 2), (b) were diagnosed with a language (\emph{n} = 1), or
-(c) did not contribute more than 80\% of valid trials (\emph{n} = 9).
+internet connection. Data collection took place from June 04th, 2020 to
+June 25th, 2020. We collected data from 71 British English native
+participants living in United Kingdom (\emph{Mean} = 21.76 years,
+\emph{SD} = 2.15, Range = 18-26, 46, 25 female).
+
+\captionsetup{labelsep=none}
+
+\begin{longtable}{l|rrrrrrc}
+
+\caption{\label{tbl-participants}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{2}{c}{Sample size} & \multicolumn{3}{c}{Age} & \multicolumn{2}{c}{L2} \\ 
+\cmidrule(lr){2-3} \cmidrule(lr){4-6} \cmidrule(lr){7-8}
+\multicolumn{1}{l}{} & N & Excluded & M & SD & Range & N knows L2? & Which L2? \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & 35 & 8 & $21.80$ & $2.08$ & $18$–$26$ & 1 & Russian \\ 
+cat-ENG & 36 & 4 & $21.72$ & $2.25$ & $18$–$25$ & 5 & German, Punjabi, French, Italian, Several \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & 33 & 12 & $21.85$ & $3.00$ & $18$–$33$ & 12 & Italian, French, German \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & 32 & 1 & $21.72$ & $2.59$ & $18$–$26$ & 3 & Japanese, German \\ 
+cat-ENG & 32 & 0 & $22.31$ & $2.39$ & $18$–$26$ & 2 & Cantonese, Irish \\ 
+\bottomrule
+
+\end{longtable}
 
@@
 
-Responses given by English participants to Catalan presented words were
-5.35 characters long on average (\emph{Median} = 5, \emph{SD} = 1.79,
-Range = 1-14), while their translations to Spanish responses were 5.57
-characters long on average (\emph{Median} = 5, \emph{SD} = 1.97, Range =
-2-21).
-
-\captionsetup{labelsep=none}
-
-\begin{longtable}{l|rrrr}
-
-\caption{\label{tbl-results}}
-
-\tabularnewline
-
-\toprule
-\multicolumn{1}{l}{} & \multicolumn{2}{c}{Correct responses} & \multicolumn{2}{c}{Incorrect responses} \\ 
-\cmidrule(lr){2-3} \cmidrule(lr){4-5}
-\multicolumn{1}{l}{} & Correct & Typo & Wrong translation & False friend \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{5}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $429$ ($16.47\%$) & $11$ ($0.42\%$) & $2,082$ ($79.95\%$) & $82$ ($3.15\%$) \\ 
-spa-ENG & $374$ ($14.37\%$) & $2$ ($0.08\%$) & $2,117$ ($81.36\%$) & $109$ ($4.19\%$) \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{5}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $505$ ($18.41\%$) & $7$ ($0.26\%$) & $2,052$ ($74.81\%$) & $179$ ($6.53\%$) \\ 
-spa-ENG & $630$ ($20.03\%$) & $6$ ($0.19\%$) & $2,352$ ($74.79\%$) & $157$ ($4.99\%$) \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{5}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & $780$ ($46.93\%$) & $20$ ($1.20\%$) & $736$ ($44.28\%$) & $126$ ($7.58\%$) \\ 
-\bottomrule
-
-\end{longtable}
-
 We modelled the probability of participants guessing the correct
@@
 participant-level random intercepts and slopes for the the main effects
-and the interaction. \textbf{?@eq-model} shows a formal description of
-the model.
+and the interaction. Eq. 1 shows a formal description of the model.
+
+\begin{figure}[H]
+
+{\centering \includegraphics{_assets/img/model.PNG}
+
+}
+
+\caption{Eq. 1}
+
+\end{figure}%
 
 To test the practical relevance of each predictor we followed
-(\citeproc{ref-krushke2018bayesian}{\textbf{krushke2018bayesian?}}). We
+(\citeproc{ref-kruschke2018bayesian}{Kruschke \& Liddell, 2018}). We
 first specified a region of practical equivalence (\emph{ROPE}) around
-zero. This area indicates the values of the regression coefficients that
-we considered as equivalent to zero. We then computed the 95\% posterior
-credible intervals (CrI) of the regression coefficients of interest.
-Finally, we calculated the proportion of the 95\% CrI that fell inside
-the ROPE. This proportion indicates the probability that the true value
-of the coefficient is equivalent to zero.
+zero ({[}-0.1, +0.1{]}, in the logit scale). This area indicates the
+values of the regression coefficients that we considered as equivalent
+to zero. We then computed the 95\% posterior credible intervals (CrI) of
+the regression coefficients of interest. Finally, we calculated the
+proportion of the 95\% CrI that fell inside the ROPE. This proportion
+indicates the probability that the true value of the coefficient is
+equivalent to zero.
 
@@
 
+We collected data for a total of 6,446 trials completed by 72 unique
+participants. Of those trials, 2,988 were provided by 36 unique
+participants who listened to Catalan words, and 3,458 trials were
+provided by 36 unique participants who listened to Spanish words. We
+excluded trials in which participants did not enter any text (\emph{n} =
+72), in which a response in a language other than English was provided
+(e.g., \texttt{agua}, \emph{n} = 51), in which participants did not
+provide a whole word (e.g., \texttt{f}, \emph{n} = 5), and in which
+participants added comments to the experimenter (e.g., \texttt{unsure},
+(\emph{n} = 13). In addition, we excluded data from participants that
+self-rated their oral and/or written skills in Catalan and Spanish, or
+any other second language as four or higher in a five-point scale
+(\emph{n} = 2), were diagnosed with a language (\emph{n} = 1), or did
+not contribute more than 80\% of valid trials (\emph{n} = 9). After
+applying trial-level and participant-level inclusion criteria, the
+resulting dataset included 5,204 trials provided by 54 unique
+participants. Of those trials, 2,602 were provided by 27 unique
+participants who listened to Catalan words, and 2,604 trials were
+provided by 32 unique participants who listened to Spanish words.
+Responses given by English participants to Catalan presented words were
+5.35 characters long on average (\emph{Median} = 5, \emph{SD} = 1.79,
+Range = 1-14), while their translations to Spanish responses were 5.57
+characters long on average (\emph{Median} = 5, \emph{SD} = 1.97, Range =
+2-21).
+
+\captionsetup{labelsep=none}
+
+\begin{longtable}{l|rrrr}
+
+\caption{\label{tbl-results}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{2}{c}{Correct responses} & \multicolumn{2}{c}{Incorrect responses} \\ 
+\cmidrule(lr){2-3} \cmidrule(lr){4-5}
+\multicolumn{1}{l}{} & Correct & Typo & Wrong translation & False friend \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $429$ ($16.47\%$) & $11$ ($0.42\%$) & $2,082$ ($79.95\%$) & $82$ ($3.15\%$) \\ 
+spa-ENG & $374$ ($14.37\%$) & $2$ ($0.08\%$) & $2,117$ ($81.36\%$) & $109$ ($4.19\%$) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & $780$ ($46.93\%$) & $20$ ($1.20\%$) & $736$ ($44.28\%$) & $126$ ($7.58\%$) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $477$ ($18.01\%$) & $7$ ($0.26\%$) & $1,986$ ($75.00\%$) & $178$ ($6.72\%$) \\ 
+spa-ENG & $590$ ($19.53\%$) & $6$ ($0.20\%$) & $2,294$ ($75.94\%$) & $131$ ($4.34\%$) \\ 
+\bottomrule
+
+\end{longtable}
+
+Table~\ref{tbl-results} shows a summary of participants' accuracy across
+Experiments 1, 2, and 3. MCMC chains in the model showed strong evidence
+of convergence (\(\hat{R}<1.01\)) (see Appendix 2 for more detailed
+model diagnostics). Participants translating Catalan words and
+participants translating Spanish words performed equivalently, as
+indicated by the regression coefficient of \emph{Group} (\(\beta\) =
+-0.199, 95\% CrI = {[}-0.512, 0.129{]}, \emph{p}(ROPE) = 0.238).
+Overall, participants responded less accurately to words with more CLPNs
+than to words with fewer CLPNs, regardless of the amount of phonological
+similarity between the presented word and its translation. This is
+indicated by the size of the regression coefficient of the two-way
+interaction between \emph{Cognateness} and \emph{CLPN} (\(\beta\) =
+-0.653, 95\% CrI = {[}-0.973, -0.313{]}, \emph{p}(ROPE) = 0). As
+anticipated, participants' performance benefited from an increase in
+\emph{Cognateness} (\(\beta\) = 7.133, 95\% CrI = {[}6.606, 7.735{]},
+\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
+effect (\(\beta\) = 0.009, 95\% CrI = {[}-0.11, 0.111{]}, \emph{p}(ROPE)
+= 0.925). Figure~\ref{fig-epreds-1} illustrates the posterior of the
+average predictions of the model for words with different values of
+\emph{Cognateness} and \emph{CLPN}. Figure~\ref{fig-coefs} shows a
+graphic summary of the posterior distribution of the regression
+coefficients of interest.
+
+\captionsetup{labelsep=none}
+
+\begin{longtable}{l|rrrrrrr}
+
+\caption{\label{tbl-dataset}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} &  & \multicolumn{3}{c}{Accuracy (\%)} & \multicolumn{3}{c}{Valid trials} \\ 
+\cmidrule(lr){3-5} \cmidrule(lr){6-8}
+\multicolumn{1}{l}{} & N & Mean ± SD & SE & Range & Mean ± SD & N trials & Range \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & $27$ & $15.86$ ± $5.20$ & $3.05$ & $8.82$–$28.71$ & $96.37$ ± $2.88$ ± $2.88$ & $2,602$ & $87$–$98$ \\ 
+cat-ENG & $32$ & $18.48$ ± $4.89$ & $3.27$ & $10.84$–$32.56$ & $81.38$ ± $3.17$ ± $3.17$ & $2,604$ & $71$–$83$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & $21$ & $48.30$ ± $5.29$ & $10.54$ & $38.27$–$58.97$ & $79.14$ ± $3.14$ ± $3.14$ & $1,662$ & $72$–$82$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{8}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & $31$ & $20.92$ ± $8.29$ & $3.76$ & $5.88$–$44.12$ & $97.45$ ± $1.80$ ± $1.80$ & $3,021$ & $88$–$98$ \\ 
+cat-ENG & $32$ & $19.74$ ± $4.94$ & $3.49$ & $10.34$–$27.91$ & $82.75$ ± $0.44$ ± $0.44$ & $2,648$ & $82$–$83$ \\ 
+\bottomrule
+
+\end{longtable}
+
+\begin{figure}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}
+
+}
+
+\caption{\label{fig-epreds-1}}
+
+\end{figure}%
+
+\subsection{Discussion}\label{discussion}
+
+In the present experiment, we investigated the extent to which
+cognateness (phonological similarity between translation equivalents) is
+sufficient for successful word translation, in the absence of semantic
+knowledge. We tested two groups of monolingual British English natives
+in a translation task that involved words in Catalan or Spanish, two
+languages participants reported having no familiarity with. Participants
+benefited strongly from cognateness, when the correct translation of the
+presented words in Catalan or Spanish had few English phonological
+neighbours with higher lexical frequency. This suggests that, in the
+absence of distractors, even naïve participants can efficiently use
+phonological similarity to succeed in a translation task. Our results
+suggest that word-to-word connections at the phonological level might
+play a strong role during L2 speech comprehension, specially in
+low-proficiency listeners.
+
+\section{Experiment 2}\label{experiment-2}
+
+\subsection{Methods}\label{methods-1}
+
+\subsubsection{Participants}\label{participants-1}
+
+Participants in Spain were contacted via announcements in Faculties, and
+were compensated 5€ or an Amazon voucher for the same value.
+Participants gave informed consent before providing any data and the
+study was conducted in accordance with ethical standards of the
+Declaration of Helsinki and the protocol was approved by the Drug
+Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar
+(2020/9080/I). Data collection took place from June 08th, 2020 to June
+28th, 2020. We collected data from 33 Spanish native participants living
+in Spain (\emph{Mean} = 21.85 years, \emph{SD} = 3, Range = 18-33, 28, 5
+female).
+
+\subsubsection{Stimuli}\label{stimuli-1}
+
+Same list of Catalan stimuli as in Experiment 1.
+
+\subsubsection{Procedure}\label{procedure-1}
+
+Same as in Experiment 1.
+
+\subsubsection{Data analysis}\label{data-analysis-1}
+
+Same approach as in Experiment 1.
+
+\subsection{Results}\label{results-1}
+
+We collected data for a total of 5,412 trials completed by 33 unique
+participants. We excluded trials in which participants did not enter any
+text (\emph{n} = 44), in which a response in a language other than
+English was provided (e.g., \texttt{agua}, \emph{n} = 51), in which
+participants did not provide a whole word (e.g., \texttt{f}, \emph{n} =
+7), and in which participants added comments to the experimenter (e.g.,
+\texttt{unsure}, (\emph{n} = 1). In addition, we excluded data from
+participants that self-rated their oral and/or written skills in Catalan
+and Spanish, or any other second language as four or higher in a
+five-point scale (\emph{n} = 22), were diagnosed with a language
+(\emph{n} = 1), or did not contribute more than 80\% of valid trials
+(\emph{n} = 9). After applying trial-level and participant-level
+inclusion criteria, the resulting dataset included 1,662 trials provided
+by 42 unique participants. Of those trials, 1,662 were provided by 21
+unique participants who listened to Catalan words, and 1,662 trials were
+provided by 21 unique participants who listened to Spanish words.
+Responses given by participants were 5.6 characters long on average
+(\emph{Median} = 5, \emph{SD} = 1.6, Range = 2-12).
+
 \textbf{?@tbl-results-1} shows a summary of participants' accuracy
-across Experiments 1, 2, and 3. Overall, participants responded less
-accurately to words with more CLPNs than to words with fewer CLPNs,
-regardless of the amount of phonological similarity between the
-presented word and its translation. This is indicated by the size of the
-regression coefficient of the two-way interaction between
-\emph{Cognateness} and \emph{CLPN} (\(\beta\) = -0.629, 95\% CrI =
-{[}-0.982, -0.344{]}, \emph{p}(ROPE) = 0). As anticipated, participants'
-performance benefited from an increase in \emph{Cognateness} (\(\beta\)
-= 7.11, 95\% CrI = {[}6.567, 7.635{]}, \emph{p}(ROPE) = 0), while the
-number of \emph{CLND} had the opposite effect (\(\beta\) = 0.007, 95\%
-CrI = {[}-0.107, 0.103{]}, \emph{p}(ROPE) = 0.927). A model including
-\emph{Group} as a main effect did not improve the predictive performance
-of the model (\(\Delta_{ELPD}\) = -1.33, \(SE\) = 1.14), indicating that
-participants in both groups performed equivalently.
-Figure~\ref{fig-epreds-1} illustrates the posterior of the average
-predictions of the model for words with different values of
-\emph{Cognateness} and \emph{CLPN}.
-
-\captionsetup{labelsep=none}
-
-\begin{longtable}{l|rrrrrrr}
-
-\caption{\label{tbl-dataset}}
-
-\tabularnewline
-
-\toprule
-\multicolumn{1}{l}{} &  & \multicolumn{3}{c}{Valid trials} & \multicolumn{3}{c}{Accuracy (\%)} \\ 
-\cmidrule(lr){3-5} \cmidrule(lr){6-8}
-\multicolumn{1}{l}{} & N & Mean ± SD & N trials & Range & Mean ± SD & SE & Range \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & $27$ & $96.37$ ± $2.88$ ± $2.88$ & $2,602$ & $87$–$98$ & $15.86$ ± $5.20$ & $3.05$ & $8.82$–$28.71$ \\ 
-cat-ENG & $32$ & $81.38$ ± $3.17$ ± $3.17$ & $2,604$ & $71$–$83$ & $18.48$ ± $4.89$ & $3.27$ & $10.84$–$32.56$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & $31$ & $101.45$ ± $1.80$ ± $1.80$ & $3,145$ & $92$–$102$ & $21.35$ ± $8.10$ & $3.83$ & $6.60$–$44.34$ \\ 
-cat-ENG & $32$ & $85.72$ ± $0.52$ ± $0.52$ & $2,743$ & $84$–$86$ & $20.06$ ± $5.37$ & $3.55$ & $10.00$–$30.00$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & $21$ & $79.14$ ± $3.14$ ± $3.14$ & $1,662$ & $72$–$82$ & $48.30$ ± $5.29$ & $10.54$ & $38.27$–$58.97$ \\ 
-\bottomrule
-
-\end{longtable}
-
-\begin{figure}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}
-
-}
-
-\caption{\label{fig-epreds-1}}
-
-\end{figure}%
-
-\subsection{Discussion}\label{discussion}
-
-\section{Experiment 2}\label{experiment-2}
-
-\subsection{Methods}\label{methods-1}
-
-\subsection{Results}\label{results-1}
-
-\textbf{?@tbl-results-1} shows a summary of participants' accuracy
-across Experiments 1, 2, and 3. Overall, participants responded less
-accurately to words with more CLPNs than to words with fewer CLPNs,
-regardless of the amount of phonological similarity between the
-presented word and its translation. This is indicated by the size of the
-regression coefficient of the two-way interaction between
-\emph{Cognateness} and \emph{CLPN} (\(\beta\) = -0.584, 95\% CrI =
-{[}-0.852, -0.331{]}, \emph{p}(ROPE) = 0). As anticipated, participants'
-performance benefited from an increase in \emph{Cognateness} (\(\beta\)
-= 6.926, 95\% CrI = {[}6.4, 7.461{]}, \emph{p}(ROPE) = 0), while the
-number of \emph{CLND} had the opposite effect (\(\beta\) = 0, 95\% CrI =
-{[}-0.113, 0.085{]}, \emph{p}(ROPE) = 0.95). A model including
-\emph{Group} as a main effect did not improve the predictive performance
-of the model (\(\Delta_{ELPD}\) = -6.56, \(SE\) = 3.21), indicating that
-participants in both groups performed equivalently.
-Figure~\ref{fig-epreds-1} illustrates the posterior of the average
-predictions of the model for words with different values of
-\emph{Cognateness} and \emph{CLPN}.
-
-\begin{figure}
-
-\centering{
-
-\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}
-
-}
-
-\caption{\label{fig-epreds-2}}
-
-\end{figure}%
-
-\subsection{Discussion}\label{discussion-1}
-
-\section{Experiment 3}\label{experiment-3}
-
-\subsection{Methods}\label{methods-2}
-
-Same as in Experiment 1.
-
-\subsection{Results}\label{results-2}
-
-\textbf{?@tbl-results-1} shows a summary of participants' accuracy
-across Experiments 1, 2, and 3. Overall, participants responded less
+across Experiments 1, 2, and 3. MCMC chains in the model showed strong
+evidence of convergence (\(\hat{R}<1.01\)) (see Appendix 2 for more
+detailed model diagnostics). Overall, participants responded less
 accurately to words with more CLPNs than to words with fewer CLPNs,
@@
 
+\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}
+
+}
+
+\caption{\label{fig-epreds-2}}
+
+\end{figure}%
+
+\subsection{Discussion}\label{discussion-1}
+
+Experiment 2 was an extension of Experiment 1 to a population of
+monolinguals whose native language is typologically similar to the
+presented language. Particularly, we presented Catalan words to Spanish
+native speakers who were reportedly unfamiliar with Catalan. Our results
+indicate a similar pattern of results as those in Experiment 1:
+participants were able to provide correct translations of presented
+Catalan words, provided that the Catalan words shared some degree of
+phonological similarity with their Spanish translation, and that the
+number of phonological neighbours with higher lexical frequency was
+reduced. In contrasts with the results in Experiment 1, the positive
+impact of phonological similarity on participants' performance in
+Experiment 2 was more resilient to the interference of phonological
+neighbours. English natives in Experiment 1 provided significantly less
+accuracy responses when more than four phonological neighbors were
+present (even when translating high-similarity words), compared to when
+only one or neighbour were present. Spanish participants in Experiment 2
+benefited from phonological similarity, even when eight neighbours were
+present. Spanish participants' performance declined after 8 neighbours,
+and was evident at 12 neighbours. Overall, this suggests that
+participants in Experiment 2, who were natives of a typologically
+similar language (Spanish) to the presented language (Catalan) benefited
+more strongly from phonological similarity than participants in
+Experiment 1, who were natives of typologically less similar language
+(English) to the presented language (Catalan, Spanish).
+
+Participants from both Experiment 1 and 2 benefited strongly from
+phonological similarity to correctly translate words from a non-native,
+reportedly unfamiliar language. This pattern of results holds for most
+of the presented stimuli, but some low-similarity Catalan and Spanish
+words were responded to surprisingly accurately by English listeners.
+Given that participants were reportedly unfamiliar with both languages,
+it was expected that participants would be very unlikely to provide
+correct translations for words sharing little to no phonological
+similarity to their correct translation. Table~\ref{tbl-surprises} a
+list of Catalan and Spanish words to which participants provided
+responses with \(\geq\) 10 average accuracy.
+
+\begin{longtable}{l|rr}
+
+\caption{\label{tbl-surprises}List of items with unexpectedly high
+accuracy: the Levenshtein similarityscore betwen th presented word (in
+Catalan or Spanish) and their correct Enlgish or Spanish translation is
+zero, but participants, who are reportedly unfamiliar with the presented
+language, were on average \textgreater10\% likely to guess the correct
+translation.}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & Accuracy (\%) & SE \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{3}{l}{Experiment 1 - cat-ENG} \\ 
+\midrule\addlinespace[2.5pt]
+cavall /kəβaʎ/ - horse /hɔs/ & $17.14$ & $6.37$ \\ 
+llibre /ʎiβɾə/ - book /bʊk/ & $17.14$ & $6.37$ \\ 
+camisa /kəmizə/ - shirt /ʃɜt/ & $16.67$ & $6.21$ \\ 
+poma /poma/ - apple /æpl/ & $16.67$ & $6.21$ \\ 
+cama /kamə/ - leg /lɛg/ & $11.11$ & $5.24$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{3}{l}{Experiment 1 - spa-ENG} \\ 
+\midrule\addlinespace[2.5pt]
+pantalón /paŋtalon/ - trousers /traʊzəz/ & $77.42$ & $7.51$ \\ 
+naranja /naɾaŋxa/ - orange /ɒrɪnʤ/ & $41.94$ & $8.86$ \\ 
+leche /leʧe/ - milk /mɪlk/ & $35.48$ & $8.59$ \\ 
+toro /toɾo/ - bull /bʊl/ & $33.33$ & $8.61$ \\ 
+libro /liβɾo/ - book /bʊk/ & $30.00$ & $8.37$ \\ 
+cebra /θebɾa/ - zebra /zibrə/ & $29.03$ & $8.15$ \\ 
+pan /pan/ - bread /brɛd/ & $29.03$ & $8.15$ \\ 
+pollo /poʎo/ - chicken /ʧɪkɪn/ & $26.67$ & $8.07$ \\ 
+jirafa /xiɾafa/ - giraffe /ʤɪrɑf/ & $20.69$ & $7.52$ \\ 
+perro /pero/ - dog /dɒɡ/ & $16.13$ & $6.61$ \\ 
+pluma /pluma/ - feather /fɛðə/ & $16.13$ & $6.61$ \\ 
+puerta /pwerta/ - door /dɔ/ & $16.13$ & $6.61$ \\ 
+pie /pje/ - foot /fʊt/ & $12.90$ & $6.02$ \\ 
+caballo /kaβaʎo/ - horse /hɔs/ & $10.34$ & $5.66$ \\ 
+bocadillo /bokadiʎo/ - sandwich /sænwɪʤ/ & $10.00$ & $5.48$ \\ 
+globo /ɡloβo/ - balloon /bəlun/ & $10.00$ & $5.48$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{3}{l}{Experiment 2 - cat-SPA} \\ 
+\midrule\addlinespace[2.5pt]
+fulla /fuʎə/ - hoja /oxa/ & $30.43$ & $9.59$ \\ 
+ull /uʎ/ - ojo /oxo/ & $21.74$ & $8.60$ \\ 
+got /gɔt/ - vaso /baso/ & $20.00$ & $8.00$ \\ 
+entrepà /entɾəpa/ - bocadillo /bokadiʎo/ & $13.04$ & $7.02$ \\ 
+mirall /miɾaʎ/ - espejo /espexo/ & $12.50$ & $6.75$ \\ 
+\bottomrule
+
+\end{longtable}
+
+It is likely that participants had prior knowledge of these words
+despite having reported little to no familiarity with the presented
+language (Catalan or Spanish). One possibility is that participants had
+previously encountered these words embedded in English linguistic input.
+Spanish words percolate English speech with relative frequency, via
+different sources such as popular culture, songs, TV programs, etc. In
+addition, words from languages other than Spanish, but with high
+similarity to the Spanish words (e.g., cognates from Italian or French)
+might appear in English speech as well. Such prior knowledge might not
+be specific to the low-similarity words highlighted before. Participants
+may also have had prior knowledge about higher-similarity words, which
+could have contributed to participants responding to such words more
+accurately than without such prior knowledge. In the case of
+higher-similarity words, it is more difficult to disentangle the extent
+to which participants' accuracy is a function of pure phonological
+similarity, or prior knowledge they had about the meaning of Spanish
+words. To investigate this issue, we run Experiment 3.
+
+{[}ADD HERE LEXICAL FREQUENCY OF CATALAN AND SPANISH WORDS IN ENGLISH,
+AND OF CATALAN WORDS IN SPANISH{]}
+
+\section{Experiment 3}\label{experiment-3}
+
+Experiment 3 is a replication of Experiment 1, in which we collected
+additional data about participants' prior familiarity with the presented
+Catalan and Spanish words, in addition to the same translation task
+presented to participants in Experiment 1.
+
+\subsection{Methods}\label{methods-2}
+
+\subsubsection{Participants}\label{participants-2}
+
+Data collection took place from October 22th, 2022 to October 23th,
+2022. We collected data from 64 British English native participants
+living in United Kingdom (\emph{Mean} = 22.02 years, \emph{SD} = 2.49,
+Range = 18-26, 36, 28 female).
+
+Participants were recruited via Prolific (5£ compensation) and SONA
+(compensation in academic credits), and gave informed consent before
+providing any data and the study was conducted in accordance with
+ethical standards of the Declaration of Helsinki and the protocol was by
+the University of Oxford Medical Sciences Inter-Divisional Research
+Ethics Committee (IDREC) (R60939/RE007). Participants were asked to
+complete the experiment using a laptop in a quiet place with good
+internet connection. We excluded data from participants that (a)
+self-rated their oral and/or written skills in Catalan and Spanish, or
+any other second language as four or higher in a five-point scale
+(\emph{n} = 2), (b) were diagnosed with a language (\emph{n} = 0), or
+(c) did not contribute more than 80\% of valid trials (\emph{n} = 1).
+
+\subsubsection{Stimuli}\label{stimuli-2}
+
+Same as in Experiment 1.
+
+\subsubsection{Procedure}\label{procedure-2}
+
+{[}TO BE COMPLETED{]}
+
+\subsubsection{Data analysis}\label{data-analysis-2}
+
+Same approach as in Experiment 1.
+
+\subsection{Results}\label{results-2}
+
+We collected data for a total of 6,016 trials completed by 64 unique
+participants. Of those trials, 2,752 were provided by 32 unique
+participants who listened to Catalan words, and 3,264 trials were
+provided by 32 unique participants who listened to Spanish words. We
+excluded trials in which participants did not enter any text (\emph{n} =
+), in which a response in a language other than English was provided
+(e.g., \texttt{agua}, \emph{n} = ), in which participants did not
+provide a whole word (e.g., \texttt{f}, \emph{n} = ), and in which
+participants added comments to the experimenter (e.g., \texttt{unsure},
+(\emph{n} = ). In addition, we excluded data from participants that
+self-rated their oral and/or written skills in Catalan and Spanish, or
+any other second language as four or higher in a five-point scale
+(\emph{n} = 2), were diagnosed with a language (\emph{n} = 0), or did
+not contribute more than 80\% of valid trials (\emph{n} = 1). After
+applying trial-level and participant-level inclusion criteria, the
+resulting dataset included 6,290 trials provided by 62 unique
+participants. Of those trials, 3,145 were provided by 31 unique
+participants who listened to Catalan words, and 2,743 trials were
+provided by 32 unique participants who listened to Spanish words.
+
+Responses given by English participants to Catalan presented words were
+5.52 characters long on average (\emph{Median} = 5, \emph{SD} = 1.75,
+Range = 1-17), while their translations to Spanish responses were 5.41
+characters long on average (\emph{Median} = 5, \emph{SD} = 1.77, Range =
+1-20).
+
+MCMC chains in the model showed strong evidence of convergence
+(\(\hat{R}<1.01\)) (see Appendix 2 for more detailed model diagnostics).
+Overall, participants reported prior knowledge more often for that
+Spanish words with unexpectedly high accuracy (see Discussion in
+Experiment 2) than for words with expected accuracy (see
+Figure~\ref{fig-knowledge}). Participants reported prior knowledge of
+Catalan words with unexpected accuracy as often as those with expected
+accuracy. This suggests that participants in Experiment 1 may have
+relied, to some extent, on their prior knowledge about form-meaning
+mappings to correctly translate some Spanish words. To isolate such an
+effect of prior Spanish knowledge, we run the same analysis as in
+Experiment 1 on the newly collected translations from Experiment 3, now
+excluding responses to words in which participants reported prior
+knowledge.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics{manuscript_files/figure-pdf/fig-knowledge-1.pdf}
+
+}
+
+\caption{\label{fig-knowledge}Catalan/Spanish prior word knowledge as
+reported by English native participants in Experiment 3. (A) Average
+proportion of participants that reported prior knowledge for words with
+surprisingly high accuracy (no phonological similarity with the correct
+translation, accuracy higher than 10\%), and for words with expected
+accuracy (low similarity-low accuracy, or high similarity-high
+accuracy). (B) Average accuracy for words with expected and unexpected
+accuracy.}
+
+\end{figure}%
+
+Participants translating Catalan words and participants translating
+Spanish words, as indicated by the regression coefficient of
+\emph{Group} (\(\beta\) = -0.049, 95\% CrI = {[}-0.388, -0.32{]},
+\emph{p}(ROPE) = 0.426). Overall, participants responded less accurately
+to words with more CLPNs than to words with fewer CLPNs, regardless of
+the amount of phonological similarity between the presented word and its
+translation. This is indicated by the size of the regression coefficient
+of the two-way interaction between \emph{Cognateness} and \emph{CLPN}
+(\(\beta\) = -0.665, 95\% CrI = {[}-1.068, -0.32{]}, \emph{p}(ROPE) =
+0.002). As anticipated, participants' performance benefited from an
+increase in \emph{Cognateness} (\(\beta\) = 7.585, 95\% CrI = {[}7.043,
+8.24{]}, \emph{p}(ROPE) = 0), while the number of \emph{CLND} had the
+opposite effect (\(\beta\) = -0.049, 95\% CrI = {[}-0.203, 0.086{]},
+\emph{p}(ROPE) = 0.753). Figure~\ref{fig-epreds-1} illustrates the
+posterior of the average predictions of the model for words with
+different values of \emph{Cognateness} and \emph{CLPN}.
+
+\begin{figure}
+
+\centering{
+
 \captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}
@@
 \subsection{Discussion}\label{discussion-2}
+
+\section{Joint analyses}\label{joint-analyses}
+
+Across Experiments 1 and 3, we found strong evidence that participants
+efficiently exploited phonological similarity to provide accurate
+translations for words in an unfamiliar language, provided that few
+phonological neighbours of higher lexical frequency were present.
+Figure~\ref{fig-coefs} summarizes the posterior distribution of the
+regression coefficients of the models in Experiments 1 to 3.
+
+\begin{figure}
+
+\centering{
+
+\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-coefs-1.pdf}
+
+}
+
+\caption{\label{fig-coefs}}
+
+\end{figure}%
+
+\section{General discussion}\label{general-discussion}
 
@@
 
+\bibitem[\citeproctext]{ref-kruschke2018bayesian}
+Kruschke, J. K., \& Liddell, T. M. (2018). The bayesian new statistics:
+Hypothesis testing, estimation, meta-analysis, and power analysis from a
+bayesian perspective. \emph{Psychonomic Bulletin \& Review}, \emph{25},
+178--206.
+
 \bibitem[\citeproctext]{ref-lecumberri2010non}

### 2024-05-08 R/datasets.R

@@ -24,12 +24,12 @@ get_dataset <- function(experiment, exp_responses, quest_responses, stimuli) {
             mutate(group = factor(group, levels = c("cat-ENG", "spa-ENG"))) 
         contrasts(out$group) <- c(0.5, -0.5)
     } else if (experiment==2) {
+        out <- dplyr::filter(dataset, group=="cat-SPA")
+    } else {
         out <- dataset |> 
             dplyr::filter(source=="Questionnaire") |> 
             mutate(group = factor(group, levels = c("cat-ENG", "spa-ENG")))
         contrasts(out$group) <- c(0.5, -0.5)
-    } else {
-        out <- dplyr::filter(dataset, group=="cat-SPA")
     }
     
     if (nrow(out)==0) {
@@ -53,7 +53,7 @@ prepare_data <- function(exp_responses, quest_responses, stimuli) {
         bind_rows(.id = "source") |> 
         dplyr::filter(valid_participant, valid_response) |> 
         select(any_of(relevant_vars)) |> 
-        left_join(stimuli_tmp, by = join_by(group, word_1)) |> 
+        inner_join(stimuli_tmp, by = join_by(group, word_1)) |> 
         mutate(across(c(freq_zipf_2, neigh_n, neigh_n_h, avg_sim, avg_sim_h, lv),
                       \(x) scale(x)[, 1],
                       .names = "{.col}_std"),

### 2024-05-08 R/questionnaire.R

@@ -14,7 +14,7 @@ get_quest_raw <- function(quest_raw_files){
     dplyr::filter(response_type=="IP Address") |> 
     rename(any_of(colnames_dict())) |> 
     select(participant_id, language, date, finished, consent, captcha,
-           age, sex, education,
+           age, gender, education,
            l1, l2, l3, l2_oral_comp, l2_writ_prod, l3_oral_comp, l3_writ_prod,
            cat_oral_comp, cat_writ_prod, spa_oral_comp, spa_writ_prod, 
            living_spa, living_cat, has_language_problems,
@@ -55,7 +55,8 @@ colnames_dict <- function() {
     city = "what_city_do_you_live_in",
     env_location = "where_are_you_completing_this_study_selected_choice",
     env_noise = "how_noisy_is_the_environment_in_which_you_are_completing_this_experiment",
-    has_vision_problems = "do_you_have_normal_or_corrected_to_normal_vision")
+    has_vision_problems = "do_you_have_normal_or_corrected_to_normal_vision",
+    gender = "sex")
 }
 
 #' Process responses from Experiment 2
@@ -130,7 +131,7 @@ get_quest_participants <- function(quest_processed,
   max_cat_trials <- 86 # max of Catalan trials
   max_spa_trials <- 103 # max number of Spanish trials
   
-  responses_coded_path <- file.path("data-raw", "questionnaire", "04_responses-coded.csv")
+  responses_coded_path <- file.path("data", "questionnaire-manual-coded.csv")
   
   quest_participants <- read_csv(responses_coded_path,
                                  col_types = "cccclicc")  |> 
@@ -149,7 +150,7 @@ get_quest_participants <- function(quest_processed,
               .by = participant_id) |>  
     inner_join(extra_info, by = join_by(participant_id)) |>
     validate_participants(min_valid_trials, blocked_languages, age_range) |> 
-    select(group, participant_id, date, age, 
+    select(group, participant_id, date, age, gender,
            l2, l2_oral_comp, l2_writ_prod,
            cat_oral_comp, cat_writ_prod,
            spa_oral_comp, spa_writ_prod,

### 2024-05-08 _targets.R

@@ -105,63 +105,33 @@ list(
     # analysis of Experiment 1
     tar_target(exp_1_m0,
                get_model_fit("exp_1_m0",
-                             correct ~ freq_zipf_2 + neigh_n_h * lv + 
-                                 (1 + freq_zipf_2 + neigh_n_h * lv | participant_id),
-                             prior = model_prior,
-                             data = dataset_1)),
-    tar_target(exp_1_m1,
-               get_model_fit("exp_1_m1",
                              correct ~ freq_zipf_2 + neigh_n_h * lv + group + 
                                  (1 + freq_zipf_2 + neigh_n_h * lv | participant_id),
                              prior = model_prior,
                              data = dataset_1)),
     
+    
     # analysis of Experiment 2
     tar_target(exp_2_m0,
                get_model_fit("exp_2_m0",
-                             correct ~ freq_zipf_2 + neigh_n_h * lv +
+                             correct ~ freq_zipf_2 + neigh_n_h * lv + 
                                  (1 + freq_zipf_2 + neigh_n_h * lv | participant_id),
                              prior = model_prior,
                              data = dataset_2)),
-    
-    tar_target(exp_2_m1,
-               get_model_fit("exp_2_m1",
-                             correct ~ freq_zipf_2 + neigh_n_h * lv * group + 
-                                 (1 + freq_zipf_2 + neigh_n_h * lv | participant_id),
-                             prior = model_prior,
-                             data = dataset_2)),
-    
-    tar_target(exp_2_m2,
-               get_model_fit("exp_2_m2",
-                             correct ~ freq_zipf_2 + neigh_n_h * lv + confidence +  
-                                 (1 + freq_zipf_2 + neigh_n_h * lv + confidence | participant_id),
-                             prior = model_prior,
-                             data = dataset_2)),
-    tar_target(exp_2_m3,
-               get_model_fit("exp_2_m3",
-                             correct ~ freq_zipf_2 + neigh_n_h * lv + knowledge +  
-                                 (1 + freq_zipf_2 + neigh_n_h + lv + knowledge | participant_id),
-                             prior = model_prior,
-                             data = dataset_2)),
-    
     
     # analysis of Experiment 3
     tar_target(exp_3_m0,
                get_model_fit("exp_3_m0",
-                             correct ~ freq_zipf_2 + neigh_n_h * lv + 
[…]

### 2024-05-09 manuscript/appendix.qmd
Add Methods section to Experiment 3

--- 2021-12-16 Rmd/appendix.Rmd
@@
----
+## Appendix 1: MCMC diagnostics
 
----
+```{r fig-appendix-diagnostics}
+#| label: fig-appendix-diagnostics
+#| fig-width: 9
+#| fig-height: 9
+var_labels <- c("b_Intercept" = "Intercept",
+                "b_freq_zipf_2" = "Frequency",
+                "b_lv" = "Cognateness",
+                "b_neigh_n_h" = "CLPN",
+                "b_neigh_n_h:lv" = "Cognateness × CLPN",
+                "b_group1" = "cat-ENG vs. spa-ENG")
 
-# Appendix 1: MCMC diagnostics
-
-```{r appendix-diagnostics, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-
-suppressWarnings({
-  bayesplot_theme_set(
-    new = theme_custom() + 
-      theme(
-        axis.text.x = element_text(size = 8), 
-        axis.text.y = element_text(size = 8),
-        strip.text = element_text(size = 10),
-        legend.key = element_rect(colour = NA, fill = "white")
-      )
-  )
-})
-
-color_scheme_set("brightblue")
-pars <- c("b_Intercept", "b_frequency_zipf", "b_pthn", "b_lv", "b_pthn:lv")
-pars_labels <- c(
-  "Intercept",
-  "Frequency (+1 SD)",
-  "PTHN \u00d7 Levenshtein", 
-  "Levenshtein (+1 SD)",
-  "PTHN (+1 SD)"
-)
-```
-
-## Pairs plot
-
-```{r appendix-mcmc, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-posterior_cp <- as.array(model_fits$fit_4)
-np_cp <- nuts_params(model_fits$fit_4)
-
-plot_pairs <- mcmc_pairs(
-  posterior_cp, pars = pars,
-  off_diag_args = list(size = 0.75),
-  diag_args = list(facet_args = list(labeller = label_value(labels = pars_labels)))
-)
-
-plot_pairs
+list(exp_1_m0, exp_2_m0, exp_3_m1) |> 
+    set_names(paste0("Experiment ", 1:3)) |> 
+    map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |> 
+    dplyr::filter(.variable != "b_Intercept") |> 
+    mutate(.value = .value/4,
+           .variable = factor(.variable, 
+                              levels = names(var_labels),
+                              labels = var_labels)) |> 
+    ggplot(aes(.iteration, .value, colour = factor(.chain))) +
+    facet_grid(.variable~exp, scales = "free") +
+    geom_line() +
+    labs(x = "Iteration",
+         y = "Value",
+         colour = "Chain",
+         fill = "Chain",
+         linetype = "Chain") +
+    scale_colour_brewer(palette = "Reds") +
+    theme(legend.position = "top",
+          legend.box = "horizontal",
+          panel.border = element_rect(fill = NA,
+                                      colour = "grey"),
+          legend.justification = c(1, 0),
+          axis.title.x = element_blank(),
+          axis.title.y = element_blank())
 ```
@@
 ```{r appendix-trace, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-(mcmc_trace(model_fits$fit_4, pars) + legend_none()) /
-  mcmc_rank_overlay(model_fits$fit_4, pars = pars) +
+(bayesplot::mcmc_trace(exp_1_m0, names(var_labels)) + bayesplot::legend_none()) /
+  bayesplot::mcmc_rank_overlay(exp_1_m0, pars = names(var_labels)) +
   plot_layout(guides = "collect") &
-  theme(
-    legend.position = "top"
-  )
-```
-
-## Gelman-Rubin convergence criterion (R-hat)
-
-```{r appendix-rhat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-rhats <- rhat(model_fits$fit_4)
-ratios <- neff_ratio(model_fits$fit_4)
-
-mcmc_rhat_hist(rhats, binwidth = 0.00025, color = "white") + 
-  mcmc_neff_hist(ratios, binwidth = 0.1) +
-  plot_layout(guides = "keep", ncol = 1) &
   theme(
@@
 
-include_graphics(here("Figures", "accuracy_spacat.png"))
+include_graphics(here("img", "accuracy_spacat.png"))
 

### 2024-05-09 manuscript/manuscript.qmd
Add Methods section to Experiment 3

@@
 
-All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
+All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/?view_only=aab7636ce1af48cf832596a7ea9101c5/) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
 
@@
               across(age, lst(mean, sd, min, max)),
-              knows_l2 = sum(l2 != "None"),
-              l2_lst = lst(unique(l2)[unique(l2)!="None"]),
-              .by = c(group, exp))
+              l2_lst = list(l2),
+              .by = c(group, exp)) |> 
+    mutate(l2_lst = map(l2_lst, 
+                        function(x) {
+                            y <- table(x)
+                            y <- y[names(y) != "None"]
+                            y <- paste0(names(y), " (", y, ")")
+                            return(y)
+                        }),
+           n_excluded = paste0(n, " (", n_excluded, ")")) |> 
+    select(-n)
 
@@
     fmt_integer(matches("min|max")) |> 
+    cols_merge_uncert(age_mean, age_sd) |> 
     cols_merge_range(age_min, age_max) |> 
     tab_spanner("Age", matches("age")) |> 
-    tab_spanner("Sample size", starts_with("n")) |> 
-    tab_spanner("L2", matches("l2")) |> 
     cols_label(group = "Group",
                exp = "Experiment",
-               n = "N",
-               n_excluded = "Excluded",
-               age_mean = "M",
-               age_sd = "SD",
+               n_excluded = "N",
+               age_mean = "Mean ± SD ",
                age_min = "Range",
-               knows_l2 = "N knows L2?",
-               l2_lst = "Which L2?") |> 
+               l2_lst = "L2") |> 
     tab_style(cell_text(style = "italic"),
@@
     tab_style(cell_text(align = "left"),
-              list(cells_body(), cells_column_labels()))
+              list(cells_body(), cells_column_labels())) |> 
+    tab_style(cell_text(weight = "bold",
+                        align = "center",
+                        style = "normal"),
+              cells_column_labels(c(l2_lst, n_excluded))) |> 
+    tab_footnote(
+        "Number of included participants (number of excluded participants.)", 
+        locations = cells_column_labels(n_excluded))
 ```
@@
 We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
-Words in the Catalan list were `r round(lengths$cat_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_flat_1_sd, 2)`, *Range* = `r round(lengths$cat_eng$ipa_flat_1_min, 2)`-`r round(lengths$cat_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations (which participants had to type), were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, *Range* = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`.
-Words in the Spanish list were `r round(lengths$spa_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_flat_1_sd, 2)`, *Range* = `r round(lengths$spa_eng$ipa_flat_1_min, 2)`-`r round(lengths$spa_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations (which participants had to type), were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, *Range* = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`.
+Words in the Catalan list were `r round(lengths$cat_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$cat_eng$ipa_flat_1_min, 2)`-`r round(lengths$cat_eng$ipa_flat_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, Range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`.
+Words in the Spanish list were `r round(lengths$spa_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$spa_eng$ipa_flat_1_min, 2)`-`r round(lengths$spa_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations (which participants had to type), were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, Range = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`.
 
@@
 Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, *Range* = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
-The average duration of the Catalan audio files was `r round(durations$spa_eng$duration_mean, 2)` (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, *Range* = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
-
-```{r}
+The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, Range = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
+The average duration of the Catalan audio files was `r round(durations$spa_eng$duration_mean, 2)` (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, Range = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
+
+```{r exc-stimuli}
 stim_exc_freq <- stimuli |> 
@@
 
-For each translation pair, we defined three predictors of interest: the lexical frequency of the correct translation (*Frequency*), the phonological similarity between the presented (Catalan or Spanish) word and their correct English translation (*Cognateness*), and presented word's number of cross-language phonological neighbours (*CLPN*).
-
-We included *Frequency* under the hypothesis that, keeping other predictors constant, participants would translate higher-frequency words more accurately and faster than lower-frequency words.
-Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores [@van2014subtlex].
+For each translation pair, we defined three predictors of interest: the lexical frequency of the correct translation (*Frequency*), the phonological similarity between the presented (Catalan or Spanish) word and their correct English translation (*Cognateness*), and the presented word's number of cross-language phonological neighbours (*CLPN*).
+
+We included *Frequency* as a nuance predictor, under the hypothesis that---keeping other predictors constant---participants would translate higher-frequency words more accurately and faster than lower-frequency words.
+Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores.
 Words in the stimuli list without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
@@
 We calculated *Cognatenesss*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
-The Levenshtein distance computes the edit distance between two character strings (in this case, phoneme transcriptions) by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
+The Levenshtein distance computes the edit distance between two character strings---in this case, phoneme transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
 We divided this edit distance by the number of characters included in the longest X-SAMPA transcription of the translation pair.
@@
 For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
-For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (trEIn)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
 Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
@@
 For each Catalan and Spanish word, we calculated the number of *CLPN* by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) different in up to one phoneme from that of the presented Catalan or Spanish word.
-Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological trnascriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
+Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
 
@@
 #| label: tbl-stimuli
+#| tbl-cap: "Stimuli properties."
 stimuli |> 
@@
     tab_style(cell_text(style = "italic"),
-              cells_column_labels())
-
+              cells_column_labels()) |> 
+    grand_summary_rows(columns = matches("mean|sd"),
+                       fns = list(label = md("**Mean**"), 
+                                  id = "totals",
+                                  fn = "mean"),
+                       fmt = ~fmt_number(., n_sigfig = 3, suffixing = TRUE))
 ```
@@
 Then, participants pressed the `RETURN/ENTER` key to confirm their answer and start and new trial.
-We excluded trials where the participant's answer was not a valid word in the target language or no answer was given.
-Trials in which the response was mistyped by only one character were counted as correct (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct word.
+
+![Schematic representation of a trial in the experimental task](_assets/img/design.png)
+
+### Data analysis
+
+#### Data processing
+
+After data collection, participants answers were manually coded into the following categories: *Correct*, *Typo*, *Wrong*, *False friend*, *Other*.
+Responses were coded as *Correct* if the provided string of characters was identical to the orthographic form of the correct translation.
+Trials were coded as *Typo* if the participant provided a string of characters only one edit distance (addition, deletion, or substitution) apart from the orthographic form of the correct translation (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct English word.
+Responses were coded as *False friend* if the participant provided a phonologically similar incorrect translation.
+Responses were coded as *Other* (see Data analysis section for more details).
+Both *Correct* and *Typo* responses were considered as correct, while *Wrong* and *False friend* responses were considered as incorrect.
+*Other* responses were excluded from data analysis.
+
 Trials in which participants took longer than 10 seconds to respond were also excluded.
@@
 
-```{r}
-#| label: procedure-figure
-#| echo: false
-#| message: false
-#| warning: false
-#| out-width: "80%"
-#| fig-cap: "Schematic representation of a trial in the experimental task."
-knitr::include_graphics(here::here("img", "design.png"))
-```
-
-### Data analysis
+#### Modelling approach and statistical inference
 
@@
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+
 After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
@@
     select(exp, group, correct, typo, wrong, false_friend,
-           matches("prop"))
+           matches("prop")) |> 
+    relocate(correct, correct_prop,
+             typo, typo_prop,
+             wrong, wrong_prop,
+             false_friend, false_friend_prop)
 
@@
     tab_spanner("Correct responses", 
-                c(correct, typo)) |> 
+                matches("correct|typo")) |> 
     tab_spanner("Incorrect responses", 
-                c(wrong, false_friend)) |> 
+                matches("wrong|false")) |> 
     fmt_integer(c(correct, typo, wrong, false_friend)) |> 
-    fmt_percent(is.double) |> 
-    cols_merge_n_pct(correct, col_pct = correct_prop) |> 
-    cols_merge_n_pct(typo, col_pct = typo_prop) |> 
-    cols_merge_n_pct(wrong, col_pct = wrong_prop) |> 
-    cols_merge_n_pct(false_friend, col_pct = false_friend_prop) |>
-
+    fmt_number(is.double, pattern = "({x})", scale_by = 100) |> 
     cols_label(correct = "Correct",
                typo = "Typo",
-               wrong = "Wrong translation",
-               false_friend = "False friend") |> 
+               wrong = "Wrong",
+               false_friend = "False friend",
+               correct_prop = "(%)",
+               typo_prop = "(%)",
+               wrong_prop = "(%)",
+               false_friend_prop = "(%)") |> 
     tab_style(cell_text(weight = "bold"), 
@@
     tab_style(cell_text(style = "italic"), 
+              cells_column_labels()) |> 
+    summary_rows(columns = is.integer,
+                 fns = list(label = md("*Sum*"),
+                            id = "totals",
+                            fn = "sum"),
+                 fmt = ~fmt_integer(.)) |>
+    summary_rows(columns = matches("prop"),
+                 fns = list(label = md("*Mean*"), 
+                            id = "means",
+                            fn = "mean"),
+                 fmt = ~fmt_number(., scale_by = 100)) |> 
+    tab_style(cell_text(align = "center"),
               cells_column_labels())
@@
 
-```{r}
+```{r coefs-1}
 c1 <- gather_draws(exp_1_m0, `b_.*`, `sd_.*`, regex = TRUE) |> 
@@
     tab_spanner("Valid trials", matches("n_trials")) |> 
-    cols_merge_uncert(n_trials_mean, n_trials_sd) |> 
     cols_merge_range(n_trials_min, n_trials_max) |> 
-    cols_merge_uncert(n_trials_mean, n_trials_sd) |> 
-    cols_merge_uncert(correct_mean, correct_sd) |> 
     cols_merge_range(correct_min, correct_max) |> 
@@
                n_participants = "N",
-               n_trials_mean = "Mean ± SD",
+               n_trials_mean = "Mean",
+               n_trials_sd = "SD",
                n_trials_min = "Range",
-               correct_mean = "Mean ± SD",
+               correct_mean = "Mean",
+               correct_sd = "SD",
                correct_sem = "SE",
                correct_min = "Range") |> 
+    summary_rows(columns = is.integer,
+                 fns = list(label = md("*Sum*"),
+                            id = "totals",
+                            fn = "sum"),
+                 fmt = ~fmt_integer(.)) |>
+    summary_rows(columns = matches("mean|sd|sem"),
+                 fns = list(label = md("*Mean*"), 
+                            id = "means",
+                            fn = "mean"),
+                 fmt = ~fmt_number(., scale_by = 100)) |> 
+    tab_style(cell_text(align = "center"),
+              cells_column_labels()) |> 
     tab_style(cell_text(style = "italic"),
@@
 #| fig-width: 10
-get_epreds <- function(model, data, n = 100, 
+#| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
+Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Cognateness* (1-100).
+Predictions are plotted separately for English participants translating Catalan words, and for English participants translating Spanish words.
+Lines indicate mean predictions, and intervals indicate 95%, 89%, 78%, 67%, and 50% credible intervals (CrI)."
+get_epreds <- function(model, data, n = 1000, 
                        lv = seq(0, 1, length.out = 100), 
@@
 In the present experiment, we investigated the extent to which cognateness (phonological similarity between translation equivalents) is sufficient for successful word translation, in the absence of semantic knowledge.
-We tested two groups of monolingual British English natives in a translation task that involved words in Catalan or Spanish, two languages participants reported having no familiarity with.
+We tested two groups of monolingual British English natives in a translation task that involved words in Catalan or Spanish, two languages participants reported having no prior familiarity with.
 Participants benefited strongly from cognateness, when the correct translation of the presented words in Catalan or Spanish had few English phonological neighbours with higher lexical frequency.
@@
 
+Results in Experiment 1 suggest that English natives were able to exploit the phonological similarity between unfamiliar words in Catalan and Spanish to provide accurate translations to English.
+English, a Germanic language, is relatively distant from Catalan and Spanish, two Romance languages.
+English shares fewer cognates with Catalan and Spanish than it does with typologically closer languages, like Dutch and German.
+In Experiment 2, we investigated whether listeners of an unfamiliar but typologically closer language benefit more strongly from phonological similarity when performing the same task as in Experiment 1.
+To this aim, we presented Spanish participants, who reported little-to-no prior familiarity with Catalan, with Catalan words.
+
 ## Methods
-
-### Participants
 
@@
 We collected data from `r n_participants` Spanish native participants living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
-
-### Stimuli
-
-Same list of Catalan stimuli as in Experiment 1.
-
-### Procedure
-
-Same as in Experiment 1.
-
-### Data analysis
-
-Same approach as in Experiment 1.
+Stimuli were the same list of Catalan stimuli as in Experiment 1.
+Procedure and data analysis were identical as in Experiment 1.
 
@@
 
-@tbl-results-1 shows a summary of participants' accuracy across Experiments 1, 2, and 3.
 MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
@@
 #| label: fig-epreds-2
+#| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
+Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Cognateness* (1-100).
+Predictions are plotted separately for English participants translating Catalan words, and for English participants translating Spanish words.
+Lines indicate mean predictions, and intervals indicate 95%, 89%, 78%, 67%, and 50% credible intervals (CrI)."
 #| fig-height: 3.5
@@
 
-### Participants
-
 ```{r participants-numbers-3}
@@
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-We excluded data from participants that (a) self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), (b) were diagnosed with a language (*n* = `r excluded_impairment`), or (c) did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
-
-### Stimuli
-
-Same as in Experiment 1.
-
-### Procedure
-
-[TO BE COMPLETED]
-
-### Data analysis
-
-Same approach as in Experiment 1.
+Stimuli were the same list of Catalan stimuli as in Experiment 1.
+
+The experiment was implemented online using Qualtrics (Qualtrics, Provo, UT).
+This platform was chosen to allow easier presentation of survey questions aimed to probe prior understanding of the presented words and participants’ confidence ratings of their answers.
+With the exception of these additional questions, we attempted to replicate the procedure of Experiment 1 as closely as possible.
+The Spanish and Catalan audio stimuli used were identical the materials in Experiment 1.
+Participants were randomly assigned to the Catalan or Spanish lists.
+The Catalan list had 83 trials and the Spanish list had 99 trials.
+Participants first completed the consent form followed by the questionnaire about demographic status, language background and set up.
+They then proceeded to the experimental task.
+
+On each trial, participants listened to the audio stimulus by clicking on the play button.
+For comparability to the PsychoPy version, participants were only allowed to play the audio one time.
+Participants were explicitly told that they would be only allowed to listen once.
+The play button vanished after one playthrough.
+Participants then had to answer three questions based on the audio they had heard on that trial.
+These questions were presented on the same page, directly below the audio player.
+They were first asked whether or not they knew the presented word (multiple choice---yes/no).
+Regardless of their answer on the first question, participants were asked what they thought the translation of the word was in English (or their best guess), and instructed to type their answer in the provided text box.
+Finally, they were asked to rate how confident they were in their answer on a scale of 0 to 7, where 7 was "very confident" and 0 was "not confident".
+There was no time limit on the response phase.
+All questions had to be answered to proceed to the next trial.
+
+Participants first completed 5 practice trials with English words as the audio stimulus (ambulance, cucumber, elephant, pear, turtle).
+The words were recorded by a female native speaker of English.
+These trials acted as attention checks, as participants should always answer "yes" to the first question on prior word knowledge and be able to accurately transcribe the word they heard.
+Following the practice phase, participants completed the test phase where they heard either Spanish words or Catalan words.
 
@@
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+
 After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
 Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
+
+INSERT PRIOR KNOWLEDGE FILTER
 
@@
 #| label: fig-epreds-3
+#| fig-cap: "Posterior model-predicted mean accuracy in Experiment 2.
+Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Cognateness* (1-100).
+Predictions are plotted separately for English participants translating Catalan words, and for English participants translating Spanish words.
+Lines indicate mean predictions, and intervals indicate 95%, 89%, 78%, 67%, and 50% credible intervals (CrI)."
 #| fig-height: 6

### 2024-05-09 manuscript/manuscript.tex
Add Methods section to Experiment 3

@@
 
-\begin{longtable}{l|rrrrrrc}
+\setlength{\LTpost}{0mm}
+
+\begin{longtable}{l|rrrc}
 
@@
 \toprule
-\multicolumn{1}{l}{} & \multicolumn{2}{c}{Sample size} & \multicolumn{3}{c}{Age} & \multicolumn{2}{c}{L2} \\ 
-\cmidrule(lr){2-3} \cmidrule(lr){4-6} \cmidrule(lr){7-8}
-\multicolumn{1}{l}{} & N & Excluded & M & SD & Range & N knows L2? & Which L2? \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & 35 & 8 & $21.80$ & $2.08$ & $18$–$26$ & 1 & Russian \\ 
-cat-ENG & 36 & 4 & $21.72$ & $2.25$ & $18$–$25$ & 5 & German, Punjabi, French, Italian, Several \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & 33 & 12 & $21.85$ & $3.00$ & $18$–$33$ & 12 & Italian, French, German \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & 32 & 1 & $21.72$ & $2.59$ & $18$–$26$ & 3 & Japanese, German \\ 
-cat-ENG & 32 & 0 & $22.31$ & $2.39$ & $18$–$26$ & 2 & Cantonese, Irish \\ 
+\multicolumn{1}{l}{} &  & \multicolumn{2}{c}{Age} &  \\ 
+\cmidrule(lr){3-4}
+\multicolumn{1}{l}{} & N\textsuperscript{\textit{1}} & Mean ± SD  & Range & L2 \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & 35 (8) & $21.80$ ± $2.08$ & $18$–$26$ & Russian (1) \\ 
+cat-ENG & 36 (4) & $21.72$ ± $2.25$ & $18$–$25$ & French (1), German (1), Italian (1), Punjabi (1), Several (1) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & 33 (12) & $21.85$ ± $3.00$ & $18$–$33$ & French (9), German (1), Italian (2) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{5}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & 32 (1) & $21.72$ ± $2.59$ & $18$–$26$ & German (2), Japanese (1) \\ 
+cat-ENG & 32 (0) & $22.31$ ± $2.39$ & $18$–$26$ & Cantonese (1), Irish (1) \\ 
 \bottomrule
@@
 \end{longtable}
+
+\begin{minipage}{\linewidth}
+\textsuperscript{\textit{1}}Number of included participants (number of excluded participants.)\\
+\end{minipage}
 
@@
 the auditory modality: one in Catalan and one in Spanish. Words in the
-Catalan list were 5.02 phonemes long on average (\emph{SD} = 1.49,
-\emph{Range} = 2-8), and the orthographic form of their English
-translations (which participants had to type), were 5.12 characters long
-on average (\emph{SD} = 1.56, \emph{Range} = 3-9. Words in the Spanish
-list were 5.52 phonemes long on average (\emph{SD} = 1.47, \emph{Range}
-= 3-9), and the orthographic form of their English translations (which
-participants had to type), were 5.29 characters long on average
-(\emph{SD} = 1.77, \emph{Range} = 3-12.
+Catalan list were 5.02 phonemes long on average (\emph{SD} = 1.49, Range
+= 2-8), and the orthographic forms of their English translations (which
+participants had to type) were 5.12 characters long on average
+(\emph{SD} = 1.56, Range = 3-9. Words in the Spanish list were 5.52
+phonemes long on average (\emph{SD} = 1.47, Range = 3-9), and the
+orthographic form of their English translations (which participants had
+to type), were 5.29 characters long on average (\emph{SD} = 1.77, Range
+= 3-12.
 
@@
 Weenink, 2021}). The average duration of the Catalan audio files was
-1.24 (\emph{SD} = 0.19, \emph{Range} = 0.8-1.58). The average duration
-of the Catalan audio files was 1.16 (\emph{SD} = 0.15, \emph{Range} =
-0.78-1.53).
+1.24 (\emph{SD} = 0.19, Range = 0.8-1.58). The average duration of the
+Catalan audio files was 1.16 (\emph{SD} = 0.15, Range = 0.78-1.53).
 
@@
 phonological similarity between the presented (Catalan or Spanish) word
-and their correct English translation (\emph{Cognateness}), and
+and their correct English translation (\emph{Cognateness}), and the
 presented word's number of cross-language phonological neighbours
@@
 
-We included \emph{Frequency} under the hypothesis that, keeping other
-predictors constant, participants would translate higher-frequency words
-more accurately and faster than lower-frequency words. Lexical
-frequencies of correct translations were extracted from SUBTLEX-UK
-(\citeproc{ref-van2014subtlex}{Van Heuven et al., 2014}), and
-transformed to Zipf scores (\citeproc{ref-van2014subtlex}{Van Heuven et
-al., 2014}). Words in the stimuli list without a lexical frequency value
-were excluded from data analysis (2 in the Catalan list, 5 in the
-Spanish list).
+We included \emph{Frequency} as a nuance predictor, under the hypothesis
+that---keeping other predictors constant---participants would translate
+higher-frequency words more accurately and faster than lower-frequency
+words. Lexical frequencies of correct translations were extracted from
+SUBTLEX-UK (\citeproc{ref-van2014subtlex}{Van Heuven et al., 2014}), and
+transformed to Zipf scores. Words in the stimuli list without a lexical
+frequency value were excluded from data analysis (2 in the Catalan list,
+5 in the Spanish list).
 
@@
 (\citeproc{ref-van2014stringdist}{van der Loo, 2014}). The Levenshtein
-distance computes the edit distance between two character strings (in
-this case, phoneme transcriptions) by counting the number of additions,
+distance computes the edit distance between two character strings---in
+this case, phoneme transcriptions---by counting the number of additions,
 deletions, and substitutions necessary to make both strings identical
@@
 (teɪbəl)-\emph{mesa} (mesa) translation pair had a 17\% similarity,
-while the \emph{train} (trEIn)-\emph{tren} (tɾEn) translation pair had a
-60\% similarity. Figure 1 summarises the lexical frequency, phonological
-neighbourhood density and phonological overlap of the words included in
-the Catalan and the Spanish lists.
+while the \emph{train} (\texttt{trEIn})-\emph{tren} (tɾEn) translation
+pair had a 60\% similarity. Figure 1 summarises the lexical frequency,
+phonological neighbourhood density and phonological overlap of the words
+included in the Catalan and the Spanish lists.
 
@@
 (\citeproc{ref-marian2012clearpond}{Marian et al., 2012})\footnote{{[}Phonological
-  trnascriptions in CLEARPOND were generated from eSPEAK,
+  transcriptions in CLEARPOND were generated from eSPEAK,
   \url{http://espeak.sourceforge.net/}{]}}.
 
-\captionsetup{labelsep=none}
-
 \begin{longtable}{l|rrrrrr}
 
-\caption{\label{tbl-stimuli}}
+\caption{\label{tbl-stimuli}Stimuli properties.}
 
@@
 cat-SPA & $5.79$ ± $0.63$ & $4.48$–$7.70$ & $0.38$ ± $0.26$ & $0.00$–$1.00$ & $0.87$ ± $2.08$ & $0$–$12$ \\ 
+\midrule 
+\midrule 
+\textbf{Mean} & $5.85$ & — & $0.225$ & — & $0.606$ & — \\ 
 \bottomrule
@@
 pressed the \texttt{RETURN/ENTER} key to confirm their answer and start
-and new trial. We excluded trials where the participant's answer was not
-a valid word in the target language or no answer was given. Trials in
-which the response was mistyped by only one character were counted as
-correct (e.g., ``pengiun'' instead of ``penguin''), as long as the
-response did not correspond to a distinct word. Trials in which
-participants took longer than 10 seconds to respond were also excluded.
-Participants contributed a total of 9152 valid trials (5694 in Catalan,
-3458 in Spanish). The task took approximately 15 minutes to be
-completed.
+and new trial.
 
@@
 
-{\centering \includegraphics[width=0.8\textwidth,height=\textheight]{../img/design.png}
+{\centering \includegraphics{_assets/img/design.png}
 
@@
 
-\caption{Schematic representation of a trial in the experimental task.}
+\caption{Schematic representation of a trial in the experimental task}
 
@@
 \subsubsection{Data analysis}\label{data-analysis}
+
+\paragraph{Data processing}\label{data-processing}
+
+After data collection, participants answers were manually coded into the
+following categories: \emph{Correct}, \emph{Typo}, \emph{Wrong},
+\emph{False friend}, \emph{Other}. Responses were coded as
+\emph{Correct} if the provided string of characters was identical to the
+orthographic form of the correct translation. Trials were coded as
+\emph{Typo} if the participant provided a string of characters only one
+edit distance (addition, deletion, or substitution) apart from the
+orthographic form of the correct translation (e.g., ``pengiun'' instead
+of ``penguin''), as long as the response did not correspond to a
+distinct English word. Responses were coded as \emph{False friend} if
+the participant provided a phonologically similar incorrect translation.
+Responses were coded as \emph{Other} (see Data analysis section for more
+details). Both \emph{Correct} and \emph{Typo} responses were considered
+as correct, while \emph{Wrong} and \emph{False friend} responses were
+considered as incorrect. \emph{Other} responses were excluded from data
+analysis.
+
+Trials in which participants took longer than 10 seconds to respond were
+also excluded. Participants contributed a total of 9152 valid trials
+(5694 in Catalan, 3458 in Spanish). The task took approximately 15
+minutes to be completed.
+
+\paragraph{Modelling approach and statistical
+inference}\label{modelling-approach-and-statistical-inference}
 
@@
 (\emph{n} = 2), were diagnosed with a language (\emph{n} = 1), or did
-not contribute more than 80\% of valid trials (\emph{n} = 9). After
-applying trial-level and participant-level inclusion criteria, the
+not contribute more than 80\% of valid trials (\emph{n} = 9).
+
+After applying trial-level and participant-level inclusion criteria, the
 resulting dataset included 5,204 trials provided by 54 unique
@@
 
-\begin{longtable}{l|rrrr}
+\begin{longtable}{l|rrrrrrrr}
 
@@
 \toprule
-\multicolumn{1}{l}{} & \multicolumn{2}{c}{Correct responses} & \multicolumn{2}{c}{Incorrect responses} \\ 
-\cmidrule(lr){2-3} \cmidrule(lr){4-5}
-\multicolumn{1}{l}{} & Correct & Typo & Wrong translation & False friend \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{5}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $429$ ($16.47\%$) & $11$ ($0.42\%$) & $2,082$ ($79.95\%$) & $82$ ($3.15\%$) \\ 
-spa-ENG & $374$ ($14.37\%$) & $2$ ($0.08\%$) & $2,117$ ($81.36\%$) & $109$ ($4.19\%$) \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{5}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & $780$ ($46.93\%$) & $20$ ($1.20\%$) & $736$ ($44.28\%$) & $126$ ($7.58\%$) \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{5}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $477$ ($18.01\%$) & $7$ ($0.26\%$) & $1,986$ ($75.00\%$) & $178$ ($6.72\%$) \\ 
-spa-ENG & $590$ ($19.53\%$) & $6$ ($0.20\%$) & $2,294$ ($75.94\%$) & $131$ ($4.34\%$) \\ 
+\multicolumn{1}{l}{} & \multicolumn{4}{c}{Correct responses} & \multicolumn{4}{c}{Incorrect responses} \\ 
+\cmidrule(lr){2-5} \cmidrule(lr){6-9}
+\multicolumn{1}{l}{} & Correct & (\%) & Typo & (\%) & Wrong & (\%) & False friend & (\%) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{9}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $429$ & ($16.47$) & $11$ & ($0.42$) & $2,082$ & ($79.95$) & $82$ & ($3.15$) \\ 
+spa-ENG & $374$ & ($14.37$) & $2$ & ($0.08$) & $2,117$ & ($81.36$) & $109$ & ($4.19$) \\ 
+\midrule 
+\emph{Sum} & $803$ & — & $13$ & — & $4,199$ & — & $191$ & — \\ 
+\emph{Mean} & — & $15.42$ & — & $0.25$ & — & $80.66$ & — & $3.67$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{9}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & $780$ & ($46.93$) & $20$ & ($1.20$) & $736$ & ($44.28$) & $126$ & ($7.58$) \\ 
+\midrule 
+\emph{Sum} & $780$ & — & $20$ & — & $736$ & — & $126$ & — \\ 
+\emph{Mean} & — & $46.93$ & — & $1.20$ & — & $44.28$ & — & $7.58$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{9}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $477$ & ($18.01$) & $7$ & ($0.26$) & $1,986$ & ($75.00$) & $178$ & ($6.72$) \\ 
+spa-ENG & $590$ & ($19.53$) & $6$ & ($0.20$) & $2,294$ & ($75.94$) & $131$ & ($4.34$) \\ 
+\midrule 
+\emph{Sum} & $1,067$ & — & $13$ & — & $4,280$ & — & $309$ & — \\ 
+\emph{Mean} & — & $18.77$ & — & $0.23$ & — & $75.47$ & — & $5.53$ \\ 
 \bottomrule
@@
 
-\begin{longtable}{l|rrrrrrr}
+\begin{longtable}{l|rrrrrrrrr}
 
@@
 \toprule
-\multicolumn{1}{l}{} &  & \multicolumn{3}{c}{Accuracy (\%)} & \multicolumn{3}{c}{Valid trials} \\ 
-\cmidrule(lr){3-5} \cmidrule(lr){6-8}
-\multicolumn{1}{l}{} & N & Mean ± SD & SE & Range & Mean ± SD & N trials & Range \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & $27$ & $15.86$ ± $5.20$ & $3.05$ & $8.82$–$28.71$ & $96.37$ ± $2.88$ ± $2.88$ & $2,602$ & $87$–$98$ \\ 
-cat-ENG & $32$ & $18.48$ ± $4.89$ & $3.27$ & $10.84$–$32.56$ & $81.38$ ± $3.17$ ± $3.17$ & $2,604$ & $71$–$83$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & $21$ & $48.30$ ± $5.29$ & $10.54$ & $38.27$–$58.97$ & $79.14$ ± $3.14$ ± $3.14$ & $1,662$ & $72$–$82$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{8}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-spa-ENG & $31$ & $20.92$ ± $8.29$ & $3.76$ & $5.88$–$44.12$ & $97.45$ ± $1.80$ ± $1.80$ & $3,021$ & $88$–$98$ \\ 
-cat-ENG & $32$ & $19.74$ ± $4.94$ & $3.49$ & $10.34$–$27.91$ & $82.75$ ± $0.44$ ± $0.44$ & $2,648$ & $82$–$83$ \\ 
+\multicolumn{1}{l}{} &  & \multicolumn{4}{c}{Accuracy (\%)} & \multicolumn{4}{c}{Valid trials} \\ 
+\cmidrule(lr){3-6} \cmidrule(lr){7-10}
+\multicolumn{1}{l}{} & N & Mean & SD & SE & Range & Mean & N trials & SD & Range \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{10}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & $27$ & $15.86$ & $5.20$ & $3.05$ & $8.82$–$28.71$ & $96.37$ & $2,602$ & $2.88$ & $87$–$98$ \\ 
+cat-ENG & $32$ & $18.48$ & $4.89$ & $3.27$ & $10.84$–$32.56$ & $81.38$ & $2,604$ & $3.17$ & $71$–$83$ \\ 
+\midrule 
+\emph{Sum} & $59$ & — & — & — & — & — & $5,206$ & — & $158$ \\ 
+\emph{Mean} & — & $17.17$ & $5.05$ & $3.16$ & — & $8,887.27$ & — & $302.71$ & — \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{10}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & $21$ & $48.30$ & $5.29$ & $10.54$ & $38.27$–$58.97$ & $79.14$ & $1,662$ & $3.14$ & $72$–$82$ \\ 
+\midrule 
+\emph{Sum} & $21$ & — & — & — & — & — & $1,662$ & — & $72$ \\ 
+\emph{Mean} & — & $48.30$ & $5.29$ & $10.54$ & — & $7,914.29$ & — & $313.51$ & — \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{10}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+spa-ENG & $31$ & $20.92$ & $8.29$ & $3.76$ & $5.88$–$44.12$ & $97.45$ & $3,021$ & $1.80$ & $88$–$98$ \\ 
+cat-ENG & $32$ & $19.74$ & $4.94$ & $3.49$ & $10.34$–$27.91$ & $82.75$ & $2,648$ & $0.44$ & $82$–$83$ \\ 
+\midrule 
+\emph{Sum} & $63$ & — & — & — & — & — & $5,669$ & — & $170$ \\ 
+\emph{Mean} & — & $20.33$ & $6.62$ & $3.62$ & — & $9,010.08$ & — & $112.22$ & — \\ 
 \bottomrule
@@
 
-\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}
+\includegraphics{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}
 
@@
 
-\caption{\label{fig-epreds-1}}
+\caption{\label{fig-epreds-1}Posterior model-predicted mean accuracy in
+Experiment 1. Predictions were generated from 4,000 posterior samples,
+extracted for different values of \emph{CLPN} (0, 2, 4, 8, 12) and
+\emph{Cognateness} (1-100). Predictions are plotted separately for
+English participants translating Catalan words, and for English
+participants translating Spanish words. Lines indicate mean predictions,
+and intervals indicate 95\%, 89\%, 78\%, 67\%, and 50\% credible
+intervals (CrI).}
 
@@
 in a translation task that involved words in Catalan or Spanish, two
-languages participants reported having no familiarity with. Participants
-benefited strongly from cognateness, when the correct translation of the
-presented words in Catalan or Spanish had few English phonological
-neighbours with higher lexical frequency. This suggests that, in the
-absence of distractors, even naïve participants can efficiently use
-phonological similarity to succeed in a translation task. Our results
-suggest that word-to-word connections at the phonological level might
-play a strong role during L2 speech comprehension, specially in
-low-proficiency listeners.
+languages participants reported having no prior familiarity with.
+Participants benefited strongly from cognateness, when the correct
+translation of the presented words in Catalan or Spanish had few English
+phonological neighbours with higher lexical frequency. This suggests
+that, in the absence of distractors, even naïve participants can
+efficiently use phonological similarity to succeed in a translation
+task. Our results suggest that word-to-word connections at the
+phonological level might play a strong role during L2 speech
+comprehension, specially in low-proficiency listeners.
 
@@
 
+Results in Experiment 1 suggest that English natives were able to
+exploit the phonological similarity between unfamiliar words in Catalan
+and Spanish to provide accurate translations to English. English, a
+Germanic language, is relatively distant from Catalan and Spanish, two
+Romance languages. English shares fewer cognates with Catalan and
+Spanish than it does with typologically closer languages, like Dutch and
+German. In Experiment 2, we investigated whether listeners of an
+unfamiliar but typologically closer language benefit more strongly from
+phonological similarity when performing the same task as in Experiment
+1. To this aim, we presented Spanish participants, who reported
+little-to-no prior familiarity with Catalan, with Catalan words.
+
 \subsection{Methods}\label{methods-1}
-
-\subsubsection{Participants}\label{participants-1}
 
@@
 in Spain (\emph{Mean} = 21.85 years, \emph{SD} = 3, Range = 18-33, 28, 5
-female).
-
-\subsubsection{Stimuli}\label{stimuli-1}
-
-Same list of Catalan stimuli as in Experiment 1.
-
-\subsubsection{Procedure}\label{procedure-1}
-
-Same as in Experiment 1.
-
-\subsubsection{Data analysis}\label{data-analysis-1}
-
-Same approach as in Experiment 1.
+female). Stimuli were the same list of Catalan stimuli as in Experiment
+1. Procedure and data analysis were identical as in Experiment 1.
 
@@
 
-\textbf{?@tbl-results-1} shows a summary of participants' accuracy
-across Experiments 1, 2, and 3. MCMC chains in the model showed strong
-evidence of convergence (\(\hat{R}<1.01\)) (see Appendix 2 for more
-detailed model diagnostics). Overall, participants responded less
-accurately to words with more CLPNs than to words with fewer CLPNs,
-regardless of the amount of phonological similarity between the
-presented word and its translation. This is indicated by the size of the
-regression coefficient of the two-way interaction between
-\emph{Cognateness} and \emph{CLPN} (\(\beta\) = -0.409, 95\% CrI =
-{[}-0.895, 0.111{]}, \emph{p}(ROPE) = 0.081). As anticipated,
-participants' performance benefited from an increase in
+MCMC chains in the model showed strong evidence of convergence
+(\(\hat{R}<1.01\)) (see Appendix 2 for more detailed model diagnostics).
+Overall, participants responded less accurately to words with more CLPNs
+than to words with fewer CLPNs, regardless of the amount of phonological
+similarity between the presented word and its translation. This is
+indicated by the size of the regression coefficient of the two-way
+interaction between \emph{Cognateness} and \emph{CLPN} (\(\beta\) =
+-0.409, 95\% CrI = {[}-0.895, 0.111{]}, \emph{p}(ROPE) = 0.081). As
+anticipated, participants' performance benefited from an increase in
 \emph{Cognateness} (\(\beta\) = 9.274, 95\% CrI = {[}8.473, 10.269{]},
@@
 
-\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}
+\includegraphics{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}
 
@@
 
-\caption{\label{fig-epreds-2}}
+\caption{\label{fig-epreds-2}Posterior model-predicted mean accuracy in
+Experiment 1. Predictions were generated from 4,000 posterior samples,
+extracted for different values of \emph{CLPN} (0, 2, 4, 8, 12) and
+\emph{Cognateness} (1-100). Predictions are plotted separately for
+English participants translating Catalan words, and for English
+participants translating Spanish words. Lines indicate mean predictions,
+and intervals indicate 95\%, 89\%, 78\%, 67\%, and 50\% credible
+intervals (CrI).}
 
@@
 
-\subsubsection{Participants}\label{participants-2}
-
 Data collection took place from October 22th, 2022 to October 23th,
@@
 complete the experiment using a laptop in a quiet place with good
-internet connection. We excluded data from participants that (a)
-self-rated their oral and/or written skills in Catalan and Spanish, or
-any other second language as four or higher in a five-point scale
-(\emph{n} = 2), (b) were diagnosed with a language (\emph{n} = 0), or
-(c) did not contribute more than 80\% of valid trials (\emph{n} = 1).
-
-\subsubsection{Stimuli}\label{stimuli-2}
-
-Same as in Experiment 1.
-
-\subsubsection{Procedure}\label{procedure-2}
-
-{[}TO BE COMPLETED{]}
-
-\subsubsection{Data analysis}\label{data-analysis-2}
-
-Same approach as in Experiment 1.
+internet connection. Stimuli were the same list of Catalan stimuli as in
+Experiment 1.
+
+The experiment was implemented online using Qualtrics (Qualtrics, Provo,
+UT). This platform was chosen to allow easier presentation of survey
+questions aimed to probe prior understanding of the presented words and
+participants' confidence ratings of their answers. With the exception of
+these additional questions, we attempted to replicate the procedure of
+Experiment 1 as closely as possible. The Spanish and Catalan audio
+stimuli used were identical the materials in Experiment 1. Participants
+were randomly assigned to the Catalan or Spanish lists. The Catalan list
+had 83 trials and the Spanish list had 99 trials. Participants first
+completed the consent form followed by the questionnaire about
+demographic status, language background and set up. They then proceeded
+to the experimental task.
+
+On each trial, participants listened to the audio stimulus by clicking
+on the play button. For comparability to the PsychoPy version,
+participants were only allowed to play the audio one time. Participants
+were explicitly told that they would be only allowed to listen once. The
+play button vanished after one playthrough. Participants then had to
+answer three questions based on the audio they had heard on that trial.
+These questions were presented on the same page, directly below the
+audio player. They were first asked whether or not they knew the
+presented word (multiple choice---yes/no). Regardless of their answer on
+the first question, participants were asked what they thought the
+translation of the word was in English (or their best guess), and
+instructed to type their answer in the provided text box. Finally, they
+were asked to rate how confident they were in their answer on a scale of
+0 to 7, where 7 was ``very confident'' and 0 was ``not confident''.
+There was no time limit on the response phase. All questions had to be
+answered to proceed to the next trial.
+
+Participants first completed 5 practice trials with English words as the
+audio stimulus (ambulance, cucumber, elephant, pear, turtle). The words
+were recorded by a female native speaker of English. These trials acted
+as attention checks, as participants should always answer ``yes'' to the
+first question on prior word knowledge and be able to accurately
+transcribe the word they heard. Following the practice phase,
+participants completed the test phase where they heard either Spanish
+words or Catalan words.
 
@@
 (\emph{n} = 2), were diagnosed with a language (\emph{n} = 0), or did
-not contribute more than 80\% of valid trials (\emph{n} = 1). After
-applying trial-level and participant-level inclusion criteria, the
+not contribute more than 80\% of valid trials (\emph{n} = 1).
+
+After applying trial-level and participant-level inclusion criteria, the
 resulting dataset included 6,290 trials provided by 62 unique
@@
 provided by 32 unique participants who listened to Spanish words.
+
+INSERT PRIOR KNOWLEDGE FILTER
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}
+\includegraphics{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}
 
@@
 
-\caption{\label{fig-epreds-3}}
+\caption{\label{fig-epreds-3}Posterior model-predicted mean accuracy in
+Experiment 2. Predictions were generated from 4,000 posterior samples,
+extracted for different values of \emph{CLPN} (0, 2, 4, 8, 12) and
+\emph{Cognateness} (1-100). Predictions are plotted separately for
+English participants translating Catalan words, and for English
+participants translating Spanish words. Lines indicate mean predictions,
+and intervals indicate 95\%, 89\%, 78\%, 67\%, and 50\% credible
+intervals (CrI).}
 

### 2024-05-10 manuscript/manuscript.qmd
Final touches to first version

@@
 One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, imagine a Spanish native, with no previous familiarity with any other language, who listens for the first time to French may encounter the word /`r knitr::asis_output("\U0070\U0254\U0281\U0074")`/ (*porte*), translation of door in French.
-The voiced uvular fricative consonant /`r knitr::asis_output("\U0281")`/ and the open-mid back rounded vowel /`r knitr::asis_output("\U0254")`/ do not exist in Spanish as phonemic categories.
-These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /r/ and /o/ [@best2001discrimination], and the non-native speech signal may engage lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ (*puerta*).
+For example, imagine a Spanish native, with no previous familiarity with any other language, who listens for the first time to French may encounter the word /pɔʁt/ (*porte*), translation of door in French.
+The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish as phonemic categories.
+These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/ [@best2001discrimination], and the non-native speech signal may engage lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /pweɾta/ (*puerta*).
 The recognition of this French, unfamiliar word relies almost entirely on its similarity with its Spanish translation.
-Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /`r knitr::asis_output("\U0064\U0254\U02D0")`/ (*door*).
-This time, phonological similarity is of little help: /`r knitr::asis_output("\U02C8\U0070\U0077\U0065\U0072\U002E\U0074\U0061")`/ and /`r knitr::asis_output("\U0064\U0254\U02D0")`/ share very few phonemes.
-It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations (which they lack, as they are unfamiliar to English).
+Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
+This time, phonological similarity is of little help: /pweɾta/ and /dɔː/ share very few phonemes.
+It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations---which they lack, as they are unfamiliar to English.
 
@@
 Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners–even low-proficiency ones–are rarely completely naïve to the language they are listening to.
-In the present study, we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates.*
+In the present study, we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
 

### 2024-05-10 manuscript/manuscript.tex
Final touches to first version

@@
 \fi
-\usepackage{lmodern}
+\usepackage[]{Times}
 \ifPDFTeX\else  
@@
 \usepackage{booktabs}
+\DeclareTextFontCommand\textphonetic{\ipafont}
 \usepackage{arxiv}
@@
 All materials, data, and code used in this study are hosted in the Open
-Science Framework \url{https://osf.io/9fjxm/} and a GitHub repository
+Science Framework
+\href{https://osf.io/9fjxm/?view_only=aab7636ce1af48cf832596a7ea9101c5/}{https://osf.io/9fjxm/}
+and a GitHub repository
 \url{https://github.com/gongcastro/translation-elicitation.git}, along

### 2024-07-01 manuscript/manuscript.qmd
New version of the introdution

@@
 ---
-title: "The role of cognateness in non-native spoken word recognition"
+title: "The role of cognateness in native spoken word recognition"
 ---
@@
 Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
-However, there are some sources of information which may make non-native speech comprehension easier.
-One of the sources of this increased difficulty stems from the possible mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
-For example, imagine a Spanish native, with no previous familiarity with any other language, who listens for the first time to French may encounter the word /pɔʁt/ (*porte*), translation of door in French.
+One of the sources of this increased difficulty is the partial mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
+For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
+This listener may encounter the word /pɔʁt/ (*porte*), translation of door in French.
 The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish as phonemic categories.
-These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/ [@best2001discrimination], and the non-native speech signal may engage lexical processing mechanisms [@weber2004lexical], leading to the activation and selection, and therefore the recognition of its translation to Spanish /pweɾta/ (*puerta*).
-The recognition of this French, unfamiliar word relies almost entirely on its similarity with its Spanish translation.
-Conversely, imagine now that the same Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
-This time, phonological similarity is of little help: /pweɾta/ and /dɔː/ share very few phonemes.
-It would come as a surprise if the listener was able to translate the English word successfully, if not by relying on previously learnt word-concept associations---which they lack, as they are unfamiliar to English.
-
-Cross-language phonological similarity seems, at face value, an important source of information for unfamiliar listeners during auditory word recognition, even if translations are not phonologically identical.
-Although this mismatch has a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners–even low-proficiency ones–are rarely completely naïve to the language they are listening to.
-In the present study, we investigate the extent to which adult listeners rely on the phonological similarity between their native and non-native languages during word recognition, capitalising on the role of *cognates*.
-
-Many languages share, to some extent, similarities at the lexical level.
+These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/ [@best2001discrimination], leading to the non-native speech signal engaging lexical processing mechanisms [@weber2004lexical].
+Even if unfamiliar with the word-form /pɔʁt/, the Spanish monolingual is likely to activate the phonological representation of its Spanish translation /pweɾta/ (*puerta*), ultimately accessing the right semantic representation.
+Although the phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners---even low-proficiency ones---are rarely completely naïve to the language they are listening to.
+
+Imagine now that the same Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
+This time, phonological similarity is of little help: /pweɾta/ and /dɔː/ do not share any phonemes.
+It would come as a surprise if the listener was able to translate the English word successfully.
+Non-native listeners rely on the phonological similarity between the word-forms presented in the input and their translations.
+Luckily for them, most languages share such kind of similarity.
 This is frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
-Cognates embody a particular case of such commonalities.
-Cognates are defined as cross-language synonyms which have similar forms whose form is similar (at the phonological level, orthographic level, signed level), commonly due to their shared etymological origin^[Some form-similar cross-language synonyms are technically not cognates.
-For example, *sun* and *sol* (in Spanish) share phonological onset but their etymology points to different origins.
-The complimentary case is equally problematic when defining cognateness: cross-language synonyms that share etymological origin may not be necessarily be recognised as cognates by the average listener.
-Such is the case of *cheese* and *queso* which share Latin root but are phonologically dissimilar.
-For simplicity, we will use the term *cognateness* in this paper to include all cross-language synonyms that fulfill a specified threshold of form similarity, regardless of etymology.
-We do not expect that etymology plays a direct role on language perception if it is not via form-similarity, since it is not necessary for participants in psycholinguistic experiments to be aware of the etymology of the words they encounter in the tasks to be subject to the effect of form-similarity.].
-For example, Romance languages such as Spanish and Catalan share many cognates [@schepens2012distributions], as in the case of *puerta* and *porta* (door in Spanish and Catalan, respectively).
-Cognates play a pivotal role in many models of bilingual lexical processing because they provide evidence that bilinguals access the lexicon in a language non-selective way.
-For instance, pictures are named faster by Spanish-Catalan bilinguals when their associated labels in both languages are cognates (e.g., *puerta*-*porta*), compared to when their labels are non-cognates (*mesa*-*taula*, [table]), which suggests that word representations of both languages are activated during word retrieval [@costa2000cognate].
-This cognate advantage is not restricted to production, but also extends to word recognition [@thierry2007brain; @midgley2011effects], word learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and word translation [@christoffels2006memory].
-
-## Lexical and conceptual routes for translation
-
-Early theoretical accounts of this 'Cognate effect' proposed that the impact of cognateness on lexical processing depends on bilinguals' proficiency in their second language (L2) [@potter1984lexical; see @chen1989patterns for similar results].
-The reasoning behind this claim was the assumption that second-language learners start acquiring words in their L2 by associating them to their translation in their L1 (lexical route), instead of directly to their shared concept (conceptual route).
-This word-to-word connectivity should be sensitive to form-similarity (i.e., cognateness): the more similar the two word- forms are, the stronger the connection.
-As learners become more proficient, the connection between L2 representations and their meaning grows stronger, and L2 word processing becomes less reliant on word-word connections between L1 and L2 representations.
-Cognates subsequently exert less impact on lexical processing in proficient speakers [see @andras2022cognate for recent experimental evidence for this claim].
-The Revised Hierarchical Model [RHM, @kroll1994category] captured this hypothesis and predicted that translating words from L1 to L2 should take longer than translating from L2 to L1.
-The rationale behind this prediction was that L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations, where L2-to-L1 translation relies more strongly on the mediation between the concept and the two word- forms.
-L1-to-L2 translation should therefore be more sensitive to the form similarity between the L1 and the L2 representations: cognate words should be retrieved faster than non-cognate words during L1-to-L2 translation.
-
-To test these predictions, @degroot1994forward and @de1992determinants asked Dutch natives with high (but non-native) English proficiency to translate Dutch words to English (L1-to-L2) or English words to Dutch (L2-to-L1).
-Although participants' performance in both conditions was roughly equivalent for cognates, non-cognates were translated more slowly from Dutch to English than from English to Dutch, suggesting that L1-to-L2 translation is more sensitive to form-similarity than L2-to-L1 translation.
-Conversely, participants' performance was more sensitive to semantic variables (e.g., concreteness) when translating English words to Dutch, compared to when translating Dutch words to English, suggesting that L2-to-L1 translation was more sensitive to the shared conceptual properties of the translation pair.
-
-These findings support a soft version of the RHM model, in which both the conceptual and the lexical translation routes are active during translation, but where L1-to-L2 translation relies more strongly on word-to-word links between L1 and L2 representations than L2-to-L1 translation does, and therefore is more sensitive to cognateness.
-Further evidence for the presence of this difference between L1-to-L2 and L2-to-L1 translations on the degree of reliance on the lexical and or the conceptual routes between L1-to-L2 and L2-to-L1 translation was provided by @phillips2006erp, who used eletrophysiological recordings of bilinguals translating words from L1 to L2 or *vice versa* after semantic or phonological priming.
-The authors reported stronger N400 responses (associated with semantic priming) in L1-to-L2 translation than in L2-to-L12 translation, and stronger phonological mismatch negativity responses (associated with phonological priming) during L2-to-L1 translation, compared to L1-to-L2 translation.
-
-## Competition from lexical neighbours
-
-Since the RHM model was proposed, later studies on monolingual populations have focused how on the role of the network of connections that words establish networks of connections with each other at the form-level (phonological or orthographic) or at the conceptual level.
-For instance, the Neighbourhood Activation Model [NAM, @luce1998recognizing] highlighted how lexical selection is affected by the number of phonological neighbours that the selected word has.
-Words surrounded by a larger number of phonological neighbours (words that only differ in one phoneme from the target word) are responded to more slowly and less accurately than words from sparser phonological neighbourhoods [@goldinger1989priming; @luce1990similarity], especially when these neighbours' lexical frequency is higher [@luce1998recognizing].
-
-The existence of phonological links between words is supported with experimental evidence from the visual world paradigm and phonological priming.
-@meyer2007activation showed that target pictures were named faster when presented together with a phonologically-related distractor, suggesting that the activation of shared phonological information allowed the target to be activated and produced faster.
-there are also findings suggesting that facilitatory or inhibitory effects on target recognition depend on the degree of similarity the competitor shares with the target [@hamburger1996phonological; @dufour2003lexical].
-When there is only mismatch of one phoneme between prime and target, inhibitory effects are exerted on target recognition [@dufour2003lexical].
-This inhibition was proposed to be due to strong lexical activation of the prime which results in lexical interference, as opposed to simple phonological activation from less similar primes that triggers facilitation.
-
-These findings of phonological links between words in the monolingual lexicon pose an important question in the field of bilingualism research: do word representations in one language form part of the phonological or orthographic neighbours of the other language?
-@weber2004lexical observed that bilinguals displayed lexical competition effects from words in their native language when performing a word-referent matching task in their L2.
-This highlights that lexical competition for bilinguals during language comprehension may not only come from the tested language but also the native language, despite the native language not being actively required for the task.
-Other studies have similarly found interference when spoken target words in one language had phonological overlap with a word in the participant's other language, even when the other language was not required by the task [@spivey1999cross; @desroches2022dynamics].
-Such patterns have even been found with young children less than 4 years old [@von2012language].
-
-The Bilingual Interactive Activation (BIA/BIA+) [@van1998orthographic; @dijkstra2002architecture] addressed this issue and suggested that lexical representations from both languages establish both excitatory and inhibitory connections with each other, resulting in an integrated lexicon for both languages.
-This is, in line with connectionist approaches to lexical processing [@mcclelland1981interactive].
-Evidence supporting such an integrated lexicon is provided by the fact that bilinguals' performance in word recognition tasks is sensitive to the orthographic neighbourhood density of the presented word's translation.
-When bilingual participants are presented with words in one language in a lexical decision task, translating them to the other language took longer when they were part of large neighbourhoods [@van1998orthographic].
-This suggests that presented words activated orthographic neighbourhoods in the other language, which competed for selection with the target word during recognition.
-
-## The present study
-
-In the present study, we tested listeners' reliance on phonological similarity when translating words from an unfamiliar language to their native language, in order to investigate the plausibility of word-to-word connections during L1-to-L2 translation.
-We tested English native speakers and Spanish native speakers in a translation elicitation task.
-This paradigm has been used by bilingualism researchers to identify easily-recognisable cognates for experimental studies.
-Using this paradigm, we aim to delve deeper into the factors affecting foreign word recognition by monolingual participants with no prior exposure to the language.
-
-Participants were presented with spoken words in an unfamiliar (non-native language) and were asked to translate them to their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
-The motivation behind testing participants in an unfamiliar language was that these participants can be considered a particular (extreme) case of unbalanced bilingualism, in which they do not know any words in L2, and therefore word-to-concept connections are, in principle, absent.
-As a consequence, these participants should only be able use word-to-word connections to succeed in the translation of words in an unfamiliar language.
-We predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
-
-Following @luce1998recognizing's NAM model, we further hypothesised that if participants are able to correctly translate words from an unfamiliar language based on their phonological similarity to the target word, similar sounding but incorrect translations (e.g., false friends) should be activated too.
-If this is the case, non-native words that have more phonological neighbours should be less likely to be correctly translated, especially if such neighbours have higher lexical frequency and are phonologically closer to the presented word than the target translation.
+For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents (also known as *cognates*) [@schepens2012distributions], as in the case of *puerta* and *porta* (door in Spanish and Catalan, respectively).
+This type of cross-linguistic similarity has been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
+This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
+This facilitation effect has been reported for comprehension [@thierry2007brain; @midgley2011effects; @dijkstra2010cross], production [@costa2000cognate], learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and translation [@christoffels2006memory].
+The role that cross-linguistic phonological similarity might play in monolingual lexical access during non-native speech comprehension is unclear.
+How much phonological similarity is enough for a monolingual to understand a word from a non-native language?
+How does phonological similarity interact with other determinants of word recognition?
+
+In the present study, we investigate the extent to which monolinguals can exploit phonological similarity when listening to words in an unfamiliar languages, in order to achieve comprehension, while taking into account other lexically-relevant variables interact with the lexical processing mechanisms involved.
+We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as *presented words*, and the correct translation for the presented words as *target translations*.
+We explored listeners' reliance on phonological similarity to succeed in the task by manipulating the amount of phonological similarity between the presented words and their target translations.
+Since participants were unfamiliar with the presented language, they could only be able to use word-to-word connections to succeed in the translation of words in an unfamiliar language.
+We therefore predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
+
+The extent to which participants can rely on phonological similarity is limited by other factors.
+In particular, the presence of phonological neighbours (word-forms than differ in one phoneme) may interfere with the activation and selection of the appropriate lexical representation in the native language.
+Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@luce1998recognizing; @goldinger1989priming; @luce1990similarity; @hamburger1996phonological; @dufour2003lexical].
+This is specially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
+The inhibitory effect of phonological neighbourhoods has also been reported across languages.
+For instance, @weber2004lexical observed that bilinguals displayed lexical competition effects from words in their native language when performing a word-referent matching task in their L2.
+
+In the case of the present task, the potential effect of cross-language phonological neighbourhood density is critical.
+The presence of phonologically similar non-translation equivalents (e.g., false friends) may act as distractor during lexical access, obstructing the selection of the appropriate target translation in the listener's lexicon.
+Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with words on the target language other than its translation, specially if any of them is more lexically frequent than the target translation.
+We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
 For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as bottom instead of as its correct translation *button* (because the lexical frequency of *bottom* is higher than that of *button* and also phonological closer).
-This combined constraint led us to calculate a specific form of phonological neighbours where a neighbour is counted only if it is higher frequency and has stronger phonological similarity than the target translation as judged by Levenshtein distance.
+To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme appart from the presented word.
 If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
 
-In Experiment 1, we collected data from two groups of British English natives.
+In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
 One group was presented with Catalan words, the other with Spanish words.
-We examined the extent to which participants were able to use the phonological similarity between the presented (Catalan or Spanish) word and its translation (i.e., cognateness) to provide accurate responses.
+We examined the extent to which participants were able to use the phonological similarity between the presented word (in Catalan or Spanish) and its target translation to provide accurate responses.
 In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
@@
 For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
-For example, the *table* (teɪbəl)-*mesa* (mesa) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+For example, the *table* (/teɪbəl/)-*mesa* (/mesa/) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
 Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.

### 2024-07-11 manuscript/manuscript.qmd
Fix PDF output

@@
 ---
-title: "The role of cognateness in native spoken word recognition"
+title: "The role of Similarity in native spoken word recognition"
 ---
@@
 This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
-This facilitation effect has been reported for comprehension [@thierry2007brain; @midgley2011effects; @dijkstra2010cross], production [@costa2000cognate], learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022cross], and translation [@christoffels2006memory].
+This facilitation effect has been reported for comprehension [@thierry2007brain; @midgley2011effects; @dijkstra2010cross], production [@costa2000cognate], learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022crosslanguage], and translation [@christoffels2006memory].
 The role that cross-linguistic phonological similarity might play in monolingual lexical access during non-native speech comprehension is unclear.
@@
 
-For each translation pair, we defined three predictors of interest: the lexical frequency of the correct translation (*Frequency*), the phonological similarity between the presented (Catalan or Spanish) word and their correct English translation (*Cognateness*), and the presented word's number of cross-language phonological neighbours (*CLPN*).
+For each translation pair, we defined three predictors of interest: the lexical frequency of the correct translation (*Frequency*), the phonological similarity between the presented (Catalan or Spanish) word and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*).
 
@@
 
-We calculated *Cognatenesss*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
+We calculated *Similaritys*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
 The Levenshtein distance computes the edit distance between two character strings---in this case, phoneme transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
@@
     tab_spanner("Frequency", matches("freq")) |> 
-    tab_spanner("Cognateness", matches("lv")) |> 
+    tab_spanner("Similarity", matches("lv")) |> 
     tab_spanner("CLPN", matches("neigh")) |> 
@@
 We modelled the probability of participants guessing the correct translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We included as fixed effects the intercept, the main effects of *Frequency*, *Cognateness*, and *CLPN*, and the two-way interaction between *Cognateness* and *CLPN*.
+We included as fixed effects the intercept, the main effects of *Frequency*, *Similarity*, and *CLPN*, and the two-way interaction between *Similarity* and *CLPN*.
 We also included participant-level random intercepts and slopes for the the main effects and the interaction.
@@
 
+::: {.content-visible when-format="docx"}
 ![Eq. 1](_assets/img/model.PNG)
+:::
+
+::: {.content-visible when-format="pdf"}
+$$
+\begin{aligned}
+&\textbf{Likelihood}  \\
+y_{i} \sim & \text{Bernoulli}(p_{i}) \\ \\
+&\textbf{Parameters}  \\
+\text{Logit}(p_{i}) = &  \beta_{0[p,w]} + \beta_{1[p]} \text{Frequency}_{i} + \beta_{2[p]} \text{PTHN}_i + \beta_{3[p]} \text{Similarity}_i + \beta_{4[p]} (\text{PTHN}_i \times \text{Similarity}_i) \\
+\beta_{0-6[p,w]} \sim & \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \text{ and  word } w \text{ in 1, ..., } W \\
+\beta_{1-6[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\ \\
+&\textbf{Prior}  \\
+\mu_{\beta_{p,w}}  \sim &  \mathcal{N}(0, 0.1) \\
+\sigma_{\beta_{p}},  \sigma_{\beta_{w}} \sim & \text{HalfCauchy}(0, 0.1) \\
+\rho_{p}, \rho_{w} \sim & \text{LKJCorr}(8) \\
+\end{aligned}
+$$
+:::
 
@@
 
-All analyses were performed in R environment [@rcore2019r].
-We used the tidyverse family of R packages [@wickham2019tidyverse] to process data and to generate figures.
+All analyses were performed in R environment [@rcoreteam2013language].
+We used the tidyverse family of R packages [@wickham2019welcome] to process data and to generate figures.
 We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
@@
 
-@tbl-results shows a summary of participants' accuracy across Experiments 1, 2, and 3.
+@tbl-dataset shows a summary of participants' accuracy across Experiments 1, 2, and 3.
 MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
@@
 Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
-This is indicated by the size of the regression coefficient of the two-way interaction between *Cognateness* and *CLPN* ($\beta$ = `r round(c1$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_lv$.lower, 3)`, `r round(c1$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_lv$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Cognateness* ($\beta$ = `r round(c1$b_lv$.median, 3)`, 95% CrI = [`r round(c1$b_lv$.lower, 3)`, `r round(c1$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h$.lower, 3)`, `r round(c1$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h$.rope, 3)`).
-@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Cognateness* and *CLPN*.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c1$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_lv$.lower, 3)`, `r round(c1$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_lv$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c1$b_lv$.median, 3)`, 95% CrI = [`r round(c1$b_lv$.lower, 3)`, `r round(c1$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h$.lower, 3)`, `r round(c1$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h$.rope, 3)`).
+@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
 @fig-coefs shows a graphic summary of the posterior distribution of the regression coefficients of interest.
@@
 #| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
-Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Cognateness* (1-100).
+Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Similarity* (1-100).
 Predictions are plotted separately for English participants translating Catalan words, and for English participants translating Spanish words.
@@
                     .width = c(0.95, 0.89, 0.78, 0.67, 0.50)) +
-    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+    labs(x = "Similarity\n(Levenshtein distance with correct translation)",
          y = "p(Correct)",
@@
 
-In the present experiment, we investigated the extent to which cognateness (phonological similarity between translation equivalents) is sufficient for successful word translation, in the absence of semantic knowledge.
+In the present experiment, we investigated the extent to which the phonological similarity between translation equivalents is sufficient for successful word translation, in the absence of semantic knowledge.
 We tested two groups of monolingual British English natives in a translation task that involved words in Catalan or Spanish, two languages participants reported having no prior familiarity with.
-Participants benefited strongly from cognateness, when the correct translation of the presented words in Catalan or Spanish had few English phonological neighbours with higher lexical frequency.
+Participants benefited strongly from phonological similarity when the correct translation of the presented words in Catalan or Spanish had few English phonological neighbours with higher lexical frequency.
 This suggests that, in the absence of distractors, even naïve participants can efficiently use phonological similarity to succeed in a translation task.
@@
 Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
-This is indicated by the size of the regression coefficient of the two-way interaction between *Cognateness* and *CLPN* ($\beta$ = `r round(c2$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_lv$.lower, 3)`, `r round(c2$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_lv$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Cognateness* ($\beta$ = `r round(c2$b_lv$.median, 3)`, 95% CrI = [`r round(c2$b_lv$.lower, 3)`, `r round(c2$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c2$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h$.lower, 3)`, `r round(c2$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h$.rope, 3)`).
-@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Cognateness* and *CLPN*.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c2$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_lv$.lower, 3)`, `r round(c2$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_lv$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c2$b_lv$.median, 3)`, 95% CrI = [`r round(c2$b_lv$.lower, 3)`, `r round(c2$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c2$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h$.lower, 3)`, `r round(c2$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h$.rope, 3)`).
+@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
 
@@
                     .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
-    labs(x = "Cognateness\n(Levenshtein distance with correct translation)",
+    labs(x = "Similarity\n(Levenshtein distance with correct translation)",
          y = "p(Correct)",
@@
 Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
-This is indicated by the size of the regression coefficient of the two-way interaction between *Cognateness* and *CLPN* ($\beta$ = `r round(c3$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h_lv$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h_lv$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Cognateness* ($\beta$ = `r round(c3$b_lv$.median, 3)`, 95% CrI = [`r round(c3$b_lv$.lower, 3)`, `r round(c3$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c3$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h$.lower, 3)`, `r round(c3$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h$.rope, 3)`).
-@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Cognateness* and *CLPN*.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c3$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h_lv$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h_lv$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c3$b_lv$.median, 3)`, 95% CrI = [`r round(c3$b_lv$.lower, 3)`, `r round(c3$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c3$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h$.lower, 3)`, `r round(c3$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h$.rope, 3)`).
+@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
 
@@
 #| fig-cap: "Posterior model-predicted mean accuracy in Experiment 2.
-Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Cognateness* (1-100).
+Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Similarity* (1-100).
 Predictions are plotted separately for English participants translating Catalan words, and for English participants translating Spanish words.
@@
                 "b_freq_zipf_2" = "Frequency",
-                "b_lv" = "Cognateness",
+                "b_lv" = "Similarity",
                 "b_neigh_n_h" = "CLPN",
-                "b_neigh_n_h:lv" = "Cognateness × CLPN",
+                "b_neigh_n_h:lv" = "Similarity × CLPN",
                 "b_group1" = "cat-ENG vs. spa-ENG")

### 2024-07-11 manuscript/manuscript.tex
Fix PDF output

@@
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
@@
 \hypersetup{
-  pdftitle={The role of cognateness in non-native spoken word recognition},
+  pdftitle={The role of cognateness in native spoken word recognition},
   pdfauthor={Gonzalo Garcia-Castro; Serene Siow; Kim Plunkett; Nuria Sebastian-Galles},
@@
 \newcommand{\runninghead}{A Preprint }
-\renewcommand{\runninghead}{Cognateness and non-native word
-recognition }
-\title{The role of cognateness in non-native spoken word recognition}
+\renewcommand{\runninghead}{Cognateness and native word recognition }
+\title{The role of cognateness in native spoken word recognition}
 \author{\textbf{Gonzalo
@@
 \begin{abstract}
-Understanding spoken words in a non-native language is easier when the
-words are phonologically similar to their translations into the
-listener's native language are phonologically similar. Phonological
-similarity plays a more important role in the language performance of
-low-proficiency bilinguals, compared to that of high-proficiency
-bilinguals., and This suggests that low-proficiency bilinguals may
-points to the latter as relying more strongly on cross-language
-word-to-word links (lexical route), while high-proficiency bilinguals
-rely more on word-concept associations (conceptual route) than on
-cross-language word-to-word links (lexical route). Disentangling
-bilinguals' reliance on the conceptual and lexical and conceptual routes
-during translation is challenging, given that both sources of
-information are present during spoken word comprehensionto some extent
-in both high- and low-proficiency bilinguals. In this study, we tested
-English and Spanish native speakers in a translation elicitation task,
-in which they listened to words in an unfamiliar language (Spanish or
-Catalan for English natives, and Catalan for Spanish natives), and then
-had to translate them to their native language. Given their lack of
-previous knowledge ion the unfamiliar tested language, phonological
-similarity between the presented words and their correct translations
-was the only cue available for participants to succeed in the task. This
-allowed us to explore the informativeness of the lexical route for word
-translation, in the absence of information from the conceptual route.
-Our results indicate that participants benefited from the phonological
-similarity between translation equivalents: the more similar the word
-they listened to and its translational equivalent, the higher the
-probability of a correct translation. These results point to
-cross-language word-to-word links as providing rich information during
-word translation.
+Understanding spoken words in a non-native language is easier when their
+translations into the listener's native language are phonologically
+similar. Phonological similarity plays a more important role in the
+language performance of low-proficiency bilinguals, compared to that of
+high-proficiency bilinguals. This suggests that low-proficiency
+bilinguals may points to the latter as relying more strongly on
+cross-language word-to-word links (lexical route), while
+high-proficiency bilinguals rely more on word-concept associations
+(conceptual route) than on cross-language word-to-word links (lexical
+route). Disentangling bilinguals' reliance on the conceptual and lexical
+and conceptual routes during translation is challenging, given that both
+sources of information are present during spoken word comprehension to
+some extent in both high- and low-proficiency bilinguals. In this study,
+we tested English and Spanish native speakers in a translation
+elicitation task, in which they listened to words in an unfamiliar
+language (Spanish or Catalan for English natives, and Catalan for
+Spanish natives), and then had to translate them to their native
+language. Given their lack of previous knowledge on the unfamiliar
+tested language, phonological similarity between the presented words and
+their correct translations was the only cue available for participants
+to succeed in the task. This allowed us to explore the informativeness
+of the lexical route for word translation, in the absence of information
+from the conceptual route. Our results indicate that participants
+benefited from the phonological similarity between translation
+equivalents: the more similar the word they listened to and its
+translational equivalent, the higher the probability of a correct
+translation. These results point to cross-language word-to-word links as
+providing rich information during word translation.
 \end{abstract}
@@
 (\citeproc{ref-lecumberri2010non}{Lecumberri et al., 2010};
-\citeproc{ref-takata1990english}{Takata \& Nábělek, 1990}). However,
-there are some sources of information which may make non-native speech
-comprehension easier. One of the sources of this increased difficulty
-stems from the possible mismatch between the phonology of the native and
-the non-native languages: some acoustic features embedded in the
-non-native speech signal do not overlap perfectly with any phonemic
-category in the listener's native language. For example, imagine a
-Spanish native, with no previous familiarity with any other language,
-who listens for the first time to French may encounter the word /pɔʁt/
-(\emph{porte}), translation of door in French. The voiced uvular
-fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not
-exist in Spanish as phonemic categories. These discrepancies can
-nonetheless be perceived as allophonic variations the native (Spanish)
-phonemes /r/ and /o/ (\citeproc{ref-best2001discrimination}{Best et al.,
-2001}), and the non-native speech signal may engage lexical processing
-mechanisms (\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}),
-leading to the activation and selection, and therefore the recognition
-of its translation to Spanish /ˈpwer.ta/ (\emph{puerta}). The
-recognition of this French, unfamiliar word relies almost entirely on
-its similarity with its Spanish translation. Conversely, imagine now
-that the same Spanish listener hears English for the first time and
-encounters the word /dɔː/ (\emph{door}). This time, phonological
-similarity is of little help: /ˈpwer.ta/ and /dɔː/ share very few
-phonemes. It would come as a surprise if the listener was able to
-translate the English word successfully, if not by relying on previously
-learnt word-concept associations (which they lack, as they are
-unfamiliar to English).
-
-Cross-language phonological similarity seems, at face value, an
-important source of information for unfamiliar listeners during auditory
-word recognition, even if translations are not phonologically identical.
-Although this mismatch has a noticeable toll on comprehension
+\citeproc{ref-takata1990english}{Takata \& Nábělek, 1990}). One of the
+sources of this increased difficulty is the partial mismatch between the
+phonology of the native and the non-native languages: some acoustic
+features embedded in the non-native speech signal do not overlap
+perfectly with any phonemic category in the listener's native language.
+For example, imagine a Spanish native with no previous familiarity with
+any other language listening to French for the first time. This listener
+may encounter the word /pɔʁt/ (\emph{porte}), translation of door in
+French. The voiced uvular fricative consonant /ʁ/ and the open-mid back
+rounded vowel /ɔ/ do not exist in Spanish as phonemic categories. These
+discrepancies can nonetheless be perceived as allophonic variations the
+native (Spanish) phonemes /ɾ/ and /o/
+(\citeproc{ref-best2001discrimination}{Best et al., 2001}), leading to
+the non-native speech signal engaging lexical processing mechanisms
+(\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}). Even if
+unfamiliar with the word-form /pɔʁt/, the Spanish monolingual is likely
+to activate the phonological representation of its Spanish translation
+/pweɾta/ (\emph{puerta}), ultimately accessing the right semantic
+representation. Although the phonetic mismatch between the non-native
+word and its translation can have a noticeable toll on comprehension
 (\citeproc{ref-cutler2004patterns}{Cutler et al., 2004}), the fact that
 word recognition can take place at all in such circumstances illustrates
-that non-native listeners--even low-proficiency ones--are rarely
-completely naïve to the language they are listening to. In the present
-study, we investigate the extent to which adult listeners rely on the
-phonological similarity between their native and non-native languages
-during word recognition, capitalising on the role of \emph{cognates.}
-
-Many languages share, to some extent, similarities at the lexical level.
-This is frequently due to their typological closeness and/or
-socio-historical events involving the speakers of these languages (e.g.,
-migration, social contact). Cognates embody a particular case of such
-commonalities. Cognates are defined as cross-language synonyms which
-have similar forms whose form is similar (at the phonological level,
-orthographic level, signed level), commonly due to their shared
-etymological origin\footnote{Some form-similar cross-language synonyms
-  are technically not cognates. For example, \emph{sun} and \emph{sol}
-  (in Spanish) share phonological onset but their etymology points to
-  different origins. The complimentary case is equally problematic when
-  defining cognateness: cross-language synonyms that share etymological
-  origin may not be necessarily be recognised as cognates by the average
-  listener. Such is the case of \emph{cheese} and \emph{queso} which
-  share Latin root but are phonologically dissimilar. For simplicity, we
-  will use the term \emph{cognateness} in this paper to include all
-  cross-language synonyms that fulfill a specified threshold of form
-  similarity, regardless of etymology. We do not expect that etymology
-  plays a direct role on language perception if it is not via
-  form-similarity, since it is not necessary for participants in
-  psycholinguistic experiments to be aware of the etymology of the words
-  they encounter in the tasks to be subject to the effect of
-  form-similarity.}. For example, Romance languages such as Spanish and
-Catalan share many cognates
+that non-native listeners---even low-proficiency ones---are rarely
+completely naïve to the language they are listening to.
+
+Imagine now that the same Spanish listener hears English for the first
+time and encounters the word /dɔː/ (\emph{door}). This time,
+phonological similarity is of little help: /pweɾta/ and /dɔː/ do not
+share any phonemes. It would come as a surprise if the listener was able
+to translate the English word successfully. Non-native listeners rely on
+the phonological similarity between the word-forms presented in the
+input and their translations. Luckily for them, most languages share
+such kind of similarity. This is frequently due to their typological
+closeness and/or socio-historical events involving the speakers of these
+languages (e.g., migration, social contact). For example, Romance
+languages such as Spanish and Catalan share many form-similar
+translation equivalents (also known as \emph{cognates})
 (\citeproc{ref-schepens2012distributions}{Schepens et al., 2012}), as in
 the case of \emph{puerta} and \emph{porta} (door in Spanish and Catalan,
-respectively). Cognates play a pivotal role in many models of bilingual
-lexical processing because they provide evidence that bilinguals access
-the lexicon in a language non-selective way. For instance, pictures are
-named faster by Spanish-Catalan bilinguals when their associated labels
-in both languages are cognates (e.g., \emph{puerta}-\emph{porta}),
-compared to when their labels are non-cognates
-(\emph{mesa}-\emph{taula}, {[}table{]}), which suggests that word
-representations of both languages are activated during word retrieval
-(\citeproc{ref-costa2000cognate}{Costa et al., 2000}). This cognate
-advantage is not restricted to production, but also extends to word
-recognition (\citeproc{ref-midgley2011effects}{Midgley et al., 2011};
-\citeproc{ref-thierry2007brain}{Thierry \& Wu, 2007}), word learning
+respectively). This type of cross-linguistic similarity has been
+extremely informative for the study of bilingual lexical processing:
+words that share high form-similarity with their translations are
+processed faster and more accurately than words sharing lower
+phonological similarity. This phenomenon has provided strong evidence in
+favour of the language-non selective hypothesis of bilingual lexical
+access, which states that bilinguals activate lexical representations in
+both languages, even during monolingual situations. This facilitation
+effect has been reported for comprehension
+(\citeproc{ref-dijkstra2010cross}{Dijkstra et al., 2010};
+\citeproc{ref-midgley2011effects}{Midgley et al., 2011};
+\citeproc{ref-thierry2007brain}{Thierry \& Wu, 2007}), production
+(\citeproc{ref-costa2000cognate}{Costa et al., 2000}), learning
 (\citeproc{ref-de2000hard}{De Groot \& Keijzer, 2000};
-\citeproc{ref-elias2022cross}{Elias \& Degani, 2022};
+\citeproc{ref-elias2022crosslanguage}{Elias \& Degani, 2022};
 \citeproc{ref-lotto1998effects}{Lotto \& De Groot, 1998};
-\citeproc{ref-valente2018does}{Valente et al., 2018}), and word
-translation (\citeproc{ref-christoffels2006memory}{Christoffels et al.,
-2006}).
-
-\subsection{Lexical and conceptual routes for
-translation}\label{lexical-and-conceptual-routes-for-translation}
-
-Early theoretical accounts of this `Cognate effect' proposed that the
-impact of cognateness on lexical processing depends on bilinguals'
-proficiency in their second language (L2)
-(\citeproc{ref-potter1984lexical}{Potter et al., 1984}; see
-\citeproc{ref-chen1989patterns}{Chen \& Leung, 1989} for similar
-results). The reasoning behind this claim was the assumption that
-second-language learners start acquiring words in their L2 by
-associating them to their translation in their L1 (lexical route),
-instead of directly to their shared concept (conceptual route). This
-word-to-word connectivity should be sensitive to form-similarity (i.e.,
-cognateness): the more similar the two word- forms are, the stronger the
-connection. As learners become more proficient, the connection between
-L2 representations and their meaning grows stronger, and L2 word
-processing becomes less reliant on word-word connections between L1 and
-L2 representations. Cognates subsequently exert less impact on lexical
-processing in proficient speakers (see
-\citeproc{ref-andras2022cognate}{Andras et al., 2022} for recent
-experimental evidence for this claim). The Revised Hierarchical Model
-(RHM, \citeproc{ref-kroll1994category}{Kroll \& Stewart, 1994}) captured
-this hypothesis and predicted that translating words from L1 to L2
-should take longer than translating from L2 to L1. The rationale behind
-this prediction was that L1-to-L2 translation relies more strongly on
-word-to-word links between L1 and L2 representations, where L2-to-L1
-translation relies more strongly on the mediation between the concept
-and the two word- forms. L1-to-L2 translation should therefore be more
-sensitive to the form similarity between the L1 and the L2
-representations: cognate words should be retrieved faster than
-non-cognate words during L1-to-L2 translation.
-
-To test these predictions, Groot et al.
-(\citeproc{ref-degroot1994forward}{1994}) and Groot
-(\citeproc{ref-de1992determinants}{1992}) asked Dutch natives with high
-(but non-native) English proficiency to translate Dutch words to English
-(L1-to-L2) or English words to Dutch (L2-to-L1). Although participants'
-performance in both conditions was roughly equivalent for cognates,
-non-cognates were translated more slowly from Dutch to English than from
-English to Dutch, suggesting that L1-to-L2 translation is more sensitive
-to form-similarity than L2-to-L1 translation. Conversely, participants'
-performance was more sensitive to semantic variables (e.g.,
-concreteness) when translating English words to Dutch, compared to when
-translating Dutch words to English, suggesting that L2-to-L1 translation
-was more sensitive to the shared conceptual properties of the
-translation pair.
-
-These findings support a soft version of the RHM model, in which both
-the conceptual and the lexical translation routes are active during
-translation, but where L1-to-L2 translation relies more strongly on
-word-to-word links between L1 and L2 representations than L2-to-L1
-translation does, and therefore is more sensitive to cognateness.
-Further evidence for the presence of this difference between L1-to-L2
-and L2-to-L1 translations on the degree of reliance on the lexical and
-or the conceptual routes between L1-to-L2 and L2-to-L1 translation was
-provided by Phillips et al. (\citeproc{ref-phillips2006erp}{2006}), who
-used eletrophysiological recordings of bilinguals translating words from
-L1 to L2 or \emph{vice versa} after semantic or phonological priming.
-The authors reported stronger N400 responses (associated with semantic
-priming) in L1-to-L2 translation than in L2-to-L12 translation, and
-stronger phonological mismatch negativity responses (associated with
-phonological priming) during L2-to-L1 translation, compared to L1-to-L2
-translation.
-
-\subsection{Competition from lexical
-neighbours}\label{competition-from-lexical-neighbours}
-
-Since the RHM model was proposed, later studies on monolingual
-populations have focused how on the role of the network of connections
-that words establish networks of connections with each other at the
-form-level (phonological or orthographic) or at the conceptual level.
-For instance, the Neighbourhood Activation Model (NAM,
-\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}) highlighted
-how lexical selection is affected by the number of phonological
-neighbours that the selected word has. Words surrounded by a larger
-number of phonological neighbours (words that only differ in one phoneme
-from the target word) are responded to more slowly and less accurately
-than words from sparser phonological neighbourhoods
-(\citeproc{ref-goldinger1989priming}{Goldinger et al., 1989};
-\citeproc{ref-luce1990similarity}{Luce et al., 1990}), especially when
-these neighbours' lexical frequency is higher
-(\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}).
-
-The existence of phonological links between words is supported with
-experimental evidence from the visual world paradigm and phonological
-priming. Meyer \& Damian (\citeproc{ref-meyer2007activation}{2007})
-showed that target pictures were named faster when presented together
-with a phonologically-related distractor, suggesting that the activation
-of shared phonological information allowed the target to be activated
-and produced faster. there are also findings suggesting that
-facilitatory or inhibitory effects on target recognition depend on the
-degree of similarity the competitor shares with the target
-(\citeproc{ref-dufour2003lexical}{Dufour \& Peereman, 2003};
-\citeproc{ref-hamburger1996phonological}{Hamburger \& Slowiaczek,
-1996}). When there is only mismatch of one phoneme between prime and
-target, inhibitory effects are exerted on target recognition
-(\citeproc{ref-dufour2003lexical}{Dufour \& Peereman, 2003}). This
-inhibition was proposed to be due to strong lexical activation of the
-prime which results in lexical interference, as opposed to simple
-phonological activation from less similar primes that triggers
-facilitation.
-
-These findings of phonological links between words in the monolingual
-lexicon pose an important question in the field of bilingualism
-research: do word representations in one language form part of the
-phonological or orthographic neighbours of the other language? Weber \&
-Cutler (\citeproc{ref-weber2004lexical}{2004}) observed that bilinguals
-displayed lexical competition effects from words in their native
-language when performing a word-referent matching task in their L2. This
-highlights that lexical competition for bilinguals during language
-comprehension may not only come from the tested language but also the
-native language, despite the native language not being actively required
-for the task. Other studies have similarly found interference when
-spoken target words in one language had phonological overlap with a word
-in the participant's other language, even when the other language was
-not required by the task (\citeproc{ref-desroches2022dynamics}{Desroches
-et al., 2022}; \citeproc{ref-spivey1999cross}{Spivey \& Marian, 1999}).
-Such patterns have even been found with young children less than 4 years
-old (\citeproc{ref-von2012language}{Von Holzen \& Mani, 2012}).
-
-The Bilingual Interactive Activation (BIA/BIA+)
-(\citeproc{ref-dijkstra2002architecture}{Dijkstra \& Van Heuven, 2002};
-\citeproc{ref-van1998orthographic}{Van Heuven et al., 1998}) addressed
-this issue and suggested that lexical representations from both
-languages establish both excitatory and inhibitory connections with each
-other, resulting in an integrated lexicon for both languages. This is,
-in line with connectionist approaches to lexical processing
-(\citeproc{ref-mcclelland1981interactive}{McClelland \& Rumelhart,
-1981}). Evidence supporting such an integrated lexicon is provided by
-the fact that bilinguals' performance in word recognition tasks is
-sensitive to the orthographic neighbourhood density of the presented
-word's translation. When bilingual participants are presented with words
-in one language in a lexical decision task, translating them to the
-other language took longer when they were part of large neighbourhoods
-(\citeproc{ref-van1998orthographic}{Van Heuven et al., 1998}). This
-suggests that presented words activated orthographic neighbourhoods in
-the other language, which competed for selection with the target word
-during recognition.
-
-\subsection{The present study}\label{the-present-study}
-
-In the present study, we tested listeners' reliance on phonological
-similarity when translating words from an unfamiliar language to their
-native language, in order to investigate the plausibility of
-word-to-word connections during L1-to-L2 translation. We tested English
-native speakers and Spanish native speakers in a translation elicitation
-task. This paradigm has been used by bilingualism researchers to
-identify easily-recognisable cognates for experimental studies. Using
-this paradigm, we aim to delve deeper into the factors affecting foreign
-word recognition by monolingual participants with no prior exposure to
-the language.
-
-Participants were presented with spoken words in an unfamiliar
-(non-native language) and were asked to translate them to their native
-language. We will henceforth refer to the auditory-presented words heard
-by participants on each trial as presented words, and the correct
-translation for the presented words as target translations. The
-motivation behind testing participants in an unfamiliar language was
-that these participants can be considered a particular (extreme) case of
-unbalanced bilingualism, in which they do not know any words in L2, and
-therefore word-to-concept connections are, in principle, absent. As a
-consequence, these participants should only be able use word-to-word
-connections to succeed in the translation of words in an unfamiliar
-language. We predicted that, if the information provided by the
+\citeproc{ref-valente2018does}{Valente et al., 2018}), and translation
+(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}). The
+role that cross-linguistic phonological similarity might play in
+monolingual lexical access during non-native speech comprehension is
+unclear. How much phonological similarity is enough for a monolingual to
+understand a word from a non-native language? How does phonological
+similarity interact with other determinants of word recognition?
+
+In the present study, we investigate the extent to which monolinguals
+can exploit phonological similarity when listening to words in an
+unfamiliar languages, in order to achieve comprehension, while taking
+into account other lexically-relevant variables interact with the
+lexical processing mechanisms involved. We used a translation
+elicitation task in which participants listened to words from an
+unfamiliar language, and then tried to guess their translation in their
+native language. We will henceforth refer to the auditory-presented
+words heard by participants on each trial as \emph{presented words}, and
+the correct translation for the presented words as \emph{target
+translations}. We explored listeners' reliance on phonological
+similarity to succeed in the task by manipulating the amount of
+phonological similarity between the presented words and their target
+translations. Since participants were unfamiliar with the presented
+language, they could only be able to use word-to-word connections to
+succeed in the translation of words in an unfamiliar language. We
+therefore predicted that, if the information provided by the
 phonological form of the unfamiliar presented words is sufficient for
@@
 
-Following Luce \& Pisoni (\citeproc{ref-luce1998recognizing}{1998})`s
-NAM model, we further hypothesised that if participants are able to
-correctly translate words from an unfamiliar language based on their
-phonological similarity to the target word, similar sounding but
-incorrect translations (e.g., false friends) should be activated too. If
-this is the case, non-native words that have more phonological
-neighbours should be less likely to be correctly translated, especially
-if such neighbours have higher lexical frequency and are phonologically
-closer to the presented word than the target translation. For instance,
+The extent to which participants can rely on phonological similarity is
+limited by other factors. In particular, the presence of phonological
+neighbours (word-forms than differ in one phoneme) may interfere with
+the activation and selection of the appropriate lexical representation
+in the native language. Words with denser phonological neighbourhoods
+are recognised more slowly and less accurately than words with sparser
+phonological neighbourhoods (\citeproc{ref-dufour2003lexical}{Dufour \&
+Peereman, 2003}; \citeproc{ref-goldinger1989priming}{Goldinger et al.,
+1989}; \citeproc{ref-hamburger1996phonological}{Hamburger \& Slowiaczek,
+1996}; \citeproc{ref-luce1990similarity}{Luce et al., 1990};
+\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}). This is
+specially true if such neighbours share higher phonological similarity
+with the presented word, or are lexically more frequent than the target
+translation (\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}).
+The inhibitory effect of phonological neighbourhoods has also been
+reported across languages. For instance, Weber \& Cutler
+(\citeproc{ref-weber2004lexical}{2004}) observed that bilinguals
+displayed lexical competition effects from words in their native
+language when performing a word-referent matching task in their L2.
+
+In the case of the present task, the potential effect of cross-language
+phonological neighbourhood density is critical. The presence of
+phonologically similar non-translation equivalents (e.g., false friends)
+may act as distractor during lexical access, obstructing the selection
+of the appropriate target translation in the listener's lexicon. Even if
+the presented word and its target translation share high phonological
+similarity, participants may still provide incorrect translations if the
+presented word also shares high phonological similarity with words on
+the target language other than its translation, specially if any of them
+is more lexically frequent than the target translation. We hypothesised
+that the size of the facilitatory role of phonological similarity would
+be inversely proportional to the amount of higher-frequency phonological
+neighbours of the presented word in the target language. For instance,
 one could expect English participants to incorrectly translate the
@@
 \emph{bottom} is higher than that of \emph{button} and also phonological
-closer). This combined constraint led us to calculate a specific form of
-phonological neighbours where a neighbour is counted only if it is
-higher frequency and has stronger phonological similarity than the
-target translation as judged by Levenshtein distance. If the
-phonological neighbourhood density of the target translation affects
-participants' performance negatively, this would suggest that non-native
-words trigger competition between words in the native language during
-translation.
+closer). To test this prediction, we developed a lexical
+frequency-dependent measure of cross-language phonological neighbourhood
+density, in which a neighbour is counted only if it is higher frequency
+and is one phoneme appart from the presented word. If the phonological
+neighbourhood density of the target translation affects participants'
+performance negatively, this would suggest that non-native words trigger
+competition between words in the native language during translation.
 
 In Experiment 1, we collected data from two groups of British English
-natives. One group was presented with Catalan words, the other with
-Spanish words. We examined the extent to which participants were able to
-use the phonological similarity between the presented (Catalan or
-Spanish) word and its translation (i.e., cognateness) to provide
-accurate responses. In Experiment 2, we tested a group of (European)
-Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are both Romance languages whose close typological
-distance is reflected in the fact that they share many cognates, where
-English is a Germanic language that shares considerably fewer cognates
-with Catalan and Spanish. By testing participants translating words from
-typologically close and distant languages, we expected to widen the
-range of the phonological similarity scores of the translation pairs
-involved in the experimental task, therefore allowing us to explore
-potential cross-language differences in participant's performance.
-Participants in Experiments 1 and 2 were surprisingly good at
-translation some words, which had low phonological similarity with their
-correct translation. In Experiment 3, we collected additional data from
-a new group of British English natives, who were presented with Catalan
-and Spanish words. In contrast with the procedure in Experiment 1,
-participants in Experiment 3 where asked, after providing a response in
-each trial, to report whether they had previous knowledge of the meaning
-of the presented word (e.g., some Spanish words are common knowledge,
-even for monolingual speakers of English). Finally, we present analyses
-on the joint data sets of Experiments 1 to 3.
+natives living in the United Kingdom. One group was presented with
+Catalan words, the other with Spanish words. We examined the extent to
+which participants were able to use the phonological similarity between
+the presented word (in Catalan or Spanish) and its target translation to
+provide accurate responses. In Experiment 2, we tested a group of
+(European) Spanish natives in the same task, who were presented with
+Catalan words. Catalan and Spanish are both Romance languages whose
+close typological distance is reflected in the fact that they share many
+cognates, where English is a Germanic language that shares considerably
+fewer cognates with Catalan and Spanish. By testing participants
+translating words from typologically close and distant languages, we
+expected to widen the range of the phonological similarity scores of the
+translation pairs involved in the experimental task, therefore allowing
+us to explore potential cross-language differences in participant's
+performance. Participants in Experiments 1 and 2 were surprisingly good
+at translation some words, which had low phonological similarity with
+their correct translation. In Experiment 3, we collected additional data
+from a new group of British English natives, who were presented with
+Catalan and Spanish words. In contrast with the procedure in Experiment
+1, participants in Experiment 3 where asked, after providing a response
+in each trial, to report whether they had previous knowledge of the
+meaning of the presented word (e.g., some Spanish words are common
+knowledge, even for monolingual speakers of English). Finally, we
+present analyses on the joint data sets of Experiments 1 to 3.
 
@@
 \emph{SD} = 2.15, Range = 18-26, 46, 25 female).
-
-\captionsetup{labelsep=none}
 
@@
 deletions, and substitutions necessary to make both strings identical
-(\citeproc{ref-levenshtein1966binary}{Levenshtein et al., 1966}). We
-divided this edit distance by the number of characters included in the
-longest X-SAMPA transcription of the translation pair. This results in a
+(\citeproc{ref-levenshtein1966binary}{Levenshtein, 1966}). We divided
+this edit distance by the number of characters included in the longest
+X-SAMPA transcription of the translation pair. This results in a
 proportion score, in which values closer to zero correspond to lower
@@
 phonological transcriptions. For example, the \emph{table}
-(teɪbəl)-\emph{mesa} (mesa) translation pair had a 17\% similarity,
+(/teɪbəl/)-\emph{mesa} (/mesa/) translation pair had a 17\% similarity,
 while the \emph{train} (\texttt{trEIn})-\emph{tren} (tɾEn) translation
@@
 
-\begin{figure}[H]
-
-{\centering \includegraphics{_assets/img/model.PNG}
-
-}
-
-\caption{Eq. 1}
-
-\end{figure}%
+\[
+\begin{aligned}
+&\textbf{Likelihood}  \\
+y_{i} \sim & Bernoulli(p_{i}) \\ \\
+&\textbf{Parameters}  \\
+\text{Logit}(p_{i}) = &  \beta_{0[p,w]} + \beta_{1[p]} \text{Frequency}_{i} + \beta_{2[p]} \text{PTHN}_i + \beta_{3[p]} \text{Similarity}_i + \beta_{4[p]} (\text{PTHN}_i \times \text{Similarity}_i) \\
+\beta_{0-6[p,w]} \sim & \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \text{ and  word } w \text{ in 1, ..., } W \\
+\beta_{1-6[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\ \\
+&\textbf{Prior}  \\
+\mu_{\beta_{p,w}}  \sim &  \mathcal{N}(0, 0.1) \\
+\sigma_{\beta_{p}},  \sigma_{\beta_{w}} \sim & \text{HalfCauchy}(0, 0.1) \\
+\rho_{p}, \rho_{w} \sim & \text{LKJCorr}(8) \\
+\end{aligned}
+\]
 
@@
 All analyses were performed in R environment
-(\citeproc{ref-rcore2019r}{RCore, 2019}). We used the tidyverse family
-of R packages (\citeproc{ref-wickham2019tidyverse}{Wickham et al.,
-2019}) to process data and to generate figures. We used the
-\texttt{brms} R package (\citeproc{ref-burkner2017brms}{Bürkner, 2017})
-using the \texttt{cmdstanr} back-end to the Stan probabilistic language
+(\citeproc{ref-rcoreteam2013language}{R Core Team, 2013}). We used the
+tidyverse family of R packages
+(\citeproc{ref-wickham2019welcome}{Wickham et al., 2019}) to process
+data and to generate figures. We used the \texttt{brms} R package
+(\citeproc{ref-burkner2017brms}{Bürkner, 2017}) using the
+\texttt{cmdstanr} back-end to the Stan probabilistic language
 (\citeproc{ref-carpenter2017stan}{Carpenter et al., 2017}) to estimate
@@
 
-\captionsetup{labelsep=none}
-
 \begin{longtable}{l|rrrrrrrr}
@@
 
-Table~\ref{tbl-results} shows a summary of participants' accuracy across
+Table~\ref{tbl-dataset} shows a summary of participants' accuracy across
 Experiments 1, 2, and 3. MCMC chains in the model showed strong evidence
@@
 coefficients of interest.
-
-\captionsetup{labelsep=none}
 
@@
 
-\captionsetup{labelsep=none}\includegraphics{manuscript_files/figure-pdf/fig-coefs-1.pdf}
+\includegraphics{manuscript_files/figure-pdf/fig-coefs-1.pdf}
 
@@
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-andras2022cognate}
-Andras, F., Rivera, M., Bajo, T., Dussias, P. E., \& Paolieri, D.
-(2022). Cognate facilitation effect during auditory comprehension of a
-second language: A visual world eye-tracking study. \emph{International
-Journal of Bilingualism}, 13670069211033359.
-
 \bibitem[\citeproctext]{ref-best2001discrimination}
@@
 Broersma, P., \& Weenink, D. (2021). \emph{Praat: Doing phonetics by
-computer {[}computer program{]}} (Version 6.1.54).
+computer {[}{Computer} program{]}} (Version 6.1.54).
 \url{http://www.praat.org/}
@@
 \bibitem[\citeproctext]{ref-burkner2017brms}
-Bürkner, P.-C. (2017). Brms: An r package for bayesian multilevel models
-using stan. \emph{Journal of Statistical Software}, \emph{80}(1), 1--28.
+Bürkner, P.-C. (2017). Brms: {An R} package for {Bayesian} multilevel
+models using {Stan}. \emph{Journal of Statistical Software},
+\emph{80}(1), 1--28.
 
@@
 Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B.,
-Betancourt, M., Brubaker, M. A., Guo, J., Li, P., \& Riddell, A. (2017).
-Stan: A probabilistic programming language. \emph{Grantee Submission},
-\emph{76}(1), 1--32.
-
-\bibitem[\citeproctext]{ref-chen1989patterns}
-Chen, H.-C., \& Leung, Y.-S. (1989). Patterns of lexical processing in a
-nonnative language. \emph{Journal of Experimental Psychology: Learning,
-Memory, and Cognition}, \emph{15}(2), 316.
+Betancourt, M., Brubaker, M., Guo, J., Li, P., \& Riddell, A. (2017).
+Stan: {A} probabilistic programming language. \emph{Journal of
+Statistical Software}, \emph{76}(1), 1--32.
 
@@
 Christoffels, I. K., De Groot, A. M., \& Kroll, J. F. (2006). Memory and
-language skills in simultaneous interpreters: The role of expertise and
-language proficiency. \emph{Journal of Memory and Language},
+language skills in simultaneous interpreters: {The} role of expertise
+and language proficiency. \emph{Journal of Memory and Language},
 \emph{54}(3), 324--345.
@@
 Cutler, A., Weber, A., Smits, R., \& Cooper, N. (2004). Patterns of
-english phoneme confusions by native and non-native listeners. \emph{The
-Journal of the Acoustical Society of America}, \emph{116}(6),
+{English} phoneme confusions by native and non-native listeners.
+\emph{The Journal of the Acoustical Society of America}, \emph{116}(6),
 3668--3678.
@@
 De Groot, A. M., \& Keijzer, R. (2000). What is hard to learn is easy to
-forget: The roles of word concreteness, cognate status, and word
+forget: {The} roles of word concreteness, cognate status, and word
 frequency in foreign-language vocabulary learning and forgetting.
@@
 
-\bibitem[\citeproctext]{ref-desroches2022dynamics}
-Desroches, A. S., Friesen, D. C., Teles, M., Korade, C. A., \& Forest,
-E. W. (2022). The dynamics of spoken word recognition in bilinguals.
-\emph{Bilingualism: Language and Cognition}, \emph{25}(4), 705--710.
-
-\bibitem[\citeproctext]{ref-dijkstra2002architecture}
-Dijkstra, T., \& Van Heuven, W. J. (2002). The architecture of the
-bilingual word recognition system: From identification to decision.
-\emph{Bilingualism: Language and Cognition}, \emph{5}(3), 175--197.
+\bibitem[\citeproctext]{ref-dijkstra2010cross}
+Dijkstra, T., Miwa, K., Brummelhuis, B., Sappelli, M., \& Baayen, H.
+(2010). How cross-language similarity and task demands affect cognate
+recognition. \emph{Journal of Memory and Language}, \emph{62}(3),
+284--301.
 
@@
 Dufour, S., \& Peereman, R. (2003). Lexical competition in phonological
-priming: Assessing the role of phonological match and mismatch lengths
+priming: {Assessing} the role of phonological match and mismatch lengths
 between primes and targets. \emph{Memory \& Cognition}, \emph{31},
@@
 
-\bibitem[\citeproctext]{ref-elias2022cross}
+\bibitem[\citeproctext]{ref-elias2022crosslanguage}
 Elias, M., \& Degani, T. (2022). Cross-language interactions during
-novel word learning: The contribution of form similarity and participant
-characteristics. \emph{Bilingualism: Language and Cognition}, 1--18.
+novel word learning: {The} contribution of form similarity and
+participant characteristics. \emph{Bilingualism: Language and
+Cognition}, 1--18.
 
@@
 Goldinger, S. D., Luce, P. A., \& Pisoni, D. B. (1989). Priming lexical
-neighbors of spoken words: Effects of competition and inhibition.
+neighbors of spoken words: {Effects} of competition and inhibition.
 \emph{Journal of Memory and Language}, \emph{28}(5), 501--518.
-
-\bibitem[\citeproctext]{ref-de1992determinants}
-Groot, A. M. de. (1992). Determinants of word translation. \emph{Journal
-of Experimental Psychology: Learning, Memory, and Cognition},
-\emph{18}(5), 1001.
-
-\bibitem[\citeproctext]{ref-degroot1994forward}
-Groot, A. M. de, Dannenburg, L., \& Vanhell, J. G. (1994). Forward and
-backward word translation by bilinguals. \emph{Journal of Memory and
-Language}, \emph{33}(5), 600--629.
+\url{https://doi.org/10.1016/0749-596X(89)90009-0}
 
@@
 
-\bibitem[\citeproctext]{ref-kroll1994category}
-Kroll, J. F., \& Stewart, E. (1994). Category interference in
-translation and picture naming: Evidence for asymmetric connections
-between bilingual memory representations. \emph{Journal of Memory and
-Language}, \emph{33}(2), 149--174.
-
 \bibitem[\citeproctext]{ref-kruschke2018bayesian}
-Kruschke, J. K., \& Liddell, T. M. (2018). The bayesian new statistics:
-Hypothesis testing, estimation, meta-analysis, and power analysis from a
-bayesian perspective. \emph{Psychonomic Bulletin \& Review}, \emph{25},
-178--206.
+Kruschke, J. K., \& Liddell, T. M. (2018). The {Bayesian New
+Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power
+analysis from a {Bayesian} perspective. \emph{Psychonomic Bulletin \&
+Review}, \emph{25}, 178--206.
 
@@
 Lecumberri, M. L. G., Cooke, M., \& Cutler, A. (2010). Non-native speech
-perception in adverse conditions: A review. \emph{Speech Communication},
-\emph{52}(11-12), 864--886.
+perception in adverse conditions: {A} review. \emph{Speech
+Communication}, \emph{52}(11-12), 864--886.
 
 \bibitem[\citeproctext]{ref-levenshtein1966binary}
-Levenshtein, V. I. et al. (1966). Binary codes capable of correcting
-deletions, insertions, and reversals. \emph{Soviet Physics Doklady},
-\emph{10}, 707--710.
+Levenshtein, V. I. (1966). Binary codes capable of correcting deletions,
+insertions, and reversals. \emph{Soviet Physics Doklady}, \emph{10},
+707--710.
 
@@
 \bibitem[\citeproctext]{ref-luce1998recognizing}
-Luce, P. A., \& Pisoni, D. B. (1998). Recognizing spoken words: The
+Luce, P. A., \& Pisoni, D. B. (1998). Recognizing spoken words: {The}
 neighborhood activation model. \emph{Ear and Hearing}, \emph{19}(1), 1.
@@
 Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012).
-\emph{CLEARPOND: Cross-linguistic easy-access resource for phonological
-and orthographic neighborhood densities}.
-
-\bibitem[\citeproctext]{ref-mcclelland1981interactive}
-McClelland, J. L., \& Rumelhart, D. E. (1981). An interactive activation
-model of context effects in letter perception: I. An account of basic
-findings. \emph{Psychological Review}, \emph{88}(5), 375.
-
-\bibitem[\citeproctext]{ref-meyer2007activation}
-Meyer, A. S., \& Damian, M. F. (2007). Activation of distractor names in
-the picture-picture interference paradigm. \emph{Memory \& Cognition},
-\emph{35}(3), 494--503.
+\emph{{CLEARPOND}: {Cross-linguistic} easy-access resource for
+phonological and orthographic neighborhood densities}.
 
@@
 Midgley, K. J., Holcomb, P. J., \& Grainger, J. (2011). Effects of
-cognate status on word comprehension in second language learners: An ERP
-investigation. \emph{Journal of Cognitive Neuroscience}, \emph{23}(7),
-1634--1647.
+cognate status on word comprehension in second language learners: {An
+ERP} investigation. \emph{Journal of Cognitive Neuroscience},
+\emph{23}(7), 1634--1647.
 
@@
 Peirce, J., Gray, J. R., Simpson, S., MacAskill, M., Höchenberger, R.,
-Sogo, H., Kastman, E., \& Lindeløv, J. K. (2019). PsychoPy2: Experiments
-in behavior made easy. \emph{Behavior Research Methods}, \emph{51}(1),
-195--203.
-
-\bibitem[\citeproctext]{ref-phillips2006erp}
-Phillips, N. A., Klein, D., Mercier, J., \& Boysson, C. de. (2006). ERP
-measures of auditory word repetition and translation priming in
-bilinguals. \emph{Brain Research}, \emph{1125}(1), 116--131.
-
-\bibitem[\citeproctext]{ref-potter1984lexical}
-Potter, M. C., So, K.-F., Von Eckardt, B., \& Feldman, L. B. (1984).
-Lexical and conceptual representation in beginning and proficient
-bilinguals. \emph{Journal of Verbal Learning and Verbal Behavior},
-\emph{23}(1), 23--38.
-
-\bibitem[\citeproctext]{ref-rcore2019r}
-RCore, T. (2019). \emph{R: A language and environment for statistical
-computing. R foundation for statistical computing, austria}.
+Sogo, H., Kastman, E., \& Lindeløv, J. K. (2019). {PsychoPy2}:
+{Experiments} in behavior made easy. \emph{Behavior Research Methods},
+\emph{51}(1), 195--203. \url{https://doi.org/10.3758/s13428-018-01193-y}
+
+\bibitem[\citeproctext]{ref-rcoreteam2013language}
+R Core Team. (2013). \emph{R: {A Language} and {Environment} for
+{Statistical Computing}}. R Foundation for Statistical Computing.
+\url{http://www.R-project.org/}
 
@@
 Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of
-cognates in europe as based on levenshtein distance. \emph{Bilingualism:
-Language and Cognition}, \emph{15}(1), 157--166.
-
-\bibitem[\citeproctext]{ref-spivey1999cross}
-Spivey, M. J., \& Marian, V. (1999). Cross talk between native and
-second languages: Partial activation of an irrelevant lexicon.
-\emph{Psychological Science}, \emph{10}(3), 281--284.
+cognates in {Europe} as based on {Levenshtein} distance.
+\emph{Bilingualism: Language and Cognition}, \emph{15}(1), 157--166.
 
@@
 Takata, Y., \& Nábělek, A. K. (1990). English consonant recognition in
-noise and in reverberation by japanese and american listeners. \emph{The
-Journal of the Acoustical Society of America}, \emph{88}(2), 663--666.
+noise and in reverberation by {Japanese} and {American} listeners.
+\emph{The Journal of the Acoustical Society of America}, \emph{88}(2),
+663--666.
 
@@
 van der Loo, M. P. J. (2014). The stringdist package for approximate
-string matching. \emph{The {R} {J}ournal}, \emph{6}, 111--122.
+string matching. \emph{The R Journal}, \emph{6}(1), 111--122.
 \url{https://CRAN.R-project.org/package=stringdist}
-
-\bibitem[\citeproctext]{ref-van1998orthographic}
-Van Heuven, W. J., Dijkstra, T., \& Grainger, J. (1998). Orthographic
-neighborhood effects in bilingual word recognition. \emph{Journal of
-Memory and Language}, \emph{39}(3), 458--483.
 
@@
 Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
-SUBTLEX-UK: A new and improved word frequency database for british
-english. \emph{Quarterly Journal of Experimental Psychology},
+{SUBTLEX-UK}: {A} new and improved word frequency database for {British
+English}. \emph{Quarterly Journal of Experimental Psychology},
 \emph{67}(6), 1176--1190.
-
-\bibitem[\citeproctext]{ref-von2012language}
-Von Holzen, K., \& Mani, N. (2012). Language nonselective lexical access
-in bilingual toddlers. \emph{Journal of Experimental Child Psychology},
-\emph{113}(4), 569--586.
 
@@
 
-\bibitem[\citeproctext]{ref-wickham2019tidyverse}
+\bibitem[\citeproctext]{ref-wickham2019welcome}
 Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
@@
 Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
-Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
+Welcome to the tidyverse. \emph{Journal of Open Source Software},
 \emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
-
-\bibitem[\citeproctext]{ref-andras2022cognate}
-Andras, F., Rivera, M., Bajo, T., Dussias, P. E., \& Paolieri, D.
-(2022). Cognate facilitation effect during auditory comprehension of a
-second language: A visual world eye-tracking study. \emph{International
-Journal of Bilingualism}, 13670069211033359.
 
@@
 Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
-Welcome to the {tidyverse}. \emph{Journal of Open Source Software},
+Welcome to the tidyverse. \emph{Journal of Open Source Software},
 \emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}

### 2024-07-22 manuscript/manuscript.qmd
New version of introduction

@@
----
-title: "The role of Similarity in native spoken word recognition"
----
-
 ```{r}
@@
 
-Listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
-One of the sources of this increased difficulty is the partial mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener's native language.
+::: {.content-visible when-format="pdf"}
+
+Learning a new language is a non-trivial task, a large part of which is acquiring a new set of word-forms to correspond to existing semantic representations in their lexicon.
+Imagine that a Spanish listener hears English for the first time and encounters the word /\textipa{"dO:}/ (*door*).
+Their knowledge of the phonological form in their native language is of little help: /\textipa{"pweR.ta}/ and /\textipa{"dO:}/ do not share any phonemes.
+It would come as a surprise if the listener was able to translate the English word successfully without having first learnt the word.
+However, there are some elements in language that may help the learner in acquiring words in a new language.
+Many languages share some degree of similarity.
+This may be due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
+For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents (also known as cognates) [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographical similarity of the former to the equivalent in their native language.
+Second language learners have been found to learn cognates more easily than non-cognates [@hall2002automatic; @otwinowska2019more].
+This type of cross-linguistic similarity has also been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
+This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
+This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
+
+Event counting on this facilitation effect of form-similarity, listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
+One of potential source of this increased difficulty is the partial mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener’s native language.
+For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
+This listener may encounter the word /\textipa{"pOKt}/ (*porte*), translation of door in French.
+When comparing the phonemic transcription to the transcription of the Spanish word /\textipa{"pweR.ta}/ (*puerta*), one may be led to infer that these two words do not have much overlap.
+But one thing to note is that the voiced uvular fricative consonant /\textipa{K}/ and the open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /\textipa{R}/ and /o/.
+A perceptual assimilation model has been proposed to account for how speakers may integrate non-native phonemes into their native phonemic categories based on goodness-of-fit [@best2001discrimination; @best1988examination].
+Although phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish monolingual may still be able to activate the phonological representation of /\textipa{"pweR.ta}/ (*puerta*) from hearing the word-form /\textipa{"pOKt}/ (*porte*), ultimately allowing them to access the right semantic representation.
+
+:::
+
+::: {.content-visible when-format="docx"}
+
+Learning a new language is a non-trivial task, a large part of which is acquiring a new set of word-forms to correspond to existing semantic representations in their lexicon.
+Imagine that a Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
+Their knowledge of the phonological form in their native language is of little help: /pweɾta/ and /dɔː/ do not share any phonemes.
+It would come as a surprise if the listener was able to translate the English word successfully without having first learnt the word.
+However, there are some elements in language that may help the learner in acquiring words in a new language.
+Many languages share some degree of similarity.
+This may be due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
+For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents (also known as cognates) [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographical similarity of the former to the equivalent in their native language.
+Second language learners have been found to learn cognates more easily than non-cognates [@hall2002automatic; @otwinowska2019more].
+This type of cross-linguistic similarity has also been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
+This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
+This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
+
+Event counting on this facilitation effect of form-similarity, listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
+One of potential source of this increased difficulty is the partial mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener’s native language.
 For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
 This listener may encounter the word /pɔʁt/ (*porte*), translation of door in French.
-The voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/ do not exist in Spanish as phonemic categories.
-These discrepancies can nonetheless be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/ [@best2001discrimination], leading to the non-native speech signal engaging lexical processing mechanisms [@weber2004lexical].
-Even if unfamiliar with the word-form /pɔʁt/, the Spanish monolingual is likely to activate the phonological representation of its Spanish translation /pweɾta/ (*puerta*), ultimately accessing the right semantic representation.
-Although the phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension [@cutler2004patterns], the fact that word recognition can take place at all in such circumstances illustrates that non-native listeners---even low-proficiency ones---are rarely completely naïve to the language they are listening to.
-
-Imagine now that the same Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
-This time, phonological similarity is of little help: /pweɾta/ and /dɔː/ do not share any phonemes.
-It would come as a surprise if the listener was able to translate the English word successfully.
-Non-native listeners rely on the phonological similarity between the word-forms presented in the input and their translations.
-Luckily for them, most languages share such kind of similarity.
-This is frequently due to their typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
-For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents (also known as *cognates*) [@schepens2012distributions], as in the case of *puerta* and *porta* (door in Spanish and Catalan, respectively).
-This type of cross-linguistic similarity has been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
-This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
-This facilitation effect has been reported for comprehension [@thierry2007brain; @midgley2011effects; @dijkstra2010cross], production [@costa2000cognate], learning [@de2000hard; @lotto1998effects; @valente2018does; @elias2022crosslanguage], and translation [@christoffels2006memory].
-The role that cross-linguistic phonological similarity might play in monolingual lexical access during non-native speech comprehension is unclear.
-How much phonological similarity is enough for a monolingual to understand a word from a non-native language?
-How does phonological similarity interact with other determinants of word recognition?
-
-In the present study, we investigate the extent to which monolinguals can exploit phonological similarity when listening to words in an unfamiliar languages, in order to achieve comprehension, while taking into account other lexically-relevant variables interact with the lexical processing mechanisms involved.
-We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as *presented words*, and the correct translation for the presented words as *target translations*.
-We explored listeners' reliance on phonological similarity to succeed in the task by manipulating the amount of phonological similarity between the presented words and their target translations.
-Since participants were unfamiliar with the presented language, they could only be able to use word-to-word connections to succeed in the translation of words in an unfamiliar language.
-We therefore predicted that, if the information provided by the phonological form of the unfamiliar presented words is sufficient for translation, participants' performance should increase when the translation pairs are phonologically more similar.
-
-The extent to which participants can rely on phonological similarity is limited by other factors.
-In particular, the presence of phonological neighbours (word-forms than differ in one phoneme) may interfere with the activation and selection of the appropriate lexical representation in the native language.
-Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@luce1998recognizing; @goldinger1989priming; @luce1990similarity; @hamburger1996phonological; @dufour2003lexical].
-This is specially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
+When comparing the phonemic transcription to the transcription of the Spanish word /pweɾta/ (*puerta*), one may be led to infer that these two words do not have much overlap.
+But one thing to note is that the voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/.
+A perceptual assimilation model has been proposed to account for how speakers may integrate non-native phonemes into their native phonemic categories based on goodness-of-fit [@best2001discrimination; @best1988examination].
+Although phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish monolingual may still be able to activate the phonological representation of /pweɾta/ (*puerta*) from hearing the word-form /pɔʁt/ (*porte*), ultimately allowing them to access the right semantic representation.
+
+:::
+
+While phonological similarity across translation equivalents can help with speech recognition, phonological similarity between words with different meanings can hinder the recognition process.
+In word retrieval, the presence of phonological neighbours (word-forms that differ from each other in one phoneme) may interfere with the activation and selection of the appropriate lexical representation in the native language.
+Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
+This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
 The inhibitory effect of phonological neighbourhoods has also been reported across languages.
 For instance, @weber2004lexical observed that bilinguals displayed lexical competition effects from words in their native language when performing a word-referent matching task in their L2.
-
-In the case of the present task, the potential effect of cross-language phonological neighbourhood density is critical.
-The presence of phonologically similar non-translation equivalents (e.g., false friends) may act as distractor during lexical access, obstructing the selection of the appropriate target translation in the listener's lexicon.
-Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with words on the target language other than its translation, specially if any of them is more lexically frequent than the target translation.
+The assimilation of non-native phonemes into native categories [@best2001discrimination; @best1988examination] was used to explain how lexical competitors in the native language can be activated even when the target words contain non-native phonemes.
+In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
+For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
+It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect cross-linguistic phonological similarity has on word recognition in a foreign language.
+This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
+
+In the present study, we investigate the extent to which monolinguals can exploit phonological similarity between translation equivalents when listening to words in an unfamiliar language, and how it interacts with competition from phonological neighbours in the native language.
+We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
+We explored listeners’ reliance on phonological similarity to succeed in the task by manipulating the amount of phonological similarity between the presented words and their target translations.
+Since participants were unfamiliar with the presented language, they could only use phonological similarity between the presented language and their native language to successfully translate the words.
+We therefore predicted that participants’ performance should increase when the translation pairs are phonologically more similar.
+We also predicted that there is a minimum threshold of phonological similarity to be sufficient for translation.
+
+We further investigated the effect of phonological competitors.
+Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language.
+The effect is predicted to be especially strong if any of the competitors is more lexically frequent than the target translation.
 We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
-For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as bottom instead of as its correct translation *button* (because the lexical frequency of *bottom* is higher than that of *button* and also phonological closer).
-To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme appart from the presented word.
-If the phonological neighbourhood density of the target translation affects participants' performance negatively, this would suggest that non-native words trigger competition between words in the native language during translation.
+For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
+To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme apart from the presented word.
+If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.We conducted a series of three experiments to test these predictions.
 
@@
 Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from typologically close and distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
-Participants in Experiments 1 and 2 were surprisingly good at translation some words, which had low phonological similarity with their correct translation.
-In Experiment 3, we collected additional data from a new group of British English natives, who were presented with Catalan and Spanish words.
-In contrast with the procedure in Experiment 1, participants in Experiment 3 where asked, after providing a response in each trial, to report whether they had previous knowledge of the meaning of the presented word (e.g., some Spanish words are common knowledge, even for monolingual speakers of English).
-Finally, we present analyses on the joint data sets of Experiments 1 to 3.
+By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant’s performance.
+One unexpected finding was that participants in Experiments 1 and 2 were surprisingly good at translating a subset of words which had low phonological similarity with their correct translation.
+We were concerned that this may be caused by some prior knowledge of specific words by our participants, as some Spanish words are commonly seen in media or product labels, making them familiar even to monolingual speakers of English.
+In Experiment 3, we collected additional data from a new group of British English natives.
+The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
+In the final section of this paper, we present analyses on the joint data sets of Experiments 1 to 3.
 
@@
 
-We calculated *Similaritys*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
+We calculated *Similarity*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
 The Levenshtein distance computes the edit distance between two character strings---in this case, phoneme transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
@@
 For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
-For example, the *table* (/teɪbəl/)-*mesa* (/mesa/) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (t`r knitr::asis_output("\U027E")`En) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
-Figure 1 summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+For example, the *table* (`teIb@l`)-*mesa* (`mesa`) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (`tREn`) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+@tbl-stimuli summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
 
@@
 *Other* responses were excluded from data analysis.
-
 Trials in which participants took longer than 10 seconds to respond were also excluded.
-Participants contributed a total of `r nrow(exp_responses)` valid trials (`r nrow(exp_responses[exp_responses$group %in% c("cat-ENG", "cat-SPA"),])` in Catalan, `r nrow(exp_responses[exp_responses$group %in% c("spa-ENG"),])` in Spanish).
+Participants contributed a total of `r nrow(exp_responses)` valid trials (`r format(nrow(exp_responses[exp_responses$group %in% c("cat-ENG", "cat-SPA"),]), big.mark = ",")` in Catalan, `r format(nrow(exp_responses[exp_responses$group %in% c("spa-ENG"),]), big.mark = ",")` in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 This proportion indicates the probability that the true value of the coefficient is equivalent to zero.
-
 All analyses were performed in R environment [@rcoreteam2013language].
@@
 Of those trials, `r format(nrow(dataset$cat_eng), big.mark = ",")` were provided by `r n_participants$cat_eng` unique participants who listened to Catalan words, and `r format(nrow(dataset$spa_eng), big.mark = ",")` trials were provided by `r n_participants$spa_eng` unique participants who listened to Spanish words.
-We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_eng["blank"]`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (e.g., `f`, *n* = `r r_validity$cat_eng["incomplete"]`), and in which participants added comments to the experimenter (e.g., `unsure`, (*n* = `r r_validity$cat_eng["other"]`).
+We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_eng["blank"]`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (e.g., `f`, *n* = `r r_validity$cat_eng["incomplete"]`), and in which participants added comments to the experimenter (e.g., `unsure`, *n* = `r r_validity$cat_eng["other"]`).
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
@@
 #| label: fig-epreds-1
-#| fig-height: 6
-#| fig-width: 10
+#| fig-height: 5
+#| fig-width: 11
 #| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
@@
 
-On each trial, participants listened to the audio stimulus by clicking on the play button.
+On each trial, participants listened to the audio stimulus by clicking on the `PLAY` button.
 For comparability to the PsychoPy version, participants were only allowed to play the audio one time.
 Participants were explicitly told that they would be only allowed to listen once.
-The play button vanished after one playthrough.
+The `PLAY` button vanished after one playthrough.
 Participants then had to answer three questions based on the audio they had heard on that trial.
 These questions were presented on the same page, directly below the audio player.
-They were first asked whether or not they knew the presented word (multiple choice---yes/no).
+They were first asked whether or not they knew the presented word (multiple choice---*yes*/*no*).
 Regardless of their answer on the first question, participants were asked what they thought the translation of the word was in English (or their best guess), and instructed to type their answer in the provided text box.

### 2024-07-22 manuscript/manuscript.tex
New version of introduction

@@
 \hypersetup{
-  pdftitle={The role of cognateness in native spoken word recognition},
+  pdftitle={The role of cognateness in recognition of spoken words in a foreign language},
   pdfauthor={Gonzalo Garcia-Castro; Serene Siow; Kim Plunkett; Nuria Sebastian-Galles},
@@
 \newcommand{\runninghead}{A Preprint }
-\renewcommand{\runninghead}{Cognateness and native word recognition }
-\title{The role of cognateness in native spoken word recognition}
+\renewcommand{\runninghead}{Cognateness and recognition of foreign
+words }
+\title{The role of cognateness in recognition of spoken words in a
+foreign language}
 \author{\textbf{Gonzalo
@@
 
-Listening to non-native speech is costlier than listening to native
-speech, even for highly proficient bilinguals, and especially in
-acoustically adverse situations
-(\citeproc{ref-lecumberri2010non}{Lecumberri et al., 2010};
-\citeproc{ref-takata1990english}{Takata \& Nábělek, 1990}). One of the
-sources of this increased difficulty is the partial mismatch between the
-phonology of the native and the non-native languages: some acoustic
-features embedded in the non-native speech signal do not overlap
-perfectly with any phonemic category in the listener's native language.
-For example, imagine a Spanish native with no previous familiarity with
-any other language listening to French for the first time. This listener
-may encounter the word /pɔʁt/ (\emph{porte}), translation of door in
-French. The voiced uvular fricative consonant /ʁ/ and the open-mid back
-rounded vowel /ɔ/ do not exist in Spanish as phonemic categories. These
-discrepancies can nonetheless be perceived as allophonic variations the
-native (Spanish) phonemes /ɾ/ and /o/
-(\citeproc{ref-best2001discrimination}{Best et al., 2001}), leading to
-the non-native speech signal engaging lexical processing mechanisms
-(\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}). Even if
-unfamiliar with the word-form /pɔʁt/, the Spanish monolingual is likely
-to activate the phonological representation of its Spanish translation
-/pweɾta/ (\emph{puerta}), ultimately accessing the right semantic
-representation. Although the phonetic mismatch between the non-native
-word and its translation can have a noticeable toll on comprehension
-(\citeproc{ref-cutler2004patterns}{Cutler et al., 2004}), the fact that
-word recognition can take place at all in such circumstances illustrates
-that non-native listeners---even low-proficiency ones---are rarely
-completely naïve to the language they are listening to.
-
-Imagine now that the same Spanish listener hears English for the first
-time and encounters the word /dɔː/ (\emph{door}). This time,
-phonological similarity is of little help: /pweɾta/ and /dɔː/ do not
-share any phonemes. It would come as a surprise if the listener was able
-to translate the English word successfully. Non-native listeners rely on
-the phonological similarity between the word-forms presented in the
-input and their translations. Luckily for them, most languages share
-such kind of similarity. This is frequently due to their typological
+Learning a new language is a non-trivial task, a large part of which is
+acquiring a new set of word-forms to correspond to existing semantic
+representations in their lexicon. Imagine that a Spanish listener hears
+English for the first time and encounters the word /\textipa{"dO:}/
+(\emph{door}). Their knowledge of the phonological form in their native
+language is of little help: /\textipa{"pweR.ta}/ and /\textipa{"dO:}/ do
+not share any phonemes. It would come as a surprise if the listener was
+able to translate the English word successfully without having first
+learnt the word. However, there are some elements in language that may
+help the learner in acquiring words in a new language. Many languages
+share some degree of similarity. This may be due to typological
 closeness and/or socio-historical events involving the speakers of these
@@
 languages such as Spanish and Catalan share many form-similar
-translation equivalents (also known as \emph{cognates})
+translation equivalents (also known as cognates)
 (\citeproc{ref-schepens2012distributions}{Schepens et al., 2012}), as in
-the case of \emph{puerta} and \emph{porta} (door in Spanish and Catalan,
-respectively). This type of cross-linguistic similarity has been
-extremely informative for the study of bilingual lexical processing:
-words that share high form-similarity with their translations are
-processed faster and more accurately than words sharing lower
-phonological similarity. This phenomenon has provided strong evidence in
-favour of the language-non selective hypothesis of bilingual lexical
-access, which states that bilinguals activate lexical representations in
-both languages, even during monolingual situations. This facilitation
-effect has been reported for comprehension
-(\citeproc{ref-dijkstra2010cross}{Dijkstra et al., 2010};
-\citeproc{ref-midgley2011effects}{Midgley et al., 2011};
+the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and
+Catalan, respectively). Given no prior knowledge of either language, a
+Spanish native speaker is likely to be much more successful at correctly
+translating Catalan \emph{porta} than English \emph{door}, due to the
+phonological and orthographical similarity of the former to the
+equivalent in their native language. Second language learners have been
+found to learn cognates more easily than non-cognates
+(\citeproc{ref-hall2002automatic}{Hall, 2002};
+\citeproc{ref-otwinowska2019more}{Otwinowska \& Szewczyk, 2019}). This
+type of cross-linguistic similarity has also been extremely informative
+for the study of bilingual lexical processing: words that share high
+form-similarity with their translations are processed faster and more
+accurately than words sharing lower phonological similarity. This
+phenomenon has provided strong evidence in favour of the language-non
+selective hypothesis of bilingual lexical access, which states that
+bilinguals activate lexical representations in both languages, even
+during monolingual situations. This facilitation effect has been
+reported for comprehension (\citeproc{ref-dijkstra2010cross}{Dijkstra et
+al., 2010}; \citeproc{ref-midgley2011effects}{Midgley et al., 2011};
 \citeproc{ref-thierry2007brain}{Thierry \& Wu, 2007}), production
@@
 \citeproc{ref-valente2018does}{Valente et al., 2018}), and translation
-(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}). The
-role that cross-linguistic phonological similarity might play in
-monolingual lexical access during non-native speech comprehension is
-unclear. How much phonological similarity is enough for a monolingual to
-understand a word from a non-native language? How does phonological
-similarity interact with other determinants of word recognition?
-
-In the present study, we investigate the extent to which monolinguals
-can exploit phonological similarity when listening to words in an
-unfamiliar languages, in order to achieve comprehension, while taking
-into account other lexically-relevant variables interact with the
-lexical processing mechanisms involved. We used a translation
-elicitation task in which participants listened to words from an
-unfamiliar language, and then tried to guess their translation in their
-native language. We will henceforth refer to the auditory-presented
-words heard by participants on each trial as \emph{presented words}, and
-the correct translation for the presented words as \emph{target
-translations}. We explored listeners' reliance on phonological
-similarity to succeed in the task by manipulating the amount of
-phonological similarity between the presented words and their target
-translations. Since participants were unfamiliar with the presented
-language, they could only be able to use word-to-word connections to
-succeed in the translation of words in an unfamiliar language. We
-therefore predicted that, if the information provided by the
-phonological form of the unfamiliar presented words is sufficient for
-translation, participants' performance should increase when the
-translation pairs are phonologically more similar.
-
-The extent to which participants can rely on phonological similarity is
-limited by other factors. In particular, the presence of phonological
-neighbours (word-forms than differ in one phoneme) may interfere with
-the activation and selection of the appropriate lexical representation
-in the native language. Words with denser phonological neighbourhoods
-are recognised more slowly and less accurately than words with sparser
-phonological neighbourhoods (\citeproc{ref-dufour2003lexical}{Dufour \&
-Peereman, 2003}; \citeproc{ref-goldinger1989priming}{Goldinger et al.,
-1989}; \citeproc{ref-hamburger1996phonological}{Hamburger \& Slowiaczek,
-1996}; \citeproc{ref-luce1990similarity}{Luce et al., 1990};
+(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}).
+
+Event counting on this facilitation effect of form-similarity, listening
+to non-native speech is costlier than listening to native speech, even
+for highly proficient bilinguals, and especially in acoustically adverse
+situations (\citeproc{ref-lecumberri2010non}{Lecumberri et al., 2010};
+\citeproc{ref-takata1990english}{Takata \& Nábělek, 1990}). One of
+potential source of this increased difficulty is the partial mismatch
+between the phonology of the native and the non-native languages: some
+acoustic features embedded in the non-native speech signal do not
+overlap perfectly with any phonemic category in the listener's native
+language. For example, imagine a Spanish native with no previous
+familiarity with any other language listening to French for the first
+time. This listener may encounter the word /\textipa{"pOKt}/
+(\emph{porte}), translation of door in French. When comparing the
+phonemic transcription to the transcription of the Spanish word
+/\textipa{"pweR.ta}/ (\emph{puerta}), one may be led to infer that these
+two words do not have much overlap. But one thing to note is that the
+voiced uvular fricative consonant /\textipa{K}/ and the open-mid back
+rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic
+categories, can be perceived as allophonic variations the native
+(Spanish) phonemes /\textipa{R}/ and /o/. A perceptual assimilation
+model has been proposed to account for how speakers may integrate
+non-native phonemes into their native phonemic categories based on
+goodness-of-fit (\citeproc{ref-best1988examination}{Best et al., 1988},
+\citeproc{ref-best2001discrimination}{2001}). Although phonetic mismatch
+between the non-native word and its translation can have a noticeable
+toll on comprehension Cutler et al.
+(\citeproc{ref-cutler2004patterns}{2004}), the Spanish monolingual may
+still be able to activate the phonological representation of
+/\textipa{"pweR.ta}/ (\emph{puerta}) from hearing the word-form
+/\textipa{"pOKt}/ (\emph{porte}), ultimately allowing them to access the
+right semantic representation.
+
+While phonological similarity across translation equivalents can help
+with speech recognition, phonological similarity between words with
+different meanings can hinder the recognition process. In word
+retrieval, the presence of phonological neighbours (word-forms that
+differ from each other in one phoneme) may interfere with the activation
+and selection of the appropriate lexical representation in the native
+language. Words with denser phonological neighbourhoods are recognised
+more slowly and less accurately than words with sparser phonological
+neighbourhoods (\citeproc{ref-dufour2003lexical}{Dufour \& Peereman,
+2003}; \citeproc{ref-goldinger1989priming}{Goldinger et al., 1989};
+\citeproc{ref-hamburger1996phonological}{Hamburger \& Slowiaczek, 1996};
+\citeproc{ref-luce1990similarity}{Luce et al., 1990};
 \citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}). This is
-specially true if such neighbours share higher phonological similarity
+especially true if such neighbours share higher phonological similarity
 with the presented word, or are lexically more frequent than the target
@@
 displayed lexical competition effects from words in their native
-language when performing a word-referent matching task in their L2.
-
-In the case of the present task, the potential effect of cross-language
-phonological neighbourhood density is critical. The presence of
-phonologically similar non-translation equivalents (e.g., false friends)
-may act as distractor during lexical access, obstructing the selection
-of the appropriate target translation in the listener's lexicon. Even if
+language when performing a word-referent matching task in their L2. The
+assimilation of non-native phonemes into native categories
+(\citeproc{ref-best1988examination}{Best et al., 1988},
+\citeproc{ref-best2001discrimination}{2001}) was used to explain how
+lexical competitors in the native language can be activated even when
+the target words contain non-native phonemes. In the case of non-native
+speech recognition, the presence of cross-linguistic pairs which are
+phonologically similar but differ in meaning (e.g., false friends) may
+act as distractors during lexical access, obstructing the selection of
+the appropriate target translation in the listener's lexicon. For
+instance, Otwinowska \& Szewczyk
+(\citeproc{ref-otwinowska2019more}{2019}) reported that false friends
+were disadvantaged relative to non-cognates by Polish second language
+learners, in contrast to cognates which were known better. It is
+therefore important to investigate the joint effect of both cognates and
+false friends when investigating the effect cross-linguistic
+phonological similarity has on word recognition in a foreign language.
+This could shed light on available strategies and challenges associated
+with the early stages of second language acquisition.
+
+In the present study, we investigate the extent to which monolinguals
+can exploit phonological similarity between translation equivalents when
+listening to words in an unfamiliar language, and how it interacts with
+competition from phonological neighbours in the native language. We used
+a translation elicitation task in which participants listened to words
+from an unfamiliar language, and then tried to guess their translation
+in their native language. We will henceforth refer to the
+auditory-presented words heard by participants on each trial as
+presented words, and the correct translation for the presented words as
+target translations. We explored listeners' reliance on phonological
+similarity to succeed in the task by manipulating the amount of
+phonological similarity between the presented words and their target
+translations. Since participants were unfamiliar with the presented
+language, they could only use phonological similarity between the
+presented language and their native language to successfully translate
+the words. We therefore predicted that participants' performance should
+increase when the translation pairs are phonologically more similar. We
+also predicted that there is a minimum threshold of phonological
+similarity to be sufficient for translation.
+
+We further investigated the effect of phonological competitors. Even if
 the presented word and its target translation share high phonological
 similarity, participants may still provide incorrect translations if the
-presented word also shares high phonological similarity with words on
-the target language other than its translation, specially if any of them
-is more lexically frequent than the target translation. We hypothesised
-that the size of the facilitatory role of phonological similarity would
-be inversely proportional to the amount of higher-frequency phonological
-neighbours of the presented word in the target language. For instance,
-one could expect English participants to incorrectly translate the
-Spanish word \emph{botón} as bottom instead of as its correct
-translation \emph{button} (because the lexical frequency of
-\emph{bottom} is higher than that of \emph{button} and also phonological
-closer). To test this prediction, we developed a lexical
-frequency-dependent measure of cross-language phonological neighbourhood
-density, in which a neighbour is counted only if it is higher frequency
-and is one phoneme appart from the presented word. If the phonological
-neighbourhood density of the target translation affects participants'
-performance negatively, this would suggest that non-native words trigger
-competition between words in the native language during translation.
+presented word also shares high phonological similarity with other words
+in the native language. The effect is predicted to be especially strong
+if any of the competitors is more lexically frequent than the target
+translation. We hypothesised that the size of the facilitatory role of
+phonological similarity would be inversely proportional to the amount of
+higher-frequency phonological neighbours of the presented word in the
+target language. For instance, one could expect English participants to
+incorrectly translate the Spanish word \emph{botón} as \emph{bottom}
+instead of as its correct translation \emph{button}, due to combined
+effect of the close phonological similarity between \emph{botón} and
+\emph{bottom} along with the high lexical frequency of \emph{bottom}
+relative to \emph{button}. To test this prediction, we developed a
+lexical frequency-dependent measure of cross-language phonological
+neighbourhood density, in which a neighbour is counted only if it is
+higher frequency and is one phoneme apart from the presented word. If
+the phonological neighbourhood density of the target translation affects
+participants' performance negatively, this would suggest that
+competitors in the native language affect recognition of non-native
+words during recognition of foreign speech.We conducted a series of
+three experiments to test these predictions.
 
@@
 fewer cognates with Catalan and Spanish. By testing participants
-translating words from typologically close and distant languages, we
+translating words from typologically close or distant languages, we
 expected to widen the range of the phonological similarity scores of the
@@
 us to explore potential cross-language differences in participant's
-performance. Participants in Experiments 1 and 2 were surprisingly good
-at translation some words, which had low phonological similarity with
-their correct translation. In Experiment 3, we collected additional data
-from a new group of British English natives, who were presented with
-Catalan and Spanish words. In contrast with the procedure in Experiment
-1, participants in Experiment 3 where asked, after providing a response
-in each trial, to report whether they had previous knowledge of the
-meaning of the presented word (e.g., some Spanish words are common
-knowledge, even for monolingual speakers of English). Finally, we
-present analyses on the joint data sets of Experiments 1 to 3.
+performance. One unexpected finding was that participants in Experiments
+1 and 2 were surprisingly good at translating a subset of words which
+had low phonological similarity with their correct translation. We were
+concerned that this may be caused by some prior knowledge of specific
+words by our participants, as some Spanish words are commonly seen in
+media or product labels, making them familiar even to monolingual
+speakers of English. In Experiment 3, we collected additional data from
+a new group of British English natives. The design was closely modelled
+after Experiment 1, except that after providing their response in each
+trial, participants reported whether they had previous knowledge of the
+presented word. In the final section of this paper, we present analyses
+on the joint data sets of Experiments 1 to 3.
 
@@
 phonological similarity between the presented (Catalan or Spanish) word
-and their correct English translation (\emph{Cognateness}), and the
+and their correct English translation (\emph{Similarity}), and the
 presented word's number of cross-language phonological neighbours
@@
 
-We calculated \emph{Cognatenesss}, our main predictor of interest, by
+We calculated \emph{Similarity}, our main predictor of interest, by
 computing the Levenshtein similarity between the X-SAMPA transcriptions
@@
 phonological transcriptions. For example, the \emph{table}
-(/teɪbəl/)-\emph{mesa} (/mesa/) translation pair had a 17\% similarity,
-while the \emph{train} (\texttt{trEIn})-\emph{tren} (tɾEn) translation
-pair had a 60\% similarity. Figure 1 summarises the lexical frequency,
-phonological neighbourhood density and phonological overlap of the words
-included in the Catalan and the Spanish lists.
+(\texttt{teIb@l})-\emph{mesa} (\texttt{mesa}) translation pair had a
+17\% similarity, while the \emph{train} (\texttt{trEIn})-\emph{tren}
+(\texttt{tREn}) translation pair had a 60\% similarity.
+Table~\ref{tbl-stimuli} summarises the lexical frequency, phonological
+neighbourhood density and phonological overlap of the words included in
+the Catalan and the Spanish lists.
 
@@
 \toprule
-\multicolumn{1}{l}{} & \multicolumn{2}{c}{Frequency} & \multicolumn{2}{c}{Cognateness} & \multicolumn{2}{c}{CLPN} \\ 
+\multicolumn{1}{l}{} & \multicolumn{2}{c}{Frequency} & \multicolumn{2}{c}{Similarity} & \multicolumn{2}{c}{CLPN} \\ 
 \cmidrule(lr){2-3} \cmidrule(lr){4-5} \cmidrule(lr){6-7}
@@
 considered as incorrect. \emph{Other} responses were excluded from data
-analysis.
-
-Trials in which participants took longer than 10 seconds to respond were
-also excluded. Participants contributed a total of 9152 valid trials
-(5694 in Catalan, 3458 in Spanish). The task took approximately 15
-minutes to be completed.
+analysis. Trials in which participants took longer than 10 seconds to
+respond were also excluded. Participants contributed a total of 9152
+valid trials (5,694 in Catalan, 3,458 in Spanish). The task took
+approximately 15 minutes to be completed.
 
@@
 included as fixed effects the intercept, the main effects of
-\emph{Frequency}, \emph{Cognateness}, and \emph{CLPN}, and the two-way
-interaction between \emph{Cognateness} and \emph{CLPN}. We also included
+\emph{Frequency}, \emph{Similarity}, and \emph{CLPN}, and the two-way
+interaction between \emph{Similarity} and \emph{CLPN}. We also included
 participant-level random intercepts and slopes for the the main effects
@@
 &\textbf{Likelihood}  \\
-y_{i} \sim & Bernoulli(p_{i}) \\ \\
+y_{i} \sim & \text{Bernoulli}(p_{i}) \\ \\
 &\textbf{Parameters}  \\
@@
 indicates the probability that the true value of the coefficient is
-equivalent to zero.
-
-All analyses were performed in R environment
+equivalent to zero. All analyses were performed in R environment
 (\citeproc{ref-rcoreteam2013language}{R Core Team, 2013}). We used the
@@
 participants added comments to the experimenter (e.g., \texttt{unsure},
-(\emph{n} = 13). In addition, we excluded data from participants that
+\emph{n} = 13). In addition, we excluded data from participants that
 self-rated their oral and/or written skills in Catalan and Spanish, or
@@
 indicated by the size of the regression coefficient of the two-way
-interaction between \emph{Cognateness} and \emph{CLPN} (\(\beta\) =
+interaction between \emph{Similarity} and \emph{CLPN} (\(\beta\) =
 -0.653, 95\% CrI = {[}-0.973, -0.313{]}, \emph{p}(ROPE) = 0). As
 anticipated, participants' performance benefited from an increase in
-\emph{Cognateness} (\(\beta\) = 7.133, 95\% CrI = {[}6.606, 7.735{]},
+\emph{Similarity} (\(\beta\) = 7.133, 95\% CrI = {[}6.606, 7.735{]},
 \emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
@@
 average predictions of the model for words with different values of
-\emph{Cognateness} and \emph{CLPN}. Figure~\ref{fig-coefs} shows a
+\emph{Similarity} and \emph{CLPN}. Figure~\ref{fig-coefs} shows a
 graphic summary of the posterior distribution of the regression
@@
 extracted for different values of \emph{CLPN} (0, 2, 4, 8, 12) and
-\emph{Cognateness} (1-100). Predictions are plotted separately for
+\emph{Similarity} (1-100). Predictions are plotted separately for
 English participants translating Catalan words, and for English
@@
 
-In the present experiment, we investigated the extent to which
-cognateness (phonological similarity between translation equivalents) is
-sufficient for successful word translation, in the absence of semantic
-knowledge. We tested two groups of monolingual British English natives
-in a translation task that involved words in Catalan or Spanish, two
+In the present experiment, we investigated the extent to which the
+phonological similarity between translation equivalents is sufficient
+for successful word translation, in the absence of semantic knowledge.
+We tested two groups of monolingual British English natives in a
+translation task that involved words in Catalan or Spanish, two
 languages participants reported having no prior familiarity with.
-Participants benefited strongly from cognateness, when the correct
-translation of the presented words in Catalan or Spanish had few English
-phonological neighbours with higher lexical frequency. This suggests
-that, in the absence of distractors, even naïve participants can
-efficiently use phonological similarity to succeed in a translation
+Participants benefited strongly from phonological similarity when the
+correct translation of the presented words in Catalan or Spanish had few
+English phonological neighbours with higher lexical frequency. This
+suggests that, in the absence of distractors, even naïve participants
+can efficiently use phonological similarity to succeed in a translation
 task. Our results suggest that word-to-word connections at the
@@
 indicated by the size of the regression coefficient of the two-way
-interaction between \emph{Cognateness} and \emph{CLPN} (\(\beta\) =
+interaction between \emph{Similarity} and \emph{CLPN} (\(\beta\) =
 -0.409, 95\% CrI = {[}-0.895, 0.111{]}, \emph{p}(ROPE) = 0.081). As
 anticipated, participants' performance benefited from an increase in
-\emph{Cognateness} (\(\beta\) = 9.274, 95\% CrI = {[}8.473, 10.269{]},
+\emph{Similarity} (\(\beta\) = 9.274, 95\% CrI = {[}8.473, 10.269{]},
 \emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
@@
 posterior of the average predictions of the model for words with
-different values of \emph{Cognateness} and \emph{CLPN}.
+different values of \emph{Similarity} and \emph{CLPN}.
 
@@
 On each trial, participants listened to the audio stimulus by clicking
-on the play button. For comparability to the PsychoPy version,
+on the \texttt{PLAY} button. For comparability to the PsychoPy version,
 participants were only allowed to play the audio one time. Participants
 were explicitly told that they would be only allowed to listen once. The
-play button vanished after one playthrough. Participants then had to
-answer three questions based on the audio they had heard on that trial.
-These questions were presented on the same page, directly below the
-audio player. They were first asked whether or not they knew the
-presented word (multiple choice---yes/no). Regardless of their answer on
-the first question, participants were asked what they thought the
-translation of the word was in English (or their best guess), and
-instructed to type their answer in the provided text box. Finally, they
-were asked to rate how confident they were in their answer on a scale of
-0 to 7, where 7 was ``very confident'' and 0 was ``not confident''.
-There was no time limit on the response phase. All questions had to be
-answered to proceed to the next trial.
+\texttt{PLAY} button vanished after one playthrough. Participants then
+had to answer three questions based on the audio they had heard on that
+trial. These questions were presented on the same page, directly below
+the audio player. They were first asked whether or not they knew the
+presented word (multiple choice---\emph{yes}/\emph{no}). Regardless of
+their answer on the first question, participants were asked what they
+thought the translation of the word was in English (or their best
+guess), and instructed to type their answer in the provided text box.
+Finally, they were asked to rate how confident they were in their answer
+on a scale of 0 to 7, where 7 was ``very confident'' and 0 was ``not
+confident''. There was no time limit on the response phase. All
+questions had to be answered to proceed to the next trial.
 
@@
 translation. This is indicated by the size of the regression coefficient
-of the two-way interaction between \emph{Cognateness} and \emph{CLPN}
+of the two-way interaction between \emph{Similarity} and \emph{CLPN}
 (\(\beta\) = -0.665, 95\% CrI = {[}-1.068, -0.32{]}, \emph{p}(ROPE) =
 0.002). As anticipated, participants' performance benefited from an
-increase in \emph{Cognateness} (\(\beta\) = 7.585, 95\% CrI = {[}7.043,
+increase in \emph{Similarity} (\(\beta\) = 7.585, 95\% CrI = {[}7.043,
 8.24{]}, \emph{p}(ROPE) = 0), while the number of \emph{CLND} had the
@@
 
+\bibitem[\citeproctext]{ref-best1988examination}
+Best, C. T., McRoberts, G. W., \& Sithole, N. M. (1988). Examination of
+perceptual reorganization for nonnative speech contrasts: Zulu click
+discrimination by english-speaking adults and infants. \emph{Journal of
+Experimental Psychology: Human Perception and Performance},
+\emph{14}(3), 345.
+
 \bibitem[\citeproctext]{ref-broersma2021praat}
@@
 De Groot, A. M., \& Keijzer, R. (2000). What is hard to learn is easy to
-forget: {The} roles of word concreteness, cognate status, and word
+forget: The roles of word concreteness, cognate status, and word
 frequency in foreign-language vocabulary learning and forgetting.
@@
 
+\bibitem[\citeproctext]{ref-hall2002automatic}
+Hall, C. J. (2002). \emph{The automatic cognate form assumption:
+Evidence for the parasitic model of vocabulary development}.
+
 \bibitem[\citeproctext]{ref-hamburger1996phonological}
@@
 \emph{23}(7), 1634--1647.
+
+\bibitem[\citeproctext]{ref-otwinowska2019more}
+Otwinowska, A., \& Szewczyk, J. M. (2019). The more similar the better?
+Factors in learning cognates, false cognates and non-cognate words.
+\emph{International Journal of Bilingual Education and Bilingualism}.
 

### 2024-07-25 manuscript/manuscript.qmd
New version of introduction

@@
 
-Learning a new language is a non-trivial task, a large part of which is acquiring a new set of word-forms to correspond to existing semantic representations in their lexicon.
-Imagine that a Spanish listener hears English for the first time and encounters the word /\textipa{"dO:}/ (*door*).
-Their knowledge of the phonological form in their native language is of little help: /\textipa{"pweR.ta}/ and /\textipa{"dO:}/ do not share any phonemes.
-It would come as a surprise if the listener was able to translate the English word successfully without having first learnt the word.
-However, there are some elements in language that may help the learner in acquiring words in a new language.
-Many languages share some degree of similarity.
-This may be due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
-For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents (also known as cognates) [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographical similarity of the former to the equivalent in their native language.
-Second language learners have been found to learn cognates more easily than non-cognates [@hall2002automatic; @otwinowska2019more].
-This type of cross-linguistic similarity has also been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
-This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
-This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
-
-Event counting on this facilitation effect of form-similarity, listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
-One of potential source of this increased difficulty is the partial mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener’s native language.
+When some German speakers listened to the song *The Power* by SNAP!
+(*Coyote Ugly*, 2000), many of them misheard the line "I've got the power" as *Agatha Bauer!* When Michael Jackson vocalised "ma-ma-coo-sah" in his song *Wanna Be Startin' Somethin'* (*Thriller*, 1982), some Dutch speakers understood *mama appelsap* [mama applejuice].
+When Spanish speakers listen to the line "Circumvent your thick ego" from the song *Pictures* by System of a Down (*Steal this Album!*, 2002), they tend to hear *Sácame de aquí, King-Kong* [take me out of here, King-Kong].
+When Japanese speakers listen to the line "I want to hold your hand" in the homonym song by The Beatles (*Meet The Beatles*, 1964), they often mishear *Aho na hōnyōhan* [stupid urinator].
+Outrageous as these examples might sound to speakers of other languages, many readers may know a few cases in their own native language.
+This auditory illusion can feel quite real, and often inevitable after the first time it happens.
+In Japanese, this phenomenon takes the name *Soramimi* (lit.
+"empty ear"), after the *Soramimi Hour*---a section of the program *Tamori Club* hosted by comedian Morita Kazuyoshi (Tamori)---in which instances of this illusion were presented with comedy purposes.
+Soramimi is a particular case of *homophonic translation*: words or phrases in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning.
+This phenomenon extends beyond song lyrics, and may occur anytime when listening to non-native speech.
+For instance, when listening to "You know me?" (\textipa{/ju: n@U mi:}/) in English, Japanese-native listeners tend to understand *yunomi* (/\textipa{jUnomi}/), which is a type of Japanese teacup.
+Overall, the occurence of Soramimi---and more generally, homophonic translations---suggests that speech in an unfamiliar language has the potential to activate lexical representations in the native language.
+
+Although homophonic translations do not necessarily take place with the preservation of meaning---which often results in outrageous mistranslations like the ones aforementioned---homophonic translations sometimes may lead to the activation of the intended meaning (i.e., a correct translation).
 For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
 This listener may encounter the word /\textipa{"pOKt}/ (*porte*), translation of door in French.
-When comparing the phonemic transcription to the transcription of the Spanish word /\textipa{"pweR.ta}/ (*puerta*), one may be led to infer that these two words do not have much overlap.
-But one thing to note is that the voiced uvular fricative consonant /\textipa{K}/ and the open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /\textipa{R}/ and /o/.
-A perceptual assimilation model has been proposed to account for how speakers may integrate non-native phonemes into their native phonemic categories based on goodness-of-fit [@best2001discrimination; @best1988examination].
-Although phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish monolingual may still be able to activate the phonological representation of /\textipa{"pweR.ta}/ (*puerta*) from hearing the word-form /\textipa{"pOKt}/ (*porte*), ultimately allowing them to access the right semantic representation.
-
-:::
-
-::: {.content-visible when-format="docx"}
-
-Learning a new language is a non-trivial task, a large part of which is acquiring a new set of word-forms to correspond to existing semantic representations in their lexicon.
-Imagine that a Spanish listener hears English for the first time and encounters the word /dɔː/ (*door*).
-Their knowledge of the phonological form in their native language is of little help: /pweɾta/ and /dɔː/ do not share any phonemes.
-It would come as a surprise if the listener was able to translate the English word successfully without having first learnt the word.
-However, there are some elements in language that may help the learner in acquiring words in a new language.
-Many languages share some degree of similarity.
-This may be due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
-For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents (also known as cognates) [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographical similarity of the former to the equivalent in their native language.
-Second language learners have been found to learn cognates more easily than non-cognates [@hall2002automatic; @otwinowska2019more].
-This type of cross-linguistic similarity has also been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
-This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
-This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
-
-Event counting on this facilitation effect of form-similarity, listening to non-native speech is costlier than listening to native speech, even for highly proficient bilinguals, and especially in acoustically adverse situations [@lecumberri2010non; @takata1990english].
-One of potential source of this increased difficulty is the partial mismatch between the phonology of the native and the non-native languages: some acoustic features embedded in the non-native speech signal do not overlap perfectly with any phonemic category in the listener’s native language.
-For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
-This listener may encounter the word /pɔʁt/ (*porte*), translation of door in French.
-When comparing the phonemic transcription to the transcription of the Spanish word /pweɾta/ (*puerta*), one may be led to infer that these two words do not have much overlap.
-But one thing to note is that the voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/.
-A perceptual assimilation model has been proposed to account for how speakers may integrate non-native phonemes into their native phonemic categories based on goodness-of-fit [@best2001discrimination; @best1988examination].
-Although phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish monolingual may still be able to activate the phonological representation of /pweɾta/ (*puerta*) from hearing the word-form /pɔʁt/ (*porte*), ultimately allowing them to access the right semantic representation.
-
-:::
-
-While phonological similarity across translation equivalents can help with speech recognition, phonological similarity between words with different meanings can hinder the recognition process.
-In word retrieval, the presence of phonological neighbours (word-forms that differ from each other in one phoneme) may interfere with the activation and selection of the appropriate lexical representation in the native language.
-Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
-This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
-The inhibitory effect of phonological neighbourhoods has also been reported across languages.
-For instance, @weber2004lexical observed that bilinguals displayed lexical competition effects from words in their native language when performing a word-referent matching task in their L2.
-The assimilation of non-native phonemes into native categories [@best2001discrimination; @best1988examination] was used to explain how lexical competitors in the native language can be activated even when the target words contain non-native phonemes.
-In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
-For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
-It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect cross-linguistic phonological similarity has on word recognition in a foreign language.
-This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
-
-In the present study, we investigate the extent to which monolinguals can exploit phonological similarity between translation equivalents when listening to words in an unfamiliar language, and how it interacts with competition from phonological neighbours in the native language.
+When comparing the phonemic transcription of the French to that of its Spanish translation /\textipa{"pweR.ta}/ (*puerta*), one may be led to infer that these two words do not have much overlap.
+But one thing to note is that the voiced uvular fricative consonant /\textipa{K}/ and the open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /\textipa{R}/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
+Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /\textipa{"pweR.ta}/ (*puerta*) from hearing the word-form /\textipa{"pOKt}/ (*porte*), ultimately allowing them to access the correct semantic representation.
+Given enough phonological similarity between the non-native word and its translation in the native language, a naïve listener may succeed at word comprehension when listening to an unfamiliar language.
+Such phonological similarity between translation equivalents---known as *cognateness*---is common across many languages, and is often due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
+For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographic similarity of the former to the equivalent in their native language.
+
+While Soramimi, and more generally homophonic translation, are popular and frequent phenomena, the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
+For instance, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi hour.
+These instances consisted of homophonic translations of English song lyrics to Japanese words and phrases.
+The vast majority of the reported instances were phrases (96%) and a few, to single words (4%).
+The phonetic features of the presented English lyrics were preserved with relatively high variability, with some Japanese resulting words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
+When analysing the few instances of homophonic translations of single words, the author identified three phonological processes that might explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /\textipa{kraI}/ to *kurai*, /\textipa{kuRai}/ [dark]), deletion (e.g., *go*, /\textipa{goU}/ to *go*, /go/ [go], and alternation (e.g., *low*, /\textipa{loU}/ to *rou*, /\textipa{Roo}/ [wax]).
+This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
+Native prosody also constrains the occurence of homophonic translations.
+For instance, @kentner2015rhythmic presented in the auditory modality French-native and German-native participants with English songs with lyrics.
+Homophonic translations followed language-specific segmentation strategies.
+Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries *before* stressed syllables (following the more frequent trochaic patter in German).
+On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries *after* stressed syllables, following the more frequent iambic pattern in French.
+
+While the previous studies have address the phonological aspects of homophonic translation, little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
+In the present work we investigated the interplay between phonological similarity and phonological neighbourhood density to examine how listeners of an unfamiliar language activate lexical representations in their native language.
+In particular, we studied how the presence of phonological neighbours in the native language lexicon affects the extent to which listeners are able to activate the correct translation of an auditorily presented word in an unfamiliar language.
 We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
@@
 Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language.
-The effect is predicted to be especially strong if any of the competitors is more lexically frequent than the target translation.
-We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
+Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
+This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
+Since the interference effects of phonological neighbourhood density has also been reported across languages [@weber2004lexical], we hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
+In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
+For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
+It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect cross-linguistic phonological similarity has on word recognition in a foreign language.
+This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
 For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
 To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme apart from the presented word.
-If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.We conducted a series of three experiments to test these predictions.
+If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.
+We conducted a series of three experiments to test these predictions.
 
@@
 In the final section of this paper, we present analyses on the joint data sets of Experiments 1 to 3.
+
+:::
+
+::: {.content-visible when-format="docx"}
+
+When some German speakers listened to the song *The Power* by SNAP!
+(*Coyote Ugly*, 2000), many of them misheard the line "I've got the power" as *Agatha Bauer!* When Michael Jackson vocalised "ma-ma-coo-sah" in his song *Wanna Be Startin' Somethin'* (*Thriller*, 1982), some Dutch speakers understood *mama appelsap* [mama applejuice].
+When Spanish speakers listen to the line "Circumvent your thick ego" from the song *Pictures* by System of a Down (*Steal this Album!*, 2002), they tend to hear *Sácame de aquí, King-Kong* [take me out of here, King-Kong].
+When Japanese speakers listen to the line "I want to hold your hand" in the homonym song by The Beatles (*Meet The Beatles*, 1964), they often mishear *Aho na hōnyōhan* [stupid urinator].
+Outrageous as these examples might sound to speakers of other languages, many readers may know a few cases in their own native language.
+This auditory illusion can feel quite real, and often inevitable after the first time it happens.
+In Japanese, this phenomenon takes the name *Soramimi* (lit.
+"empty ear"), after the *Soramimi Hour*---a section of the program *Tamori Club* hosted by comedian Morita Kazuyoshi (Tamori)---in which instances of this illusion were presented with comedy purposes.
+Soramimi is a particular case of *homophonic translation*: words or phrases in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning.
+This phenomenon extends beyond song lyrics, and may occur anytime when listening to non-native speech.
+For instance, when listening to "You know me?" (/ju: nəʊ mi:/) in English, Japanese-native listeners tend to understand *yunomi* (/jʊnomi/), which is a type of Japanese teacup.
+Overall, the occurence of Soramimi---and more generally, homophonic translations---suggests that speech in an unfamiliar language has the potential to activate lexical representations in the native language.
+
+Although homophonic translations do not necessarily take place with the preservation of meaning---which often results in outrageous mistranslations like the ones aforementioned---homophonic translations sometimes may lead to the activation of the intended meaning (i.e., a correct translation).
+For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
+This listener may encounter the word /pɔKt/ (*porte*), translation of door in French.
+When comparing the phonemic transcription of the French to that of its Spanish translation /pweɾta/ (*puerta*), one may be led to infer that these two words do not have much overlap.
+But one thing to note is that the voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /ʁ/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
+Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /pweɾta}/ (*puerta*) from hearing the word-form /pɔʁt/ (*porte*), ultimately allowing them to access the correct semantic representation.
+Given enough phonological similarity between the non-native word and its translation in the native language, a naïve listener may succeed at word comprehension when listening to an unfamiliar language.
+Such phonological similarity between translation equivalents---known as *cognateness*---is common across many languages, and is often due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
+For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographic similarity of the former to the equivalent in their native language.
+
+While Soramimi, and more generally homophonic translation, are popular and frequent phenomena, the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
+For instance, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi hour.
+These instances consisted of homophonic translations of English song lyrics to Japanese words and phrases.
+The vast majority of the reported instances were phrases (96%) and a few, to single words (4%).
+The phonetic features of the presented English lyrics were preserved with relatively high variability, with some Japanese resulting words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
+When analysing the few instances of homophonic translations of single words, the author identified three phonological processes that might explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /krʌɪ/ to *kurai*, /kuɾai/ [dark]), deletion (e.g., *go*, /goʊ/ to *go*, /go/ [go], and alternation (e.g., *low*, /loʊ/ to *rou*, /ɾoo/ [wax]).
+This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
+Native prosody also constrains the occurence of homophonic translations.
+For instance, @kentner2015rhythmic presented in the auditory modality French-native and German-native participants with English songs with lyrics.
+Homophonic translations followed language-specific segmentation strategies.
+Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries *before* stressed syllables (following the more frequent trochaic patter in German).
+On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries *after* stressed syllables, following the more frequent iambic pattern in French.
+
+While the previous studies have address the phonological aspects of homophonic translation, little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
+In the present work we investigated the interplay between phonological similarity and phonological neighbourhood density to examine how listeners of an unfamiliar language activate lexical representations in their native language.
+In particular, we studied how the presence of phonological neighbours in the native language lexicon affects the extent to which listeners are able to activate the correct translation of an auditorily presented word in an unfamiliar language.
+We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
+We explored listeners’ reliance on phonological similarity to succeed in the task by manipulating the amount of phonological similarity between the presented words and their target translations.
+Since participants were unfamiliar with the presented language, they could only use phonological similarity between the presented language and their native language to successfully translate the words.
+We therefore predicted that participants’ performance should increase when the translation pairs are phonologically more similar.
+We also predicted that there is a minimum threshold of phonological similarity to be sufficient for translation.
+
+We further investigated the effect of phonological competitors.
+Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language.
+Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
+This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
+Since the interference effects of phonological neighbourhood density has also been reported across languages [@weber2004lexical], we hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
+In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
+For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
+It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect cross-linguistic phonological similarity has on word recognition in a foreign language.
+This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
+For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
+To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme apart from the presented word.
+If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.
+We conducted a series of three experiments to test these predictions.
+
+In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
+One group was presented with Catalan words, the other with Spanish words.
+We examined the extent to which participants were able to use the phonological similarity between the presented word (in Catalan or Spanish) and its target translation to provide accurate responses.
+In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
+Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant’s performance.
+One unexpected finding was that participants in Experiments 1 and 2 were surprisingly good at translating a subset of words which had low phonological similarity with their correct translation.
+We were concerned that this may be caused by some prior knowledge of specific words by our participants, as some Spanish words are commonly seen in media or product labels, making them familiar even to monolingual speakers of English.
+In Experiment 3, we collected additional data from a new group of British English natives.
+The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
+In the final section of this paper, we present analyses on the joint data sets of Experiments 1 to 3.
+
+:::
 
@@
 
+This type of cross-linguistic similarity has also been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
+This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
+This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
+
 # References

### 2024-07-25 manuscript/manuscript.tex
New version of introduction

@@
 \DeclareTextFontCommand\textphonetic{\ipafont}
+\usepackage[utf8]{inputenc}
+\usepackage{CJKutf8}
 \usepackage{arxiv}
@@
 \hypersetup{
-  pdftitle={The role of cognateness in recognition of spoken words in a foreign language},
+  pdftitle={The role of cognateness in native spoken word recognition},
   pdfauthor={Gonzalo Garcia-Castro; Serene Siow; Kim Plunkett; Nuria Sebastian-Galles},
@@
 words }
-\title{The role of cognateness in recognition of spoken words in a
-foreign language}
+\title{The role of cognateness in native spoken word recognition}
 \author{\textbf{Gonzalo
@@
 \maketitle
-\begin{abstract}
-Understanding spoken words in a non-native language is easier when their
-translations into the listener's native language are phonologically
-similar. Phonological similarity plays a more important role in the
-language performance of low-proficiency bilinguals, compared to that of
-high-proficiency bilinguals. This suggests that low-proficiency
-bilinguals may points to the latter as relying more strongly on
-cross-language word-to-word links (lexical route), while
-high-proficiency bilinguals rely more on word-concept associations
-(conceptual route) than on cross-language word-to-word links (lexical
-route). Disentangling bilinguals' reliance on the conceptual and lexical
-and conceptual routes during translation is challenging, given that both
-sources of information are present during spoken word comprehension to
-some extent in both high- and low-proficiency bilinguals. In this study,
-we tested English and Spanish native speakers in a translation
-elicitation task, in which they listened to words in an unfamiliar
-language (Spanish or Catalan for English natives, and Catalan for
-Spanish natives), and then had to translate them to their native
-language. Given their lack of previous knowledge on the unfamiliar
-tested language, phonological similarity between the presented words and
-their correct translations was the only cue available for participants
-to succeed in the task. This allowed us to explore the informativeness
-of the lexical route for word translation, in the absence of information
-from the conceptual route. Our results indicate that participants
-benefited from the phonological similarity between translation
-equivalents: the more similar the word they listened to and its
-translational equivalent, the higher the probability of a correct
-translation. These results point to cross-language word-to-word links as
-providing rich information during word translation.
-\end{abstract}
 {\bfseries \emph Keywords}
@@
 
-Learning a new language is a non-trivial task, a large part of which is
-acquiring a new set of word-forms to correspond to existing semantic
-representations in their lexicon. Imagine that a Spanish listener hears
-English for the first time and encounters the word /\textipa{"dO:}/
-(\emph{door}). Their knowledge of the phonological form in their native
-language is of little help: /\textipa{"pweR.ta}/ and /\textipa{"dO:}/ do
-not share any phonemes. It would come as a surprise if the listener was
-able to translate the English word successfully without having first
-learnt the word. However, there are some elements in language that may
-help the learner in acquiring words in a new language. Many languages
-share some degree of similarity. This may be due to typological
-closeness and/or socio-historical events involving the speakers of these
-languages (e.g., migration, social contact). For example, Romance
-languages such as Spanish and Catalan share many form-similar
-translation equivalents (also known as cognates)
+When some German speakers listened to the song \emph{The Power} by SNAP!
+(\emph{Coyote Ugly}, 2000), many of them misheard the line ``I've got
+the power'' as \emph{Agatha Bauer!} When Michael Jackson vocalised
+``ma-ma-coo-sah'' in his song \emph{Wanna Be Startin' Somethin'}
+(\emph{Thriller}, 1982), some Dutch speakers understood \emph{mama
+appelsap} {[}mama applejuice{]}. When Spanish speakers listen to the
+line ``Circumvent your thick ego'' from the song \emph{Pictures} by
+System of a Down (\emph{Steal this Album!}, 2002), they tend to hear
+\emph{Sácame de aquí, King-Kong} {[}take me out of here, King-Kong{]}.
+When Japanese speakers listen to the line ``I want to hold your hand''
+in the homonym song by The Beatles (\emph{Meet The Beatles}, 1964), they
+often mishear \emph{Aho na hōnyōhan} {[}stupid urinator{]}. Outrageous
+as these examples might sound to speakers of other languages, many
+readers may know a few cases in their own native language. This auditory
+illusion can feel quite real, and often inevitable after the first time
+it happens. In Japanese, this phenomenon takes the name \emph{Soramimi}
+(lit. ``empty ear''), after the \emph{Soramimi Hour}---a section of the
+program \emph{Tamori Club} hosted by comedian Morita Kazuyoshi
+(Tamori)---in which instances of this illusion were presented with
+comedy purposes. Soramimi is a particular case of \emph{homophonic
+translation}: words or phrases in one language are translated into
+similar-sounding words and phrases in a different language, without
+necessarily preserving the meaning. This phenomenon extends beyond song
+lyrics, and may occur anytime when listening to non-native speech. For
+instance, when listening to ``You know me?'' (\textipa{/ju: n@U mi:}/)
+in English, Japanese-native listeners tend to understand \emph{yunomi}
+(/\textipa{jUnomi}/), which is a type of Japanese teacup. Overall, the
+occurence of Soramimi---and more generally, homophonic
+translations---suggests that speech in an unfamiliar language has the
+potential to activate lexical representations in the native language.
+
+Although homophonic translations do not necessarily take place with the
+preservation of meaning---which often results in outrageous
+mistranslations like the ones aforementioned---homophonic translations
+sometimes may lead to the activation of the intended meaning (i.e., a
+correct translation). For example, imagine a Spanish native with no
+previous familiarity with any other language listening to French for the
+first time. This listener may encounter the word /\textipa{"pOKt}/
+(\emph{porte}), translation of door in French. When comparing the
+phonemic transcription of the French to that of its Spanish translation
+/\textipa{"pweR.ta}/ (\emph{puerta}), one may be led to infer that these
+two words do not have much overlap. But one thing to note is that the
+voiced uvular fricative consonant /\textipa{K}/ and the open-mid back
+rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic
+categories, can be perceived as allophonic variations the native
+(Spanish) phonemes /\textipa{R}/ and /o/, as described by assimilation
+models of phoneme perception (\citeproc{ref-best1988examination}{Best et
+al., 1988}, \citeproc{ref-best2001discrimination}{2001}). Although this
+phonetic mismatch between the non-native word and its translation can
+have a noticeable toll on comprehension Cutler et al.
+(\citeproc{ref-cutler2004patterns}{2004}), the Spanish listener may
+still be able to activate the phonological representation of
+/\textipa{"pweR.ta}/ (\emph{puerta}) from hearing the word-form
+/\textipa{"pOKt}/ (\emph{porte}), ultimately allowing them to access the
+correct semantic representation. Given enough phonological similarity
+between the non-native word and its translation in the native language,
+a naïve listener may succeed at word comprehension when listening to an
+unfamiliar language. Such phonological similarity between translation
+equivalents---known as \emph{cognateness}---is common across many
+languages, and is often due to typological closeness and/or
+socio-historical events involving the speakers of these languages (e.g.,
+migration, social contact). For example, Romance languages such as
+Spanish and Catalan share many form-similar translation equivalents
 (\citeproc{ref-schepens2012distributions}{Schepens et al., 2012}), as in
@@
 translating Catalan \emph{porta} than English \emph{door}, due to the
-phonological and orthographical similarity of the former to the
-equivalent in their native language. Second language learners have been
-found to learn cognates more easily than non-cognates
-(\citeproc{ref-hall2002automatic}{Hall, 2002};
-\citeproc{ref-otwinowska2019more}{Otwinowska \& Szewczyk, 2019}). This
-type of cross-linguistic similarity has also been extremely informative
-for the study of bilingual lexical processing: words that share high
-form-similarity with their translations are processed faster and more
-accurately than words sharing lower phonological similarity. This
-phenomenon has provided strong evidence in favour of the language-non
-selective hypothesis of bilingual lexical access, which states that
-bilinguals activate lexical representations in both languages, even
-during monolingual situations. This facilitation effect has been
-reported for comprehension (\citeproc{ref-dijkstra2010cross}{Dijkstra et
-al., 2010}; \citeproc{ref-midgley2011effects}{Midgley et al., 2011};
-\citeproc{ref-thierry2007brain}{Thierry \& Wu, 2007}), production
-(\citeproc{ref-costa2000cognate}{Costa et al., 2000}), learning
-(\citeproc{ref-de2000hard}{De Groot \& Keijzer, 2000};
-\citeproc{ref-elias2022crosslanguage}{Elias \& Degani, 2022};
-\citeproc{ref-lotto1998effects}{Lotto \& De Groot, 1998};
-\citeproc{ref-valente2018does}{Valente et al., 2018}), and translation
-(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}).
-
-Event counting on this facilitation effect of form-similarity, listening
-to non-native speech is costlier than listening to native speech, even
-for highly proficient bilinguals, and especially in acoustically adverse
-situations (\citeproc{ref-lecumberri2010non}{Lecumberri et al., 2010};
-\citeproc{ref-takata1990english}{Takata \& Nábělek, 1990}). One of
-potential source of this increased difficulty is the partial mismatch
-between the phonology of the native and the non-native languages: some
-acoustic features embedded in the non-native speech signal do not
-overlap perfectly with any phonemic category in the listener's native
-language. For example, imagine a Spanish native with no previous
-familiarity with any other language listening to French for the first
-time. This listener may encounter the word /\textipa{"pOKt}/
-(\emph{porte}), translation of door in French. When comparing the
-phonemic transcription to the transcription of the Spanish word
-/\textipa{"pweR.ta}/ (\emph{puerta}), one may be led to infer that these
-two words do not have much overlap. But one thing to note is that the
-voiced uvular fricative consonant /\textipa{K}/ and the open-mid back
-rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic
-categories, can be perceived as allophonic variations the native
-(Spanish) phonemes /\textipa{R}/ and /o/. A perceptual assimilation
-model has been proposed to account for how speakers may integrate
-non-native phonemes into their native phonemic categories based on
-goodness-of-fit (\citeproc{ref-best1988examination}{Best et al., 1988},
-\citeproc{ref-best2001discrimination}{2001}). Although phonetic mismatch
-between the non-native word and its translation can have a noticeable
-toll on comprehension Cutler et al.
-(\citeproc{ref-cutler2004patterns}{2004}), the Spanish monolingual may
-still be able to activate the phonological representation of
-/\textipa{"pweR.ta}/ (\emph{puerta}) from hearing the word-form
-/\textipa{"pOKt}/ (\emph{porte}), ultimately allowing them to access the
-right semantic representation.
-
-While phonological similarity across translation equivalents can help
-with speech recognition, phonological similarity between words with
-different meanings can hinder the recognition process. In word
-retrieval, the presence of phonological neighbours (word-forms that
-differ from each other in one phoneme) may interfere with the activation
-and selection of the appropriate lexical representation in the native
-language. Words with denser phonological neighbourhoods are recognised
-more slowly and less accurately than words with sparser phonological
-neighbourhoods (\citeproc{ref-dufour2003lexical}{Dufour \& Peereman,
-2003}; \citeproc{ref-goldinger1989priming}{Goldinger et al., 1989};
-\citeproc{ref-hamburger1996phonological}{Hamburger \& Slowiaczek, 1996};
-\citeproc{ref-luce1990similarity}{Luce et al., 1990};
-\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}). This is
-especially true if such neighbours share higher phonological similarity
-with the presented word, or are lexically more frequent than the target
-translation (\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}).
-The inhibitory effect of phonological neighbourhoods has also been
-reported across languages. For instance, Weber \& Cutler
-(\citeproc{ref-weber2004lexical}{2004}) observed that bilinguals
-displayed lexical competition effects from words in their native
-language when performing a word-referent matching task in their L2. The
-assimilation of non-native phonemes into native categories
-(\citeproc{ref-best1988examination}{Best et al., 1988},
-\citeproc{ref-best2001discrimination}{2001}) was used to explain how
-lexical competitors in the native language can be activated even when
-the target words contain non-native phonemes. In the case of non-native
-speech recognition, the presence of cross-linguistic pairs which are
-phonologically similar but differ in meaning (e.g., false friends) may
-act as distractors during lexical access, obstructing the selection of
-the appropriate target translation in the listener's lexicon. For
-instance, Otwinowska \& Szewczyk
-(\citeproc{ref-otwinowska2019more}{2019}) reported that false friends
-were disadvantaged relative to non-cognates by Polish second language
-learners, in contrast to cognates which were known better. It is
-therefore important to investigate the joint effect of both cognates and
-false friends when investigating the effect cross-linguistic
-phonological similarity has on word recognition in a foreign language.
-This could shed light on available strategies and challenges associated
-with the early stages of second language acquisition.
-
-In the present study, we investigate the extent to which monolinguals
-can exploit phonological similarity between translation equivalents when
-listening to words in an unfamiliar language, and how it interacts with
-competition from phonological neighbours in the native language. We used
-a translation elicitation task in which participants listened to words
+phonological and orthographic similarity of the former to the equivalent
+in their native language.
+
+While Soramimi, and more generally homophonic translation, are popular
+and frequent phenomena, the psycholinguistic literature on the topic is
+relatively scarce, and has mostly focused on the phonological variables
+involved in the occurence of Soramimi. For instance, Otake
+(\citeproc{ref-otake2007interlingual}{2007}) analysed 194 instances of
+Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi
+hour. These instances consisted of homophonic translations of English
+song lyrics to Japanese words and phrases. The vast majority of the
+reported instances were phrases (96\%) and a few, to single words (4\%).
+The phonetic features of the presented English lyrics were preserved
+with relatively high variability, with some Japanese resulting words or
+phrases sharing high overlap with the original English strings of
+sounds, and some sharing very little overlap. When analysing the few
+instances of homophonic translations of single words, the author
+identified three phonological processes that might explain how Japanese
+listeners reconstructed the English input to generate the Japanese
+words: insertion (e.g., \emph{cry} /\textipa{kraI}/ to \emph{kurai},
+/\textipa{kuRai}/ {[}dark{]}), deletion (e.g., \emph{go},
+/\textipa{goU}/ to \emph{go}, /go/ {[}go{]}, and alternation (e.g.,
+\emph{low}, /\textipa{loU}/ to \emph{rou}, /\textipa{Roo}/ {[}wax{]}).
+This suggests that the reported homophonic translations can be
+explained, to some extent, as the result of the Japanese listeners
+accommodating the strings of English sounds to the Japanese phoneme
+inventory and Japanese phonotactics
+(\citeproc{ref-dupoux1999epenthetic}{Dupoux et al., 1999};
+\citeproc{ref-peperkamp2008perceptual}{Peperkamp et al., 2008}). Native
+prosody also constrains the occurence of homophonic translations. For
+instance, Kentner (\citeproc{ref-kentner2015rhythmic}{2015}) presented
+in the auditory modality French-native and German-native participants
+with English songs with lyrics. Homophonic translations followed
+language-specific segmentation strategies. Most homophonic translations
+reported by German-native participants resulted from inserting phrase
+and word boundaries \emph{before} stressed syllables (following the more
+frequent trochaic patter in German). On the contrary, most homophonic
+translations reported by French-native participants resulted from
+inserting phrase and word boundaries \emph{after} stressed syllables,
+following the more frequent iambic pattern in French.
+
+While the previous studies have address the phonological aspects of
+homophonic translation, little attention has been paid to the dynamics
+of lexical activation and competition that underlay this phenomenon. In
+the present work we investigated the interplay between phonological
+similarity and phonological neighbourhood density to examine how
+listeners of an unfamiliar language activate lexical representations in
+their native language. In particular, we studied how the presence of
+phonological neighbours in the native language lexicon affects the
+extent to which listeners are able to activate the correct translation
+of an auditorily presented word in an unfamiliar language. We used a
+translation elicitation task in which participants listened to words
 from an unfamiliar language, and then tried to guess their translation
@@
 presented word also shares high phonological similarity with other words
-in the native language. The effect is predicted to be especially strong
-if any of the competitors is more lexically frequent than the target
-translation. We hypothesised that the size of the facilitatory role of
-phonological similarity would be inversely proportional to the amount of
+in the native language. Words with denser phonological neighbourhoods
+are recognised more slowly and less accurately than words with sparser
+phonological neighbourhoods (\citeproc{ref-dufour2003lexical}{Dufour \&
+Peereman, 2003}; \citeproc{ref-goldinger1989priming}{Goldinger et al.,
+1989}; \citeproc{ref-hamburger1996phonological}{Hamburger \& Slowiaczek,
+1996}; \citeproc{ref-luce1990similarity}{Luce et al., 1990};
+\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}). This is
+especially true if such neighbours share higher phonological similarity
+with the presented word, or are lexically more frequent than the target
+translation (\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}).
+Since the interference effects of phonological neighbourhood density has
+also been reported across languages
+(\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}), we
+hypothesised that the size of the facilitatory role of phonological
+similarity would be inversely proportional to the amount of
 higher-frequency phonological neighbours of the presented word in the
-target language. For instance, one could expect English participants to
-incorrectly translate the Spanish word \emph{botón} as \emph{bottom}
-instead of as its correct translation \emph{button}, due to combined
-effect of the close phonological similarity between \emph{botón} and
-\emph{bottom} along with the high lexical frequency of \emph{bottom}
-relative to \emph{button}. To test this prediction, we developed a
-lexical frequency-dependent measure of cross-language phonological
-neighbourhood density, in which a neighbour is counted only if it is
-higher frequency and is one phoneme apart from the presented word. If
-the phonological neighbourhood density of the target translation affects
-participants' performance negatively, this would suggest that
-competitors in the native language affect recognition of non-native
-words during recognition of foreign speech.We conducted a series of
-three experiments to test these predictions.
+target language. In the case of non-native speech recognition, the
+presence of cross-linguistic pairs which are phonologically similar but
+differ in meaning (e.g., false friends) may act as distractors during
+lexical access, obstructing the selection of the appropriate target
+translation in the listener's lexicon. For instance, Otwinowska \&
+Szewczyk (\citeproc{ref-otwinowska2019more}{2019}) reported that false
+friends were disadvantaged relative to non-cognates by Polish second
+language learners, in contrast to cognates which were known better. It
+is therefore important to investigate the joint effect of both cognates
+and false friends when investigating the effect cross-linguistic
+phonological similarity has on word recognition in a foreign language.
+This could shed light on available strategies and challenges associated
+with the early stages of second language acquisition. For instance, one
+could expect English participants to incorrectly translate the Spanish
+word \emph{botón} as \emph{bottom} instead of as its correct translation
+\emph{button}, due to combined effect of the close phonological
+similarity between \emph{botón} and \emph{bottom} along with the high
+lexical frequency of \emph{bottom} relative to \emph{button}. To test
+this prediction, we developed a lexical frequency-dependent measure of
+cross-language phonological neighbourhood density, in which a neighbour
+is counted only if it is higher frequency and is one phoneme apart from
+the presented word. If the phonological neighbourhood density of the
+target translation affects participants' performance negatively, this
+would suggest that competitors in the native language affect recognition
+of non-native words during recognition of foreign speech. We conducted a
+series of three experiments to test these predictions.
 
@@
 
+This type of cross-linguistic similarity has also been extremely
+informative for the study of bilingual lexical processing: words that
+share high form-similarity with their translations are processed faster
+and more accurately than words sharing lower phonological similarity.
+This phenomenon has provided strong evidence in favour of the
+language-non selective hypothesis of bilingual lexical access, which
+states that bilinguals activate lexical representations in both
+languages, even during monolingual situations. This facilitation effect
+has been reported for comprehension
+(\citeproc{ref-dijkstra2010cross}{Dijkstra et al., 2010};
+\citeproc{ref-midgley2011effects}{Midgley et al., 2011};
+\citeproc{ref-thierry2007brain}{Thierry \& Wu, 2007}), production
+(\citeproc{ref-costa2000cognate}{Costa et al., 2000}), learning
+(\citeproc{ref-de2000hard}{De Groot \& Keijzer, 2000};
+\citeproc{ref-elias2022crosslanguage}{Elias \& Degani, 2022};
+\citeproc{ref-lotto1998effects}{Lotto \& De Groot, 1998};
+\citeproc{ref-valente2018does}{Valente et al., 2018}), and translation
+(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}).
+
 \section{References}\label{references}
@@
 
+\bibitem[\citeproctext]{ref-dupoux1999epenthetic}
+Dupoux, E., Kakehi, K., Hirose, Y., Pallier, C., \& Mehler, J. (1999).
+Epenthetic vowels in japanese: A perceptual illusion? \emph{Journal of
+Experimental Psychology: Human Perception and Performance},
+\emph{25}(6), 1568.
+
 \bibitem[\citeproctext]{ref-elias2022crosslanguage}
@@
 
-\bibitem[\citeproctext]{ref-hall2002automatic}
-Hall, C. J. (2002). \emph{The automatic cognate form assumption:
-Evidence for the parasitic model of vocabulary development}.
-
 \bibitem[\citeproctext]{ref-hamburger1996phonological}
@@
 \emph{3}(4), 520--525.
+
+\bibitem[\citeproctext]{ref-kentner2015rhythmic}
+Kentner, G. (2015). Rhythmic segmentation in auditory illusions-evidence
+from cross-linguistic mondegreens. \emph{ICPhS}.
 
@@
 
-\bibitem[\citeproctext]{ref-lecumberri2010non}
-Lecumberri, M. L. G., Cooke, M., \& Cutler, A. (2010). Non-native speech
-perception in adverse conditions: {A} review. \emph{Speech
-Communication}, \emph{52}(11-12), 864--886.
-
 \bibitem[\citeproctext]{ref-levenshtein1966binary}
@@
 
+\bibitem[\citeproctext]{ref-otake2007interlingual}
+Otake, T. (2007). Interlingual near homophonic words and phrases in L2
+listening: Evidence from misheard song lyrics. \emph{Proceedings of the
+16th International Congress of Phonetic Sciences (ICPhS 2007)},
+777--780.
+
 \bibitem[\citeproctext]{ref-otwinowska2019more}
@@
 
+\bibitem[\citeproctext]{ref-peperkamp2008perceptual}
+Peperkamp, S., Vendelin, I., \& Nakamura, K. (2008). On the perceptual
+origin of loanword adaptations: Experimental evidence from japanese.
+\emph{Phonology}, \emph{25}(1), 129--164.
+
 \bibitem[\citeproctext]{ref-rcoreteam2013language}
@@
 \emph{Bilingualism: Language and Cognition}, \emph{15}(1), 157--166.
-
-\bibitem[\citeproctext]{ref-takata1990english}
-Takata, Y., \& Nábělek, A. K. (1990). English consonant recognition in
-noise and in reverberation by {Japanese} and {American} listeners.
-\emph{The Journal of the Acoustical Society of America}, \emph{88}(2),
-663--666.
 

### 2024-10-08 manuscript/manuscript.qmd
Add discussion and data about knowledge filter

@@
 # import objects
-tar_config_set(store = here::here("_targets"),
-               script = here::here("_targets.R"))
+tar_config_set(
+  store = here::here("_targets"),
+  script = here::here("_targets.R")
+)
 tar_load_globals()
@@
 theme_set(theme_ggdist() +
-              theme(panel.grid.major.y = element_line(colour = "grey",
-                                                      linetype = "dotted")))
+  theme(panel.grid.major.y = element_line(
+    colour = "grey",
+    linetype = "dotted"
+  )))
 
@@
 For instance, when listening to "You know me?" (\textipa{/ju: n@U mi:}/) in English, Japanese-native listeners tend to understand *yunomi* (/\textipa{jUnomi}/), which is a type of Japanese teacup.
-Overall, the occurence of Soramimi---and more generally, homophonic translations---suggests that speech in an unfamiliar language has the potential to activate lexical representations in the native language.
-
-Although homophonic translations do not necessarily take place with the preservation of meaning---which often results in outrageous mistranslations like the ones aforementioned---homophonic translations sometimes may lead to the activation of the intended meaning (i.e., a correct translation).
+Overall, the occurrence of Soramimi---and more generally, homophonic translations---suggests that speech in an unfamiliar language has the potential to activate lexical representations in the native language.
+
+Although homophonic translations do not necessarily take place with the preservation of meaning---which often results in outrageous mistranslations like the ones aforementioned---homophonic translations sometimes may lead to correct translations.
 For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
@@
 
-While Soramimi, and more generally homophonic translation, are popular and frequent phenomena, the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
+While Soramimi (and more generally homophonic translations) are popular and frequent phenomena, the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
 For instance, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi hour.
@@
 
-While the previous studies have address the phonological aspects of homophonic translation, little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
-In the present work we investigated the interplay between phonological similarity and phonological neighbourhood density to examine how listeners of an unfamiliar language activate lexical representations in their native language.
-In particular, we studied how the presence of phonological neighbours in the native language lexicon affects the extent to which listeners are able to activate the correct translation of an auditorily presented word in an unfamiliar language.
+These studies have addressed the phonological aspects of homophonic translation, but little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
+In the present work, we investigated the interplay between phonological similarity and phonological neighbourhood density in a non-native speech perception task.
+In particular, we studied how the presence of phonological neighbours in the native language lexicon affects the extent to which listeners are able to activate the correct translation of an auditorily presented word in thier native language.
 We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as *presented words*, and the correct translation for the presented words as *target translations*.
 We explored listeners’ reliance on phonological similarity to succeed in the task by manipulating the amount of phonological similarity between the presented words and their target translations.
@@
 In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+Catalan and Spanish are two Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
 By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant’s performance.
@@
 The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
-In the final section of this paper, we present analyses on the joint data sets of Experiments 1 to 3.
+In the final section of this article, we present analyses on the joint data sets of Experiments 1 to 3.
 
@@
 participants_1 <- exp_participants |>
-    filter(group!="cat-SPA")
+  filter(group != "cat-SPA")
 
@@
 ```
-
-Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+We collected data from `r n_participants` British English-native adults living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female[1]` female).
+Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
+Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
+Participants gave informed consent before providing any data.
+The study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
-We collected data from `r n_participants` British English native participants living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
 
@@
 #| label: tbl-participants
+#| tbl-cap: "Participant details."
 tbl_data <- participants |>
   mutate(exp = case_when(
-        source=="Experiment" & group!= "cat-SPA" ~ "Experiment 1",
-        source=="Experiment" & group== "cat-SPA" ~ "Experiment 2",
-        source=="Questionnaire" ~ "Experiment 3",
+    source == "Experiment" & group != "cat-SPA" ~ "Experiment 1",
+    source == "Experiment" & group == "cat-SPA" ~ "Experiment 2",
+    source == "Questionnaire" ~ "Experiment 3",
   )) |>
-    summarise(n = n(),
-              n_excluded = sum(!valid_status=="Valid"),
+  summarise(
+    n = n(),
+    n_excluded = sum(!valid_status == "Valid"),
     across(age, lst(mean, sd, min, max)),
     l2_lst = list(l2),
-              .by = c(group, exp)) |> 
-    mutate(l2_lst = map(l2_lst, 
+    .by = c(group, exp)
+  ) |>
+  mutate(
+    l2_lst = map(
+      l2_lst,
       function(x) {
@@
         return(y)
-                        }),
-           n_excluded = paste0(n, " (", n_excluded, ")")) |> 
+      }
+    ),
+    n_excluded = paste0(n, " (", n_excluded, ")")
+  ) |>
   select(-n)
@@
   tab_spanner("Age", matches("age")) |>
-    cols_label(group = "Group",
+  cols_label(
+    group = "Group",
     exp = "Experiment",
@@
     age_min = "Range",
-               l2_lst = "L2") |> 
-    tab_style(cell_text(style = "italic"),
-              cells_column_labels()) |> 
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) |> 
-    tab_style(cell_text(align = "left"),
-              list(cells_body(), cells_column_labels())) |> 
-    tab_style(cell_text(weight = "bold",
+    l2_lst = "L2"
+  ) |>
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels()
+  ) |>
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners()
+  ) |>
+  tab_style(
+    cell_text(align = "left"),
+    list(cells_body(), cells_column_labels())
+  ) |>
+  tab_style(
+    cell_text(
+      weight = "bold",
       align = "center",
-                        style = "normal"),
-              cells_column_labels(c(l2_lst, n_excluded))) |> 
+      style = "normal"
+    ),
+    cells_column_labels(c(l2_lst, n_excluded))
+  ) |>
   tab_footnote(
     "Number of included participants (number of excluded participants.)",
-        locations = cells_column_labels(n_excluded))
+    locations = cells_column_labels(n_excluded)
+  )
 ```
@@
   mutate(across(ipa_flat_1:word_2, nchar)) |>
-    summarise(across(c(ipa_flat_1, ipa_flat_2, word_1, word_2), 
-                     lst(mean, sd, min, max)),
-              .by = group)
+  summarise(
+    across(
+      c(ipa_flat_1, ipa_flat_2, word_1, word_2),
+      lst(mean, sd, min, max)
+    ),
+    .by = group
+  )
 lengths <- split(lengths, lengths$group)
@@
   summarise(across(duration, lst(mean, sd, min, max)),
-              .by = group) 
+    .by = group
+  )
 durations <- split(durations, durations$group)
@@
 
-We arranged two lists of input words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
-Words in the Catalan list were `r round(lengths$cat_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$cat_eng$ipa_flat_1_min, 2)`-`r round(lengths$cat_eng$ipa_flat_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, Range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`.
-Words in the Spanish list were `r round(lengths$spa_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$spa_eng$ipa_flat_1_min, 2)`-`r round(lengths$spa_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations (which participants had to type), were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, Range = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`.
-
-Participants listened to one audio file in each trial, each containing a single word presented in isolation.
+We created two lists of input words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
+Words in the Catalan list were `r round(lengths$cat_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$cat_eng$ipa_flat_1_min, 2)`-`r round(lengths$cat_eng$ipa_flat_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, Range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`).
+Words in the Spanish list were `r round(lengths$spa_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$spa_eng$ipa_flat_1_min, 2)`-`r round(lengths$spa_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, Range = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`).
+
+In each trial, participants listened to one audio file, wich contained a single word.
 The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
@@
 Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, Range = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
-The average duration of the Catalan audio files was `r round(durations$spa_eng$duration_mean, 2)` (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, Range = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
+The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, Range = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
+The average duration of the Spanish audio files was `r round(durations$spa_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, Range = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
 
@@
 
-For each translation pair, we defined three predictors of interest: the lexical frequency of the correct translation (*Frequency*), the phonological similarity between the presented (Catalan or Spanish) word and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*).
-
-We included *Frequency* as a nuance predictor, under the hypothesis that---keeping other predictors constant---participants would translate higher-frequency words more accurately and faster than lower-frequency words.
+For each word in the Catalan and Spanish lists pair, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word (in Catalan or Spanish) and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
+*Frequency* was included as a nuance predictor, under the hypothesis that---keeping other predictors constant---participants would translate higher-frequency words more accurately and faster than lower-frequency words.
 Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores.
@@
 We calculated *Similarity*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
-The Levenshtein distance computes the edit distance between two character strings---in this case, phoneme transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
+The Levenshtein distance computes the edit distance between two character strings---in this case, two phonological transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
 We divided this edit distance by the number of characters included in the longest X-SAMPA transcription of the translation pair.
 This results in a proportion score, in which values closer to zero correspond to lower Levenshtein distances between phonological transcriptions (i.e., higher similarity), and values closer to 1 correspond to higher Levenshtein distances (i.e., lower similarity).
-This transformation also help account for the fact that the Levenshtein distance tends to increase with the length of the transcriptions.
+This transformation accounts for the fact that the Levenshtein distance tends to increase with the length of the transcriptions.
 For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
@@
 #| label: tbl-stimuli
-#| tbl-cap: "Stimuli properties."
+#| tbl-cap: "Stimuli details."
 stimuli |>
-    summarise(across(c(freq_zipf_2, lv, neigh_n_h),
+  summarise(
+    across(c(freq_zipf_2, lv, neigh_n_h),
       lst(mean, sd, min, max),
-                     na.rm = TRUE),
-              .by = c(group)) |> 
+      na.rm = TRUE
+    ),
+    .by = c(group)
+  ) |>
   gt(rowname_col = "group") |>
@@
   fmt_number(is.numeric) |>
-    fmt_integer(c(neigh_n_h_min,
-                  neigh_n_h_max)) |> 
+  fmt_integer(c(
+    neigh_n_h_min,
+    neigh_n_h_max
+  )) |>
   cols_merge_uncert(freq_zipf_2_mean, freq_zipf_2_sd) |>
@@
   cols_merge_range(neigh_n_h_min, neigh_n_h_max) |>
-    cols_label(freq_zipf_2_mean = "Mean ± SD",
+  cols_label(
+    freq_zipf_2_mean = "Mean ± SD",
     freq_zipf_2_sd = "SD",
@@
     neigh_n_h_sd = "SD",
-               neigh_n_h_min = "Range") |> 
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) |> 
-    tab_style(cell_text(style = "italic"),
-              cells_column_labels()) |> 
-    grand_summary_rows(columns = matches("mean|sd"),
-                       fns = list(label = md("**Mean**"), 
+    neigh_n_h_min = "Range"
+  ) |>
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners()
+  ) |>
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels()
+  ) |>
+  grand_summary_rows(
+    columns = matches("mean|sd"),
+    fns = list(
+      label = md("**Mean**"),
       id = "totals",
-                                  fn = "mean"),
-                       fmt = ~fmt_number(., n_sigfig = 3, suffixing = TRUE))
+      fn = "mean"
+    ),
+    fmt = ~ fmt_number(., n_sigfig = 3, suffixing = TRUE)
+  )
 ```
@@
 
-![Schematic representation of a trial in the experimental task](_assets/img/design.png)
+![Schematic representation of a trial in the experimental task.
+The trial stars with the presentation of a fixation point in the centre of the screen (yellow dot).
+After 1,000 ms, the auditory word was presented while the fixation point remained on screen.
+After the offset of the audio, the fixation point disappeared and a visual prompt (`>`) was presented.
+Participants then wrote their and clicked `RETURN`, making the end of the trial.](_assets/img/design.png)
 
@@
 
-After data collection, participants answers were manually coded into the following categories: *Correct*, *Typo*, *Wrong*, *False friend*, *Other*.
-Responses were coded as *Correct* if the provided string of characters was identical to the orthographic form of the correct translation.
-Trials were coded as *Typo* if the participant provided a string of characters only one edit distance (addition, deletion, or substitution) apart from the orthographic form of the correct translation (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct English word.
-Responses were coded as *False friend* if the participant provided a phonologically similar incorrect translation.
-Responses were coded as *Other* (see Data analysis section for more details).
+After data collection, participants' answers were manually coded into the following categories: *Correct*, *Typo*, *Wrong*, *False friend*, *Other*.
+A response was coded as *Correct* if the provided string of characters was identical to the orthographic form of the correct translation.
+A response was coded as *Typo* if the participant provided a string of characters only one edit distance (addition, deletion, or substitution) apart from the orthographic form of the correct translation (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct English word.
+A response was coded as *False friend* if the participant provided a phonologically similar incorrect translation.
+Responses not meeting the criteria for previous categories were labelled as *Wrong* or *Other* (see Data analysis section for more details).
 Both *Correct* and *Typo* responses were considered as correct, while *Wrong* and *False friend* responses were considered as incorrect.
@@
 $$
-:::
-
-To test the practical relevance of each predictor we followed [@kruschke2018bayesian].
+::: {eq-1}
+
+To test the practical relevance of each predictor we followed @kruschke2018bayesian.
 We first specified a region of practical equivalence (*ROPE*) around zero ([-0.1, +0.1], in the logit scale).
@@
 dataset <- exp_responses |>
-    filter(group!="cat-SPA") |> 
+  filter(group != "cat-SPA") |>
   group_split(group) |>
@@
 names(dataset) <- janitor::make_clean_names(names(dataset))
-r_validity <- map(dataset, 
-                  \(x) janitor::clean_names(table(x$response_type)))
+r_validity <- map(
+  dataset,
+  \(x) janitor::clean_names(table(x$response_type))
+)
 n_participants <- map(dataset, \(x) n_distinct(x$participant_id))
@@
 excluded_impairment <- participants_1 |>
-    filter(valid_status=="Language impairment") |> 
+  filter(valid_status == "Language impairment") |>
   nrow()
@@
 excluded_ntrials <- participants_1 |>
-    filter(valid_status=="Insufficient trials") |> 
+  filter(valid_status == "Insufficient trials") |>
   nrow()
@@
 dataset_clean <- exp_responses |>
-    filter(group!="cat-SPA") |> 
+  filter(group != "cat-SPA") |>
   dplyr::filter(valid_participant, valid_response) |>
@@
   mutate(len = nchar(response)) |>
-    summarise(across(len,
-                     lst(mean, median, sd, min, max)),
-              .by = c(group)) 
+  summarise(
+    across(
+      len,
+      lst(mean, median, sd, min, max)
+    ),
+    .by = c(group)
+  )
 lengths <- split(lengths, lengths$group)
@@
 
-```{r tbl-results}
-#| label: tbl-results
-tbl_data <- list("Experiment 1" = dataset_1,
-                 "Experiment 2" = dataset_2,
-                 "Experiment 3" = dataset_3) |> 
-    bind_rows(.id = "exp") |> 
-    count(exp, group, response_type) |> 
-    pivot_wider(names_from = response_type,
-                values_from = n,
-                names_repair = janitor::make_clean_names) |>
-    relocate(exp, group, correct, typo, wrong, false_friend) |> 
-    mutate(n_correct = rowSums(cbind(correct, typo)),
-           n_incorrect = rowSums(cbind(wrong, false_friend)),
-           n_total = rowSums(cbind(n_correct, n_incorrect)),
-           across(correct:false_friend, 
-                  \(x) x / n_total,
-                  .names = "{col}_prop")) |> 
-    select(exp, group, correct, typo, wrong, false_friend,
-           matches("prop")) |> 
-    relocate(correct, correct_prop,
-             typo, typo_prop,
-             wrong, wrong_prop,
-             false_friend, false_friend_prop)
-
-tbl_data |> 
-    gt(rowname_col = "group", groupname_col = "exp") |> 
-    fmt_missing(everything(), everything(), "--") |> 
-    tab_spanner("Correct responses", 
-                matches("correct|typo")) |> 
-    tab_spanner("Incorrect responses", 
-                matches("wrong|false")) |> 
-    fmt_integer(c(correct, typo, wrong, false_friend)) |> 
-    fmt_number(is.double, pattern = "({x})", scale_by = 100) |> 
-    cols_label(correct = "Correct",
-               typo = "Typo",
-               wrong = "Wrong",
-               false_friend = "False friend",
-               correct_prop = "(%)",
-               typo_prop = "(%)",
-               wrong_prop = "(%)",
-               false_friend_prop = "(%)") |> 
-    tab_style(cell_text(weight = "bold"), 
-              cells_column_spanners()) |> 
-    tab_style(cell_text(style = "italic"), 
-              cells_column_labels()) |> 
-    summary_rows(columns = is.integer,
-                 fns = list(label = md("*Sum*"),
-                            id = "totals",
-                            fn = "sum"),
-                 fmt = ~fmt_integer(.)) |>
-    summary_rows(columns = matches("prop"),
-                 fns = list(label = md("*Mean*"), 
-                            id = "means",
-                            fn = "mean"),
-                 fmt = ~fmt_number(., scale_by = 100)) |> 
-    tab_style(cell_text(align = "center"),
-              cells_column_labels())
-
-```
-
 ```{r coefs-1}
 c1 <- gather_draws(exp_1_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
-    summarise(across(.value, lst(.median = median, 
+  summarise(across(.value, lst(
+    .median = median,
     .lower = \(x) hdi(x)[1],
     .upper = \(x) hdi(x)[2],
-                                 .rope = \(x) mean(between(x, -0.1, 0.1))),
-                     .names = "{fn}"))
+    .rope = \(x) mean(between(x, -0.1, 0.1))
+  ),
+  .names = "{fn}"
+  ))
 c1 <- split(c1, c1$.variable)
@@
 #'
-prop_adj <- function(x, n){
-    e <- (x+2)/(n+4)
+prop_adj <- function(x, n) {
+  e <- (x + 2) / (n + 4)
   return(e)
@@
 prop_adj_se <- function(x, n) {
-    e <- (x+2)/(n+4)
-    se <- sqrt(e*(1-e)/(n+4))
+  e <- (x + 2) / (n + 4)
+  se <- sqrt(e * (1 - e) / (n + 4))
   return(se)
@@
 prop_adj_ci <- function(x, n, .width = 0.95) {
-    e <- (x+2)/(n+4)
-    se <- sqrt(e*(1-e)/(n+4))
-    ci <-  e + qnorm(c((1-.width)/2, (1-(1-.width)/2)))*se
-    ci[1] <- ifelse(ci[1]<0, 0, ci[1]) # truncate at 0
-    ci[2] <- ifelse(ci[2]>1, 1, ci[2]) # truncate at 1
+  e <- (x + 2) / (n + 4)
+  se <- sqrt(e * (1 - e) / (n + 4))
+  ci <- e + qnorm(c((1 - .width) / 2, (1 - (1 - .width) / 2))) * se
+  ci[1] <- ifelse(ci[1] < 0, 0, ci[1]) # truncate at 0
+  ci[2] <- ifelse(ci[2] > 1, 1, ci[2]) # truncate at 1
   return(ci)
@@
 
-tbl_data <- list("Experiment 1" = dataset_1,
+tbl_data <- list(
+  "Experiment 1" = dataset_1,
   "Experiment 2" = dataset_2,
-                 "Experiment 3" = dataset_3) |> 
+  "Experiment 3" = dataset_3
+) |>
   bind_rows(.id = "exp") |>
   add_count(exp, group, participant_id, name = "trials") |>
-    summarise(n_trials = n(),
+  summarise(
+    n_trials = n(),
     correct = sum(correct),
-              .by = c(group, exp, participant_id)) |> 
+    .by = c(group, exp, participant_id)
+  ) |>
   mutate(correct = prop_adj(correct, n_trials)) |>
@@
     n_participants = n_distinct(participant_id),
-              .by = c(group, exp)) |> 
+    .by = c(group, exp)
+  ) |>
   relocate(n_participants, matches("correct"))
@@
 tbl_data |>
-    gt(rowname_col = "group",
-       groupname_col = "exp") |> 
+  gt(
+    rowname_col = "group",
+    groupname_col = "exp"
+  ) |>
   fmt_number(is.numeric, decimals = 2) |>
-    #gtExtras::gt_plt_dist(correct_list, type = "density") |> 
+  # gtExtras::gt_plt_dist(correct_list, type = "density") |>
   fmt_integer(c(matches("sum|min|max"), n_participants)) |>
@@
   cols_merge_range(correct_min, correct_max) |>
-    cols_label(n_trials_sum = "N trials",
+  cols_label(
+    n_trials_sum = "N trials",
     n_participants = "N",
@@
     correct_sem = "SE",
-               correct_min = "Range") |> 
-    summary_rows(columns = is.integer,
-                 fns = list(label = md("*Sum*"),
+    correct_min = "Range"
+  ) |>
+  summary_rows(
+    columns = is.integer,
+    fns = list(
+      label = md("*Sum*"),
       id = "totals",
-                            fn = "sum"),
-                 fmt = ~fmt_integer(.)) |>
-    summary_rows(columns = matches("mean|sd|sem"),
-                 fns = list(label = md("*Mean*"), 
+      fn = "sum"
+    ),
+    fmt = ~ fmt_integer(.)
+  ) |>
+  summary_rows(
+    columns = matches("mean|sd|sem"),
+    fns = list(
+      label = md("*Mean*"),
       id = "means",
-                            fn = "mean"),
-                 fmt = ~fmt_number(., scale_by = 100)) |> 
-    tab_style(cell_text(align = "center"),
-              cells_column_labels()) |> 
-    tab_style(cell_text(style = "italic"),
-              cells_column_labels()) |> 
-    tab_style(cell_text(weight = "bold"),
-              cells_column_spanners()) |> 
-    tab_style(cell_text(align = "left"),
-              list(cells_body(), cells_column_labels()))
+      fn = "mean"
+    ),
+    fmt = ~ fmt_number(., scale_by = 100)
+  ) |>
+  tab_style(
+    cell_text(align = "center"),
+    cells_column_labels()
+  ) |>
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels()
+  ) |>
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners()
+  ) |>
+  tab_style(
+    cell_text(align = "left"),
+    list(cells_body(), cells_column_labels())
+  )
 ```
@@
                        ...) {
-
   knowledge <- unique(model$data$knowledge)
@@
 get_epreds(exp_1_m0, dataset_1) |>
-    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+  mutate(
+    neigh_n_h = paste0(neigh_n_h, " neighbours"),
     neigh_n_h = factor(neigh_n_h,
-                              levels = paste0(c(0, 2, 4, 8, 12),
-                                              " neighbours"),
-                              ordered = TRUE)) |> 
+      levels = paste0(
+        c(0, 2, 4, 8, 12),
+        " neighbours"
+      ),
+      ordered = TRUE
+    )
+  ) |>
   ggplot(aes(lv, .epred)) +
-    facet_grid(group~neigh_n_h) +
+  facet_grid(group ~ neigh_n_h) +
   stat_lineribbon(aes(fill_ramp = after_stat(level)),
-                    linewidth = 1/2,
-                    .width = c(0.95, 0.89, 0.78, 0.67, 0.50)) +
-    labs(x = "Similarity\n(Levenshtein distance with correct translation)",
+    linewidth = 1 / 2,
+    .width = c(0.95, 0.89, 0.78, 0.67, 0.50)
+  ) +
+  labs(
+    x = "Similarity\n(Levenshtein distance with correct translation)",
     y = "p(Correct)",
@@
     linetype = "Cross-language neighbourhood density",
-         fill_ramp = "CrI (%)") +
+    fill_ramp = "CrI (%)"
+  ) +
   scale_x_continuous(labels = scales::percent) +
@@
   scale_fill_brewer(palette = "Reds") +
-    theme(legend.position = "none",
-          legend.box = "vertical") 
+  theme(
+    legend.position = "none",
+    legend.box = "vertical"
+  )
 ```
@@
 
-In the present experiment, we investigated the extent to which the phonological similarity between translation equivalents is sufficient for successful word translation, in the absence of semantic knowledge.
-We tested two groups of monolingual British English natives in a translation task that involved words in Catalan or Spanish, two languages participants reported having no prior familiarity with.
+In Experiment 1, we investigated the extent to which the phonological similarity between translation equivalents is sufficient for successful word translation, in the absence of conceptual knowledge about the presented word.
+We tested two groups of monolingual British English-native adults in a translation task that involved words in Catalan or Spanish, two languages participants reported having no prior familiarity with.
 Participants benefited strongly from phonological similarity when the correct translation of the presented words in Catalan or Spanish had few English phonological neighbours with higher lexical frequency.
-This suggests that, in the absence of distractors, even naïve participants can efficiently use phonological similarity to succeed in a translation task.
-Our results suggest that word-to-word connections at the phonological level might play a strong role during L2 speech comprehension, specially in low-proficiency listeners.
+This suggests that word-forms in an unfamiliar language have a strong potential to activate their translation equivalents in the native language, provided some phonological similarity between both words, and the absence of more frequent phonological neighbors.
+
+Participants in Experiment 1 were surprisingly good at translating words from Catalan and Spanish (two unfamiliar languages) to their native language.
+If English participants were likely to activate the correct English translations of the presented words in Catalan and Spanish, it is possible that speakers of typologically closer languages to Catalan and Spanish may benefit even more strongly from phonological similarity in the same task.
+English is a Germanic language (like Dutch or German), while Catalan and Spanish are Romance languages (like Italian, French, Portuguese).
+English shares fewer phonologically similar translations with Romance languages than Romance languages share with each other.
+It is possible that the probability of homophonic translations is higher when listening to an unfamiliar language from the same typological family as the native language.
+We tested this hypothesis in Experiment 2.
 
@@
 participants_2 <- exp_participants |>
-    filter(group=="cat-SPA")
+  filter(group == "cat-SPA")
 
@@
 
-Participants in Spain were contacted via announcements in Faculties, and were compensated 5€ or an Amazon voucher for the same value.
+We collected data from `r n_participants` Spanish native adults living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
+Participants in Spain were contacted via announcements at the University campus(es), and were compensated 5€ or an Amazon voucher for the same value.
 Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
-Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
-We collected data from `r n_participants` Spanish native participants living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
-Stimuli were the same list of Catalan stimuli as in Experiment 1.
-Procedure and data analysis were identical as in Experiment 1.
+
+Stimuli were the same list of Catalan stimuli as in Experiment 1. Procedure and data analysis were identical as in Experiment 1.
 
@@
 dataset <- exp_responses |>
-    filter(group=="cat-SPA") |> 
+  filter(group == "cat-SPA") |>
   group_split(group) |>
@@
 dataset_clean <- exp_responses |>
-    filter(group=="cat-SPA") |> 
+  filter(group == "cat-SPA") |>
   dplyr::filter(valid_participant, valid_response) |>
@@
   mutate(len = nchar(response)) |>
-    summarise(across(len,
-                     lst(mean, median, sd, min, max)),
-              .by = c(group)) 
+  summarise(
+    across(
+      len,
+      lst(mean, median, sd, min, max)
+    ),
+    .by = c(group)
+  )
 ```
@@
 We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_spa["blank"]`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r r_validity$cat_spa["language"]`), in which participants did not provide a whole word (e.g., `f`, *n* = `r r_validity$cat_spa["incomplete"]`), and in which participants added comments to the experimenter (e.g., `unsure`, (*n* = `r r_validity$cat_spa["other"]`).
-In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a developmental language disorder (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
 After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_spa$participant_id) + n_distinct(dataset_clean$cat_spa$participant_id)` unique participants.
@@
 c2 <- gather_draws(exp_2_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
-    summarise(across(.value, lst(.median = median, 
+  summarise(across(.value, lst(
+    .median = median,
     .lower = \(x) hdi(x)[1],
     .upper = \(x) hdi(x)[2],
-                                 .rope = \(x) mean(between(x, -0.1, 0.1))),
-                     .names = "{fn}"))
+    .rope = \(x) mean(between(x, -0.1, 0.1))
+  ),
+  .names = "{fn}"
+  ))
 c2 <- split(c2, c2$.variable)
@@
 get_epreds(exp_2_m0, dataset_2) |>
-    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+  mutate(
+    neigh_n_h = paste0(neigh_n_h, " neighbours"),
     neigh_n_h = factor(neigh_n_h,
-                              levels = paste0(c(0, 2, 4, 8, 12),
-                                              " neighbours"),
-                              ordered = TRUE)) |> 
+      levels = paste0(
+        c(0, 2, 4, 8, 12),
+        " neighbours"
+      ),
+      ordered = TRUE
+    )
+  ) |>
   ggplot(aes(lv, .epred)) +
@@
   stat_lineribbon(aes(fill_ramp = after_stat(level)),
-                    linewidth = 1/2,
-                    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
-    labs(x = "Similarity\n(Levenshtein distance with correct translation)",
+    linewidth = 1 / 2,
+    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)
+  ) +
+  labs(
+    x = "Similarity\n(Levenshtein distance with correct translation)",
     y = "p(Correct)",
@@
 
-Experiment 2 was an extension of Experiment 1 to a population of monolinguals whose native language is typologically similar to the presented language.
-Particularly, we presented Catalan words to Spanish native speakers who were reportedly unfamiliar with Catalan.
+Experiment 2 was an extension of Experiment 1 to a population of monolinguals whose native language is typologically close to the presented language.
+We presented Catalan words to Spanish native adults who were reportedly unfamiliar with Catalan.
 Our results indicate a similar pattern of results as those in Experiment 1: participants were able to provide correct translations of presented Catalan words, provided that the Catalan words shared some degree of phonological similarity with their Spanish translation, and that the number of phonological neighbours with higher lexical frequency was reduced.
-In contrasts with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 was more resilient to the interference of phonological neighbours.
+In contrasts with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 was more resilient to the interference of phonological neighbourhood size.
 English natives in Experiment 1 provided significantly less accuracy responses when more than four phonological neighbors were present (even when translating high-similarity words), compared to when only one or neighbour were present.
@@
 #| label: tbl-surprises
-#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen th presented word (in Catalan or Spanish) and their correct Enlgish or Spanish translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
-tbl_data <- list("Experiment 1" = dataset_1,
-                 "Experiment 2" = dataset_2) |>
+#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct Enlgish translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+tbl_data <- list(
+  "Experiment 1" = dataset_1,
+  "Experiment 2" = dataset_2
+) |>
   bind_rows(.id = "exp") |>
-    summarise(n = n(),
+  summarise(
+    n = n(),
     correct = sum(correct),
-              .by = c(exp, group, lv, translation)) |> 
-    mutate(accuracy = prop_adj(correct, n),
-           accuracy_se = prop_adj_se(correct, n)) |>    
-    dplyr::filter(lv==0, accuracy >= 0.1) |> 
+    .by = c(exp, group, lv, translation)
+  ) |>
+  mutate(
+    accuracy = prop_adj(correct, n),
+    accuracy_se = prop_adj_se(correct, n)
+  ) |>
+  dplyr::filter(lv == 0, accuracy >= 0.1) |>
   select(exp, translation, group, accuracy, accuracy_se) |>
@@
 tbl_data |>
-    gt(rowname_col = "translation",
-       groupname_col = c("exp", "group")) |> 
+  gt(
+    rowname_col = "translation",
+    groupname_col = c("exp", "group")
+  ) |>
   fmt_number(is.numeric, scale_by = 100) |>
@@
 Spanish words percolate English speech with relative frequency, via different sources such as popular culture, songs, TV programs, etc.
-In addition, words from languages other than Spanish, but with high similarity to the Spanish words (e.g., cognates from Italian or French) might appear in English speech as well.
+In addition, words from languages other than Spanish or Catalan, but with high similarity to the Spanish or Catalan words (e.g., cognates from Italian or French) might appear in English speech as well.
 Such prior knowledge might not be specific to the low-similarity words highlighted before.
@@
 In the case of higher-similarity words, it is more difficult to disentangle the extent to which participants' accuracy is a function of pure phonological similarity, or prior knowledge they had about the meaning of Spanish words.
-To investigate this issue, we run Experiment 3.
+Experiment 3 was addressed at investigating this issue.
 
@@
 participants_3 <- quest_participants |>
-    filter(group!="cat-SPA")
+  filter(group != "cat-SPA")
 
@@
 
+We collected data from `r n_participants` British English native adults living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
-We collected data from `r n_participants` British English native participants living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
+Stimuli were the same list of Catalan stimuli as in Experiment 1.
 
@@
 excluded_impairment <- participants_3 |>
-    filter(valid_status=="Language impairment") |> 
+  filter(valid_status == "Language impairment") |>
   nrow()
@@
 excluded_ntrials <- participants_3 |>
-    filter(valid_status=="Insufficient trials") |> 
+  filter(valid_status == "Insufficient trials") |>
   nrow()
 ```
-
-Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
-Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-Stimuli were the same list of Catalan stimuli as in Experiment 1.
 
@@
 
-On each trial, participants listened to the audio stimulus by clicking on the `PLAY` button.
+In each trial, participants listened to the audio stimulus by clicking on the `PLAY` button.
 For comparability to the PsychoPy version, participants were only allowed to play the audio one time.
@@
 dataset <- quest_responses |>
-    filter(group!="cat-SPA") |> 
+  filter(group != "cat-SPA") |>
   group_split(group) |>
@@
 excluded_ntrials <- participants_3 |>
-    filter(valid_status=="Insufficient trials") |> 
+  filter(valid_status == "Insufficient trials") |>
   nrow()
@@
 dataset_clean <- quest_responses |>
-    filter(group!="cat-SPA") |> 
+  filter(group != "cat-SPA") |>
   dplyr::filter(valid_participant, valid_response) |>
@@
 
-After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$spa_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$spa_eng$participant_id)` unique participants.
 Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
 
-INSERT PRIOR KNOWLEDGE FILTER
+```{r knowledge}
+knowledge_cateng <- dataset_clean$cat_eng |> count(knowledge)
+fam_cateng <- dataset_clean$cat_eng |> 
+  filter(knowledge) |> 
+  summarise(confidence_mean = mean(confidence),
+  confidence_sd = sd(confidence))
+
+nofam_cateng <- dataset_clean$cat_eng |> 
+  filter(!knowledge) |> 
+  summarise(confidence_mean = mean(confidence),
+  confidence_sd = sd(confidence))
+
+knowledge_spaeng <- dataset_clean$spa_eng |> count(knowledge)
+fam_spaeng <- dataset_clean$spa_eng |> 
+  filter(knowledge) |> 
+  summarise(confidence_mean = mean(confidence),
+  confidence_sd = sd(confidence))
+
+nofam_spaeng <- dataset_clean$spa_eng |> 
+  filter(!knowledge) |> 
+  summarise(confidence_mean = mean(confidence),
+  confidence_sd = sd(confidence))
+```
+
+From the `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` total responses provided by English participants who listened to Catalan words, participants reported having prior knowledge of the presented Catalan words in `r knowledge_cateng$n[2]` (`r round(100*knowledge_cateng$n[2] / sum(knowledge_cateng$n), 2)`%) of them.
+In those responses, participants reported an average of `r round(fam_cateng$confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(fam_cateng$confidence_sd, 2)`).
+In responses in which no prior knowledge of the presented word was reported, average confidence was `r round(nofam_cateng $confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(nofam_cateng$confidence_sd, 2)`).
+From the `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` total responses provided by English participants who listened to Spanish words, participants reported having prior knowledge of the presented Spanish words in `r knowledge_spaeng$n[2]` (`r round(100*knowledge_spaeng$n[2] / sum(knowledge_spaeng$n), 2)`%) of them.
+In those responses, participants reported an average of `r round(fam_spaeng$confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(fam_spaeng$confidence_sd, 2)`).
+In responses in which no prior knowledge of the presented word was reported, average confidence was `r round(nofam_spaeng$confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(nofam_spaeng $confidence_sd, 2)`).
+Before data analysis, responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word were excluded from the dataset.
 
@@
 lengths <- dataset_3 |>
+  filter(!knowledge) |> 
   mutate(len = map_int(strsplit(response, ""), length)) |>
-    summarise(across(len,
-                     lst(mean, median, sd, min, max)),
-              .by = c(group)) 
+  summarise(
+    across(
+      len,
+      lst(mean, median, sd, min, max)
+    ),
+    .by = c(group)
+  )
 lengths <- split(lengths, lengths$group)
@@
 (B) Average accuracy for words with expected and unexpected accuracy."
-dataset_3 |> 
+plot_1_data <- dataset_3 |>
   mutate(surprise = translation %in% tbl_data$translation) |>
-    summarise(knowledge = mean(knowledge),
-              .by = c(translation, group, surprise)) |> 
+  summarise(
+    knowledge = mean(knowledge),
+    .by = c(translation, group, surprise)
+  ) |>
   select(translation, group, surprise, knowledge) |>
@@
     "Unexpectedly high accuracy\n(no phonological similarity)",
-        "Expected accuracy")) |> 
-    ggplot(aes(knowledge, group, shape = group)) +
-    facet_wrap(~surprise, ncol = 1, switch = "y") +
-    stat_summary(fun = mean,
-                 size = 2,
-                 geom = "point") +
-    stat_summary(fun.data = mean_se,
-                 geom = "errorbar",
-                 width = 0.1) +
-    labs(x = "Prior knowledge reported",
-         y = "Group") +
-    scale_x_continuous(labels = scales::percent) +
-
-    dataset_3 |> 
+    "Expected accuracy"
+  ))
+
+plot_1 <- plot_1_data |> 
+  filter(surprise=="Unexpectedly high accuracy\n(no phonological similarity)") |> 
+  ggplot(aes(1, knowledge, colour = group, fill = group)) +
+  facet_wrap(~group, ncol = 1) +
+  geom_dots(aes(group = NA), side = "both", shape = 1, layout = "swarm", stackratio = 1.5, dotsize = 1, binwidth = 0.05, show.legend = FALSE) +
+  geom_rect(
+    data = filter(plot_1_data, surprise=="Expected accuracy") |> 
+      summarise(
+      knowledge_mean = mean(knowledge), 
+      knowledge_std = sd(knowledge),
+      n = n(),
+      .by = group
+    ) |> 
+    mutate(knowledge_se = knowledge_std / sqrt(n)),
+   aes(xmin = 0.5, xmax = 1.5, 
+   ymin = knowledge_mean - knowledge_se, 
+   ymax = knowledge_mean + knowledge_se, 
+   fill = group, 
+   colour = group
+   ), 
+       inherit.aes = FALSE,
+   alpha = 1/4,
+   color = NA) +
+  geom_hline(
+    data = filter(plot_1_data, surprise=="Expected accuracy") |>
+      summarise(knowledge_mean = mean(knowledge), .by = group),
+    aes(yintercept = knowledge_mean, color = group)
+  ) +
+  labs(
+    y = "Prior knowledge reported",
+    x = "Group",
+    colour = "Group",
+    fill = "Group"
+  ) +
+  scale_y_continuous(labels = scales::percent) +
+  theme(axis.text.x = element_blank(),
+  axis.title.x = element_blank())
+
+plot_2_data <- dataset_3 |>
   mutate(surprise = translation %in% tbl_data$translation) |>
@@
     "Unexpectedly high accuracy\n(no phonological similarity)",
-        "Expected accuracy")) |> 
-    summarise(correct = sum(correct),
+    "Expected accuracy"
+  )) |>
+  summarise(
+    correct = sum(correct),
     n = n(),
-              .by = c(translation, group, surprise, knowledge)) |> 
+    .by = c(translation, group, surprise, knowledge)
+  ) |>
   mutate(accuracy = prop_adj(correct, n)) |>
-    select(translation, group, accuracy, surprise, knowledge) |> 
-    ggplot(aes(accuracy, group, shape = group)) +
-    facet_wrap(~surprise, ncol = 1, switch = "y") +
-    stat_summary(fun = mean,
-                 size = 2,
-                 position = position_dodge(width = 0.75),
-                 geom = "point") +
-    stat_summary(fun.data = mean_se,
-                 geom = "errorbar",
-                 width = 0.1,
-                 position = position_dodge(width = 0.75)) +
-    labs(x = "Accuracy",
-         y = "Group") +
-    scale_x_continuous(labels = scales::percent,
-                       limits = c(0, 1)) +
-    theme(axis.title.y = element_blank(),
-          axis.text.y = element_blank(),
-          strip.text = element_blank()) +
-
+  select(translation, group, accuracy, surprise, knowledge)
+
+plot_2 <- plot_2_data |> 
+  ggplot(aes(1, accuracy, colour = group, fill = group)) +
+  facet_wrap(~group, ncol = 1) +
+geom_dots(aes(group = NA), side = "both", shape = 1, layout = "swarm", stackratio = 0.7, dotsize = 0.5, binwidth = 0.05, show.legend = FALSE) +
+  geom_rect(
+    data = filter(plot_2_data, surprise=="Expected accuracy") |> 
+      summarise(
+      knowledge_mean = mean(knowledge), 
+      knowledge_std = sd(knowledge),
+      n = n(),
+      .by = group
+    ) |> 
+    mutate(knowledge_se = knowledge_std / sqrt(n)),
+   aes(xmin = 0.5, xmax = 1.5, 
+   ymin = knowledge_mean - knowledge_se, 
+   ymax = knowledge_mean + knowledge_se, 
+   fill = group, 
+   colour = group
+   ), 
+       inherit.aes = FALSE,
+   alpha = 1/4,
+   color = NA) +
+  geom_hline(
+    data = filter(plot_2_data, surprise=="Expected accuracy") |>
+      summarise(knowledge_mean = mean(knowledge), .by = group),
+    aes(yintercept = knowledge_mean, color = group)
+  ) +
+  labs(
+    y = "Accuracy",
+    x = "Group",
+    colour = "Group",
+    fill = "Group"
+  ) +
+  scale_y_continuous(
+    labels = scales::percent
+  ) +
+  theme(
+    axis.title.x = element_blank(),
+    axis.text.x = element_blank(),
+  )
+
+plot_1 | plot_2 +
   plot_layout(nrow = 1, guides = "collect") &
   plot_annotation(tag_levels = "A") &
-    labs(colour = "Group",
-         shape = "Group") &
-    theme(panel.border = element_rect(fill = NA,
+  labs(
+    colour = "Group",
+    shape = "Group"
+  ) &
+  theme(
+    panel.border = element_rect(
+      fill = NA,
       colour = "grey",
-                                      linewidth = 1),
+      linewidth = 1
+    ),
     axis.line = element_blank(),
-          axis.text.y = element_blank(),
-          axis.title.y = element_blank(),
-          axis.ticks.y = element_blank(),
-          legend.position = "top",
-          panel.grid.major.x = element_line(colour = "grey",
-                                            linetype = "dotted"),
-          panel.grid.major.y = element_blank()) 
+    axis.text.x = element_blank(),
+    axis.title.x = element_blank(),
+    axis.ticks.x = element_blank(),
+    panel.grid.major.y = element_line(
+      colour = "grey",
+      linetype = "dotted"
+    ),
+    panel.grid.major.x = element_blank()
+  )
 ```
@@
 c3 <- gather_draws(exp_3_m1, `b_.*`, `sd_.*`, regex = TRUE) |>
-    summarise(across(.value, lst(.median = median, 
+  summarise(across(.value, lst(
+    .median = median,
     .lower = \(x) hdi(x)[1],
     .upper = \(x) hdi(x)[2],
-                                 .rope = \(x) mean(between(x, -0.1, 0.1))),
-                     .names = "{fn}"))
+    .rope = \(x) mean(between(x, -0.1, 0.1))
+  ),
+  .names = "{fn}"
+  ))
 c3 <- split(c3, c3$.variable)
@@
 
-Participants translating Catalan words and participants translating Spanish words, as indicated by the regression coefficient of *Group* ($\beta$ = `r round(c3$b_group1$.median, 3)`, 95% CrI = [`r round(c3$b_group1$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_group1$.rope, 3)`).
-Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
+Participants translating Catalan words and participants translating Spanish words performed similarly, as indicated by the regression coefficient of *Group* ($\beta$ = `r round(c3$b_group1$.median, 3)`, 95% CrI = [`r round(c3$b_group1$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_group1$.rope, 3)`).
+Overall, both groups of participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
 This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c3$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h_lv$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h_lv$.rope, 3)`).
@@
 get_epreds(exp_3_m1, dataset_3) |>
-    mutate(neigh_n_h = paste0(neigh_n_h, " neighbours"),
+  mutate(
+    neigh_n_h = paste0(neigh_n_h, " neighbours"),
     neigh_n_h = factor(neigh_n_h,
-                              levels = paste0(c(0, 2, 4, 8, 12),
-                                              " neighbours"),
-                              ordered = TRUE)) |> 
+      levels = paste0(
+        c(0, 2, 4, 8, 12),
+        " neighbours"
+      ),
+      ordered = TRUE
+    )
+  ) |>
   ggplot(aes(lv, .epred)) +
-    facet_grid(group~neigh_n_h) +
+  facet_grid(group ~ neigh_n_h) +
   stat_lineribbon(aes(fill_ramp = after_stat(level)),
-                    linewidth = 1/2,
-                    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)) +
-    labs(x = "Similarity\n(Levenshtein distance with correct translation)",
+    linewidth = 1 / 2,
+    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)
+  ) +
+  labs(
+    x = "Similarity\n(Levenshtein distance with correct translation)",
     y = "p(Correct)",
-         fill_ramp = "CrI (%)") +
+    fill_ramp = "CrI (%)"
+  ) +
   scale_x_continuous(labels = scales::percent) +
@@
   scale_fill_brewer(palette = "Reds") +
-    theme(legend.position = "top",
-          legend.box = "vertical") 
+  theme(
+    legend.position = "top",
+    legend.box = "vertical"
+  )
 ```
@@
 
-# Joint analyses
+# Pooled analyses of Experiments 1 and 3
 
@@
 #| fig-height: 5
-var_labels <- c("b_Intercept" = "Intercept",
+var_labels <- c(
+  "b_Intercept" = "Intercept",
   "b_freq_zipf_2" = "Frequency",
@@
   "b_neigh_n_h:lv" = "Similarity × CLPN",
-                "b_group1" = "cat-ENG vs. spa-ENG")
+  "b_group1" = "cat-ENG vs. spa-ENG"
+)
 
@@
   dplyr::filter(.variable != "b_Intercept") |>
-    mutate(.value = .value/4,
+  mutate(
+    .value = .value / 4,
     .variable = factor(.variable,
       levels = names(var_labels),
-                              labels = var_labels)) |> 
+      labels = var_labels
+    )
+  ) |>
   ggplot(aes(.value, reorder(exp, desc(exp)))) +
   facet_wrap(~.variable, scales = "free_x") +
-    annotate(geom = "rect",
-             xmin = -0.1/4,
-             xmax = 0.1/4,
+  annotate(
+    geom = "rect",
+    xmin = -0.1 / 4,
+    xmax = 0.1 / 4,
     ymin = -Inf,
@@
     fill = "grey",
-             alpha = 3/4,
-             colour = NA) +
-    geom_vline(xintercept = 0, 
+    alpha = 3 / 4,
+    colour = NA
+  ) +
+  geom_vline(
+    xintercept = 0,
     colour = "grey",
-               linewidth = 3/4) +
+    linewidth = 3 / 4
+  ) +
   stat_slab(aes(fill = after_stat(level)),
@@
     breaks = 20,
-              normalize = "panels") +
-    stat_spike(at = "median",
+    normalize = "panels"
+  ) +
+  stat_spike(
+    at = "median",
     size = 1,
     aes(linetype = "Median"),
-               normalize = "panels") +
-    labs(x = "Estimate", y = "Experiment",
+    normalize = "panels"
+  ) +
+  labs(
+    x = "Estimate", y = "Experiment",
     colour = "CrI",
     fill = "CrI",
-         linetype = "") +
+    linetype = ""
+  ) +
   scale_fill_brewer(palette = "Reds") +
-    theme(legend.position = c(1, 0),
+  theme(
+    legend.position = c(1, 0),
     legend.box = "horizontal",
@@
     axis.title.x = element_blank(),
-          axis.title.y = element_blank())
+    axis.title.y = element_blank()
+  )
+```
+
+```{r tbl-results}
+#| label: tbl-results
+tbl_data <- list(
+  "Experiment 1" = dataset_1,
+  "Experiment 2" = dataset_2,
+  "Experiment 3" = dataset_3
+) |>
+  bind_rows(.id = "exp") |>
+  count(exp, group, response_type) |>
+  pivot_wider(
+    names_from = response_type,
+    values_from = n,
+    names_repair = janitor::make_clean_names
+  ) |>
+  relocate(exp, group, correct, typo, wrong, false_friend) |>
+  mutate(
+    n_correct = rowSums(cbind(correct, typo)),
+    n_incorrect = rowSums(cbind(wrong, false_friend)),
+    n_total = rowSums(cbind(n_correct, n_incorrect)),
+    across(correct:false_friend,
+      \(x) x / n_total,
+      .names = "{col}_prop"
+    )
+  ) |>
+  select(
+    exp, group, correct, typo, wrong, false_friend,
+    matches("prop")
+  ) |>
+  relocate(
+    correct, correct_prop,
+    typo, typo_prop,
+    wrong, wrong_prop,
+    false_friend, false_friend_prop
+  )
+
+tbl_data |>
+  gt(rowname_col = "group", groupname_col = "exp") |>
+  fmt_missing(everything(), everything(), "--") |>
+  tab_spanner(
+    "Correct responses",
+    matches("correct|typo")
+  ) |>
+  tab_spanner(
+    "Incorrect responses",
+    matches("wrong|false")
+  ) |>
+  fmt_integer(c(correct, typo, wrong, false_friend)) |>
+  fmt_number(is.double, pattern = "({x})", scale_by = 100) |>
+  cols_label(
+    correct = "Correct",
+    typo = "Typo",
+    wrong = "Wrong",
+    false_friend = "False friend",
+    correct_prop = "(%)",
+    typo_prop = "(%)",
+    wrong_prop = "(%)",
+    false_friend_prop = "(%)"
+  ) |>
+  tab_style(
+    cell_text(weight = "bold"),
+    cells_column_spanners()
+  ) |>
+  tab_style(
+    cell_text(style = "italic"),
+    cells_column_labels()
+  ) |>
+  summary_rows(
+    columns = is.integer,
+    fns = list(
+      label = md("*Sum*"),
+      id = "totals",
+      fn = "sum"
+    ),
+    fmt = ~ fmt_integer(.)
+  ) |>
+  summary_rows(
+    columns = matches("prop"),
+    fns = list(
+      label = md("*Mean*"),
+      id = "means",
+      fn = "mean"
+    ),
+    fmt = ~ fmt_number(., scale_by = 100)
+  ) |>
+  tab_style(
+    cell_text(align = "center"),
+    cells_column_labels()
+  )
+
 ```
@@
 # General discussion
+
+The present work investigated the dynamics of lexical activation and selection dring non-native speech listening.
+Using a translation elicitation task, we collected data from a group of English native adults (Experiment 1) and a group of Spanish native adults (Experiment 2).
+Neither group was familiar with Catalan, Spanish, or any other Romance language.
+In each trial, participants were presented with Spanish words (Experiment 1) or Catalan words (Experiments 1 and 2), and then asked to type their best-guess translation for the presented word in their native language.
+Participants provided correct translations more often if the presented word and its translation shared phonological similarity, only if the number of phonological neighbours around the presented word was small.
+These results suggest that listening to words in an unfamiliar language can trigger dynamics of lexical activation and selection similar to those triggered by listening to speech in the native language.
+
+In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
+One group was presented with Catalan words, the other with Spanish words.
+We examined the extent to which participants were able to use the phonological similarity between the presented word (in Catalan or Spanish) and its target translation to provide accurate responses, despite reporting no prior relevant experience with the Catalan or Spanish.
+In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
+Catalan and Spanish are two Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
+Some participants in Experiments 1 and 2 were surprisingly good at translating a subset of words which had low phonological similarity with their correct translation.
+We were concerned that this may be caused by some prior knowledge of specific words by our participants, as some Spanish words are commonly seen in media or product labels, making them familiar even to monolingual speakers of English.
+In Experiment 3, we collected additional data from a new group of British English natives.
+The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
+After removing responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we found similar resutls as in Experiment 1, suggesting that participants did not simply rely on prior familiarity with the word-forms presented but rather exclusively on their phonological similarity with the correct English translation.
+
+Our findings provide insights into the mechanisms underlying homophonic translations.
+This phenomenon is frequently experienced by listeners of foreign speech, when strings of sounds in the input are perceived as strings of words or phrases in the native language, without necessarily preserving the meaning.
 

### 2024-10-08 manuscript/manuscript.tex
Add discussion and data about knowledge filter

@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
@@
 \makeatother
+
 \ifLuaTeX
@@
 \ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
+  \usepackage[english]{selnolig} % disable illegal ligatures
 \fi
@@
 (/\textipa{jUnomi}/), which is a type of Japanese teacup. Overall, the
-occurence of Soramimi---and more generally, homophonic
+occurrence of Soramimi---and more generally, homophonic
 translations---suggests that speech in an unfamiliar language has the
@@
 mistranslations like the ones aforementioned---homophonic translations
-sometimes may lead to the activation of the intended meaning (i.e., a
-correct translation). For example, imagine a Spanish native with no
-previous familiarity with any other language listening to French for the
-first time. This listener may encounter the word /\textipa{"pOKt}/
-(\emph{porte}), translation of door in French. When comparing the
-phonemic transcription of the French to that of its Spanish translation
-/\textipa{"pweR.ta}/ (\emph{puerta}), one may be led to infer that these
-two words do not have much overlap. But one thing to note is that the
-voiced uvular fricative consonant /\textipa{K}/ and the open-mid back
-rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic
-categories, can be perceived as allophonic variations the native
-(Spanish) phonemes /\textipa{R}/ and /o/, as described by assimilation
-models of phoneme perception (\citeproc{ref-best1988examination}{Best et
-al., 1988}, \citeproc{ref-best2001discrimination}{2001}). Although this
-phonetic mismatch between the non-native word and its translation can
-have a noticeable toll on comprehension Cutler et al.
+sometimes may lead to correct translations. For example, imagine a
+Spanish native with no previous familiarity with any other language
+listening to French for the first time. This listener may encounter the
+word /\textipa{"pOKt}/ (\emph{porte}), translation of door in French.
+When comparing the phonemic transcription of the French to that of its
+Spanish translation /\textipa{"pweR.ta}/ (\emph{puerta}), one may be led
+to infer that these two words do not have much overlap. But one thing to
+note is that the voiced uvular fricative consonant /\textipa{K}/ and the
+open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish
+as phonemic categories, can be perceived as allophonic variations the
+native (Spanish) phonemes /\textipa{R}/ and /o/, as described by
+assimilation models of phoneme perception
+(\citeproc{ref-best1988examination}{Best et al., 1988},
+\citeproc{ref-best2001discrimination}{2001}). Although this phonetic
+mismatch between the non-native word and its translation can have a
+noticeable toll on comprehension Cutler et al.
 (\citeproc{ref-cutler2004patterns}{2004}), the Spanish listener may
@@
 
-While Soramimi, and more generally homophonic translation, are popular
+While Soramimi (and more generally homophonic translations) are popular
 and frequent phenomena, the psycholinguistic literature on the topic is
@@
 
-While the previous studies have address the phonological aspects of
-homophonic translation, little attention has been paid to the dynamics
-of lexical activation and competition that underlay this phenomenon. In
-the present work we investigated the interplay between phonological
-similarity and phonological neighbourhood density to examine how
-listeners of an unfamiliar language activate lexical representations in
-their native language. In particular, we studied how the presence of
+These studies have addressed the phonological aspects of homophonic
+translation, but little attention has been paid to the dynamics of
+lexical activation and competition that underlay this phenomenon. In the
+present work, we investigated the interplay between phonological
+similarity and phonological neighbourhood density in a non-native speech
+perception task. In particular, we studied how the presence of
 phonological neighbours in the native language lexicon affects the
 extent to which listeners are able to activate the correct translation
-of an auditorily presented word in an unfamiliar language. We used a
+of an auditorily presented word in thier native language. We used a
 translation elicitation task in which participants listened to words
@@
 auditory-presented words heard by participants on each trial as
-presented words, and the correct translation for the presented words as
-target translations. We explored listeners' reliance on phonological
-similarity to succeed in the task by manipulating the amount of
-phonological similarity between the presented words and their target
-translations. Since participants were unfamiliar with the presented
-language, they could only use phonological similarity between the
-presented language and their native language to successfully translate
-the words. We therefore predicted that participants' performance should
-increase when the translation pairs are phonologically more similar. We
-also predicted that there is a minimum threshold of phonological
-similarity to be sufficient for translation.
+\emph{presented words}, and the correct translation for the presented
+words as \emph{target translations}. We explored listeners' reliance on
+phonological similarity to succeed in the task by manipulating the
+amount of phonological similarity between the presented words and their
+target translations. Since participants were unfamiliar with the
+presented language, they could only use phonological similarity between
+the presented language and their native language to successfully
+translate the words. We therefore predicted that participants'
+performance should increase when the translation pairs are
+phonologically more similar. We also predicted that there is a minimum
+threshold of phonological similarity to be sufficient for translation.
 
@@
 (European) Spanish natives in the same task, who were presented with
-Catalan words. Catalan and Spanish are both Romance languages whose
-close typological distance is reflected in the fact that they share many
+Catalan words. Catalan and Spanish are two Romance languages whose close
+typological distance is reflected in the fact that they share many
 cognates, where English is a Germanic language that shares considerably
@@
 trial, participants reported whether they had previous knowledge of the
-presented word. In the final section of this paper, we present analyses
-on the joint data sets of Experiments 1 to 3.
+presented word. In the final section of this article, we present
+analyses on the joint data sets of Experiments 1 to 3.
 
@@
 
-Participants were recruited via Prolific (5£ compensation) and SONA
-(compensation in academic credits), and gave informed consent before
-providing any data and the study was conducted in accordance with
-ethical standards of the Declaration of Helsinki and the protocol was by
-the University of Oxford Medical Sciences Inter-Divisional Research
-Ethics Committee (IDREC) (R60939/RE007). Participants were asked to
-complete the experiment using a laptop in a quiet place with good
-internet connection. Data collection took place from June 04th, 2020 to
-June 25th, 2020. We collected data from 71 British English native
-participants living in United Kingdom (\emph{Mean} = 21.76 years,
-\emph{SD} = 2.15, Range = 18-26, 46, 25 female).
+We collected data from 71 British English-native adults living in United
+Kingdom (\emph{Mean} = 21.76 years, \emph{SD} = 2.15, Range = 18-26, 46
+female). Data collection took place from June 04th, 2020 to June 25th,
+2020. Participants were recruited via Prolific (5£ compensation) and
+SONA (compensation in academic credits). Participants gave informed
+consent before providing any data. The study was conducted in accordance
+with ethical standards of the Declaration of Helsinki and the protocol
+was by the University of Oxford Medical Sciences Inter-Divisional
+Research Ethics Committee (IDREC) (R60939/RE007). Participants were
+asked to complete the experiment using a laptop in a quiet place with
+good internet connection.
 
@@
 
-\caption{\label{tbl-participants}}
+\caption{\label{tbl-participants}Participant details.}
 
@@
 
-We arranged two lists of input words to be presented to participants in
+We created two lists of input words to be presented to participants in
 the auditory modality: one in Catalan and one in Spanish. Words in the
@@
 participants had to type) were 5.12 characters long on average
-(\emph{SD} = 1.56, Range = 3-9. Words in the Spanish list were 5.52
+(\emph{SD} = 1.56, Range = 3-9). Words in the Spanish list were 5.52
 phonemes long on average (\emph{SD} = 1.47, Range = 3-9), and the
-orthographic form of their English translations (which participants had
-to type), were 5.29 characters long on average (\emph{SD} = 1.77, Range
-= 3-12.
-
-Participants listened to one audio file in each trial, each containing a
-single word presented in isolation. The audio files were the same ones
-used in child experiments conducted in the Laboratori de Recerca en
-Infància of Universitat Pompeu Fabra (Barcelona, Spain). These audio
-files were recorded by a proficient Catalan-Spanish female bilingual
-from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files
-in the same session, and then de-noised using Audacity and normalised at
-peak intensity using Praat (\citeproc{ref-broersma2021praat}{Broersma \&
+orthographic form of their English translations were 5.29 characters
+long on average (\emph{SD} = 1.77, Range = 3-12).
+
+In each trial, participants listened to one audio file, wich contained a
+single word. The audio files were the same ones used in child
+experiments conducted in the Laboratori de Recerca en Infància of
+Universitat Pompeu Fabra (Barcelona, Spain). These audio files were
+recorded by a proficient Catalan-Spanish female bilingual from the
+Metropolitan Area of Barcelona in a child-directed manner. Catalan and
+Spanish words were recorded at 44,100 Hz in separate files in the same
+session, and then de-noised using Audacity and normalised at peak
+intensity using Praat (\citeproc{ref-broersma2021praat}{Broersma \&
 Weenink, 2021}). The average duration of the Catalan audio files was
-1.24 (\emph{SD} = 0.19, Range = 0.8-1.58). The average duration of the
-Catalan audio files was 1.16 (\emph{SD} = 0.15, Range = 0.78-1.53).
-
-For each translation pair, we defined three predictors of interest: the
-lexical frequency of the correct translation (\emph{Frequency}), the
-phonological similarity between the presented (Catalan or Spanish) word
-and their correct English translation (\emph{Similarity}), and the
-presented word's number of cross-language phonological neighbours
-(\emph{CLPN}).
-
-We included \emph{Frequency} as a nuance predictor, under the hypothesis
-that---keeping other predictors constant---participants would translate
-higher-frequency words more accurately and faster than lower-frequency
-words. Lexical frequencies of correct translations were extracted from
-SUBTLEX-UK (\citeproc{ref-van2014subtlex}{Van Heuven et al., 2014}), and
-transformed to Zipf scores. Words in the stimuli list without a lexical
-frequency value were excluded from data analysis (2 in the Catalan list,
-5 in the Spanish list).
+1.24 seconds (\emph{SD} = 0.19, Range = 0.8-1.58). The average duration
+of the Spanish audio files was 1.16 seconds (\emph{SD} = 0.15, Range =
+0.78-1.53).
+
+For each word in the Catalan and Spanish lists pair, we defined three
+predictors of interest: the lexical frequency of the correct English
+translation (\emph{Frequency}), the phonological similarity between the
+presented word (in Catalan or Spanish) and their correct English
+translation (\emph{Similarity}), and the presented word's number of
+cross-language phonological neighbours (\emph{CLPN}) in English.
+\emph{Frequency} was included as a nuance predictor, under the
+hypothesis that---keeping other predictors constant---participants would
+translate higher-frequency words more accurately and faster than
+lower-frequency words. Lexical frequencies of correct translations were
+extracted from SUBTLEX-UK (\citeproc{ref-van2014subtlex}{Van Heuven et
+al., 2014}), and transformed to Zipf scores. Words in the stimuli list
+without a lexical frequency value were excluded from data analysis (2 in
+the Catalan list, 5 in the Spanish list).
 
@@
 distance computes the edit distance between two character strings---in
-this case, phoneme transcriptions---by counting the number of additions,
-deletions, and substitutions necessary to make both strings identical
-(\citeproc{ref-levenshtein1966binary}{Levenshtein, 1966}). We divided
-this edit distance by the number of characters included in the longest
-X-SAMPA transcription of the translation pair. This results in a
+this case, two phonological transcriptions---by counting the number of
+additions, deletions, and substitutions necessary to make both strings
+identical (\citeproc{ref-levenshtein1966binary}{Levenshtein, 1966}). We
+divided this edit distance by the number of characters included in the
+longest X-SAMPA transcription of the translation pair. This results in a
 proportion score, in which values closer to zero correspond to lower
@@
 similarity), and values closer to 1 correspond to higher Levenshtein
-distances (i.e., lower similarity). This transformation also help
-account for the fact that the Levenshtein distance tends to increase
-with the length of the transcriptions. For interpretability, we
-subtracted this proportion from one, so that values closer to one
-correspond to higher similarity between phonological transcriptions, and
-values closer to zero correspond to lower similarity between
-phonological transcriptions. For example, the \emph{table}
-(\texttt{teIb@l})-\emph{mesa} (\texttt{mesa}) translation pair had a
-17\% similarity, while the \emph{train} (\texttt{trEIn})-\emph{tren}
-(\texttt{tREn}) translation pair had a 60\% similarity.
-Table~\ref{tbl-stimuli} summarises the lexical frequency, phonological
-neighbourhood density and phonological overlap of the words included in
-the Catalan and the Spanish lists.
+distances (i.e., lower similarity). This transformation accounts for the
+fact that the Levenshtein distance tends to increase with the length of
+the transcriptions. For interpretability, we subtracted this proportion
+from one, so that values closer to one correspond to higher similarity
+between phonological transcriptions, and values closer to zero
+correspond to lower similarity between phonological transcriptions. For
+example, the \emph{table} (\texttt{teIb@l})-\emph{mesa} (\texttt{mesa})
+translation pair had a 17\% similarity, while the \emph{train}
+(\texttt{trEIn})-\emph{tren} (\texttt{tREn}) translation pair had a 60\%
+similarity. Table~\ref{tbl-stimuli} summarises the lexical frequency,
+phonological neighbourhood density and phonological overlap of the words
+included in the Catalan and the Spanish lists.
 
@@
 
-\caption{\label{tbl-stimuli}Stimuli properties.}
+\caption{\label{tbl-stimuli}Stimuli details.}
 
@@
 
-{\centering \includegraphics{_assets/img/design.png}
+{\centering \pandocbounded{\includegraphics[keepaspectratio]{_assets/img/design.png}}
 
@@
 
-\caption{Schematic representation of a trial in the experimental task}
+\caption{Schematic representation of a trial in the experimental task.
+The trial stars with the presentation of a fixation point in the centre
+of the screen (yellow dot). After 1,000 ms, the auditory word was
+presented while the fixation point remained on screen. After the offset
+of the audio, the fixation point disappeared and a visual prompt
+(\texttt{\textgreater{}}) was presented. Participants then wrote their
+and clicked \texttt{RETURN}, making the end of the trial.}
 
@@
 
-After data collection, participants answers were manually coded into the
-following categories: \emph{Correct}, \emph{Typo}, \emph{Wrong},
-\emph{False friend}, \emph{Other}. Responses were coded as
+After data collection, participants' answers were manually coded into
+the following categories: \emph{Correct}, \emph{Typo}, \emph{Wrong},
+\emph{False friend}, \emph{Other}. A response was coded as
 \emph{Correct} if the provided string of characters was identical to the
-orthographic form of the correct translation. Trials were coded as
+orthographic form of the correct translation. A response was coded as
 \emph{Typo} if the participant provided a string of characters only one
@@
 of ``penguin''), as long as the response did not correspond to a
-distinct English word. Responses were coded as \emph{False friend} if
+distinct English word. A response was coded as \emph{False friend} if
 the participant provided a phonologically similar incorrect translation.
-Responses were coded as \emph{Other} (see Data analysis section for more
+Responses not meeting the criteria for previous categories were labelled
+as \emph{Wrong} or \emph{Other} (see Data analysis section for more
 details). Both \emph{Correct} and \emph{Typo} responses were considered
@@
 \end{aligned}
-\]
-
-To test the practical relevance of each predictor we followed
-(\citeproc{ref-kruschke2018bayesian}{Kruschke \& Liddell, 2018}). We
-first specified a region of practical equivalence (\emph{ROPE}) around
-zero ({[}-0.1, +0.1{]}, in the logit scale). This area indicates the
-values of the regression coefficients that we considered as equivalent
-to zero. We then computed the 95\% posterior credible intervals (CrI) of
-the regression coefficients of interest. Finally, we calculated the
+\] ::: \{eq-1\}
+
+To test the practical relevance of each predictor we followed Kruschke
+\& Liddell (\citeproc{ref-kruschke2018bayesian}{2018}). We first
+specified a region of practical equivalence (\emph{ROPE}) around zero
+({[}-0.1, +0.1{]}, in the logit scale). This area indicates the values
+of the regression coefficients that we considered as equivalent to zero.
+We then computed the 95\% posterior credible intervals (CrI) of the
+regression coefficients of interest. Finally, we calculated the
 proportion of the 95\% CrI that fell inside the ROPE. This proportion
@@
 2-21).
-
-\begin{longtable}{l|rrrrrrrr}
-
-\caption{\label{tbl-results}}
-
-\tabularnewline
-
-\toprule
-\multicolumn{1}{l}{} & \multicolumn{4}{c}{Correct responses} & \multicolumn{4}{c}{Incorrect responses} \\ 
-\cmidrule(lr){2-5} \cmidrule(lr){6-9}
-\multicolumn{1}{l}{} & Correct & (\%) & Typo & (\%) & Wrong & (\%) & False friend & (\%) \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{9}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $429$ & ($16.47$) & $11$ & ($0.42$) & $2,082$ & ($79.95$) & $82$ & ($3.15$) \\ 
-spa-ENG & $374$ & ($14.37$) & $2$ & ($0.08$) & $2,117$ & ($81.36$) & $109$ & ($4.19$) \\ 
-\midrule 
-\emph{Sum} & $803$ & — & $13$ & — & $4,199$ & — & $191$ & — \\ 
-\emph{Mean} & — & $15.42$ & — & $0.25$ & — & $80.66$ & — & $3.67$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{9}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & $780$ & ($46.93$) & $20$ & ($1.20$) & $736$ & ($44.28$) & $126$ & ($7.58$) \\ 
-\midrule 
-\emph{Sum} & $780$ & — & $20$ & — & $736$ & — & $126$ & — \\ 
-\emph{Mean} & — & $46.93$ & — & $1.20$ & — & $44.28$ & — & $7.58$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{9}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $477$ & ($18.01$) & $7$ & ($0.26$) & $1,986$ & ($75.00$) & $178$ & ($6.72$) \\ 
-spa-ENG & $590$ & ($19.53$) & $6$ & ($0.20$) & $2,294$ & ($75.94$) & $131$ & ($4.34$) \\ 
-\midrule 
-\emph{Sum} & $1,067$ & — & $13$ & — & $4,280$ & — & $309$ & — \\ 
-\emph{Mean} & — & $18.77$ & — & $0.23$ & — & $75.47$ & — & $5.53$ \\ 
-\bottomrule
-
-\end{longtable}
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}}
 
@@
 
-In the present experiment, we investigated the extent to which the
-phonological similarity between translation equivalents is sufficient
-for successful word translation, in the absence of semantic knowledge.
-We tested two groups of monolingual British English natives in a
-translation task that involved words in Catalan or Spanish, two
-languages participants reported having no prior familiarity with.
-Participants benefited strongly from phonological similarity when the
-correct translation of the presented words in Catalan or Spanish had few
-English phonological neighbours with higher lexical frequency. This
-suggests that, in the absence of distractors, even naïve participants
-can efficiently use phonological similarity to succeed in a translation
-task. Our results suggest that word-to-word connections at the
-phonological level might play a strong role during L2 speech
-comprehension, specially in low-proficiency listeners.
+In Experiment 1, we investigated the extent to which the phonological
+similarity between translation equivalents is sufficient for successful
+word translation, in the absence of conceptual knowledge about the
+presented word. We tested two groups of monolingual British
+English-native adults in a translation task that involved words in
+Catalan or Spanish, two languages participants reported having no prior
+familiarity with. Participants benefited strongly from phonological
+similarity when the correct translation of the presented words in
+Catalan or Spanish had few English phonological neighbours with higher
+lexical frequency. This suggests that word-forms in an unfamiliar
+language have a strong potential to activate their translation
+equivalents in the native language, provided some phonological
+similarity between both words, and the absence of more frequent
+phonological neighbors.
+
+Participants in Experiment 1 were surprisingly good at translating words
+from Catalan and Spanish (two unfamiliar languages) to their native
+language. If English participants were likely to activate the correct
+English translations of the presented words in Catalan and Spanish, it
+is possible that speakers of typologically closer languages to Catalan
+and Spanish may benefit even more strongly from phonological similarity
+in the same task. English is a Germanic language (like Dutch or German),
+while Catalan and Spanish are Romance languages (like Italian, French,
+Portuguese). English shares fewer phonologically similar translations
+with Romance languages than Romance languages share with each other. It
+is possible that the probability of homophonic translations is higher
+when listening to an unfamiliar language from the same typological
+family as the native language. We tested this hypothesis in Experiment
+2.
 
@@
 
-Participants in Spain were contacted via announcements in Faculties, and
-were compensated 5€ or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the
-study was conducted in accordance with ethical standards of the
+We collected data from 33 Spanish native adults living in Spain
+(\emph{Mean} = 21.85 years, \emph{SD} = 3, Range = 18-33, 28, 5 female).
+Data collection took place from June 08th, 2020 to June 28th, 2020.
+Participants in Spain were contacted via announcements at the University
+campus(es), and were compensated 5€ or an Amazon voucher for the same
+value. Participants gave informed consent before providing any data and
+the study was conducted in accordance with ethical standards of the
 Declaration of Helsinki and the protocol was approved by the Drug
 Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar
-(2020/9080/I). Data collection took place from June 08th, 2020 to June
-28th, 2020. We collected data from 33 Spanish native participants living
-in Spain (\emph{Mean} = 21.85 years, \emph{SD} = 3, Range = 18-33, 28, 5
-female). Stimuli were the same list of Catalan stimuli as in Experiment
-1. Procedure and data analysis were identical as in Experiment 1.
+(2020/9080/I).
+
+Stimuli were the same list of Catalan stimuli as in Experiment 1.
+Procedure and data analysis were identical as in Experiment 1.
 
@@
 and Spanish, or any other second language as four or higher in a
-five-point scale (\emph{n} = 22), were diagnosed with a language
-(\emph{n} = 1), or did not contribute more than 80\% of valid trials
-(\emph{n} = 9). After applying trial-level and participant-level
-inclusion criteria, the resulting dataset included 1,662 trials provided
-by 42 unique participants. Of those trials, 1,662 were provided by 21
-unique participants who listened to Catalan words, and 1,662 trials were
-provided by 21 unique participants who listened to Spanish words.
-Responses given by participants were 5.6 characters long on average
-(\emph{Median} = 5, \emph{SD} = 1.6, Range = 2-12).
+five-point scale (\emph{n} = 22), were diagnosed with a developmental
+language disorder (\emph{n} = 1), or did not contribute more than 80\%
+of valid trials (\emph{n} = 9). After applying trial-level and
+participant-level inclusion criteria, the resulting dataset included
+1,662 trials provided by 42 unique participants. Of those trials, 1,662
+were provided by 21 unique participants who listened to Catalan words,
+and 1,662 trials were provided by 21 unique participants who listened to
+Spanish words. Responses given by participants were 5.6 characters long
+on average (\emph{Median} = 5, \emph{SD} = 1.6, Range = 2-12).
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}}
 
@@
 Experiment 2 was an extension of Experiment 1 to a population of
-monolinguals whose native language is typologically similar to the
-presented language. Particularly, we presented Catalan words to Spanish
-native speakers who were reportedly unfamiliar with Catalan. Our results
-indicate a similar pattern of results as those in Experiment 1:
-participants were able to provide correct translations of presented
-Catalan words, provided that the Catalan words shared some degree of
-phonological similarity with their Spanish translation, and that the
-number of phonological neighbours with higher lexical frequency was
-reduced. In contrasts with the results in Experiment 1, the positive
-impact of phonological similarity on participants' performance in
-Experiment 2 was more resilient to the interference of phonological
-neighbours. English natives in Experiment 1 provided significantly less
-accuracy responses when more than four phonological neighbors were
-present (even when translating high-similarity words), compared to when
-only one or neighbour were present. Spanish participants in Experiment 2
-benefited from phonological similarity, even when eight neighbours were
-present. Spanish participants' performance declined after 8 neighbours,
-and was evident at 12 neighbours. Overall, this suggests that
-participants in Experiment 2, who were natives of a typologically
-similar language (Spanish) to the presented language (Catalan) benefited
-more strongly from phonological similarity than participants in
-Experiment 1, who were natives of typologically less similar language
-(English) to the presented language (Catalan, Spanish).
+monolinguals whose native language is typologically close to the
+presented language. We presented Catalan words to Spanish native adults
+who were reportedly unfamiliar with Catalan. Our results indicate a
+similar pattern of results as those in Experiment 1: participants were
+able to provide correct translations of presented Catalan words,
+provided that the Catalan words shared some degree of phonological
+similarity with their Spanish translation, and that the number of
+phonological neighbours with higher lexical frequency was reduced. In
+contrasts with the results in Experiment 1, the positive impact of
+phonological similarity on participants' performance in Experiment 2 was
+more resilient to the interference of phonological neighbourhood size.
+English natives in Experiment 1 provided significantly less accuracy
+responses when more than four phonological neighbors were present (even
+when translating high-similarity words), compared to when only one or
+neighbour were present. Spanish participants in Experiment 2 benefited
+from phonological similarity, even when eight neighbours were present.
+Spanish participants' performance declined after 8 neighbours, and was
+evident at 12 neighbours. Overall, this suggests that participants in
+Experiment 2, who were natives of a typologically similar language
+(Spanish) to the presented language (Catalan) benefited more strongly
+from phonological similarity than participants in Experiment 1, who were
+natives of typologically less similar language (English) to the
+presented language (Catalan, Spanish).
 
@@
 \caption{\label{tbl-surprises}List of items with unexpectedly high
-accuracy: the Levenshtein similarityscore betwen th presented word (in
-Catalan or Spanish) and their correct Enlgish or Spanish translation is
-zero, but participants, who are reportedly unfamiliar with the presented
-language, were on average \textgreater10\% likely to guess the correct
+accuracy: the Levenshtein similarityscore betwen the presented word (in
+Catalan or Spanish) and their correct Enlgish translation is zero, but
+participants, who are reportedly unfamiliar with the presented language,
+were on average \textgreater10\% likely to guess the correct
 translation.}
@@
 different sources such as popular culture, songs, TV programs, etc. In
-addition, words from languages other than Spanish, but with high
-similarity to the Spanish words (e.g., cognates from Italian or French)
-might appear in English speech as well. Such prior knowledge might not
-be specific to the low-similarity words highlighted before. Participants
-may also have had prior knowledge about higher-similarity words, which
-could have contributed to participants responding to such words more
-accurately than without such prior knowledge. In the case of
-higher-similarity words, it is more difficult to disentangle the extent
-to which participants' accuracy is a function of pure phonological
-similarity, or prior knowledge they had about the meaning of Spanish
-words. To investigate this issue, we run Experiment 3.
+addition, words from languages other than Spanish or Catalan, but with
+high similarity to the Spanish or Catalan words (e.g., cognates from
+Italian or French) might appear in English speech as well. Such prior
+knowledge might not be specific to the low-similarity words highlighted
+before. Participants may also have had prior knowledge about
+higher-similarity words, which could have contributed to participants
+responding to such words more accurately than without such prior
+knowledge. In the case of higher-similarity words, it is more difficult
+to disentangle the extent to which participants' accuracy is a function
+of pure phonological similarity, or prior knowledge they had about the
+meaning of Spanish words. Experiment 3 was addressed at investigating
+this issue.
 
@@
 
-Data collection took place from October 22th, 2022 to October 23th,
-2022. We collected data from 64 British English native participants
-living in United Kingdom (\emph{Mean} = 22.02 years, \emph{SD} = 2.49,
-Range = 18-26, 36, 28 female).
-
-Participants were recruited via Prolific (5£ compensation) and SONA
-(compensation in academic credits), and gave informed consent before
-providing any data and the study was conducted in accordance with
-ethical standards of the Declaration of Helsinki and the protocol was by
-the University of Oxford Medical Sciences Inter-Divisional Research
-Ethics Committee (IDREC) (R60939/RE007). Participants were asked to
-complete the experiment using a laptop in a quiet place with good
-internet connection. Stimuli were the same list of Catalan stimuli as in
-Experiment 1.
+We collected data from 64 British English native adults living in United
+Kingdom (\emph{Mean} = 22.02 years, \emph{SD} = 2.49, Range = 18-26, 36,
+28 female). Data collection took place from October 22th, 2022 to
+October 23th, 2022. Participants were recruited via Prolific (5£
+compensation) and SONA (compensation in academic credits), and gave
+informed consent before providing any data and the study was conducted
+in accordance with ethical standards of the Declaration of Helsinki and
+the protocol was by the University of Oxford Medical Sciences
+Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+Participants were asked to complete the experiment using a laptop in a
+quiet place with good internet connection. Stimuli were the same list of
+Catalan stimuli as in Experiment 1.
 
@@
 
-On each trial, participants listened to the audio stimulus by clicking
+In each trial, participants listened to the audio stimulus by clicking
 on the \texttt{PLAY} button. For comparability to the PsychoPy version,
@@
 After applying trial-level and participant-level inclusion criteria, the
-resulting dataset included 6,290 trials provided by 62 unique
+resulting dataset included 5,888 trials provided by 63 unique
 participants. Of those trials, 3,145 were provided by 31 unique
@@
 
-INSERT PRIOR KNOWLEDGE FILTER
+From the 3,145 total responses provided by English participants who
+listened to Catalan words, participants reported having prior knowledge
+of the presented Catalan words in 446 (14.18\%) of them. In those
+responses, participants reported an average of 5.05 confidence in the
+0-8 scale (\emph{SD} = 1.94). In responses in which no prior knowledge
+of the presented word was reported, average confidence was 1.13
+confidence in the 0-8 scale (\emph{SD} = 1.38). From the 2,743 total
+responses provided by English participants who listened to Spanish
+words, participants reported having prior knowledge of the presented
+Spanish words in 446 (14.18\%) of them. In those responses, participants
+reported an average of 4.79 confidence in the 0-8 scale (\emph{SD} =
+1.82). In responses in which no prior knowledge of the presented word
+was reported, average confidence was 1.25 confidence in the 0-8 scale
+(\emph{SD} = 1.66). Before data analysis, responses in which
+participants reported prior knowledge about the meaning of the presented
+Catalan or Spanish word were excluded from the dataset.
 
 Responses given by English participants to Catalan presented words were
-5.52 characters long on average (\emph{Median} = 5, \emph{SD} = 1.75,
-Range = 1-17), while their translations to Spanish responses were 5.41
-characters long on average (\emph{Median} = 5, \emph{SD} = 1.77, Range =
+5.49 characters long on average (\emph{Median} = 5, \emph{SD} = 1.71,
+Range = 1-14), while their translations to Spanish responses were 5.39
+characters long on average (\emph{Median} = 5, \emph{SD} = 1.75, Range =
 1-20).
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-knowledge-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-knowledge-1.pdf}}
 
@@
 Participants translating Catalan words and participants translating
-Spanish words, as indicated by the regression coefficient of
-\emph{Group} (\(\beta\) = -0.049, 95\% CrI = {[}-0.388, -0.32{]},
-\emph{p}(ROPE) = 0.426). Overall, participants responded less accurately
-to words with more CLPNs than to words with fewer CLPNs, regardless of
-the amount of phonological similarity between the presented word and its
-translation. This is indicated by the size of the regression coefficient
-of the two-way interaction between \emph{Similarity} and \emph{CLPN}
-(\(\beta\) = -0.665, 95\% CrI = {[}-1.068, -0.32{]}, \emph{p}(ROPE) =
-0.002). As anticipated, participants' performance benefited from an
-increase in \emph{Similarity} (\(\beta\) = 7.585, 95\% CrI = {[}7.043,
-8.24{]}, \emph{p}(ROPE) = 0), while the number of \emph{CLND} had the
-opposite effect (\(\beta\) = -0.049, 95\% CrI = {[}-0.203, 0.086{]},
+Spanish words performed similarly, as indicated by the regression
+coefficient of \emph{Group} (\(\beta\) = -0.049, 95\% CrI = {[}-0.388,
+-0.32{]}, \emph{p}(ROPE) = 0.426). Overall, both groups of participants
+responded less accurately to words with more CLPNs than to words with
+fewer CLPNs, regardless of the amount of phonological similarity between
+the presented word and its translation. This is indicated by the size of
+the regression coefficient of the two-way interaction between
+\emph{Similarity} and \emph{CLPN} (\(\beta\) = -0.665, 95\% CrI =
+{[}-1.068, -0.32{]}, \emph{p}(ROPE) = 0.002). As anticipated,
+participants' performance benefited from an increase in
+\emph{Similarity} (\(\beta\) = 7.585, 95\% CrI = {[}7.043, 8.24{]},
+\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
+effect (\(\beta\) = -0.049, 95\% CrI = {[}-0.203, 0.086{]},
 \emph{p}(ROPE) = 0.753). Figure~\ref{fig-epreds-1} illustrates the
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}}
 
@@
 
-\section{Joint analyses}\label{joint-analyses}
+\section{Pooled analyses of Experiments 1 and
+3}\label{pooled-analyses-of-experiments-1-and-3}
 
@@
 
-\includegraphics{manuscript_files/figure-pdf/fig-coefs-1.pdf}
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-coefs-1.pdf}}
 
@@
 
+\begin{longtable}{l|rrrrrrrr}
+
+\caption{\label{tbl-results}}
+
+\tabularnewline
+
+\toprule
+\multicolumn{1}{l}{} & \multicolumn{4}{c}{Correct responses} & \multicolumn{4}{c}{Incorrect responses} \\ 
+\cmidrule(lr){2-5} \cmidrule(lr){6-9}
+\multicolumn{1}{l}{} & Correct & (\%) & Typo & (\%) & Wrong & (\%) & False friend & (\%) \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{9}{l}{Experiment 1} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $429$ & ($16.47$) & $11$ & ($0.42$) & $2,082$ & ($79.95$) & $82$ & ($3.15$) \\ 
+spa-ENG & $374$ & ($14.37$) & $2$ & ($0.08$) & $2,117$ & ($81.36$) & $109$ & ($4.19$) \\ 
+\midrule 
+\emph{Sum} & $803$ & — & $13$ & — & $4,199$ & — & $191$ & — \\ 
+\emph{Mean} & — & $15.42$ & — & $0.25$ & — & $80.66$ & — & $3.67$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{9}{l}{Experiment 2} \\ 
+\midrule\addlinespace[2.5pt]
+cat-SPA & $780$ & ($46.93$) & $20$ & ($1.20$) & $736$ & ($44.28$) & $126$ & ($7.58$) \\ 
+\midrule 
+\emph{Sum} & $780$ & — & $20$ & — & $736$ & — & $126$ & — \\ 
+\emph{Mean} & — & $46.93$ & — & $1.20$ & — & $44.28$ & — & $7.58$ \\ 
+\midrule\addlinespace[2.5pt]
+\multicolumn{9}{l}{Experiment 3} \\ 
+\midrule\addlinespace[2.5pt]
+cat-ENG & $477$ & ($18.01$) & $7$ & ($0.26$) & $1,986$ & ($75.00$) & $178$ & ($6.72$) \\ 
+spa-ENG & $590$ & ($19.53$) & $6$ & ($0.20$) & $2,294$ & ($75.94$) & $131$ & ($4.34$) \\ 
+\midrule 
+\emph{Sum} & $1,067$ & — & $13$ & — & $4,280$ & — & $309$ & — \\ 
+\emph{Mean} & — & $18.77$ & — & $0.23$ & — & $75.47$ & — & $5.53$ \\ 
+\bottomrule
+
+\end{longtable}
+
 \section{General discussion}\label{general-discussion}
+
+The present work investigated the dynamics of lexical activation and
+selection dring non-native speech listening. Using a translation
+elicitation task, we collected data from a group of English native
+adults (Experiment 1) and a group of Spanish native adults (Experiment
+2). Neither group was familiar with Catalan, Spanish, or any other
+Romance language. In each trial, participants were presented with
+Spanish words (Experiment 1) or Catalan words (Experiments 1 and 2), and
+then asked to type their best-guess translation for the presented word
+in their native language. Participants provided correct translations
+more often if the presented word and its translation shared phonological
+similarity, only if the number of phonological neighbours around the
+presented word was small. These results suggest that listening to words
+in an unfamiliar language can trigger dynamics of lexical activation and
+selection similar to those triggered by listening to speech in the
+native language.
+
+In Experiment 1, we collected data from two groups of British English
+natives living in the United Kingdom. One group was presented with
+Catalan words, the other with Spanish words. We examined the extent to
+which participants were able to use the phonological similarity between
+the presented word (in Catalan or Spanish) and its target translation to
+provide accurate responses, despite reporting no prior relevant
+experience with the Catalan or Spanish. In Experiment 2, we tested a
+group of (European) Spanish natives in the same task, who were presented
+with Catalan words. Catalan and Spanish are two Romance languages whose
+close typological distance is reflected in the fact that they share many
+cognates, where English is a Germanic language that shares considerably
+fewer cognates with Catalan and Spanish. By testing participants
+translating words from typologically close or distant languages, we
+expected to widen the range of the phonological similarity scores of the
+translation pairs involved in the experimental task, therefore allowing
+us to explore potential cross-language differences in participant's
+performance. Some participants in Experiments 1 and 2 were surprisingly
+good at translating a subset of words which had low phonological
+similarity with their correct translation. We were concerned that this
+may be caused by some prior knowledge of specific words by our
+participants, as some Spanish words are commonly seen in media or
+product labels, making them familiar even to monolingual speakers of
+English. In Experiment 3, we collected additional data from a new group
+of British English natives. The design was closely modelled after
+Experiment 1, except that after providing their response in each trial,
+participants reported whether they had previous knowledge of the
+presented word. After removing responses in which participants reported
+prior knowledge about the meaning of the presented Catalan or Spanish
+word, we found similar resutls as in Experiment 1, suggesting that
+participants did not simply rely on prior familiarity with the
+word-forms presented but rather exclusively on their phonological
+similarity with the correct English translation.
+
+Our findings provide insights into the mechanisms underlying homophonic
+translations. This phenomenon is frequently experienced by listeners of
+foreign speech, when strings of sounds in the input are perceived as
+strings of words or phrases in the native language, without necessarily
+preserving the meaning.
 

### 2024-10-15 manuscript/manuscript.qmd
Add General discussion

@@
 
-The present work investigated the dynamics of lexical activation and selection dring non-native speech listening.
-Using a translation elicitation task, we collected data from a group of English native adults (Experiment 1) and a group of Spanish native adults (Experiment 2).
-Neither group was familiar with Catalan, Spanish, or any other Romance language.
-In each trial, participants were presented with Spanish words (Experiment 1) or Catalan words (Experiments 1 and 2), and then asked to type their best-guess translation for the presented word in their native language.
-Participants provided correct translations more often if the presented word and its translation shared phonological similarity, only if the number of phonological neighbours around the presented word was small.
-These results suggest that listening to words in an unfamiliar language can trigger dynamics of lexical activation and selection similar to those triggered by listening to speech in the native language.
-
-In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
-One group was presented with Catalan words, the other with Spanish words.
-We examined the extent to which participants were able to use the phonological similarity between the presented word (in Catalan or Spanish) and its target translation to provide accurate responses, despite reporting no prior relevant experience with the Catalan or Spanish.
-In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are two Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
-Some participants in Experiments 1 and 2 were surprisingly good at translating a subset of words which had low phonological similarity with their correct translation.
-We were concerned that this may be caused by some prior knowledge of specific words by our participants, as some Spanish words are commonly seen in media or product labels, making them familiar even to monolingual speakers of English.
-In Experiment 3, we collected additional data from a new group of British English natives.
-The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
-After removing responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we found similar resutls as in Experiment 1, suggesting that participants did not simply rely on prior familiarity with the word-forms presented but rather exclusively on their phonological similarity with the correct English translation.
-
-Our findings provide insights into the mechanisms underlying homophonic translations.
-This phenomenon is frequently experienced by listeners of foreign speech, when strings of sounds in the input are perceived as strings of words or phrases in the native language, without necessarily preserving the meaning.
-
-This type of cross-linguistic similarity has also been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
+The present work explored the psycholinguistic bases of homophonic translation, a phenomenon in which listening to speech in a non-native---perhaps unfamiliar---language leads to the activation of lexical representations in the native language, without necessarily preserving the meaning.
+We investigated how phonological similarity and its interaction with phonological neighbourhood density impact the dynamics of lexical activation and selection during non-native word processing.
+We designed a translation elicitation task in which participants listened to individual words in a non-native, unfamiliar language.
+After listening to each word, participants were asked to provide their best-guess translation for each word in their native language.
+We run three experiments.
+In Experiment 1, British English-native adults listened to a list of words in Catalan or Spanish.
+Participants reported no prior familiarity with Catalan, Spanish, or any other Romance language; yet, participants provided accurate translations for words that shared some degree of phonological similarity with their correct English translation.
+Participants only benefited from phonological similarity when the number of English phonological neighbours of the presented word was small.
+Experiment 2 was aimed at extending our findings to a population of participants whose native language was typologically closer to the unfamiliar presented language.
+We tested a group of Spanish native adults who listened to a list of Catalan words.
+Again, participants reported no prior familiarity with Catalan, or any other Romance language other than Spanish.
+In line with the results in English natives, we also found an interaction between phonological similarity and phonological neighbourhood size.
+The facilitation effect of phonological similarity in the Spanish participants was more robust to the interferring effect of phonological neighbourhood density, providing correct translations for words with high neighbourhood densities more often than English participants.
+In Experiment 3, we replicated the results from Experiment 1, while additionally controlling for participants' prior familiarity with the Catalan and Spanish words.
+We suspected that some English participants might have managed to surprisingly provide correct translations for some low-phonological similarity words---despite having reported being unfamiliar with Catalan, Spanish, or any other Romance language---thanks to prior experience with those words.
+We collected additional data from a new group of British English native adults.
+The experimental design in Experiment 3 was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
+After removing responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we found similar results as in Experiment 1, suggesting that participants did not simply rely on prior familiarity with the word-forms presented but rather exclusively on their phonological similarity with the correct English translation.
+
+Our findings provide novel psycholinguistic insights into homophonic translation from an experimental perspective.
+Homophonic is a frequent, often spontaneous phenomenon that has a has been described across many cultures and  languages.
+The most famous term coined for it is *Soramimi*, in Japanese.
+*Soramimi* refers specifically to homophonic translation in a musical context, when particular sections of the lyrics of a song in a non-native language make listeners think of particular words and phrases in their own native language, with possibly a completely different meaning.
+While *Somoramimi* refers to the spontaneous ocurrence of homophonic translations, this phenomenon is often intentionally exploited by translators to preserve the meter and "sound structure" (e.g., rhymes) of the original poem in the target language [e.g.,@levick2016translating;  @pilshchikov2016semiotics], in such way that the translator "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
+For isntance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).
+
+The underlying mechanisms that lead to spontaneous homophonic translations are informative of the undeerlying mechanisms of native speech processing, yet remain relatively unexplored.
+How does the lexical system deal with the non-native speech signal, in which form-meaning associations are scarce, if not absent?
+The present work addressed this issue by exploring the conditions under which non-native speech is likely to result in correct translations.
+To account for the fact that prior familiarity with the language (e.g., form-meaning associations) may contribute to correct translations, we tested participants in a language that they reported being unfamiliar with.
+In such conditions, participants can almost exlusively rely on the acoustic properties of the speech signal to provide correct translations.
+
+Our findings suggest that participants were able to exploit the phonological similarity between the presented words and their correct translation, even when both words shared less than 50% of their phonological form (as measured by their Levenshtein distance).
+How the perceptual system deals with non-native phonemes is a vastly studied subject.
+The same cannot be said about how the perceptual system deals with non-native *word-forms*.
+The present study provides experimental evidence suggesting that participants accommodate non-native phonological forms to their native phoneme inventory, leading to the activation of native lexical representations.
+One of the limitations of the present study is that Levenshtein similarity underestimates the actual phonetic similarity between the signal and the phonological form of the correct translations.
+Phonological transcriptions like X-SAMPA---which we used to calculate Levenshtein similarity----are abstract representations that ignore non-phonological contrasts (e.g., phones), and consider different symbols as completely different phonemes disregarding the fact that some phonemes share more phonetic features than others.
+Altogether, it is likely that participants exploited a more fine grained, raw source of information to provide correct answers.
+
+Our findings also suggest that, in line with previous studies of word-processing, phonological neighbourhood size interfered participants' ability to produce correct translations, even for words with high phonological similarity with their correct translation.
+We calculated a measure of cross-linguistic neighbourhood size by counting, for each presented word (in Catalan or Spanish in Experiments 1 and 3, Catalan in Experiment 2) the number of phonologically related words (with a Levenshtein distance of one) in the target language (English in Experiments 1 and 3, Spanish in Experiment 2).
+To account for the fact that competition between phonological neighbours is sensitive to lexical frequency (high-freuqency neighbours produce stronger interference effects), we only counted neighbours with a lexical freuqency higher than the correct translation.
+Using this measure, we found that, across the three experiments, participants ability to exploit phonological similarity to produce correct translations declined as the number of phonological neighbours increased.
+
+A comparison between results in Experiments 1 and 2 suggested that the performance of Spanish participants translating Catalan was more resilient to the interfering effects of phonological neighbourhood density than the performance of English participants translating Catalan or Spanish.
+As highlihted before, Catalan and Spanish (both Romance languages) are typologically closer to each other than to English (a Germanic language).
+Catalan and Spanish share a higher proportion of cognates (i.e., form-similar translation equivalents) than Catalan and English, or Spanish and English.
+Previous studies in bilinguals show that this kind of cross-linguistic similarity impacts lexical processing:  words that share high phonological similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
 This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
 This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
+To our knowledge, this present study is the first to investigate the role of cross-linguistic phonological similarity in monolinguals.
+In particular, our participants were reportedly unfamiliar with the presented language.
+As such, they had to complete our translation elicitation task in the absence of form-meaning mappings, therefore relying exclusively on the phonological similarity between the presented words and their correct translations.
+This provides a convenient case to investigate the role of phonological similarity in lexical processing, disantangling the contribution of phonological similarity from the contribution of conceptual equivalence.
+Future studies might expand the methods used in the present study to elaborate in this issue.
+
+Two populations of speakers were tested in the present study: British English native adults living in the United Kindgom, and Spanish native adults living in Spain.
+Our findings suggest that the typological distance between the presented and target language is associated with a more robust facilitation effect of phonological similarity.
+Whether such association is linear (i.e., robustness of facilitation increases linearly with typological distance), proportional, or even extends beyond the specific pair of languages involved in the present work remains unclear.
+The generalisability of these findings can also be tested by increasing the repertoire of words involved in th translation elicitation task.
+In Experiments 1 to 3, participants answered to a maximum of 105 words.
+All words were of high-frequency, and had a low age-of-acquisition.
+In order to characterise listeners' ability to translate correctly words from an unfamiliar language, words with varying levels of difficulty should be included.
+
+In summary, the present paper provides insights into the psycholinguistic mechanisms underlying homophonic translation.
+English and Spanish native adults were tested in a translation elicitation task in which they had to guess the English or Spanish translation of a series of words in an unfamiliar language.
+Participants successfully exploited phonological similarity between the presented words and their correct translations to provide correct answers.
+Participants' performance in the task only benefited from phonological similarity when the presented word had few higher-frequency phonological neighbours in the target language.
+Finally, the faciltiation effect of phonological similarity was more robust in the Spanish native participants, who translated words from a typologically closer language than English participants.
+Overall, the findings presented in the present paper suggest that the processing of words in a non-native, unfamiliar language recruits mechanisms of lexical activation, selection, and interference parallel to those recruited by listening to words in a native language.
 

### 2024-10-15 manuscript/manuscript.tex
Add General discussion

@@
 words, participants reported having prior knowledge of the presented
-Spanish words in 446 (14.18\%) of them. In those responses, participants
+Spanish words in 197 (7.18\%) of them. In those responses, participants
 reported an average of 4.79 confidence in the 0-8 scale (\emph{SD} =
@@
 
-The present work investigated the dynamics of lexical activation and
-selection dring non-native speech listening. Using a translation
-elicitation task, we collected data from a group of English native
-adults (Experiment 1) and a group of Spanish native adults (Experiment
-2). Neither group was familiar with Catalan, Spanish, or any other
-Romance language. In each trial, participants were presented with
-Spanish words (Experiment 1) or Catalan words (Experiments 1 and 2), and
-then asked to type their best-guess translation for the presented word
-in their native language. Participants provided correct translations
-more often if the presented word and its translation shared phonological
-similarity, only if the number of phonological neighbours around the
-presented word was small. These results suggest that listening to words
-in an unfamiliar language can trigger dynamics of lexical activation and
-selection similar to those triggered by listening to speech in the
-native language.
-
-In Experiment 1, we collected data from two groups of British English
-natives living in the United Kingdom. One group was presented with
-Catalan words, the other with Spanish words. We examined the extent to
-which participants were able to use the phonological similarity between
-the presented word (in Catalan or Spanish) and its target translation to
-provide accurate responses, despite reporting no prior relevant
-experience with the Catalan or Spanish. In Experiment 2, we tested a
-group of (European) Spanish natives in the same task, who were presented
-with Catalan words. Catalan and Spanish are two Romance languages whose
-close typological distance is reflected in the fact that they share many
-cognates, where English is a Germanic language that shares considerably
-fewer cognates with Catalan and Spanish. By testing participants
-translating words from typologically close or distant languages, we
-expected to widen the range of the phonological similarity scores of the
-translation pairs involved in the experimental task, therefore allowing
-us to explore potential cross-language differences in participant's
-performance. Some participants in Experiments 1 and 2 were surprisingly
-good at translating a subset of words which had low phonological
-similarity with their correct translation. We were concerned that this
-may be caused by some prior knowledge of specific words by our
-participants, as some Spanish words are commonly seen in media or
-product labels, making them familiar even to monolingual speakers of
-English. In Experiment 3, we collected additional data from a new group
-of British English natives. The design was closely modelled after
-Experiment 1, except that after providing their response in each trial,
-participants reported whether they had previous knowledge of the
+The present work explored the psycholinguistic bases of homophone
+translations, a phenomenon in which listening to speech in a
+non-native---perhaps unfamiliar---language leads to the activation of
+lexical representations in the native language. We investigated how
+phonological similarity and its interaction with phonological
+neighbourhood density impact the dynamics of lexical activation and
+selection during non-native word processing. We designed a translation
+elicitation task in which participants listened to individual words in a
+non-native, unfamiliar language, and then were asked to provide their
+best-guess translation for each word in their native language. We run
+three experiments. In Experiment 1, British English-native adults
+listened to a list of words in Catalan or Spanish. Participants reported
+no prior familiarity with Catalan, Spanish, or any other Romance
+language; yet, participants provided accurate translations for words
+that shared some degree of phonological similarity with their correct
+English translation. Participants only benefited from phonological
+similarity when the number of English phonological neighbours of the
+presented word was small. Experiment 2 was aimed at extending our
+findings to a population of participants whose native language was
+typologically closer, to the (still unfamiliar) presented language. We
+tested a group of Spanish native adults who listened to a list of
+Catalan words. Again, participants reported no prior familiarity with
+Catalan, or any other Romance language other than Spanish. In line with
+the results in English natives, Spanish natives also showed an
+interaction between phonological similarity and phonological
+neighbourhood size. However, the facilitation effect of phonological
+similarity in the Spanish participants was more robust to the
+interferring effect of phonological neighbourhood density, providing
+correct translations for high-neighbourhood density words more often
+than participants in Experiment 1. In Experiment 3, we replicated the
+results from Experiment 1, while controlling for participants' prior
+familiarity with the Catalan and Spanish words. We suspected that some
+English participants might have managed to provide correct translations
+for some low-phonological similarity words despite being reportedly
+unfamiliar with Catalan, Spanish, or any other Romance language, thanks
+to previous exposure to those words. We collected additional data from a
+new group of British English natives. The design was closely modelled
+after Experiment 1, except that after providing their response in each
+trial, participants reported whether they had previous knowledge of the
 presented word. After removing responses in which participants reported
@@
 
-Our findings provide insights into the mechanisms underlying homophonic
-translations. This phenomenon is frequently experienced by listeners of
-foreign speech, when strings of sounds in the input are perceived as
-strings of words or phrases in the native language, without necessarily
-preserving the meaning.
-
-This type of cross-linguistic similarity has also been extremely
-informative for the study of bilingual lexical processing: words that
-share high form-similarity with their translations are processed faster
-and more accurately than words sharing lower phonological similarity.
-This phenomenon has provided strong evidence in favour of the
-language-non selective hypothesis of bilingual lexical access, which
-states that bilinguals activate lexical representations in both
-languages, even during monolingual situations. This facilitation effect
-has been reported for comprehension
+The results in the present work tap into homophonic translation, a
+frequent phenomenon to many speakers to occurs when speech in a
+non-native language activates lexical representations in the native
+language withhout necessarily preserving the meaning. This phenomenon
+has been described across many languages, and coined with different
+terms, the most famous being the \emph{Soramimi} in Japanese.
+\emph{Soramimi} refers specifically to homophonic translation in the
+context of music, when particular sections of the lyrics of a song in a
+non-native language make listeners think of particular words and phrases
+in their own native language, with possibly a completely different
+meaning. While \emph{Somoramimi} refers to the spontaneous ocurrence of
+homophonic translations, homophonic translation is often exploited by
+translators to preserve the meter and ``sound structure'' (e.g., rhymes)
+of the original poem in the target language (e.g.,
+\citeproc{ref-levick2016translating}{Levick, 2016};
+\citeproc{ref-pilshchikov2016semiotics}{Pilshchikov, 2016}), in such way
+that the translator ``sacrifice{[}s{]} fidelity to the image rather than
+the melodiousness of verse'' (\citeproc{ref-briusov1911ot}{Briusov,
+1911}). In the same line, Edith Grossman's translation of \emph{El
+Quijote} (\citeproc{ref-cervantes}{\textbf{cervantes?}}) present a
+similar case, in which formal aspects of the original text in Spanish
+are favoured to some degree over equivalence to English, even sometimes
+keeping Spanish words from the original text in the English translation
+(e.g., \emph{Señor}, \emph{ínsula}).
+
+The underlying mechanisms that lead to spontaneous homophonic
+translations are informative of the undeerlying mechanisms of native
+speech processing: how does the lexical system deal with the non-native
+speech signal, in which form-meaning associations are scarce, if not
+absent? This question remains relatively explored from a
+psycholinguistic standpoint. The present work was addressed at exploring
+the conditions under which non-native speech is likely to result in
+correct translations. To account for the fact that prior familiarity
+with the language (e.g., form-meaning associations) may contribute to
+correct translations, we tested participants in a language that they
+reported being unfamiliar with. In such conditions, participants can
+almost exlusively rely on the acoustic properties of the speech signal
+to provide correct translations.
+
+Our findings suggest that participants were able to exploit the
+phonological similarity between the presented words and their correct
+translation, even when both words shared less than 50\% of their
+phonological form (as measured by their Levenshtein distance). How the
+perceptual system deals with non-native phonemes is a vastly studied
+subject. The same cannot be said about how the perceptual system deals
+with non-native word-forms. We provide new experimental evidence
+suggesting that participants accommodate non-native phonological forms
+to their native phoneme inventory, leading to the activation of native
+lexical representations. One of the limitations of the present study is
+that Levenshtein similarity underestimates the actual phonetic
+similarity between the signal and the phonological form of the correct
+translations. Phonological transcriptions like X-SAMPA---which we used
+to calculate Levenshtein similarity----are abstract representations that
+ignore non-phonological contrasts (e.g., phones), and consider different
+symbols as completely different phonemes disregarding the fact that some
+phonemes share more phonetic features than others. Altogether, it is
+likely that participants exploited a more fine grained, raw source of
+information to provide correct answers.
+
+Our findings also suggest that, in line with previous studies of
+word-processing, phonological neighbourhood size interfered
+participants' ability to produce correct translations, even for words
+with high phonological similarity with their correct translation. We
+calculated a measure of cross-linguistic neighbourhood size by counting,
+for each presented word (in Catalan or Spanish in Experiments 1 and 3,
+Catalan in Experiment 2) the number of phonologically related words
+(with a Levenshtein distance of one) in the target language (English in
+Experiments 1 and 3, Spanish in Experiment 2). To account for the fact
+that competition between phonological neighbours is sensitive to lexical
+frequency (high-freuqency neighbours produce stronger interference
+effects), we only counted neighbours with a lexical freuqency higher
+than the correct translation. Using this measure, we found that, across
+the three experiments, participants ability to exploit phonological
+similarity to produce correct translations declined as the number of
+phonological neighbours increased.
+
+A comparison between results in Experiments 1 and 2 suggested that the
+performance of Spanish participants translating Catalan was more
+resilient to the interfering effects of phonological neighbourhood
+density than the performance of English participants translating Catalan
+or Spanish. As highlihted before, Catalan and Spanish (both Romance
+languages) are typologically closer to each other than to English (a
+Germanic language). Catalan and Spanish share a higher proportion of
+cognates (i.e., form-similar translation equivalents) than Catalan and
+English, or Spanish and English. Previous studies in bilinguals show
+that this kind of cross-linguistic similarity impacts lexical
+processing: words that share high phonological similarity with their
+translations are processed faster and more accurately than words sharing
+lower phonological similarity. This phenomenon has provided strong
+evidence in favour of the language-non selective hypothesis of bilingual
+lexical access, which states that bilinguals activate lexical
+representations in both languages, even during monolingual situations.
+This facilitation effect has been reported for comprehension
 (\citeproc{ref-dijkstra2010cross}{Dijkstra et al., 2010};
@@
 \citeproc{ref-valente2018does}{Valente et al., 2018}), and translation
-(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}).
+(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}). To
+our knowledge, this present study is the first to investigate the role
+of cross-linguistic phonological similarity in monolinguals. In
+particular, our participants were reportedly unfamiliar with the
+presented language. As such, they had to complete our translation
+elicitation task in the absence of form-meaning mappings, therefore
+relying exclusively on the phonological similarity between the presented
+words and their correct translations. This provides a convenient case to
+investigate the role of phonological similarity in lexical processing,
+disantangling the contribution of phonological similarity from the
+contribution of conceptual equivalence. Future studies might expand the
+methods used in the present study to elaborate in this issue.
+
+Two populations of speakers were tested in the present study: British
+English native adults living in the United Kindgom, and Spanish native
+adults living in Spain. Our findings suggest that the typological
+distance between the presented and target language is associated with a
+more robust facilitation effect of phonological similarity. Whether such
+association is linear (i.e., robustness of facilitation increases
+linearly with typological distance), proportional, or even extends
+beyond the specific pair of languages involved in the present work
+remains unclear. The generalisability of these findings can also be
+tested by increasing the repertoire of words involved in th translation
+elicitation task. In Experiments 1 to 3, participants answered to a
+maximum of 105 words. All words were of high-frequency, and had a low
+age-of-acquisition. In order to characterise listeners' ability to
+translate correctly words from an unfamiliar language, words with
+varying levels of difficulty should be included.
+
+In summary, the present paper provides insights into the
+psycholinguistic mechanisms underlying homophonic translation. English
+and Spanish native adults were tested in a translation elicitation task
+in which they had to guess the English or Spanish translation of a
+series of words in an unfamiliar language. Participants successfully
+exploited phonological similarity between the presented words and their
+correct translations to provide correct answers. Participants'
+performance in the task only benefited from phonological similarity when
+the presented word had few higher-frequency phonological neighbours in
+the target language. Finally, the faciltiation effect of phonological
+similarity was more robust in the Spanish native participants, who
+translated words from a typologically closer language than English
+participants. Overall, the findings presented in the present paper
+suggest that the processing of words in a non-native, unfamiliar
+language recruits mechanisms of lexical activation, selection, and
+interference parallel to those recruited by listening to words in a
+native language.
 
@@
 
+\bibitem[\citeproctext]{ref-briusov1911ot}
+Briusov, V. Y. (1911). \emph{Ot perevodchika}. Skorpion.
+
 \bibitem[\citeproctext]{ref-broersma2021praat}
@@
 
+\bibitem[\citeproctext]{ref-levick2016translating}
+Levick, T. (2016). Translating homophonic wordplay in patrick goujon's
+moi non: A case study. \emph{Sound/Writing: On Homophonic Translation}.
+
 \bibitem[\citeproctext]{ref-lotto1998effects}
@@
 \emph{Phonology}, \emph{25}(1), 129--164.
+
+\bibitem[\citeproctext]{ref-pilshchikov2016semiotics}
+Pilshchikov, I. (2016). The semiotics of phonetic translation.
+\emph{Studia Metrica Et Poetica}, \emph{3}(1), 53--104.
 

### 2024-10-16 manuscript/appendix.qmd
Fix figures

@@
 ```
+
+## Appendix 3: pooled analyses of Experiments 1 and 3
+
+Across Experiments 1 and 3, we found strong evidence that participants efficiently exploited phonological similarity to provide accurate translations for words in an unfamiliar language, provided that few phonological neighbours of higher lexical frequency were present.
+@fig-coefs summarizes the posterior distribution of the regression coefficients of the models in Experiments 1 to 3.
+
+```{r fig-coefs}
+#| label: fig-coefs
+#| fig-width: 9
+#| fig-height: 5
+var_labels <- c(
+  "b_Intercept" = "Intercept",
+  "b_freq_zipf_2" = "Frequency",
+  "b_lv" = "Similarity",
+  "b_neigh_n_h" = "CLPN",
+  "b_neigh_n_h:lv" = "Similarity × CLPN",
+  "b_group1" = "cat-ENG vs. spa-ENG"
+)
+
+list(exp_1_m0, exp_2_m0, exp_3_m1) |>
+  set_names(paste0("Experiment ", 1:3)) |>
+  map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |>
+  dplyr::filter(.variable != "b_Intercept") |>
+  mutate(
+    .value = .value / 4,
+    .variable = factor(.variable,
+      levels = names(var_labels),
+      labels = var_labels
+    )
+  ) |>
+  ggplot(aes(.value, reorder(exp, desc(exp)))) +
+  facet_wrap(~.variable, scales = "free_x") +
+  annotate(
+    geom = "rect",
+    xmin = -0.1 / 4,
+    xmax = 0.1 / 4,
+    ymin = -Inf,
+    ymax = Inf,
+    fill = "grey",
+    alpha = 3 / 4,
+    colour = NA
+  ) +
+  geom_vline(
+    xintercept = 0,
+    colour = "grey",
+    linewidth = 3 / 4
+  ) +
+  stat_slab(aes(fill = after_stat(level)),
+    point_interval = mode_hdi,
+    .width = c(0.95, 0.89, 0.67, 1),
+    outline_bars = TRUE,
+    colour = "white",
+    linewidth = 0.1,
+    breaks = 20,
+    normalize = "panels"
+  ) +
+  stat_spike(
+    at = "median",
+    size = 1,
+    aes(linetype = "Median"),
+    normalize = "panels"
+  ) +
+  labs(
+    x = "Estimate", y = "Experiment",
+    colour = "CrI",
+    fill = "CrI",
+    linetype = ""
+  ) +
+  scale_fill_brewer(palette = "Reds") +
+  theme(
+    legend.position = c(1, 0),
+    legend.box = "horizontal",
+    legend.justification = c(1, 0),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+```

### 2024-10-16 manuscript/manuscript.qmd
Fix figures

--- 2024-10-08 manuscript/manuscript.qmd
@@
 library(tidybayes)
+library(ggExtra)
+library(tinytable)
 
@@
 "empty ear"), after the *Soramimi Hour*---a section of the program *Tamori Club* hosted by comedian Morita Kazuyoshi (Tamori)---in which instances of this illusion were presented with comedy purposes.
-Soramimi is a particular case of *homophonic translation*: words or phrases in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning.
+Soramimi is a particular case of *homophonic translation*: words or phrases in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
 This phenomenon extends beyond song lyrics, and may occur anytime when listening to non-native speech.
@@
 Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
-This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
+This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing; @dufour2010phonological].
 Since the interference effects of phonological neighbourhood density has also been reported across languages [@weber2004lexical], we hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
@@
 Trials in which participants took longer than 10 seconds to respond were also excluded.
-Participants contributed a total of `r nrow(exp_responses)` valid trials (`r format(nrow(exp_responses[exp_responses$group %in% c("cat-ENG", "cat-SPA"),]), big.mark = ",")` in Catalan, `r format(nrow(exp_responses[exp_responses$group %in% c("spa-ENG"),]), big.mark = ",")` in Spanish).
+Participants contributed a total of `r format(nrow(exp_responses), big.mark = ",")` valid trials (`r format(nrow(exp_responses[exp_responses$group %in% c("cat-ENG", "cat-SPA"),]), big.mark = ",")` in Catalan, `r format(nrow(exp_responses[exp_responses$group %in% c("spa-ENG"),]), big.mark = ",")` in Spanish).
 The task took approximately 15 minutes to be completed.
@@
 \end{aligned}
-$$
-::: {eq-1}
+$$ {#eq-1}
+
+:::
 
 To test the practical relevance of each predictor we followed @kruschke2018bayesian.
-We first specified a region of practical equivalence (*ROPE*) around zero ([-0.1, +0.1], in the logit scale).
+We first specified a region of practical equivalence (ROPE) around zero ([-0.1, +0.1], in the logit scale).
 This area indicates the values of the regression coefficients that we considered as equivalent to zero.
@@
 Of those trials, `r format(nrow(dataset$cat_eng), big.mark = ",")` were provided by `r n_participants$cat_eng` unique participants who listened to Catalan words, and `r format(nrow(dataset$spa_eng), big.mark = ",")` trials were provided by `r n_participants$spa_eng` unique participants who listened to Spanish words.
-We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_eng["blank"]`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (e.g., `f`, *n* = `r r_validity$cat_eng["incomplete"]`), and in which participants added comments to the experimenter (e.g., `unsure`, *n* = `r r_validity$cat_eng["other"]`).
+We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_eng["blank"]`), in which a response in a language other than English was provided (e.g., "`agua`", *n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (e.g., "`f`", *n* = `r r_validity$cat_eng["incomplete"]`), and in which participants added comments to the experimenter (e.g., "`unsure`", *n* = `r r_validity$cat_eng["other"]`).
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
@@
 @fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
-@fig-coefs shows a graphic summary of the posterior distribution of the regression coefficients of interest.
 
@@
 We collected data for a total of `r format(nrow(dataset$cat_spa) + nrow(dataset$cat_spa), big.mark = ",")` trials completed by `r n_participants$cat_spa` unique participants.
-We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_spa["blank"]`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r r_validity$cat_spa["language"]`), in which participants did not provide a whole word (e.g., `f`, *n* = `r r_validity$cat_spa["incomplete"]`), and in which participants added comments to the experimenter (e.g., `unsure`, (*n* = `r r_validity$cat_spa["other"]`).
+We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_spa["blank"]`), in which a response in a language other than Spanish was provided (*n* = `r r_validity$cat_spa["language"]`), in which participants did not provide a whole word (*n* = `r r_validity$cat_spa["incomplete"]`), and in which participants added comments to the experimenter (*n* = `r r_validity$cat_spa["other"]`).
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a developmental language disorder (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
@@
 
-```{r}
+```{r posterior-2}
 c2 <- gather_draws(exp_2_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
@@
   bind_rows(.id = "exp") |>
+  mutate(translation = paste0())
   summarise(
@@
 tbl_data |>
-  gt(
-    rowname_col = "translation",
-    groupname_col = c("exp", "group")
-  ) |>
-  fmt_number(is.numeric, scale_by = 100) |>
-  cols_label(accuracy = "Accuracy (%)", accuracy_se = "SE")
+  select(translation, accuracy, accuracy_se) |> 
+  mutate(accuracy = round(accuracy * 100, 2),
+         accuracy_se = round(accuracy_se * 100, 2)) |> 
+  set_names(c("Translation", "Accuracy (%)", "SE")) |> 
+  tt() |> 
+  group_tt(i = list(
+    "Experiment 1 (cat-ENG)" = 5,
+    "Experiment 2 (spa-ENG)" = 21,
+    "Experiment 3 (cat-SPA)" = 26))
 ```
@@
 
-[ADD HERE LEXICAL FREQUENCY OF CATALAN AND SPANISH WORDS IN ENGLISH, AND OF CATALAN WORDS IN SPANISH]
-
 # Experiment 3
@@
 
-```{r}
+```{r participants-excluded-3}
 #| label: participants-excluded-3
@@
 #| label: dataset-3
+dataset <- split(quest_responses, quest_responses$group)
+names(dataset) <- janitor::make_clean_names(names(dataset))
+r_validity <- map(
+  dataset,
+  \(x) janitor::clean_names(table(x$response_type))
+)
+n_participants <- map(dataset, \(x) n_distinct(x$participant_id))
+
 dataset <- quest_responses |>
@@
   janitor::clean_names()
-
-dataset <- split(quest_responses, quest_responses$group)
-names(dataset) <- janitor::make_clean_names(names(dataset))
-r_validity <- map(
-  dataset,
-  \(x) janitor::clean_names(table(x$response_type))
-)
-n_participants <- map(dataset, \(x) n_distinct(x$participant_id))
 
@@
 Of those trials, `r format(nrow(dataset$cat_eng), big.mark = ",")` were provided by `r n_participants$cat_eng` unique participants who listened to Catalan words, and `r format(nrow(dataset$spa_eng), big.mark = ",")` trials were provided by `r n_participants$spa_eng` unique participants who listened to Spanish words.
-We excluded trials in which participants did not enter any text (*n* = `r dataset$cat_eng$blank`), in which a response in a language other than English was provided (e.g., `agua`, *n* = `r dataset$cat_eng$language`), in which participants did not provide a whole word (e.g., `f`, *n* = `r dataset$cat_eng$incomplete`), and in which participants added comments to the experimenter (e.g., `unsure`, (*n* = `r dataset$cat_eng$other`).
+We excluded trials in which participants did not enter any text (*n* = 0), in which a response in a language other than English was provided (*n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (*n* = 0), and in which participants added comments to the experimenter (*n* = 0).
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
@@
 
-```{r}
+```{r lengths-2}
 lengths <- dataset_3 |>
@@
 #| fig-height: 5
-#| fig-cap: "Catalan/Spanish prior word knowledge as reported by English native participants in Experiment 3.
-(A) Average proportion of participants that reported prior knowledge for words with surprisingly high accuracy (no phonological similarity with the correct translation, accuracy higher than 10%), and for words with expected accuracy (low similarity-low accuracy, or high similarity-high accuracy).
-(B) Average accuracy for words with expected and unexpected accuracy."
-plot_1_data <- dataset_3 |>
+#| fig-cap: "Catalan and Spanish prior word knowledge, as reported by English native participants in Experiment 3.
+The X-axis indicates the proportion of participants that reported prior knowledge with the word.
+The Y-axis indicates participants' accuracy (the proportion of correct responses by participants).
+Each colour indicates whether the data point corresponds to a word that participants answered to with expected accuracy (red) or with unexpectedly high accuracy (blue), as defined by (no phonological similarity with the correct translation, by accuracy higher than 10%)."
+p <- dataset_3 |>#| 
   mutate(surprise = translation %in% tbl_data$translation) |>
@@
     knowledge = mean(knowledge),
+    correct = sum(correct),
+    n = n(),
     .by = c(translation, group, surprise)
   ) |>
-  select(translation, group, surprise, knowledge) |>
-  mutate(surprise = if_else(
+  mutate(accuracy = prop_adj(correct, n),
+  surprise = if_else(
     surprise,
-    "Unexpectedly high accuracy\n(no phonological similarity)",
-    "Expected accuracy"
-  ))
-
-plot_1 <- plot_1_data |> 
-  filter(surprise=="Unexpectedly high accuracy\n(no phonological similarity)") |> 
-  ggplot(aes(1, knowledge, colour = group, fill = group)) +
-  facet_wrap(~group, ncol = 1) +
-  geom_dots(aes(group = NA), side = "both", shape = 1, layout = "swarm", stackratio = 1.5, dotsize = 1, binwidth = 0.05, show.legend = FALSE) +
-  geom_rect(
-    data = filter(plot_1_data, surprise=="Expected accuracy") |> 
-      summarise(
-      knowledge_mean = mean(knowledge), 
-      knowledge_std = sd(knowledge),
-      n = n(),
-      .by = group
-    ) |> 
-    mutate(knowledge_se = knowledge_std / sqrt(n)),
-   aes(xmin = 0.5, xmax = 1.5, 
-   ymin = knowledge_mean - knowledge_se, 
-   ymax = knowledge_mean + knowledge_se, 
-   fill = group, 
-   colour = group
-   ), 
-       inherit.aes = FALSE,
-   alpha = 1/4,
-   color = NA) +
-  geom_hline(
-    data = filter(plot_1_data, surprise=="Expected accuracy") |>
-      summarise(knowledge_mean = mean(knowledge), .by = group),
-    aes(yintercept = knowledge_mean, color = group)
-  ) +
-  labs(
-    y = "Prior knowledge reported",
-    x = "Group",
-    colour = "Group",
-    fill = "Group"
-  ) +
-  scale_y_continuous(labels = scales::percent) +
-  theme(axis.text.x = element_blank(),
-  axis.title.x = element_blank())
-
-plot_2_data <- dataset_3 |>
-  mutate(surprise = translation %in% tbl_data$translation) |>
-  mutate(surprise = if_else(
-    surprise,
-    "Unexpectedly high accuracy\n(no phonological similarity)",
+    "Unexpected high accuracy\n(no phonological similarity)",
     "Expected accuracy"
   )) |> 
-  summarise(
-    correct = sum(correct),
-    n = n(),
-    .by = c(translation, group, surprise, knowledge)
-  ) |>
-  mutate(accuracy = prop_adj(correct, n)) |>
-  select(translation, group, accuracy, surprise, knowledge)
-
-plot_2 <- plot_2_data |> 
-  ggplot(aes(1, accuracy, colour = group, fill = group)) +
-  facet_wrap(~group, ncol = 1) +
-geom_dots(aes(group = NA), side = "both", shape = 1, layout = "swarm", stackratio = 0.7, dotsize = 0.5, binwidth = 0.05, show.legend = FALSE) +
-  geom_rect(
-    data = filter(plot_2_data, surprise=="Expected accuracy") |> 
-      summarise(
-      knowledge_mean = mean(knowledge), 
-      knowledge_std = sd(knowledge),
-      n = n(),
-      .by = group
-    ) |> 
-    mutate(knowledge_se = knowledge_std / sqrt(n)),
-   aes(xmin = 0.5, xmax = 1.5, 
-   ymin = knowledge_mean - knowledge_se, 
-   ymax = knowledge_mean + knowledge_se, 
-   fill = group, 
-   colour = group
-   ), 
-       inherit.aes = FALSE,
-   alpha = 1/4,
-   color = NA) +
-  geom_hline(
-    data = filter(plot_2_data, surprise=="Expected accuracy") |>
-      summarise(knowledge_mean = mean(knowledge), .by = group),
-    aes(yintercept = knowledge_mean, color = group)
+  ggplot(aes(knowledge, accuracy, colour = surprise, fill = surprise, shape = surprise)) +
+  facet_wrap(~group) +
+  geom_point(stroke = 1, size = 2) +
+  #geom_rug(outside = TRUE, sides = "tr") +
+  labs(
+    x = "Reported knowledge",
+    y = "Accuracy"
   ) +
-  labs(
-    y = "Accuracy",
-    x = "Group",
-    colour = "Group",
-    fill = "Group"
-  ) +
-  scale_y_continuous(
-    labels = scales::percent
-  ) +
+  scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
+  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
+  scale_shape_manual(values = c(1, 2)) +
+  scale_colour_manual(values = c(
+    RColorBrewer::brewer.pal(3, "Reds")[3], 
+    RColorBrewer::brewer.pal(3, "Blues")[3]
+  )) +
   theme(
-    axis.title.x = element_blank(),
-    axis.text.x = element_blank(),
-  )
-
-plot_1 | plot_2 +
-  plot_layout(nrow = 1, guides = "collect") &
-  plot_annotation(tag_levels = "A") &
-  labs(
-    colour = "Group",
-    shape = "Group"
-  ) &
-  theme(
-    panel.border = element_rect(
-      fill = NA,
-      colour = "grey",
-      linewidth = 1
-    ),
-    axis.line = element_blank(),
-    axis.text.x = element_blank(),
-    axis.title.x = element_blank(),
-    axis.ticks.x = element_blank(),
-    panel.grid.major.y = element_line(
-      colour = "grey",
-      linetype = "dotted"
-    ),
-    panel.grid.major.x = element_blank()
-  )
-```
-
-```{r}
+    legend.position = "top",
+    legend.title = element_blank(),
+  )
+```
+
+```{r draws-3}
 c3 <- gather_draws(exp_3_m1, `b_.*`, `sd_.*`, regex = TRUE) |>
@@
 Lines indicate mean predictions, and intervals indicate 95%, 89%, 78%, 67%, and 50% credible intervals (CrI)."
-#| fig-height: 6
+#| fig-height: 5
 #| fig-width: 10
@@
   theme(
-    legend.position = "top",
+    legend.position = "none",
     legend.box = "vertical"
@@
 ## Discussion
-
-# Pooled analyses of Experiments 1 and 3
-
-Across Experiments 1 and 3, we found strong evidence that participants efficiently exploited phonological similarity to provide accurate translations for words in an unfamiliar language, provided that few phonological neighbours of higher lexical frequency were present.
-@fig-coefs summarizes the posterior distribution of the regression coefficients of the models in Experiments 1 to 3.
-
-```{r fig-coefs}
-#| label: fig-coefs
-#| fig-width: 9
-#| fig-height: 5
-var_labels <- c(
-  "b_Intercept" = "Intercept",
-  "b_freq_zipf_2" = "Frequency",
-  "b_lv" = "Similarity",
-  "b_neigh_n_h" = "CLPN",
-  "b_neigh_n_h:lv" = "Similarity × CLPN",
-  "b_group1" = "cat-ENG vs. spa-ENG"
-)
-
-list(exp_1_m0, exp_2_m0, exp_3_m1) |>
-  set_names(paste0("Experiment ", 1:3)) |>
-  map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |>
-  dplyr::filter(.variable != "b_Intercept") |>
-  mutate(
-    .value = .value / 4,
-    .variable = factor(.variable,
-      levels = names(var_labels),
-      labels = var_labels
-    )
-  ) |>
-  ggplot(aes(.value, reorder(exp, desc(exp)))) +
-  facet_wrap(~.variable, scales = "free_x") +
-  annotate(
-    geom = "rect",
-    xmin = -0.1 / 4,
-    xmax = 0.1 / 4,
-    ymin = -Inf,
-    ymax = Inf,
-    fill = "grey",
-    alpha = 3 / 4,
-    colour = NA
-  ) +
-  geom_vline(
-    xintercept = 0,
-    colour = "grey",
-    linewidth = 3 / 4
-  ) +
-  stat_slab(aes(fill = after_stat(level)),
-    point_interval = mode_hdi,
-    .width = c(0.95, 0.89, 0.67, 1),
-    outline_bars = TRUE,
-    colour = "white",
-    linewidth = 0.1,
-    breaks = 20,
-    normalize = "panels"
-  ) +
-  stat_spike(
-    at = "median",
-    size = 1,
-    aes(linetype = "Median"),
-    normalize = "panels"
-  ) +
-  labs(
-    x = "Estimate", y = "Experiment",
-    colour = "CrI",
-    fill = "CrI",
-    linetype = ""
-  ) +
-  scale_fill_brewer(palette = "Reds") +
-  theme(
-    legend.position = c(1, 0),
-    legend.box = "horizontal",
-    legend.justification = c(1, 0),
-    axis.title.x = element_blank(),
-    axis.title.y = element_blank()
-  )
-```
 
@@
 
-The present work investigated the dynamics of lexical activation and selection dring non-native speech listening.
-Using a translation elicitation task, we collected data from a group of English native adults (Experiment 1) and a group of Spanish native adults (Experiment 2).
-Neither group was familiar with Catalan, Spanish, or any other Romance language.
-In each trial, participants were presented with Spanish words (Experiment 1) or Catalan words (Experiments 1 and 2), and then asked to type their best-guess translation for the presented word in their native language.
-Participants provided correct translations more often if the presented word and its translation shared phonological similarity, only if the number of phonological neighbours around the presented word was small.
-These results suggest that listening to words in an unfamiliar language can trigger dynamics of lexical activation and selection similar to those triggered by listening to speech in the native language.
-
-In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
-One group was presented with Catalan words, the other with Spanish words.
-We examined the extent to which participants were able to use the phonological similarity between the presented word (in Catalan or Spanish) and its target translation to provide accurate responses, despite reporting no prior relevant experience with the Catalan or Spanish.
-In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are two Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant's performance.
-Some participants in Experiments 1 and 2 were surprisingly good at translating a subset of words which had low phonological similarity with their correct translation.
-We were concerned that this may be caused by some prior knowledge of specific words by our participants, as some Spanish words are commonly seen in media or product labels, making them familiar even to monolingual speakers of English.
-In Experiment 3, we collected additional data from a new group of British English natives.
-The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
-After removing responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we found similar resutls as in Experiment 1, suggesting that participants did not simply rely on prior familiarity with the word-forms presented but rather exclusively on their phonological similarity with the correct English translation.
-
-Our findings provide insights into the mechanisms underlying homophonic translations.
-This phenomenon is frequently experienced by listeners of foreign speech, when strings of sounds in the input are perceived as strings of words or phrases in the native language, without necessarily preserving the meaning.
-
-This type of cross-linguistic similarity has also been extremely informative for the study of bilingual lexical processing: words that share high form-similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
+The present work explored the psycholinguistic bases of homophonic translation, a phenomenon in which listening to speech in a non-native---perhaps unfamiliar---language leads to the activation of lexical representations in the native language, without necessarily preserving the meaning.
+We investigated how phonological similarity and its interaction with phonological neighbourhood density impact the dynamics of lexical activation and selection during non-native word processing.
+We designed a translation elicitation task in which participants listened to individual words in a non-native, unfamiliar language.
+After listening to each word, participants were asked to provide their best-guess translation for each word in their native language.
+We run three experiments.
+In Experiment 1, British English-native adults listened to a list of words in Catalan or Spanish.
+Participants reported no prior familiarity with Catalan, Spanish, or any other Romance language, yet provided accurate translations for words that shared some degree of phonological similarity with their correct English translation.
+Participants only benefited from phonological similarity when the number of English phonological neighbours of the presented Catalan or Spanish word was small.
+Experiment 2 was aimed at extending our findings to a population of participants whose native language was typologically closer to the unfamiliar language presented.
+We tested a group of Spanish native adults who listened to a list of Catalan words.
+Again, participants reported no prior familiarity with Catalan, or any other Romance language other than Spanish.
+In line with the results in English natives, we found an interaction between phonological similarity and phonological neighbourhood size.
+The facilitation effect of phonological similarity in the Spanish participants was more robust to the interferring effect of phonological neighbourhood density: Spanish participants provided correct translations for words with high neighbourhood densities more often than English participants.
+We suspected that participants in Experiments 1 and 2 might have managed to provide correct translations for some low-phonological similarity words thanks to having prior experience with those words, despite having reported being unfamiliar with Catalan, Spanish, or any other Romance language.
+Experiment 3 was a replica of Experiment 1, in which we collected additional information about participants' prior familiarity with the presented Catalan and Spanish words.
+The design of Experiment 3 was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
+After removing responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we run the same analyses as in Experiment 1, finding parallel results.
+Results in Experiment 3 suggest participants in the present study did not simply rely on prior familiarity with some of the word-forms presented.
+Rather, participants relied on the phonological similarity between the presented Catalan and Spanish words and their correct translation.
+
+Homophonic translation is an often spontaneous phenomenon that occurs commonly when listening to non-native speech, and is prevalent across many cultures and languages [@dembeck2015oberflachenubersetzung; @efimova2018homophonic].
+The most famous case of homophonic translation is *Soramimi*, which refers to homophonic translation in the context of music listening, when particular sections of the lyrics of a song in a non-native language make listeners think of particular words and phrases in their own native language.
+This sometimes result in a completely different meaning than thhe originally intended.
+Homophonic translation occurs outside the context of music.
+For instance, it is often intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
+For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
+Although homophonic translation has mostly received attention as a perceptual curiosity or as a literary figure, little attention has been paid to homophonic translation as a spontanous *psycholinguistic* event.
+The psycholiguistic mechanisms behind spontaneous homophonic translations are informative of the mechanisms underlying native speech processing.
+How does the lexical system deal with a non-native speech signal in which form-meaning associations are scarce, if not absent?
+
+How the perceptual system deals with non-native speech is a vastly studied subject [e.g., @best1988examination; @miyawaki1975effect; @pallier2001influence; @weber2004lexical; @van2015learning; see @baese2020perception for review].
+The same cannot be said about how the perceptual system deals with speech in an *unfamiliar* language.
+Previous research had explored how bilinguals recognise, translate, and produce words as a function of their proficiency in each of their languages [e.g., @kroll1994category; @dijkstra2019multilink; @dijkstra2002architecture; @christoffels2006memory].
+The case of a very low-proficiency bilingual may be close to that of a listener of an unfamiliar language, but being bilingual implies some form of proficiency in both languages, such as some knowledge about the meaning of some words.
+Even low-proficiency bilinguals are likely to have already acquired many form-meaning mappings in their lower proficiency language.
+Characterising the impact of phonological similarity on word processing is therefore challenging in bilinguals, due to the simultaneous contribution of phonological *and* conceptual links in the lexicon.
+By testing participants with *no* prior familiarity with the presented language (i.e., no proficiency), we ensured that no form-meaning associations were available for participants to correctly translate the unfamiliar words, if not by relying excluusively on phonological similarity.
+
+We calculated phonological similarity between word pairs as the Levenshtein similarity between the X-SAMPA transcription of their phonological forms [see @floccia2018introduction for a similar approach].
+Using this measure, we found that participants were able to exploit the phonological similarity between the presented words and their correct translation, even when both words-forms shared few phonemes.
+This points to listeners accommodating non-native phonological forms to their native phoneme inventory, resulting in the activation of native lexical representations.
+Future studies may consider comparing the suitability of other measures of phonological similarity that take into account finer-grained phonetic or prosodic cues present in the acoustic signal presented to participants.
+Phonological transcriptions like X-SAMPA are abstract representations that ignore non-phonological contrasts (e.g., phones), and consider different symbols as completely different phonemes, disregarding the fact that some phonemes share more phonetic features than others.
+Additionally, prosodic cues such as lexical stress, which our measure of phonological similarity does not include, might also provide participants further information about the correct translation of the presented word-forms.
+Altogether, it is likely that participants in the present study were able to exploit additional cues in the acoustic signal to provide correct translations.
+Future studies may explore the relative contribution of such cues in the occurrence of homophonic translation.
+Overall, the present study provides some insights into the role of phonological similarity in the auditory presentation of words in an unfamiliar language.
+
+Our findings also suggest that the facilitation effect of phonological familiarity was moderated by phonological neighbourghood density.
+This is in line which previous studies suggesting that the presence of (high-frequency) phonological neighbours interferes with the lexical selection in word comprehension tasks [e.g., @dufour2010phonological; @luce1998recognizing; @vitevitch1999probabilistic; @vitevitch1998words; @grainger1990word].
+Our translation elicitation task can be understood as a particular instance of an auditory word recognition task, in which the presented unfamiliar words have the potential to activate phonological neighbours in participants' lexicon.
+We calculated a measure of cross-linguistic neighbourhood size by counting, for each presented word (in Catalan or Spanish in Experiments 1 and 3, Catalan in Experiment 2) the number of phonologically related words (with a Levenshtein distance of one) in the target language (English in Experiments 1 and 3, Spanish in Experiment 2).
+To account for the fact that competition between phonological neighbours is sensitive to lexical frequency [high-frequency neighbours produce stronger interference effects, @luce1998recognizing; @dufour2010phonological], we only counted neighbours with a lexical frequency higher than the correct translation.
+Using this measure, we found that participants' ability to exploit phonological similarity to produce correct translations declined as the number of phonological neighbours increased.
+Overall, this suggests that listening to words in an unfamiliar language trigger the same mechanisms of lexical activation and selection that listening to words in a native language does.
+
+A comparison between results in Experiments 1 and 2 suggested that the performance of Spanish participants translating Catalan was more resilient to the interfering effects of phonological neighbourhood density than the performance of English participants translating Catalan or Spanish.
+As highlighted before, Catalan and Spanish (both Romance languages) are typologically closer to each other than to English (a Germanic language).
+Catalan and Spanish share a higher proportion of cognates (i.e., form-similar translation equivalents) than Catalan and English, or Spanish and English.
+Previous studies in bilinguals show that this kind of cross-linguistic similarity impacts lexical processing: words that share high phonological similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
 This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
 This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
+To our knowledge, this present study is the first to investigate the role of cross-linguistic phonological similarity in a monolingual population, in which the role of conceptual links in the lexicon is ruled out, given participants lack of familiarity with the presented language.
+This provided a convenient case to investigate the role of phonological similarity in lexical processing, disantangling the contribution of phonological similarity from the contribution of conceptual equivalence.
+Overall, our findings suggest that the typological distance between the presented and target language is associated with a more robust facilitation effect of phonological similarity.
+
+The specific mechanisms behind this effect are unclear.
+One possibility is that, as aforementioned, Catalan provided additional cues to Spanish participants in Experiment 2 compared to English participants in Experiment 1.
+For instance, it is likely that subphonemic cues not accounted by our measure of phonological similarity were more informative to Spanish participants than to English participants.
+In addition, it remains to be tested whether such association is linear (i.e., robustness of facilitation increases linearly with typological distance), proportional, or even extends beyond the specific pair of languages involved in the present work remains unclear.
+Do other close or distant language pairs show the same pattern of results?
+The generalisability of these findings can also be tested by increasing the repertoire of words involved in the translation elicitation task.
+In Experiments 1 to 3, participants answered to a maximum of 105 words.
+All words were of high-frequency, and had a low age-of-acquisition.
+In order to characterise listeners' ability to translate correctly words from an unfamiliar language, words with varying levels of difficulty should be included.
+
+In summary, the present paper provides insights into the psycholinguistic mechanisms underlying homophonic translation.
+English and Spanish native adults were tested in a translation elicitation task in which they had to guess the English or Spanish translation of a series of words in an unfamiliar language.
+Participants successfully exploited phonological similarity between the presented words and their correct translations to provide correct answers.
+Participants' performance in the task only benefited from phonological similarity when the presented word had few higher-frequency phonological neighbours in the target language.
+Finally, the facilitation effect of phonological similarity was more robust in the Spanish native participants, who translated words from a typologically closer language than English participants.
+Overall, the findings presented in the present paper suggest that the processing of words in a non-native, unfamiliar language recruits mechanisms of lexical activation, selection, and interference parallel to those recruited by listening to words in a native language.
 

### 2024-10-16 manuscript/manuscript.tex
Fix figures

--- 2024-10-08 manuscript/manuscript.tex
@@
 \usepackage{array}
+\usepackage{float}
+\usepackage{tabularray}
+\usepackage[normalem]{ulem}
+\usepackage{graphicx}
+\UseTblrLibrary{booktabs}
+\UseTblrLibrary{rotating}
+\UseTblrLibrary{siunitx}
+\NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
+\newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
+\newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
 \usepackage{tipa}
@@
 similar-sounding words and phrases in a different language, without
-necessarily preserving the meaning. This phenomenon extends beyond song
-lyrics, and may occur anytime when listening to non-native speech. For
-instance, when listening to ``You know me?'' (\textipa{/ju: n@U mi:}/)
-in English, Japanese-native listeners tend to understand \emph{yunomi}
-(/\textipa{jUnomi}/), which is a type of Japanese teacup. Overall, the
-occurrence of Soramimi---and more generally, homophonic
-translations---suggests that speech in an unfamiliar language has the
-potential to activate lexical representations in the native language.
+necessarily preserving the meaning
+(\citeproc{ref-gasparov2006semen}{Gasparov, 2006}). This phenomenon
+extends beyond song lyrics, and may occur anytime when listening to
+non-native speech. For instance, when listening to ``You know me?''
+(\textipa{/ju: n@U mi:}/) in English, Japanese-native listeners tend to
+understand \emph{yunomi} (/\textipa{jUnomi}/), which is a type of
+Japanese teacup. Overall, the occurrence of Soramimi---and more
+generally, homophonic translations---suggests that speech in an
+unfamiliar language has the potential to activate lexical
+representations in the native language.
 
@@
 analysis. Trials in which participants took longer than 10 seconds to
-respond were also excluded. Participants contributed a total of 9152
+respond were also excluded. Participants contributed a total of 9,152
 valid trials (5,694 in Catalan, 3,458 in Spanish). The task took
@@
 
-\[
+\begin{equation}\phantomsection\label{eq-1}{
 \begin{aligned}
@@
 \end{aligned}
-\] ::: \{eq-1\}
+}\end{equation}
 
@@
 \& Liddell (\citeproc{ref-kruschke2018bayesian}{2018}). We first
-specified a region of practical equivalence (\emph{ROPE}) around zero
-({[}-0.1, +0.1{]}, in the logit scale). This area indicates the values
-of the regression coefficients that we considered as equivalent to zero.
-We then computed the 95\% posterior credible intervals (CrI) of the
+specified a region of practical equivalence (ROPE) around zero ({[}-0.1,
++0.1{]}, in the logit scale). This area indicates the values of the
+regression coefficients that we considered as equivalent to zero. We
+then computed the 95\% posterior credible intervals (CrI) of the
 regression coefficients of interest. Finally, we calculated the
@@
 72), in which a response in a language other than English was provided
-(e.g., \texttt{agua}, \emph{n} = 51), in which participants did not
-provide a whole word (e.g., \texttt{f}, \emph{n} = 5), and in which
-participants added comments to the experimenter (e.g., \texttt{unsure},
-\emph{n} = 13). In addition, we excluded data from participants that
-self-rated their oral and/or written skills in Catalan and Spanish, or
-any other second language as four or higher in a five-point scale
-(\emph{n} = 2), were diagnosed with a language (\emph{n} = 1), or did
-not contribute more than 80\% of valid trials (\emph{n} = 9).
+(e.g., ``\texttt{agua}'', \emph{n} = 51), in which participants did not
+provide a whole word (e.g., ``\texttt{f}'', \emph{n} = 5), and in which
+participants added comments to the experimenter (e.g.,
+``\texttt{unsure}'', \emph{n} = 13). In addition, we excluded data from
+participants that self-rated their oral and/or written skills in Catalan
+and Spanish, or any other second language as four or higher in a
+five-point scale (\emph{n} = 2), were diagnosed with a language
+(\emph{n} = 1), or did not contribute more than 80\% of valid trials
+(\emph{n} = 9).
 
@@
 average predictions of the model for words with different values of
-\emph{Similarity} and \emph{CLPN}. Figure~\ref{fig-coefs} shows a
-graphic summary of the posterior distribution of the regression
-coefficients of interest.
+\emph{Similarity} and \emph{CLPN}.
 
@@
 text (\emph{n} = 44), in which a response in a language other than
-English was provided (e.g., \texttt{agua}, \emph{n} = 51), in which
-participants did not provide a whole word (e.g., \texttt{f}, \emph{n} =
-7), and in which participants added comments to the experimenter (e.g.,
-\texttt{unsure}, (\emph{n} = 1). In addition, we excluded data from
-participants that self-rated their oral and/or written skills in Catalan
-and Spanish, or any other second language as four or higher in a
-five-point scale (\emph{n} = 22), were diagnosed with a developmental
-language disorder (\emph{n} = 1), or did not contribute more than 80\%
-of valid trials (\emph{n} = 9). After applying trial-level and
-participant-level inclusion criteria, the resulting dataset included
-1,662 trials provided by 42 unique participants. Of those trials, 1,662
-were provided by 21 unique participants who listened to Catalan words,
-and 1,662 trials were provided by 21 unique participants who listened to
-Spanish words. Responses given by participants were 5.6 characters long
-on average (\emph{Median} = 5, \emph{SD} = 1.6, Range = 2-12).
+Spanish was provided (\emph{n} = 51), in which participants did not
+provide a whole word (\emph{n} = 7), and in which participants added
+comments to the experimenter (\emph{n} = 1). In addition, we excluded
+data from participants that self-rated their oral and/or written skills
+in Catalan and Spanish, or any other second language as four or higher
+in a five-point scale (\emph{n} = 22), were diagnosed with a
+developmental language disorder (\emph{n} = 1), or did not contribute
+more than 80\% of valid trials (\emph{n} = 9). After applying
+trial-level and participant-level inclusion criteria, the resulting
+dataset included 1,662 trials provided by 42 unique participants. Of
+those trials, 1,662 were provided by 21 unique participants who listened
+to Catalan words, and 1,662 trials were provided by 21 unique
+participants who listened to Spanish words. Responses given by
+participants were 5.6 characters long on average (\emph{Median} = 5,
+\emph{SD} = 1.6, Range = 2-12).
 
@@
 
-\begin{longtable}{l|rr}
+\begin{table}
 
@@
 
-\tabularnewline
-
+\centering{
+
+\centering
+\begin{tblr}[         %% tabularray outer open
+]                     %% tabularray outer close
+{                     %% tabularray inner open
+colspec={Q[]Q[]Q[]},
+cell{6}{1}={c=3}{},cell{23}{1}={c=3}{},cell{29}{1}={c=3}{},
+cell{2}{1}={preto={\hspace{1em}}},
+cell{3}{1}={preto={\hspace{1em}}},
+cell{4}{1}={preto={\hspace{1em}}},
+cell{5}{1}={preto={\hspace{1em}}},
+cell{7}{1}={preto={\hspace{1em}}},
+cell{8}{1}={preto={\hspace{1em}}},
+cell{9}{1}={preto={\hspace{1em}}},
+cell{10}{1}={preto={\hspace{1em}}},
+cell{11}{1}={preto={\hspace{1em}}},
+cell{12}{1}={preto={\hspace{1em}}},
+cell{13}{1}={preto={\hspace{1em}}},
+cell{14}{1}={preto={\hspace{1em}}},
+cell{15}{1}={preto={\hspace{1em}}},
+cell{16}{1}={preto={\hspace{1em}}},
+cell{17}{1}={preto={\hspace{1em}}},
+cell{18}{1}={preto={\hspace{1em}}},
+cell{19}{1}={preto={\hspace{1em}}},
+cell{20}{1}={preto={\hspace{1em}}},
+cell{21}{1}={preto={\hspace{1em}}},
+cell{22}{1}={preto={\hspace{1em}}},
+cell{24}{1}={preto={\hspace{1em}}},
+cell{25}{1}={preto={\hspace{1em}}},
+cell{26}{1}={preto={\hspace{1em}}},
+cell{27}{1}={preto={\hspace{1em}}},
+cell{28}{1}={preto={\hspace{1em}}},
+cell{30}{1}={preto={\hspace{1em}}},
+cell{31}{1}={preto={\hspace{1em}}},
+}                     %% tabularray inner close
 \toprule
-\multicolumn{1}{l}{} & Accuracy (\%) & SE \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{3}{l}{Experiment 1 - cat-ENG} \\ 
-\midrule\addlinespace[2.5pt]
-cavall /kəβaʎ/ - horse /hɔs/ & $17.14$ & $6.37$ \\ 
-llibre /ʎiβɾə/ - book /bʊk/ & $17.14$ & $6.37$ \\ 
-camisa /kəmizə/ - shirt /ʃɜt/ & $16.67$ & $6.21$ \\ 
-poma /poma/ - apple /æpl/ & $16.67$ & $6.21$ \\ 
-cama /kamə/ - leg /lɛg/ & $11.11$ & $5.24$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{3}{l}{Experiment 1 - spa-ENG} \\ 
-\midrule\addlinespace[2.5pt]
-pantalón /paŋtalon/ - trousers /traʊzəz/ & $77.42$ & $7.51$ \\ 
-naranja /naɾaŋxa/ - orange /ɒrɪnʤ/ & $41.94$ & $8.86$ \\ 
-leche /leʧe/ - milk /mɪlk/ & $35.48$ & $8.59$ \\ 
-toro /toɾo/ - bull /bʊl/ & $33.33$ & $8.61$ \\ 
-libro /liβɾo/ - book /bʊk/ & $30.00$ & $8.37$ \\ 
-cebra /θebɾa/ - zebra /zibrə/ & $29.03$ & $8.15$ \\ 
-pan /pan/ - bread /brɛd/ & $29.03$ & $8.15$ \\ 
-pollo /poʎo/ - chicken /ʧɪkɪn/ & $26.67$ & $8.07$ \\ 
-jirafa /xiɾafa/ - giraffe /ʤɪrɑf/ & $20.69$ & $7.52$ \\ 
-perro /pero/ - dog /dɒɡ/ & $16.13$ & $6.61$ \\ 
-pluma /pluma/ - feather /fɛðə/ & $16.13$ & $6.61$ \\ 
-puerta /pwerta/ - door /dɔ/ & $16.13$ & $6.61$ \\ 
-pie /pje/ - foot /fʊt/ & $12.90$ & $6.02$ \\ 
-caballo /kaβaʎo/ - horse /hɔs/ & $10.34$ & $5.66$ \\ 
-bocadillo /bokadiʎo/ - sandwich /sænwɪʤ/ & $10.00$ & $5.48$ \\ 
-globo /ɡloβo/ - balloon /bəlun/ & $10.00$ & $5.48$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{3}{l}{Experiment 2 - cat-SPA} \\ 
-\midrule\addlinespace[2.5pt]
-fulla /fuʎə/ - hoja /oxa/ & $30.43$ & $9.59$ \\ 
-ull /uʎ/ - ojo /oxo/ & $21.74$ & $8.60$ \\ 
-got /gɔt/ - vaso /baso/ & $20.00$ & $8.00$ \\ 
-entrepà /entɾəpa/ - bocadillo /bokadiʎo/ & $13.04$ & $7.02$ \\ 
-mirall /miɾaʎ/ - espejo /espexo/ & $12.50$ & $6.75$ \\ 
+Translation & Accuracy (%) & SE \\ \midrule %% TinyTableHeader
+cavall /kəβaʎ/ - horse /hɔs/             & 17.14 & 6.37 \\
+llibre /ʎiβɾə/ - book /bʊk/              & 17.14 & 6.37 \\
+camisa /kəmizə/ - shirt /ʃɜt/            & 16.67 & 6.21 \\
+poma /poma/ - apple /æpl/                & 16.67 & 6.21 \\
+Experiment 1 (cat-ENG) && \\
+cama /kamə/ - leg /lɛg/                  & 11.11 & 5.24 \\
+pantalón /paŋtalon/ - trousers /traʊzəz/ & 77.42 & 7.51 \\
+naranja /naɾaŋxa/ - orange /ɒrɪnʤ/       & 41.94 & 8.86 \\
+leche /leʧe/ - milk /mɪlk/               & 35.48 & 8.59 \\
+toro /toɾo/ - bull /bʊl/                 & 33.33 & 8.61 \\
+libro /liβɾo/ - book /bʊk/               & 30.00 & 8.37 \\
+cebra /θebɾa/ - zebra /zibrə/            & 29.03 & 8.15 \\
+pan /pan/ - bread /brɛd/                 & 29.03 & 8.15 \\
+pollo /poʎo/ - chicken /ʧɪkɪn/           & 26.67 & 8.07 \\
+jirafa /xiɾafa/ - giraffe /ʤɪrɑf/        & 20.69 & 7.52 \\
+perro /pero/ - dog /dɒɡ/                 & 16.13 & 6.61 \\
+pluma /pluma/ - feather /fɛðə/           & 16.13 & 6.61 \\
+puerta /pwerta/ - door /dɔ/              & 16.13 & 6.61 \\
+pie /pje/ - foot /fʊt/                   & 12.90 & 6.02 \\
+caballo /kaβaʎo/ - horse /hɔs/           & 10.34 & 5.66 \\
+bocadillo /bokadiʎo/ - sandwich /sænwɪʤ/ & 10.00 & 5.48 \\
+Experiment 2 (spa-ENG) && \\
+globo /ɡloβo/ - balloon /bəlun/          & 10.00 & 5.48 \\
+fulla /fuʎə/ - hoja /oxa/                & 30.43 & 9.59 \\
+ull /uʎ/ - ojo /oxo/                     & 21.74 & 8.60 \\
+got /gɔt/ - vaso /baso/                  & 20.00 & 8.00 \\
+entrepà /entɾəpa/ - bocadillo /bokadiʎo/ & 13.04 & 7.02 \\
+Experiment 3 (cat-SPA) && \\
+mirall /miɾaʎ/ - espejo /espexo/         & 12.50 & 6.75 \\
 \bottomrule
-
-\end{longtable}
+\end{tblr}
+
+}
+
+\end{table}%
 
@@
 
-{[}ADD HERE LEXICAL FREQUENCY OF CATALAN AND SPANISH WORDS IN ENGLISH,
-AND OF CATALAN WORDS IN SPANISH{]}
-
 \section{Experiment 3}\label{experiment-3}
@@
 We collected data for a total of 6,016 trials completed by 64 unique
-participants. Of those trials, 2,752 were provided by 32 unique
-participants who listened to Catalan words, and 3,264 trials were
+participants. Of those trials, 3,264 were provided by 32 unique
+participants who listened to Catalan words, and 2,752 trials were
 provided by 32 unique participants who listened to Spanish words. We
 excluded trials in which participants did not enter any text (\emph{n} =
-), in which a response in a language other than English was provided
-(e.g., \texttt{agua}, \emph{n} = ), in which participants did not
-provide a whole word (e.g., \texttt{f}, \emph{n} = ), and in which
-participants added comments to the experimenter (e.g., \texttt{unsure},
-(\emph{n} = ). In addition, we excluded data from participants that
-self-rated their oral and/or written skills in Catalan and Spanish, or
-any other second language as four or higher in a five-point scale
-(\emph{n} = 2), were diagnosed with a language (\emph{n} = 0), or did
-not contribute more than 80\% of valid trials (\emph{n} = 1).
+0), in which a response in a language other than English was provided
+(\emph{n} = 9), in which participants did not provide a whole word
+(\emph{n} = 0), and in which participants added comments to the
+experimenter (\emph{n} = 0). In addition, we excluded data from
+participants that self-rated their oral and/or written skills in Catalan
+and Spanish, or any other second language as four or higher in a
+five-point scale (\emph{n} = 2), were diagnosed with a language
+(\emph{n} = 0), or did not contribute more than 80\% of valid trials
+(\emph{n} = 1).
 
@@
 Experiment 2) than for words with expected accuracy (see
-Figure~\ref{fig-knowledge}). Participants reported prior knowledge of
+\textbf{?@fig-knowledge}). Participants reported prior knowledge of
 Catalan words with unexpected accuracy as often as those with expected
@@
 knowledge.
-
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-knowledge-1.pdf}}
-
-}
-
-\caption{\label{fig-knowledge}Catalan/Spanish prior word knowledge as
-reported by English native participants in Experiment 3. (A) Average
-proportion of participants that reported prior knowledge for words with
-surprisingly high accuracy (no phonological similarity with the correct
-translation, accuracy higher than 10\%), and for words with expected
-accuracy (low similarity-low accuracy, or high similarity-high
-accuracy). (B) Average accuracy for words with expected and unexpected
-accuracy.}
-
-\end{figure}%
 
@@
 \subsection{Discussion}\label{discussion-2}
-
-\section{Pooled analyses of Experiments 1 and
-3}\label{pooled-analyses-of-experiments-1-and-3}
-
-Across Experiments 1 and 3, we found strong evidence that participants
-efficiently exploited phonological similarity to provide accurate
-translations for words in an unfamiliar language, provided that few
-phonological neighbours of higher lexical frequency were present.
-Figure~\ref{fig-coefs} summarizes the posterior distribution of the
-regression coefficients of the models in Experiments 1 to 3.
-
-\begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-coefs-1.pdf}}
-
-}
-
-\caption{\label{fig-coefs}}
-
-\end{figure}%
 
@@
 
-The present work investigated the dynamics of lexical activation and
-selection dring non-native speech listening. Using a translation
-elicitation task, we collected data from a group of English native
-adults (Experiment 1) and a group of Spanish native adults (Experiment
-2). Neither group was familiar with Catalan, Spanish, or any other
-Romance language. In each trial, participants were presented with
-Spanish words (Experiment 1) or Catalan words (Experiments 1 and 2), and
-then asked to type their best-guess translation for the presented word
-in their native language. Participants provided correct translations
-more often if the presented word and its translation shared phonological
-similarity, only if the number of phonological neighbours around the
-presented word was small. These results suggest that listening to words
-in an unfamiliar language can trigger dynamics of lexical activation and
-selection similar to those triggered by listening to speech in the
-native language.
-
-In Experiment 1, we collected data from two groups of British English
-natives living in the United Kingdom. One group was presented with
-Catalan words, the other with Spanish words. We examined the extent to
-which participants were able to use the phonological similarity between
-the presented word (in Catalan or Spanish) and its target translation to
-provide accurate responses, despite reporting no prior relevant
-experience with the Catalan or Spanish. In Experiment 2, we tested a
-group of (European) Spanish natives in the same task, who were presented
-with Catalan words. Catalan and Spanish are two Romance languages whose
-close typological distance is reflected in the fact that they share many
-cognates, where English is a Germanic language that shares considerably
-fewer cognates with Catalan and Spanish. By testing participants
-translating words from typologically close or distant languages, we
-expected to widen the range of the phonological similarity scores of the
-translation pairs involved in the experimental task, therefore allowing
-us to explore potential cross-language differences in participant's
-performance. Some participants in Experiments 1 and 2 were surprisingly
-good at translating a subset of words which had low phonological
-similarity with their correct translation. We were concerned that this
-may be caused by some prior knowledge of specific words by our
-participants, as some Spanish words are commonly seen in media or
-product labels, making them familiar even to monolingual speakers of
-English. In Experiment 3, we collected additional data from a new group
-of British English natives. The design was closely modelled after
-Experiment 1, except that after providing their response in each trial,
+The present work explored the psycholinguistic bases of homophonic
+translation, a phenomenon in which listening to speech in a
+non-native---perhaps unfamiliar---language leads to the activation of
+lexical representations in the native language, without necessarily
+preserving the meaning. We investigated how phonological similarity and
+its interaction with phonological neighbourhood density impact the
+dynamics of lexical activation and selection during non-native word
+processing. We designed a translation elicitation task in which
+participants listened to individual words in a non-native, unfamiliar
+language. After listening to each word, participants were asked to
+provide their best-guess translation for each word in their native
+language. We run three experiments. In Experiment 1, British
+English-native adults listened to a list of words in Catalan or Spanish.
+Participants reported no prior familiarity with Catalan, Spanish, or any
+other Romance language, yet provided accurate translations for words
+that shared some degree of phonological similarity with their correct
+English translation. Participants only benefited from phonological
+similarity when the number of English phonological neighbours of the
+presented Catalan or Spanish word was small. Experiment 2 was aimed at
+extending our findings to a population of participants whose native
+language was typologically closer to the unfamiliar language presented.
+We tested a group of Spanish native adults who listened to a list of
+Catalan words. Again, participants reported no prior familiarity with
+Catalan, or any other Romance language other than Spanish. In line with
+the results in English natives, we found an interaction between
+phonological similarity and phonological neighbourhood size. The
+facilitation effect of phonological similarity in the Spanish
+participants was more robust to the interferring effect of phonological
+neighbourhood density: Spanish participants provided correct
+translations for words with high neighbourhood densities more often than
+English participants. We suspected that participants in Experiments 1
+and 2 might have managed to provide correct translations for some
+low-phonological similarity words thanks to having prior experience with
+those words, despite having reported being unfamiliar with Catalan,
+Spanish, or any other Romance language. Experiment 3 was a replica of
+Experiment 1, in which we collected additional information about
+participants' prior familiarity with the presented Catalan and Spanish
+words. The design of Experiment 3 was closely modelled after Experiment
+1, except that after providing their response in each trial,
 participants reported whether they had previous knowledge of the
@@
 prior knowledge about the meaning of the presented Catalan or Spanish
-word, we found similar resutls as in Experiment 1, suggesting that
-participants did not simply rely on prior familiarity with the
-word-forms presented but rather exclusively on their phonological
-similarity with the correct English translation.
-
-Our findings provide insights into the mechanisms underlying homophonic
-translations. This phenomenon is frequently experienced by listeners of
-foreign speech, when strings of sounds in the input are perceived as
-strings of words or phrases in the native language, without necessarily
-preserving the meaning.
-
-This type of cross-linguistic similarity has also been extremely
-informative for the study of bilingual lexical processing: words that
-share high form-similarity with their translations are processed faster
-and more accurately than words sharing lower phonological similarity.
-This phenomenon has provided strong evidence in favour of the
-language-non selective hypothesis of bilingual lexical access, which
-states that bilinguals activate lexical representations in both
-languages, even during monolingual situations. This facilitation effect
-has been reported for comprehension
+word, we run the same analyses as in Experiment 1, finding parallel
+results. Results in Experiment 3 suggest participants in the present
+study did not simply rely on prior familiarity with some of the
+word-forms presented. Rather, participants relied on the phonological
+similarity between the presented Catalan and Spanish words and their
+correct translation.
+
+Homophonic translation is an often spontaneous phenomenon that occurs
+commonly when listening to non-native speech, and is prevalent across
+many cultures and languages
+(\citeproc{ref-dembeck2015oberflachenubersetzung}{Dembeck, 2015};
+\citeproc{ref-efimova2018homophonic}{Efimova et al., 2018}). The most
+famous case of homophonic translation is \emph{Soramimi}, which refers
+to homophonic translation in the context of music listening, when
+particular sections of the lyrics of a song in a non-native language
+make listeners think of particular words and phrases in their own native
+language. This sometimes result in a completely different meaning than
+thhe originally intended. Homophonic translation occurs outside the
+context of music. For instance, it is often intentionally employed by
+translators to preserve the meter and ``sound structure'' (e.g., rhymes)
+of an original text in the language of translation (e.g.,
+\citeproc{ref-levick2016translating}{Levick, 2016};
+\citeproc{ref-pilshchikov2016semiotics}{Pilshchikov, 2016})\footnote{The
+  rationale behind this translation approach is to ``sacrifice{[}s{]}
+  fidelity to the image rather than the melodiousness of verse''
+  (\citeproc{ref-briusov1911ot}{Briusov, 1911}). For instance, Edith
+  Grossman's translation of \emph{El Quijote}
+  (\citeproc{ref-cervantes}{Cervantes \& Grossman, 2002}) present a
+  similar case, in which formal aspects of the original text in Spanish
+  are favoured to some degree over equivalence to English, even
+  sometimes keeping Spanish words from the original text in the English
+  translation (e.g., \emph{Señor}, \emph{ínsula}).}. Although homophonic
+translation has mostly received attention as a perceptual curiosity or
+as a literary figure, little attention has been paid to homophonic
+translation as a spontanous \emph{psycholinguistic} event. The
+psycholiguistic mechanisms behind spontaneous homophonic translations
+are informative of the mechanisms underlying native speech processing.
+How does the lexical system deal with a non-native speech signal in
+which form-meaning associations are scarce, if not absent?
+
+How the perceptual system deals with non-native speech is a vastly
+studied subject (e.g., \citeproc{ref-best1988examination}{Best et al.,
+1988}; \citeproc{ref-miyawaki1975effect}{Miyawaki et al., 1975};
+\citeproc{ref-pallier2001influence}{Pallier et al., 2001};
+\citeproc{ref-van2015learning}{Van Leussen \& Escudero, 2015};
+\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}; see
+\citeproc{ref-baese2020perception}{Baese-Berk et al., 2020} for review).
+The same cannot be said about how the perceptual system deals with
+speech in an \emph{unfamiliar} language. Previous research had explored
+how bilinguals recognise, translate, and produce words as a function of
+their proficiency in each of their languages
+(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006};
+\citeproc{ref-dijkstra2019multilink}{Dijkstra et al., 2019};
+\citeproc{ref-dijkstra2002architecture}{Dijkstra \& Van Heuven, 2002};
+e.g., \citeproc{ref-kroll1994category}{Kroll \& Stewart, 1994}). The
+case of a very low-proficiency bilingual may be close to that of a
+listener of an unfamiliar language, but being bilingual implies some
+form of proficiency in both languages, such as some knowledge about the
+meaning of some words. Even low-proficiency bilinguals are likely to
+have already acquired many form-meaning mappings in their lower
+proficiency language. Characterising the impact of phonological
+similarity on word processing is therefore challenging in bilinguals,
+due to the simultaneous contribution of phonological \emph{and}
+conceptual links in the lexicon. By testing participants with \emph{no}
+prior familiarity with the presented language (i.e., no proficiency), we
+ensured that no form-meaning associations were available for
+participants to correctly translate the unfamiliar words, if not by
+relying excluusively on phonological similarity.
+
+We calculated phonological similarity between word pairs as the
+Levenshtein similarity between the X-SAMPA transcription of their
+phonological forms (see \citeproc{ref-floccia2018introduction}{Floccia
+et al., 2018} for a similar approach). Using this measure, we found that
+participants were able to exploit the phonological similarity between
+the presented words and their correct translation, even when both
+words-forms shared few phonemes. This points to listeners accommodating
+non-native phonological forms to their native phoneme inventory,
+resulting in the activation of native lexical representations. Future
+studies may consider comparing the suitability of other measures of
+phonological similarity that take into account finer-grained phonetic or
+prosodic cues present in the acoustic signal presented to participants.
+Phonological transcriptions like X-SAMPA are abstract representations
+that ignore non-phonological contrasts (e.g., phones), and consider
+different symbols as completely different phonemes, disregarding the
+fact that some phonemes share more phonetic features than others.
+Additionally, prosodic cues such as lexical stress, which our measure of
+phonological similarity does not include, might also provide
+participants further information about the correct translation of the
+presented word-forms. Altogether, it is likely that participants in the
+present study were able to exploit additional cues in the acoustic
+signal to provide correct translations. Future studies may explore the
+relative contribution of such cues in the occurrence of homophonic
+translation. Overall, the present study provides some insights into the
+role of phonological similarity in the auditory presentation of words in
+an unfamiliar language.
+
+Our findings also suggest that the facilitation effect of phonological
+familiarity was moderated by phonological neighbourghood density. This
+is in line which previous studies suggesting that the presence of
+(high-frequency) phonological neighbours interferes with the lexical
+selection in word comprehension tasks (e.g.,
+\citeproc{ref-dufour2010phonological}{Dufour \& Frauenfelder, 2010};
+\citeproc{ref-grainger1990word}{Grainger, 1990};
+\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998};
+\citeproc{ref-vitevitch1998words}{Vitevitch \& Luce, 1998};
+\citeproc{ref-vitevitch1999probabilistic}{Vitevitch \& Luce, 1999}). Our
+translation elicitation task can be understood as a particular instance
+of an auditory word recognition task, in which the presented unfamiliar
+words have the potential to activate phonological neighbours in
+participants' lexicon. We calculated a measure of cross-linguistic
+neighbourhood size by counting, for each presented word (in Catalan or
+Spanish in Experiments 1 and 3, Catalan in Experiment 2) the number of
+phonologically related words (with a Levenshtein distance of one) in the
+target language (English in Experiments 1 and 3, Spanish in Experiment
+2). To account for the fact that competition between phonological
+neighbours is sensitive to lexical frequency
+(\citeproc{ref-dufour2010phonological}{Dufour \& Frauenfelder, 2010};
+high-frequency neighbours produce stronger interference effects,
+\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}), we only
+counted neighbours with a lexical frequency higher than the correct
+translation. Using this measure, we found that participants' ability to
+exploit phonological similarity to produce correct translations declined
+as the number of phonological neighbours increased. Overall, this
+suggests that listening to words in an unfamiliar language trigger the
+same mechanisms of lexical activation and selection that listening to
+words in a native language does.
+
+A comparison between results in Experiments 1 and 2 suggested that the
+performance of Spanish participants translating Catalan was more
+resilient to the interfering effects of phonological neighbourhood
+density than the performance of English participants translating Catalan
+or Spanish. As highlighted before, Catalan and Spanish (both Romance
+languages) are typologically closer to each other than to English (a
+Germanic language). Catalan and Spanish share a higher proportion of
+cognates (i.e., form-similar translation equivalents) than Catalan and
+English, or Spanish and English. Previous studies in bilinguals show
+that this kind of cross-linguistic similarity impacts lexical
+processing: words that share high phonological similarity with their
+translations are processed faster and more accurately than words sharing
+lower phonological similarity. This phenomenon has provided strong
+evidence in favour of the language-non selective hypothesis of bilingual
+lexical access, which states that bilinguals activate lexical
+representations in both languages, even during monolingual situations.
+This facilitation effect has been reported for comprehension
 (\citeproc{ref-dijkstra2010cross}{Dijkstra et al., 2010};
@@
 \citeproc{ref-valente2018does}{Valente et al., 2018}), and translation
-(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}).
+(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}). To
+our knowledge, this present study is the first to investigate the role
+of cross-linguistic phonological similarity in a monolingual population,
+in which the role of conceptual links in the lexicon is ruled out, given
+participants lack of familiarity with the presented language. This
+provided a convenient case to investigate the role of phonological
+similarity in lexical processing, disantangling the contribution of
+phonological similarity from the contribution of conceptual equivalence.
+Overall, our findings suggest that the typological distance between the
+presented and target language is associated with a more robust
+facilitation effect of phonological similarity.
+
+The specific mechanisms behind this effect are unclear. One possibility
+is that, as aforementioned, Catalan provided additional cues to Spanish
+participants in Experiment 2 compared to English participants in
+Experiment 1. For instance, it is likely that subphonemic cues not
+accounted by our measure of phonological similarity were more
+informative to Spanish participants than to English participants. In
+addition, it remains to be tested whether such association is linear
+(i.e., robustness of facilitation increases linearly with typological
+distance), proportional, or even extends beyond the specific pair of
+languages involved in the present work remains unclear. Do other close
+or distant language pairs show the same pattern of results? The
+generalisability of these findings can also be tested by increasing the
+repertoire of words involved in the translation elicitation task. In
+Experiments 1 to 3, participants answered to a maximum of 105 words. All
+words were of high-frequency, and had a low age-of-acquisition. In order
+to characterise listeners' ability to translate correctly words from an
+unfamiliar language, words with varying levels of difficulty should be
+included.
+
+In summary, the present paper provides insights into the
+psycholinguistic mechanisms underlying homophonic translation. English
+and Spanish native adults were tested in a translation elicitation task
+in which they had to guess the English or Spanish translation of a
+series of words in an unfamiliar language. Participants successfully
+exploited phonological similarity between the presented words and their
+correct translations to provide correct answers. Participants'
+performance in the task only benefited from phonological similarity when
+the presented word had few higher-frequency phonological neighbours in
+the target language. Finally, the facilitation effect of phonological
+similarity was more robust in the Spanish native participants, who
+translated words from a typologically closer language than English
+participants. Overall, the findings presented in the present paper
+suggest that the processing of words in a non-native, unfamiliar
+language recruits mechanisms of lexical activation, selection, and
+interference parallel to those recruited by listening to words in a
+native language.
 
@@
 \begin{CSLReferences}{1}{0}
+\bibitem[\citeproctext]{ref-baese2020perception}
+Baese-Berk, M. M., McLaughlin, D. J., \& McGowan, K. B. (2020).
+Perception of non-native speech. \emph{Language and Linguistics
+Compass}, \emph{14}(7), e12375.
+
 \bibitem[\citeproctext]{ref-best2001discrimination}
@@
 
+\bibitem[\citeproctext]{ref-cervantes}
+Cervantes, M. de, \& Grossman, E. (2002). First part of the ingenious
+nobleman don quixote of la mancha. \emph{Conjunctions}, \emph{38},
+207--225. \url{http://www.jstor.org/stable/24518146}
+
 \bibitem[\citeproctext]{ref-christoffels2006memory}
@@
 
+\bibitem[\citeproctext]{ref-dembeck2015oberflachenubersetzung}
+Dembeck, T. (2015). Oberfl{ä}chen{ü}bersetzung: The poetics and cultural
+politics of homophonic translation. \emph{Critical Multilingualism
+Studies}, \emph{3}(1), 7--25.
+
 \bibitem[\citeproctext]{ref-dijkstra2010cross}
@@
 
+\bibitem[\citeproctext]{ref-dijkstra2002architecture}
+Dijkstra, T., \& Van Heuven, W. J. (2002). The architecture of the
+bilingual word recognition system: {From} identification to decision.
+\emph{Bilingualism: Language and Cognition}, \emph{5}(3), 175--197.
+
+\bibitem[\citeproctext]{ref-dijkstra2019multilink}
+Dijkstra, T., Wahl, A., Buytenhuijs, F., Halem, N. V., Al-Jibouri, Z.,
+Korte, M. D., \& Rekké, S. (2019). Multilink: A computational model for
+bilingual word recognition and word translation. \emph{Bilingualism:
+Language and Cognition}, \emph{22}(4), 657--679.
+\url{https://doi.org/10.1017/S1366728918000287}
+
+\bibitem[\citeproctext]{ref-dufour2010phonological}
+Dufour, S., \& Frauenfelder, U. H. (2010). Phonological neighbourhood
+effects in french spoken-word recognition. \emph{Quarterly Journal of
+Experimental Psychology}, \emph{63}(2), 226--238.
+
 \bibitem[\citeproctext]{ref-dufour2003lexical}
@@
 
+\bibitem[\citeproctext]{ref-efimova2018homophonic}
+Efimova, N. N., Ruzhnikova, M. L., \& Violina, M. I. (2018). Homophonic
+translation as humpty-dumpty's choice. \emph{SHS Web of Conferences},
+\emph{50}, 01049.
+
 \bibitem[\citeproctext]{ref-elias2022crosslanguage}
@@
 
+\bibitem[\citeproctext]{ref-floccia2018introduction}
+Floccia, C., Sambrook, T. D., Delle Luche, C., Kwok, R., Goslin, J.,
+White, L., Cattani, A., Sullivan, E., Abbot-Smith, K., Krott, A., et al.
+(2018). I: {Introduction}. \emph{Monographs of the Society for Research
+in Child Development}, \emph{83}(1), 7--29.
+
+\bibitem[\citeproctext]{ref-gasparov2006semen}
+Gasparov, M. L. (2006). \emph{Semen kirsanov}. Akademicheskij proekt.
+
 \bibitem[\citeproctext]{ref-goldinger1989priming}
@@
 
+\bibitem[\citeproctext]{ref-grainger1990word}
+Grainger, J. (1990). Word frequency and neighborhood frequency effects
+in lexical decision and naming. \emph{Journal of Memory and Language},
+\emph{29}(2), 228--244.
+
 \bibitem[\citeproctext]{ref-hamburger1996phonological}
@@
 from cross-linguistic mondegreens. \emph{ICPhS}.
+
+\bibitem[\citeproctext]{ref-kroll1994category}
+Kroll, J. F., \& Stewart, E. (1994). Category interference in
+translation and picture naming: {Evidence} for asymmetric connections
+between bilingual memory representations. \emph{Journal of Memory and
+Language}, \emph{33}(2), 149--174.
 
@@
 
+\bibitem[\citeproctext]{ref-miyawaki1975effect}
+Miyawaki, K., Jenkins, J. J., Strange, W., Liberman, A. M., Verbrugge,
+R., \& Fujimura, O. (1975). An effect of linguistic experience: The
+discrimination of {[}r{]} and {[}l{]} by native speakers of japanese and
+english. \emph{Perception \& Psychophysics}, \emph{18}(5), 331--340.
+
 \bibitem[\citeproctext]{ref-otake2007interlingual}
@@
 
+\bibitem[\citeproctext]{ref-pallier2001influence}
+Pallier, C., Colomé, A., \& Sebastián-Gallés, N. (2001). The influence
+of native-language phonology on lexical access: Exemplar-based versus
+abstract lexical entries. \emph{Psychological Science}, \emph{12}(6),
+445--449.
+
 \bibitem[\citeproctext]{ref-peirce2019psychopy2}
@@
 
+\bibitem[\citeproctext]{ref-pilshchikov2016semiotics}
+Pilshchikov, I. (2016). The semiotics of phonetic translation.
+\emph{Studia Metrica Et Poetica}, \emph{3}(1), 53--104.
+
 \bibitem[\citeproctext]{ref-rcoreteam2013language}
@@
 \emph{67}(6), 1176--1190.
+
+\bibitem[\citeproctext]{ref-van2015learning}
+Van Leussen, J.-W., \& Escudero, P. (2015). Learning to perceive and
+recognize a second language: The L2LP model revised. \emph{Frontiers in
+Psychology}, \emph{6}, 1000.
+
+\bibitem[\citeproctext]{ref-vitevitch1998words}
+Vitevitch, M. S., \& Luce, P. A. (1998). When words compete: Levels of
+processing in perception of spoken words. \emph{Psychological Science},
+\emph{9}(4), 325--329.
+
+\bibitem[\citeproctext]{ref-vitevitch1999probabilistic}
+Vitevitch, M. S., \& Luce, P. A. (1999). Probabilistic {Phonotactics}
+and {Neighborhood Activation} in {Spoken Word Recognition}.
+\emph{Journal of Memory and Language}, \emph{40}(3), 374--408.
+\url{https://doi.org/10.1006/jmla.1998.2618}
 
@@
 
+\bibitem[\citeproctext]{ref-baese2020perception}
+Baese-Berk, M. M., McLaughlin, D. J., \& McGowan, K. B. (2020).
+Perception of non-native speech. \emph{Language and Linguistics
+Compass}, \emph{14}(7), e12375.
+
 \bibitem[\citeproctext]{ref-best2001discrimination}

### 2024-10-17 manuscript/manuscript.qmd
Fix table of surprises

@@
 #| label: tbl-surprises
-#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct Enlgish translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
 tbl_data <- list(
@@
   bind_rows(.id = "exp") |>
-  mutate(translation = paste0())
+  mutate(
+    translation = paste0(word_1, " - ",  word_2),
+    xsampa = paste0("/", sampa_1, "/ - /", sampa_2, "/")
+  ) |> 
   summarise(
@@
     correct = sum(correct),
-    .by = c(exp, group, lv, translation)
+    .by = c(exp, group, lv, translation, xsampa)
   ) |>
@@
   dplyr::filter(lv == 0, accuracy >= 0.1) |>
-  select(exp, translation, group, accuracy, accuracy_se) |>
-  arrange(exp, group, desc(accuracy))
-
-tbl_data |>
-  select(translation, accuracy, accuracy_se) |> 
+  select(exp, translation, xsampa, group, accuracy, accuracy_se) |>
+  arrange(exp, group, desc(accuracy)) |> 
+  select(translation, xsampa, accuracy, accuracy_se) |> 
   mutate(accuracy = round(accuracy * 100, 2),
-         accuracy_se = round(accuracy_se * 100, 2)) |> 
-  set_names(c("Translation", "Accuracy (%)", "SE")) |> 
-  tt() |> 
+         accuracy_se = round(accuracy_se * 100, 2))
+
+names(tbl_data) <- c("Translation", "X-SAMPA", "Accuracy (\\%)", "SE")
+
+tt(tbl_data) |> 
   group_tt(i = list(
-    "Experiment 1 (cat-ENG)" = 5,
-    "Experiment 2 (spa-ENG)" = 21,
-    "Experiment 3 (cat-SPA)" = 26))
+    "Experiment 1 (cat-ENG)" = 1,
+    "Experiment 2 (spa-ENG)" = 6,
+    "Experiment 3 (cat-SPA)" = 22)) |> 
+  style_tt(i = 0, bold = TRUE, line = "tb") |> 
+  style_tt(i = 1, bold = TRUE, line = "b") |> 
+  style_tt(i = 7, bold = TRUE, line = "tb") |> 
+  style_tt(i = 24, bold = TRUE, line = "tb")
+
 ```

### 2024-10-17 manuscript/manuscript.tex
Fix table of surprises

@@
 accuracy: the Levenshtein similarityscore betwen the presented word (in
-Catalan or Spanish) and their correct Enlgish translation is zero, but
+Catalan or Spanish) and their correct English translation is zero, but
 participants, who are reportedly unfamiliar with the presented language,
@@
 {                     %% tabularray inner open
-colspec={Q[]Q[]Q[]},
-cell{6}{1}={c=3}{},cell{23}{1}={c=3}{},cell{29}{1}={c=3}{},
-cell{2}{1}={preto={\hspace{1em}}},
+colspec={Q[]Q[]Q[]Q[]},
+cell{2}{1}={c=4}{},cell{8}{1}={c=4}{},cell{25}{1}={c=4}{},
 cell{3}{1}={preto={\hspace{1em}}},
@@
 cell{5}{1}={preto={\hspace{1em}}},
+cell{6}{1}={preto={\hspace{1em}}},
 cell{7}{1}={preto={\hspace{1em}}},
-cell{8}{1}={preto={\hspace{1em}}},
 cell{9}{1}={preto={\hspace{1em}}},
@@
 cell{22}{1}={preto={\hspace{1em}}},
+cell{23}{1}={preto={\hspace{1em}}},
 cell{24}{1}={preto={\hspace{1em}}},
-cell{25}{1}={preto={\hspace{1em}}},
 cell{26}{1}={preto={\hspace{1em}}},
@@
 cell{28}{1}={preto={\hspace{1em}}},
+cell{29}{1}={preto={\hspace{1em}}},
 cell{30}{1}={preto={\hspace{1em}}},
 cell{31}{1}={preto={\hspace{1em}}},
+row{1}={,cmd=\bfseries,},
+hline{1,2}={1,2,3,4}{solid, 0.1em, black},
+row{2}={,cmd=\bfseries,},
+hline{3}={1,2,3,4}{solid, 0.1em, black},
+row{8}={,cmd=\bfseries,},
+hline{8,9}={1,2,3,4}{solid, 0.1em, black},
+row{25}={,cmd=\bfseries,},
+hline{25,26}={1,2,3,4}{solid, 0.1em, black},
 }                     %% tabularray inner close
 \toprule
-Translation & Accuracy (%) & SE \\ \midrule %% TinyTableHeader
-cavall /kəβaʎ/ - horse /hɔs/             & 17.14 & 6.37 \\
-llibre /ʎiβɾə/ - book /bʊk/              & 17.14 & 6.37 \\
-camisa /kəmizə/ - shirt /ʃɜt/            & 16.67 & 6.21 \\
-poma /poma/ - apple /æpl/                & 16.67 & 6.21 \\
-Experiment 1 (cat-ENG) && \\
-cama /kamə/ - leg /lɛg/                  & 11.11 & 5.24 \\
-pantalón /paŋtalon/ - trousers /traʊzəz/ & 77.42 & 7.51 \\
-naranja /naɾaŋxa/ - orange /ɒrɪnʤ/       & 41.94 & 8.86 \\
-leche /leʧe/ - milk /mɪlk/               & 35.48 & 8.59 \\
-toro /toɾo/ - bull /bʊl/                 & 33.33 & 8.61 \\
-libro /liβɾo/ - book /bʊk/               & 30.00 & 8.37 \\
-cebra /θebɾa/ - zebra /zibrə/            & 29.03 & 8.15 \\
-pan /pan/ - bread /brɛd/                 & 29.03 & 8.15 \\
-pollo /poʎo/ - chicken /ʧɪkɪn/           & 26.67 & 8.07 \\
-jirafa /xiɾafa/ - giraffe /ʤɪrɑf/        & 20.69 & 7.52 \\
-perro /pero/ - dog /dɒɡ/                 & 16.13 & 6.61 \\
-pluma /pluma/ - feather /fɛðə/           & 16.13 & 6.61 \\
-puerta /pwerta/ - door /dɔ/              & 16.13 & 6.61 \\
-pie /pje/ - foot /fʊt/                   & 12.90 & 6.02 \\
-caballo /kaβaʎo/ - horse /hɔs/           & 10.34 & 5.66 \\
-bocadillo /bokadiʎo/ - sandwich /sænwɪʤ/ & 10.00 & 5.48 \\
-Experiment 2 (spa-ENG) && \\
-globo /ɡloβo/ - balloon /bəlun/          & 10.00 & 5.48 \\
-fulla /fuʎə/ - hoja /oxa/                & 30.43 & 9.59 \\
-ull /uʎ/ - ojo /oxo/                     & 21.74 & 8.60 \\
-got /gɔt/ - vaso /baso/                  & 20.00 & 8.00 \\
-entrepà /entɾəpa/ - bocadillo /bokadiʎo/ & 13.04 & 7.02 \\
-Experiment 3 (cat-SPA) && \\
-mirall /miɾaʎ/ - espejo /espexo/         & 12.50 & 6.75 \\
+Translation & X-SAMPA & Accuracy (\%) & SE \\ \midrule %% TinyTableHeader
+Experiment 1 (cat-ENG) &&& \\
+cavall - horse       & /k5BaL/ - /hOs/          & 17.14 & 6.37 \\
+llibre - book        & /LiB45/ - /bUk/          & 17.14 & 6.37 \\
+camisa - shirt       & /k5miz5/ - /S36t/        & 16.67 & 6.21 \\
+poma - apple         & /poma/ - /1pl/           & 16.67 & 6.21 \\
+cama - leg           & /kam5/ - /lEg/           & 11.11 & 5.24 \\
+Experiment 2 (spa-ENG) &&& \\
+pantalon - trousers  & /pantalon/ - /tr0aUs56z/ & 77.42 & 7.51 \\
+naranja - orange     & /na4aNxa/ - /Or0I0ndZ/   & 41.94 & 8.86 \\
+leche - milk         & /letSe/ - /mIlk/         & 35.48 & 8.59 \\
+toro - bull          & /to4o/ - /bUl/           & 33.33 & 8.61 \\
+libro - book         & /liB4o/ - /bUk/          & 30.00 & 8.37 \\
+cebra - zebra        & /TeB4a/ - /zibr05/       & 29.03 & 8.15 \\
+pan - bread          & /pan/ - /br0Ed/          & 29.03 & 8.15 \\
+pollo - chicken      & /poLo/ - /tSIkIn/        & 26.67 & 8.07 \\
+jirafa - giraffe     & /xi4afa/ - /dZI0r01f/    & 20.69 & 7.52 \\
+perro - dog          & /pErro/ - /dAg/          & 16.13 & 6.61 \\
+pluma - feather      & /pluma/ - /fED56/        & 16.13 & 6.61 \\
+puerta - door        & /pwE4ta/ - /dor/         & 16.13 & 6.61 \\
+pie - foot           & /pje/ - /fUt/            & 12.90 & 6.02 \\
+caballo - horse      & /kaBaLo/ - /hors/        & 10.34 & 5.66 \\
+bocadillo - sandwich & /bokaDiLo/ - /s1ndwItS/  & 10.00 & 5.48 \\
+globo - balloon      & /gloBo/ - /b5lun/        & 10.00 & 5.48 \\
+Experiment 3 (cat-SPA) &&& \\
+fulla - hoja         & /fuL5/ - /oxa/           & 30.43 & 9.59 \\
+ull - ojo            & /uL/ - /oxo/             & 21.74 & 8.60 \\
+got - vaso           & /gOt/ - /baso/           & 20.00 & 8.00 \\
+entrepa - bocadillo  & /ent45pa/ - /bokaDiLo/   & 13.04 & 7.02 \\
+mirall - espejo      & /mi4aL/ - /espexo/       & 12.50 & 6.75 \\
 \bottomrule

### 2024-10-18 manuscript/manuscript.qmd
Fix preprint

--- 2024-10-16 manuscript/manuscript.qmd
@@
 library(tinytable)
+library(stringdist)
 
@@
 
-n_participants <- n_distinct(participants_1)
+n_participants <- dplyr::n_distinct(participants_1)
 n_participants_group <- table(participants_1$group)
@@
 lengths <- stimuli |>
-  select(group, ipa_flat_1, ipa_flat_2, word_1, word_2) |>
-  mutate(across(ipa_flat_1:word_2, nchar)) |>
+  select(group, ipa_1, ipa_2, word_1, word_2) |>
+  mutate(across(ipa_1:word_2, nchar)) |>
   summarise(
     across(
-      c(ipa_flat_1, ipa_flat_2, word_1, word_2),
+      c(ipa_1, ipa_2, word_1, word_2),
       lst(mean, sd, min, max)
@@
 We created two lists of input words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
-Words in the Catalan list were `r round(lengths$cat_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$cat_eng$ipa_flat_1_min, 2)`-`r round(lengths$cat_eng$ipa_flat_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, Range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`).
-Words in the Spanish list were `r round(lengths$spa_eng$ipa_flat_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_flat_1_sd, 2)`, Range = `r round(lengths$spa_eng$ipa_flat_1_min, 2)`-`r round(lengths$spa_eng$ipa_flat_1_max, 2)`), and the orthographic form of their English translations were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, Range = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`).
+Words in the Catalan list were `r round(lengths$cat_eng$ipa_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_1_sd, 2)`, Range = `r round(lengths$cat_eng$ipa_1_min, 2)`-`r round(lengths$cat_eng$ipa_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, Range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`).
+Words in the Spanish list were `r round(lengths$spa_eng$ipa_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_1_sd, 2)`, Range = `r round(lengths$spa_eng$ipa_1_min, 2)`-`r round(lengths$spa_eng$ipa_1_max, 2)`), and the orthographic form of their English translations were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, Range = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`).
 
@@
 For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
-For example, the *table* (`teIb@l`)-*mesa* (`mesa`) translation pair had a `r scales::percent(stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (`tREn`) translation pair had a `r scales::percent(stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+For example, the *table* (`teIb@l`)-*mesa* (`mesa`) translation pair had a `r scales::percent(stringdist::stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (`tREn`) translation pair had a `r scales::percent(stringdist::stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
 @tbl-stimuli summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
@@
 )
-n_participants <- map(dataset, \(x) n_distinct(x$participant_id))
+n_participants <- map(dataset, \(x) dplyr::n_distinct(x$participant_id))
 
@@
 
-After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
-Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id) + dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
+Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
 Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, Range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, Range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
@@
     across(n_trials, lst(mean, sum, sd, min, max)),
-    n_participants = n_distinct(participant_id),
+    n_participants = dplyr::n_distinct(participant_id),
     .by = c(group, exp)
@@
 #| label: fig-epreds-1
+#| cache: true
 #| fig-height: 5
@@
 
-n_participants <- n_distinct(participants_2)
+n_participants <- dplyr::n_distinct(participants_2)
 n_participants_group <- table(participants_2$group)
@@
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a developmental language disorder (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
-After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_spa$participant_id) + n_distinct(dataset_clean$cat_spa$participant_id)` unique participants.
-Of those trials, `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Spanish words.
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id) + dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` unique participants.
+Of those trials, `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Spanish words.
 Responses given by participants were `r round(lengths$len_mean, 2)` characters long on average (*Median* = `r round(lengths$len_median, 2)`, *SD* = `r round(lengths$len_sd, 2)`, Range = `r round(lengths$len_min, 2)`-`r round(lengths$len_max, 2)`).
@@
 #| label: fig-epreds-2
+#| cache: true
 #| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
@@
 #| label: tbl-surprises
-#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct Enlgish translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+ipa_to_tipa <- function(ipa) {
+  dict <- c(
+    "ɑ" = "A", 
+    "ɒ" = "6",
+    "æ" = "\\ae", 
+    "ʌ" = "2", 
+    "β" = "B",
+    "ʤ" = "}\\\\textdyoghlig\\\\textipa{",
+    "d͡ʒ" = "}\\\\textdyoghlig\\\\textipa{",
+    "dʒ" = "}\\\\textdyoghlig\\\\textipa{", 
+    "ð" = "D",
+    "ə" = "@",
+    "ɛ" = "E",
+    "ɣ" = "G",
+    "ɡ" = "g",
+    "ɫ" = "\\|~l", 
+    "ʎ" = "}\\\\textlambda\\\\textipa{",
+    "ŋ" = "N",
+    "ɲ" = "}\\\\textltailn\\\\textipa{", 
+    "ɔ" = "O",
+    "ɾ" = "R",
+    "ʧ" = "}\\\\textteshlig\\\\textipa{", 
+    "tʃ" = "}\\\\textteshlig\\\\textipa{", 
+    "t͡ʃ" = "}\\\\textteshlig\\\\textipa{", 
+    "θ" = "T", 
+    "t͡s" = "}\\\\texttslig\\\\textipa{", 
+    "ts" = "}\\\\texttslig\\\\textipa{", 
+    "ʃ" = "S", 
+    "ʒ" = "Z", 
+    "ɪ" = "I", 
+    "ʊ" = "U", 
+    "ɜ" = "3", 
+    "ɹ" = "\\*r", 
+    "ː" = ":", 
+    "ˈ" = '"'
+  )
+  tipa <- stringr::str_replace_all(ipa, dict)
+  tipa <- paste0("\\textipa{", tipa, "}")
+  return(tipa)
+}
+
 tbl_data <- list(
@@
   bind_rows(.id = "exp") |>
-  mutate(translation = paste0())
+  mutate(
+    translation = paste0(word_1, " - ",  word_2),
+    ipa = ipa_to_tipa(paste0("/", ipa_1, "/ - /", ipa_2, "/"))
+  ) |> 
   summarise(
@@
     correct = sum(correct),
-    .by = c(exp, group, lv, translation)
+    .by = c(exp, group, lv, translation, ipa)
   ) |>
@@
   dplyr::filter(lv == 0, accuracy >= 0.1) |>
-  select(exp, translation, group, accuracy, accuracy_se) |>
-  arrange(exp, group, desc(accuracy))
-
-tbl_data |>
-  select(translation, accuracy, accuracy_se) |> 
+  select(exp, translation, ipa, group, accuracy, accuracy_se) |>
+  arrange(exp, group, desc(accuracy)) |> 
+  select(translation, ipa, accuracy, accuracy_se) |> 
   mutate(accuracy = round(accuracy * 100, 2),
-         accuracy_se = round(accuracy_se * 100, 2)) |> 
-  set_names(c("Translation", "Accuracy (%)", "SE")) |> 
-  tt() |> 
+         accuracy_se = round(accuracy_se * 100, 2))
+
+names(tbl_data) <- c("Translation", "IPA", "Accuracy (\\%)", "SE")
+
+tt(tbl_data) |> 
+  style_tt(i = 0, bold = TRUE, line = "tb") |> 
   group_tt(i = list(
-    "Experiment 1 (cat-ENG)" = 5,
-    "Experiment 2 (spa-ENG)" = 21,
-    "Experiment 3 (cat-SPA)" = 26))
+    "Experiment 1 (cat-ENG)" = 1,
+    "Experiment 2 (spa-ENG)" = 6,
+    "Experiment 3 (cat-SPA)" = 22)) |> 
+  style_tt(i = 1, bold = TRUE, line = "b") |> 
+  style_tt(i = 7, bold = TRUE, line = "tb") |> 
+  style_tt(i = 24, bold = TRUE, line = "tb")
 ```
@@
 
-n_participants <- n_distinct(participants_3)
+n_participants <- dplyr::n_distinct(participants_3)
 n_participants_group <- table(participants_3$group)
@@
 
-After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$spa_eng), big.mark = ",")` trials provided by `r n_distinct(dataset_clean$cat_eng$participant_id) + n_distinct(dataset_clean$spa_eng$participant_id)` unique participants.
-Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$spa_eng), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id) + dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` unique participants.
+Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
 
@@
 #| label: fig-epreds-3
+#| cache: true
 #| fig-cap: "Posterior model-predicted mean accuracy in Experiment 2.

### 2024-10-18 manuscript/manuscript.tex
Fix preprint

@@
 \fi
-\usepackage[]{Times}
+\usepackage{lmodern}
 \ifPDFTeX\else  
@@
 \usepackage{booktabs}
-\DeclareTextFontCommand\textphonetic{\ipafont}
-\usepackage[utf8]{inputenc}
-\usepackage{CJKutf8}
 \usepackage{arxiv}
@@
 \title{The role of cognateness in native spoken word recognition}
+\def\asep{\\\\\\ } % default: all authors on same column
+\def\asep{\And }
 \author{\textbf{Gonzalo
@@
 Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\texttt{\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}}\And\textbf{Serene
+08005\\\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}\asep\textbf{Serene
 Siow}~\orcidlink{0000-0001-6482-2191}\\Department of Experimental
 Psychology\\University of Oxford\\Oxford, OX2
-6GG\\\texttt{\href{mailto:siow.serene@gmail.com}{siow.serene@gmail.com}}\And\textbf{Kim
+6GG\\\href{mailto:siow.serene@gmail.com}{siow.serene@gmail.com}\asep\textbf{Kim
 Plunkett}~\orcidlink{0000-0003-0216-7480}\\Department of Experimental
 Psychology\\University of Oxford\\Oxford, OX2
-6GG\\\texttt{\href{mailto:kim.plunkett@psy.ox.ac.uk}{kim.plunkett@psy.ox.ac.uk}}\And\textbf{Nuria
+6GG\\\href{mailto:kim.plunkett@psy.ox.ac.uk}{kim.plunkett@psy.ox.ac.uk}\asep\textbf{Nuria
 Sebastian-Galles}~\orcidlink{0000-0001-6938-2498}\\Center for Brain and
 Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\texttt{\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}}
+08005\\\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}
 \date{}
@@
 \maketitle
+\begin{abstract}
+Insert abstract here.
+\end{abstract}
 {\bfseries \emph Keywords}
@@
 the auditory modality: one in Catalan and one in Spanish. Words in the
-Catalan list were 5.02 phonemes long on average (\emph{SD} = 1.49, Range
-= 2-8), and the orthographic forms of their English translations (which
+Catalan list were 6.67 phonemes long on average (\emph{SD} = 2.06, Range
+= 2-11), and the orthographic forms of their English translations (which
 participants had to type) were 5.12 characters long on average
-(\emph{SD} = 1.56, Range = 3-9). Words in the Spanish list were 5.52
-phonemes long on average (\emph{SD} = 1.47, Range = 3-9), and the
+(\emph{SD} = 1.56, Range = 3-9). Words in the Spanish list were 7.27
+phonemes long on average (\emph{SD} = 2.05, Range = 3-13), and the
 orthographic form of their English translations were 5.29 characters
@@
 \toprule
-Translation & X-SAMPA & Accuracy (\%) & SE \\ \midrule %% TinyTableHeader
+Translation & IPA & Accuracy (\%) & SE \\ \midrule %% TinyTableHeader
 Experiment 1 (cat-ENG) &&& \\
-cavall - horse       & /k5BaL/ - /hOs/          & 17.14 & 6.37 \\
-llibre - book        & /LiB45/ - /bUk/          & 17.14 & 6.37 \\
-camisa - shirt       & /k5miz5/ - /S36t/        & 16.67 & 6.21 \\
-poma - apple         & /poma/ - /1pl/           & 16.67 & 6.21 \\
-cama - leg           & /kam5/ - /lEg/           & 11.11 & 5.24 \\
+cavall - horse       & \textipa{/k@'ba}\textlambda\textipa{/ - /hO:s/}                                  & 17.14 & 6.37 \\
+llibre - book        & \textipa{/"}\textlambda\textipa{i.BR@/ - /bUk/}                                  & 17.14 & 6.37 \\
+camisa - shirt       & \textipa{/ka.'mi.za/ - /S3:t/}                                                     & 16.67 & 6.21 \\
+poma - apple         & \textipa{/"po.ma/ - /"aepl/}                                                       & 16.67 & 6.21 \\
+cama - leg           & \textipa{/"ka.m@/ - /lEg/}                                                         & 11.11 & 5.24 \\
 Experiment 2 (spa-ENG) &&& \\
-pantalon - trousers  & /pantalon/ - /tr0aUs56z/ & 77.42 & 7.51 \\
-naranja - orange     & /na4aNxa/ - /Or0I0ndZ/   & 41.94 & 8.86 \\
-leche - milk         & /letSe/ - /mIlk/         & 35.48 & 8.59 \\
-toro - bull          & /to4o/ - /bUl/           & 33.33 & 8.61 \\
-libro - book         & /liB4o/ - /bUk/          & 30.00 & 8.37 \\
-cebra - zebra        & /TeB4a/ - /zibr05/       & 29.03 & 8.15 \\
-pan - bread          & /pan/ - /br0Ed/          & 29.03 & 8.15 \\
-pollo - chicken      & /poLo/ - /tSIkIn/        & 26.67 & 8.07 \\
-jirafa - giraffe     & /xi4afa/ - /dZI0r01f/    & 20.69 & 7.52 \\
-perro - dog          & /pErro/ - /dAg/          & 16.13 & 6.61 \\
-pluma - feather      & /pluma/ - /fED56/        & 16.13 & 6.61 \\
-puerta - door        & /pwE4ta/ - /dor/         & 16.13 & 6.61 \\
-pie - foot           & /pje/ - /fUt/            & 12.90 & 6.02 \\
-caballo - horse      & /kaBaLo/ - /hors/        & 10.34 & 5.66 \\
-bocadillo - sandwich & /bokaDiLo/ - /s1ndwItS/  & 10.00 & 5.48 \\
-globo - balloon      & /gloBo/ - /b5lun/        & 10.00 & 5.48 \\
+pantalon - trousers  & \textipa{/paN.ta"lon/ - /"traUz@z/}                                                & 77.42 & 7.51 \\
+naranja - orange     & \textipa{/na"RaN.xa/ - /"6rIn}\textdyoghlig\textipa{/}                           & 41.94 & 8.86 \\
+leche - milk         & \textipa{/"le.}\textteshlig\textipa{e/ - /mIlk/}                                 & 35.48 & 8.59 \\
+toro - bull          & \textipa{/"to.Ro/ - /bUl/}                                                         & 33.33 & 8.61 \\
+libro - book         & \textipa{/"li.BRo/ - /bUk/}                                                        & 30.00 & 8.37 \\
+cebra - zebra        & \textipa{/"Te.bRa/ - /"zi:br@/}                                                    & 29.03 & 8.15 \\
+pan - bread          & \textipa{/pan/ - /brEd/}                                                           & 29.03 & 8.15 \\
+pollo - chicken      & \textipa{/"po.}\textlambda\textipa{o/ - /"}\textteshlig\textipa{IkIn/}         & 26.67 & 8.07 \\
+jirafa - giraffe     & \textipa{/xi'Ra.fa/ - /}\textdyoghlig\textipa{I"rA:f/}                           & 20.69 & 7.52 \\
+perro - dog          & \textipa{/pe.ro/ - /d6g/}                                                          & 16.13 & 6.61 \\
+pluma - feather      & \textipa{/plu.ma/ - /"fED@/}                                                       & 16.13 & 6.61 \\
+puerta - door        & \textipa{/pwer.ta/ - /dO:/}                                                        & 16.13 & 6.61 \\
+pie - foot           & \textipa{/pje/ - /fUt/}                                                            & 12.90 & 6.02 \\
+caballo - horse      & \textipa{/ka"Ba.}\textlambda\textipa{o/ - /hO:s/}                                & 10.34 & 5.66 \\
+bocadillo - sandwich & \textipa{/bo.ka"di.}\textlambda\textipa{o/ - /"saenwI}\textdyoghlig\textipa{/} & 10.00 & 5.48 \\
+globo - balloon      & \textipa{/"glo.Bo/ - /b@"lu:n/}                                                    & 10.00 & 5.48 \\
 Experiment 3 (cat-SPA) &&& \\
-fulla - hoja         & /fuL5/ - /oxa/           & 30.43 & 9.59 \\
-ull - ojo            & /uL/ - /oxo/             & 21.74 & 8.60 \\
-got - vaso           & /gOt/ - /baso/           & 20.00 & 8.00 \\
-entrepa - bocadillo  & /ent45pa/ - /bokaDiLo/   & 13.04 & 7.02 \\
-mirall - espejo      & /mi4aL/ - /espexo/       & 12.50 & 6.75 \\
+fulla - hoja         & \textipa{/"fu.}\textlambda\textipa{@/ - /"o.xa/}                                 & 30.43 & 9.59 \\
+ull - ojo            & \textipa{/u}\textlambda\textipa{/ - /"o.xo/}                                     & 21.74 & 8.60 \\
+got - vaso           & \textipa{/"gOt/ - /"ba.so/}                                                        & 20.00 & 8.00 \\
+entrepa - bocadillo  & \textipa{/ˌen.tR@"pa/ - /bo.ka"di.}\textlambda\textipa{o/}                       & 13.04 & 7.02 \\
+mirall - espejo      & \textipa{/mi"Ra}\textlambda\textipa{/ - /es'pe.xo/}                              & 12.50 & 6.75 \\
 \bottomrule

### 2024-11-15 manuscript/manuscript.qmd
Fix epreds

@@
 library(stringdist)
-
 set.seed(1234)
@@
 tar_load(exp_3_m1)
+tar_load(exp_12_m0)
 
@@
 When some German speakers listened to the song *The Power* by SNAP!
-(*Coyote Ugly*, 2000), many of them misheard the line "I've got the power" as *Agatha Bauer!* When Michael Jackson vocalised "ma-ma-coo-sah" in his song *Wanna Be Startin' Somethin'* (*Thriller*, 1982), some Dutch speakers understood *mama appelsap* [mama applejuice].
+(*Coyote Ugly*, 2000), many of them misheard the line "I've got the power" as *Agatha Bauer*.
 When Spanish speakers listen to the line "Circumvent your thick ego" from the song *Pictures* by System of a Down (*Steal this Album!*, 2002), they tend to hear *Sácame de aquí, King-Kong* [take me out of here, King-Kong].
-When Japanese speakers listen to the line "I want to hold your hand" in the homonym song by The Beatles (*Meet The Beatles*, 1964), they often mishear *Aho na hōnyōhan* [stupid urinator].
-Outrageous as these examples might sound to speakers of other languages, many readers may know a few cases in their own native language.
-This auditory illusion can feel quite real, and often inevitable after the first time it happens.
+Outrageous as these examples might sound, the reader may know a few cases in their own native language.
+This auditory illusion is common across languages and cultures [@dembeck2015oberflachenubersetzung; @efimova2018homophonic] and can feel quite real, and often inevitable.
 In Japanese, this phenomenon takes the name *Soramimi* (lit.
-"empty ear"), after the *Soramimi Hour*---a section of the program *Tamori Club* hosted by comedian Morita Kazuyoshi (Tamori)---in which instances of this illusion were presented with comedy purposes.
-Soramimi is a particular case of *homophonic translation*: words or phrases in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
-This phenomenon extends beyond song lyrics, and may occur anytime when listening to non-native speech.
-For instance, when listening to "You know me?" (\textipa{/ju: n@U mi:}/) in English, Japanese-native listeners tend to understand *yunomi* (/\textipa{jUnomi}/), which is a type of Japanese teacup.
-Overall, the occurrence of Soramimi---and more generally, homophonic translations---suggests that speech in an unfamiliar language has the potential to activate lexical representations in the native language.
-
-Although homophonic translations do not necessarily take place with the preservation of meaning---which often results in outrageous mistranslations like the ones aforementioned---homophonic translations sometimes may lead to correct translations.
-For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
-This listener may encounter the word /\textipa{"pOKt}/ (*porte*), translation of door in French.
-When comparing the phonemic transcription of the French to that of its Spanish translation /\textipa{"pweR.ta}/ (*puerta*), one may be led to infer that these two words do not have much overlap.
-But one thing to note is that the voiced uvular fricative consonant /\textipa{K}/ and the open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /\textipa{R}/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
-Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /\textipa{"pweR.ta}/ (*puerta*) from hearing the word-form /\textipa{"pOKt}/ (*porte*), ultimately allowing them to access the correct semantic representation.
+"empty ear").
+Soramimi is a particular case of homophonic translation in which words or phrases that appear in the lyrics of a song in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
+Homophonic translation also occurs outside the context of music.
+For instance, it is often intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
+For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
+Although homophonic translation is mostly seen as a perceptual curiosity or as a literary figure, little attention has been paid to homophonic translation as a spontanous psycholinguistic event.
+
+@otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show *Soramimi hour*, in which examples of Soramimi were presented the the audience for comedy purposes.
+These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
+The vast majority of the reported instances were phrases (96%) and few were single words (4%).
+There was relatively high variability in the degree to which phonetic features of the presented English lyrics were preserved, with some resulting Japanese words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
+An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /\textipa{kraI}/ to *kurai*, /\textipa{kuRai}/ [dark]), deletion (e.g., *go*, /\textipa{goU}/ to *go*, /go/ [go], and alternation (e.g., *low*, /\textipa{loU}/ to *rou*, /\textipa{Roo}/ [wax]).
+This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
+
+In the present work, we investigated the interplay between phonological similarity and phonological neighbourhood density to examine how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
+Although homophonic translations do not necessarily take place with the preservation of meaning---which results in outrageous mistranslations like the aforementioned ones---homophonic translations sometimes may lead to correct translations.
+For example, imagine an English native listening to Dutch for the first time.
+This listener may encounter the word /\textipa{"pINgUIn}/ (*pinguin*), which sounds similar to its English translation /\textipa{"pENgwIn}/ (penguin).
 Given enough phonological similarity between the non-native word and its translation in the native language, a naïve listener may succeed at word comprehension when listening to an unfamiliar language.
@@
 
-While Soramimi (and more generally homophonic translations) are popular and frequent phenomena, the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
-For instance, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi hour.
-These instances consisted of homophonic translations of English song lyrics to Japanese words and phrases.
-The vast majority of the reported instances were phrases (96%) and a few, to single words (4%).
-The phonetic features of the presented English lyrics were preserved with relatively high variability, with some Japanese resulting words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
-When analysing the few instances of homophonic translations of single words, the author identified three phonological processes that might explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /\textipa{kraI}/ to *kurai*, /\textipa{kuRai}/ [dark]), deletion (e.g., *go*, /\textipa{goU}/ to *go*, /go/ [go], and alternation (e.g., *low*, /\textipa{loU}/ to *rou*, /\textipa{Roo}/ [wax]).
-This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
-Native prosody also constrains the occurence of homophonic translations.
+When comparing the phonemic transcription of the French to that of its Spanish translation /\textipa{"pweR.ta}/ (*puerta*), one may be led to infer that these two words do not have much overlap.
+But one thing to note is that the voiced uvular fricative consonant /\textipa{K}/ and the open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /\textipa{R}/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
+Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /\textipa{"pweR.ta}/ (*puerta*) from hearing the word-form /\textipa{"pOKt}/ (*porte*), ultimately allowing them to access the correct semantic representation.
+Native prosody also constrains the occurrence of homophonic translations.
 For instance, @kentner2015rhythmic presented in the auditory modality French-native and German-native participants with English songs with lyrics.
 Homophonic translations followed language-specific segmentation strategies.
-Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries *before* stressed syllables (following the more frequent trochaic patter in German).
-On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries *after* stressed syllables, following the more frequent iambic pattern in French.
-
-These studies have addressed the phonological aspects of homophonic translation, but little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
-In the present work, we investigated the interplay between phonological similarity and phonological neighbourhood density in a non-native speech perception task.
-In particular, we studied how the presence of phonological neighbours in the native language lexicon affects the extent to which listeners are able to activate the correct translation of an auditorily presented word in thier native language.
-We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
+Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries before stressed syllables (following the more frequent trochaic pattern in German).
+On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries after stressed syllables, following the more frequent iambic pattern in French.
+
+To explore the lexical mechanisms triggered by unfamiliar speech listening, we used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess the words translations in their native language.
 We will henceforth refer to the auditory-presented words heard by participants on each trial as *presented words*, and the correct translation for the presented words as *target translations*.
-We explored listeners’ reliance on phonological similarity to succeed in the task by manipulating the amount of phonological similarity between the presented words and their target translations.
+By manipulating the amount of phonological similarity between the presented words and their target translations, we explored listeners' reliance on phonological similarity to succeed in the task.
 Since participants were unfamiliar with the presented language, they could only use phonological similarity between the presented language and their native language to successfully translate the words.
@@
 Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
-This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing].
-Since the interference effects of phonological neighbourhood density has also been reported across languages [@weber2004lexical], we hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
+This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target word [@luce1998recognizing].
+Such interference effect of phonological neighbourhood density has also been reported across languages [@weber2004lexical].
+We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
 In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
@@
 
-While Soramimi, and more generally homophonic translation, are popular and frequent phenomena, the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
+Soramimi, and more generally homophonic translation, are popular and frequent phenomena, but the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
 For instance, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi hour.
@@
 
-While the previous studies have address the phonological aspects of homophonic translation, little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
+While the previous studies have addressed the phonological mechanisms behind homophonic translation, little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
 In the present work we investigated the interplay between phonological similarity and phonological neighbourhood density to examine how listeners of an unfamiliar language activate lexical representations in their native language.
 In particular, we studied how the presence of phonological neighbours in the native language lexicon affects the extent to which listeners are able to activate the correct translation of an auditorily presented word in an unfamiliar language.
-We used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
+
+Non-native speech perception is a vastly studied subject [@best1988examination; @miyawaki1975effect; @pallier2001influence; @weber2004lexical; @van2015learning; see @baese2020perception for review], but most of this research considers the case of participants with some degree of familiarity with the non-native language, characterising how bilinguals recognise, translate, and produce words as a function of their proficiency in each of their languages [e.g., @kroll1994category; @dijkstra2019multilink; @dijkstra2002architecture; @christoffels2006memory].
+Even low-proficiency bilinguals are likely to have already acquired many form-meaning mappings in their lower proficiency language.
+Characterising the impact of phonological similarity on word processing is therefore challenging in bilinguals, due to the simultaneous contribution of phonological *and* conceptual links in the lexicon.
+By testing participants with *no* prior familiarity with the presented language (i.e., no proficiency), we ensured that no form-meaning associations were available for participants to correctly translate the unfamiliar words, if not by relying exclusively on phonological similarity.
+
+In this study, we used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
 We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
@@
 This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing; @dufour2010phonological].
-Since the interference effects of phonological neighbourhood density has also been reported across languages [@weber2004lexical], we hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
-In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
+Since the interference effects of phonological neighbourhood density has also been reported across languages [@weber2004lexical].
+The presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
 For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
@@
 This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
+
+We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours that the the presented word has in the target language.
 For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
@@
 If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.
+
 We conducted a series of three experiments to test these predictions.
-
 In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
@@
 In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, while English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
 By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant’s performance.
@@
 
-In each trial, participants listened to one audio file, wich contained a single word.
+In each trial, participants listened to one audio file, which contained a single word.
 The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
@@
 
-For each word in the Catalan and Spanish lists pair, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word (in Catalan or Spanish) and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
-*Frequency* was included as a nuance predictor, under the hypothesis that---keeping other predictors constant---participants would translate higher-frequency words more accurately and faster than lower-frequency words.
+For each word in the Catalan and Spanish lists pair, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
+*Frequency* was included as a nuance predictor, under the hypothesis that---keeping other predictors constant---participants would provide correct responses more often when the correct translations are high frequency words in their native language.
 Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores.
-Words in the stimuli list without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
+Translations without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
 
@@
 The Levenshtein distance computes the edit distance between two character strings---in this case, two phonological transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
-We divided this edit distance by the number of characters included in the longest X-SAMPA transcription of the translation pair.
-This results in a proportion score, in which values closer to zero correspond to lower Levenshtein distances between phonological transcriptions (i.e., higher similarity), and values closer to 1 correspond to higher Levenshtein distances (i.e., lower similarity).
-This transformation accounts for the fact that the Levenshtein distance tends to increase with the length of the transcriptions.
+We divided this edit distance by the number of characters in the longest X-SAMPA transcription of the translation pair.
+This transformation accounts for the fact that Levenshtein distance tends to increase with the length of the transcriptions.
 For interpretability, we subtracted this proportion from one, so that values closer to one correspond to higher similarity between phonological transcriptions, and values closer to zero correspond to lower similarity between phonological transcriptions.
-For example, the *table* (`teIb@l`)-*mesa* (`mesa`) translation pair had a `r scales::percent(stringdist::stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (`tREn`) translation pair had a `r scales::percent(stringdist::stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+For example, the *table* (`teIb@l`)-*mesa* (`mesa`) translation pair had `r scales::percent(stringdist::stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (`tREn`) translation pair had `r scales::percent(stringdist::stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
+
+For each Catalan and Spanish word, we calculated the number of *CLPN* by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) differed in up to one phoneme from that of the presented Catalan or Spanish word.
+Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
 @tbl-stimuli summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
-
-For each Catalan and Spanish word, we calculated the number of *CLPN* by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) different in up to one phoneme from that of the presented Catalan or Spanish word.
-Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
 
@@
 Then, participants completed the experimental task.
-Participants were informed that they would listen to a series of pre-recorded words in either Catalan or Spanish (English participants) or Catalan (Spanish participants).
-They were instructed to listen to each word, guess its meaning in English (English participants) or Spanish (Spanish participants), and type their answer as soon as possible.
-English participants were randomly assigned to the Catalan or Spanish lists.
+Participants were informed that they would listen to a series of pre-recorded words in either Catalan or Spanish.
+They were instructed to listen to each word, guess its meaning in English, and type their answer as soon as possible.
+Participants were randomly assigned to the Catalan or Spanish lists.
 Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="cat-ENG",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="spa-ENG",])` trials.
@@
 Upon the offset of the fixation point and audio, participants were prompted to write their answer by a `>` symbol.
-Typed letters were displayed in the screen in real time to provide visual feed-back to participants.
+Typed letters were displayed on the screen in real time to provide visual feed-back to participants.
 Participants were allowed to correct their answer.
-Then, participants pressed the `RETURN/ENTER` key to confirm their answer and start and new trial.
+Then, participants pressed the `RETURN/ENTER` key to confirm their answer and start a new trial.
 
@@
 After the offset of the audio, the fixation point disappeared and a visual prompt (`>`) was presented.
-Participants then wrote their and clicked `RETURN`, making the end of the trial.](_assets/img/design.png)
+Participants then wrote their answer and clicked `RETURN`, making the end of the trial.](_assets/img/design.png)
 
@@
 A response was coded as *Correct* if the provided string of characters was identical to the orthographic form of the correct translation.
-A response was coded as *Typo* if the participant provided a string of characters only one edit distance (addition, deletion, or substitution) apart from the orthographic form of the correct translation (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct English word.
-A response was coded as *False friend* if the participant provided a phonologically similar incorrect translation.
+A response was coded as *Typo* if the participant provided a string of characters that was only one edit distance (addition, deletion, or substitution) apart from the orthographic form of the correct translation (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct English word.
+A response was coded as *False friend* if the participant's response was incorrect, but phonologically similar to the presented word.
 Responses not meeting the criteria for previous categories were labelled as *Wrong* or *Other* (see Data analysis section for more details).
@@
 Trials in which participants took longer than 10 seconds to respond were also excluded.
-Participants contributed a total of `r format(nrow(exp_responses), big.mark = ",")` valid trials (`r format(nrow(exp_responses[exp_responses$group %in% c("cat-ENG", "cat-SPA"),]), big.mark = ",")` in Catalan, `r format(nrow(exp_responses[exp_responses$group %in% c("spa-ENG"),]), big.mark = ",")` in Spanish).
-The task took approximately 15 minutes to be completed.
+Participants contributed a total of `r format(nrow(dplyr::filter(exp_responses, valid_status=="Valid", valid_response, group %in% c("cat-ENG", "spa-ENG"))), big.mark = ",")` valid trials (`r format(nrow(dplyr::filter(exp_responses, valid_status=="Valid", valid_response, group %in% c("cat-ENG"))), big.mark = ",")` in Catalan, `r format(nrow(dplyr::filter(exp_responses, valid_status=="Valid", valid_response, group %in% c("spa-ENG"))), big.mark = ",")` in Spanish).
+The task took approximately 15 minutes to complete.
 
@@
 We included as fixed effects the intercept, the main effects of *Frequency*, *Similarity*, and *CLPN*, and the two-way interaction between *Similarity* and *CLPN*.
-We also included participant-level random intercepts and slopes for the the main effects and the interaction.
+We also included participant-level random intercepts and slopes for the main effects and the interaction.
 Eq.
@@
 We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` unique participants.
-Of those trials, `r format(nrow(dataset$cat_eng), big.mark = ",")` were provided by `r n_participants$cat_eng` unique participants who listened to Catalan words, and `r format(nrow(dataset$spa_eng), big.mark = ",")` trials were provided by `r n_participants$spa_eng` unique participants who listened to Spanish words.
 We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_eng["blank"]`), in which a response in a language other than English was provided (e.g., "`agua`", *n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (e.g., "`f`", *n* = `r r_validity$cat_eng["incomplete"]`), and in which participants added comments to the experimenter (e.g., "`unsure`", *n* = `r r_validity$cat_eng["other"]`).
@@
 ```{r coefs-1}
+#| label: coefs-1
 c1 <- gather_draws(exp_1_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
@@
 Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
-This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c1$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_lv$.lower, 3)`, `r round(c1$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_lv$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c1$b_lv$.median, 3)`, 95% CrI = [`r round(c1$b_lv$.lower, 3)`, `r round(c1$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h$.lower, 3)`, `r round(c1$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h$.rope, 3)`).
+This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c1$b_neigh_n_h_std_lv_std$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_std_lv_std$.lower, 3)`, `r round(c1$b_neigh_n_h_std_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_std_lv_std$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c1$b_lv_std$.median, 3)`, 95% CrI = [`r round(c1$b_lv_std$.lower, 3)`, `r round(c1$b_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv_std$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h_std$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_std$.lower, 3)`, `r round(c1$b_neigh_n_h_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_std$.rope, 3)`).
 @fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
@@
 #| label: fig-epreds-1
-#| cache: true
+#| cache: false
 #| fig-height: 5
@@
 Lines indicate mean predictions, and intervals indicate 95%, 89%, 78%, 67%, and 50% credible intervals (CrI)."
-get_epreds <- function(model, data, n = 1000,
+get_epreds <- function(model, data, n = 100,
                        lv = seq(0, 1, length.out = 100),
@@
                        ...) {
+  lv_std <- (lv - mean(data$lv, na.rm = TRUE)) / sd(data$lv, na.rm = TRUE)
+  neigh_n_h_std <- (neigh_n_h - mean(data$neigh_n_h, na.rm = TRUE)) / sd(data$neigh_n_h, na.rm = TRUE) 
   knowledge <- unique(model$data$knowledge)
   confidence <- unique(model$data$confidence)
-  freq_zipf_2 <- mean(data$freq_zipf_2, na.rm = TRUE)
+  freq_zipf_2_std <- 0
   group <- unique(model$data$group)
-  nd <- expand_grid(freq_zipf_2, neigh_n_h, lv, knowledge, confidence, group)
+  experiment <- unique(model$data$experiment)
+  nd <- expand_grid(freq_zipf_2_std, neigh_n_h_std, lv_std, knowledge, confidence, group, experiment)
   epreds <- tidybayes::add_epred_draws(nd, model, re_formula = NA, ...)
-
   return(epreds)
@@
 
-get_epreds(exp_1_m0, dataset_1) |>
-  mutate(
-    neigh_n_h = paste0(neigh_n_h, " neighbours"),
-    neigh_n_h = factor(neigh_n_h,
-      levels = paste0(
-        c(0, 2, 4, 8, 12),
-        " neighbours"
-      ),
-      ordered = TRUE
-    )
-  ) |>
-  ggplot(aes(lv, .epred)) +
-  facet_grid(group ~ neigh_n_h) +
+lv <- seq(0, 1, length.out = 100)
+lv_std <- (lv - mean(dataset_1$lv_std, na.rm = TRUE)) / sd(dataset_1$lv_std, na.rm = TRUE)
+neigh_n_h <- c(0, 2, 4, 8, 12)
+epreds_1 <- get_epreds(exp_1_m0, dataset_1, neigh_n_h = neigh_n_h, lv = lv) 
+epreds_1$neigh_n_h_std <- factor(epreds_1$neigh_n_h_std, levels = unique(epreds_1$neigh_n_h_std), labels = paste0(neigh_n_h, " neighbours"), ordered = TRUE)
+
+epreds_1 |> 
+  ggplot(aes(lv_std, .epred)) +
+  facet_grid(group ~ neigh_n_h_std) +
   stat_lineribbon(aes(fill_ramp = after_stat(level)),
@@
   labs(
-    x = "Similarity\n(Levenshtein distance with correct translation)",
+    x = "Similarity\n(Levenshtein similarity with correct translation)",
     y = "p(Correct)",
@@
   ) +
-  scale_x_continuous(labels = scales::percent) +
+  scale_x_continuous(
+    labels = \(x) scales::percent((x * sd(dataset_1$lv, na.rm = TRUE)) + mean(dataset_1$lv, na.rm = TRUE)), 
+    breaks = (seq(0, 1, 0.2) - mean(dataset_1$lv, na.rm = TRUE)) / sd(dataset_1$lv, na.rm = TRUE)) +
   scale_y_continuous(labels = scales::percent) +
@@
 English, a Germanic language, is relatively distant from Catalan and Spanish, two Romance languages.
-English shares fewer cognates with Catalan and Spanish than it does with typologically closer languages, like Dutch and German.
+In comparison, Catalan and Spanish are typologically close languages that share many more cognates.
 In Experiment 2, we investigated whether listeners of an unfamiliar but typologically closer language benefit more strongly from phonological similarity when performing the same task as in Experiment 1.
@@
 Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
-This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c2$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_lv$.lower, 3)`, `r round(c2$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_lv$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c2$b_lv$.median, 3)`, 95% CrI = [`r round(c2$b_lv$.lower, 3)`, `r round(c2$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c2$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c2$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h$.lower, 3)`, `r round(c2$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h$.rope, 3)`).
-@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c2$b_neigh_n_h_std_lv_std$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_std_lv_std$.lower, 3)`, `r round(c2$b_neigh_n_h_std_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_std_lv_std$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c2$b_lv_std$.median, 3)`, 95% CrI = [`r round(c2$b_lv_std$.lower, 3)`, `r round(c2$b_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_lv_std$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c2$b_neigh_n_h_std$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_std$.lower, 3)`, `r round(c2$b_neigh_n_h_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_std$.rope, 3)`).
+@fig-epreds-2 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
 
@@
 #| label: fig-epreds-2
-#| cache: true
+#| cache: false
 #| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
 Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Similarity* (1-100).
-Predictions are plotted separately for English participants translating Catalan words, and for English participants translating Spanish words.
 Lines indicate mean predictions, and intervals indicate 95%, 89%, 78%, 67%, and 50% credible intervals (CrI)."
@@
 #| fig-width: 10
-get_epreds(exp_2_m0, dataset_2) |>
-  mutate(
-    neigh_n_h = paste0(neigh_n_h, " neighbours"),
-    neigh_n_h = factor(neigh_n_h,
-      levels = paste0(
-        c(0, 2, 4, 8, 12),
-        " neighbours"
-      ),
-      ordered = TRUE
-    )
-  ) |>
-  ggplot(aes(lv, .epred)) +
-  facet_wrap(~neigh_n_h, nrow = 1) +
+lv <- seq(0, 1, length.out = 100)
+lv_std <- (lv - mean(dataset_2$lv_std, na.rm = TRUE)) / sd(dataset_2$lv_std, na.rm = TRUE)
+neigh_n_h <- c(0, 2, 4, 8, 12)
+epreds_2 <- get_epreds(exp_2_m0, dataset_2, neigh_n_h = neigh_n_h, lv = lv) 
+epreds_2$neigh_n_h_std <- factor(epreds_2$neigh_n_h_std, levels = unique(epreds_2$neigh_n_h_std), labels = paste0(neigh_n_h, " neighbours"), ordered = TRUE)
+
+epreds_2 |> 
+  ggplot(aes(lv_std, .epred)) +
+  facet_grid(~ neigh_n_h_std) +
   stat_lineribbon(aes(fill_ramp = after_stat(level)),
     linewidth = 1 / 2,
-    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)
+    .width = c(0.95, 0.89, 0.78, 0.67, 0.50)
   ) +
   labs(
-    x = "Similarity\n(Levenshtein distance with correct translation)",
+    x = "Similarity\n(Levenshtein similarity with correct translation)",
     y = "p(Correct)",
@@
   ) +
-  scale_x_continuous(labels = scales::percent) +
+  scale_x_continuous(
+    labels = \(x) scales::percent((x * sd(dataset_2$lv, na.rm = TRUE)) + mean(dataset_2$lv, na.rm = TRUE)), 
+    breaks = (seq(0, 1, 0.2) - mean(dataset_2$lv, na.rm = TRUE)) / sd(dataset_2$lv, na.rm = TRUE)) +
   scale_y_continuous(labels = scales::percent) +
@@
 
+```{r posterior-12}
+c12 <- gather_draws(exp_12_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
+  summarise(across(.value, lst(
+    .median = median,
+    .lower = \(x) hdi(x)[1],
+    .upper = \(x) hdi(x)[2],
+    .rope = \(x) mean(between(x, -0.1, 0.1))
+  ),
+  .names = "{fn}"
+  ))
+c12 <- split(c12, c12$.variable)
+names(c12) <- janitor::make_clean_names(names(c12))
+```
+
+In order to compare the results from Experiments 1 and 2 directly, we fit a model on the joint datasets from both experiments.
+This model included all predictors from the models presented in Experiments 1 and 2, and the predictor *Experiment* with levels `"Experiment 1"` and `"Experiment 2"`, and a three-way interaction between *Experiment*, *Similarity*, and *CLPN*.
+The *Experiment* variable was sum-coded as `-0.5` = `"Experiment 1"` and `+0.5` = `"Experiment 2"` [@schad2020capitalize].
+All two-way interactions between the three predictors were also included.
+The effect of *CLPN* on *Similarity* was mostly equivalent in both groups, as indicated by the coefficient of the three-way way interaction ($\beta$ = `r round(c12$b_neigh_n_h_std_lv_std_experiment1$.median, 3)`, 95% CrI = [`r round(c12$b_neigh_n_h_std_lv_std_experiment1$.lower, 3)`, `r round(c12$b_neigh_n_h_std_lv_std$.upper, 3)`]).
+The posterior probability of this coefficient being equivalent to zero (*p*(ROPE)) wwas `r round(c12$b_neigh_n_h_std_lv_std_experiment1$.rope, 3)`), that is, inconclusive.
+The coefficient of the interaction term between *Experiment* and *CLPN* was also equivalent to zero ($\beta$ = `r round(c12$b_neigh_n_h_std_experiment1$.median, 3)`, 95% CrI = [`r round(c12$b_neigh_n_h_std_experiment1$.lower, 3)`, `r round(c12$b_neigh_n_h_std_experiment1$.upper, 3)`], *p*(ROPE) = `r round(c12$b_neigh_n_h_std_experiment1$.rope, 3)`), suggesting that participants from Experiments 1 and 2 were affected by CLPN on a similar basis.
+Finally, the coefficient of the interaction between *Experiment* and *Similarity* was different from zero ($\beta$ = `r round(c12$b_neigh_n_h_std_experiment1$.median, 3)`, 95% CrI = [`r round(c12$b_lv_std_experiment1$.lower, 3)`, `r round(c12$b_lv_std_experiment1$.upper, 3)`], *p*(ROPE) = `r round(c12$b_lv_std_experiment1$.rope, 3)`), with a positive sign indicating that the increments in *Similarity* were associated with a larger increment in probability of correct translation in Experiment 2, compared to in Experiment 1: for an average value of *CLPN*, participants in Experiment 2 benefited more strongly from *Similarity* than participants in Experiment 1.
+
 ## Discussion
@@
 Our results indicate a similar pattern of results as those in Experiment 1: participants were able to provide correct translations of presented Catalan words, provided that the Catalan words shared some degree of phonological similarity with their Spanish translation, and that the number of phonological neighbours with higher lexical frequency was reduced.
-In contrasts with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 was more resilient to the interference of phonological neighbourhood size.
-English natives in Experiment 1 provided significantly less accuracy responses when more than four phonological neighbors were present (even when translating high-similarity words), compared to when only one or neighbour were present.
-Spanish participants in Experiment 2 benefited from phonological similarity, even when eight neighbours were present.
-Spanish participants' performance declined after 8 neighbours, and was evident at 12 neighbours.
-Overall, this suggests that participants in Experiment 2, who were natives of a typologically similar language (Spanish) to the presented language (Catalan) benefited more strongly from phonological similarity than participants in Experiment 1, who were natives of typologically less similar language (English) to the presented language (Catalan, Spanish).
+In contrast with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 was more resilient to the interference of phonological neighbourhood size.
+English natives in Experiment 1 provided significantly less accurate responses when more than four phonological neighbors were present (even when translating high-similarity words), compared to when less than three neighbours were present.
+Spanish participants in Experiment 2 benefited from phonological similarity even when eight neighbours were present.
+Spanish participants' performance declined after eight neighbours, and was most evident at 12 neighbours.
+Overall, this suggests that participants in Experiment 2, who were natives of a typologically similar language (Spanish) to the presented language (Catalan), benefited more strongly from phonological similarity than participants in Experiment 1, who were natives of typologically less similar language (English) to the presented language (Catalan, Spanish).
 
@@
 Given that participants were reportedly unfamiliar with both languages, it was expected that participants would be very unlikely to provide correct translations for words sharing little to no phonological similarity to their correct translation.
-@tbl-surprises a list of Catalan and Spanish words to which participants provided responses with $\geq$ 10 average accuracy.
+@tbl-surprises shows a list of Catalan and Spanish words to which participants provided responses with $\geq$ 10 average accuracy.
+
+::: {.content-visible when-format="pdf"}
 
@@
 #| label: tbl-surprises
-#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarity score between the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
 ipa_to_tipa <- function(ipa) {
@@
 ```
+:::
+
+::: {.content-visible when-format="docx"}
+
+```{r tbl-surprises-docx}
+#| label: tbl-surprises-docx
+#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+tbl_data <- list(
+  "Experiment 1" = dataset_1,
+  "Experiment 2" = dataset_2
+) |>
+  bind_rows(.id = "exp") |>
+  mutate(
+    translation = paste0(word_1, " - ",  word_2),
+    ipa = paste0(ipa_1, " - ",  ipa_2),
+  ) |> 
+  summarise(
+    n = n(),
+    correct = sum(correct),
+    .by = c(exp, group, lv, translation, ipa)
+  ) |>
+  mutate(
+    accuracy = prop_adj(correct, n),
+    accuracy_se = prop_adj_se(correct, n)
+  ) |>
+  dplyr::filter(lv == 0, accuracy >= 0.1) |>
+  select(exp, translation, ipa, group, accuracy, accuracy_se) |>
+  arrange(exp, group, desc(accuracy)) |> 
+  select(translation, ipa, accuracy, accuracy_se) |> 
+  mutate(accuracy = round(accuracy * 100, 2),
+         accuracy_se = round(accuracy_se * 100, 2))
+
+names(tbl_data) <- c("Translation", "IPA", "Accuracy (\\%)", "SE")
+
+tt(tbl_data) |> 
+  style_tt(i = 0, bold = TRUE, line = "tb") |> 
+  group_tt(i = list(
+    "Experiment 1 (cat-ENG)" = 1,
+    "Experiment 2 (spa-ENG)" = 6,
+    "Experiment 3 (cat-SPA)" = 22)) |> 
+  style_tt(i = 1, bold = TRUE, line = "b") |> 
+  style_tt(i = 7, bold = TRUE, line = "tb") |> 
+  style_tt(i = 24, bold = TRUE, line = "tb")
+```
+
+:::
 
@@
 In the case of higher-similarity words, it is more difficult to disentangle the extent to which participants' accuracy is a function of pure phonological similarity, or prior knowledge they had about the meaning of Spanish words.
-Experiment 3 was addressed at investigating this issue.
+Experiment 3 was addressed at investigating the issue of prior knowledge.
 
@@
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
-Stimuli were the same list of Catalan stimuli as in Experiment 1.
+Stimuli were the same lists of Catalan and Spanish stimuli as in Experiment 1.
 
@@
 These questions were presented on the same page, directly below the audio player.
-They were first asked whether or not they knew the presented word (multiple choice---*yes*/*no*).
+They were first asked whether they knew the presented word (multiple choice---*yes*/*no*).
 Regardless of their answer on the first question, participants were asked what they thought the translation of the word was in English (or their best guess), and instructed to type their answer in the provided text box.
@@
 Participants first completed 5 practice trials with English words as the audio stimulus (ambulance, cucumber, elephant, pear, turtle).
-The words were recorded by a female native speaker of English.
+The words were recorded by a female native speaker of British English.
 These trials acted as attention checks, as participants should always answer "yes" to the first question on prior word knowledge and be able to accurately transcribe the word they heard.
@@
 We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` unique participants.
-Of those trials, `r format(nrow(dataset$cat_eng), big.mark = ",")` were provided by `r n_participants$cat_eng` unique participants who listened to Catalan words, and `r format(nrow(dataset$spa_eng), big.mark = ",")` trials were provided by `r n_participants$spa_eng` unique participants who listened to Spanish words.
-We excluded trials in which participants did not enter any text (*n* = 0), in which a response in a language other than English was provided (*n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (*n* = 0), and in which participants added comments to the experimenter (*n* = 0).
+We excluded trials in which participants did not enter any text (*n* = 0), provided a response in a language other than English (*n* = `r r_validity$cat_eng["language"]`), did not provide a whole word (*n* = 0), or added comments to the experimenter (*n* = 0).
 In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
@@
 In those responses, participants reported an average of `r round(fam_cateng$confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(fam_cateng$confidence_sd, 2)`).
-In responses in which no prior knowledge of the presented word was reported, average confidence was `r round(nofam_cateng $confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(nofam_cateng$confidence_sd, 2)`).
+For responses where no prior knowledge of the presented word was reported, average confidence was `r round(nofam_cateng $confidence_mean, 2)` (*SD* = `r round(nofam_cateng$confidence_sd, 2)`).
 From the `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` total responses provided by English participants who listened to Spanish words, participants reported having prior knowledge of the presented Spanish words in `r knowledge_spaeng$n[2]` (`r round(100*knowledge_spaeng$n[2] / sum(knowledge_spaeng$n), 2)`%) of them.
 In those responses, participants reported an average of `r round(fam_spaeng$confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(fam_spaeng$confidence_sd, 2)`).
-In responses in which no prior knowledge of the presented word was reported, average confidence was `r round(nofam_spaeng$confidence_mean, 2)` confidence in the 0-8 scale (*SD* = `r round(nofam_spaeng $confidence_sd, 2)`).
-Before data analysis, responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word were excluded from the dataset.
+For responses where no prior knowledge of the presented word was reported, average confidence was `r round(nofam_spaeng$confidence_mean, 2)` (*SD* = `r round(nofam_spaeng $confidence_sd, 2)`).
+Before data analysis, responses where participants reported prior knowledge about the meaning of the presented Catalan or Spanish word were excluded from the dataset.
 
@@
 MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
-Overall, participants reported prior knowledge more often for that Spanish words with unexpectedly high accuracy (see Discussion in Experiment 2) than for words with expected accuracy (see @fig-knowledge).
-Participants reported prior knowledge of Catalan words with unexpected accuracy as often as those with expected accuracy.
+Overall, participants reported prior knowledge more often for the Spanish words that showed unexpectedly high accuracy in Experiment 1 (see Discussion in Experiment 2) than for words with expected accuracy (see @fig-knowledge).
+Participants reported prior knowledge of Catalan words with unexpected accuracy in Experiment 1 as often as those with expected accuracy.
 This suggests that participants in Experiment 1 may have relied, to some extent, on their prior knowledge about form-meaning mappings to correctly translate some Spanish words.
-To isolate such an effect of prior Spanish knowledge, we run the same analysis as in Experiment 1 on the newly collected translations from Experiment 3, now excluding responses to words in which participants reported prior knowledge.
+To isolate such an effect of prior Spanish knowledge, we run the same analysis as in Experiment 1 on the newly collected translations from Experiment 3, now excluding responses to words for which participants reported prior knowledge.
 
@@
 
-Participants translating Catalan words and participants translating Spanish words performed similarly, as indicated by the regression coefficient of *Group* ($\beta$ = `r round(c3$b_group1$.median, 3)`, 95% CrI = [`r round(c3$b_group1$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_group1$.rope, 3)`).
+Participants translating Catalan words and participants translating Spanish words performed similarly, as indicated by the regression coefficient of *Group* ($\beta$ = `r round(c3$b_group1$.median, 3)`, 95% CrI = [`r round(c3$b_group1$.lower, 3)`, `r round(c3$b_neigh_n_h_std_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c3$b_group1$.rope, 3)`).
 Overall, both groups of participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
-This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c3$b_neigh_n_h_lv$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h_lv$.lower, 3)`, `r round(c3$b_neigh_n_h_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h_lv$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c3$b_lv$.median, 3)`, 95% CrI = [`r round(c3$b_lv$.lower, 3)`, `r round(c3$b_lv$.upper, 3)`], *p*(ROPE) = `r round(c3$b_lv$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c3$b_neigh_n_h$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h$.lower, 3)`, `r round(c3$b_neigh_n_h$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h$.rope, 3)`).
-@fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
+This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c3$b_neigh_n_h_std_lv_std$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h_std_lv_std$.lower, 3)`, `r round(c3$b_neigh_n_h_std_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h_std_lv_std$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c3$b_lv_std$.median, 3)`, 95% CrI = [`r round(c3$b_lv_std$.lower, 3)`, `r round(c3$b_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c3$b_lv_std$.rope, 3)`), while the number of *CLPN* had the opposite effect ($\beta$ = `r round(c3$b_neigh_n_h_std$.median, 3)`, 95% CrI = [`r round(c3$b_neigh_n_h_std$.lower, 3)`, `r round(c3$b_neigh_n_h_std$.upper, 3)`], *p*(ROPE) = `r round(c3$b_neigh_n_h_std$.rope, 3)`).
+@fig-epreds-3 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
 
@@
 #| label: fig-epreds-3
-#| cache: true
-#| fig-cap: "Posterior model-predicted mean accuracy in Experiment 2.
+#| cache: false
+#| fig-cap: "Posterior model-predicted mean accuracy in Experiment 3.
 Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Similarity* (1-100).
@@
 #| fig-width: 10
-get_epreds(exp_3_m1, dataset_3) |>
-  mutate(
-    neigh_n_h = paste0(neigh_n_h, " neighbours"),
-    neigh_n_h = factor(neigh_n_h,
-      levels = paste0(
-        c(0, 2, 4, 8, 12),
-        " neighbours"
-      ),
-      ordered = TRUE
-    )
-  ) |>
-  ggplot(aes(lv, .epred)) +
-  facet_grid(group ~ neigh_n_h) +
+lv <- seq(0, 1, length.out = 100)
+lv_std <- (lv - mean(dataset_3$lv_std, na.rm = TRUE)) / sd(dataset_3$lv_std, na.rm = TRUE)
+neigh_n_h <- c(0, 2, 4, 8, 12)
+epreds_3 <- get_epreds(exp_3_m1, dataset_3, neigh_n_h = neigh_n_h, lv = lv) 
+epreds_3$neigh_n_h_std <- factor(epreds_3$neigh_n_h_std, levels = unique(epreds_3$neigh_n_h_std), labels = paste0(neigh_n_h, " neighbours"), ordered = TRUE)
+
+epreds_3 |> 
+  ggplot(aes(lv_std, .epred)) +
+  facet_grid(group ~ neigh_n_h_std) +
   stat_lineribbon(aes(fill_ramp = after_stat(level)),
     linewidth = 1 / 2,
-    .width = c(0.99, 0.95, 0.89, 0.78, 0.67, 0.50)
+    .width = c(0.95, 0.89, 0.78, 0.67, 0.50)
   ) +
   labs(
-    x = "Similarity\n(Levenshtein distance with correct translation)",
+    x = "Similarity\n(Levenshtein similarity with correct translation)",
     y = "p(Correct)",
+    colour = "Cross-language neighbourhood density",
+    fill = "Cross-language neighbourhood density",
+    linetype = "Cross-language neighbourhood density",
     fill_ramp = "CrI (%)"
   ) +
-  scale_x_continuous(labels = scales::percent) +
+  scale_x_continuous(
+    labels = \(x) scales::percent((x * sd(dataset_3$lv, na.rm = TRUE)) + mean(dataset_3$lv, na.rm = TRUE)), 
+    breaks = (seq(0, 1, 0.2) - mean(dataset_3$lv, na.rm = TRUE)) / sd(dataset_3$lv, na.rm = TRUE)) +
   scale_y_continuous(labels = scales::percent) +
@@
   )
+
 ```
@@
 After listening to each word, participants were asked to provide their best-guess translation for each word in their native language.
-We run three experiments.
 In Experiment 1, British English-native adults listened to a list of words in Catalan or Spanish.
 Participants reported no prior familiarity with Catalan, Spanish, or any other Romance language, yet provided accurate translations for words that shared some degree of phonological similarity with their correct English translation.
-Participants only benefited from phonological similarity when the number of English phonological neighbours of the presented Catalan or Spanish word was small.
+We calculated phonological similarity between word pairs as the Levenshtein similarity between the X-SAMPA transcription of their phonological forms [see @floccia2018introduction for a similar approach].
+Using this measure, we found that participants were able to exploit the phonological similarity between the presented words and their correct translation, even when both words-forms shared few phonemes.
+This points to listeners accommodating non-native phonological forms to their native phoneme inventory, resulting in the activation of native lexical representations.
+
 Experiment 2 was aimed at extending our findings to a population of participants whose native language was typologically closer to the unfamiliar language presented.
@@
 Again, participants reported no prior familiarity with Catalan, or any other Romance language other than Spanish.
-In line with the results in English natives, we found an interaction between phonological similarity and phonological neighbourhood size.
-The facilitation effect of phonological similarity in the Spanish participants was more robust to the interferring effect of phonological neighbourhood density: Spanish participants provided correct translations for words with high neighbourhood densities more often than English participants.
-We suspected that participants in Experiments 1 and 2 might have managed to provide correct translations for some low-phonological similarity words thanks to having prior experience with those words, despite having reported being unfamiliar with Catalan, Spanish, or any other Romance language.
-Experiment 3 was a replica of Experiment 1, in which we collected additional information about participants' prior familiarity with the presented Catalan and Spanish words.
+In line with the results with English natives, we found an interaction between phonological similarity and phonological neighbourhood size.
+The facilitation effect of phonological similarity in the Spanish participants was more robust to the interfering effect of phonological neighbourhood density: Spanish participants provided correct translations for words with high neighbourhood densities more often than English participants.
+Experiment 3 was a replication of Experiment 1, in which we collected additional information about participants' prior familiarity with the presented Catalan and Spanish words.
+This follow-up experiment was designed to address concerns that participants in Experiments 1 and 2 might have managed to provide correct translations for some words thanks to having prior experience with those words.
 The design of Experiment 3 was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
-After removing responses in which participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we run the same analyses as in Experiment 1, finding parallel results.
-Results in Experiment 3 suggest participants in the present study did not simply rely on prior familiarity with some of the word-forms presented.
-Rather, participants relied on the phonological similarity between the presented Catalan and Spanish words and their correct translation.
-
-Homophonic translation is an often spontaneous phenomenon that occurs commonly when listening to non-native speech, and is prevalent across many cultures and languages [@dembeck2015oberflachenubersetzung; @efimova2018homophonic].
-The most famous case of homophonic translation is *Soramimi*, which refers to homophonic translation in the context of music listening, when particular sections of the lyrics of a song in a non-native language make listeners think of particular words and phrases in their own native language.
-This sometimes result in a completely different meaning than thhe originally intended.
-Homophonic translation occurs outside the context of music.
-For instance, it is often intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
-For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
-Although homophonic translation has mostly received attention as a perceptual curiosity or as a literary figure, little attention has been paid to homophonic translation as a spontanous *psycholinguistic* event.
-The psycholiguistic mechanisms behind spontaneous homophonic translations are informative of the mechanisms underlying native speech processing.
-How does the lexical system deal with a non-native speech signal in which form-meaning associations are scarce, if not absent?
-
-How the perceptual system deals with non-native speech is a vastly studied subject [e.g., @best1988examination; @miyawaki1975effect; @pallier2001influence; @weber2004lexical; @van2015learning; see @baese2020perception for review].
-The same cannot be said about how the perceptual system deals with speech in an *unfamiliar* language.
-Previous research had explored how bilinguals recognise, translate, and produce words as a function of their proficiency in each of their languages [e.g., @kroll1994category; @dijkstra2019multilink; @dijkstra2002architecture; @christoffels2006memory].
-The case of a very low-proficiency bilingual may be close to that of a listener of an unfamiliar language, but being bilingual implies some form of proficiency in both languages, such as some knowledge about the meaning of some words.
-Even low-proficiency bilinguals are likely to have already acquired many form-meaning mappings in their lower proficiency language.
-Characterising the impact of phonological similarity on word processing is therefore challenging in bilinguals, due to the simultaneous contribution of phonological *and* conceptual links in the lexicon.
-By testing participants with *no* prior familiarity with the presented language (i.e., no proficiency), we ensured that no form-meaning associations were available for participants to correctly translate the unfamiliar words, if not by relying excluusively on phonological similarity.
-
-We calculated phonological similarity between word pairs as the Levenshtein similarity between the X-SAMPA transcription of their phonological forms [see @floccia2018introduction for a similar approach].
-Using this measure, we found that participants were able to exploit the phonological similarity between the presented words and their correct translation, even when both words-forms shared few phonemes.
-This points to listeners accommodating non-native phonological forms to their native phoneme inventory, resulting in the activation of native lexical representations.
+After removing responses where participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we ran the same analyses as in Experiment 1, and found parallel results.
+Overall, our results suggest that participants were able to rely on the phonological similarity between the presented Catalan and Spanish words and their correct translation to guide word recognition in an unfamiliar language.
+
 Future studies may consider comparing the suitability of other measures of phonological similarity that take into account finer-grained phonetic or prosodic cues present in the acoustic signal presented to participants.
@@
 
-Our findings also suggest that the facilitation effect of phonological familiarity was moderated by phonological neighbourghood density.
-This is in line which previous studies suggesting that the presence of (high-frequency) phonological neighbours interferes with the lexical selection in word comprehension tasks [e.g., @dufour2010phonological; @luce1998recognizing; @vitevitch1999probabilistic; @vitevitch1998words; @grainger1990word].
+Our findings also suggest that the facilitation effect of phonological similarity was moderated by phonological neighbourhood density.
+This is in line with previous studies suggesting that the presence of (high-frequency) phonological neighbours interferes with the lexical selection in word comprehension tasks [e.g., @dufour2010phonological; @luce1998recognizing; @vitevitch1999probabilistic; @vitevitch1998words; @grainger1990word].
 Our translation elicitation task can be understood as a particular instance of an auditory word recognition task, in which the presented unfamiliar words have the potential to activate phonological neighbours in participants' lexicon.
-We calculated a measure of cross-linguistic neighbourhood size by counting, for each presented word (in Catalan or Spanish in Experiments 1 and 3, Catalan in Experiment 2) the number of phonologically related words (with a Levenshtein distance of one) in the target language (English in Experiments 1 and 3, Spanish in Experiment 2).
-To account for the fact that competition between phonological neighbours is sensitive to lexical frequency [high-frequency neighbours produce stronger interference effects, @luce1998recognizing; @dufour2010phonological], we only counted neighbours with a lexical frequency higher than the correct translation.
+We calculated a measure of cross-linguistic neighbourhood size by counting, for each presented word the number of phonologically related words (with a Levenshtein distance of one) in the target language (English in Experiments 1 and 3, Spanish in Experiment 2).
+To account for the fact that competition between phonological neighbours is sensitive to lexical frequency---high-frequency neighbours produce stronger interference effects [@luce1998recognizing; @dufour2010phonological]---we only counted neighbours with a lexical frequency higher than the correct translation.
 Using this measure, we found that participants' ability to exploit phonological similarity to produce correct translations declined as the number of phonological neighbours increased.
-Overall, this suggests that listening to words in an unfamiliar language trigger the same mechanisms of lexical activation and selection that listening to words in a native language does.
-
-A comparison between results in Experiments 1 and 2 suggested that the performance of Spanish participants translating Catalan was more resilient to the interfering effects of phonological neighbourhood density than the performance of English participants translating Catalan or Spanish.
+This suggests that listening to words in an unfamiliar language triggered similar mechanisms of lexical activation and selection that listening to words in a native language does.
+The generalisability of these findings can be tested by increasing the repertoire of words involved in the translation elicitation task.
+In Experiments 1 to 3, participants answered to a maximum of 105 words.
+All words had high frequency and low age-of-acquisition.
+To better characterise the factors that guide listeners’ ability to match words from an unfamiliar language with words in their native language lexicon, words with varying levels of difficulty should be included in future studies.
+
+A comparison between the results in Experiments 1 and 2 suggested that the performance of Spanish participants translating Catalan was more resilient to the interfering effects of phonological neighbourhood density than the performance of English participants translating Catalan or Spanish.
 As highlighted before, Catalan and Spanish (both Romance languages) are typologically closer to each other than to English (a Germanic language).
-Catalan and Spanish share a higher proportion of cognates (i.e., form-similar translation equivalents) than Catalan and English, or Spanish and English.
-Previous studies in bilinguals show that this kind of cross-linguistic similarity impacts lexical processing: words that share high phonological similarity with their translations are processed faster and more accurately than words sharing lower phonological similarity.
-This phenomenon has provided strong evidence in favour of the language-non selective hypothesis of bilingual lexical access, which states that bilinguals activate lexical representations in both languages, even during monolingual situations.
-This facilitation effect has been reported for comprehension [@dijkstra2010cross; @midgley2011effects; @thierry2007brain], production [@costa2000cognate], learning [@de2000hard; @elias2022crosslanguage; @lotto1998effects; @valente2018does], and translation [@christoffels2006memory].
-To our knowledge, this present study is the first to investigate the role of cross-linguistic phonological similarity in a monolingual population, in which the role of conceptual links in the lexicon is ruled out, given participants lack of familiarity with the presented language.
-This provided a convenient case to investigate the role of phonological similarity in lexical processing, disantangling the contribution of phonological similarity from the contribution of conceptual equivalence.
+Catalan and Spanish share a higher proportion of cognates than Catalan and English, or Spanish and English.
 Overall, our findings suggest that the typological distance between the presented and target language is associated with a more robust facilitation effect of phonological similarity.
@@
 The specific mechanisms behind this effect are unclear.
-One possibility is that, as aforementioned, Catalan provided additional cues to Spanish participants in Experiment 2 compared to English participants in Experiment 1.
-For instance, it is likely that subphonemic cues not accounted by our measure of phonological similarity were more informative to Spanish participants than to English participants.
-In addition, it remains to be tested whether such association is linear (i.e., robustness of facilitation increases linearly with typological distance), proportional, or even extends beyond the specific pair of languages involved in the present work remains unclear.
-Do other close or distant language pairs show the same pattern of results?
-The generalisability of these findings can also be tested by increasing the repertoire of words involved in the translation elicitation task.
-In Experiments 1 to 3, participants answered to a maximum of 105 words.
-All words were of high-frequency, and had a low age-of-acquisition.
-In order to characterise listeners' ability to translate correctly words from an unfamiliar language, words with varying levels of difficulty should be included.
+One possibility is that there were subphonemic features of similarity between Catalan and Spanish words that were not adequately captured by our relatively coarse measure of similarity.
+Phonological transcriptions like X-SAMPA are abstract representations that ignore non-phonological contrasts (e.g., phones), and consider different symbols as completely different phonemes, disregarding the fact that some phonemes share more phonetic features than others.
+Additionally, prosodic cues such as lexical stress, which our measure of phonological similarity does not include, might also provide participants with further information about the correct translation of the presented word-forms.
+There may be more of these subphonemic cues available between the typologically close Spanish and Catalan than between the more distant English and Spanish, or English and Catalan.
+In addition, it remains to be tested whether the effect of typological distance is linear (i.e., robustness of facilitation increases linearly with typological distance), proportional, and whether other close or distant language pairs show the same pattern of results.
+Future studies may consider comparing the suitability of other measures of phonological similarity that take into account finer-grained phonetic or prosodic cues present in the acoustic signal, and explore the relative contribution of such cues in the occurrence of homophonic translation.
 

### 2024-11-15 manuscript/manuscript.tex
Fix epreds

@@
 (\emph{Coyote Ugly}, 2000), many of them misheard the line ``I've got
-the power'' as \emph{Agatha Bauer!} When Michael Jackson vocalised
-``ma-ma-coo-sah'' in his song \emph{Wanna Be Startin' Somethin'}
-(\emph{Thriller}, 1982), some Dutch speakers understood \emph{mama
-appelsap} {[}mama applejuice{]}. When Spanish speakers listen to the
+the power'' as \emph{Agatha Bauer}. When Spanish speakers listen to the
 line ``Circumvent your thick ego'' from the song \emph{Pictures} by
@@
 \emph{Sácame de aquí, King-Kong} {[}take me out of here, King-Kong{]}.
-When Japanese speakers listen to the line ``I want to hold your hand''
-in the homonym song by The Beatles (\emph{Meet The Beatles}, 1964), they
-often mishear \emph{Aho na hōnyōhan} {[}stupid urinator{]}. Outrageous
-as these examples might sound to speakers of other languages, many
-readers may know a few cases in their own native language. This auditory
-illusion can feel quite real, and often inevitable after the first time
-it happens. In Japanese, this phenomenon takes the name \emph{Soramimi}
-(lit. ``empty ear''), after the \emph{Soramimi Hour}---a section of the
-program \emph{Tamori Club} hosted by comedian Morita Kazuyoshi
-(Tamori)---in which instances of this illusion were presented with
-comedy purposes. Soramimi is a particular case of \emph{homophonic
-translation}: words or phrases in one language are translated into
-similar-sounding words and phrases in a different language, without
-necessarily preserving the meaning
-(\citeproc{ref-gasparov2006semen}{Gasparov, 2006}). This phenomenon
-extends beyond song lyrics, and may occur anytime when listening to
-non-native speech. For instance, when listening to ``You know me?''
-(\textipa{/ju: n@U mi:}/) in English, Japanese-native listeners tend to
-understand \emph{yunomi} (/\textipa{jUnomi}/), which is a type of
-Japanese teacup. Overall, the occurrence of Soramimi---and more
-generally, homophonic translations---suggests that speech in an
-unfamiliar language has the potential to activate lexical
-representations in the native language.
-
-Although homophonic translations do not necessarily take place with the
-preservation of meaning---which often results in outrageous
-mistranslations like the ones aforementioned---homophonic translations
-sometimes may lead to correct translations. For example, imagine a
-Spanish native with no previous familiarity with any other language
-listening to French for the first time. This listener may encounter the
-word /\textipa{"pOKt}/ (\emph{porte}), translation of door in French.
+Outrageous as these examples might sound, the reader may know a few
+cases in their own native language. This auditory illusion is common
+across languages and cultures
+(\citeproc{ref-dembeck2015oberflachenubersetzung}{Dembeck, 2015};
+\citeproc{ref-efimova2018homophonic}{Efimova et al., 2018}) and can feel
+quite real, and often inevitable. In Japanese, this phenomenon takes the
+name \emph{Soramimi} (lit. ``empty ear''). Soramimi is a particular case
+of homophonic translation in which words or phrases that appear in the
+lyrics of a song in one language are translated into similar-sounding
+words and phrases in a different language, without necessarily
+preserving the meaning (\citeproc{ref-gasparov2006semen}{Gasparov,
+2006}). Homophonic translation also occurs outside the context of music.
+For instance, it is often intentionally employed by translators to
+preserve the meter and ``sound structure'' (e.g., rhymes) of an original
+text in the language of translation (e.g.,
+\citeproc{ref-levick2016translating}{Levick, 2016};
+\citeproc{ref-pilshchikov2016semiotics}{Pilshchikov, 2016})\footnote{The
+  rationale behind this translation approach is to ``sacrifice{[}s{]}
+  fidelity to the image rather than the melodiousness of verse''
+  (\citeproc{ref-briusov1911ot}{Briusov, 1911}). For instance, Edith
+  Grossman's translation of \emph{El Quijote}
+  (\citeproc{ref-cervantes}{Cervantes \& Grossman, 2002}) present a
+  similar case, in which formal aspects of the original text in Spanish
+  are favoured to some degree over equivalence to English, even
+  sometimes keeping Spanish words from the original text in the English
+  translation (e.g., \emph{Señor}, \emph{ínsula}).}. Although homophonic
+translation is mostly seen as a perceptual curiosity or as a literary
+figure, little attention has been paid to homophonic translation as a
+spontanous psycholinguistic event.
+
+Otake (\citeproc{ref-otake2007interlingual}{2007}) analysed 194
+instances of Soramimi, broadcasted between 1992 and 2007 by the TV show
+\emph{Soramimi hour}, in which examples of Soramimi were presented the
+the audience for comedy purposes. These instances consisted of
+homophonic translations of English song lyrics to words and phrases in
+Japanese. The vast majority of the reported instances were phrases
+(96\%) and few were single words (4\%). There was relatively high
+variability in the degree to which phonetic features of the presented
+English lyrics were preserved, with some resulting Japanese words or
+phrases sharing high overlap with the original English strings of
+sounds, and some sharing very little overlap. An analysis of the few
+instances of homophonic translations of single words revealed three main
+phonological processes that could explain how Japanese listeners
+reconstructed the English input to generate the Japanese words:
+insertion (e.g., \emph{cry} /\textipa{kraI}/ to \emph{kurai},
+/\textipa{kuRai}/ {[}dark{]}), deletion (e.g., \emph{go},
+/\textipa{goU}/ to \emph{go}, /go/ {[}go{]}, and alternation (e.g.,
+\emph{low}, /\textipa{loU}/ to \emph{rou}, /\textipa{Roo}/ {[}wax{]}).
+This suggests that the reported homophonic translations can be
+explained, to some extent, as the result of the Japanese listeners
+accommodating the strings of English sounds to the Japanese phoneme
+inventory and Japanese phonotactics
+(\citeproc{ref-dupoux1999epenthetic}{Dupoux et al., 1999};
+\citeproc{ref-peperkamp2008perceptual}{Peperkamp et al., 2008}).
+
+In the present work, we investigated the interplay between phonological
+similarity and phonological neighbourhood density to examine how
+listeners may activate lexical representations in their native language
+when listening to an unfamiliar language. Although homophonic
+translations do not necessarily take place with the preservation of
+meaning---which results in outrageous mistranslations like the
+aforementioned ones---homophonic translations sometimes may lead to
+correct translations. For example, imagine an English native listening
+to Dutch for the first time. This listener may encounter the word
+/\textipa{"pINgUIn}/ (\emph{pinguin}), which sounds similar to its
+English translation /\textipa{"pENgwIn}/ (penguin). Given enough
+phonological similarity between the non-native word and its translation
+in the native language, a naïve listener may succeed at word
+comprehension when listening to an unfamiliar language. Such
+phonological similarity between translation equivalents---known as
+\emph{cognateness}---is common across many languages, and is often due
+to typological closeness and/or socio-historical events involving the
+speakers of these languages (e.g., migration, social contact). For
+example, Romance languages such as Spanish and Catalan share many
+form-similar translation equivalents
+(\citeproc{ref-schepens2012distributions}{Schepens et al., 2012}), as in
+the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and
+Catalan, respectively). Given no prior knowledge of either language, a
+Spanish native speaker is likely to be much more successful at correctly
+translating Catalan \emph{porta} than English \emph{door}, due to the
+phonological and orthographic similarity of the former to the equivalent
+in their native language.
+
 When comparing the phonemic transcription of the French to that of its
@@
 /\textipa{"pOKt}/ (\emph{porte}), ultimately allowing them to access the
-correct semantic representation. Given enough phonological similarity
-between the non-native word and its translation in the native language,
-a naïve listener may succeed at word comprehension when listening to an
-unfamiliar language. Such phonological similarity between translation
-equivalents---known as \emph{cognateness}---is common across many
-languages, and is often due to typological closeness and/or
-socio-historical events involving the speakers of these languages (e.g.,
-migration, social contact). For example, Romance languages such as
-Spanish and Catalan share many form-similar translation equivalents
-(\citeproc{ref-schepens2012distributions}{Schepens et al., 2012}), as in
-the case of \emph{puerta} and \emph{porta} (\emph{door} in Spanish and
-Catalan, respectively). Given no prior knowledge of either language, a
-Spanish native speaker is likely to be much more successful at correctly
-translating Catalan \emph{porta} than English \emph{door}, due to the
-phonological and orthographic similarity of the former to the equivalent
-in their native language.
-
-While Soramimi (and more generally homophonic translations) are popular
-and frequent phenomena, the psycholinguistic literature on the topic is
-relatively scarce, and has mostly focused on the phonological variables
-involved in the occurence of Soramimi. For instance, Otake
-(\citeproc{ref-otake2007interlingual}{2007}) analysed 194 instances of
-Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi
-hour. These instances consisted of homophonic translations of English
-song lyrics to Japanese words and phrases. The vast majority of the
-reported instances were phrases (96\%) and a few, to single words (4\%).
-The phonetic features of the presented English lyrics were preserved
-with relatively high variability, with some Japanese resulting words or
-phrases sharing high overlap with the original English strings of
-sounds, and some sharing very little overlap. When analysing the few
-instances of homophonic translations of single words, the author
-identified three phonological processes that might explain how Japanese
-listeners reconstructed the English input to generate the Japanese
-words: insertion (e.g., \emph{cry} /\textipa{kraI}/ to \emph{kurai},
-/\textipa{kuRai}/ {[}dark{]}), deletion (e.g., \emph{go},
-/\textipa{goU}/ to \emph{go}, /go/ {[}go{]}, and alternation (e.g.,
-\emph{low}, /\textipa{loU}/ to \emph{rou}, /\textipa{Roo}/ {[}wax{]}).
-This suggests that the reported homophonic translations can be
-explained, to some extent, as the result of the Japanese listeners
-accommodating the strings of English sounds to the Japanese phoneme
-inventory and Japanese phonotactics
-(\citeproc{ref-dupoux1999epenthetic}{Dupoux et al., 1999};
-\citeproc{ref-peperkamp2008perceptual}{Peperkamp et al., 2008}). Native
-prosody also constrains the occurence of homophonic translations. For
-instance, Kentner (\citeproc{ref-kentner2015rhythmic}{2015}) presented
-in the auditory modality French-native and German-native participants
-with English songs with lyrics. Homophonic translations followed
-language-specific segmentation strategies. Most homophonic translations
-reported by German-native participants resulted from inserting phrase
-and word boundaries \emph{before} stressed syllables (following the more
-frequent trochaic patter in German). On the contrary, most homophonic
+correct semantic representation. Native prosody also constrains the
+occurrence of homophonic translations. For instance, Kentner
+(\citeproc{ref-kentner2015rhythmic}{2015}) presented in the auditory
+modality French-native and German-native participants with English songs
+with lyrics. Homophonic translations followed language-specific
+segmentation strategies. Most homophonic translations reported by
+German-native participants resulted from inserting phrase and word
+boundaries before stressed syllables (following the more frequent
+trochaic pattern in German). On the contrary, most homophonic
 translations reported by French-native participants resulted from
-inserting phrase and word boundaries \emph{after} stressed syllables,
-following the more frequent iambic pattern in French.
-
-These studies have addressed the phonological aspects of homophonic
-translation, but little attention has been paid to the dynamics of
-lexical activation and competition that underlay this phenomenon. In the
-present work, we investigated the interplay between phonological
-similarity and phonological neighbourhood density in a non-native speech
-perception task. In particular, we studied how the presence of
-phonological neighbours in the native language lexicon affects the
-extent to which listeners are able to activate the correct translation
-of an auditorily presented word in thier native language. We used a
-translation elicitation task in which participants listened to words
-from an unfamiliar language, and then tried to guess their translation
-in their native language. We will henceforth refer to the
-auditory-presented words heard by participants on each trial as
-\emph{presented words}, and the correct translation for the presented
-words as \emph{target translations}. We explored listeners' reliance on
-phonological similarity to succeed in the task by manipulating the
+inserting phrase and word boundaries after stressed syllables, following
+the more frequent iambic pattern in French.
+
+To explore the lexical mechanisms triggered by unfamiliar speech
+listening, we used a translation elicitation task in which participants
+listened to words from an unfamiliar language, and then tried to guess
+the words translations in their native language. We will henceforth
+refer to the auditory-presented words heard by participants on each
+trial as \emph{presented words}, and the correct translation for the
+presented words as \emph{target translations}. By manipulating the
 amount of phonological similarity between the presented words and their
-target translations. Since participants were unfamiliar with the
-presented language, they could only use phonological similarity between
-the presented language and their native language to successfully
+target translations, we explored listeners' reliance on phonological
+similarity to succeed in the task. Since participants were unfamiliar
+with the presented language, they could only use phonological similarity
+between the presented language and their native language to successfully
 translate the words. We therefore predicted that participants'
@@
 with the presented word, or are lexically more frequent than the target
-translation (\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}).
-Since the interference effects of phonological neighbourhood density has
-also been reported across languages
-(\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}), we
-hypothesised that the size of the facilitatory role of phonological
-similarity would be inversely proportional to the amount of
-higher-frequency phonological neighbours of the presented word in the
+word (\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}). Such
+interference effect of phonological neighbourhood density has also been
+reported across languages (\citeproc{ref-weber2004lexical}{Weber \&
+Cutler, 2004}). We hypothesised that the size of the facilitatory role
+of phonological similarity would be inversely proportional to the amount
+of higher-frequency phonological neighbours of the presented word in the
 target language. In the case of non-native speech recognition, the
@@
 
-In each trial, participants listened to one audio file, wich contained a
-single word. The audio files were the same ones used in child
+In each trial, participants listened to one audio file, which contained
+a single word. The audio files were the same ones used in child
 experiments conducted in the Laboratori de Recerca en Infància of
@@
 translation (\emph{Frequency}), the phonological similarity between the
-presented word (in Catalan or Spanish) and their correct English
-translation (\emph{Similarity}), and the presented word's number of
-cross-language phonological neighbours (\emph{CLPN}) in English.
-\emph{Frequency} was included as a nuance predictor, under the
-hypothesis that---keeping other predictors constant---participants would
-translate higher-frequency words more accurately and faster than
-lower-frequency words. Lexical frequencies of correct translations were
+presented word and their correct English translation
+(\emph{Similarity}), and the presented word's number of cross-language
+phonological neighbours (\emph{CLPN}) in English. \emph{Frequency} was
+included as a nuance predictor, under the hypothesis that---keeping
+other predictors constant---participants would provide correct responses
+more often when the correct translations are high frequency words in
+their native language. Lexical frequencies of correct translations were
 extracted from SUBTLEX-UK (\citeproc{ref-van2014subtlex}{Van Heuven et
-al., 2014}), and transformed to Zipf scores. Words in the stimuli list
-without a lexical frequency value were excluded from data analysis (2 in
-the Catalan list, 5 in the Spanish list).
+al., 2014}), and transformed to Zipf scores. Translations without a
+lexical frequency value were excluded from data analysis (2 in the
+Catalan list, 5 in the Spanish list).
 
@@
 identical (\citeproc{ref-levenshtein1966binary}{Levenshtein, 1966}). We
-divided this edit distance by the number of characters included in the
-longest X-SAMPA transcription of the translation pair. This results in a
-proportion score, in which values closer to zero correspond to lower
-Levenshtein distances between phonological transcriptions (i.e., higher
-similarity), and values closer to 1 correspond to higher Levenshtein
-distances (i.e., lower similarity). This transformation accounts for the
-fact that the Levenshtein distance tends to increase with the length of
-the transcriptions. For interpretability, we subtracted this proportion
-from one, so that values closer to one correspond to higher similarity
-between phonological transcriptions, and values closer to zero
-correspond to lower similarity between phonological transcriptions. For
-example, the \emph{table} (\texttt{teIb@l})-\emph{mesa} (\texttt{mesa})
-translation pair had a 17\% similarity, while the \emph{train}
-(\texttt{trEIn})-\emph{tren} (\texttt{tREn}) translation pair had a 60\%
-similarity. Table~\ref{tbl-stimuli} summarises the lexical frequency,
-phonological neighbourhood density and phonological overlap of the words
-included in the Catalan and the Spanish lists.
+divided this edit distance by the number of characters in the longest
+X-SAMPA transcription of the translation pair. This transformation
+accounts for the fact that Levenshtein distance tends to increase with
+the length of the transcriptions. For interpretability, we subtracted
+this proportion from one, so that values closer to one correspond to
+higher similarity between phonological transcriptions, and values closer
+to zero correspond to lower similarity between phonological
+transcriptions. For example, the \emph{table}
+(\texttt{teIb@l})-\emph{mesa} (\texttt{mesa}) translation pair had 17\%
+similarity, while the \emph{train} (\texttt{trEIn})-\emph{tren}
+(\texttt{tREn}) translation pair had 60\% similarity.
 
@@
 lexical frequency, and whose phonological transcription (in X-SAMPA
-format) different in up to one phoneme from that of the presented
-Catalan or Spanish word. Lexical frequencies and phonological
-transcriptions were extracted from the multilingual database CLEARPOND
+format) differed in up to one phoneme from that of the presented Catalan
+or Spanish word. Lexical frequencies and phonological transcriptions
+were extracted from the multilingual database CLEARPOND
 (\citeproc{ref-marian2012clearpond}{Marian et al., 2012})\footnote{{[}Phonological
   transcriptions in CLEARPOND were generated from eSPEAK,
-  \url{http://espeak.sourceforge.net/}{]}}.
+  \url{http://espeak.sourceforge.net/}{]}}. Table~\ref{tbl-stimuli}
+summarises the lexical frequency, phonological neighbourhood density and
+phonological overlap of the words included in the Catalan and the
+Spanish lists.
 
@@
 informed that they would listen to a series of pre-recorded words in
-either Catalan or Spanish (English participants) or Catalan (Spanish
-participants). They were instructed to listen to each word, guess its
-meaning in English (English participants) or Spanish (Spanish
-participants), and type their answer as soon as possible. English
-participants were randomly assigned to the Catalan or Spanish lists.
+either Catalan or Spanish. They were instructed to listen to each word,
+guess its meaning in English, and type their answer as soon as possible.
+Participants were randomly assigned to the Catalan or Spanish lists.
 Participants in the Catalan list were presented with 83 trials, and
@@
 participants were prompted to write their answer by a
-\texttt{\textgreater{}} symbol. Typed letters were displayed in the
+\texttt{\textgreater{}} symbol. Typed letters were displayed on the
 screen in real time to provide visual feed-back to participants.
@@
 pressed the \texttt{RETURN/ENTER} key to confirm their answer and start
-and new trial.
+a new trial.
 
@@
 (\texttt{\textgreater{}}) was presented. Participants then wrote their
-and clicked \texttt{RETURN}, making the end of the trial.}
+answer and clicked \texttt{RETURN}, making the end of the trial.}
 
@@
 orthographic form of the correct translation. A response was coded as
-\emph{Typo} if the participant provided a string of characters only one
-edit distance (addition, deletion, or substitution) apart from the
-orthographic form of the correct translation (e.g., ``pengiun'' instead
-of ``penguin''), as long as the response did not correspond to a
+\emph{Typo} if the participant provided a string of characters that was
+only one edit distance (addition, deletion, or substitution) apart from
+the orthographic form of the correct translation (e.g., ``pengiun''
+instead of ``penguin''), as long as the response did not correspond to a
 distinct English word. A response was coded as \emph{False friend} if
-the participant provided a phonologically similar incorrect translation.
-Responses not meeting the criteria for previous categories were labelled
-as \emph{Wrong} or \emph{Other} (see Data analysis section for more
-details). Both \emph{Correct} and \emph{Typo} responses were considered
-as correct, while \emph{Wrong} and \emph{False friend} responses were
-considered as incorrect. \emph{Other} responses were excluded from data
-analysis. Trials in which participants took longer than 10 seconds to
-respond were also excluded. Participants contributed a total of 9,152
-valid trials (5,694 in Catalan, 3,458 in Spanish). The task took
-approximately 15 minutes to be completed.
+the participant's response was incorrect, but phonologically similar to
+the presented word. Responses not meeting the criteria for previous
+categories were labelled as \emph{Wrong} or \emph{Other} (see Data
+analysis section for more details). Both \emph{Correct} and \emph{Typo}
+responses were considered as correct, while \emph{Wrong} and \emph{False
+friend} responses were considered as incorrect. \emph{Other} responses
+were excluded from data analysis. Trials in which participants took
+longer than 10 seconds to respond were also excluded. Participants
+contributed a total of 5,206 valid trials (2,604 in Catalan, 2,602 in
+Spanish). The task took approximately 15 minutes to complete.
 
@@
 interaction between \emph{Similarity} and \emph{CLPN}. We also included
-participant-level random intercepts and slopes for the the main effects
-and the interaction. Eq. 1 shows a formal description of the model.
+participant-level random intercepts and slopes for the main effects and
+the interaction. Eq. 1 shows a formal description of the model.
 
@@
 We collected data for a total of 6,446 trials completed by 72 unique
-participants. Of those trials, 2,988 were provided by 36 unique
-participants who listened to Catalan words, and 3,458 trials were
-provided by 36 unique participants who listened to Spanish words. We
-excluded trials in which participants did not enter any text (\emph{n} =
-72), in which a response in a language other than English was provided
-(e.g., ``\texttt{agua}'', \emph{n} = 51), in which participants did not
-provide a whole word (e.g., ``\texttt{f}'', \emph{n} = 5), and in which
-participants added comments to the experimenter (e.g.,
-``\texttt{unsure}'', \emph{n} = 13). In addition, we excluded data from
-participants that self-rated their oral and/or written skills in Catalan
-and Spanish, or any other second language as four or higher in a
-five-point scale (\emph{n} = 2), were diagnosed with a language
-(\emph{n} = 1), or did not contribute more than 80\% of valid trials
-(\emph{n} = 9).
+participants. We excluded trials in which participants did not enter any
+text (\emph{n} = 72), in which a response in a language other than
+English was provided (e.g., ``\texttt{agua}'', \emph{n} = 51), in which
+participants did not provide a whole word (e.g., ``\texttt{f}'',
+\emph{n} = 5), and in which participants added comments to the
+experimenter (e.g., ``\texttt{unsure}'', \emph{n} = 13). In addition, we
+excluded data from participants that self-rated their oral and/or
+written skills in Catalan and Spanish, or any other second language as
+four or higher in a five-point scale (\emph{n} = 2), were diagnosed with
+a language (\emph{n} = 1), or did not contribute more than 80\% of valid
+trials (\emph{n} = 9).
 
@@
 indicated by the regression coefficient of \emph{Group} (\(\beta\) =
--0.199, 95\% CrI = {[}-0.512, 0.129{]}, \emph{p}(ROPE) = 0.238).
+-0.165, 95\% CrI = {[}-0.482, 0.125{]}, \emph{p}(ROPE) = 0.302).
 Overall, participants responded less accurately to words with more CLPNs
@@
 interaction between \emph{Similarity} and \emph{CLPN} (\(\beta\) =
--0.653, 95\% CrI = {[}-0.973, -0.313{]}, \emph{p}(ROPE) = 0). As
+-0.236, 95\% CrI = {[}-0.362, -0.118{]}, \emph{p}(ROPE) = 0.015). As
 anticipated, participants' performance benefited from an increase in
-\emph{Similarity} (\(\beta\) = 7.133, 95\% CrI = {[}6.606, 7.735{]},
+\emph{Similarity} (\(\beta\) = 1.437, 95\% CrI = {[}1.314, 1.544{]},
 \emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
-effect (\(\beta\) = 0.009, 95\% CrI = {[}-0.11, 0.111{]}, \emph{p}(ROPE)
-= 0.925). Figure~\ref{fig-epreds-1} illustrates the posterior of the
-average predictions of the model for words with different values of
-\emph{Similarity} and \emph{CLPN}.
+effect (\(\beta\) = -0.167, 95\% CrI = {[}-0.352, -0.01{]},
+\emph{p}(ROPE) = 0.2). Figure~\ref{fig-epreds-1} illustrates the
+posterior of the average predictions of the model for words with
+different values of \emph{Similarity} and \emph{CLPN}.
 
@@
 Germanic language, is relatively distant from Catalan and Spanish, two
-Romance languages. English shares fewer cognates with Catalan and
-Spanish than it does with typologically closer languages, like Dutch and
-German. In Experiment 2, we investigated whether listeners of an
-unfamiliar but typologically closer language benefit more strongly from
-phonological similarity when performing the same task as in Experiment
-1. To this aim, we presented Spanish participants, who reported
-little-to-no prior familiarity with Catalan, with Catalan words.
+Romance languages. In comparison, Catalan and Spanish are typologically
+close languages that share many more cognates. In Experiment 2, we
+investigated whether listeners of an unfamiliar but typologically closer
+language benefit more strongly from phonological similarity when
+performing the same task as in Experiment 1. To this aim, we presented
+Spanish participants, who reported little-to-no prior familiarity with
+Catalan, with Catalan words.
 
@@
 interaction between \emph{Similarity} and \emph{CLPN} (\(\beta\) =
--0.409, 95\% CrI = {[}-0.895, 0.111{]}, \emph{p}(ROPE) = 0.081). As
+-0.162, 95\% CrI = {[}-0.369, 0.032{]}, \emph{p}(ROPE) = 0.268). As
 anticipated, participants' performance benefited from an increase in
-\emph{Similarity} (\(\beta\) = 9.274, 95\% CrI = {[}8.473, 10.269{]},
+\emph{Similarity} (\(\beta\) = 1.906, 95\% CrI = {[}1.712, 2.086{]},
 \emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
-effect (\(\beta\) = -0.069, 95\% CrI = {[}-0.353, 0.192{]},
-\emph{p}(ROPE) = 0.492). Figure~\ref{fig-epreds-1} illustrates the
+effect (\(\beta\) = -0.274, 95\% CrI = {[}-0.613, 0.088{]},
+\emph{p}(ROPE) = 0.14). Figure~\ref{fig-epreds-2} illustrates the
 posterior of the average predictions of the model for words with
@@
 extracted for different values of \emph{CLPN} (0, 2, 4, 8, 12) and
-\emph{Similarity} (1-100). Predictions are plotted separately for
-English participants translating Catalan words, and for English
-participants translating Spanish words. Lines indicate mean predictions,
-and intervals indicate 95\%, 89\%, 78\%, 67\%, and 50\% credible
-intervals (CrI).}
+\emph{Similarity} (1-100). Lines indicate mean predictions, and
+intervals indicate 95\%, 89\%, 78\%, 67\%, and 50\% credible intervals
+(CrI).}
 
 \end{figure}%
+
+In order to compare the results from Experiments 1 and 2 directly, we
+fit a model on the joint datasets from both experiments. This model
+included all predictors from the models presented in Experiments 1 and
+2, and the predictor \emph{Experiment} with levels
+\texttt{"Experiment\ 1"} and \texttt{"Experiment\ 2"}, and a three-way
+interaction between \emph{Experiment}, \emph{Similarity}, and
+\emph{CLPN}. The \emph{Experiment} variable was sum-coded as
+\texttt{-0.5} = \texttt{"Experiment\ 1"} and \texttt{+0.5} =
+\texttt{"Experiment\ 2"} (\citeproc{ref-schad2020capitalize}{Schad et
+al., 2020}). All two-way interactions between the three predictors were
+also included. The effect of \emph{CLPN} on \emph{Similarity} was mostly
+equivalent in both groups, as indicated by the coefficient of the
+three-way way interaction (\(\beta\) = 0.105, 95\% CrI = {[}-0.151,
+-0.043{]}). The posterior probability of this coefficient being
+equivalent to zero (\emph{p}(ROPE)) wwas 0.422), that is, inconclusive.
+The coefficient of the interaction term between \emph{Experiment} and
+\emph{CLPN} was also equivalent to zero (\(\beta\) = -0.09, 95\% CrI =
+{[}-0.458, 0.346{]}, \emph{p}(ROPE) = 0.361), suggesting that
+participants from Experiments 1 and 2 were affected by CLPN on a similar
+basis. Finally, the coefficient of the interaction between
+\emph{Experiment} and \emph{Similarity} was different from zero
+(\(\beta\) = -0.09, 95\% CrI = {[}0.326, 0.785{]}, \emph{p}(ROPE) = 0),
+with a positive sign indicating that the increments in \emph{Similarity}
+were associated with a larger increment in probability of correct
+translation in Experiment 2, compared to in Experiment 1: for an average
+value of \emph{CLPN}, participants in Experiment 2 benefited more
+strongly from \emph{Similarity} than participants in Experiment 1.
 
@@
 phonological neighbours with higher lexical frequency was reduced. In
-contrasts with the results in Experiment 1, the positive impact of
+contrast with the results in Experiment 1, the positive impact of
 phonological similarity on participants' performance in Experiment 2 was
 more resilient to the interference of phonological neighbourhood size.
-English natives in Experiment 1 provided significantly less accuracy
+English natives in Experiment 1 provided significantly less accurate
 responses when more than four phonological neighbors were present (even
-when translating high-similarity words), compared to when only one or
-neighbour were present. Spanish participants in Experiment 2 benefited
-from phonological similarity, even when eight neighbours were present.
-Spanish participants' performance declined after 8 neighbours, and was
-evident at 12 neighbours. Overall, this suggests that participants in
-Experiment 2, who were natives of a typologically similar language
-(Spanish) to the presented language (Catalan) benefited more strongly
-from phonological similarity than participants in Experiment 1, who were
-natives of typologically less similar language (English) to the
-presented language (Catalan, Spanish).
+when translating high-similarity words), compared to when less than
+three neighbours were present. Spanish participants in Experiment 2
+benefited from phonological similarity even when eight neighbours were
+present. Spanish participants' performance declined after eight
+neighbours, and was most evident at 12 neighbours. Overall, this
+suggests that participants in Experiment 2, who were natives of a
+typologically similar language (Spanish) to the presented language
+(Catalan), benefited more strongly from phonological similarity than
+participants in Experiment 1, who were natives of typologically less
+similar language (English) to the presented language (Catalan, Spanish).
 
@@
 correct translations for words sharing little to no phonological
-similarity to their correct translation. Table~\ref{tbl-surprises} a
-list of Catalan and Spanish words to which participants provided
+similarity to their correct translation. Table~\ref{tbl-surprises} shows
+a list of Catalan and Spanish words to which participants provided
 responses with \(\geq\) 10 average accuracy.
@@
 \caption{\label{tbl-surprises}List of items with unexpectedly high
-accuracy: the Levenshtein similarityscore betwen the presented word (in
-Catalan or Spanish) and their correct English translation is zero, but
-participants, who are reportedly unfamiliar with the presented language,
-were on average \textgreater10\% likely to guess the correct
+accuracy: the Levenshtein similarity score between the presented word
+(in Catalan or Spanish) and their correct English translation is zero,
+but participants, who are reportedly unfamiliar with the presented
+language, were on average \textgreater10\% likely to guess the correct
 translation.}
@@
 meaning of Spanish words. Experiment 3 was addressed at investigating
-this issue.
+the issue of prior knowledge.
 
@@
 Participants were asked to complete the experiment using a laptop in a
-quiet place with good internet connection. Stimuli were the same list of
-Catalan stimuli as in Experiment 1.
+quiet place with good internet connection. Stimuli were the same lists
+of Catalan and Spanish stimuli as in Experiment 1.
 
@@
 trial. These questions were presented on the same page, directly below
-the audio player. They were first asked whether or not they knew the
-presented word (multiple choice---\emph{yes}/\emph{no}). Regardless of
-their answer on the first question, participants were asked what they
-thought the translation of the word was in English (or their best
-guess), and instructed to type their answer in the provided text box.
-Finally, they were asked to rate how confident they were in their answer
-on a scale of 0 to 7, where 7 was ``very confident'' and 0 was ``not
-confident''. There was no time limit on the response phase. All
-questions had to be answered to proceed to the next trial.
+the audio player. They were first asked whether they knew the presented
+word (multiple choice---\emph{yes}/\emph{no}). Regardless of their
+answer on the first question, participants were asked what they thought
+the translation of the word was in English (or their best guess), and
+instructed to type their answer in the provided text box. Finally, they
+were asked to rate how confident they were in their answer on a scale of
+0 to 7, where 7 was ``very confident'' and 0 was ``not confident''.
+There was no time limit on the response phase. All questions had to be
+answered to proceed to the next trial.
 
@@
 audio stimulus (ambulance, cucumber, elephant, pear, turtle). The words
-were recorded by a female native speaker of English. These trials acted
-as attention checks, as participants should always answer ``yes'' to the
-first question on prior word knowledge and be able to accurately
-transcribe the word they heard. Following the practice phase,
+were recorded by a female native speaker of British English. These
+trials acted as attention checks, as participants should always answer
+``yes'' to the first question on prior word knowledge and be able to
+accurately transcribe the word they heard. Following the practice phase,
 participants completed the test phase where they heard either Spanish
@@
 We collected data for a total of 6,016 trials completed by 64 unique
-participants. Of those trials, 3,264 were provided by 32 unique
-participants who listened to Catalan words, and 2,752 trials were
-provided by 32 unique participants who listened to Spanish words. We
-excluded trials in which participants did not enter any text (\emph{n} =
-0), in which a response in a language other than English was provided
-(\emph{n} = 9), in which participants did not provide a whole word
-(\emph{n} = 0), and in which participants added comments to the
-experimenter (\emph{n} = 0). In addition, we excluded data from
-participants that self-rated their oral and/or written skills in Catalan
-and Spanish, or any other second language as four or higher in a
-five-point scale (\emph{n} = 2), were diagnosed with a language
-(\emph{n} = 0), or did not contribute more than 80\% of valid trials
-(\emph{n} = 1).
+participants. We excluded trials in which participants did not enter any
+text (\emph{n} = 0), provided a response in a language other than
+English (\emph{n} = 9), did not provide a whole word (\emph{n} = 0), or
+added comments to the experimenter (\emph{n} = 0). In addition, we
+excluded data from participants that self-rated their oral and/or
+written skills in Catalan and Spanish, or any other second language as
+four or higher in a five-point scale (\emph{n} = 2), were diagnosed with
+a language (\emph{n} = 0), or did not contribute more than 80\% of valid
+trials (\emph{n} = 1).
 
@@
 responses, participants reported an average of 5.05 confidence in the
-0-8 scale (\emph{SD} = 1.94). In responses in which no prior knowledge
-of the presented word was reported, average confidence was 1.13
-confidence in the 0-8 scale (\emph{SD} = 1.38). From the 2,743 total
-responses provided by English participants who listened to Spanish
-words, participants reported having prior knowledge of the presented
-Spanish words in 197 (7.18\%) of them. In those responses, participants
-reported an average of 4.79 confidence in the 0-8 scale (\emph{SD} =
-1.82). In responses in which no prior knowledge of the presented word
-was reported, average confidence was 1.25 confidence in the 0-8 scale
-(\emph{SD} = 1.66). Before data analysis, responses in which
-participants reported prior knowledge about the meaning of the presented
-Catalan or Spanish word were excluded from the dataset.
+0-8 scale (\emph{SD} = 1.94). For responses where no prior knowledge of
+the presented word was reported, average confidence was 1.13 (\emph{SD}
+= 1.38). From the 2,743 total responses provided by English participants
+who listened to Spanish words, participants reported having prior
+knowledge of the presented Spanish words in 197 (7.18\%) of them. In
+those responses, participants reported an average of 4.79 confidence in
+the 0-8 scale (\emph{SD} = 1.82). For responses where no prior knowledge
+of the presented word was reported, average confidence was 1.25
+(\emph{SD} = 1.66). Before data analysis, responses where participants
+reported prior knowledge about the meaning of the presented Catalan or
+Spanish word were excluded from the dataset.
 
@@
 (\(\hat{R}<1.01\)) (see Appendix 2 for more detailed model diagnostics).
-Overall, participants reported prior knowledge more often for that
-Spanish words with unexpectedly high accuracy (see Discussion in
-Experiment 2) than for words with expected accuracy (see
-\textbf{?@fig-knowledge}). Participants reported prior knowledge of
-Catalan words with unexpected accuracy as often as those with expected
-accuracy. This suggests that participants in Experiment 1 may have
-relied, to some extent, on their prior knowledge about form-meaning
-mappings to correctly translate some Spanish words. To isolate such an
-effect of prior Spanish knowledge, we run the same analysis as in
-Experiment 1 on the newly collected translations from Experiment 3, now
-excluding responses to words in which participants reported prior
-knowledge.
+Overall, participants reported prior knowledge more often for the
+Spanish words that showed unexpectedly high accuracy in Experiment 1
+(see Discussion in Experiment 2) than for words with expected accuracy
+(see \textbf{?@fig-knowledge}). Participants reported prior knowledge of
+Catalan words with unexpected accuracy in Experiment 1 as often as those
+with expected accuracy. This suggests that participants in Experiment 1
+may have relied, to some extent, on their prior knowledge about
+form-meaning mappings to correctly translate some Spanish words. To
+isolate such an effect of prior Spanish knowledge, we run the same
+analysis as in Experiment 1 on the newly collected translations from
+Experiment 3, now excluding responses to words for which participants
+reported prior knowledge.
 
@@
 Spanish words performed similarly, as indicated by the regression
-coefficient of \emph{Group} (\(\beta\) = -0.049, 95\% CrI = {[}-0.388,
--0.32{]}, \emph{p}(ROPE) = 0.426). Overall, both groups of participants
+coefficient of \emph{Group} (\(\beta\) = -0.065, 95\% CrI = {[}-0.389,
+-0.118{]}, \emph{p}(ROPE) = 0.422). Overall, both groups of participants
 responded less accurately to words with more CLPNs than to words with
@@
 the regression coefficient of the two-way interaction between
-\emph{Similarity} and \emph{CLPN} (\(\beta\) = -0.665, 95\% CrI =
-{[}-1.068, -0.32{]}, \emph{p}(ROPE) = 0.002). As anticipated,
+\emph{Similarity} and \emph{CLPN} (\(\beta\) = -0.252, 95\% CrI =
+{[}-0.385, -0.118{]}, \emph{p}(ROPE) = 0.015). As anticipated,
 participants' performance benefited from an increase in
-\emph{Similarity} (\(\beta\) = 7.585, 95\% CrI = {[}7.043, 8.24{]},
-\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
-effect (\(\beta\) = -0.049, 95\% CrI = {[}-0.203, 0.086{]},
-\emph{p}(ROPE) = 0.753). Figure~\ref{fig-epreds-1} illustrates the
+\emph{Similarity} (\(\beta\) = 1.532, 95\% CrI = {[}1.41, 1.657{]},
+\emph{p}(ROPE) = 0), while the number of \emph{CLPN} had the opposite
+effect (\(\beta\) = -0.288, 95\% CrI = {[}-0.523, -0.082{]},
+\emph{p}(ROPE) = 0.023). Figure~\ref{fig-epreds-3} illustrates the
 posterior of the average predictions of the model for words with
@@
 \caption{\label{fig-epreds-3}Posterior model-predicted mean accuracy in
-Experiment 2. Predictions were generated from 4,000 posterior samples,
+Experiment 3. Predictions were generated from 4,000 posterior samples,
 extracted for different values of \emph{CLPN} (0, 2, 4, 8, 12) and
@@
 provide their best-guess translation for each word in their native
-language. We run three experiments. In Experiment 1, British
-English-native adults listened to a list of words in Catalan or Spanish.
-Participants reported no prior familiarity with Catalan, Spanish, or any
-other Romance language, yet provided accurate translations for words
-that shared some degree of phonological similarity with their correct
-English translation. Participants only benefited from phonological
-similarity when the number of English phonological neighbours of the
-presented Catalan or Spanish word was small. Experiment 2 was aimed at
-extending our findings to a population of participants whose native
-language was typologically closer to the unfamiliar language presented.
-We tested a group of Spanish native adults who listened to a list of
-Catalan words. Again, participants reported no prior familiarity with
-Catalan, or any other Romance language other than Spanish. In line with
-the results in English natives, we found an interaction between
-phonological similarity and phonological neighbourhood size. The
-facilitation effect of phonological similarity in the Spanish
-participants was more robust to the interferring effect of phonological
-neighbourhood density: Spanish participants provided correct
-translations for words with high neighbourhood densities more often than
-English participants. We suspected that participants in Experiments 1
-and 2 might have managed to provide correct translations for some
-low-phonological similarity words thanks to having prior experience with
-those words, despite having reported being unfamiliar with Catalan,
-Spanish, or any other Romance language. Experiment 3 was a replica of
-Experiment 1, in which we collected additional information about
-participants' prior familiarity with the presented Catalan and Spanish
-words. The design of Experiment 3 was closely modelled after Experiment
-1, except that after providing their response in each trial,
-participants reported whether they had previous knowledge of the
-presented word. After removing responses in which participants reported
-prior knowledge about the meaning of the presented Catalan or Spanish
-word, we run the same analyses as in Experiment 1, finding parallel
-results. Results in Experiment 3 suggest participants in the present
-study did not simply rely on prior familiarity with some of the
-word-forms presented. Rather, participants relied on the phonological
-similarity between the presented Catalan and Spanish words and their
-correct translation.
-
-Homophonic translation is an often spontaneous phenomenon that occurs
-commonly when listening to non-native speech, and is prevalent across
-many cultures and languages
-(\citeproc{ref-dembeck2015oberflachenubersetzung}{Dembeck, 2015};
-\citeproc{ref-efimova2018homophonic}{Efimova et al., 2018}). The most
-famous case of homophonic translation is \emph{Soramimi}, which refers
-to homophonic translation in the context of music listening, when
-particular sections of the lyrics of a song in a non-native language
-make listeners think of particular words and phrases in their own native
-language. This sometimes result in a completely different meaning than
-thhe originally intended. Homophonic translation occurs outside the
-context of music. For instance, it is often intentionally employed by
-translators to preserve the meter and ``sound structure'' (e.g., rhymes)
-of an original text in the language of translation (e.g.,
-\citeproc{ref-levick2016translating}{Levick, 2016};
-\citeproc{ref-pilshchikov2016semiotics}{Pilshchikov, 2016})\footnote{The
-  rationale behind this translation approach is to ``sacrifice{[}s{]}
-  fidelity to the image rather than the melodiousness of verse''
-  (\citeproc{ref-briusov1911ot}{Briusov, 1911}). For instance, Edith
-  Grossman's translation of \emph{El Quijote}
-  (\citeproc{ref-cervantes}{Cervantes \& Grossman, 2002}) present a
-  similar case, in which formal aspects of the original text in Spanish
-  are favoured to some degree over equivalence to English, even
-  sometimes keeping Spanish words from the original text in the English
-  translation (e.g., \emph{Señor}, \emph{ínsula}).}. Although homophonic
-translation has mostly received attention as a perceptual curiosity or
-as a literary figure, little attention has been paid to homophonic
-translation as a spontanous \emph{psycholinguistic} event. The
-psycholiguistic mechanisms behind spontaneous homophonic translations
-are informative of the mechanisms underlying native speech processing.
-How does the lexical system deal with a non-native speech signal in
-which form-meaning associations are scarce, if not absent?
-
-How the perceptual system deals with non-native speech is a vastly
-studied subject (e.g., \citeproc{ref-best1988examination}{Best et al.,
-1988}; \citeproc{ref-miyawaki1975effect}{Miyawaki et al., 1975};
-\citeproc{ref-pallier2001influence}{Pallier et al., 2001};
-\citeproc{ref-van2015learning}{Van Leussen \& Escudero, 2015};
-\citeproc{ref-weber2004lexical}{Weber \& Cutler, 2004}; see
-\citeproc{ref-baese2020perception}{Baese-Berk et al., 2020} for review).
-The same cannot be said about how the perceptual system deals with
-speech in an \emph{unfamiliar} language. Previous research had explored
-how bilinguals recognise, translate, and produce words as a function of
-their proficiency in each of their languages
-(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006};
-\citeproc{ref-dijkstra2019multilink}{Dijkstra et al., 2019};
-\citeproc{ref-dijkstra2002architecture}{Dijkstra \& Van Heuven, 2002};
-e.g., \citeproc{ref-kroll1994category}{Kroll \& Stewart, 1994}). The
-case of a very low-proficiency bilingual may be close to that of a
-listener of an unfamiliar language, but being bilingual implies some
-form of proficiency in both languages, such as some knowledge about the
-meaning of some words. Even low-proficiency bilinguals are likely to
-have already acquired many form-meaning mappings in their lower
-proficiency language. Characterising the impact of phonological
-similarity on word processing is therefore challenging in bilinguals,
-due to the simultaneous contribution of phonological \emph{and}
-conceptual links in the lexicon. By testing participants with \emph{no}
-prior familiarity with the presented language (i.e., no proficiency), we
-ensured that no form-meaning associations were available for
-participants to correctly translate the unfamiliar words, if not by
-relying excluusively on phonological similarity.
-
-We calculated phonological similarity between word pairs as the
-Levenshtein similarity between the X-SAMPA transcription of their
-phonological forms (see \citeproc{ref-floccia2018introduction}{Floccia
-et al., 2018} for a similar approach). Using this measure, we found that
-participants were able to exploit the phonological similarity between
-the presented words and their correct translation, even when both
-words-forms shared few phonemes. This points to listeners accommodating
-non-native phonological forms to their native phoneme inventory,
-resulting in the activation of native lexical representations. Future
-studies may consider comparing the suitability of other measures of
-phonological similarity that take into account finer-grained phonetic or
-prosodic cues present in the acoustic signal presented to participants.
-Phonological transcriptions like X-SAMPA are abstract representations
-that ignore non-phonological contrasts (e.g., phones), and consider
-different symbols as completely different phonemes, disregarding the
-fact that some phonemes share more phonetic features than others.
-Additionally, prosodic cues such as lexical stress, which our measure of
-phonological similarity does not include, might also provide
-participants further information about the correct translation of the
-presented word-forms. Altogether, it is likely that participants in the
-present study were able to exploit additional cues in the acoustic
-signal to provide correct translations. Future studies may explore the
-relative contribution of such cues in the occurrence of homophonic
-translation. Overall, the present study provides some insights into the
-role of phonological similarity in the auditory presentation of words in
-an unfamiliar language.
+language. In Experiment 1, British English-native adults listened to a
+list of words in Catalan or Spanish. Participants reported no prior
+familiarity with Catalan, Spanish, or any other Romance language, yet
+provided accurate translations for words that shared some degree of
+phonological similarity with their correct English translation. We
+calculated phonological similarity between word pairs as the Levenshtein
+similarity between the X-SAMPA transcription of their phonological forms
+(see \citeproc{ref-floccia2018introduction}{Floccia et al., 2018} for a
+similar approach). Using this measure, we found that participants were
+able to exploit the phonological similarity between the presented words
+and their correct translation, even when both words-forms shared few
+phonemes. This points to listeners accommodating non-native phonological
+forms to their native phoneme inventory, resulting in the activation of
+native lexical representations.
+
+Experiment 2 was aimed at extending our findings to a population of
+participants whose native language was typologically closer to the
+unfamiliar language presented. We tested a group of Spanish native
+adults who listened to a list of Catalan words. Again, participants
+reported no prior familiarity with Catalan, or any other Romance
+language other than Spanish. In line with the results with English
+natives, we found an interaction between phonological similarity and
+phonological neighbourhood size. The facilitation effect of phonological
+similarity in the Spanish participants was more robust to the
+interfering effect of phonological neighbourhood density: Spanish
+participants provided correct translations for words with high
+neighbourhood densities more often than English participants. Experiment
+3 was a replication of Experiment 1, in which we collected additional
+information about participants' prior familiarity with the presented
+Catalan and Spanish words. This follow-up experiment was designed to
+address concerns that participants in Experiments 1 and 2 might have
+managed to provide correct translations for some words thanks to having
+prior experience with those words. The design of Experiment 3 was
+closely modelled after Experiment 1, except that after providing their
+response in each trial, participants reported whether they had previous
+knowledge of the presented word. After removing responses where
+participants reported prior knowledge about the meaning of the presented
+Catalan or Spanish word, we ran the same analyses as in Experiment 1,
+and found parallel results. Overall, our results suggest that
+participants were able to rely on the phonological similarity between
+the presented Catalan and Spanish words and their correct translation to
+guide word recognition in an unfamiliar language.
+
+Future studies may consider comparing the suitability of other measures
+of phonological similarity that take into account finer-grained phonetic
+or prosodic cues present in the acoustic signal presented to
+participants. Phonological transcriptions like X-SAMPA are abstract
+representations that ignore non-phonological contrasts (e.g., phones),
+and consider different symbols as completely different phonemes,
+disregarding the fact that some phonemes share more phonetic features
+than others. Additionally, prosodic cues such as lexical stress, which
+our measure of phonological similarity does not include, might also
+provide participants further information about the correct translation
+of the presented word-forms. Altogether, it is likely that participants
+in the present study were able to exploit additional cues in the
+acoustic signal to provide correct translations. Future studies may
+explore the relative contribution of such cues in the occurrence of
+homophonic translation. Overall, the present study provides some
+insights into the role of phonological similarity in the auditory
+presentation of words in an unfamiliar language.
 
 Our findings also suggest that the facilitation effect of phonological
-familiarity was moderated by phonological neighbourghood density. This
-is in line which previous studies suggesting that the presence of
+similarity was moderated by phonological neighbourhood density. This is
+in line with previous studies suggesting that the presence of
 (high-frequency) phonological neighbours interferes with the lexical
@@
 participants' lexicon. We calculated a measure of cross-linguistic
-neighbourhood size by counting, for each presented word (in Catalan or
-Spanish in Experiments 1 and 3, Catalan in Experiment 2) the number of
+neighbourhood size by counting, for each presented word the number of
 phonologically related words (with a Levenshtein distance of one) in the
@@
 2). To account for the fact that competition between phonological
-neighbours is sensitive to lexical frequency
+neighbours is sensitive to lexical frequency---high-frequency neighbours
+produce stronger interference effects
 (\citeproc{ref-dufour2010phonological}{Dufour \& Frauenfelder, 2010};
-high-frequency neighbours produce stronger interference effects,
-\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998}), we only
+\citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998})---we only
 counted neighbours with a lexical frequency higher than the correct
@@
 exploit phonological similarity to produce correct translations declined
-as the number of phonological neighbours increased. Overall, this
-suggests that listening to words in an unfamiliar language trigger the
-same mechanisms of lexical activation and selection that listening to
-words in a native language does.
-
-A comparison between results in Experiments 1 and 2 suggested that the
-performance of Spanish participants translating Catalan was more
+as the number of phonological neighbours increased. This suggests that
+listening to words in an unfamiliar language triggered similar
+mechanisms of lexical activation and selection that listening to words
+in a native language does. The generalisability of these findings can be
+tested by increasing the repertoire of words involved in the translation
+elicitation task. In Experiments 1 to 3, participants answered to a
+maximum of 105 words. All words had high frequency and low
+age-of-acquisition. To better characterise the factors that guide
+listeners' ability to match words from an unfamiliar language with words
+in their native language lexicon, words with varying levels of
+difficulty should be included in future studies.
+
+A comparison between the results in Experiments 1 and 2 suggested that
+the performance of Spanish participants translating Catalan was more
 resilient to the interfering effects of phonological neighbourhood
@@
 Germanic language). Catalan and Spanish share a higher proportion of
-cognates (i.e., form-similar translation equivalents) than Catalan and
-English, or Spanish and English. Previous studies in bilinguals show
-that this kind of cross-linguistic similarity impacts lexical
-processing: words that share high phonological similarity with their
-translations are processed faster and more accurately than words sharing
-lower phonological similarity. This phenomenon has provided strong
-evidence in favour of the language-non selective hypothesis of bilingual
-lexical access, which states that bilinguals activate lexical
-representations in both languages, even during monolingual situations.
-This facilitation effect has been reported for comprehension
-(\citeproc{ref-dijkstra2010cross}{Dijkstra et al., 2010};
-\citeproc{ref-midgley2011effects}{Midgley et al., 2011};
-\citeproc{ref-thierry2007brain}{Thierry \& Wu, 2007}), production
-(\citeproc{ref-costa2000cognate}{Costa et al., 2000}), learning
-(\citeproc{ref-de2000hard}{De Groot \& Keijzer, 2000};
-\citeproc{ref-elias2022crosslanguage}{Elias \& Degani, 2022};
-\citeproc{ref-lotto1998effects}{Lotto \& De Groot, 1998};
-\citeproc{ref-valente2018does}{Valente et al., 2018}), and translation
-(\citeproc{ref-christoffels2006memory}{Christoffels et al., 2006}). To
-our knowledge, this present study is the first to investigate the role
-of cross-linguistic phonological similarity in a monolingual population,
-in which the role of conceptual links in the lexicon is ruled out, given
-participants lack of familiarity with the presented language. This
-provided a convenient case to investigate the role of phonological
-similarity in lexical processing, disantangling the contribution of
-phonological similarity from the contribution of conceptual equivalence.
-Overall, our findings suggest that the typological distance between the
-presented and target language is associated with a more robust
-facilitation effect of phonological similarity.
+cognates than Catalan and English, or Spanish and English. Overall, our
+findings suggest that the typological distance between the presented and
+target language is associated with a more robust facilitation effect of
+phonological similarity.
 
 The specific mechanisms behind this effect are unclear. One possibility
-is that, as aforementioned, Catalan provided additional cues to Spanish
-participants in Experiment 2 compared to English participants in
-Experiment 1. For instance, it is likely that subphonemic cues not
-accounted by our measure of phonological similarity were more
-informative to Spanish participants than to English participants. In
-addition, it remains to be tested whether such association is linear
-(i.e., robustness of facilitation increases linearly with typological
-distance), proportional, or even extends beyond the specific pair of
-languages involved in the present work remains unclear. Do other close
-or distant language pairs show the same pattern of results? The
-generalisability of these findings can also be tested by increasing the
-repertoire of words involved in the translation elicitation task. In
-Experiments 1 to 3, participants answered to a maximum of 105 words. All
-words were of high-frequency, and had a low age-of-acquisition. In order
-to characterise listeners' ability to translate correctly words from an
-unfamiliar language, words with varying levels of difficulty should be
-included.
+is that there were subphonemic features of similarity between Catalan
+and Spanish words that were not adequately captured by our relatively
+coarse measure of similarity. Phonological transcriptions like X-SAMPA
+are abstract representations that ignore non-phonological contrasts
+(e.g., phones), and consider different symbols as completely different
+phonemes, disregarding the fact that some phonemes share more phonetic
+features than others. Additionally, prosodic cues such as lexical
+stress, which our measure of phonological similarity does not include,
+might also provide participants with further information about the
+correct translation of the presented word-forms. There may be more of
+these subphonemic cues available between the typologically close Spanish
+and Catalan than between the more distant English and Spanish, or
+English and Catalan. In addition, it remains to be tested whether the
+effect of typological distance is linear (i.e., robustness of
+facilitation increases linearly with typological distance),
+proportional, and whether other close or distant language pairs show the
+same pattern of results. Future studies may consider comparing the
+suitability of other measures of phonological similarity that take into
+account finer-grained phonetic or prosodic cues present in the acoustic
+signal, and explore the relative contribution of such cues in the
+occurrence of homophonic translation.
 
@@
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-baese2020perception}
-Baese-Berk, M. M., McLaughlin, D. J., \& McGowan, K. B. (2020).
-Perception of non-native speech. \emph{Language and Linguistics
-Compass}, \emph{14}(7), e12375.
-
 \bibitem[\citeproctext]{ref-best2001discrimination}
@@
 
-\bibitem[\citeproctext]{ref-christoffels2006memory}
-Christoffels, I. K., De Groot, A. M., \& Kroll, J. F. (2006). Memory and
-language skills in simultaneous interpreters: {The} role of expertise
-and language proficiency. \emph{Journal of Memory and Language},
-\emph{54}(3), 324--345.
-
-\bibitem[\citeproctext]{ref-costa2000cognate}
-Costa, A., Caramazza, A., \& Sebastian-Galles, N. (2000). The cognate
-facilitation effect: Implications for models of lexical access.
-\emph{Journal of Experimental Psychology: Learning, Memory, and
-Cognition}, \emph{26}(5), 1283.
-
 \bibitem[\citeproctext]{ref-cutler2004patterns}
@@
 
-\bibitem[\citeproctext]{ref-de2000hard}
-De Groot, A. M., \& Keijzer, R. (2000). What is hard to learn is easy to
-forget: The roles of word concreteness, cognate status, and word
-frequency in foreign-language vocabulary learning and forgetting.
-\emph{Language Learning}, \emph{50}(1), 1--56.
-
 \bibitem[\citeproctext]{ref-dembeck2015oberflachenubersetzung}
@@
 Studies}, \emph{3}(1), 7--25.
-
-\bibitem[\citeproctext]{ref-dijkstra2010cross}
-Dijkstra, T., Miwa, K., Brummelhuis, B., Sappelli, M., \& Baayen, H.
-(2010). How cross-language similarity and task demands affect cognate
-recognition. \emph{Journal of Memory and Language}, \emph{62}(3),
-284--301.
-
-\bibitem[\citeproctext]{ref-dijkstra2002architecture}
-Dijkstra, T., \& Van Heuven, W. J. (2002). The architecture of the
-bilingual word recognition system: {From} identification to decision.
-\emph{Bilingualism: Language and Cognition}, \emph{5}(3), 175--197.
-
-\bibitem[\citeproctext]{ref-dijkstra2019multilink}
-Dijkstra, T., Wahl, A., Buytenhuijs, F., Halem, N. V., Al-Jibouri, Z.,
-Korte, M. D., \& Rekké, S. (2019). Multilink: A computational model for
-bilingual word recognition and word translation. \emph{Bilingualism:
-Language and Cognition}, \emph{22}(4), 657--679.
-\url{https://doi.org/10.1017/S1366728918000287}
 
@@
 
-\bibitem[\citeproctext]{ref-elias2022crosslanguage}
-Elias, M., \& Degani, T. (2022). Cross-language interactions during
-novel word learning: {The} contribution of form similarity and
-participant characteristics. \emph{Bilingualism: Language and
-Cognition}, 1--18.
-
 \bibitem[\citeproctext]{ref-floccia2018introduction}
@@
 
-\bibitem[\citeproctext]{ref-kroll1994category}
-Kroll, J. F., \& Stewart, E. (1994). Category interference in
-translation and picture naming: {Evidence} for asymmetric connections
-between bilingual memory representations. \emph{Journal of Memory and
-Language}, \emph{33}(2), 149--174.
-
 \bibitem[\citeproctext]{ref-kruschke2018bayesian}
@@
 
-\bibitem[\citeproctext]{ref-lotto1998effects}
-Lotto, L., \& De Groot, A. M. (1998). Effects of learning method and
-word type on acquiring vocabulary in an unfamiliar language.
-\emph{Language Learning}, \emph{48}(1), 31--69.
-
 \bibitem[\citeproctext]{ref-luce1998recognizing}
@@
 phonological and orthographic neighborhood densities}.
-
-\bibitem[\citeproctext]{ref-midgley2011effects}
-Midgley, K. J., Holcomb, P. J., \& Grainger, J. (2011). Effects of
-cognate status on word comprehension in second language learners: {An
-ERP} investigation. \emph{Journal of Cognitive Neuroscience},
-\emph{23}(7), 1634--1647.
-
-\bibitem[\citeproctext]{ref-miyawaki1975effect}
-Miyawaki, K., Jenkins, J. J., Strange, W., Liberman, A. M., Verbrugge,
-R., \& Fujimura, O. (1975). An effect of linguistic experience: The
-discrimination of {[}r{]} and {[}l{]} by native speakers of japanese and
-english. \emph{Perception \& Psychophysics}, \emph{18}(5), 331--340.
 
@@
 
-\bibitem[\citeproctext]{ref-pallier2001influence}
-Pallier, C., Colomé, A., \& Sebastián-Gallés, N. (2001). The influence
-of native-language phonology on lexical access: Exemplar-based versus
-abstract lexical entries. \emph{Psychological Science}, \emph{12}(6),
-445--449.
-
 \bibitem[\citeproctext]{ref-peirce2019psychopy2}
@@
 
+\bibitem[\citeproctext]{ref-schad2020capitalize}
+Schad, D. J., Vasishth, S., Hohenstein, S., \& Kliegl, R. (2020). How to
+capitalize on a priori contrasts in linear (mixed) models: {A} tutorial.
+\emph{Journal of Memory and Language}, \emph{110}, 104038.
+
 \bibitem[\citeproctext]{ref-schepens2012distributions}
@@
 \emph{Bilingualism: Language and Cognition}, \emph{15}(1), 157--166.
-
-\bibitem[\citeproctext]{ref-thierry2007brain}
-Thierry, G., \& Wu, Y. J. (2007). Brain potentials reveal unconscious
-translation during foreign-language comprehension. \emph{Proceedings of
-the National Academy of Sciences}, \emph{104}(30), 12530--12535.
-
-\bibitem[\citeproctext]{ref-valente2018does}
-Valente, D., Ferré, P., Soares, A., Rato, A., \& Comesaña, M. (2018).
-Does phonological overlap of cognate words modulate cognate acquisition
-and processing in developing and skilled readers? \emph{Language
-Acquisition}, \emph{25}(4), 438--453.
 
@@
 \emph{67}(6), 1176--1190.
-
-\bibitem[\citeproctext]{ref-van2015learning}
-Van Leussen, J.-W., \& Escudero, P. (2015). Learning to perceive and
-recognize a second language: The L2LP model revised. \emph{Frontiers in
-Psychology}, \emph{6}, 1000.
 
@@
 
-\bibitem[\citeproctext]{ref-baese2020perception}
-Baese-Berk, M. M., McLaughlin, D. J., \& McGowan, K. B. (2020).
-Perception of non-native speech. \emph{Language and Linguistics
-Compass}, \emph{14}(7), e12375.
-
 \bibitem[\citeproctext]{ref-best2001discrimination}

### 2025-01-13 manuscript/manuscript.qmd
Address NSG comments

@@
 Outrageous as these examples might sound, the reader may know a few cases in their own native language.
-This auditory illusion is common across languages and cultures [@dembeck2015oberflachenubersetzung; @efimova2018homophonic] and can feel quite real, and often inevitable.
+This auditory illusion is common across languages and cultures and can feel quite real, often inevitable [@dembeck2015oberflachenubersetzung; @efimova2018homophonic].
 In Japanese, this phenomenon takes the name *Soramimi* (lit.
 "empty ear").
-Soramimi is a particular case of homophonic translation in which words or phrases that appear in the lyrics of a song in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
-Homophonic translation also occurs outside the context of music.
-For instance, it is often intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
+From a linguistic point of view, Soramimi is a particular case of homophonic translation: words or phrases that appear in the speech stream in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
+Homophonic translation has mostly received attention as a perceptual curiosity, and as a literary figure, intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
 For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
-Although homophonic translation is mostly seen as a perceptual curiosity or as a literary figure, little attention has been paid to homophonic translation as a spontanous psycholinguistic event.
-
-@otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show *Soramimi hour*, in which examples of Soramimi were presented the the audience for comedy purposes.
+Little attention has been paid to homophonic translation as a psycholinguistic event.
+
+In one of the few available studies concerning homophonic translation as a psycholinguistic phenomenon, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the Japanese TV show *Soramimi hour*, in which examples of Soramimi were presented to the audience for comedy purposes.
 These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
@@
 This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
-
-In the present work, we investigated the interplay between phonological similarity and phonological neighbourhood density to examine how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
+The phonological processes described by @otake2007interlingual do not take place in a vaccum, but rather in the context of a lexicon.
+As words or phrases in the native lexicon are activated by their acoustic similarity with the non-native speech stream, they trigger a series of mechanisms that result in the selection of one of the candidate word or phrase in the native language.
+In the present work, we investigated the interplay between phonological similarity and one of the main mechanisms involved in the dynamics of lexical selection: phonological neighbourhood density.
+The phonological neighbourhood density of a given word-form refers to the number of words-forms that differ in a single phoneme from the former.
+For instance, the English word *cat* hbelongs to a dense phonological neighbourhood, with 50 English phonological neighbours (e.g., hat, bat, cap), while the word *charger* belongs to a sparse phonological neighbourhood, with six neighbours (e.g., charged,larger, charter, charmer).
+Phonological neighbourhood density affects lexical processing.
+For instance, words from dense neighbourhoods are recognized more slowly and less accuractely in the auditory modality than words from sparse neighbourhoods [e.g., @luce1998recognizing; @vitevitch1998words; @dufour2010phonological; @vitevitch2006clustering].
+In this study, we examined how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
+
 Although homophonic translations do not necessarily take place with the preservation of meaning---which results in outrageous mistranslations like the aforementioned ones---homophonic translations sometimes may lead to correct translations.
@@
 For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographic similarity of the former to the equivalent in their native language.
-
-When comparing the phonemic transcription of the French to that of its Spanish translation /\textipa{"pweR.ta}/ (*puerta*), one may be led to infer that these two words do not have much overlap.
-But one thing to note is that the voiced uvular fricative consonant /\textipa{K}/ and the open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /\textipa{R}/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
-Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /\textipa{"pweR.ta}/ (*puerta*) from hearing the word-form /\textipa{"pOKt}/ (*porte*), ultimately allowing them to access the correct semantic representation.
-Native prosody also constrains the occurrence of homophonic translations.
-For instance, @kentner2015rhythmic presented in the auditory modality French-native and German-native participants with English songs with lyrics.
-Homophonic translations followed language-specific segmentation strategies.
-Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries before stressed syllables (following the more frequent trochaic pattern in German).
-On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries after stressed syllables, following the more frequent iambic pattern in French.
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and (orthographic similarity) of the former to the equivalent in their native language.
 
@@
 We therefore predicted that participants’ performance should increase when the translation pairs are phonologically more similar.
-We also predicted that there is a minimum threshold of phonological similarity to be sufficient for translation.
+We also predicted that there is a minimum threshold of phonological similarity to be sufficient for correct translation.
 
@@
 When some German speakers listened to the song *The Power* by SNAP!
-(*Coyote Ugly*, 2000), many of them misheard the line "I've got the power" as *Agatha Bauer!* When Michael Jackson vocalised "ma-ma-coo-sah" in his song *Wanna Be Startin' Somethin'* (*Thriller*, 1982), some Dutch speakers understood *mama appelsap* [mama applejuice].
+(*Coyote Ugly*, 2000), many of them misheard the line "I've got the power" as *Agatha Bauer*.
 When Spanish speakers listen to the line "Circumvent your thick ego" from the song *Pictures* by System of a Down (*Steal this Album!*, 2002), they tend to hear *Sácame de aquí, King-Kong* [take me out of here, King-Kong].
-When Japanese speakers listen to the line "I want to hold your hand" in the homonym song by The Beatles (*Meet The Beatles*, 1964), they often mishear *Aho na hōnyōhan* [stupid urinator].
-Outrageous as these examples might sound to speakers of other languages, many readers may know a few cases in their own native language.
-This auditory illusion can feel quite real, and often inevitable after the first time it happens.
+Outrageous as these examples might sound, the reader may know a few cases in their own native language.
+This auditory illusion is common across languages and cultures and can feel quite real, often inevitable [@dembeck2015oberflachenubersetzung; @efimova2018homophonic].
 In Japanese, this phenomenon takes the name *Soramimi* (lit.
-"empty ear"), after the *Soramimi Hour*---a section of the program *Tamori Club* hosted by comedian Morita Kazuyoshi (Tamori)---in which instances of this illusion were presented with comedy purposes.
-Soramimi is a particular case of *homophonic translation*: words or phrases in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning.
-This phenomenon extends beyond song lyrics, and may occur anytime when listening to non-native speech.
-For instance, when listening to "You know me?" (/ju: nəʊ mi:/) in English, Japanese-native listeners tend to understand *yunomi* (/jʊnomi/), which is a type of Japanese teacup.
-Overall, the occurence of Soramimi---and more generally, homophonic translations---suggests that speech in an unfamiliar language has the potential to activate lexical representations in the native language.
-
-Although homophonic translations do not necessarily take place with the preservation of meaning---which often results in outrageous mistranslations like the ones aforementioned---homophonic translations sometimes may lead to the activation of the intended meaning (i.e., a correct translation).
-For example, imagine a Spanish native with no previous familiarity with any other language listening to French for the first time.
-This listener may encounter the word /pɔKt/ (*porte*), translation of door in French.
-When comparing the phonemic transcription of the French to that of its Spanish translation /pweɾta/ (*puerta*), one may be led to infer that these two words do not have much overlap.
-But one thing to note is that the voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /ʁ/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
-Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /pweɾta}/ (*puerta*) from hearing the word-form /pɔʁt/ (*porte*), ultimately allowing them to access the correct semantic representation.
+"empty ear").
+From a linguistic point of view, Soramimi is a particular case of homophonic translation: words or phrases that appear in the speech stream in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
+Homophonic translation has mostly received attention as a perceptual curiosity, and as a literary figure, intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
+For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
+Little attention has been paid to homophonic translation as a psycholinguistic event.
+
+In one of the few available studies concerning homophonic translation as a language processing phenomenon, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show *Soramimi hour*, in which examples of Soramimi were presented the the audience for comedy purposes.
+These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
+The vast majority of the reported instances were phrases (96%) and few were single words (4%).
+There was relatively high variability in the degree to which phonetic features of the presented English lyrics were preserved, with some resulting Japanese words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
+An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /krʌɪ/ to *kurai*, /kuɾai/ [dark]), deletion (e.g., *go*, /goʊ/ to *go*, /go/ [go], and alternation (e.g., *low*, /loʊ/ to *rou*, /ɾoo/ [wax]).
+This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
+
+The phonological processes described by @otake2007interlingual do not take place in a vaccum, but rather in the context of a lexicon.
+As words or phrases in the native lexicon are activated by their acoustic similarity with the non-native speech stream, they trigger a series of mechanisms that result in the selection of one of the candidate word or phrase in the native language.
+In the present work, we investigated the interplay between phonological similarity and one of the main mechanisms involved in the dynamics of lexical selection: phonological neighbourhood density.
+In particular, we examined how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
+Although homophonic translations do not necessarily take place with the preservation of meaning---which results in outrageous mistranslations like the aforementioned ones---homophonic translations sometimes may lead to correct translations.
+For example, imagine an English native listening to Dutch for the first time.
+This listener may encounter the word /ˈpɪŋ.guɪn/ (*pinguin*), which sounds similar to its English translation /ˈpɛŋ.ɡwɪn/ (penguin).
 Given enough phonological similarity between the non-native word and its translation in the native language, a naïve listener may succeed at word comprehension when listening to an unfamiliar language.
@@
 
-Soramimi, and more generally homophonic translation, are popular and frequent phenomena, but the psycholinguistic literature on the topic is relatively scarce, and has mostly focused on the phonological variables involved in the occurence of Soramimi.
-For instance, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show Soramimi hour.
-These instances consisted of homophonic translations of English song lyrics to Japanese words and phrases.
-The vast majority of the reported instances were phrases (96%) and a few, to single words (4%).
-The phonetic features of the presented English lyrics were preserved with relatively high variability, with some Japanese resulting words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
-When analysing the few instances of homophonic translations of single words, the author identified three phonological processes that might explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /krʌɪ/ to *kurai*, /kuɾai/ [dark]), deletion (e.g., *go*, /goʊ/ to *go*, /go/ [go], and alternation (e.g., *low*, /loʊ/ to *rou*, /ɾoo/ [wax]).
-This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
-Native prosody also constrains the occurence of homophonic translations.
+When comparing the phonemic transcription of the French to that of its Spanish translation /ˈpweɾ.ta/ (*puerta*), one may be led to infer that these two words do not have much overlap.
+But one thing to note is that the voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
+Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /ˈpweɾ.ta/ (*puerta*) from hearing the word-form /ˈpɔʁt/ (*porte*), ultimately allowing them to access the correct semantic representation.
+Native prosody also constrains the occurrence of homophonic translations.
 For instance, @kentner2015rhythmic presented in the auditory modality French-native and German-native participants with English songs with lyrics.
 Homophonic translations followed language-specific segmentation strategies.
-Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries *before* stressed syllables (following the more frequent trochaic patter in German).
-On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries *after* stressed syllables, following the more frequent iambic pattern in French.
-
-While the previous studies have addressed the phonological mechanisms behind homophonic translation, little attention has been paid to the dynamics of lexical activation and competition that underlay this phenomenon.
-In the present work we investigated the interplay between phonological similarity and phonological neighbourhood density to examine how listeners of an unfamiliar language activate lexical representations in their native language.
-In particular, we studied how the presence of phonological neighbours in the native language lexicon affects the extent to which listeners are able to activate the correct translation of an auditorily presented word in an unfamiliar language.
-
-Non-native speech perception is a vastly studied subject [@best1988examination; @miyawaki1975effect; @pallier2001influence; @weber2004lexical; @van2015learning; see @baese2020perception for review], but most of this research considers the case of participants with some degree of familiarity with the non-native language, characterising how bilinguals recognise, translate, and produce words as a function of their proficiency in each of their languages [e.g., @kroll1994category; @dijkstra2019multilink; @dijkstra2002architecture; @christoffels2006memory].
-Even low-proficiency bilinguals are likely to have already acquired many form-meaning mappings in their lower proficiency language.
-Characterising the impact of phonological similarity on word processing is therefore challenging in bilinguals, due to the simultaneous contribution of phonological *and* conceptual links in the lexicon.
-By testing participants with *no* prior familiarity with the presented language (i.e., no proficiency), we ensured that no form-meaning associations were available for participants to correctly translate the unfamiliar words, if not by relying exclusively on phonological similarity.
-
-In this study, we used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess their translation in their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as presented words, and the correct translation for the presented words as target translations.
-We explored listeners’ reliance on phonological similarity to succeed in the task by manipulating the amount of phonological similarity between the presented words and their target translations.
+Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries before stressed syllables (following the more frequent trochaic pattern in German).
+On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries after stressed syllables, following the more frequent iambic pattern in French.
+
+To explore the lexical mechanisms triggered by unfamiliar speech listening, we used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess the words translations in their native language.
+We will henceforth refer to the auditory-presented words heard by participants on each trial as *presented words*, and the correct translation for the presented words as *target translations*.
+By manipulating the amount of phonological similarity between the presented words and their target translations, we explored listeners' reliance on phonological similarity to succeed in the task.
 Since participants were unfamiliar with the presented language, they could only use phonological similarity between the presented language and their native language to successfully translate the words.
@@
 Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
-This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target translation [@luce1998recognizing; @dufour2010phonological].
-Since the interference effects of phonological neighbourhood density has also been reported across languages [@weber2004lexical].
-The presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
+This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target word [@luce1998recognizing].
+Such interference effect of phonological neighbourhood density has also been reported across languages [@weber2004lexical].
+We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
+In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
 For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
@@
 This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
-
-We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours that the the presented word has in the target language.
 For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
@@
 If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.
-
 We conducted a series of three experiments to test these predictions.
+
 In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
@@
 In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are both Romance languages whose close typological distance is reflected in the fact that they share many cognates, while English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
+Catalan and Spanish are two Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
 By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant’s performance.
@@
 
-In each trial, participants listened to one audio file, which contained a single word.
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
-These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, Range = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
-The average duration of the Spanish audio files was `r round(durations$spa_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, Range = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
-
 ```{r exc-stimuli}
@@
 
-For each word in the Catalan and Spanish lists pair, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
+In each trial, participants listened to one audio file, which contained a single word.
+The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
+Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
+The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, Range = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
+The average duration of the Spanish audio files was `r round(durations$spa_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, Range = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
+
+For each word pair in the Catalan and Spanish lists, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
 *Frequency* was included as a nuance predictor, under the hypothesis that---keeping other predictors constant---participants would provide correct responses more often when the correct translations are high frequency words in their native language.
@@
 Translations without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
-
-We calculated *Similarity*, our main predictor of interest, by computing the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
+*Similarity*, our main predictor of interest, was calculated as the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
 The Levenshtein distance computes the edit distance between two character strings---in this case, two phonological transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
@@
 For example, the *table* (`teIb@l`)-*mesa* (`mesa`) translation pair had `r scales::percent(stringdist::stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (`tREn`) translation pair had `r scales::percent(stringdist::stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
-
-For each Catalan and Spanish word, we calculated the number of *CLPN* by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) differed in up to one phoneme from that of the presented Catalan or Spanish word.
+We calculated the number of *CLPN* for each Catalan and Spanish word by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) differed in up to one phoneme from that of the presented Catalan or Spanish word.
 Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
@@
 We modelled the probability of participants guessing the correct translation of each presented word using a generalised multilevel Bayesian regression model with a Bernoulli logit link distribution.
-We included as fixed effects the intercept, the main effects of *Frequency*, *Similarity*, and *CLPN*, and the two-way interaction between *Similarity* and *CLPN*.
+We included as fixed effects the intercept, the main effects of *Frequency*, *Similarity*, *CLPN*, and *Group* [sum-coded as `cat-ENG = -0`, `spa-ENG = +0.5`, @schad2020capitalize] and the three-way interaction between *Similarity*, *CLPN*, and *Group*.
+The predictor *Group* was included to account for any differences in task performance between English participants translating Catalan and Spanish words.
 We also included participant-level random intercepts and slopes for the main effects and the interaction.
@@
 &\textbf{Parameters}  \\
-\text{Logit}(p_{i}) = &  \beta_{0[p,w]} + \beta_{1[p]} \text{Frequency}_{i} + \beta_{2[p]} \text{PTHN}_i + \beta_{3[p]} \text{Similarity}_i + \beta_{4[p]} (\text{PTHN}_i \times \text{Similarity}_i) \\
-\beta_{0-6[p,w]} \sim & \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \text{ and  word } w \text{ in 1, ..., } W \\
-\beta_{1-6[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\ \\
+\text{Logit}(p_{i}) = &  \beta_{0[p,w]} + \beta_{1[p]} \text{Frequency}_{i} + \beta_{2[p]} \text{Similarity}_i +  \beta_{3[p]} \text{CLPN}_i + \beta_{4[p]} \text{Group}_i + \\
+&\beta_{5[p]} (\text{Similarity}_i \times \text{CLPN}_i) + \beta_{6[p]} (\text{Similarity}_i \times \text{Group}_i) + \\
+& \beta_{7[p]} (\text{CLPN}_i \times \text{Group}_i) + \beta_{8[p]} (\text{Similarity}_i \times \text{CLPN}_i \times \text{Group}_i) \\ \\
+\beta_{0-8[p,w]} \sim & \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \text{ and  word } w \text{ in 1, ..., } W \\
+\beta_{1-8[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\ \\
 &\textbf{Prior}  \\
@@
 We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+Numeric predictors were standardised before entering the model by subtracting the mean and dividing by the standard deviation.
+Categorical predictors were sum-coded as `-0.5` and `+0.5` [@schad2020capitalize].
 
@@
 #| label: tbl-dataset
+#| tbl-cap: "Summary of participants' accuracy in the translation elicitation task across Experiments 1, 2, and 3."
 sem <- function(x) mean(x) / (sqrt(length(x)))
@@
 
-Stimuli were the same list of Catalan stimuli as in Experiment 1. Procedure and data analysis were identical as in Experiment 1.
+Stimuli were the same list of Catalan stimuli as in Experiment 1.
+Procedure and data analysis were identical as in Experiment 1, with the only exception that the statistical model did not include the *Group* predictor, given that only one group (`cat-SPA`) participated in this experiment.
 
@@
 This model included all predictors from the models presented in Experiments 1 and 2, and the predictor *Experiment* with levels `"Experiment 1"` and `"Experiment 2"`, and a three-way interaction between *Experiment*, *Similarity*, and *CLPN*.
-The *Experiment* variable was sum-coded as `-0.5` = `"Experiment 1"` and `+0.5` = `"Experiment 2"` [@schad2020capitalize].
+The *Experiment* variable was sum-coded as `-0.5 = Experiment 1` and `+0.5 = Experiment 2`.
 All two-way interactions between the three predictors were also included.
-The effect of *CLPN* on *Similarity* was mostly equivalent in both groups, as indicated by the coefficient of the three-way way interaction ($\beta$ = `r round(c12$b_neigh_n_h_std_lv_std_experiment1$.median, 3)`, 95% CrI = [`r round(c12$b_neigh_n_h_std_lv_std_experiment1$.lower, 3)`, `r round(c12$b_neigh_n_h_std_lv_std$.upper, 3)`]).
-The posterior probability of this coefficient being equivalent to zero (*p*(ROPE)) wwas `r round(c12$b_neigh_n_h_std_lv_std_experiment1$.rope, 3)`), that is, inconclusive.
+The effect of *CLPN* on *Similarity* was equivalent in both groups, as indicated by the coefficient of the three-way way interaction ($\beta$ = `r round(c12$b_neigh_n_h_std_lv_std_experiment1$.median, 3)`, 95% CrI = [`r round(c12$b_neigh_n_h_std_lv_std_experiment1$.lower, 3)`, `r round(c12$b_neigh_n_h_std_lv_std$.upper, 3)`]).
+The posterior probability of this coefficient being equivalent to zero---*p*(ROPE)---was `r round(c12$b_neigh_n_h_std_lv_std_experiment1$.rope, 3)`), that is, inconclusive.
 The coefficient of the interaction term between *Experiment* and *CLPN* was also equivalent to zero ($\beta$ = `r round(c12$b_neigh_n_h_std_experiment1$.median, 3)`, 95% CrI = [`r round(c12$b_neigh_n_h_std_experiment1$.lower, 3)`, `r round(c12$b_neigh_n_h_std_experiment1$.upper, 3)`], *p*(ROPE) = `r round(c12$b_neigh_n_h_std_experiment1$.rope, 3)`), suggesting that participants from Experiments 1 and 2 were affected by CLPN on a similar basis.
@@
 Our results indicate a similar pattern of results as those in Experiment 1: participants were able to provide correct translations of presented Catalan words, provided that the Catalan words shared some degree of phonological similarity with their Spanish translation, and that the number of phonological neighbours with higher lexical frequency was reduced.
-In contrast with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 was more resilient to the interference of phonological neighbourhood size.
-English natives in Experiment 1 provided significantly less accurate responses when more than four phonological neighbors were present (even when translating high-similarity words), compared to when less than three neighbours were present.
-Spanish participants in Experiment 2 benefited from phonological similarity even when eight neighbours were present.
-Spanish participants' performance declined after eight neighbours, and was most evident at 12 neighbours.
+In contrast with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 larger.
+Spanish natives in Experiment 2 provided exploited phonological similarity more strongly than English natives in Experiment 1.
 Overall, this suggests that participants in Experiment 2, who were natives of a typologically similar language (Spanish) to the presented language (Catalan), benefited more strongly from phonological similarity than participants in Experiment 1, who were natives of typologically less similar language (English) to the presented language (Catalan, Spanish).
-
 Participants from both Experiment 1 and 2 benefited strongly from phonological similarity to correctly translate words from a non-native, reportedly unfamiliar language.
@@
 
+Experiment 3 was a conceptual replication of Experiment 1 in which we gathered additional information about English native participants' prior familiarity with the meaning of the Catalan and Spanish words presented.
+After each trial, participants reported whether they had prior knowledge about the word meaning, and how confident they were about such information in a numeric scale from 0 to 7.
+To control for the effect of participants' prior knowledge abbout form-meaning mappings, we excluded from the analyses those words in which participants reported such prior knowledge.
+We found equivalent results to those in Experiment 1, indicating that participants' surprisingly good performance in the translation elicitation task was not due to being familiar with the meaning of words they were presented with.
+
 ```{r tbl-results}
@@
 
-The present work explored the psycholinguistic bases of homophonic translation, a phenomenon in which listening to speech in a non-native---perhaps unfamiliar---language leads to the activation of lexical representations in the native language, without necessarily preserving the meaning.
+The present work explored the lexical processing bases of homophonic translation, a phenomenon in which listening to speech in a non-native---perhaps unfamiliar---language leads to the activation of lexical representations in the native language, without necessarily preserving the meaning.
 We investigated how phonological similarity and its interaction with phonological neighbourhood density impact the dynamics of lexical activation and selection during non-native word processing.
@@
 In line with the results with English natives, we found an interaction between phonological similarity and phonological neighbourhood size.
-The facilitation effect of phonological similarity in the Spanish participants was more robust to the interfering effect of phonological neighbourhood density: Spanish participants provided correct translations for words with high neighbourhood densities more often than English participants.
+Spanish natives in Experiment 2 exploited phonological similarity more strongly than English natives in Experiment 1: the association between phonological similarity and the probability of correct translation was stronger in Experiment 2, compared to in Experiment 1.
+
 Experiment 3 was a replication of Experiment 1, in which we collected additional information about participants' prior familiarity with the presented Catalan and Spanish words.
@@
 
-In summary, the present paper provides insights into the psycholinguistic mechanisms underlying homophonic translation.
+In summary, the present paper provides insights into the processing mechanisms underlying homophonic translation.
 English and Spanish native adults were tested in a translation elicitation task in which they had to guess the English or Spanish translation of a series of words in an unfamiliar language.
@@
 Participants' performance in the task only benefited from phonological similarity when the presented word had few higher-frequency phonological neighbours in the target language.
-Finally, the facilitation effect of phonological similarity was more robust in the Spanish native participants, who translated words from a typologically closer language than English participants.
+Finally, the facilitation effect of phonological similarity was stronger in the Spanish native participants, who translated words from a typologically closer language than English participants.
 Overall, the findings presented in the present paper suggest that the processing of words in a non-native, unfamiliar language recruits mechanisms of lexical activation, selection, and interference parallel to those recruited by listening to words in a native language.

### 2025-01-13 manuscript/manuscript.tex
Address NSG comments

@@
 words }
-\title{The role of cognateness in native spoken word recognition}
+\title{The role of cognateness in native spoken word
+recognition\thanks{The authors declare no conflicts of interest with
+regard to the funding source of this study. This study was supported by
+the Spanish Ministry for Science and Innovation and State Research
+Agency (Project PID2021- 123416NB-I00 financed by MCIN/ AEI/
+10.13039/501100011033 / FEDER, UE) and the Economic and Social Research
+Council (ESRC) (ES/S010947/1, UK). GGC was supported by a FPI research
+contract (PRE2019-088165). We are grateful to Irina Lepadatu and Nicola
+Gillen, and to the Speech Acquisition and Perception research group for
+their helpful feedback. We thank Xavier Mayoral, Silvia Blanch, Pamela
+Miller, and Cristina Cuadrado for their technical support. We also thank
+all families and infants who participated in the experiments. This study
+was conducted according to guidelines laid down in the Declaration of
+Helsinki, and was approved by the Drug Research Ethical Committee (CEIm)
+of the IMIM Parc de Salut Mar, reference 2020/9080/I. The data and code
+necessary to reproduce the analyses presented here are publicly
+accessible, as are the materials necessary to attempt to replicate the
+findings. The analyses presented here were not preregistered. Data,
+code, and materials for this research are available at the following
+URLs: https://osf.io/hy984/,
+https://github.com/gongcastro/translation-elicitation.}}
 \def\asep{\\\\\\ } % default: all authors on same column
@@
 \begin{abstract}
-Insert abstract here.
+While listening to speech in an unfamiliar language, words and phrases
+in the native language are activated, leading to common auditory
+illusions like Soramimi, a particular case of homophonic translation.
+Homophonic translation has received little attention as a
+psycholinguistic phenomenon. In this study, we explored the mechanisms
+underlying lexical activation and selection in homoponic translation.
+Across three studies, English and Spanish native adults completed a
+translation elicitation task, in which they listened to a series of
+Catalan or Spanish words (to which participants were unfamiliar), and
+then had to type their best-guess translation in their native language.
+Both English and Spanish natives were surprisingly good at translating
+unfamiliar words, efficiently exploiting the phonological similarity
+between the presented (unfamiliar) words and their correct translation.
+When the correct translation belonged to high-density phonological
+neighbours, participants' ability to benefit from phonological
+similarity decreased. Finally, Spanish natives exploited phonological
+similarity more efficiently: provided identical phonological similarity,
+Spanish natives were more likely to guess the correct translation of the
+unfamiliar word than English natives. Overall, we show that speech in an
+unfamiliar language triggers equivalent dynamics of lexical selection
+than native speech, providing a psycholinguistic account for homophonic
+translation.
 \end{abstract}
@@
 cases in their own native language. This auditory illusion is common
-across languages and cultures
+across languages and cultures and can feel quite real, often inevitable
 (\citeproc{ref-dembeck2015oberflachenubersetzung}{Dembeck, 2015};
-\citeproc{ref-efimova2018homophonic}{Efimova et al., 2018}) and can feel
-quite real, and often inevitable. In Japanese, this phenomenon takes the
-name \emph{Soramimi} (lit. ``empty ear''). Soramimi is a particular case
-of homophonic translation in which words or phrases that appear in the
-lyrics of a song in one language are translated into similar-sounding
-words and phrases in a different language, without necessarily
-preserving the meaning (\citeproc{ref-gasparov2006semen}{Gasparov,
-2006}). Homophonic translation also occurs outside the context of music.
-For instance, it is often intentionally employed by translators to
-preserve the meter and ``sound structure'' (e.g., rhymes) of an original
-text in the language of translation (e.g.,
+\citeproc{ref-efimova2018homophonic}{Efimova et al., 2018}). In
+Japanese, this phenomenon takes the name \emph{Soramimi} (lit. ``empty
+ear''). From a linguistic point of view, Soramimi is a particular case
+of homophonic translation: words or phrases that appear in the speech
+stream in one language are translated into similar-sounding words and
+phrases in a different language, without necessarily preserving the
+meaning (\citeproc{ref-gasparov2006semen}{Gasparov, 2006}). Homophonic
+translation has mostly received attention as a perceptual curiosity, and
+as a literary figure, intentionally employed by translators to preserve
+the meter and ``sound structure'' (e.g., rhymes) of an original text in
+the language of translation (e.g.,
 \citeproc{ref-levick2016translating}{Levick, 2016};
@@
   sometimes keeping Spanish words from the original text in the English
-  translation (e.g., \emph{Señor}, \emph{ínsula}).}. Although homophonic
-translation is mostly seen as a perceptual curiosity or as a literary
-figure, little attention has been paid to homophonic translation as a
-spontanous psycholinguistic event.
-
-Otake (\citeproc{ref-otake2007interlingual}{2007}) analysed 194
-instances of Soramimi, broadcasted between 1992 and 2007 by the TV show
-\emph{Soramimi hour}, in which examples of Soramimi were presented the
+  translation (e.g., \emph{Señor}, \emph{ínsula}).}. Little attention
+has been paid to homophonic translation as a psycholinguistic event.
+
+In one of the few available studies concerning homophonic translation as
+a psycholinguistic phenomenon, Otake
+(\citeproc{ref-otake2007interlingual}{2007}) analysed 194 instances of
+Soramimi, broadcasted between 1992 and 2007 by the Japanese TV show
+\emph{Soramimi hour}, in which examples of Soramimi were presented to
 the audience for comedy purposes. These instances consisted of
@@
 (\citeproc{ref-dupoux1999epenthetic}{Dupoux et al., 1999};
-\citeproc{ref-peperkamp2008perceptual}{Peperkamp et al., 2008}).
-
-In the present work, we investigated the interplay between phonological
-similarity and phonological neighbourhood density to examine how
-listeners may activate lexical representations in their native language
-when listening to an unfamiliar language. Although homophonic
-translations do not necessarily take place with the preservation of
-meaning---which results in outrageous mistranslations like the
-aforementioned ones---homophonic translations sometimes may lead to
-correct translations. For example, imagine an English native listening
-to Dutch for the first time. This listener may encounter the word
-/\textipa{"pINgUIn}/ (\emph{pinguin}), which sounds similar to its
+\citeproc{ref-peperkamp2008perceptual}{Peperkamp et al., 2008}). The
+phonological processes described by Otake
+(\citeproc{ref-otake2007interlingual}{2007}) do not take place in a
+vaccum, but rather in the context of a lexicon. As words or phrases in
+the native lexicon are activated by their acoustic similarity with the
+non-native speech stream, they trigger a series of mechanisms that
+result in the selection of one of the candidate word or phrase in the
+native language. In the present work, we investigated the interplay
+between phonological similarity and one of the main mechanisms involved
+in the dynamics of lexical selection: phonological neighbourhood
+density. The phonological neighbourhood density of a given word-form
+refers to the number of words-forms that differ in a single phoneme from
+the former. For instance, the English word \emph{cat} hbelongs to a
+dense phonological neighbourhood, with 50 English phonological
+neighbours (e.g., hat, bat, cap), while the word \emph{charger} belongs
+to a sparse phonological neighbourhood, with six neighbours (e.g.,
+charged,larger, charter, charmer). Phonological neighbourhood density
+affects lexical processing. For instance, words from dense
+neighbourhoods are recognized more slowly and less accuractely in the
+auditory modality than words from sparse neighbourhoods
+(\citeproc{ref-dufour2010phonological}{Dufour \& Frauenfelder, 2010};
+e.g., \citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998};
+\citeproc{ref-vitevitch2006clustering}{M. Vitevitch, 2006};
+\citeproc{ref-vitevitch1998words}{Vitevitch \& Luce, 1998}). In this
+study, we examined how listeners may activate lexical representations in
+their native language when listening to an unfamiliar language.
+
+Although homophonic translations do not necessarily take place with the
+preservation of meaning---which results in outrageous mistranslations
+like the aforementioned ones---homophonic translations sometimes may
+lead to correct translations. For example, imagine an English native
+listening to Dutch for the first time. This listener may encounter the
+word /\textipa{"pINgUIn}/ (\emph{pinguin}), which sounds similar to its
 English translation /\textipa{"pENgwIn}/ (penguin). Given enough
@@
 translating Catalan \emph{porta} than English \emph{door}, due to the
-phonological and orthographic similarity of the former to the equivalent
-in their native language.
-
-When comparing the phonemic transcription of the French to that of its
-Spanish translation /\textipa{"pweR.ta}/ (\emph{puerta}), one may be led
-to infer that these two words do not have much overlap. But one thing to
-note is that the voiced uvular fricative consonant /\textipa{K}/ and the
-open-mid back rounded vowel /\textipa{O}/, which do not exist in Spanish
-as phonemic categories, can be perceived as allophonic variations the
-native (Spanish) phonemes /\textipa{R}/ and /o/, as described by
-assimilation models of phoneme perception
-(\citeproc{ref-best1988examination}{Best et al., 1988},
-\citeproc{ref-best2001discrimination}{2001}). Although this phonetic
-mismatch between the non-native word and its translation can have a
-noticeable toll on comprehension Cutler et al.
-(\citeproc{ref-cutler2004patterns}{2004}), the Spanish listener may
-still be able to activate the phonological representation of
-/\textipa{"pweR.ta}/ (\emph{puerta}) from hearing the word-form
-/\textipa{"pOKt}/ (\emph{porte}), ultimately allowing them to access the
-correct semantic representation. Native prosody also constrains the
-occurrence of homophonic translations. For instance, Kentner
-(\citeproc{ref-kentner2015rhythmic}{2015}) presented in the auditory
-modality French-native and German-native participants with English songs
-with lyrics. Homophonic translations followed language-specific
-segmentation strategies. Most homophonic translations reported by
-German-native participants resulted from inserting phrase and word
-boundaries before stressed syllables (following the more frequent
-trochaic pattern in German). On the contrary, most homophonic
-translations reported by French-native participants resulted from
-inserting phrase and word boundaries after stressed syllables, following
-the more frequent iambic pattern in French.
+phonological and (orthographic similarity) of the former to the
+equivalent in their native language.
 
@@
 phonologically more similar. We also predicted that there is a minimum
-threshold of phonological similarity to be sufficient for translation.
+threshold of phonological similarity to be sufficient for correct
+translation.
 
@@
 
-For each word in the Catalan and Spanish lists pair, we defined three
+For each word pair in the Catalan and Spanish lists, we defined three
 predictors of interest: the lexical frequency of the correct English
@@
 lexical frequency value were excluded from data analysis (2 in the
-Catalan list, 5 in the Spanish list).
-
-We calculated \emph{Similarity}, our main predictor of interest, by
-computing the Levenshtein similarity between the X-SAMPA transcriptions
-of each pair of translations using the \texttt{stringdist} R package
-(\citeproc{ref-van2014stringdist}{van der Loo, 2014}). The Levenshtein
-distance computes the edit distance between two character strings---in
-this case, two phonological transcriptions---by counting the number of
-additions, deletions, and substitutions necessary to make both strings
-identical (\citeproc{ref-levenshtein1966binary}{Levenshtein, 1966}). We
-divided this edit distance by the number of characters in the longest
-X-SAMPA transcription of the translation pair. This transformation
-accounts for the fact that Levenshtein distance tends to increase with
-the length of the transcriptions. For interpretability, we subtracted
-this proportion from one, so that values closer to one correspond to
-higher similarity between phonological transcriptions, and values closer
-to zero correspond to lower similarity between phonological
-transcriptions. For example, the \emph{table}
-(\texttt{teIb@l})-\emph{mesa} (\texttt{mesa}) translation pair had 17\%
-similarity, while the \emph{train} (\texttt{trEIn})-\emph{tren}
-(\texttt{tREn}) translation pair had 60\% similarity.
-
-For each Catalan and Spanish word, we calculated the number of
-\emph{CLPN} by counting the number of English words with same or higher
+Catalan list, 5 in the Spanish list). \emph{Similarity}, our main
+predictor of interest, was calculated as the Levenshtein similarity
+between the X-SAMPA transcriptions of each pair of translations using
+the \texttt{stringdist} R package (\citeproc{ref-van2014stringdist}{van
+der Loo, 2014}). The Levenshtein distance computes the edit distance
+between two character strings---in this case, two phonological
+transcriptions---by counting the number of additions, deletions, and
+substitutions necessary to make both strings identical
+(\citeproc{ref-levenshtein1966binary}{Levenshtein, 1966}). We divided
+this edit distance by the number of characters in the longest X-SAMPA
+transcription of the translation pair. This transformation accounts for
+the fact that Levenshtein distance tends to increase with the length of
+the transcriptions. For interpretability, we subtracted this proportion
+from one, so that values closer to one correspond to higher similarity
+between phonological transcriptions, and values closer to zero
+correspond to lower similarity between phonological transcriptions. For
+example, the \emph{table} (\texttt{teIb@l})-\emph{mesa} (\texttt{mesa})
+translation pair had 17\% similarity, while the \emph{train}
+(\texttt{trEIn})-\emph{tren} (\texttt{tREn}) translation pair had 60\%
+similarity. We calculated the number of \emph{CLPN} for each Catalan and
+Spanish word by counting the number of English words with same or higher
 lexical frequency, and whose phonological transcription (in X-SAMPA
@@
 included as fixed effects the intercept, the main effects of
-\emph{Frequency}, \emph{Similarity}, and \emph{CLPN}, and the two-way
-interaction between \emph{Similarity} and \emph{CLPN}. We also included
-participant-level random intercepts and slopes for the main effects and
-the interaction. Eq. 1 shows a formal description of the model.
+\emph{Frequency}, \emph{Similarity}, \emph{CLPN}, and \emph{Group}
+(sum-coded as \texttt{cat-ENG\ =\ -0}, \texttt{spa-ENG\ =\ +0.5},
+\citeproc{ref-schad2020capitalize}{Schad et al., 2020}) and the
+three-way interaction between \emph{Similarity}, \emph{CLPN}, and
+\emph{Group}. The predictor \emph{Group} was included to account for any
+differences in task performance between English participants translating
+Catalan and Spanish words. We also included participant-level random
+intercepts and slopes for the main effects and the interaction. Eq. 1
+shows a formal description of the model.
 
@@
 and compare the models (see Appendix 1 for mode details on the models).
+Numeric predictors were standardised before entering the model by
+subtracting the mean and dividing by the standard deviation. Categorical
+predictors were sum-coded as \texttt{-0.5} and \texttt{+0.5}
+(\citeproc{ref-schad2020capitalize}{Schad et al., 2020}).
 
@@
 
-\caption{\label{tbl-dataset}}
+\caption{\label{tbl-dataset}Summary of participants' accuracy in the
+translation elicitation task across Experiments 1, 2, and 3.}
 
@@
 Stimuli were the same list of Catalan stimuli as in Experiment 1.
-Procedure and data analysis were identical as in Experiment 1.
+Procedure and data analysis were identical as in Experiment 1, with the
+only exception that the statistical model did not include the
+\emph{Group} predictor, given that only one group (\texttt{cat-SPA})
+participated in this experiment.
 
@@
 \emph{CLPN}. The \emph{Experiment} variable was sum-coded as
-\texttt{-0.5} = \texttt{"Experiment\ 1"} and \texttt{+0.5} =
-\texttt{"Experiment\ 2"} (\citeproc{ref-schad2020capitalize}{Schad et
-al., 2020}). All two-way interactions between the three predictors were
-also included. The effect of \emph{CLPN} on \emph{Similarity} was mostly
-equivalent in both groups, as indicated by the coefficient of the
-three-way way interaction (\(\beta\) = 0.105, 95\% CrI = {[}-0.151,
--0.043{]}). The posterior probability of this coefficient being
-equivalent to zero (\emph{p}(ROPE)) wwas 0.422), that is, inconclusive.
-The coefficient of the interaction term between \emph{Experiment} and
+\texttt{-0.5\ =\ Experiment\ 1} and \texttt{+0.5\ =\ Experiment\ 2}. All
+two-way interactions between the three predictors were also included.
+The effect of \emph{CLPN} on \emph{Similarity} was equivalent in both
+groups, as indicated by the coefficient of the three-way way interaction
+(\(\beta\) = 0.105, 95\% CrI = {[}-0.151, -0.043{]}). The posterior
+probability of this coefficient being equivalent to
+zero---\emph{p}(ROPE)---was 0.422), that is, inconclusive. The
+coefficient of the interaction term between \emph{Experiment} and
 \emph{CLPN} was also equivalent to zero (\(\beta\) = -0.09, 95\% CrI =
@@
 contrast with the results in Experiment 1, the positive impact of
-phonological similarity on participants' performance in Experiment 2 was
-more resilient to the interference of phonological neighbourhood size.
-English natives in Experiment 1 provided significantly less accurate
-responses when more than four phonological neighbors were present (even
-when translating high-similarity words), compared to when less than
-three neighbours were present. Spanish participants in Experiment 2
-benefited from phonological similarity even when eight neighbours were
-present. Spanish participants' performance declined after eight
-neighbours, and was most evident at 12 neighbours. Overall, this
-suggests that participants in Experiment 2, who were natives of a
+phonological similarity on participants' performance in Experiment 2
+larger. Spanish natives in Experiment 2 provided exploited phonological
+similarity more strongly than English natives in Experiment 1. Overall,
+this suggests that participants in Experiment 2, who were natives of a
 typologically similar language (Spanish) to the presented language
@@
 similar language (English) to the presented language (Catalan, Spanish).
-
 Participants from both Experiment 1 and 2 benefited strongly from
@@
 
+Experiment 3 was a conceptual replication of Experiment 1 in which we
+gathered additional information about English native participants' prior
+familiarity with the meaning of the Catalan and Spanish words presented.
+After each trial, participants reported whether they had prior knowledge
+about the word meaning, and how confident they were about such
+information in a numeric scale from 0 to 7. To control for the effect of
+participants' prior knowledge abbout form-meaning mappings, we excluded
+from the analyses those words in which participants reported such prior
+knowledge. We found equivalent results to those in Experiment 1,
+indicating that participants' surprisingly good performance in the
+translation elicitation task was not due to being familiar with the
+meaning of words they were presented with.
+
 \begin{longtable}{l|rrrrrrrr}
@@
 
-The present work explored the psycholinguistic bases of homophonic
+The present work explored the lexical processing bases of homophonic
 translation, a phenomenon in which listening to speech in a
@@
 natives, we found an interaction between phonological similarity and
-phonological neighbourhood size. The facilitation effect of phonological
-similarity in the Spanish participants was more robust to the
-interfering effect of phonological neighbourhood density: Spanish
-participants provided correct translations for words with high
-neighbourhood densities more often than English participants. Experiment
-3 was a replication of Experiment 1, in which we collected additional
-information about participants' prior familiarity with the presented
-Catalan and Spanish words. This follow-up experiment was designed to
-address concerns that participants in Experiments 1 and 2 might have
-managed to provide correct translations for some words thanks to having
-prior experience with those words. The design of Experiment 3 was
-closely modelled after Experiment 1, except that after providing their
-response in each trial, participants reported whether they had previous
-knowledge of the presented word. After removing responses where
+phonological neighbourhood size. Spanish natives in Experiment 2
+exploited phonological similarity more strongly than English natives in
+Experiment 1: the association between phonological similarity and the
+probability of correct translation was stronger in Experiment 2,
+compared to in Experiment 1.
+
+Experiment 3 was a replication of Experiment 1, in which we collected
+additional information about participants' prior familiarity with the
+presented Catalan and Spanish words. This follow-up experiment was
+designed to address concerns that participants in Experiments 1 and 2
+might have managed to provide correct translations for some words thanks
+to having prior experience with those words. The design of Experiment 3
+was closely modelled after Experiment 1, except that after providing
+their response in each trial, participants reported whether they had
+previous knowledge of the presented word. After removing responses where
 participants reported prior knowledge about the meaning of the presented
@@
 
-In summary, the present paper provides insights into the
-psycholinguistic mechanisms underlying homophonic translation. English
-and Spanish native adults were tested in a translation elicitation task
-in which they had to guess the English or Spanish translation of a
-series of words in an unfamiliar language. Participants successfully
-exploited phonological similarity between the presented words and their
-correct translations to provide correct answers. Participants'
-performance in the task only benefited from phonological similarity when
-the presented word had few higher-frequency phonological neighbours in
-the target language. Finally, the facilitation effect of phonological
-similarity was more robust in the Spanish native participants, who
-translated words from a typologically closer language than English
-participants. Overall, the findings presented in the present paper
-suggest that the processing of words in a non-native, unfamiliar
-language recruits mechanisms of lexical activation, selection, and
-interference parallel to those recruited by listening to words in a
-native language.
+In summary, the present paper provides insights into the processing
+mechanisms underlying homophonic translation. English and Spanish native
+adults were tested in a translation elicitation task in which they had
+to guess the English or Spanish translation of a series of words in an
+unfamiliar language. Participants successfully exploited phonological
+similarity between the presented words and their correct translations to
+provide correct answers. Participants' performance in the task only
+benefited from phonological similarity when the presented word had few
+higher-frequency phonological neighbours in the target language.
+Finally, the facilitation effect of phonological similarity was stronger
+in the Spanish native participants, who translated words from a
+typologically closer language than English participants. Overall, the
+findings presented in the present paper suggest that the processing of
+words in a non-native, unfamiliar language recruits mechanisms of
+lexical activation, selection, and interference parallel to those
+recruited by listening to words in a native language.
 
@@
 \begin{CSLReferences}{1}{0}
-\bibitem[\citeproctext]{ref-best2001discrimination}
-Best, C. T., McRoberts, G. W., \& Goodell, E. (2001). Discrimination of
-non-native consonant contrasts varying in perceptual assimilation to the
-listener's native phonological system. \emph{The Journal of the
-Acoustical Society of America}, \emph{109}(2), 775--794.
-
-\bibitem[\citeproctext]{ref-best1988examination}
-Best, C. T., McRoberts, G. W., \& Sithole, N. M. (1988). Examination of
-perceptual reorganization for nonnative speech contrasts: Zulu click
-discrimination by english-speaking adults and infants. \emph{Journal of
-Experimental Psychology: Human Perception and Performance},
-\emph{14}(3), 345.
-
 \bibitem[\citeproctext]{ref-briusov1911ot}
@@
 
-\bibitem[\citeproctext]{ref-cutler2004patterns}
-Cutler, A., Weber, A., Smits, R., \& Cooper, N. (2004). Patterns of
-{English} phoneme confusions by native and non-native listeners.
-\emph{The Journal of the Acoustical Society of America}, \emph{116}(6),
-3668--3678.
-
 \bibitem[\citeproctext]{ref-dembeck2015oberflachenubersetzung}
@@
 
-\bibitem[\citeproctext]{ref-kentner2015rhythmic}
-Kentner, G. (2015). Rhythmic segmentation in auditory illusions-evidence
-from cross-linguistic mondegreens. \emph{ICPhS}.
-
 \bibitem[\citeproctext]{ref-kruschke2018bayesian}
@@
 \emph{67}(6), 1176--1190.
+
+\bibitem[\citeproctext]{ref-vitevitch2006clustering}
+Vitevitch, M. (2006). The clustering coefficient of phonological
+neighborhoods influences spoken word recognition. \emph{The Journal of
+the Acoustical Society of America}, \emph{120}(5), 3252--3252.
+\url{https://doi.org/10.1121/1.4788314}
 
@@
 
-\bibitem[\citeproctext]{ref-best2001discrimination}
-Best, C. T., McRoberts, G. W., \& Goodell, E. (2001). Discrimination of
-non-native consonant contrasts varying in perceptual assimilation to the
-listener's native phonological system. \emph{The Journal of the
-Acoustical Society of America}, \emph{109}(2), 775--794.
-
-\bibitem[\citeproctext]{ref-best1988examination}
-Best, C. T., McRoberts, G. W., \& Sithole, N. M. (1988). Examination of
-perceptual reorganization for nonnative speech contrasts: Zulu click
-discrimination by english-speaking adults and infants. \emph{Journal of
-Experimental Psychology: Human Perception and Performance},
-\emph{14}(3), 345.
-
 \bibitem[\citeproctext]{ref-briusov1911ot}

### 2025-02-07 manuscript/manuscript.qmd
Minor styling changes

@@
 
-When some German speakers listened to the song *The Power* by SNAP!
-(*Coyote Ugly*, 2000), many of them misheard the line "I've got the power" as *Agatha Bauer*.
+When some German speakers listen to the song *The Power* by SNAP!
+(*Coyote Ugly*, 2000), many of them mishear the line "I've got the power" as *Agatha Bauer*.
 When Spanish speakers listen to the line "Circumvent your thick ego" from the song *Pictures* by System of a Down (*Steal this Album!*, 2002), they tend to hear *Sácame de aquí, King-Kong* [take me out of here, King-Kong].
@@
 For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
-Little attention has been paid to homophonic translation as a psycholinguistic event.
-
-In one of the few available studies concerning homophonic translation as a psycholinguistic phenomenon, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the Japanese TV show *Soramimi hour*, in which examples of Soramimi were presented to the audience for comedy purposes.
+Little attention has been paid to homophonic translation from a language processing perspective.
+
+One exception is the study by @otake2007interlingual, who analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the Japanese TV show *Soramimi hour*, in which examples of Soramimi were presented to the audience for comedy purposes.
 These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
-The vast majority of the reported instances were phrases (96%) and few were single words (4%).
+The vast majority of the reported instances were phrases (96%) and a few, single words (4%).
 There was relatively high variability in the degree to which phonetic features of the presented English lyrics were preserved, with some resulting Japanese words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
-An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /\textipa{kraI}/ to *kurai*, /\textipa{kuRai}/ [dark]), deletion (e.g., *go*, /\textipa{goU}/ to *go*, /go/ [go], and alternation (e.g., *low*, /\textipa{loU}/ to *rou*, /\textipa{Roo}/ [wax]).
+An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /\textipa{kraI}/ to *kurai* /\textipa{kuRai}/ [dark]), deletion (e.g., *go* /\textipa{goU}/ to *go* /go/ [*Go*, a board game], and alternation (e.g., *low* /\textipa{loU}/ to *rou* /\textipa{Roo}/ [wax]).
 This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
+
 The phonological processes described by @otake2007interlingual do not take place in a vaccum, but rather in the context of a lexicon.
@@
 In the present work, we investigated the interplay between phonological similarity and one of the main mechanisms involved in the dynamics of lexical selection: phonological neighbourhood density.
-The phonological neighbourhood density of a given word-form refers to the number of words-forms that differ in a single phoneme from the former.
-For instance, the English word *cat* hbelongs to a dense phonological neighbourhood, with 50 English phonological neighbours (e.g., hat, bat, cap), while the word *charger* belongs to a sparse phonological neighbourhood, with six neighbours (e.g., charged,larger, charter, charmer).
+The phonological neighbourhood density of a given word-form refers to the number of words-forms that differ in a single phoneme from the it.
+For instance, the English word *cat* belongs to a dense phonological neighbourhood, with 50 English phonological neighbours (e.g., hat, bat, cap), while the word *charger* belongs to a sparse phonological neighbourhood, with six neighbours (e.g., charged, larger, charter, charmer) [@marian2012clearpond].
 Phonological neighbourhood density affects lexical processing.
-For instance, words from dense neighbourhoods are recognized more slowly and less accuractely in the auditory modality than words from sparse neighbourhoods [e.g., @luce1998recognizing; @vitevitch1998words; @dufour2010phonological; @vitevitch2006clustering].
+For instance, words from dense neighbourhoods are recognized more slowly and less accuractely in the auditory modality than words from sparse neighbourhoods [@luce1998recognizing; @vitevitch1998words; e.g.,  @dufour2010phonological; @vitevitch2006clustering].
 In this study, we examined how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
 
-Although homophonic translations do not necessarily take place with the preservation of meaning---which results in outrageous mistranslations like the aforementioned ones---homophonic translations sometimes may lead to correct translations.
+Although homophonic translations rarely involve preservation of meaning, they sometimes lead to correct translations.
 For example, imagine an English native listening to Dutch for the first time.
@@
 
-We further investigated the effect of phonological competitors.
+We further investigated the effect of phonological competition.
 Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language.
@@
 For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
-It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect cross-linguistic phonological similarity has on word recognition in a foreign language.
+It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect that cross-linguistic phonological similarity has on word recognition in a foreign language.
 This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
-For instance, one could expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
+For instance, one might expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to the combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
 To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme apart from the presented word.
@@
 For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
-Little attention has been paid to homophonic translation as a psycholinguistic event.
-
-In one of the few available studies concerning homophonic translation as a language processing phenomenon, @otake2007interlingual analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show *Soramimi hour*, in which examples of Soramimi were presented the the audience for comedy purposes.
+Little attention has been paid to homophonic translation from a language processing perspective.
+
+One exception is the study by @otake2007interlingual, who analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show *Soramimi hour*, in which examples of Soramimi were presented the the audience for comedy purposes.
 These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
-The vast majority of the reported instances were phrases (96%) and few were single words (4%).
+The vast majority of the reported instances were phrases (96%) and a few, single words (4%).
 There was relatively high variability in the degree to which phonetic features of the presented English lyrics were preserved, with some resulting Japanese words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
-An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /krʌɪ/ to *kurai*, /kuɾai/ [dark]), deletion (e.g., *go*, /goʊ/ to *go*, /go/ [go], and alternation (e.g., *low*, /loʊ/ to *rou*, /ɾoo/ [wax]).
+An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /krʌɪ/ to *kurai*, /kuɾai/ [dark]), deletion (e.g., *go* /goʊ/ to *go* /go/ [*Go*, a board game], and alternation (e.g., *low* /loʊ/ to *rou* /ɾoo/ [wax]).
 This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
@@
 In the present work, we investigated the interplay between phonological similarity and one of the main mechanisms involved in the dynamics of lexical selection: phonological neighbourhood density.
-In particular, we examined how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
-Although homophonic translations do not necessarily take place with the preservation of meaning---which results in outrageous mistranslations like the aforementioned ones---homophonic translations sometimes may lead to correct translations.
+The phonological neighbourhood density of a given word-form refers to the number of words-forms that differ in a single phoneme from the it.
+For instance, the English word *cat* belongs to a dense phonological neighbourhood, with 50 English phonological neighbours (e.g., hat, bat, cap), while the word *charger* belongs to a sparse phonological neighbourhood, with six neighbours (e.g., charged, larger, charter, charmer) [@marian2012clearpond].
+Phonological neighbourhood density affects lexical processing.
+For instance, words from dense neighbourhoods are recognized more slowly and less accuractely in the auditory modality than words from sparse neighbourhoods [@luce1998recognizing; @vitevitch1998words; e.g.,  @dufour2010phonological; @vitevitch2006clustering].
+In this study, we examined how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
+
+Although homophonic translations rarely involve preservation of meaning, they sometimes lead to correct translations.
 For example, imagine an English native listening to Dutch for the first time.
-This listener may encounter the word /ˈpɪŋ.guɪn/ (*pinguin*), which sounds similar to its English translation /ˈpɛŋ.ɡwɪn/ (penguin).
+This listener may encounter the word /\textipa{"pINgUIn}/ (*pinguin*), which sounds similar to its English translation /\textipa{"pENgwIn}/ (penguin).
 Given enough phonological similarity between the non-native word and its translation in the native language, a naïve listener may succeed at word comprehension when listening to an unfamiliar language.
@@
 For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and orthographic similarity of the former to the equivalent in their native language.
-
-When comparing the phonemic transcription of the French to that of its Spanish translation /ˈpweɾ.ta/ (*puerta*), one may be led to infer that these two words do not have much overlap.
-But one thing to note is that the voiced uvular fricative consonant /ʁ/ and the open-mid back rounded vowel /ɔ/, which do not exist in Spanish as phonemic categories, can be perceived as allophonic variations the native (Spanish) phonemes /ɾ/ and /o/, as described by assimilation models of phoneme perception [@best2001discrimination; @best1988examination].
-Although this phonetic mismatch between the non-native word and its translation can have a noticeable toll on comprehension @cutler2004patterns, the Spanish listener may still be able to activate the phonological representation of /ˈpweɾ.ta/ (*puerta*) from hearing the word-form /ˈpɔʁt/ (*porte*), ultimately allowing them to access the correct semantic representation.
-Native prosody also constrains the occurrence of homophonic translations.
-For instance, @kentner2015rhythmic presented in the auditory modality French-native and German-native participants with English songs with lyrics.
-Homophonic translations followed language-specific segmentation strategies.
-Most homophonic translations reported by German-native participants resulted from inserting phrase and word boundaries before stressed syllables (following the more frequent trochaic pattern in German).
-On the contrary, most homophonic translations reported by French-native participants resulted from inserting phrase and word boundaries after stressed syllables, following the more frequent iambic pattern in French.
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and (orthographic similarity) of the former to the equivalent in their native language.
 
@@
 We therefore predicted that participants’ performance should increase when the translation pairs are phonologically more similar.
-We also predicted that there is a minimum threshold of phonological similarity to be sufficient for translation.
-
-We further investigated the effect of phonological competitors.
+We also predicted that there is a minimum threshold of phonological similarity to be sufficient for correct translation.
+
+We further investigated the effect of phonological competition.
 Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language.
@@
 In the final section of this article, we present analyses on the joint data sets of Experiments 1 to 3.
-
 :::
@@
   format("%B %dth, %Y")
+collection_dates <- gsub(" 0", " ", collection_dates)
 
@@
 ```
-We collected data from `r n_participants` British English-native adults living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female[1]` female).
+We collected data from `r n_participants` British English-native adults living in the United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female[1]` female).
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
-Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits).
+Participants were recruited via Prolific (£5 compensation) and SONA (compensation in academic credits).
 Participants gave informed consent before providing any data.
@@
     age_min = "Range",
-    l2_lst = "L2"
+    l2_lst = "Second language"
   ) |>
@@
 
-We created two lists of input words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
+We created two lists of words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
 Words in the Catalan list were `r round(lengths$cat_eng$ipa_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_1_sd, 2)`, Range = `r round(lengths$cat_eng$ipa_1_min, 2)`-`r round(lengths$cat_eng$ipa_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, Range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`).
@@
 
-For each word pair in the Catalan and Spanish lists, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word and their correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
-*Frequency* was included as a nuance predictor, under the hypothesis that---keeping other predictors constant---participants would provide correct responses more often when the correct translations are high frequency words in their native language.
+For each word pair in the Catalan and Spanish lists, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word and its correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
+*Frequency* was included as a nuisance predictor, under the hypothesis that---keeping other predictors constant---participants would provide correct responses more often when the correct translations are high frequency words in their native language.
 Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores.
@@
 For example, the *table* (`teIb@l`)-*mesa* (`mesa`) translation pair had `r scales::percent(stringdist::stringsim(enc2utf8("teɪbəl"), enc2utf8("mesa")))` similarity, while the *train* (`trEIn`)-*tren* (`tREn`) translation pair had `r scales::percent(stringdist::stringsim(enc2utf8("trEIn"), enc2utf8("tɾEn")))` similarity.
-We calculated the number of *CLPN* for each Catalan and Spanish word by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) differed in up to one phoneme from that of the presented Catalan or Spanish word.
-Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
+We calculated the number of *CLPN* for each Catalan and Spanish word by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) differed by up to one phoneme from that of the presented Catalan or Spanish word.
+Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
 @tbl-stimuli summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
@@
 Participants were randomly assigned to the Catalan or Spanish lists.
-Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="cat-ENG",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="spa-ENG",])` trials.
+Participants in the Catalan list were presented with `r nrow(stimuli[stimuli$group=="cat-ENG",])` trials, and participants in the Spanish list were presented with `r nrow(stimuli[stimuli$group=="spa-ENG",])` trials^[Stimuli lists were built from a pool of audio recordings made for a different study.
+The different number of trials in the Catalan and Spanish lists is due to more Spanish audios being available than Catalan audios.].
 
@@
 Then, participants pressed the `RETURN/ENTER` key to confirm their answer and start a new trial.
+@fig-design illustrates the time course of a trial in the translation elicitation task.
 
@@
 After the offset of the audio, the fixation point disappeared and a visual prompt (`>`) was presented.
-Participants then wrote their answer and clicked `RETURN`, making the end of the trial.](_assets/img/design.png)
+Participants then wrote their answer and clicked `RETURN`, making the end of the trial.](_assets/img/design.png){#fig-design}
 
@@
 
-To test the practical relevance of each predictor we followed @kruschke2018bayesian.
+We followed @kruschke2018bayesian's guidelines for statistical inference.
 We first specified a region of practical equivalence (ROPE) around zero ([-0.1, +0.1], in the logit scale).
-This area indicates the values of the regression coefficients that we considered as equivalent to zero.
-We then computed the 95% posterior credible intervals (CrI) of the regression coefficients of interest.
-Finally, we calculated the proportion of the 95% CrI that fell inside the ROPE.
+This area indicates the values of the regression coefficients that we considered equivalent to zero.
+We then computed the 95% posterior credible intervals (CrI) of the regression coefficients of interest, which indicates the most likely range of values that contains the true value of the coefficient with 95% probability.
+Finally, for each regression coefficient we calculated the proportion of the 95% CrI that fell inside the ROPE, noted as *p*(ROPE).
 This proportion indicates the probability that the true value of the coefficient is equivalent to zero.
-All analyses were performed in R environment [@rcoreteam2013language].
+For instance, if the *p*(ROPE) of a regression coefficient $\beta$ is 0.5, this indicates that there is a 50% probability that the true value of $\beta$ is zero or equivalent to zero, given the data.
+If *p*(ROPE)=0.01, this indicates that there is a 1% probability that the true value of the regression coefficient is zero or equivalent.
+If *p*(ROPE)=0.99, this indicates that there is a 99% probability that the true value of the regression coefficient is zero or equivalent.
+
+All analyses were performed in the R environment [@rcoreteam2013language].
 We used the tidyverse family of R packages [@wickham2019welcome] to process data and to generate figures.
-We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for mode details on the models).
+We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for more details on the models).
 Numeric predictors were standardised before entering the model by subtracting the mean and dividing by the standard deviation.
-Categorical predictors were sum-coded as `-0.5` and `+0.5` [@schad2020capitalize].
 
@@
 
-We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` unique participants.
+We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` participants.
 We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_eng["blank"]`), in which a response in a language other than English was provided (e.g., "`agua`", *n* = `r r_validity$cat_eng["language"]`), in which participants did not provide a whole word (e.g., "`f`", *n* = `r r_validity$cat_eng["incomplete"]`), and in which participants added comments to the experimenter (e.g., "`unsure`", *n* = `r r_validity$cat_eng["other"]`).
-In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
-
-After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id) + dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` unique participants.
-Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
+In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language disorder (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+
+After applying trial-level and participant-level exclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$cat_eng), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id) + dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` participants.
+Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` participants who listened to Spanish words.
 Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, Range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, Range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
@@
 @tbl-dataset shows a summary of participants' accuracy across Experiments 1, 2, and 3.
-MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
 Participants translating Catalan words and participants translating Spanish words performed equivalently, as indicated by the regression coefficient of *Group* ($\beta$ = `r round(c1$b_group1$.median, 3)`, 95% CrI = [`r round(c1$b_group1$.lower, 3)`, `r round(c1$b_group1$.upper, 3)`], *p*(ROPE) = `r round(c1$b_group1$.rope, 3)`).
@@
 This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c1$b_neigh_n_h_std_lv_std$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_std_lv_std$.lower, 3)`, `r round(c1$b_neigh_n_h_std_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_std_lv_std$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c1$b_lv_std$.median, 3)`, 95% CrI = [`r round(c1$b_lv_std$.lower, 3)`, `r round(c1$b_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv_std$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h_std$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_std$.lower, 3)`, `r round(c1$b_neigh_n_h_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_std$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c1$b_lv_std$.median, 3)`, 95% CrI = [`r round(c1$b_lv_std$.lower, 3)`, `r round(c1$b_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_lv_std$.rope, 3)`), while the number of *CLPN* had the opposite effect ($\beta$ = `r round(c1$b_neigh_n_h_std$.median, 3)`, 95% CrI = [`r round(c1$b_neigh_n_h_std$.lower, 3)`, `r round(c1$b_neigh_n_h_std$.upper, 3)`], *p*(ROPE) = `r round(c1$b_neigh_n_h_std$.rope, 3)`).
 @fig-epreds-1 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
@@
     correct_min = "Range"
-  ) |>
-  summary_rows(
-    columns = is.integer,
-    fns = list(
-      label = md("*Sum*"),
-      id = "totals",
-      fn = "sum"
-    ),
-    fmt = ~ fmt_integer(.)
-  ) |>
-  summary_rows(
-    columns = matches("mean|sd|sem"),
-    fns = list(
-      label = md("*Mean*"),
-      id = "means",
-      fn = "mean"
-    ),
-    fmt = ~ fmt_number(., scale_by = 100)
   ) |>
@@
 Participants benefited strongly from phonological similarity when the correct translation of the presented words in Catalan or Spanish had few English phonological neighbours with higher lexical frequency.
+As the number of higher-frequency English phonological neighbours increased, both participants' accuracy and the impact of phonological similarity approached zero.
 This suggests that word-forms in an unfamiliar language have a strong potential to activate their translation equivalents in the native language, provided some phonological similarity between both words, and the absence of more frequent phonological neighbors.
@@
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
-Participants in Spain were contacted via announcements at the University campus(es), and were compensated 5€ or an Amazon voucher for the same value.
+Participants in Spain were contacted via announcements at the University campus(es), and were compensated €5 or an Amazon voucher for the same value.
 Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
@@
 
-We collected data for a total of `r format(nrow(dataset$cat_spa) + nrow(dataset$cat_spa), big.mark = ",")` trials completed by `r n_participants$cat_spa` unique participants.
+We collected data for a total of `r format(nrow(dataset$cat_spa) + nrow(dataset$cat_spa), big.mark = ",")` trials completed by `r n_participants$cat_spa` participants.
 We excluded trials in which participants did not enter any text (*n* = `r r_validity$cat_spa["blank"]`), in which a response in a language other than Spanish was provided (*n* = `r r_validity$cat_spa["language"]`), in which participants did not provide a whole word (*n* = `r r_validity$cat_spa["incomplete"]`), and in which participants added comments to the experimenter (*n* = `r r_validity$cat_spa["other"]`).
-In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a developmental language disorder (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
-After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id) + dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` unique participants.
-Of those trials, `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` unique participants who listened to Spanish words.
+In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a developmental language disorder (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id) + dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` participants.
 Responses given by participants were `r round(lengths$len_mean, 2)` characters long on average (*Median* = `r round(lengths$len_median, 2)`, *SD* = `r round(lengths$len_sd, 2)`, Range = `r round(lengths$len_min, 2)`-`r round(lengths$len_max, 2)`).
@@
 
-MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
 Overall, participants responded less accurately to words with more CLPNs than to words with fewer CLPNs, regardless of the amount of phonological similarity between the presented word and its translation.
 This is indicated by the size of the regression coefficient of the two-way interaction between *Similarity* and *CLPN* ($\beta$ = `r round(c2$b_neigh_n_h_std_lv_std$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_std_lv_std$.lower, 3)`, `r round(c2$b_neigh_n_h_std_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_std_lv_std$.rope, 3)`).
-As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c2$b_lv_std$.median, 3)`, 95% CrI = [`r round(c2$b_lv_std$.lower, 3)`, `r round(c2$b_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_lv_std$.rope, 3)`), while the number of *CLND* had the opposite effect ($\beta$ = `r round(c2$b_neigh_n_h_std$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_std$.lower, 3)`, `r round(c2$b_neigh_n_h_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_std$.rope, 3)`).
+As anticipated, participants' performance benefited from an increase in *Similarity* ($\beta$ = `r round(c2$b_lv_std$.median, 3)`, 95% CrI = [`r round(c2$b_lv_std$.lower, 3)`, `r round(c2$b_lv_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_lv_std$.rope, 3)`), while the number of *CLPN* had the opposite effect ($\beta$ = `r round(c2$b_neigh_n_h_std$.median, 3)`, 95% CrI = [`r round(c2$b_neigh_n_h_std$.lower, 3)`, `r round(c2$b_neigh_n_h_std$.upper, 3)`], *p*(ROPE) = `r round(c2$b_neigh_n_h_std$.rope, 3)`).
 @fig-epreds-2 illustrates the posterior of the average predictions of the model for words with different values of *Similarity* and *CLPN*.
@@
   format("%B %dth, %Y") 
+collection_dates <- gsub(" 0", " ", collection_dates)
 
@@
 
-We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` unique participants.
+We collected data for a total of `r format(nrow(dataset$cat_eng) + nrow(dataset$spa_eng), big.mark = ",")` trials completed by `r n_participants$cat_eng + n_participants$spa_eng` participants.
 We excluded trials in which participants did not enter any text (*n* = 0), provided a response in a language other than English (*n* = `r r_validity$cat_eng["language"]`), did not provide a whole word (*n* = 0), or added comments to the experimenter (*n* = 0).
-In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
-
-After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$spa_eng), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id) + dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` unique participants.
-Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` unique participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` unique participants who listened to Spanish words.
+In addition, we excluded data from participants that self-rated their oral and/or written skills in Catalan and Spanish, or any other second language as four or higher in a five-point scale (*n* = `r excluded_lang`), were diagnosed with a language disorder (*n* = `r excluded_impairment`), or did not contribute more than 80% of valid trials (*n* = `r excluded_ntrials`).
+
+After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_eng) + nrow(dataset_clean$spa_eng), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id) + dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` participants.
+Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` participants who listened to Spanish words.
 
@@
 
-MCMC chains in the model showed strong evidence of convergence ($\hat{R}<1.01$) (see Appendix 2 for more detailed model diagnostics).
 Overall, participants reported prior knowledge more often for the Spanish words that showed unexpectedly high accuracy in Experiment 1 (see Discussion in Experiment 2) than for words with expected accuracy (see @fig-knowledge).
@@
 #| fig-width: 7
-#| fig-height: 5
+#| fig-height: 4
 #| fig-cap: "Catalan and Spanish prior word knowledge, as reported by English native participants in Experiment 3.
@@
 Each colour indicates whether the data point corresponds to a word that participants answered to with expected accuracy (red) or with unexpectedly high accuracy (blue), as defined by (no phonological similarity with the correct translation, by accuracy higher than 10%)."
-p <- dataset_3 |>#| 
+dataset_3 |>
   mutate(surprise = translation %in% tbl_data$translation) |>
@@
   )) |> 
-  ggplot(aes(knowledge, accuracy, colour = surprise, fill = surprise, shape = surprise)) +
+  ggplot(aes(knowledge, accuracy)) +
   facet_wrap(~group) +
-  geom_point(stroke = 1, size = 2) +
-  #geom_rug(outside = TRUE, sides = "tr") +
+  geom_point(
+    stroke = 1, 
+    size = 2, 
+    alpha = 0.25, 
+    color = RColorBrewer::brewer.pal(3, "Reds")[3], 
+    shape = 1
+  ) +
   labs(
@@
   scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_shape_manual(values = c(1, 2)) +
-  scale_colour_manual(values = c(
-    RColorBrewer::brewer.pal(3, "Reds")[3], 
-    RColorBrewer::brewer.pal(3, "Blues")[3]
-  )) +
   theme(
@@
 #| label: tbl-results
+#| eval: false
 tbl_data <- list(

### 2025-02-07 manuscript/manuscript.tex
Minor styling changes

@@
 the Spanish Ministry for Science and Innovation and State Research
-Agency (Project PID2021- 123416NB-I00 financed by MCIN/ AEI/
-10.13039/501100011033 / FEDER, UE) and the Economic and Social Research
-Council (ESRC) (ES/S010947/1, UK). GGC was supported by a FPI research
-contract (PRE2019-088165). We are grateful to Irina Lepadatu and Nicola
-Gillen, and to the Speech Acquisition and Perception research group for
-their helpful feedback. We thank Xavier Mayoral, Silvia Blanch, Pamela
-Miller, and Cristina Cuadrado for their technical support. We also thank
-all families and infants who participated in the experiments. This study
-was conducted according to guidelines laid down in the Declaration of
-Helsinki, and was approved by the Drug Research Ethical Committee (CEIm)
-of the IMIM Parc de Salut Mar, reference 2020/9080/I. The data and code
-necessary to reproduce the analyses presented here are publicly
-accessible, as are the materials necessary to attempt to replicate the
-findings. The analyses presented here were not preregistered. Data,
-code, and materials for this research are available at the following
-URLs: https://osf.io/hy984/,
-https://github.com/gongcastro/translation-elicitation.}}
+Agency (Project PGC2018-101831-B-I00 financed by Spanish Ministry of
+Science Innovation and Universities (MCIU), Spanish Research Agency
+(AEI) and the European Regional Development Fund (ERDF), and by the
+Economic and Social Research Council (ESRC) (ES/S010947/1, UK). GGC was
+supported by a FPI research contract (PRE2019-088165). SS was supported
+by XXXXXXXXXXXXX. KP was supported by XXXXXXXXXXXXXXXXX. NSG was
+supported by an ICREA Academia 2019 award from the Catalan Institution
+for Research and Advanced Studies (ICREA). We are grateful to Irina
+Lepadatu and Nicola Gillen, and to the Speech Acquisition and Perception
+research group for their helpful feedback. We thank Xavier Mayoral,
+Silvia Blanch, Pamela Miller, and Cristina Cuadrado for their technical
+support. We also thank all families and infants who participated in the
+experiments. This study was conducted according to guidelines laid down
+in the Declaration of Helsinki, and was approved by the Drug Research
+Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference
+2020/9080/I. The data and code necessary to reproduce the analyses
+presented here are publicly accessible, as are the materials necessary
+to attempt to replicate the findings. The analyses presented here were
+not preregistered. Data, code, and materials for this research are
+available at the following URLs: \url{https://osf.io/hy984/},
+\url{https://github.com/gongcastro/translation-elicitation}.}}
 \def\asep{\\\\\\ } % default: all authors on same column
@@
 
-When some German speakers listened to the song \emph{The Power} by SNAP!
-(\emph{Coyote Ugly}, 2000), many of them misheard the line ``I've got
-the power'' as \emph{Agatha Bauer}. When Spanish speakers listen to the
-line ``Circumvent your thick ego'' from the song \emph{Pictures} by
-System of a Down (\emph{Steal this Album!}, 2002), they tend to hear
-\emph{Sácame de aquí, King-Kong} {[}take me out of here, King-Kong{]}.
-Outrageous as these examples might sound, the reader may know a few
-cases in their own native language. This auditory illusion is common
-across languages and cultures and can feel quite real, often inevitable
+When some German speakers listen to the song \emph{The Power} by SNAP!
+(\emph{Coyote Ugly}, 2000), many of them mishear the line ``I've got the
+power'' as \emph{Agatha Bauer}. When Spanish speakers listen to the line
+``Circumvent your thick ego'' from the song \emph{Pictures} by System of
+a Down (\emph{Steal this Album!}, 2002), they tend to hear \emph{Sácame
+de aquí, King-Kong} {[}take me out of here, King-Kong{]}. Outrageous as
+these examples might sound, the reader may know a few cases in their own
+native language. This auditory illusion is common across languages and
+cultures and can feel quite real, often inevitable
 (\citeproc{ref-dembeck2015oberflachenubersetzung}{Dembeck, 2015};
@@
   translation (e.g., \emph{Señor}, \emph{ínsula}).}. Little attention
-has been paid to homophonic translation as a psycholinguistic event.
-
-In one of the few available studies concerning homophonic translation as
-a psycholinguistic phenomenon, Otake
-(\citeproc{ref-otake2007interlingual}{2007}) analysed 194 instances of
-Soramimi, broadcasted between 1992 and 2007 by the Japanese TV show
+has been paid to homophonic translation from a language processing
+perspective.
+
+One exception is the study by Otake
+(\citeproc{ref-otake2007interlingual}{2007}), who analysed 194 instances
+of Soramimi, broadcasted between 1992 and 2007 by the Japanese TV show
 \emph{Soramimi hour}, in which examples of Soramimi were presented to
@@
 Japanese. The vast majority of the reported instances were phrases
-(96\%) and few were single words (4\%). There was relatively high
+(96\%) and a few, single words (4\%). There was relatively high
 variability in the degree to which phonetic features of the presented
@@
 reconstructed the English input to generate the Japanese words:
-insertion (e.g., \emph{cry} /\textipa{kraI}/ to \emph{kurai},
-/\textipa{kuRai}/ {[}dark{]}), deletion (e.g., \emph{go},
-/\textipa{goU}/ to \emph{go}, /go/ {[}go{]}, and alternation (e.g.,
-\emph{low}, /\textipa{loU}/ to \emph{rou}, /\textipa{Roo}/ {[}wax{]}).
+insertion (e.g., \emph{cry} /\textipa{kraI}/ to \emph{kurai}
+/\textipa{kuRai}/ {[}dark{]}), deletion (e.g., \emph{go} /\textipa{goU}/
+to \emph{go} /go/ {[}\emph{Go}, a board game{]}, and alternation (e.g.,
+\emph{low} /\textipa{loU}/ to \emph{rou} /\textipa{Roo}/ {[}wax{]}).
 This suggests that the reported homophonic translations can be
@@
 (\citeproc{ref-dupoux1999epenthetic}{Dupoux et al., 1999};
-\citeproc{ref-peperkamp2008perceptual}{Peperkamp et al., 2008}). The
-phonological processes described by Otake
+\citeproc{ref-peperkamp2008perceptual}{Peperkamp et al., 2008}).
+
+The phonological processes described by Otake
 (\citeproc{ref-otake2007interlingual}{2007}) do not take place in a
@@
 refers to the number of words-forms that differ in a single phoneme from
-the former. For instance, the English word \emph{cat} hbelongs to a
-dense phonological neighbourhood, with 50 English phonological
-neighbours (e.g., hat, bat, cap), while the word \emph{charger} belongs
-to a sparse phonological neighbourhood, with six neighbours (e.g.,
-charged,larger, charter, charmer). Phonological neighbourhood density
-affects lexical processing. For instance, words from dense
-neighbourhoods are recognized more slowly and less accuractely in the
-auditory modality than words from sparse neighbourhoods
-(\citeproc{ref-dufour2010phonological}{Dufour \& Frauenfelder, 2010};
-e.g., \citeproc{ref-luce1998recognizing}{Luce \& Pisoni, 1998};
-\citeproc{ref-vitevitch2006clustering}{M. Vitevitch, 2006};
-\citeproc{ref-vitevitch1998words}{Vitevitch \& Luce, 1998}). In this
+the it. For instance, the English word \emph{cat} belongs to a dense
+phonological neighbourhood, with 50 English phonological neighbours
+(e.g., hat, bat, cap), while the word \emph{charger} belongs to a sparse
+phonological neighbourhood, with six neighbours (e.g., charged, larger,
+charter, charmer) (\citeproc{ref-marian2012clearpond}{Marian et al.,
+2012}). Phonological neighbourhood density affects lexical processing.
+For instance, words from dense neighbourhoods are recognized more slowly
+and less accuractely in the auditory modality than words from sparse
+neighbourhoods (\citeproc{ref-luce1998recognizing}{Luce \& Pisoni,
+1998}; \citeproc{ref-vitevitch1998words}{Vitevitch \& Luce, 1998}; e.g.,
+\citeproc{ref-dufour2010phonological}{Dufour \& Frauenfelder, 2010};
+\citeproc{ref-vitevitch2006clustering}{M. Vitevitch, 2006}). In this
 study, we examined how listeners may activate lexical representations in
@@
 
-Although homophonic translations do not necessarily take place with the
-preservation of meaning---which results in outrageous mistranslations
-like the aforementioned ones---homophonic translations sometimes may
-lead to correct translations. For example, imagine an English native
-listening to Dutch for the first time. This listener may encounter the
-word /\textipa{"pINgUIn}/ (\emph{pinguin}), which sounds similar to its
-English translation /\textipa{"pENgwIn}/ (penguin). Given enough
-phonological similarity between the non-native word and its translation
-in the native language, a naïve listener may succeed at word
+Although homophonic translations rarely involve preservation of meaning,
+they sometimes lead to correct translations. For example, imagine an
+English native listening to Dutch for the first time. This listener may
+encounter the word /\textipa{"pINgUIn}/ (\emph{pinguin}), which sounds
+similar to its English translation /\textipa{"pENgwIn}/ (penguin). Given
+enough phonological similarity between the non-native word and its
+translation in the native language, a naïve listener may succeed at word
 comprehension when listening to an unfamiliar language. Such
@@
 
-We further investigated the effect of phonological competitors. Even if
+We further investigated the effect of phonological competition. Even if
 the presented word and its target translation share high phonological
@@
 is therefore important to investigate the joint effect of both cognates
-and false friends when investigating the effect cross-linguistic
+and false friends when investigating the effect that cross-linguistic
 phonological similarity has on word recognition in a foreign language.
@@
 with the early stages of second language acquisition. For instance, one
-could expect English participants to incorrectly translate the Spanish
+might expect English participants to incorrectly translate the Spanish
 word \emph{botón} as \emph{bottom} instead of as its correct translation
-\emph{button}, due to combined effect of the close phonological
+\emph{button}, due to the combined effect of the close phonological
 similarity between \emph{botón} and \emph{bottom} along with the high
@@
 
-We collected data from 71 British English-native adults living in United
-Kingdom (\emph{Mean} = 21.76 years, \emph{SD} = 2.15, Range = 18-26, 46
-female). Data collection took place from June 04th, 2020 to June 25th,
-2020. Participants were recruited via Prolific (5£ compensation) and
-SONA (compensation in academic credits). Participants gave informed
-consent before providing any data. The study was conducted in accordance
-with ethical standards of the Declaration of Helsinki and the protocol
-was by the University of Oxford Medical Sciences Inter-Divisional
-Research Ethics Committee (IDREC) (R60939/RE007). Participants were
-asked to complete the experiment using a laptop in a quiet place with
-good internet connection.
+We collected data from 71 British English-native adults living in the
+United Kingdom (\emph{Mean} = 21.76 years, \emph{SD} = 2.15, Range =
+18-26, 46 female). Data collection took place from June 4th, 2020 to
+June 25th, 2020. Participants were recruited via Prolific (£5
+compensation) and SONA (compensation in academic credits). Participants
+gave informed consent before providing any data. The study was conducted
+in accordance with ethical standards of the Declaration of Helsinki and
+the protocol was by the University of Oxford Medical Sciences
+Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+Participants were asked to complete the experiment using a laptop in a
+quiet place with good internet connection.
 
@@
 \cmidrule(lr){3-4}
-\multicolumn{1}{l}{} & N\textsuperscript{\textit{1}} & Mean ± SD  & Range & L2 \\ 
+\multicolumn{1}{l}{} & N\textsuperscript{\textit{1}} & Mean ± SD  & Range & Second language \\ 
 \midrule\addlinespace[2.5pt]
@@
 
-We created two lists of input words to be presented to participants in
-the auditory modality: one in Catalan and one in Spanish. Words in the
+We created two lists of words to be presented to participants in the
+auditory modality: one in Catalan and one in Spanish. Words in the
 Catalan list were 6.67 phonemes long on average (\emph{SD} = 2.06, Range
@@
 translation (\emph{Frequency}), the phonological similarity between the
-presented word and their correct English translation
-(\emph{Similarity}), and the presented word's number of cross-language
-phonological neighbours (\emph{CLPN}) in English. \emph{Frequency} was
-included as a nuance predictor, under the hypothesis that---keeping
-other predictors constant---participants would provide correct responses
-more often when the correct translations are high frequency words in
-their native language. Lexical frequencies of correct translations were
-extracted from SUBTLEX-UK (\citeproc{ref-van2014subtlex}{Van Heuven et
-al., 2014}), and transformed to Zipf scores. Translations without a
-lexical frequency value were excluded from data analysis (2 in the
-Catalan list, 5 in the Spanish list). \emph{Similarity}, our main
-predictor of interest, was calculated as the Levenshtein similarity
-between the X-SAMPA transcriptions of each pair of translations using
-the \texttt{stringdist} R package (\citeproc{ref-van2014stringdist}{van
-der Loo, 2014}). The Levenshtein distance computes the edit distance
-between two character strings---in this case, two phonological
+presented word and its correct English translation (\emph{Similarity}),
+and the presented word's number of cross-language phonological
+neighbours (\emph{CLPN}) in English. \emph{Frequency} was included as a
+nuisance predictor, under the hypothesis that---keeping other predictors
+constant---participants would provide correct responses more often when
+the correct translations are high frequency words in their native
+language. Lexical frequencies of correct translations were extracted
+from SUBTLEX-UK (\citeproc{ref-van2014subtlex}{Van Heuven et al.,
+2014}), and transformed to Zipf scores. Translations without a lexical
+frequency value were excluded from data analysis (2 in the Catalan list,
+5 in the Spanish list). \emph{Similarity}, our main predictor of
+interest, was calculated as the Levenshtein similarity between the
+X-SAMPA transcriptions of each pair of translations using the
+\texttt{stringdist} R package (\citeproc{ref-van2014stringdist}{van der
+Loo, 2014}). The Levenshtein distance computes the edit distance between
+two character strings---in this case, two phonological
 transcriptions---by counting the number of additions, deletions, and
@@
 lexical frequency, and whose phonological transcription (in X-SAMPA
-format) differed in up to one phoneme from that of the presented Catalan
+format) differed by up to one phoneme from that of the presented Catalan
 or Spanish word. Lexical frequencies and phonological transcriptions
 were extracted from the multilingual database CLEARPOND
-(\citeproc{ref-marian2012clearpond}{Marian et al., 2012})\footnote{{[}Phonological
+(\citeproc{ref-marian2012clearpond}{Marian et al., 2012})\footnote{Phonological
   transcriptions in CLEARPOND were generated from eSPEAK,
-  \url{http://espeak.sourceforge.net/}{]}}. Table~\ref{tbl-stimuli}
+  \url{http://espeak.sourceforge.net/}}{]}. Table~\ref{tbl-stimuli}
 summarises the lexical frequency, phonological neighbourhood density and
@@
 Participants in the Catalan list were presented with 83 trials, and
-participants in the Spanish list were presented with 99 trials.
+participants in the Spanish list were presented with 99
+trials\footnote{Stimuli lists were built from a pool of audio recordings
+  made for a different study. The different number of trials in the
+  Catalan and Spanish lists is due to more Spanish audios being
+  available than Catalan audios.}.
 
@@
 pressed the \texttt{RETURN/ENTER} key to confirm their answer and start
-a new trial.
-
-\begin{figure}[H]
-
-{\centering \pandocbounded{\includegraphics[keepaspectratio]{_assets/img/design.png}}
+a new trial. Figure~\ref{fig-design} illustrates the time course of a
+trial in the translation elicitation task.
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{_assets/img/design.png}}
 
@@
 
-\caption{Schematic representation of a trial in the experimental task.
-The trial stars with the presentation of a fixation point in the centre
-of the screen (yellow dot). After 1,000 ms, the auditory word was
-presented while the fixation point remained on screen. After the offset
-of the audio, the fixation point disappeared and a visual prompt
-(\texttt{\textgreater{}}) was presented. Participants then wrote their
-answer and clicked \texttt{RETURN}, making the end of the trial.}
+\caption{\label{fig-design}Schematic representation of a trial in the
+experimental task. The trial stars with the presentation of a fixation
+point in the centre of the screen (yellow dot). After 1,000 ms, the
+auditory word was presented while the fixation point remained on screen.
+After the offset of the audio, the fixation point disappeared and a
+visual prompt (\texttt{\textgreater{}}) was presented. Participants then
+wrote their answer and clicked \texttt{RETURN}, making the end of the
+trial.}
 
@@
 
-To test the practical relevance of each predictor we followed Kruschke
-\& Liddell (\citeproc{ref-kruschke2018bayesian}{2018}). We first
-specified a region of practical equivalence (ROPE) around zero ({[}-0.1,
-+0.1{]}, in the logit scale). This area indicates the values of the
-regression coefficients that we considered as equivalent to zero. We
-then computed the 95\% posterior credible intervals (CrI) of the
-regression coefficients of interest. Finally, we calculated the
-proportion of the 95\% CrI that fell inside the ROPE. This proportion
-indicates the probability that the true value of the coefficient is
-equivalent to zero. All analyses were performed in R environment
+We followed Kruschke \& Liddell
+(\citeproc{ref-kruschke2018bayesian}{2018})'s guidelines for statistical
+inference. We first specified a region of practical equivalence (ROPE)
+around zero ({[}-0.1, +0.1{]}, in the logit scale). This area indicates
+the values of the regression coefficients that we considered equivalent
+to zero. We then computed the 95\% posterior credible intervals (CrI) of
+the regression coefficients of interest, which indicates the most likely
+range of values that contains the true value of the coefficient with
+95\% probability. Finally, for each regression coefficient we calculated
+the proportion of the 95\% CrI that fell inside the ROPE, noted as
+\emph{p}(ROPE). This proportion indicates the probability that the true
+value of the coefficient is equivalent to zero. For instance, if the
+\emph{p}(ROPE) of a regression coefficient \(\beta\) is 0.5, this
+indicates that there is a 50\% probability that the true value of
+\(\beta\) is zero or equivalent to zero, given the data. If
+\emph{p}(ROPE)=0.01, this indicates that there is a 1\% probability that
+the true value of the regression coefficient is zero or equivalent. If
+\emph{p}(ROPE)=0.99, this indicates that there is a 99\% probability
+that the true value of the regression coefficient is zero or equivalent.
+
+All analyses were performed in the R environment
 (\citeproc{ref-rcoreteam2013language}{R Core Team, 2013}). We used the
@@
 (\citeproc{ref-carpenter2017stan}{Carpenter et al., 2017}) to estimate
-and compare the models (see Appendix 1 for mode details on the models).
+and compare the models (see Appendix 1 for more details on the models).
 Numeric predictors were standardised before entering the model by
-subtracting the mean and dividing by the standard deviation. Categorical
-predictors were sum-coded as \texttt{-0.5} and \texttt{+0.5}
-(\citeproc{ref-schad2020capitalize}{Schad et al., 2020}).
+subtracting the mean and dividing by the standard deviation.
 
@@
 
-We collected data for a total of 6,446 trials completed by 72 unique
+We collected data for a total of 6,446 trials completed by 72
 participants. We excluded trials in which participants did not enter any
@@
 four or higher in a five-point scale (\emph{n} = 2), were diagnosed with
-a language (\emph{n} = 1), or did not contribute more than 80\% of valid
-trials (\emph{n} = 9).
-
-After applying trial-level and participant-level inclusion criteria, the
-resulting dataset included 5,204 trials provided by 54 unique
-participants. Of those trials, 2,602 were provided by 27 unique
-participants who listened to Catalan words, and 2,604 trials were
-provided by 32 unique participants who listened to Spanish words.
-Responses given by English participants to Catalan presented words were
-5.35 characters long on average (\emph{Median} = 5, \emph{SD} = 1.79,
-Range = 1-14), while their translations to Spanish responses were 5.57
-characters long on average (\emph{Median} = 5, \emph{SD} = 1.97, Range =
-2-21).
+a language disorder (\emph{n} = 1), or did not contribute more than 80\%
+of valid trials (\emph{n} = 9).
+
+After applying trial-level and participant-level exclusion criteria, the
+resulting dataset included 5,204 trials provided by 54 participants. Of
+those trials, 2,602 were provided by 27 participants who listened to
+Catalan words, and 2,604 trials were provided by 32 participants who
+listened to Spanish words. Responses given by English participants to
+Catalan presented words were 5.35 characters long on average
+(\emph{Median} = 5, \emph{SD} = 1.79, Range = 1-14), while their
+translations to Spanish responses were 5.57 characters long on average
+(\emph{Median} = 5, \emph{SD} = 1.97, Range = 2-21).
 
 Table~\ref{tbl-dataset} shows a summary of participants' accuracy across
-Experiments 1, 2, and 3. MCMC chains in the model showed strong evidence
-of convergence (\(\hat{R}<1.01\)) (see Appendix 2 for more detailed
-model diagnostics). Participants translating Catalan words and
+Experiments 1, 2, and 3. Participants translating Catalan words and
 participants translating Spanish words performed equivalently, as
@@
 \emph{Similarity} (\(\beta\) = 1.437, 95\% CrI = {[}1.314, 1.544{]},
-\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
+\emph{p}(ROPE) = 0), while the number of \emph{CLPN} had the opposite
 effect (\(\beta\) = -0.167, 95\% CrI = {[}-0.352, -0.01{]},
@@
 cat-ENG & $32$ & $18.48$ & $4.89$ & $3.27$ & $10.84$–$32.56$ & $81.38$ & $2,604$ & $3.17$ & $71$–$83$ \\ 
-\midrule 
-\emph{Sum} & $59$ & — & — & — & — & — & $5,206$ & — & $158$ \\ 
-\emph{Mean} & — & $17.17$ & $5.05$ & $3.16$ & — & $8,887.27$ & — & $302.71$ & — \\ 
 \midrule\addlinespace[2.5pt]
@@
 cat-SPA & $21$ & $48.30$ & $5.29$ & $10.54$ & $38.27$–$58.97$ & $79.14$ & $1,662$ & $3.14$ & $72$–$82$ \\ 
-\midrule 
-\emph{Sum} & $21$ & — & — & — & — & — & $1,662$ & — & $72$ \\ 
-\emph{Mean} & — & $48.30$ & $5.29$ & $10.54$ & — & $7,914.29$ & — & $313.51$ & — \\ 
 \midrule\addlinespace[2.5pt]
@@
 cat-ENG & $32$ & $19.74$ & $4.94$ & $3.49$ & $10.34$–$27.91$ & $82.75$ & $2,648$ & $0.44$ & $82$–$83$ \\ 
-\midrule 
-\emph{Sum} & $63$ & — & — & — & — & — & $5,669$ & — & $170$ \\ 
-\emph{Mean} & — & $20.33$ & $6.62$ & $3.62$ & — & $9,010.08$ & — & $112.22$ & — \\ 
 \bottomrule
@@
 Catalan or Spanish had few English phonological neighbours with higher
-lexical frequency. This suggests that word-forms in an unfamiliar
-language have a strong potential to activate their translation
-equivalents in the native language, provided some phonological
-similarity between both words, and the absence of more frequent
-phonological neighbors.
+lexical frequency. As the number of higher-frequency English
+phonological neighbours increased, both participants' accuracy and the
+impact of phonological similarity approached zero. This suggests that
+word-forms in an unfamiliar language have a strong potential to activate
+their translation equivalents in the native language, provided some
+phonological similarity between both words, and the absence of more
+frequent phonological neighbors.
 
@@
 (\emph{Mean} = 21.85 years, \emph{SD} = 3, Range = 18-33, 28, 5 female).
-Data collection took place from June 08th, 2020 to June 28th, 2020.
+Data collection took place from June 8th, 2020 to June 28th, 2020.
 Participants in Spain were contacted via announcements at the University
-campus(es), and were compensated 5€ or an Amazon voucher for the same
+campus(es), and were compensated €5 or an Amazon voucher for the same
 value. Participants gave informed consent before providing any data and
@@
 
-We collected data for a total of 5,412 trials completed by 33 unique
+We collected data for a total of 5,412 trials completed by 33
 participants. We excluded trials in which participants did not enter any
@@
 data from participants that self-rated their oral and/or written skills
-in Catalan and Spanish, or any other second language as four or higher
-in a five-point scale (\emph{n} = 22), were diagnosed with a
-developmental language disorder (\emph{n} = 1), or did not contribute
-more than 80\% of valid trials (\emph{n} = 9). After applying
-trial-level and participant-level inclusion criteria, the resulting
-dataset included 1,662 trials provided by 42 unique participants. Of
-those trials, 1,662 were provided by 21 unique participants who listened
-to Catalan words, and 1,662 trials were provided by 21 unique
-participants who listened to Spanish words. Responses given by
+in Catalan or any other second language as four or higher in a
+five-point scale (\emph{n} = 22), were diagnosed with a developmental
+language disorder (\emph{n} = 1), or did not contribute more than 80\%
+of valid trials (\emph{n} = 9). After applying trial-level and
+participant-level inclusion criteria, the resulting dataset included
+1,662 trials provided by 42 participants. Responses given by
 participants were 5.6 characters long on average (\emph{Median} = 5,
@@
 
-MCMC chains in the model showed strong evidence of convergence
-(\(\hat{R}<1.01\)) (see Appendix 2 for more detailed model diagnostics).
 Overall, participants responded less accurately to words with more CLPNs
@@
 \emph{Similarity} (\(\beta\) = 1.906, 95\% CrI = {[}1.712, 2.086{]},
-\emph{p}(ROPE) = 0), while the number of \emph{CLND} had the opposite
+\emph{p}(ROPE) = 0), while the number of \emph{CLPN} had the opposite
 effect (\(\beta\) = -0.274, 95\% CrI = {[}-0.613, 0.088{]},
@@
 
-We collected data for a total of 6,016 trials completed by 64 unique
+We collected data for a total of 6,016 trials completed by 64
 participants. We excluded trials in which participants did not enter any
@@
 four or higher in a five-point scale (\emph{n} = 2), were diagnosed with
-a language (\emph{n} = 0), or did not contribute more than 80\% of valid
-trials (\emph{n} = 1).
+a language disorder (\emph{n} = 0), or did not contribute more than 80\%
+of valid trials (\emph{n} = 1).
 
 After applying trial-level and participant-level inclusion criteria, the
-resulting dataset included 5,888 trials provided by 63 unique
-participants. Of those trials, 3,145 were provided by 31 unique
-participants who listened to Catalan words, and 2,743 trials were
-provided by 32 unique participants who listened to Spanish words.
+resulting dataset included 5,888 trials provided by 63 participants. Of
+those trials, 3,145 were provided by 31 participants who listened to
+Catalan words, and 2,743 trials were provided by 32 participants who
+listened to Spanish words.
 
@@
 
-MCMC chains in the model showed strong evidence of convergence
-(\(\hat{R}<1.01\)) (see Appendix 2 for more detailed model diagnostics).
 Overall, participants reported prior knowledge more often for the
@@
 (see Discussion in Experiment 2) than for words with expected accuracy
-(see \textbf{?@fig-knowledge}). Participants reported prior knowledge of
-Catalan words with unexpected accuracy in Experiment 1 as often as those
-with expected accuracy. This suggests that participants in Experiment 1
-may have relied, to some extent, on their prior knowledge about
-form-meaning mappings to correctly translate some Spanish words. To
-isolate such an effect of prior Spanish knowledge, we run the same
+(see Figure~\ref{fig-knowledge}). Participants reported prior knowledge
+of Catalan words with unexpected accuracy in Experiment 1 as often as
+those with expected accuracy. This suggests that participants in
+Experiment 1 may have relied, to some extent, on their prior knowledge
+about form-meaning mappings to correctly translate some Spanish words.
+To isolate such an effect of prior Spanish knowledge, we run the same
 analysis as in Experiment 1 on the newly collected translations from
@@
 reported prior knowledge.
+
+\begin{figure}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-knowledge-1.pdf}}
+
+}
+
+\caption{\label{fig-knowledge}Catalan and Spanish prior word knowledge,
+as reported by English native participants in Experiment 3. The X-axis
+indicates the proportion of participants that reported prior knowledge
+with the word. The Y-axis indicates participants' accuracy (the
+proportion of correct responses by participants). Each colour indicates
+whether the data point corresponds to a word that participants answered
+to with expected accuracy (red) or with unexpectedly high accuracy
+(blue), as defined by (no phonological similarity with the correct
+translation, by accuracy higher than 10\%).}
+
+\end{figure}%
 
@@
 meaning of words they were presented with.
-
-\begin{longtable}{l|rrrrrrrr}
-
-\caption{\label{tbl-results}}
-
-\tabularnewline
-
-\toprule
-\multicolumn{1}{l}{} & \multicolumn{4}{c}{Correct responses} & \multicolumn{4}{c}{Incorrect responses} \\ 
-\cmidrule(lr){2-5} \cmidrule(lr){6-9}
-\multicolumn{1}{l}{} & Correct & (\%) & Typo & (\%) & Wrong & (\%) & False friend & (\%) \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{9}{l}{Experiment 1} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $429$ & ($16.47$) & $11$ & ($0.42$) & $2,082$ & ($79.95$) & $82$ & ($3.15$) \\ 
-spa-ENG & $374$ & ($14.37$) & $2$ & ($0.08$) & $2,117$ & ($81.36$) & $109$ & ($4.19$) \\ 
-\midrule 
-\emph{Sum} & $803$ & — & $13$ & — & $4,199$ & — & $191$ & — \\ 
-\emph{Mean} & — & $15.42$ & — & $0.25$ & — & $80.66$ & — & $3.67$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{9}{l}{Experiment 2} \\ 
-\midrule\addlinespace[2.5pt]
-cat-SPA & $780$ & ($46.93$) & $20$ & ($1.20$) & $736$ & ($44.28$) & $126$ & ($7.58$) \\ 
-\midrule 
-\emph{Sum} & $780$ & — & $20$ & — & $736$ & — & $126$ & — \\ 
-\emph{Mean} & — & $46.93$ & — & $1.20$ & — & $44.28$ & — & $7.58$ \\ 
-\midrule\addlinespace[2.5pt]
-\multicolumn{9}{l}{Experiment 3} \\ 
-\midrule\addlinespace[2.5pt]
-cat-ENG & $477$ & ($18.01$) & $7$ & ($0.26$) & $1,986$ & ($75.00$) & $178$ & ($6.72$) \\ 
-spa-ENG & $590$ & ($19.53$) & $6$ & ($0.20$) & $2,294$ & ($75.94$) & $131$ & ($4.34$) \\ 
-\midrule 
-\emph{Sum} & $1,067$ & — & $13$ & — & $4,280$ & — & $309$ & — \\ 
-\emph{Mean} & — & $18.77$ & — & $0.23$ & — & $75.47$ & — & $5.53$ \\ 
-\bottomrule
-
-\end{longtable}
 

### 2025-03-22 manuscript/manuscript.qmd
Finish manuscript, appendix, clean repo, and add Docker image

@@
 
-::: {.content-visible when-format="pdf"}
-
 When some German speakers listen to the song *The Power* by SNAP!
@@
 In the final section of this article, we present analyses on the joint data sets of Experiments 1 to 3.
-
-:::
-
-::: {.content-visible when-format="docx"}
-
-When some German speakers listen to the song *The Power* by SNAP!
-(*Coyote Ugly*, 2000), many of them mishear the line "I've got the power" as *Agatha Bauer*.
-When Spanish speakers listen to the line "Circumvent your thick ego" from the song *Pictures* by System of a Down (*Steal this Album!*, 2002), they tend to hear *Sácame de aquí, King-Kong* [take me out of here, King-Kong].
-Outrageous as these examples might sound, the reader may know a few cases in their own native language.
-This auditory illusion is common across languages and cultures and can feel quite real, often inevitable [@dembeck2015oberflachenubersetzung; @efimova2018homophonic].
-In Japanese, this phenomenon takes the name *Soramimi* (lit.
-"empty ear").
-From a linguistic point of view, Soramimi is a particular case of homophonic translation: words or phrases that appear in the speech stream in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
-Homophonic translation has mostly received attention as a perceptual curiosity, and as a literary figure, intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
-For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
-Little attention has been paid to homophonic translation from a language processing perspective.
-
-One exception is the study by @otake2007interlingual, who analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the TV show *Soramimi hour*, in which examples of Soramimi were presented the the audience for comedy purposes.
-These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
-The vast majority of the reported instances were phrases (96%) and a few, single words (4%).
-There was relatively high variability in the degree to which phonetic features of the presented English lyrics were preserved, with some resulting Japanese words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
-An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /krʌɪ/ to *kurai*, /kuɾai/ [dark]), deletion (e.g., *go* /goʊ/ to *go* /go/ [*Go*, a board game], and alternation (e.g., *low* /loʊ/ to *rou* /ɾoo/ [wax]).
-This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
-
-The phonological processes described by @otake2007interlingual do not take place in a vaccum, but rather in the context of a lexicon.
-As words or phrases in the native lexicon are activated by their acoustic similarity with the non-native speech stream, they trigger a series of mechanisms that result in the selection of one of the candidate word or phrase in the native language.
-In the present work, we investigated the interplay between phonological similarity and one of the main mechanisms involved in the dynamics of lexical selection: phonological neighbourhood density.
-The phonological neighbourhood density of a given word-form refers to the number of words-forms that differ in a single phoneme from the it.
-For instance, the English word *cat* belongs to a dense phonological neighbourhood, with 50 English phonological neighbours (e.g., hat, bat, cap), while the word *charger* belongs to a sparse phonological neighbourhood, with six neighbours (e.g., charged, larger, charter, charmer) [@marian2012clearpond].
-Phonological neighbourhood density affects lexical processing.
-For instance, words from dense neighbourhoods are recognized more slowly and less accuractely in the auditory modality than words from sparse neighbourhoods [@luce1998recognizing; @vitevitch1998words; e.g.,  @dufour2010phonological; @vitevitch2006clustering].
-In this study, we examined how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
-
-Although homophonic translations rarely involve preservation of meaning, they sometimes lead to correct translations.
-For example, imagine an English native listening to Dutch for the first time.
-This listener may encounter the word /\textipa{"pINgUIn}/ (*pinguin*), which sounds similar to its English translation /\textipa{"pENgwIn}/ (penguin).
-Given enough phonological similarity between the non-native word and its translation in the native language, a naïve listener may succeed at word comprehension when listening to an unfamiliar language.
-Such phonological similarity between translation equivalents---known as *cognateness*---is common across many languages, and is often due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
-For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and (orthographic similarity) of the former to the equivalent in their native language.
-
-To explore the lexical mechanisms triggered by unfamiliar speech listening, we used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess the words translations in their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as *presented words*, and the correct translation for the presented words as *target translations*.
-By manipulating the amount of phonological similarity between the presented words and their target translations, we explored listeners' reliance on phonological similarity to succeed in the task.
-Since participants were unfamiliar with the presented language, they could only use phonological similarity between the presented language and their native language to successfully translate the words.
-We therefore predicted that participants’ performance should increase when the translation pairs are phonologically more similar.
-We also predicted that there is a minimum threshold of phonological similarity to be sufficient for correct translation.
-
-We further investigated the effect of phonological competition.
-Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language.
-Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
-This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target word [@luce1998recognizing].
-Such interference effect of phonological neighbourhood density has also been reported across languages [@weber2004lexical].
-We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
-In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
-For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
-It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect that cross-linguistic phonological similarity has on word recognition in a foreign language.
-This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
-For instance, one might expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to the combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
-To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme apart from the presented word.
-If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.
-We conducted a series of three experiments to test these predictions.
-
-In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
-One group was presented with Catalan words, the other with Spanish words.
-We examined the extent to which participants were able to use the phonological similarity between the presented word (in Catalan or Spanish) and its target translation to provide accurate responses.
-In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are two Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant’s performance.
-One unexpected finding was that participants in Experiments 1 and 2 were surprisingly good at translating a subset of words which had low phonological similarity with their correct translation.
-We were concerned that this may be caused by some prior knowledge of specific words by our participants, as some Spanish words are commonly seen in media or product labels, making them familiar even to monolingual speakers of English.
-In Experiment 3, we collected additional data from a new group of British English natives.
-The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
-In the final section of this article, we present analyses on the joint data sets of Experiments 1 to 3.
-:::
 
@@
 Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
-@tbl-stimuli summarises the lexical frequency, phonological neighbourhood density and phonological overlap of the words included in the Catalan and the Spanish lists.
+@tbl-stimuli summarises the lexical frequency, *CLPN* and phonological overlap of the words included in the Catalan and the Spanish lists.
 
@@
 
-::: {.content-visible when-format="docx"}
-![Eq. 1](_assets/img/model.PNG)
-:::
-
-::: {.content-visible when-format="pdf"}
 $$
@@
 
-:::
-
 We followed @kruschke2018bayesian's guidelines for statistical inference.
@@
 We used the tidyverse family of R packages [@wickham2019welcome] to process data and to generate figures.
-We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for more details on the models).
+We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for model diagnostics).
 Numeric predictors were standardised before entering the model by subtracting the mean and dividing by the standard deviation.
@@
   fmt_number(is.numeric, decimals = 2) |>
-  # gtExtras::gt_plt_dist(correct_list, type = "density") |>
   fmt_integer(c(matches("sum|min|max"), n_participants)) |>
@@
     y = "p(Correct)",
-    colour = "Cross-language neighbourhood density",
-    fill = "Cross-language neighbourhood density",
-    linetype = "Cross-language neighbourhood density",
+    colour = "Cross-language Phonological Neighbours",
+    fill = "Cross-language Phonological Neighbours",
+    linetype = "Cross-language Phonological Neighbours",
     fill_ramp = "CrI (%)"
@@
 In contrast with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 larger.
-Spanish natives in Experiment 2 provided exploited phonological similarity more strongly than English natives in Experiment 1.
-Overall, this suggests that participants in Experiment 2, who were natives of a typologically similar language (Spanish) to the presented language (Catalan), benefited more strongly from phonological similarity than participants in Experiment 1, who were natives of typologically less similar language (English) to the presented language (Catalan, Spanish).
+Spanish natives in Experiment 2 exploited phonological similarity more strongly than English natives in Experiment 1.
+Overall, this suggests that participants in Experiment 2, who were natives of a typologically similar language (Spanish) to the presented language (Catalan), benefited more strongly from phonological similarity than participants in Experiment 1, who were natives of a typologically less similar language (English) to the presented language (Catalan, Spanish).
 Participants from both Experiment 1 and 2 benefited strongly from phonological similarity to correctly translate words from a non-native, reportedly unfamiliar language.
@@
 
-::: {.content-visible when-format="pdf"}
-
-```{r tbl-surprises}
+```{r}
 #| label: tbl-surprises
@@
 ```
-:::
-
-::: {.content-visible when-format="docx"}
-
-```{r tbl-surprises-docx}
-#| label: tbl-surprises-docx
-#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarityscore betwen the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
-tbl_data <- list(
-  "Experiment 1" = dataset_1,
-  "Experiment 2" = dataset_2
-) |>
-  bind_rows(.id = "exp") |>
-  mutate(
-    translation = paste0(word_1, " - ",  word_2),
-    ipa = paste0(ipa_1, " - ",  ipa_2),
-  ) |> 
-  summarise(
-    n = n(),
-    correct = sum(correct),
-    .by = c(exp, group, lv, translation, ipa)
-  ) |>
-  mutate(
-    accuracy = prop_adj(correct, n),
-    accuracy_se = prop_adj_se(correct, n)
-  ) |>
-  dplyr::filter(lv == 0, accuracy >= 0.1) |>
-  select(exp, translation, ipa, group, accuracy, accuracy_se) |>
-  arrange(exp, group, desc(accuracy)) |> 
-  select(translation, ipa, accuracy, accuracy_se) |> 
-  mutate(accuracy = round(accuracy * 100, 2),
-         accuracy_se = round(accuracy_se * 100, 2))
-
-names(tbl_data) <- c("Translation", "IPA", "Accuracy (\\%)", "SE")
-
-tt(tbl_data) |> 
-  style_tt(i = 0, bold = TRUE, line = "tb") |> 
-  group_tt(i = list(
-    "Experiment 1 (cat-ENG)" = 1,
-    "Experiment 2 (spa-ENG)" = 6,
-    "Experiment 3 (cat-SPA)" = 22)) |> 
-  style_tt(i = 1, bold = TRUE, line = "b") |> 
-  style_tt(i = 7, bold = TRUE, line = "tb") |> 
-  style_tt(i = 24, bold = TRUE, line = "tb")
-```
-
-:::
 
@@
 This suggests that participants in Experiment 1 may have relied, to some extent, on their prior knowledge about form-meaning mappings to correctly translate some Spanish words.
-To isolate such an effect of prior Spanish knowledge, we run the same analysis as in Experiment 1 on the newly collected translations from Experiment 3, now excluding responses to words for which participants reported prior knowledge.
+To isolate such an effect of prior Spanish knowledge, we ran the same analysis as in Experiment 1 on the newly collected translations from Experiment 3, now excluding responses to words for which participants reported prior knowledge.
 
@@
 #| label: fig-knowledge
-#| fig-width: 7
+#| fig-width: 9
 #| fig-height: 4
-#| fig-cap: "Catalan and Spanish prior word knowledge, as reported by English native participants in Experiment 3.
+#| fig-cap: "Catalan and Spanish prior word knowledge reported by English native participants in Experiment 3.
 The X-axis indicates the proportion of participants that reported prior knowledge with the word.
-The Y-axis indicates participants' accuracy (the proportion of correct responses by participants).
-Each colour indicates whether the data point corresponds to a word that participants answered to with expected accuracy (red) or with unexpectedly high accuracy (blue), as defined by (no phonological similarity with the correct translation, by accuracy higher than 10%)."
+The Y-axis indicates participants' accuracy (the proportion of participants that translated the word correctly).
+For visualisation purposes, data points have been aggregated so that numbers and color coding show the number of data points with identical accuracy and proportion of reported knowledge."
+
+bin_interval <- seq(0, 1, 0.05)
 dataset_3 |>
-  mutate(surprise = translation %in% tbl_data$translation) |>
+  mutate(surprise = translation %in% tbl_data$Translation) |>
   summarise(
@@
   ) |>
-  mutate(accuracy = prop_adj(correct, n),
-  surprise = if_else(
-    surprise,
-    "Unexpected high accuracy\n(no phonological similarity)",
-    "Expected accuracy"
-  )) |> 
-  ggplot(aes(knowledge, accuracy)) +
+  mutate(
+    accuracy = prop_adj(correct, n),
+  accuracy_binned = cut(accuracy, bin_interval, labels = bin_interval[-length(bin_interval)], include.lowest = TRUE, right = TRUE),
+    knowledge_binned = cut(knowledge, bin_interval, labels = as.numeric(bin_interval[-length(bin_interval)]), include.lowest = TRUE, right = TRUE)
+  ) |> 
+    summarise(
+      n = n(),
+      .by = c(accuracy_binned, knowledge_binned, group)
+    ) |> 
+      mutate(is_gt_20 = n > 20,
+      knowledge_binned = as.double(as.character(knowledge_binned)),
+      accuracy_binned = as.double(as.character(accuracy_binned))) |> 
+  ggplot(aes(knowledge_binned, accuracy_binned, fill = n, label = n, color = is_gt_20)) +
   facet_wrap(~group) +
-  geom_point(
-    stroke = 1, 
-    size = 2, 
-    alpha = 0.25, 
-    color = RColorBrewer::brewer.pal(3, "Reds")[3], 
-    shape = 1
-  ) +
+  geom_tile(color = NA) +
+  geom_text(size = 3) +
   labs(
@@
   ) +
-  scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
-  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
+  scale_fill_distiller(palette = "Reds", direction=1, limits = c(0, 37)) +
+  scale_color_manual(values = c("black", "white")) +
+  scale_x_continuous(labels = scales::percent, limits = c(-0.05, 1)) +
+  scale_y_continuous(labels = scales::percent, limits = c(-0.05, 1), breaks = seq(0, 1, 0.1)) +
   theme(
-    legend.position = "top",
-    legend.title = element_blank(),
+    legend.position = "none",
   )
@@
 After each trial, participants reported whether they had prior knowledge about the word meaning, and how confident they were about such information in a numeric scale from 0 to 7.
-To control for the effect of participants' prior knowledge abbout form-meaning mappings, we excluded from the analyses those words in which participants reported such prior knowledge.
+To control for the effect of participants' prior knowledge about form-meaning mappings, we excluded from the analyses those words in which participants reported such prior knowledge.
 We found equivalent results to those in Experiment 1, indicating that participants' surprisingly good performance in the translation elicitation task was not due to being familiar with the meaning of words they were presented with.
-
-```{r tbl-results}
-#| label: tbl-results
-#| eval: false
-tbl_data <- list(
-  "Experiment 1" = dataset_1,
-  "Experiment 2" = dataset_2,
-  "Experiment 3" = dataset_3
-) |>
-  bind_rows(.id = "exp") |>
-  count(exp, group, response_type) |>
-  pivot_wider(
-    names_from = response_type,
-    values_from = n,
-    names_repair = janitor::make_clean_names
-  ) |>
-  relocate(exp, group, correct, typo, wrong, false_friend) |>
-  mutate(
-    n_correct = rowSums(cbind(correct, typo)),
-    n_incorrect = rowSums(cbind(wrong, false_friend)),
-    n_total = rowSums(cbind(n_correct, n_incorrect)),
-    across(correct:false_friend,
-      \(x) x / n_total,
-      .names = "{col}_prop"
-    )
-  ) |>
-  select(
-    exp, group, correct, typo, wrong, false_friend,
-    matches("prop")
-  ) |>
-  relocate(
-    correct, correct_prop,
-    typo, typo_prop,
-    wrong, wrong_prop,
-    false_friend, false_friend_prop
-  )
-
-tbl_data |>
-  gt(rowname_col = "group", groupname_col = "exp") |>
-  fmt_missing(everything(), everything(), "--") |>
-  tab_spanner(
-    "Correct responses",
-    matches("correct|typo")
-  ) |>
-  tab_spanner(
-    "Incorrect responses",
-    matches("wrong|false")
-  ) |>
-  fmt_integer(c(correct, typo, wrong, false_friend)) |>
-  fmt_number(is.double, pattern = "({x})", scale_by = 100) |>
-  cols_label(
-    correct = "Correct",
-    typo = "Typo",
-    wrong = "Wrong",
-    false_friend = "False friend",
-    correct_prop = "(%)",
-    typo_prop = "(%)",
-    wrong_prop = "(%)",
-    false_friend_prop = "(%)"
-  ) |>
-  tab_style(
-    cell_text(weight = "bold"),
-    cells_column_spanners()
-  ) |>
-  tab_style(
-    cell_text(style = "italic"),
-    cells_column_labels()
-  ) |>
-  summary_rows(
-    columns = is.integer,
-    fns = list(
-      label = md("*Sum*"),
-      id = "totals",
-      fn = "sum"
-    ),
-    fmt = ~ fmt_integer(.)
-  ) |>
-  summary_rows(
-    columns = matches("prop"),
-    fns = list(
-      label = md("*Mean*"),
-      id = "means",
-      fn = "mean"
-    ),
-    fmt = ~ fmt_number(., scale_by = 100)
-  ) |>
-  tab_style(
-    cell_text(align = "center"),
-    cells_column_labels()
-  )
-
-```
 

### 2025-03-22 src/clearpond.R

@@ -1,42 +1,0 @@
-#' Download CLEARPOND
-#' 
-download_clearpond <- function(.language = c("english", "dutch", "french", "spanish", "german")){
-    
-    urls <- tribble(
-        ~lang, ~url, ~data, ~header,
-        "english", "https://clearpond.northwestern.edu/englishCPdatabase2.zip", "englishCPdatabase2.txt", "clearpondHeaders_EN.txt",
-        "dutch", "https://clearpond.northwestern.edu/dutchCPdatabase2.zip", "dutchCPdatabase2.txt", "clearpondHeaders_NL.txt",
-        "french", "https://clearpond.northwestern.edu/frenchCPdatabase2.zip", "frenchCPdatabase2.txt", "clearpondHeaders_FR.txt",
-        "german", "https://clearpond.northwestern.edu/germanCPdatabase2.zip", "germanCPdatabase2.txt", "clearpondHeaders_DE.txt",
-        "spanish", "https://clearpond.northwestern.edu/spanishCPdatabase2.zip", "spanishCPdatabase2.txt", "clearpondHeaders_SP.txt"
-    ) %>%
-        dplyr::filter(lang %in% .language)
-    
-    dir <- tempdir()
-    files <- replicate(tempfile(), n = length(urls$lang))
-    d <- pmap(.l = list(url = as.list(urls$url), 
-                        file = as.list(files),
-                        data = as.list(urls$data),
-                        header = as.list(urls$header)),
-              .f = function(url = .l[[1]],
-                            file = .l[[2]],
-                            data = .l[[3]], 
-                            header = .l[[4]]) {
-                  download.file(url, destfile = file)
-                  unzip(zipfile = file, exdir = dir)
-                  headers <- c("word", read.delim(paste0(dir, .Platform$file.sep, header))[,1])
-                  d <- read.delim(
-                      paste0(dir, .Platform$file.sep, data)) %>%
-                      `colnames<-`(., headers) %>%
-                      as_tibble() %>%
-                      mutate_at(vars(ends_with("W")), ~str_split(., pattern = ";")
-                      )
-                  return(d)
-              }
-    ) %>%
-        set_names(language) %>%
-        bind_rows(.id = "language") %>%
-        janitor::clean_names()
-    
-    return(d)
-}

### 2025-03-25 manuscript/manuscript.qmd
Mention Docker image in manuscript

@@
+---
+abstract: |
+  When listening to speech in an unfamiliar language, words and phrases in the native language are frequently activated due to their acoustic similarity with some parts of the speech stream.
+This study explored this phenomenon, for the first time, from a language processing perspective.
+Across three studies, English- and Spanish-native adults completed a translation elicitation task, in which they listened to a series of words from an unfamiliar language (Catalan or Spanish for English speakers, Catalan for Spanish speakers).
+For each presented word, participants had to type their best-guess translation in their native language.
+Both English and Spanish natives were surprisingly good at translating unfamiliar words, efficiently exploiting the phonological similarity between the presented (unfamiliar) words and their correct translation.
+When the correct translation belonged to high-density phonological neighbourhood, participants' ability to benefit from phonological similarity decreased.
+Spanish participants, whose native language was typologically closer to the presented language, benefited more strongly from phonological similarity than English participants.
+Overall, we show that speech in an unfamiliar language triggers equivalent dynamics of lexical selection than native speech, providing a psycholinguistic account for homophonic translation.
+---
+
 ```{r}
@@
 All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/?view_only=aab7636ce1af48cf832596a7ea9101c5/) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
+For reproducibility, a Docker image of the RStudio session is available on DockerHub ([https://hub.docker.com/repository/docker/gongcastro/translation-elicitation/general](https://hub.docker.com/repository/docker/gongcastro/translation-elicitation/general)).
 

### 2025-04-08 article/appendix.qmd
Rename manuscript folder to article

```{r}
#| label: setup
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(patchwork)
library(gt)
library(gtExtras)
library(brms)
library(tidybayes)
library(ggExtra)
library(tinytable)
library(stringdist)
set.seed(1234)

# import objects
tar_config_set(
  store = here::here("_targets"),
  script = here::here("_targets.R")
)
tar_load_globals()
tar_load(participants)
tar_load(exp_participants)
tar_load(exp_responses)
tar_load(quest_responses)
tar_load(quest_participants)
tar_load(stimuli)
tar_load(epreds)
tar_load(dataset_1)
tar_load(dataset_2)
tar_load(dataset_3)
tar_load(exp_1_m0)
tar_load(exp_2_m0)
tar_load(exp_3_m1)
tar_load(exp_12_m0)

theme_set(theme_ggdist() +
  theme(panel.grid.major.y = element_line(
    colour = "grey",
    linetype = "dotted"
  )))

```

## Appendix 1: Model diagnostics

One way to diagnose the behaviour of Hamiltonian Monte Carlot (HMC, the algorithm used by Stan to explore the posterior distribution of a model) is to check whether the MCMC chains have converged. @fig-appendix-diagnostics shows the values sampled by the MCMC chains of each of the fixed coefficients of each model reported in the manuscript.  Evidence of chain convergence is provided by the same region of values being sampled across the final interations of the chain, as it is the case for the three models depicted.

```{r fig-appendix-diagnostics}
#| label: fig-appendix-diagnostics
#| fig-width: 9
#| fig-height: 9
var_labels <- c(
  "b_Intercept" = "Intercept",
  "b_freq_zipf_2_std" = "Frequency",
  "b_lv_std" = "Cognateness",
  "b_neigh_n_h_std" = "CLPN",
  "b_neigh_n_h_std:lv_std" = "Cognateness × CLPN",
  "b_group1" = "cat-ENG vs. spa-ENG"
)

list(exp_1_m0, exp_2_m0, exp_3_m1) |>
  set_names(paste0("Experiment ", 1:3)) |>
  map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |>
  dplyr::filter(.variable != "b_Intercept") |>
  mutate(
    .value = .value / 4,
    .variable = factor(.variable,
      levels = names(var_labels),
      labels = var_labels
    )
  ) |>
  ggplot(aes(.iteration, .value, colour = factor(.chain))) +
  facet_grid(.variable ~ exp, scales = "free") +
  geom_line() +
  labs(
    x = "Iteration",
    y = "Value",
    colour = "Chain",
    fill = "Chain",
    linetype = "Chain"
  ) +
  scale_colour_brewer(palette = "Reds") +
  theme(
    legend.position = "top",
    legend.box = "horizontal",
    panel.border = element_rect(
      fill = NA,
      colour = "grey"
    ),
    legend.justification = c(1, 0),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

## Appendix 2: Pooled analyses of Experiments 1 and 3

Across Experiments 1 and 3, we found strong evidence that participants efficiently exploited phonological similarity to provide accurate translations for words in an unfamiliar language, provided that few phonological neighbours of higher lexical frequency were present. @fig-coefs summarizes the posterior distribution of the regression coefficients of the models in Experiments 1 to 3.

```{r fig-coefs}
#| label: fig-coefs
#| fig-width: 9
#| fig-height: 5
var_labels <- c(
  "b_Intercept" = "Intercept",
  "b_freq_zipf_2_std" = "Frequency",
  "b_lv_std" = "Similarity",
  "b_neigh_n_h_std" = "CLPN",
  "b_neigh_n_h_std:lv_std" = "Similarity × CLPN",
  "b_group1" = "cat-ENG vs. spa-ENG"
)

list(exp_1_m0, exp_2_m0, exp_3_m1) |>
  set_names(paste0("Experiment ", 1:3)) |>
  map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |>
  dplyr::filter(.variable != "b_Intercept") |>
  mutate(
    .value = .value / 4,
    .variable = factor(.variable,
      levels = names(var_labels),
      labels = var_labels
    )
  ) |>
  ggplot(aes(.value, reorder(exp, desc(exp)))) +
  facet_wrap(~.variable, scales = "free_x") +
  annotate(
    geom = "rect",
    xmin = -0.1 / 4,
    xmax = 0.1 / 4,
    ymin = -Inf,
    ymax = Inf,
    fill = "grey",
    alpha = 3 / 4,
    colour = NA
  ) +
  geom_vline(
    xintercept = 0,
    colour = "grey",
    linewidth = 3 / 4
  ) +
  stat_slab(aes(fill = after_stat(level)),
    point_interval = mode_hdi,
    .width = c(0.95, 0.89, 0.67, 1),
    outline_bars = TRUE,
    colour = "white",
    linewidth = 0.1,
    breaks = 20,
    normalize = "panels"
  ) +
  stat_spike(
    at = "median",
    size = 1,
    aes(linetype = "Median"),
    normalize = "panels"
  ) +
  labs(
    x = "Estimate", y = "Experiment",
    colour = "CrI",
    fill = "CrI",
    linetype = ""
  ) +
  scale_fill_brewer(palette = "Reds") +
  theme(
    legend.position = c(1, 0),
    legend.box = "horizontal",
    legend.justification = c(1, 0),
    axis.title.x = element_blank(),
    axis.title.y = element_blank()
  )
```

### 2025-04-08 docs/manuscript.qmd
Rename article folder to docs

--- 2025-03-25 manuscript/manuscript.qmd
@@
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
-Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.

### 2025-07-24 docs/manuscript.qmd
Update

@@
 library(stringdist)
+
 set.seed(1234)
@@
 )
+
 tar_load_globals()
@@
   )))
-
 ```
@@
 
-```{r tbl-stimuli}
+```{r}
 #| label: tbl-stimuli
@@
 We used the tidyverse family of R packages [@wickham2019welcome] to process data and to generate figures.
-We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for model diagnostics).
+We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see #apx-diagnostics for model diagnostics).
 Numeric predictors were standardised before entering the model by subtracting the mean and dividing by the standard deviation.
@@
 
-```{r tbl-dataset}
+```{r}
 #| label: tbl-dataset
@@
 
-```{r fig-knowledge}
+```{r}
 #| label: fig-knowledge
@@
 For visualisation purposes, data points have been aggregated so that numbers and color coding show the number of data points with identical accuracy and proportion of reported knowledge."
-
 bin_interval <- seq(0, 1, 0.05)
@@
 
-```{r fig-epreds-3}
+```{r}
 #| label: fig-epreds-3
@@
 
-# References
+# References {.unnumbered}
+
+::: {#refs}
+:::
+
+## Appendix 1: Model diagnostics {#apx-diagnostics}
+
+One way to diagnose the behaviour of Hamiltonian Monte Carlot (HMC, the algorithm used by Stan to explore the posterior distribution of a model) is to check whether the MCMC chains have converged.
+@fig-appendix-diagnostics shows the values sampled by the MCMC chains of each of the fixed coefficients of each model reported in the manuscript.
+Evidence of chain convergence is provided by the same region of values being sampled across the final interations of the chain, as it is the case for the three models depicted.
+
+```{r}
+#| label: apxfig-appendix-diagnostics
+#| fig-cap: "Traceplots of fixed regression coefficients of models in Experiments 1, 2, and 3."
+#| fig-width: 9
+#| fig-height: 9
+var_labels <- c(
+  "b_Intercept" = "Intercept",
+  "b_freq_zipf_2_std" = "Frequency",
+  "b_lv_std" = "Cognateness",
+  "b_neigh_n_h_std" = "CLPN",
+  "b_neigh_n_h_std:lv_std" = "Cognateness × CLPN",
+  "b_group1" = "cat-ENG vs. spa-ENG"
+)
+
+list(exp_1_m0, exp_2_m0, exp_3_m1) |>
+  set_names(paste0("Experiment ", 1:3)) |>
+  map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |>
+  dplyr::filter(.variable != "b_Intercept") |>
+  mutate(
+    .value = .value / 4,
+    .variable = factor(.variable,
+      levels = names(var_labels),
+      labels = var_labels
+    )
+  ) |>
+  ggplot(aes(.iteration, .value, colour = factor(.chain))) +
+  facet_grid(.variable ~ exp, scales = "free") +
+  geom_line() +
+  labs(
+    x = "Iteration",
+    y = "Value",
+    colour = "Chain",
+    fill = "Chain",
+    linetype = "Chain"
+  ) +
+  scale_colour_brewer(palette = "Reds") +
+  theme(
+    legend.position = "top",
+    legend.box = "horizontal",
+    panel.border = element_rect(
+      fill = NA,
+      colour = "grey"
+    ),
+    legend.justification = c(1, 0),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+```
+
+## Appendix 2: Pooled analyses of Experiments 1 and 3 {#apx-pooled}
+
+Across Experiments 1 and 3, we found strong evidence that participants efficiently exploited phonological similarity to provide accurate translations for words in an unfamiliar language, provided that few phonological neighbours of higher lexical frequency were present.
+@fig-coefs summarizes the posterior distribution of the regression coefficients of the models in Experiments 1 to 3.
+
+```{r}
+#| label: apxfig-coefs
+#| fig-cap: "Regression coefficients across Experiments 1, 2, and 3."
+#| fig-width: 9
+#| fig-height: 5
+var_labels <- c(
+  "b_Intercept" = "Intercept",
+  "b_freq_zipf_2_std" = "Frequency",
+  "b_lv_std" = "Similarity",
+  "b_neigh_n_h_std" = "CLPN",
+  "b_neigh_n_h_std:lv_std" = "Similarity × CLPN",
+  "b_group1" = "cat-ENG vs. spa-ENG"
+)
+
+list(exp_1_m0, exp_2_m0, exp_3_m1) |>
+  set_names(paste0("Experiment ", 1:3)) |>
+  map_dfr(gather_draws, `b_.*`, regex = TRUE, .id = "exp") |>
+  dplyr::filter(.variable != "b_Intercept") |>
+  mutate(
+    .value = .value / 4,
+    .variable = factor(.variable,
+      levels = names(var_labels),
+      labels = var_labels
+    )
+  ) |>
+  ggplot(aes(.value, reorder(exp, desc(exp)))) +
+  facet_wrap(~.variable, scales = "free_x") +
+  annotate(
+    geom = "rect",
+    xmin = -0.1 / 4,
+    xmax = 0.1 / 4,
+    ymin = -Inf,
+    ymax = Inf,
+    fill = "grey",
+    alpha = 3 / 4,
+    colour = NA
+  ) +
+  geom_vline(
+    xintercept = 0,
+    colour = "grey",
+    linewidth = 3 / 4
+  ) +
+  stat_slab(aes(fill = after_stat(level)),
+    point_interval = mode_hdi,
+    .width = c(0.95, 0.89, 0.67, 1),
+    outline_bars = TRUE,
+    colour = "white",
+    linewidth = 0.1,
+    breaks = 20,
+    normalize = "panels"
+  ) +
+  stat_spike(
+    at = "median",
+    size = 1,
+    aes(linetype = "Median"),
+    normalize = "panels"
+  ) +
+  labs(
+    x = "Estimate", y = "Experiment",
+    colour = "CrI",
+    fill = "CrI",
+    linetype = ""
+  ) +
+  scale_fill_brewer(palette = "Reds") +
+  theme(
+    legend.position = c(1, 0),
+    legend.box = "horizontal",
+    legend.justification = c(1, 0),
+    axis.title.x = element_blank(),
+    axis.title.y = element_blank()
+  )
+```

### 2025-07-24 docs/manuscript.tex
Update

--- 2025-02-07 manuscript/manuscript.tex
@@
-% Options for packages loaded elsewhere
-\PassOptionsToPackage{unicode}{hyperref}
-\PassOptionsToPackage{hyphens}{url}
-\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
-%
 \documentclass[
-]{article}
-
-\usepackage{amsmath,amssymb}
-\usepackage{iftex}
-\ifPDFTeX
-  \usepackage[T1]{fontenc}
-  \usepackage[utf8]{inputenc}
-  \usepackage{textcomp} % provide euro and other symbols
-\else % if luatex or xetex
-  \usepackage{unicode-math}
-  \defaultfontfeatures{Scale=MatchLowercase}
-  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
-\fi
-\usepackage{lmodern}
-\ifPDFTeX\else  
-    % xetex/luatex font selection
-    \setmainfont[]{Latin Modern Roman}
-  \setmathfont[]{Latin Modern Math}
-\fi
-% Use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\IfFileExists{microtype.sty}{% use microtype if available
-  \usepackage[]{microtype}
-  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
-}{}
+  man,
+  longtable,
+  nolmodern,
+  notxfonts,
+  notimes,
+  colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue]{apa7}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
+
+\usepackage[bidi=default]{babel}
+\babelprovide[main,import]{english}
+
+% get rid of language-specific shorthands (see #6817):
+\let\LanguageShortHands\languageshorthands
+\def\languageshorthands#1{}
+
+\RequirePackage{longtable}
+\RequirePackage{threeparttablex}
+
 \makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
+\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
+	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+	{-.5em}%
+	{\normalfont\normalsize\bfseries\typesectitle}}
+
+\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{0.5em}%
+	{0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
+	{-\z@\relax}%
+	{\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
 \makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
-% Make \paragraph and \subparagraph free-standing
-\makeatletter
-\ifx\paragraph\undefined\else
-  \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
-\fi
-\ifx\subparagraph\undefined\else
-  \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
-\fi
-\makeatother
+
+\usepackage{longtable, booktabs, multirow, multicol, colortbl, hhline, caption, array, float, xpatch}
+\usepackage{subcaption}
+
+\renewcommand\thesubfigure{\Alph{subfigure}}
+\setcounter{topnumber}{2}
+\setcounter{bottomnumber}{2}
+\setcounter{totalnumber}{4}
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\bottomfraction}{0.85}
+\renewcommand{\textfraction}{0.15}
+\renewcommand{\floatpagefraction}{0.7}
+
+\usepackage{tcolorbox}
+\tcbuselibrary{listings,theorems, breakable, skins}
+\usepackage{fontawesome5}
+
+\definecolor{quarto-callout-color}{HTML}{909090}
+\definecolor{quarto-callout-note-color}{HTML}{0758E5}
+\definecolor{quarto-callout-important-color}{HTML}{CC1914}
+\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
+\definecolor{quarto-callout-tip-color}{HTML}{00A047}
+\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
+\definecolor{quarto-callout-color-frame}{HTML}{ACACAC}
+\definecolor{quarto-callout-note-color-frame}{HTML}{4582EC}
+\definecolor{quarto-callout-important-color-frame}{HTML}{D9534F}
+\definecolor{quarto-callout-warning-color-frame}{HTML}{F0AD4E}
+\definecolor{quarto-callout-tip-color-frame}{HTML}{02B875}
+\definecolor{quarto-callout-caution-color-frame}{HTML}{FD7E14}
+
+%\newlength\Oldarrayrulewidth
+%\newlength\Oldtabcolsep
+
+\usepackage{hyperref}
 
 \providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \makesavenoteenv{longtable}
+
 \usepackage{graphicx}
@@
 \makeatother
+
 % definitions for citeproc citations
@@
 
+\usepackage{newtx}
+
+\defaultfontfeatures{Scale=MatchLowercase}
+\defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+
+\title{The role of cognateness in native spoken word recognition}
+
+\shorttitle{The role of cognateness in native spoken word recognition}
+
+\usepackage{etoolbox}
+
+\authorsnames[{1},{2},{2},{1}]{Gonzalo Garcia-Castro,Serene Siow,Kim
+Plunkett,Nuria Sebastian-Galles}
+
+\authorsaffiliations{
+{Center for Brain and Cognition, Universitat Pompeu Fabra},{Department
+of Experimental Psychology, University of Oxford}}
+
+\leftheader{Garcia-Castro, Siow, Plunkett and Sebastian-Galles}
+
+\abstract{When listening to speech in an unfamiliar language, words and
+phrases in the native language are frequently activated due to their
+acoustic similarity with some parts of the speech stream. This study
+explored this phenomenon, for the first time, from a language processing
+perspective. Across three studies, English- and Spanish-native adults
+completed a translation elicitation task, in which they listened to a
+series of words from an unfamiliar language (Catalan or Spanish for
+English speakers, Catalan for Spanish speakers). For each presented
+word, participants had to type their best-guess translation in their
+native language. Both English and Spanish natives were surprisingly good
+at translating unfamiliar words, efficiently exploiting the phonological
+similarity between the presented (unfamiliar) words and their correct
+translation. When the correct translation belonged to high-density
+phonological neighbourhood, participants' ability to benefit from
+phonological similarity decreased. Spanish participants, whose native
+language was typologically closer to the presented language, benefited
+more strongly from phonological similarity than English participants.
+Overall, we show that speech in an unfamiliar language triggers
+equivalent dynamics of lexical selection than native speech, providing a
+psycholinguistic account for homophonic translation. }
+
+\keywords{cognateness, spoken word recognition, phonology, speech
+processing, non-native speech, lexical access}
+
+\authornote{\par{\addORCIDlink{Gonzalo
+Garcia-Castro}{0000-0002-8553-4209}}\par{\addORCIDlink{Serene
+Siow}{0000-0001-6482-2191}}\par{\addORCIDlink{Kim
+Plunkett}{0000-0003-0216-7480}}\par{\addORCIDlink{Nuria
+Sebastian-Galles}{0000-0001-6938-2498}}
+
+\par{       }
+\par{Correspondence concerning this article should be addressed
+to Gonzalo
+Garcia-Castro, Email: \href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}}
+}
+
+\makeatletter
+\let\endoldlt\endlongtable
+\def\endlongtable{
+\hline
+\endoldlt
+}
+\makeatother
+\RequirePackage{longtable}
+\DeclareDelayedFloatFlavor{longtable}{table}
+
+\urlstyle{same}
+
 \usepackage{booktabs}
+
 \usepackage{caption}
+
 \usepackage{longtable}
+
 \usepackage{colortbl}
+
 \usepackage{array}
+
 \usepackage{float}
+
 \usepackage{tabularray}
+
 \usepackage[normalem]{ulem}
+
 \usepackage{graphicx}
+
 \UseTblrLibrary{booktabs}
+
 \UseTblrLibrary{rotating}
+
 \UseTblrLibrary{siunitx}
+
 \NewTableCommand{\tinytableDefineColor}[3]{\definecolor{#1}{#2}{#3}}
+
 \newcommand{\tinytableTabularrayUnderline}[1]{\underline{#1}}
+
 \newcommand{\tinytableTabularrayStrikeout}[1]{\sout{#1}}
-\usepackage{tipa}
-\usepackage{booktabs}
-\usepackage{arxiv}
-\usepackage{orcidlink}
-\usepackage{amsmath}
-\usepackage[T1]{fontenc}
+
+\makeatletter
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{plain}
+\@ifundefined{c@chapter}{\newfloat{apxfig}{h}{loapxfig}}{\newfloat{apxfig}{h}{loapxfig}[chapter]}
+\floatname{apxfig}{Appendix Fig.}
+\newcommand*\listofapxfigs{\listof{apxfig}{List of Appendix Fig.s}}
+\makeatother
 \makeatletter
@@
 
-\ifLuaTeX
-\usepackage[bidi=basic]{babel}
-\else
-\usepackage[bidi=default]{babel}
-\fi
-\babelprovide[main,import]{english}
-\ifPDFTeX
-\else
-\babelfont{rm}[]{Latin Modern Roman}
-\fi
-% get rid of language-specific shorthands (see #6817):
-\let\LanguageShortHands\languageshorthands
-\def\languageshorthands#1{}
-\ifLuaTeX
-  \usepackage[english]{selnolig} % disable illegal ligatures
-\fi
-\usepackage{bookmark}
-
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
-\hypersetup{
-  pdftitle={The role of cognateness in native spoken word recognition},
-  pdfauthor={Gonzalo Garcia-Castro; Serene Siow; Kim Plunkett; Nuria Sebastian-Galles},
-  pdflang={en},
-  pdfkeywords={cognate, spoken word
-recognition, phonology, bilingualism, non-native speech
-processing, lexical access},
-  colorlinks=true,
-  linkcolor={blue},
-  filecolor={Maroon},
-  citecolor={Blue},
-  urlcolor={Blue},
-  pdfcreator={LaTeX via pandoc}}
-
-\usepackage{lineno}
-\linenumbers
-\newcommand{\runninghead}{A Preprint }
-\renewcommand{\runninghead}{Cognateness and recognition of foreign
-words }
-\title{The role of cognateness in native spoken word
-recognition\thanks{The authors declare no conflicts of interest with
-regard to the funding source of this study. This study was supported by
-the Spanish Ministry for Science and Innovation and State Research
-Agency (Project PGC2018-101831-B-I00 financed by Spanish Ministry of
-Science Innovation and Universities (MCIU), Spanish Research Agency
-(AEI) and the European Regional Development Fund (ERDF), and by the
-Economic and Social Research Council (ESRC) (ES/S010947/1, UK). GGC was
-supported by a FPI research contract (PRE2019-088165). SS was supported
-by XXXXXXXXXXXXX. KP was supported by XXXXXXXXXXXXXXXXX. NSG was
-supported by an ICREA Academia 2019 award from the Catalan Institution
-for Research and Advanced Studies (ICREA). We are grateful to Irina
-Lepadatu and Nicola Gillen, and to the Speech Acquisition and Perception
-research group for their helpful feedback. We thank Xavier Mayoral,
-Silvia Blanch, Pamela Miller, and Cristina Cuadrado for their technical
-support. We also thank all families and infants who participated in the
-experiments. This study was conducted according to guidelines laid down
-in the Declaration of Helsinki, and was approved by the Drug Research
-Ethical Committee (CEIm) of the IMIM Parc de Salut Mar, reference
-2020/9080/I. The data and code necessary to reproduce the analyses
-presented here are publicly accessible, as are the materials necessary
-to attempt to replicate the findings. The analyses presented here were
-not preregistered. Data, code, and materials for this research are
-available at the following URLs: \url{https://osf.io/hy984/},
-\url{https://github.com/gongcastro/translation-elicitation}.}}
-\def\asep{\\\\\\ } % default: all authors on same column
-\def\asep{\And }
-\author{\textbf{Gonzalo
-Garcia-Castro}~\orcidlink{0000-0002-8553-4209}\\Center for Brain and
-Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\href{mailto:gonzalo.garciadecastro@upf.edu}{gonzalo.garciadecastro@upf.edu}\asep\textbf{Serene
-Siow}~\orcidlink{0000-0001-6482-2191}\\Department of Experimental
-Psychology\\University of Oxford\\Oxford, OX2
-6GG\\\href{mailto:siow.serene@gmail.com}{siow.serene@gmail.com}\asep\textbf{Kim
-Plunkett}~\orcidlink{0000-0003-0216-7480}\\Department of Experimental
-Psychology\\University of Oxford\\Oxford, OX2
-6GG\\\href{mailto:kim.plunkett@psy.ox.ac.uk}{kim.plunkett@psy.ox.ac.uk}\asep\textbf{Nuria
-Sebastian-Galles}~\orcidlink{0000-0001-6938-2498}\\Center for Brain and
-Cognition\\Universitat Pompeu Fabra\\Barcelona,
-08005\\\href{mailto:nuria.sebastian@upf.edu}{nuria.sebastian@upf.edu}}
-\date{}
+% From https://tex.stackexchange.com/a/645996/211326
+%%% apa7 doesn't want to add appendix section titles in the toc
+%%% let's make it do it
+\makeatletter
+\xpatchcmd{\appendix}
+  {\par}
+  {\addcontentsline{toc}{section}{\@currentlabelname}\par}
+  {}{}
+\makeatother
+
+%% Disable longtable counter
+%% https://tex.stackexchange.com/a/248395/211326
+
+\usepackage{etoolbox}
+
+\makeatletter
+\patchcmd{\LT@caption}
+  {\bgroup}
+  {\bgroup\global\LTpatch@captiontrue}
+  {}{}
+\patchcmd{\longtable}
+  {\par}
+  {\par\global\LTpatch@captionfalse}
+  {}{}
+\apptocmd{\endlongtable}
+  {\ifLTpatch@caption\else\addtocounter{table}{-1}\fi}
+  {}{}
+\newif\ifLTpatch@caption
+\makeatother
+
 \begin{document}
+
 \maketitle
-\begin{abstract}
-While listening to speech in an unfamiliar language, words and phrases
-in the native language are activated, leading to common auditory
-illusions like Soramimi, a particular case of homophonic translation.
-Homophonic translation has received little attention as a
-psycholinguistic phenomenon. In this study, we explored the mechanisms
-underlying lexical activation and selection in homoponic translation.
-Across three studies, English and Spanish native adults completed a
-translation elicitation task, in which they listened to a series of
-Catalan or Spanish words (to which participants were unfamiliar), and
-then had to type their best-guess translation in their native language.
-Both English and Spanish natives were surprisingly good at translating
-unfamiliar words, efficiently exploiting the phonological similarity
-between the presented (unfamiliar) words and their correct translation.
-When the correct translation belonged to high-density phonological
-neighbours, participants' ability to benefit from phonological
-similarity decreased. Finally, Spanish natives exploited phonological
-similarity more efficiently: provided identical phonological similarity,
-Spanish natives were more likely to guess the correct translation of the
-unfamiliar word than English natives. Overall, we show that speech in an
-unfamiliar language triggers equivalent dynamics of lexical selection
-than native speech, providing a psycholinguistic account for homophonic
-translation.
-\end{abstract}
-{\bfseries \emph Keywords}
-\def\sep{\textbullet\ }
-cognate \sep spoken word
-recognition \sep phonology \sep bilingualism \sep non-native speech
-processing \sep 
-lexical access
+
+\setlength\LTleft{0pt}
 
@@
 lexical access, obstructing the selection of the appropriate target
-translation in the listener's lexicon. For instance, Otwinowska \&
+translation in the listener's lexicon. For instance, Otwinowska and
 Szewczyk (\citeproc{ref-otwinowska2019more}{2019}) reported that false
@@
 \url{https://github.com/gongcastro/translation-elicitation.git}, along
-with additional notes.
+with additional notes. For reproducibility, a Docker image of the
+RStudio session is available on DockerHub
+(\url{https://hub.docker.com/repository/docker/gongcastro/translation-elicitation/general}).
 
@@
 
+\begin{table}
+
+{\caption{{Participant details.}{\label{tbl-participants}}}
+\vspace{-20pt}}
+
 \setlength{\LTpost}{0mm}
-
-\begin{longtable}{l|rrrc}
-
-\caption{\label{tbl-participants}Participant details.}
-
-\tabularnewline
-
+\begin{longtable*}{l|rrrc}
 \toprule
@@
 \bottomrule
-
-\end{longtable}
-
+\end{longtable*}
 \begin{minipage}{\linewidth}
@@
 \end{minipage}
+
+\end{table}
 
@@
   \url{http://espeak.sourceforge.net/}}{]}. Table~\ref{tbl-stimuli}
-summarises the lexical frequency, phonological neighbourhood density and
-phonological overlap of the words included in the Catalan and the
-Spanish lists.
-
-\begin{longtable}{l|rrrrrr}
-
-\caption{\label{tbl-stimuli}Stimuli details.}
-
-\tabularnewline
-
+summarises the lexical frequency, \emph{CLPN} and phonological overlap
+of the words included in the Catalan and the Spanish lists.
+
+\begin{table}
+
+{\caption{{Stimuli details.}{\label{tbl-stimuli}}}
+\vspace{-20pt}}
+
+\begin{longtable*}{l|rrrrrr}
 \toprule
@@
 \bottomrule
-
-\end{longtable}
+\end{longtable*}
+
+\end{table}
 
@@
 
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{_assets/img/design.png}}
-
-}
-
 \caption{\label{fig-design}Schematic representation of a trial in the
@@
 
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{_assets/img/design.png}}
+
+}
+
 \end{figure}%
@@
 
-\paragraph{Data processing}\label{data-processing}
+\paragraph{Data processing.}\label{data-processing}
 
@@
 \paragraph{Modelling approach and statistical
-inference}\label{modelling-approach-and-statistical-inference}
+inference.}\label{modelling-approach-and-statistical-inference}
 
@@
 
-We followed Kruschke \& Liddell
+We followed Kruschke and Liddell
 (\citeproc{ref-kruschke2018bayesian}{2018})'s guidelines for statistical
@@
 (\citeproc{ref-carpenter2017stan}{Carpenter et al., 2017}) to estimate
-and compare the models (see Appendix 1 for more details on the models).
+and compare the models (see \#apx-diagnostics for model diagnostics).
 Numeric predictors were standardised before entering the model by
@@
 
-\begin{longtable}{l|rrrrrrrrr}
-
-\caption{\label{tbl-dataset}Summary of participants' accuracy in the
-translation elicitation task across Experiments 1, 2, and 3.}
-
-\tabularnewline
-
+\begin{table}
+
+{\caption{{Summary of participants' accuracy in the translation
+elicitation task across Experiments 1, 2, and 3.}{\label{tbl-dataset}}}
+\vspace{-20pt}}
+
+\begin{longtable*}{l|rrrrrrrrr}
 \toprule
@@
 \bottomrule
-
-\end{longtable}
+\end{longtable*}
+
+\end{table}
 
 \begin{figure}
-
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}}
-
-}
 
@@
 intervals (CrI).}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-1-1.pdf}}
+
+}
 
@@
 
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}}
-
-}
-
 \caption{\label{fig-epreds-2}Posterior model-predicted mean accuracy in
@@
 (CrI).}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-2-1.pdf}}
+
+}
 
@@
 phonological similarity on participants' performance in Experiment 2
-larger. Spanish natives in Experiment 2 provided exploited phonological
+larger. Spanish natives in Experiment 2 exploited phonological
 similarity more strongly than English natives in Experiment 1. Overall,
@@
 (Catalan), benefited more strongly from phonological similarity than
-participants in Experiment 1, who were natives of typologically less
+participants in Experiment 1, who were natives of a typologically less
 similar language (English) to the presented language (Catalan, Spanish).
@@
 
-\caption{\label{tbl-surprises}List of items with unexpectedly high
-accuracy: the Levenshtein similarity score between the presented word
-(in Catalan or Spanish) and their correct English translation is zero,
-but participants, who are reportedly unfamiliar with the presented
-language, were on average \textgreater10\% likely to guess the correct
-translation.}
-
-\centering{
+{\caption{{List of items with unexpectedly high accuracy: the
+Levenshtein similarity score between the presented word (in Catalan or
+Spanish) and their correct English translation is zero, but
+participants, who are reportedly unfamiliar with the presented language,
+were on average \textgreater10\% likely to guess the correct
+translation.}{\label{tbl-surprises}}}
+\vspace{-20pt}}
 
@@
 
-}
-
-\end{table}%
+\end{table}
 
@@
 in accordance with ethical standards of the Declaration of Helsinki and
-the protocol was by the University of Oxford Medical Sciences
+the protocol was approved by the University of Oxford Medical Sciences
 Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
@@
 about form-meaning mappings to correctly translate some Spanish words.
-To isolate such an effect of prior Spanish knowledge, we run the same
+To isolate such an effect of prior Spanish knowledge, we ran the same
 analysis as in Experiment 1 on the newly collected translations from
@@
 
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-knowledge-1.pdf}}
-
-}
-
-\caption{\label{fig-knowledge}Catalan and Spanish prior word knowledge,
-as reported by English native participants in Experiment 3. The X-axis
+\caption{\label{fig-knowledge}Catalan and Spanish prior word knowledge
+reported by English native participants in Experiment 3. The X-axis
 indicates the proportion of participants that reported prior knowledge
 with the word. The Y-axis indicates participants' accuracy (the
-proportion of correct responses by participants). Each colour indicates
-whether the data point corresponds to a word that participants answered
-to with expected accuracy (red) or with unexpectedly high accuracy
-(blue), as defined by (no phonological similarity with the correct
-translation, by accuracy higher than 10\%).}
+proportion of participants that translated the word correctly). For
+visualisation purposes, data points have been aggregated so that numbers
+and color coding show the number of data points with identical accuracy
+and proportion of reported knowledge.}
+
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-knowledge-1.pdf}}
+
+}
 
@@
 
-\centering{
-
-\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}}
-
-}
-
 \caption{\label{fig-epreds-3}Posterior model-predicted mean accuracy in
@@
 
+\centering{
+
+\pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/fig-epreds-3-1.pdf}}
+
+}
+
 \end{figure}%
@@
 information in a numeric scale from 0 to 7. To control for the effect of
-participants' prior knowledge abbout form-meaning mappings, we excluded
+participants' prior knowledge about form-meaning mappings, we excluded
 from the analyses those words in which participants reported such prior
@@
 
-\section{References}\label{references}
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 
@@
 
-\bibitem[\citeproctext]{ref-briusov1911ot}
-Briusov, V. Y. (1911). \emph{Ot perevodchika}. Skorpion.
-
-\bibitem[\citeproctext]{ref-broersma2021praat}
-Broersma, P., \& Weenink, D. (2021). \emph{Praat: Doing phonetics by
-computer {[}{Computer} program{]}} (Version 6.1.54).
-\url{http://www.praat.org/}
-
-\bibitem[\citeproctext]{ref-burkner2017brms}
-Bürkner, P.-C. (2017). Brms: {An R} package for {Bayesian} multilevel
-models using {Stan}. \emph{Journal of Statistical Software},
-\emph{80}(1), 1--28.
-
-\bibitem[\citeproctext]{ref-carpenter2017stan}
-Carpenter, B., Gelman, A., Hoffman, M. D., Lee, D., Goodrich, B.,
-Betancourt, M., Brubaker, M., Guo, J., Li, P., \& Riddell, A. (2017).
-Stan: {A} probabilistic programming language. \emph{Journal of
-Statistical Software}, \emph{76}(1), 1--32.
-
-\bibitem[\citeproctext]{ref-cervantes}
-Cervantes, M. de, \& Grossman, E. (2002). First part of the ingenious
-nobleman don quixote of la mancha. \emph{Conjunctions}, \emph{38},
-207--225. \url{http://www.jstor.org/stable/24518146}
-
-\bibitem[\citeproctext]{ref-dembeck2015oberflachenubersetzung}
-Dembeck, T. (2015). Oberfl{ä}chen{ü}bersetzung: The poetics and cultural
-politics of homophonic translation. \emph{Critical Multilingualism
-Studies}, \emph{3}(1), 7--25.
-
-\bibitem[\citeproctext]{ref-dufour2010phonological}
-Dufour, S., \& Frauenfelder, U. H. (2010). Phonological neighbourhood
-effects in french spoken-word recognition. \emph{Quarterly Journal of
-Experimental Psychology}, \emph{63}(2), 226--238.
-
-\bibitem[\citeproctext]{ref-dufour2003lexical}
-Dufour, S., \& Peereman, R. (2003). Lexical competition in phonological
-priming: {Assessing} the role of phonological match and mismatch lengths
-between primes and targets. \emph{Memory \& Cognition}, \emph{31},
-1271--1283.
-
-\bibitem[\citeproctext]{ref-dupoux1999epenthetic}
-Dupoux, E., Kakehi, K., Hirose, Y., Pallier, C., \& Mehler, J. (1999).
-Epenthetic vowels in japanese: A perceptual illusion? \emph{Journal of
-Experimental Psychology: Human Perception and Performance},
-\emph{25}(6), 1568.
-
-\bibitem[\citeproctext]{ref-efimova2018homophonic}
-Efimova, N. N., Ruzhnikova, M. L., \& Violina, M. I. (2018). Homophonic
-translation as humpty-dumpty's choice. \emph{SHS Web of Conferences},
-\emph{50}, 01049.
-
-\bibitem[\citeproctext]{ref-floccia2018introduction}
-Floccia, C., Sambrook, T. D., Delle Luche, C., Kwok, R., Goslin, J.,
-White, L., Cattani, A., Sullivan, E., Abbot-Smith, K., Krott, A., et al.
-(2018). I: {Introduction}. \emph{Monographs of the Society for Research
-in Child Development}, \emph{83}(1), 7--29.
-
-\bibitem[\citeproctext]{ref-gasparov2006semen}
-Gasparov, M. L. (2006). \emph{Semen kirsanov}. Akademicheskij proekt.
-
-\bibitem[\citeproctext]{ref-goldinger1989priming}
-Goldinger, S. D., Luce, P. A., \& Pisoni, D. B. (1989). Priming lexical
-neighbors of spoken words: {Effects} of competition and inhibition.
-\emph{Journal of Memory and Language}, \emph{28}(5), 501--518.
-\url{https://doi.org/10.1016/0749-596X(89)90009-0}
-
-\bibitem[\citeproctext]{ref-grainger1990word}
-Grainger, J. (1990). Word frequency and neighborhood frequency effects
-in lexical decision and naming. \emph{Journal of Memory and Language},
-\emph{29}(2), 228--244.
-
-\bibitem[\citeproctext]{ref-hamburger1996phonological}
-Hamburger, M., \& Slowiaczek, L. M. (1996). Phonological priming
-reflects lexical competition. \emph{Psychonomic Bulletin \& Review},
-\emph{3}(4), 520--525.
-
-\bibitem[\citeproctext]{ref-kruschke2018bayesian}
-Kruschke, J. K., \& Liddell, T. M. (2018). The {Bayesian New
-Statistics}: {Hypothesis} testing, estimation, meta-analysis, and power
-analysis from a {Bayesian} perspective. \emph{Psychonomic Bulletin \&
-Review}, \emph{25}, 178--206.
-
-\bibitem[\citeproctext]{ref-levenshtein1966binary}
-Levenshtein, V. I. (1966). Binary codes capable of correcting deletions,
-insertions, and reversals. \emph{Soviet Physics Doklady}, \emph{10},
-707--710.
-
-\bibitem[\citeproctext]{ref-levick2016translating}
-Levick, T. (2016). Translating homophonic wordplay in patrick goujon's
-moi non: A case study. \emph{Sound/Writing: On Homophonic Translation}.
-
-\bibitem[\citeproctext]{ref-luce1998recognizing}
-Luce, P. A., \& Pisoni, D. B. (1998). Recognizing spoken words: {The}
-neighborhood activation model. \emph{Ear and Hearing}, \emph{19}(1), 1.
-
-\bibitem[\citeproctext]{ref-luce1990similarity}
-Luce, P. A., Pisoni, D. B., \& Goldinger, S. D. (1990). \emph{Similarity
-neighborhoods of spoken words.}
-
-\bibitem[\citeproctext]{ref-marian2012clearpond}
-Marian, V., Bartolotti, J., Chabal, S., \& Shook, A. (2012).
-\emph{{CLEARPOND}: {Cross-linguistic} easy-access resource for
-phonological and orthographic neighborhood densities}.
-
-\bibitem[\citeproctext]{ref-otake2007interlingual}
-Otake, T. (2007). Interlingual near homophonic words and phrases in L2
-listening: Evidence from misheard song lyrics. \emph{Proceedings of the
-16th International Congress of Phonetic Sciences (ICPhS 2007)},
-777--780.
-
-\bibitem[\citeproctext]{ref-otwinowska2019more}
-Otwinowska, A., \& Szewczyk, J. M. (2019). The more similar the better?
-Factors in learning cognates, false cognates and non-cognate words.
-\emph{International Journal of Bilingual Education and Bilingualism}.
-
-\bibitem[\citeproctext]{ref-peirce2019psychopy2}
-Peirce, J., Gray, J. R., Simpson, S., MacAskill, M., Höchenberger, R.,
-Sogo, H., Kastman, E., \& Lindeløv, J. K. (2019). {PsychoPy2}:
-{Experiments} in behavior made easy. \emph{Behavior Research Methods},
-\emph{51}(1), 195--203. \url{https://doi.org/10.3758/s13428-018-01193-y}
-
-\bibitem[\citeproctext]{ref-peperkamp2008perceptual}
-Peperkamp, S., Vendelin, I., \& Nakamura, K. (2008). On the perceptual
-origin of loanword adaptations: Experimental evidence from japanese.
-\emph{Phonology}, \emph{25}(1), 129--164.
-
-\bibitem[\citeproctext]{ref-pilshchikov2016semiotics}
-Pilshchikov, I. (2016). The semiotics of phonetic translation.
-\emph{Studia Metrica Et Poetica}, \emph{3}(1), 53--104.
-
-\bibitem[\citeproctext]{ref-rcoreteam2013language}
-R Core Team. (2013). \emph{R: {A Language} and {Environment} for
-{Statistical Computing}}. R Foundation for Statistical Computing.
-\url{http://www.R-project.org/}
-
-\bibitem[\citeproctext]{ref-schad2020capitalize}
-Schad, D. J., Vasishth, S., Hohenstein, S., \& Kliegl, R. (2020). How to
-capitalize on a priori contrasts in linear (mixed) models: {A} tutorial.
-\emph{Journal of Memory and Language}, \emph{110}, 104038.
-
-\bibitem[\citeproctext]{ref-schepens2012distributions}
-Schepens, J., Dijkstra, T., \& Grootjen, F. (2012). Distributions of
-cognates in {Europe} as based on {Levenshtein} distance.
-\emph{Bilingualism: Language and Cognition}, \emph{15}(1), 157--166.
-
-\bibitem[\citeproctext]{ref-van2014stringdist}
-van der Loo, M. P. J. (2014). The stringdist package for approximate
-string matching. \emph{The R Journal}, \emph{6}(1), 111--122.
-\url{https://CRAN.R-project.org/package=stringdist}
-
-\bibitem[\citeproctext]{ref-van2014subtlex}
-Van Heuven, W. J., Mandera, P., Keuleers, E., \& Brysbaert, M. (2014).
-{SUBTLEX-UK}: {A} new and improved word frequency database for {British
-English}. \emph{Quarterly Journal of Experimental Psychology},
-\emph{67}(6), 1176--1190.
-
-\bibitem[\citeproctext]{ref-vitevitch2006clustering}
-Vitevitch, M. (2006). The clustering coefficient of phonological
-neighborhoods influences spoken word recognition. \emph{The Journal of
-the Acoustical Society of America}, \emph{120}(5), 3252--3252.
-\url{https://doi.org/10.1121/1.4788314}
-
-\bibitem[\citeproctext]{ref-vitevitch1998words}
-Vitevitch, M. S., \& Luce, P. A. (1998). When words compete: Levels of
-processing in perception of spoken words. \emph{Psychological Science},
-\emph{9}(4), 325--329.
-
-\bibitem[\citeproctext]{ref-vitevitch1999probabilistic}
-Vitevitch, M. S., \& Luce, P. A. (1999). Probabilistic {Phonotactics}
-and {Neighborhood Activation} in {Spoken Word Recognition}.
-\emph{Journal of Memory and Language}, \emph{40}(3), 374--408.
-\url{https://doi.org/10.1006/jmla.1998.2618}
-
-\bibitem[\citeproctext]{ref-weber2004lexical}
-Weber, A., \& Cutler, A. (2004). Lexical competition in non-native
-spoken-word recognition. \emph{Journal of Memory and Language},
-\emph{50}(1), 1--25.
-
-\bibitem[\citeproctext]{ref-wickham2019welcome}
-Wickham, H., Averick, M., Bryan, J., Chang, W., McGowan, L. D.,
-François, R., Grolemund, G., Hayes, A., Henry, L., Hester, J., Kuhn, M.,
-Pedersen, T. L., Miller, E., Bache, S. M., Müller, K., Ooms, J.,
-Robinson, D., Seidel, D. P., Spinu, V., \ldots{} Yutani, H. (2019).
-Welcome to the tidyverse. \emph{Journal of Open Source Software},
-\emph{4}(43), 1686. \url{https://doi.org/10.21105/joss.01686}
-
 \end{CSLReferences}
 
+\appendix
+
+\section{}\label{}
+
+\subsection{Appendix 1: Model diagnostics}\label{apx-diagnostics}
+
+One way to diagnose the behaviour of Hamiltonian Monte Carlot (HMC, the
+algorithm used by Stan to explore the posterior distribution of a model)
+is to check whether the MCMC chains have converged.
+\textbf{?@fig-appendix-diagnostics} shows the values sampled by the MCMC
+chains of each of the fixed coefficients of each model reported in the
+manuscript. Evidence of chain convergence is provided by the same region
+of values being sampled across the final interations of the chain, as it
+is the case for the three models depicted.
+
+\begin{figure}[H]
+
+\caption{Traceplots of fixed regression coefficients of models in
+Experiments 1, 2, and 3.}
+
+{\centering \pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/apxfig-appendix-diagnostics-1.pdf}}
+
+}
+
+\end{figure}%
+
+\section{}\label{}
+
+\subsection{Appendix 2: Pooled analyses of Experiments 1 and
+3}\label{apx-pooled}
+
+Across Experiments 1 and 3, we found strong evidence that participants
+efficiently exploited phonological similarity to provide accurate
+translations for words in an unfamiliar language, provided that few
+phonological neighbours of higher lexical frequency were present.
+\textbf{?@fig-coefs} summarizes the posterior distribution of the
+regression coefficients of the models in Experiments 1 to 3.
+
+\begin{figure}[H]
+
+\caption{Regression coefficients across Experiments 1, 2, and 3.}
+
+{\centering \pandocbounded{\includegraphics[keepaspectratio]{manuscript_files/figure-pdf/apxfig-coefs-1.pdf}}
+
+}
+
+\end{figure}%
+
 \end{document}

### 2025-07-24 R/clearpond.R

@@ -1,42 +1,69 @@
 #' Download CLEARPOND
-#' 
-download_clearpond <- function(.language = c("english", "dutch", "french", "spanish", "german")){
-    
-    urls <- tribble(
-        ~lang, ~url, ~data, ~header,
-        "english", "https://clearpond.northwestern.edu/englishCPdatabase2.zip", "englishCPdatabase2.txt", "clearpondHeaders_EN.txt",
-        "dutch", "https://clearpond.northwestern.edu/dutchCPdatabase2.zip", "dutchCPdatabase2.txt", "clearpondHeaders_NL.txt",
-        "french", "https://clearpond.northwestern.edu/frenchCPdatabase2.zip", "frenchCPdatabase2.txt", "clearpondHeaders_FR.txt",
-        "german", "https://clearpond.northwestern.edu/germanCPdatabase2.zip", "germanCPdatabase2.txt", "clearpondHeaders_DE.txt",
-        "spanish", "https://clearpond.northwestern.edu/spanishCPdatabase2.zip", "spanishCPdatabase2.txt", "clearpondHeaders_SP.txt"
-    ) %>%
-        dplyr::filter(lang %in% .language)
-    
-    dir <- tempdir()
-    files <- replicate(tempfile(), n = length(urls$lang))
-    d <- pmap(.l = list(url = as.list(urls$url), 
-                        file = as.list(files),
-                        data = as.list(urls$data),
-                        header = as.list(urls$header)),
-              .f = function(url = .l[[1]],
-                            file = .l[[2]],
-                            data = .l[[3]], 
-                            header = .l[[4]]) {
-                  download.file(url, destfile = file)
-                  unzip(zipfile = file, exdir = dir)
-                  headers <- c("word", read.delim(paste0(dir, .Platform$file.sep, header))[,1])
-                  d <- read.delim(
-                      paste0(dir, .Platform$file.sep, data)) %>%
-                      `colnames<-`(., headers) %>%
-                      as_tibble() %>%
-                      mutate_at(vars(ends_with("W")), ~str_split(., pattern = ";")
-                      )
-                  return(d)
-              }
-    ) %>%
-        set_names(language) %>%
-        bind_rows(.id = "language") %>%
-        janitor::clean_names()
-    
-    return(d)
[…]

### 2025-07-24 R/datasets.R

@@ -1,9 +1,9 @@
 #' Get participants
 #'
-get_participants <- function(exp_participants, quest_participants) {
+get_participants <- function(exp_participants, que_participants) {
   out <- list(
     Experiment = exp_participants,
-    Questionnaire = quest_participants
+    Questionnaire = que_participants
   ) |>
     bind_rows(.id = "source")
 
@@ -12,12 +12,12 @@ get_participants <- function(exp_participants, quest_participants) {
 
 #' Get dataset for analysis
 #'
-get_dataset <- function(experiment, exp_responses, quest_responses, stimuli) {
+get_dataset <- function(experiment, exp_responses, que_responses, stimuli) {
   if (!(experiment %in% 1:3)) {
     cli_abort("{.code experiment} must be one of 1, 2, or 3")
   }
 
-  dataset <- prepare_data(exp_responses, quest_responses, stimuli)
+  dataset <- prepare_data(exp_responses, que_responses, stimuli)
 
   if (experiment == 1) {
     out <- dataset |>
@@ -41,25 +41,32 @@ get_dataset <- function(experiment, exp_responses, quest_responses, stimuli) {
 
 #' Merge data from Experiment and from Questionnaire and select relevant variables
 #'
-prepare_data <- function(exp_responses, quest_responses, stimuli) {
+prepare_data <- function(exp_responses, que_responses, stimuli) {
   stimuli_tmp <- clean_stimuli(stimuli)
 
   relevant_vars <- c(
-    "source", "participant_id", "group", "word_1",
-    "response", "correct", "response_type",
-    "confidence", "knowledge"
+    "source",
+    "participant_id",
+    "group",
+    "word_1",
+    "response",
+    "correct",
+    "response_type",
+    "confidence",
+    "knowledge"
   )
 
   out <- list(
     Experiment = exp_responses,
-    Questionnaire = quest_responses
+    Questionnaire = que_responses
   ) |>
     bind_rows(.id = "source") |>
     dplyr::filter(valid_participant, valid_response) |>
     select(any_of(relevant_vars)) |>
     inner_join(stimuli_tmp, by = join_by(group, word_1)) |>
     mutate(
-      across(c(freq_zipf_2, neigh_n, neigh_n_h, avg_sim, avg_sim_h, lv),
+      across(
[…]

### 2025-07-24 _targets.R

@@ -1,21 +1,19 @@
-suppressWarnings({
-  suppressPackageStartupMessages({
-    library(tarchetypes)
-    library(dplyr)
-    library(tidyr)
-    library(stringr)
-    library(ggplot2)
-    library(ggdist)
-    library(beeswarm)
-    library(readxl)
-    library(brms)
-    library(cmdstanr)
-    library(stringdist)
-    library(cli)
-    library(readr)
-    library(purrr)
-  })
-})
+library(tarchetypes)
+library(here)
+library(dplyr)
+library(tidyr)
+library(stringr)
+library(ggplot2)
+library(ggdist)
+library(beeswarm)
+library(readxl)
+library(brms)
+library(cmdstanr)
+library(stringdist)
+library(cli)
+library(quarto)
+library(readr)
+library(purrr)
 
 # load functions
 invisible({
@@ -31,24 +29,24 @@ options(
   brms.backend = "cmdstanr"
 )
 
+clearpond_path <- here("data", "raw", "cleapond")
+exp_data_path <- here("data", "raw", "experiment")
+que_data_path <- here("data", "raw", "questionnaire")
 
 list(
-  # CLEARPOND database
-  tar_target(cp_c_path_eng, "data-raw/clearpond/englishCPdatabase2.txt", format = "file"),
-  tar_target(cp_h_path_eng, "data-raw/clearpond/clearpondHeaders_EN.txt", format = "file"),
-  tar_target(cp_c_path_spa, "data-raw/clearpond/spanishCPdatabase2.txt", format = "file"),
-  tar_target(cp_h_path_spa, "data-raw/clearpond/clearpondHeaders_SP.txt", format = "file"),
-  tar_target(clearpond, import_clearpond(
-    cp_c_path_eng,
-    cp_c_path_spa,
-    cp_h_path_eng,
-    cp_h_path_spa
-  )),
-
-  # stimuli ------------------------------------------------------------------
-  tar_target(trial_list_path, file.path("stimuli", "trials.xlsx"),
-    format = "file"
+  # CLEARPOND database -----
+  tar_target(
+    clearpond,
+    import_clearpond(
+      here(clearpond_path, "englishCPdatabase2.txt"),
+      here(clearpond_path, "spanishCPdatabase2.txt"),
+      here(clearpond_path, , "clearpondHeaders_EN.txt"),
+      here(clearpond_path, "clearpondHeaders_SP.txt"),
+    )
   ),
+
+  # stimuli -----
[…]

### 2025-07-25 docs/appendix.qmd
Update repo

--- 2025-04-08 article/appendix.qmd
@@
 
-theme_set(theme_ggdist() +
-  theme(panel.grid.major.y = element_line(
+theme_set(
+  theme_ggdist() +
+    theme(
+      panel.grid.major.y = element_line(
         colour = "grey",
         linetype = "dotted"
-  )))
+      )
+    )
+)
 
@@
     .value = .value / 4,
-    .variable = factor(.variable,
+    .variable = factor(
+      .variable,
       levels = names(var_labels),
@@
   ) +
-  stat_slab(aes(fill = after_stat(level)),
+  stat_slab(
+    aes(fill = after_stat(level)),
     point_interval = mode_hdi,
@@
   labs(
-    x = "Estimate", y = "Experiment",
+    x = "Estimate",
+    y = "Experiment",
     colour = "CrI",

### 2025-07-25 docs/manuscript.qmd
Update repo

--- 2025-04-08 docs/manuscript.qmd
@@
 ---
-abstract: |
-  When listening to speech in an unfamiliar language, words and phrases in the native language are frequently activated due to their acoustic similarity with some parts of the speech stream.
-This study explored this phenomenon, for the first time, from a language processing perspective.
-Across three studies, English- and Spanish-native adults completed a translation elicitation task, in which they listened to a series of words from an unfamiliar language (Catalan or Spanish for English speakers, Catalan for Spanish speakers).
-For each presented word, participants had to type their best-guess translation in their native language.
-Both English and Spanish natives were surprisingly good at translating unfamiliar words, efficiently exploiting the phonological similarity between the presented (unfamiliar) words and their correct translation.
-When the correct translation belonged to high-density phonological neighbourhood, participants' ability to benefit from phonological similarity decreased.
-Spanish participants, whose native language was typologically closer to the presented language, benefited more strongly from phonological similarity than English participants.
-Overall, we show that speech in an unfamiliar language triggers equivalent dynamics of lexical selection than native speech, providing a psycholinguistic account for homophonic translation.
+abstract:  "The present study characterises homophonic translation---a phenomenon by which non-native speech activates words or phrases in a listener's native language---from a language processing perspective.
+Across three studies, English and Spanish-native monolingual adults completed a translation elicitation task in which participants listened to words in an unfamiliar language (Catalan or Spanish), and were asked to type their best-guess translation in their native language.
+Our findings suggest that that listening to speech in an unfamiliar language triggers equivalent dynamics of lexical selection than native speech.
+In particular, phonological similarity between the presented words and their correct translations interacted with phonological neighbourhood density: participants correctly translated words with high phonological similarity with their correct translation only when few other words in the native language shared phonological similarity with the presented word."
 ---
-
 ```{r}
 #| label: setup
-library(targets)
+library(here)
+library(readr)
 library(dplyr)
 library(tidyr)
+library(purrr)
+library(readxl)
 library(ggplot2)
+library(ggExtra)
 library(patchwork)
-library(gt)
-library(gtExtras)
+library(tinytable)
 library(brms)
 library(tidybayes)
-library(ggExtra)
-library(tinytable)
 library(stringdist)
+
 set.seed(1234)
 
-# import objects
-tar_config_set(
-  store = here::here("_targets"),
-  script = here::here("_targets.R")
-)
-tar_load_globals()
-tar_load(participants)
-tar_load(exp_participants)
-tar_load(exp_responses)
-tar_load(quest_responses)
-tar_load(quest_participants)
-tar_load(stimuli)
-tar_load(epreds)
-tar_load(dataset_1)
-tar_load(dataset_2)
-tar_load(dataset_3)
-tar_load(exp_1_m0)
-tar_load(exp_2_m0)
-tar_load(exp_3_m1)
-tar_load(exp_12_m0)
-
-theme_set(theme_ggdist() +
-  theme(panel.grid.major.y = element_line(
+options(readr.show_col_types = FALSE, tinytable_html_mathjax = TRUE)
+
+set_here("../")
+
+participants <- read_csv(here("out", "participants.csv"))
+exp_participants <- read_csv(here("out", "exp_participants.csv"))
+que_participants <- read_csv(here("out", "que_participants.csv"))
+stimuli <- read_csv(here("out", "stimuli.csv"))
+exp_responses <- read_csv(here("out", "exp_responses.csv"))
+que_responses <- read_csv(here("out", "que_responses.csv"))
+dataset_1 <- read_csv(here("out", "dataset-1.csv"))
+dataset_2 <- read_csv(here("out", "dataset-2.csv"))
+dataset_3 <- read_csv(here("out", "dataset-3.csv"))
+exp_1_m0 <- readRDS(here("out", "exp_1_m0.rds"))
+exp_2_m0 <- readRDS(here("out", "exp_2_m0.rds"))
+exp_3_m0 <- readRDS(here("out", "exp_3_m0.rds"))
+exp_3_m1 <- readRDS(here("out", "exp_3_m1.rds"))
+exp_12_m0 <- readRDS(here("out", "exp_12_m0.rds"))
+
+theme_set(
+  theme_ggdist() +
+    theme(
+      panel.grid.major.y = element_line(
         colour = "grey",
         linetype = "dotted"
-  )))
-
+      )
+    )
+)
 ```
@@
 When some German speakers listen to the song *The Power* by SNAP!
-(*Coyote Ugly*, 2000), many of them mishear the line "I've got the power" as *Agatha Bauer*.
-When Spanish speakers listen to the line "Circumvent your thick ego" from the song *Pictures* by System of a Down (*Steal this Album!*, 2002), they tend to hear *Sácame de aquí, King-Kong* [take me out of here, King-Kong].
-Outrageous as these examples might sound, the reader may know a few cases in their own native language.
-This auditory illusion is common across languages and cultures and can feel quite real, often inevitable [@dembeck2015oberflachenubersetzung; @efimova2018homophonic].
-In Japanese, this phenomenon takes the name *Soramimi* (lit.
-"empty ear").
-From a linguistic point of view, Soramimi is a particular case of homophonic translation: words or phrases that appear in the speech stream in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
-Homophonic translation has mostly received attention as a perceptual curiosity, and as a literary figure, intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics]^[The rationale behind this translation approach is to "sacrifice[s] fidelity to the image rather than the melodiousness of verse" [@briusov1911ot].
-For instance, Edith Grossman's translation of *El Quijote* [@cervantes] present a similar case, in which formal aspects of the original text in Spanish are favoured to some degree over equivalence to English, even sometimes keeping Spanish words from the original text in the English translation (e.g., *Señor*, *ínsula*).].
+(*Coyote Ugly*, 2000) many of them mishear the line "I've got the power" as *Agatha Bauer*.
+This auditory illusion, which takes the name *Soramimi* (lit.
+"empty ear") in Japanese is common across languages and cultures [@dembeck2015oberflachenubersetzung; @efimova2018homophonic].
+Soramimi is a particular case of homophonic translation: words or phrases that appear in the speech stream in one language are translated into similar-sounding words and phrases in a different language, without necessarily preserving the meaning [@gasparov2006semen].
+Homophonic translation has mostly received attention as a perceptual curiosity or as a literary figure, intentionally employed by translators to preserve the meter and "sound structure" (e.g., rhymes) of an original text in the language of translation [e.g., @levick2016translating; @pilshchikov2016semiotics].
 Little attention has been paid to homophonic translation from a language processing perspective.
 
-One exception is the study by @otake2007interlingual, who analysed 194 instances of Soramimi, broadcasted between 1992 and 2007 by the Japanese TV show *Soramimi hour*, in which examples of Soramimi were presented to the audience for comedy purposes.
+One exception is the study by @otake2007interlingual, who analysed 194 instances of Soramimi from the Japanese TV show *Soramimi hour*, broadcasted between 1992 and 2007, in which examples of Soramimi were presented to the audience for comedy purposes.
 These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
-The vast majority of the reported instances were phrases (96%) and a few, single words (4%).
-There was relatively high variability in the degree to which phonetic features of the presented English lyrics were preserved, with some resulting Japanese words or phrases sharing high overlap with the original English strings of sounds, and some sharing very little overlap.
-An analysis of the few instances of homophonic translations of single words revealed three main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /\textipa{kraI}/ to *kurai* /\textipa{kuRai}/ [dark]), deletion (e.g., *go* /\textipa{goU}/ to *go* /go/ [*Go*, a board game], and alternation (e.g., *low* /\textipa{loU}/ to *rou* /\textipa{Roo}/ [wax]).
-This suggests that the reported homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to the Japanese phoneme inventory and Japanese phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
-
-The phonological processes described by @otake2007interlingual do not take place in a vaccum, but rather in the context of a lexicon.
-As words or phrases in the native lexicon are activated by their acoustic similarity with the non-native speech stream, they trigger a series of mechanisms that result in the selection of one of the candidate word or phrase in the native language.
+While the degree to which phonetic features of the presented English lyrics were in the translation varied considerably, the author identified tree main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /kɹaɪ̯/ to *kurai* /kɯ̟ɾa̠i/ [dark]), deletion (e.g., *go* /ɡoʊ/ to *go* /go/ [*Go*, a board game], and alternation (e.g., *low* /loʊ/ to *rou* /ɾo̞ː/ [wax]).
+These findings suggests that homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to Japanese phonology and phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
+
+Though rarely, homophonic translations sometimes preserve the meaning of the original word or phrase, even if the listener is completely unfamiliar with the language being spoken3.
+For example, an English native listening to Dutch for the first time might encounter the Dutch word /ˈpɪŋ.(ɡ)ʋɪn/ (*pinguin*) and be able to activate its correct English translation /ˈpɛŋɡwɪn/ (penguin), given how phonologically similar both word-forms are.
+This phonological similarity between translation equivalents---known as *cognateness*---is common across many languages, and is often due to typological closeness or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
+For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan than English words by capitalising solely on phonological similarity.
+
 In the present work, we investigated the interplay between phonological similarity and one of the main mechanisms involved in the dynamics of lexical selection: phonological neighbourhood density.
-The phonological neighbourhood density of a given word-form refers to the number of words-forms that differ in a single phoneme from the it.
+The phonological neighbourhood density of a given word-form refers to the number of words-forms that differ in a single phoneme from it.
 For instance, the English word *cat* belongs to a dense phonological neighbourhood, with 50 English phonological neighbours (e.g., hat, bat, cap), while the word *charger* belongs to a sparse phonological neighbourhood, with six neighbours (e.g., charged, larger, charter, charmer) [@marian2012clearpond].
-Phonological neighbourhood density affects lexical processing.
-For instance, words from dense neighbourhoods are recognized more slowly and less accuractely in the auditory modality than words from sparse neighbourhoods [@luce1998recognizing; @vitevitch1998words; e.g.,  @dufour2010phonological; @vitevitch2006clustering].
-In this study, we examined how listeners may activate lexical representations in their native language when listening to an unfamiliar language.
-
-Although homophonic translations rarely involve preservation of meaning, they sometimes lead to correct translations.
-For example, imagine an English native listening to Dutch for the first time.
-This listener may encounter the word /\textipa{"pINgUIn}/ (*pinguin*), which sounds similar to its English translation /\textipa{"pENgwIn}/ (penguin).
-Given enough phonological similarity between the non-native word and its translation in the native language, a naïve listener may succeed at word comprehension when listening to an unfamiliar language.
-Such phonological similarity between translation equivalents---known as *cognateness*---is common across many languages, and is often due to typological closeness and/or socio-historical events involving the speakers of these languages (e.g., migration, social contact).
-For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan *porta* than English *door*, due to the phonological and (orthographic similarity) of the former to the equivalent in their native language.
-
-To explore the lexical mechanisms triggered by unfamiliar speech listening, we used a translation elicitation task in which participants listened to words from an unfamiliar language, and then tried to guess the words translations in their native language.
-We will henceforth refer to the auditory-presented words heard by participants on each trial as *presented words*, and the correct translation for the presented words as *target translations*.
-By manipulating the amount of phonological similarity between the presented words and their target translations, we explored listeners' reliance on phonological similarity to succeed in the task.
-Since participants were unfamiliar with the presented language, they could only use phonological similarity between the presented language and their native language to successfully translate the words.
-We therefore predicted that participants’ performance should increase when the translation pairs are phonologically more similar.
-We also predicted that there is a minimum threshold of phonological similarity to be sufficient for correct translation.
-
-We further investigated the effect of phonological competition.
-Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language.
-Words with denser phonological neighbourhoods are recognised more slowly and less accurately than words with sparser phonological neighbourhoods [@dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing].
+Words from dense neighbourhoods are processed (e.g., recognised) more slowly and less accurately in the auditory modality than words from sparse neighbourhoods [e.g., @dufour2003lexical; @goldinger1989priming; @hamburger1996phonological; @luce1990similarity; @luce1998recognizing; @vitevitch1998words; @dufour2010phonological; @vitevitch2006clustering].
 This is especially true if such neighbours share higher phonological similarity with the presented word, or are lexically more frequent than the target word [@luce1998recognizing].
 Such interference effect of phonological neighbourhood density has also been reported across languages [@weber2004lexical].
-We hypothesised that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
-In the case of non-native speech recognition, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon.
-For instance, @otwinowska2019more reported that false friends were disadvantaged relative to non-cognates by Polish second language learners, in contrast to cognates which were known better.
-It is therefore important to investigate the joint effect of both cognates and false friends when investigating the effect that cross-linguistic phonological similarity has on word recognition in a foreign language.
-This could shed light on available strategies and challenges associated with the early stages of second language acquisition.
-For instance, one might expect English participants to incorrectly translate the Spanish word *botón* as *bottom* instead of as its correct translation *button*, due to the combined effect of the close phonological similarity between *botón* and *bottom* along with the high lexical frequency of *bottom* relative to *button*.
-To test this prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme apart from the presented word.
+In the case of non-native speech processing, the presence of cross-linguistic pairs which are phonologically similar but differ in meaning (e.g., false friends) may act as distractors during lexical access, obstructing the selection of the appropriate target translation in the listener’s lexicon [@otwinowska2019more].
+
+We designed a translation elicitation task in which participants listened to words from an unfamiliar language (*presented words*), and then tried to guess the words translations in their native language (*target translations*).
+Since participants were unfamiliar with the presented language, they could only rely on phonological similarity between the presented language and their native language to successfully translate the words, instead of on previously learned word-meaning associations.
+We predicted that participants’ performance should increase when the translation pairs are phonologically more similar.
+A second prediction was that the size of the facilitatory role of phonological similarity would be inversely proportional to the amount of higher-frequency phonological neighbours of the presented word in the target language.
+Even if the presented word and its target translation share high phonological similarity, participants may still provide incorrect translations if the presented word also shares high phonological similarity with other words in the native language, paralelling the dynamics of lexical selection engaged during native speech processing.
+To test this second prediction, we developed a lexical frequency-dependent measure of cross-language phonological neighbourhood density, in which a neighbour is counted only if it is higher frequency and is one phoneme apart from the presented word.
 If the phonological neighbourhood density of the target translation affects participants’ performance negatively, this would suggest that competitors in the native language affect recognition of non-native words during recognition of foreign speech.
-We conducted a series of three experiments to test these predictions.
-
-In Experiment 1, we collected data from two groups of British English natives living in the United Kingdom.
+
+# Experiment 1
+
+This study was aimed at exploring the interplay between cross-linguistic similarity and phonological neighbourhood and their impact in the performance of two groups of British English natives in the translation elicitation task.
 One group was presented with Catalan words, the other with Spanish words.
 We examined the extent to which participants were able to use the phonological similarity between the presented word (in Catalan or Spanish) and its target translation to provide accurate responses.
-In Experiment 2, we tested a group of (European) Spanish natives in the same task, who were presented with Catalan words.
-Catalan and Spanish are two Romance languages whose close typological distance is reflected in the fact that they share many cognates, where English is a Germanic language that shares considerably fewer cognates with Catalan and Spanish.
-By testing participants translating words from typologically close or distant languages, we expected to widen the range of the phonological similarity scores of the translation pairs involved in the experimental task, therefore allowing us to explore potential cross-language differences in participant’s performance.
-One unexpected finding was that participants in Experiments 1 and 2 were surprisingly good at translating a subset of words which had low phonological similarity with their correct translation.
-We were concerned that this may be caused by some prior knowledge of specific words by our participants, as some Spanish words are commonly seen in media or product labels, making them familiar even to monolingual speakers of English.
-In Experiment 3, we collected additional data from a new group of British English natives.
-The design was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
-In the final section of this article, we present analyses on the joint data sets of Experiments 1 to 3.
-
-# Experiment 1
 
@@
 
-All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/](https://osf.io/9fjxm/?view_only=aab7636ce1af48cf832596a7ea9101c5/) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
-For reproducibility, a Docker image of the RStudio session is available on DockerHub ([https://hub.docker.com/repository/docker/gongcastro/translation-elicitation/general](https://hub.docker.com/repository/docker/gongcastro/translation-elicitation/general)).
+All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/?view_only=751d6ab27ffd4dd4856e4c07c4372991](https://osf.io/9fjxm/?view_only=751d6ab27ffd4dd4856e4c07c4372991) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
+For reproducibility, a Docker image of the RStudio session is available on DockerHub ([LINK REDACTED FOR ANONIMITY]).
 
@@
 ```
-We collected data from `r n_participants` British English-native adults living in the United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female[1]` female).
+
+We collected data from `r n_participants` British English-native adults living in the United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female[1]` female).
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
@@
 
-```{r tbl-participants}
+```{r}
 #| label: tbl-participants
-#| tbl-cap: "Participant details."
+#| tbl-cap: "Participant details. *N* indicates the number of included participants, with excluded participants in parenthesis."
 tbl_data <- participants |>
-  mutate(exp = case_when(
+  mutate(
+    exp = case_when(
       source == "Experiment" & group != "cat-SPA" ~ "Experiment 1",
@@
       source == "Questionnaire" ~ "Experiment 3",
-  )) |>
+    )
+  ) |>
   summarise(
@@
     ),
-    n_excluded = paste0(n, " (", n_excluded, ")")
-  ) |>
-  select(-n)
+    n_excluded = paste0(n, " (", n_excluded, ")"),
+    age_mean_sd = paste0(round(age_mean, 2), " (", round(age_sd, 2), ")"),
+    age_min_max = paste0(age_min, "--", age_max)
+  ) |>
+  select(group, n_excluded, age_mean_sd, age_min_max, l2_lst)
+
+colnames(tbl_data) <- c(
+  "",
+  "N",
+  "Age: Mean (SD)",
+  "Age: Range",
+  "Second language"
+)
 
 tbl_data |>
-  gt(groupname_col = "exp", rowname_col = "group") |>
-  fmt_number(matches("age")) |>
-  fmt_integer(matches("min|max")) |>
-  cols_merge_uncert(age_mean, age_sd) |>
-  cols_merge_range(age_min, age_max) |>
-  tab_spanner("Age", matches("age")) |>
-  cols_label(
-    group = "Group",
-    exp = "Experiment",
-    n_excluded = "N",
-    age_mean = "Mean ± SD ",
-    age_min = "Range",
-    l2_lst = "Second language"
-  ) |>
-  tab_style(
-    cell_text(style = "italic"),
-    cells_column_labels()
-  ) |>
-  tab_style(
-    cell_text(weight = "bold"),
-    cells_column_spanners()
-  ) |>
-  tab_style(
-    cell_text(align = "left"),
-    list(cells_body(), cells_column_labels())
-  ) |>
-  tab_style(
-    cell_text(
-      weight = "bold",
-      align = "center",
-      style = "normal"
+  tt(width = c(0.175, 0.15, 0.2, 0.15, 0.325)) |>
+  format_tt(
+    j = 3:4,
+    digits = 2,
+    num_zero = TRUE,
+    num_fmt = "decimal"
+  ) |>
+  group_tt(
+    i = list(
+      "Experiment 1" = 1,
+      "Experiment 2" = 3,
+      "Experiment 3" = 4
     ),
-    cells_column_labels(c(l2_lst, n_excluded))
-  ) |>
-  tab_footnote(
-    "Number of included participants (number of excluded participants.)",
-    locations = cells_column_labels(n_excluded)
   )
@@
 durations <- stimuli |>
-  summarise(across(duration, lst(mean, sd, min, max)),
-    .by = group
-  )
+  summarise(across(duration, lst(mean, sd, min, max)), .by = group)
 durations <- split(durations, durations$group)
@@
 We created two lists of words to be presented to participants in the auditory modality: one in Catalan and one in Spanish.
-Words in the Catalan list were `r round(lengths$cat_eng$ipa_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_1_sd, 2)`, Range = `r round(lengths$cat_eng$ipa_1_min, 2)`-`r round(lengths$cat_eng$ipa_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, Range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`).
-Words in the Spanish list were `r round(lengths$spa_eng$ipa_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_1_sd, 2)`, Range = `r round(lengths$spa_eng$ipa_1_min, 2)`-`r round(lengths$spa_eng$ipa_1_max, 2)`), and the orthographic form of their English translations were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, Range = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`).
-
-```{r exc-stimuli}
+Words in the Catalan list were `r round(lengths$cat_eng$ipa_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$cat_eng$ipa_1_sd, 2)`, range = `r round(lengths$cat_eng$ipa_1_min, 2)`-`r round(lengths$cat_eng$ipa_1_max, 2)`), and the orthographic forms of their English translations (which participants had to type) were `r round(lengths$cat_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$cat_eng$word_2_sd, 2)`, range = `r round(lengths$cat_eng$word_2_min, 2)`-`r round(lengths$cat_eng$word_2_max, 2)`).
+Words in the Spanish list were `r round(lengths$spa_eng$ipa_1_mean, 2)` phonemes long on average (*SD* = `r round(lengths$spa_eng$ipa_1_sd, 2)`, range = `r round(lengths$spa_eng$ipa_1_min, 2)`-`r round(lengths$spa_eng$ipa_1_max, 2)`), and the orthographic form of their English translations were `r round(lengths$spa_eng$word_2_mean, 2)` characters long on average (*SD* = `r round(lengths$spa_eng$word_2_sd, 2)`, range = `r round(lengths$spa_eng$word_2_min, 2)`-`r round(lengths$spa_eng$word_2_max, 2)`).
+
+```{r}
+#| label: exc-stimuli
 stim_exc_freq <- stimuli |>
@@
 In each trial, participants listened to one audio file, which contained a single word.
-The audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+Audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
 These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
-Catalan and Spanish words were recorded at 44,100 Hz in separate files in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
-The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, Range = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
-The average duration of the Spanish audio files was `r round(durations$spa_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, Range = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
-
-For each word pair in the Catalan and Spanish lists, we defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word and its correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
-*Frequency* was included as a nuisance predictor, under the hypothesis that---keeping other predictors constant---participants would provide correct responses more often when the correct translations are high frequency words in their native language.
-Lexical frequencies of correct translations were extracted from SUBTLEX-UK [@van2014subtlex], and transformed to Zipf scores.
+Catalan and Spanish words were recorded at 44,100 Hz in the same session, and then de-noised using Audacity and normalised at peak intensity using Praat [@broersma2021praat].
+The average duration of the Catalan audio files was `r round(durations$cat_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$cat_eng$duration_sd, 2)`, range = `r round(durations$cat_eng$duration_min, 2)`-`r round(durations$cat_eng$duration_max, 2)`).
+The average duration of the Spanish audio files was `r round(durations$spa_eng$duration_mean, 2)` seconds (*SD* = `r round(durations$spa_eng$duration_sd, 2)`, range = `r round(durations$spa_eng$duration_min, 2)`-`r round(durations$spa_eng$duration_max, 2)`).
+
+We defined three predictors of interest: the lexical frequency of the correct English translation (*Frequency*), the phonological similarity between the presented word and its correct English translation (*Similarity*), and the presented word's number of cross-language phonological neighbours (*CLPN*) in English.
+*Frequency* was included as a nuisance predictor, under the hypothesis that participants would provide correct responses more often when the correct translations are high frequency words in their native language, keeping other predictors constant.
+Lexical frequencies of correct English and Spanish translations were extracted from SUBTLEX-UK [@van2014subtlex] and SUBTLEX-ESP [@cuetos2011subtlexesp], respectively, and transformed to Zipf scores.
 Translations without a lexical frequency value were excluded from data analysis (`r stim_exc_freq$cat_eng$n` in the Catalan list, `r stim_exc_freq$spa_eng$n` in the Spanish list).
-*Similarity*, our main predictor of interest, was calculated as the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
+*Similarity* was calculated as the Levenshtein similarity between the X-SAMPA transcriptions of each pair of translations using the `stringdist` R package [@van2014stringdist].
 The Levenshtein distance computes the edit distance between two character strings---in this case, two phonological transcriptions---by counting the number of additions, deletions, and substitutions necessary to make both strings identical [@levenshtein1966binary].
@@
 We calculated the number of *CLPN* for each Catalan and Spanish word by counting the number of English words with same or higher lexical frequency, and whose phonological transcription (in X-SAMPA format) differed by up to one phoneme from that of the presented Catalan or Spanish word.
-Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)]].
+Lexical frequencies and phonological transcriptions were extracted from the multilingual database CLEARPOND [@marian2012clearpond]^[Phonological transcriptions in CLEARPOND were generated from eSPEAK, [http://espeak.sourceforge.net/](http://espeak.sourceforge.net/)].
 @tbl-stimuli summarises the lexical frequency, *CLPN* and phonological overlap of the words included in the Catalan and the Spanish lists.
 
-```{r tbl-stimuli}
+```{r}
 #| label: tbl-stimuli
 #| tbl-cap: "Stimuli details."
-stimuli |>
+tbl_data <- stimuli |>
+  drop_na(freq_zipf_2, lv, neigh_n_h) |>
   summarise(
-    across(c(freq_zipf_2, lv, neigh_n_h),
-      lst(mean, sd, min, max),
-      na.rm = TRUE
+    across(c(freq_zipf_2, lv, neigh_n_h), list(mean, sd, min, max)),
+    .by = c(group)
+  )
+
+colnames(tbl_data) <- c(
+  "",
+  "M",
+  "SD",
+  "Min",
+  "Max",
+  "M",
+  "SD",
+  "Min",
+  "Max",
+  "M",
+  "SD",
+  "Min",
+  "Max"
+)
+
+tbl_data |>
+  tt(
+    width = c(
+      0.16,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07
+    )
+  ) |>
+  format_tt(
+    j = 2:ncol(tbl_data),
+    digits = 2,
+    num_zero = TRUE,
+    num_fmt = "decimal"
+  ) |>
+  group_tt(
+    j = list(
+      "Lexical frequency" = 2:5,
+      "Phonological simlarity" = 6:9,
+      "CLPN" = 10:13
     ),
-    .by = c(group)
-  ) |>
-  gt(rowname_col = "group") |>
-  tab_spanner("Frequency", matches("freq")) |>
-  tab_spanner("Similarity", matches("lv")) |>
-  tab_spanner("CLPN", matches("neigh")) |>
-  fmt_number(is.numeric) |>
-  fmt_integer(c(
-    neigh_n_h_min,
-    neigh_n_h_max
-  )) |>
-  cols_merge_uncert(freq_zipf_2_mean, freq_zipf_2_sd) |>
-  cols_merge_uncert(lv_mean, lv_sd) |>
-  cols_merge_uncert(neigh_n_h_mean, neigh_n_h_sd) |>
-  cols_merge_range(freq_zipf_2_min, freq_zipf_2_max) |>
-  cols_merge_range(lv_min, lv_max) |>
-  cols_merge_range(neigh_n_h_min, neigh_n_h_max) |>
-  cols_label(
-    freq_zipf_2_mean = "Mean ± SD",
-    freq_zipf_2_sd = "SD",
-    freq_zipf_2_min = "Range",
-    lv_mean = "Mean ± SD",
-    lv_sd = "SD",
-    lv_min = "Range",
-    neigh_n_h_mean = "Mean ± SD",
-    neigh_n_h_sd = "SD",
-    neigh_n_h_min = "Range"
-  ) |>
-  tab_style(
-    cell_text(weight = "bold"),
-    cells_column_spanners()
-  ) |>
-  tab_style(
-    cell_text(style = "italic"),
-    cells_column_labels()
-  ) |>
-  grand_summary_rows(
-    columns = matches("mean|sd"),
-    fns = list(
-      label = md("**Mean**"),
-      id = "totals",
-      fn = "mean"
-    ),
-    fmt = ~ fmt_number(., n_sigfig = 3, suffixing = TRUE)
   )
@@
 
-Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
+@fig-design illustrates the time course of a trial in the translation elicitation task.Each trial started with a yellow fixation point presented during one second on the centre of the screen over a black background.
 After one second, the audio started playing while the dot remained being displayed until the audio ended.
-Upon the offset of the fixation point and audio, participants were prompted to write their answer by a `>` symbol.
+Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol.
 Typed letters were displayed on the screen in real time to provide visual feed-back to participants.
@@
 Then, participants pressed the `RETURN/ENTER` key to confirm their answer and start a new trial.
-@fig-design illustrates the time course of a trial in the translation elicitation task.
 
@@
 After the offset of the audio, the fixation point disappeared and a visual prompt (`>`) was presented.
-Participants then wrote their answer and clicked `RETURN`, making the end of the trial.](_assets/img/design.png){#fig-design}
+Participants then wrote their answer and clicked `RETURN`, making the end of the trial.](../img/design.png){#fig-design}
 
@@
 
-After data collection, participants' answers were manually coded into the following categories: *Correct*, *Typo*, *Wrong*, *False friend*, *Other*.
+After data collection, participants' answers were manually coded into the following categories: *Correct*, *Typo*, *Wrong*, *Other*.
 A response was coded as *Correct* if the provided string of characters was identical to the orthographic form of the correct translation.
-A response was coded as *Typo* if the participant provided a string of characters that was only one edit distance (addition, deletion, or substitution) apart from the orthographic form of the correct translation (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to a distinct English word.
-A response was coded as *False friend* if the participant's response was incorrect, but phonologically similar to the presented word.
+A response was coded as *Typo* if the participant provided a string of characters that was only one edit distance (addition, deletion, or substitution) apart from the orthographic form of the correct translation (e.g., "pengiun" instead of "penguin"), as long as the response did not correspond to an extant English word.
 Responses not meeting the criteria for previous categories were labelled as *Wrong* or *Other* (see Data analysis section for more details).
-Both *Correct* and *Typo* responses were considered as correct, while *Wrong* and *False friend* responses were considered as incorrect.
+Both *Correct* and *Typo* responses were considered as correct, while *Wrong* responses were considered as incorrect.
 *Other* responses were excluded from data analysis.
@@
 We also included participant-level random intercepts and slopes for the main effects and the interaction.
-Eq.
-1 shows a formal description of the model.
+@eq-1 shows a formal description of the model.
+Numeric predictors were standardised before entering the model by subtracting the mean and dividing by the standard deviation.
 
@@
 &\textbf{Likelihood}  \\
-y_{i} \sim & \text{Bernoulli}(p_{i}) \\ \\
+y_{i} \sim & \text{Bernoulli}(p_{i}) \\
 &\textbf{Parameters}  \\
@@
 &\beta_{5[p]} (\text{Similarity}_i \times \text{CLPN}_i) + \beta_{6[p]} (\text{Similarity}_i \times \text{Group}_i) + \\
-& \beta_{7[p]} (\text{CLPN}_i \times \text{Group}_i) + \beta_{8[p]} (\text{Similarity}_i \times \text{CLPN}_i \times \text{Group}_i) \\ \\
+& \beta_{7[p]} (\text{CLPN}_i \times \text{Group}_i) + \beta_{8[p]} (\text{Similarity}_i \times \text{CLPN}_i \times \text{Group}_i) \\
 \beta_{0-8[p,w]} \sim & \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \text{ and  word } w \text{ in 1, ..., } W \\
-\beta_{1-8[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\ \\
+\beta_{1-8[p]} \sim &  \mathcal{N}(\mu_{\beta_{j}}, \sigma_{\beta_{j}}) \text{, for participant } p \text{ in 1, ..., } P \\
 &\textbf{Prior}  \\
@@
 We used the tidyverse family of R packages [@wickham2019welcome] to process data and to generate figures.
-We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models (see Appendix 1 for model diagnostics).
-Numeric predictors were standardised before entering the model by subtracting the mean and dividing by the standard deviation.
+We used the `brms` R package [@burkner2017brms] using the `cmdstanr` back-end to the Stan probabilistic language [@carpenter2017stan] to estimate and compare the models.
 
@@
 
-```{r dataset-1}
+```{r}
 #| label: dataset-1
@@
 excluded_lang <- participants_1 |>
-  filter(l2_writ_prod > 3 |
+  filter(
+    l2_writ_prod > 3 |
       l2_oral_comp > 3 |
@@
       spa_oral_comp > 3 |
-    spa_writ_prod > 3) |>
+      spa_writ_prod > 3
+  ) |>
   nrow()
@@
 Of those trials, `r format(nrow(dataset_clean$cat_eng), big.mark = ",")` were provided by `r dplyr::n_distinct(dataset_clean$cat_eng$participant_id)` participants who listened to Catalan words, and `r format(nrow(dataset_clean$spa_eng), big.mark = ",")` trials were provided by `r dplyr::n_distinct(dataset_clean$spa_eng$participant_id)` participants who listened to Spanish words.
-Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, Range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, Range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
-
-```{r coefs-1}
+Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
+
+```{r}
 #| label: coefs-1
 c1 <- gather_draws(exp_1_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
-  summarise(across(.value, lst(
+  summarise(across(
+    .value,
+    lst(
       .median = median,
@@
   mutate(correct = prop_adj(correct, n_trials)) |>
-  summarise(across(correct, lst(mean, sd, sem, min, max)),
+  summarise(
+    across(correct, lst(mean, sd, sem, min, max)),
     across(n_trials, lst(mean, sum, sd, min, max)),
@@
   ) |>
-  relocate(n_participants, matches("correct"))
+  relocate(exp, group, n_participants, matches("correct")) |>
+  select(-exp)
+
+colnames(tbl_data) <- c(
+  "",
+  "N",
+  "M",
+  "SD",
+  "SE",
+  "Min",
+  "Max",
+  "M",
+  "N trials",
+  "SD",
+  "Min",
+  "Max"
+)
 
 tbl_data |>
-  gt(
-    rowname_col = "group",
-    groupname_col = "exp"
-  ) |>
-  fmt_number(is.numeric, decimals = 2) |>
-  fmt_integer(c(matches("sum|min|max"), n_participants)) |>
-  fmt_number(matches("correct"), scale_by = 100) |>
-  tab_spanner("Accuracy (%)", matches("correct")) |>
-  tab_spanner("Valid trials", matches("n_trials")) |>
-  cols_merge_range(n_trials_min, n_trials_max) |>
-  cols_merge_range(correct_min, correct_max) |>
-  cols_label(
-    n_trials_sum = "N trials",
-    n_participants = "N",
-    n_trials_mean = "Mean",
-    n_trials_sd = "SD",
-    n_trials_min = "Range",
-    correct_mean = "Mean",
-    correct_sd = "SD",
-    correct_sem = "SE",
-    correct_min = "Range"
-  ) |>
-  tab_style(
-    cell_text(align = "center"),
-    cells_column_labels()
-  ) |>
-  tab_style(
-    cell_text(style = "italic"),
-    cells_column_labels()
-  ) |>
-  tab_style(
-    cell_text(weight = "bold"),
-    cells_column_spanners()
-  ) |>
-  tab_style(
-    cell_text(align = "left"),
-    list(cells_body(), cells_column_labels())
-  )
-```
-
-```{r fig-epreds-1}
+  tt(
+    width = c(
+      0.19,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.07,
+      0.11,
+      0.07,
+      0.07,
+      0.07
+    )
+  ) |>
+  group_tt(
+    i = list("Experiment 1" = 1, "Experiment 2" = 3, "Experiment 3" = 5),
+    j = list("Accuracy (%)" = 3:7, "Valid trials" = 8:12)
+  ) |>
+  format_tt(
+    j = 3:10,
+    digits = 2,
+    num_zero = TRUE,
+    num_mark_big = ",",
+    num_fmt = "decimal"
+  )
+```
+
+```{r}
 #| label: fig-epreds-1
@@
 Lines indicate mean predictions, and intervals indicate 95%, 89%, 78%, 67%, and 50% credible intervals (CrI)."
-get_epreds <- function(model, data, n = 100,
+get_epreds <- function(
+  model,
+  data,
+  n = 100,
   lv = seq(0, 1, length.out = 100),
   neigh_n_h = c(0, 2, 4, 8, 12),
-                       ...) {
+  ...
+) {
   lv_std <- (lv - mean(data$lv, na.rm = TRUE)) / sd(data$lv, na.rm = TRUE)
-  neigh_n_h_std <- (neigh_n_h - mean(data$neigh_n_h, na.rm = TRUE)) / sd(data$neigh_n_h, na.rm = TRUE) 
+  neigh_n_h_std <- (neigh_n_h - mean(data$neigh_n_h, na.rm = TRUE)) /
+    sd(data$neigh_n_h, na.rm = TRUE)
   knowledge <- unique(model$data$knowledge)
@@
   experiment <- unique(model$data$experiment)
-  nd <- expand_grid(freq_zipf_2_std, neigh_n_h_std, lv_std, knowledge, confidence, group, experiment)
+  nd <- expand_grid(
+    freq_zipf_2_std,
+    neigh_n_h_std,
+    lv_std,
+    knowledge,
+    confidence,
+    group,
+    experiment
+  )
   epreds <- tidybayes::add_epred_draws(nd, model, re_formula = NA, ...)
@@
 lv <- seq(0, 1, length.out = 100)
-lv_std <- (lv - mean(dataset_1$lv_std, na.rm = TRUE)) / sd(dataset_1$lv_std, na.rm = TRUE)
+lv_std <- (lv - mean(dataset_1$lv_std, na.rm = TRUE)) /
+  sd(dataset_1$lv_std, na.rm = TRUE)
 neigh_n_h <- c(0, 2, 4, 8, 12)
 epreds_1 <- get_epreds(exp_1_m0, dataset_1, neigh_n_h = neigh_n_h, lv = lv)
-epreds_1$neigh_n_h_std <- factor(epreds_1$neigh_n_h_std, levels = unique(epreds_1$neigh_n_h_std), labels = paste0(neigh_n_h, " neighbours"), ordered = TRUE)
+epreds_1$neigh_n_h_std <- factor(
+  epreds_1$neigh_n_h_std,
+  levels = unique(epreds_1$neigh_n_h_std),
+  labels = paste0(neigh_n_h, " neighbours"),
+  ordered = TRUE
+)
 
@@
   facet_grid(group ~ neigh_n_h_std) +
-  stat_lineribbon(aes(fill_ramp = after_stat(level)),
+  stat_lineribbon(
     linewidth = 1 / 2,
@@
     y = "p(Correct)",
-    colour = "Cross-language Phonological Neighbours",
-    fill = "Cross-language Phonological Neighbours",
-    linetype = "Cross-language Phonological Neighbours",
-    fill_ramp = "CrI (%)"
   ) +
   scale_x_continuous(
-    labels = \(x) scales::percent((x * sd(dataset_1$lv, na.rm = TRUE)) + mean(dataset_1$lv, na.rm = TRUE)), 
-    breaks = (seq(0, 1, 0.2) - mean(dataset_1$lv, na.rm = TRUE)) / sd(dataset_1$lv, na.rm = TRUE)) +
+    labels = \(x) {
+      scales::percent(
+        (x * sd(dataset_1$lv, na.rm = TRUE)) + mean(dataset_1$lv, na.rm = TRUE)
+      )
+    },
+    breaks = (seq(0, 1, 0.25) - mean(dataset_1$lv, na.rm = TRUE)) /
+      sd(dataset_1$lv, na.rm = TRUE)
+  ) +
   scale_y_continuous(labels = scales::percent) +
@@
 
-## Discussion
-
-In Experiment 1, we investigated the extent to which the phonological similarity between translation equivalents is sufficient for successful word translation, in the absence of conceptual knowledge about the presented word.
-We tested two groups of monolingual British English-native adults in a translation task that involved words in Catalan or Spanish, two languages participants reported having no prior familiarity with.
-Participants benefited strongly from phonological similarity when the correct translation of the presented words in Catalan or Spanish had few English phonological neighbours with higher lexical frequency.
-As the number of higher-frequency English phonological neighbours increased, both participants' accuracy and the impact of phonological similarity approached zero.
-This suggests that word-forms in an unfamiliar language have a strong potential to activate their translation equivalents in the native language, provided some phonological similarity between both words, and the absence of more frequent phonological neighbors.
-
-Participants in Experiment 1 were surprisingly good at translating words from Catalan and Spanish (two unfamiliar languages) to their native language.
-If English participants were likely to activate the correct English translations of the presented words in Catalan and Spanish, it is possible that speakers of typologically closer languages to Catalan and Spanish may benefit even more strongly from phonological similarity in the same task.
-English is a Germanic language (like Dutch or German), while Catalan and Spanish are Romance languages (like Italian, French, Portuguese).
-English shares fewer phonologically similar translations with Romance languages than Romance languages share with each other.
-It is possible that the probability of homophonic translations is higher when listening to an unfamiliar language from the same typological family as the native language.
-We tested this hypothesis in Experiment 2.
-
 # Experiment 2
@@
 
-```{r participants-numbers-2}
+```{r}
 #| label: participants-numbers-2
@@
 
-We collected data from `r n_participants` Spanish native adults living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+We collected data from `r n_participants` Spanish native adults living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
@@
 
-```{r dataset-2}
+```{r}
 #| label: dataset-2
@@
 excluded_lang <- participants_2 |>
-  filter(l2_writ_prod > 3 |
+  filter(
+    l2_writ_prod > 3 |
       l2_oral_comp > 3 |
@@
 After applying trial-level and participant-level inclusion criteria, the resulting dataset included `r format(nrow(dataset_clean$cat_spa), big.mark = ",")` trials provided by `r dplyr::n_distinct(dataset_clean$cat_spa$participant_id) + dplyr::n_distinct(dataset_clean$cat_spa$participant_id)` participants.
-Responses given by participants were `r round(lengths$len_mean, 2)` characters long on average (*Median* = `r round(lengths$len_median, 2)`, *SD* = `r round(lengths$len_sd, 2)`, Range = `r round(lengths$len_min, 2)`-`r round(lengths$len_max, 2)`).
-
-```{r posterior-2}
+Responses given by participants were `r round(lengths$len_mean, 2)` characters long on average (*Median* = `r round(lengths$len_median, 2)`, *SD* = `r round(lengths$len_sd, 2)`, range = `r round(lengths$len_min, 2)`-`r round(lengths$len_max, 2)`).
+
+```{r}
+#| label: posterior-2
 c2 <- gather_draws(exp_2_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
-  summarise(across(.value, lst(
+  summarise(across(
+    .value,
+    lst(
       .median = median,
@@
 
-```{r fig-epreds-2}
+```{r}
 #| label: fig-epreds-2
@@
 #| fig-width: 10
-lv <- seq(0, 1, length.out = 100)
-lv_std <- (lv - mean(dataset_2$lv_std, na.rm = TRUE)) / sd(dataset_2$lv_std, na.rm = TRUE)
+lv_std <- (lv - mean(dataset_2$lv_std, na.rm = TRUE)) /
+  sd(dataset_2$lv_std, na.rm = TRUE)
 neigh_n_h <- c(0, 2, 4, 8, 12)
 epreds_2 <- get_epreds(exp_2_m0, dataset_2, neigh_n_h = neigh_n_h, lv = lv)
-epreds_2$neigh_n_h_std <- factor(epreds_2$neigh_n_h_std, levels = unique(epreds_2$neigh_n_h_std), labels = paste0(neigh_n_h, " neighbours"), ordered = TRUE)
+epreds_2$neigh_n_h_std <- factor(
+  epreds_2$neigh_n_h_std,
+  levels = unique(epreds_2$neigh_n_h_std),
+  labels = paste0(neigh_n_h, " neighbours"),
+  ordered = TRUE
+)
 
@@
   ggplot(aes(lv_std, .epred)) +
-  facet_grid(~ neigh_n_h_std) +
-  stat_lineribbon(aes(fill_ramp = after_stat(level)),
+  facet_grid(~neigh_n_h_std) +
+  stat_lineribbon(
     linewidth = 1 / 2,
@@
     y = "p(Correct)",
-    colour = "Cross-language Phonological Neighbours",
-    fill = "Cross-language Phonological Neighbours",
-    linetype = "Cross-language Phonological Neighbours",
-    fill_ramp = "CrI (%)"
   ) +
   scale_x_continuous(
-    labels = \(x) scales::percent((x * sd(dataset_2$lv, na.rm = TRUE)) + mean(dataset_2$lv, na.rm = TRUE)), 
-    breaks = (seq(0, 1, 0.2) - mean(dataset_2$lv, na.rm = TRUE)) / sd(dataset_2$lv, na.rm = TRUE)) +
+    labels = \(x) {
+      scales::percent(
+        (x * sd(dataset_2$lv, na.rm = TRUE)) + mean(dataset_2$lv, na.rm = TRUE)
+      )
+    },
+    breaks = (seq(0, 1, 0.25) - mean(dataset_2$lv, na.rm = TRUE)) /
+      sd(dataset_2$lv, na.rm = TRUE)
+  ) +
   scale_y_continuous(labels = scales::percent) +
@@
 
-```{r posterior-12}
+```{r}
+#| label: posterior-12
 c12 <- gather_draws(exp_12_m0, `b_.*`, `sd_.*`, regex = TRUE) |>
-  summarise(across(.value, lst(
+  summarise(across(
+    .value,
+    lst(
       .median = median,
@@
 
-## Discussion
-
-Experiment 2 was an extension of Experiment 1 to a population of monolinguals whose native language is typologically close to the presented language.
-We presented Catalan words to Spanish native adults who were reportedly unfamiliar with Catalan.
-Our results indicate a similar pattern of results as those in Experiment 1: participants were able to provide correct translations of presented Catalan words, provided that the Catalan words shared some degree of phonological similarity with their Spanish translation, and that the number of phonological neighbours with higher lexical frequency was reduced.
-In contrast with the results in Experiment 1, the positive impact of phonological similarity on participants' performance in Experiment 2 larger.
-Spanish natives in Experiment 2 exploited phonological similarity more strongly than English natives in Experiment 1.
-Overall, this suggests that participants in Experiment 2, who were natives of a typologically similar language (Spanish) to the presented language (Catalan), benefited more strongly from phonological similarity than participants in Experiment 1, who were natives of a typologically less similar language (English) to the presented language (Catalan, Spanish).
+```{r}
+#| label: tbl-surprises
+#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarity score between the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
+tbl_data <- list(
+  "Experiment 1" = dataset_1,
+  "Experiment 2" = dataset_2
+) |>
+  bind_rows(.id = "exp") |>
+  mutate(
+    translation = paste0(word_1, " - ", word_2),
+    ipa = paste0("/", ipa_1, "/ - /", ipa_2, "/")
+  ) |>
+  summarise(
+    n = n(),
+    correct = sum(correct),
+    .by = c(exp, group, lv, translation, ipa)
+  ) |>
+  mutate(
+    accuracy = prop_adj(correct, n),
+    accuracy_se = prop_adj_se(correct, n)
+  ) |>
+  dplyr::filter(lv == 0, accuracy >= 0.1) |>
+  select(exp, translation, ipa, group, accuracy, accuracy_se) |>
+  arrange(exp, group, desc(accuracy)) |>
+  select(translation, ipa, accuracy, accuracy_se) |>
+  mutate(
+    accuracy = round(accuracy * 100, 2),
+    accuracy_se = round(accuracy_se * 100, 2)
+  )
+
+names(tbl_data) <- c("Translation", "IPA", "Accuracy (\\%)", "SE")
+
+tt(tbl_data) |>
+  group_tt(
+    i = list(
+      "Experiment 1 (cat-ENG)" = 1,
+      "Experiment 2 (spa-ENG)" = 6,
+      "Experiment 3 (cat-SPA)" = 22
+    )
+  )
+```
+
+# Experiment 3
+
 Participants from both Experiment 1 and 2 benefited strongly from phonological similarity to correctly translate words from a non-native, reportedly unfamiliar language.
@@
 @tbl-surprises shows a list of Catalan and Spanish words to which participants provided responses with $\geq$ 10 average accuracy.
-
-```{r}
-#| label: tbl-surprises
-#| tbl-cap: "List of items with unexpectedly high accuracy: the Levenshtein similarity score between the presented word (in Catalan or Spanish) and their correct English translation is zero, but participants, who are reportedly unfamiliar with the presented language, were on average >10% likely to guess the correct translation."
-ipa_to_tipa <- function(ipa) {
-  dict <- c(
-    "ɑ" = "A", 
-    "ɒ" = "6",
-    "æ" = "\\ae", 
-    "ʌ" = "2", 
-    "β" = "B",
-    "ʤ" = "}\\\\textdyoghlig\\\\textipa{",
-    "d͡ʒ" = "}\\\\textdyoghlig\\\\textipa{",
-    "dʒ" = "}\\\\textdyoghlig\\\\textipa{", 
-    "ð" = "D",
-    "ə" = "@",
-    "ɛ" = "E",
-    "ɣ" = "G",
-    "ɡ" = "g",
-    "ɫ" = "\\|~l", 
-    "ʎ" = "}\\\\textlambda\\\\textipa{",
-    "ŋ" = "N",
-    "ɲ" = "}\\\\textltailn\\\\textipa{", 
-    "ɔ" = "O",
-    "ɾ" = "R",
-    "ʧ" = "}\\\\textteshlig\\\\textipa{", 
-    "tʃ" = "}\\\\textteshlig\\\\textipa{", 
-    "t͡ʃ" = "}\\\\textteshlig\\\\textipa{", 
-    "θ" = "T", 
-    "t͡s" = "}\\\\texttslig\\\\textipa{", 
-    "ts" = "}\\\\texttslig\\\\textipa{", 
-    "ʃ" = "S", 
-    "ʒ" = "Z", 
-    "ɪ" = "I", 
-    "ʊ" = "U", 
-    "ɜ" = "3", 
-    "ɹ" = "\\*r", 
-    "ː" = ":", 
-    "ˈ" = '"'
-  )
-  tipa <- stringr::str_replace_all(ipa, dict)
-  tipa <- paste0("\\textipa{", tipa, "}")
-  return(tipa)
-}
-
-tbl_data <- list(
-  "Experiment 1" = dataset_1,
-  "Experiment 2" = dataset_2
-) |>
-  bind_rows(.id = "exp") |>
-  mutate(
-    translation = paste0(word_1, " - ",  word_2),
-    ipa = ipa_to_tipa(paste0("/", ipa_1, "/ - /", ipa_2, "/"))
-  ) |> 
-  summarise(
-    n = n(),
-    correct = sum(correct),
-    .by = c(exp, group, lv, translation, ipa)
-  ) |>
-  mutate(
-    accuracy = prop_adj(correct, n),
-    accuracy_se = prop_adj_se(correct, n)
-  ) |>
-  dplyr::filter(lv == 0, accuracy >= 0.1) |>
-  select(exp, translation, ipa, group, accuracy, accuracy_se) |>
-  arrange(exp, group, desc(accuracy)) |> 
-  select(translation, ipa, accuracy, accuracy_se) |> 
-  mutate(accuracy = round(accuracy * 100, 2),
-         accuracy_se = round(accuracy_se * 100, 2))
-
-names(tbl_data) <- c("Translation", "IPA", "Accuracy (\\%)", "SE")
-
-tt(tbl_data) |> 
-  style_tt(i = 0, bold = TRUE, line = "tb") |> 
-  group_tt(i = list(
-    "Experiment 1 (cat-ENG)" = 1,
-    "Experiment 2 (spa-ENG)" = 6,
-    "Experiment 3 (cat-SPA)" = 22)) |> 
-  style_tt(i = 1, bold = TRUE, line = "b") |> 
-  style_tt(i = 7, bold = TRUE, line = "tb") |> 
-  style_tt(i = 24, bold = TRUE, line = "tb")
-```
-
-It is likely that participants had prior knowledge of these words despite having reported little to no familiarity with the presented language (Catalan or Spanish).
+It is possible that participants had prior knowledge of these words despite having reported little to no familiarity with the presented language (Catalan or Spanish).
 One possibility is that participants had previously encountered these words embedded in English linguistic input.
@@
 In the case of higher-similarity words, it is more difficult to disentangle the extent to which participants' accuracy is a function of pure phonological similarity, or prior knowledge they had about the meaning of Spanish words.
-Experiment 3 was addressed at investigating the issue of prior knowledge.
-
-# Experiment 3
-
-Experiment 3 is a replication of Experiment 1, in which we collected additional data about participants' prior familiarity with the presented Catalan and Spanish words, in addition to the same translation task presented to participants in Experiment 1.
+Experiment 3 is a replication of Experiment 1 in which we collected additional data about participants' prior familiarity with the presented Catalan and Spanish words.
+We used this information to remove from analyses any responses in which participants reported prior knowledge with the meaning of the presented word.
 
@@
 
-```{r participants-numbers-3}
+```{r}
 #| label: participants-numbers-3
-participants_3 <- quest_participants |>
+participants_3 <- que_participants |>
   filter(group != "cat-SPA")
@@
 
-We collected data from `r n_participants` British English native adults living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, Range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+We collected data from `r n_participants` British English native adults living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
@@
 
-```{r participants-excluded-3}
+```{r}
 #| label: participants-excluded-3
 excluded_lang <- participants_3 |>
-  filter(l2_writ_prod > 3 |
+  filter(
+    l2_writ_prod > 3 |
       l2_oral_comp > 3 |
@@
 They were first asked whether they knew the presented word (multiple choice---*yes*/*no*).
-Regardless of their answer on the first question, participants were asked what they thought the translation of the word was in English (or their best guess), and instructed to type their answer in the provided text box.
-Finally, they were asked to rate how confident they were in their answer on a scale of 0 to 7, where 7 was "very confident" and 0 was "not confident".
-There was no time limit on the response phase.
-All questions had to be answered to proceed to the next trial.
-
 Participants first completed 5 practice trials with English words as the audio stimulus (ambulance, cucumber, elephant, pear, turtle).
@@
 
-```{r dataset-3}
+```{r}
 #| label: dataset-3
-dataset <- split(quest_responses, quest_responses$group)
+dataset <- split(que_responses, que_responses$group)
 names(dataset) <- janitor::make_clean_names(names(dataset))
@@
 
-dataset <- quest_responses |>
+dataset <- que_responses |>
   filter(group != "cat-SPA") |>
@@
 excluded_lang <- participants_3 |>
-  filter(l2_writ_prod > 3 |
+  filter(
+    l2_writ_prod > 3 |
       l2_oral_comp > 3 |
@@
 
-dataset_clean <- quest_responses |>
+dataset_clean <- que_responses |>
   filter(group != "cat-SPA") |>
@@
 
-```{r knowledge}
+```{r}
+#| label: knowledge
 knowledge_cateng <- dataset_clean$cat_eng |> count(knowledge)
@@
   filter(knowledge) |>
-  summarise(confidence_mean = mean(confidence),
-  confidence_sd = sd(confidence))
+  summarise(confidence_mean = mean(confidence), confidence_sd = sd(confidence))
 
@@
   filter(!knowledge) |>
-  summarise(confidence_mean = mean(confidence),
-  confidence_sd = sd(confidence))
+  summarise(confidence_mean = mean(confidence), confidence_sd = sd(confidence))
 
@@
   filter(!knowledge) |>
-  summarise(confidence_mean = mean(confidence),
-  confidence_sd = sd(confidence))
+  summarise(confidence_mean = mean(confidence), confidence_sd = sd(confidence))
 ```
@@
 
-```{r lengths-2}
+```{r}
+#| label: lengths-2
 lengths <- dataset_3 |>
@@
 
-Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, Range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, Range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
-
-Overall, participants reported prior knowledge more often for the Spanish words that showed unexpectedly high accuracy in Experiment 1 (see Discussion in Experiment 2) than for words with expected accuracy (see @fig-knowledge).
+Responses given by English participants to Catalan presented words were `r round(lengths$cat_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$cat_eng$len_median, 2)`, *SD* = `r round(lengths$cat_eng$len_sd, 2)`, range = `r round(lengths$cat_eng$len_min, 2)`-`r round(lengths$cat_eng$len_max, 2)`), while their translations to Spanish responses were `r round(lengths$spa_eng$len_mean, 2)` characters long on average (*Median* = `r round(lengths$spa_eng$len_median, 2)`, *SD* = `r round(lengths$spa_eng$len_sd, 2)`, range = `r round(lengths$spa_eng$len_min, 2)`-`r round(lengths$spa_eng$len_max, 2)`).
+
+Overall, participants reported prior knowledge more often for the Spanish words that showed unexpectedly high accuracy in Experiment 1 than for words with expected accuracy (see @fig-knowledge).
 Participants reported prior knowledge of Catalan words with unexpected accuracy in Experiment 1 as often as those with expected accuracy.
@@
     accuracy = prop_adj(correct, n),
-  accuracy_binned = cut(accuracy, bin_interval, labels = bin_interval[-length(bin_interval)], include.lowest = TRUE, right = TRUE),
-    knowledge_binned = cut(knowledge, bin_interval, labels = as.numeric(bin_interval[-length(bin_interval)]), include.lowest = TRUE, right = TRUE)
+    accuracy_binned = cut(
+      accuracy,
+      bin_interval,
+      labels = bin_interval[-length(bin_interval)],
+      include.lowest = TRUE,
+      right = TRUE
+    ),
+    knowledge_binned = cut(
+      knowledge,
+      bin_interval,
+      labels = as.numeric(bin_interval[-length(bin_interval)]),
+      include.lowest = TRUE,
+      right = TRUE
+    )
   ) |>
@@
   ) |>
-      mutate(is_gt_20 = n > 20,
+  mutate(
+    is_gt_20 = n > 20,
     knowledge_binned = as.double(as.character(knowledge_binned)),
-      accuracy_binned = as.double(as.character(accuracy_binned))) |> 
-  ggplot(aes(knowledge_binned, accuracy_binned, fill = n, label = n, color = is_gt_20)) +
+    accuracy_binned = as.double(as.character(accuracy_binned))
+  ) |>
+  ggplot(aes(
+    knowledge_binned,
+    accuracy_binned,
+    fill = n,
+    label = n,
+    color = is_gt_20
+  )) +
   facet_wrap(~group) +
@@
   ) +
-  scale_fill_distiller(palette = "Reds", direction=1, limits = c(0, 37)) +
+  scale_fill_distiller(palette = "Reds", direction = 1, limits = c(0, 37)) +
   scale_color_manual(values = c("black", "white")) +
-  scale_x_continuous(labels = scales::percent, limits = c(-0.05, 1)) +
-  scale_y_continuous(labels = scales::percent, limits = c(-0.05, 1), breaks = seq(0, 1, 0.1)) +
+  scale_x_continuous(
+    labels = scales::percent,
+    limits = c(-0.05, 1),
+    breaks = seq(0, 1, 0.25)
+  ) +
+  scale_y_continuous(
+    labels = scales::percent,
+    limits = c(-0.05, 1),
+    breaks = seq(0, 1, 0.1)
+  ) +
   theme(
@@
 
-```{r draws-3}
+```{r}
+#| label:  draws-3
 c3 <- gather_draws(exp_3_m1, `b_.*`, `sd_.*`, regex = TRUE) |>
-  summarise(across(.value, lst(
+  summarise(across(
+    .value,
+    lst(
       .median = median,
@@
 lv <- seq(0, 1, length.out = 100)
-lv_std <- (lv - mean(dataset_3$lv_std, na.rm = TRUE)) / sd(dataset_3$lv_std, na.rm = TRUE)
+lv_std <- (lv - mean(dataset_3$lv_std, na.rm = TRUE)) /
+  sd(dataset_3$lv_std, na.rm = TRUE)
 neigh_n_h <- c(0, 2, 4, 8, 12)
 epreds_3 <- get_epreds(exp_3_m1, dataset_3, neigh_n_h = neigh_n_h, lv = lv)
-epreds_3$neigh_n_h_std <- factor(epreds_3$neigh_n_h_std, levels = unique(epreds_3$neigh_n_h_std), labels = paste0(neigh_n_h, " neighbours"), ordered = TRUE)
+epreds_3$neigh_n_h_std <- factor(
+  epreds_3$neigh_n_h_std,
+  levels = unique(epreds_3$neigh_n_h_std),
+  labels = paste0(neigh_n_h, " neighbours"),
+  ordered = TRUE
+)
 
@@
     y = "p(Correct)",
-    colour = "Cross-language Phonological Neighbours",
-    fill = "Cross-language Phonological Neighbours",
-    linetype = "Cross-language Phonological Neighbours",
-    fill_ramp = "CrI (%)"
   ) +
   scale_x_continuous(
-    labels = \(x) scales::percent((x * sd(dataset_3$lv, na.rm = TRUE)) + mean(dataset_3$lv, na.rm = TRUE)), 
-    breaks = (seq(0, 1, 0.2) - mean(dataset_3$lv, na.rm = TRUE)) / sd(dataset_3$lv, na.rm = TRUE)) +
+    labels = \(x) {
+      scales::percent(
+        (x * sd(dataset_3$lv, na.rm = TRUE)) + mean(dataset_3$lv, na.rm = TRUE)
+      )
+    },
+    breaks = (seq(0, 1, 0.25) - mean(dataset_3$lv, na.rm = TRUE)) /
+      sd(dataset_3$lv, na.rm = TRUE)
+  ) +
   scale_y_continuous(labels = scales::percent) +
@@
 
-## Discussion
-
-Experiment 3 was a conceptual replication of Experiment 1 in which we gathered additional information about English native participants' prior familiarity with the meaning of the Catalan and Spanish words presented.
-After each trial, participants reported whether they had prior knowledge about the word meaning, and how confident they were about such information in a numeric scale from 0 to 7.
-To control for the effect of participants' prior knowledge about form-meaning mappings, we excluded from the analyses those words in which participants reported such prior knowledge.
-We found equivalent results to those in Experiment 1, indicating that participants' surprisingly good performance in the translation elicitation task was not due to being familiar with the meaning of words they were presented with.
-
 # General discussion
 
-The present work explored the lexical processing bases of homophonic translation, a phenomenon in which listening to speech in a non-native---perhaps unfamiliar---language leads to the activation of lexical representations in the native language, without necessarily preserving the meaning.
-We investigated how phonological similarity and its interaction with phonological neighbourhood density impact the dynamics of lexical activation and selection during non-native word processing.
-We designed a translation elicitation task in which participants listened to individual words in a non-native, unfamiliar language.
-After listening to each word, participants were asked to provide their best-guess translation for each word in their native language.
-In Experiment 1, British English-native adults listened to a list of words in Catalan or Spanish.
-Participants reported no prior familiarity with Catalan, Spanish, or any other Romance language, yet provided accurate translations for words that shared some degree of phonological similarity with their correct English translation.
-We calculated phonological similarity between word pairs as the Levenshtein similarity between the X-SAMPA transcription of their phonological forms [see @floccia2018introduction for a similar approach].
-Using this measure, we found that participants were able to exploit the phonological similarity between the presented words and their correct translation, even when both words-forms shared few phonemes.
-This points to listeners accommodating non-native phonological forms to their native phoneme inventory, resulting in the activation of native lexical representations.
-
-Experiment 2 was aimed at extending our findings to a population of participants whose native language was typologically closer to the unfamiliar language presented.
-We tested a group of Spanish native adults who listened to a list of Catalan words.
-Again, participants reported no prior familiarity with Catalan, or any other Romance language other than Spanish.
-In line with the results with English natives, we found an interaction between phonological similarity and phonological neighbourhood size.
-Spanish natives in Experiment 2 exploited phonological similarity more strongly than English natives in Experiment 1: the association between phonological similarity and the probability of correct translation was stronger in Experiment 2, compared to in Experiment 1.
-
-Experiment 3 was a replication of Experiment 1, in which we collected additional information about participants' prior familiarity with the presented Catalan and Spanish words.
-This follow-up experiment was designed to address concerns that participants in Experiments 1 and 2 might have managed to provide correct translations for some words thanks to having prior experience with those words.
-The design of Experiment 3 was closely modelled after Experiment 1, except that after providing their response in each trial, participants reported whether they had previous knowledge of the presented word.
-After removing responses where participants reported prior knowledge about the meaning of the presented Catalan or Spanish word, we ran the same analyses as in Experiment 1, and found parallel results.
-Overall, our results suggest that participants were able to rely on the phonological similarity between the presented Catalan and Spanish words and their correct translation to guide word recognition in an unfamiliar language.
+Using a translation elicitation task, we investigated how phonological similarity and its interaction with phonological neighbourhood density impact the dynamics of lexical activation and selection during non-native word processing.
+Our findings support an account of homophonic translation in which thhihs phenomenon results from listeners accommodating non-native phonological forms to their native phoneme inventory, resulting in the activation of native lexical representations.
+In Experiment 1, we found that English natives activate English word-forms that share some phonological similarity with auditorily presented Catalan and Spanish words (both unfamiliar languages to participants), and later engage dynamics of lexical competition and selection similar to those engaged in native speech processing.
+In Experiment 2 we extended these findings to a population of participants whose native language was typologically closer to the unfamiliar language presented.
+Spanish native adults who were unfamiliar with Catalan listened to a list of Catalan words.
+Spanish natives in Experiment 2 exploited phonological similarity even more strongly than English natives in Experiment 1, suggesting that the typological distance between the presented and target language is associated with a more robust facilitation effect of phonological similarity.
+In Experiment 3, we replicated the results from Experiment 1 after removing responses in which participants reported some prior knowlegment about the meaning of the presented word, showing that participants' surprisingly good performance in the translation elicitation task was not driven by unreported prior experience with the Catalan and Spanish words.
 
@@
 Future studies may explore the relative contribution of such cues in the occurrence of homophonic translation.
-Overall, the present study provides some insights into the role of phonological similarity in the auditory presentation of words in an unfamiliar language.
-
-Our findings also suggest that the facilitation effect of phonological similarity was moderated by phonological neighbourhood density.
-This is in line with previous studies suggesting that the presence of (high-frequency) phonological neighbours interferes with the lexical selection in word comprehension tasks [e.g., @dufour2010phonological; @luce1998recognizing; @vitevitch1999probabilistic; @vitevitch1998words; @grainger1990word].
-Our translation elicitation task can be understood as a particular instance of an auditory word recognition task, in which the presented unfamiliar words have the potential to activate phonological neighbours in participants' lexicon.
-We calculated a measure of cross-linguistic neighbourhood size by counting, for each presented word the number of phonologically related words (with a Levenshtein distance of one) in the target language (English in Experiments 1 and 3, Spanish in Experiment 2).
-To account for the fact that competition between phonological neighbours is sensitive to lexical frequency---high-frequency neighbours produce stronger interference effects [@luce1998recognizing; @dufour2010phonological]---we only counted neighbours with a lexical frequency higher than the correct translation.
-Using this measure, we found that participants' ability to exploit phonological similarity to produce correct translations declined as the number of phonological neighbours increased.
-This suggests that listening to words in an unfamiliar language triggered similar mechanisms of lexical activation and selection that listening to words in a native language does.
-The generalisability of these findings can be tested by increasing the repertoire of words involved in the translation elicitation task.
-In Experiments 1 to 3, participants answered to a maximum of 105 words.
-All words had high frequency and low age-of-acquisition.
-To better characterise the factors that guide listeners’ ability to match words from an unfamiliar language with words in their native language lexicon, words with varying levels of difficulty should be included in future studies.
-
-A comparison between the results in Experiments 1 and 2 suggested that the performance of Spanish participants translating Catalan was more resilient to the interfering effects of phonological neighbourhood density than the performance of English participants translating Catalan or Spanish.
-As highlighted before, Catalan and Spanish (both Romance languages) are typologically closer to each other than to English (a Germanic language).
-Catalan and Spanish share a higher proportion of cognates than Catalan and English, or Spanish and English.
-Overall, our findings suggest that the typological distance between the presented and target language is associated with a more robust facilitation effect of phonological similarity.
-
-The specific mechanisms behind this effect are unclear.
-One possibility is that there were subphonemic features of similarity between Catalan and Spanish words that were not adequately captured by our relatively coarse measure of similarity.
-Phonological transcriptions like X-SAMPA are abstract representations that ignore non-phonological contrasts (e.g., phones), and consider different symbols as completely different phonemes, disregarding the fact that some phonemes share more phonetic features than others.
-Additionally, prosodic cues such as lexical stress, which our measure of phonological similarity does not include, might also provide participants with further information about the correct translation of the presented word-forms.
-There may be more of these subphonemic cues available between the typologically close Spanish and Catalan than between the more distant English and Spanish, or English and Catalan.
-In addition, it remains to be tested whether the effect of typological distance is linear (i.e., robustness of facilitation increases linearly with typological distance), proportional, and whether other close or distant language pairs show the same pattern of results.
-Future studies may consider comparing the suitability of other measures of phonological similarity that take into account finer-grained phonetic or prosodic cues present in the acoustic signal, and explore the relative contribution of such cues in the occurrence of homophonic translation.
+Another possible limitation of the present study is the limited size and diversity of the all of which had high lexical frequency and low age-of-acquisition.
+To better characterise the factors that guide listeners’ ability to match words from an unfamiliar language with words in their native language lexicon, words with more varying levels of difficulty should be included in future studies.
 

### 2025-07-25 _targets.R

@@ -1,8 +1,11 @@
+library(targets)
 library(tarchetypes)
 library(here)
 library(dplyr)
 library(tidyr)
 library(stringr)
+library(readr)
+library(purrr)
 library(ggplot2)
 library(ggdist)
 library(beeswarm)
@@ -12,41 +15,26 @@ library(cmdstanr)
 library(stringdist)
 library(cli)
 library(quarto)
-library(readr)
-library(purrr)
+
 
 # load functions
 invisible({
   lapply(
-    list.files("R", pattern = ".R$", full.names = TRUE),
+    list.files("scripts", pattern = ".R$", full.names = TRUE),
     \(x) source(x, encoding = "UTF-8")
   )
 })
 
 # set number of cores to use with brms
-options(
-  mc.cores = 4,
-  brms.backend = "cmdstanr"
-)
+options(mc.cores = 4, brms.backend = "cmdstanr")
 
-clearpond_path <- here("data", "raw", "cleapond")
-exp_data_path <- here("data", "raw", "experiment")
-que_data_path <- here("data", "raw", "questionnaire")
 
 list(
   # CLEARPOND database -----
-  tar_target(
-    clearpond,
-    import_clearpond(
-      here(clearpond_path, "englishCPdatabase2.txt"),
-      here(clearpond_path, "spanishCPdatabase2.txt"),
-      here(clearpond_path, , "clearpondHeaders_EN.txt"),
-      here(clearpond_path, "clearpondHeaders_SP.txt"),
-    )
-  ),
+  tar_target(clearpond, import_clearpond()),
 
   # stimuli -----
-  tar_target(trial_list_path, here("stimuli", "trials.xlsx"), format = "file"),
+  tar_target(trial_list_path, here("data", "trials.xlsx"), format = "file"),
   tar_target(trial_list, get_trial_list(trial_list_path)),
   # exclude some stimuli:
   # corona: problematic after COVID-19
@@ -63,7 +51,7 @@ list(
   tar_target(levenshtein, get_levenshtein(trial_list)),
 
   # get audio durations
-  tar_target(audios_path, here("stimuli", "sounds")),
+  tar_target(audios_path, here("sounds")),
   tar_target(durations, get_duration(trial_list, audios_path)),
 
   # join all stimuli data
@@ -77,23 +65,34 @@ list(
       stimuli_exclude = stimuli_exclude
     )
   ),
+  tar_target(stimuli_file, write_csv(stimuli, here("out", "stimuli.csv"))),
 
[…]

### 2025-07-25 scripts/stimuli.R

@@ -8,7 +8,7 @@ library(cli)
 #'
 import_clearpond <- function(
   cp_c_path_eng = here("data", "raw", "clearpond", "englishCPdatabase2.txt"),
-  cp_c_path_spa = here("data", "raw", "clearpond", "englishCPdatabase2.txt"),
+  cp_c_path_spa = here("data", "raw", "clearpond", "spanishCPdatabase2.txt"),
   cp_h_path_eng = here("data", "raw", "clearpond", "clearpondHeaders_EN.txt"),
   cp_h_path_spa = here("data", "raw", "clearpond", "clearpondHeaders_SP.txt")
 ) {

### 2025-07-28 docs/appendix.qmd
Mask preprint

@@
 
-```{r fig-appendix-diagnostics}
+```{r}
 #| label: fig-appendix-diagnostics
+#| fig-cap: "Traceplot of MCMC posterior samples of the fixed regression coefficients of the models in Experiments 1, 2, and 3."
 #| fig-width: 9
@@
 
-```{r fig-coefs}
+```{r}
 #| label: fig-coefs
+#| fig-cap: "Posterior distribution of fixed regression coefficients of the models in Experiments 1, 2, and 3."
 #| fig-width: 9

### 2025-07-28 docs/manuscript.qmd
Mask preprint

@@
 ---
-abstract:  "The present study characterises homophonic translation---a phenomenon by which non-native speech activates words or phrases in a listener's native language---from a language processing perspective.
+abstract:  "Homophonic translation is a popular perceptual phenomenon by which non-native speech activates words or phrases in a listener's native language.
+In this study, we characterise homophonic translation from a language processing perspective.
 Across three studies, English and Spanish-native monolingual adults completed a translation elicitation task in which participants listened to words in an unfamiliar language (Catalan or Spanish), and were asked to type their best-guess translation in their native language.
@@
 These instances consisted of homophonic translations of English song lyrics to words and phrases in Japanese.
-While the degree to which phonetic features of the presented English lyrics were in the translation varied considerably, the author identified tree main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /kɹaɪ̯/ to *kurai* /kɯ̟ɾa̠i/ [dark]), deletion (e.g., *go* /ɡoʊ/ to *go* /go/ [*Go*, a board game], and alternation (e.g., *low* /loʊ/ to *rou* /ɾo̞ː/ [wax]).
+While the degree to which phonetic features of the presented English lyrics were preserved in the translation varied considerably, Otake identified tree main phonological processes that could explain how Japanese listeners reconstructed the English input to generate the Japanese words: insertion (e.g., *cry* /kɹaɪ̯/ to *kurai* /kɯ̟ɾa̠i/ [dark]), deletion (e.g., *go* /ɡoʊ/ to *go* /go/ [*Go*, a board game], and alternation (e.g., *low* /loʊ/ to *rou* /ɾo̞ː/ [wax]).
 These findings suggests that homophonic translations can be explained, to some extent, as the result of the Japanese listeners accommodating the strings of English sounds to Japanese phonology and phonotactics [@peperkamp2008perceptual; @dupoux1999epenthetic].
 
-Though rarely, homophonic translations sometimes preserve the meaning of the original word or phrase, even if the listener is completely unfamiliar with the language being spoken3.
+Though rarely, homophonic translations sometimes preserve the meaning of the original word or phrase, even if the listener is completely unfamiliar with the language being spoken.
 For example, an English native listening to Dutch for the first time might encounter the Dutch word /ˈpɪŋ.(ɡ)ʋɪn/ (*pinguin*) and be able to activate its correct English translation /ˈpɛŋɡwɪn/ (penguin), given how phonologically similar both word-forms are.
@@
 For example, Romance languages such as Spanish and Catalan share many form-similar translation equivalents  [@schepens2012distributions], as in the case of *puerta* and *porta* (*door* in Spanish and Catalan, respectively).
-Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan than English words by capitalising solely on phonological similarity.
+Given no prior knowledge of either language, a Spanish native speaker is likely to be much more successful at correctly translating Catalan words than English words via homophonic translation, that is, by capitalising solely on phonological similarity.
 
@@
 
-All materials, data, and code used in this study are hosted in the Open Science Framework [https://osf.io/9fjxm/?view_only=751d6ab27ffd4dd4856e4c07c4372991](https://osf.io/9fjxm/?view_only=751d6ab27ffd4dd4856e4c07c4372991) and a GitHub repository [https://github.com/gongcastro/translation-elicitation.git](https://github.com/gongcastro/translation-elicitation.git), along with additional notes.
+All materials, data, and code used in this study are hosted in the Open Science Framework ([https://osf.io/9fjxm/?view_only=751d6ab27ffd4dd4856e4c07c4372991](https://osf.io/9fjxm/?view_only=751d6ab27ffd4dd4856e4c07c4372991)) and a GitHub repository ([https://anonymous.4open.science/r/translation-elicitation-8BCE](https://anonymous.4open.science/r/translation-elicitation-8BCE) [FOR ANONYMITY, LINK TEMPORARILY REDIRECTED TO AN ANONYMOUS MIRROR OF THE GITHUB REPOSITORY]), along with additional notes.
 For reproducibility, a Docker image of the RStudio session is available on DockerHub ([LINK REDACTED FOR ANONIMITY]).
@@
 Participants gave informed consent before providing any data.
-The study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+The study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was by the [ETHICAL COMMITTEE REDACTED FOR ANONYMITY].
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
@@
 #| label: tbl-participants
-#| tbl-cap: "Participant details. *N* indicates the number of included participants, with excluded participants in parenthesis."
+#| tbl-cap: "Participant information. *N* indicates the number of included participants, with excluded participants between parentheses."
 tbl_data <- participants |>
@@
 In each trial, participants listened to one audio file, which contained a single word.
-Audio files were the same ones used in child experiments conducted in the Laboratori de Recerca en Infància of Universitat Pompeu Fabra (Barcelona, Spain).
+Audio files were the same ones used in child experiments conducted in the [LABORATORY NAME REDACTED FOR ANONYMITY].
 These audio files were recorded by a proficient Catalan-Spanish female bilingual from the Metropolitan Area of Barcelona in a child-directed manner.
@@
 #| label: tbl-stimuli
-#| tbl-cap: "Stimuli details."
+#| tbl-cap: "Stimuli information. *M*: mean, *SD*: standard deviation, *Min*: minimum, *Max*: maximum."
 tbl_data <- stimuli |>
@@
       "Lexical frequency" = 2:5,
-      "Phonological simlarity" = 6:9,
+      "Phonological similarity" = 6:9,
       "CLPN" = 10:13
@@
 \mu_{\beta_{p,w}}  \sim &  \mathcal{N}(0, 0.1) \\
-\sigma_{\beta_{p}},  \sigma_{\beta_{w}} \sim & \text{HalfCauchy}(0, 0.1) \\
-\rho_{p}, \rho_{w} \sim & \text{LKJCorr}(8) \\
+\sigma_{\beta_{p}},  \sigma_{\beta_{w}} \sim & \text{Exponential}(3) \\
+\rho_{p}, \rho_{w} \sim & \text{LKJCorr}(5) \\
 \end{aligned}
@@
 #| fig-height: 5
-#| fig-width: 11
+#| fig-width: 10
 #| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
@@
 
-We collected data from `r n_participants` Spanish native adults living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+We collected data from `r n_participants` Spanish native adults living in Spain (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female[1]` female).
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
 Participants in Spain were contacted via announcements at the University campus(es), and were compensated €5 or an Amazon voucher for the same value.
-Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the Drug Research Ethical Committee (CEIm) of the IMIM Parc de Salut Mar (2020/9080/I).
+Participants gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the [ETHICAL COMMITTEE REDACTED FOR ANONYMITY].
 
@@
 #| cache: false
-#| fig-cap: "Posterior model-predicted mean accuracy in Experiment 1.
+#| fig-cap: "Posterior model-predicted mean accuracy in Experiment 2.
 Predictions were generated from 4,000 posterior samples, extracted for different values of *CLPN* (0, 2, 4, 8, 12) and *Similarity* (1-100).
@@
 
-We collected data from `r n_participants` British English native adults living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female` female).
+We collected data from `r n_participants` British English native adults living in United Kingdom (*Mean* = `r round(participants_age$age_mean, 2)` years, *SD* = `r round(participants_age$age_sd, 2)`, range = `r round(participants_age$age_min, 2)`-`r round(participants_age$age_max, 2)`, `r n_female[1]` female).
 Data collection took place from `r collection_dates[1]` to `r collection_dates[2]`.
-Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the University of Oxford Medical Sciences Inter-Divisional Research Ethics Committee (IDREC) (R60939/RE007).
+Participants were recruited via Prolific (5£ compensation) and SONA (compensation in academic credits), and gave informed consent before providing any data and the study was conducted in accordance with ethical standards of the Declaration of Helsinki and the protocol was approved by the [ETHICAL COMMITTEE REDACTED FOR ANONYMITY].
 Participants were asked to complete the experiment using a laptop in a quiet place with good internet connection.
@@
 
-Future studies may consider comparing the suitability of other measures of phonological similarity that take into account finer-grained phonetic or prosodic cues present in the acoustic signal presented to participants.
 Phonological transcriptions like X-SAMPA are abstract representations that ignore non-phonological contrasts (e.g., phones), and consider different symbols as completely different phonemes, disregarding the fact that some phonemes share more phonetic features than others.
@@
 Altogether, it is likely that participants in the present study were able to exploit additional cues in the acoustic signal to provide correct translations.
-Future studies may explore the relative contribution of such cues in the occurrence of homophonic translation.
+Future studies may consider a measure of phonological similarity that captures finer-grained phonetic or prosodic cues present in the acoustic signal presented to participants.
+explore the relative contribution of such cues in the occurrence of homophonic translation.
 Another possible limitation of the present study is the limited size and diversity of the all of which had high lexical frequency and low age-of-acquisition.
